### 2016-01-05 doc/invt.tex
initial commit

\documentclass[reprint]{revtex4-1}
\usepackage{amsmath}
\begin{document}

\title{Optimal updating factor in Wang-Landau and metadynamics simulations}

\maketitle

\section{Background}

\subsection{Flat-distribution sampling}

Given a system,
consider the problem of computing
the distribution, $p^*_i$
along a discrete quantity $i$.
%
The discrete nature of $i$ is natural
in discrete models.
%
For example, $i$ can be the energy $E$
in lattice models; or the temperature index
in a simulated tempering simulation.
%
For a continuous quantity, $z$,
which is typical in molecular systems,
we can also discretize $z$
such that the integer $i$ represents
the index of a small interval, or a bin, $(z, z + dz)$.
%
In both cases,
the distribution can be normalized as
$\sum_{i = 1}^n p^*_i = 1$.

For a large system,
the distribution $p^*_i$ is often
localized around some $i_0$,
%
and to find out the global property,
it is often desirable to carry out
a biased sampling that targets
a wider distribution $p_i$.
%
%Here, we refer to simulations that target
%a flat or nearly flat distribution
%as entropic or multicanonical sampling.

To do so, we need to introduce a bias potential $\phi_i$,
such that the target distribution is changed to
%
\begin{equation}
  \pi_i \propto p^*_i \, e^{-\phi_i}.
  \label{eq:pi_p_phi1}
\end{equation}
%
Upon normalization $\sum_{i = 1}^n \pi_i = 1$,
we get
%
\begin{equation}
  \pi_i =
  \frac{ p^*_i \, e^{-\phi_i} }
  { \sum_{j = 1}^n p^*_j \, e^{-\phi_j} }.
  \label{eq:pi_p_phi}
\end{equation}
%
Particularly,
to achieve a flat distribution $\pi_i$,
the bias potential $\phi_i$
must coincide with $\log p^*_i$,
up to an additive constant.

Generally, we may wish to target
a slowly-varying distribution, $p_i$,
by adjusting the bias potential $\phi_i$
on the fly.
%
For example,
we can monitor the histogram accumulated over
a period of time and use it for $\pi_i$ in
Eq. \eqref{eq:pi_p_phi1},
and then update
$\phi_i$ according to
$$
\phi_i^{\mathrm{new}}
=
\phi_i^{\mathrm{old}}
-
\log \frac{ p_i } { \pi_i }.
$$
%
%From Eq. \eqref{eq:pi_p_phi1},
%it is clear that to achieve the target distribution
%$\pi_i = p_i$ requires
%%
%\begin{equation}
%\log p^*_i
%=\phi_i + \ln p_i + c
%\label{eq:bias_pmf}
%\end{equation}
%%
%%
%In the end of the simulation,
%the PMF, $\phi^*_i$,
%can be deduced from Eq. \eqref{eq:bias_pmf}.
%
The above inversion process can be inconvenient
in practice,
for we need to make the period
as long as a multiple of
the autocorrelation time,
which is difficult to estimate for an unknown system.

\subsection{Wang-Landau algorithm}

The Wang-Landau (WL) algorithm is a convenient technique
of obtaining the desired bias potential $\phi_i$
on the fly.
%
To simplify the notation, we introduce
a shifted bias potential
%
\begin{equation}
  v_i \equiv \phi_i - \ln p^*_i + \ln p_i.
  \label{eq:v_def}
\end{equation}
%
%
and in $v_i$, Eq. \eqref{eq:pi_p_phi}
becomes
%
\begin{equation}
  \pi_i = \frac{ p_i e^{-v_i} }
  { \sum_{j = 1}^n p_j e^{-v_j} }
  \propto p_i \, e^{-v_i}.
  \label{eq:pi_p_v}
\end{equation}
%
According to Eq. \eqref{eq:pi_p_phi1},
$v_i$ should approach a constant of $i$
upon convergence (i.e., when $\pi_i \approx p_i$).

In the WL algorithm, $v_i$ are updated
in each MC step $t$,
and $v_i(t)$ as a function of $t$
is updated as
%
\begin{equation}
  v_i(t+1)
  =
  v_i(t)
  +
  \delta_{i, \, i(t)}
  \frac{ \alpha(t) } { p_i }.
  \label{eq:wl_update}
\end{equation}
%
where $i(t)$ is the bin at step $t$,
and $\alpha(t)$ is the updating magnitude.
%
Note that the updating scheme only applies
to the $v_i$ at the current bin.
%The values of $v_j(t)$ at $j \ne i(t)$
%are kept unchanged.
%
One can show that with a constant $\alpha(t) = \alpha > 0$,
the distribution collected from
the trajectory is identical to $p^*_i$.
%
However, the updating scheme
makes $v_i(t)$ a fluctuating quantity,
and Eq. \eqref{eq:pi_p_v} no longer holds
at all times.
%%
%The source of the deviation is two-fold.
%%
%First, since $v_i(t)$ is updated continuously,
%there is a random noise that is proportional
%to $\sqrt \alpha$.
%%
%Second, there is a systematic error
%that comes from the updating dynamics itself,
%as it breaks the Markovian nature
%that underlies Eq. \eqref{eq:pi_p_v}.
%%
%
Thus, one has to decrease $\alpha$ over time
to reduce the error of $v_i$,
in order to get the unbiased distribution
$\ln p_i^*$ via Eq. \eqref{eq:v_def}.

In the original WL scheme,
the updating magnitude $\alpha$ (or $\ln f$
in the original paper) is kept as a constant
for a period of time,
which is referred to as a stage below.
%
The histogram collected in the stage is monitored.
%
Once the histogram is sufficiently flat,
we are allowed to enter a new stage
with a reduced $\alpha$
(usually half as large as
that in the previous stage).
%
This scheme works well for early stages.
%
However, in later stages, it tends to reduce $\alpha$
too quickly, making the asymptotic error
saturate.

\subsection{$1/t$ formula}

A more effective way
of updating the $\alpha(t)$
is to discard the stage-wise setup,
and to follow the formula
%
\begin{equation}
  \alpha(t) = \frac{1}{t},
  \label{eq:alpha_invt}
\end{equation}
%
where $t$ is the number of steps
from the beginning of the MC simulation,
which shall be referred to as the ``time'' below.
%
This surprisingly simple formula has attracted
several studies, notably the one by Zhou.
%
There are, however, some reservations about
the constant of proportionality:
the optimal formula for $\alpha(t)$
might be $C/t$ with a constant $C$
different from $1.0$.

Here we shall derive the optimal $\alpha(t)$
for the WL algorithm, Eq. \eqref{eq:wl_update}.
%
Further, we shall generalize the result
to a class of multiple-bin updating schemes,
in which an update affects
not only the current bin $i(t)$,
but also a few neighboring ones.
%(such schemes are designed
%to maintain a smooth bias potential).
%
We shall show that
the constant of proportionality, $1$,
in Eq. \eqref{eq:alpha_invt}
is optimal for the single-bin updating scheme,
Eq. \eqref{eq:wl_update}
(as in the WL algorithm),
but not necessarily so
for a multiple-bin updating scheme.
%
Below, we discuss the single-bin updating scheme
in Section \ref{sec:single-bin},
and the multiple-bin case
in Section \ref{sec:multiple-bin}.

\section{\label{sec:single-bin}
Single-bin updating scheme}

In this section,
we shall derive the optimal $\alpha(t)$
for the single-bin updating scheme,
Eq. \eqref{eq:wl_update}.
%
To do so,
we shall first express the error of $v(t)$
as a functional of $\alpha(t)$,
and then find the optimal value
by functional variation.

\subsection{Differential equation}

We first approximate Eq. \eqref{eq:wl_update}
by a differential equation
%
\begin{equation}
  \dot v_i(t)
  =
  h_i(t) \frac{ \alpha(t) } { p_i },
  \label{eq:vt_diffeq}
\end{equation}
%
where
$\dot v_i(t) \equiv dv_i(t)/dt$,
%
and $h_i(t) = \delta_{i, i(t)}$
is the instantaneous histogram,
which is equal to $1.0$
for the current bin $i(t)$
or zero otherwise.

Next, we split $h_i(t)$ into a deterministic part
and a noise
%
\begin{equation}
  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
  \label{eq:h_split}
\end{equation}
%
Here, the deterministic part can be related
to the ``ensemble average'' of $h_i$.
%
The ensemble consists of many similar simulation copies
that have experienced the same schedule $\alpha(t)$
and have reached the same $v_i(t)$
at time $t$.
%
The initial states and random numbers of the copies
may, however, be different.

For sufficiently small $\alpha(t)$,
the sampling process is approximately
Markovian of a finite order,
and we may assume Eq. \eqref{eq:pi_p_v}
for the deterministic part
%
\begin{equation}
  \langle h_i(t) \rangle
  \approx
  \frac{ p_i \, e^{-v_i} }
  { \sum_{j = 1}^n p_j \, e^{-v_j} }.
  \label{eq:h_ave}
\end{equation}

For the noise part, we have
$\langle \zeta_i(t) \rangle = 0$.
%
%The noise is not necessarily white.
Unless specified otherwise, we assume
a general noise with the correlation
function depending on the time difference:
%
\begin{equation}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  \sigma_{ij}(t - t'),
  \label{eq:zeta_zeta_correlation}
\end{equation}
%
where $\sigma_{ij}(t)$ is an even function of $t$
that vanishes at large $t$.

\subsection{Linear approximation}

To proceed, we first observe that
the average $\bar v = \sum_{i = 1}^n p_i \, v_i$
increases steadily over time:
%
\begin{equation}
\frac{ d \bar v } { d t }
=
\sum_{i = 1}^n p_i \dot v_i
=
\alpha(t) \sum_{i = 1}^n h_i(t) = \alpha(t).
\label{eq:dvbardt}
\end{equation}
%
However, the difference between $v_i$ and $\bar v$,
%
\begin{equation}
  x_i \equiv v_i - \bar v = v_i - \sum_{j = 1}^n p_j \, v_j,
  \label{eq:x_def}
\end{equation}
%
is expected to be small in the asymptotic regime,
and we can expand Eq. \eqref{eq:h_ave} as
\begin{align}
\langle h_i(t) \rangle
&\approx
\frac{ p_i \, e^{- x_i} }
{ \sum_{ j = 1}^n p_j \, e^{- x_j} }
\approx
\frac{ p_i ( 1 - x_i ) }
{ \sum_{ j = 1}^n p_j (1 - x_j) }
\notag \\
&\approx
p_i \, \left(
  1 - x_i + \sum_{j=1}^n p_j \, x_j
\right)
=
p_i \, (1 - x_i),
\label{eq:hdet_linearize}
\end{align}
where we have used $\sum_{j=1}^n p_j = 1$,
and
%
\begin{equation}
  \sum_{i = 1}^n p_i \, x_i = 0,
  \label{eq:px_sum}
\end{equation}
which follows directly from Eq. \eqref{eq:x_def}.
%
With Eqs.
\eqref{eq:vt_diffeq},
\eqref{eq:h_split},
\eqref{eq:dvbardt},
and
\eqref{eq:hdet_linearize},
we get a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  -\alpha(t) \, \left[ x_i(t) - \frac{ \zeta_i(t) } { p_i } \right].
  \label{eq:dxdt_WL}
\end{equation}

The total error of the bias potentials can be written in
terms of $x_i$,
\begin{equation}
\mathcal E
=
\sum_{i = 1}^n p_i \, \langle x_i^2 \rangle
=
\sum_{i = 1}^n p_i \, \left\langle (v_i - \bar v)^2 \right\rangle,
\label{eq:error_sum}
\end{equation}
where
$\bar v = \sum_{i = 1}^n p_i v_i$.
%
Our aim is to find the $\alpha(t)$
that minimizes Eq. \eqref{eq:error_sum}.
%
To do so, we shall first consider
a simpler one-variable problem.

\subsection{One-variable problem}

Consider the following equation
of a single variable:
%
\begin{equation}
\dot x(t) = -\alpha(t) \, \lambda \left[ x(t) - \xi(t) \right],
\label{eq:dxdt_alpha}
\end{equation}
%
where $\xi(t)$ is a generalized noise
that is invariant under time translation:
%
\begin{equation}
\left\langle \xi(t) \, \xi(t') \right\rangle
=
\kappa(t - t').
\label{eq:noise_correlation}
\end{equation}
%
Additionally, we require
\begin{equation}
  \lim_{t \rightarrow \pm\infty} \kappa(t) = 0.
  \label{eq:kappat_limit}
\end{equation}
%
For example, for a white noise,
$\kappa(t)$ is proportional to
Dirac's $\delta$-function, $\delta(t)$.
%
We wish to show that the optimal $\alpha(t)$
of minimizing $\langle x^2(t) \rangle$ at long times
is given by
%
\begin{equation}
  \alpha(t) = \frac{1}{\lambda \, (t + t_0)}.
\label{eq:alpha_opt}
\end{equation}

To do so, we first recall
the formal solution of Eq. \eqref{eq:dxdt_alpha}:
%
\begin{equation}
x(t) = x(0) \, e^{-\lambda \, q(t)}
+ \int_0^t \dot u(t') \, \xi(t') \, dt',
\label{eq:xt_solution}
\end{equation}
%
where,
%
\begin{equation}
q(t) \equiv \int_0^t \alpha(t') \, dt',
\label{eq:qt_definition}
\end{equation}
%
and
%
\begin{align}
u(q(t'))
&\equiv
e^{-\lambda \, q(t) + \lambda \, q(t')}.
\label{eq:ut_definition}
\end{align}

We shall further demand that
%
\begin{equation}
  \lim_{t \to \infty} q(t) \to \infty.
  \label{eq:qt_limit}
\end{equation}
%
Then, for a large time $t$,
the first term on the right-hand side
of Eq. \eqref{eq:xt_solution} can be neglected, and
%
\begin{align}
\left\langle x^2(t) \right\rangle
%&=
%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%\notag
%\\
&=
\int_0^t \int_0^t
  \dot u(q(t')) \, \dot u(q(t'')) \,
  \kappa(t' - t'') \, dt'' \, dt'.
\label{eq:x2t_average}
\end{align}

Variating this expression and
using the Euler-Lagrange equation yields
$$
\begin{aligned}
0
&=
\frac{d}{d\tau} \int_0^t \dot u(q(t')) \, \kappa(t' - \tau) \, dt'
= \int_0^t \ddot u(q(t')) \, \kappa(t' - \tau) \, dt',
\end{aligned}
$$
where we have dropped the boundary terms
by using Eq. \eqref{eq:kappat_limit}.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
In order for this to hold for any $\tau$,
we must have
%
\begin{equation}
\ddot u(q(t')) = 0,
\qquad
\mathrm{or}
\;\;
\dot u(q(t')) = c,
\label{eq:ddu_eq_0}
\end{equation}
%
where $c$ is a constant of $t'$.
%
Using Eq. \eqref{eq:ut_definition}
we get
$$
e^{-\lambda \, q(t) + \lambda \, q(t')}
=
c \, (t' + t_0),
$$
where $t_0$ is a constant.
%
Taking the logarithm, and differentiating this with respect to $t'$
yields Eq. \eqref{eq:alpha_opt}.

\subsection{Optimal $\alpha(t)$ for the single-bin updating scheme}

Going back to the problem of
minimizing Eq. \eqref{eq:error_sum}
for the single-bin updating scheme,
we find from Eq. \eqref{eq:dxdt_WL}
that with $\lambda = 1$,
Eq. \eqref{eq:alpha_opt}
will optimize each term $\langle x_i^2 \rangle$,
and hence the sum.
%
Thus, Eq. \eqref{eq:alpha_invt} is optimal
up to a constant $t_0$ defining the origin of $t$.

\section{\label{sec:multiple-bin}
Multiple-bin update}

We now extend the above analysis to
the multiple-bin updating scheme.
%
In this case,
upon a visit to bin $j$,
we update not only $v_j$,
but also $v_i$ for a few neighboring bins $i$.
%
The update magnitudes are given by a matrix $\mathbf w$
with elements $w_{ij}$.
%
Then, Eq. \eqref{eq:vt_diffeq} is generalized to
\begin{equation}
  \dot v_i(t) =
  \sum_{j=1}^n \alpha(t) \, h_j(t) \frac{ w_{ij} } { p_j }.
  \label{eq:vt_diffeq_mbin}
\end{equation}

The procedure of optimization is largely
the same as the single-bin case,
whereas the matrix nature brings about
some difficulties that require
certain restrictions and approximations.

\subsection{Updating matrix $\mathbf w$}

The matrix $\mathbf w$ is not arbitrary.
%
A necessary condition to sample desired distribution
$\mathbf p = (p_1, \dots, p_n)$
is that in Eq. \eqref{eq:vt_diffeq_mbin}
when $h_j(t)$ coincides with $p_j$,
the rate of change $\dot v_i(t)$
should be independent of $i$.
%
This allows $v_i(t)$ to approach a constant
asymptotically.
%
Thus, by a proper scaling of $\alpha(t)$,
we may write this condition as
%
\begin{equation}
  \sum_{j = 1}^n w_{ij} = 1.
  \label{eq:w_sumj}
\end{equation}
%
This means $(1, \dots, 1)^T$
is a right eigenvector of $\mathbf w$
with eigenvalue $1$.
%
Thus, the transpose $\mathbf w^T$
resembles a transition matrix,
although the matrix elements can be negative.

To simplify the following discussion,
we shall further impose the
detailed balance condition:
%
\begin{equation}
  p_i \, w_{ij} = p_j \, w_{ji}.
  \label{eq:w_detailedbalance}
\end{equation}
%
It follows that
\begin{equation}
  \sum_{i = 1}^n p_i \, w_{ij}
  =
  \sum_{i = 1}^n p_j \, w_{ji}
  = p_j,
  \label{eq:w_balance}
\end{equation}
%
i.e., $\mathbf p$ is a left eigenvector of
$\mathbf w$ with eigenvalue $1$.

Further, by Eq. \eqref{eq:w_detailedbalance},
we can define a symmetric matrix $\hat{\mathbf w}$
as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$
that can be diagonalized
with a set of orthonormal eigenvectors:
%
$\sum_{i = 1}^n U_{ki} \, \hat w_{ij} = \lambda_k \, U_{kj}$.
%
Thus,
in terms of diagonalizing $\mathbf w$, we have
\begin{equation}
  \sum_{i = 1}^n T_{ki} \, w_{ij} = \lambda_k \, T_{kj},
  \label{eq:T_w}
\end{equation}
where
$T_{ij} = \sqrt{p_j} \, U_{ij}$ satisfies
the orthonormal conditions:
%
\begin{align}
\sum_{k = 1}^n T_{ki} \, T_{kj}
&= \delta_{ij} \, p_i,
\label{eq:T_orthonormal_cols}
\\
\sum_{k = 1}^n \frac{ T_{ik} \, T_{jk} }{ p_k }
&= \delta_{ij}.
\label{eq:T_orthonormal_rows}
\end{align}

We emphasize that, in a stable updating scheme,
all eigenvalues $\lambda_k$ must be nonnegative,
for otherwise there is a fluctuating mode
that can increase indefinitely.

\subsection{Linearization and decoupling}

We can now simplify Eq. \eqref{eq:vt_diffeq_mbin}.
%
By using Eq. \eqref{eq:w_balance},
we can show that
Eqs. \eqref{eq:dvbardt} and \eqref{eq:px_sum}
remain true.
%
%Thus, for Eq. \eqref{eq:x_def}, we have
%
$$
\begin{aligned}
\dot x_i
&= \alpha(t) \sum_{j=1}^n w_{ij}
\left( \frac{ h_j } { p_j }  - 1 \right)
\\
&=
-\alpha(t) \sum_{j = 1}^n
w_{ij} \left[ x_j(t) - \frac{\zeta_j (t)}{p_j} \right],
\end{aligned}
$$
where
we have expanded the right-hand side
in linear terms of $x_j$ in the second step.

Now by diagonalizing the matrix $\mathbf w$
using matrix defined in Eqs. \eqref{eq:T_w},
\eqref{eq:T_orthonormal_cols},
and
\eqref{eq:T_orthonormal_rows},
we get a set of decoupled equations
%
\begin{equation}
\dot y_i(t)
=
-\alpha(t) \, \lambda_i
[y_i(t) - \eta_i(t)],
\label{eq:yt_diffeq}
\end{equation}
%
where
\begin{align}
  y_i &= \sum_{j=1}^n T_{ij} \, x_j,
  \label{eq:y_def}
  \\
  \eta_i &= \sum_{j=1}^n T_{ij} \frac{ \zeta_j}{ p_j}.
  \label{eq:eta_def}
\end{align}

\subsection{Minimization of the error function}

The error function can be obtained
from Eq. \eqref{eq:T_orthonormal_cols}
as
\begin{align}
  \mathcal E
  &\equiv
  \sum_{i = 1}^n p_i \, x_i^2
  =
  \sum_{i, j, k=1}^n T_{ki} \, T_{kj} \, x_i \, x_j
  =
  \sum_{k = 1}^n y_k^2.
  \label{eq:y2_sum}
\end{align}
%
If we arrange $y_k$ such that
the mode $y_1 = \sum_{j=1}^n p_j \, x_j$,
(corresponding to $\lambda_1 = 1$),
then $y_1$ vanishes identically
by Eq. \eqref{eq:px_sum},
and the sum of $y_k$ in Eq. \eqref{eq:y2_sum}
can start from $k = 2$.

In analogous to Eq. \eqref{eq:x2t_average},
we can rewrite Eq. \eqref{eq:y2_sum} as
%
\begin{align}
  \mathcal E
  %&= \sum_{i = 2}^n \langle y_i^2 \rangle
  %\notag
  %\\
  &=
  \int_0^t \int_0^t
  \sum_{i = 2}^n
  \dot u_i(q(t')) \,
  \dot u_i(q(t'')) \,
  \kappa_i(t' - t'') \, dt' \, dt'',
  \notag
  %\label{eq:y2_int}
\end{align}
%
where
\begin{align*}
  u_i(q(t'))
  &\equiv
  e^{\lambda_i \, [q(t') - \, q(t)]},
  \\
  \kappa_i(t - t')
  &\equiv
  \left\langle \eta_i(t) \, \eta_i(t') \right\rangle,
\end{align*}
%
and the extremal condition of $\mathcal E$ is given by
%
\begin{align}
\sum_{i=2}^n
u_i'(q(\tau))
\int_0^t
\ddot u_i(q(t')) \, \kappa_i(t' - \tau) \, dt' = 0.
\label{eq:optimal_mbin1}
\end{align}
%
If all $\lambda_i$, hence all $u_i$, are the same,
the above condition can be satisfied
by Eq. \eqref{eq:ddu_eq_0},
which is the single-bin case.
%
The general solution of Eq. \eqref{eq:optimal_mbin1}
is harder to obtain.
%

\subsection{Optimal $\alpha(t)$}

To proceed, we shall further assume perfect sampling,
%
\begin{equation}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  p_i \, \delta_{ij} \, \delta(t - t').
  \label{eq:zz_perfect}
\end{equation}
%
Thus, from Eqs. \eqref{eq:T_orthonormal_rows} and \eqref{eq:eta_def},
we get
%
\begin{equation}
  \kappa_i(t - t')
  =
  \sum_{j,k = 1}^n
  \frac{ T_{ij} } { p_j }
  \frac{ T_{ik} } { p_k }
  \langle \zeta_j(t) \, \zeta_k(t') \rangle
  =
  \delta(t - t').
  \label{eq:kappa_perfect}
\end{equation}
%
Thus, Eq. \eqref{eq:optimal_mbin1}
is simplified as
%
\begin{align}
  \sum_{i=1}^n
  \ddot u_i(q(\tau)) \, u'_i(q(\tau)) = 0.
  \label{eq:optimal_mbin2}
\end{align}
%
%We can study two limiting cases.
%
For $\tau \ll t$,
the dominant contribution comes from
the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
and
$$
\ddot u_n(\tau) = 0,
\quad
\dot u_n(\tau) = c,
\quad
\mathrm{and}\;
u_n(\tau) = c \, (\tau + t_0),
$$
which means
$$
\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
$$

\subsection{Comparison of updating schemes}

A more direct, and often more accurate, approach
is to assume the functional form
\begin{equation}
\alpha(\tau) = \frac{1}{\lambda \, (\tau + t_0) },
\label{eq:alpha_trial}
\end{equation}
and find the optimal choice of $\lambda$.
%
Using Eqs. \eqref{eq:kappa_perfect}, \eqref{eq:alpha_trial}
in Eq. \eqref{eq:y2_sum}, we get
%
\begin{equation}
\mathcal E
=
\int_0^t
\sum_{i=0}^n \dot u_i^2(q(t')) \, dt'
\stackrel{t \to \infty}
{=\joinrel=\joinrel=}
\frac{1}{t}
\sum_{i = 2}^n
\frac{ \lambda_i^2 } { \lambda \, (2 \, \lambda_i - \lambda) },
\label{eq:y2_sum_perfect_invt}
\end{equation}
%
and the optimal $\lambda$
can be found from the smallest root of
%
\begin{equation}
\sum_{i = 2}^n
\frac{ \lambda_i - \lambda }
{ \left(2 - \lambda/ \lambda_i \right)^2 }
= 0.
\end{equation}

Under the approximation, we have
$$
\frac{ \lambda_i^2 }{ \lambda \, (2 \lambda_i - \lambda) } \ge 1,
$$
with the equality achieved only at $\lambda = \lambda_i$,
we see the from Eq. \eqref{eq:y2_sum_perfect_invt}
that the minimal error is achieved with
\begin{equation}
  \lambda_2 = \dots = \lambda_n = \lambda.
\end{equation}
%
and the updating matrix $\mathbf w$ satisfying this condition
is essentially a multiple of the identity matrix
(since $\lambda_1$ corresponds to the mode
of a uniform shift of all $v_i$,
we can freely set it to $\lambda_2$
without changing the nature of the updating scheme).
%
This shows that in terms of convergence,
the single-bin updating scheme
(as adopted in the WL algorithm)
is the most efficient candidate
of all multiple-bin schemes.

%\section{Numerical results}

%Here

\end{document}

### 2016-01-06 doc/invt.tex
add invtpar.h

@@
 \usepackage{amsmath}
+\usepackage{upgreek}
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{tikz}
+\usepackage{hyperref}
+
+\definecolor{DarkBlue}{RGB}{0,0,64}
+\definecolor{DarkBrown}{RGB}{64,20,10}
+\definecolor{DarkGreen}{RGB}{0,64,0}
+\definecolor{DarkPurple}{RGB}{64,0,42}
+% annotation macros
+\newcommand{\repl}[2]{{\color{gray} [#1] }{\color{blue} #2}}
+\newcommand{\add}[1]{{\color{blue} #1}}
+\newcommand{\del}[1]{{\color{gray} [#1]}}
+\newcommand{\note}[1]{{\color{DarkGreen}\footnotesize \textsc{Note.} #1}}
+\newcommand{\answer}[1]{{\color{DarkBlue}\footnotesize \textsc{Answer.} #1}}
+\newcommand{\summary}[1]{{\color{DarkPurple}\footnotesize \textsc{Summary.} #1}}
+
 \begin{document}
@@
 the distribution collected from
-the trajectory is identical to $p^*_i$.
+the trajectory is identical to $p_i$.
 %
@@
 
-The total error of the bias potentials can be written in
+The total error of the bias potential can be written in
 terms of $x_i$,
@@
 we have expanded the right-hand side
-in linear terms of $x_j$ in the second step.
+in linear terms of $x_j$.
 
@@
 \end{equation}
-and find the optimal choice of $\lambda$.
+and find the optimal value of $\lambda$.
 %
@@
 
-%\section{Numerical results}
+\subsection{Example}
 
-%Here
+As an example, consider the following matrix $\mathbf w$
+%
+\begin{equation}
+  w_{ij} = m_{|i-j|} + m_{i+j-1} + m_{2n+1-i-j},
+  \label{eq:w_band}
+\end{equation}
+%
+where $m_0, \dots, m_b \, (b < n)$
+are a set of numbers satisfying
+%
+\begin{equation}
+m_0 + 2 \, m_1 + \cdots + 2 \, m_b = 1,
+\label{eq:m_normalization}
+\end{equation}
+%
+and $m_l = 0$ for $l > b$.
+
+This matrix is clearly symmetric, and hence
+satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
+for the flat distribution $p_i \equiv 1/n$.
+%
+It admits $n$ eigenvectors,
+$\pmb\theta^{(1)}, \dots, \pmb\theta^{(n)}$,
+given by
+%
+\begin{equation}
+\theta^{(k)}_i
+= \cos \frac{ (2 \, i - 1) k \pi}{2 \, n}.
+\label{eq:wband_eigenvector}
+\end{equation}
+%
+with eigenvalues
+\note{To show Eqs.
+  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue},
+  we first define $m_l = 0$, for $l \le 0$, then
+$$
+\begin{aligned}
+&\sum_{i = 1}^n
+  \theta^{(k)}_i
+  \, w_{ij}
+=
+- \theta^{(k)}_j \, m_0
+\\
+&
++\sum_{i=1}^n
+  \theta^{(k)}_i
+  \left(
+    m_{i - j} + m_{j - i} + m_{i + j - 1} + m_{2n+1-i-j}
+  \right)
+.
+\end{aligned}
+$$
+
+Now
+$$
+\begin{aligned}
+  &
+  \sum_{i=1}^n
+  \theta^{(k)}_i \,
+  \left(
+    m_{i - j} + m_{2n+1-i-j}
+  \right)
+  \\
+  &=
+  \sum_{l=1-j}^{n-j} \theta^{(k)}_{j+l} \, m_l
+  +
+  \sum_{l=n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, m_l
+  \\
+  &=
+  \sum_{l=1-j}^{2n-j} \theta^{(k)}_{j+l} \, m_l
+  =
+  \sum_{l=0}^{b} m_l \, \cos\frac{(j+l-\frac12)k\pi}{n},
+\end{aligned}
+$$
+where we have used
+$$
+\theta^{(k)}_i = \theta^{(k)}_{1-i} = \theta^{(k)}_{2n+1-i}.
+$$
+Similarly,
+$$
+\begin{aligned}
+  \sum_{i=1}^n
+  \theta^{(k)}_i
+  \left(
+    m_{j - i} + m_{i+j-1}
+  \right)
+  =
+  \sum_{l=0}^{b} m_l \, \cos\frac{(j-l-\frac12)k\pi}{n}.
+\end{aligned}
+$$
+%
+So
+$$
+\begin{aligned}
+\sum_{i = 1}^n \theta^{(k)}_i \, w_{ij}
+&=
+- \theta^{(k)}_j \, m_0
++\sum_{l=0}^{b} m_l \, 2 \, \cos\frac{lk\pi}{n} \, \cos\frac{(j-\frac12)k\pi}{n}
+\\
+&= \theta^{(k)}_j \, \lambda_k,
+\end{aligned}
+$$
+with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}.
+}
+%
+\begin{equation}
+\lambda_k = m_0 + \sum_{l = 1}^b 2 \, m_l \cos \frac{k l \pi}{n}.
+\label{eq:wband_eigenvalue}
+\end{equation}
+
+\section{Numerical results}
+
+1. Single-bin update.
+Show the error versus $\lambda$.
+
+2. Multiple-bin update scheme.
 

### 2016-01-06 prog/invt.c

@@ -1,22 +1,16 @@
 #include "mtrand.h"
 #include "util.h"
+#include "invtpar.h"
 
 
 
 /* return the root squared error of the inverse time scheme
- *
- * the updating magnitude is `c/t` for t > `nequil`, or `a0` otherwise
- * `n` is the number of bins
- * `tau` is the correlation time
- * `nsteps` is the number of steps
  * */
-static double comperr(double c, double alpha0, double nequil,
-    int n, double tau, long nsteps,
-    double frac,
-    int verbose)
+static double comperr(const invtpar_t *p)
 {
   double *v, a, dv, err, t0;
-  int i, j, t;
+  int i, j, n = p->n;
+  long t;
 
   xnew(v, n);
   for ( i = 0; i < n; i++ ) {
@@ -24,11 +18,11 @@ static double comperr(double c, double alpha0, double nequil,
   }
 
   /* constant */
-  t0 = c / alpha0;
+  t0 = p->c / p->alpha0;
 
-  for ( t = 0; t < nsteps + nequil; t++ ) {
+  for ( t = 0; t < p->nsteps + p->nequil; t++ ) {
     /* MCMC sampling */
-    if ( i < 0 || rand01() * tau < 1 ) {
+    if ( i < 0 || p->tcorr <= 0 || rand01() * p->tcorr < 1 ) {
       j = (int) ( n * rand01() );
       dv = v[j] - v[i];
       if ( dv < 0 || rand01() < exp(-dv) ) {
@@ -39,27 +33,29 @@ static double comperr(double c, double alpha0, double nequil,
     /* compute the updating magnitude
      * the distribution density is p = 1/n
      * this is why we have to multiply by n */
-    if ( t >= nequil ) {
+    if ( t >= p->nequil ) {
       /* the constant t0 makes the transition
        * of alpha at t = t0 smooth */
-      a = c * n / (t - nequil + t0);
+      a = p->c * n / (t - p->nequil + t0);
     } else {
-      a = alpha0 * n;
+      a = p->alpha0 * n;
     }
 
-    if ( frac > 0 ) {
-      double rem = 1;
+    if ( p->nbn > 1 ) {
+      double rem = 1, z;
 
       if ( i > 0 ) {
-        v[i - 1] += a * frac;
-        rem -= frac;
+        z = p->nbs[1];
+        v[i - 1] += a * z;
[…]

### 2016-01-06 prog/invtpar.h

@@ -1,0 +1,269 @@
+#ifndef INVTPAR_H__
+#define INVTPAR_H__
+
+
+
+/* parameters for the invt.c */
+
+
+
+
+#include "util.h"
+
+
+
+/* maximal number of neighbors
+ * in the multiple-bin updating scheme */
+#define NBMAX 127
+
+
+typedef struct {
+  double c; /* constant for the updating magnitude */
+  int n; /* number of bins */
+  double alpha0; /* initial updating magnitude */
+  long nequil; /* equilibration time */
+  long nsteps; /* number of steps */
+  int nbn; /* number of neighboring bins */
+  double nbs[NBMAX + 1]; /* magnitudes of neighbors */
+  int sampmethod; /* sampling method */
+  double tcorr; /* correlation time */
+  int ntrials; /* number of trials */
+  int verbose; /* verbose level */
+  const char *prog; /* name of the program */
+} invtpar_t;
+
+
+
+enum {
+  SAMPMETHOD_PERFECT = 0,
+  SAMPMETHOD_NNMC,
+  SAMPMETHOD_COUNT
+};
+
+const char *sampmethod_names[][8] = {
+  {"perfect", "p"},
+  {"near-neighbor Monte Carlo", "nnmc", "nn"},
+  {""}
+};
+
+
+
+static void invtpar_init(invtpar_t *p)
+{
+  p->c = 1.0;
+  p->n = 10;
+  p->alpha0 = 0.1;
+  p->nequil = p->n * 10000L;
+  p->nsteps = 100000000L;
+  p->tcorr = 0.0; /* perfect sampling */
+  p->ntrials = 100;
+  p->nbn = 1; /* single bin update */
+  p->prog = "invt";
+}
+
+
+
+static int invtpar_load(invtpar_t *m, const char *fn)
+{
+  FILE *fp;
+  char buf[800], *p, *key, *val;
+  int inpar;
+
+  if ( (fp = fopen(fn, "r")) == NULL ) {
+    fprintf(stderr, "cannot load %s\n", fn);
+    return -1;
+  }
+
+  while ( fgets(buf, sizeof buf, fp) ) {
+    strstrip(buf); /* remove trailing spaces */
+    if ( buf[0] == '\0' || buf[0] == '#' ) continue;
+
+    /* parse the line to a key and a value */
+    /* find the end of the key */
+    inpar = 0; /* within (...) */
+    for ( p = buf; *p; p++ ) {
[…]

### 2016-01-07 doc/invt.tex
add discussion on the tridiagonal matrix

@@
 x(t) = x(0) \, e^{-\lambda \, q(t)}
-+ \int_0^t \dot u(t') \, \xi(t') \, dt',
++ \int_0^t \dot u\bigl( q(t') \bigr) \, \xi(t') \, dt',
 \label{eq:xt_solution}
@@
 \begin{align}
-u(q(t'))
+u(z)
 &\equiv
-e^{-\lambda \, q(t) + \lambda \, q(t')}.
+e^{-\lambda \, q(t) + \lambda \, z}.
 \label{eq:ut_definition}
@@
 \int_0^t \int_0^t
-  \dot u(q(t')) \, \dot u(q(t'')) \,
+  \dot u\bigl( q(t') \bigr) \,
+  \dot u\bigl( q(t'') \bigr) \,
   \kappa(t' - t'') \, dt'' \, dt'.
@@
 &=
-\frac{d}{d\tau} \int_0^t \dot u(q(t')) \, \kappa(t' - \tau) \, dt'
-= \int_0^t \ddot u(q(t')) \, \kappa(t' - \tau) \, dt',
+\frac{d}{d\tau} \int_0^t
+  \dot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt'
+= \int_0^t
+  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt',
 \end{aligned}
@@
 \begin{equation}
-\ddot u(q(t')) = 0,
+\ddot u\bigl( q(t') \bigr) = 0,
 \qquad
@@
 \;\;
-\dot u(q(t')) = c,
+\dot u\bigl( q(t') \bigr) = c,
 \label{eq:ddu_eq_0}
@@
   \sum_{i = 2}^n
-  \dot u_i(q(t')) \,
-  \dot u_i(q(t'')) \,
+  \dot u_i\bigl( q(t') \bigr) \,
+  \dot u_i\bigl( q(t'') \bigr) \,
   \kappa_i(t' - t'') \, dt' \, dt'',
-  \notag
-  %\label{eq:y2_int}
+  \label{eq:error_mbin}
 \end{align}
@@
 \begin{align*}
-  u_i(q(t'))
+  u_i(z)
   &\equiv
-  e^{\lambda_i \, [q(t') - \, q(t)]},
+  e^{\lambda_i \, [q(t') - z]},
   \\
@@
   &\equiv
-  \left\langle \eta_i(t) \, \eta_i(t') \right\rangle,
+  \left\langle
+    \eta_i(t) \, \eta_i(t')
+  \right\rangle,
 \end{align*}
@@
 \sum_{i=2}^n
-u_i'(q(\tau))
+\lambda_i \, u_i\bigl( q(\tau) \bigr)
 \int_0^t
-\ddot u_i(q(t')) \, \kappa_i(t' - \tau) \, dt' = 0.
-\label{eq:optimal_mbin1}
+\ddot u_i\bigl( q(t') \bigr) \, \kappa_i(t' - \tau) \, dt' = 0.
+\label{eq:optimal_mbin}
 \end{align}
@@
 %
-The general solution of Eq. \eqref{eq:optimal_mbin1}
+The general solution of Eq. \eqref{eq:optimal_mbin}
 is harder to obtain.
-%
 
@@
 %
-Thus, Eq. \eqref{eq:optimal_mbin1}
-is simplified as
+Then, Eq. \eqref{eq:error_mbin} is simplified as
+%
+\begin{align}
+  \mathcal E
+  &=
+  \int_0^t
+  \sum_{i = 2}^n
+  \dot u_i^2\bigl( q(t') \bigr) \, dt',
+  \label{eq:error_mbin_perfect}
+\end{align}
+%
+and Eq. \eqref{eq:optimal_mbin} becomes
 %
@@
   \sum_{i=1}^n
-  \ddot u_i(q(\tau)) \, u'_i(q(\tau)) = 0.
-  \label{eq:optimal_mbin2}
+  \lambda_i \, u_i\bigl( q(\tau) \bigr) \, \ddot u_i\bigl( q(\tau) \bigr) = 0.
+  \label{eq:optimal_mbin_perfect}
 \end{align}
+%
+\note{If we treat $u_i$ as the position,
+  and $\lambda_i$ as the mass,
+  then $\lambda_i \, \ddot u_i$ is the force.
+  So the left-hand side of Eq. \eqref{eq:optimal_mbin_perfect}
+  is the virial.
+  %
+  Thus this condition states that the total virial is zero.
+}
+
 %
@@
 %
-Using Eqs. \eqref{eq:kappa_perfect}, \eqref{eq:alpha_trial}
-in Eq. \eqref{eq:y2_sum}, we get
+Using Eq. \eqref{eq:alpha_trial}
+in Eq. \eqref{eq:error_mbin_perfect} yield
 %
@@
 \mathcal E
-=
-\int_0^t
-\sum_{i=0}^n \dot u_i^2(q(t')) \, dt'
 \stackrel{t \to \infty}
@@
 and the optimal $\lambda$
-can be found from the smallest root of
+can be found from the smallest real root of
 %
@@
 = 0.
+\label{eq:optimal_lambda_approx}
 \end{equation}
@@
 
-\subsection{Example}
+\subsection{Uniform band matrix}
 
@@
 \theta^{(k)}_i
-= \cos \frac{ (2 \, i - 1) k \pi}{2 \, n}.
+= \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n}.
 \label{eq:wband_eigenvector}
@@
 with eigenvalues
+\begin{equation}
+  \lambda_k = m_0 + \sum_{l = 1}^b 2 \, m_l \cos \frac{(k - 1)  \, l \pi}{n}.
+\label{eq:wband_eigenvalue}
+\end{equation}
+
 \note{To show Eqs.
@@
   =
-  \sum_{l=0}^{b} m_l \, \cos\frac{(j+l-\frac12)k\pi}{n},
+  \sum_{l=0}^{b} m_l \, \cos\frac{(j+l-\frac12)(k-1)\pi}{n},
 \end{aligned}
@@
   =
-  \sum_{l=0}^{b} m_l \, \cos\frac{(j-l-\frac12)k\pi}{n}.
+  \sum_{l=0}^{b} m_l \, \cos\frac{(j-l-\frac12)(k-1)\pi}{n}.
 \end{aligned}
@@
 - \theta^{(k)}_j \, m_0
-+\sum_{l=0}^{b} m_l \, 2 \, \cos\frac{lk\pi}{n} \, \cos\frac{(j-\frac12)k\pi}{n}
++\sum_{l=0}^{b} m_l \, 2 \, \cos\frac{(k-1)l\pi}{n} \, \cos\frac{(j-\frac12)(k-1)\pi}{n}
 \\
@@
 %
-\begin{equation}
-\lambda_k = m_0 + \sum_{l = 1}^b 2 \, m_l \cos \frac{k l \pi}{n}.
-\label{eq:wband_eigenvalue}
+
+\subsubsection{Tridiagonal matrix}
+
+Particularly, if $m_2 = \dots = m_b = 0$,
+the matrix $\mathbf w$ is tridiagonal,
+%
+\begin{equation}
+\arraycolsep=1.4pt\def\arraystretch{1.4}
+\mathbf w
+=
+\left(
+  \begin{array}{cccccccc}
+    1 - m_1   & m_1 & 0 & \dots & 0 \\
+    m_1 & 1 - 2 \, m_1  & m_1 & \dots & 0 \\
+    \vdots & &  & & \vdots \\
+    0 & \dots & m_1 & 1 - 2 \, m_1  & m_1 \\
+    0 & \dots & 0 & m_1 & 1 - m_1
+  \end{array}
+\right).
+\label{eq:wtridiag}
+\end{equation}
+%
+We have
+\begin{equation}
+  \lambda_k = 1 - 4 \, m_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
+\label{eq:wtridiag_eigenvalue}
+\end{equation}
+
+Thus, this updating matrix requires
+$$
+m_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
+$$
+
+For a large $n \gg 1$,
+we can convert Eq.
+\eqref{eq:y2_sum_perfect_invt}
+%and
+%\eqref{eq:optimal_lambda_approx}
+to an integral,
+$$
+\begin{aligned}
+\mathcal E
+&=
+\frac{2 \, n}{\pi \, t}
+\int_0^{\pi/2}
+\frac{ \left(1 - 4 \, m_1 \, \cos^2 p \right)^2 }
+{ \lambda \, \left(2 - 8 \, m_1 \, \cos^2 p - \lambda \right) }
+\, dp
+\notag \\
+&=
+\frac{n}{4 \, t}
+\left(
+  \frac{2 - 4 \, m_1 + \lambda}{ \lambda }
+  +
+  \frac{ \lambda }
+  { \sqrt{ (2 - \lambda) (2 - 8 \, m_1 - \lambda) } }
+\right).
+\end{aligned}
+$$
+\note{
+  The integral is done by contour integration.
+  %
+$$
+\begin{aligned}
+\mathcal E
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\int_0^{2 \, \pi}
+\frac{ \left(1 - 4 \, m_1 \, \cos^2 p \right)^2 }
+{ 2 - 8 \, m_1 \, \cos^2 p - \lambda }
+\, dp
+\\
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\oint
+\frac{ \left[1 - m_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
+{ 2 - 2 \, m_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
+\, \frac{dz}{z},
+\end{aligned}
+$$
+where the contour is around the unit circle in the complex plane.
+%
+This integral has five poles, one at the origin,
+which produces a residue $\frac{2 - 4 \, m_1 - \lambda}{4}$
+(from series expansion of small $z$),
+and the other four at
+$$
+z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, m_1},
+$$
+with two of them inside the unit circle:
+$$
+z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, m_1 - \lambda}}
+{\sqrt{8 m_1}}.
+$$
+Thus,
+$$
+\begin{aligned}
+  \mathcal E
+&=
+\frac{n}{\lambda \, t}
+\left(
+ \frac{2 - 4 \, m_1 - \lambda}{4}
+ +
+ \sum_{z = z_{\pm} }
+ \frac{ \left(z^2 - m_1 (1 + z^2)^2 \right)^2 }
+ { 2 z^4 (2 - 4 m_1 - \lambda - 4 m_1 z^2) }
+\right)
+\\
+&=
+\frac{n}{\lambda \, t}
+\left(
+  \frac{2 - 4 \, m_1 - \lambda}{4}
+ +
+ \frac{ \left(1 - m_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
+ { 2 - 4 m_1 - \lambda - 4 m_1 z_{\pm}^2 }
+\right)
+\\
+&=
+\frac{n}{\lambda \, t}
+\left(
+  \frac{2 - 4 \, m_1 - \lambda}{4}
+ +
+ \frac{ (\lambda/2)^2 }
+ { \sqrt{(2-\lambda) (2 - 8 m_1 -\lambda)} }
+\right).
+\end{aligned}
+$$
+}
+%
+Minimizing the function yields
+%
+\begin{equation}
+\lambda = \frac{1 - 4 \, m_1} { 1 - 2 \, m_1 },
+\label{eq:lambda_tridiag}
+\end{equation}
+%
+and the minimal error is given by
+%
+\begin{equation}
+\mathcal E
+=
+\frac{n}{t}
+\left(
+  1+ \frac{2 \, m_1^2}{1-4 \, m_1}
+\right),
+\label{eq:error_tridiag}
+\end{equation}
+%
+which increases rapidly with $m_1$.
+
+\subsubsection{Gaussian matrix}
+
+Another common updating scheme
+in metadynamics is a Gaussian wave.
+%
+With $b = n - 1 \gg 1$,
+we can take the continuous limit:
+$x = l \, \Delta x$,
+$\Delta x = \pi/n$,
+and
+$m_l = m(x) \, \Delta x$.
+%
+Then
+Eq. \eqref{eq:wband_eigenvalue}
+can be approximated by an integral:
+%
+\begin{equation}
+\lambda_k
+=
+2 \int_0^{\pi}
+m(x) \, \cos\bigl((k-1) \, x\bigr) \, dx,
+\label{eq:lambda_int}
+\end{equation}
+%
+with the normalization
+%
+\begin{equation}
+1 = 2 \int_0^{\pi} m(x) \, dx.
+\label{eq:mx_normalization}
+\end{equation}
+
+Particularly, for the Gaussian wave
+%
+\begin{equation}
+m(x) = \frac{1}{\sqrt{2\pi\sigma^2}}
+\exp\left( -\frac{ x^2 } { 2 \, \sigma^2 } \right),
+\end{equation}
+%
+with $\Delta x \ll \sigma \ll 1$,
+we can extend the upper limit of the integrals
+in Eqs. \eqref{eq:lambda_int}
+and \eqref{eq:mx_normalization} to infinity,
+and
+\begin{equation}
+\lambda_k = \exp\left(
+  -\frac{ (k-1)^2 \, \sigma^2 } { 2 }
+\right).
+\label{eq:lambda_Gaussian}
+\end{equation}
+
+Owing to singularity,
+we can show that the optimal $\lambda$
+in Eq. \eqref{eq:alpha_trial}
+\begin{equation}
+  \lambda \approx 2 \lambda_{n}.
 \end{equation}
@@
 \section{Numerical results}
+
+\subsection{Optimal $\alpha(t)$}
+
+To determine the optimal schedule,
+we shall test a model system
+in the following fashion.
+%
+Initially $\alpha$ is set to $\alpha_0 = 1/n$
+for a sufficiently long time.
+%
+Then we reset the simulation time $t = 0$,
+and switch to
+$$
+\alpha(t) = \min\left\{ \alpha_0, \frac{ 1 } { \lambda \, t } \right\}.
+$$
 
@@
 2. Multiple-bin update scheme.
+Error vs. $\lambda$.
+
+3. Optimality of the Single-bin update.
+Error vs. $m_1$.
 

### 2016-01-07 prog/invt.c

@@ -1,15 +1,15 @@
-#include "mtrand.h"
-#include "util.h"
-#include "invtpar.h"
+
+
+#include "invt.h"
 
 
 
 /* return the root squared error of the inverse time scheme
  * */
-static double comperr(const invtpar_t *p)
+static double comperr(const invtpar_t *m)
 {
-  double *v, a, dv, err, t0;
-  int i, j, n = p->n;
+  double *v, a, err;
+  int i, n = m->n;
   long t;
 
   xnew(v, n);
@@ -17,40 +17,35 @@ static double comperr(const invtpar_t *p)
     v[i] = 0;
   }
 
-  /* constant */
-  t0 = p->c / p->alpha0;
-
-  for ( t = 0; t < p->nsteps + p->nequil; t++ ) {
+  i = 0;
+  for ( t = 0; t < m->nsteps + m->nequil; t++ ) {
     /* MCMC sampling */
-    if ( i < 0 || p->tcorr <= 0 || rand01() * p->tcorr < 1 ) {
-      j = (int) ( n * rand01() );
-      dv = v[j] - v[i];
-      if ( dv < 0 || rand01() < exp(-dv) ) {
-        i = j;
-      }
+    if ( m->tcorr <= 0 || rand01() * m->tcorr < 1 )
+    {
+      i = mc_metro_g(v, n, i);
     }
 
-    /* compute the updating magnitude
-     * the distribution density is p = 1/n
-     * this is why we have to multiply by n */
-    if ( t >= p->nequil ) {
-      /* the constant t0 makes the transition
-       * of alpha at t = t0 smooth */
-      a = p->c * n / (t - p->nequil + t0);
+    /* compute the updating magnitude */
+    if ( t >= m->nequil + m->c / m->alpha0 - m->t0 ) {
+      a = m->c / (t - m->nequil + m->t0);
     } else {
-      a = p->alpha0 * n;
+      a = m->alpha0;
     }
 
-    if ( p->nbn > 1 ) {
+    /* the distribution density is p = 1/n
+     * this is why we have to multiply by n */
+    a *= n;
+
+    if ( m->nbn > 1 ) {
       double rem = 1, z;
 
       if ( i > 0 ) {
-        z = p->nbs[1];
+        z = m->nbs[1];
         v[i - 1] += a * z;
         rem -= z;
       }
       if ( i < n - 1 ) {
[…]

### 2016-01-07 prog/invt.h

@@ -1,0 +1,95 @@
+#ifndef INVT_H__
+#define INVT_H__
+
+
+
+#include "mtrand.h"
+#include "util.h"
+#include "invtpar.h"
+
+
+/* global Metropolis move */
+static int mc_metro_g(double *v, int n, int i)
+{
+  int j, acc;
+  double dv, r;
+
+  j = (int) ( (i + 1 + (n - 1) * rand01()) ) % n;
+  dv = v[j] - v[i];
+  if ( dv <= 0 ) {
+    acc = 1;
+  } else {
+    r = rand01();
+    acc = ( r < exp(-dv) );
+  }
+
+  if ( acc ) {
+    i = j;
+  }
+
+  return i;
+}
+
+
+
+/* local Metropolis move */
+static int mc_metro_l(double *v, int n, int i)
+{
+  int j, acc;
+  double dv, r;
+
+  j = (rand01() > 0.5) ? i + 1 : i - 1;
+  /* periodic boundary condition */
+  j = (j + n) % n;
+  // if ( j < 0 || j >= n ) return i;
+
+  dv = v[j] - v[i];
+  if ( dv <= 0 ) {
+    acc = 1;
+  } else {
+    r = rand01();
+    acc = ( r < exp(-dv) );
+  }
+
+  if ( acc ) {
+    i = j;
+  }
+
+  return i;
+}
+
+
+
+
+/* normalize the potential */
+static void normalize(double *v, int n)
+{
+  double s = 0;
+  int i;
+
+  for ( i = 0; i < n; i++ ) {
+    s += v[i];
+  }
+  s /= n;
+  for ( i = 0; i < n; i++ ) {
+    v[i] -= s;
+  }
+}
+
+
+/* compute the root-mean-squared error */
+static double geterror(double *v, int n)
+{
+  double err = 0;
+  int i;
+
+  for ( i = 0; i < n; i++ ) {
+    err += v[i] * v[i];
+  }
+  return sqrt(err / n);
+}
+
+
+
+#endif /* INVT_H__ */
+

### 2016-01-08 doc/invt.tex
editing invt.tex, add abstract

@@
 \usepackage{amsmath}
+\usepackage{mathtools}
 \usepackage{upgreek}
@@
 
+\begin{abstract}
+  Wang-Landau (WL) and metadynamics algorithms
+  are widely-used techniques in free energy simulations.
+  %
+  These techniques allow one to sample a flat distribution
+  along a quantity of interest
+  by constructing a bias potential on the fly,
+  and to compute the negative potential of mean force (PMF),
+  which is the negative of the bias potential.
+  %
+  However, the asymptotic error of the resulting PMF
+  is determined by the magnitude of updating the bias potential,
+  and in what fashion the magnitude is reduced
+  over the simulation time.
+  %
+  In the Wang-Landau case,
+  an effective prescription of determining the magnitude
+  is to set it to be the inverse of the number of simulation steps.
+  %
+  Here, we shall reexamine the optimality of this prescription,
+  and attempt to extend it to the metadynamics case,
+  in which the bias potential at several neighboring bins
+  are updating according to a prescribed function,
+  upon each visit.
+  %
+  Our analysis also suggests that in terms of the convergence
+  the WL algorithm represents one of the best of the class
+  updating schemes among more sophisticated candidates.
+\end{abstract}
+
 \maketitle
@@
 %
-The discrete nature of $i$ is natural
-in discrete models.
+%The discreteness of $i$ is natural
+%in discrete models.
 %
 For example, $i$ can be the energy $E$
-in lattice models; or the temperature index
+in lattice spin models; or the temperature index
 in a simulated tempering simulation.
 %
-For a continuous quantity, $z$,
+For the quantity of interest, $z$, is continuous,
 which is typical in molecular systems,
-we can also discretize $z$
+we can discretize $z$
 such that the integer $i$ represents
@@
 For a large system,
-the distribution $p^*_i$ is often
-localized around some $i_0$,
+the distribution $p^*_i$ tends to
+be localized around some $i_0$,
 %
@@
 
-Generally, we may wish to target
-a slowly-varying distribution, $p_i$,
+Generally, we may wish to sample
+a nearly flat distribution, $p_i$,
 by adjusting the bias potential $\phi_i$
@@
 Next, we split $h_i(t)$ into a deterministic part
-and a noise
+and a noise part,
 %
@@
 %The noise is not necessarily white.
-Unless specified otherwise, we assume
-a general noise with the correlation
+We consider a general noise with the correlation
 function depending on the time difference:
@@
   \dot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt'
-= \int_0^t
+\\
+&= \int_0^t
   \ddot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt',
@@
 Thus, Eq. \eqref{eq:alpha_invt} is optimal
-up to a constant $t_0$ defining the origin of $t$.
+up to a constant $t_0$. %defining the origin of $t$.
 
@@
 
-We now extend the above analysis to
-the multiple-bin updating scheme.
+We now consider the multiple-bin updating scheme.
 %
@@
 upon a visit to bin $j$,
-we update not only $v_j$,
-but also $v_i$ for a few neighboring bins $i$.
+we update not only the bias potential there,
+but also those for a few neighboring bins $i$.
 %
@@
 whereas the matrix nature brings about
-some difficulties that require
+some technical difficulties that require
 certain restrictions and approximations.
@@
 
-\subsection{Comparison of updating schemes}
+\subsection{Optimal coefficient of proportionality}
 
@@
 is to assume the functional form
+%
 \begin{equation}
@@
 \end{equation}
-and find the optimal value of $\lambda$.
+%
+and to find the optimal value of $\lambda$.
 %
 Using Eq. \eqref{eq:alpha_trial}
-in Eq. \eqref{eq:error_mbin_perfect} yield
-%
-\begin{equation}
+in Eq. \eqref{eq:error_mbin_perfect} yields
+%
+\begin{align}
 \mathcal E
-\stackrel{t \to \infty}
-{=\joinrel=\joinrel=}
-\frac{1}{t}
+&=
+\frac{1}{t+t_0}
+\sum_{i=2}^n \frac{ \lambda_i^2 } { \lambda \, (2 \lambda_i - \lambda) }
+\left[
+  1 - \left(
+    1 + \frac{t}{t_0}
+  \right)^{1 - 2 \, \lambda_i/\lambda}
+\right]
+\notag
+\\
+&\stackrel{      t \gg t_0      }
+          { =\joinrel=\joinrel= }
+%
 \sum_{i = 2}^n
-\frac{ \lambda_i^2 } { \lambda \, (2 \, \lambda_i - \lambda) },
+\begin{dcases}
+  \frac{ 1 }
+       { t }
+  \frac{ \lambda_i^2 }
+       { \lambda \, (2 \, \lambda_i - \lambda) }
+  &
+  \mathrm{if \;} \lambda < 2 \, \lambda_i,
+  \\%[1em]
+  %
+  %
+  \frac{   1    }
+       { 4 \, t }
+  \ln \frac{ t + t_0 }
+           {  t_0    }
+  &
+  \mathrm{if \;} \lambda = 2 \, \lambda_i,
+  \\%[1em]
+  %
+  %
+  \frac{ t_0^{ 2 \, \lambda_i / \lambda  - 1} }
+       { t^{ 2 \, \lambda_i / \lambda } }
+  \frac{ \lambda_i^2 }
+       { \lambda \, (\lambda - 2 \, \lambda_i) }
+  &
+  \mathrm{if \;} \lambda > 2 \, \lambda_i.
+\end{dcases}
 \label{eq:y2_sum_perfect_invt}
-\end{equation}
-%
-and the optimal $\lambda$
-can be found from the smallest real root of
+\end{align}
+%
+\note{In the above expression,
+$t + t_0$ are replaced by $t$.
+%
+Adding back $t_0$, we get
+$$
+\begin{aligned}
+  \mathcal E
+  \stackrel{         t \to \infty         }
+           { =\joinrel=\joinrel=\joinrel= }
+  \sum_{i = 2}^n
+  \begin{dcases}
+    \frac{ 1 }{ t + t_0 }
+    \frac{ \lambda_i^2 } { \lambda \, (2 \, \lambda_i - \lambda) }
+    &
+    \mathrm{if \;} \lambda < 2 \, \lambda_i,
+    \\%[1em]
+    %
+    %
+    \frac{1}{4 \, (t + t_0) } \ln\frac{t + t_0}{t_0}
+    &
+    \mathrm{if \;} \lambda = 2 \, \lambda_i,
+    \\
+    %
+    %
+    \frac{ t_0^{ 2 \, \lambda_i / \lambda  - 1} }
+         { (t + t_0)^{ 2 \, \lambda_i / \lambda } }
+    \frac{ \lambda_i^2 }
+         { \lambda \, (\lambda - 2 \, \lambda_i) }
+    &
+    \mathrm{if \;} \lambda > 2 \, \lambda_i.
+  \end{dcases}
+\end{aligned}
+$$
+}
+%
+The last two cases, $\lambda \ge 2\lambda_i$,
+are asymptotically suboptimal
+as they decay slower than $1/t$.
+%
+Thus, the optimal $\lambda$
+satisfies $\lambda < 2 \min \lambda_i$,
+and in this case,
+we can seek that
+the optimal value from
+the smallest real root of
 %
@@
 
-This matrix is clearly symmetric, and hence
+This matrix, being symmetric,
 satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
@@
 %
+We discuss two special cases below.
 
@@
 \end{equation}
-
-Thus, this updating matrix requires
+%
+Thus, for this updating matrix to be stable,
+we need
+$\min \lambda_k = \lambda_n > 0$,
+and
 $$
@@
 
-For a large $n \gg 1$,
-we can convert Eq.
+We can find an analytical expression
+of the optimal value of $\lambda$
+in the limit of large $n$.
+%
+In this case,
+we convert Eq.
 \eqref{eq:y2_sum_perfect_invt}
@@
 $$
-\note{
-  The integral is done by contour integration.
+\note{The integral is evaluated by contour integration.
   %
@@
 $$
-where the contour is around the unit circle in the complex plane.
+where the contour is the unit circle
+in the complex plane of $z$.
 %
@@
 %
-and the minimal error is given by
+and the error in the optimal is given by
 %
@@
   1+ \frac{2 \, m_1^2}{1-4 \, m_1}
-\right),
+\right).
 \label{eq:error_tridiag}
@@
 %
-which increases rapidly with $m_1$.
+From Eq. \eqref{eq:error_tridiag},
+it is clear that the error
+increases rapidly with magnitude of $m_1$,
+no matter its sign,
+and optimal value is given by $m_1 = 0$,
+which is the single-bin update scheme.
 
@@
 %
-With $b = n - 1 \gg 1$,
-we can take the continuous limit:
+Below we show that the updating scheme
+is stable, although
+it might be less effective
+in reducing local noises.
+
+We start by defining
+a continuous limit
+for $b = n - 1 \gg 1$:
 $x = l \, \Delta x$,
@@
 =
-2 \int_0^{\pi}
-m(x) \, \cos\bigl((k-1) \, x\bigr) \, dx,
+2 \int_0^\pi
+  m(x) \, \cos \bigl( (k-1) \, x \bigr) \, dx,
 \label{eq:lambda_int}
@@
 \begin{equation}
-1 = 2 \int_0^{\pi} m(x) \, dx.
+1 = 2 \int_0^\pi m(x) \, dx.
 \label{eq:mx_normalization}
@@
 
-Particularly, for the Gaussian wave
-%
-\begin{equation}
-m(x) = \frac{1}{\sqrt{2\pi\sigma^2}}
-\exp\left( -\frac{ x^2 } { 2 \, \sigma^2 } \right),
-\end{equation}
-%
-with $\Delta x \ll \sigma \ll 1$,
-we can extend the upper limit of the integrals
+For the Gaussian wave,
+$m(x) \propto \exp(-\frac{1}{2} x^2/\sigma^2)$.
+%
+If $\Delta x \ll \sigma \ll 1$,
+we can extended
+the upper limit of the integrals
 in Eqs. \eqref{eq:lambda_int}
-and \eqref{eq:mx_normalization} to infinity,
-and
-\begin{equation}
-\lambda_k = \exp\left(
-  -\frac{ (k-1)^2 \, \sigma^2 } { 2 }
+and \eqref{eq:mx_normalization}
+to infinity, and
+%
+\begin{equation}
+m(x)
+=
+\frac{            1            }
+     { \sqrt{ 2 \pi \sigma^2 } }
+%
+\exp\left(
+      -\frac{       x^2     }
+            { 2 \, \sigma^2 }
+    \right),
+\end{equation}
+%
+%
+The eigenvalues are given by
+%
+\begin{equation}
+\lambda_k
+=
+\exp\left(
+      -\frac{ (k - 1)^2 \, \sigma^2 }
+            {           2           }
     \right).
@@
 \end{equation}
-
-Owing to singularity,
-we can show that the optimal $\lambda$
-in Eq. \eqref{eq:alpha_trial}
-\begin{equation}
-  \lambda \approx 2 \lambda_{n}.
-\end{equation}
+%
+The eigenvalues are positive,
+which means that the updating scheme
+is stable.
+
+However,
+as $\sigma/\Delta x \gg 1$,
+the smallest eigenvalue
+%
+$$
+\lambda_n
+\approx
+\exp\left(
+      -\frac{ n^2 \, \sigma^2 }
+            {        2        }
+    \right)
+=
+\exp\left[
+      -\frac{ \pi^2 }{ 2 }
+      \left(
+        \frac{  \sigma }
+             { \Delta x }
+      \right)^2
+    \right]
+.
+$$
+%
+is exponentially small.
+%
+This means that it will be difficult
+to use the optimal value of $\lambda$
+given by Eq. \eqref{eq:optimal_lambda_approx},
+which leads to a value
+$\lambda < 2 \, \lambda_n$
+that can be too small in practice.
+%
+In other words,
+with a reasonable $\lambda$,
+the error of the last few fluctuation modes
+will always decay suboptimally,
+accordingly to Eq. \eqref{eq:y2_sum_perfect_invt}.
+%
+However, since these modes
+represent high-frequency fluctuations,
+they may be unimportant.
+%
+Thus,
+we may truncate the sum of the error function
+in Eq. \eqref{eq:y2_sum_perfect_invt}
+at some $k_{\max}$,
+which reflects a minimal length scale
+$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$,
+that is greater than or
+at least comparable to $\sigma$.
+%
+Then, the optimal $\lambda$ is given by
+$$
+\lambda
+\approx
+\lambda_{ k_{\max} }
+\approx
+\exp\left(
+      -\frac{ k_{\max}^2 \, \sigma^2 }
+            {           2            }
+    \right)
+=
+\exp\left(
+      -\frac{  \pi^2 \, \sigma^2    }
+            {    2 \, l_{\min}^2  }
+    \right).
+$$
 

### 2016-01-08 doc/invt.tex
small corrections

@@
   along a quantity of interest
-  by constructing a bias potential on the fly,
-  and to compute the negative potential of mean force (PMF),
-  which is the negative of the bias potential.
+  by adaptively constructing a bias potential,
+  as the negative potential of mean force (PMF).
   %
@@
   In the Wang-Landau case,
-  an effective prescription of determining the magnitude
-  is to set it to be the inverse of the number of simulation steps.
+  a simple and effective formula for the magnitude
+  is given by the inverse of the number of simulation steps.
   %
@@
   in which the bias potential at several neighboring bins
-  are updating according to a prescribed function,
+  are updated according to a predetermined window function,
   upon each visit.
@@
 %
-If we arrange $y_k$ such that
-the mode $y_1 = \sum_{j=1}^n p_j \, x_j$,
+We note that there is one mode, say
+$y_1 = \sum_{j=1}^n p_j \, x_j$
 (corresponding to $\lambda_1 = 1$),
-then $y_1$ vanishes identically
-by Eq. \eqref{eq:px_sum},
-and the sum of $y_k$ in Eq. \eqref{eq:y2_sum}
-can start from $k = 2$.
+vanishes identically by Eq. \eqref{eq:px_sum}.
+%
+Thus, we can start the sum in Eq. \eqref{eq:y2_sum}
+from $k = 2$.
 
@@
 %
-From Eq. \eqref{eq:error_tridiag},
-it is clear that the error
+%From Eq. \eqref{eq:error_tridiag}, it is clear that
+Clearly, the error
 increases rapidly with magnitude of $m_1$,
@@
 %
-The eigenvalues are positive,
-which means that the updating scheme
-is stable.
+The eigenvalues are all positive,
+suggesting stability of the updating scheme.
 
@@
 %
-This means that it will be difficult
-to use the optimal value of $\lambda$
-given by Eq. \eqref{eq:optimal_lambda_approx},
-which leads to a value
-$\lambda < 2 \, \lambda_n$
-that can be too small in practice.
+This means that
+the optimal value of $\lambda$
+given by Eq. \eqref{eq:optimal_lambda_approx}
+would be too small in practice,
+for it requires $\lambda < 2 \, \lambda_n$.
 %
@@
 Thus,
-we may truncate the sum of the error function
+we can truncate the sum of the error function
 in Eq. \eqref{eq:y2_sum_perfect_invt}
 at some $k_{\max}$,
-which reflects a minimal length scale
-$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$,
-that is greater than or
-at least comparable to $\sigma$.
+which corresponds to a minimal length scale
+$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
+comparable to $\sigma$.
 %

### 2016-01-09 doc/invt.tex
invt.c: multiple-bin update

@@
   %
-  Our analysis also suggests that in terms of the convergence
+  Our analysis also suggests that in terms of convergence
   the WL algorithm represents one of the best of the class

### 2016-01-09 prog/invtpar.h

@@ -25,6 +25,7 @@ typedef struct {
   double alpha0; /* initial updating magnitude */
   int nbn; /* width of the updating window function */
   double nbs[NBMAX + 1]; /* shape of the window function */
+  double wgaus; /* width of the Gaussian window */
   int sampmethod; /* sampling method */
   double tcorr; /* correlation time */
   long nequil; /* equilibration time */
@@ -126,6 +127,7 @@ static void invtpar_init(invtpar_t *m)
     m->nbs[i] = 0;
   }
   m->nbs[0] = 1;
+  m->wgaus = 0;
   m->sampmethod = 0;
   m->prog = "invt";
   m->verbose = 0;
@@ -152,10 +154,37 @@ static void invtpar_compute(invtpar_t *m)
   }
 */
 
-  for ( x = 0, i = 1; i < m->nbn; i++ ) {
-    x += m->nbs[i];
+  /* construct the Gaussian window */
+  if ( m->wgaus > 0 ) {
+    m->nbn = NBMAX;
+
+    if ( m->nbn >= m->n / 2 ) {
+      m->nbn = m->n / 2;
+    }
+
+    /* truncate the Gaussian at 5 sigma */
+    if ( m->nbn >= m->wgaus * 5 ) {
+      m->nbn = m->wgaus * 5;
+    }
+
+    x = 1;
+    for ( i = 1; i < m->nbn; i++ ) {
+      m->nbs[i] = exp(-0.5*i*i/m->wgaus/m->wgaus);
+      x += m->nbs[i] * 2;
+    }
+
+    /* normalize */
+    for ( i = 0; i < m->nbn; i++ ) {
+      m->nbs[i] /= x;
+    }
+
+  } else {
+
+    for ( x = 0, i = 1; i < m->nbn; i++ ) {
+      x += m->nbs[i];
+    }
+    m->nbs[0] = 1 - 2 * x;
   }
-  m->nbs[0] = 1 - 2 * x;
 }
 
 
@@ -213,6 +242,11 @@ static int invtpar_load(invtpar_t *m, const char *fn)
              || strstartswith(key, "neighbo")
              || strcmpfuzzy(key, "window") == 0 ) {
       m->nbn = 1 + readarray(m->nbs + 1, NBMAX, val);
+    } else if ( strcmpfuzzy(key, "wgaus") == 0
+             || strcmpfuzzy(key, "width-Gaussian") == 0
+             || strcmpfuzzy(key, "sigma") == 0
[…]

### 2016-01-14 doc/invt.tex
invt.tex, add the proof of optimality of WL scheme

@@
 
-To do so, we need to introduce a bias potential $\phi_i$,
+To do so, we need to introduce a bias potential $V_i$,
 such that the target distribution is changed to
@@
 \begin{equation}
-  \pi_i \propto p^*_i \, e^{-\phi_i}.
+  \pi_i \propto p^*_i \, e^{-V_i}.
   \label{eq:pi_p_phi1}
@@
   \pi_i =
-  \frac{ p^*_i \, e^{-\phi_i} }
-  { \sum_{j = 1}^n p^*_j \, e^{-\phi_j} }.
+  \frac{ p^*_i \, e^{-V_i} }
+  { \sum_{j = 1}^n p^*_j \, e^{-V_j} }.
   \label{eq:pi_p_phi}
@@
 to achieve a flat distribution $\pi_i$,
-the bias potential $\phi_i$
+the bias potential $V_i$
 must coincide with $\log p^*_i$,
@@
 a nearly flat distribution, $p_i$,
-by adjusting the bias potential $\phi_i$
+by adjusting the bias potential $V_i$
 on the fly.
@@
 and then update
-$\phi_i$ according to
-$$
-\phi_i^{\mathrm{new}}
-=
-\phi_i^{\mathrm{old}}
+$V_i$ according to
+$$
+V_i^{\mathrm{new}}
+=
+V_i^{\mathrm{old}}
 -
@@
 %\log p^*_i
-%=\phi_i + \ln p_i + c
+%=V_i + \ln p_i + c
 %\label{eq:bias_pmf}
@@
 %In the end of the simulation,
-%the PMF, $\phi^*_i$,
+%the PMF, $-\log p^*_i$,
 %can be deduced from Eq. \eqref{eq:bias_pmf}.
 %
-The above inversion process can be inconvenient
-in practice,
+The above inversion process,
+adopted by early multicanonical or entropic sampling,
+can be inconvenient in practice,
 for we need to make the period
@@
 The Wang-Landau (WL) algorithm is a convenient technique
-of obtaining the desired bias potential $\phi_i$
+of obtaining the desired bias potential $V_i$
 on the fly.
@@
 \begin{equation}
-  v_i \equiv \phi_i - \ln p^*_i + \ln p_i.
+  v_i \equiv V_i - \ln p^*_i + \ln p_i.
   \label{eq:v_def}
@@
 \begin{equation}
-  \pi_i = \frac{ p_i e^{-v_i} }
-  { \sum_{j = 1}^n p_j e^{-v_j} }
-  \propto p_i \, e^{-v_i}.
+  \pi_i
+  =
+  \frac{                p_i \, e^{-v_i} }
+       { \sum_{j = 1}^n p_j \, e^{-v_j} }
+  \propto
+  p_i \, e^{-v_i}.
   \label{eq:pi_p_v}
@@
 %
-The initial states and random numbers of the copies
-may, however, be different.
+The initial states and the stochastic forces
+during the process may, however, be different.
 
 For sufficiently small $\alpha(t)$,
-the sampling process is approximately
-Markovian of a finite order,
+the $v_i$'s remain roughly the same for a short period.
+%
+Then,
+the sampling process is approximately Markovian, % of a finite order,
 and we may assume Eq. \eqref{eq:pi_p_v}
@@
   \approx
+  \pi_i
+  =
   \frac{ p_i \, e^{-v_i} }
@@
 \end{equation}
-
+%
+%
+%
 For the noise part, we have
-$\langle \zeta_i(t) \rangle = 0$.
+$\langle \zeta_i(t) \rangle = 0$,
 %
 %The noise is not necessarily white.
-We consider a general noise with the correlation
-function depending on the time difference:
+and the correlation functions
+depend only on the time difference
 %
@@
 that vanishes at large $t$.
+%
+More explicitly,
+if the transition matrix of a period $\tau$
+is $T^\tau_{ij}$,
+then
+$
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  =
+  \langle h_i(t) \, h_j(t') \rangle
+  -
+  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
+  =
+  T^{t - t'}_{ij} \pi_j - \pi_i \, \pi_j.
+$
 
@@
 %
-$\sum_{i = 1}^n U_{ki} \, \hat w_{ij} = \lambda_k \, U_{kj}$.
+$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
 %
@@
 \begin{equation}
-  \sum_{i = 1}^n T_{ki} \, w_{ij} = \lambda_k \, T_{kj},
-  \label{eq:T_w}
+  \sum_{i = 1}^n \phi_{ki} \, w_{ij} = \lambda_k \, \phi_{kj},
+  \label{eq:eig_w}
 \end{equation}
 where
-$T_{ij} = \sqrt{p_j} \, U_{ij}$ satisfies
+$\phi_{ij} = \sqrt{p_j} \, \varphi_{ij}$ satisfies
 the orthonormal conditions:
@@
 \begin{align}
-\sum_{k = 1}^n T_{ki} \, T_{kj}
+\sum_{k = 1}^n \phi_{ki} \, \phi_{kj}
 &= \delta_{ij} \, p_i,
-\label{eq:T_orthonormal_cols}
-\\
-\sum_{k = 1}^n \frac{ T_{ik} \, T_{jk} }{ p_k }
+\label{eq:eig_orthonormal_cols}
+\\
+\sum_{k = 1}^n \frac{ \phi_{ik} \, \phi_{jk} }{ p_k }
 &= \delta_{ij}.
-\label{eq:T_orthonormal_rows}
+\label{eq:eig_orthonormal_rows}
 \end{align}
 
-We emphasize that, in a stable updating scheme,
+%It is worth pointing out
+Note that, in a stable updating scheme,
 all eigenvalues $\lambda_k$ must be nonnegative,
@@
 Now by diagonalizing the matrix $\mathbf w$
-using matrix defined in Eqs. \eqref{eq:T_w},
-\eqref{eq:T_orthonormal_cols},
+using matrix defined in Eqs. \eqref{eq:eig_w},
+\eqref{eq:eig_orthonormal_cols},
 and
-\eqref{eq:T_orthonormal_rows},
+\eqref{eq:eig_orthonormal_rows},
 we get a set of decoupled equations
@@
 \begin{align}
-  y_i &= \sum_{j=1}^n T_{ij} \, x_j,
+  y_i &= \sum_{j=1}^n \phi_{ij} \, x_j,
   \label{eq:y_def}
   \\
-  \eta_i &= \sum_{j=1}^n T_{ij} \frac{ \zeta_j}{ p_j}.
+  \eta_i &= \sum_{j=1}^n \phi_{ij} \frac{ \zeta_j}{ p_j}.
   \label{eq:eta_def}
@@
 The error function can be obtained
-from Eq. \eqref{eq:T_orthonormal_cols}
+from Eq. \eqref{eq:eig_orthonormal_cols}
 as
@@
   =
-  \sum_{i, j, k=1}^n T_{ki} \, T_{kj} \, x_i \, x_j
+  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \, x_i \, x_j
   =
@@
     \eta_i(t) \, \eta_i(t')
-  \right\rangle,
+  \right\rangle.
 \end{align*}
 %
-and the extremal condition of $\mathcal E$ is given by
+\note{The extremal condition of $\mathcal E$ is given by
 %
@@
 %
-The general solution of Eq. \eqref{eq:optimal_mbin}
-is harder to obtain.
-
-\subsection{Optimal $\alpha(t)$}
-
-To proceed, we shall further assume perfect sampling,
-%
-\begin{equation}
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  =
-  p_i \, \delta_{ij} \, \delta(t - t').
-  \label{eq:zz_perfect}
-\end{equation}
-%
-Thus, from Eqs. \eqref{eq:T_orthonormal_rows} and \eqref{eq:eta_def},
-we get
-%
-\begin{equation}
-  \kappa_i(t - t')
-  =
-  \sum_{j,k = 1}^n
-  \frac{ T_{ij} } { p_j }
-  \frac{ T_{ik} } { p_k }
-  \langle \zeta_j(t) \, \zeta_k(t') \rangle
-  =
-  \delta(t - t').
-  \label{eq:kappa_perfect}
-\end{equation}
+Unfortunately, the above expression
+appears to be useless in a general setting.
+}
+
+\subsection{White-noise approximation}
+
+In the asymptotically regime,
+where $t \gg 1/\alpha(t)$,
+we can approximate $\kappa_i(t)$ as
+%
+\begin{equation}
+  \kappa_i(t) \approx \Gamma_i \, \delta(t)
+  \label{eq:kappa_delta}
+\end{equation}
+%
+where, $\Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt$.
 %
@@
   \sum_{i = 2}^n
-  \dot u_i^2\bigl( q(t') \bigr) \, dt',
-  \label{eq:error_mbin_perfect}
+  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt',
+  \label{eq:error_mbin1}
 \end{align}
 %
-and Eq. \eqref{eq:optimal_mbin} becomes
+%
+\note{This is equivalent to saying that the Fourier transform
+  $$
+  \tilde \kappa_i(\omega) = \Gamma_i
+  $$
+  is a roughly constant.
+  %
+  For the next order correction we would have
+  $$
+  \tilde \kappa_i(\omega) = \Gamma_i + \Gamma^{(2)}_i \omega^2 + \dots,
+  $$
+  which leads to a correction to $\mathcal E$
+  $$
+  \Gamma^{(2)}_i
+  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'.
+  $$
+  This correction is of order $\alpha^2$
+  compared to the main term in Eq. \eqref{eq:error_mbin1},
+  and hence asymptotically unimportant.
+
+  The extremal property, Eq. \eqref{eq:optimal_mbin},
+  now becomes
   %
@@
     \sum_{i=1}^n
-  \lambda_i \, u_i\bigl( q(\tau) \bigr) \, \ddot u_i\bigl( q(\tau) \bigr) = 0.
-  \label{eq:optimal_mbin_perfect}
+      \Gamma_i \, \lambda_i \,
+      u_i\bigl( q(\tau) \bigr) \,
+      \ddot u_i\bigl( q(\tau) \bigr) = 0.
+    \label{eq:optimal_mbin1}
   \end{align}
   %
-\note{If we treat $u_i$ as the position,
-  and $\lambda_i$ as the mass,
-  then $\lambda_i \, \ddot u_i$ is the force.
-  So the left-hand side of Eq. \eqref{eq:optimal_mbin_perfect}
-  is the virial.
-  %
-  Thus this condition states that the total virial is zero.
+  If we interpret $u_i$ as the position,
+  and $\Gamma_i \, \lambda_i$ as the mass,
+  then the right-hand side give the virial.
+  %
+  Thus, the extremal condition demands
+  the total virial to be zero.
+  $$\,$$
 }
+%
+For example,
+if we assume perfect sampling,
+%
+\begin{equation}
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  =
+  p_i \, (1 - p_i) \, \delta_{ij} \, \delta(t - t').
+  \label{eq:zz_perfect}
+\end{equation}
+%
+Thus, from Eqs. \eqref{eq:eig_orthonormal_rows} and \eqref{eq:eta_def},
+we get
+%
+\begin{equation}
+  \kappa_i(t - t')
+  =
+  \sum_{j,k = 1}^n
+  \frac{ \phi_{ij} } { p_j }
+  \frac{ \phi_{ik} } { p_k }
+  \langle \zeta_j(t) \, \zeta_k(t') \rangle
+  %
+  \approx \delta(t - t').
+  \label{eq:kappa_perfect}
+\end{equation}
+%
+where we have assumed the large $n$ limit
+such that $p_i \ll 1$
+in the last expression.
 
@@
 
+\subsection{\label{sec:optWL}
+Optimality of the single-bin updating scheme}
+
+We can show from Eq. \eqref{eq:error_mbin1}
+that the single-bin updating scheme
+is an optimal updating scheme.
+%
+Using the Cauchy-Schwarz inequality, we have,
+for any set of $\beta_i \ge 0$
+%
+\begin{align}
+&
+\left( \int_0^t
+  dt' \sum_{i = 2}^n \Gamma_i \,
+                     \dot u_i^2\bigl( q(t') \bigr)
+\right)
+\left(
+\int_0^t
+  dt' \sum_{i = 2}^n \Gamma_i \, c_i^2
+\right)
+\notag
+\\
+&
+\ge
+\left(
+\int_0^t
+  dt' \sum_{i = 2}^n \Gamma_i \, c_i \,
+                     \dot u_i\bigl( q(t') \bigr)
+\right)^2
+\notag
+\\
+&
+=
+\left(
+  \sum_{i = 2}^n \Gamma_k \, c_i
+    \left[
+      1 - e^{ -\lambda_i q(t) }
+    \right]
+\right)^2.
+\notag
+\end{align}
+%
+\note{A quick proof of the above version
+  of the Cauchy-Schwarz inequality
+  is the following.
+  %
+  First, we have
+$$
+\int_0^t
+  dt' \sum_{i = 2}^n \Gamma_i \,
+    \left( \dot u_i \, X - c_i \right)^2
+\ge 0,
+$$
+  for any $X$.
+  %
+  Thus, the right-hand side,
+  as a quadratic polynomial of $X$,
+  must have a non-positive discriminant.
+  %
+  This condition gives the desired inequality.
+  \\%
+}
+%
+Note that the last expression depends on
+$q(t')$ only through the end point value,
+which is fixed in the variation process.
+%
+The equality is achieved
+if $\dot u_i\left( q(t') \right) = c_i$
+for all $i = 2, \dots, n$ at any $t'$,
+up to a multiplicative factor.
+%
+Solving this set of equations yields
+$$
+\alpha(t') = \frac{              1              }
+                  { \lambda_i \, (t' + t_{i0} ) },
+\quad
+\mathrm{with\;\;}
+t_{i0} = \frac{             t            }
+              { e^{ \lambda_i q(t) } - 1 }.
+$$
+Such a solution is possible only if
+\begin{equation}
+\lambda_2 = \dots = \lambda_n.
+\end{equation}
+%
+\note{
+  We add some steps for solution.
+  Integrating $\dot u_i = c_i$ yields
+  \begin{equation}
+  u_i(t') = c_i \left(t' + t_{i0} \right).
+  \label{eq:ui_solution}
+  \end{equation}
+  Given that $u_i(t) = 1$ and $u_i(0) = e^{-\lambda_i \, q(t)}$,
+  we have
+  $$
+  c_i = \frac{ 1 }{ t + t_{i0} },
+  \quad
+  \mathrm{and\;}
+  \frac{ t_{i0} } { t + t_{i0} }
+  =
+  e^{ -\lambda_k \, q(t) }.
+  $$
+  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
+  $$
+  -\lambda_i \, q(t) + \lambda_i \, q(t')
+  = \log c_i + \log\left( t' + t_{i0} \right).
+  $$
+  Differentiating this with respect to $t'$ yields
+  $$
+  \lambda_i \, \alpha(t') = \frac{ 1 } { t' + t_{i0} }.
+  $$
+}
+
+Now an updating matrix $\mathbf w$ satisfying this condition
+is essentially a multiple of the identity matrix
+(since $\lambda_1$ corresponds to the mode
+of a uniform shift of all $v_i$,
+we can freely set it to $\lambda_2$
+without changing the nature of the updating scheme).
+%
+Thus, in terms of convergence,
+the single-bin updating scheme
+(as adopted in the WL algorithm)
+is generally the most efficient
+updating scheme.
+
 \subsection{Optimal coefficient of proportionality}
 
-A more direct, and often more accurate, approach
-is to assume the functional form
+If we wish to find,
+for a given updating matrix $\mathbf w$,
+a near-optimal $\alpha(t)$,
+%
+then a good candidate is
 %
@@
 Using Eq. \eqref{eq:alpha_trial}
-in Eq. \eqref{eq:error_mbin_perfect} yields
+in Eq. \eqref{eq:error_mbin1} yields
 %
@@
 that the minimal error is achieved with
-\begin{equation}
-  \lambda_2 = \dots = \lambda_n = \lambda.
-\end{equation}
-%
-and the updating matrix $\mathbf w$ satisfying this condition
-is essentially a multiple of the identity matrix
-(since $\lambda_1$ corresponds to the mode
-of a uniform shift of all $v_i$,
-we can freely set it to $\lambda_2$
-without changing the nature of the updating scheme).
-%
-This shows that in terms of convergence,
-the single-bin updating scheme
-(as adopted in the WL algorithm)
-is the most efficient candidate
-of all multiple-bin schemes.
+$\lambda_2 = \dots = \lambda_n = \lambda$,
+agreeing with the general conclusion
+in Sec. \ref{sec:optWL}.
 
@@
 \begin{equation}
-  w_{ij} = m_{|i-j|} + m_{i+j-1} + m_{2n+1-i-j},
+  w_{ij} = \mu_{|i-j|} + \mu_{i+j-1} + \mu_{2n+1-i-j},
   \label{eq:w_band}
@@
 %
-where $m_0, \dots, m_b \, (b < n)$
+where $\mu_0, \dots, \mu_b \, (b < n)$
 are a set of numbers satisfying
@@
 \begin{equation}
-m_0 + 2 \, m_1 + \cdots + 2 \, m_b = 1,
+\mu_0 + 2 \, \mu_1 + \cdots + 2 \, \mu_b = 1,
 \label{eq:m_normalization}
@@
 %
-and $m_l = 0$ for $l > b$.
+and $\mu_l = 0$ for $l > b$.
 
@@
 %
-It admits $n$ eigenvectors,
-$\pmb\theta^{(1)}, \dots, \pmb\theta^{(n)}$,
+It has $n$ orthonormal eigenvectors,
+$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
 given by
@@
 \begin{equation}
+\varphi^{(k)}_i
+= \sqrt{
+    \frac{ 2 - \delta_{k, 1} }
+         {       n           }
+       }
+  \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n}.
+\label{eq:wband_eigenvector}
+\end{equation}
+%
+with eigenvalues
+\begin{equation}
+  \lambda_k = \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}.
+\label{eq:wband_eigenvalue}
+\end{equation}
+%
+\note{We derive Eqs.
+  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue}
+  below.
+
+  To simplify the calculation,
+  let us first consider the unnormalized eigenvectors,
+  $$
   \theta^{(k)}_i
-= \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n}.
-\label{eq:wband_eigenvector}
-\end{equation}
-%
-with eigenvalues
-\begin{equation}
-  \lambda_k = m_0 + \sum_{l = 1}^b 2 \, m_l \cos \frac{(k - 1)  \, l \pi}{n}.
-\label{eq:wband_eigenvalue}
-\end{equation}
-
-\note{To show Eqs.
-  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue},
-  we first define $m_l = 0$, for $l \le 0$, then
+  =
+  \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n},
+  $$
+  which satisfies the following reflection symmetry,
+  \begin{equation}
+    \theta^{(k)}_i = \theta^{(k)}_{1-i} = \theta^{(k)}_{2n+1-i}.
+    \label{eq:wband_reflection_symmetry}
+  \end{equation}
+  %
+  We shall also define $\mu_l = 0$, for $l \le 0$, then
 $$
@@
 =
-- \theta^{(k)}_j \, m_0
+- \theta^{(k)}_j \, \mu_0
 \\
@@
   \left(
-    m_{i - j} + m_{j - i} + m_{i + j - 1} + m_{2n+1-i-j}
+    \mu_{i - j} + \mu_{j - i} + \mu_{i + j - 1} + \mu_{2n+1-i-j}
   \right)
@@
   \left(
-    m_{i - j} + m_{2n+1-i-j}
+    \mu_{i - j} + \mu_{2n+1-i-j}
   \right)
@@
   &=
-  \sum_{l=1-j}^{n-j} \theta^{(k)}_{j+l} \, m_l
+  \sum_{l=1-j}^{n-j} \theta^{(k)}_{j+l} \, \mu_l
   +
-  \sum_{l=n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, m_l
-  \\
-  &=
-  \sum_{l=1-j}^{2n-j} \theta^{(k)}_{j+l} \, m_l
-  =
-  \sum_{l=0}^{b} m_l \, \cos\frac{(j+l-\frac12)(k-1)\pi}{n},
+  \sum_{l=n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, \mu_l
+  \\
+  &=
+  \sum_{l=1-j}^{2n-j} \theta^{(k)}_{j+l} \, \mu_l
+  =
+  \sum_{l=0}^{b} \mu_l \, \cos\frac{(j+l-\frac12)(k-1)\pi}{n},
 \end{aligned}
 $$
-where we have used
-$$
-\theta^{(k)}_i = \theta^{(k)}_{1-i} = \theta^{(k)}_{2n+1-i}.
-$$
+where we have used Eq. \eqref{eq:wband_reflection_symmetry}.
+%
 Similarly,
@@
   \left(
-    m_{j - i} + m_{i+j-1}
+    \mu_{j - i} + \mu_{i+j-1}
   \right)
   =
-  \sum_{l=0}^{b} m_l \, \cos\frac{(j-l-\frac12)(k-1)\pi}{n}.
+  \sum_{l=0}^{b} \mu_l \, \cos\frac{(j-l-\frac12)(k-1)\pi}{n}.
 \end{aligned}
@@
 &=
-- \theta^{(k)}_j \, m_0
-+\sum_{l=0}^{b} m_l \, 2 \, \cos\frac{(k-1)l\pi}{n} \, \cos\frac{(j-\frac12)(k-1)\pi}{n}
+- \theta^{(k)}_j \, \mu_0
++\sum_{l=0}^{b} \mu_l \, 2 \, \cos\frac{(k-1)l\pi}{n} \, \cos\frac{(j-\frac12)(k-1)\pi}{n}
 \\
@@
 with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}.
+
+Next, let us compute the normalization factor.
+%
+If $k = 1$, $\theta^{(1)}_i = 1$, and
+$
+\sum_{i = 1}^n \left( \theta^{(1)}_i \right)^2 = n.
+$
+For $k \ge 2$,
+$$
+\begin{aligned}
+  \sum_{i = 1}^n \cos^2 \left[\left(i - \frac1 2 \right) a\right]
+  &=
+  \frac n 2
+  +
+  \frac 1 2
+  \sum_{i = 1}^n \cos\left[(2 i - 1)\, a \right]
+  =
+  \frac n 2,
+\end{aligned}
+$$
+where $a = \frac{k-1}{n} \pi$.
+Thus, the normalization factor $\sqrt{(2 - \delta_{k1})/n}$
+encompasses both cases.
+
+We can also show that column-wise orthogonality.
+%
+For this we need a lemma, for $a = q \, \pi/n$,
+with $q$ a positive integer.
+$$
+\begin{aligned}
+\sum_{k = 1}^n \cos[(k - 1) \, a]
+&=
+1 + \cos a + \dots + \cos[(n - 1) \, a]
+\\
+&=
+\frac{
+      \sin\frac a 2
+    + \sin \left[ \left( n - \frac 1 2 \right) a \right]
     }
+    {
+      2 \, \sin \frac a 2
+    }
+\\
+&=
+\frac{ 1 - (-1)^q } { 2 }
+= \operatorname{odd}(q),
+\end{aligned}
+$$
+%
+where, we have used
+$\sin \left[ \left( n - \frac 1 2 \right) a \right]
+= \sin \left( q \, \pi - \frac a 2 \right)
+= -(-)^q\sin\frac a 2.$
+%
+and $\operatorname{odd}(q)$
+yields $1$ for odd $q$ or $0$ otherwise.
+%
+Adding the case of $q = 0$, where the sum yields $n$,
+we get
+$$
+\sum_{k = 1}^n \cos\left[(k - 1) \, \frac { q \, \pi } { n }  \right]
+= n \delta_{q, 0} + \operatorname{odd}(q).
+$$
+
+Now, for the orthogonality,
+$$
+\begin{aligned}
+  \sum_{k = 1}^n
+  \theta^{(k)}_i \, \theta^{(k)}_j
+  &=
+  \frac 1 2
+  \sum_{k = 1}^n
+  \left[
+    \cos \tfrac{ (i - j) (k - 1) \, \pi }
+               {         n              }
+    +
+    \cos \tfrac{ (i + j - 1) (k - 1) \, \pi }
+               {             n              }
+  \right]
+  \\
+  &=
+  \frac 1 2
+  \left[
+    n \, \delta_{i, j}
+    +
+    \operatorname{odd}(i - j)
+    +
+    \operatorname{odd}(i + j - 1)
+  \right]
+  \\
+  &=
+  \frac n 2 \, \delta_{i, j}
+  + \frac 1 2.
+\end{aligned}
+$$
+where we have used the fact
+that $i - j$ shares the parity with $i + j$
+in the last step.
+%
+So
+$$
+  \sum_{k = 1}^n
+  \varphi^{(k)}_i \, \varphi^{(k)}_j
+  =
+  \frac 2 n
+  \sum_{k = 1}^n
+  \theta^{(k)}_i \, \theta^{(k)}_j
+  -
+  \frac 1 n
+  =
+  \delta_{i, j}.
+$$
+}
 %
@@
 
-Particularly, if $m_2 = \dots = m_b = 0$,
+Particularly, if $\mu_2 = \dots = \mu_b = 0$,
 the matrix $\mathbf w$ is tridiagonal,
@@
 \begin{equation}
-\arraycolsep=1.4pt\def\arraystretch{1.4}
+\arraycolsep=3.6pt\def\arraystretch{1.4}
 \mathbf w
@@
   \begin{array}{cccccccc}
-    1 - m_1   & m_1 & 0 & \dots & 0 \\
-    m_1 & 1 - 2 \, m_1  & m_1 & \dots & 0 \\
+    1 - \mu_1   & \mu_1 & 0 & \dots & 0 \\
+    \mu_1 & 1 - 2 \, \mu_1  & \mu_1 & \dots & 0 \\
     \vdots & &  & & \vdots \\
-    0 & \dots & m_1 & 1 - 2 \, m_1  & m_1 \\
-    0 & \dots & 0 & m_1 & 1 - m_1
+    0 & \dots & \mu_1 & 1 - 2 \, \mu_1  & \mu_1 \\
+    0 & \dots & 0 & \mu_1 & 1 - \mu_1
   \end{array}
@@
 \begin{equation}
-  \lambda_k = 1 - 4 \, m_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
+  \lambda_k = 1 - 4 \, \mu_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
 \label{eq:wtridiag_eigenvalue}
@@
 $$
-m_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
+\mu_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
 $$
@@
 \int_0^{\pi/2}
-\frac{ \left(1 - 4 \, m_1 \, \cos^2 p \right)^2 }
-{ \lambda \, \left(2 - 8 \, m_1 \, \cos^2 p - \lambda \right) }
+\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
+{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
 \, dp
@@
 \left(
-  \frac{2 - 4 \, m_1 + \lambda}{ \lambda }
+  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
   +
   \frac{ \lambda }
-  { \sqrt{ (2 - \lambda) (2 - 8 \, m_1 - \lambda) } }
+  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
 \right).
@@
 \int_0^{2 \, \pi}
-\frac{ \left(1 - 4 \, m_1 \, \cos^2 p \right)^2 }
-{ 2 - 8 \, m_1 \, \cos^2 p - \lambda }
+\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
+{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
 \, dp
@@
 \oint
-\frac{ \left[1 - m_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
-{ 2 - 2 \, m_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
+\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
+{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
 \, \frac{dz}{z},
@@
 This integral has five poles, one at the origin,
-which produces a residue $\frac{2 - 4 \, m_1 - \lambda}{4}$
+which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
 (from series expansion of small $z$),
@@
 $$
-z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, m_1},
+z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
 $$
@@
 $$
-z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, m_1 - \lambda}}
-{\sqrt{8 m_1}}.
+z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
+{\sqrt{8 \mu_1}}.
 $$
@@
 \left(
- \frac{2 - 4 \, m_1 - \lambda}{4}
+ \frac{2 - 4 \, \mu_1 - \lambda}{4}
  +
  \sum_{z = z_{\pm} }
- \frac{ \left(z^2 - m_1 (1 + z^2)^2 \right)^2 }
- { 2 z^4 (2 - 4 m_1 - \lambda - 4 m_1 z^2) }
+ \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
+ { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
 \right)
@@
 \left(
-  \frac{2 - 4 \, m_1 - \lambda}{4}
+  \frac{2 - 4 \, \mu_1 - \lambda}{4}
  +
- \frac{ \left(1 - m_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
- { 2 - 4 m_1 - \lambda - 4 m_1 z_{\pm}^2 }
+ \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
+ { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2 }
 \right)
@@
 \left(
-  \frac{2 - 4 \, m_1 - \lambda}{4}
+  \frac{2 - 4 \, \mu_1 - \lambda}{4}
  +
  \frac{ (\lambda/2)^2 }
- { \sqrt{(2-\lambda) (2 - 8 m_1 -\lambda)} }
+ { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
 \right).
@@
 \begin{equation}
-\lambda = \frac{1 - 4 \, m_1} { 1 - 2 \, m_1 },
+\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
 \label{eq:lambda_tridiag}
@@
 \left(
-  1+ \frac{2 \, m_1^2}{1-4 \, m_1}
+  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
 \right).
@@
 Clearly, the error
-increases rapidly with magnitude of $m_1$,
+increases rapidly with magnitude of $\mu_1$,
 no matter its sign,
-and optimal value is given by $m_1 = 0$,
+and optimal value is given by $\mu_1 = 0$,
 which is the single-bin update scheme.
@@
 and
-$m_l = m(x) \, \Delta x$.
+$\mu_l = \mu(x) \, \Delta x$.
 %
@@
 2 \int_0^\pi
-  m(x) \, \cos \bigl( (k-1) \, x \bigr) \, dx,
+  \mu(x) \, \cos \bigl( (k-1) \, x \bigr) \, dx,
 \label{eq:lambda_int}
@@
 \begin{equation}
-1 = 2 \int_0^\pi m(x) \, dx.
+1 = 2 \int_0^\pi \mu(x) \, dx.
 \label{eq:mx_normalization}
@@
 For the Gaussian wave,
-$m(x) \propto \exp(-\frac{1}{2} x^2/\sigma^2)$.
+$\mu(x) \propto \exp(-\frac{1}{2} x^2/\sigma^2)$.
 %
@@
 \begin{equation}
-m(x)
+\mu(x)
 =
@@
 However, since these modes
-represent high-frequency fluctuations,
-they may be unimportant.
+represent short wavelength fluctuations,
+and these modes may be in
+the system under consideration.
 %
@@
 $l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
-comparable to $\sigma$.
-%
-Then, the optimal $\lambda$ is given by
-$$
-\lambda
-\approx
+greater than $\sigma$.
+%
+Then,
+$$
 \lambda_{ k_{\max} }
@@
 3. Optimality of the Single-bin update.
-Error vs. $m_1$.
+Error vs. $\mu_1$.
 

### 2016-01-14 prog/invt.c
add modtest.c

@@ -4,8 +4,7 @@
 
 
 
-/* return the root squared error of the inverse time scheme
- * */
+/* return the root-mean-squared error of the inverse time scheme */
 static double comperr(const invtpar_t *m, double *err0)
 {
   double *v, *vac, a, err;
@@ -37,7 +36,7 @@ static double comperr(const invtpar_t *m, double *err0)
     }
 
     /* try to start production */
-    if ( !prod && t >= m->nequil + m->c / m->alpha0 - m->t0 ) {
+    if ( !prod && t >= m->nequil ) {
       prod = 1;
       *err0 = geterror(v, n);
       if ( m->verbose >= 1 ) {
@@ -53,10 +52,7 @@ static double comperr(const invtpar_t *m, double *err0)
       a = m->alpha0;
     }
 
-
-    /* the distribution density is p = 1/n
-     * this is why we have to multiply by n */
-    a *= n;
+    a /= m->p[i];
 
     if ( m->nbn > 1 ) {
       mbin_update(v, n, i, a, m->nbs, m->nbn);
@@ -84,7 +80,7 @@ static double comperr(const invtpar_t *m, double *err0)
 
 
 
-static double invt_test(invtpar_t *m)
+static double invt_run(invtpar_t *m)
 {
   double err, see = 0;
   double err0, see0 = 0;
@@ -119,7 +115,8 @@ int main(int argc, char **argv)
   invtpar_init(m);
   invtpar_doargs(m, argc, argv);
   invtpar_dump(m);
-  invt_test(m);
+  invt_run(m);
+  invtpar_finish(m);
 
   return 0;
 }

### 2016-01-14 prog/invtpar.h

@@ -22,6 +22,7 @@ typedef struct {
   double c; /* constant for the updating magnitude */
   double t0; /* offset of equation c/(t + t0) */
   int n; /* number of bins */
+  double *p; /* target distribution */
   double alpha0; /* initial updating magnitude */
   int nbn; /* width of the updating window function */
   double nbs[NBMAX + 1]; /* shape of the window function */
@@ -118,6 +119,7 @@ static void invtpar_init(invtpar_t *m)
   m->c = 1.0;
   m->t0 = 0;
   m->n = 10;
+  m->p = NULL;
   m->alpha0 = 0.0;
   m->nequil = m->n * 10000L;
   m->nsteps = 100000000L;
@@ -150,11 +152,12 @@ static void invtpar_compute(invtpar_t *m)
     m->alpha0 = 0.01 / m->n;
   }
 
-/*
   if ( m->t0 <= 0 ) {
+    /* set the default t0 such that
+     * alpha is continuous at the beginning
+     * of the production run */
     m->t0 = m->c / m->alpha0;
   }
-*/
 
   /* construct the Gaussian window */
   if ( m->wgaus > 0 ) {
@@ -166,7 +169,7 @@ static void invtpar_compute(invtpar_t *m)
 
     /* truncate the Gaussian at 5 sigma */
     if ( m->nbn >= m->wgaus * 5 ) {
-      m->nbn = m->wgaus * 5;
+      m->nbn = (int) (m->wgaus * 5 + 0.5);
     }
 
     x = 1;
@@ -187,6 +190,25 @@ static void invtpar_compute(invtpar_t *m)
     }
     m->nbs[0] = 1 - 2 * x;
   }
+
+  /* initialize the target distribution */
+  if ( m->p ) {
+    free(m->p);
+  }
+  xnew(m->p, m->n);
+
+  for ( i = 0; i < m->n; i++ ) {
+    m->p[i] = 1.0 / m->n;
+  }
+}
+
+
+
+/* clean up */
+static void invtpar_finish(invtpar_t *m)
+{
+  free(m->p);
+  m->p = NULL;
 }
 
 

### 2016-01-14 prog/modetest.c

@@ -1,0 +1,86 @@
+/* test the eigenmode decomposition code */
+#include "invt.h"
+
+
+static void printarr(const double *a, int n, const char *delim)
+{
+  int i;
+
+  for ( i = 0; i < n; i++ ) {
+    printf("%10.4f", a[i]);
+    if ( i < n - 1 ) printf(delim);
+  }
+  printf("\n");
+}
+
+
+static int output(const double *v1, const double *v2, int n,
+    const char *fn)
+{
+  int i;
+  FILE *fp;
+
+
+  if ( (fp = fopen(fn, "w")) == NULL ) {
+    fprintf(stderr, "cannot write %s\n", fn);
+    return -1;
+  }
+
+  for ( i = 0; i < n; i++ ) {
+    fprintf(fp, "%d\t%g\t%g\n", i, v1[i], v2[i]);
+  }
+
+  fclose(fp);
+
+  return 0;
+}
+
+
+int main(void)
+{
+  int n = 100, i, j;
+  double *u, *u1, *v, *v1, *v2, *costab;
+
+  xnew(u, n);
+  xnew(u1, n);
+  xnew(v, n);
+  xnew(v1, n);
+  xnew(v2, n);
+  costab = makecostab(n);
+
+  /* random noise */
+  for ( i = 0; i < n; i++ ) {
+    v[i] = randgaus();
+  }
+
+  /* convert them to modes */
+  getcosmodes(v, n, u, costab);
+
+  /* convert back */
+  fromcosmodes(v1, n, u, costab);
+
+  /* truncate high frequency modes */
+  for ( i = 0; i < n; i++ ) {
+    u1[i] = ( i <= 10 ) ? u[i] : 0;
+  }
+
+  /* convert back */
+  fromcosmodes(v2, n, u1, costab);
+
+  /* print out u and v */
+  if ( n <= 12 ) {
+    printarr(v, n, "  ");
+    printarr(u, n, "  ");
+    printarr(v1, n, "  ");
+    printarr(v2, n, "  ");
+  }
+
+  output(v1, v2, n, "out.dat");
+
+  free(u);
+  free(u1);
+  free(v);
+  free(v1);
+  free(v2);
+  free(costab);
+}

### 2016-01-15 doc/invt.tex
add modtest.c
add corr.h to compute correlation functions

@@
 In the asymptotically regime,
-where $t \gg 1/\alpha(t)$,
+where $t \gg 1/\max \alpha(t)$,
 we can approximate $\kappa_i(t)$ as
@@
   \Gamma^{(2)}_i
-  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'.
+  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'
+  \propto
+  \alpha^2 \, \mathcal E.
   $$
   This correction is of order $\alpha^2$
-  compared to the main term in Eq. \eqref{eq:error_mbin1},
-  and hence asymptotically unimportant.
-
+  compared to the main contribution in Eq. \eqref{eq:error_mbin1}.
+  %
+  Thus, we can assume Eq. \eqref{eq:kappa_delta},
+  once $\alpha(t)$ has dropped below a certain level.
+  \\
+}
+\note{
   The extremal property, Eq. \eqref{eq:optimal_mbin},
@@
 %
-\note{A quick proof of the above version
+\note{A short proof of the above version
   of the Cauchy-Schwarz inequality
@@
 
-We can also show that column-wise orthogonality.
+We can also show that columnwise orthogonality.
 %

### 2016-01-15 prog/corr.h

@@ -1,0 +1,194 @@
+#ifndef CORR_H__
+#define CORR_H__
+
+
+
+/* module to compute autocorrelation function */
+
+
+
+#include "util.h"
+
+
+
+#define CORR_BLKSZ  1024
+
+
+
+typedef struct {
+  int n;
+  int cnt; /* number of frames */
+  int cap; /* maximal number of frames */
+  double *arr; /* data array */
+} corr_t;
+
+
+
+/* open an correction function object */
+static corr_t *corr_open(int n, int cap)
+{
+  corr_t *c;
+
+  xnew(c, 1);
+  c->n = n;
+  c->cnt = 0;
+  if ( cap <= 0 ) {
+    cap = CORR_BLKSZ;
+  }
+  c->cap = cap;
+  xnew(c->arr, c->cap * c->n);
+
+  return c;
+}
+
+
+
+/* close the correction function object */
+static void corr_close(corr_t *c)
+{
+  free(c->arr);
+  free(c);
+}
+
+
+
+/* add a frame */
+static void corr_add(corr_t *c, double *u)
+{
+  int i, n = c->n;
+
+  if ( c->cnt >= c->cap ) {
+    c->cap += CORR_BLKSZ;
+    /* make the array larger */
+    xrenew(c->arr, c->cap * c->n);
+  }
+
+  /* add the new entry */
+  for ( i = 0; i < n; i++ ) {
+    c->arr[c->cnt * n + i] = u[i];
+  }
+  c->cnt += 1;
+}
+
+
+
+static void corr_getave(corr_t *c, double *uave)
+{
+  int i, k, n = c->n;
+
+  for ( i = 0; i < n; i++ ) {
+    uave[i] = 0;
+  }
+
+  for ( k = 0; k < c->cnt; k++ ) {
+    for ( i = 0; i < n; i++ ) {
+      uave[i] += c->arr[k * n + i];
+    }
+  }
+
+  for ( i = 0; i < n; i++ ) {
+    uave[i] /= c->cnt;
+  }
+}
+
+
+
+/* compute autocorrelation functions
+ * for a separation of j frames */
+static int corr_compute(corr_t *c, double *uu, int j,
+    const double *uave)
+{
+  int i, k, k2, n = c->n, tmax = c->cnt;
+  double u1, u2;
+
+  /* reset the accumulators */
+  for ( i = 0; i < n; i++ ) {
+    uu[i] = 0;
+  }
+
+  /* loop over the starting point of the window
[…]

### 2016-01-16 doc/invt.tex
invt.tex: introduction

@@
 \begin{abstract}
-  Wang-Landau (WL) and metadynamics algorithms
-  are widely-used techniques in free energy simulations.
-  %
-  These techniques allow one to sample a flat distribution
-  along a quantity of interest
+  The Wang-Landau (WL) algorithm and metadynamics
+  are two closely-related techniques in free energy simulations.
+  %
+  They allow one to sample a flat distribution
+  along a quantity of interest, $z$,
   by adaptively constructing a bias potential,
-  as the negative potential of mean force (PMF).
-  %
-  However, the asymptotic error of the resulting PMF
-  is determined by the magnitude of updating the bias potential,
-  and in what fashion the magnitude is reduced
+  which is the negative potential of mean force (PMF).
+  %
+  The two algorithms differ by
+  the manner of updating the bias potential:
+  %
+  in the WL case, each simulation step triggers
+  an update of the bias potential to only
+  the bin containing the current $z$,
+  while in the metadynamics case,
+  the bias potential of several neighboring bins
+  are affected as well.
+  %
+  The asymptotic error of the resulting PMF
+  is determined by the updating magnitude,
+  and the rate of the reduction of the magnitude
   over the simulation time.
   %
-  In the Wang-Landau case,
+  In the WL case,
   a simple and effective formula for the magnitude
@@
   %
-  Here, we shall reexamine the optimality of this prescription,
-  and attempt to extend it to the metadynamics case,
-  in which the bias potential at several neighboring bins
-  are updated according to a predetermined window function,
-  upon each visit.
-  %
-  Our analysis also suggests that in terms of convergence
-  the WL algorithm represents one of the best of the class
-  updating schemes among more sophisticated candidates.
+  In this study, we shall reexamine the optimality of this prescription,
+  and attempt to extend it to the metadynamics case.
+  %
+  Further, we compared the two algorithms, and found that
+  although metadynamics produces a smoother PMF profile,
+  it is less effective in reducing errors with the shorter wavelengths.
+  %
+  Thus, it relies more heavily on the assumption of
+  the smoothness of the PMF profile,
+  especially with a wide updating window.
 \end{abstract}
@@
 
-\section{Background}
-
-\subsection{Flat-distribution sampling}
+\section{Introduction}
+
+Free energy calculation\cite{frenkel} is a central theme
+in computational physics and chemistry.
+%
+Given a system,
+the problem is to compute a distribution, $p^*(z)$,
+along a quantity of interest $z$, and to see how
+the distribution changes with the external conditions,
+such as temperatures and pressures.
+%
+The negative logarithm $-\log p^*(z)$,
+or the potential of mean force (PMF),
+defines a free energy.
+%
+A straightforward approach of computing the PMF
+is to run multiple independent simulations
+under different conditions,
+and then piece together the information
+gathered there.
+%
+This solution is often unsatisfactory
+for a complex and glassy system,
+because the system can be trapped
+in a local free energy minimum
+for a long time.
+
+A better solution is to artificially alter
+the target distribution,
+such that we can, in a single simulation, sample
+a flat distribution\cite{mezei1987, berg1992, lee1993,
+wang2001, wang2001pre, laio2002}
+along $z$ over a wide range.
+%
+This alleviates the above problem of local trapping
+as the system is now forced to travel along the $z$ direction,
+which often represents a slow reaction coordinate.
+%
+It, however, requires one to efficiently construct a
+bias potential for an often unknown system.
+%
+Note that the bias potential is directly related to the PMF
+as the former must precisely offset the latter
+to achieve a flat distribution.
+
+The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
+and metadynamics\cite{laio2002}
+are two widely-used techniques for this purpose,
+in which a bias potential is actively build-up
+based on the simulation history.
+%
+In essence, these techniques
+encourage the system to leave a region that has been visited
+by elevating the bias potential there.
+%
+To be effective, the updates to bias potential
+are made at every few steps,
+while a side effect is that the underlying dynamics
+is no longer a equilibrium sampling
+required by statistical mechanics.
+%
+Thus, one has to reduce the magnitude of updating
+over the course of simulation.
+
+Naturally, the manner of reducing
+the updating magnitude\cite{
+belardinelli2007, belardinelli2007jcp, belardinelli2008,
+morozov2007, zhou2008, morozov2009,
+komura2012, caparica2012, caparica2014}
+determines the precision of the final bias potential,
+hence that of PMF.
+%
+For the WL algorithm, it is found
+that asymptotically the optimal magnitude of updating $\ln f$
+should be given by the
+inverse time\cite{
+belardinelli2007, belardinelli2007jcp, belardinelli2008,
+morozov2007, zhou2008},
+where ``time'' is defined as
+the number of simulation steps
+divided by the number of bins.
+
+In this study,
+we shall give a proof of the inverse-time formula,
+and try to generalize it to
+updating schemes involving neighboring bins,
+such as those used in metadynamics.
+%
+As a byproduct, we shall show that
+the single-bin scheme used by the WL algorithm
+is generally one of the best
+updating schemes in terms of convergence.
+%
+The article is organized as follows.
+%
+We present the analytical results in Sec. \ref{sec:theory},
+numerically verify some key aspects
+in Sec. \ref{sec:results},
+and conclude the article
+in Sec. \ref{sec:conclusion}.
+
+\section{\label{sec:theory}
+Theory}
+
+In this section,
+we first review the basics and fix notations
+in Sec. \ref{sec:background}.
+%
+We then prove the optimality
+of the inverse-time formula
+for the single-bin WL algorithm
+in Sec. \ref{sec:single-bin}.
+%
+Next, we discuss the generalized
+multiple-bin schemes
+in Secs. \ref{sec:multiple-bin}
+and \ref{sec:band-matrix}.
+%
+
+\subsection{\label{sec:background}
+Background}
+
+\subsubsection{Flat-distribution sampling}
 
@@
 In both cases,
-the distribution can be normalized as
+the distribution is normalized as
 $\sum_{i = 1}^n p^*_i = 1$.
@@
 
-To do so, we need to introduce a bias potential $V_i$,
-such that the target distribution is changed to
+To do so, we need to introduce a bias potential $V_i$
+to modify the target distribution to
 %
@@
 %
-Upon normalization $\sum_{i = 1}^n \pi_i = 1$,
+Upon normalization, $\sum_{i = 1}^n \pi_i = 1$,
 we get
@@
 Generally, we may wish to sample
-a nearly flat distribution, $p_i$,
+a nearly flat distribution\cite{dayal2004, trebst2004}, $p_i$,
 by adjusting the bias potential $V_i$
@@
 The above inversion process,
-adopted by early multicanonical or entropic sampling,
+adopted by early multicanonical or
+entropic sampling\cite{berg1992, lee1993},
 can be inconvenient in practice,
@@
 
-\subsection{Wang-Landau algorithm}
-
-The Wang-Landau (WL) algorithm is a convenient technique
-of obtaining the desired bias potential $V_i$
-on the fly.
-%
-To simplify the notation, we introduce
-a shifted bias potential
+For later convenience, we introduce
 %
@@
 %
-%
-and in $v_i$, Eq. \eqref{eq:pi_p_phi}
+Since the last two terms of Eq. \eqref{eq:v_def}
+are constant during the course of simulation,
+updates to $V_i$ are equivalent to those to $v_i$.
+%
+In terms of $v_i$, Eq. \eqref{eq:pi_p_phi}
 becomes
@@
 
-In the WL algorithm, $v_i$ are updated
+\subsubsection{Updating schemes}
+
+In the WL algorithm\cite{wang2001, wang2001pre},
+$v_i$ are updated
 in each MC step $t$,
@@
 %
-Note that the updating scheme only applies
-to the $v_i$ at the current bin.
+We refer to the above scheme as the single-bin scheme,
+for it applies only to the bias potential
+at the current bin, $i(t)$,
+%
+By contrast, a multiple-bin scheme
+means the general case in which
+several neighboring bins are updated.
+
 %The values of $v_j(t)$ at $j \ne i(t)$
 %are kept unchanged.
-%
+
 One can show that with a constant $\alpha(t) = \alpha > 0$,
@@
 %
-However, the updating scheme
+However, the continuous update
 makes $v_i(t)$ a fluctuating quantity,
@@
 
-\subsection{$1/t$ formula}
+\subsubsection{$1/t$ formula}
 
@@
 of updating the $\alpha(t)$
-is to discard the stage-wise setup,
-and to follow the formula
+is to follow the formula
 %
@@
 
-Here we shall derive the optimal $\alpha(t)$
-for the WL algorithm, Eq. \eqref{eq:wl_update}.
-%
-Further, we shall generalize the result
-to a class of multiple-bin updating schemes,
-in which an update affects
-not only the current bin $i(t)$,
-but also a few neighboring ones.
-%(such schemes are designed
-%to maintain a smooth bias potential).
-%
-We shall show that
-the constant of proportionality, $1$,
-in Eq. \eqref{eq:alpha_invt}
-is optimal for the single-bin updating scheme,
-Eq. \eqref{eq:wl_update}
-(as in the WL algorithm),
-but not necessarily so
-for a multiple-bin updating scheme.
-%
-Below, we discuss the single-bin updating scheme
-in Section \ref{sec:single-bin},
-and the multiple-bin case
-in Section \ref{sec:multiple-bin}.
-
-\section{\label{sec:single-bin}
-Single-bin updating scheme}
+%Here we shall derive the optimal $\alpha(t)$
+%For the WL algorithm, Eq. \eqref{eq:wl_update}.
+%%
+%Further, we shall generalize the result
+%To a class of multiple-bin schemes,
+%In which an update affects
+%Not only the current bin $i(t)$,
+%But also a few neighboring ones.
+%%(such schemes are designed
+%%to maintain a smooth bias potential).
+%%
+%We shall show that
+%The constant of proportionality, $1$,
+%In Eq. \eqref{eq:alpha_invt}
+%Is optimal for the single-bin scheme,
+%Eq. \eqref{eq:wl_update}
+%(as in the WL algorithm),
+%But not necessarily so
+%For a multiple-bin scheme.
+%%
+%Below, we discuss the single-bin scheme
+%In Section \ref{sec:single-bin},
+%And the multiple-bin case
+%In Section \ref{sec:multiple-bin}.
+
+\subsection{\label{sec:single-bin}
+Single-bin scheme}
 
@@
 we shall derive the optimal $\alpha(t)$
-for the single-bin updating scheme,
+for the single-bin scheme,
 Eq. \eqref{eq:wl_update}.
@@
 
-\subsection{Differential equation}
+\subsubsection{Differential equation}
 
@@
 %
-%The noise is not necessarily white.
+%The noise is not necessarily white,
 and the correlation functions
@@
 
-\subsection{Linear approximation}
+\subsubsection{Linear approximation}
 
@@
 
-\subsection{One-variable problem}
+\subsubsection{One-variable problem}
 
@@
 
-\subsection{Optimal $\alpha(t)$ for the single-bin updating scheme}
+\subsubsection{Optimal $\alpha(t)$ for the single-bin scheme}
 
@@
 minimizing Eq. \eqref{eq:error_sum}
-for the single-bin updating scheme,
+for the single-bin scheme,
 we find from Eq. \eqref{eq:dxdt_WL}
@@
 
-\section{\label{sec:multiple-bin}
-Multiple-bin update}
-
-We now consider the multiple-bin updating scheme.
+\subsection{\label{sec:multiple-bin}
+Multiple-bin scheme}
+
+We now consider the multiple-bin scheme.
 %
@@
 
-\subsection{Updating matrix $\mathbf w$}
+\subsubsection{Updating matrix}
 
@@
 
-\subsection{Linearization and decoupling}
+\subsubsection{Linearization and decoupling}
 
@@
 
-\subsection{Minimization of the error function}
+\subsubsection{Minimization of the error function}
 
@@
 
-\subsection{White-noise approximation}
-
-In the asymptotically regime,
-where $t \gg 1/\max \alpha(t)$,
+\subsubsection{White-noise approximation}
+
+In the asymptotic regime,
 we can approximate $\kappa_i(t)$ as
@@
   \sum_{i = 2}^n
-  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt',
+  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt'.
   \label{eq:error_mbin1}
@@
 %
+The next order correction to Eq. \eqref{eq:kappa_delta}
+would lead to a correction to $\mathcal E$
+that is about $\alpha_{\max}^2$ as large.
 %
@@
 
-\subsection{\label{sec:optWL}
-Optimality of the single-bin updating scheme}
-
-We can show from Eq. \eqref{eq:error_mbin1}
-that the single-bin updating scheme
-is an optimal updating scheme.
+\subsubsection{\label{sec:optWL}
+Optimality of the single-bin scheme}
+
+We now show
+that the single-bin scheme
+is an optimal scheme.
+%
+Consider a class of $\mathbf w$ matrices
+with the same set of eigenvectors,
+hence same $\Gamma_i$'s,
+but with different eigenvalues, $\lambda_i$'s.
 %
@@
 %
+Thus, the inequality sets a lower bound
+of the error $\mathcal E$
+expressed in Eq. \eqref{eq:error_mbin1},
+%
 The equality is achieved
@@
 Thus, in terms of convergence,
-the single-bin updating scheme
+the single-bin scheme
 (as adopted in the WL algorithm)
@@
 
-\subsection{Optimal coefficient of proportionality}
+\subsubsection{Optimal coefficient of proportionality}
 
@@
 
-\subsection{Uniform band matrix}
+\subsection{\label{sec:band-matrix}
+Uniform band matrix}
 
@@
 and optimal value is given by $\mu_1 = 0$,
-which is the single-bin update scheme.
+which is the single-bin scheme.
 
@@
 %
-Below we show that the updating scheme
+Below we show that this scheme
 is stable, although
@@
 
-\section{Numerical results}
+\section{\label{sec:results}
+Numerical results}
 
@@
 %
-Initially $\alpha$ is set to $\alpha_0 = 1/n$
+Initially $\alpha$ is set to some $\alpha_0$
 for a sufficiently long time.
@@
 $$
-\alpha(t) = \min\left\{ \alpha_0, \frac{ 1 } { \lambda \, t } \right\}.
-$$
-
-1. Single-bin update.
+\alpha(t) = \frac{ 1 } { \lambda \, (t + t_0) },
+$$
+where $t_0 = 1/(\lambda \, \alpha_0)$
+to make $\alpha(t)$ continuous.
+
+1. Single-bin scheme.
 Show the error versus $\lambda$.
 
-2. Multiple-bin update scheme.
+Change upon using local sampling scheme.
+
+2. Multiple-bin scheme.
 Error vs. $\lambda$.
 
-3. Optimality of the Single-bin update.
+3. Optimality of the Single-bin scheme.
 Error vs. $\mu_1$.
 
+\section{\label{sec:conclusion}
+Conclusions and Discussions}
+
+We verified the optimality of the inverse-time formula.
+
+We have shown that
+without a priori information,
+the single-bin scheme is asymptotically
+the best among all updating scheme
+in term of convergence.
+
+\bibliography{simul}
 \end{document}

### 2016-01-18 prog/corr.h
invt.c: add analytical error code

@@ -139,11 +139,12 @@ static int corr_compute(corr_t *c, double *uu, int j,
  * if `subave` is true, the time-average value is subtracted
  * from c->arr[] before computing the correlation function
  * otherwise, zero is assumed as the average */
-static int corr_save(corr_t *c, int dt, double tol,
+static int corr_save(corr_t *c, int dt,
+    int tmax, double tol,
     int subave, const char *fn)
 {
-  int i, j, n = c->n;
-  double *uu0, *uu, *uave;
+  int i, j, n = c->n, stop;
+  double *uu0, *uu, *uave = NULL;
   FILE *fp;
 
   if ( (fp = fopen(fn, "w")) == NULL ) {
@@ -163,7 +164,9 @@ static int corr_save(corr_t *c, int dt, double tol,
   /* save the heading */
   fprintf(fp, "# %d %d %d\n", n, dt, c->cnt);
 
-  for ( j = 0; j < c->cnt; j++ ) {
+  if ( tmax <= 0 ) tmax = c->cnt;
+
+  for ( j = 0; j < tmax; j++ ) {
     /* compute correlation functions
      * at a separation of j frames */
     corr_compute(c, uu, j, uave);
@@ -174,8 +177,6 @@ static int corr_save(corr_t *c, int dt, double tol,
         uu0[i] = uu[i];
       }
     } else {
-      int stop = 1;
-
       /* decide if to stop the calculation */
       for ( i = 0; i < n; i++ ) {
         if ( uu[i] > uu0[i] * tol ) {

### 2016-01-18 prog/invtpar.h

@@ -25,14 +25,15 @@ typedef struct {
   double *p; /* target distribution */
   double alpha0; /* initial updating magnitude */
   int fixa; /* fix alpha during the entire process */
-  int nbn; /* width of the updating window function */
-  double nbs[NBMAX + 1]; /* shape of the window function */
-  double wgaus; /* width of the Gaussian window */
+  int winn; /* width of the updating window function */
+  double win[NBMAX + 1]; /* shape of the window function */
+  double wingaus; /* width of the Gaussian window */
   double initrand; /* magnitude of the initial Gaussian noise */
   int kcutoff; /* cutoff wave number of the initial noise */
   int sampmethod; /* sampling method */
   double tcorr; /* correlation time for the sampling method */
 
+  int docorr; /* compute correlation functions */
   int nstcorr; /* time interval of computing autocorrelation function */
   double corrtol; /* error tolerance of the autocorrelation function */
   char fncorr[FILENAME_MAX]; /* file name for the autocorrelation function */
@@ -129,17 +130,18 @@ static void invtpar_init(invtpar_t *m)
   m->p = NULL;
   m->alpha0 = 0.0;
   m->fixa = 0;
-  m->nbn = 1; /* single bin update */
-  for ( i = 1; i < m->nbn; i++ ) {
-    m->nbs[i] = 0;
+  m->winn = 1; /* single bin update */
+  for ( i = 1; i < m->winn; i++ ) {
+    m->win[i] = 0;
   }
-  m->nbs[0] = 1;
-  m->wgaus = 0;
+  m->win[0] = 1;
+  m->wingaus = 0;
   m->initrand = 0;
   m->kcutoff = 0;
   m->sampmethod = 0;
   m->tcorr = 0.0; /* perfect sampling */
 
+  m->docorr = 0; /* compute correlation functions */
   m->nstcorr = 0; /* don't do correlation functions */
   m->corrtol = 1e-8;
   m->fncorr[0] = '\0';
@@ -173,48 +175,58 @@ static void invtpar_compute(invtpar_t *m)
     m->t0 = m->c / m->alpha0;
   }
 
-  if ( m->fixa ) {
-    /* if a fixed alpha is used,
-     * prepare parameters for the
[…]

### 2016-01-19 prog/invt.c
invtpar.h: add invtpar_keymatch

@@ -126,7 +126,8 @@ static double simulmeta(const invtpar_t *m, double *err0)
 static double invt_run(invtpar_t *m)
 {
   double err, see = 0, averr, errref;
-  double err0, see0 = 0, averr0, err0ref;
+  double err0, see0 = 0, averr0, err0ref; /* initial */
+  double err1ref; /* final saturated */
   int i;
 
   mtscramble( time(NULL) );
@@ -139,8 +140,13 @@ static double invt_run(invtpar_t *m)
     /* do multiple runs to compute the average error */
 
     /* compute the prediction from the analytical result */
+    /* initial saturated error */
     err0ref = esterror0_ez(m->alpha0,
-        m->n, m->winn, m->win, m->sampmethod, m->verbose);
+        m->n, m->winn, m->win, m->sampmethod,
+        "initial", m->verbose);
+    err1ref = esterror0_ez(m->c / (m->t0 + (double) m->nsteps),
+        m->n, m->winn, m->win, m->sampmethod,
+        "final", m->verbose);
     errref = esterror_ez(m->c, (double) m->nsteps, m->t0,
         m->n, m->winn, m->win, m->sampmethod, m->verbose);
 
@@ -160,10 +166,12 @@ static double invt_run(invtpar_t *m)
 
     averr = sqrt( see / m->ntrials );
     averr0 = sqrt( see0 / m->ntrials );
-    printf("average error: %10.8f -> %10.8f, sqr %e -> %e, "
-           "ref: %10.8f --> %10.8f, sqr %e -> %e\n",
-        averr0, averr, averr0 * averr0, averr * averr,
+    printf("average error: %10.8f -> %10.8f, sqr %e -> %e\n",
+        averr0, averr, averr0 * averr0, averr * averr);
+    printf("predicted val: %10.8f -> %10.8f, sqr %e -> %e\n",
         err0ref, errref, err0ref * err0ref, errref * errref);
+    printf("saturated val: %10.8f -> %10.8f, sqr %e -> %e\n",
+        err0ref, err1ref, err0ref * err0ref, err1ref * err1ref);
   }
 
   return err;

### 2016-01-19 prog/invt.h

@@ -285,9 +285,9 @@ static double *estgamma(int n, int sampmethod)
     } else if ( sampmethod == SAMPMETHOD_HEATBATH ) {
       gamma[i] = 1.0;
     } else {
-      if ( i == 1 ) {
-        /* complain once */
-        fprintf(stderr, "Error: unknown sampling method, %d\n", sampmethod);
+      if ( i == 1 ) { /* complain only once */
+        fprintf(stderr, "Error: unknown sampling method, %d\n",
+            sampmethod);
       }
       gamma[i] = 1.0;
     }
@@ -296,6 +296,63 @@ static double *estgamma(int n, int sampmethod)
   return gamma;
 }
 
+
+
+
+/* print out the values of lambda_i and gamma_i */
+static void dumplambdagamma(int n, const double *lamarr,
+    const double *gamma, const double *xerr)
+{
+  int i;
+
+  fprintf(stderr, "   i:   lambda_i      gamma_i\n");
+  for ( i = 0; i < n; i++ ) {
+    fprintf(stderr, "%4d: %10.6f %10.3f %20.6e\n",
+        i + 1, lamarr[i], gamma[i], xerr[i]);
+  }
+}
+
+
+
+/* estimate the error of a constant alpha
+ * according to the analytical prediction */
+static double esterror0_ez(double alpha,
+   int n, int winn, double *win, int sampmethod,
+   const char *name, int verbose)
+{
+  int i;
+  double *lamarr, *gamma, *xerr, err;
+
+  xnew(xerr, n);
+
+  /* estimate the eigenvalues of the w matrix,
+   * for the updating scheme */
+  lamarr = esteigvals(n, winn, win);
+
+  /* estimate the correlation integrals
+   * of the eigenmodes of the w matrix,
+   * for the updating scheme */
+  gamma = estgamma(n, sampmethod);
+
+  err = 0;
+  for ( i = 0; i < n; i++ ) {
+    xerr[i] = alpha * lamarr[i] / (alpha * lamarr[i] + 1 / gamma[i]);
+    err += xerr[i];
+  }
+
+  /* print out the values of lambda_i and gamma_i */
+  if ( verbose >= 2 ) {
+    dumplambdagamma(n, lamarr, gamma, xerr);
+  }
+  fprintf(stderr, "estimated %s saturated error %g, sqr: %e\n",
[…]

### 2016-01-19 prog/invtpar.h

@@ -67,11 +67,15 @@ const char *sampmethod_names[][MAX_OPT_ALIASES] = {
 
 
 /* parse a string into an array */
-static int readarray(double *arr, int n, char *s)
+static int readarray(double *arr, int n, const char *s0)
 {
   const char *delims = ",;:";
-  char *p;
+  char *p, *s;
   int i = 0;
+
+  /* make a copy of s0 */
+  xnew(s, strlen(s0) + 1);
+  strcpy(s, s0);
 
   p = strtok(s, delims);
   while ( p != NULL ) {
@@ -80,6 +84,9 @@ static int readarray(double *arr, int n, char *s)
     if ( i >= n )
       break;
   }
+
+  free(s);
+
   return i;
 }
 
@@ -126,7 +133,7 @@ static void invtpar_init(invtpar_t *m)
 
   m->c = 1.0;
   m->t0 = 0;
-  m->n = 10;
+  m->n = 100;
   m->p = NULL;
   m->alpha0 = 0.0;
   m->fixa = 0;
@@ -255,6 +262,76 @@ static void invtpar_finish(invtpar_t *m)
 
 
 
+/* match string key and value pairs */
+static int invtpar_keymatch(invtpar_t *m,
+    const char *key, const char *val)
+{
+  if ( strcmp(key, "n") == 0 ) {
+    m->n = atoi(val);
+  } else if ( strcmpfuzzy(key, "c") == 0
+           || strcmpfuzzy(key, "invt-c") ) {
+    m->c = atof(val);
+  } else if ( strcmpfuzzy(key, "t0") == 0 ) {
+    m->t0 = atof(val);
+  } else if ( strcmpfuzzy(key, "alpha0") == 0
+           || strcmpfuzzy(key, "a0") == 0 ) {
+    m->alpha0 = atof(val);
+  } else if ( strstartswith(key, "fixa") ) {
+    m->fixa = 1;
+  } else if ( strstartswith(key, "nb")
+           || strstartswith(key, "neighbo")
+           || strcmpfuzzy(key, "win") == 0
+           || strcmpfuzzy(key, "window") == 0 ) {
+    m->winn = 1 + readarray(m->win + 1, NBMAX, val);
+  } else if ( strcmpfuzzy(key, "wgaus") == 0
+           || strcmpfuzzy(key, "wingaus") == 0
+           || strcmpfuzzy(key, "width-Gaussian") == 0
+           || strcmpfuzzy(key, "sigma") == 0
[…]

### 2016-01-19 doc/invt.tex
invt.c: add analytical error code
invtpar.h: add invtpar_keymatch
correcting mistakes in invt.tex

@@
   along a quantity of interest, $z$,
-  by adaptively constructing a bias potential,
-  which is the negative potential of mean force (PMF).
+  by adaptively constructing a bias potential
+  that offsets the potential of mean force (PMF).
   %
@@
   although metadynamics produces a smoother PMF profile,
+  which is necessary for stable numerical differentiation
+  in computing the force,
   it is less effective in reducing errors with the shorter wavelengths.
@@
 along a quantity of interest $z$, and to see how
-the distribution changes with the external conditions,
+the distribution changes with external conditions,
 such as temperatures and pressures.
@@
 This alleviates the above problem of local trapping
-as the system is now forced to travel along the $z$ direction,
+as the system is now forced to travel more flexibly
+along the $z$ direction,
 which often represents a slow reaction coordinate.
@@
 %
-Note that the bias potential is directly related to the PMF
-as the former must precisely offset the latter
+Note that the bias potential must exactly cancel the PMF
 to achieve a flat distribution.
@@
 are two widely-used techniques for this purpose,
-in which a bias potential is actively build-up
-based on the simulation history.
+in which a bias potential is actively built-up on-the-fly
+during simulation.
 %
@@
 by elevating the bias potential there.
+%
+A main difference of the two techniques
+is that in the WL case, an update is limited to the bin
+containing the current $z$,
+while metadynamics adopts a more widespread window
+covering several neighboring bins.
+%
+The latter ensures a smooth profile
+suitable for numerical differentiation for force calculation,
+and hence is more often used molecular dynamics simulations.
 %
@@
 and try to generalize it to
-updating schemes involving neighboring bins,
+updating schemes involving several neighboring bins,
 such as those used in metadynamics.
@@
 entropic sampling\cite{berg1992, lee1993},
-can be inconvenient in practice,
-for we need to make the period
-as long as a multiple of
-the autocorrelation time,
-which is difficult to estimate for an unknown system.
-
-For later convenience, we introduce
+can be inconvenient in practice.
+%
+For example, a reasonably precise the histogram
+requires an accumulation period
+longer than the autocorrelation time,
+whereas the latter is difficult to
+estimate beforehand for an unknown system.
+
+For later convenience, we introduce a shifted bias potential
 %
@@
 Since the last two terms of Eq. \eqref{eq:v_def}
-are constant during the course of simulation,
-updates to $V_i$ are equivalent to those to $v_i$.
-%
-In terms of $v_i$, Eq. \eqref{eq:pi_p_phi}
+remain constant during the course of simulation,
+updates to $V_i$'s are equivalent to those to $v_i$'s.
+%
+In terms of $v_i$'s, Eq. \eqref{eq:pi_p_phi}
 becomes
@@
 %
-We refer to the above scheme as the single-bin scheme,
+We refer to the above scheme as the \emph{single-bin scheme},
 for it applies only to the bias potential
@@
 %
-By contrast, a multiple-bin scheme
-means the general case in which
-several neighboring bins are updated.
-
-%The values of $v_j(t)$ at $j \ne i(t)$
-%are kept unchanged.
+In a more general \emph{multiple-bin scheme}, however,
+several neighboring bins are updated:
+%
+\begin{equation}
+  v_i(t+1)
+  =
+  v_i(t)
+  +
+  w_{i, \, i(t)}
+  \frac{ \alpha(t) } { p_i },
+  \label{eq:mbin_update}
+\end{equation}
+%
+where $w_{ij}$ are a set of coefficients satisfying
+$\sum_{j=1}^n w_{ij} = 1$.
+%
+We shall discuss other restrictions on $w_{ij}$
+in Sec. \ref{sec:updating-matrix}.
 
@@
 %
-However, the continuous update
-makes $v_i(t)$ a fluctuating quantity,
+However, the above updates
+change $v_i(t)$ continuously,
 and Eq. \eqref{eq:pi_p_v} no longer holds
@@
 we are allowed to enter a new stage
-with a reduced $\alpha$
+with a reduced $\alpha$\cite{wang2001, wang2001pre}
 (usually half as large as
@@
 too quickly, making the asymptotic error
-saturate.
+saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
 
@@
 This surprisingly simple formula has attracted
-several studies, notably the one by Zhou.
-%
-There are, however, some reservations about
+several studies\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008,
+morozov2007, zhou2008, morozov2009,
+komura2012, caparica2012, caparica2014}.
+%
+There are, however, some reservation about
 the constant of proportionality:
@@
 might be $C/t$ with a constant $C$
-different from $1.0$.
+different from $1.0$\cite{
+morozov2007, zhou2008, morozov2009}.
 
@@
 as a functional of $\alpha(t)$,
-and then find the optimal value
-by functional variation.
+and then minimize it by variation.
 
@@
 %
-However, the difference between $v_i$ and $\bar v$,
+However, we expect the difference between $v_i$ and $\bar v$,
 %
@@
 %
-is expected to be small in the asymptotic regime,
-and we can expand Eq. \eqref{eq:h_ave} as
+to be small in the asymptotic regime,
+and expand Eq. \eqref{eq:h_ave} as
 \begin{align}
@@
 %
-With Eqs.
+From Eqs.
 \eqref{eq:vt_diffeq},
 \eqref{eq:h_split},
-\eqref{eq:dvbardt},
-and
-\eqref{eq:hdet_linearize},
+\eqref{eq:dvbardt}-\eqref{eq:hdet_linearize},
 we get a set of decoupled equations
@@
 \end{equation}
+\note{The derivation:
+$$
+\begin{aligned}
+  \dot x_i
+  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
+            {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \dot v_i - \dot{ \bar v }
+  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:dvbardt}}}
+            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \alpha(t) \left( \frac{ h_i } { p_i } - 1 \right)
+  \\
+  &\stackrel{\mathrm{Eq.\; \eqref{eq:h_split}}}
+            {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \alpha(t) \left( \frac{ \langle h_i \rangle } { p_i } + \frac{ \zeta_i } { p_i } - 1 \right)
+  \stackrel{\mathrm{Eq.\; \eqref{eq:hdet_linearize}}}
+            {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \alpha(t) \left[ -x_i(t) + \frac{ \zeta_i } { p_i } \right].
+\end{aligned}
+$$
+}
 
@@
 %
-To do so, we shall first consider
-a simpler one-variable problem.
+Fortunately, we shall find that it suffices to consider
+a simpler one-variable problem
+in this case.
 
@@
 %
-We wish to show that the optimal $\alpha(t)$
+We wish to show that the $\alpha(t)$
 of minimizing $\langle x^2(t) \rangle$ at long times
@@
 \end{align}
+So
+\begin{align}
+\left\langle x^2(t) \right\rangle
+&=
+\int_0^t \int_0^t
+  \dot u\bigl( q(t') \bigr) \,
+  \dot u\bigl( q(t'') \bigr) \,
+  \kappa(t' - t'') \, dt'' \, dt'
+\notag
+\\
+&\hphantom{=}
++\langle x^2(0) \rangle e^{-2 \, \lambda \, q(t)}.
+\label{eq:x2t_average}
+\end{align}
 
@@
 %
-Then, for a large time $t$,
-the first term on the right-hand side
+Then, for a large $t$,
+the second term on the right-hand side
 of Eq. \eqref{eq:xt_solution} can be neglected, and
@@
   \kappa(t' - t'') \, dt'' \, dt'.
-\label{eq:x2t_average}
+\label{eq:x2t_average_asym}
 \end{align}
@@
 minimizing Eq. \eqref{eq:error_sum}
-for the single-bin scheme,
-we find from Eq. \eqref{eq:dxdt_WL}
-that with $\lambda = 1$,
+for the single-bin scheme.
+%
+Note that Eq. \eqref{eq:dxdt_WL}
+is of the form of Eq. \eqref{eq:dxdt_alpha}.
+%
+Thus, with $\lambda = 1$,
 Eq. \eqref{eq:alpha_opt}
@@
 %
-Thus, Eq. \eqref{eq:alpha_invt} is optimal
-up to a constant $t_0$. %defining the origin of $t$.
+In other words, the $\alpha(t)$ given by
+Eq. \eqref{eq:alpha_invt} is optimal
+up to a constant $t_0$.
 
@@
 
-We now consider the multiple-bin scheme.
+We now consider the multiple-bin scheme,
+Eq. \eqref{eq:mbin_update}.
 %
@@
 
-The procedure of optimization is largely
-the same as the single-bin case,
+The optimization procedure is similar to the single-bin case,
 whereas the matrix nature brings about
-some technical difficulties that require
+some technical complexities that require
 certain restrictions and approximations.
 
-\subsubsection{Updating matrix}
+\subsubsection{\label{sec:updating-matrix}
+Updating matrix}
 
@@
 $\mathbf p = (p_1, \dots, p_n)$
-is that in Eq. \eqref{eq:vt_diffeq_mbin}
-when $h_j(t)$ coincides with $p_j$,
-the rate of change $\dot v_i(t)$
+is the following.
+%
+In Eq. \eqref{eq:vt_diffeq_mbin},
+if $h_j(t)$ were the same as $p_j$,
+the rates of change $\dot v_i(t)$'s
 should be independent of $i$.
 %
-This allows $v_i(t)$ to approach a constant
-asymptotically.
+This allows $v_i(t)$'s to grow uniformly
+in the asymptotic regime.
 %
@@
 %
-This means $(1, \dots, 1)^T$
+The above condition means that $(1, \dots, 1)^T$
 is a right eigenvector of $\mathbf w$
@@
 as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$
-that can be diagonalized
+that can be diagonalized\cite{vankampen}
 with a set of orthonormal eigenvectors:
@@
 $\phi_{ij} = \sqrt{p_j} \, \varphi_{ij}$ satisfies
-the orthonormal conditions:
+the orthonormal conditions\cite{vankampen}:
 %
@@
 Note that, in a stable updating scheme,
-all eigenvalues $\lambda_k$ must be nonnegative,
+all eigenvalues $\lambda_k$ of $\mathbf w$
+must be nonnegative,
 for otherwise there is a fluctuating mode
@@
 Eqs. \eqref{eq:dvbardt} and \eqref{eq:px_sum}
-remain true.
-%
-%Thus, for Eq. \eqref{eq:x_def}, we have
+remain true, and
 %
@@
 \\
-&=
+&\approx
 -\alpha(t) \sum_{j = 1}^n
@@
 
+\note{The proof of Eq. \eqref{eq:dvbardt},
+$$
+\begin{aligned}
+  \dot{ \bar v }
+  &= \sum_{i = 1}^n p_i \, \dot v_i
+  \stackrel{\mathrm{Eq.\; \eqref{eq:vt_diffeq_mbin}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \sum_{j = 1}^n \alpha(t) \, \frac{ h_j(t) } { p_j } \sum_{ i = 1 }^{n} p_i \, w_{ij}
+  \\
+  &
+  \stackrel{\mathrm{Eq.\; \eqref{eq:w_balance}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \sum_{j = 1}^n \alpha(t) \, h_j(t) = \alpha(t).
+\end{aligned}
+$$
+Equation \eqref{eq:px_sum} follows directly from the definition
+Eq. \eqref{eq:x_def}.
+}
+
 Now by diagonalizing the matrix $\mathbf w$
-using matrix defined in Eqs. \eqref{eq:eig_w},
+using the eigenvectors defined in Eqs. \eqref{eq:eig_w},
 \eqref{eq:eig_orthonormal_cols},
@@
 %
-We note that there is one mode, say
+Note that there is always one mode, say
 $y_1 = \sum_{j=1}^n p_j \, x_j$
@@
 would lead to a correction to $\mathcal E$
-that is about $\alpha_{\max}^2$ as large.
-%
+that is about $\alpha_{\max}^2$ times as large.
+
 \note{This is equivalent to saying that the Fourier transform
@@
   %
-  \approx \delta(t - t').
+  \approx \delta(t - t'),
   \label{eq:kappa_perfect}
@@
 %
+or
+$\Gamma_i \approx 1$,
 where we have assumed the large $n$ limit
@@
 %
-\note{A short proof of the above version
-  of the Cauchy-Schwarz inequality
+\note{A short proof of the above inequality
   is the following.
@@
   %
-  Thus, the right-hand side,
-  as a quadratic polynomial of $X$,
-  must have a non-positive discriminant.
+  Thus, the left-hand side,
+  expanded as a quadratic polynomial of $X$,
+  $A \, X^2 + B \, X + C$,
+  must have a non-positive discriminant,
+  $B^2 - 4 \, A \, C$.
   %
@@
 \frac{1}{t+t_0}
-\sum_{i=2}^n \frac{ \lambda_i^2 } { \lambda \, (2 \lambda_i - \lambda) }
+\sum_{i=2}^n
+  \frac{       \Gamma_i \, \lambda_i^2      }
+       { \lambda \, (2 \lambda_i - \lambda) }
 \left[
   1 - \left(
-    1 + \frac{t}{t_0}
+        1 + \frac{ t }{ t_0 }
       \right)^{1 - 2 \, \lambda_i/\lambda}
@@
        { t }
-  \frac{ \lambda_i^2 }
+  \frac{         \Gamma_i \, \lambda_i^2       }
        { \lambda \, (2 \, \lambda_i - \lambda) }
@@
   %
-  \frac{   1    }
+  \frac{ \Gamma_i }
        {  4 \, t  }
@@
   %
-  \frac{ t_0^{ 2 \, \lambda_i / \lambda  - 1} }
+  \frac{ t_0^{ 2 \, \lambda_i / \lambda  - 1 } }
        {   t^{ 2 \, \lambda_i / \lambda      } }
-  \frac{ \lambda_i^2 }
+  \frac{      \Gamma_i \, \lambda_i^2          }
        { \lambda \, (\lambda - 2 \, \lambda_i) }
@@
   \begin{dcases}
-    \frac{ 1 }{ t + t_0 }
-    \frac{ \lambda_i^2 } { \lambda \, (2 \, \lambda_i - \lambda) }
+    \frac{    1    }
+         { t + t_0 }
+    \frac{       \Gamma_i \, \lambda_i^2         }
+         { \lambda \, (2 \, \lambda_i - \lambda) }
     &
@@
     %
-    \frac{1}{4 \, (t + t_0) } \ln\frac{t + t_0}{t_0}
+    \frac{    \Gamma_i    }
+         { 4 \, (t + t_0) }
+    \ln \frac{ t + t_0 }
+             {   t_0  }
     &
@@
          { (t + t_0)^{ 2 \, \lambda_i / \lambda } }
-    \frac{ \lambda_i^2 }
+    \frac{      \Gamma_i \, \lambda_i^2          }
          { \lambda \, (\lambda - 2 \, \lambda_i) }
@@
 and $\mu_l = 0$ for $l > b$.
+%
+On the right-hand side of Eq. \eqref{eq:w_band},
+the first term represents the relative updating magnitude
+to a neighbor that is $|i - j|$ bins away,
+whereas the next two terms are added to avoid
+boundary artifacts.
 
@@
 with eigenvalues
-\begin{equation}
-  \lambda_k = \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}.
+\begin{align}
+  \lambda_k
+  &=
+  \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
+  \notag \\
+  &=
+  1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
+  .
 \label{eq:wband_eigenvalue}
-\end{equation}
+\end{align}
 %
@@
 %\eqref{eq:optimal_lambda_approx}
-to an integral,
+to an integral.
+%
+Particularly,
+for perfect sampling, $\Gamma_i \approx 1$,
+we get
 $$
@@
 no matter its sign,
-and optimal value is given by $\mu_1 = 0$,
-which is the single-bin scheme.
+and the minimal value, obtained at $\mu_1 = 0$,
+corresponds to the single-bin scheme.
 

### 2016-01-19 prog/invt.h

@@ -364,7 +364,7 @@ static double esterr1(double lambda, double t, double t0,
     double lambda_i, double gamma_i)
 {
   const double tol = 1e-10;
-  double del, lammax, r;
+  double del, lammax, r, err1;
 
   del = fabs(lambda_i * 2 - lambda);
   /* lammax is the larger of lambda and lambda_i */
@@ -375,7 +375,9 @@ static double esterr1(double lambda, double t, double t0,
   } else {
     r = lambda_i / lambda;
     //printf("lambda %g, %g, r %g, gamma_i %g, t %g, t0 %g\n", lambda, lambda_i, r, gamma_i, t, t0); getchar();
-    return gamma_i / (t + t0) * ( r * r / ( 2 * r - 1 ) )
+    /* remainder error */
+    err1 = r / t0 * gamma_i * pow(t0/(t+t0), r);
+    return err1 + gamma_i / (t + t0) * ( r * r / ( 2 * r - 1 ) )
            * ( 1 - pow(t0 / (t + t0), 2 * r - 1) );
   }
 }

### 2016-01-20 prog/corr.h
add langtest.c

@@ -140,10 +140,10 @@ static int corr_compute(corr_t *c, double *uu, int j,
  * from c->arr[] before computing the correlation function
  * otherwise, zero is assumed as the average */
 static int corr_save(corr_t *c, int dt,
-    int tmax, double tol,
+    double tmax, double tol,
     int subave, const char *fn)
 {
-  int i, j, n = c->n, stop;
+  int i, j, jmax, n = c->n, stop;
   double *uu0, *uu, *uave = NULL;
   FILE *fp;
 
@@ -164,9 +164,13 @@ static int corr_save(corr_t *c, int dt,
   /* save the heading */
   fprintf(fp, "# %d %d %d\n", n, dt, c->cnt);
 
-  if ( tmax <= 0 ) tmax = c->cnt;
+  if ( tmax <= 0 ) {
+    jmax = c->cnt;
+  } else {
+    jmax = (int) ( tmax / dt );
+  }
 
-  for ( j = 0; j < tmax; j++ ) {
+  for ( j = 0; j < jmax; j++ ) {
     /* compute correlation functions
      * at a separation of j frames */
     corr_compute(c, uu, j, uave);

### 2016-01-20 prog/invt.h

@@ -18,7 +18,10 @@ static int mc_metro_g(const double *v, int n, int i)
   int j, acc;
   double dv, r;
 
-  j = (int) ( (i + 1 + (n - 1) * rand01()) ) % n;
+  /*
+   * j = (int) ( (i + 1 + (n - 1) * rand01()) ) % n;
+   * */
+  j = (int) ( n * rand01() );
   dv = v[j] - v[i];
   if ( dv <= 0 ) {
     acc = 1;
@@ -150,6 +153,23 @@ static double geterror(double *v, int n, const double *p)
     x = v[i] * v[i];
     /* multiply the normalization factor */
     x *= ( p != NULL ) ? p[i] : 1.0 / n;
+    err += x;
+  }
+
+  return sqrt(err);
+}
+
+
+
+/* compute the root-mean-squared error of the modes `u`
+ * This routine here is demonstrate the normalization */
+__inline static double getuerror(double *u, int n)
+{
+  double err = 0, x;
+  int i;
+
+  for ( i = 0; i < n; i++ ) {
+    x = u[i] * u[i];
     err += x;
   }
 
@@ -336,7 +356,7 @@ static double esterror0_ez(double alpha,
 
   err = 0;
   for ( i = 0; i < n; i++ ) {
-    xerr[i] = alpha * lamarr[i] / (alpha * lamarr[i] + 1 / gamma[i]);
+    xerr[i] = 0.5 * alpha * lamarr[i] * gamma[i];
     err += xerr[i];
   }
 
@@ -376,7 +396,7 @@ static double esterr1(double lambda, double t, double t0,
     r = lambda_i / lambda;
     //printf("lambda %g, %g, r %g, gamma_i %g, t %g, t0 %g\n", lambda, lambda_i, r, gamma_i, t, t0); getchar();
     /* remainder error */
-    err1 = r / t0 * gamma_i * pow(t0/(t+t0), r);
+    err1 = 0.5 * r / t0 * gamma_i * pow(t0/(t+t0), 2 * r);
     return err1 + gamma_i / (t + t0) * ( r * r / ( 2 * r - 1 ) )
            * ( 1 - pow(t0 / (t + t0), 2 * r - 1) );
   }
@@ -407,7 +427,7 @@ static double esterrn(double lambda, double t, double t0,
 
 
 
-/* estimate the error of the updating schedule
+/* estimate the final error of the updating schedule
  * alpha(t) = c / (t + t0)
  * according to the analytical prediction */
 static double esterror_ez(double c, double t, double t0,

### 2016-01-20 doc/invt.tex
add langtest.c
invt.tex: add discussion on the correlation integrals

@@
 and expand Eq. \eqref{eq:h_ave} as
+%
 \begin{align}
@@
 \approx
-\frac{ p_i ( 1 - x_i ) }
+\frac{                 p_i (1 - x_i) }
      { \sum_{ j = 1}^n p_j (1 - x_j) }
-\notag \\
+\notag
+\\
 &\approx
@@
 \end{align}
+%
 where we have used $\sum_{j=1}^n p_j = 1$,
@@
 where
-$\phi_{ij} = \sqrt{p_j} \, \varphi_{ij}$ satisfies
+$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$ satisfies
 the orthonormal conditions\cite{vankampen}:
@@
 %
-where, $\Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt$.
+where,
+\begin{equation}
+  \Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt,
+  \notag
+  %\label{eq:Gamma_integral}
+\end{equation}
+or in discrete time,
+\begin{equation}
+  \Gamma_i = \sum_{t = -\infty}^\infty \kappa_i(t).
+  \label{eq:Gamma_sum}
+\end{equation}
 %
@@
   once $\alpha(t)$ has dropped below a certain level.
-  \\
 }
+
 \note{
@@
   and $\Gamma_i \, \lambda_i$ as the mass,
-  then the right-hand side give the virial.
+  then the right-hand side gives the virial.
   %
@@
 }
-%
-For example,
-if we assume perfect sampling,
+
+\subsubsection{Integrals of the correlation functions}
+
+We can evaluate the $\Gamma_i$'s
+in two special case.
+%
+If the sampling is perfect,
+Eq. \eqref{eq:kappa_delta} is exact, and
+%
+\begin{equation}
+  \Gamma_i = 1.
+  \label{eq:Gamma_perfect}
+\end{equation}
+%
+In this case,
+the bin index $i$ is an independent random variable
+at each time step $t$, and
 %
@@
   =
-  p_i \, (1 - p_i) \, \delta_{ij} \, \delta(t - t').
+  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \, \delta(t - t'),
   \label{eq:zz_perfect}
@@
 %
+where $\delta(t - t')$ is understood as an adaption of $\delta_{t, t'}$
+from the discrete time case to the continuous limit. % hereinafter.
+%
+\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
+  with $t \ne t'$ (going back to the discrete time case),
+  $$
+  \begin{aligned}
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  &=
+  \langle \zeta_i(t)  \rangle
+  \langle \zeta_j(t') \rangle
+  \\
+  &=
+  \bigl\langle h_i(t)  - \langle h_i(t)  \rangle \bigr\rangle
+  \bigl\langle h_j(t') - \langle h_j(t') \rangle \bigr\rangle
+  =
+  0.
+  \end{aligned}
+  $$
+  Next, for $t = t'$, we have
+  $$
+  \zeta_i(t)
+  = h_i(t) - \langle h_i(t) \rangle
+  = \delta_{i, i(t)} - p_i,
+  $$
+  and
+  $$
+  \begin{aligned}
+  \left\langle
+    \zeta_i(t) \, \zeta_j(t)
+  \right\rangle
+  &=
+  \left\langle
+    \delta_{i, i(t)} \, \delta_{j, i(t)}
+  \right\rangle
+  -
+  p_i \, p_j
+  \\
+  &=
+  \left\langle
+    \delta_{i, i(t)}
+  \right\rangle
+  \delta_{i, j}
+  -
+  p_i \, p_j
+  =
+  p_i \, \delta_{ij}
+  - p_i \, p_j.
+  \end{aligned}
+  $$
+  %
+  The average
+  $\left\langle \delta_{i, i(t)} \right\rangle$ is $p_i$,
+  because as $i(t)$ goes through $1$ to $n$,
+  $\delta_{i, i(t)}$ gives $1$
+  if $i(t)$ matches $i$ with probability $p_i$,
+  or $0$ otherwise.
+}
+%
 Thus, from Eqs. \eqref{eq:eig_orthonormal_rows} and \eqref{eq:eta_def},
-we get
+we get, for $i \ge 2$,
 %
@@
   %
-  \approx \delta(t - t'),
+  = \delta(t - t'),
   \label{eq:kappa_perfect}
@@
 %
-or
-$\Gamma_i \approx 1$,
-where we have assumed the large $n$ limit
-such that $p_i \ll 1$
-in the last expression.
+\note{To show Eq. \eqref{eq:kappa_perfect},
+  $$
+  \begin{aligned}
+  \kappa_i(t - t')
+  &=
+  \sum_{j,k = 1}^n
+  \frac{ \phi_{ij} } { p_j }
+  \frac{ \phi_{ik} } { p_k }
+  \langle \zeta_j(t) \, \zeta_k(t') \rangle
+  \\
+  %
+  &=
+  \sum_{j,k = 1}^n
+  \frac{ \phi_{ij} } { p_j }
+  \frac{ \phi_{ik} } { p_k }
+  \left(
+    p_j \, \delta_{j, k} - p_j \, p_k
+  \right)
+  \, \delta(t - t')
+  \\
+  &=
+  \left(
+    \sum_{j = 1}^n
+    \frac{ \phi_{ij}^2 } { p_j }
+    -
+    \sum_{j = 1}^n \phi_{ij}
+    \sum_{k = 1}^n \phi_{ik}
+  \right)
+  \, \delta(t - t')
+  \\
+  &=
+  (1 - 0 \cdot 0) \, \delta(t - t').
+  \end{aligned}
+  $$
+  where,
+  we have used
+  Eq. \eqref{eq:eta_def}
+  on the first line,
+  %
+  Eq. \eqref{eq:zz_perfect}
+  on the second line,
+  and
+  Eq. \eqref{eq:eig_orthonormal_rows}
+  on the last line.
+  %
+  Note for $i \ge 2$,
+  $$
+  \sum_{j = 1}^n \phi_{ij}
+  =
+  \sum_{j = 1}^n \frac{ \phi_{ij} \, \phi_{1j} } { p_j }
+  =
+  \delta_{i1}
+  =
+  0.
+  $$
+  as the mode $1$, or the first row, $\phi_{1j}$,
+  is given by the equilibrium distribution $p_j$.
+}
+
+Another special case is the one-step process\cite{vankampen},
+in each step, the system hops to the previous $i - 1$ bin
+or the next bin $i + 1$ with equal probability $1/2$,
+ignoring out-of-boundary transitions.
+%
+We shall further assume that
+the eigenvectors of the $\mathbf w$ matrix
+coincide with the those of the transition matrix
+of the above sampling process
+(cf. Sec. \ref{sec:band-matrix}).
+%
+Since the transition matrix is tridiagonal,
+as given by Eq. \eqref{eq:wtridiag} with $\mu_1 = 1/2$,
+the eigenvalues are given
+by Eq. \eqref{eq:wband_eigenvalue} as
+$$
+\Lambda_i = 1 - 2 \sin^2 \frac{ (i - 1) \, \pi } { 2 \, n }.
+$$
+Note, however, these $\Lambda_i$'s are the eigenvalues of
+the transition matrix of sampling scheme,
+not to be confused with the eigenvalues, $\lambda_i$'s,
+of the $\mathbf w$ matrix.
+%
+The correlation function $\kappa_i(t)$ of
+decays as $\kappa_i(0) \, \Lambda_i^t$,
+with the same-time value $\kappa_i(0) = 1$ being the same as
+that in perfect sampling.
+%
+Thus, from Eq. \eqref{eq:Gamma_sum}, we get
+%
+\begin{equation}
+\Gamma_i
+=
+1 + 2 \, \sum_{t = 1}^\infty \Lambda_i^t
+=
+\frac{ 1 + \Lambda_i } { 1 - \Lambda_i }
+=
+\cot^2 \left( \frac{ i \, \pi } { 2 \, n } \right).
+\label{eq:Gamma_onestep}
+\end{equation}
+%
+We can use this case to model local sampling processes.
 
@@
 %
-For $\tau \ll t$,
-the dominant contribution comes from
-the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
-and
-$$
-\ddot u_n(\tau) = 0,
-\quad
-\dot u_n(\tau) = c,
-\quad
-\mathrm{and}\;
-u_n(\tau) = c \, (\tau + t_0),
-$$
-which means
-$$
-\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
-$$
+%For $\tau \ll t$,
+%the dominant contribution comes from
+%the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
+%and
+%$$
+%\ddot u_n(\tau) = 0,
+%\quad
+%\dot u_n(\tau) = c,
+%\quad
+%\mathrm{and}\;
+%u_n(\tau) = c \, (\tau + t_0),
+%$$
+%which means
+%$$
+%\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
+%$$
 
@@
 Consider a class of $\mathbf w$ matrices
-with the same set of eigenvectors,
+sharing the same set of eigenvectors,
 hence same $\Gamma_i$'s,
@@
 &
-\left( \int_0^t
-  dt' \sum_{i = 2}^n \Gamma_i \,
-                     \dot u_i^2\bigl( q(t') \bigr)
+\left(
+  \int_0^t dt'
+    \sum_{i = 2}^n
+      \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr)
 \right)
+%
 \left(
-\int_0^t
-  dt' \sum_{i = 2}^n \Gamma_i \, c_i^2
+  \int_0^t dt'
+    \sum_{i = 2}^n
+      \Gamma_i \, c_i^2
 \right)
+%
 \notag
@@
 \left(
-\int_0^t
-  dt' \sum_{i = 2}^n \Gamma_i \, c_i \,
-                     \dot u_i\bigl( q(t') \bigr)
+  \int_0^t dt'
+    \sum_{i = 2}^n
+      \Gamma_i \, c_i \, \dot u_i\bigl( q(t') \bigr)
 \right)^2
@@
 with eigenvalues
+%
 \begin{align}
@@
 Particularly,
-for perfect sampling, $\Gamma_i \approx 1$,
-we get
+for perfect sampling,
+we get from Eq. \eqref{eq:Gamma_perfect},
 $$

### 2016-01-20 prog/invt.h

@@ -12,6 +12,7 @@
 #include "invtpar.h"
 
 
+
 /* global Metropolis move */
 static int mc_metro_g(const double *v, int n, int i)
 {
@@ -295,16 +296,17 @@ static double *esteigvals(int n, int winn, const double *win)
 static double *estgamma(int n, int sampmethod)
 {
   int i;
-  double *gamma;
+  double *gamma, x;
 
   xnew(gamma, n);
 
   gamma[0] = 0;
   for ( i = 1; i < n; i++ ) {
     if ( sampmethod == SAMPMETHOD_METROGLOBAL ) {
-      gamma[i] = n / (n - 1.0);
+      gamma[i] = 1.0; /* n / (n - 1.0); */
     } else if ( sampmethod == SAMPMETHOD_METROLOCAL ) {
-      gamma[i] = 1.0 / (1 - cos(i*M_PI/n));
+      x = sin( i * M_PI * 0.5 / n );
+      gamma[i] = 1.0 / (x * x) - 1;
     } else if ( sampmethod == SAMPMETHOD_HEATBATH ) {
       gamma[i] = 1.0;
     } else {

### 2016-01-21 doc/invt.tex
add a template configuration file

@@
 
-Another special case is the one-step process\cite{vankampen},
-in each step, the system hops to the previous $i - 1$ bin
-or the next bin $i + 1$ with equal probability $1/2$,
+Another special case is the one-step process\cite{vankampen}:
+in each step, the system hops to the previous bin $i - 1$
+or the next bin $i + 1$ with equal probability $\frac 1 2$,
 ignoring out-of-boundary transitions.
 %
+The transition matrix is
+%
+\begin{equation}
+\arraycolsep=3.6pt\def\arraystretch{1.4}
+\mathbf T
+=
+\left(
+  \begin{array}{cccccccc}
+    \frac 1 2 & \frac 1 2 & 0 & \dots & 0 \\
+    \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
+    \vdots & &  & & \vdots \\
+    0 & \dots & \frac 1 2 & 0  & \frac 1 2 \\
+    0 & \dots & 0 & \frac 1 2 & \frac 1 2
+  \end{array}
+\right).
+\label{eq:T_tridiag}
+\end{equation}
+%
 We shall further assume that
-the eigenvectors of the $\mathbf w$ matrix
-coincide with the those of the transition matrix
-of the above sampling process
-(cf. Sec. \ref{sec:band-matrix}).
-%
-Since the transition matrix is tridiagonal,
-as given by Eq. \eqref{eq:wtridiag} with $\mu_1 = 1/2$,
-the eigenvalues are given
-by Eq. \eqref{eq:wband_eigenvalue} as
+the eigenvectors of $\mathbf w$
+coincide with those of $\mathbf T$,
+while the eigenvalues may differ
+(this is true for a class of matrices
+discussed in Sec. \ref{sec:band-matrix}).
+%
+The eigenvalues of $\mathbf T$ are
 $$
@@
 $$
-Note, however, these $\Lambda_i$'s are the eigenvalues of
-the transition matrix of sampling scheme,
-not to be confused with the eigenvalues, $\lambda_i$'s,
-of the $\mathbf w$ matrix.
-%
-The correlation function $\kappa_i(t)$ of
+Note that these $\Lambda_i$'s are
+not to be confused with the eigenvalues of $\mathbf w$,
+$\lambda_i$'s.
+%
+The correlation function $\kappa_i(t)$
 decays as $\kappa_i(0) \, \Lambda_i^t$,
@@
 =
-\cot^2 \left( \frac{ i \, \pi } { 2 \, n } \right).
+\cot^2 \left[ \frac{ (i - 1) \, \pi } { 2 \, n } \right].
 \label{eq:Gamma_onestep}
@@
 \note{
-  We add some steps for solution.
+  We add some steps for the above solution.
   Integrating $\dot u_i = c_i$ yields
@@
 %
-and to find the optimal value of $\lambda$.
+for some optimal value of $\lambda$.
 %

### 2016-01-21 prog/invt.h

@@ -305,8 +305,8 @@ static double *estgamma(int n, int sampmethod)
     if ( sampmethod == SAMPMETHOD_METROGLOBAL ) {
       gamma[i] = 1.0; /* n / (n - 1.0); */
     } else if ( sampmethod == SAMPMETHOD_METROLOCAL ) {
-      x = sin( i * M_PI * 0.5 / n );
-      gamma[i] = 1.0 / (x * x) - 1;
+      x = tan( i * M_PI * 0.5 / n );
+      gamma[i] = 1.0 / (x * x);
     } else if ( sampmethod == SAMPMETHOD_HEATBATH ) {
       gamma[i] = 1.0;
     } else {

### 2016-01-22 prog/invtpar.h
add predict.c

@@ -43,6 +43,12 @@ typedef struct {
   long ntrials; /* number of trials */
   int verbose; /* verbose level */
   const char *prog; /* name of the program */
+
+#ifdef CSCAN /* for predict.c */
+  double cmin;
+  double cdel;
+  double cmax;
+#endif /* CSCAN */
 } invtpar_t;
 
 
@@ -98,6 +104,12 @@ static void invtpar_init(invtpar_t *m)
   m->ntrials = 100;
   m->prog = "invt";
   m->verbose = 0;
+
+#ifdef CSCAN /* for predict.c */
+  m->cmin = 0.1;
+  m->cdel = 0.01;
+  m->cmax = 2.0;
+#endif /* CSCAN */
 }
 
 
@@ -226,6 +238,11 @@ static void invtpar_help(const invtpar_t *m)
   fprintf(stderr, "  --try=:        set the number of trials, default %ld\n", m->ntrials);
   fprintf(stderr, "  --step=:       set the number of simulation steps, default %ld\n", m->nsteps);
   fprintf(stderr, "  --equil=:      set the number of equilibration steps, default %ld\n", m->nequil);
+#ifdef CSCAN /* for predict.c */
+  fprintf(stderr, "  --cmin=:       set the minimal c in c-scan, default %g\n", m->cmin);
+  fprintf(stderr, "  --cdel=:       set the increment of c in c-scan, default %g\n", m->cdel);
+  fprintf(stderr, "  --cmax=:       set the maximal c in c-scan, default %g\n", m->cmax);
+#endif /* CSCAN */
   fprintf(stderr, "  -v:            be verbose, -vv to be more verbose, etc.\n");
   fprintf(stderr, "  -h, --help:    display this message\n");
   exit(1);
@@ -445,6 +462,20 @@ static int invtpar_keymatch(invtpar_t *m,
   {
     m->verbose = val ? atoi(val) : 1;
   }
+#ifdef CSCAN /* for predict.c */
+  else if ( strcmpfuzzy(key, "cmin") == 0 )
+  {
+    m->cmin = invtpar_getdouble(m, key, val);
+  }
+  else if ( strcmpfuzzy(key, "cdel") == 0 )
+  {
+    m->cdel = invtpar_getdouble(m, key, val);
+  }
+  else if ( strcmpfuzzy(key, "cmax") == 0 )
+  {
+    m->cmax = invtpar_getdouble(m, key, val);
[…]

### 2016-01-22 prog/predict.c

@@ -1,0 +1,53 @@
+/* predict the error over the range */
+
+
+
+/* allow additional parameters for c scan */
+#define CSCAN
+#include "invt.h"
+
+
+
+static void invt_predict(invtpar_t *m)
+{
+  double c, t, t0;
+  double err, err0, err1;
+
+  t = (double) m->nsteps;
+  for ( c = m->cmin; c < m->cmax + 0.001 * m->cdel; c += m->cdel ) {
+    t0 = c / m->alpha0;
+
+    /* initial equilibrium value */
+    err0 = esterror0_ez(m->alpha0,
+        m->n, m->winn, m->win, m->sampmethod,
+        "initial", 0);
+
+    /* c / (t + t0) */
+    err = esterror_ez(c, t, t0,
+        m->n, m->winn, m->win, m->sampmethod,
+        0);
+
+    /* final equilibrium value */
+    err1 = esterror0_ez(c / (t0 + t),
+        m->n, m->winn, m->win, m->sampmethod,
+        "final", 0);
+
+    printf("%g\t%g\t%g\t%g\n",
+        c, err0, err, err1);
+  }
+}
+
+
+
+int main(int argc, char **argv)
+{
+  invtpar_t m[1];
+
+  invtpar_init(m);
+  invtpar_doargs(m, argc, argv);
+  invtpar_dump(m);
+  invt_predict(m);
+  invtpar_finish(m);
+
+  return 0;
+}

### 2016-01-22 doc/invt.tex
add predict.c
add errsbin.gp

@@
 \newcommand{\summary}[1]{{\color{DarkPurple}\footnotesize \textsc{Summary.} #1}}
+
+\newcommand{\Err}{E}
 
@@
 \begin{equation}
-\mathcal E
+\Err
 =
@@
 \begin{align}
-  \mathcal E
+  \Err
   &\equiv
@@
 \begin{align}
-  \mathcal E
+  \Err
   %&= \sum_{i = 2}^n \langle y_i^2 \rangle
@@
   &\equiv
-  e^{\lambda_i \, [q(t') - z]},
+  e^{\lambda_i \, [z - q(t)]},
   \\
@@
 %
-\note{The extremal condition of $\mathcal E$ is given by
+\note{The extremal condition of $\Err$ is given by
 %
@@
 \begin{align}
-  \mathcal E
+  \Err
   &=
@@
 The next order correction to Eq. \eqref{eq:kappa_delta}
-would lead to a correction to $\mathcal E$
+would lead to a correction to $\Err$
 that is about $\alpha_{\max}^2$ times as large.
@@
   $$
-  which leads to a correction to $\mathcal E$
+  which leads to a correction to $\Err$
   $$
@@
   \propto
-  \alpha^2 \, \mathcal E.
+  \alpha^2 \, \Err.
   $$
@@
 Thus, the inequality sets a lower bound
-of the error $\mathcal E$
+of the error $\Err$
 expressed in Eq. \eqref{eq:error_mbin1},
@@
 \begin{align}
-\mathcal E
+\Err
 &=
@@
 \begin{aligned}
-  \mathcal E
+  \Err
   \stackrel{         t \to \infty         }
@@
 \begin{aligned}
-\mathcal E
+\Err
 &=
@@
 \begin{aligned}
-  \mathcal E
+\Err
 &=

### 2016-01-25 doc/invt.tex
adding a few references

@@
 The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
-and metadynamics\cite{laio2002}
+and metadynamics\cite{laio2002, laio2008, barducci2011, sutto2012}
 are two widely-used techniques for this purpose,
@@
 containing the current $z$,
-while metadynamics adopts a more widespread window
+while metadynamics adopts an extended window
 covering several neighboring bins.
@@
 Naturally, the manner of reducing
-the updating magnitude\cite{
+the updating magnitude\cite{liang2007,
 belardinelli2007, belardinelli2007jcp, belardinelli2008,
 morozov2007, zhou2008, morozov2009,
-komura2012, caparica2012, caparica2014}
+komura2012, caparica2012, caparica2014,
+barducci2008, dickson2011, dama2014, dickson2015}
 determines the precision of the final bias potential,
@@
 should be given by the
-inverse time\cite{
+inverse time\cite{liang2007,
 belardinelli2007, belardinelli2007jcp, belardinelli2008,
@@
 Generally, we may wish to sample
-a nearly flat distribution\cite{dayal2004, trebst2004}, $p_i$,
+a nearly flat distribution\cite{dayal2004, trebst2004, singh2011},
+$p_i$,
 by adjusting the bias potential $V_i$
@@
 in Sec. \ref{sec:updating-matrix}.
+%
+The multiple-bin scheme is more commonly
+used in metadynamics.
 
@@
 at all times.
+%
+In other words,
+the error in $v_i(t)$
+depends on the updating magnitude\cite{
+  zhou2005, liang2005, bussi2006, liang2007,
+  morozov2007, zhou2008, morozov2009, crespo2010,
+  atchade2011, fort2015}.
 %%
@@
 saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
+%
+Similarly, for metadynamics,
+there are attempts to reduce the updating magnitude\cite{
+  barducci2008, dickson2011, dama2014, dickson2015}.
 
@@
 the matrix $\mathbf w$ is tridiagonal,
-%
-\begin{equation}
-\arraycolsep=3.6pt\def\arraystretch{1.4}
-\mathbf w
-=
-\left(
-  \begin{array}{cccccccc}
-    1 - \mu_1   & \mu_1 & 0 & \dots & 0 \\
-    \mu_1 & 1 - 2 \, \mu_1  & \mu_1 & \dots & 0 \\
-    \vdots & &  & & \vdots \\
-    0 & \dots & \mu_1 & 1 - 2 \, \mu_1  & \mu_1 \\
-    0 & \dots & 0 & \mu_1 & 1 - \mu_1
-  \end{array}
-\right).
-\label{eq:wtridiag}
-\end{equation}
-%
-We have
+%%
+%\begin{equation}
+%\arraycolsep=3.6pt\def\arraystretch{1.4}
+%\mathbf w
+%=
+%\left(
+%  \begin{array}{cccccccc}
+%    1 - \mu_1   & \mu_1 & 0 & \dots & 0 \\
+%    \mu_1 & 1 - 2 \, \mu_1  & \mu_1 & \dots & 0 \\
+%    \vdots & &  & & \vdots \\
+%    0 & \dots & \mu_1 & 1 - 2 \, \mu_1  & \mu_1 \\
+%    0 & \dots & 0 & \mu_1 & 1 - \mu_1
+%  \end{array}
+%\right).
+%\label{eq:wtridiag}
+%\end{equation}
+%%
+%We have
+and the eigenvalues are
 \begin{equation}
@@
 %\eqref{eq:optimal_lambda_approx}
-to an integral.
-%
-Particularly,
-for perfect sampling,
-we get from Eq. \eqref{eq:Gamma_perfect},
+to an integral,
+$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
 $$
 \begin{aligned}
-\Err
+\Err_2
+&=
+\frac{2 \, n}{\pi \, t}
+\int_0^{\pi/2}
+\frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
+{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
+\, dp.
+\end{aligned}
+$$
+
+For perfect sampling,
+we have Eq. \eqref{eq:Gamma_perfect}, and
+$$
+\begin{aligned}
+\Err_2
 &=
@@
 \begin{aligned}
-\Err
+\Err_2
 &=
@@
 \lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
-\label{eq:lambda_tridiag}
+\label{eq:lambda_tridiag_perfect}
 \end{equation}
@@
 \right).
-\label{eq:error_tridiag}
+\label{eq:error_tridiag_prefect}
 \end{equation}
@@
 corresponds to the single-bin scheme.
+
+For one-step sampling,
+we have Eq. \eqref{eq:Gamma_onestep}, and
+$$
+\begin{aligned}
+\Err_2
+&=
+\frac{   2 \, n }
+     { \pi \, t }
+\int_0^{\pi/2}
+    \frac{          \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2           }
+         { \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
+    \frac{ \cos^2 p }
+         { \sin^2 p + \delta^2 }
+\, dp
+\notag \\
+&
+\stackrel{    \delta \ll 1     }
+         { =\joinrel=\joinrel= }
+\frac{  2 \, n  }
+     { \pi \, t }
+\frac{          (1 - 4 \, \mu_1)^2            }
+     {  \lambda \, (2 - 8 \, \mu_1 - \lambda) }
+\frac{    1   }
+     { \delta }
+,
+\end{aligned}
+$$
+%
+where $\delta \propto 1/n$ is a small parameter to keep the integral finite.
+%
+The optimal $\lambda$ is therefore
+%
+\begin{equation}
+\lambda = 1 - 4 \, \mu_1,
+\label{eq:lambda_tridiag_onestep}
+\end{equation}
 
@@
 %
-\begin{equation}
+\begin{align}
 \lambda_k
-=
-\exp\left(
+&=
+\exp\left[
       -\frac{ (k - 1)^2 \, \sigma^2 }
             {           2           }
-    \right).
-\label{eq:lambda_Gaussian}
-\end{equation}
-%
-The eigenvalues are all positive,
-suggesting stability of the updating scheme.
-
-However,
-as $\sigma/\Delta x \gg 1$,
-the smallest eigenvalue
-%
-$$
-\lambda_n
-\approx
-\exp\left(
-      -\frac{ n^2 \, \sigma^2 }
-            {        2        }
-    \right)
-=
+    \right]
+\notag
+\\
+&=
 \exp\left[
       -\frac{ \pi^2 }{ 2 }
+      \left(
+        \frac{ k - 1 }
+             {   n   }
+      \right)^2
       \left(
@@
 .
-$$
-%
-is exponentially small.
-%
-This means that
-the optimal value of $\lambda$
-given by Eq. \eqref{eq:optimal_lambda_approx}
-would be too small in practice,
-for it requires $\lambda < 2 \, \lambda_n$.
-%
-In other words,
-with a reasonable $\lambda$,
-the error of the last few fluctuation modes
-will always decay suboptimally,
-accordingly to Eq. \eqref{eq:y2_sum_perfect_invt}.
-%
-However, since these modes
-represent short wavelength fluctuations,
-and these modes may be in
-the system under consideration.
-%
-Thus,
-we can truncate the sum of the error function
-in Eq. \eqref{eq:y2_sum_perfect_invt}
-at some $k_{\max}$,
-which corresponds to a minimal length scale
-$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
-greater than $\sigma$.
-%
-Then,
-$$
-\lambda_{ k_{\max} }
-\approx
-\exp\left(
-      -\frac{ k_{\max}^2 \, \sigma^2 }
-            {           2            }
-    \right)
-=
-\exp\left(
-      -\frac{  \pi^2 \, \sigma^2    }
-            {    2 \, l_{\min}^2  }
-    \right).
-$$
+\label{eq:lambda_Gaussian}
+\end{align}
+%
+The eigenvalues are all positive,
+suggesting stability of the updating scheme.
+
+%However,
+%as $\sigma/\Delta x \gg 1$,
+%the smallest eigenvalue
+%%
+%$$
+%\lambda_n
+%\approx
+%\exp\left(
+%      -\frac{ n^2 \, \sigma^2 }
+%            {        2        }
+%    \right)
+%=
+%\exp\left[
+%      -\frac{ \pi^2 }{ 2 }
+%      \left(
+%        \frac{  \sigma }
+%             { \Delta x }
+%      \right)^2
+%    \right]
+%.
+%$$
+%%
+%is exponentially small.
+%%
+%This means that
+%the optimal value of $\lambda$
+%given by Eq. \eqref{eq:optimal_lambda_approx}
+%would be too small in practice,
+%for it requires $\lambda < 2 \, \lambda_n$.
+%%
+%In other words,
+%with a reasonable $\lambda$,
+%the error of the last few fluctuation modes
+%will always decay suboptimally,
+%accordingly to Eq. \eqref{eq:y2_sum_perfect_invt}.
+%%
+%However, since these modes
+%represent short wavelength fluctuations,
+%and these modes may be in
+%the system under consideration.
+%%
+%Thus,
+%we can truncate the sum of the error function
+%in Eq. \eqref{eq:y2_sum_perfect_invt}
+%at some $k_{\max}$,
+%which corresponds to a minimal length scale
+%$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
+%greater than $\sigma$.
+%%
+%Then,
+%$$
+%\lambda_{ k_{\max} }
+%\approx
+%\exp\left(
+%      -\frac{ k_{\max}^2 \, \sigma^2 }
+%            {           2            }
+%    \right)
+%=
+%\exp\left(
+%      -\frac{  \pi^2 \, \sigma^2    }
+%            {    2 \, l_{\min}^2  }
+%    \right).
+%$$
 
@@
 
-\subsection{Optimal $\alpha(t)$}
+\subsection{General simulation protocol}
 
@@
 
-1. Single-bin scheme.
+\subsection{Single-bin (WL) updating scheme}
+
 Show the error versus $\lambda$.
 
-Change upon using local sampling scheme.
-
-2. Multiple-bin scheme.
-Error vs. $\lambda$.
-
-3. Optimality of the Single-bin scheme.
+Good agreement between theory and simulation.
+
+The optimal $\lambda$ is near $1.0$,
+independent of the underlying sampling scheme.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsbin.pdf}
+  }
+  \caption{
+    \label{fig:err_singlebin}
+    Error, $E$, versus the proportionality constant, $1/\lambda$,
+    in Eq. \eqref{eq:alpha_trial}
+    for the single-bin (WL) scheme.
+    %
+  }
+\end{center}
+\end{figure}
+
+\subsection{Nearest neighbor updating scheme}
+
+Normalized error vs. $\lambda$.
+
+The optimal $\lambda$ depends on the simulation length.
+This dependence is more significant in the global sampling scheme.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/errnnbr.pdf}
+  }
+  \caption{
+    \label{fig:err_nnbr}
+    Normalized error, $(t + t_0) \, E$,
+    versus the proportionality constant, $1/\lambda$,
+    in Eq. \eqref{eq:alpha_trial}
+    for the nearest-neighbor updating scheme with $\mu_1 = 0.24$.
+    %
+  }
+\end{center}
+\end{figure}
+
+\subsection{Gaussian (metadynamics) updating scheme}
+
+\subsection{Optimality of the Single-bin scheme}
+
+Randomize the initial error?
+
 Error vs. $\mu_1$.
+
+Error vs. $\sigma$.
 

### 2016-01-26 doc/invt.tex
invt.tex: add error of the nearest neighbor scheme for the local sampling

@@
 %
-For the quantity of interest, $z$, is continuous,
+For a continuous quantity, $z$,
 which is typical in molecular systems,
@@
 depends on the updating magnitude\cite{
-  zhou2005, liang2005, bussi2006, liang2007,
+  zhou2005, liang2005, laio2005, bussi2006, poulain2006, liang2007,
   morozov2007, zhou2008, morozov2009, crespo2010,
@@
 Similarly, for metadynamics,
-there are attempts to reduce the updating magnitude\cite{
-  barducci2008, dickson2011, dama2014, dickson2015}.
+there are various ways of reducing the updating magnitude\cite{
+  marsili2006, barducci2008, dickson2011, dama2014, dickson2015}.
 
@@
 to a neighbor that is $|i - j|$ bins away,
-whereas the next two terms are added to avoid
-boundary artifacts.
+whereas the next two terms are added to
+adopt the open boundary condition.
+%
+If $i$ represents a periodic variable,
+the last two terms can be dropped under
+the minimum image convention for $|i-j|$\cite{dama2014}.
 
@@
 
-\subsubsection{Tridiagonal matrix}
+\subsubsection{Nearest-neighbor updating scheme}
 
@@
 \int_0^{\pi/2}
-\frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
-{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
+  \frac{    \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
+       { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
 \, dp.
@@
 For perfect sampling,
-we have Eq. \eqref{eq:Gamma_perfect}, and
+we have, from Eq. \eqref{eq:Gamma_perfect},
+$\Gamma(p) = 1$, and
 $$
@@
 \int_0^{\pi/2}
-\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
-{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
+\frac{ \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2 }
+{ \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
 \, dp
@@
 %
+It is somewhat more convenient to change variable $p \to \frac{ \pi } { 2 } - p$,
+and
 $$
@@
 For one-step sampling,
-we have Eq. \eqref{eq:Gamma_onestep}, and
+we have, from Eq. \eqref{eq:Gamma_onestep},
+$\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2)$,
+where we have introduced
+$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
+as a small parameter to avoid the divergence
+around the origin.
+%
+Then
 $$
@@
 \int_0^{\pi/2}
-    \frac{          \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2           }
-         { \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
+    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
+         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
     \frac{ \cos^2 p }
@@
      { \pi \, t }
-\frac{          (1 - 4 \, \mu_1)^2            }
-     {  \lambda \, (2 - 8 \, \mu_1 - \lambda) }
+\frac{             1             }
+     {  \lambda \, (2 - \lambda) }
 \frac{    1   }
      { \delta }
+.
+\end{aligned}
+$$
+%
+The optimal $\lambda$ is therefore
+%
+\begin{equation}
+\lambda = 1,
+\label{eq:lambda_tridiag_onestep}
+\end{equation}
+%
+which is the same of the global sampling scheme.
+
+\note{The integral is evaluated as follows.
+%
+We again change variable $p \to \frac{ \pi } { 2 } - p$,
+and
+$$
+\begin{aligned}
+\Err_2
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\int_0^{2 \, \pi}
+\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
+     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
+\cdot
+\frac{ \sin^2 p            }
+     { \cos^2 p + \delta^2 }
+\, dp
+\\
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\oint
+\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
+     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
+\cdot
+\frac{ 4 - \left( z + \frac 1 z \right)^2 }
+     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
+\, \frac{dz}{z}
+\\
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\oint
+\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
+     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
+\cdot
+\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
+     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
+\, \frac{ dz }{ z^3 }
+\\
+&=
+\frac{ n } { \lambda \, t }
+\sum_i \operatorname{Res}_i
 ,
@@
 $$
-%
-where $\delta \propto 1/n$ is a small parameter to keep the integral finite.
-%
-The optimal $\lambda$ is therefore
-%
-\begin{equation}
-\lambda = 1 - 4 \, \mu_1,
-\label{eq:lambda_tridiag_onestep}
-\end{equation}
-
-\subsubsection{Gaussian matrix}
+where the contour is the unit circle
+in the complex plane of $z$,
+and the integral is reduced to a sum of residues
+around the poles within the unit circle.
+
+The poles of the integral come from three factors,
+namely,
+1: at $z = 0$,
+2: at the zeros of
+$(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2$,
+3: at the zeros of
+$( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2$.
+
+For the first pole around the origin,
+we need to expand the integrand as
+$$
+\frac{ A + B \, z^2 + \cdots }
+     {          z^3          },
+$$
+then the residue is simply $B$.
+%
+Since
+$$
+\begin{aligned}
+\left[ z^2 - \mu_1 (z^2 + 1)^2 \right]^2
+&=
+[ - \mu_1 + (1 - 2 \, \mu_1) \, z^2]^2
+\\
+&=
+\mu_1^2
+\left[
+  1 + \left(4 - \tfrac { 2  } { \mu_1 } \right) z^2
+\right],
+\\
+%
+(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2
+&=
+-2 \, \mu_1 \,
+\left[
+  1 + \left(
+        2 + \tfrac{ \lambda - 2 } { 2 \, \mu_1 }
+      \right)
+\right],
+\\
+%
+4 \, z - ( z^2 + 1 )^2
+&=
+-( 1 - z^2 ),
+\\
+%
+( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2
+&=
+1 + (2 + 4 \, \delta^2) \, z^2,
+\end{aligned}
+$$
+the integrand is
+$$
+\begin{aligned}
+\frac{ \mu_1 }
+     {   2   }
+\left[
+  1 + \left(
+        - \frac{ 2 + \lambda } { 2 \, \mu_1 }
+        - 2 - 4 \, \delta^2
+      \right) \, z^2
+\right].
+\end{aligned}
+$$
+and the residue is
+$$
+\operatorname{Res}_1
+=
+  - \frac{ 2 + \lambda } { 4 }
+  - \mu_1 \, \left( 1 +  \, \delta^2 \right).
+$$
+
+For the second part,
+we have four roots of
+$$
+(2 - \lambda) \, z^2 = 2 \, \mu_1 \, (z^2 + 1)^2,
+$$
+which are
+$$
+z = \pm \sqrt { \frac{ 2 - \lambda }
+                     { 8 \, \mu_1 } }
+\pm \sqrt { \frac{ 2 - \lambda }
+                 { 8 \, \mu_1 } - 1 }.
+$$
+Two of them lie in the unit circle, namely
+$$
+z = \pm \left(
+  \sqrt { \frac{ 2 - \lambda }
+               { 8 \, \mu_1 } }
+- \sqrt { \frac{ 2 - \lambda }
+               { 8 \, \mu_1 } - 1 }
+        \right),
+$$
+and
+$$
+4 \, \mu_1 \, ( z^2 + 1)
+=
+2 - \lambda
+-
+\sqrt{ ( 2 - \lambda ) ( 2 - \lambda - 8 \, \mu_1 ) }.
+$$
+Thus,
+$$
+\begin{aligned}
+\left[ z^2 - \mu_1 \, (1 + z^2)^2 \right]^2
+&=
+\left( z^2 - \tfrac{ 2 - \lambda } { 2 } z^2 \right)^2
+=
+\frac{ \lambda^2 } { 4 } z^4.
+\\
+%
+4 z^2 - (1 + z^2)^2
+&=
+-\frac{ 2 - \lambda - 8 \, \mu_1  } { 2 \, \mu_1 } z^2,
+\\
+%
+(z^2 + 1)^2 + 4 \, \delta^2 \, z^2
+&=
+\frac{ 2 - \lambda + 8 \, \mu_1 \, \delta^2 }
+     { 2 \, \mu_1 }
+     z^2,
+\\
+%
+(2 - \lambda) \, 2 \, z
+-
+8 \, \mu_1 \, (z^2 + 1) \, z
+&=
+2 \, z \, \left[2 - \lambda - 4 \,  \mu_1 \, (z^2 + 1) \right]
+\\
+&= 2 \, z \, \sqrt{ (2 - \lambda) ( 2 - \lambda - 8 \, \mu_1 ) },
+\end{aligned}
+$$
+where we have used the L'H\^{o}pital's rule
+and taken the derivative for the last factor.
+%
+Note that the two roots within the unit share the same
+residue and the sum for this part is
+$$
+\operatorname{Res}_2
+=
+-\frac{                 \lambda^2                     }
+      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
+\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }.
+$$
+
+For the last part,
+we have four roots of
+$(z^2 + 1)^2 = - 4 \, \delta^2 \, z^2$,
+with two of them lying in the unit circle:
+$$
+z = \pm i \left( \sqrt{ 1 + \delta^2 } - \delta \right),
+$$
+such that
+$$
+z^2 + 1 + 2 \, \delta^2 = 2 \, \delta \, \sqrt{1 + \delta^2}.
+$$
+Then,
+$$
+\begin{aligned}
+z^2 - \mu_1 (1 + z^2)^2
+&= z^2 \, (1 + 4 \, \mu_1 \, \delta^2),
+\\
+4 \, z^2 - (1 + z^2)^2
+&=
+4 \, z^2 \, (1 + \delta^2)
+\\
+%
+(2 - \lambda) \, z^2
+- 2 \, \mu_1 (1 + z^2)^2
+&=
+\left(
+  2 - \lambda + 8 \, \mu_1 \, \delta^2
+\right) \, z^2
+\\
+%
+4 \, z \, (1 + z^2)
++
+8 \, \delta^2 \, z
+&=
+4 \, z \, (z^2 + 1 + \delta^2)
+=
+4 \, z \, 2 \, \delta \, \sqrt{ 1 + \delta^2 }.
+\end{aligned}
+$$
+Note that the two roots within the unit share the same
+residue and the sum for this part is
+$$
+\operatorname{Res}_3
+=
+\frac{ (1 + 4 \, \mu_1 \, \delta^2 )^2 }
+     { (2 + 8 \, \mu_1 \, \delta^2 - \lambda }
+\frac{ \sqrt{ 1 + \delta^2 } }
+     {        \delta         }.
+$$
+
+Now as $\delta \to 0$, the last contribution will dominate,
+and we have
+$$
+\Err_2
+\approx
+\frac{ n } { t }
+\frac{ (1 + 4 \, \mu_1 \, \delta^2)^2 }
+{ \lambda \, (2 + 8 \, \mu_1 \, \delta^2 - \lambda) }
+\approx
+\frac{ n } { t }
+\frac{ 1 }
+{ \lambda \, (2 - \lambda) },
+$$
+and the minimum is reached at $\lambda \approx 1$.
+}
+
+\subsubsection{Gaussian updating scheme}
 
@@
 
-The optimal $\lambda$ depends on the simulation length.
+In the global sampling scheme,
+the optimal $\lambda$ depends on the simulation length.
 This dependence is more significant in the global sampling scheme.

### 2016-01-28 doc/invt.tex
editing invt.tex

@@
   =
-  h_i(t) \frac{ \alpha(t) } { p_i },
+  \alpha(t) \, \frac{ h_i(t) } { p_i },
   \label{eq:vt_diffeq}
@@
 
-\subsubsection{One-variable problem}
+\subsubsection{Optimization of a one-variable problem}
 
@@
 \begin{align}
-\left\langle x^2(t) \right\rangle
-&=
+&\left\langle x^2(t) \right\rangle
+=
+\langle x^2(0) \rangle e^{-2 \, \lambda \, q(t)}
+\notag
+\\
+&\qquad
++
 \int_0^t \int_0^t
@@
   \dot u\bigl( q(t'') \bigr) \,
-  \kappa(t' - t'') \, dt'' \, dt'
-\notag
-\\
-&\hphantom{=}
-+\langle x^2(0) \rangle e^{-2 \, \lambda \, q(t)}.
+  \kappa(t' - t'') \, dt'' \, dt'.
 \label{eq:x2t_average}
@@
 \begin{equation}
-  \dot v_i(t) =
-  \sum_{j=1}^n \alpha(t) \, h_j(t) \frac{ w_{ij} } { p_j }.
+  \dot v_i(t)
+  =
+  \alpha(t) \sum_{j=1}^n \, w_{ij} \frac{ h_j(t) } { p_j }.
   \label{eq:vt_diffeq_mbin}
@@
 
-\subsubsection{Linearization and decoupling}
+\subsubsection{Linearization and decoupling to eigenmodes}
 
@@
   \dot{ \bar v }
-  &= \sum_{i = 1}^n p_i \, \dot v_i
+  &=
+  \sum_{i = 1}^n p_i \, \dot v_i
+  %
   \stackrel{\mathrm{Eq.\; \eqref{eq:vt_diffeq_mbin}}}
   {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \sum_{j = 1}^n \alpha(t) \, \frac{ h_j(t) } { p_j } \sum_{ i = 1 }^{n} p_i \, w_{ij}
+  %
+  \sum_{j = 1}^n \alpha(t) \, \frac{ h_j(t) } { p_j }
+                 \sum_{ i = 1 }^{n} p_i \, w_{ij}
   \\
@@
   {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \sum_{j = 1}^n \alpha(t) \, h_j(t) = \alpha(t).
+  \sum_{j = 1}^n \alpha(t) \, h_j(t)
+  =
+  \alpha(t).
 \end{aligned}
@@
 \begin{equation}
-\dot y_i(t)
-=
--\alpha(t) \, \lambda_i
-[y_i(t) - \eta_i(t)],
+\dot y_k(t)
+=
+-\alpha(t) \, \lambda_k \, [y_k(t) - \eta_k(t)],
 \label{eq:yt_diffeq}
@@
 \begin{align}
-  y_i &= \sum_{j=1}^n \phi_{ij} \, x_j,
+  y_k &= \sum_{j=1}^n \phi_{kj} \, x_j,
   \label{eq:y_def}
   \\
-  \eta_i &= \sum_{j=1}^n \phi_{ij} \frac{ \zeta_j}{ p_j}.
+  \eta_k &= \sum_{j=1}^n \phi_{kj} \frac{ \zeta_j}{ p_j}.
   \label{eq:eta_def}
@@
 
-\subsubsection{Minimization of the error function}
+\subsubsection{Error function}
 
@@
   \Err
-  &\equiv
-  \sum_{i = 1}^n p_i \, x_i^2
-  =
-  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \, x_i \, x_j
-  =
-  \sum_{k = 1}^n y_k^2.
+  \equiv
+  \sum_{i = 1}^n p_i \,
+                 \langle x_i^2 \rangle
+  =
+  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \,
+                     \langle x_i \, x_j \rangle
+  =
+  \sum_{k = 1}^n \langle y_k^2 \rangle,
   \label{eq:y2_sum}
 \end{align}
+%
+which is a superposition of the components of the modes.
 %
@@
 In analogous to Eq. \eqref{eq:x2t_average},
-we can rewrite Eq. \eqref{eq:y2_sum} as
-%
-\begin{align}
+we have, from Eq. \eqref{eq:y2_sum},
+%
+\begin{equation}
   \Err
-  %&= \sum_{i = 2}^n \langle y_i^2 \rangle
-  %\notag
-  %\\
-  &=
+  =
+  \Err_R + \Err_A,
+  \label{eq:error_split}
+\end{equation}
+%
+which is the sum of the residual error
+%
+\begin{equation}
+  \Err_R
+  =
+  \sum_{k = 2}^n
+    \left\langle y_k^2(0) \right\rangle \,
+    e^{ - 2 \, \lambda_k  \, q(t) },
+  \label{eq:error_res}
+\end{equation}
+%
+from the decay of the initial errors,
+and the asymptotic error
+\begin{equation}
+  \Err_A
+  =
   \int_0^t \int_0^t
-  \sum_{i = 2}^n
-  \dot u_i\bigl( q(t') \bigr) \,
-  \dot u_i\bigl( q(t'') \bigr) \,
-  \kappa_i(t' - t'') \, dt' \, dt'',
-  \label{eq:error_mbin}
-\end{align}
-%
-where
+  \sum_{k = 2}^n
+    \dot u_k\bigl( q(t') \bigr) \,
+    \dot u_k\bigl( q(t'') \bigr) \,
+    \kappa_k(t' - t'') \, dt' \, dt'',
+  \label{eq:error_asym}
+\end{equation}
+%
+where,
 \begin{align*}
-  u_i(z)
+  u_k(z)
   &\equiv
-  e^{\lambda_i \, [z - q(t)]},
-  \\
-  \kappa_i(t - t')
+  e^{\lambda_k \, [z - q(t)]},
+  \\
+  \kappa_k(t - t')
   &\equiv
   \left\langle
-    \eta_i(t) \, \eta_i(t')
+    \eta_k(t) \, \eta_k(t')
   \right\rangle.
@@
 %
-\note{The extremal condition of $\Err$ is given by
+\note{The extremal condition of $\Err_A$ is given by
 %
 \begin{align}
-\sum_{i=2}^n
-\lambda_i \, u_i\bigl( q(\tau) \bigr)
+\sum_{k = 2}^n
+\lambda_k \, u_k\bigl( q(\tau) \bigr)
 \int_0^t
-\ddot u_i\bigl( q(t') \bigr) \, \kappa_i(t' - \tau) \, dt' = 0.
+\ddot u_k \bigl( q(t') \bigr) \, \kappa_k(t' - \tau) \, dt' = 0.
 \label{eq:optimal_mbin}
@@
 %
-If all $\lambda_i$, hence all $u_i$, are the same,
+If all $\lambda_k$, hence all $u_k$, are the same,
 the above condition can be satisfied
@@
 In the asymptotic regime,
-we can approximate $\kappa_i(t)$ as
-%
-\begin{equation}
-  \kappa_i(t) \approx \Gamma_i \, \delta(t)
+we can approximate $\kappa_k(t)$ as
+%
+\begin{equation}
+  \kappa_k(t) \approx \Gamma_k \, \delta(t)
   \label{eq:kappa_delta}
@@
 \begin{equation}
-  \Gamma_i = \int_{-\infty}^\infty \kappa_i(t) \, dt,
+  \Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt,
   \notag
@@
 \begin{equation}
-  \Gamma_i = \sum_{t = -\infty}^\infty \kappa_i(t).
+  \Gamma_k = \sum_{t = -\infty}^\infty \kappa_k(t).
   \label{eq:Gamma_sum}
@@
 %
-Then, Eq. \eqref{eq:error_mbin} is simplified as
+Then, Eq. \eqref{eq:error_asym} is simplified as
 %
 \begin{align}
-  \Err
+  \Err_A
   &=
   \int_0^t
-  \sum_{i = 2}^n
-  \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr) \, dt'.
-  \label{eq:error_mbin1}
+  \sum_{k = 2}^n
+  \Gamma_k \, \dot u_k^2\bigl( q(t') \bigr) \, dt'.
+  \label{eq:error_asym1}
 \end{align}
@@
   $$
-  \tilde \kappa_i(\omega) = \Gamma_i
+  \tilde \kappa_k(\omega) = \Gamma_k
   $$
@@
   $$
-  \tilde \kappa_i(\omega) = \Gamma_i + \Gamma^{(2)}_i \omega^2 + \dots,
+  \tilde \kappa_k(\omega) = \Gamma_k + \Gamma^{(2)}_k \omega^2 + \dots,
   $$
@@
   $$
-  \Gamma^{(2)}_i
-  \int_0^t \ddot u_i^2\bigl( q(t') \bigr) \, dt'
+  \Gamma^{(2)}_k
+  \int_0^t \ddot u_k^2\bigl( q(t') \bigr) \, dt'
   \propto
@@
   This correction is of order $\alpha^2$
-  compared to the main contribution in Eq. \eqref{eq:error_mbin1}.
+  compared to the main contribution in Eq. \eqref{eq:error_asym1}.
   %
@@
   \begin{align}
-    \sum_{i=1}^n
-      \Gamma_i \, \lambda_i \,
-      u_i\bigl( q(\tau) \bigr) \,
-      \ddot u_i\bigl( q(\tau) \bigr) = 0.
+    \sum_{k = 1}^n
+      \Gamma_k \, \lambda_k \,
+      u_k\bigl( q(\tau) \bigr) \,
+      \ddot u_k\bigl( q(\tau) \bigr) = 0.
     \label{eq:optimal_mbin1}
@@
   %
-  If we interpret $u_i$ as the position,
-  and $\Gamma_i \, \lambda_i$ as the mass,
+  If we interpret $u_k$ as the position,
+  and $\Gamma_k \, \lambda_k$ as the mass,
   then the right-hand side gives the virial.
@@
 
-We can evaluate the $\Gamma_i$'s
+We can evaluate the $\Gamma_k$'s
 in two special case.
@@
 \begin{equation}
-  \Gamma_i = 1.
+  \Gamma_k = 1.
   \label{eq:Gamma_perfect}
@@
 In this case,
-the bin index $i$ is an independent random variable
+the bin index $k$ is an independent random variable
 at each time step $t$, and
@@
 \begin{equation}
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  \left\langle
+    \zeta_i(t) \, \zeta_j(t')
+  \right\rangle
   =
@@
   \begin{aligned}
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  &=
-  \langle \zeta_i(t)  \rangle
-  \langle \zeta_j(t') \rangle
+  \left\langle
+    \zeta_i(t) \, \zeta_j(t')
+  \right\rangle
+  &=
+  \left\langle
+    \zeta_i(t)
+  \right\rangle
+  \left\langle
+    \zeta_j(t')
+  \right\rangle
   \\
@@
 Thus, from Eqs. \eqref{eq:eig_orthonormal_rows} and \eqref{eq:eta_def},
-we get, for $i \ge 2$,
-%
-\begin{equation}
-  \kappa_i(t - t')
-  =
-  \sum_{j,k = 1}^n
-  \frac{ \phi_{ij} } { p_j }
-  \frac{ \phi_{ik} } { p_k }
-  \langle \zeta_j(t) \, \zeta_k(t') \rangle
+we get, for $k \ge 2$,
+%
+\begin{equation}
+  \kappa_k(t - t')
+  =
+  \sum_{j,l = 1}^n
+  \frac{ \phi_{kj} } { p_j }
+  \frac{ \phi_{kl} } { p_l }
+  \langle \zeta_j(t) \, \zeta_l(t') \rangle
   %
@@
   \begin{aligned}
-  \kappa_i(t - t')
-  &=
-  \sum_{j,k = 1}^n
-  \frac{ \phi_{ij} } { p_j }
-  \frac{ \phi_{ik} } { p_k }
-  \langle \zeta_j(t) \, \zeta_k(t') \rangle
-  \\
-  %
-  &=
-  \sum_{j,k = 1}^n
-  \frac{ \phi_{ij} } { p_j }
-  \frac{ \phi_{ik} } { p_k }
+  \kappa_k(t - t')
+  &=
+  \sum_{j,l = 1}^n
+  \frac{ \phi_{kj} } { p_j }
+  \frac{ \phi_{kl} } { p_l }
+  \langle \zeta_j(t) \, \zeta_l(t') \rangle
+  \\
+  %
+  &=
+  \sum_{j,l = 1}^n
+  \frac{ \phi_{kj} } { p_j }
+  \frac{ \phi_{kl} } { p_l }
   \left(
-    p_j \, \delta_{j, k} - p_j \, p_k
+    p_j \, \delta_{j, l} - p_j \, p_l
   \right)
@@
     \sum_{j = 1}^n
-    \frac{ \phi_{ij}^2 } { p_j }
+    \frac{ \phi_{kj}^2 } { p_j }
     -
-    \sum_{j = 1}^n \phi_{ij}
-    \sum_{k = 1}^n \phi_{ik}
+    \sum_{j = 1}^n \phi_{kj}
+    \sum_{l = 1}^n \phi_{kl}
   \right)
@@
   %
-  Note for $i \ge 2$,
-  $$
-  \sum_{j = 1}^n \phi_{ij}
-  =
-  \sum_{j = 1}^n \frac{ \phi_{ij} \, \phi_{1j} } { p_j }
-  =
-  \delta_{i1}
+  Note for $k \ge 2$,
+  $$
+  \sum_{j = 1}^n \phi_{kj}
+  =
+  \sum_{j = 1}^n \frac{ \phi_{kj} \, \phi_{1j} } { p_j }
+  =
+  \delta_{k1}
   =
@@
 $$
-\Lambda_i = 1 - 2 \sin^2 \frac{ (i - 1) \, \pi } { 2 \, n }.
-$$
-Note that these $\Lambda_i$'s are
+\Lambda_k = 1 - 2 \sin^2 \frac{ (k - 1) \, \pi } { 2 \, n }.
+$$
+Note that these $\Lambda_k$'s are
 not to be confused with the eigenvalues of $\mathbf w$,
-$\lambda_i$'s.
-%
-The correlation function $\kappa_i(t)$
-decays as $\kappa_i(0) \, \Lambda_i^t$,
-with the same-time value $\kappa_i(0) = 1$ being the same as
+$\lambda_k$'s.
+%
+The correlation function $\kappa_k(t)$
+decays as $\kappa_k(0) \, \Lambda_k^t$,
+with the same-time value $\kappa_k(0) = 1$ being the same as
 that in perfect sampling.
 %
-Thus, from Eq. \eqref{eq:Gamma_sum}, we get
-%
-\begin{equation}
-\Gamma_i
-=
-1 + 2 \, \sum_{t = 1}^\infty \Lambda_i^t
-=
-\frac{ 1 + \Lambda_i } { 1 - \Lambda_i }
-=
-\cot^2 \left[ \frac{ (i - 1) \, \pi } { 2 \, n } \right].
+Thus,
+$\Gamma_k$ is roughly twice
+the autocorrelation time,
+and from Eq. \eqref{eq:Gamma_sum}, we get
+%
+\begin{equation}
+\Gamma_k
+=
+1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
+=
+\frac{ 1 + \Lambda_k } { 1 - \Lambda_k }
+=
+\cot^2 \left[ \frac{ (k - 1) \, \pi } { 2 \, n } \right].
 \label{eq:Gamma_onestep}
@@
 We can use this case to model local sampling processes.
+
+\subsubsection{\label{eq:eqlerr}
+Equilibrium error under a fixed updating magnitude
+}
+
+Before addressing the optimal $\alpha(t)$,
+let us first compute the equilibrium value of $\Err$,
+under a fixed $\alpha(t) \equiv \alpha_0$.
+%
+In this case, we can drop $\Err_R$, and
+%
+\begin{align}
+  \Err
+  &=
+  \Err_A
+  =
+  \int_0^t
+    \sum_{k = 2}^n
+      \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
+      e^{ 2 \, \lambda_k \, \alpha_0 \, (t' - t) }
+    \, dt
+  \notag
+  \\
+  &
+  \stackrel{ t \to \infty }
+  { =\joinrel=\joinrel=\joinrel= }
+  %
+  \sum_{ k = 2 }^n
+    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0,
+  \label{eq:error_eql}
+\end{align}
+%
+or, in terms of the components,
+%
+\begin{equation}
+  \left\langle
+    y_k^2
+  \right\rangle
+  =
+  \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0,
+  \label{eq:y2_eql}
+\end{equation}
+%
+
+If we can assume that before entering the asymptotic regime,
+the system has entered the equilibrium under a fixed $\alpha_0$,
+we can then use Eq. \eqref{eq:y2_eql}
+for the $\langle y_k^2(0) \rangle$'s in Eq. \eqref{eq:error_res}.
+
+\note{If we can interpret $\Gamma_k$ as twice the autocorrelation time,
+  and the kernel is exponential
+  $$
+  \kappa_k = \exp\left( - \frac{2}{\Gamma_k} |t| \right),
+  $$
+  then a slightly more accurate value is given by\cite{vankampen}
+  $$
+  \left\langle
+    y_k^2
+  \right\rangle
+  =
+  \frac{      ( \lambda_k \, \alpha_0 )^2     }
+       { \lambda_k \, \alpha_0 + 2 / \Gamma_k }.
+  $$
+  But it appears that the main error of Eq. \eqref{eq:y2_eql}
+  is an underestimate instead of an overestimate.
+  And a possible source of error might be from
+  converting a discrete sum to a continuous integral.
+}
 
@@
 %the dominant contribution comes from
-%the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
+%the term(s) with the smallest eigenvalue $\lambda_k = \lambda_n$,
 %and
@@
 sharing the same set of eigenvectors,
-hence same $\Gamma_i$'s,
-but with different eigenvalues, $\lambda_i$'s.
+hence same $\Gamma_k$'s,
+but with different eigenvalues, $\lambda_k$'s.
 %
 Using the Cauchy-Schwarz inequality, we have,
-for any set of $\beta_i \ge 0$
+for any set of $\beta_k \ge 0$
 %
@@
   \int_0^t dt'
-    \sum_{i = 2}^n
-      \Gamma_i \, \dot u_i^2\bigl( q(t') \bigr)
+    \sum_{k = 2}^n
+      \Gamma_k \, \dot u_k^2\bigl( q(t') \bigr)
 \right)
@@
   \int_0^t dt'
-    \sum_{i = 2}^n
-      \Gamma_i \, c_i^2
+    \sum_{k = 2}^n
+      \Gamma_k \, c_k^2
 \right)
@@
 &
+\qquad \qquad
 \ge
@@
   \int_0^t dt'
-    \sum_{i = 2}^n
-      \Gamma_i \, c_i \, \dot u_i\bigl( q(t') \bigr)
+    \sum_{k = 2}^n
+      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t') \bigr)
 \right)^2
@@
 &
+\qquad \qquad
 =
 \left(
-  \sum_{i = 2}^n \Gamma_k \, c_i
+  \sum_{k = 2}^n \Gamma_k \, c_k
     \left[
-      1 - e^{ -\lambda_i q(t) }
+      1 - e^{ -\lambda_k q(t) }
     \right]
@@
 %
-\note{A short proof of the above inequality
-  is the following.
-  %
-  First, we have
+To see this, we first observe that
 $$
 \int_0^t
-  dt' \sum_{i = 2}^n \Gamma_i \,
-    \left( \dot u_i \, X - c_i \right)^2
+  dt' \sum_{k = 2}^n \Gamma_k \,
+    \left( \dot u_k \, Y - c_k \right)^2
 \ge 0,
 $$
-  for any $X$.
-  %
-  Thus, the left-hand side,
-  expanded as a quadratic polynomial of $X$,
-  $A \, X^2 + B \, X + C$,
+holds for any $Y$.
+%
+So the left-hand side,
+as a quadratic polynomial of $Y$,
+$A \, Y^2 + B \, Y + C$,
 must have a non-positive discriminant,
@@
 This condition gives the desired inequality.
-  \\%
-}
 %
@@
 of the error $\Err$
-expressed in Eq. \eqref{eq:error_mbin1},
+expressed in Eq. \eqref{eq:error_asym1}.
 %
 The equality is achieved
-if $\dot u_i\left( q(t') \right) = c_i$
-for all $i = 2, \dots, n$ at any $t'$,
+if $\dot u_k\left( q(t') \right) = c_k$
+for all $k = 2, \dots, n$ at any $t'$,
 up to a multiplicative factor.
@@
 \alpha(t') = \frac{              1              }
-                  { \lambda_i \, (t' + t_{i0} ) },
+                  { \lambda_k \, (t' + t_{k0} ) },
 \quad
 \mathrm{with\;\;}
-t_{i0} = \frac{             t            }
-              { e^{ \lambda_i q(t) } - 1 }.
+t_{k0} = \frac{             t            }
+              { e^{ \lambda_k q(t) } - 1 }.
 $$
@@
   We add some steps for the above solution.
-  Integrating $\dot u_i = c_i$ yields
-  \begin{equation}
-  u_i(t') = c_i \left(t' + t_{i0} \right).
+  Integrating $\dot u_k = c_k$ yields
+  \begin{equation}
+  u_k(t') = c_k \left(t' + t_{k0} \right).
   \label{eq:ui_solution}
   \end{equation}
-  Given that $u_i(t) = 1$ and $u_i(0) = e^{-\lambda_i \, q(t)}$,
+  Given that $u_k(t) = 1$ and $u_k(0) = e^{-\lambda_k \, q(t)}$,
   we have
   $$
-  c_i = \frac{ 1 }{ t + t_{i0} },
+  c_k = \frac{ 1 }{ t + t_{k0} },
   \quad
   \mathrm{and\;}
-  \frac{ t_{i0} } { t + t_{i0} }
+  \frac{ t_{k0} } { t + t_{k0} }
   =
@@
   $$
-  -\lambda_i \, q(t) + \lambda_i \, q(t')
-  = \log c_i + \log\left( t' + t_{i0} \right).
+  -\lambda_k \, q(t) + \lambda_k \, q(t')
+  = \log c_k + \log\left( t' + t_{k0} \right).
   $$
@@
   $$
-  \lambda_i \, \alpha(t') = \frac{ 1 } { t' + t_{i0} }.
+  \lambda_k \, \alpha(t') = \frac{ 1 } { t' + t_{k0} }.
   $$
@@
 (since $\lambda_1$ corresponds to the mode
-of a uniform shift of all $v_i$,
+of a uniform shift of all $v_k$,
 we can freely set it to $\lambda_2$
@@
 
-\subsubsection{Optimal coefficient of proportionality}
-
-If we wish to find,
-for a given updating matrix $\mathbf w$,
-a near-optimal $\alpha(t)$,
-%
-then a good candidate is
+\subsubsection{Variationally optimal schedule, $\alpha(t)$}
+
+From Eq. \eqref{eq:alpha_opt},
+it is reasonable to conjecture that
+a near-optimal $\alpha(t)$ for
+a given updating matrix $\mathbf w$ would be
 %
@@
 %
-for some optimal value of $\lambda$.
+for some value of $\lambda$.
+%
+Below we shall find the optimal value
+in some special cases.
 %
 Using Eq. \eqref{eq:alpha_trial}
-in Eq. \eqref{eq:error_mbin1} yields
+in Eq. \eqref{eq:error_asym1} yields
 %
 \begin{align}
-\Err
-&=
-\frac{1}{t+t_0}
-\sum_{i=2}^n
-  \frac{       \Gamma_i \, \lambda_i^2      }
-       { \lambda \, (2 \lambda_i - \lambda) }
+\Err_A
+&=
+\frac{    1    }
+     { t + t_0 }
+\sum_{k = 2}^n
+  \frac{ \Gamma_k \, \nu_k^2 }
+       {    2 \, \nu_k - 1   }
 \left[
@@
         1 + \frac{ t }{ t_0 }
-      \right)^{1 - 2 \, \lambda_i/\lambda}
-\right]
-\notag
-\\
-&\stackrel{      t \gg t_0      }
-          { =\joinrel=\joinrel= }
-%
-\sum_{i = 2}^n
-\begin{dcases}
-  \frac{ 1 }
-       { t }
-  \frac{         \Gamma_i \, \lambda_i^2       }
-       { \lambda \, (2 \, \lambda_i - \lambda) }
-  &
-  \mathrm{if \;} \lambda < 2 \, \lambda_i,
-  \\%[1em]
-  %
-  %
-  \frac{ \Gamma_i }
-       {  4 \, t  }
-  \ln \frac{ t + t_0 }
-           {  t_0    }
-  &
-  \mathrm{if \;} \lambda = 2 \, \lambda_i,
-  \\%[1em]
-  %
-  %
-  \frac{ t_0^{ 2 \, \lambda_i / \lambda  - 1 } }
-       {   t^{ 2 \, \lambda_i / \lambda      } }
-  \frac{      \Gamma_i \, \lambda_i^2          }
-       { \lambda \, (\lambda - 2 \, \lambda_i) }
-  &
-  \mathrm{if \;} \lambda > 2 \, \lambda_i.
-\end{dcases}
-\label{eq:y2_sum_perfect_invt}
+      \right)^{1 - 2 \, \nu_k}
+\right],
+\label{eq:error_asym_invt}
 \end{align}
 %
-\note{In the above expression,
-$t + t_0$ are replaced by $t$.
-%
-Adding back $t_0$, we get
+where $\nu_k \equiv \lambda_k / \lambda$.
+%
+At long times, we get
 $$
 \begin{aligned}
-  \Err
+  \Err_A
   \stackrel{         t \to \infty         }
            { =\joinrel=\joinrel=\joinrel= }
-  \sum_{i = 2}^n
+  \sum_{k = 2}^n
   \begin{dcases}
@@
          { t + t_0 }
-    \frac{       \Gamma_i \, \lambda_i^2         }
-         { \lambda \, (2 \, \lambda_i - \lambda) }
+    \frac{ \Gamma_k \, \nu_k^2 }
+         {   2 \, \nu_k - 1    }
     &
-    \mathrm{if \;} \lambda < 2 \, \lambda_i,
+    \mathrm{if \;} 2 \, \nu_k > 1,
     \\%[1em]
@@
     %
-    \frac{    \Gamma_i    }
+    \frac{    \Gamma_k    }
          { 4 \, (t + t_0) }
@@
     &
-    \mathrm{if \;} \lambda = 2 \, \lambda_i,
-    \\
-    %
-    %
-    \frac{  t_0^{ 2 \, \lambda_i / \lambda  - 1}  }
-         { (t + t_0)^{ 2 \, \lambda_i / \lambda } }
-    \frac{      \Gamma_i \, \lambda_i^2          }
-         { \lambda \, (\lambda - 2 \, \lambda_i) }
+    \mathrm{if \;} 2 \, \nu_k = 1,
+    \\
+    %
+    %
+    \frac{  t_0^{ 2 \, \nu_k  - 1}  }
+         { (t + t_0)^{ 2 \, \nu_k } }
+    \frac{ \Gamma_k \, \nu_k^2 }
+         {   1 - 2 \, \nu_k    }
     &
-    \mathrm{if \;} \lambda > 2 \, \lambda_i.
+    \mathrm{if \;} 2 \, \nu_k < 1.
   \end{dcases}
@@
 $$
-}
-%
-The last two cases, $\lambda \ge 2\lambda_i$,
+%
+The last two cases, $\lambda \ge 2\lambda_k$,
 are asymptotically suboptimal
@@
 Thus, the optimal $\lambda$
-satisfies $\lambda < 2 \min \lambda_i$,
+satisfies $\lambda < 2 \min \lambda_k$,
 and in this case,
@@
 \begin{equation}
-\sum_{i = 2}^n
-\frac{ \lambda_i - \lambda }
-{ \left(2 - \lambda/ \lambda_i \right)^2 }
+\sum_{k = 2}^n
+\frac{ \lambda_k - \lambda }
+{ \left(2 - \lambda/ \lambda_k \right)^2 }
 = 0.
@@
 $$
-\frac{ \lambda_i^2 }{ \lambda \, (2 \lambda_i - \lambda) } \ge 1,
-$$
-with the equality achieved only at $\lambda = \lambda_i$,
-we see the from Eq. \eqref{eq:y2_sum_perfect_invt}
+\frac{ \lambda_k^2 }{ \lambda \, (2 \, \lambda_k - \lambda) } \ge 1,
+$$
+with the equality achieved only at $\lambda = \lambda_k$,
+we see the from Eq. \eqref{eq:error_asym_invt}
 that the minimal error is achieved with
@@
 
+In practice, however, we may wish to optimize
+relative short simulations that have not entered
+the asymptotic regime.
+%
+Then the residual error, $\Err_R$, is not negligible.
+%
+From Eq. \eqref{eq:error_res}, we get
+%
+\begin{equation}
+\Err_R
+=
+\sum_{k = 2}^n
+  \frac{ \Gamma_k \, \nu_k }
+       {        2 \, t_0   }
+  \left(
+      \frac{   t_0   }
+           { t + t_0 }
+   \right)^{ 2 \, \nu_k }
+\label{eq:error_res_invt}
+\end{equation}
+%
+where we have assumed that at $t = 0$
+the system has achieved an equilibrium
+at a constant $\alpha_0$,
+such that the values of
+$\left\langle y_k^2(0) \right\rangle$
+can be computed from Eq. \eqref{eq:y2_eql}.
+%
+The total error is given by the sum of
+Eqs. \eqref{eq:error_res_invt} and \eqref{eq:error_asym_invt}.
+%
+\begin{align}
+\Err
+&=
+\Err_R + \Err_A
+\notag
+\\
+&=
+\sum_{ k = 2 }^n
+  E^f_k \,
+  \left\{
+    1
+    +
+    \frac{ 1 } { 2 \, \nu_k - 1 }
+    \left[
+      1 -
+      \left(
+        \frac{   t_0   }
+             { t + t_0 }
+      \right)^{2 \, \nu_k - 1}
+    \right]
+  \right\}
+,
+\label{eq:error_invt}
+\end{align}
+%
+where
+$$
+E^f_k
+\equiv
+\left\langle
+  y_k^2
+\right\rangle_{ \alpha(t) }
+=
+\frac{  \Gamma_k \, \nu_k   }
+     {  2 \, (t + t_0)      }
+,
+$$
+is the equilibrium error of mode $k$
+under the constant terminal value
+of $\alpha(t)$.
+
+Two remarks on Eq. \eqref{eq:error_invt} are in order.
+%
+First, the final error is always greater than
+the saturated equilibrium value at the terminal $\alpha(t)$:
+$$
+E > E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
+$$
+since the second term in the braces
+is nonnegative no matter the sign of $2 \, \nu_k - 1$.
+%
+Second, in the optimal case,
+$\nu_k = 1$,
+and the final error is twice $E_\mathrm{sat}$
+at long times.
+
 \subsection{\label{sec:band-matrix}
-Uniform band matrix}
-
-As an example, consider the following matrix $\mathbf w$
-%
-\begin{equation}
-  w_{ij} = \mu_{|i-j|} + \mu_{i+j-1} + \mu_{2n+1-i-j},
+Homogeneous updating schemes}
+
+In practice, we usually use a homogeneous updating scheme
+that is translational invariant.
+%
+We shall study these schemes below.
+
+\subsubsection{
+Eigenvalues and stability}
+
+Consider the following matrix $\mathbf w$
+%
+\begin{equation}
+  w_{ij}
+  =
+  \mu_{ |i - j| }
+  + \mu_{ i + j - 1 }
+  + \mu_{ 2 n + 1 - i - j },
   \label{eq:w_band}
@@
 and $\mu_l = 0$ for $l > b$.
+%
+The matrix is translational invariant under the reflective
+boundary condition\cite{bussi2006}.
 %
@@
 to a neighbor that is $|i - j|$ bins away,
-whereas the next two terms are added to
-adopt the open boundary condition.
-%
-If $i$ represents a periodic variable,
-the last two terms can be dropped under
-the minimum image convention for $|i-j|$\cite{dama2014}.
-
-This matrix, being symmetric,
+whereas the next two terms are added
+%according to the boundary condition
+to avoid unintended distortion
+to the equilibrium distribution $p_i = 1/n$\cite{bussi2006}.
+%
+Similarly, if $i$ represents a periodic variable,
+the last two terms can be dropped, and
+$|i-j|$ is evaluated
+under the minimum image convention\cite{dama2014}.
+This case is, however, not considered further.
+
+This matrix, being symmetric to the indices $i$ and $j$,
 satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
-for the flat distribution $p_i \equiv 1/n$.
+for the flat distribution $p_i = 1/n$.
 %
@@
        }
-  \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n}.
+  \cos \frac{ \left( i - \frac 1 2 \right) (k - 1) \, \pi}
+            {                  n                         }.
 \label{eq:wband_eigenvector}
@@
   \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
-  \notag \\
+  \label{eq:wband_eigenvalue1}
+  \\
   &=
@@
 \end{align}
-%
+
 \note{We derive Eqs.
@@
   =
-  \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n},
+  \cos \frac{ \left( i - \frac 1 2 \right) (k - 1) \, \pi}{n},
   $$
@@
   \begin{equation}
-    \theta^{(k)}_i = \theta^{(k)}_{1-i} = \theta^{(k)}_{2n+1-i}.
+    \theta^{(k)}_i = \theta^{(k)}_{1 - i} = \theta^{(k)}_{ 2 n + 1 - i }.
     \label{eq:wband_reflection_symmetry}
@@
 &
-+\sum_{i=1}^n
++\sum_{i = 1}^n
   \theta^{(k)}_i
   \left(
-    \mu_{i - j} + \mu_{j - i} + \mu_{i + j - 1} + \mu_{2n+1-i-j}
+    \mu_{i - j} + \mu_{j - i} + \mu_{i + j - 1} + \mu_{2 \, n + 1 - i - j}
   \right)
@@
   &
-  \sum_{i=1}^n
+  \sum_{i = 1}^n
   \theta^{(k)}_i \,
   \left(
-    \mu_{i - j} + \mu_{2n+1-i-j}
+    \mu_{i - j} + \mu_{ 2 n + 1 - i - j }
   \right)
@@
   +
-  \sum_{l=n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, \mu_l
+  \sum_{l = n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, \mu_l
   \\
@@
 \begin{aligned}
-  \sum_{i=1}^n
+  \sum_{i = 1}^n
   \theta^{(k)}_i
@@
 
-We can also show that columnwise orthogonality.
+We can also show that column-wise orthogonality.
 %
@@
 %
+
+We can define a discrete kernel function from the set of numbers as
+$$
+\hat \mu_i = \mu_{ |i| }.
+$$
+Then, according to Eq. \eqref{eq:vt_diffeq_mbin},
+the rate of change of the bias potential, $\dot v_i(t)$,
+is given by a convolution of the above kernel, $\hat \mu_j$,
+and the normalized histogram, $h_j/p_j$\cite{bussi2006},
+and Eq. \eqref{eq:wband_eigenvalue1}
+shows that the eigenvalues, $\lambda_k$, of the updating scheme
+is the Fourier transform of the kernel.
+%
+So, for the updating scheme to be stable,
+the Fourier transform of the kernel
+must be nonnegative.
+
 We discuss two special cases below.
@@
 we convert Eq.
-\eqref{eq:y2_sum_perfect_invt}
+\eqref{eq:error_asym_invt}
 %and
@@
 \frac{ n } { \lambda \, t }
-\sum_i \operatorname{Res}_i
+\sum_l \operatorname{Res}_l
 ,
@@
 
-Another common updating scheme
-in metadynamics is a Gaussian wave.
-%
-Below we show that this scheme
-is stable, although
-it might be less effective
+The Gaussian kernel is commonly
+used in metadynamics.
+%
+Below we show that this
+leads to a stable updating scheme,
+although it can be less effective
 in reducing local noises.
@@
 %will always decay suboptimally,
-%accordingly to Eq. \eqref{eq:y2_sum_perfect_invt}.
+%accordingly to Eq. \eqref{eq:error_asym_invt}.
 %%
@@
 %we can truncate the sum of the error function
-%in Eq. \eqref{eq:y2_sum_perfect_invt}
+%in Eq. \eqref{eq:error_asym_invt}
 %at some $k_{\max}$,
@@
 the best among all updating scheme
-in term of convergence.
+in terms of convergence.
 

### 2016-01-28 doc/invt.tex
replace the thermostat

@@
 \sum_{ k = 2 }^n
-  E^f_k \,
+  \Err^f_k \,
   \left\{
@@
     +
-    \frac{ 1 } { 2 \, \nu_k - 1 }
-    \left[
-      1 -
-      \left(
-        \frac{   t_0   }
-             { t + t_0 }
-      \right)^{2 \, \nu_k - 1}
-    \right]
+    \frac{ 1 - \left[ t_0 / (t + t_0) \right]^{2 \, \nu_k - 1} }
+         {                 2 \, \nu_k - 1                      }
+%    \frac{       1        }
+%         { 2 \, \nu_k - 1 }
+%    \left[
+%      1 -
+%      \left(
+%        \tfrac{   t_0   }
+%              { t + t_0 }
+%      \right)^{2 \, \nu_k - 1}
+%    \right]
   \right\}
@@
 $$
-E^f_k
+\Err^f_k
 \equiv
@@
 
-Two remarks on Eq. \eqref{eq:error_invt} are in order.
+\note{Derivation of Eq. \eqref{eq:error_invt}:
+$$
+\begin{aligned}
+\Err_R
+&=
+\sum_{ k = 2 }^n
+  \Err^f_k \left( \frac{   t_0   }
+                       { t + t_0 }
+          \right)^{ 2 \nu_k - 1 }
+\\
+\Err_A
+&=
+\sum_{ k = 2 }^n
+  \frac{ 2 \, \Err^f_k \, \nu_k }
+       {    2 \, \nu_k - 1      }
+  \left[
+    1 -
+    \left(
+      \tfrac{   t_0   }
+            { t + t_0 }
+      \right)^{2 \, \nu_k - 1}
+  \right]
+\\
+&=
+\sum_{ k = 2 }^n
+  \Err^f_k \,
+  \left[
+    1 -
+    \left(
+      \tfrac{   t_0   }
+            { t + t_0 }
+      \right)^{2 \, \nu_k - 1}
+  \right]
+  +
+  \frac{   \Err^f_k     }
+       { 2 \, \nu_k - 1 }
+  \left[
+    1 -
+    \left(
+      \tfrac{   t_0   }
+           { t + t_0 }
+      \right)^{2 \, \nu_k - 1}
+  \right].
+\end{aligned}
+$$
+Adding the two yields Eq. \eqref{eq:error_invt}.
+}
+
+We give two remarks on Eq. \eqref{eq:error_invt}.
 %
@@
 $$
-since the second term in the braces
+because the second term in the braces
 is nonnegative no matter the sign of $2 \, \nu_k - 1$.
@@
 =
-4 \, z \, 2 \, \delta \, \sqrt{ 1 + \delta^2 }.
+8 \, z \, \delta \, \sqrt{ 1 + \delta^2 }.
 \end{aligned}

### 2016-01-28 prog/invt.c

@@ -36,8 +36,7 @@ static double simulmeta(const invtpar_t *m, double *err0)
 
   if ( m->sampmethod == SAMPMETHOD_MD ) {
     invtmd_init(invtmd, n,
-        m->mddt, m->tp, m->thermdt,
-        m->dwa, m->dwb, v);
+        m->mddt, m->tp, m->thermdt, v);
   }
 
   /* space for the accumulative distribution function */
@@ -144,7 +143,7 @@ static double invt_run(invtpar_t *m)
   long i;
 
   /* clock() is probably better than time(NULL) */
-  mtscramble( clock() );
+  //mtscramble( clock() );
 
   if ( m->docorr ) {
     /* compute correlation functions

### 2016-01-29 doc/invt.tex
add asymptotic error discussion and errcomp.gp

@@
 
+For an arbitrary updating matrix $\mathbf w$,
+optimizing Eq. \eqref{eq:error_asym1}
+can lead to an algebraically complex expression for $\alpha(t)$.
+%
+In practice, it is more useful to find a simple expression for $\alpha(t)$
+that is near optimal.
+%
 From Eq. \eqref{eq:alpha_opt},
-it is reasonable to conjecture that
-a near-optimal $\alpha(t)$ for
-a given updating matrix $\mathbf w$ would be
+a reasonable candidate would be
 %
@@
 %
-Below we shall find the optimal value
-in some special cases.
-%
+Below we shall assume this functional form
+and find the optimal value
+in some cases.
+%
+We shall show that the optimal $\lambda$
+generally depends on not only the spectrum
+of eigenvalues and the sampling method,
+but also other parameters,
+such as the simulation length, $t$.
+
 Using Eq. \eqref{eq:alpha_trial}
@@
 
-In practice, however, we may wish to optimize
-relative short simulations that have not entered
+In practice, however,
+real simulations
+may be relative short
+and have not entered
 the asymptotic regime.
 %
-Then the residual error, $\Err_R$, is not negligible.
+Then, the residual error, $\Err_R$, is not negligible.
 %
@@
            { t + t_0 }
-   \right)^{ 2 \, \nu_k }
+   \right)^{ 2 \, \nu_k },
 \label{eq:error_res_invt}
@@
 %
-where we have assumed that at $t = 0$
+assuming that at $t = 0$
 the system has achieved an equilibrium
@@
 under the constant terminal value
-of $\alpha(t)$.
+of $\alpha(t) = 1/[\lambda (t + t_0)]$.
 
@@
 
-We give two remarks on Eq. \eqref{eq:error_invt}.
+We give several remarks on Eq. \eqref{eq:error_invt}.
 %
@@
 because the second term in the braces
-is nonnegative no matter the sign of $2 \, \nu_k - 1$.
-%
-Second, in the optimal case,
-$\nu_k = 1$,
-and the final error is twice $E_\mathrm{sat}$
-at long times.
+is nonnegative no matter the sign of $2 \, \nu_k - 1$
+(in fact, it is a decreasing function
+that vanishes only at $\nu_k \to +\infty$).
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/errcomp.pdf}
+  }
+  \caption{
+    \label{fig:err_component}
+    Component of the error, $G(\nu_k, r)$,
+    versus $\nu_k = \lambda_k / \lambda$,
+    where $r = t_0 / (t + t_0)$.
+    %
+  }
+\end{center}
+\end{figure}
+
+Second, we shall show that the optimal $\lambda$
+depends on the simulation length,
+and it tends to decrease as simulation lengthens.
+%
+Let us first rewrite Eq. \eqref{eq:error_invt} as
+%
+$$
+\Err
+=
+\sum_{ k = 2 }^n
+  \frac{    \Gamma_k    }
+       { 2 \, (t + t_0) }
+  G\left( \nu_k, \frac{ t_0 } { t + t_0} \right),
+$$
+%
+where the $k$th component is proportional to
+%
+$$
+G\left( \nu_k, r \right)
+=
+\nu_k
+\,
+\left(
+  1
+  +
+  \frac{ 1 - r^{2 \, \nu_k - 1} }
+       {   2 \, \nu_k - 1       }
+\right).
+$$
+%
+As shown in Fig. \ref{fig:err_component},
+the function $G(\nu_k, r)$ depends nontrivial on $\nu_k$.
+%
+For a fixed and sufficiently small $r = t_0 / (t + t_0)$,
+$G(\nu_k, r)$ has two local minima along $\nu_k$,
+one at $\nu_k = 0$, and the other around $\nu_k = 1$,
+with the former being deeper albeit narrower.
+%
+This shows that for a relatively short simulation,
+hence relative large value of $r$,
+the minimal error may be achieved at $\nu_k \approx 0$,
+corresponding to a large value of $\lambda$
+($\lambda \gg \lambda_k$).
+%
+However, this minimum is unstable as simulation lengthens,
+for $G(\nu_k, r)$ increases rapidly
+with decreasing $r$ under a fixed $\nu_k$.
+%
+In other words,
+asymptotically, the minimal error is always achieved at
+$\nu_k = 1$ or $\lambda = \lambda_k$.
+%
+In this case,
+$G(\nu_k, r) \to G(1, 0) = 2$
+and the total error is roughly twice $E_\mathrm{sat}$.
 
@@
 
-The Gaussian kernel is commonly
-used in metadynamics.
-%
-Below we show that this
-leads to a stable updating scheme,
-although it can be less effective
-in reducing local noises.
+The Gaussian updating kernel is commonly
+adopted in metadynamics.
+%
+Below we show that the resulting updating scheme
+is stable, albeit less effective
+in reducing short-wavelength noises.
 
@@
 $x = l \, \Delta x$,
-$\Delta x = \pi/n$,
 and
-$\mu_l = \mu(x) \, \Delta x$.
+$\mu_l = \mu(x) \, \Delta x$,
+where
+$\Delta x = \pi/n$
+is the bin size.
 %
@@
 
-For the Gaussian wave,
-$\mu(x) \propto \exp(-\frac{1}{2} x^2/\sigma^2)$.
-%
-If $\Delta x \ll \sigma \ll 1$,
+Particularly,
+for the Gaussian updating kernel,
+if $\Delta x \ll \sigma \ll 1$,
 we can extended

### 2016-01-29 doc/invt.tex
editing invt.tex

@@
   although metadynamics produces a smoother PMF profile,
-  which is necessary for stable numerical differentiation
-  in computing the force,
   it is less effective in reducing errors with the shorter wavelengths.
   %
-  Thus, it relies more heavily on the assumption of
+  Thus, the latter relies more heavily on the assumption of
   the smoothness of the PMF profile,
@@
 Given a system,
-the problem is to compute a distribution, $p^*(z)$,
+the problem is to compute,
+via either Monte Carlo (MC) and
+molecular dynamics (MD) simulations,
+a distribution, $p^*(z)$,
 along a quantity of interest $z$, and to see how
@@
 a flat distribution\cite{mezei1987, berg1992, lee1993,
-wang2001, wang2001pre, laio2002}
+wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}
 along $z$ over a wide range.
@@
 
-\subsubsection{Integrals of the correlation functions}
+\subsubsection{\label{sec:Gamma}
+Integrals of the correlation functions}
 
@@
 \right).
-\label{eq:T_tridiag}
+\label{eq:T_nn}
 \end{equation}
@@
 For an arbitrary updating matrix $\mathbf w$,
-optimizing Eq. \eqref{eq:error_asym1}
+direct optimization of Eq. \eqref{eq:error_asym1}
 can lead to an algebraically complex expression for $\alpha(t)$.
 %
-In practice, it is more useful to find a simple expression for $\alpha(t)$
+In practice, it is more useful to find
+a simple expression for $\alpha(t)$
 that is near optimal.
@@
 \alpha(\tau) = \frac{1}{\lambda \, (\tau + t_0) },
-\label{eq:alpha_trial}
+\label{eq:alpha_invtlambda}
 \end{equation}
@@
 
-Using Eq. \eqref{eq:alpha_trial}
+Using Eq. \eqref{eq:alpha_invtlambda}
 in Eq. \eqref{eq:error_asym1} yields
@@
 
-We give several remarks on Eq. \eqref{eq:error_invt}.
-%
-First, the final error is always greater than
+%We give several remarks on Eq. \eqref{eq:error_invt}.
+%
+
+Clearly, the final error is always greater than
 the saturated equilibrium value at the terminal $\alpha(t)$:
@@
 that vanishes only at $\nu_k \to +\infty$).
+
+\subsubsection{\label{sec:optlambda}
+Optimal parameter $\lambda$}
+
+We shall show that the optimal $\lambda$
+depends on the simulation length,
+and it tends to decrease as simulation lengthens.
+%
+Let us first rewrite Eq. \eqref{eq:error_invt} as
+%
+$$
+\Err
+=
+\sum_{ k = 2 }^n
+  \frac{    \Gamma_k    }
+       { 2 \, (t + t_0) }
+  G\left( \nu_k, \frac{ t_0 } { t + t_0} \right),
+$$
+%
+where the $k$th component is proportional to
+%
+$$
+G\left( \nu_k, r \right)
+=
+\nu_k
+\,
+\left(
+  1
+  +
+  \frac{ 1 - r^{2 \, \nu_k - 1} }
+       {   2 \, \nu_k - 1       }
+\right).
+$$
+%
+As shown in Fig. \ref{fig:err_component},
+the function $G(\nu_k, r)$ depends non-trivially on $\nu_k$.
+%
+For a fixed and sufficiently small $r = t_0 / (t + t_0)$,
+$G(\nu_k, r)$ has two local minima along $\nu_k$,
+one at $\nu_k = 0$, and the other around $\nu_k = 1$,
+with the former being deeper albeit narrower.
+%
+This means that for a relatively short simulation,
+hence a relatively large value of $r$,
+the minimal error may be achieved at $\nu_k \approx 0$,
+corresponding to a large value of $\lambda$
+($\lambda \gg \lambda_k$).
+%
+However, this minimum is unstable as simulation lengthens,
+for $G(\nu_k, r)$ increases rapidly
+with decreasing $r$ under a fixed $\nu_k$.
+%
+In other words,
+asymptotically, the minimal error is always achieved at
+$\nu_k = 1$ or $\lambda = \lambda_k$.
+%
+In this case,
+$G(\nu_k, r) \to G(1, 0) = 2$
+and the total error is roughly twice $E_\mathrm{sat}$.
+
+For an arbitrary spectrum of eigenvalues,
+$\lambda_2, \dots, \lambda_n$,
+we expect that the optimal $\lambda$ to shift gradually to
+smaller values as the simulation length, $t$, increasing.
+%
+Asymptotically, we expect the optimal $\lambda$
+to be around $\min_k \lambda_k$.
 
@@
 \end{figure}
-
-Second, we shall show that the optimal $\lambda$
-depends on the simulation length,
-and it tends to decrease as simulation lengthens.
-%
-Let us first rewrite Eq. \eqref{eq:error_invt} as
-%
-$$
-\Err
-=
-\sum_{ k = 2 }^n
-  \frac{    \Gamma_k    }
-       { 2 \, (t + t_0) }
-  G\left( \nu_k, \frac{ t_0 } { t + t_0} \right),
-$$
-%
-where the $k$th component is proportional to
-%
-$$
-G\left( \nu_k, r \right)
-=
-\nu_k
-\,
-\left(
-  1
-  +
-  \frac{ 1 - r^{2 \, \nu_k - 1} }
-       {   2 \, \nu_k - 1       }
-\right).
-$$
-%
-As shown in Fig. \ref{fig:err_component},
-the function $G(\nu_k, r)$ depends nontrivial on $\nu_k$.
-%
-For a fixed and sufficiently small $r = t_0 / (t + t_0)$,
-$G(\nu_k, r)$ has two local minima along $\nu_k$,
-one at $\nu_k = 0$, and the other around $\nu_k = 1$,
-with the former being deeper albeit narrower.
-%
-This shows that for a relatively short simulation,
-hence relative large value of $r$,
-the minimal error may be achieved at $\nu_k \approx 0$,
-corresponding to a large value of $\lambda$
-($\lambda \gg \lambda_k$).
-%
-However, this minimum is unstable as simulation lengthens,
-for $G(\nu_k, r)$ increases rapidly
-with decreasing $r$ under a fixed $\nu_k$.
-%
-In other words,
-asymptotically, the minimal error is always achieved at
-$\nu_k = 1$ or $\lambda = \lambda_k$.
-%
-In this case,
-$G(\nu_k, r) \to G(1, 0) = 2$
-and the total error is roughly twice $E_\mathrm{sat}$.
 
@@
 %\right).
-%\label{eq:wtridiag}
+%\label{eq:wnn}
 %\end{equation}
@@
   \lambda_k = 1 - 4 \, \mu_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
-\label{eq:wtridiag_eigenvalue}
+\label{eq:wnn_eigenvalue}
 \end{equation}
@@
 $\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
-$$
-\begin{aligned}
-\Err_2
-&=
+%
+\begin{align}
+\Err_A
+=
 \frac{2 \, n}{\pi \, t}
@@
 \, dp.
-\end{aligned}
-$$
+\notag
+%\label{eq:error_nn_asym_int}
+\end{align}
 
@@
 \begin{aligned}
-\Err_2
+\Err_A
 &=
@@
 \lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
-\label{eq:lambda_tridiag_perfect}
-\end{equation}
-%
-and the error in the optimal is given by
-%
-\begin{equation}
-\Err
+\label{eq:lambda_nn_perfect}
+\end{equation}
+%
+and the asymptotic error in the optimal case is given by
+%
+\begin{equation}
+\Err_A
 =
@@
 \right).
-\label{eq:error_tridiag_prefect}
-\end{equation}
-%
-%From Eq. \eqref{eq:error_tridiag}, it is clear that
+\label{eq:error_nn_prefect}
+\end{equation}
+%
+%From Eq. \eqref{eq:error_nn}, it is clear that
 Clearly, the error
-increases rapidly with magnitude of $\mu_1$,
+increases with the magnitude of $\mu_1$
 no matter its sign,
@@
 corresponds to the single-bin scheme.
-
-For one-step sampling,
-we have, from Eq. \eqref{eq:Gamma_onestep},
-$\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2)$,
+%
+Note that we have ignored the residual error
+and assumed $\lambda < 2 \, \min \lambda_k = 2 - 8 \, \mu_1$
+for all $k$ in the above solution.
+%
+Fortunately,
+the $\lambda$ from Eq. \eqref{eq:lambda_nn_perfect}
+satisfies this condition
+by the stability condition,
+$4 \, \mu_1 < 1$.
+
+For the one-step sampling,
+we find from Eq. \eqref{eq:Gamma_onestep}
+that $\Gamma(p) = \cot^2 p$
+is peaked around $p \approx 0$,
+%
+Then
+$$
+\begin{aligned}
+\Err_A
+&
+\propto
+\frac{   2 \, n }
+     { \pi \, t }
+\left.
+\frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
+     { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
+\right|_{ p \approx 0 }
+\notag \\
+&
+\approx
+\frac{   2 \, n }
+     { \pi \, t }
+\frac{             1             }
+     {  \lambda \, (2 - \lambda) }
+.
+\end{aligned}
+$$
+%
+The optimal $\lambda$ is therefore
+%
+\begin{equation}
+\lambda \approx 1.
+\label{eq:lambda_nn_onestep}
+\end{equation}
+%
+
+\note{We can be a little more precise here.
+%
+From Eq. \eqref{eq:Gamma_onestep}, we get
+%
+$$
+\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
+$$
+%
 where we have introduced
@@
 %
-Then
 $$
 \begin{aligned}
-\Err_2
+\Err_A
 &=
@@
 $$
-%
-The optimal $\lambda$ is therefore
-%
-\begin{equation}
-\lambda = 1,
-\label{eq:lambda_tridiag_onestep}
-\end{equation}
-%
-which is the same of the global sampling scheme.
-
-\note{The integral is evaluated as follows.
+The integral is evaluated as follows.
 %
@@
 and the integral is reduced to a sum of residues
-around the poles within the unit circle.
-
-The poles of the integral come from three factors,
+around the poles within.
+
+The poles of the integral come from three sources,
 namely,
-1: at $z = 0$,
-2: at the zeros of
+% 1.
+$z = 0$,
+% 2.
+the zeros of
 $(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2$,
-3: at the zeros of
+% 3.
+and the zeros of
 $( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2$.
@@
 $$
-then the residue is simply $B$.
+then the residue is $B$.
 %
@@
 \left[ z^2 - \mu_1 (z^2 + 1)^2 \right]^2
-&=
+&
+\approx
 [ - \mu_1 + (1 - 2 \, \mu_1) \, z^2]^2
 \\
-&=
+&
+\approx
 \mu_1^2
@@
 %
+4 \, z^2 - ( z^2 + 1 )^2
+&
+\approx
+-( 1 - 2 \, z^2 ),
+\\
+%
 (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2
-&=
+&
+\approx
 -2 \, \mu_1 \,
@@
       \right)
+      \, z^2
 \right],
@@
 %
-4 \, z - ( z^2 + 1 )^2
-&=
--( 1 - z^2 ),
-\\
-%
 ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2
-&=
+&
+\approx
 1 + (2 + 4 \, \delta^2) \, z^2,
@@
   - \frac{ 2 + \lambda } { 4 }
-  - \mu_1 \, \left( 1 +  \, \delta^2 \right).
+  - \mu_1 \, \left( 1 + 2 \, \delta^2 \right).
 $$
@@
 which are
-$$
-z = \pm \sqrt { \frac{ 2 - \lambda }
-                     { 8 \, \mu_1 } }
-\pm \sqrt { \frac{ 2 - \lambda }
-                 { 8 \, \mu_1 } - 1 }.
-$$
-Two of them lie in the unit circle, namely
-$$
-z = \pm \left(
+\begin{equation}
+z
+=
+\pm
 \sqrt { \frac{ 2 - \lambda }
              { 8 \, \mu_1  } }
-- \sqrt { \frac{ 2 - \lambda }
+\pm
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } - 1 }.
+\label{eq:nnint_local_zeros_part2}
+\end{equation}
+We share distinguish two cases.
+\textbf{Case A.}
+If $\lambda < 2 - 8 \, \mu_1$,
+two of the roots lie in the unit circle, namely
+$$
+z
+=
+\pm
+\left(
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } }
+-
+\sqrt { \frac{ 2 - \lambda }
              { 8 \, \mu_1  } - 1 }
@@
 $$
-and
+which satisfy
 $$
@@
 \begin{aligned}
-\left[ z^2 - \mu_1 \, (1 + z^2)^2 \right]^2
+\left[ z^2 - \mu_1 \, \left( z^2 + 1 \right)^2 \right]^2
 &=
@@
 %
-4 z^2 - (1 + z^2)^2
+4 z^2 - \left( z^2 + 1 \right)^2
 &=
@@
 %
-(z^2 + 1)^2 + 4 \, \delta^2 \, z^2
+\left( z^2 + 1 \right)^2 + 4 \, \delta^2 \, z^2
 &=
@@
 -
-8 \, \mu_1 \, (z^2 + 1) \, z
+8 \, \mu_1 \, \left( z^2 + 1 \right) \, z
 &=
@@
 residue and the sum for this part is
+%
 $$
@@
 \sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }.
+$$
+%
+\textbf{Case B.}
+If $2 - 8 \, \mu_1 \le \lambda < 2$,
+the four zeros in Eq. \eqref{eq:nnint_local_zeros_part2}:
+%
+\begin{equation}
+z
+=
+\pm
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } }
+\pm i \,
+\sqrt { 1 - \frac{ 2 - \lambda }
+                 { 8 \, \mu_1  } },
+\label{eq:nnint_local_zeros_part2b}
+\end{equation}
+%
+all satisfy $|z| = 1$, and hence
+all lie on the border of the unit circle.
+%
+Thus, each contributes half of the residue there.
+%
+However, since the four zeros satisfy
+$$
+2 - \lambda - 4 \, \mu_1 \, ( z^2 + 1)
+=
+\pm i
+\sqrt{ ( 2 - \lambda ) ( 8 \, \mu_1 - 2 + \lambda ) },
+$$
+the sum of the four residues gives zero in this case:
+$$
+\operatorname{Res}_2 = 0.
+$$
+In summary,
+$$
+\operatorname{Res}_2
+=
+\begin{dcases}
+-\frac{                 \lambda^2                     }
+      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
+\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }
+&
+\mathrm{if \;}
+\lambda < 2 - 8 \, \mu_1,
+\\
+0
+&
+\mathrm{otherwise}.
+\end{dcases}
 $$
@@
 $$
-such that
+satisfying
 $$
@@
 \begin{aligned}
-z^2 - \mu_1 (1 + z^2)^2
+z^2 - \mu_1 \, \left( z^2 + 1 \right)^2
 &= z^2 \, (1 + 4 \, \mu_1 \, \delta^2),
 \\
-4 \, z^2 - (1 + z^2)^2
+4 \, z^2 - \left( z^2 + 1 \right)^2
 &=
@@
 (2 - \lambda) \, z^2
-- 2 \, \mu_1 (1 + z^2)^2
+- 2 \, \mu_1 \, \left( z^2 + 1 \right)^2
 &=
@@
 %
-4 \, z \, (1 + z^2)
+4 \, z \, \left( z^2 + 1 \right)
 +
@@
 $$
-Note that the two roots within the unit share the same
+Note that the above two roots share the same
 residue and the sum for this part is
@@
 =
-\frac{ (1 + 4 \, \mu_1 \, \delta^2 )^2 }
-     { (2 + 8 \, \mu_1 \, \delta^2 - \lambda }
+\frac{ \left( 1 + 4 \, \mu_1 \, \delta^2 \right)^2 }
+     {        2 + 8 \, \mu_1 \, \delta^2 - \lambda }
 \frac{ \sqrt{ 1 + \delta^2 } }
@@
 $$
-\Err_2
+\Err_A
 \approx
@@
 { \lambda \, (2 + 8 \, \mu_1 \, \delta^2 - \lambda) }
+\frac{ 1 } { \delta }
 \approx
@@
 \frac{ 1 }
-{ \lambda \, (2 - \lambda) },
+{ \lambda \, (2 - \lambda) }
+\frac{ 1 } { \delta },
 $$
 and the minimum is reached at $\lambda \approx 1$.
+
+The weakness of the above calculation is two fold.
+%
+First, it ignores the residual error.
+%
+Second, the simplified expression for the asymptotic error
+$E_A$ is also misused.
+%
+Namely, with $\lambda \approx 1$,
+we cannot assume $\lambda < 2 \, \lambda_k$ in general.
+%
+So it only works for a very small value of $\mu_1$.
 }
 
+Note that Eq. \eqref{eq:lambda_nn_onestep} is only good
+for sufficiently short simulations.
+%
+For very longer simulations,
+the optimal $\lambda$ tends to shift
+toward the smallest eigenvalue,
+$\min \lambda_k = 1 - 4 \, \mu_1$
+according to the discussion in Sec. \ref{sec:optlambda}.
+%
+
 \subsubsection{Gaussian updating scheme}
@@
 The Gaussian updating kernel is commonly
-adopted in metadynamics.
+adopted by metadynamics simulations.
 %
@@
 
-\subsection{General simulation protocol}
-
-To determine the optimal schedule,
-we shall test a model system
-in the following fashion.
-%
-Initially $\alpha$ is set to some $\alpha_0$
-for a sufficiently long time.
-%
-Then we reset the simulation time $t = 0$,
-and switch to
-$$
-\alpha(t) = \frac{ 1 } { \lambda \, (t + t_0) },
-$$
-where $t_0 = 1/(\lambda \, \alpha_0)$
-to make $\alpha(t)$ continuous.
+\subsection{Test system and simulation protocol}
+
+We tested the above theory on a one-dimensional system
+of $n = 100$ bins,
+with a flat target distribution
+$p_i = 1/n$.
+%
+Since in the asymptotic regime,
+the distribution converges to $p_i$
+no matter the intrinsic distribution, $p^*_i$,
+we also set the latter to be flat for convenience.
+%
+Our goal is to observe how the error
+of the bias potential defined in Eq. \eqref{eq:error_sum}
+depends on the schedule function $\alpha(t)$
+at the end of the simulation.
+
+Initially,
+we set the bias potential to zero,
+$v_i \equiv 0$,
+%
+and let the system undergo a period of equilibration
+during which the updating magnitude is fixed at $\alpha_0$.
+%
+After the equilibration,
+we reset the origin of the time, $t$,
+and started the schedule specified by
+Eq. \eqref{eq:alpha_invtlambda}.
+%
+The goal is then to observe how the final error
+depends on the parameter, $\lambda$.
+%
+Further, we choose $t_0 = 1/(\lambda \, \alpha_0)$
+to smooth the transition from equilibration.
+
+Since the error depends on the underlying sampling scheme,
+we have applied the above protocol to two
+sampling schemes.
+%
+In both cases, we used the
+Metropolis algorithm\cite{metropolis1953, newman, frenkel}
+to sample the state space.
+%
+In each step, a new bin index $j$ was proposed
+and then accepted with probability
+%
+$
+A(i \to j) = \min\{ 1, \exp(v_i - v_j) \}.
+$
+In the first sampling scheme,
+we choose $j$ randomly out of the $n$ bins
+to emulate the perfect sampling scheme
+discussed in Sec. \ref{sec:Gamma}.
+%
+We refer to this scheme as the global sampling scheme.
+%
+In the second sampling scheme,
+we limit $j$ to the two neighbors of $i$,
+$i \pm 1$, with equal probability $1/2$,
+to mimic the one-step sampling scheme
+in Sec. \ref{sec:Gamma}.
+%
+This scheme models local sampling processes
+typical in MD and straightforward MC simulations.
+%
+We refer to this as the local sampling scheme,
 
@@
 
-Show the error versus $\lambda$.
-
-Good agreement between theory and simulation.
-
-The optimal $\lambda$ is near $1.0$,
-independent of the underlying sampling scheme.
+We first tested the single-bin (WL) updating scheme
+using the above protocol.
+%
+Here, we chose $\alpha_0 = 0.0001$
+and ran the simulation for $t = 10^8$ steps.
+%
+As shown in Fig. \ref{fig:err_singlebin},
+the numerical results of the terminal error agreed well with
+the analytical prediction by Eq. \eqref{eq:error_invt}.
+%
+For both global and local sampling schemes,
+the optimal $\lambda$ occurred around $1.0$,
+the optimal value in the asymptotic regime.
+%
+Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
+represents the relative updating magnitude.
+%
+Figure \ref{fig:err_singlebin} also shows that
+a smaller than optimal value of $1/\lambda$
+tends to increase the error more rapidly
+than a larger than optimal value,
+suggesting that the latter is the safer choice
+when the optimal value is unknown.
+
+Our final observation is that the error function
+in the two cases differed only by a multiplicative constant.
+%
+This again agreed with the theory.
+%
+Since all eigenvalues of the single-bin scheme were
+the same, $\lambda_2 = \cdots = \lambda_n = 1$,
+the error in Eq. \eqref{eq:error_invt}
+depends on the sampling scheme only through
+the sum $\sum_{ k = 2 }^n \Gamma_k$,
+which gives rise to the multiplicative constant mentioned above.
 
@@
     \label{fig:err_singlebin}
-    Error, $E$, versus the proportionality constant, $1/\lambda$,
-    in Eq. \eqref{eq:alpha_trial}
+    Error, $E$, versus the
+    relative updating magnitude,
+    %proportionality constant,
+    $1/\lambda$,
+    in Eq. \eqref{eq:alpha_invtlambda}
     for the single-bin (WL) scheme.
     %
+    The results have been averaged over $1000$ independent runs.
   }
@@
 
-\subsection{Nearest neighbor updating scheme}
-
-Normalized error vs. $\lambda$.
-
-In the global sampling scheme,
-the optimal $\lambda$ depends on the simulation length.
-This dependence is more significant in the global sampling scheme.
+\subsection{Nearest-neighbor updating scheme}
+
+We now turn to the nearest-neighbor updating scheme.
+%
+We again find good agreement between simulation and analytical results.
+%
+A new feature is that the optimal $\lambda$
+depends on the simulation length, $t$.
+%
+As shown in Fig. \ref{fig:err_nnbr},
+the normalized error, $(t + t_0) \, \Err$,
+hence the optimal value of $1/\lambda$,
+depended non-trivially on the simulation length
+in the global sampling case.
+%
+The optimal $1/\lambda$ gradually increased
+with simulation length, which is in agreement with
+the discussion in Sec. \ref{sec:optlambda}.
+%
+For the local sampling case, however,
+the normalized error roughly maintained the sample shape,
+and the optimal $\lambda$ is around $1.0$,
+as suggested by Eq. \eqref{eq:lambda_nn_onestep}.
+
+To understand this behavior,
+we observe that for shorter simulations,
+long-wavelength modes,
+(with larger $\lambda_k$ and smaller $k$),
+dominate the error function,
+according to Eq. \eqref{eq:error_invt},
+as the equilibrium error under a fixed updating magnitude
+is proportional to $\lambda_k \, \Gamma_k$.
+%
+For longer simulations,
+short-wavelength modes
+(with smaller $\lambda_k$ and larger $k$)
+become more important,
+for $\lambda_k$ also gives the decay rate
+of the $k$th mode,
+and errors in smaller $\lambda_k$ modes
+are harder to damp out.
+%
+This causes $\lambda$ to shift toward
+a smaller $\lambda_k$ value.
+%
+Note also that while increasing $\lambda$
+also increases the error of the long-wavelength modes,
+this rate of increase is much smaller,
+as shown in the $\nu_k > 1$ branch
+in Fig. \ref{fig:err_component}.
+%
+The shift of $\lambda$, however, is less noticeable
+under the local sampling scheme,
+because the weight of a longer-wavelength mode
+is further multiplied by much larger value of $\Gamma_k$.
 
@@
     Normalized error, $(t + t_0) \, E$,
-    versus the proportionality constant, $1/\lambda$,
-    in Eq. \eqref{eq:alpha_trial}
+    versus the %proportionality constant,
+    relative updating magnitude,
+    $1/\lambda$,
+    in Eq. \eqref{eq:alpha_invtlambda}
     for the nearest-neighbor updating scheme with $\mu_1 = 0.24$.
     %
+    The results have been averaged over $1000$ independent runs.
   }
@@
 
-We verified the optimality of the inverse-time formula.
+We verified the optimality of the inverse-time formula
+for the Wang-Landau algorithm.
 
@@
 
+For metadynamics, and other multiple-bin updating schemes,
+we found formulas to predict the error for finite length
+simulations and the optimal relative updating magnitude,
+$1/\lambda$.
+
+\section{Acknowledgments}
+
+We thanks Dr. Y. Mei and Dr. J. Ma for helpful discussions.
+
 \bibliography{simul}

### 2016-02-02 doc/invt.tex
small corrections and edits to invt.tex

@@
 and metadynamics\cite{laio2002, laio2008, barducci2011, sutto2012}
-are two widely-used techniques for this purpose,
+are two widely-used and closely-related\cite{micheletti2004}
+techniques for this purpose,
 in which a bias potential is actively built-up on-the-fly
@@
 In this study,
-we shall give a proof of the inverse-time formula,
-and try to generalize it to
-updating schemes involving several neighboring bins,
+we shall give another proof of the inverse-time formula.
+%
+We then adapt the formula
+to updating schemes involving several neighboring bins,
 such as those used in metadynamics.
+%
+We shall show that the optimal updating magnitude
+depends nontrivially on the simulation length
+and other parameters.
 %
@@
 the single-bin scheme used by the WL algorithm
-is generally one of the best
-updating schemes in terms of convergence.
+is best in asymptotic convergence.
 %
@@
 %The noise is not necessarily white,
-and the correlation functions
+and the autocorrelation functions
 depend only on the time difference
@@
 \begin{align}
-u(z)
+u(q')
 &\equiv
-e^{-\lambda \, q(t) + \lambda \, z}.
+e^{-\lambda \, q(t) + \lambda \, q'}.
 \label{eq:ut_definition}
@@
 =
-\langle x^2(0) \rangle e^{-2 \, \lambda \, q(t)}
+\langle x^2(0) \, \rangle e^{-2 \, \lambda \, q(t)}
 \notag
@@
 
-We shall further demand that
-%
-\begin{equation}
-  \lim_{t \to \infty} q(t) \to \infty.
-  \label{eq:qt_limit}
-\end{equation}
-%
-Then, for a large $t$,
-the second term on the right-hand side
-of Eq. \eqref{eq:xt_solution} can be neglected, and
-%
-\begin{align}
-\left\langle x^2(t) \right\rangle
+We now fix the endpoint value of $q(t)$
+and variate the curve.
+%
+This is equivalent to variating $u$
+with fixing endpoint values,
+$u\bigl( q(0) \bigr)  = e^{-\lambda \, q(t)}$
+and
+$u\bigl( q(t) \bigr) = 1$.
+%
+Note also that the first term on the right-hand side
+of Eq. \eqref{eq:x2t_average} is fixed during
+the variation process.
+%
+%We shall further demand that
+%%
+%\begin{equation}
+%  \lim_{t \to \infty} q(t) \to \infty.
+%  \label{eq:qt_limit}
+%\end{equation}
+%%
+%Then, for a large $t$,
+%the second term on the right-hand side
+%of Eq. \eqref{eq:xt_solution} can be neglected, and
+%%
+%\begin{align}
+%\left\langle x^2(t) \right\rangle
+%%&=
+%%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
+%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
+%%\notag
+%%\\
 %&=
-%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
-%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
-%\notag
-%\\
-&=
-\int_0^t \int_0^t
-  \dot u\bigl( q(t') \bigr) \,
-  \dot u\bigl( q(t'') \bigr) \,
-  \kappa(t' - t'') \, dt'' \, dt'.
-\label{eq:x2t_average_asym}
-\end{align}
-
-Variating this expression and
-using the Euler-Lagrange equation yields
+%\int_0^t \int_0^t
+%  \dot u\bigl( q(t') \bigr) \,
+%  \dot u\bigl( q(t'') \bigr) \,
+%  \kappa(t' - t'') \, dt'' \, dt'.
+%\label{eq:x2t_average_asym}
+%\end{align}
+%
+By the Euler-Lagrange equation, we get
 $$
@@
 \subsubsection{\label{sec:Gamma}
-Integrals of the correlation functions}
+Integrals of the autocorrelation functions}
 
@@
 %
-The correlation function $\kappa_k(t)$
+The autocorrelation function $\kappa_k(t)$
 decays as $\kappa_k(0) \, \Lambda_k^t$,
@@
 
-\note{If we can interpret $\Gamma_k$ as twice the autocorrelation time,
+\note{Another possible approximation
+  is to interpret $\Gamma_k$ as twice the autocorrelation time,
   and the kernel is exponential
@@
 
-%
-%We can study two limiting cases.
+%\subsubsection{\label{twolimits}
+%Two limits of the optimal schedule, $\alpha(t)$}
+%
+%
+%The optimal $\alpha(t)$ that minimizes the error
+%Eq. \eqref{eq:error_split}
+%is still difficult to find analytically.
+%%
+%However, we can study two limiting cases,
+%for the asymptotic part, Eq. \eqref{eq:error_asym1}.
 %
@@
 %$$
-%\ddot u_n(\tau) = 0,
-%\quad
 %\dot u_n(\tau) = c,
@@
 %$$
+%
+%
+%For $\tau \approx t$,
+%$u_k(q) \approx 1$,
 
 \subsubsection{\label{sec:optWL}
-Optimality of the single-bin scheme}
-
-We now show
-that the single-bin scheme
-is an optimal scheme.
+Asymptotic optimality of the single-bin (WL) scheme}
+
+We now give an interesting deduction
+following from the error functional
+given in Eq. \eqref{eq:error_asym1}:
+the single-bin scheme is asymptotically optimal.
 %
@@
 Using the Cauchy-Schwarz inequality, we have,
-for any set of $\beta_k \ge 0$
+for any set of numbers, $c_k \ge 0$,
 %
@@
 \right)^2.
-\notag
+\label{eq:CSineq}
 \end{align}
 %
-To see this, we first observe that
+This inequality follows from the fact
+that the following quadratic polynomial of $Y$,
 $$
@@
     \left( \dot u_k \, Y - c_k \right)^2
-\ge 0,
-$$
-holds for any $Y$.
-%
-So the left-hand side,
-as a quadratic polynomial of $Y$,
-$A \, Y^2 + B \, Y + C$,
-must have a non-positive discriminant,
+  \equiv
+  A \, Y^2 + B \, Y + C
+%\ge 0,
+$$
+is always nonnegative, and hence
+has a non-positive discriminant,
 $B^2 - 4 \, A \, C$.
 %
-This condition gives the desired inequality.
-%
-Note that the last expression depends on
-$q(t')$ only through the end point value,
+This condition furnishes the desired inequality.
+%
+Note that the last expression of \eqref{eq:CSineq}
+depends on $q(t')$ only through the end point value,
 which is fixed in the variation process.
@@
 of the error $\Err$
-expressed in Eq. \eqref{eq:error_asym1}.
+given in Eq. \eqref{eq:error_asym1}.
 %
@@
 Solving this set of equations yields
+%
 $$
@@
 $$
+%
 Such a solution is possible only if
+%
 \begin{equation}
@@
 %
-\note{
-  We add some steps for the above solution.
+\note{We add some steps for the above solution.
   Integrating $\dot u_k = c_k$ yields
@@
 %
-Thus, in terms of convergence,
+Thus, in terms of asymptotic convergence,
 the single-bin scheme
@@
 
+We must, however, emphasize that the above result
+holds only under the long time limit,
+and it is not necessarily true
+for finite length simulations.
+%
+For example, consider a system that is initially
+prepared under a fixed $\alpha_0$
+before starting the schedule of, $\alpha(t)$.
+%
+Then for shorter times, $t \approx 0$,
+the error is given by Eq. \eqref{eq:error_eql},
+which is clearly smaller if some eigenvalues
+are less than $1.0$ (the single-bin case).
+%
+In practice, real simulations
+may have long achieved the desired precision
+before reaching the asymptotic regime,
+especially with some $\lambda_k$'s close to $0$.
+
 \subsubsection{Variationally optimal schedule, $\alpha(t)$}
 
-For an arbitrary updating matrix $\mathbf w$,
+Generally, for an arbitrary updating matrix $\mathbf w$,
 direct optimization of Eq. \eqref{eq:error_asym1}
-can lead to an algebraically complex expression for $\alpha(t)$.
-%
-In practice, it is more useful to find
+often leads to a complicated expression for $\alpha(t)$.
+%
+In practice, it is also useful to find
 a simple expression for $\alpha(t)$
-that is near optimal.
+that is optimal with respect to some free parameter(s).
 %
@@
 $$
-E > E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
+E \ge E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
 $$
@@
 that vanishes only at $\nu_k \to +\infty$).
+%
+For a set of nonidentical $\lambda_k$'s,
+the equality is reached only at $t = 0$.
 
@@
 2 \int_0^\pi
-  \mu(x) \, \cos \bigl( (k-1) \, x \bigr) \, dx,
+  \mu(x) \, \cos \bigl[ (k-1) \, x \bigr] \, dx,
 \label{eq:lambda_int}
@@
 Metropolis algorithm\cite{metropolis1953, newman, frenkel}
-to sample the state space.
-%
-In each step, a new bin index $j$ was proposed
+to sample the state space:
+%
+in each step, a new bin index $j$ was proposed
 and then accepted with probability
@@
 $
-In the first sampling scheme,
+The difference is that,
+in the first sampling scheme,
 we choose $j$ randomly out of the $n$ bins
@@
 %
-We refer to this scheme as the global sampling scheme.
-%
-In the second sampling scheme,
+We refer to this as the global sampling scheme.
+%
+In the second scheme,
 we limit $j$ to the two neighbors of $i$,
@@
 than a larger than optimal value,
-suggesting that the latter is the safer choice
+suggesting that the latter (over-updating)
+is generally preferable to the latter (under-updating)
 when the optimal value is unknown.
 
-Our final observation is that the error function
-in the two cases differed only by a multiplicative constant.
+Our final observation is that the error functions
+in the two cases shared the same shape,
+differing only by a multiplicative constant.
 %
@@
 the sum $\sum_{ k = 2 }^n \Gamma_k$,
-which gives rise to the multiplicative constant mentioned above.
+which gives rise to the above multiplicative constant.
 
@@
 %
-We again find good agreement between simulation and analytical results.
+We again find good agreement
+between simulation and analytical results,
+although there are subtle differences.
 %
@@
 
-To understand this behavior,
+To understand this dependence,
 we observe that for shorter simulations,
-long-wavelength modes,
+long-wavelength modes
 (with larger $\lambda_k$ and smaller $k$),
-dominate the error function,
-according to Eq. \eqref{eq:error_invt},
-as the equilibrium error under a fixed updating magnitude
-is proportional to $\lambda_k \, \Gamma_k$.
+dominate the error function in Eq. \eqref{eq:error_invt}
+because of the prefactor $\lambda_k \, \Gamma_k$,
+which is inherited from
+the equilibrium error under a fixed updating magnitude,
+Eq. \eqref{eq:error_eql}.
 %
@@
 for $\lambda_k$ also gives the decay rate
-of the $k$th mode,
-and errors in smaller $\lambda_k$ modes
-are harder to damp out.
-%
-This causes $\lambda$ to shift toward
-a smaller $\lambda_k$ value.
+of the $k$th mode.
+%and errors in smaller $\lambda_k$ modes
+%are harder to damp out.
+%
+This drives the optimal $\lambda$ to shift toward
+a smaller $\lambda_k$ value
+as the simulation lengthens.
 %
@@
 under the local sampling scheme,
-because the weight of a longer-wavelength mode
-is further multiplied by much larger value of $\Gamma_k$.
+because longer-wavelength modes more heavily
+weighted by much larger values of $\Gamma_k$.
 
@@
 
-\subsection{Optimality of the Single-bin scheme}
+The error of Gaussian updating scheme is similar to
+that of the nearest neighbor scheme,
+as shown in Fig \ref{fig:err_sig5}.
+%
+The dependency of the simulation length is more prominent.
+%
+Even for local sampling,
+there is now a visible shift of the optimal $\lambda$.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsig5.pdf}
+  }
+  \caption{
+    \label{fig:err_sig5}
+    Normalized error, $(t + t_0) \, E$,
+    versus the %proportionality constant,
+    relative updating magnitude,
+    $1/\lambda$,
+    in Eq. \eqref{eq:alpha_invtlambda}
+    for the Gaussian updating scheme with $\sigma = 5$ bins.
+    %
+    The results have been averaged over $1000$ independent runs.
+  }
+\end{center}
+\end{figure}
+
+\subsection{Optimality of the single-bin scheme}
 

### 2016-02-02 prog/invt.c

@@ -173,12 +173,12 @@ static double invt_run(invtpar_t *m)
           m->n, m->winn, m->win, m->sampmethod,
           "final", m->verbose + 1);
 
-      errref = esterror_ez(m->c, t, m->t0,
+      errref = esterror_ez(m->c, t, m->t0, m->alpha0,
           m->n, m->winn, m->win, m->sampmethod,
           m->verbose + 1);
 
       /* compute the optimal c */
-      optc = estbestc(t, m->t0,
+      optc = estbestc(t, 0, m->alpha0,
           m->n, m->winn, m->win, m->sampmethod,
           1e-8, &errmin, m->verbose);
       printf("predicted optimal c %g, err %g\n", optc, errmin);

### 2016-02-02 prog/invt.h

@@ -365,7 +365,7 @@ static double esterrn(double lambda, double t, double t0,
  * currently, assuming equilibrium values of < x^2 >
  * of alpha(0) = c / t0 at t = 0
  * */
-static double esterror_ez(double c, double t, double t0,
+static double esterror_ez(double c, double t, double t0, double a0,
    int n, int winn, double *win, int sampmethod,
    int verbose)
 {
@@ -382,6 +382,9 @@ static double esterror_ez(double c, double t, double t0,
    * for the updating scheme */
   gamma = estgamma(n, sampmethod);
 
+  if ( fabs(t0) <= 0 ) {
+    t0 = c / a0;
+  }
   err = esterrn(1.0 / c, t, t0, n, lamarr, gamma, xerr);
 
   /* print out the values of lambda_i and gamma_i */
@@ -405,13 +408,16 @@ static double esterror_ez(double c, double t, double t0,
 /* estimate the best parameter c for the updating schedule
  * alpha(t) = c / (t + t0)
  * according to the analytical prediction
+ * if t0 <= 0, t0 is set to 1 / (c a0)
+ * assuming a single local minimum
  * */
-static double estbestc(double t, double t0,
+static double estbestc(double t, double t0, double a0,
    int n, int winn, double *win, int sampmethod,
    double prec, double *err, int verbose)
 {
   double cl, cm, cr, cn;
   double el, em, er, en;
+  int it;
 
   /* specify the initial bracket */
   cl = 0.5;
@@ -419,14 +425,18 @@ static double estbestc(double t, double t0,
   cr = 2.0;
 
   /* compute the values at the initial bracket */
-  el = esterror_ez(cl, t, t0, n, winn, win, sampmethod, 0);
-  em = esterror_ez(cm, t, t0, n, winn, win, sampmethod, 0);
-  er = esterror_ez(cr, t, t0, n, winn, win, sampmethod, 0);
+  el = esterror_ez(cl, t, t0, a0, n, winn, win, sampmethod, 0);
+  em = esterror_ez(cm, t, t0, a0, n, winn, win, sampmethod, 0);
+  er = esterror_ez(cr, t, t0, a0, n, winn, win, sampmethod, 0);
 
-  while ( 1 ) {
+  for ( it = 1; ; it++ ) {
     if ( verbose ) {
-      fprintf(stderr, "%g (%g) - %g (%g) - %g (%g)\n",
-          cl, el, cm, em, cr, er);
[…]

### 2016-02-02 prog/predict.c

@@ -18,7 +18,7 @@ static void invt_predict(invtpar_t *m)
     t0 = c / m->alpha0;
 
     /* c / (t + t0) */
-    err = esterror_ez(c, t, t0,
+    err = esterror_ez(c, t, t0, m->alpha0,
         m->n, m->winn, m->win, m->sampmethod,
         0);
 

### 2016-02-03 doc/invt.tex
invt.tex: add asymptotic limits

@@
 %
-We then adapt the formula
-to updating schemes involving several neighboring bins,
+We then try to derive a similar result
+for the optimal updating magnitude
+for updating schemes involving several neighboring bins,
 such as those used in metadynamics.
 %
-We shall show that the optimal updating magnitude
+It turns out that the optimal updating magnitude
 depends nontrivially on the simulation length
-and other parameters.
-%
-As a byproduct, we shall show that
+and other parameters in the latter case.
+%
+We also shall show that
 the single-bin scheme used by the WL algorithm
-is best in asymptotic convergence.
+is best in asymptotic convergence,
+although practical simulations may not need
+to reach this regime.
 %
@@
 
-%\subsubsection{\label{twolimits}
-%Two limits of the optimal schedule, $\alpha(t)$}
-%
-%
-%The optimal $\alpha(t)$ that minimizes the error
-%Eq. \eqref{eq:error_split}
-%is still difficult to find analytically.
-%%
-%However, we can study two limiting cases,
-%for the asymptotic part, Eq. \eqref{eq:error_asym1}.
-%
-%For $\tau \ll t$,
-%the dominant contribution comes from
-%the term(s) with the smallest eigenvalue $\lambda_k = \lambda_n$,
-%and
-%$$
-%\dot u_n(\tau) = c,
-%\quad
-%\mathrm{and}\;
-%u_n(\tau) = c \, (\tau + t_0),
-%$$
-%which means
-%$$
-%\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
-%$$
-%
-%
-%For $\tau \approx t$,
-%$u_k(q) \approx 1$,
+\subsubsection{\label{twolimits}
+Asymptotically optimal schedule}
+
+The asymptotic error given by Eq. \eqref{eq:error_asym1}
+can be rewritten as
+%
+\begin{equation}
+  \Err_A
+  =
+  \int_0^t
+    {\mathcal L} \bigl[ q(\tau)\bigr]
+    \, d\tau,
+  \notag
+  %\label{eq:error_asym_action}
+\end{equation}
+%
+where, we have defined the Lagrangian,
+%
+\begin{align}
+  {\mathcal L} \bigl[ q(\tau) \bigr]
+  &=
+  \sum_{ k = 2 }^n
+    \Gamma_k \, \dot u_k^2\bigl( q(\tau) \bigr)
+  \notag
+  \\
+  &=
+  \sum_{ k = 2 }^n
+    \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl( q(\tau) \bigr)
+  \; \dot q^2( \tau ).
+  \label{eq:error_asym_lagr}
+\end{align}
+%
+To minimize the functional,
+we notice that the Lagrangian given by Eq. \eqref{eq:error_asym_lagr}
+does not explicitly depend on time, $t$.
+%
+Thus, the Hamiltonian,
+%
+\begin{equation}
+  \mathcal H
+  =
+  \frac{ \partial \mathcal L }
+       { \partial \dot q     }
+  \, \partial \dot q
+  -
+  \mathcal L
+  =
+  2 \, \mathcal L
+  - \mathcal L
+  =
+  \mathcal L,
+  %=
+  %\mathrm{const.}
+  \notag
+  %\label{eq:error_asym_Hamiltonian}
+\end{equation}
+%
+is conserved, and in this case,
+the Hamiltonian and the Lagrangian coincide.
+%
+So we may write
+%
+\begin{equation}
+  \mathcal L
+  =
+    \textstyle\sum_{ k = 2 }^n
+      \Gamma_k \, \lambda_k^2
+      \, u_k^2 \bigl[ q(\tau) \bigr]
+  \;
+  \dot q^2(\tau)
+  =
+  c^2
+  \quad
+  (c > 0).
+  \label{eq:Lagrangian_const}
+\end{equation}
+%
+Integrating integrate the square root of the left-hand side,
+%
+\begin{equation}
+  \int_0^{ q(\tau) }
+    \sqrt{
+      \textstyle\sum_{ k = 2 }^n
+        \Gamma_k \, \lambda_k^2
+        \, u_k^2( q' )
+    }
+    \;
+    d q'
+  =
+  c \, d\tau
+  ,
+  \label{eq:q_opt}
+\end{equation}
+%
+we get an implicit equation for the optimal $\alpha(t')$
+via $q(\tau) = \int_0^\tau \alpha(t') \, dt'$.
+
+Although the explicit solution of Eq. \eqref{eq:q_opt} is difficult,
+we can obtain an expression regarding $\alpha(\tau)$
+by differentiating Eq. \eqref{eq:Lagrangian_const},
+which yields
+%
+\begin{equation}
+  \frac{ d      }
+       { d \tau }
+  \left(
+    \frac{       1        }
+         { \alpha( \tau ) }
+  \right)
+  =
+  \frac{
+    \sum_{ k = 2 }^n
+      \Gamma_k \, \lambda_k^3
+      \, u_k^2 \bigl( q(\tau) \bigr)
+  }
+  {
+    \sum_{ k = 2 }^n
+      \Gamma_k \, \lambda_k^2
+      \, u_k^2 \bigl( q(\tau) \bigr)
+  }
+  .
+  \label{eq:dinvadt}
+\end{equation}
+%
+
+Consider examine two limiting cases.
+%
+First, during the beginning of a long simulation,
+we have $\tau \ll t$ and $q(t) - q(\tau) \to \infty$,
+the sums on the right-hand side of Eq. \eqref{eq:dinvadt}
+is dominated by the term with the largest
+$u_k = e^{ \lambda_k \, [q(\tau) - q(t)] }$,
+or, equivalently, with the smallest $\lambda_k$.
+Then
+\begin{equation}
+  \frac{ d      }
+       { d \tau }
+  \left(
+    \frac{       1        }
+         { \alpha( \tau ) }
+  \right)
+  =
+  \lambda_{ \min }
+  ,
+  \notag
+  %\label{eq:dinvadt_limit1}
+\end{equation}
+%
+and
+%
+\begin{equation}
+  \alpha( \tau )
+  =
+  \frac{               1                  }
+       { \lambda_{ \min } ( \tau + t_1 )  }
+  .
+  \label{eq:alpha_limit1}
+\end{equation}
+%
+
+Second, near the end of simulation,
+we have $\tau \to t$, $u_k \approx 1$,
+%
+\begin{equation}
+  \frac{ d      }
+       { d \tau }
+  \left(
+    \frac{       1        }
+         { \alpha( \tau ) }
+  \right)
+  =
+  \frac{
+    \sum_{ k = 2 }^n
+      \Gamma_k \, \lambda_k^3
+  }
+  {
+    \sum_{ k = 2 }^n
+      \Gamma_k \, \lambda_k^2
+  }
+  =
+  \bar \lambda,
+  \notag
+  %\label{eq:dinvadt_limit2}
+\end{equation}
+%
+where,
+$\bar \lambda$
+is the average of $\lambda_k$ weighted by $\Gamma_k \, \lambda_k^2$,
+and it tends to be dominated
+by the largest $\lambda_k$.
+%
+Thus, in this regime, we have
+%
+\begin{equation}
+  \alpha( \tau )
+  =
+  \frac{             1                }
+       { \bar \lambda ( \tau + t_2 )  }
+  .
+  \label{eq:alpha_limit2}
+\end{equation}
+%
+Thus, we may characterize the optimal schedule as
+%
+\begin{equation}
+  \alpha(\tau)
+  =
+  \frac{                      1                          }
+       { \lambda(\tau) \, \bigl[ \tau + t_0(\tau) \bigr] }
+,
+\notag
+\end{equation}
+%
+where $\lambda(\tau)$ and $t_0(\tau)$ are two
+slowly-varying functions,
+with the former increases gradually from $\lambda_{\min}$
+to $\bar \lambda$.
+
+%Both Eqs. \eqref{eq:alpha_limit1} and \eqref{eq:alpha_limit2}
+%resemble Eq. \eqref{eq:alpha_opt}.
 

### 2016-02-03 prog/invt.h
invt.h adding code for the exact schedule

@@ -260,7 +260,8 @@ static void dumplambdagamma(int n, const double *lamarr,
 
 
 
-/* estimate the error of a constant alpha
+/* estimate the square-root error
+ * under a constant updating magnitude alpha
  * according to the analytical prediction */
 static double esterror0_ez(double alpha,
    int n, int winn, double *win, int sampmethod,
@@ -377,7 +378,7 @@ static double esterror_ez(double c, double t, double t0, double a0,
    * for the updating scheme */
   lamarr = esteigvals(n, winn, win);
 
-  /* estimate the correlation integrals
+  /* estimate the autocorrelation integrals
    * of the eigenmodes of the w matrix,
    * for the updating scheme */
   gamma = estgamma(n, sampmethod);
@@ -505,6 +506,178 @@ static double estbestc(double t, double t0, double a0,
   *err = em;
 
   return cm;
+}
+
+
+
+/* compute the integrand for `intqt` */
+static double intqt_getvel(int n, const double *lamarr, const double *gamma,
+    double dq)
+{
+  int i;
+  double lam, y = 0;
+
+  for ( i = 1; i < n; i++ ) {
+    lam = lamarr[i];
+    y += gamma[i] * lam * lam * exp( 2 * lam * dq );
+  }
+
+  return sqrt( y );
+}
+
+
+
+/* compute the analytically optimal protocol
+ * for a period t, and fixed
+ * qt = Integral {from 0 to t} alpha(t') dt'
+ * the results are saved to tarr[0..m], qarr[0..m], and aarr[0..m]
+ * return the square-root error
+ * */
+static double intqt(double t, double qt,
+    int m, double *tarr, double *aarr, double *qarr,
+    int n, const double *lamarr, const double *gamma)
+{
+  double c, dq, err, y, dt;
+  int j;
+
+  /* integrate over q over uniform grid */
+  dq = qt / m;
+  tarr[0] = 0;
+  qarr[0] = 0;
+  y = intqt_getvel(n, lamarr, gamma, qarr[0] - qt);
+  for ( j = 1; j <= m; j++ ) {
+    qarr[j] = j * dq;
+    tarr[j] = tarr[j - 1] + y * 0.5 * dq;
[…]

### 2016-02-04 doc/invt.tex
invt.h adding code for the exact schedule
add errsigscan.gp and alphacmp.gp

@@
 for the $\langle y_k^2(0) \rangle$'s in Eq. \eqref{eq:error_res}.
-
+%
 \note{Another possible approximation
@@
 }
+%
+In this case, we have
+%
+\begin{equation}
+  \Err_R
+  =
+  \sum_{k = 2}^n
+      \frac 1 2 \, \alpha_0 \, \Gamma_k \, \lambda_k \,
+      e^{  - 2 \, \lambda_k \, q(t) }
+  .
+  \label{eq:error_res1}
+\end{equation}
+%
+It follows that the residue error
+depends on the schedule $\alpha(\tau)$
+only through the integral $q(t)$.
+%
+Thus, if we variate the schedule under a fixed $q(t)$,
+then this part can be ignored,
+and the optimization can be done exclusively
+on the asymptotic part, $\Err_A$.
 
 \subsubsection{\label{twolimits}
-Asymptotically optimal schedule}
-
-The asymptotic error given by Eq. \eqref{eq:error_asym1}
-can be rewritten as
+Optimal schedule}
+
+As shown above, with fixed endpoint value of $q(t)$,
+the optimization of the error is equivalent to
+that of the asymptotic error.
+%
+We now rewrite Eq. \eqref{eq:error_asym1} as
 %
@@
 %
-we get an implicit equation for the optimal $\alpha(t')$
-via $q(\tau) = \int_0^\tau \alpha(t') \, dt'$.
+we get an implicit equation for the optimal $\alpha(\tau)$
+via its integral $q(\tau)$.
 
@@
 such as the simulation length, $t$.
+%
+Note, however,
+since Eq. \eqref{eq:alpha_invtlambda}
+does not fix the endpoint value,
+%
+\begin{equation}
+q(t) = \frac{ 1 } { \lambda }
+\log\left(
+  \frac{ t + t_0 } { t_0 }
+\right),
+\label{eq:qt_invtlambda}
+\end{equation}
+%
+we have to include the residual error
+during the optimization.
 
@@
 %
-From Eq. \eqref{eq:error_res}, we get
+From Eqs. \eqref{eq:error_res1} and \eqref{eq:qt_invtlambda},
+we get
 %
@@
 
-\subsection{Optimality of the single-bin scheme}
-
-Randomize the initial error?
-
-Error vs. $\mu_1$.
-
-Error vs. $\sigma$.
+\subsection{Accuracy of the inverse-time schedule}
+
+In the above, we have assumed the inverse-time schedule,
+Eq. \eqref{eq:alpha_invtlambda}.
+%
+In Fig. \ref{fig:alpha_cmp},
+we compared the inverse-time schedule
+with the optimal $\lambda$,
+and the optimal schedule from integrating
+Eq. \eqref{eq:q_opt}.
+%
+The difference was larger for the global sampling scheme.
+%
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/alphacmp.pdf}
+  }
+  \caption{
+    \label{fig:alpha_cmp}
+    Optimal schedules, $alpha(t)$, from Eq. \eqref{eq:q_opt},
+    versus the inverse-time schedule,
+    Eq. \eqref{eq:alpha_invtlambda},
+    with the optimal parameter, $\lambda$.
+    %
+    For fair comparison,
+    the $q(t) = \int_0^t \alpha(\tau) \, d\tau$
+    were set to be the same
+    for both schedules.
+  }
+\end{center}
+\end{figure}
+
+In Fig. \ref{fig:err_sigscan},
+we compared the error from the inverse-time schedule
+with the optimal $\lambda$,
+and that from the asymptotically optimal schedule,
+for the Gaussian updating scheme.
+%
+The errors from the two schedules appeared to be reasonable close,
+unless for large width case.
+%
+This suggests that very long metadynamics simulations
+may benefit from optimal schedule,
+instead of the simple formula, given by
+Eq. \eqref{eq:alpha_invtlambda}.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsigscan.pdf}
+  }
+  \caption{
+    \label{fig:err_sigscan}
+    Normalized error, $(t + t_0) \, E$,
+    versus the width of the Gaussian updating scheme,
+    $\sigma$,
+    according to the schedule,
+    Eq. \eqref{eq:alpha_invtlambda},
+    with the optimal $\lambda$,
+    and or the asymptotically optimal schedule.
+  }
+\end{center}
+\end{figure}
+
+%\subsection{Optimality of the single-bin scheme}
+
+The above comparison also shows the asymptotic superiority
+of the single-bin scheme.
+It is clear that for short times, 
+the single-bin (WL) may not be optimal,
+but it ultimately wins out as simulation lengthens.
+
+%Randomize the initial error?
+%
+%Error vs. $\mu_1$.
+%
+%Error vs. $\sigma$.
 
@@
 
-\section{Acknowledgments}
-
-We thanks Dr. Y. Mei and Dr. J. Ma for helpful discussions.
+%\section{Acknowledgments}
+%
+%We thanks Dr. Y. Mei and Dr. J. Ma for helpful discussions.
 

### 2016-02-04 prog/invtpar.h

@@ -25,6 +25,10 @@ typedef struct {
   double *p; /* target distribution */
   double alpha0; /* initial updating magnitude */
   int fixa; /* fix alpha during the entire process */
+
+  int alpha_nint; /* number of integration points for the exactly optimal alpha(t) */
+  char fnalpha[FILENAME_MAX]; /* output file for the exactly optimal alpha */
+
   int pbc; /* periodic boundary condition */
   int winn; /* width of the updating window function */
   double win[NBMAX + 1]; /* shape of the window function */
@@ -110,6 +114,9 @@ static void invtpar_init(invtpar_t *m)
   m->p = NULL;
   m->alpha0 = 0.0;
   m->fixa = 0;
+
+  m->alpha_nint = 1000;
+  m->fnalpha[0] = '\0';;
 
   m->pbc = 0;
   m->winn = 1; /* single bin update */
@@ -290,6 +297,8 @@ static void invtpar_help(const invtpar_t *m)
   fprintf(stderr, "  --a0=:         set the initial alpha during equilibration, default %g\n", m->alpha0);
   fprintf(stderr, "  --t0=:         set t0 in alpha = c/(t + t0), if unset, t0 = c/a0, default %g\n", m->t0);
   fprintf(stderr, "  --fixa:        fix the alpha during the entire process, default %d\n", m->fixa);
+  fprintf(stderr, "  --nint:        set the number of integration points for the exact optimal schedule alpha(t), default %d\n", m->fixa);
+  fprintf(stderr, "  --fnalpha:     set the output file to output the exact optimal schedule, alpha(t), default %s\n", m->fnalpha);
   fprintf(stderr, "  --pbc:         use periodic boundary condition, default %d\n", m->pbc);
   fprintf(stderr, "  --nb=:         explicitly set the update window parameters, separated by comma, like --nb=0.2,0.4\n");
   fprintf(stderr, "  --sig=:        set the Gaussian window width, default %g\n", m->wingaus);
@@ -470,9 +479,19 @@ static int invtpar_keymatch(invtpar_t *m,
   {
     m->alpha0 = invtpar_getdouble(m, key, val);
   }
-  else if ( strstartswith(key, "fixa") )
+  else if ( strcmpfuzzy(key, "fixa") == 0
+         || strcmpfuzzy(key, "fixalpha") == 0 )
   {
[…]

### 2016-02-05 doc/invt.tex
reorganize the error code

@@
 In practice, it is also useful to find
-a simple expression for $\alpha(t)$
-that is optimal with respect to some free parameter(s).
+a reasonably optimal $\alpha(t)$ with a simpler expression.
 %
 From Eq. \eqref{eq:alpha_opt},
-a reasonable candidate would be
+we consider the following invert-time formula
 %

### 2016-02-05 prog/invt.h

@@ -7,6 +7,7 @@
 
 
 
+#include <stdarg.h>
 #include "mtrand.h"
 #include "util.h"
 #include "invtpar.h"
@@ -91,21 +92,21 @@ __inline static double normalize(double *v, int n, double sig,
 static double *esteigvals(int n, int winn, const double *win)
 {
   int i, j;
-  double *lamarr, x;
+  double *lambda, x;
 
-  xnew(lamarr, n);
+  xnew(lambda, n);
 
   /* loop over eigenvalues */
   for ( i = 0; i < n; i++ ) {
-    lamarr[i] = 1;
+    lambda[i] = 1;
     /* loop over windows */
     for ( j = 1; j < winn; j++ ) {
       x = sin(i * j * M_PI * 0.5 / n);
-      lamarr[i] -= 4 * win[j] * x * x;
+      lambda[i] -= 4 * win[j] * x * x;
     }
   }
 
-  return lamarr;
+  return lambda;
 }
 
 
@@ -156,16 +157,30 @@ static double *estgamma(int n, int sampmethod)
 
 
 
-/* print out the values of lambda_i and gamma_i */
-static void dumplambdagamma(int n, const double *lamarr,
-    const double *gamma, const double *xerr)
+/* print out the values of lambda and gamma and errors */
+__inline static void dumperror(int n, const double *lambda,
+    const double *gamma, int nerr, ...)
 {
-  int i;
+  int i, j;
+  va_list ap;
+  const double *xerr;
 
-  fprintf(stderr, "   i:   lambda_i      gamma_i\n");
-  for ( i = 0; i < n; i++ ) {
-    fprintf(stderr, "%4d: %10.6f %10.3f %20.6e\n",
-        i + 1, lamarr[i], gamma[i], xerr[i]);
+  fprintf(stderr, "   i:   lambda        gamma    ");
+  for ( j = 0; j < nerr; j++ ) {
+    fprintf(stderr, " err_%-10d", j);
+  }
+  fprintf(stderr, "\n");
+
+  for ( i = 1; i < n; i++ ) {
+    fprintf(stderr, "%4d: %10.6f %10.3f",
+        i + 1, lambda[i], gamma[i]);
+    va_start(ap, nerr);
+    for ( j = 0; j < nerr; j++ ) {
+      xerr = va_arg(ap, const double *);
+      fprintf(stderr, " %14.4e", xerr[i]);
+    }
+    va_end(ap);
[…]

### 2016-02-05 prog/invtpar.h

@@ -125,7 +125,7 @@ static void invtpar_init(invtpar_t *m)
   for ( i = 1; i < m->winn; i++ ) {
     m->win[i] = 0;
   }
-  m->win[0] = 1;
+  m->win[0] = 1; /* single-bin case */
   m->wingaus = 0;
   m->initrand = 0;
   m->kcutoff = 0;
@@ -216,7 +216,7 @@ static void invtpar_compute(invtpar_t *m)
     m->alpha0 = 0.01 / m->n;
   }
 
-  if ( m->t0 <= 0 ) {
+  if ( fabs(m->t0) <= 0 ) { /* means t0 == 0 */
     /* set the default t0 such that
      * alpha is continuous at the beginning
      * of the production run */
@@ -244,22 +244,30 @@ static void invtpar_compute(invtpar_t *m)
     m->ntrials = 1;
   }
 
-  /* construct the Gaussian window */
+  /* initialize the window function */
   if ( m->wingaus > 0 ) {
+
+    /* construct the Gaussian window */
     invtpar_mkgauswin(m);
 
   } else {
 
+    /* normalize the window function
+     * such that it sums to 1.0 */
     for ( x = 0, i = 1; i < m->winn; i++ ) {
       x += m->win[i];
     }
     m->win[0] = 1 - 2 * x;
+
   }
 
   /* initialize the target distribution */
   if ( m->p ) {
+    /* in case this function has been called twice
+     * release the previous memory */
     free(m->p);
   }
+  /* initialize a normalized distribution */
   xnew(m->p, m->n);
   for ( i = 0; i < m->n; i++ ) {
     m->p[i] = 1.0 / m->n;
@@ -270,7 +278,7 @@ static void invtpar_compute(invtpar_t *m)
     m->kcutoff = m->n;
   }
 
-  /* Currently, turn on PBC for MD automatically */
+  /* currently, turn on PBC for MD automatically */
   if ( m->sampmethod == SAMPMETHOD_MD ) {
     m->pbc = 1;
   }

### 2016-02-05 prog/predict.c

@@ -9,28 +9,30 @@
 
 
 
-static void invt_geterr(invtpar_t *m)
+static void invt_geterr(invtpar_t *m,
+    const double *lambda, const double *gamma)
 {
   double t, err1, err2;
 
   t = (double) m->nsteps;
 
   /* compute the optimal c and the error for 1/t formula */
-  m->c = estbestc(t, 0, m->alpha0, m->n, m->winn, m->win,
-      m->sampmethod, 0, &err1, 0);
+  m->c = estbestc_invt(t, m->alpha0, m->n, lambda, gamma,
+      0, &err1, 0);
 
   /* compute the exact minimal error under the same condition */
-  err2 = opterror_ez(m->c, t, m->alpha0,
-      m->alpha_nint, m->fnalpha,
-      m->n, m->winn, m->win, m->sampmethod,
-      NULL, 0);
+  err2 = esterror_opt(t,
+      intq_getqt(t, m->c, m->c / m->alpha0), m->alpha0,
+      m->alpha_nint, NULL, m->n, NULL,
+      lambda, gamma);
 
   printf("c %g, err %g (invt), %g (exact)\n", m->c, err1, err2);
 }
 
 
 
-static void invt_scanc(invtpar_t *m)
+static void invt_scanc(invtpar_t *m,
+    const double *lambda, const double *gamma)
 {
   double c, t, t0;
   double err, err0, err1, err2;
@@ -40,25 +42,22 @@ static void invt_scanc(invtpar_t *m)
     t0 = c / m->alpha0;
 
     /* c / (t + t0) */
-    err = esterror_ez(c, t, t0, m->alpha0,
-        m->n, m->winn, m->win, m->sampmethod,
-        0);
+    err = esterror_invt(t, c, m->alpha0, m->n, NULL,
+        lambda, gamma);
 
     /* initial equilibrium value */
-    err0 = esterror0_ez(m->alpha0,
-        m->n, m->winn, m->win, m->sampmethod,
-        "initial", 0);
+    err0 = esterror_eql(m->alpha0, m->n, NULL,
+        lambda, gamma);
 
     /* final equilibrium value */
-    err1 = esterror0_ez(c / (t0 + t),
-        m->n, m->winn, m->win, m->sampmethod,
-        "final", 0);
+    err1 = esterror_eql(c / (t0 + t), m->n, NULL,
+        lambda, gamma);
 
[…]

### 2016-02-05 doc/invt.tex
add a few references

@@
 
-Free energy calculation\cite{frenkel} is a central theme
+Free energy calculation\cite{frenkel, newman} is a central theme
 in computational physics and chemistry.
@@
 the problem is to compute,
-via either Monte Carlo (MC) and
-molecular dynamics (MD) simulations,
+via either Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
+and molecular dynamics\cite{frenkel, karplus2002} (MD) simulations,
 a distribution, $p^*(z)$,
@@
 The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
-and metadynamics\cite{laio2002, laio2008, barducci2011, sutto2012}
+and metadynamics\cite{huber1994, laio2002, laio2008, barducci2011, sutto2012}
 are two widely-used and closely-related\cite{micheletti2004}
@@
 komura2012, caparica2012, caparica2014,
-barducci2008, dickson2011, dama2014, dickson2015}
+barducci2008, dickson2011, dama2014}
 determines the precision of the final bias potential,
@@
 Similarly, for metadynamics,
-there are various ways of reducing the updating magnitude\cite{
-  marsili2006, barducci2008, dickson2011, dama2014, dickson2015}.
+there are various ways of reducing the updating magnitude,
+particularly in the development of
+self-healing umbrella sampling
+and well-tempered metadynamics\cite{
+  marsili2006, barducci2008, dickson2011, dama2014}.
 
@@
 %according to the boundary condition
-to avoid unintended distortion
+to avoid unintended distortion\cite{dickson2011, mcgovern2013}
 to the equilibrium distribution $p_i = 1/n$\cite{bussi2006}.
@@
 In both cases, we used the
-Metropolis algorithm\cite{metropolis1953, newman, frenkel}
+Metropolis algorithm\cite{metropolis1953, newman, frenkel, landau_binder}
 to sample the state space:

### 2016-02-09 doc/invt.tex
vertical layout for figures

@@
 
-Generally, for an arbitrary updating matrix $\mathbf w$,
-direct optimization of Eq. \eqref{eq:error_asym1}
+Direct optimization of Eq. \eqref{eq:error_asym1}
+for an arbitrary updating matrix $\mathbf w$
 often leads to a complicated expression for $\alpha(t)$.
@@
 We again find good agreement
-between simulation and analytical results,
-although there are subtle differences.
+between simulation and analytical results.
 %
@@
     \label{fig:alpha_cmp}
-    Optimal schedules, $alpha(t)$, from Eq. \eqref{eq:q_opt},
+    Optimal schedules, $\alpha(t)$, from Eq. \eqref{eq:q_opt},
     versus the inverse-time schedule,

### 2016-02-11 doc/invt.tex
add backup figure scripts

@@
-\documentclass[reprint]{revtex4-1}
+\documentclass[reprint, floatfix]{revtex4-1}
 \usepackage{amsmath}
@@
 
-We now give an interesting deduction
-following from the error functional
-given in Eq. \eqref{eq:error_asym1}:
+The error functional given in Eq. \eqref{eq:error_asym1}
+leads to an interesting deduction:
 the single-bin scheme is asymptotically optimal.
@@
 %
-This condition furnishes the desired inequality.
-%
 Note that the last expression of \eqref{eq:CSineq}
@@
 (as adopted in the WL algorithm)
-is generally the most efficient
-updating scheme.
+is most efficient among updating schemes.
 
@@
 before reaching the asymptotic regime,
-especially with some $\lambda_k$'s close to $0$.
+especially if the updating matrix
+has some $\lambda_k$'s close to $0$.
 
@@
 such as the simulation length, $t$.
-%
-Note, however,
-since Eq. \eqref{eq:alpha_invtlambda}
-does not fix the endpoint value,
-%
-\begin{equation}
-q(t) = \frac{ 1 } { \lambda }
-\log\left(
-  \frac{ t + t_0 } { t_0 }
-\right),
-\label{eq:qt_invtlambda}
-\end{equation}
-%
-we have to include the residual error
-during the optimization.
 
@@
 with the equality achieved only at $\lambda = \lambda_k$,
-we see the from Eq. \eqref{eq:error_asym_invt}
-that the minimal error is achieved with
+and the minimal error from \eqref{eq:error_asym_invt}
+is achieved with
 $\lambda_2 = \dots = \lambda_n = \lambda$,
@@
 
-In practice, however,
-real simulations
-may be relative short
-and have not entered
+Real simulations, however,
+can be relatively short
+and need not enter
 the asymptotic regime.
@@
 Then, the residual error, $\Err_R$, is not negligible.
+%
+Note, however,
+since Eq. \eqref{eq:alpha_invtlambda}
+does not fix the endpoint value,
+%
+\begin{equation}
+q(t) = \frac{ 1 } { \lambda }
+\log\left(
+  \frac{ t + t_0 } { t_0 }
+\right),
+\label{eq:qt_invtlambda}
+\end{equation}
+%
+we have to include the residual error
+during the optimization.
 %
@@
 $$
-is the equilibrium error of mode $k$
-under the constant terminal value
+is the equilibrium, or saturated, error of mode $k$
+at the terminal value
 of $\alpha(t) = 1/[\lambda (t + t_0)]$.
@@
 
-We shall show that the optimal $\lambda$
-depends on the simulation length,
-and it tends to decrease as simulation lengthens.
+We now show that the optimal $\lambda$
+tends to decrease with the simulation length.
 %
@@
 $G(\nu_k, r)$ has two local minima along $\nu_k$,
-one at $\nu_k = 0$, and the other around $\nu_k = 1$,
-with the former being deeper albeit narrower.
+one at $\nu_k = 0$,
+and the other shallower one around $\nu_k = 1$.
 %
@@
 %
-However, this minimum is unstable as simulation lengthens,
+However, this minimum is unstable,
 for $G(\nu_k, r)$ increases rapidly
-with decreasing $r$ under a fixed $\nu_k$.
+with decreasing $r$, or increasing simulation length,
+under a fixed $\nu_k$.
 %
@@
 Asymptotically, we expect the optimal $\lambda$
-to be around $\min_k \lambda_k$.
+to be around $\min \lambda_k$.
 
@@
 
-In practice, we usually use a homogeneous updating scheme
-that is translational invariant.
-%
-We shall study these schemes below.
+We now consider a class of commonly-used updating scheme
+for sampling a uniform distribution, $p_i = 1/n$.
+%
+These schemes are homogeneous,
+or translational invariant with respect to $z$.
+%
+This invariance is natural for a periodic variable\cite{dama2014}.
+%
+Below, however, we focus on a variable
+under the reflective boundary conditions.
 
@@
 %
-Similarly, if $i$ represents a periodic variable,
+\note{Similarly, if $i$ represents a periodic variable,
 the last two terms can be dropped, and
@@
 under the minimum image convention\cite{dama2014}.
-This case is, however, not considered further.
+This case is, however, not considered further.}
 
@@
 the rate of change of the bias potential, $\dot v_i(t)$,
-is given by a convolution of the above kernel, $\hat \mu_j$,
+is given by the convolution of the above kernel, $\hat \mu_j$,
 and the normalized histogram, $h_j/p_j$\cite{bussi2006},
@@
 $\min \lambda_k = \lambda_n > 0$,
-and
+or
 $$
@@
 &=
-\frac{n}{\lambda \, t}
+\frac{      n       }
+     { \lambda \, t }
 \left(
- \frac{2 - 4 \, \mu_1 - \lambda}{4}
+ \frac{ 2 - 4 \, \mu_1 - \lambda }
+      {          4               }
  +
@@
 &=
-\frac{n}{\lambda \, t}
+\frac{       n      }
+     { \lambda \, t }
 \left(
-  \frac{2 - 4 \, \mu_1 - \lambda}{4}
+  \frac{ 2 - 4 \, \mu_1 - \lambda }
+       {          4               }
  +
@@
 $4 \, \mu_1 < 1$.
+\note{This is because,
+$$
+\frac{ 1 - 4 \, \mu_1 } { 1 - 2 \, \mu_1 } < 2 - 8 \, \mu_1
+\quad \leftrightarrow \quad
+1 < 2 - 4 \, \mu_1.
+$$
+}
 
@@
 Second, the simplified expression for the asymptotic error
-$E_A$ is also misused.
+$E_A$ may also be misused.
 %
@@
 according to the discussion in Sec. \ref{sec:optlambda}.
-%
 
@@
 
+\section{\label{eq:stabilize_wband}
+Stabilization}
+
 \section{\label{sec:results}

### 2016-02-11 doc/invt.tex
invt.tex: add section for stabilization

@@
 \usepackage{hyperref}
+
+\hypersetup{
+  colorlinks,
+  linkcolor={red!30!black},
+  citecolor={green!20!black},
+  urlcolor={blue!80!black}
+}
 
@@
   in the WL case, each simulation step triggers
-  an update of the bias potential to only
+  an update to only
   the bin containing the current $z$,
   while in the metadynamics case,
-  the bias potential of several neighboring bins
+  several neighboring bins
   are affected as well.
@@
   The asymptotic error of the resulting PMF
-  is determined by the updating magnitude,
-  and the rate of the reduction of the magnitude
-  over the simulation time.
+  is determined by the updating schedule,
+  or the updating magnitude
+  as a function of simulation time.
   %
@@
   In this study, we shall reexamine the optimality of this prescription,
-  and attempt to extend it to the metadynamics case.
-  %
-  Further, we compared the two algorithms, and found that
-  although metadynamics produces a smoother PMF profile,
-  it is less effective in reducing errors with the shorter wavelengths.
-  %
-  Thus, the latter relies more heavily on the assumption of
-  the smoothness of the PMF profile,
-  especially with a wide updating window.
+  and attempt to give similar prescriptions for the metadynamics case.
+  %
+  Our results show that the optimal schedule
+  and the effectiveness of metadynamics
+  are sensitive to the simulation length and other parameters.
+  %Further, we compared the two algorithms, and found that
+  %although metadynamics produces a smoother PMF profile,
+  %it is less effective in reducing errors with the shorter wavelengths.
+  %%
+  %Thus, the latter relies more heavily on the assumption of
+  %the smoothness of the PMF profile,
+  %especially with a wide updating window.
 \end{abstract}
@@
 along the $z$ direction,
-which often represents a slow reaction coordinate.
+which often represents
+a direction of slow motion.
 %
@@
 containing the current $z$,
-while metadynamics adopts an extended window
+while metadynamics adopts an extended updating window
 covering several neighboring bins.
@@
 %
-Thus, one has to reduce the magnitude of updating
+Thus, one often has to reduce the magnitude of updating
 over the course of simulation.
 
-Naturally, the manner of reducing
+Naturally, the updating schedule,
+or the manner of reducing
 the updating magnitude\cite{liang2007,
@@
 In this study,
-we shall give another proof of the inverse-time formula.
-%
-We then try to derive a similar result
-for the optimal updating magnitude
+we shall reexamine the inverse-time formula,
+and similarly optimize schedules
 for updating schemes involving several neighboring bins,
@@
 %
-It turns out that the optimal updating magnitude
-depends nontrivially on the simulation length
-and other parameters in the latter case.
+It turns out that the optimal updating schedule
+can be sensitive the simulation length.
 %
@@
 the single-bin scheme used by the WL algorithm
-is best in asymptotic convergence,
-although practical simulations may not need
-to reach this regime.
+is best in asymptotic convergence.
+%
+For finite-length simulations, however,
+metadynamics may deliver better performance.
 %
@@
 %
-Below, however, we focus on a variable
+Below, however, we focus on a $z$ that is
 under the reflective boundary conditions.
@@
   \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
-  \label{eq:wband_eigenvalue1}
-  \\
-  &=
-  1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
-  .
   \label{eq:wband_eigenvalue}
+  %\\
+  %&=
+  %1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
+  %.
+  %\label{eq:wband_eigenvalue2}
 \end{align}
@@
 and the normalized histogram, $h_j/p_j$\cite{bussi2006},
-and Eq. \eqref{eq:wband_eigenvalue1}
+and Eq. \eqref{eq:wband_eigenvalue}
 shows that the eigenvalues, $\lambda_k$, of the updating scheme
@@
 must be nonnegative.
+%
+In Sec. \ref{sec:stabilize_wband},
+we show how to stabilize an arbitrary kernel.
 
@@
 %
-
-\note{We can be a little more precise here.
+\note{We can be a bit more precise here.
 %
@@
 
-\section{\label{eq:stabilize_wband}
+\subsubsection{\label{sec:stabilize_wband}
 Stabilization}
+
+The updating scheme determined by an arbitrary set of numbers,
+$\mu_1, \dots, \mu_b$, is not necessarily stable,
+i.e., the eigenvalues from Eq. \eqref{eq:wband_eigenvalue}
+can be negative.
+%
+Below we show how to minimally modify these numbers
+to make the eigenvalues nonnegative.
+
+The key of this modification is the inversion formula
+%
+\begin{align}
+  \mu_l
+  =
+  \frac 1 n
+  \left[
+    \frac{ \lambda_1 + (-1)^l \, \lambda_{n+1} }
+         {               2                     }
+    +
+    \sum_{ k = 1 }^{ n - 1 }
+      \lambda_{ k + 1 } \,
+      \cos \frac{ k \, l \, \pi }
+                {      n        }
+  \right],
+  \label{eq:mu_from_lambda}
+\end{align}
+%
+where
+%
+\begin{align}
+  \lambda_{ n + 1 }
+  =
+  (-1)^{ n - 1 }
+  \left[
+    \lambda_1
+    +
+    2 \, \sum_{ k = 1 }^{ n - 1 }
+            (-1)^k \, \lambda_{k + 1}
+  \right],
+  \label{eq:lambdan}
+\end{align}
+%
+which ensures that the $\mu_n$
+computed from Eq. \eqref{eq:mu_from_lambda}
+vanishes.
+%
+\note{The inversion formula is derived from the cosine transform.
+  %
+  If we define $\mu_n = 0$ and
+  %
+  \begin{align}
+    \mu_{ 2 n - l } = \mu_l
+    \quad
+    \mathrm{for\;} l = 1, \dots, n - 1,
+    \label{eq:mu_reflection}
+  \end{align}
+  %
+  then Eq. \eqref{eq:wband_eigenvalue} can be rewritten as
+  %
+  \begin{align}
+    \lambda_{k+1}
+    =
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
+    \label{eq:lambda_cosine_sum}
+  \end{align}
+  %
+  This formula for $\lambda_{k+1}$
+  can be readily extended to $k = 2 \, n - 1$,
+  and we have
+  %
+  \begin{align}
+    \lambda_{ 2 \, n + 1 - k } = \lambda_k.
+    \label{eq:lambda_reflection}
+  \end{align}
+  %
+  Thus,
+  %
+  $$
+  \begin{aligned}
+    \sum_{ k = 0 }^{ 2 \, n - 1 }
+      \lambda_{ k + 1 } \,
+      \cos \frac{ k \, p \, \pi }
+                {      n        }
+    &=
+    \sum_{ k = 0 }^{ 2 \, n - 1 }
+      \sum_{ l = 0 }^{ 2 \, n - 1 }
+        \mu_l \,
+        \cos \frac{ k \, l \, \pi }
+                  {      n        }
+        \cos \frac{ k \, p \, \pi }
+                  {      n        }
+    \\
+    &=
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+      \frac{ \mu_l } { 2 }
+      \sum_{ k = 0 }^{ 2 \, n - 1 }
+        \cos \frac{ k \, (p + l) \, \pi }
+                  {      n        }
+                  +
+        \cos \frac{ k \, (p - l) \, \pi }
+                  {      n        }
+    \\
+    &=
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+      \mu_l \, n \left(
+        \delta_{ p + l - 2 \, n, 0 }
+        +
+        \delta_{ p - l, 0 }
+      \right)
+    \\
+    &=
+    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
+    =
+    2 \, n \, \mu_p.
+  \end{aligned}
+  $$
+  This entails
+  $$
+  \begin{aligned}
+    \mu_l
+    &=
+    \frac{    1   }
+         { 2 \, n }
+    \sum_{ k = 0 }^{ 2 \, n - 1 }
+      \lambda_{ k + 1 } \,
+      \cos \frac{ k \, l \, \pi }
+                {      n        }
+              \\
+    &=
+    \frac{    1   }
+         { 2 \, n }
+    \left[
+      \lambda_1
+      +
+      (-1)^l \, \lambda_{n + 1}
+      +
+      2 \sum_{ k = 1 }^{ n - 1 }
+        \lambda_{ k + 1 } \,
+        \cos \frac{ k \, l \, \pi }
+                  {      n        }
+    \right],
+  \end{aligned}
+  $$
+  where we have used Eq. \eqref{eq:lambda_reflection}
+  in the last step.
+
+  However, we are usually given only $\lambda_1, \dots, \lambda_n$
+  without $\lambda_{n + 1}$.
+  %
+  Fortunately, we can deduce the latter from
+  the condition that $\mu_n = 0$, which means
+  $$
+  \begin{aligned}
+    0 = \mu_n
+    =
+    \frac{1}{n}
+    \left[
+      \frac{ \lambda_1 + (-1)^n \, \lambda_{n+1} }
+           {               2                     }
+      +
+      \sum_{ k = 1 }^{ n - 1 }
+      \lambda_{k+1} (-1)^k
+    \right].
+  \end{aligned}
+  $$
+  This allows us to solve for $\lambda_{ n + 1 }$,
+  yielding Eq. \eqref{eq:lambdan}
+
+  Some examples for checking.
+  For $n = 2$,
+  $\lambda_3
+  = \mu_0 - 2 \, \mu_1
+  = -(\lambda_1 - 2 \, \lambda_2)$,
+  and
+  $$
+  \begin{aligned}
+  \mu_0
+  &=
+  \lambda_2,
+  \\
+  \mu_1
+  &=
+  \frac{ \lambda_1 - \lambda_2 }
+       {           2           }.
+  \end{aligned}
+  $$
+  For $n = 3$,
+  $\lambda_4
+  = \mu_0 - 2 \, \mu_1 + 2 \, \mu_2
+  = \lambda_1 - 2 \, \lambda_2 + 2 \, \lambda_3$,
+  and
+  $$
+  \begin{aligned}
+  \mu_0
+  &=
+  \frac{ 1 } { 3 }
+  \left(
+    \lambda_1 + 2 \, \lambda_3
+  \right),
+  \\
+  \mu_1
+  &=
+  \frac{ 1 } { 3 }
+  \left(
+    \frac 3 2 \lambda_2
+    -
+    \frac 3 2 \lambda_3
+  \right)
+  \\
+  \mu_2
+  &=
+  \frac{ 1 } { 3 }
+  \left(
+    \lambda_1
+    -
+    \frac 3 2 \lambda_2
+    +
+    \frac 1 2 \lambda_3
+  \right)
+  \end{aligned}
+  $$
+}
+Thus, the algorithm of stabilization consists of three steps.
+%
+\begin{enumerate}
+  \item
+    Compute the eigenvalues from $\lambda_0, \dots, \lambda_{n-1}$
+    from the given $\mu_0, \dots, \mu_b$ with $b < n$
+    according to Eq. \eqref{eq:wband_eigenvalue}.
+  \item
+    Replace the negative eigenvalues by zeros.
+  \item
+    Compute the new $\mu_l$ $(0 \le l < n)$ from
+    the modified eigenvalues according to
+    Eq. \eqref{eq:mu_from_lambda}.
+\end{enumerate}
+%
+Note that, after the stabilization,
+the window is usually widened.
 

### 2016-02-12 doc/invt.tex
add dinvacmp.gp

@@
 wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}
-along $z$ over a wide range.
+along $z$.
 %
@@
 divided by the number of bins.
+%
+For metadynamics, ...
 
@@
 We also shall show that
-the single-bin scheme used by the WL algorithm
+the single-bin updating scheme used by the WL algorithm
 is best in asymptotic convergence.
@@
 of the inverse-time formula
-for the single-bin WL algorithm
+for the single-bin updating WL algorithm
 in Sec. \ref{sec:single-bin}.
@@
 
-\subsubsection{$1/t$ formula}
+\subsubsection{Updating magnitude and the inverse-time formula}
 
@@
 
+\subsection{\label{sec:mbin_opta}
+Optimal schedule for the multiple-bin updating scheme}
+
 \subsubsection{\label{eq:eqlerr}
@@
 %
-Integrating integrate the square root of the left-hand side,
+Integrating the square root of the left-hand side, we get
 %
@@
 
-\subsubsection{Variationally optimal schedule, $\alpha(t)$}
+\subsubsection{Inverse-time schedule}
 
@@
 \subsubsection{\label{sec:optlambda}
-Optimal parameter $\lambda$}
+Optimal proportionality constant}
 
@@
 
-We now consider a class of commonly-used updating scheme
-for sampling a uniform distribution, $p_i = 1/n$.
-%
-These schemes are homogeneous,
-or translational invariant with respect to $z$.
-%
-This invariance is natural for a periodic variable\cite{dama2014}.
-%
-Below, however, we focus on a $z$ that is
-under the reflective boundary conditions.
+We now consider a class of updating schemes
+that are homogeneous,
+or translationally invariant. % with respect to $z$,
+%
+As we shall see,
+these schemes include the Gaussian updating scheme
+commonly used in metadynamics.
+%
+Although the translational invariance is
+natural for a periodic variable\cite{dama2014},
+it can be also extended to an open-end variable
+by imposing the reflective boundary conditions.
+%
+For simplicity, we shall assume that our target
+distribution is uniform, or $p_i = 1/n$, below.
 
@@
 %
+Compare with Eqs. \eqref{eq:alpha_limit1} and \eqref{eq:alpha_limit2}.
+
 \begin{figure}[h]
@@
   \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/alphacmp.pdf}
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/dinvacmp.pdf}
   }
@@
     \label{fig:alpha_cmp}
-    Optimal schedules, $\alpha(t)$, from Eq. \eqref{eq:q_opt},
+    Optimal schedules, in terms of $\dot \alpha(t)$, from Eq. \eqref{eq:q_opt},
     versus the inverse-time schedule,

### 2016-02-12 prog/intq.h

@@ -14,6 +14,7 @@ typedef struct {
   double *tarr;
   double *qarr;
   double *aarr;
+  double *dinva;
   int n;
   const double *lambda; /* eigenvalues of the updating magnitude */
   const double *gamma; /* autocorrelation integrals */
@@ -32,6 +33,7 @@ static intq_t *intq_open(double t, double qt, int m,
   xnew(intq->tarr, m + 1);
   xnew(intq->aarr, m + 1);
   xnew(intq->qarr, m + 1);
+  xnew(intq->dinva, m + 1);
   intq->t = t;
   intq->qt = qt;
   intq->n = n;
@@ -45,10 +47,13 @@ static intq_t *intq_open(double t, double qt, int m,
 
 static void intq_close(intq_t *intq)
 {
-  free(intq->tarr);
-  free(intq->aarr);
-  free(intq->qarr);
-  free(intq);
+  if ( intq != NULL ) {
+    free(intq->tarr);
+    free(intq->aarr);
+    free(intq->qarr);
+    free(intq->dinva);
+    free(intq);
+  }
 }
 
 
@@ -109,6 +114,18 @@ static void intq_geta(intq_t *intq)
     intq->aarr[j] += 0.5 * y;
   }
   intq->aarr[m] = y;
+
+  /* differentiate 1/a(t) */
+  y = ( 1.0 / intq->aarr[1] - 1.0 / intq->aarr[0] )
+    / (       intq->tarr[1] -       intq->tarr[0] );
+  intq->dinva[0] = y;
+  for ( j = 1; j < m; j++ ) {
+    intq->dinva[j] += 0.5 * y;
+    y = ( 1.0 / intq->aarr[j + 1] - 1.0 / intq->aarr[j] )
+      / (       intq->tarr[j + 1] -       intq->tarr[j] );
+    intq->dinva[j] += 0.5 * y;
+  }
+  intq->dinva[m] = y;
 }
 
 
@@ -298,8 +315,9 @@ __inline static int intq_save(intq_t *intq,
   for ( i = 0; i < m; i++ ) {
     a1 = c / (intq->tarr[i] + t0);
     q1 = c * log( 1 + intq->tarr[i] / t0 );
-    fprintf(fp, "%g\t%20.8e\t%12.6f\t%20.8e\t%12.6f\n",
-        intq->tarr[i], intq->aarr[i], intq->qarr[i], a1, q1);
[…]

### 2016-02-12 prog/predict.c

@@ -13,6 +13,7 @@ static void invt_geterr(invtpar_t *m,
     const double *lambda, const double *gamma)
 {
   double t, err1, err2;
+  intq_t *intq;
 
   t = (double) m->nsteps;
 
@@ -23,10 +24,15 @@ static void invt_geterr(invtpar_t *m,
   /* compute the exact minimal error under the same condition */
   err2 = esterror_opt(t,
       intq_getqt(t, m->c, m->c / m->alpha0), m->alpha0,
-      m->alpha_nint, NULL, m->n, NULL,
+      m->alpha_nint, &intq, m->n, NULL,
       lambda, gamma);
 
-  printf("c %g, err %g (invt), %g (exact)\n", m->c, err1, err2);
+  /* save the optimal schedule to file */
+  intq_save(intq, m->c, m->c / m->alpha0, m->fnalpha);
+
+  printf("c %g, err %g (invt), %g (exact), %s\n", m->c, err1, err2, m->fnalpha);
+
+  intq_close(intq);
 }
 
 

### 2016-03-20 doc/invt.tex
update doc

@@
   The Wang-Landau (WL) algorithm and metadynamics
-  are two closely-related techniques in free energy simulations.
-  %
-  They allow one to sample a flat distribution
+  are two closely-related techniques
+  in free energy calculations.
+  %
+  Both methods sample a flat distribution
   along a quantity of interest, $z$,
-  by adaptively constructing a bias potential
-  that offsets the potential of mean force (PMF).
-  %
-  The two algorithms differ by
+  by constructing a bias potential that offsets
+  the potential of mean force (PMF) on the fly.
+  %
+  The two, however, differ by
   the manner of updating the bias potential:
   %
-  in the WL case, each simulation step triggers
-  an update to only
+  in the WL case,
+  each update applies to only
   the bin containing the current $z$,
   while in the metadynamics case,
-  several neighboring bins
-  are affected as well.
+  a Gaussian wave packet
+  involving several neighboring bins
+  is used.
   %
@@
   In the WL case,
-  a simple and effective formula for the magnitude
-  is given by the inverse of the number of simulation steps.
-  %
-  In this study, we shall reexamine the optimality of this prescription,
-  and attempt to give similar prescriptions for the metadynamics case.
-  %
-  Our results show that the optimal schedule
-  and the effectiveness of metadynamics
-  are sensitive to the simulation length and other parameters.
+  the optimal schedule
+  is known to be given by
+  the inverse simulation time.
+  %
+  In this study, 
+  we give a formula of computing the optimal schedule
+  for a general updating scheme
+  involving multiple bins.
+  %
+  We show that while metadynamics can be advantageous
+  for relatively short simulations,
+  the WL algorithm is best for asymptotic convergences.
+  %the optimal schedule
+  %and the effectiveness of metadynamics
+  %are sensitive to the simulation length and other parameters.
+  %
   %Further, we compared the two algorithms, and found that
@@
 a distribution, $p^*(z)$,
-along a quantity of interest $z$, and to see how
-the distribution changes with external conditions,
-such as temperatures and pressures.
-%
-The negative logarithm $-\log p^*(z)$,
-or the potential of mean force (PMF),
-defines a free energy.
-%
-A straightforward approach of computing the PMF
-is to run multiple independent simulations
-under different conditions,
-and then piece together the information
-gathered there.
-%
-This solution is often unsatisfactory
+along a quantity of interest $z$.
+%
+The negative logarithm,
+$-\log p^*(z)$,
+defines a free energy,
+or the potential of mean force (PMF).
+%
+Usually, the distribution $p^*(z)$ is
+localized in one or a few small $z$ regions.
+%
+Therefore, to capture the global shape
+of the PMF,
+one has to run multiple independent simulations
+under different umbrella or bias potentials
+along $z$\cite{mezei1987, berg1992, lee1993},
+and then piece together the information.
+%
+This solution is clumsy in practice
+and often suffers from a convergence problem
 for a complex and glassy system,
@@
 A better solution is to artificially alter
-the target distribution,
-such that we can, in a single simulation, sample
-a flat distribution\cite{mezei1987, berg1992, lee1993,
-wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}
-along $z$.
+the target distribution along $z$
+to a flat one\cite{mezei1987, berg1992, lee1993,
+wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
 %
 This alleviates the above problem of local trapping
-as the system is now forced to travel more flexibly
-along the $z$ direction,
+by forcing the system to travel more 
+deliberately along $z$,
 which often represents
@@
 %
-It, however, requires one to efficiently construct a
-bias potential for an often unknown system.
-%
-Note that the bias potential must exactly cancel the PMF
-to achieve a flat distribution.
+To achieve a flat distribution, however,
+the method must be able to construct a
+bias potential that exactly cancels the PMF.
 
 The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
-and metadynamics\cite{huber1994, laio2002, laio2008, barducci2011, sutto2012}
+and meta\-dy\-nam\-ics\cite{huber1994, laio2002,
+laio2008, barducci2011, sutto2012}
 are two widely-used and closely-related\cite{micheletti2004}
@@
 %
-In essence, these techniques
-encourage the system to leave a region that has been visited
-by elevating the bias potential there.
-%
-A main difference of the two techniques
-is that in the WL case, an update is limited to the bin
+These techniques employ updating loops
+that discourage the system
+to sample configurations
+sharing the current $z$ value
+by elevating the bias potential here.
+%
+The main difference of the two
+is that in the WL case,
+each updating step is limited to the bin
 containing the current $z$,
-while metadynamics adopts an extended updating window
+while metadynamics adopts an extended
+updating wave packet
 covering several neighboring bins.
 %
-The latter ensures a smooth profile
-suitable for numerical differentiation for force calculation,
-and hence is more often used molecular dynamics simulations.
-%
-To be effective, the updates to bias potential
-are made at every few steps,
-while a side effect is that the underlying dynamics
-is no longer a equilibrium sampling
-required by statistical mechanics.
-%
-Thus, one often has to reduce the magnitude of updating
-over the course of simulation.
-
-Naturally, the updating schedule,
+The latter is more often used in MD simulations
+as it guarantees a smooth profile
+of the bias potential
+suitable for numerical differentiation
+in deriving the bias force.
+%
+However, in both cases, the bias potential
+is updated frequently (usually every few MC or MD steps),
+which brings about a side effect 
+of disrupting the underlying
+equilibrium sampling dynamics\cite{laio2002}.
+%
+Thus, one often has to reduce
+the updating magnitude
+over the course of simulation
+to decrease the error.
+
+The updating schedule,
 or the manner of reducing
@@
 komura2012, caparica2012, caparica2014,
-barducci2008, dickson2011, dama2014}
+barducci2008, dickson2011, dama2014},
+therefore,
 determines the precision of the final bias potential,
-hence that of PMF.
-%
-For the WL algorithm, it is found
-that asymptotically the optimal magnitude of updating $\ln f$
-should be given by the
+hence that of the PMF.
+%
+For the WL algorithm, it is known
+that the optimal magnitude of updating $\ln f$
+is given by the
 inverse time\cite{liang2007,
@@
 morozov2007, zhou2008},
-where ``time'' is defined as
+where ``time'' means
 the number of simulation steps
@@
 %
-For metadynamics, ...
+The optimal schedule for metadynamics,
+however, is less studied.
 
 In this study,
-we shall reexamine the inverse-time formula,
-and similarly optimize schedules
-for updating schemes involving several neighboring bins,
-such as those used in metadynamics.
-%
-It turns out that the optimal updating schedule
-can be sensitive the simulation length.
-%
-We also shall show that
-the single-bin updating scheme used by the WL algorithm
-is best in asymptotic convergence.
-%
-For finite-length simulations, however,
-metadynamics may deliver better performance.
+we present a method of computing
+the optimal schedule
+for a general updating scheme,
+including those used in WL and metadynamics simulations.
+%
+Although the method recovers the inverse-time formula
+for the WL case,
+it predicts, for metadynamics,
+a more algebraically complex schedule,
+which is sensitive to the simulation length
+and the width of the Gaussian wave packet.
+%
+Further, by comparing
+different updating schemes,
+we show that
+the single-bin updating scheme
+used by the WL algorithm
+is best for infinitely long simulations.
+%
+%For finite-length simulations, however,
+%metadynamics may deliver better performance.
 %
@@
 
-In this section,
-we first review the basics and fix notations
-in Sec. \ref{sec:background}.
-%
-We then prove the optimality
-of the inverse-time formula
-for the single-bin updating WL algorithm
-in Sec. \ref{sec:single-bin}.
-%
-Next, we discuss the generalized
-multiple-bin schemes
+We develop the theory
+in the following order.
+%
+In Sec. \ref{sec:background},
+we first review the basics of
+flat-distribution sampling
+and the development on the optimal schedule.
+%
+This also helps fixing the notations.
+%
+Then, in Sec. \ref{sec:single-bin},
+we illustrate the method of
+computing the optimal schedule
+on the simplest case of
+single-bin updating scheme.
+%
+We shall prove the optimality
+of the known inverse-time formula.
+%
+Next, we turn to the general case
+of multiple-bin updating schemes
 in Secs. \ref{sec:multiple-bin}
-and \ref{sec:band-matrix}.
-%
+to \ref{sec:band-matrix}.
+%
+Finally, we compare different updating schemes
+in the long time limit
+in Sec. \ref{sec:cmpschemes}.
 
@@
 
-\subsubsection{Flat-distribution sampling}
+\subsubsection{\label{sec:FDS}
+Flat-distribution sampling}
 
@@
 such that the integer $i$ represents
-the index of a small interval, or a bin, $(z, z + dz)$.
+the index of a small interval, or a bin,
+$(z, z + dz)$.
 %
@@
 the distribution $p^*_i$ tends to
-be localized around some $i_0$,
-%
-and to find out the global property,
-it is often desirable to carry out
+be localized,
+%
+and to find out the global shape,
+it is often advantageous to carry out
 a biased sampling that targets
@@
 
-Generally, we may wish to sample
+Here we shall call a simulation that targets
 a nearly flat distribution\cite{dayal2004, trebst2004, singh2011},
 $p_i$,
-by adjusting the bias potential $V_i$
-on the fly.
-%
-For example,
-we can monitor the histogram accumulated over
-a period of time and use it for $\pi_i$ in
-Eq. \eqref{eq:pi_p_phi1},
-and then update
-$V_i$ according to
-$$
-V_i^{\mathrm{new}}
-=
-V_i^{\mathrm{old}}
--
-\log \frac{ p_i } { \pi_i }.
-$$
-%
-%From Eq. \eqref{eq:pi_p_phi1},
-%it is clear that to achieve the target distribution
-%$\pi_i = p_i$ requires
-%%
-%\begin{equation}
-%\log p^*_i
-%=V_i + \ln p_i + c
-%\label{eq:bias_pmf}
-%\end{equation}
-%%
-%%
-%In the end of the simulation,
-%the PMF, $-\log p^*_i$,
-%can be deduced from Eq. \eqref{eq:bias_pmf}.
-%
-The above inversion process,
-adopted by early multicanonical or
-entropic sampling\cite{berg1992, lee1993},
-can be inconvenient in practice.
-%
-For example, a reasonably precise the histogram
+a flat-distribution sampling (FDS) simulation.
+%
+Roughly speaking, there are two types of FDS methods.
+%
+The first type,
+referred to as the equilibrium FDS % (E-FDS) methods hereinafter,
+estimates the bias potential, $V_i$,
+and fixes it during simulation.
+%
+Examples include the original
+entropic or multicanonical sampling\cite{berg1992, lee1993}
+and simulated tempering\cite{marinari1992, lyubartsev1992}
+with $i$ being the index of energy
+and that of temperature, respectively.
+%
+The estimated bias potential, however,
+can often be improved in the end of simulation.
+%
+To do so,
+we accumulate a histogram along $i$
+during simulation and use it for $\pi_i$ in
+Eq. \eqref{eq:pi_p_phi1}.
+%
+This allows us to correct $V_i$ as
+%
+\begin{equation}
+V_i^{\mathrm{corr}}
+=
+V_i^{\mathrm{}}
++
+\log \frac{ \pi_i } { p_i }.
+\label{eq:vcorr_equil}
+\end{equation}
+%
+For an unknown system,
+the main difficulty of the process
+is that it is hard to guess a sufficiently accurate
+$V_i$ to start with.
+%
+Thus, 
+it usually takes several iterative rounds of simulations
+to achieve the desired flat distribution.
+%
+Further, a reasonably precise histogram
 requires an accumulation period
 longer than the autocorrelation time,
-whereas the latter is difficult to
-estimate beforehand for an unknown system.
+which can also be hard to estimate beforehand.
+
+For this reason, the second type of methods,
+referred to as the adaptive FDS methods,
+were invented.
+%
+Examples include
+the WL and metadynamics methods.
+%
+These methods update the bias potential more frequently
+and are more capable of achieving a flat distribution.
+%
+However, they are less well understood theoretically,
+the updating loops used there
+break the microscopic reversibility
+of the underlying sampling mechanism.
+%
+%Thus, for a conservative practitioner,
+%the E-FDS methods appear more appealing
+%once a sufficiently accurate bias potential is available.
+%
+To minimize this effect,
+many adaptive methods gradually
+reduce the updating magnitude
+over the course of simulation\cite{marsili2006,
+liang2007,
+belardinelli2007, belardinelli2007jcp, belardinelli2008}.
+%
+Thus, in late stages of a long simulation,
+the updating is negligible and
+such an adaptive simulation is rather similar
+to the corresponding equilibrium simulation.
+%
+If the bias potential were converged to the exact value,
+the equilibrium simulation can be thought as
+the optimal asymptotic limit of the adaptive counterpart,
+delivering the maximal possible efficiency.
 
@@
 for it applies only to the bias potential
-at the current bin, $i(t)$,
+at the current bin, $i(t)$.
 %
@@
 %
-The multiple-bin scheme is more commonly
-used in metadynamics.
+The Gaussian updating scheme
+used by metadynamics
+is a multiple-bin scheme.
 
@@
 (usually half as large as
-that in the previous stage).
+the previous value).
 %
@@
 
-A more effective way
-of updating the $\alpha(t)$
+For the single-bin scheme
+used in WL algorithm,
+it was found that
+a more effective way
+of updating the schedule $\alpha(t)$
+in late stages
 is to follow the formula
@@
 \begin{equation}
-x(t) = x(0) \, e^{-\lambda \, q(t)}
-+ \int_0^t \dot u\bigl( q(t') \bigr) \, \xi(t') \, dt',
+x(T)
+=
+x(0) \, e^{-\lambda \, q(T)}
++
+\int_0^T
+  \dot u\bigl( q(t) \bigr) \, \xi(t) \, dt,
 \label{eq:xt_solution}
@@
 &\equiv
-e^{-\lambda \, q(t) + \lambda \, q'}.
+e^{-\lambda \, q(T) + \lambda \, q'}.
 \label{eq:ut_definition}
 \end{align}
-So
+At the end of period $T$, we have
 \begin{align}
-&\left\langle x^2(t) \right\rangle
-=
-\langle x^2(0) \, \rangle e^{-2 \, \lambda \, q(t)}
+&\left\langle x^2(T) \right\rangle
+=
+\langle x^2(0) \,
+\rangle e^{-2 \, \lambda \, q(T)}
 \notag
@@
 +
-\int_0^t \int_0^t
+\int_0^T \int_0^T
+  \dot u\bigl( q(t) \bigr) \,
   \dot u\bigl( q(t') \bigr) \,
-  \dot u\bigl( q(t'') \bigr) \,
-  \kappa(t' - t'') \, dt'' \, dt'.
+  \kappa(t - t') \, dt' \, dt.
 \label{eq:x2t_average}
@@
 
-We now fix the endpoint value of $q(t)$
+We now fix the endpoint value of $q(T)$
 and variate the curve.
@@
 with fixing endpoint values,
-$u\bigl( q(0) \bigr)  = e^{-\lambda \, q(t)}$
+$u\bigl( q(0) \bigr)  = e^{-\lambda \, q(T)}$
 and
-$u\bigl( q(t) \bigr) = 1$.
+$u\bigl( q(T) \bigr) = 1$.
 %
@@
 &=
-\frac{d}{d\tau} \int_0^t
-  \dot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt'
-\\
-&= \int_0^t
-  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt',
+\frac{d}{dt} \int_0^T
+  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
+\\
+&= \int_0^T
+  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
 \end{aligned}
@@
 %
-In order for this to hold for any $\tau$,
+In order for this to hold for any $t$,
 we must have
@@
 %
-Using Eq. \eqref{eq:ut_definition}
+Using Eq. \eqref{eq:ut_definition},
 we get
@@
 %
-Taking the logarithm, and differentiating this with respect to $t'$
+Taking the logarithm,
+and differentiating both sides
+with respect to $t'$
 yields Eq. \eqref{eq:alpha_opt}.
@@
 %
-In other words, the $\alpha(t)$ given by
-Eq. \eqref{eq:alpha_invt} is optimal
-up to a constant $t_0$.
+In other words, the optimal $\alpha(t)$
+is given by
+\begin{equation}
+\alpha(t)
+=
+\frac{    1     }
+     {  t + t_0 }
+,
+\label{eq:alpha_invt1}
+\end{equation}
+agreeing with Eq. \eqref{eq:alpha_invt}.
 
@@
 %
-The update magnitudes are given by a matrix $\mathbf w$
+The update magnitudes are given by a matrix
+$\mathbf w$
 with elements $w_{ij}$.
@@
 %
-A necessary condition to sample desired distribution
+A necessary condition to sample
+the desired distribution
 $\mathbf p = (p_1, \dots, p_n)$
@@
 
-To simplify the following discussion,
+To simplify the ensuing discussion,
 we shall further impose the
@@
 \end{equation}
-where
-$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$ satisfies
-the orthonormal conditions\cite{vankampen}:
+where the eigenvectors
+$$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$$ 
+satisfy the orthonormal conditions\cite{vankampen}:
 %
@@
 
+Particularly, there is always an eigenvector 
+corresponding to the equilibrium distribution.
+%
+We shall label this eigenvector as the first one,
+and
+%
+\begin{equation}
+\phi_{1k} = p_k,
+\label{eq:eigenmode1}
+\end{equation}
+%
+which satisfies Eq. \eqref{eq:eig_orthonormal_rows}.
+%
+The corresponding eigenvalue is $\lambda_1 = 1$.
+
 %It is worth pointing out
-Note that, in a stable updating scheme,
+Further, in a stable updating scheme,
 all eigenvalues $\lambda_k$ of $\mathbf w$
@@
 %
-Note that there is always one mode, say
-$y_1 = \sum_{j=1}^n p_j \, x_j$
-(corresponding to $\lambda_1 = 1$),
+By Eq. \eqref{eq:eigenmode1},
+the first eigenmode, $y_1 = \sum_{j=1}^n p_j \, x_j$,
 vanishes identically by Eq. \eqref{eq:px_sum}.
@@
 we have, from Eq. \eqref{eq:y2_sum},
+the errors at a period of $T$,
 %
@@
     \left\langle y_k^2(0) \right\rangle \,
-    e^{ - 2 \, \lambda_k  \, q(t) },
+    e^{ - 2 \, \lambda_k  \, q(T) },
   \label{eq:error_res}
@@
 %
-from the decay of the initial errors,
+from the decay of the initial error,
 and the asymptotic error
@@
   =
-  \int_0^t \int_0^t
+  \int_0^T \int_0^T
   \sum_{k = 2}^n
+    \dot u_k\bigl( q(t) \bigr) \,
     \dot u_k\bigl( q(t') \bigr) \,
-    \dot u_k\bigl( q(t'') \bigr) \,
-    \kappa_k(t' - t'') \, dt' \, dt'',
+    \kappa_k(t - t') \, dt \, dt',
   \label{eq:error_asym}
@@
 \begin{align*}
-  u_k(z)
+  u_k(q')
   &\equiv
-  e^{\lambda_k \, [z - q(t)]},
+  e^{\lambda_k \, [q' - q(T)]},
   \\
@@
 \sum_{k = 2}^n
-\lambda_k \, u_k\bigl( q(\tau) \bigr)
-\int_0^t
-\ddot u_k \bigl( q(t') \bigr) \, \kappa_k(t' - \tau) \, dt' = 0.
+\lambda_k \, u_k\bigl( q(t) \bigr)
+\int_0^T
+  \ddot u_k \bigl( q(t') \bigr) \,
+  \kappa_k(t' - t) \, dt' = 0.
 \label{eq:optimal_mbin}
@@
   &=
-  \int_0^t
+  \int_0^T
   \sum_{k = 2}^n
-  \Gamma_k \, \dot u_k^2\bigl( q(t') \bigr) \, dt'.
+  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt.
   \label{eq:error_asym1}
@@
 %
-The next order correction to Eq. \eqref{eq:kappa_delta}
+
+\note{The next order correction to Eq. \eqref{eq:kappa_delta}
 would lead to a correction to $\Err$
-that is about $\alpha_{\max}^2$ times as large.
-
-\note{This is equivalent to saying that the Fourier transform
+that is about $\alpha_{\max}^2$ times as large,
+where $\alpha_{\max} = \max \alpha(t)$.
+%
+This is equivalent to saying that the Fourier transform
   $$
@@
   \Gamma^{(2)}_k
-  \int_0^t \ddot u_k^2\bigl( q(t') \bigr) \, dt'
+  \int_0^T \ddot u_k^2\bigl( q(t) \bigr) \, dt
   \propto
@@
       \Gamma_k \, \lambda_k \,
-      u_k\bigl( q(\tau) \bigr) \,
-      \ddot u_k\bigl( q(\tau) \bigr) = 0.
+      u_k\bigl( q(t) \bigr) \,
+      \ddot u_k\bigl( q(t) \bigr) = 0.
     \label{eq:optimal_mbin1}
@@
 
-We can evaluate the $\Gamma_k$'s
-in two special case.
+We can evaluate the values of $\Gamma_k$'s
+in two special cases.
 %
@@
 \arraycolsep=3.6pt\def\arraystretch{1.4}
-\mathbf T
+\mathbf Q
 =
@@
 %
-We shall further assume that
+For a large class of updating schemes
+(cf. Sec. \ref{sec:band-matrix}),
+we can further assume that
 the eigenvectors of $\mathbf w$
-coincide with those of $\mathbf T$,
-while the eigenvalues may differ
-(this is true for a class of matrices
-discussed in Sec. \ref{sec:band-matrix}).
-%
-The eigenvalues of $\mathbf T$ are
+coincide with those of $\mathbf Q$,
+while the eigenvalues may differ.
+%
+The eigenvalues of $\mathbf Q$ are
 $$
@@
 $$
-Note that these $\Lambda_k$'s are
-not to be confused with the eigenvalues of $\mathbf w$,
-$\lambda_k$'s.
-%
-The autocorrelation function $\kappa_k(t)$
+%Note that these $\Lambda_k$'s are
+%not to be confused with the eigenvalues of $\mathbf w$,
+%$\lambda_k$'s.
+%
+In this case,
+the autocorrelation function $\kappa_k(t)$
 decays as $\kappa_k(0) \, \Lambda_k^t$,
-with the same-time value $\kappa_k(0) = 1$ being the same as
+with the same-time value
+$\kappa_k(0) = 1$
+being the same as
 that in perfect sampling.
@@
 Before addressing the optimal $\alpha(t)$,
-let us first compute the equilibrium value of $\Err$,
-under a fixed $\alpha(t) \equiv \alpha_0$.
-%
-In this case, we can drop $\Err_R$, and
+let us first compute the equilibrium value of $\Err$
+under a fixed schedule $\alpha(t) \equiv \alpha_0$.
+%
+In this case, we drop $\Err_R$, and
+the error at the end of a long period $T'$
+is given by
 %
@@
   =
-  \int_0^t
+  \int_0^{T'}
     \sum_{k = 2}^n
       \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
-      e^{ 2 \, \lambda_k \, \alpha_0 \, (t' - t) }
+      e^{ 2 \, \lambda_k \, \alpha_0 \, (t - T') }
     \, dt
@@
   &
-  \stackrel{ t \to \infty }
+  \stackrel{ T' \to \infty }
   { =\joinrel=\joinrel=\joinrel= }
@@
   \sum_{ k = 2 }^n
-    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0,
+    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0.
   \label{eq:error_eql}
@@
 %
-or, in terms of the components,
+In terms of the components, we have
 %
@@
 
-If we can assume that before entering the asymptotic regime,
-the system has entered the equilibrium under a fixed $\alpha_0$,
-we can then use Eq. \eqref{eq:y2_eql}
+Below we shall assume that before starting
+the schedule $\alpha(t)$ to be optimized,
+the system has entered the equilibrium
+under a fixed $\alpha_0$.
+%
+That is, for $t \le 0$,
+$\alpha(t) = \alpha_0$.
+%
+We can then use Eq. \eqref{eq:y2_eql}
 for the $\langle y_k^2(0) \rangle$'s in Eq. \eqref{eq:error_res}.
@@
   $$
-  \kappa_k = \exp\left( - \frac{2}{\Gamma_k} |t| \right),
+  \kappa_k(t) = \exp\left( - \frac{2}{\Gamma_k} |t| \right),
   $$
@@
 %
-In this case, we have
+In this case,
+the residual error at the end of period $T$
+is given by
 %
@@
       \frac 1 2 \, \alpha_0 \, \Gamma_k \, \lambda_k \,
-      e^{  - 2 \, \lambda_k \, q(t) }
+      e^{  - 2 \, \lambda_k \, q(T) }
   .
@@
 It follows that the residue error
-depends on the schedule $\alpha(\tau)$
-only through the integral $q(t)$.
-%
-Thus, if we variate the schedule under a fixed $q(t)$,
+depends on the schedule $\alpha(t)$
+only through endpoint value of
+the integral $q(T)$.
+%
+Thus, if we variate the schedule under a fixed $q(T)$,
 then this part can be ignored,
@@
 
-\subsubsection{\label{twolimits}
+\subsubsection{\label{sec:optschedule}
 Optimal schedule}
@@
   =
-  \int_0^t
-    {\mathcal L} \bigl[ q(\tau)\bigr]
-    \, d\tau,
+  \int_0^T
+    {\mathcal L} \bigl[ q(t)\bigr]
+    \, dt,
   \notag
@@
 \begin{align}
-  {\mathcal L} \bigl[ q(\tau) \bigr]
+  {\mathcal L} \bigl[ q(t) \bigr]
   &=
   \sum_{ k = 2 }^n
-    \Gamma_k \, \dot u_k^2\bigl( q(\tau) \bigr)
+    \Gamma_k \, \dot u_k^2\bigl[ q(t) \bigr]
   \notag
@@
   \sum_{ k = 2 }^n
-    \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl( q(\tau) \bigr)
-  \; \dot q^2( \tau ).
+    \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
+  \; \dot q^2( t ).
   \label{eq:error_asym_lagr}
@@
 %
-To minimize the functional,
-we notice that the Lagrangian given by Eq. \eqref{eq:error_asym_lagr}
-does not explicitly depend on time, $t$.
-%
-Thus, the Hamiltonian,
+Note that the Lagrangian
+does not explicitly depend on time, $t$,
+so the Hamiltonian,
 %
@@
 %
-is conserved, and in this case,
-the Hamiltonian and the Lagrangian coincide.
+is conserved,
+and coincides with the Lagrangian.
 %
@@
       \Gamma_k \, \lambda_k^2
-      \, u_k^2 \bigl[ q(\tau) \bigr]
+      \, u_k^2 \bigl[ q(t) \bigr]
   \;
-  \dot q^2(\tau)
+  \dot q^2(t)
   =
@@
 \begin{equation}
-  \int_0^{ q(\tau) }
+  \int_{ 0 }^{ q(t) }
     \sqrt{
@@
   =
-  c \, d\tau
+  c \, t
   ,
@@
 %
-we get an implicit equation for the optimal $\alpha(\tau)$
-via its integral $q(\tau)$.
-
-Although the explicit solution of Eq. \eqref{eq:q_opt} is difficult,
-we can obtain an expression regarding $\alpha(\tau)$
+where $c$ is determined by
+the boundary condition at $t = T$.
+%
+Thus, we have an implicit equation for $q(t)$,
+and hence the optimal schedule, $\alpha(t)$.
+
+Although Eq. \eqref{eq:q_opt} is hard to solve explicitly,
+we can obtain an expression regarding $\alpha(t)$
 by differentiating Eq. \eqref{eq:Lagrangian_const},
@@
   \frac{ d   }
-       { d \tau }
-  \left(
+       { d t }
+  \left[
     \frac{       1     }
-         { \alpha( \tau ) }
-  \right)
+         { \alpha( t ) }
+  \right]
   =
@@
       \Gamma_k \, \lambda_k^3
-      \, u_k^2 \bigl( q(\tau) \bigr)
+      \, u_k^2 \bigl[ q(t) \bigr]
   }
@@
       \Gamma_k \, \lambda_k^2
-      \, u_k^2 \bigl( q(\tau) \bigr)
+      \, u_k^2 \bigl[ q(t) \bigr]
   }
@@
 %
-
-Consider examine two limiting cases.
-%
-First, during the beginning of a long simulation,
-we have $\tau \ll t$ and $q(t) - q(\tau) \to \infty$,
+The right-hand side is a weighted average
+of the eigenvalues, $\lambda_k$'s,
+and we shall refer to this as the instantaneous
+eigenvalue, $\lambda(t)$.
+
+For the single-bin scheme, $\lambda_k = 1$,
+and the instantaneous eigenvalue is also unity.
+Thus,
+$$
+\frac { d   }
+      { d t }
+\left[
+  \frac{      1    }
+	   { \alpha(t) }
+\right]
+=
+1,
+$$
+which again leads to Eq. \eqref{eq:alpha_invt1}.
+
+For a general multiple-bin scheme,
+we consider two limiting cases.
+%
+First, in the beginning of a long simulation,
+we have $t \ll T$ and $q(T) - q(t) \to \infty$,
 the sums on the right-hand side of Eq. \eqref{eq:dinvadt}
 is dominated by the term with the largest
-$u_k = e^{ \lambda_k \, [q(\tau) - q(t)] }$,
-or, equivalently, with the smallest $\lambda_k$.
+$u_k = e^{ \lambda_k \, [q(t) - q(T)] }$,
+or, equivalently, with the smallest nonzero
+$\lambda_k$.
 Then
@@
   ,
-  \notag
-  %\label{eq:dinvadt_limit1}
+  %\notag
+  \label{eq:dinvadt_limit1}
 \end{equation}
@@
 \begin{equation}
-  \alpha( \tau )
+  \alpha( t )
   =
   \frac{               1               }
-       { \lambda_{ \min } ( \tau + t_1 )  }
+       { \lambda_{ \min } ( t + t_1 )  }
   .
@@
 Second, near the end of simulation,
-we have $\tau \to t$, $u_k \approx 1$,
+we have $t \to T$, $u_k \approx 1$,
 %
@@
   \bar \lambda,
-  \notag
-  %\label{eq:dinvadt_limit2}
+  %\notag
+  \label{eq:dinvadt_limit2}
 \end{equation}
@@
 $\bar \lambda$
-is the average of $\lambda_k$ weighted by $\Gamma_k \, \lambda_k^2$,
+is the average of $\lambda_k$ weighted
+by $\Gamma_k \, \lambda_k^2$,
 and it tends to be dominated
@@
 \begin{equation}
-  \alpha( \tau )
+  \alpha( t )
   =
   \frac{             1             }
-       { \bar \lambda ( \tau + t_2 )  }
+       { \bar \lambda ( t + t_2 )  }
   .
@@
 \begin{equation}
-  \alpha(\tau)
+  \alpha(t)
   =
   \frac{                  1                     }
-       { \lambda(\tau) \, \bigl[ \tau + t_0(\tau) \bigr] }
+       { \lambda(t) \, \bigl[ t + t_0(t) \bigr] }
 ,
@@
 %
-where $\lambda(\tau)$ and $t_0(\tau)$ are two
-slowly-varying functions,
-with the former increases gradually from $\lambda_{\min}$
-to $\bar \lambda$.
-
-%Both Eqs. \eqref{eq:alpha_limit1} and \eqref{eq:alpha_limit2}
-%resemble Eq. \eqref{eq:alpha_opt}.
-
-\subsubsection{\label{sec:optWL}
-Asymptotic optimality of the single-bin (WL) scheme}
-
-The error functional given in Eq. \eqref{eq:error_asym1}
-leads to an interesting deduction:
-the single-bin scheme is asymptotically optimal.
-%
-Consider a class of $\mathbf w$ matrices
-sharing the same set of eigenvectors,
-hence same $\Gamma_k$'s,
-but with different eigenvalues, $\lambda_k$'s.
-%
-Using the Cauchy-Schwarz inequality, we have,
-for any set of numbers, $c_k \ge 0$,
+where the instantaneous eigenvalue, $\lambda(t)$,
+increases gradually from $\lambda_{\min}$ to $\bar \lambda$,
+and $t_0(t)$ is a slowly-varying function.
+
+\subsubsection{\label{sec:measure_gamma}
+Measurement the integrals
+of the autocorrelation functions}
+
+To use Eq. \eqref{eq:q_opt} in a real simulation,
+we need to know both $\lambda_k$'s and $\Gamma_k$'s.
+%
+While the updating scheme alone
+determines the eigenmodes and $\lambda_k$'s,
+the values of $\Gamma_k$'s depend also on the system
+and the underlying sampling scheme (MC or MD).
+%
+Below is a technique of measuring the latter.
+
+The idea is that we can measure $\Gamma_k$'s
+via a constant schedule, $\alpha(t) = \alpha_0$,
+simulation using Eq. \eqref{eq:y2_eql}.
+%
+A practical difficulty is that
+we often do not know the intrinsic distribution, $p^*_i$,
+beforehand, and thus cannot compute
+the shifted bias potential, $v_i$,
+from the original one, $V_i$
+(cf. Sec. \ref{sec:FDS}).
+%
+To circumvent the problem, we shall use
+the following modification:
+%
+\begin{equation}
+\operatorname{var} Y_k
+=
+\frac{1}{2}
+\alpha \, \Gamma_k \, \lambda_k,
+\label{eq:varY}
+\end{equation}
+%
+where
+``$\operatorname{var}$''
+means the variance from time averaging,
+and
 %
 \begin{align}
-&
-\left(
-  \int_0^t dt'
-    \sum_{k = 2}^n
-      \Gamma_k \, \dot u_k^2\bigl( q(t') \bigr)
-\right)
-%
-\left(
-  \int_0^t dt'
-    \sum_{k = 2}^n
-      \Gamma_k \, c_k^2
-\right)
-%
-\notag
-\\
-&
-\qquad \qquad
-\ge
-\left(
-  \int_0^t dt'
-    \sum_{k = 2}^n
-      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t') \bigr)
-\right)^2
-\notag
-\\
-&
-\qquad \qquad
-=
-\left(
-  \sum_{k = 2}^n \Gamma_k \, c_k
-    \left[
-      1 - e^{ -\lambda_k q(t) }
-    \right]
-\right)^2.
-\label{eq:CSineq}
+Y_k
+&=
+\sum_{ i = 1 }^n
+  \phi_{k i} \, (V_i - \bar V),
+\label{eq:Y_def}
+\\
+\bar V
+&=
+\sum_{ j = 1 }^n p_j \, V_j.
 \end{align}
-%
-This inequality follows from the fact
-that the following quadratic polynomial of $Y$,
-$$
-\int_0^t
-  dt' \sum_{k = 2}^n \Gamma_k \,
-    \left( \dot u_k \, Y - c_k \right)^2
-  \equiv
-  A \, Y^2 + B \, Y + C
-%\ge 0,
-$$
-is always nonnegative, and hence
-has a non-positive discriminant,
-$B^2 - 4 \, A \, C$.
-%
-Note that the last expression of \eqref{eq:CSineq}
-depends on $q(t')$ only through the end point value,
-which is fixed in the variation process.
-%
-Thus, the inequality sets a lower bound
-of the error $\Err$
-given in Eq. \eqref{eq:error_asym1}.
-%
-The equality is achieved
-if $\dot u_k\left( q(t') \right) = c_k$
-for all $k = 2, \dots, n$ at any $t'$,
-up to a multiplicative factor.
-%
-Solving this set of equations yields
-%
-$$
-\alpha(t') = \frac{              1              }
-                  { \lambda_k \, (t' + t_{k0} ) },
-\quad
-\mathrm{with\;\;}
-t_{k0} = \frac{             t            }
-              { e^{ \lambda_k q(t) } - 1 }.
-$$
-%
-Such a solution is possible only if
-%
-\begin{equation}
-\lambda_2 = \dots = \lambda_n.
-\end{equation}
-%
-\note{We add some steps for the above solution.
-  Integrating $\dot u_k = c_k$ yields
-  \begin{equation}
-  u_k(t') = c_k \left(t' + t_{k0} \right).
-  \label{eq:ui_solution}
-  \end{equation}
-  Given that $u_k(t) = 1$ and $u_k(0) = e^{-\lambda_k \, q(t)}$,
+
+To show Eq. \eqref{eq:varY},
+we first use Eq. \eqref{eq:v_def}
+to rewrite Eq. \eqref{eq:x_def}
+as
+$$
+\begin{aligned}
+x_i
+&=
+(V_i - \bar V) + s_i
+\\
+s_i
+&=
+\ln \frac{ p_i } { p^*_i }
+-
+\sum_{ j = 1 }^n
+  p_j \, \ln \frac{ p_j } { p^*_j }.
+\end{aligned}
+$$
+Thus, by Eqs. \eqref{eq:y_def} and \eqref{eq:Y_def},
 we have
 $$
-  c_k = \frac{ 1 }{ t + t_{k0} },
-  \quad
-  \mathrm{and\;}
-  \frac{ t_{k0} } { t + t_{k0} }
-  =
-  e^{ -\lambda_k \, q(t) }.
-  $$
-  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
-  $$
-  -\lambda_k \, q(t) + \lambda_k \, q(t')
-  = \log c_k + \log\left( t' + t_{k0} \right).
-  $$
-  Differentiating this with respect to $t'$ yields
-  $$
-  \lambda_k \, \alpha(t') = \frac{ 1 } { t' + t_{k0} }.
-  $$
-}
-
-Now an updating matrix $\mathbf w$ satisfying this condition
-is essentially a multiple of the identity matrix
-(since $\lambda_1$ corresponds to the mode
-of a uniform shift of all $v_k$,
-we can freely set it to $\lambda_2$
-without changing the nature of the updating scheme).
-%
-Thus, in terms of asymptotic convergence,
-the single-bin scheme
-(as adopted in the WL algorithm)
-is most efficient among updating schemes.
-
-We must, however, emphasize that the above result
-holds only under the long time limit,
-and it is not necessarily true
-for finite length simulations.
-%
-For example, consider a system that is initially
-prepared under a fixed $\alpha_0$
-before starting the schedule of, $\alpha(t)$.
-%
-Then for shorter times, $t \approx 0$,
-the error is given by Eq. \eqref{eq:error_eql},
-which is clearly smaller if some eigenvalues
-are less than $1.0$ (the single-bin case).
-%
-In practice, real simulations
-may have long achieved the desired precision
-before reaching the asymptotic regime,
-especially if the updating matrix
-has some $\lambda_k$'s close to $0$.
-
-\subsubsection{Inverse-time schedule}
-
-Direct optimization of Eq. \eqref{eq:error_asym1}
-for an arbitrary updating matrix $\mathbf w$
-often leads to a complicated expression for $\alpha(t)$.
-%
-In practice, it is also useful to find
-a reasonably optimal $\alpha(t)$ with a simpler expression.
-%
-From Eq. \eqref{eq:alpha_opt},
-we consider the following invert-time formula
-%
-\begin{equation}
-\alpha(\tau) = \frac{1}{\lambda \, (\tau + t_0) },
+y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, s_i.
+$$
+Since the second term on the right-hand side
+is a constant, and the average of $y_k$ is zero,
+we have
+$$
+\operatorname{var} Y_k
+=
+\operatorname{var} y_k
+=
+\langle y_k^2 \rangle.
+$$
+Finally, by using Eq. \eqref{eq:y2_eql},
+we get Eq. \eqref{eq:varY}.
+
+\subsubsection{\label{sec:invt_alpha}
+Inverse-time schedules}
+
+The optimal schedule predicted
+by the direct solution of Eq. \eqref{eq:q_opt}
+is complicated for a general
+multiple-bin updating scheme.
+%
+Several previous studies\cite{marsili2006, barducci2008, dickson2011}
+have been focused on analytically simpler schedules.
+%
+As an example of these approaches,
+we consider here a variant of
+the inverse-time formula, Eq. \eqref{eq:alpha_invt1}
+%
+\begin{equation}
+\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
 \label{eq:alpha_invtlambda}
@@
 %
-for some value of $\lambda$.
-%
-Below we shall assume this functional form
-and find the optimal value
-in some cases.
-%
-We shall show that the optimal $\lambda$
-generally depends on not only the spectrum
-of eigenvalues and the sampling method,
-but also other parameters,
-such as the simulation length, $t$.
+with an optimized proportionality constant, $\lambda$.
+%
+We shall compute the error computed from
+this variationally optimized schedule
+and compare it with that from the optimal schedule.
+%
+We shall show that the optimal parameter, $\lambda$,
+is sensitive to the simulation time, $t$,
+which parallels the shift of
+the instantaneous eigenvalue,
+$\frac{d}{dt}\alpha^{-1}(t)$,
+as discussed in Sec. \ref{sec:optschedule}.
 
@@
 \frac{    1    }
-     { t + t_0 }
+     { T + t_0 }
 \sum_{k = 2}^n
@@
   1 - \left(
-        1 + \frac{ t }{ t_0 }
+        1 + \frac{ T }{ t_0 }
       \right)^{1 - 2 \, \nu_k}
@@
     \frac{    1    }
-         { t + t_0 }
+         { T + t_0 }
     \frac{ \Gamma_k \, \nu_k^2 }
@@
     \frac{    \Gamma_k    }
-         { 4 \, (t + t_0) }
-    \ln \frac{ t + t_0 }
+         { 4 \, (T + t_0) }
+    \ln \frac{ T + t_0 }
              {   t_0   }
@@
     \frac{  t_0^{ 2 \, \nu_k  - 1}  }
-         { (t + t_0)^{ 2 \, \nu_k } }
+         { (T + t_0)^{ 2 \, \nu_k } }
     \frac{ \Gamma_k \, \nu_k^2 }
@@
 are asymptotically suboptimal
-as they decay slower than $1/t$.
+as they decay slower than $1/T$.
 %
@@
 \begin{equation}
-q(t) = \frac{ 1 } { \lambda }
+q(T) = \frac{ 1 } { \lambda }
 \log\left(
-  \frac{ t + t_0 } { t_0 }
+  \frac{ T + t_0 } { t_0 }
 \right),
@@
       \frac{   t_0   }
-           { t + t_0 }
+           { T + t_0 }
    \right)^{ 2 \, \nu_k },
@@
     +
-    \frac{ 1 - \left[ t_0 / (t + t_0) \right]^{2 \, \nu_k - 1} }
+    \frac{ 1 - \left[ t_0 / (T + t_0) \right]^{2 \, \nu_k - 1} }
          {                 2 \, \nu_k - 1                      }
@@
 %        \tfrac{   t_0   }
-%              { t + t_0 }
+%              { T + t_0 }
 %      \right)^{2 \, \nu_k - 1}
@@
 \frac{  \Gamma_k \, \nu_k   }
-     {    2 \, (t + t_0)    }
+     {    2 \, (T + t_0)    }
 ,
@@
 at the terminal value
-of $\alpha(t) = 1/[\lambda (t + t_0)]$.
+of $\alpha(t) = 1/[\lambda (T + t_0)]$.
 
@@
   \Err^f_k \left( \frac{   t_0   }
-                       { t + t_0 }
+                       { T + t_0 }
           \right)^{ 2 \nu_k - 1 }
@@
       \tfrac{   t_0   }
-            { t + t_0 }
+            { T + t_0 }
       \right)^{2 \, \nu_k - 1}
@@
       \tfrac{   t_0   }
-           { t + t_0 }
+           { T + t_0 }
       \right)^{2 \, \nu_k - 1}
@@
 For a set of nonidentical $\lambda_k$'s,
-the equality is reached only at $t = 0$.
+the equality is reached only at $T = 0$.
 
@@
   \frac{    \Gamma_k    }
-       { 2 \, (t + t_0) }
-  G\left( \nu_k, \frac{ t_0 } { t + t_0} \right),
+       { 2 \, (T + t_0) }
+  G\left( \nu_k, \frac{ t_0 } { T + t_0} \right),
 $$
@@
 As we shall see,
-these schemes include the Gaussian updating scheme
-commonly used in metadynamics.
+these schemes include the Gaussian updating scheme,
+which is widely used in metadynamics.
 %
@@
 natural for a periodic variable\cite{dama2014},
-it can be also extended to an open-end variable
+it can be also extended to open-end variables
 by imposing the reflective boundary conditions.
@@
 \varphi^{(k)}_i
-= \sqrt{
+=
+\varphi_{k i}
+=
+\sqrt{
     \frac{ 2 - \delta_{k, 1} }
@@
 
-We can define a discrete kernel function from the set of numbers as
+We can define a window kernel from the set of numbers as
 $$
@@
 or
-$$
+\begin{equation}
 \mu_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
-$$
+\label{eq:nn_stable}
+\end{equation}
 
@@
 =
-\frac{2 \, n}{\pi \, t}
+\frac{2 \, n}{\pi \, T}
 \int_0^{\pi/2}
@@
 &=
-\frac{2 \, n}{\pi \, t}
+\frac{2 \, n}{\pi \, T}
 \int_0^{\pi/2}
@@
 &=
-\frac{n}{4 \, t}
+\frac{n}{4 \, T}
 \left(
@@
 &=
-\frac{n}{\lambda \, t}
+\frac{n}{\lambda \, T}
 \frac{1}{2 \pi i}
@@
 This integral has five poles, one at the origin,
-which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
+which produces a residue
+$\frac{2 - 4 \, \mu_1 - \lambda}{4}$
 (from series expansion of small $z$),
@@
 \frac{      n       }
-     { \lambda \, t }
+     { \lambda \, T }
 \left(
@@
 \frac{       n      }
-     { \lambda \, t }
+     { \lambda \, T }
 \left(
@@
 =
-\frac{n}{t}
+\frac{n}{T}
 \left(
@@
 \frac{   2 \, n }
-     { \pi \, t }
+     { \pi \, T }
 \left.
@@
 \frac{   2 \, n }
-     { \pi \, t }
+     { \pi \, T }
 \frac{             1             }
@@
 \approx
-\frac{ n } { t }
+\frac{ n } { T }
 \frac{ (1 + 4 \, \mu_1 \, \delta^2)^2 }
@@
 \approx
-\frac{ n } { t }
+\frac{ n } { T }
 \frac{ 1 }
@@
 
-The updating scheme determined by an arbitrary set of numbers,
-$\mu_1, \dots, \mu_b$, is not necessarily stable,
-i.e., the eigenvalues from Eq. \eqref{eq:wband_eigenvalue}
-can be negative.
-%
-Below we show how to minimally modify these numbers
-to make the eigenvalues nonnegative.
-
-The key of this modification is the inversion formula
+A practical problem of using homogeneous updating scheme
+is the following.
+%
+A homogeneous updating scheme
+specified by an arbitrary set of numbers
+$\mu_1, \dots, \mu_b$, ($b < n - 1$),
+is not necessarily stable
+with all nonnegative eigenvalues.
+%
+%
+For example, the nearest-neighbor updating scheme
+is unstable if \eqref{eq:nn_stable} is violated.
+%
+Even the Gaussian updating scheme can be unstable
+if the window kernel is truncated.
+%
+Below we show how to minimally modify
+the window kernel function
+to meet the stability criteria.
+%
+\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
+}
+
+The key is the observation of Eq. \eqref{eq:wband_eigenvalue}
+that the eigenvalues are related to the window kernel
+by a cosine transform,
+and thus the relationship can be readily inverted.
+%
+Since we impose $\mu_n = 0$,
+the explicit inversion formula is
 %
@@
             (-1)^k \, \lambda_{k + 1}
-  \right],
+  \right].
   \label{eq:lambdan}
@@
 %
-which ensures that the $\mu_n$
+\note{Eq. \eqref{eq:lambdan}
+ensures that the $\mu_n$
 computed from Eq. \eqref{eq:mu_from_lambda}
 vanishes.
-%
-\note{The inversion formula is derived from the cosine transform.
+
+The inversion formula is derived from the cosine transform.
   %
@@
 }
-Thus, the algorithm of stabilization consists of three steps.
+
+Now given a window kernel, $\mu_1, \dots, \mu_n$,
+we can compute the eigenvalues from \eqref{eq:wband_eigenvalue}.
+%
+By replacing negative eigenvalues by zeros and
+using Eq. \eqref{eq:mu_from_lambda},
+we get a new window kernel
+$\mu^\mathrm{new}_1, \dots, \mu^\mathrm{new}_{n - 1}$,
+which is stable by construction.
+%
+However, the new window may be wider than
+the original one if any of
+$\mu^\mathrm{new}_{b+1}, \dots, \mu^\mathrm{new}_{n-1}$
+is nonzero,
+and if we truncate the window kernel at $b$ bins,
+the stability problem is still existent
+albeit usually less severe
+(in terms of the most negative eigenvalue).
+%
+Thus, to obtain a stable window kernel of width $b$,
+we have to iterate the above process many times.
+The resulting algorithm is the following.
+
 %
@@
   \item
-    Compute the eigenvalues from $\lambda_0, \dots, \lambda_{n-1}$
-    from the given $\mu_0, \dots, \mu_b$ with $b < n$
-    according to Eq. \eqref{eq:wband_eigenvalue}.
+    Given a window kernel $mu_1, \dots, \mu_b$,
+    compute the eigenvalues
+    $\lambda_0, \dots, \lambda_{n-1}$
+	from Eq. \eqref{eq:wband_eigenvalue}.
   \item
-    Replace the negative eigenvalues by zeros.
+    If all eigenvalues are nonnegative,
+    we are done, are break the loop.
   \item
-    Compute the new $\mu_l$ $(0 \le l < n)$ from
-    the modified eigenvalues according to
+    Otherwise, replace negative eigenvalues by zeros,
+    and compute a new window kernel,
+    $\mu_1, \dots, \mu_b$ from
     Eq. \eqref{eq:mu_from_lambda}.
+    Set $\mu_l = 0$ for $l > b$.
+    Go to step 1.
 \end{enumerate}
-%
-Note that, after the stabilization,
-the window is usually widened.
+
+\subsection{\label{sec:cmpschemes}
+Comparison of updating schemes}
+
+\subsubsection{\label{sec:optWL}
+Asymptotic optimality of the single-bin (WL) scheme}
+
+The error functional given in Eq. \eqref{eq:error_asym1}
+leads to an interesting deduction:
+the single-bin scheme is asymptotically optimal.
+%
+Consider a class of $\mathbf w$ matrices
+sharing the same set of eigenvectors,
+hence same $\Gamma_k$'s,
+but with different eigenvalues, $\lambda_k$'s.
+%
+Using the Cauchy-Schwarz inequality, we have,
+for any set of numbers, $c_k \ge 0$,
+%
+\begin{align}
+&
+\left(
+  \int_0^t dt'
+    \sum_{k = 2}^n
+      \Gamma_k \, \dot u_k^2\bigl( q(t') \bigr)
+\right)
+%
+\left(
+  \int_0^t dt'
+    \sum_{k = 2}^n
+      \Gamma_k \, c_k^2
+\right)
+%
+\notag
+\\
+&
+\qquad \qquad
+\ge
+\left(
+  \int_0^t dt'
+    \sum_{k = 2}^n
+      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t') \bigr)
+\right)^2
+\notag
+\\
+&
+\qquad \qquad
+=
+\left(
+  \sum_{k = 2}^n \Gamma_k \, c_k
+    \left[
+      1 - e^{ -\lambda_k q(t) }
+    \right]
+\right)^2.
+\label{eq:CSineq}
+\end{align}
+%
+This inequality follows from the non-negativity of 
+the following quadratic polynomial of $Y$,
+$$
+\int_0^t
+  dt' \sum_{k = 2}^n \Gamma_k \,
+    \left( \dot u_k \, Y - c_k \right)^2
+  \equiv
+  A \, Y^2 + B \, Y + C
+%\ge 0,
+$$
+which implies a non-positive discriminant,
+$B^2 - 4 \, A \, C$.
+%
+Note that the last expression of \eqref{eq:CSineq}
+depends on $q(t')$ only through the end point value,
+which is fixed in the variation process.
+%
+Thus, the inequality sets a lower bound
+of the error $\Err$
+given in Eq. \eqref{eq:error_asym1}.
+%
+The equality is achieved
+if $\dot u_k\left( q(t') \right) = c_k$
+for all $k = 2, \dots, n$ at any $t'$,
+up to a multiplicative factor.
+%
+Solving this set of equations yields
+%
+$$
+\alpha(t') = \frac{              1              }
+                  { \lambda_k \, (t' + t_{k0} ) },
+\quad
+\mathrm{with\;\;}
+t_{k0} = \frac{             t            }
+              { e^{ \lambda_k q(t) } - 1 }.
+$$
+%
+Such a solution is possible only if
+%
+\begin{equation}
+\lambda_2 = \dots = \lambda_n.
+\end{equation}
+%
+\note{We add some steps for the above solution.
+  Integrating $\dot u_k = c_k$ yields
+  \begin{equation}
+  u_k(t') = c_k \left(t' + t_{k0} \right).
+  \label{eq:ui_solution}
+  \end{equation}
+  Given that $u_k(t) = 1$ and $u_k(0) = e^{-\lambda_k \, q(t)}$,
+  we have
+  $$
+  c_k = \frac{ 1 }{ t + t_{k0} },
+  \quad
+  \mathrm{and\;}
+  \frac{ t_{k0} } { t + t_{k0} }
+  =
+  e^{ -\lambda_k \, q(t) }.
+  $$
+  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
+  $$
+  -\lambda_k \, q(t) + \lambda_k \, q(t')
+  = \log c_k + \log\left( t' + t_{k0} \right).
+  $$
+  Differentiating this with respect to $t'$ yields
+  $$
+  \lambda_k \, \alpha(t') = \frac{ 1 } { t' + t_{k0} }.
+  $$
+}
+
+Now an updating matrix $\mathbf w$ satisfying this condition
+is essentially a multiple of the identity matrix
+(since $\lambda_1$ corresponds to the mode
+of a uniform shift of all $v_k$,
+we can freely set it to $\lambda_2$
+without changing the nature of the updating scheme).
+%
+Thus, in terms of asymptotic convergence,
+the single-bin scheme
+(as adopted in the WL algorithm)
+is most efficient among updating schemes.
+%
+In the single-bin case,
+$c_k = 1/(t + t_0)$,
+and the error is
+\begin{align}
+E
+=
+\frac{       t     }
+     { (t + t_0)^2 }
+\sum_{ k = 1 }^n
+  \Gamma_k
+.
+\label{eq:error_singlebin}
+\end{align}
+
+We emphasize that the above result
+holds only under the long time limit,
+and it is not necessarily true
+for finite length simulations.
+%
+For example, consider a system that is initially
+prepared under a fixed $\alpha_0$
+at the beginning of the schedule $\alpha(t)$.
+%
+Then for shorter times, $t \approx 0$,
+the error, given by Eq. \eqref{eq:error_eql},
+is clearly smaller if some eigenvalues
+are less than $1.0$ (the single-bin case).
+%
+In practice, real simulations
+may have long achieved the desired precision
+before reaching the asymptotic regime,
+especially if the updating matrix
+has some $\lambda_k$'s close to $0$.
+
+\subsubsection{\label{sec:equilerr}
+Comparison to the error of the equilibrium FDS method}
+
+As discussed in Sec. \ref{sec:FDS},
+the asymptotic limit of an adaptive FDS method
+is the corresponding equilibrium FDS method
+that fixes the bias potential to the exact value.
+%
+Such a simulation delivers the maximal possible efficiency.
+
+A practical question now arises.
+%
+In late stages of an adaptive FDS simulation,
+will we be better off switching to the equilibrium FDS method
+to avoid possible efficiency loss due to the adaptive updates?
+%
+Below, we shall show that this is unnecessary
+if the simulation uses the single-bin (WL) updating scheme
+with the optimal schedule, Eq. \eqref{eq:alpha_invt}.
+%
+That is, the bias potential obtained from
+an optimal single-bin scheme FDS simulation
+will be similar in precision to that from
+the ideal equilibrium FDS simulation
+in long times.
+%
+Thus, our demonstration below serves as an alternative proof
+of the optimality of the single-bin scheme.
+
+The error of the equilibrium FDS method
+depends on the precision of the histogram.
+%
+The error is minimized
+when $v_i$'s assume the exact values,
+%
+and in this case,
+the error came from the second term
+on the right-hand side of Eq. \eqref{eq:vcorr_equil},
+%
+\begin{align}
+E
+&=
+\left\langle
+  \sum_{ i = 1 }^ n
+    p_i \,
+    \left(
+      \log \frac { \bar h_i }
+                 {      p_i }
+    \right)^2
+\right\rangle
+\notag
+\\
+&\approx
+\sum_{ i = 1 }^ n
+  p_i \,
+  \left\langle
+    \left(
+      \frac { \bar h_i }
+            {      p_i }
+      - 1
+    \right)^2
+  \right\rangle.
+\end{align}
+
+To make better comparison with
+the error of the single-bin case,
+we shall rewrite the error as
+a sum over the eigenmodes.
+We define
+\begin{align}
+\bar \eta_k
+=
+\frac{ 1 } { t }
+\sum_{ t_1 = 1 } ^ t
+  \eta_k( t_1 ),
+\end{align}
+%
+with $\eta_k(t_1)$ defined by Eq. \eqref{eq:eta_def}
+and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
+%\begin{align}
+%\eta_k( t_1 )
+%=
+%\sum_{ i = 1 }^ n
+%\phi_{k i} \left[
+%             \frac{ h_i(t_1) }
+%                  { p_i      }
+%             - 1
+%           \right],
+%\end{align}
+%and
+%\begin{align}
+%\sum_{ k = 1 }^n
+%  \phi_{k i} \, \phi_{k j}
+%  = p_i \, \delta_{i j}.
+%\end{align}
+%
+Then,
+%
+\begin{align}
+E
+=
+\sum_{ k = 1 }^n
+  \left\langle
+    \bar \eta_k^2
+  \right\rangle.
+  \notag
+\label{eq:error_sbin_equil}
+\end{align}
+%
+\note{Because
+%
+\begin{align*}
+\sum_{ k = 1 }^n
+  \bar \eta_k^2
+&=
+\frac{ 1 } { t^2 }
+\sum_{ t_1, t_2 = 1 }^n
+  \sum_{ k = 1 }^n
+    \eta_k( t_1 ) \, \eta_k( t_2 )
+\\
+&=
+\frac{ 1 } { t^2 }
+\sum_{ t_1, t_2 = 1 }^n
+  \sum_{ i, j = 1 }^n
+    \left[
+      \frac{ h_i(t_1) }
+           { p_i      }
+      - 1
+    \right]
+    \left[
+      \frac{ h_j(t_2) }
+           { p_j      }
+      - 1
+    \right]
+    \sum_{ k = 1 }^n
+      \phi_{k \, i} \, \phi_{k \, j}
+\\
+&=
+\frac{ 1 } { t^2 }
+\sum_{ t_1, t_2 = 1 }^n
+  \sum_{ i, j = 1 }^n
+    \left[
+      \frac{ h_i(t_1) }
+           { p_i      }
+      - 1
+    \right]
+    \left[
+      \frac{ h_j(t_2) }
+           { p_j      }
+      - 1
+    \right]
+    p_i \, \delta_{i, j}
+\\
+&=
+\sum_{ i = 1 }^n
+\frac{ 1 } { t }
+\sum_{ t_1 = 1 }^n
+    \left[
+      \frac{ h_i(t_1) }
+           { p_i      }
+      - 1
+    \right]
+\frac{ 1 } { t }
+\sum_{ t_2 = 1 }^n
+    \left[
+      \frac{ h_i(t_2) }
+           { p_i      }
+      - 1
+    \right]
+\\
+&=
+\sum_{ i = 1 }^n
+  p_i \,
+    \left(
+      \frac{ \bar h_i }
+           { p_i      }
+      - 1
+    \right)^2
+.
+\end{align*}
+}
+The sum can start from $k = 2$, for
+by Eq. \eqref{eq:eigenmode1}, we have
+$$
+\eta_1(t_1)
+=
+\sum_{ i = 1 }^n
+  \phi_{1i} \, \frac{ h_i(t_1) - p_i } { p_i }
+=
+\sum_{ i = 1 }^n h_i(t_1)
+- \sum_{ i = 1 }^n p_i = 0.
+$$
+
+In the long-time limit,
+\begin{align*}
+\left\langle
+  \bar \eta_k^2
+\right\rangle
+&=
+\frac{1}{t^2}
+\sum_{ t_1, t_2 = 1 }^n
+\langle \eta_k(t_1) \, \eta_k(t_2) \rangle
+\\
+&\approx
+\frac{1}{t}
+\sum_{ \tau = -\infty }^{ \infty }
+\langle \eta_k(t_1) \, \eta_k(t_1 + \tau) \rangle
+\\
+&=
+\frac{1 }{t}
+\int_{-\infty}^\infty
+\kappa_n(\tau) \, d\tau
+=
+\frac{1}{t}
+\Gamma_k.
+\end{align*}
+%
+Thus,
+\begin{align}
+E
+=
+\frac{ 1 } { t }
+\sum_{ k = 2 }^n \Gamma_k
+\end{align}
+
+Now comparing this with the error of the single-bin scheme
+with the optimal schedule,
+Eq. \eqref{eq:error_singlebin},
+we find that the difference is negligible
+in the long time limit.
 
@@
 %
-Compare with Eqs. \eqref{eq:alpha_limit1} and \eqref{eq:alpha_limit2}.
+As predicted by
+Eqs. \eqref{eq:dinvadt_limit1} and \eqref{eq:dinvadt_limit2},
+the rate of change of $1/\alpha(t)$
+indeed increases from $\lambda_{\min}$
+to $\bar \lambda \approx 1$.
 
@@
     For fair comparison,
-    the $q(t) = \int_0^t \alpha(\tau) \, d\tau$
+    the $q(T) = \int_0^t \alpha(t) \, dt$
     were set to be the same

### 2016-03-20 doc/notes/equilerr.tex
add notes

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{hyperref}
\begin{document}

\title{Error of the static flat-distribution sampling}
\author{}
\date{\vspace{-10ex}}
\maketitle

[The content of this note is absorbed
in Secs. II.A.1 and II.F.2]

\section{Introduction}

There are two types of flat-distribution sampling (FDS) methods.
%
The first type of methods,
referred to as the equilibrium FDS (E-FDS) method hereinafter,
fixes the bias potential, $v_i$, during simulation
and only corrects it in the end.
%
Examples include the original
entropic or multicanonical sampling and simulated tempering.
%
The main difficulty is that if a sufficiently accurate bias potential
is unknown before simulation,
it can be hard to achieve the desired flat distribution.

For this reason, the second type of methods,
exemplified by the Wang-Landau (WL) algorithm and metadynamics,
are invented.
%
These methods construct the bias potential on the fly
and achieve a flat distribution in the meantime.
%
Although powerful, they are less well understood,
in that the adaptive updates to the bias potential
breaks the microscopic reversibility
of the underlying sampling mechanism.
%
Thus, for a conservative practitioner,
the E-FDS methods again become appealing
once a sufficiently accurate bias potential is available.
%
Further, with the bias potential fixed to the exact value,
the equilibrium methods can be thought as
the asymptotic limit of the adaptive methods upon convergence,
and thus deliver the maximal possible efficiency.

A practical question now arises.
%
In later stages of an adaptive FDS simulation,
will we be better off switching to the equilibrium FDS
in order to avoid possible efficiency loss due to the adaptive updates?
%
Below, we shall show that it is unnecessary to do so
if the simulation is using the single-bin (WL) updating scheme
with the optimal schedule for controlling the updating magnitude.
%
That is, the bias potential obtained from
an optimal single-bin scheme FDS simulation
will be similar in precision to that from
an E-FDS simulation in long times.
%
Thus, our demonstration below serves as an alternative proof
of the optimality of the single-bin scheme.

\section{Error}

Here is a quick review of E-FDS.
%
In E-FDS, we fix the bias potential, $v_i$, during simulation,
and collect a histogram, $\bar h_i$, via trajectory averaging:
%
\begin{align}
\bar h_i
=
\frac 1 t
\sum_{ t_1 = 1 }^{ t }
  h_i( t_1 )
=
\frac 1 t
\sum_{ t_1 = 1 }^{ t }
  \delta_{ i, i(t_1) },
\label{eq:hbar_def}
\end{align}
%
where $h_i(t_1) = \delta_{ i, i(t_1) }$
is the instantaneous histogram.
%
In the end,
we use this histogram to correct the bias potential as
%
\begin{align}
v_i^\mathrm{corr}
=
v_i
+
\log \frac { \bar h_i }
           {      p_i }.
\end{align}

The error of the method depends on the precision of the histogram.
%
Below we shall evaluate the error in the ideal case
in which $v_i$'s assume the exact values.
Then,
%
\begin{align}
E
=
\left\langle
  \sum_{ i = 1 }^ n
    p_i \,
    \left(
      \log \frac { \bar h_i }
                 {      p_i }
    \right)^2
\right\rangle
\approx
\sum_{ i = 1 }^ n
  p_i \,
  \left\langle
    \left(
      \frac { \bar h_i }
            {      p_i }
      - 1
    \right)^2
  \right\rangle.
\end{align}

To make better comparison with
the error of the single-bin case,
we shall rewrite the error as
a sum of the errors of the eigenmodes.
We define
\begin{align}
\bar \eta_k
=
\frac{ 1 } { t }
\sum_{ t_1 = 1 } ^ t
  \eta_k( t_1 ),
\end{align}
%
with
\begin{align}
\eta_k( t_1 )
=
\sum_{ i = 1 }^ n
\phi_{k i} \left[
             \frac{ h_i(t_1) }
                  { p_i      }
             - 1
           \right],
\end{align}
and
\begin{align}
\sum_{ k = 1 }^n
  \phi_{k i} \, \phi_{k j}
  = p_i \, \delta_{i j}.
\end{align}
%
Then,
%
\begin{align}
E
=
\sum_{ k = 1 }^n
  \left\langle
    \bar \eta_k^2
  \right\rangle,
\label{eq:error_sbin_equil}
\end{align}
%
because
%
\begin{align*}
\sum_{ k = 1 }^n
  \bar \eta_k^2
&=
\frac{ 1 } { t^2 }
\sum_{ t_1, t_2 = 1 }^n
  \sum_{ k = 1 }^n
    \eta_k( t_1 ) \, \eta_k( t_2 )
\\
&=
\frac{ 1 } { t^2 }
\sum_{ t_1, t_2 = 1 }^n
  \sum_{ i, j = 1 }^n
    \left[
      \frac{ h_i(t_1) }
           { p_i      }
      - 1
    \right]
    \left[
      \frac{ h_j(t_2) }
           { p_j      }
      - 1
    \right]
    \sum_{ k = 1 }^n
      \phi_{k \, i} \, \phi_{k \, j}
\\
&=
\frac{ 1 } { t^2 }
\sum_{ t_1, t_2 = 1 }^n
  \sum_{ i, j = 1 }^n
    \left[
      \frac{ h_i(t_1) }
           { p_i      }
      - 1
    \right]
    \left[
      \frac{ h_j(t_2) }
           { p_j      }
      - 1
    \right]
    p_i \, \delta_{i, j}
\\
&=
\sum_{ i = 1 }^n
\frac{ 1 } { t }
\sum_{ t_1 = 1 }^n
    \left[
      \frac{ h_i(t_1) }
           { p_i      }
      - 1
    \right]
\frac{ 1 } { t }
\sum_{ t_2 = 1 }^n
    \left[
      \frac{ h_i(t_2) }
           { p_i      }
      - 1
    \right]
\\
&=
\sum_{ i = 1 }^n
  p_i \,
    \left(
      \frac{ \bar h_i }
           { p_i      }
      - 1
    \right)^2
.
\end{align*}
%
The sum can start from $k = 2$,
because by $\phi_{1i} = p_i$,
we have
$$
\eta_1(t_1)
=
\sum_{ i = 1 }^n
  \phi_{1i} \, \frac{ h_i(t_1) - p_i } { p_i }
=
\sum_{ i = 1 }^n h_i(t_1)
- \sum_{ i = 1 }^n p_i = 0.
$$

Finally, we shall evaluate Eq. \eqref{eq:error_sbin_equil}
in the long-time limit:
\begin{align*}
\left\langle
  \bar \eta_k^2
\right\rangle
&=
\frac{1}{t^2}
\sum_{ t_1, t_2 = 1 }^n
\langle \eta_k(t_1) \, \eta_k(t_2) \rangle
\\
&\approx
\frac{1}{t}
\sum_{ \tau = -\infty }^{ \infty }
\langle \eta_k(t_1) \, \eta_k(t_1 + \tau) \rangle
\\
&=
\frac{1 }{t}
\int_{-\infty}^\infty
\kappa_n(\tau) \, d\tau
=
\frac{1}{t}
\Gamma_k.
\end{align*}
%
Thus,
\begin{align}
E
=
\frac{ 1 } { t }
\sum_{ k = 2 }^n
\Gamma_k
\end{align}

Now comparing this with the error of the single-bin scheme
with the optimal schedule,
$$
E_{\mathrm{sb}}
=
\frac{    t    }
     { (t + t_0)^2 }
\sum_{ k = 2 }^n
  \Gamma_k,
$$
we find that the difference is negligible in the long time limit.

\end{document}

### 2016-03-21 doc/invt.tex
small edits to invt.tex

@@
 Thus, we have an implicit equation for $q(t)$,
-and hence the optimal schedule, $\alpha(t)$.
+and hence the optimal schedule,
+$\alpha(t) = \dot q(t)$.
 
@@
 %
-Since in the asymptotic regime,
-the distribution converges to $p_i$
-no matter the intrinsic distribution, $p^*_i$,
-we also set the latter to be flat for convenience.
-%
-Our goal is to observe how the error
+%Since in the asymptotic regime,
+%the distribution converges to $p_i$
+%no matter the intrinsic distribution, $p^*_i$,
+%we also set the latter to be flat for convenience.
+For convenience,
+the intrinsic distribution, $p^*_i$,
+is a set to be flat as well.
+%
+Our goal is to observe how the fluctuating error
 of the bias potential defined in Eq. \eqref{eq:error_sum}
-depends on the schedule function $\alpha(t)$
+depends on the schedule, $\alpha(t)$,
 at the end of the simulation.
@@
 %
-and let the system undergo a period of equilibration
-during which the updating magnitude is fixed at $\alpha_0$.
-%
-After the equilibration,
+and let the system undergo an equilibration period
+of constant updating magnitude,
+$\alpha(t) = \alpha_0$.
+%
+After this,
 we reset the origin of the time, $t$,
 and started the schedule specified by
+either
+Eq. \eqref{eq:q_opt}
+or
 Eq. \eqref{eq:alpha_invtlambda}.
 %
-The goal is then to observe how the final error
-depends on the parameter, $\lambda$.
-%
-Further, we choose $t_0 = 1/(\lambda \, \alpha_0)$
+We use the latter case to verify our theory.
+%
+We wish to see how the final error
+depends on the parameter, $\lambda$,
+and if it matches the predicted value.
+%
+Here, we used $t_0 = 1/(\lambda \, \alpha_0)$
 to smooth the transition from equilibration.
@@
 we have applied the above protocol to two
-sampling schemes.
+sampling schemes,
+a global one and a local one.
 %
@@
 $
-The difference is that,
+However,
 in the first sampling scheme,
-we choose $j$ randomly out of the $n$ bins
-to emulate the perfect sampling scheme
-discussed in Sec. \ref{sec:Gamma}.
-%
-We refer to this as the global sampling scheme.
+we choose the destination $j$
+randomly out of the $n$ bins.
+%
+This sampling scheme,
+referred to this as the global sampling scheme,
+emulates the perfect sampling scheme
+discussed in Sec. \ref{sec:Gamma}
+in the long time limit.
 %
 In the second scheme,
-we limit $j$ to the two neighbors of $i$,
+we limit the destination $j$
+to the two neighbors of $i$,
 $i \pm 1$, with equal probability $1/2$,
@@
 %
-This scheme models local sampling processes
+This scheme,
+referred to this as the local sampling scheme,
+models local sampling processes
 typical in MD and straightforward MC simulations.
 %
-We refer to this as the local sampling scheme,
 
@@
 
-Our final observation is that the error functions
-in the two cases shared the same shape,
-differing only by a multiplicative constant.
+Another observation is that the two error functions
+differed only by a multiplicative constant.
 %
@@
 %
-Since all eigenvalues of the single-bin scheme were
+Since all eigenvalues of the single-bin scheme are
 the same, $\lambda_2 = \cdots = \lambda_n = 1$,
@@
 
-We verified the optimality of the inverse-time formula
-for the Wang-Landau algorithm.
-
-We have shown that
-without a priori information,
-the single-bin scheme is asymptotically
-the best among all updating scheme
-in terms of convergence.
-
-For metadynamics, and other multiple-bin updating schemes,
-we found formulas to predict the error for finite length
-simulations and the optimal relative updating magnitude,
-$1/\lambda$.
+In conclusion,
+we have proposed a method of computing
+the optimal schedule of the updating magnitude
+for a general adaptive FDS simulation.
+%
+Adaptive FDS methods differ by the updating schemes,
+which specify how the bias potential is modified
+in each updating step.
+%
+In the single-bin or WL case,
+the update is limited to the current bin;
+while in the metadynamics case,
+the bias potential of a few neighboring bins to the current one
+is updated according to a Gaussian wave packet.
+
+By this method,
+we have verified the optimality of
+the inverse-time schedule,
+Eq. \eqref{eq:alpha_invt},
+for the single-bin (WL) case.
+%
+However,
+the optimal schedule,
+given by the solution of Eq. \eqref{eq:q_opt},
+is algebraically more involved
+for a general multiple-bin scheme,
+including the Gaussian one used in metadynamics.
+%
+In the latter case,
+both the optimal schedule and the resulting error
+are sensitive to the simulation length.
+
+We have also compared different updating schemes.
+%
+We have shown that in the long time limit
+the single-bin scheme belongs to a class of
+optimal updating schemes.
+%
+That is, the single-bin scheme
+is asymptotically as efficient as
+an equilibrium FDS simulation
+with the exact bias potential.
 

### 2016-03-25 prog/invtpar.h
implement reflective boundary conditions for MD

@@ -150,9 +150,9 @@ static void invtpar_init(invtpar_t *m)
   m->fngamma[0] = '\0';
 
   /* molecular dynamics parameters */
-  m->mddt = 0.01;
+  m->mddt = 0.005;
   m->tp = 1.0;
-  m->thermdt = m->mddt;
+  m->thermdt = 0.1;
 #if 0
   m->dwa = 1.0;
   m->dwb = 2.0;
@@ -200,9 +200,9 @@ static void invtpar_mkgauswin(invtpar_t *m)
     m->winn = m->n;
   }
 
-  /* truncate the Gaussian at 5 * sigma */
-  if ( m->winn >= sig * 5 ) {
-    m->winn = (int) (sig * 5 + 0.5);
+  /* truncate the Gaussian at 10 * sigma */
+  if ( m->winn >= sig * 10 ) {
+    m->winn = (int) (sig * 10 + 0.5);
     if ( m->winn < 1 ) {
       m->winn = 1;
     }
@@ -317,11 +317,6 @@ static void invtpar_compute(invtpar_t *m)
   /* set the initial cutoff wave number */
   if ( m->kcutoff <= 0 ) {
     m->kcutoff = m->n;
-  }
-
-  /* currently, turn on PBC for MD automatically */
-  if ( m->sampmethod == SAMPMETHOD_MD ) {
-    m->pbc = 1;
   }
 
   if ( m->localg > 0.5 ) {

### 2016-03-25 doc/invt.tex
implement reflective boundary conditions for MD
invt.tex: edit abstract and introduction

@@
   %
-  in the WL case,
-  each update applies to only
+  the WL algorithm updates only
   the bin containing the current $z$,
-  while in the metadynamics case,
-  a Gaussian wave packet
-  involving several neighboring bins
-  is used.
-  %
-  The asymptotic error of the resulting PMF
+  while metadynamics updates also
+  several neighboring bins according to
+  a Gaussian wave packet.
+  %
+  The error of the resulting PMF
   is determined by the updating schedule,
@@
   for relatively short simulations,
-  the WL algorithm is best for asymptotic convergences.
+  the WL algorithm is best for asymptotic convergence.
   %the optimal schedule
@@
 %
-Usually, the distribution $p^*(z)$ is
-localized in one or a few small $z$ regions.
+Normally, the distribution $p^*(z)$ is
+limited in one or a few narrow $z$ regions.
 %
@@
 along $z$\cite{mezei1987, berg1992, lee1993},
-and then piece together the information.
+and then piece together the entire curve.
 %
 This solution is clumsy in practice
-and often suffers from a convergence problem
+and often fails to converge quickly
 for a complex and glassy system,
@@
 
-A better solution is to artificially alter
-the target distribution along $z$
-to a flat one\cite{mezei1987, berg1992, lee1993,
+A better solution is to artificially flatten
+the target distribution along $z$\cite{mezei1987, berg1992, lee1993,
 wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
@@
 techniques for this purpose,
-in which a bias potential is actively built-up on-the-fly
+in which a bias potential is actively built up on-the-fly
 during simulation.
@@
 These techniques employ updating loops
-that discourage the system
-to sample configurations
-sharing the current $z$ value
+that discourage future visits to configurations
+with the current $z$ value
 by elevating the bias potential here.
 %
-The main difference of the two
+The main difference between the two
 is that in the WL case,
@@
 while metadynamics adopts an extended
-updating wave packet
+Gaussian wave packet
 covering several neighboring bins.
@@
 that the optimal magnitude of updating $\ln f$
-is given by the
+is given by the product of the
 inverse time\cite{liang2007,
 belardinelli2007, belardinelli2007jcp, belardinelli2008,
-morozov2007, zhou2008},
-where ``time'' means
-the number of simulation steps
-divided by the number of bins.
-%
-The optimal schedule for metadynamics,
+morozov2007, zhou2008}
+(with ``time'' being
+the number of simulation steps)
+and the number of bins.
+%
+The optimal schedule for the metadynamics case,
 however, is less studied.
@@
 %
-Although the method recovers the inverse-time formula
-for the WL case,
-it predicts, for metadynamics,
+The method reproduces the inverse-time formula
+for the WL case.
+%
+It also predicts, for metadynamics,
 a more algebraically complex schedule,
@@
 flat-distribution sampling
-and the development on the optimal schedule.
+and some known aspects on the optimal schedule.
 %
@@
 on the simplest case of
-single-bin updating scheme.
+single-bin (WL) updating scheme.
 %
@@
 consider the problem of computing
-the distribution, $p^*_i$
+the distribution, $p^*_i$,
 along a discrete quantity $i$.
@@
 For example, $i$ can be the energy $E$
-in lattice spin models; or the temperature index
+in a lattice spin model; or the temperature index
 in a simulated tempering simulation.
@@
 The first type,
-referred to as the equilibrium FDS % (E-FDS) methods hereinafter,
+referred to as the equilibrium FDS, % (E-FDS) methods hereinafter,
 estimates the bias potential, $V_i$,
+before simulation,
 and fixes it during simulation.
@@
 %
-The estimated bias potential, however,
-can often be improved in the end of simulation.
-%
-To do so,
-we accumulate a histogram along $i$
-during simulation and use it for $\pi_i$ in
-Eq. \eqref{eq:pi_p_phi1}.
-%
-This allows us to correct $V_i$ as
-%
-\begin{equation}
-V_i^{\mathrm{corr}}
-=
-V_i^{\mathrm{}}
+The estimated bias potential, $V_i$,
+can often be improved
+by the histogram along $i$ accumulated
+during simulation
+Using the histogram as $\pi_i$ in
+Eq. \eqref{eq:pi_p_phi1}
+allows us to correct $V_i$ as
+%
+\begin{equation}
+V^\mathrm{corr}_i
+=
+V_i
 +
-\log \frac{ \pi_i } { p_i }.
+\log \frac{ \pi_i }
+          {   p_i }.
 \label{eq:vcorr_equil}
@@
 %
+\note{This follows from
+  $$
+  \begin{aligned}
+    \pi_i &\propto p^*_i \, e^{-V_i}, \\
+    p_i &\propto p^*_i \, e^{-V^\mathrm{corr}_i}.
+  \end{aligned}
+  $$
+}
 For an unknown system,
@@
 Thus,
-it usually takes several iterative rounds of simulations
+it can take many iterative rounds of simulations
 to achieve the desired flat distribution.
@@
 These methods update the bias potential more frequently
-and are more capable of achieving a flat distribution.
+and can easily achieve a flat distribution.
 %
 However, they are less well understood theoretically,
-the updating loops used there
+as the updating loops used there
 break the microscopic reversibility
@@
 If the bias potential were converged to the exact value,
-the equilibrium simulation can be thought as
+this particular equilibrium simulation can be thought as
 the optimal asymptotic limit of the adaptive counterpart,
-delivering the maximal possible efficiency.
+delivering the maximal efficiency.
 
@@
 $v_i$ should approach a constant of $i$
-upon convergence (i.e., when $\pi_i \approx p_i$).
+upon convergence (i.e., when $\pi_i \approx p_i$)
+of an adaptive FDS.
 

### 2016-03-25 doc/invt.tex
invt.tex: editing theory A-C.

@@
 \begin{equation}
-  \pi_i =
+  \pi_i
+  =
   \frac{                p^*_i \, e^{-V_i} }
-  { \sum_{j = 1}^n p^*_j \, e^{-V_j} }.
+       { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
+  .
   \label{eq:pi_p_phi}
@@
 In the WL algorithm\cite{wang2001, wang2001pre},
-$v_i$ are updated
-in each MC step $t$,
-and $v_i(t)$ as a function of $t$
-is updated as
+the bias potential, $v_i(t)$, is updated
+in each MC step $t$ as
 %
@@
 the distribution collected from
-the trajectory is identical to $p_i$.
-%
-However, the above updates
+the trajectory is identical to $p_i$ in long times.
+%
+However, as the above updates
 change $v_i(t)$ continuously,
-and Eq. \eqref{eq:pi_p_v} no longer holds
+Eq. \eqref{eq:pi_p_v} no longer holds
 at all times.
@@
 for a period of time,
-which is referred to as a stage below.
+referred to as a stage below.
 %
@@
 %
-However, in later stages, it tends to reduce $\alpha$
-too quickly, making the asymptotic error
+However, it tends to reduce $\alpha$
+too quickly in late stages, making the asymptotic error
 saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
 %
-Similarly, for metadynamics,
-there are various ways of reducing the updating magnitude,
-particularly in the development of
-self-healing umbrella sampling
-and well-tempered metadynamics\cite{
-  marsili2006, barducci2008, dickson2011, dama2014}.
 
@@
 
-For the single-bin scheme
-used in WL algorithm,
+%For the single-bin scheme
+%used in WL algorithm,
+Later,
 it was found that
-a more effective way
+a better way
 of updating the schedule $\alpha(t)$
@@
 %
-This surprisingly simple formula has attracted
+This simple yet effective formula has attracted
 several studies\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008,
@@
 
-%Here we shall derive the optimal $\alpha(t)$
-%For the WL algorithm, Eq. \eqref{eq:wl_update}.
-%%
-%Further, we shall generalize the result
-%To a class of multiple-bin schemes,
-%In which an update affects
-%Not only the current bin $i(t)$,
-%But also a few neighboring ones.
-%%(such schemes are designed
-%%to maintain a smooth bias potential).
-%%
-%We shall show that
-%The constant of proportionality, $1$,
-%In Eq. \eqref{eq:alpha_invt}
-%Is optimal for the single-bin scheme,
-%Eq. \eqref{eq:wl_update}
-%(as in the WL algorithm),
-%But not necessarily so
-%For a multiple-bin scheme.
-%%
-%Below, we discuss the single-bin scheme
-%In Section \ref{sec:single-bin},
-%And the multiple-bin case
-%In Section \ref{sec:multiple-bin}.
+As Eq. \eqref{eq:alpha_invt} was being developed for the WL algorithm,
+several similar ways of
+reducing the updating magnitude
+were developed for umbrella sampling and metadynamics,
+particularly the
+self-healing umbrella sampling
+and well-tempered metadynamics\cite{
+  marsili2006, barducci2008, dickson2011, dama2014}.
+%
+Our aim below is to the optimal schedule $\alpha(t)$,
+for a general updating scheme.
 
@@
 To do so,
-we shall first express the error of $v(t)$
+we shall first express the error of
+the shifted bias potential, $v_i(t)$,
 as a functional of $\alpha(t)$,
@@
 %
-Here, the deterministic part can be related
-to the ``ensemble average'' of $h_i$.
-%
-The ensemble consists of many similar simulation copies
+The deterministic part can be related
+to the ``ensemble average'' of $h_i$,
+where the ensemble consists of many similar simulation copies
 that have experienced the same schedule $\alpha(t)$
@@
   \approx
-  \pi_i
-  =
-  \frac{ p_i \, e^{-v_i} }
-  { \sum_{j = 1}^n p_j \, e^{-v_j} }.
+  \pi_i(t)
+  =
+  \frac{                p_i \, e^{-v_i(t)} }
+       { \sum_{j = 1}^n p_j \, e^{-v_j(t)} }.
   \label{eq:h_ave}
@@
   \sigma_{ij}(t - t'),
-  \label{eq:zeta_zeta_correlation}
+  \notag
+  %\label{eq:zeta_zeta_correlation}
 \end{equation}
@@
 %
-to be small in the asymptotic regime,
-and expand Eq. \eqref{eq:h_ave} as
+to be small in the asymptotic regime.
+Expanding Eq. \eqref{eq:h_ave} in $x_i$ yields
 %
@@
 %
-Fortunately, we shall find that it suffices to consider
+Fortunately, we find that it suffices to consider
 a simpler one-variable problem
@@
 \kappa(t - t').
-\label{eq:noise_correlation}
+\notag
+%\label{eq:noise_correlation}
 \end{equation}
@@
   \lim_{t \rightarrow \pm\infty} \kappa(t) = 0.
-  \label{eq:kappat_limit}
+  \label{eq:kappa_limit}
 \end{equation}
@@
   \dot u\bigl( q(t) \bigr) \, \xi(t) \, dt,
-\label{eq:xt_solution}
+\notag
+%\label{eq:xt_solution}
 \end{equation}
@@
 q(t) \equiv \int_0^t \alpha(t') \, dt',
-\label{eq:qt_definition}
+\notag
+%\label{eq:q_def}
 \end{equation}
@@
 e^{-\lambda \, q(T) + \lambda \, q'}.
-\label{eq:ut_definition}
+\label{eq:u_def}
 \end{align}
@@
 
-We now fix the endpoint value of $q(T)$
-and variate the curve.
+We now fix the endpoint value $q(T)$
+and variate the curve, $q(t)$ ($0 < t < T$).
 %
 This is equivalent to variating $u$
-with fixing endpoint values,
+with fixed endpoint values,
 $u\bigl( q(0) \bigr)  = e^{-\lambda \, q(T)}$
@@
 where we have dropped the boundary terms
-by using Eq. \eqref{eq:kappat_limit}.
+by using Eq. \eqref{eq:kappa_limit}.
 %
@@
 %
-Using Eq. \eqref{eq:ut_definition},
+Using Eq. \eqref{eq:u_def},
 we get
 $$
-e^{-\lambda \, q(t) + \lambda \, q(t')}
+e^{-\lambda \, q(T) + \lambda \, q(t')}
 =
@@
 \end{equation}
-agreeing with Eq. \eqref{eq:alpha_invt}.
+and Eq. \eqref{eq:alpha_invt}
+is the special case of $t_0 = 0$.
 
@@
 upon a visit to bin $j$,
-we update not only the bias potential there,
-but also those for a few neighboring bins $i$.
+we update not only the bias potential at bin $j$,
+but also those at a few neighboring bins, $i$'s.
 %
@@
 %It is worth pointing out
-Further, in a stable updating scheme,
-all eigenvalues $\lambda_k$ of $\mathbf w$
-must be nonnegative,
-for otherwise there is a fluctuating mode
-that can increase indefinitely.
+Finally, there is a stability condition.
+%
+In a \emph{stable} updating scheme,
+no fluctuating mode can increase indefinitely,
+which means all eigenvalues $\lambda_k$ of $\mathbf w$
+must be nonnegative.
 
@@
 -\alpha(t) \, \lambda_k \, [y_k(t) - \eta_k(t)],
-\label{eq:yt_diffeq}
+\notag
+%\label{eq:yt_diffeq}
 \end{equation}
@@
 %
-By Eq. \eqref{eq:eigenmode1},
+From Eq. \eqref{eq:eigenmode1},
 the first eigenmode, $y_1 = \sum_{j=1}^n p_j \, x_j$,
@@
 %
-Thus, we can start the sum in Eq. \eqref{eq:y2_sum}
+Thus, we can start the last sum in Eq. \eqref{eq:y2_sum}
 from $k = 2$.
@@
 we have, from Eq. \eqref{eq:y2_sum},
-the errors at a period of $T$,
+the error after a period of $T$,
 %
@@
   \Err_R + \Err_A,
-  \label{eq:error_split}
+  \notag
+  %\label{eq:error_split}
 \end{equation}
@@
 
-In the asymptotic regime,
+In the asymptotic regime
+(i.e., if the simulation length, $T$, is sufficiently long),
 we can approximate $\kappa_k(t)$ as
@@
 %
-
 \note{The next order correction to Eq. \eqref{eq:kappa_delta}
@@
 }
-
+%
 \note{
@@
 }
+In this way,
+the underlying sampling process
+affects the error only through the set of numbers, $\Gamma_i$'s.
 
@@
 
-We can evaluate the values of $\Gamma_k$'s
+We evaluate the integrals of the autocorrelation functions,
+$\Gamma_k$'s,
 in two special cases.
 %
-If the sampling is perfect,
+First, if the sampling is perfect,
 Eq. \eqref{eq:kappa_delta} is exact, and
@@
 In this case,
-the bin index $k$ is an independent random variable
+the bin index $i$ is an independent random variable
 at each time step $t$, and
@@
 %
-where $\delta(t - t')$ is understood as an adaption of $\delta_{t, t'}$
-from the discrete time case to the continuous limit. % hereinafter.
+where $\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
+in the continuous limit.
 %
@@
 %
-Thus, from Eqs. \eqref{eq:eig_orthonormal_rows} and \eqref{eq:eta_def},
-we get, for $k \ge 2$,
+Thus,
+for $k \ge 2$, we get
 %
@@
 \end{equation}
+%
+where we have used
+Eqs. \eqref{eq:eig_orthonormal_rows},
+\eqref{eq:eigenmode1},
+\eqref{eq:eta_def},
+and
+$$
+  \sum_{j = 1}^n \phi_{kj}
+  =
+  \sum_{j = 1}^n \frac{ \phi_{kj} \, \phi_{1j} } { p_j }
+  =
+  \delta_{k1}
+  =
+  0.
+$$
 %
@@
   on the last line.
-  %
-  Note for $k \ge 2$,
-  $$
-  \sum_{j = 1}^n \phi_{kj}
-  =
-  \sum_{j = 1}^n \frac{ \phi_{kj} \, \phi_{1j} } { p_j }
-  =
-  \delta_{k1}
-  =
-  0.
-  $$
-  as the mode $1$, or the first row, $\phi_{1j}$,
-  is given by the equilibrium distribution $p_j$.
 }
@@
 in each step, the system hops to the previous bin $i - 1$
-or the next bin $i + 1$ with equal probability $\frac 1 2$,
-ignoring out-of-boundary transitions.
+or the next bin $i + 1$ with equal probability $1/2$.
 %
@@
   \begin{array}{cccccccc}
-    \frac 1 2 & \frac 1 2 & 0 & \dots & 0 \\
+    \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
     \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
@@
     0 & \dots & \frac 1 2 & 0  & \frac 1 2 \\
-    0 & \dots & 0 & \frac 1 2 & \frac 1 2
+    \frac {1 - \sigma} 2 & \dots & 0 & \frac 1 2 & \frac \sigma 2
   \end{array}
-\right).
+\right),
 \label{eq:T_nn}
 \end{equation}
+%
+where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
 %
@@
 we can further assume that
-the eigenvectors of $\mathbf w$
-coincide with those of $\mathbf Q$,
-while the eigenvalues may differ.
+the matrices $\mathbf Q$ and $\mathbf w$
+share eigvenvectors, but not eigenvalues.
 %
@@
 $$
-\Lambda_k = 1 - 2 \sin^2 \frac{ (k - 1) \, \pi } { 2 \, n }.
-$$
+\Lambda_k
+=
+1 -
+2 \sin^2 \frac{ (k - 1) \, g \, \pi }
+              {       2 \, n        }
+,
+$$
+where
+$g = 2$ for a periodic variable, or $1$ otherwise.
 %Note that these $\Lambda_k$'s are
@@
 Thus,
-$\Gamma_k$ is roughly twice
-the autocorrelation time,
-and from Eq. \eqref{eq:Gamma_sum}, we get
+%$\Gamma_k$ is roughly twice the autocorrelation time, and
+from Eq. \eqref{eq:Gamma_sum}, we get
 %
@@
 %
-We can use this case to model local sampling processes.
+This case is intended to model a local sampling process.
 
@@
 \subsubsection{\label{eq:eqlerr}
-Equilibrium error under a fixed updating magnitude
+Equilibrium error under a constant updating magnitude
 }

### 2016-03-29 doc/invt.tex
invt.tex: updates theory

@@
 \newcommand{\Err}{E}
+\newcommand{\ii}{\mathrm{i}}
 
@@
   for relatively short simulations,
-  the WL algorithm is best for asymptotic convergence.
+  the single-bin updating scheme
+  used in the WL algorithm
+  belongs to a class of updating schemes
+  optimal for asymptotic convergence.
   %the optimal schedule
@@
 %
-This solution is clumsy in practice
-and often fails to converge quickly
+This solution can be clumsy in practice
+and it often fails to converge quickly
 for a complex and glassy system,
@@
 with the current $z$ value
-by elevating the bias potential here.
+by elevating the bias potential there.
 %
@@
 For the WL algorithm, it is known
-that the optimal magnitude of updating $\ln f$
+that the optimal updating magnitude, $\ln f$,
 is given by the product of the
@@
 morozov2007, zhou2008}
+and the number of bins
 (with ``time'' being
-the number of simulation steps)
-and the number of bins.
+the number of simulation steps).
 %
@@
   \Err_R + \Err_A,
-  \notag
-  %\label{eq:error_split}
+  \label{eq:error_split}
 \end{equation}
@@
 Before addressing the optimal $\alpha(t)$,
-let us first compute the equilibrium value of $\Err$
-under a fixed schedule $\alpha(t) \equiv \alpha_0$.
-%
-In this case, we drop $\Err_R$, and
-the error at the end of a long period $T'$
-is given by
+we shall first consider a constant schedule
+$\alpha(t) \equiv \alpha_0$,
+and compute the equilibrium error, $\Err$,
+in this case.
+%
+After a long period $T'$,
+the residual error becomes negligible, and
 %
@@
   =
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0,
+  \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0.
   \label{eq:y2_eql}
@@
 
-Below we shall assume that before starting
-the schedule $\alpha(t)$ to be optimized,
+We shall use this result to model the initial error
+before starting the schedule $\alpha(t)$ to be optimized.
+%
+That is, for $t < 0$, we assume
+$\alpha(t) = \alpha_0$, and
 the system has entered the equilibrium
-under a fixed $\alpha_0$.
-%
-That is, for $t \le 0$,
-$\alpha(t) = \alpha_0$.
+under a fixed $\alpha_0$ at $t = 0$.
 %
@@
 %
-In this case,
+Then,
 the residual error at the end of period $T$
@@
 %
-It follows that the residue error
+Note that the residual error
 depends on the schedule $\alpha(t)$
 only through endpoint value of
-the integral $q(T)$.
+the integral $q(T) = \int_0^T \alpha(t) \, dt$.
 %
 Thus, if we variate the schedule under a fixed $q(T)$,
-then this part can be ignored,
+this residual part can be ignored,
 and the optimization can be done exclusively
@@
 
-As shown above, with fixed endpoint value of $q(t)$,
-the optimization of the error is equivalent to
-that of the asymptotic error.
-%
-We now rewrite Eq. \eqref{eq:error_asym1} as
+As shown above, with fixed endpoint value, $q(T)$,
+the optimization of the total error is equivalent to
+that of the asymptotic part.
+%
+We can rewrite Eq. \eqref{eq:error_asym1} as an action
 %
@@
 %
-where, we have defined the Lagrangian,
+where the Lagrangian,
 %
@@
     \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
-  \; \dot q^2( t ).
+  \; \dot q^2( t ),
   \label{eq:error_asym_lagr}
@@
 %
-Note that the Lagrangian
+represents a free particle
+with a position-dependent mass.
+%
+Since the Lagrangian
 does not explicitly depend on time, $t$,
-so the Hamiltonian,
+the Hamiltonian,
 %
@@
 %
-is conserved,
-and coincides with the Lagrangian.
-%
-So we may write
+which coincides with the Lagrangian,
+is conserved.
+%
+So we may set
 %
@@
 %
-where $c$ is determined by
-the boundary condition at $t = T$.
+where $c$ is to be determined by
+the boundary condition at $t = T$ as
+%
+\begin{equation}
+  c =
+  \frac 1 T
+  \int_{ 0 }^{ q(T) }
+    \sqrt{
+      \textstyle\sum_{ k = 2 }^n
+        \Gamma_k \, \lambda_k^2
+        \, u_k^2( q' )
+    }
+    \;
+    d q'
+  .
+  \notag
+  %\label{eq:q_opt}
+\end{equation}
 %
@@
 
-Although Eq. \eqref{eq:q_opt} is hard to solve explicitly,
-we can obtain an expression regarding $\alpha(t)$
+Unfortunately,
+Eq. \eqref{eq:q_opt} does not provide much information
+on the behavior of $\alpha(t)$,
+and the explicit solution
+can be formidable in the general case.
+%
+We can nonetheless obtain an expression regarding $\alpha(t)$
 by differentiating Eq. \eqref{eq:Lagrangian_const},
@@
   }
+  \equiv
+  \lambda(t)
   .
@@
 %
-The right-hand side is a weighted average
+The expression in the middle is a weighted average
 of the eigenvalues, $\lambda_k$'s,
-and we shall refer to this as the instantaneous
+and we shall refer to it as the instantaneous
 eigenvalue, $\lambda(t)$.
+%
+The instantaneous eigenvalue
+gives the rate of change of
+the inverse updating magnitude.
 
 For the single-bin scheme, $\lambda_k = 1$,
-and the instantaneous eigenvalue is also unity.
+and the instantaneous eigenvalue is always unity.
 Thus,
@@
 For a general multiple-bin scheme,
-we consider two limiting cases.
-%
-First, in the beginning of a long simulation,
-we have $t \ll T$ and $q(T) - q(t) \to \infty$,
-the sums on the right-hand side of Eq. \eqref{eq:dinvadt}
-is dominated by the term with the largest
-$u_k = e^{ \lambda_k \, [q(t) - q(T)] }$,
-or, equivalently, with the smallest nonzero
-$\lambda_k$.
-Then
-\begin{equation}
-  \frac{ d   }
-       { d t }
-  \left[
+we can characterize the solution of
+Eq. \eqref{eq:dinvadt} as
+\begin{equation}
+  \alpha(t)
+  =
   \frac{                  1                     }
-         { \alpha( t ) }
-  \right]
-  =
-  \lambda_{ \min }
+       { \lambda(t) \, \bigl[ t + t_0(t) \bigr] }
 ,
-  %\notag
-  \label{eq:dinvadt_limit1}
-\end{equation}
-%
-and
-%
-\begin{equation}
-  \alpha( t )
-  =
-  \frac{               1               }
-       { \lambda_{ \min } ( t + t_1 )  }
-  .
-  \label{eq:alpha_limit1}
-\end{equation}
-%
-
-Second, near the end of simulation,
-we have $t \to T$, $u_k \approx 1$,
-%
-\begin{equation}
-  \frac{ d   }
-       { d t }
-  \left[
-    \frac{       1     }
-         { \alpha( t ) }
-  \right]
+  \label{eq:alpha_approxinvt}
+\end{equation}
+%
+where
+$t_0(t)$ is a slowly-varying function.
+
+By differentiating Eq. \eqref{eq:dinvadt},
+we can show that instantaneous eigenvalue,
+$\lambda(t)$, is an increasing function.
+%
+\note{Proof that $\lambda(t)$ is an increasing function.
+  Using $\dot u_k = \lambda_k \, u_k \, \alpha(t)$,
+  we have
+  $$
+  \begin{aligned}
+    \dot \lambda
+    &=
+    \frac
+    {
+      \sum_{k = 2}^n 2 \, \Gamma_k \, \lambda_k^3 \, u_k \, \dot u_k
+    }
+    {
+      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
+    }
+    -
+    \frac
+    {
+      \sum_{k = 2}^n \Gamma_k \, \lambda_k^3 \, u_k^2
+      \sum_{k = 2}^n 2 \, \Gamma_k \, \lambda_k^2 \, u_k \, \dot u_k
+    }
+    {
+      \left(
+        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
+      \right)^2
+    }
+    \\
+    &=
+    \frac
+    {
+      \sum_{k = 2}^n \Gamma_k \, \lambda_k^4 \, u_k^2
+      \,
+      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
+      -
+      \left(
+        \sum_{k = 2}^n \Gamma_k \, \lambda_k^3 \, u_k^2
+      \right)^2
+    }
+    {
+      \left(
+        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
+      \right)^2
+    }
+    2 \alpha(t).
+  \end{aligned}
+  $$
+  The numerator of the fraction of the last expression
+  is nonnegative by the Cauchy-Schwarz inequality,
+  or by the nonpositiveness of the discriminant
+  of the following nonnegative definite quadratic form of $X$:
+  $$
+  \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2 \, (X - \lambda_k)^2 \ge 0.
+  $$
+  Thus, $\dot \lambda \ge 0$, and $\lambda(t)$ is monotonically increasing.
+}
+%
+Below we shall study the initial and final values
+of $\lambda(t)$.
+%
+At the end of a simulation,
+we have $t = T$, and $u_k = 1$.
+%
+Thus,
+the instantaneous eigenvalue
+%
+\begin{equation}
+  \lambda(T)
   =
@@
   }
-  =
-  \bar \lambda,
-  %\notag
+  ,
   \label{eq:dinvadt_limit2}
 \end{equation}
-%
-where,
-$\bar \lambda$
-is the average of $\lambda_k$ weighted
+is an average of $\lambda_k$'s weighted
 by $\Gamma_k \, \lambda_k^2$,
-and it tends to be dominated
-by the largest $\lambda_k$.
-%
-Thus, in this regime, we have
-%
-\begin{equation}
-  \alpha( t )
-  =
-  \frac{             1             }
-       { \bar \lambda ( t + t_2 )  }
-  .
-  \label{eq:alpha_limit2}
-\end{equation}
-%
-Thus, we may characterize the optimal schedule as
-%
-\begin{equation}
-  \alpha(t)
-  =
-  \frac{                  1                     }
-       { \lambda(t) \, \bigl[ t + t_0(t) \bigr] }
-,
-\notag
-\end{equation}
-%
-where the instantaneous eigenvalue, $\lambda(t)$,
-increases gradually from $\lambda_{\min}$ to $\bar \lambda$,
-and $t_0(t)$ is a slowly-varying function.
+and it is independent of the simulation length, $T$.
+%
+Further, it often occurs that $\Gamma_k$
+increases with $\lambda_k$,
+such that the relative weight $\Gamma_k \, \lambda_k^2$
+also increases with $\lambda_k$.
+%
+It follows that $\lambda(T)$ will be dominated
+the larger values of $\lambda_k$'s.
+
+By contrast,
+at the beginning of a simulation,
+the relative weight, $\Gamma_k \, \lambda_k^2 \, u_k^2(0)$,
+is heavily influenced by the exponential factor
+$u_k^2(0) = e^{ -2 \, \lambda_k \, q(T) }$,
+and hence is sensitive to the simulation length, $T$.
+%
+A longer simulation means a smaller $\lambda(0)$,
+and that $\lambda(t)$ would be small
+for longer near the beginning of the simulation.
 
@@
 the values of $\Gamma_k$'s depend also on the system
-and the underlying sampling scheme (MC or MD).
-%
-Below is a technique of measuring the latter.
-
-The idea is that we can measure $\Gamma_k$'s
-via a constant schedule, $\alpha(t) = \alpha_0$,
-simulation using Eq. \eqref{eq:y2_eql}.
+and the underlying sampling process (MC or MD).
+%
+Below is a technique of measuring $\Gamma_k$'s.
+
+The idea is to measure $\Gamma_k$'s
+via a simulation with a constant schedule,
+$\alpha(t) = \alpha_0$,
+using Eq. \eqref{eq:y2_eql}.
 %
@@
 \sum_{ j = 1 }^n p_j \, V_j.
+\notag
 \end{align}
@@
 
-\subsubsection{\label{sec:invt_alpha}
-Inverse-time schedules}
-
-The optimal schedule predicted
-by the direct solution of Eq. \eqref{eq:q_opt}
-is complicated for a general
-multiple-bin updating scheme.
-%
-Several previous studies\cite{marsili2006, barducci2008, dickson2011}
-have been focused on analytically simpler schedules.
-%
-As an example of these approaches,
-we consider here a variant of
-the inverse-time formula, Eq. \eqref{eq:alpha_invt1}
-%
-\begin{equation}
-\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
-\label{eq:alpha_invtlambda}
-\end{equation}
-%
-with an optimized proportionality constant, $\lambda$.
-%
-We shall compute the error computed from
-this variationally optimized schedule
-and compare it with that from the optimal schedule.
-%
-We shall show that the optimal parameter, $\lambda$,
-is sensitive to the simulation time, $t$,
-which parallels the shift of
-the instantaneous eigenvalue,
-$\frac{d}{dt}\alpha^{-1}(t)$,
-as discussed in Sec. \ref{sec:optschedule}.
-
-Using Eq. \eqref{eq:alpha_invtlambda}
-in Eq. \eqref{eq:error_asym1} yields
-%
-\begin{align}
-\Err_A
-&=
-\frac{    1    }
-     { T + t_0 }
-\sum_{k = 2}^n
-  \frac{ \Gamma_k \, \nu_k^2 }
-       {    2 \, \nu_k - 1   }
-\left[
-  1 - \left(
-        1 + \frac{ T }{ t_0 }
-      \right)^{1 - 2 \, \nu_k}
-\right],
-\label{eq:error_asym_invt}
-\end{align}
-%
-where $\nu_k \equiv \lambda_k / \lambda$.
-%
-At long times, we get
-$$
-\begin{aligned}
-  \Err_A
-  \stackrel{         t \to \infty         }
-           { =\joinrel=\joinrel=\joinrel= }
-  \sum_{k = 2}^n
-  \begin{dcases}
-    \frac{    1    }
-         { T + t_0 }
-    \frac{ \Gamma_k \, \nu_k^2 }
-         {   2 \, \nu_k - 1    }
-    &
-    \mathrm{if \;} 2 \, \nu_k > 1,
-    \\%[1em]
-    %
-    %
-    \frac{    \Gamma_k    }
-         { 4 \, (T + t_0) }
-    \ln \frac{ T + t_0 }
-             {   t_0   }
-    &
-    \mathrm{if \;} 2 \, \nu_k = 1,
-    \\
-    %
-    %
-    \frac{  t_0^{ 2 \, \nu_k  - 1}  }
-         { (T + t_0)^{ 2 \, \nu_k } }
-    \frac{ \Gamma_k \, \nu_k^2 }
-         {   1 - 2 \, \nu_k    }
-    &
-    \mathrm{if \;} 2 \, \nu_k < 1.
-  \end{dcases}
-\end{aligned}
-$$
-%
-The last two cases, $\lambda \ge 2\lambda_k$,
-are asymptotically suboptimal
-as they decay slower than $1/T$.
-%
-Thus, the optimal $\lambda$
-satisfies $\lambda < 2 \min \lambda_k$,
-and in this case,
-we can seek that
-the optimal value from
-the smallest real root of
-%
-\begin{equation}
-\sum_{k = 2}^n
-\frac{ \lambda_k - \lambda }
-{ \left(2 - \lambda/ \lambda_k \right)^2 }
-= 0.
-\label{eq:optimal_lambda_approx}
-\end{equation}
-
-Under the approximation, we have
-$$
-\frac{ \lambda_k^2 }{ \lambda \, (2 \, \lambda_k - \lambda) } \ge 1,
-$$
-with the equality achieved only at $\lambda = \lambda_k$,
-and the minimal error from \eqref{eq:error_asym_invt}
-is achieved with
-$\lambda_2 = \dots = \lambda_n = \lambda$,
-agreeing with the general conclusion
-in Sec. \ref{sec:optWL}.
-
-Real simulations, however,
-can be relatively short
-and need not enter
-the asymptotic regime.
-%
-Then, the residual error, $\Err_R$, is not negligible.
-%
-Note, however,
-since Eq. \eqref{eq:alpha_invtlambda}
-does not fix the endpoint value,
-%
-\begin{equation}
-q(T) = \frac{ 1 } { \lambda }
-\log\left(
-  \frac{ T + t_0 } { t_0 }
-\right),
-\label{eq:qt_invtlambda}
-\end{equation}
-%
-we have to include the residual error
-during the optimization.
-%
-From Eqs. \eqref{eq:error_res1} and \eqref{eq:qt_invtlambda},
-we get
-%
-\begin{equation}
-\Err_R
-=
-\sum_{k = 2}^n
-  \frac{ \Gamma_k \, \nu_k }
-       {        2 \, t_0   }
-  \left(
-      \frac{   t_0   }
-           { T + t_0 }
-   \right)^{ 2 \, \nu_k },
-\label{eq:error_res_invt}
-\end{equation}
-%
-assuming that at $t = 0$
-the system has achieved an equilibrium
-at a constant $\alpha_0$,
-such that the values of
-$\left\langle y_k^2(0) \right\rangle$
-can be computed from Eq. \eqref{eq:y2_eql}.
-%
-The total error is given by the sum of
-Eqs. \eqref{eq:error_res_invt} and \eqref{eq:error_asym_invt}.
-%
-\begin{align}
-\Err
-&=
-\Err_R + \Err_A
-\notag
-\\
-&=
-\sum_{ k = 2 }^n
-  \Err^f_k \,
-  \left\{
-    1
-    +
-    \frac{ 1 - \left[ t_0 / (T + t_0) \right]^{2 \, \nu_k - 1} }
-         {                 2 \, \nu_k - 1                      }
-%    \frac{       1        }
-%         { 2 \, \nu_k - 1 }
-%    \left[
-%      1 -
-%      \left(
-%        \tfrac{   t_0   }
-%              { T + t_0 }
-%      \right)^{2 \, \nu_k - 1}
-%    \right]
-  \right\}
-,
-\label{eq:error_invt}
-\end{align}
-%
-where
-$$
-\Err^f_k
-\equiv
-\left\langle
-  y_k^2
-\right\rangle_{ \alpha(t) }
-=
-\frac{  \Gamma_k \, \nu_k   }
-     {    2 \, (T + t_0)    }
-,
-$$
-is the equilibrium, or saturated, error of mode $k$
-at the terminal value
-of $\alpha(t) = 1/[\lambda (T + t_0)]$.
-
-\note{Derivation of Eq. \eqref{eq:error_invt}:
-$$
-\begin{aligned}
-\Err_R
-&=
-\sum_{ k = 2 }^n
-  \Err^f_k \left( \frac{   t_0   }
-                       { T + t_0 }
-          \right)^{ 2 \nu_k - 1 }
-\\
-\Err_A
-&=
-\sum_{ k = 2 }^n
-  \frac{ 2 \, \Err^f_k \, \nu_k }
-       {    2 \, \nu_k - 1      }
-  \left[
-    1 -
-    \left(
-      \tfrac{   t_0   }
-            { T + t_0 }
-      \right)^{2 \, \nu_k - 1}
-  \right]
-\\
-&=
-\sum_{ k = 2 }^n
-  \Err^f_k \,
-  \left[
-    1 -
-    \left(
-      \tfrac{   t_0   }
-            { T + t_0 }
-      \right)^{2 \, \nu_k - 1}
-  \right]
-  +
-  \frac{   \Err^f_k     }
-       { 2 \, \nu_k - 1 }
-  \left[
-    1 -
-    \left(
-      \tfrac{   t_0   }
-           { T + t_0 }
-      \right)^{2 \, \nu_k - 1}
-  \right].
-\end{aligned}
-$$
-Adding the two yields Eq. \eqref{eq:error_invt}.
-}
-
-%We give several remarks on Eq. \eqref{eq:error_invt}.
-%
-
-Clearly, the final error is always greater than
-the saturated equilibrium value at the terminal $\alpha(t)$:
-$$
-E \ge E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
-$$
-because the second term in the braces
-is nonnegative no matter the sign of $2 \, \nu_k - 1$
-(in fact, it is a decreasing function
-that vanishes only at $\nu_k \to +\infty$).
-%
-For a set of nonidentical $\lambda_k$'s,
-the equality is reached only at $T = 0$.
-
-\subsubsection{\label{sec:optlambda}
-Optimal proportionality constant}
-
-We now show that the optimal $\lambda$
-tends to decrease with the simulation length.
-%
-Let us first rewrite Eq. \eqref{eq:error_invt} as
-%
-$$
-\Err
-=
-\sum_{ k = 2 }^n
-  \frac{    \Gamma_k    }
-       { 2 \, (T + t_0) }
-  G\left( \nu_k, \frac{ t_0 } { T + t_0} \right),
-$$
-%
-where the $k$th component is proportional to
-%
-$$
-G\left( \nu_k, r \right)
-=
-\nu_k
-\,
-\left(
-  1
-  +
-  \frac{ 1 - r^{2 \, \nu_k - 1} }
-       {   2 \, \nu_k - 1       }
-\right).
-$$
-%
-As shown in Fig. \ref{fig:err_component},
-the function $G(\nu_k, r)$ depends non-trivially on $\nu_k$.
-%
-For a fixed and sufficiently small $r = t_0 / (t + t_0)$,
-$G(\nu_k, r)$ has two local minima along $\nu_k$,
-one at $\nu_k = 0$,
-and the other shallower one around $\nu_k = 1$.
-%
-This means that for a relatively short simulation,
-hence a relatively large value of $r$,
-the minimal error may be achieved at $\nu_k \approx 0$,
-corresponding to a large value of $\lambda$
-($\lambda \gg \lambda_k$).
-%
-However, this minimum is unstable,
-for $G(\nu_k, r)$ increases rapidly
-with decreasing $r$, or increasing simulation length,
-under a fixed $\nu_k$.
-%
-In other words,
-asymptotically, the minimal error is always achieved at
-$\nu_k = 1$ or $\lambda = \lambda_k$.
-%
-In this case,
-$G(\nu_k, r) \to G(1, 0) = 2$
-and the total error is roughly twice $E_\mathrm{sat}$.
-
-For an arbitrary spectrum of eigenvalues,
-$\lambda_2, \dots, \lambda_n$,
-we expect that the optimal $\lambda$ to shift gradually to
-smaller values as the simulation length, $t$, increasing.
-%
-Asymptotically, we expect the optimal $\lambda$
-to be around $\min \lambda_k$.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/errcomp.pdf}
-  }
-  \caption{
-    \label{fig:err_component}
-    Component of the error, $G(\nu_k, r)$,
-    versus $\nu_k = \lambda_k / \lambda$,
-    where $r = t_0 / (t + t_0)$.
-    %
-  }
-\end{center}
-\end{figure}
-
 \subsection{\label{sec:band-matrix}
@@
 
-We now consider a class of updating schemes
-that are homogeneous,
-or translationally invariant. % with respect to $z$,
-%
-As we shall see,
-these schemes include the Gaussian updating scheme,
-which is widely used in metadynamics.
-%
-Although the translational invariance is
-natural for a periodic variable\cite{dama2014},
-it can be also extended to open-end variables
-by imposing the reflective boundary conditions.
-%
-For simplicity, we shall assume that our target
-distribution is uniform, or $p_i = 1/n$, below.
-
-\subsubsection{
-Eigenvalues and stability}
-
-Consider the following matrix $\mathbf w$
+Many commonly-used updating schemes
+can be characterized by a rigid wave packet
+whose center drifts with the current bin.
+%
+We call these schemes homogeneous,
+or translationally invariant.
+%
+An example is the Gaussian updating scheme
+used in metadynamics.
+%
+Homogeneous updating schemes have significantly fewer
+parameters than the general case,
+and the multiplication of the updating matrix,
+$\mathbf w$,
+can be reduced to a convolution with the wave packet,
+or the updating kernel.
+%
+Thus,
+the eigenvalues, $\lambda_k$'s, of $\mathbf w$
+can be readily computed and manipulated
+via the shape of the updating kernel.
+
+While the translational invariance of
+the updating scheme
+naturally suits a periodic variable\cite{dama2014},
+it can also be extended to a non-periodic variable
+by imposing the reflective boundary condition.
+%
+For simplicity, we shall assume that the target
+distribution is flat, or $p_i = 1/n$, below.
+
+\subsubsection{\label{sec:bandkernel}
+Window kernel and eigenvalues}
+
+It is natural to introduce a homogeneous updating scheme
+for a periodic variable\cite{dama2014}.
+%
+Consider the following matrix, $\mathbf w$,
 %
@@
   =
-  \mu_{ |i - j| }
-  + \mu_{ i + j - 1 }
-  + \mu_{ 2 n + 1 - i - j },
-  \label{eq:w_band}
-\end{equation}
-%
-where $\mu_0, \dots, \mu_b \, (b < n)$
+  \mu_{i-j}
+  +
+  \mu_{i-j+n}
+  +
+  \mu_{i-j-n}
+  ,
+\label{eq:w_band_pbc}
+\end{equation}
+%
+where
+$\mu_{-b}, \dots, \mu_b$ ($b < n/2$)
 are a set of numbers satisfying
@@
 \begin{equation}
-\mu_0 + 2 \, \mu_1 + \cdots + 2 \, \mu_b = 1,
-\label{eq:m_normalization}
+  \mu_{-b} + \cdots + \mu_b = 1
+  ,
+\label{eq:mu_normalization}
 \end{equation}
@@
 %
-The matrix is translational invariant under the reflective
-boundary condition\cite{bussi2006}.
-%
-On the right-hand side of Eq. \eqref{eq:w_band},
-the first term represents the relative updating magnitude
-to a neighbor that is $|i - j|$ bins away,
-whereas the next two terms are added
-%according to the boundary condition
-to avoid unintended distortion\cite{dickson2011, mcgovern2013}
-to the equilibrium distribution $p_i = 1/n$\cite{bussi2006}.
-%
-\note{Similarly, if $i$ represents a periodic variable,
-the last two terms can be dropped, and
-$|i-j|$ is evaluated
-under the minimum image convention\cite{dama2014}.
-This case is, however, not considered further.}
-
-This matrix, being symmetric to the indices $i$ and $j$,
+Particularly,
+we shall only consider symmetric kernels
+that satisfy
+%
+\begin{equation}
+  \mu_i = \mu_{-i}
+  .
+\label{eq:mu_symm}
+\end{equation}
+%
+These numbers constitute a \emph{updating kernel}.
+
+This above matrix, $\mathbf w$,
+being symmetric to the indices $i$ and $j$,
 satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
@@
 %
-It has $n$ orthonormal eigenvectors,
+To find the eigenvectors,
+we note that for a periodic variable $\varphi_i$,
+the out-of-boundary values are defined by
+%
+\begin{equation}
+  \varphi_i = \varphi_{i \pm n},
+\label{eq:phi_pbc}
+\end{equation}
+%
+such that $i \pm n$ lies in between $1$ and $n$.
+%
+We then have
+%
+\begin{equation}
+  \sum_{ j = 1 }^n
+    w_{ij} \, \varphi_j
+  =
+  \sum_{ j = 1 - n }^{ 2 \, n }
+    \mu_{i - j} \, \varphi_j
+  =
+  \sum_{ l = -b }^{ b }
+    \mu_l \, \varphi_{ i - l}
+  .
+\label{eq:wmul_to_convol}
+\end{equation}
+%
+\note{The derivation of the first step of
+  Eq. \eqref{eq:wmul_to_convol}:
+$$
+\begin{aligned}
+  \sum_{j = 1}^n
+    w_{ij} \, \varphi_j
+  &=
+  \sum_{j = 1}^n
+    \mu_{i - j} \, \varphi_j
+  +
+  \sum_{j = 1}^n
+    \mu_{i - j - n} \, \varphi_j
+  +
+  \sum_{j = 1}^n
+    \mu_{i - j + n} \, \varphi_j
+  \\
+  &=
+  \sum_{j = 1}^n
+    \mu_{i - j} \, \varphi_j
+  +
+  \sum_{j = 1+n}^{2 \, n}
+    \mu_{i - l} \, \varphi_l
+  +
+  \sum_{j = 1-n}^0
+    \mu_{i - j + n} \, \varphi_j
+  \\
+  &=
+  \sum_{j = 1-n}^{2 \, n}
+    \mu_{i - j} \, \varphi_j
+  .
+\end{aligned}
+$$
+The second step of Eq. \eqref{eq:wmul_to_convol}
+follows from the constraint $\mu_l = 0$ for $|l| > b$.
+
+Similarly,
+for the left eigenvector, we have
+$$
+  \sum_{ i = 1 }^n
+    \varphi_i \, w_{ij}
+  =
+  \sum_{ l = -b }^b
+    \varphi_{j - l} \, \mu_{-l}
+  .
+$$
+But due to the symmetry Eq. \eqref{eq:mu_symm},
+it is identical to \eqref{eq:wmul_to_convol}.
+}
+
+Thus,
+we may construct a set of orthonormal eigenvectors,
 $\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
-given by
+as
+\begin{equation}
+  \phi^{(k)}_j
+  =
+  \phi_{kj}
+  =
+  \frac{1}{\sqrt n}
+  \exp
+  \left[
+    \ii
+    \frac{ (k - 1) \, j \, 2 \, \pi }
+         {            n             }
+  \right]
+  ,
+  \notag
+\end{equation}
+%
+with the eigenvalues being
+%
+\begin{equation}
+  \lambda_k
+  =
+  \mu_0
+  +
+  2 \,
+  \sum_{ l = 1 }^b
+  \mu_l
+  \cos
+  \frac{ (k - 1) \, l \, 2 \, \pi }
+       {            n             }
+  .
+  \label{eq:wband_eigenvalue_pbc}
+\end{equation}
+%
+\note{Consider the unnormalized version
+  $$
+  \Phi^{(k)}_j =
+  \exp\left[
+    \frac{ ( k - 1 ) \, j \, 2 \, \pi }
+         {              n             }
+    \ii
+  \right]
+  ,
+  $$
+  we have
+  $$
+  \begin{aligned}
+  \sum_{j = 1}^n
+    w_{ij} \, \Phi^{(k)}_j
+  &=
+  \sum_{l = -b}^b
+    \mu_l \, \Phi^{(k)}_{i - l}
+  \\
+  &=
+  \mu_0 \, \Phi^{(k)}_i
+  +
+  \sum_{l = 1}^b
+    \mu_l \,
+    \left[ \Phi^{(k)}_{i - l} + \Phi^{(k)}_{i + l} \right]
+  \\
+  &=
+  \Phi^{(k)}_i \,
+  \left[
+    \mu_0
+    +
+    2 \sum_{l = 1}^b
+      \mu_l \, \cos
+      \frac{ (k - 1) \, l \, 2 \, \pi }
+           {            n             }
+  \right]
+  .
+  \end{aligned}
+  $$
+  The term in the square brackets is the eigenvalue given by
+  Eq. \eqref{eq:wband_eigenvalue_pbc}.
+}
+%
+Note, however, the two-fold degeneracy as
+$\lambda_{n - k + 2} = \lambda_k$.
+
+We now adapt the above formulae to a non-periodic variable.
+%
+We shall use the reflective boundary condition\cite{bussi2006},
+and modifies the form of the updating matrix, $\mathbf w$, as
+%
+%
+\begin{equation}
+  w_{ij}
+  =
+  \mu_{ i - j }
+  +
+  \mu_{ i + j - 1 }
+  +
+  \mu_{ i + j - 2 n - 1 },
+  \label{eq:w_band}
+\end{equation}
+%
+Again the updating kernel satisfies
+Eqs. \eqref{eq:mu_normalization}
+and
+\eqref{eq:mu_symm}.
+%
+However,
+the radius $b$ only need to be less than $n$
+instead of $n/2$.
+%
+The last two terms
+on the right-hand side of \eqref{eq:w_band} are added
+to avoid unintended distortion\cite{dickson2011, mcgovern2013}
+to the equilibrium distribution $p_i = 1/n$\cite{bussi2006},
+such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
+are satisfied.
+
+To compute the eigenvectors,
+we shall replace Eq. \eqref{eq:phi_pbc}
+and redefine the out-of-boundary values
+of $\varphi_j$ as
+%
+\begin{equation}
+  \varphi_j
+  =
+  \begin{dcases}
+    \varphi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
+    \varphi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
+  \end{dcases}
+\label{eq:phi_refl}
+\end{equation}
+%
+One can show that Eq. \eqref{eq:wmul_to_convol}
+still holds.
+%
+\note{The derivation of Eq. \eqref{eq:wmul_to_convol}
+  in this case is similar,
+  $$
+  \begin{aligned}
+    \sum_{j = 1}^n w_{ij} \, \varphi_j
+    &=
+    \sum_{j = 1}^n
+      \mu_{i - j} \, \varphi_j
+    +
+    \sum_{j = 1}^n
+      \mu_{i + j - 1} \, \varphi_j
+    +
+    \sum_{j = 1}^n
+      \mu_{i + j - 2 \, n - 1} \, \varphi_j
+    \\
+    &=
+    \sum_{j = 1}^n
+      \mu_{i - j} \, \varphi_j
+    +
+    \sum_{l = 1 - n}^0
+      \mu_{i - l} \, \varphi_l
+    +
+    \sum_{l = n + 1}^{ 2 \, n }
+      \mu_{i - l} \, \varphi_l
+    \\
+    &=
+    \sum_{j = 1 - n}^{ 2 \, n}
+      \mu_{i - j} \, \varphi_j.
+  \end{aligned}
+  $$
+}
+%
+Thus, it is also possible to construct
+$n$ orthonormal eigenvectors,
+$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
+as
 %
@@
      }
-  \cos \frac{ \left( i - \frac 1 2 \right) (k - 1) \, \pi}
+   \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, \pi}
             {                  n                         }.
-\label{eq:wband_eigenvector}
+\notag
+%\label{eq:wband_eigenvector_refl}
 \end{equation}
@@
   &=
-  \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
-  \label{eq:wband_eigenvalue}
-  %\\
-  %&=
-  %1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
-  %.
-  %\label{eq:wband_eigenvalue2}
+  \mu_0
+  +
+  2
+  \sum_{l = 1}^b
+    \mu_l
+    \cos \frac{(k - 1)  \, l \pi}{n}
+  \label{eq:wband_eigenvalue_refl}
 \end{align}
-
-\note{We derive Eqs.
-  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue}
-  below.
+%
+\note{We derive the above two equations below.
 
@@
   $$
-  \theta^{(k)}_i
+  \Phi^{(k)}_i
   =
@@
   $$
-  which satisfies the following reflection symmetry,
-  \begin{equation}
-    \theta^{(k)}_i = \theta^{(k)}_{1 - i} = \theta^{(k)}_{ 2 n + 1 - i }.
-    \label{eq:wband_reflection_symmetry}
-  \end{equation}
-  %
-  We shall also define $\mu_l = 0$, for $l \le 0$, then
-$$
-\begin{aligned}
-&\sum_{i = 1}^n
-  \theta^{(k)}_i
-  \, w_{ij}
-=
-- \theta^{(k)}_j \, \mu_0
-\\
-&
-+\sum_{i = 1}^n
-  \theta^{(k)}_i
-  \left(
-    \mu_{i - j} + \mu_{j - i} + \mu_{i + j - 1} + \mu_{2 \, n + 1 - i - j}
-  \right)
-.
-\end{aligned}
-$$
-
-Now
-$$
-\begin{aligned}
-  &
-  \sum_{i = 1}^n
-  \theta^{(k)}_i \,
-  \left(
-    \mu_{i - j} + \mu_{ 2 n + 1 - i - j }
-  \right)
-  \\
-  &=
-  \sum_{l=1-j}^{n-j} \theta^{(k)}_{j+l} \, \mu_l
-  +
-  \sum_{l = n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, \mu_l
-  \\
-  &=
-  \sum_{l=1-j}^{2n-j} \theta^{(k)}_{j+l} \, \mu_l
-  =
-  \sum_{l=0}^{b} \mu_l \, \cos\frac{(j+l-\frac12)(k-1)\pi}{n},
-\end{aligned}
-$$
-where we have used Eq. \eqref{eq:wband_reflection_symmetry}.
-%
-Similarly,
+  which satisfies the reflective boundary condition,
+  Eq. \eqref{eq:phi_refl}.
+  %
+  So
   $$
@@
   \sum_{i = 1}^n
-  \theta^{(k)}_i
-  \left(
-    \mu_{j - i} + \mu_{i+j-1}
-  \right)
-  =
-  \sum_{l=0}^{b} \mu_l \, \cos\frac{(j-l-\frac12)(k-1)\pi}{n}.
+    \Phi^{(k)}_i \, w_{ij}
+  &=
+  \sum_{l = -b}^b
+    \Phi^{(k)}_{j - l} \, \mu_l
+  \\
+  &=
+    \Phi^{(k)}_j \, \mu_0
+  + \sum_{l=0}^{b}
+    \mu_l \,
+    \left[
+      \Phi^{(k)}_{j-l}
+      +
+      \Phi^{(k)}_{j+l}
+    \right]
+  \\
+  &= \Phi^{(k)}_j \, \lambda_k,
   \end{aligned}
   $$
-%
-So
-$$
-\begin{aligned}
-\sum_{i = 1}^n \theta^{(k)}_i \, w_{ij}
-&=
-- \theta^{(k)}_j \, \mu_0
-+\sum_{l=0}^{b} \mu_l \, 2 \, \cos\frac{(k-1)l\pi}{n} \, \cos\frac{(j-\frac12)(k-1)\pi}{n}
-\\
-&= \theta^{(k)}_j \, \lambda_k,
-\end{aligned}
-$$
-with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}.
+  with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue_refl}.
+
+  \hrulefill
 
@@
   %
-If $k = 1$, $\theta^{(1)}_i = 1$, and
+  If $k = 1$, $\Phi^{(1)}_i = 1$, and
   $
-\sum_{i = 1}^n \left( \theta^{(1)}_i \right)^2 = n.
+  \sum_{i = 1}^n \left( \Phi^{(1)}_i \right)^2 = n.
   $
@@
 
-We can also show that column-wise orthogonality.
-%
-For this we need a lemma, for $a = q \, \pi/n$,
+  \hrulefill
+
+  To show the column-wise orthogonality,
+  we need a lemma, for $a = q \, \pi/n$,
   with $q$ a positive integer.
@@
   $$
-\sum_{k = 1}^n \cos\left[(k - 1) \, \frac { q \, \pi } { n }  \right]
-= n \delta_{q, 0} + \operatorname{odd}(q).
+  \sum_{k = 1}^n
+    \cos\left[(k - 1) \, \frac { q \, \pi } { n }  \right]
+  = n \, \delta_{q, 0}
+  + \operatorname{odd}(q).
   $$
@@
     \sum_{k = 1}^n
-  \theta^{(k)}_i \, \theta^{(k)}_j
+    \Phi^{(k)}_i \, \Phi^{(k)}_j
     &=
@@
     \sum_{k = 1}^n
-  \theta^{(k)}_i \, \theta^{(k)}_j
+    \Phi^{(k)}_i \, \Phi^{(k)}_j
     -
@@
 
-We can define a window kernel from the set of numbers as
-$$
-\hat \mu_i = \mu_{ |i| }.
-$$
-Then, according to Eq. \eqref{eq:vt_diffeq_mbin},
-the rate of change of the bias potential, $\dot v_i(t)$,
-is given by the convolution of the above kernel, $\hat \mu_j$,
-and the normalized histogram, $h_j/p_j$\cite{bussi2006},
-and Eq. \eqref{eq:wband_eigenvalue}
-shows that the eigenvalues, $\lambda_k$, of the updating scheme
-is the Fourier transform of the kernel.
+Note that Eq. \eqref{eq:wband_eigenvalue_refl}
+differs from Eq. \eqref{eq:wband_eigenvalue_pbc}
+by a factor of $2$
+in the argument of the cosine function.
+%
+We may unify the two as
+%
+\begin{equation}
+  \lambda_k
+  =
+  \mu_0
+  +
+  2
+  \sum_{ l = 1 }^b
+    \mu_l \,
+    \cos \frac{ (k - 1) \, l \, g \, \pi }
+              {            n             }
+  ,
+  \label{eq:wband_eigenvalue}
+\end{equation}
+%
+where the degeneracy $g$ is $2$
+for the periodic case,
+or $1$ otherwise.
+%
+Equation \eqref{eq:wband_eigenvalue}
+shows that the eigenvalues, $\lambda_k$,
+of the updating scheme
+is the cosine transform of
+the updating kernel.
 %
 So, for the updating scheme to be stable,
-the Fourier transform of the kernel
+the cosine transform of the kernel
 must be nonnegative.
 %
-In Sec. \ref{sec:stabilize_wband},
-we show how to stabilize an arbitrary kernel.
+If this is not so,
+we can modify the kernel
+using the technique described
+in Sec. \ref{sec:stabilize_wband}.
 
@@
 
-\subsubsection{Nearest-neighbor updating scheme}
-
-Particularly, if $\mu_2 = \dots = \mu_b = 0$,
+\subsubsection{\label{sec:nnscheme}
+Nearest-neighbor updating scheme}
+
+If $\mu_2 = \dots = \mu_b = 0$,
 the matrix $\mathbf w$ is tridiagonal,
@@
 \begin{equation}
-  \lambda_k = 1 - 4 \, \mu_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
+  \lambda_k
+  =
+  1 -
+  4 \, \mu_1 \sin^2
+  \frac{ (k - 1) \, g \, \pi }
+       {       2 \, n        }
+  .
 \label{eq:wnn_eigenvalue}
@@
 \begin{equation}
-\mu_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
+  \mu_1 <
+  \begin{dcases}
+    \frac 1 4
+    & \mathrm{periodic, \;} n \mathrm{ \; even,}
+    \\
+    \frac 1 4
+    \cos^{-2}\frac{  \pi   }
+                  { 2 \, n }
+    & \mathrm{otherwise}.
+  \end{dcases}
 \label{eq:nn_stable}
 \end{equation}
-
-We can find an analytical expression
-of the optimal value of $\lambda$
-in the limit of large $n$.
-%
-In this case,
-we convert Eq.
-\eqref{eq:error_asym_invt}
-%and
-%\eqref{eq:optimal_lambda_approx}
-to an integral,
-$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
-%
-\begin{align}
-\Err_A
-=
-\frac{2 \, n}{\pi \, T}
-\int_0^{\pi/2}
-  \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
-       {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-\, dp.
-\notag
-%\label{eq:error_nn_asym_int}
-\end{align}
-
-For perfect sampling,
-we have, from Eq. \eqref{eq:Gamma_perfect},
-$\Gamma(p) = 1$, and
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{2 \, n}{\pi \, T}
-\int_0^{\pi/2}
-\frac{ \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2 }
-{ \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-\, dp
-\notag \\
-&=
-\frac{n}{4 \, T}
-\left(
-  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
-  +
-  \frac{ \lambda }
-  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
-\right).
-\end{aligned}
-$$
-\note{The integral is evaluated by contour integration.
-%
-It is somewhat more convenient to change variable $p \to \frac{ \pi } { 2 } - p$,
-and
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{n}{\lambda \, T}
-\frac{1}{2 \pi i}
-\int_0^{2 \, \pi}
-\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
-{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
-\, dp
-\\
-&=
-\frac{n}{\lambda \, T}
-\frac{1}{2 \pi i}
-\oint
-\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
-{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
-\, \frac{dz}{z},
-\end{aligned}
-$$
-where the contour is the unit circle
-in the complex plane of $z$.
-%
-This integral has five poles, one at the origin,
-which produces a residue
-$\frac{2 - 4 \, \mu_1 - \lambda}{4}$
-(from series expansion of small $z$),
-and the other four at
-$$
-z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
-$$
-with two of them inside the unit circle:
-$$
-z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
-{\sqrt{8 \mu_1}}.
-$$
-Thus,
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{      n       }
-     { \lambda \, T }
-\left(
- \frac{ 2 - 4 \, \mu_1 - \lambda }
-      {          4               }
- +
- \sum_{z = z_{\pm} }
- \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
- { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
-\right)
-\\
-&=
-\frac{       n      }
-     { \lambda \, t }
-\left(
-  \frac{ 2 - 4 \, \mu_1 - \lambda }
-       {          4               }
- +
- \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
-      { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2                 }
-\right)
-\\
-&=
-\frac{       n      }
-     { \lambda \, T }
-\left(
-  \frac{ 2 - 4 \, \mu_1 - \lambda }
-       {          4               }
- +
- \frac{ (\lambda/2)^2 }
- { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
-\right).
-\end{aligned}
-$$
-}
-%
-Minimizing the function yields
-%
-\begin{equation}
-\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
-\label{eq:lambda_nn_perfect}
-\end{equation}
-%
-and the asymptotic error in the optimal case is given by
-%
-\begin{equation}
-\Err_A
-=
-\frac{n}{T}
-\left(
-  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
-\right).
-\label{eq:error_nn_prefect}
-\end{equation}
-%
-%From Eq. \eqref{eq:error_nn}, it is clear that
-Clearly, the error
-increases with the magnitude of $\mu_1$
-no matter its sign,
-and the minimal value, obtained at $\mu_1 = 0$,
-corresponds to the single-bin scheme.
-%
-Note that we have ignored the residual error
-and assumed $\lambda < 2 \, \min \lambda_k = 2 - 8 \, \mu_1$
-for all $k$ in the above solution.
-%
-Fortunately,
-the $\lambda$ from Eq. \eqref{eq:lambda_nn_perfect}
-satisfies this condition
-by the stability condition,
-$4 \, \mu_1 < 1$.
-\note{This is because,
-$$
-\frac{ 1 - 4 \, \mu_1 } { 1 - 2 \, \mu_1 } < 2 - 8 \, \mu_1
-\quad \leftrightarrow \quad
-1 < 2 - 4 \, \mu_1.
-$$
-}
-
-For the one-step sampling,
-we find from Eq. \eqref{eq:Gamma_onestep}
-that $\Gamma(p) = \cot^2 p$
-is peaked around $p \approx 0$,
-%
-Then
-$$
-\begin{aligned}
-\Err_A
-&
-\propto
-\frac{   2 \, n }
-     { \pi \, T }
-\left.
-\frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
-     { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-\right|_{ p \approx 0 }
-\notag \\
-&
-\approx
-\frac{   2 \, n }
-     { \pi \, T }
-\frac{             1             }
-     {  \lambda \, (2 - \lambda) }
-.
-\end{aligned}
-$$
-%
-The optimal $\lambda$ is therefore
-%
-\begin{equation}
-\lambda \approx 1.
-\label{eq:lambda_nn_onestep}
-\end{equation}
-%
-\note{We can be a bit more precise here.
-%
-From Eq. \eqref{eq:Gamma_onestep}, we get
-%
-$$
-\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
-$$
-%
-where we have introduced
-$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
-as a small parameter to avoid the divergence
-around the origin.
-%
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{   2 \, n }
-     { \pi \, t }
-\int_0^{\pi/2}
-    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
-         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-    \frac{ \cos^2 p }
-         { \sin^2 p + \delta^2 }
-\, dp
-\notag \\
-&
-\stackrel{    \delta \ll 1     }
-         { =\joinrel=\joinrel= }
-\frac{   2 \, n }
-     { \pi \, t }
-\frac{             1             }
-     {  \lambda \, (2 - \lambda) }
-\frac{    1   }
-     { \delta }
-.
-\end{aligned}
-$$
-The integral is evaluated as follows.
-%
-We again change variable $p \to \frac{ \pi } { 2 } - p$,
-and
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{n}{\lambda \, t}
-\frac{1}{2 \pi i}
-\int_0^{2 \, \pi}
-\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
-     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
-\cdot
-\frac{ \sin^2 p            }
-     { \cos^2 p + \delta^2 }
-\, dp
-\\
-&=
-\frac{n}{\lambda \, t}
-\frac{1}{2 \pi i}
-\oint
-\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
-     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
-\cdot
-\frac{ 4 - \left( z + \frac 1 z \right)^2 }
-     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
-\, \frac{dz}{z}
-\\
-&=
-\frac{n}{\lambda \, t}
-\frac{1}{2 \pi i}
-\oint
-\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
-     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
-\cdot
-\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
-     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
-\, \frac{ dz }{ z^3 }
-\\
-&=
-\frac{ n } { \lambda \, t }
-\sum_l \operatorname{Res}_l
-,
-\end{aligned}
-$$
-where the contour is the unit circle
-in the complex plane of $z$,
-and the integral is reduced to a sum of residues
-around the poles within.
-
-The poles of the integral come from three sources,
-namely,
-% 1.
-$z = 0$,
-% 2.
-the zeros of
-$(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2$,
-% 3.
-and the zeros of
-$( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2$.
-
-For the first pole around the origin,
-we need to expand the integrand as
-$$
-\frac{ A + B \, z^2 + \cdots }
-     {          z^3          },
-$$
-then the residue is $B$.
-%
-Since
-$$
-\begin{aligned}
-\left[ z^2 - \mu_1 (z^2 + 1)^2 \right]^2
-&
-\approx
-[ - \mu_1 + (1 - 2 \, \mu_1) \, z^2]^2
-\\
-&
-\approx
-\mu_1^2
-\left[
-  1 + \left(4 - \tfrac { 2  } { \mu_1 } \right) z^2
-\right],
-\\
-%
-4 \, z^2 - ( z^2 + 1 )^2
-&
-\approx
--( 1 - 2 \, z^2 ),
-\\
-%
-(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2
-&
-\approx
--2 \, \mu_1 \,
-\left[
-  1 + \left(
-        2 + \tfrac{ \lambda - 2 } { 2 \, \mu_1 }
-      \right)
-      \, z^2
-\right],
-\\
-%
-( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2
-&
-\approx
-1 + (2 + 4 \, \delta^2) \, z^2,
-\end{aligned}
-$$
-the integrand is
-$$
-\begin{aligned}
-\frac{ \mu_1 }
-     {   2   }
-\left[
-  1 + \left(
-        - \frac{ 2 + \lambda } { 2 \, \mu_1 }
-        - 2 - 4 \, \delta^2
-      \right) \, z^2
-\right].
-\end{aligned}
-$$
-and the residue is
-$$
-\operatorname{Res}_1
-=
-  - \frac{ 2 + \lambda } { 4 }
-  - \mu_1 \, \left( 1 + 2 \, \delta^2 \right).
-$$
-
-For the second part,
-we have four roots of
-$$
-(2 - \lambda) \, z^2 = 2 \, \mu_1 \, (z^2 + 1)^2,
-$$
-which are
-\begin{equation}
-z
-=
-\pm
-\sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } }
-\pm
-\sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } - 1 }.
-\label{eq:nnint_local_zeros_part2}
-\end{equation}
-We share distinguish two cases.
-\textbf{Case A.}
-If $\lambda < 2 - 8 \, \mu_1$,
-two of the roots lie in the unit circle, namely
-$$
-z
-=
-\pm
-\left(
-\sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } }
--
-\sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } - 1 }
-\right),
-$$
-which satisfy
-$$
-4 \, \mu_1 \, ( z^2 + 1)
-=
-2 - \lambda
--
-\sqrt{ ( 2 - \lambda ) ( 2 - \lambda - 8 \, \mu_1 ) }.
-$$
-Thus,
-$$
-\begin{aligned}
-\left[ z^2 - \mu_1 \, \left( z^2 + 1 \right)^2 \right]^2
-&=
-\left( z^2 - \tfrac{ 2 - \lambda } { 2 } z^2 \right)^2
-=
-\frac{ \lambda^2 } { 4 } z^4.
-\\
-%
-4 z^2 - \left( z^2 + 1 \right)^2
-&=
--\frac{ 2 - \lambda - 8 \, \mu_1  } { 2 \, \mu_1 } z^2,
-\\
-%
-\left( z^2 + 1 \right)^2 + 4 \, \delta^2 \, z^2
-&=
-\frac{ 2 - \lambda + 8 \, \mu_1 \, \delta^2 }
-     { 2 \, \mu_1 }
-     z^2,
-\\
-%
-(2 - \lambda) \, 2 \, z
--
-8 \, \mu_1 \, \left( z^2 + 1 \right) \, z
-&=
-2 \, z \, \left[2 - \lambda - 4 \,  \mu_1 \, (z^2 + 1) \right]
-\\
-&= 2 \, z \, \sqrt{ (2 - \lambda) ( 2 - \lambda - 8 \, \mu_1 ) },
-\end{aligned}
-$$
-where we have used the L'H\^{o}pital's rule
-and taken the derivative for the last factor.
-%
-Note that the two roots within the unit share the same
-residue and the sum for this part is
-%
-$$
-\operatorname{Res}_2
-=
--\frac{                 \lambda^2                     }
-      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
-\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }.
-$$
-%
-\textbf{Case B.}
-If $2 - 8 \, \mu_1 \le \lambda < 2$,
-the four zeros in Eq. \eqref{eq:nnint_local_zeros_part2}:
-%
-\begin{equation}
-z
-=
-\pm
-\sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } }
-\pm i \,
-\sqrt { 1 - \frac{ 2 - \lambda }
-                 { 8 \, \mu_1  } },
-\label{eq:nnint_local_zeros_part2b}
-\end{equation}
-%
-all satisfy $|z| = 1$, and hence
-all lie on the border of the unit circle.
-%
-Thus, each contributes half of the residue there.
-%
-However, since the four zeros satisfy
-$$
-2 - \lambda - 4 \, \mu_1 \, ( z^2 + 1)
-=
-\pm i
-\sqrt{ ( 2 - \lambda ) ( 8 \, \mu_1 - 2 + \lambda ) },
-$$
-the sum of the four residues gives zero in this case:
-$$
-\operatorname{Res}_2 = 0.
-$$
-In summary,
-$$
-\operatorname{Res}_2
-=
-\begin{dcases}
--\frac{                 \lambda^2                     }
-      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
-\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }
-&
-\mathrm{if \;}
-\lambda < 2 - 8 \, \mu_1,
-\\
-0
-&
-\mathrm{otherwise}.
-\end{dcases}
-$$
-
-For the last part,
-we have four roots of
-$(z^2 + 1)^2 = - 4 \, \delta^2 \, z^2$,
-with two of them lying in the unit circle:
-$$
-z = \pm i \left( \sqrt{ 1 + \delta^2 } - \delta \right),
-$$
-satisfying
-$$
-z^2 + 1 + 2 \, \delta^2 = 2 \, \delta \, \sqrt{1 + \delta^2}.
-$$
-Then,
-$$
-\begin{aligned}
-z^2 - \mu_1 \, \left( z^2 + 1 \right)^2
-&= z^2 \, (1 + 4 \, \mu_1 \, \delta^2),
-\\
-4 \, z^2 - \left( z^2 + 1 \right)^2
-&=
-4 \, z^2 \, (1 + \delta^2)
-\\
-%
-(2 - \lambda) \, z^2
-- 2 \, \mu_1 \, \left( z^2 + 1 \right)^2
-&=
-\left(
-  2 - \lambda + 8 \, \mu_1 \, \delta^2
-\right) \, z^2
-\\
-%
-4 \, z \, \left( z^2 + 1 \right)
-+
-8 \, \delta^2 \, z
-&=
-4 \, z \, (z^2 + 1 + \delta^2)
-=
-8 \, z \, \delta \, \sqrt{ 1 + \delta^2 }.
-\end{aligned}
-$$
-Note that the above two roots share the same
-residue and the sum for this part is
-$$
-\operatorname{Res}_3
-=
-\frac{ \left( 1 + 4 \, \mu_1 \, \delta^2 \right)^2 }
-     {        2 + 8 \, \mu_1 \, \delta^2 - \lambda }
-\frac{ \sqrt{ 1 + \delta^2 } }
-     {        \delta         }.
-$$
-
-Now as $\delta \to 0$, the last contribution will dominate,
-and we have
-$$
-\Err_A
-\approx
-\frac{ n } { T }
-\frac{ (1 + 4 \, \mu_1 \, \delta^2)^2 }
-{ \lambda \, (2 + 8 \, \mu_1 \, \delta^2 - \lambda) }
-\frac{ 1 } { \delta }
-\approx
-\frac{ n } { T }
-\frac{ 1 }
-{ \lambda \, (2 - \lambda) }
-\frac{ 1 } { \delta },
-$$
-and the minimum is reached at $\lambda \approx 1$.
-
-The weakness of the above calculation is two fold.
-%
-First, it ignores the residual error.
-%
-Second, the simplified expression for the asymptotic error
-$E_A$ may also be misused.
-%
-Namely, with $\lambda \approx 1$,
-we cannot assume $\lambda < 2 \, \lambda_k$ in general.
-%
-So it only works for a very small value of $\mu_1$.
-}
-
-Note that Eq. \eqref{eq:lambda_nn_onestep} is only good
-for sufficiently short simulations.
-%
-For very longer simulations,
-the optimal $\lambda$ tends to shift
-toward the smallest eigenvalue,
-$\min \lambda_k = 1 - 4 \, \mu_1$
-according to the discussion in Sec. \ref{sec:optlambda}.
 
@@
 %
-Below we show that the resulting updating scheme
+Below we show that the corresponding updating scheme
 is stable, albeit less effective
@@
     \right),
+\notag
 \end{equation}
@@
 .
-\label{eq:lambda_Gaussian}
+\notag
+%\label{eq:lambda_Gaussian}
 \end{align}
@@
 For example, the nearest-neighbor updating scheme
-is unstable if \eqref{eq:nn_stable} is violated.
+is unstable if Eq. \eqref{eq:nn_stable} is violated.
 %
 Even the Gaussian updating scheme can be unstable
-if the window kernel is truncated.
+if the updating kernel is truncated.
 %
 Below we show how to minimally modify
-the window kernel function
+the updating kernel
 to meet the stability criteria.
@@
 The key is the observation of Eq. \eqref{eq:wband_eigenvalue}
-that the eigenvalues are related to the window kernel
+that the eigenvalues are related to the updating kernel
 by a cosine transform,
@@
 %
-Since we impose $\mu_n = 0$,
+For the periodic case,
+the inversion formula is
+%
+\begin{equation}
+  \mu_l
+  =
+  \frac 1 n
+  \sum_{ k = 1 }^n
+  \lambda_k
+  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
+            {            n             }
+  .
+\label{eq:mu_from_lambda_pbc}
+\end{equation}
+
+For a non-periodic variable,
 the explicit inversion formula is
@@
   =
-  \frac 1 n
-  \left[
-    \frac{ \lambda_1 + (-1)^l \, \lambda_{n+1} }
-         {               2                     }
-    +
-    \sum_{ k = 1 }^{ n - 1 }
-      \lambda_{ k + 1 } \,
-      \cos \frac{ k \, l \, \pi }
+  \frac 1 { 2 \, n }
+  \sum_{ k = 1 }^{ 2 \, n }
+    \lambda_{ k } \,
+    \cos \frac{ (k - 1) \, l \, \pi }
               {            n        }
-  \right],
-  \label{eq:mu_from_lambda}
+  ,
+\label{eq:mu_from_lambda_refl}
 \end{align}
@@
 where
+we have defined
+$\lambda_k \equiv \lambda_{n + 2 - k}$
+for $k = n + 2, \dots, 2 \, n$,
+as well as
 %
@@
     +
-    2 \, \sum_{ k = 1 }^{ n - 1 }
-            (-1)^k \, \lambda_{k + 1}
-  \right].
+    2 \, \sum_{ k = 2 }^{ n }
+      (-1)^{k - 1} \, \lambda_k
+  \right]
+  ,
 \label{eq:lambdan}
 \end{align}
+to satisfy the constraint $\mu_n = 0$.
 %
@@
 ensures that the $\mu_n$
-computed from Eq. \eqref{eq:mu_from_lambda}
+computed from Eq. \eqref{eq:mu_from_lambda_refl}
 vanishes.
@@
     \mathrm{for\;} l = 1, \dots, n - 1,
-    \label{eq:mu_reflection}
+  \notag
+  %\label{eq:mu_reflection}
   \end{align}
   %
-  then Eq. \eqref{eq:wband_eigenvalue} can be rewritten as
+  then Eq. \eqref{eq:wband_eigenvalue_refl} can be rewritten as
   %
@@
     \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
-    \label{eq:lambda_cosine_sum}
+  \notag
+  %\label{eq:lambda_cosine_sum}
   \end{align}
@@
   \frac{ \lambda_1 - \lambda_2 }
-       {           2           }.
+       {           2           }
+  .
   \end{aligned}
@@
     \lambda_1 + 2 \, \lambda_3
-  \right),
+  \right)
+  ,
   \\
@@
   \right)
+  ,
   \\
@@
   \right)
+  .
   \end{aligned}
@@
 
-Now given a window kernel, $\mu_1, \dots, \mu_n$,
-we can compute the eigenvalues from \eqref{eq:wband_eigenvalue}.
-%
-By replacing negative eigenvalues by zeros and
-using Eq. \eqref{eq:mu_from_lambda},
-we get a new window kernel
+Now given a updating kernel, $\mu_1, \dots, \mu_n$,
+we can compute the eigenvalues from
+Eq. \eqref{eq:wband_eigenvalue}.
+%
+By replacing negative eigenvalues with zeros and
+using Eq. \eqref{eq:mu_from_lambda_pbc}
+or \eqref{eq:mu_from_lambda_refl},
+we get a new updating kernel
 $\mu^\mathrm{new}_1, \dots, \mu^\mathrm{new}_{n - 1}$,
@@
 %
-However, the new window may be wider than
-the original one if any of
-$\mu^\mathrm{new}_{b+1}, \dots, \mu^\mathrm{new}_{n-1}$
-is nonzero,
-and if we truncate the window kernel at $b$ bins,
-the stability problem is still existent
-albeit usually less severe
+However, the new kernel
+is widened
+as it contains $n/g - 1$ instead of $b$ entries.
+If we truncate the new updating kernel
+at a radius of $b$ bins,
+the stability problem remains
+albeit, hopefully, becomes less severe
 (in terms of the most negative eigenvalue).
 %
-Thus, to obtain a stable window kernel of width $b$,
+Thus, to obtain a stable updating kernel
+without increasing the radius $b$,
 we have to iterate the above process many times.
@@
   \item
-    Given a window kernel $mu_1, \dots, \mu_b$,
+    Given a updating kernel $\mu_1, \dots, \mu_b$,
     compute the eigenvalues
@@
     If all eigenvalues are nonnegative,
-    we are done, are break the loop.
+    we are done, and exit the loop.
   \item
     Otherwise, replace negative eigenvalues by zeros,
-    and compute a new window kernel,
+    and compute a new updating kernel,
     $\mu_1, \dots, \mu_b$ from
-    Eq. \eqref{eq:mu_from_lambda}.
+    Eq. \eqref{eq:mu_from_lambda_pbc}
+    or
+    \eqref{eq:mu_from_lambda_refl}.
     Set $\mu_l = 0$ for $l > b$.
@@
 
+Now that we can compute the optimal schedule
+for a given updating scheme,
+can we also compare different updating schemes
+and pick the best one?
+%
+Below we address this question
+in the asymptotic, or long simulation time,
+limit.
+
 \subsubsection{\label{sec:optWL}
@@
 
-The error functional given in Eq. \eqref{eq:error_asym1}
-leads to an interesting deduction:
-the single-bin scheme is asymptotically optimal.
+We shall first show,
+directly from Eq. \eqref{eq:error_asym1},
+that the single-bin scheme is asymptotically optimal.
 %
@@
 Using the Cauchy-Schwarz inequality, we have,
-for any set of numbers, $c_k \ge 0$,
+for any set of nonnegative numbers, $c_k$'s,
 %
@@
 \left(
-  \int_0^t dt'
+  \int_0^T dt
     \sum_{k = 2}^n
-      \Gamma_k \, \dot u_k^2\bigl( q(t') \bigr)
+      \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr)
 \right)
@@
 \left(
-  \int_0^t dt'
+  \int_0^T dt
     \sum_{k = 2}^n
@@
 \left(
-  \int_0^t dt'
+  \int_0^T dt
     \sum_{k = 2}^n
-      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t') \bigr)
+      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t) \bigr)
 \right)^2
@@
     \left[
-      1 - e^{ -\lambda_k q(t) }
+      1 - e^{ -\lambda_k \, q(T) }
     \right]
@@
 $$
-\int_0^t
-  dt' \sum_{k = 2}^n \Gamma_k \,
+\int_0^T
+  dt \sum_{k = 2}^n \Gamma_k \,
     \left( \dot u_k \, Y - c_k \right)^2
@@
 Note that the last expression of \eqref{eq:CSineq}
-depends on $q(t')$ only through the end point value,
+depends on the curve, $q(t)$ ($0 < t < T$),
+only through the endpoint value, $q(T)$,
 which is fixed in the variation process.
@@
 The equality is achieved
-if $\dot u_k\left( q(t') \right) = c_k$
-for all $k = 2, \dots, n$ at any $t'$,
+if $\dot u_k\left( q(t) \right) = c_k$
+for all $k = 2, \dots, n$ at any $t$,
 up to a multiplicative factor.
@@
 $$
-\alpha(t') = \frac{              1              }
-                  { \lambda_k \, (t' + t_{k0} ) },
+\alpha(t) = \frac{              1             }
+                 { \lambda_k \, (t + t_{k0} ) },
 \quad
 \mathrm{with\;\;}
-t_{k0} = \frac{             t            }
-              { e^{ \lambda_k q(t) } - 1 }.
+t_{k0} = \frac{             T            }
+              { e^{ \lambda_k q(T) } - 1 }.
 $$
@@
   \begin{equation}
-  u_k(t') = c_k \left(t' + t_{k0} \right).
+  u_k(t) = c_k \left(t + t_{k0} \right).
   \label{eq:ui_solution}
   \end{equation}
-  Given that $u_k(t) = 1$ and $u_k(0) = e^{-\lambda_k \, q(t)}$,
+  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
   we have
   $$
-  c_k = \frac{ 1 }{ t + t_{k0} },
+  c_k = \frac{ 1 }{ T + t_{k0} },
   \quad
   \mathrm{and\;}
-  \frac{ t_{k0} } { t + t_{k0} }
-  =
-  e^{ -\lambda_k \, q(t) }.
+  \frac{ t_{k0} } { T + t_{k0} }
+  =
+  e^{ -\lambda_k \, q(T) }.
   $$
@@
   $$
-  -\lambda_k \, q(t) + \lambda_k \, q(t')
-  = \log c_k + \log\left( t' + t_{k0} \right).
-  $$
-  Differentiating this with respect to $t'$ yields
-  $$
-  \lambda_k \, \alpha(t') = \frac{ 1 } { t' + t_{k0} }.
-  $$
-}
-
-Now an updating matrix $\mathbf w$ satisfying this condition
-is essentially a multiple of the identity matrix
-(since $\lambda_1$ corresponds to the mode
+  -\lambda_k \, q(T) + \lambda_k \, q(t)
+  = \log c_k + \log\left( t + t_{k0} \right).
+  $$
+  Differentiating this with respect to $t$ yields
+  $$
+  \lambda_k \, \alpha(t)
+  =
+  \frac {     1      }
+        { t + t_{k0} }.
+  $$
+}
+
+Since $\lambda_1$ corresponds to the mode
 of a uniform shift of all $v_k$,
-we can freely set it to $\lambda_2$
-without changing the nature of the updating scheme).
+we can freely set $\lambda_1$ to $\lambda_2$
+without changing the nature of the updating scheme.
+%
+Now an updating matrix $\mathbf w$
+with identical eigenvalues
+is essentially a multiple of the identity matrix.
 %
@@
 In the single-bin case,
-$c_k = 1/(t + t_0)$,
+$c_k = 1/(T + t_0)$,
 and the error is
@@
   =
-\frac{       t     }
-     { (t + t_0)^2 }
-\sum_{ k = 1 }^n
+  \frac{       T     }
+       { (T + t_0)^2 }
+  \sum_{ k = 2 }^n
     \Gamma_k
@@
 
-We emphasize that the above result
+Note that the above result
 holds only under the long time limit,
@@
 %
-Then for shorter times, $t \approx 0$,
+Then for shorter times, $T \approx 0$,
 the error, given by Eq. \eqref{eq:error_eql},
@@
 %
-In practice, real simulations
-may have long achieved the desired precision
-before reaching the asymptotic regime,
-especially if the updating matrix
-has some $\lambda_k$'s close to $0$.
+%In practice, real simulations
+%may have long achieved the desired precision
+%before reaching the asymptotic regime,
+%especially if the updating matrix
+%has some $\lambda_k$'s close to $0$.
+
+We can also generalize
+the single-bin scheme to a class of
+optimal updating schemes.
+%
+For example,
+if we know in advance that the intrinsic
+PMF is sufficiently smooth such that
+the error modes can be limited to $k \le K$.
+%
+Then we can manually construct
+an updating scheme with
+$$
+\lambda_1 = \cdots = \lambda_K = 1,
+\mathrm{\; and \;}
+\lambda_{K+1} = \cdots = \lambda_n = 0.
+$$
+For homogeneous updating schemes,
+the updating kernel can be inversely
+found from Eq. \eqref{eq:mu_from_lambda_pbc}
+or \eqref{eq:mu_from_lambda_refl}.
+%
+Then,
+Eq. \eqref{eq:error_singlebin} is modified to
+$$
+  E
+  =
+  \frac {       T     }
+        { (T + t_0)^2 }
+  \sum_{ k = 2 }^K
+    \Gamma_k.
+$$
 
@@
 is the corresponding equilibrium FDS method
-that fixes the bias potential to the exact value.
-%
-Such a simulation delivers the maximal possible efficiency.
+with the bias potential fixed to the exact value.
+%
+Such a simulation delivers the maximal efficiency.
 
@@
 In late stages of an adaptive FDS simulation,
+if we are confident that the bias potential
+is sufficiently accurate,
 will we be better off switching to the equilibrium FDS method
-to avoid possible efficiency loss due to the adaptive updates?
+to prevent possible efficiency loss
+due to the adaptive updates?
 %
@@
     \right\rangle.
+\notag
 \end{align}
@@
   =
-\frac{ 1 } { t }
-\sum_{ t_1 = 1 } ^ t
-  \eta_k( t_1 ),
+  \frac{ 1 } { T }
+  \sum_{ t = 1 } ^ T
+    \eta_k( t )
+  ,
+\notag
 \end{align}
@@
 \notag
-\label{eq:error_sbin_equil}
+%\label{eq:error_sbin_equil}
 \end{align}
@@
 &=
-\frac{ 1 } { t^2 }
-\sum_{ t_1, t_2 = 1 }^n
+\frac{ 1 } { T^2 }
+\sum_{ t, t' = 1 }^T
   \sum_{ k = 1 }^n
-    \eta_k( t_1 ) \, \eta_k( t_2 )
-\\
-&=
-\frac{ 1 } { t^2 }
-\sum_{ t_1, t_2 = 1 }^n
+    \eta_k( t ) \, \eta_k( t' )
+\\
+&=
+\frac{ 1 } { T^2 }
+\sum_{ t, t' = 1 }^T
   \sum_{ i, j = 1 }^n
     \left[
-      \frac{ h_i(t_1) }
+      \frac{ h_i(t) }
            { p_i    }
@@
     \left[
-      \frac{ h_j(t_2) }
+      \frac{ h_j(t') }
            { p_j     }
@@
 &=
-\frac{ 1 } { t^2 }
-\sum_{ t_1, t_2 = 1 }^n
+\frac{ 1 } { T^2 }
+\sum_{ t, t' = 1 }^T
   \sum_{ i, j = 1 }^n
     \left[
-      \frac{ h_i(t_1) }
+      \frac{ h_i(t) }
            { p_i    }
@@
 \sum_{ i = 1 }^n
-\frac{ 1 } { t }
-\sum_{ t_1 = 1 }^n
+\frac{ 1 } { T }
+\sum_{ t = 1 }^T
     \left[
-      \frac{ h_i(t_1) }
+      \frac{ h_i(t) }
            { p_i    }
@@
     \right]
-\frac{ 1 } { t }
-\sum_{ t_2 = 1 }^n
+\frac{ 1 } { T }
+\sum_{ t' = 1 }^T
     \left[
-      \frac{ h_i(t_2) }
+      \frac{ h_i(t') }
            { p_i     }
@@
 $$
-\eta_1(t_1)
+\eta_1(t)
 =
 \sum_{ i = 1 }^n
-  \phi_{1i} \, \frac{ h_i(t_1) - p_i } { p_i }
-=
-\sum_{ i = 1 }^n h_i(t_1)
-- \sum_{ i = 1 }^n p_i = 0.
+  \phi_{1i} \, \frac{ h_i(t) - p_i } { p_i }
+=
+\sum_{ i = 1 }^n h_i(t)
+-
+\sum_{ i = 1 }^n p_i = 0.
 $$
@@
 &=
-\frac{1}{t^2}
-\sum_{ t_1, t_2 = 1 }^n
-\langle \eta_k(t_1) \, \eta_k(t_2) \rangle
+\frac{1}{T^2}
+\sum_{ t, t' = 1 }^T
+\langle \eta_k(t) \, \eta_k(t') \rangle
 \\
 &\approx
-\frac{1}{t}
+\frac{1}{T}
 \sum_{ \tau = -\infty }^{ \infty }
-\langle \eta_k(t_1) \, \eta_k(t_1 + \tau) \rangle
-\\
-&=
-\frac{1 }{t}
+\langle \eta_k(t) \, \eta_k(t + \tau) \rangle
+\\
+&=
+\frac{1}{T}
 \int_{-\infty}^\infty
@@
 =
-\frac{1}{t}
+\frac{1}{T}
 \Gamma_k.
@@
 Thus,
-\begin{align}
+\begin{equation}
   E
   =
-\frac{ 1 } { t }
+  \frac{ 1 } { T }
   \sum_{ k = 2 }^n \Gamma_k
-\end{align}
+  .
+\label{eq:error_equil}
+\end{equation}
 
@@
 in the long time limit.
+
+Again, if we are certain that
+the intrinsic PMF is sufficiently smooth,
+then we can explicitly filter some error modes,
+say those for $k > K$,
+in the histogram, and the sum in
+Eq. \eqref{eq:error_equil}
+can be truncated at $k = K$.
 
@@
     \label{fig:err_nnbr}
-    Normalized error, $(t + t_0) \, E$,
+    Normalized error, $(T + t_0) \, E$,
     versus the %proportionality constant,
@@
     \label{fig:err_sig5}
-    Normalized error, $(t + t_0) \, E$,
+    Normalized error, $(T + t_0) \, E$,
     versus the %proportionality constant,
@@
     \label{fig:err_sigscan}
-    Normalized error, $(t + t_0) \, E$,
+    Normalized error, $(T + t_0) \, E$,
     versus the width of the Gaussian updating scheme,
@@
 
-%\section{Acknowledgments}
-%
-%We thanks Dr. Y. Mei and Dr. J. Ma for helpful discussions.
+\section{Acknowledgments}
+
+We thanks Dr. Y. Mei and Dr. J. Ma for helpful discussions.
+
+\appendix
+
+\section{\label{sec:invt_alpha}
+Inverse-time schedules}
+
+The optimal schedule predicted
+by the direct solution of Eq. \eqref{eq:q_opt}
+is complicated for a general
+multiple-bin updating scheme.
+%
+Several previous studies\cite{marsili2006, barducci2008, dickson2011}
+have been focused on analytically simpler schedules.
+%
+In a similar spirit,
+we consider here a variant of
+the inverse-time formula, Eq. \eqref{eq:alpha_invt1}
+%
+\begin{equation}
+\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
+\label{eq:alpha_invtlambda}
+\end{equation}
+%
+with the proportionality constant, $\lambda$,
+to be optimized.
+%
+We shall compute the error computed from
+this variationally optimized schedule
+and compare it with that from the optimal schedule.
+%
+We shall show that the optimal parameter, $\lambda$,
+is sensitive to the simulation time, $T$,
+which parallels the gradual shift of
+the instantaneous eigenvalue,
+$\lambda(t)$,
+as discussed in Sec. \ref{sec:optschedule}.
+
+\subsection{\label{sec:invt_error}
+Error
+}
+
+Using Eq. \eqref{eq:alpha_invtlambda}
+in Eq. \eqref{eq:error_asym1} yields
+%
+\begin{align}
+\Err_A
+&=
+\frac{    1    }
+     { T + t_0 }
+\sum_{k = 2}^n
+  \frac{ \Gamma_k \, \nu_k^2 }
+       {    2 \, \nu_k - 1   }
+\left[
+  1 - \left(
+        1 + \frac{ T }{ t_0 }
+      \right)^{1 - 2 \, \nu_k}
+\right],
+\label{eq:error_asym_invt}
+\end{align}
+%
+where $\nu_k \equiv \lambda_k / \lambda$.
+%
+At long times, $T \to \infty$, we get
+$$
+\begin{aligned}
+  \Err_A
+  =
+  \sum_{k = 2}^n
+  \begin{dcases}
+    \frac{    1    }
+         { T + t_0 }
+    \frac{ \Gamma_k \, \nu_k^2 }
+         {   2 \, \nu_k - 1    }
+    &
+    \mathrm{if \;} 2 \, \nu_k > 1,
+    \\%[1em]
+    %
+    %
+    \frac{    \Gamma_k    }
+         { 4 \, (T + t_0) }
+    \ln \frac{ T + t_0 }
+             {   t_0   }
+    &
+    \mathrm{if \;} 2 \, \nu_k = 1,
+    \\
+    %
+    %
+    \frac{  t_0^{ 2 \, \nu_k  - 1}  }
+         { (T + t_0)^{ 2 \, \nu_k } }
+    \frac{ \Gamma_k \, \nu_k^2 }
+         {   1 - 2 \, \nu_k    }
+    &
+    \mathrm{if \;} 2 \, \nu_k < 1.
+  \end{dcases}
+\end{aligned}
+$$
+%
+The last two cases
+are asymptotically suboptimal
+as they decay slower than $1/T$.
+%
+Thus, by assuming exclusively the first case
+$\lambda < 2 \lambda_k$,
+the optimal $\lambda$
+satisfies $\lambda < 2 \min \lambda_k$,
+and can be identified as
+the smallest real root of
+%
+\begin{equation}
+\sum_{k = 2}^n
+\frac{ \lambda_k - \lambda }
+{ \left(2 - \lambda/ \lambda_k \right)^2 }
+= 0.
+\label{eq:optimal_lambda_approx}
+\end{equation}
+
+Under the approximation, we have
+$$
+\frac{ \lambda_k^2 }
+     { \lambda \, (2 \, \lambda_k - \lambda) }
+\ge 1,
+$$
+with the equality achieved only at $\lambda = \lambda_k$.
+For the equality to hold for all $k$'s,
+we must have
+$\lambda_2 = \dots = \lambda_n = \lambda$.
+%
+Now a matrix with uniform eigenvalues
+is a multiple of the identity matrix,
+which corresponds to the single-bin updating scheme.
+%
+The above result means that
+the single-bin updating scheme is best
+for the asymptotic convergence.
+%
+This result is discussed in more details in
+Sec. \ref{sec:optWL}.
+
+Real simulations, however,
+can be relatively short
+and need not enter
+the asymptotic regime.
+%
+Then, the residual error, $\Err_R$, is not negligible.
+%
+Note, however,
+since Eq. \eqref{eq:alpha_invtlambda}
+does not fix the endpoint value,
+%
+\begin{equation}
+q(T) = \frac{ 1 } { \lambda }
+\log\left(
+  \frac{ T + t_0 } { t_0 }
+\right),
+\label{eq:qt_invtlambda}
+\end{equation}
+%
+we have to include the residual error
+during the optimization.
+%
+From Eqs. \eqref{eq:error_res1} and \eqref{eq:qt_invtlambda},
+we get
+%
+\begin{equation}
+\Err_R
+=
+\sum_{k = 2}^n
+  \frac{ \Gamma_k \, \nu_k }
+       {        2 \, t_0   }
+  \left(
+      \frac{   t_0   }
+           { T + t_0 }
+   \right)^{ 2 \, \nu_k },
+\label{eq:error_res_invt}
+\end{equation}
+%
+assuming that at $t = 0$
+the system has achieved an equilibrium
+at a constant $\alpha_0$,
+such that the values of
+$\left\langle y_k^2(0) \right\rangle$
+can be computed from Eq. \eqref{eq:y2_eql}.
+%
+The total error is given by the sum of
+Eqs. \eqref{eq:error_res_invt} and \eqref{eq:error_asym_invt}.
+%
+\begin{align}
+\Err
+&=
+\Err_R + \Err_A
+\notag
+\\
+&=
+\sum_{ k = 2 }^n
+  \Err^f_k \,
+  \left\{
+    1
+    +
+    \frac{ 1 - \left[ t_0 / (T + t_0) \right]^{2 \, \nu_k - 1} }
+         {                 2 \, \nu_k - 1                      }
+%    \frac{       1        }
+%         { 2 \, \nu_k - 1 }
+%    \left[
+%      1 -
+%      \left(
+%        \tfrac{   t_0   }
+%              { T + t_0 }
+%      \right)^{2 \, \nu_k - 1}
+%    \right]
+  \right\}
+,
+\label{eq:error_invt}
+\end{align}
+%
+where
+$$
+\Err^f_k
+\equiv
+\left\langle
+  y_k^2
+\right\rangle_{ \alpha(t) }
+=
+\frac{  \Gamma_k \, \nu_k   }
+     {    2 \, (T + t_0)    }
+,
+$$
+is the equilibrium, or saturated, error of mode $k$
+at the terminal updating magnitude
+of $\alpha(T) = 1/[\lambda (T + t_0)]$.
+
+\note{Derivation of Eq. \eqref{eq:error_invt}:
+$$
+\begin{aligned}
+\Err_R
+&=
+\sum_{ k = 2 }^n
+  \Err^f_k \left( \frac{   t_0   }
+                       { T + t_0 }
+          \right)^{ 2 \nu_k - 1 }
+\\
+\Err_A
+&=
+\sum_{ k = 2 }^n
+  \frac{ 2 \, \Err^f_k \, \nu_k }
+       {    2 \, \nu_k - 1      }
+  \left[
+    1 -
+    \left(
+      \tfrac{   t_0   }
+            { T + t_0 }
+      \right)^{2 \, \nu_k - 1}
+  \right]
+\\
+&=
+\sum_{ k = 2 }^n
+  \Err^f_k \,
+  \left[
+    1 -
+    \left(
+      \tfrac{   t_0   }
+            { T + t_0 }
+      \right)^{2 \, \nu_k - 1}
+  \right]
+  +
+  \frac{   \Err^f_k     }
+       { 2 \, \nu_k - 1 }
+  \left[
+    1 -
+    \left(
+      \tfrac{   t_0   }
+           { T + t_0 }
+      \right)^{2 \, \nu_k - 1}
+  \right].
+\end{aligned}
+$$
+Adding the two yields Eq. \eqref{eq:error_invt}.
+}
+
+%We give several remarks on Eq. \eqref{eq:error_invt}.
+%
+
+Clearly, the final error is always greater than
+the saturated equilibrium error
+at the terminal $\alpha(T)$:
+$$
+E \ge E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
+$$
+because the second term in the braces
+of Eq. \eqref{eq:error_invt}
+is nonnegative no matter the sign of $2 \, \nu_k - 1$
+(in fact, it is a decreasing function
+that vanishes only at $\nu_k \to +\infty$).
+%
+For a set of nonidentical $\lambda_k$'s,
+the equality is reached only at $T = 0$.
+
+\subsection{\label{sec:optlambda}
+Optimal proportionality constant
+}
+
+As shown in Sec. \ref{sec:optschedule},
+for the optimal schedule,
+the instantaneous eigenvalue, $\lambda(t)$,
+is on average smaller
+in a longer simulation.
+%
+Intuitively,
+we expect the optimal $\lambda$
+that minimizes the error of the inverse-time schedule,
+Eq. \eqref{eq:error_invt},
+to be a kind of time average of $\lambda(t)$,
+and thus also to decrease with
+the simulation length, $T$.
+%
+Below we show this is indeed the case.
+
+Let us first rewrite Eq. \eqref{eq:error_invt} as
+%
+$$
+\Err
+=
+\sum_{ k = 2 }^n
+  \frac{    \Gamma_k    }
+       { 2 \, (T + t_0) }
+  G\left( \nu_k, \frac{ t_0 } { T + t_0} \right),
+$$
+%
+where the $k$th component is proportional to
+%
+$$
+G\left( \nu_k, r \right)
+=
+\nu_k
+\,
+\left(
+  1
+  +
+  \frac{ 1 - r^{2 \, \nu_k - 1} }
+       {   2 \, \nu_k - 1       }
+\right).
+$$
+%
+As shown in Fig. \ref{fig:err_component},
+the function $G(\nu_k, r)$ depends non-trivially on $\nu_k$.
+%
+For a fixed and sufficiently small $r = t_0 / (T + t_0)$,
+$G(\nu_k, r)$ has two local minima along $\nu_k$,
+one at $\nu_k = 0$,
+and the other shallower one around $\nu_k = 1$.
+%
+For a short simulation, or a large $r$,
+the minimal error may be achieved at $\nu_k \approx 0$,
+corresponding to a large value of $\lambda$
+($\lambda \gg \lambda_k$).
+%
+However, this minimum is unstable,
+for $G(\nu_k, r)$ increases rapidly
+with decreasing $r$, or increasing simulation length,
+under a fixed $\nu_k$.
+%
+In other words,
+asymptotically, the minimal error
+is always achieved at the other minimum,
+$\nu_k = 1$ or $\lambda = \lambda_k$.
+%
+In this case,
+$G(\nu_k, r) \to G(1, 0) = 2$
+and the total error is roughly twice $E_\mathrm{sat}$.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/errcomp.pdf}
+  }
+  \caption{
+    \label{fig:err_component}
+    Component of the error, $G(\nu_k, r)$,
+    versus $\nu_k = \lambda_k / \lambda$,
+    where $r = t_0 / (T + t_0)$.
+    %
+  }
+\end{center}
+\end{figure}
+
+For a set of nonidentical eigenvalues,
+$\lambda_2, \dots, \lambda_n$,
+it is impossible to achieve $\lambda = \lambda_k$
+for every $k$,
+and we shall argue that
+the optimal $\lambda$
+tends to be biased toward
+the low end of the spectrum
+in a long simulation.
+%
+From Fig. \ref{fig:err_component},
+it is seen that around the minimum at $\nu_k = 1$,
+the error function $G(\nu_k, r)$
+has a steeper slope for the $\nu_k < 1$
+(or $\lambda > \lambda_k$) side
+than for the $\nu_k > 1$ (or $\lambda < \lambda_k$) side.
+%
+This means a smaller $\lambda$ is preferred
+in terms of minimizing the error.
+%
+Further, the difference in slopes
+between the two sides increases
+with the simulation length, $T$
+(or as $r$ decreases).
+%
+We thus expect the optimal $\lambda$
+to shift toward the smaller values of $\lambda_k$'s,
+and ultimately to approach $\min \lambda_k$
+in the infinite $T$ limit.
+
+The shift of the optimal $\lambda$
+is also affected by the values of $\Gamma_k$'s.
+%
+If large $\lambda_k$'s are associated with
+large $\Gamma_k$'s,
+these modes weight more in the error function,
+and the above shift toward smaller $\lambda$
+slows down.
+%
+For example,
+in the nearest-neighbor or Gaussian updating scheme
+discussed in Sec. \ref{sec:band-matrix},
+the index $k$ represents the wave number
+(or the inverse wave length)
+and $\lambda_k$ decreases with $k$.
+%
+As shown in Eq. \eqref{eq:Gamma_onestep},
+in the one-step sampling process,
+$\Gamma_k$ also decreases with $k$,
+and thus, $\lambda_k$ and $\Gamma_k$
+are positively correlated.
+%
+Therefore, we expect the shift of the optimal $\lambda$
+in the above case
+to be slower than in that in the case of perfect sampling,
+where $\Gamma_k \equiv 1$ is a constant.
+%
+If $\Gamma_k$ decreases too quickly with $k$,
+the shift can be hard to detect
+[see Eq. \eqref{eq:lambda_nn_onestep} for an example].
+
+\subsection{Nearest-neighbor updating scheme}
+
+For the nearest-neighbor updating scheme
+introduced in Sec. \ref{sec:nnscheme},
+we can find an analytical expression
+of the optimal value of $\lambda$
+in the limit of large $n$.
+%
+In this case,
+we convert Eq.
+\eqref{eq:error_asym_invt}
+%and
+%\eqref{eq:optimal_lambda_approx}
+to an integral,
+$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
+%
+\begin{align}
+\Err_A
+=
+\frac{2 \, n}{\pi \, T}
+\int_0^{\pi/2}
+  \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
+       {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
+\, dp.
+\notag
+%\label{eq:error_nn_asym_int}
+\end{align}
+
+For perfect sampling,
+we have, from Eq. \eqref{eq:Gamma_perfect},
+$\Gamma(p) = 1$, and
+$$
+\begin{aligned}
+\Err_A
+&=
+\frac{2 \, n}{\pi \, T}
+\int_0^{\pi/2}
+\frac{ \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2 }
+{ \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
+\, dp
+\notag \\
+&=
+\frac{n}{4 \, T}
+\left(
+  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
+  +
+  \frac{ \lambda }
+  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
+\right).
+\end{aligned}
+$$
+\note{The integral is evaluated by contour integration.
+%
+It is somewhat more convenient to change variable $p \to \frac{ \pi } { 2 } - p$,
+and
+$$
+\begin{aligned}
+\Err_A
+&=
+\frac{n}{\lambda \, T}
+\frac{1}{2 \pi i}
+\int_0^{2 \, \pi}
+\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
+{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
+\, dp
+\\
+&=
+\frac{n}{\lambda \, T}
+\frac{1}{2 \pi i}
+\oint
+\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
+{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
+\, \frac{dz}{z},
+\end{aligned}
+$$
+where the contour is the unit circle
+in the complex plane of $z$.
+%
+This integral has five poles, one at the origin,
+which produces a residue
+$\frac{2 - 4 \, \mu_1 - \lambda}{4}$
+(from series expansion of small $z$),
+and the other four at
+$$
+z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
+$$
+with two of them inside the unit circle:
+$$
+z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
+{\sqrt{8 \mu_1}}.
+$$
+Thus,
+$$
+\begin{aligned}
+\Err_A
+&=
+\frac{      n       }
+     { \lambda \, T }
+\left(
+ \frac{ 2 - 4 \, \mu_1 - \lambda }
+      {          4               }
+ +
+ \sum_{z = z_{\pm} }
+ \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
+ { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
+\right)
+\\
+&=
+\frac{       n      }
+     { \lambda \, t }
+\left(
+  \frac{ 2 - 4 \, \mu_1 - \lambda }
+       {          4               }
+ +
+ \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
+      { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2                 }
+\right)
+\\
+&=
+\frac{       n      }
+     { \lambda \, T }
+\left(
+  \frac{ 2 - 4 \, \mu_1 - \lambda }
+       {          4               }
+ +
+ \frac{ (\lambda/2)^2 }
+ { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
+\right).
+\end{aligned}
+$$
+}
+%
+Minimizing the function yields
+%
+\begin{equation}
+\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
+\label{eq:lambda_nn_perfect}
+\end{equation}
+%
+and the asymptotic error in the optimal case is given by
+%
+\begin{equation}
+\Err_A
+=
+\frac{n}{T}
+\left(
+  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
+\right).
+\label{eq:error_nn_prefect}
+\end{equation}
+%
+%From Eq. \eqref{eq:error_nn}, it is clear that
+Clearly, the error
+increases with the magnitude of $\mu_1$
+no matter its sign,
+and the minimal value, obtained at $\mu_1 = 0$,
+corresponds to the single-bin scheme.
+%
+Note that we have ignored the residual error
+and assumed $\lambda < 2 \, \min \lambda_k = 2 - 8 \, \mu_1$
+for all $k$ in the above solution.
+%
+Fortunately,
+the $\lambda$ from Eq. \eqref{eq:lambda_nn_perfect}
+satisfies this condition
+by the stability condition,
+$4 \, \mu_1 < 1$.
+\note{This is because,
+$$
+\frac{ 1 - 4 \, \mu_1 } { 1 - 2 \, \mu_1 } < 2 - 8 \, \mu_1
+\quad \leftrightarrow \quad
+1 < 2 - 4 \, \mu_1.
+$$
+}
+
+For the one-step sampling,
+we find from Eq. \eqref{eq:Gamma_onestep}
+that $\Gamma(p) = \cot^2 p$
+is peaked around $p \approx 0$,
+%
+Then
+$$
+\begin{aligned}
+\Err_A
+&
+\propto
+\frac{   2 \, n }
+     { \pi \, T }
+\left.
+\frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
+     { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
+\right|_{ p \approx 0 }
+\notag \\
+&
+\approx
+\frac{   2 \, n }
+     { \pi \, T }
+\frac{             1             }
+     {  \lambda \, (2 - \lambda) }
+.
+\end{aligned}
+$$
+%
+The optimal $\lambda$ is therefore
+%
+\begin{equation}
+\lambda \approx 1.
+\label{eq:lambda_nn_onestep}
+\end{equation}
+%
+\note{We can be a bit more precise here.
+%
+From Eq. \eqref{eq:Gamma_onestep}, we get
+%
+$$
+\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
+$$
+%
+where we have introduced
+$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
+as a small parameter to avoid the divergence
+around the origin.
+%
+$$
+\begin{aligned}
+\Err_A
+&=
+\frac{   2 \, n }
+     { \pi \, t }
+\int_0^{\pi/2}
+    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
+         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
+    \frac{ \cos^2 p }
+         { \sin^2 p + \delta^2 }
+\, dp
+\notag \\
+&
+\stackrel{    \delta \ll 1     }
+         { =\joinrel=\joinrel= }
+\frac{   2 \, n }
+     { \pi \, t }
+\frac{             1             }
+     {  \lambda \, (2 - \lambda) }
+\frac{    1   }
+     { \delta }
+.
+\end{aligned}
+$$
+The integral is evaluated as follows.
+%
+We again change variable $p \to \frac{ \pi } { 2 } - p$,
+and
+$$
+\begin{aligned}
+\Err_A
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\int_0^{2 \, \pi}
+\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
+     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
+\cdot
+\frac{ \sin^2 p            }
+     { \cos^2 p + \delta^2 }
+\, dp
+\\
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\oint
+\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
+     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
+\cdot
+\frac{ 4 - \left( z + \frac 1 z \right)^2 }
+     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
+\, \frac{dz}{z}
+\\
+&=
+\frac{n}{\lambda \, t}
+\frac{1}{2 \pi i}
+\oint
+\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
+     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
+\cdot
+\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
+     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
+\, \frac{ dz }{ z^3 }
+\\
+&=
+\frac{ n } { \lambda \, t }
+\sum_l \operatorname{Res}_l
+,
+\end{aligned}
+$$
+where the contour is the unit circle
+in the complex plane of $z$,
+and the integral is reduced to a sum of residues
+around the poles within.
+
+The poles of the integral come from three sources,
+namely,
+% 1.
+$z = 0$,
+% 2.
+the zeros of
+$(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2$,
+% 3.
+and the zeros of
+$( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2$.
+
+For the first pole around the origin,
+we need to expand the integrand as
+$$
+\frac{ A + B \, z^2 + \cdots }
+     {          z^3          },
+$$
+then the residue is $B$.
+%
+Since
+$$
+\begin{aligned}
+\left[ z^2 - \mu_1 (z^2 + 1)^2 \right]^2
+&
+\approx
+[ - \mu_1 + (1 - 2 \, \mu_1) \, z^2]^2
+\\
+&
+\approx
+\mu_1^2
+\left[
+  1 + \left(4 - \tfrac { 2  } { \mu_1 } \right) z^2
+\right],
+\\
+%
+4 \, z^2 - ( z^2 + 1 )^2
+&
+\approx
+-( 1 - 2 \, z^2 ),
+\\
+%
+(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2
+&
+\approx
+-2 \, \mu_1 \,
+\left[
+  1 + \left(
+        2 + \tfrac{ \lambda - 2 } { 2 \, \mu_1 }
+      \right)
+      \, z^2
+\right],
+\\
+%
+( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2
+&
+\approx
+1 + (2 + 4 \, \delta^2) \, z^2,
+\end{aligned}
+$$
+the integrand is
+$$
+\begin{aligned}
+\frac{ \mu_1 }
+     {   2   }
+\left[
+  1 + \left(
+        - \frac{ 2 + \lambda } { 2 \, \mu_1 }
+        - 2 - 4 \, \delta^2
+      \right) \, z^2
+\right].
+\end{aligned}
+$$
+and the residue is
+$$
+\operatorname{Res}_1
+=
+  - \frac{ 2 + \lambda } { 4 }
+  - \mu_1 \, \left( 1 + 2 \, \delta^2 \right).
+$$
+
+For the second part,
+we have four roots of
+$$
+(2 - \lambda) \, z^2 = 2 \, \mu_1 \, (z^2 + 1)^2,
+$$
+which are
+\begin{equation}
+z
+=
+\pm
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } }
+\pm
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } - 1 }.
+\label{eq:nnint_local_zeros_part2}
+\end{equation}
+We share distinguish two cases.
+\textbf{Case A.}
+If $\lambda < 2 - 8 \, \mu_1$,
+two of the roots lie in the unit circle, namely
+$$
+z
+=
+\pm
+\left(
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } }
+-
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } - 1 }
+\right),
+$$
+which satisfy
+$$
+4 \, \mu_1 \, ( z^2 + 1)
+=
+2 - \lambda
+-
+\sqrt{ ( 2 - \lambda ) ( 2 - \lambda - 8 \, \mu_1 ) }.
+$$
+Thus,
+$$
+\begin{aligned}
+\left[ z^2 - \mu_1 \, \left( z^2 + 1 \right)^2 \right]^2
+&=
+\left( z^2 - \tfrac{ 2 - \lambda } { 2 } z^2 \right)^2
+=
+\frac{ \lambda^2 } { 4 } z^4.
+\\
+%
+4 z^2 - \left( z^2 + 1 \right)^2
+&=
+-\frac{ 2 - \lambda - 8 \, \mu_1  } { 2 \, \mu_1 } z^2,
+\\
+%
+\left( z^2 + 1 \right)^2 + 4 \, \delta^2 \, z^2
+&=
+\frac{ 2 - \lambda + 8 \, \mu_1 \, \delta^2 }
+     { 2 \, \mu_1 }
+     z^2,
+\\
+%
+(2 - \lambda) \, 2 \, z
+-
+8 \, \mu_1 \, \left( z^2 + 1 \right) \, z
+&=
+2 \, z \, \left[2 - \lambda - 4 \,  \mu_1 \, (z^2 + 1) \right]
+\\
+&= 2 \, z \, \sqrt{ (2 - \lambda) ( 2 - \lambda - 8 \, \mu_1 ) },
+\end{aligned}
+$$
+where we have used the L'H\^{o}pital's rule
+and taken the derivative for the last factor.
+%
+Note that the two roots within the unit share the same
+residue and the sum for this part is
+%
+$$
+\operatorname{Res}_2
+=
+-\frac{                 \lambda^2                     }
+      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
+\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }.
+$$
+%
+\textbf{Case B.}
+If $2 - 8 \, \mu_1 \le \lambda < 2$,
+the four zeros in Eq. \eqref{eq:nnint_local_zeros_part2}:
+%
+\begin{equation}
+z
+=
+\pm
+\sqrt { \frac{ 2 - \lambda }
+             { 8 \, \mu_1  } }
+\pm i \,
+\sqrt { 1 - \frac{ 2 - \lambda }
+                 { 8 \, \mu_1  } },
+\label{eq:nnint_local_zeros_part2b}
+\end{equation}
+%
+all satisfy $|z| = 1$, and hence
+all lie on the border of the unit circle.
+%
+Thus, each contributes half of the residue there.
+%
+However, since the four zeros satisfy
+$$
+2 - \lambda - 4 \, \mu_1 \, ( z^2 + 1)
+=
+\pm i
+\sqrt{ ( 2 - \lambda ) ( 8 \, \mu_1 - 2 + \lambda ) },
+$$
+the sum of the four residues gives zero in this case:
+$$
+\operatorname{Res}_2 = 0.
+$$
+In summary,
+$$
+\operatorname{Res}_2
+=
+\begin{dcases}
+-\frac{                 \lambda^2                     }
+      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
+\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }
+&
+\mathrm{if \;}
+\lambda < 2 - 8 \, \mu_1,
+\\
+0
+&
+\mathrm{otherwise}.
+\end{dcases}
+$$
+
+For the last part,
+we have four roots of
+$(z^2 + 1)^2 = - 4 \, \delta^2 \, z^2$,
+with two of them lying in the unit circle:
+$$
+z = \pm i \left( \sqrt{ 1 + \delta^2 } - \delta \right),
+$$
+satisfying
+$$
+z^2 + 1 + 2 \, \delta^2 = 2 \, \delta \, \sqrt{1 + \delta^2}.
+$$
+Then,
+$$
+\begin{aligned}
+z^2 - \mu_1 \, \left( z^2 + 1 \right)^2
+&= z^2 \, (1 + 4 \, \mu_1 \, \delta^2),
+\\
+4 \, z^2 - \left( z^2 + 1 \right)^2
+&=
+4 \, z^2 \, (1 + \delta^2)
+\\
+%
+(2 - \lambda) \, z^2
+- 2 \, \mu_1 \, \left( z^2 + 1 \right)^2
+&=
+\left(
+  2 - \lambda + 8 \, \mu_1 \, \delta^2
+\right) \, z^2
+\\
+%
+4 \, z \, \left( z^2 + 1 \right)
++
+8 \, \delta^2 \, z
+&=
+4 \, z \, (z^2 + 1 + \delta^2)
+=
+8 \, z \, \delta \, \sqrt{ 1 + \delta^2 }.
+\end{aligned}
+$$
+Note that the above two roots share the same
+residue and the sum for this part is
+$$
+\operatorname{Res}_3
+=
+\frac{ \left( 1 + 4 \, \mu_1 \, \delta^2 \right)^2 }
+     {        2 + 8 \, \mu_1 \, \delta^2 - \lambda }
+\frac{ \sqrt{ 1 + \delta^2 } }
+     {        \delta         }.
+$$
+
+Now as $\delta \to 0$, the last contribution will dominate,
+and we have
+$$
+\Err_A
+\approx
+\frac{ n } { T }
+\frac{ (1 + 4 \, \mu_1 \, \delta^2)^2 }
+{ \lambda \, (2 + 8 \, \mu_1 \, \delta^2 - \lambda) }
+\frac{ 1 } { \delta }
+\approx
+\frac{ n } { T }
+\frac{ 1 }
+{ \lambda \, (2 - \lambda) }
+\frac{ 1 } { \delta },
+$$
+and the minimum is reached at $\lambda \approx 1$.
+
+The weakness of the above calculation is two fold.
+%
+First, it ignores the residual error.
+%
+Second, the simplified expression for the asymptotic error
+$E_A$ may also be misused.
+%
+Namely, with $\lambda \approx 1$,
+we cannot assume $\lambda < 2 \, \lambda_k$ in general.
+%
+So it only works for a very small value of $\mu_1$.
+}
+
+Note that Eq. \eqref{eq:lambda_nn_onestep}
+is an approximation good only
+for sufficiently short simulations.
+%
+For very longer simulations,
+the optimal $\lambda$
+will ultimately approach
+the smallest eigenvalue,
+$\min \lambda_k = 1 - 4 \, \mu_1$
+according to the discussion in Sec. \ref{sec:optlambda}.
 

### 2016-03-30 doc/invt.tex
intq.h: variate qt for the optimal schedule

@@
 
-We tested the above theory on a one-dimensional system
+Our test system is
+a model one-dimensional system
 of $n = 100$ bins,
 with a flat target distribution
-$p_i = 1/n$.
+$p_i = 1/n$
+under the reflective boundary condition.
 %
@@
 %
-and let the system undergo an equilibration period
-of constant updating magnitude,
+and equilibrate the system under
+a constant updating magnitude,
 $\alpha(t) = \alpha_0$.
@@
 we reset the origin of the time, $t$,
-and started the schedule specified by
-either
-Eq. \eqref{eq:q_opt}
-or
-Eq. \eqref{eq:alpha_invtlambda}.
-%
-We use the latter case to verify our theory.
-%
-We wish to see how the final error
-depends on the parameter, $\lambda$,
-and if it matches the predicted value.
-%
-Here, we used $t_0 = 1/(\lambda \, \alpha_0)$
-to smooth the transition from equilibration.
-
-Since the error depends on the underlying sampling scheme,
-we have applied the above protocol to two
-sampling schemes,
+and started the target schedule.
+%specified by either
+%Eq. \eqref{eq:q_opt}
+%or
+%Eq. \eqref{eq:alpha_invtlambda}.
+
+For the Gaussian updating scheme,
+the kernel is truncated at
+$\min\{10 \, \sigma, n - 1\}$
+and numerically stabilized
+using the technique
+in Sec. \ref{sec:stabilize_wband}.
+
+Since the error depends on
+the underlying sampling process,
+we have devised two sampling processes,
 a global one and a local one.
 %
-In both cases, we used the
-Metropolis algorithm\cite{metropolis1953, newman, frenkel, landau_binder}
-to sample the state space:
+Both cases were based on
+Metropolis algorithm\cite{metropolis1953, newman, frenkel,
+landau_binder}:
 %
@@
 However,
-in the first sampling scheme,
+in the first and global sampling process,
 we choose the destination $j$
@@
 %
-This sampling scheme,
-referred to this as the global sampling scheme,
-emulates the perfect sampling scheme
+This sampling process
+emulates the perfect sampling process
 discussed in Sec. \ref{sec:Gamma}
@@
 %
-In the second scheme,
+In the second or local process,
 we limit the destination $j$
@@
 $i \pm 1$, with equal probability $1/2$,
-to mimic the one-step sampling scheme
+to mimic the one-step sampling process
 in Sec. \ref{sec:Gamma}.
 %
-This scheme,
-referred to this as the local sampling scheme,
-models local sampling processes
-typical in MD and straightforward MC simulations.
-%
-
-\subsection{Single-bin (WL) updating scheme}
-
-We first tested the single-bin (WL) updating scheme
-using the above protocol.
-%
-Here, we chose $\alpha_0 = 0.0001$
+
+\subsection{Inverse-time schedule}
+
+We first tested the theory using
+the inverse-time schedule,
+%
+\begin{equation}
+\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
+\label{eq:alpha_invtlambda}
+\end{equation}
+%
+where $\lambda$ is a parameter to be optimized.
+and $t_0 = 1/(\lambda \, \alpha_0)$
+to smooth the transition from equilibration.
+%
+This schedule is analytically simpler
+than the optimal schedule,
+and it is studied in Appendix \ref{sec:invt_schedule}.
+%
+We wish to see how the final error
+depends on the parameter, $\lambda$,
+and if it matches the predicted value,
+given by Eq. \eqref{eq:error_invt}.
+
+\subsubsection{Single-bin (WL) updating scheme}
+
+For the single-bin updating scheme,
+we chose $\alpha_0 = 0.0001$
 and ran the simulation for $t = 10^8$ steps.
@@
 %
-For both global and local sampling schemes,
+For both global and local sampling processes,
 the optimal $\lambda$ occurred around $1.0$,
-the optimal value in the asymptotic regime.
-%
-Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
-represents the relative updating magnitude.
-%
-Figure \ref{fig:err_singlebin} also shows that
-a smaller than optimal value of $1/\lambda$
-tends to increase the error more rapidly
-than a larger than optimal value,
-suggesting that the latter (over-updating)
-is generally preferable to the latter (under-updating)
-when the optimal value is unknown.
-
-Another observation is that the two error functions
-differed only by a multiplicative constant.
-%
-This again agreed with the theory.
-%
-Since all eigenvalues of the single-bin scheme are
-the same, $\lambda_2 = \cdots = \lambda_n = 1$,
+verifying Eq. \eqref{eq:alpha_invt}.
+%
+%Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
+%represents the relative updating magnitude.
+%%
+%Figure \ref{fig:err_singlebin} also shows
+%an asymmetry of the error curve.
+%A smaller than optimal value of
+%the updating magnitude, $1/\lambda$,
+%tends to increase the error more rapidly
+%than a larger than optimal value,
+%suggesting that the over-updating
+%is generally preferable to the under-updating
+%when the optimal value is unknown.
+
+We also observe
+that the two error functions
+differed only by a multiplicative constant,
+again again agreeing with the theory.
+%
+With identical eigenvalues,
+$\lambda_1 = \cdots = \lambda_n = 1$,
 the error in Eq. \eqref{eq:error_invt}
-depends on the sampling scheme only through
+depends on the sampling process only through
 the sum $\sum_{ k = 2 }^n \Gamma_k$,
@@
 
-\subsection{Nearest-neighbor updating scheme}
-
-We now turn to the nearest-neighbor updating scheme.
+\subsubsection{Nearest-neighbor updating scheme}
+
+As an example of the multiple-bin schemes,
+we chose the nearest-neighbor updating scheme
+described in Sec. \eqref{sec:nnscheme}.
 %
@@
 A new feature is that the optimal $\lambda$
-depends on the simulation length, $t$.
+depends on the simulation length, $T$.
 %
 As shown in Fig. \ref{fig:err_nnbr},
-the normalized error, $(t + t_0) \, \Err$,
+the normalized error, $(T + t_0) \, \Err$,
 hence the optimal value of $1/\lambda$,
-depended non-trivially on the simulation length
+depended non-trivially on the simulation length, $T$,
 in the global sampling case.
@@
 %
-For the local sampling case, however,
-the normalized error roughly maintained the sample shape,
-and the optimal $\lambda$ is around $1.0$,
+For the local sampling case,
+the $T$-dependence was hardly visible
+because of the predominantly large $\Gamma_k$'s
+of the small $k$ modes.
+%
+The optimal $\lambda$ remained around $1.0$,
 as suggested by Eq. \eqref{eq:lambda_nn_onestep}.
 
-To understand this dependence,
-we observe that for shorter simulations,
-long-wavelength modes
-(with larger $\lambda_k$ and smaller $k$),
-dominate the error function in Eq. \eqref{eq:error_invt}
-because of the prefactor $\lambda_k \, \Gamma_k$,
-which is inherited from
-the equilibrium error under a fixed updating magnitude,
-Eq. \eqref{eq:error_eql}.
-%
-For longer simulations,
-short-wavelength modes
-(with smaller $\lambda_k$ and larger $k$)
-become more important,
-for $\lambda_k$ also gives the decay rate
-of the $k$th mode.
-%and errors in smaller $\lambda_k$ modes
-%are harder to damp out.
-%
-This drives the optimal $\lambda$ to shift toward
-a smaller $\lambda_k$ value
-as the simulation lengthens.
-%
-Note also that while increasing $\lambda$
-also increases the error of the long-wavelength modes,
-this rate of increase is much smaller,
-as shown in the $\nu_k > 1$ branch
-in Fig. \ref{fig:err_component}.
-%
-The shift of $\lambda$, however, is less noticeable
-under the local sampling scheme,
-because longer-wavelength modes more heavily
-weighted by much larger values of $\Gamma_k$.
+%To understand this dependence,
+%we observe that for shorter simulations,
+%long-wavelength modes
+%(with larger $\lambda_k$ and smaller $k$),
+%dominate the error function in Eq. \eqref{eq:error_invt}
+%because of the prefactor $\lambda_k \, \Gamma_k$,
+%which is inherited from
+%the equilibrium error under a fixed updating magnitude,
+%Eq. \eqref{eq:error_eql}.
+%%
+%For longer simulations,
+%short-wavelength modes
+%(with smaller $\lambda_k$ and larger $k$)
+%become more important,
+%for $\lambda_k$ also gives the decay rate
+%of the $k$th mode.
+%%and errors in smaller $\lambda_k$ modes
+%%are harder to damp out.
+%%
+%This drives the optimal $\lambda$ to shift toward
+%a smaller $\lambda_k$ value
+%as the simulation lengthens.
+%%
+%Note also that while increasing $\lambda$
+%also increases the error of the long-wavelength modes,
+%this rate of increase is much smaller,
+%as shown in the $\nu_k > 1$ branch
+%in Fig. \ref{fig:err_component}.
+%%
+%The shift of $\lambda$, however, is less noticeable
+%under the local sampling scheme,
+%because longer-wavelength modes more heavily
+%weighted by much larger values of $\Gamma_k$.
 
@@
 
-\subsection{Gaussian (metadynamics) updating scheme}
-
-The error of Gaussian updating scheme is similar to
-that of the nearest neighbor scheme,
-as shown in Fig \ref{fig:err_sig5}.
-%
-The dependency of the simulation length is more prominent.
-%
-Even for local sampling,
-there is now a visible shift of the optimal $\lambda$.
+%\subsection{Gaussian (metadynamics) updating scheme}
+%
+%
+%The error of Gaussian updating scheme is similar to
+%that of the nearest neighbor scheme,
+%as shown in Fig \ref{fig:err_sig5}.
+%%
+%The dependency of the simulation length is more prominent.
+%%
+%Even for local sampling,
+%there is now a visible shift of the optimal $\lambda$.
+%
+%
+%\begin{figure}[h]
+%\begin{center}
+%  \makebox[\linewidth][c]{
+%    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsig5.pdf}
+%  }
+%  \caption{
+%    \label{fig:err_sig5}
+%    Normalized error, $(T + t_0) \, E$,
+%    versus the %proportionality constant,
+%    relative updating magnitude,
+%    $1/\lambda$,
+%    in Eq. \eqref{eq:alpha_invtlambda}
+%    for the Gaussian updating scheme with $\sigma = 5$ bins.
+%    %
+%    The results have been averaged over $1000$ independent runs.
+%  }
+%\end{center}
+%\end{figure}
+%
+
+\subsection{Optimal schedule}
+
+We now turn to the optimal schedule
+predicted by Eq. \eqref{eq:q_opt}.
+%
+For the Gaussian updating scheme
+with $\sigma = 10$,
+we computed the schedule,
+and compared it with the inverse-time schedule
+Eq. \eqref{eq:alpha_invtlambda}
+with the optimal $\lambda$.
 
@@
   \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsig5.pdf}
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/optacmp.pdf}
   }
   \caption{
-    \label{fig:err_sig5}
-    Normalized error, $(T + t_0) \, E$,
-    versus the %proportionality constant,
-    relative updating magnitude,
-    $1/\lambda$,
-    in Eq. \eqref{eq:alpha_invtlambda}
-    for the Gaussian updating scheme with $\sigma = 5$ bins.
-    %
-    The results have been averaged over $1000$ independent runs.
-  }
-\end{center}
-\end{figure}
-
-\subsection{Accuracy of the inverse-time schedule}
-
-In the above, we have assumed the inverse-time schedule,
-Eq. \eqref{eq:alpha_invtlambda}.
-%
-In Fig. \ref{fig:alpha_cmp},
-we compared the inverse-time schedule
-with the optimal $\lambda$,
-and the optimal schedule from integrating
-Eq. \eqref{eq:q_opt}.
-%
-The difference was larger for the global sampling scheme.
-%
-As predicted by
-Eqs. \eqref{eq:dinvadt_limit1} and \eqref{eq:dinvadt_limit2},
-the rate of change of $1/\alpha(t)$
-indeed increases from $\lambda_{\min}$
-to $\bar \lambda \approx 1$.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/dinvacmp.pdf}
-  }
-  \caption{
-    \label{fig:alpha_cmp}
-    Optimal schedules, in terms of $\dot \alpha(t)$, from Eq. \eqref{eq:q_opt},
-    versus the inverse-time schedule,
-    Eq. \eqref{eq:alpha_invtlambda},
-    with the optimal parameter, $\lambda$.
-    %
-    For fair comparison,
-    the $q(T) = \int_0^t \alpha(t) \, dt$
+    \label{fig:optacmp}
+    Optimal schedules from Eq. \eqref{eq:q_opt},
+    versus the optimized inverse-time schedule
+    Eq. \eqref{eq:alpha_invtlambda}
+    for the Gaussian updating scheme
+    with $\sigma = 10$.
+    %
+    $q(T) = \int_0^t \alpha(t) \, dt$
     were set to be the same
@@
 \end{figure}
+
+As shown in Fig. \ref{fig:optacmp}(a),
+the optimal schedule has a more complex curvature
+than the optimized inverse-time schedule.
+%
+In Fig. \ref{fig:optacmp}(b),
+we show that in the former case,
+the rate of change of $\alpha^{-1}(t)$
+increases gradually
+from $\lambda(0)$ to $\lambda(T) \approx 1$,
+as discussed in Sec. \ref{sec:optschedule}.
+%
+The errors in the two cases were
+$4.0 \times 10^{-7}$
+and
+$5.9 \times 10^{-7}$,
+respectively,
+for the global sampling process,
+%
+or
+$1.67 \times 10^{-4}$
+and
+$1.26 \times 10^{-4}$,
+respectively
+for the local sampling process,
+suggesting a slight edge of
+the optimal schedule from Eq. \eqref{eq:q_opt}.
+
+\subsection{Comparison of updating schemes}
 
@@
 
-\section{\label{sec:invt_alpha}
+\section{\label{sec:invt_schedule}
 Inverse-time schedules}
@@
 In a similar spirit,
-we consider here a variant of
-the inverse-time formula, Eq. \eqref{eq:alpha_invt1}
-%
-\begin{equation}
-\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
-\label{eq:alpha_invtlambda}
-\end{equation}
+we consider here
+the inverse-time formula,
+Eq. \eqref{eq:alpha_invtlambda},
 %

### 2016-03-30 prog/intq.h

@@ -9,12 +9,11 @@
 
 typedef struct {
   int m; /* number of integration point */
-  double t; /* number of times */
-  double qt;
-  double *tarr;
+  double t; /* total simulation time */
+  double *tarr; /* time grid */
   double *qarr;
   double *aarr;
-  double *dinva;
+  double *dinva; /* instantaneous eigenvalue */
   int n;
   const double *lambda; /* eigenvalues of the updating magnitude */
   const double *gamma; /* autocorrelation integrals */
@@ -23,7 +22,7 @@ typedef struct {
 
 
 
-static intq_t *intq_open(double t, double qt, int m,
+static intq_t *intq_open(double t, int m,
     int n, const double *lambda, const double *gamma)
 {
   intq_t *intq;
@@ -35,7 +34,6 @@ static intq_t *intq_open(double t, double qt, int m,
   xnew(intq->qarr, m + 1);
   xnew(intq->dinva, m + 1);
   intq->t = t;
-  intq->qt = qt;
   intq->n = n;
   intq->lambda = lambda;
   intq->gamma = gamma;
@@ -59,10 +57,10 @@ static void intq_close(intq_t *intq)
 
 
 /* compute the integrand for `intq` */
-static double intq_getvel(intq_t *intq, double q)
+static double intq_getvel(intq_t *intq, double dq)
 {
   int i, n = intq->n;
-  double lam, y = 0, dq = q - intq->qt;
+  double lam, y = 0;
 
   for ( i = 1; i < n; i++ ) {
     lam = intq->lambda[i];
@@ -74,26 +72,23 @@ static double intq_getvel(intq_t *intq, double q)
 
 
 
-/* compute the optimal protocol
- * for a period t, and fixed
- * qt = Integral {from 0 to t} alpha(t') dt'
- * the results are saved to tarr[0..m], qarr[0..m], and aarr[0..m]
- * return the square-root asymptotic error
+/* compute the optimal schedule, in terms of q(t)
+ * the results are saved to tarr[0..m], qarr[0..m]
  * */
-static void intq_geta(intq_t *intq)
+static void intq_getq(intq_t *intq, double qt)
 {
   double c, dq, y;
   int j, m = intq->m;
 
   /* integrate over q over uniform grid */
-  dq = intq->qt / m;
[…]

### 2016-03-30 prog/invt.h

@@ -472,9 +472,9 @@ static double estbestc_invt(double t, double a0,
       }
     }
 
-    /* find the minimal value */
+    /* update the bracket */
     if ( el < em && el < er ) {
-      /* el is the minimal of the three, extend to the left */
+      /* el is the least of the three, extend to the left */
       cr = cm;
       cm = cl;
       cl = cl * 0.5; /* make sure cl > 0 */
@@ -482,7 +482,7 @@ static double estbestc_invt(double t, double a0,
       em = el;
       el = esterror_invt(t, cl, a0, n, NULL, lambda, gamma);
     } else if ( er < em && er < el ) {
-      /* er is the minimal of the three, extend to the right */
+      /* er is the least of the three, extend to the right */
       cl = cm;
       cm = cr;
       cr = cr * 2.0;
@@ -495,9 +495,9 @@ static double estbestc_invt(double t, double a0,
         break;
       }
 
-      /* em is the minimal of the three */
+      /* em is the least of the three */
       if ( cm - cl > cr - cm ) {
-        /* refine the left half */
+        /* refine the left side */
         cn = (cl + cm) * 0.5;
         en = esterror_invt(t, cn, a0, n, NULL, lambda, gamma);
         if ( en > em ) {
@@ -512,7 +512,7 @@ static double estbestc_invt(double t, double a0,
           em = en;
         }
       } else {
-        /* refine the right half */
+        /* refine the right side */
         cn = (cm + cr) * 0.5;
         en = esterror_invt(t, cn, a0, n, NULL, lambda, gamma);
         if ( en > em ) {

### 2016-03-30 prog/invtpar.h

@@ -26,11 +26,12 @@ typedef struct {
   double alpha0; /* initial updating magnitude */
   int fixa; /* fix alpha during the entire process */
 
-  int optc; /* use the optimal c */
+  int optc; /* use the optimal proportionality constant for the 1/t schedule */
 
   int opta; /* use the analytically optimal alpha(t) */
   int alpha_nint; /* number of integration points for the exactly optimal alpha(t) */
   char fnalpha[FILENAME_MAX]; /* output file for the exactly optimal alpha */
+  double qprec; /* target precision of the integral of alpha(t) */
 
   int pbc; /* periodic boundary condition */
   int winn; /* width of the updating window function */
@@ -127,8 +128,9 @@ static void invtpar_init(invtpar_t *m)
   m->optc = 0;
 
   m->opta = 0;
-  m->alpha_nint = 1000;
-  m->fnalpha[0] = '\0';;
+  m->alpha_nint = 10000;
+  m->fnalpha[0] = '\0';
+  m->qprec = 1e-5;
 
   m->pbc = 0;
   m->winn = 1; /* single bin update */
@@ -348,10 +350,11 @@ static void invtpar_help(const invtpar_t *m)
   fprintf(stderr, "  --a0=:         set the initial alpha during equilibration, default %g\n", m->alpha0);
   fprintf(stderr, "  --t0=:         set t0 in alpha = c/(t + t0), if unset, t0 = c/a0, default %g\n", m->t0);
   fprintf(stderr, "  --fixa:        fix the alpha during the entire process, default %d\n", m->fixa);
-  fprintf(stderr, "  --optc:        use the optimal c for alpha(t), default %d\n", m->optc);
+  fprintf(stderr, "  --optc:        use the optimal proportionality constant c in the inverse-time schedule for alpha(t), default %d\n", m->optc);
   fprintf(stderr, "  --opta:        use the exact optimal schedule alpha(t), default %d\n", m->opta);
   fprintf(stderr, "  --nint:        set the number of integration points for the exact optimal schedule alpha(t), default %d\n", m->alpha_nint);
   fprintf(stderr, "  --fnalpha=:    set the output file to output the exact optimal schedule, alpha(t), default %s\n", m->fnalpha);
+  fprintf(stderr, "  --qprec:       set the precision of the integral of the alpha(t), default %g\n", m->qprec);
[…]

### 2016-03-30 prog/predict.c

@@ -12,7 +12,7 @@
 static void invt_geterr(invtpar_t *m,
     const double *lambda, const double *gamma)
 {
-  double t, err1, err2;
+  double t, qt = 0, err1, err2;
   intq_t *intq;
 
   t = (double) m->nsteps;
@@ -22,15 +22,15 @@ static void invt_geterr(invtpar_t *m,
       0, &err1, 0);
 
   /* compute the exact minimal error under the same condition */
-  err2 = esterror_opt(t,
-      intq_getqt(t, m->c, m->c / m->alpha0), m->alpha0,
+  err2 = esterror_opt(t, m->alpha0, &qt, m->qprec,
       m->alpha_nint, &intq, m->n, NULL,
-      lambda, gamma);
+      lambda, gamma, m->verbose);
 
   /* save the optimal schedule to file */
   intq_save(intq, m->c, m->c / m->alpha0, m->fnalpha);
 
-  printf("c %g, err %g (invt), %g (exact), %s\n", m->c, err1, err2, m->fnalpha);
+  printf("c %g, err %g, sqr %g (invt), %g, sqr %g (exact), %s\n",
+      m->c, err1, err1 * err1, err2, err2 * err2, m->fnalpha);
 
   intq_close(intq);
 }
@@ -40,7 +40,7 @@ static void invt_geterr(invtpar_t *m,
 static void invt_scanc(invtpar_t *m,
     const double *lambda, const double *gamma)
 {
-  double c, t, t0;
+  double c, qt = 0, t, t0;
   double err, err0, err1, err2;
 
   t = (double) m->nsteps;
@@ -60,10 +60,9 @@ static void invt_scanc(invtpar_t *m,
         lambda, gamma);
 
     /* compute the exact minimal error under the same condition */
-    err2 = esterror_opt(t,
-        intq_getqt(t, m->c, m->t0), m->alpha0,
+    err2 = esterror_opt(t, m->alpha0, &qt, m->qprec,
         m->alpha_nint, NULL, m->n, NULL,
-        lambda, gamma);
+        lambda, gamma, m->verbose);
 
     printf("%g\t%g\t%g\t%g\t%g\n",
         c, err, err0, err1, err2);
@@ -76,7 +75,7 @@ static void invt_scannb(invtpar_t *m,
     const double *gamma)
 {
   double t, t0;
-  double nb, c, err1, err1norm, err2, err2norm;
+  double nb, c, qt = 0, err1, err1norm, err2, err2norm;
[…]

### 2016-04-01 doc/invt.tex
small edits

@@
   In this study,
-  we give a formula of computing the optimal schedule
+  we give a method of computing the optimal schedule
   for a general updating scheme
@@
   %
-  We show that while metadynamics can be advantageous
-  for relatively short simulations,
+  We show that the optimal schedule of metadynamics
+  depends on the simulation length
+  and the width of the Gaussian.
+  %is solved from the equation of motion
+  %of a free particle with a position-dependent mass.
+  %
+  Further,
   the single-bin updating scheme
   used in the WL algorithm
-  belongs to a class of updating schemes
-  optimal for asymptotic convergence.
-  %the optimal schedule
-  %and the effectiveness of metadynamics
-  %are sensitive to the simulation length and other parameters.
-  %
-  %Further, we compared the two algorithms, and found that
-  %although metadynamics produces a smoother PMF profile,
-  %it is less effective in reducing errors with the shorter wavelengths.
-  %%
-  %Thus, the latter relies more heavily on the assumption of
-  %the smoothness of the PMF profile,
-  %especially with a wide updating window.
+  belongs to a class of multiple-bin updating schemes
+  optimal for asymptotic convergence,
+  and the optimal schedules of these schemes
+  are all given by the inverse-time formula.
 \end{abstract}
@@
 used by the WL algorithm
-is best for infinitely long simulations.
+can be generalized to class of multiple-bin updating schemes
+that are partially optimal for infinitely long simulations,
+and the inverse-time schedule is optimal for these schemes.
 %
@@
 during simulation
-Using the histogram as $\pi_i$ in
+Using the histogram $H_i$ as $\pi_i$ in
 Eq. \eqref{eq:pi_p_phi1}
@@
 +
-\log \frac{ \pi_i }
+\log \frac{ H_i }
           { p_i }.
@@
 %
-and $h_i(t) = \delta_{i, i(t)}$
+and
+\begin{equation}
+  h_i(t) = \delta_{i, i(t)}
+  ,
+  \label{eq:h_def}
+\end{equation}
 is the instantaneous histogram,
@@
 =
--\alpha(t) \, \lambda_k \, [y_k(t) - \eta_k(t)],
+-\alpha(t) \, \lambda_k \, [y_k(t) - \theta_k(t)],
 \notag
@@
   \\
-  \eta_k &= \sum_{j=1}^n \phi_{kj} \frac{ \zeta_j}{ p_j}.
+  \theta_k &= \sum_{j=1}^n \phi_{kj} \frac{ \zeta_j}{ p_j}.
   \label{eq:eta_def}
@@
   \left\langle
-    \eta_k(t) \, \eta_k(t')
+    \theta_k(t) \, \theta_k(t')
   \right\rangle.
@@
   \; \dot q^2( t ),
-  \label{eq:error_asym_lagr}
+  \label{eq:error_asym_Lagrangian}
 \end{align}
@@
 
+In practice, Eq. \eqref{eq:varY}
+may fail if some eigenvalue $\lambda_k$ is zero.
+%
+However, the corresponding modes can be ignored
+as they do not contribute to the Lagrangian defined by
+Eq. \eqref{eq:error_asym_Lagrangian}.
+
 \subsection{\label{sec:band-matrix}
@@
 \subsubsection{\label{sec:bandkernel}
-Window kernel and eigenvalues}
+Updating kernel and eigenvalues}
 
@@
 where
-$\mu_{-b}, \dots, \mu_b$ ($b < n/2$)
+$\mu_{-b}, \dots, \mu_b$ ($b \le n/2$)
 are a set of numbers satisfying
@@
 \end{equation}
+%
+If $b = n/2$, then $\mu_{-b}$ is dropped
+from Eq. \eqref{eq:normalization}.
 %
@@
 %
-Thus, in terms of asymptotic convergence,
-the single-bin scheme
-(as adopted in the WL algorithm)
-is most efficient among updating schemes.
+Thus, if no eigenvalue $\lambda_k$ is zero,
+then in terms of asymptotic convergence,
+the single-bin scheme adopted in the WL algorithm
+is most efficient.
 %
@@
 
+\subsubsection{\label{sec:optscheme}
+Optimal schemes}
+
 We can also generalize
@@
 $$
+
+The optimal schedule of these schemes
+remains to be the inverse-time schedule,
+given by Eq. \eqref{eq:alpha_invt1}.
 
@@
 if the simulation uses the single-bin (WL) updating scheme
-with the optimal schedule, Eq. \eqref{eq:alpha_invt}.
+with the optimal schedule, Eq. \eqref{eq:alpha_invt1}.
 %
@@
       \left(
-        \log \frac { \bar h_i }
+        \log \frac { H_i }
                    { p_i }
@@
       \left(
-        \frac { \bar h_i }
+        \frac { H_i }
               { p_i }
@@
       \right)^2
-    \right\rangle.
+    \right\rangle,
 \notag
 \end{align}
+where
+$H_i = \frac{1}{T} \sum_{t = 1}^T h_i(t)$
+with
+$h_i(t)$ defined by Eq. \eqref{eq:h_def}.
 
@@
 \begin{align}
-  \bar \eta_k
+  \Theta_k
   =
@@
   \sum_{ t = 1 } ^ T
-    \eta_k( t )
+    \theta_k( t )
   ,
@@
 %
-with $\eta_k(t_1)$ defined by Eq. \eqref{eq:eta_def}
+with $\theta_k(t_1)$ defined by Eq. \eqref{eq:eta_def}
 and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
 %\begin{align}
-%\eta_k( t_1 )
+%\theta_k( t_1 )
 %=
@@
     \left\langle
-      \bar \eta_k^2
+      \Theta_k^2
     \right\rangle.
@@
 \sum_{ k = 1 }^n
-  \bar \eta_k^2
+  \Theta_k^2
 &=
@@
   \sum_{ k = 1 }^n
-    \eta_k( t ) \, \eta_k( t' )
+    \theta_k( t ) \, \theta_k( t' )
 \\
@@
 $$
-\eta_1(t)
+\theta_1(t)
 =
@@
 \left\langle
-  \bar \eta_k^2
+  \Theta_k^2
 \right\rangle
@@
 \sum_{ t, t' = 1 }^T
-\langle \eta_k(t) \, \eta_k(t') \rangle
+\langle \theta_k(t) \, \theta_k(t') \rangle
 \\
@@
 \sum_{ \tau = -\infty }^{ \infty }
-\langle \eta_k(t) \, \eta_k(t + \tau) \rangle
+\langle \theta_k(t) \, \theta_k(t + \tau) \rangle
 \\
@@
 the optimal $\lambda$ occurred around $1.0$,
-verifying Eq. \eqref{eq:alpha_invt}.
+verifying Eq. \eqref{eq:alpha_invt1}.
 %
@@
     with $\sigma = 10$.
-    %
-    $q(T) = \int_0^t \alpha(t) \, dt$
-    were set to be the same
-    for both schedules.
   }
@@
 The errors in the two cases were
-$4.0 \times 10^{-7}$
+$2.7 \times 10^{-7}$
 and
@@
 or
-$1.67 \times 10^{-4}$
+$9.1 \times 10^{-5}$
 and
-$1.26 \times 10^{-4}$,
+$1.67 \times 10^{-4}$,
 respectively
@@
 the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt},
+Eq. \eqref{eq:alpha_invt1},
 for the single-bin (WL) case.

### 2016-04-01 prog/invt.h

@@ -138,7 +138,8 @@ static double *geteigvals(int n,
 /* modify the window function
  * such that no eigenvalue is negative */
 __inline static double *trimwindow(int n,
-    int *winn, double *win, int pbc, double tol)
+    int *winn, double *win,
+    int pbc, double tol, int verbose)
 {
   const int itmax = 10000;
   int i, j, err, it;
@@ -232,7 +233,7 @@ __inline static double *trimwindow(int n,
         nwin[j] /= n;
       }
 
-      if ( err == 0 || it == itmax - 1 ) {
+      if ( verbose > 0 && (err == 0 || it == itmax - 1) ) {
         fprintf(stderr, "it %d: win %4d: %22.10e -> %22.10e\n",
             it, j, win[j], nwin[j]);
       }

### 2016-04-01 prog/predict.c

@@ -127,7 +127,7 @@ static void invt_scansig(invtpar_t *m,
     m->gaussig = sig;
     invtpar_mkgauswin(m);
     //lambda = geteigvals(m->n, m->winn, m->win, m->pbc, 0, NULL, 1);
-    lambda = trimwindow(m->n, &m->winn, m->win, m->pbc, 0);
+    lambda = trimwindow(m->n, &m->winn, m->win, m->pbc, 0, m->verbose);
 
     /* find the optimal c, according to the inverse time schedule */
     c = estbestc_invt(t, m->alpha0, m->n, lambda, gamma,

### 2016-04-01 doc/invt.tex
invt.tex: update the theory

@@
 %
-Further, by comparing
-different updating schemes,
-we show that
+Further, we show that
 the single-bin updating scheme
@@
 %metadynamics may deliver better performance.
-%
+
 The article is organized as follows.
@@
 $\alpha(t) = \dot q(t)$.
+%
+Since we only optimized the asymptotic error $E_A$,
+the parameter $q(T)$ can be further optimized
+such that total error, $E = E_R + E_A$, is minimal.
 
@@
 
+\subsubsection{\label{sec:program}
+Program}
+
+We can now give a complete program of computing
+the optimal schedule for a given updating scheme.
+
+\begin{enumerate}
+
+\item
+Given an updating matrix, $\mathbf w$,
+compute the eigenvalues, $\lambda_k$'s.
+
+\item
+Run a preliminary adaptive FDS simulation
+using the updating scheme with a constant
+updating magnitude, $\alpha_0$.
+Use Eq. \eqref{eq:varY} to compute $\Gamma_k$'s.
+
+\item
+Select an initial value of $q(T)$,
+e.g., $\ln(1 + \alpha_0 \, T)$.
+
+\item
+Compute the optimal schedule from
+numerically integrating Eq. \eqref{eq:q_opt}.
+
+\item
+Compute the error $E$ from
+Eqs. \eqref{eq:error_split},
+  \eqref{eq:error_asym1} and
+  \eqref{eq:error_res1}.
+
+\item
+Modify the value of $q(T)$ and repeat the above two steps,
+until the minimal error $E$ is found.
+
+\end{enumerate}
+
+The dependence of variables are shown in Fig. \ref{fig:vardep}.
+
+\begin{figure}
+\begin{tikzpicture}
+  \node[draw, text width={width("Updating ")}]
+    (w) at (0, 3) {Updating \\matrix, $w$};
+
+  \node[draw, text width={width("System  11")}]
+    (sys) at (0, 1) {System \& \\sampling \\process};
+
+  \node[draw, text width={width("Eigenvalues, ")}]
+    (lambda) at (2.5, 3) {Eigenvalues, \\$\lambda_2, \dots, \lambda_n$};
+
+  \node[draw, text width={width("Integrals, ")}]
+    (gamma) at (2.5, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
+
+  \node[draw, text width={width("Simulation ")}]
+    (T) at (2.5, -0.5) {Simulation \\time, $T$};
+
+  \node[draw, text width={width("Schedule ")}]
+    (alpha) at (5, 1) {Optimal schedule, \\ $\alpha(t)$};
+
+  \node[draw, text width={width("Error ")}]
+    (err) at (7, 1) {Error, \\ $\Err$};
+
+  \draw[->] (w) edge (lambda)
+            (w) edge (gamma)
+            (sys) edge (gamma)
+            (lambda) edge (alpha)
+            (gamma) edge (alpha)
+            (T) edge (alpha)
+            (alpha) edge (err);
+
+\end{tikzpicture}
+\caption{
+  \label{fig:vardep}
+  Dependence of variables in the method of computing
+  the optimal schedule.
+}
+\end{figure}
+
 \subsection{\label{sec:band-matrix}
@@
 If $b = n/2$, then $\mu_{-b}$ is dropped
-from Eq. \eqref{eq:normalization}.
-%
-These numbers constitute a \emph{updating kernel}.
+from Eq. \eqref{eq:mu_normalization}.
+%
+These numbers constitutes a \emph{updating kernel}.
 
@@
 hence same $\Gamma_k$'s,
-but with different eigenvalues, $\lambda_k$'s.
+but with different positive eigenvalues,
+$\lambda_k$'s.
 %
@@
 %
-$$
+\begin{equation}
   \alpha(t) = \frac{              1             }
                    { \lambda_k \, (t + t_{k0} ) },
-\quad
-\mathrm{with\;\;}
+\label{eq:alpha_invtk}
+\end{equation}
+with
+\begin{equation}
   t_{k0} = \frac{             T            }
                 { e^{ \lambda_k q(T) } - 1 }.
-$$
+\label{eq:tk0}
+\end{equation}
 %
@@
 %
-\note{We add some steps for the above solution.
-  Integrating $\dot u_k = c_k$ yields
-  \begin{equation}
-  u_k(t) = c_k \left(t + t_{k0} \right).
-  \label{eq:ui_solution}
-  \end{equation}
-  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
-  we have
-  $$
-  c_k = \frac{ 1 }{ T + t_{k0} },
-  \quad
-  \mathrm{and\;}
-  \frac{ t_{k0} } { T + t_{k0} }
-  =
-  e^{ -\lambda_k \, q(T) }.
-  $$
-  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
-  $$
-  -\lambda_k \, q(T) + \lambda_k \, q(t)
-  = \log c_k + \log\left( t + t_{k0} \right).
-  $$
-  Differentiating this with respect to $t$ yields
-  $$
-  \lambda_k \, \alpha(t)
-  =
-  \frac {     1      }
-        { t + t_{k0} }.
-  $$
-}
-
-Since $\lambda_1$ corresponds to the mode
-of a uniform shift of all $v_k$,
-we can freely set $\lambda_1$ to $\lambda_2$
-without changing the nature of the updating scheme.
-%
-Now an updating matrix $\mathbf w$
-with identical eigenvalues
-is essentially a multiple of the identity matrix.
-%
-Thus, if no eigenvalue $\lambda_k$ is zero,
-then in terms of asymptotic convergence,
-the single-bin scheme adopted in the WL algorithm
-is most efficient.
-%
-In the single-bin case,
-$c_k = 1/(T + t_0)$,
-and the error is
+In this case,
+all $t_{k0}$'s also share the same value,
+$t_0$ by Eq. \eqref{eq:tk0},
+such that
+$u_k = (t + t_0) / (T + t_0)$,
+and
+$c_k \propto 1/(T + t_0)$
+for $k \ge 2$.
+%
+The asymptotic error is then
 \begin{align}
-  E
+  E_A
   =
@@
 \end{align}
-
-Note that the above result
-holds only under the long time limit,
-and it is not necessarily true
-for finite length simulations.
+%
+If we further assume that $\lambda_2 = 1$,
+the optimal schedule given by Eq. \eqref{eq:alpha_invtk}
+recovers Eq. \eqref{eq:alpha_invt1}.
+%
+\note{We add some steps for the above solution.
+  Integrating $\dot u_k = c_k$ yields
+  \begin{equation}
+  u_k(t) = c_k \left(t + t_{k0} \right).
+  \label{eq:ui_solution}
+  \end{equation}
+  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
+  we have
+  $$
+  c_k = \frac{ 1 }{ T + t_{k0} },
+  \quad
+  \mathrm{and\;}
+  \frac{ t_{k0} } { T + t_{k0} }
+  =
+  e^{ -\lambda_k \, q(T) }.
+  $$
+  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
+  $$
+  -\lambda_k \, q(T) + \lambda_k \, q(t)
+  = \log c_k + \log\left( t + t_{k0} \right).
+  $$
+  Differentiating this with respect to $t$ yields
+  $$
+  \lambda_k \, \alpha(t)
+  =
+  \frac {     1      }
+        { t + t_{k0} }.
+  $$
+}
+
+Since $\lambda_1$ corresponds to the mode
+of a uniform shift of all $v_k$,
+we can freely set $\lambda_1$ to $\lambda_2$
+without changing the nature of the updating scheme.
+%
+Now an updating matrix $\mathbf w$
+with identical eigenvalues
+is essentially a multiple of the identity matrix.
+%
+Thus, if no eigenvalue $\lambda_k$ is zero,
+then in terms of asymptotic convergence,
+the single-bin scheme adopted in the WL algorithm
+is most efficient.
+%
+
+Note that
+since we only optimized the asymptotic error,
+the inverse-time schedule is optimal
+only in the long time limit,
+but not necessarily so for a finite-length simulation.
 %
@@
 Then for shorter times, $T \approx 0$,
-the error, given by Eq. \eqref{eq:error_eql},
+the error, dominated by the residual error
+given by Eq. \eqref{eq:error_res1},
 is clearly smaller if some eigenvalues
@@
 \subsubsection{\label{sec:optscheme}
-Optimal schemes}
+Cardinal updating schemes}
 
@@
 the single-bin scheme to a class of
-optimal updating schemes.
+optimal updating schemes
+that allow zero eigenvalues.
 %
 For example,
-if we know in advance that the intrinsic
-PMF is sufficiently smooth such that
-the error modes can be limited to $k \le K$.
-%
-Then we can manually construct
-an updating scheme with
+if we can assume that the PMF is sufficiently smooth
+such that the error modes can be limited to $k \le K$.
+%
+Then we can construct an updating scheme with
 $$
@@
 $$
-For homogeneous updating schemes,
-the updating kernel can be inversely
-found from Eq. \eqref{eq:mu_from_lambda_pbc}
-or \eqref{eq:mu_from_lambda_refl}.
 %
@@
 $$
-
 The optimal schedule of these schemes
@@
 given by Eq. \eqref{eq:alpha_invt1}.
+%
+Below, we construct these schemes for
+the homogeneous case.
+
+For a periodic variable, if we demand
+%
+$$
+\begin{aligned}
+&
+\lambda_1 = \cdots = \lambda_{K+1}
+= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
+\\
+&
+\lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
+\end{aligned}
+$$
+for $K \ge 1$, then by using
+Eq. \eqref{eq:mu_from_lambda_pbc},
+we get
+\begin{equation}
+  \mu_l
+  =
+  \frac{
+    \sin
+    \dfrac{ (2 K + 1) \, l \, \pi }
+         {              n        }
+  }
+  {
+    n \, \sin \dfrac{ l \, \pi } { n }
+  }
+  .
+  \label{eq:mu_cardinal_pbc}
+\end{equation}
+\note{Derivation.
+$$
+\begin{aligned}
+\mu_l
+&=
+\frac 1 n \sum_{k = 0}^{n-1} \lambda_{k+1} \cos \frac{ k \, l \, 2 \, \pi } { n }
+\\
+&=
+\frac{1}{n}
+\left(
+  1 +
+  \sum_{k=1}^K
+  \cos \frac { k \, l \, 2 \, \pi } { n }
+  +
+  \sum_{k=n-K}^{n-1}
+  \cos \frac { k \, l \, 2 \, \pi } { n }
+\right)
+\\
+&=
+\frac 1 n
+\sum_{k=-K}^K
+\cos \frac { k \, l \, 2 \, \pi } { n }
+=
+  \frac{
+    \sin
+    \dfrac{ (2 K + 1) \, l \, \pi }
+         {              n        }
+  }
+  {
+    n \, \sin \dfrac{ l \, \pi } { n }
+  }
+.
+\end{aligned}
+$$
+}
+
+For a non-periodic variable, we demand
+$$
+\lambda_1 = \cdots = \lambda_{K+1} = 1,
+\qquad
+\lambda_{K+2} = \cdots = \lambda_n = 0.
+$$
+Then Eq. \eqref{eq:lambdan} gives
+$\lambda_n = (-1)^{n+K-1}$,
+and from Eq. \eqref{eq:mu_from_lambda_refl},
+we have
+\begin{equation}
+  \mu_l
+  =
+  \frac{1}{2 \, n}
+  \left[
+    (-1)^{n+K+l-1}
+    +
+    \frac{
+      \sin
+      \dfrac{ (2 K + 1) \, l \, \pi }
+           {         2 \, n        }
+    }
+    {
+      n \, \sin \dfrac{ l \, \pi } { 2 \, n }
+    }
+  \right]
+  .
+\end{equation}
+\note{Derivation.
+$$
+\begin{aligned}
+  \lambda_{n+1}
+  &=
+  (-1)^{n-1}
+  \left[
+    \lambda_1
+    + 2 (-\lambda_2 + \lambda_3 - \cdots + (-1)^K \lambda_{K+1})
+  \right]
+  \\
+  &=
+  \begin{dcases}
+    (-1)^{n-1} & K \mathrm{\; even,} \\
+    (-1)^n     & K \mathrm{\; odd.}
+  \end{dcases}
+\end{aligned}
+$$
+So
+$$
+\begin{aligned}
+  \lambda_l
+  &=
+  \frac{1}{2\,n}
+  \left[
+    1 +
+    2 \sum_{k=2}^{K+1}
+    \cos \frac { (k - 1) \, l \pi } { n }
+    +
+    (-1)^{n+K-1} (-1)^l
+  \right]
+  \\
+  &=
+  \frac{1}{2\,n}
+  \left[
+    (-1)^{n+K+l-1}
+    +
+    \sum_{k=-K}^{K}
+    \cos \frac { k \, l \pi } { n }
+  \right]
+  .
+\end{aligned}
+$$
+
+\hrulefill
+
+Example 1. If $n = 3, K = 1$,
+then $\lambda_1 = \lambda_2 = 1$, $\lambda_3 = 0$,
+and $\lambda_4 = \lambda_1 - 2 \, \lambda_2 + 2 \, \lambda_3 = -1$.
+%
+We can verify that $\mu_0 = 1/3, \mu_1 = 1/2, \mu_2 = -1/6$.
+%
+Although $\mu_1 > \mu_0$, the system will not be trapped
+in the current bin because of the reflective boundary condition.
+%
+Upon a visit to the middle bin, $j = 2$,
+$v_i$ at bin $i = 1$ or $3$ is updated by $1/2 - 1/6 = 1/3$,
+same as the update on bin $2$.
+Upon a visit to bin $i = 1$, the three bins are updated
+by $1/2 + 1/3 = 5/6$, $1/2 - 1/6 = 1/3$, and $-1/6$,
+respectively.
+
+\hrulefill
+
+Example 2. If $n = 3, K = 2$,
+then $\lambda_1 = \lambda_2 = \lambda_3 = 1$, $\lambda_4 = 1$.
+We have $\mu_0 = 1, \mu_1 = \mu_2 = 0$,
+recovering the single-bin scheme.
+
+\hrulefill
+
+Example 3. If $n = 4, K = 1$,
+then $\lambda_1 = \lambda_2 = 1, \lambda_3 = \lambda_4 = 0$,
+$\lambda_5 = (-1)(\lambda_1 - 2 \, \lambda_2 + 2 \, \lambda_3 - 2 \, \lambda_4 = 1$.
+The kernel is given by
+$\mu_0 = 1/2, \mu_1 = \sqrt{2}/8, \mu_2 = 1/4, \mu_3 = -\sqrt{2}/8$.
+}
 
@@
 %
-with $\theta_k(t_1)$ defined by Eq. \eqref{eq:eta_def}
+with $\theta_k(t)$ defined by Eq. \eqref{eq:eta_def}
 and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
 %\begin{align}
-%\theta_k( t_1 )
+%\theta_k( t )
 %=
@@
 %\phi_{k i} \left[
-%             \frac{ h_i(t_1) }
+%             \frac{ h_i(t) }
 %                  { p_i    }
@@
 %
-Adaptive FDS methods differ by the updating schemes,
-which specify how the bias potential is modified
-in each updating step.
-%
-In the single-bin or WL case,
-the update is limited to the current bin;
-while in the metadynamics case,
-the bias potential of a few neighboring bins to the current one
-is updated according to a Gaussian wave packet.
-
-By this method,
-we have verified the optimality of
-the inverse-time schedule,
+Adaptive FDS methods
+effectively sample a flat distribution
+along a quantity of interest, $z$,
+by constructing a bias potential along $z$
+that offsets the PMF.
+%
+The optimal schedule ensures fastest convergence
+of the bias potential to the PMF.
+%They differ by the updating schemes,
+%which specify how the bias potential is modified
+%in each updating step.
+%%
+%In the single-bin or WL case,
+%the update is limited to the current bin;
+%while in the metadynamics case,
+%the bias potential of a neighborhood of the current one
+%is updated with relative weight specified by a Gaussian wave packet.
+
+The optimal schedule computed from our method
+is identical to the inverse-time schedule,
 Eq. \eqref{eq:alpha_invt1},
@@
 %
-However,
-the optimal schedule,
-given by the solution of Eq. \eqref{eq:q_opt},
-is algebraically more involved
-for a general multiple-bin scheme,
-including the Gaussian one used in metadynamics.
+For a general multiple-bin scheme,
+including the Gaussian one used in metadynamics,
+the optimal schedule does not have a closed form,
+but is given by an equation of motion
+of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
 %
 In the latter case,
-both the optimal schedule and the resulting error
+the optimal schedule and error
 are sensitive to the simulation length.
@@
 the single-bin scheme belongs to a class of
-optimal updating schemes.
-%
-That is, the single-bin scheme
-is asymptotically as efficient as
-an equilibrium FDS simulation
-with the exact bias potential.
+asymptotically optimal updating schemes.
+%
+%That is, the single-bin scheme
+%is asymptotically as efficient as
+%an equilibrium FDS simulation
+%with the exact bias potential.
 
@@
 
-We thanks Dr. Y. Mei and Dr. J. Ma for helpful discussions.
+We thanks
+Dr. Y. Mei, Dr. J. Ma,
+J. Drake, O. Nassar, Dr. C. Lai and Dr. S. Ou
+for helpful discussions.
 

### 2016-04-02 prog/invt.h
invt.c: correct mu for pbc

@@ -96,7 +96,7 @@ static double *geteigvals(int n,
     double tol, int *err, int verbose)
 {
   int i, j, nerr = 0;
-  double *lambda, x;
+  double *lambda, x, fac;
 
   if ( tol <= 0 ) { /* use the default value */
     tol = 100 * DBL_EPSILON;
@@ -114,7 +114,9 @@ static double *geteigvals(int n,
         x *= 0.5;
       }
       x = sin(x);
-      lambda[i] -= 4 * win[j] * x * x;
+      fac = 4;
+      if ( pbc && j * 2 == n ) fac = 2;
+      lambda[i] -= fac * win[j] * x * x;
     }
     if ( lambda[i] < -tol ) {
       nerr += 1;
@@ -131,6 +133,27 @@ static double *geteigvals(int n,
   }
 
   return lambda;
+}
+
+
+
+/* save the window function */
+__inline static int savewin(int winn, double *win,
+    const char *fn)
+{
+  FILE *fp;
+  int i;
+
+  if ( (fp = fopen(fn, "w")) == NULL ) {
+    fprintf(stderr, "cannot open %s\n", fn);
+    return -1;
+  }
+  fprintf(fp, "# %d\n", winn);
+  for ( i = 0; i < winn; i++ ) {
+    fprintf(fp, "%d %g\n", i, win[i]);
+  }
+  fclose(fp);
+  return 0;
 }
 
 

### 2016-04-04 doc/invt.tex
invt.c: correct mu for pbc
optimal t_0 = 2/alpha_0

@@
 is most efficient.
-%
-
-Note that
-since we only optimized the asymptotic error,
-the inverse-time schedule is optimal
-only in the long time limit,
-but not necessarily so for a finite-length simulation.
-%
-For example, consider a system that is initially
-prepared under a fixed $\alpha_0$
-at the beginning of the schedule $\alpha(t)$.
-%
-Then for shorter times, $T \approx 0$,
-the error, dominated by the residual error
-given by Eq. \eqref{eq:error_res1},
-is clearly smaller if some eigenvalues
-are less than $1.0$ (the single-bin case).
-%
-%In practice, real simulations
-%may have long achieved the desired precision
-%before reaching the asymptotic regime,
-%especially if the updating matrix
-%has some $\lambda_k$'s close to $0$.
 
 \subsubsection{\label{sec:optscheme}
-Cardinal updating schemes}
+Brick-wall band-pass updating schemes}
 
@@
 the single-bin scheme to a class of
-optimal updating schemes
+\emph{brick-wall band-pass} updating schemes
 that allow zero eigenvalues.
@@
 Then,
-Eq. \eqref{eq:error_singlebin} is modified to
-$$
-  E
+Eq. \eqref{eq:error_singlebin} is changed to
+%
+\begin{equation}
+  E_A
   =
@@
     \Gamma_k.
-$$
+\label{eq:error_asym_sinc}
+\end{equation}
+%
 The optimal schedule of these schemes
@@
   .
-  \label{eq:mu_cardinal_pbc}
+  \label{eq:mu_sinc_pbc}
 \end{equation}
@@
   .
+\label{eq:mu_sinc_refl}
 \end{equation}
@@
 }
+
+\subsubsection{\label{sec:finlencorr}
+Finite-length correction}
+
+In the above, we have been focusing on
+the asymptotic error.
+%
+For a finite-length simulation,
+we should include the residual part,
+Eq. \eqref{eq:error_res}
+and minimize the total error.
+%
+For the brick-wall band-pass updating schemes,
+this allows to determine the optimal value
+of the free parameter, $t_0$.
+
+Assuming the system is initially equilibrated
+under a constant updating magnitude, $\alpha_0$,
+we may use Eq. \eqref{eq:error_res1} with
+$$
+q(T)
+=
+\int \frac{ dt } { t + t_0 }
+=
+\ln \frac{ T + t_0 } { t_0 }.
+$$
+%
+Thus, from Eqs. \eqref{eq:error_split}
+and \eqref{eq:error_asym_sinc}, we have
+%
+\begin{equation}
+  E
+  =
+  E_R + E_A
+  =
+  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
+        {          ( T + t_0 )^2          }
+  \sum_{ k = 2 }^K \Gamma_k
+  .
+\notag
+%\label{eq:error_sinc0}
+\end{equation}
+%
+The optimal error is reached at
+%
+\begin{equation}
+  t_0
+  =
+  \frac{     2    }
+       { \alpha_0 }
+  ,
+\label{eq:t0_sinc}
+\end{equation}
+%
+and at this point
+%
+\begin{equation}
+  E
+  =
+  \frac{        1         }
+       { T + 2 / \alpha_0 }
+  \sum_{ k = 2 }^K
+    \Gamma_k
+  .
+\label{eq:error_sinc}
+\end{equation}
+%
+This means that at the beginning of the schedule,
+the updating magnitude is
+$\alpha(0^+) = \alpha_0 / 2$,
+half of the previous equilibrium value.
+%
+Interestingly, this factor of $1/2$
+happens to coincide with the suggested
+reduction factor of updating magnitude
+in the original WL algorithm\cite{wang2001, wang2001pre}
+during stage transitions.
+
+%
+%For example, consider a system that is initially
+%prepared under a fixed $\alpha_0$
+%at the beginning of the schedule $\alpha(t)$.
+%%
+%Then for shorter times, $T \approx 0$,
+%the error, dominated by the residual error
+%given by Eq. \eqref{eq:error_res1},
+%is clearly smaller if some eigenvalues
+%are less than $1.0$ (the single-bin case).
+%
+%In practice, real simulations
+%may have long achieved the desired precision
+%before reaching the asymptotic regime,
+%especially if the updating matrix
+%has some $\lambda_k$'s close to $0$.
 

### 2016-04-04 doc/invt.tex
add sinc.gp

@@
 in the original WL algorithm\cite{wang2001, wang2001pre}
-during stage transitions.
+at stage transitions.
 
@@
 %
-The total error is given by the sum of
-Eqs. \eqref{eq:error_res_invt} and \eqref{eq:error_asym_invt}.
+The total error is
 %
@@
 \sum_{ k = 2 }^n
-  \Err^f_k \,
-  \left\{
-    1
-    +
-    \frac{ 1 - \left[ t_0 / (T + t_0) \right]^{2 \, \nu_k - 1} }
-         {                 2 \, \nu_k - 1                      }
-%    \frac{       1        }
+  \frac
+  {
+    \Gamma_k \, \nu_k \,
+    \left[
+      \nu_k
+      -
+      (\nu_k - 1)
+      \left(
+        \frac{ t_0 } { T + t_0 }
+      \right)^{ 2 \, \nu_k - 1 }
+    \right]
+  }
+  {
+    (T + t_0) \, (2 \, \nu_k - 1)
+  }
+  .
+\label{eq:error_invt}
+\end{align}
+%
+%where
+%$$
+%\Err^f_k
+%\equiv
+%\left\langle
+%  y_k^2
+%\right\rangle_{ \alpha(t) }
+%=
+%\frac{  \Gamma_k \, \nu_k   }
+%     {    2 \, (T + t_0)    }
+%,
+%$$
+%is the equilibrium, or saturated, error of mode $k$
+%at the terminal updating magnitude
+%of $\alpha(T) = 1/[\lambda (T + t_0)]$.
+
+%\note{Derivation of Eq. \eqref{eq:error_invt}:
+%$$
+%\begin{aligned}
+%\Err_R
+%&=
+%\sum_{ k = 2 }^n
+%  \Err^f_k \left( \frac{   t_0   }
+%                       { T + t_0 }
+%          \right)^{ 2 \nu_k - 1 }
+%\\
+%\Err_A
+%&=
+%\sum_{ k = 2 }^n
+%  \frac{ 2 \, \Err^f_k \, \nu_k }
 %       {    2 \, \nu_k - 1      }
@@
 %  \right]
-  \right\}
-,
-\label{eq:error_invt}
-\end{align}
-%
-where
-$$
-\Err^f_k
-\equiv
-\left\langle
-  y_k^2
-\right\rangle_{ \alpha(t) }
-=
-\frac{  \Gamma_k \, \nu_k   }
-     {    2 \, (T + t_0)    }
-,
-$$
-is the equilibrium, or saturated, error of mode $k$
-at the terminal updating magnitude
-of $\alpha(T) = 1/[\lambda (T + t_0)]$.
-
-\note{Derivation of Eq. \eqref{eq:error_invt}:
-$$
-\begin{aligned}
-\Err_R
-&=
-\sum_{ k = 2 }^n
-  \Err^f_k \left( \frac{   t_0   }
-                       { T + t_0 }
-          \right)^{ 2 \nu_k - 1 }
-\\
-\Err_A
-&=
-\sum_{ k = 2 }^n
-  \frac{ 2 \, \Err^f_k \, \nu_k }
-       {    2 \, \nu_k - 1      }
-  \left[
-    1 -
-    \left(
-      \tfrac{   t_0   }
-            { T + t_0 }
-      \right)^{2 \, \nu_k - 1}
-  \right]
-\\
-&=
-\sum_{ k = 2 }^n
-  \Err^f_k \,
-  \left[
-    1 -
-    \left(
-      \tfrac{   t_0   }
-            { T + t_0 }
-      \right)^{2 \, \nu_k - 1}
-  \right]
-  +
-  \frac{   \Err^f_k     }
-       { 2 \, \nu_k - 1 }
-  \left[
-    1 -
-    \left(
-      \tfrac{   t_0   }
-           { T + t_0 }
-      \right)^{2 \, \nu_k - 1}
-  \right].
-\end{aligned}
-$$
-Adding the two yields Eq. \eqref{eq:error_invt}.
-}
+%\\
+%&=
+%\sum_{ k = 2 }^n
+%  \Err^f_k \,
+%  \left[
+%    1 -
+%    \left(
+%      \tfrac{   t_0   }
+%            { T + t_0 }
+%      \right)^{2 \, \nu_k - 1}
+%  \right]
+%  +
+%  \frac{   \Err^f_k     }
+%       { 2 \, \nu_k - 1 }
+%  \left[
+%    1 -
+%    \left(
+%      \tfrac{   t_0   }
+%           { T + t_0 }
+%      \right)^{2 \, \nu_k - 1}
+%  \right].
+%\end{aligned}
+%$$
+%Adding the two yields Eq. \eqref{eq:error_invt}.
+%}
 
@@
 
-Clearly, the final error is always greater than
-the saturated equilibrium error
-at the terminal $\alpha(T)$:
-$$
-E \ge E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
-$$
-because the second term in the braces
-of Eq. \eqref{eq:error_invt}
-is nonnegative no matter the sign of $2 \, \nu_k - 1$
-(in fact, it is a decreasing function
-that vanishes only at $\nu_k \to +\infty$).
-%
-For a set of nonidentical $\lambda_k$'s,
-the equality is reached only at $T = 0$.
+%Clearly, the final error is always greater than
+%the saturated equilibrium error
+%at the terminal $\alpha(T)$:
+%$$
+%E \ge E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
+%$$
+%because the second term in the braces
+%of Eq. \eqref{eq:error_invt}
+%is nonnegative no matter the sign of $2 \, \nu_k - 1$
+%(in fact, it is a decreasing function
+%that vanishes only at $\nu_k \to +\infty$).
+%%
+%For a set of nonidentical $\lambda_k$'s,
+%the equality is reached only at $T = 0$.
 
@@
   \frac{ \Gamma_k  }
-       { 2 \, (T + t_0) }
+       { (T + t_0) }
   G\left( \nu_k, \frac{ t_0 } { T + t_0} \right),
@@
 =
+\frac { \nu_k }
+      { 2 \, \nu_k - 1 }
+\,
+\left[
   \nu_k
-\,
-\left(
-  1
-  +
-  \frac{ 1 - r^{2 \, \nu_k - 1} }
-       {   2 \, \nu_k - 1       }
-\right).
+  -
+  (\nu_k - 1) \,
+  r^{2 \, \nu_k - 1}
+\right].
 $$

### 2016-04-05 doc/invt.tex
figures: improve sinc.gp

@@
 
-We now adapt the above formulae to a non-periodic variable.
-%
-We shall use the reflective boundary condition\cite{bussi2006},
-and modifies the form of the updating matrix, $\mathbf w$, as
+To adapt the above formulae to a non-periodic variable,
+we use the reflective boundary condition\cite{bussi2006},
+and change the form of the updating matrix to
 %
@@
   \mu_{ i + j - 2 n - 1 },
-  \label{eq:w_band}
+  \label{eq:w_band_refl}
 \end{equation}
@@
 Eqs. \eqref{eq:mu_normalization}
-and
-\eqref{eq:mu_symm}.
+and \eqref{eq:mu_symm}.
 %
 However,
-the radius $b$ only need to be less than $n$
-instead of $n/2$.
+the radius $b$ only need to be less than $n$.
 %
 The last two terms
-on the right-hand side of \eqref{eq:w_band} are added
+on the right-hand side of \eqref{eq:w_band_refl} are added
 to avoid unintended distortion\cite{dickson2011, mcgovern2013}
@@
 
-To compute the eigenvectors,
-we shall replace Eq. \eqref{eq:phi_pbc}
-and redefine the out-of-boundary values
+Similar to Eq. \eqref{eq:phi_pbc},
+we define the out-of-boundary values
 of $\varphi_j$ as
@@
 %
-One can show that Eq. \eqref{eq:wmul_to_convol}
+such that Eq. \eqref{eq:wmul_to_convol}
 still holds.
@@
 %
-Thus, it is also possible to construct
-$n$ orthonormal eigenvectors,
+Thus, we also have $n$ orthonormal eigenvectors,
 $\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
@@
 \subsubsection{\label{sec:optscheme}
-Brick-wall band-pass updating schemes}
+Bandpass updating schemes}
 
@@
 the single-bin scheme to a class of
-\emph{brick-wall band-pass} updating schemes
+perfect (brick-wall) \emph{bandpass} updating schemes
 that allow zero eigenvalues.
@@
 %
-For the brick-wall band-pass updating schemes,
+For the bandpass updating schemes,
 this allows to determine the optimal value
@@
 
-Again, if we are certain that
+Again, the above argument can be extended
+to the bandpass updating schemes.
+%
+If we are certain that
 the intrinsic PMF is sufficiently smooth,
@@
 with a flat target distribution
-$p_i = 1/n$
+$p_i = 1/n$.
+%
+Unless specified otherwise,
+we assume $i$ is a non-periodic variable
 under the reflective boundary condition.
@@
 a constant updating magnitude,
-$\alpha(t) = \alpha_0$.
+$\alpha_0$.
 %
@@
 %
-where $\lambda$ is a parameter to be optimized.
-and $t_0 = 1/(\lambda \, \alpha_0)$
-to smooth the transition from equilibration.
-%
-This schedule is analytically simpler
-than the optimal schedule,
-and it is studied in Appendix \ref{sec:invt_schedule}.
-%
-We wish to see how the final error
-depends on the parameter, $\lambda$,
-and if it matches the predicted value,
-given by Eq. \eqref{eq:error_invt}.
+where $\lambda$ is a parameter to be optimized
+with $t_0 = 2/(\lambda \, \alpha_0)$.
+%
+This simple schedule
+allows us to predict the error analytically
+[cf. Appendix \ref{sec:invt_schedule}],
+and it serves as a model of verifying our theory.
 
@@
 
-We also observe
-that the two error functions
+The two error functions
 differed only by a multiplicative constant,
-again again agreeing with the theory.
+again agreeing with the theory.
 %
@@
 
+While the optimal schedules of the Gaussian updating schemes
+are rather complex,
+those of the bandpass updating schemes
+are always given by the inverse-time formula,
+Eq. \eqref{eq:alpha_invt1},
+with the optimal $t_0$
+given by Eq. \eqref{eq:t0_sinc},
+as shown in Fig. \ref{fig:sinc}(a).
+
+In Fig. \ref{fig:sinc}(b),
+we show that the updating kernels
+of the bandpass updating schemes
+are similar to Gaussian ones
+except the oscillatory tails.
+%
+The equivalent width, $\sigma$, is roughly given
+\begin{equation}
+  \sigma
+  \approx
+  \frac
+  {
+    n
+  }
+  {
+    \sqrt{ 2 \, \pi } \, g \, K
+  }
+  ,
+\label{eq:sigma_equiv}
+\end{equation}
+where $g = 2$ for the periodic case
+or $g = 1$ otherwise.
+%
+The kernel is a zigzag in the non-periodic case
+because of the term $(-1)^{n+K+l-1}$
+in Eq. \eqref{eq:mu_sinc_refl}.
+%
+However, the ruggedness disappears
+in the actual updating matrix, Eq. \eqref{eq:w_band_refl},
+as the above oscillatory term for $l = i - j$
+is canceled by either the term for $l = i + j - 1$
+or that for $l = i + j - 2 \, n - 1$,
+as shown in the inset
+of Fig. \ref{fig:sinc}(b).
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
+  }
+  \caption{
+    \label{fig:sinc}
+    (a) Optimal schedules of the bandpass
+    updating schemes for a periodic variable
+    with $K = 4$,
+    and for a non-periodic variable
+    with $K = 8$.
+    (b) Corresponding updating kernels
+    compared to a Gaussian with a similar width,
+    $\sigma = 5$.
+    The lines are a guide to the eyes.
+    Inset. Updating matrix according to Eq. \eqref{eq:w_band_refl}.
+  }
+\end{center}
+\end{figure}
+
 \subsection{Comparison of updating schemes}
 
-In Fig. \ref{fig:err_sigscan},
-we compared the error from the inverse-time schedule
-with the optimal $\lambda$,
-and that from the asymptotically optimal schedule,
-for the Gaussian updating scheme.
-%
-The errors from the two schedules appeared to be reasonable close,
-unless for large width case.
-%
-This suggests that very long metadynamics simulations
-may benefit from optimal schedule,
-instead of the simple formula, given by
-Eq. \eqref{eq:alpha_invtlambda}.
+For a finite-length simulation,
+the optimal updating scheme
+depends on the simulation length, $T$,
+and other parameters.
+%
+To illustrate this,
+we compare Gaussian updating schemes
+with different width, $\sigma$.
+%
+For the global sampling process,
+we see from Fig. \ref{fig:errsigscan}(a)
+that the lowest error was achieved
+by the widest window for the shorter simulation
+of $T = 10^8$,
+while the minimum moved to $\sigma = 0$
+as the simulation lengthens to $T = 10^{10}$.
+%
+However,
+the shift was hard to detect
+for the local sampling scheme,
+as shown in Fig. \ref{fig:errsigscan}(b).
+%
+In all cases,
+the bandpass scheme
+appeared to be a superior
+alternative to the Gaussian updating scheme.
 
@@
   \caption{
-    \label{fig:err_sigscan}
+    \label{fig:errsigscan}
     Normalized error, $(T + t_0) \, E$,
-    versus the width of the Gaussian updating scheme,
-    $\sigma$,
-    according to the schedule,
-    Eq. \eqref{eq:alpha_invtlambda},
-    with the optimal $\lambda$,
-    and or the asymptotically optimal schedule.
+    versus the width, $\sigma$,
+    for the Gaussian
+    and the brick-wall bandpass updating schemes.
+    %
+    In the latter case,
+    $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
   }
@@
 \end{figure}
-
-%\subsection{Optimality of the single-bin scheme}
-
-The above comparison also shows the asymptotic superiority
-of the single-bin scheme.
-It is clear that for short times,
-the single-bin (WL) may not be optimal,
-but it ultimately wins out as simulation lengthens.
-
-%Randomize the initial error?
-%
-%Error vs. $\mu_1$.
-%
-%Error vs. $\sigma$.
 
@@
 the single-bin scheme belongs to a class of
-asymptotically optimal updating schemes.
+brick-wall bandpass updating schemes that are
+asymptotically optimal.
 %
@@
   \frac{ \Gamma_k \, \nu_k }
-       {        2 \, t_0   }
+       {        t_0   }
   \left(
@@
 %
-assuming that at $t = 0$
-the system has achieved an equilibrium
+assuming that
+the system was initially equilibrated
 at a constant $\alpha_0$,
@@
 $\left\langle y_k^2(0) \right\rangle$
-can be computed from Eq. \eqref{eq:y2_eql}.
+can be computed from Eq. \eqref{eq:y2_eql},
+as well as Eq. \eqref{eq:t0_sinc}.
 %
@@
 is an approximation good only
-for sufficiently short simulations.
-%
-For very longer simulations,
+for sufficiently short simulations,
+since we have taken the limit of $n \to \infty$
+for a fixed $T$.
+%
+For very long simulations,
 the optimal $\lambda$

### 2016-04-05 prog/invt.h

@@ -137,7 +137,7 @@ static double *geteigvals(int n,
 
 
 
-/* save the window function */
+/* save the updating window function */
 __inline static int savewin(int winn, double *win,
     const char *fn)
 {
@@ -151,6 +151,52 @@ __inline static int savewin(int winn, double *win,
   fprintf(fp, "# %d\n", winn);
   for ( i = 0; i < winn; i++ ) {
     fprintf(fp, "%d %g\n", i, win[i]);
+  }
+  fprintf(fp, "%d 0\n", winn);
+  fclose(fp);
+  return 0;
+}
+
+
+/* save the updating window function
+ * in matrix form */
+__inline static int savewinmat(int winn, double *win,
+    int n, int pbc, const char *fn)
+{
+  FILE *fp;
+  int i, j, ks[3], k, l;
+
+  if ( (fp = fopen(fn, "w")) == NULL ) {
+    fprintf(stderr, "cannot open %s\n", fn);
+    return -1;
+  }
+  fprintf(fp, "# %d %d\n", winn, n);
+  for ( i = 1; i <= n; i++ ) {
+    for ( j = 1; j <= n; j++ ) {
+      double y = 0;
+
+      if ( pbc ) {
+        ks[0] = i - j;
+        ks[1] = i - j - n;
+        ks[2] = i - j + n;
+      } else {
+        ks[0] = i - j;
+        ks[1] = i + j - 1;
+        ks[2] = i + j - 2 * n - 1;
+      }
+      for ( l = 0; l < 3; l++ ) {
+        k = ks[l];
+        if ( k < winn && k > -winn ) {
+          if ( k >= 0 ) {
+            y += win[k];
+          } else {
+            y += win[-k];
+          }
+        }
+      }
+      fprintf(fp, "%d %d %g\n", i, j, y);
+    }
+    fprintf(fp, "\n");
   }
   fclose(fp);
   return 0;

### 2016-04-05 doc/invt.tex
code t0 = 2*c/a0, small changes

@@
 The estimated bias potential, $V_i$,
-can often be improved
-by the histogram along $i$ accumulated
-during simulation
-Using the histogram $H_i$ as $\pi_i$ in
+are often improved
+by the histogram, $H_i$, along $i$ accumulated
+during simulation.
+%
+Using $H_i$ as $\pi_i$ in
 Eq. \eqref{eq:pi_p_phi1}
@@
 we are allowed to enter a new stage
-with a reduced $\alpha$\cite{wang2001, wang2001pre}
-(usually half as large as
-the previous value).
+with $\alpha$
+reduced by a factor of one
+half\cite{wang2001, wang2001pre}.
 %
@@
 and Eq. \eqref{eq:alpha_invt}
-is the special case of $t_0 = 0$.
+is the case of $t_0 = 0$.
 
@@
 %
-The instantaneous eigenvalue
-gives the rate of change of
+It gives the rate of change of
 the inverse updating magnitude.
@@
   \label{fig:vardep}
-  Dependence of variables in the method of computing
+  Dependence of variables in computing
   the optimal schedule.
@@
 However,
-the radius $b$ only need to be less than $n$.
+the cutoff $b$ only need to be less than $n$.
 %
@@
 \subsubsection{\label{sec:stabilize_wband}
-Stabilization}
+Inversion and Stabilization}
 
@@
   \right\rangle
-\notag
-\\
-  &\approx
+%\notag
+%\\
+  %&
+  \approx
   \sum_{ i = 1 }^ n
@@
   \langle \theta_k(t) \, \theta_k(t') \rangle
-\\
-&\approx
-\frac{1}{T}
-\sum_{ \tau = -\infty }^{ \infty }
-\langle \theta_k(t) \, \theta_k(t + \tau) \rangle
-\\
-&=
+  %\\
+  %&\approx
+  %\frac{1}{T}
+  %\sum_{ \tau = -\infty }^{ \infty }
+  %\langle \theta_k(t) \, \theta_k(t + \tau) \rangle
+  %\\
+  %&=
+  =
   \frac{1}{T}
@@
   =
-\frac{1}{T}
-\Gamma_k.
+  \frac{ \Gamma_k }{ T }
+.
 \end{align*}
@@
 the intrinsic PMF is sufficiently smooth,
-then we can explicitly filter some error modes,
-say those for $k > K$,
-in the histogram, and the sum in
-Eq. \eqref{eq:error_equil}
-can be truncated at $k = K$.
+then we can explicitly filter some error modes
+in the histogram, and truncate the sum in
+Eq. \eqref{eq:error_equil} at $k = K$.
 
@@
 %
-This simple schedule
-allows us to predict the error analytically
-[cf. Appendix \ref{sec:invt_schedule}],
-and it serves as a model of verifying our theory.
+The error of this schedule
+can be computed analytically
+(cf. Appendix \ref{sec:invt_schedule}),
+and thus be tested against the theory.
 
@@
 The optimal $1/\lambda$ gradually increased
-with simulation length, which is in agreement with
-the discussion in Sec. \ref{sec:optlambda}.
+with simulation length,
+which parallel the gradual shift of
+the instantaneous eigenvalue $\lambda(t)$
+as discussed in Sec. \ref{sec:optschedule}.
+%which is in agreement with
+%the discussion in Sec. \ref{sec:optlambda}.
 %
@@
 
-While the optimal schedules of the Gaussian updating schemes
-are rather complex,
-those of the bandpass updating schemes
+On the other hand,
+the optimal schedules of the bandpass updating schemes
 are always given by the inverse-time formula,
@@
 %
-However, the ruggedness disappears
+However, as shown in the inset of Fig. \ref{fig:sinc}(b),
+the ruggedness disappears
 in the actual updating matrix, Eq. \eqref{eq:w_band_refl},
@@
 or that for $l = i + j - 2 \, n - 1$,
-as shown in the inset
-of Fig. \ref{fig:sinc}(b).
 
@@
     \label{fig:sinc}
-    (a) Optimal schedules of the bandpass
-    updating schemes for a periodic variable
-    with $K = 4$,
-    and for a non-periodic variable
-    with $K = 8$.
-    (b) Corresponding updating kernels
+    Bandpass updating schemes.
+    (a) Optimal schedule for a periodic variable
+    with $K = 3$,
+    and that for a non-periodic variable
+    with $K = 6$.
+    (b) Updating kernels
     compared to a Gaussian with a similar width,
@@
     The lines are a guide to the eyes.
-    Inset. Updating matrix according to Eq. \eqref{eq:w_band_refl}.
+    Inset. Updating matrix in the non-periodic case
+    is free of the zigzaggedness shown in the kernel.
   }
@@
 we see from Fig. \ref{fig:errsigscan}(a)
-that the lowest error was achieved
+that the minimal error was achieved
 by the widest window for the shorter simulation
 of $T = 10^8$,
-while the minimum moved to $\sigma = 0$
-as the simulation lengthens to $T = 10^{10}$.
+while the minimum around $\sigma = 0$
+became more favorable as the simulation lengthens.
 %
@@
 as shown in Fig. \ref{fig:errsigscan}(b).
-%
-In all cases,
-the bandpass scheme
-appeared to be a superior
-alternative to the Gaussian updating scheme.
+
+In contrast,
+the normalized error of the bandpass updating scheme
+was insensitive to the simulation length,
+in agreement with Eq. \eqref{eq:error_sinc}.
+%
+Further, in all cases the bandpass scheme
+appeared to be superior to the Gaussian updating scheme.
 
@@
     for the Gaussian
-    and the brick-wall bandpass updating schemes.
+    and the bandpass updating schemes.
     %
@@
 effectively sample a flat distribution
-along a quantity of interest, $z$,
-by constructing a bias potential along $z$
+by constructing a bias potential
 that offsets the PMF.
@@
 The optimal schedule ensures fastest convergence
-of the bias potential to the PMF.
+of the bias potential.
 %They differ by the updating schemes,
@@
 The optimal schedule computed from our method
-is identical to the inverse-time schedule,
+recovers the known inverse-time formula,
 Eq. \eqref{eq:alpha_invt1},
 for the single-bin (WL) case.
+%
+The origin of time, $t_0$, is better changed
+to the value given by Eq. \eqref{eq:t0_sinc},
+such that the initial updating magnitude
+is half of the previous equilibrium value.
 %
@@
 %
-In the latter case,
+In this case,
 the optimal schedule and error
@@
 
-We have also compared different updating schemes.
-%
-We have shown that in the long time limit
-the single-bin scheme belongs to a class of
-brick-wall bandpass updating schemes that are
+We have also shown that in the long time limit
+the single-bin scheme can be generalized
+to a class of bandpass updating schemes that are
 asymptotically optimal.
+%
+The bandpass updating schemes
+eliminate local noise in the bias potential
+without affecting the global convergence,
+and they appeared to outperform
+the Gaussian updating schemes.
+%
+Thus, the bandpass schemes
+may be useful in real metadynamics simulations.
 %
@@
 
-The optimal schedule predicted
-by the direct solution of Eq. \eqref{eq:q_opt}
-is complicated for a general
-multiple-bin updating scheme.
-%
-Several previous studies\cite{marsili2006, barducci2008, dickson2011}
-have been focused on analytically simpler schedules.
-%
-In a similar spirit,
-we consider here
-the inverse-time formula,
+As an alternative to the
+optimal schedule from Eq. \eqref{eq:q_opt},
+%
+here we consider the inverse-time formula,
 Eq. \eqref{eq:alpha_invtlambda},
-%
-with the proportionality constant, $\lambda$,
-to be optimized.
-%
-We shall compute the error computed from
-this variationally optimized schedule
-and compare it with that from the optimal schedule.
-%
-We shall show that the optimal parameter, $\lambda$,
-is sensitive to the simulation time, $T$,
-which parallels the gradual shift of
-the instantaneous eigenvalue,
-$\lambda(t)$,
-as discussed in Sec. \ref{sec:optschedule}.
+with $\lambda$ optimized variationally.
+%
+This approach allows to obtain
+an explicit formula for the error,
+which can be tested against simulation.
+%
+%Several previous studies\cite{marsili2006, barducci2008, dickson2011}
+%have been focused on analytically simpler schedules.
+%%
+%In a similar spirit,
+%we consider here
+%the inverse-time formula,
+%Eq. \eqref{eq:alpha_invtlambda},
+%%
+%%
+%We shall compute the error computed from
+%this variationally optimized schedule
+%and compare it with that from the optimal schedule.
+%%
+%We shall show that the optimal parameter, $\lambda$,
+%is sensitive to the simulation time, $T$,
+%which parallels the gradual shift of
+%the instantaneous eigenvalue,
+%$\lambda(t)$,
+%as discussed in Sec. \ref{sec:optschedule}.
 
@@
   1 - \left(
-        1 + \frac{ T }{ t_0 }
-      \right)^{1 - 2 \, \nu_k}
+        \frac {     t_0 }
+              { T + t_0 }
+      \right)^{ 2 \, \nu_k - 1 }
 \right],
@@
 %
-At long times, $T \to \infty$, we get
+The error is asymptotically minimal
+only if $2\,\nu_k > 1$ for every $k$.
+\note{At long times, $T \to \infty$, we get
 $$
@@
 as they decay slower than $1/T$.
-%
-Thus, by assuming exclusively the first case
-$\lambda < 2 \lambda_k$,
-the optimal $\lambda$
-satisfies $\lambda < 2 \min \lambda_k$,
-and can be identified as
+}
+%
+In this case,
+we can drop the second term
+in the square brackets,
+and identify the optimal $\lambda$ as
 the smallest real root of
@@
   \sum_{k = 2}^n
-\frac{ \lambda_k - \lambda }
+    \frac { \lambda_k - \lambda }
           { \left(2 - \lambda/ \lambda_k \right)^2 }
   = 0.
-\label{eq:optimal_lambda_approx}
-\end{equation}
-
-Under the approximation, we have
-$$
+\notag
+%\label{eq:optimal_lambda_approx}
+\end{equation}
+%
+Since
+$$
+\frac{ \nu_k^2        }
+     { 2 \, \nu_k - 1 }
+=
 \frac{ \lambda_k^2 }
@@
 $$
-with the equality achieved only at $\lambda = \lambda_k$.
-For the equality to hold for all $k$'s,
-we must have
+with the equality achieved only at $\lambda = \lambda_k$,
+the optimal scheme must satisfy
 $\lambda_2 = \dots = \lambda_n = \lambda$.
 %
-Now a matrix with uniform eigenvalues
-is a multiple of the identity matrix,
-which corresponds to the single-bin updating scheme.
-%
-The above result means that
-the single-bin updating scheme is best
-for the asymptotic convergence.
-%
-This result is discussed in more details in
+%Now a matrix with uniform eigenvalues
+%is a multiple of the identity matrix,
+%which corresponds to the single-bin updating scheme.
+%
+%The above result means that
+%the single-bin updating scheme is best
+%for the asymptotic convergence.
+%
+%This result is discussed in more details in
+This again reflects the asymptotic optimality
+of the single-bin scheme, as discussed in
 Sec. \ref{sec:optWL}.
 
-Real simulations, however,
-can be relatively short
-and need not enter
-the asymptotic regime.
-%
-Then, the residual error, $\Err_R$, is not negligible.
-%
-Note, however,
-since Eq. \eqref{eq:alpha_invtlambda}
-does not fix the endpoint value,
-%
-\begin{equation}
-q(T) = \frac{ 1 } { \lambda }
+For a finite-length simulation,
+we need to include the residual error, $\Err_R$.
+%
+By Eq. \eqref{eq:alpha_invtlambda}, we have
+%
+\begin{equation}
+  q(T)
+  =
+  \frac{ 1 } { \lambda }
   \log\left(
     \frac{ T + t_0 } { t_0 }
-\right),
-\label{eq:qt_invtlambda}
-\end{equation}
-%
-we have to include the residual error
-during the optimization.
-%
-From Eqs. \eqref{eq:error_res1} and \eqref{eq:qt_invtlambda},
-we get
+  \right)
+  ,
+\notag
+%\label{eq:qt_invtlambda}
+\end{equation}
+%
+so that Eq. \eqref{eq:error_res1} becomes
 %
@@
 
+\note{
+
 \subsection{\label{sec:optlambda}
@@
 
-As shown in Sec. \ref{sec:optschedule},
-for the optimal schedule,
-the instantaneous eigenvalue, $\lambda(t)$,
-is on average smaller
-in a longer simulation.
-%
 Intuitively,
@@
 Eq. \eqref{eq:error_invt},
-to be a kind of time average of $\lambda(t)$,
-and thus also to decrease with
+to be a kind of time average of
+the instantaneous eigenvalue $\lambda(t)$
+in Sec. \ref{sec:optschedule}.
+%
+Thus, it should also decrease with
 the simulation length, $T$.
 %
-Below we show this is indeed the case.
+Below we show this is indeed so.
 
@@
 [see Eq. \eqref{eq:lambda_nn_onestep} for an example].
+}
 
@@
 
-For the one-step sampling,
+For the one-step sampling process,
 we find from Eq. \eqref{eq:Gamma_onestep}
@@
 \right|_{ p \approx 0 }
-\notag \\
-&
+%\notag \\
+%&
 \approx
@@
 
-Note that Eq. \eqref{eq:lambda_nn_onestep}
+However, Eq. \eqref{eq:lambda_nn_onestep}
 is an approximation good only
@@
 the optimal $\lambda$
-will ultimately approach
+will ultimately drift to
 the smallest eigenvalue,
-$\min \lambda_k = 1 - 4 \, \mu_1$
-according to the discussion in Sec. \ref{sec:optlambda}.
+$\min \lambda_k = 1 - 4 \, \mu_1$.
+%as discussed in Sec. \ref{sec:optlambda}.
 

### 2016-04-05 prog/invtpar.h

@@ -308,10 +308,8 @@ static void invtpar_compute(invtpar_t *m)
   }
 
   if ( fabs(m->t0) <= 0 ) { /* means t0 == 0 */
-    /* set the default t0 such that
-     * alpha is continuous at the beginning
-     * of the production run */
-    m->t0 = m->c / m->alpha0;
+    /* set the default t0 to 2 * c / alpha0 */
+    m->t0 = 2 * m->c / m->alpha0;
   }
 
   /* turn on correlation function computation
@@ -1081,7 +1079,7 @@ static void invtpar_dump(const invtpar_t *m)
     fprintf(stderr, "sig-scan window: %g:%g:%g\n",
         m->sigmin, m->sigdel, m->sigmax);
   }
-  
+
   if ( m->okscan ) {
     fprintf(stderr, "okmax-scan window: 0:%d:%d\n",
         m->okdel, m->okmax);

### 2016-04-06 doc/invt.tex
predict.c: merging the scan code

@@
 %
-As shown in Fig. \ref{fig:err_singlebin},
+As shown in Fig. \ref{fig:errsbin},
 the numerical results of the terminal error agreed well with
@@
 %%
-%Figure \ref{fig:err_singlebin} also shows
+%Figure \ref{fig:errsbin} also shows
 %an asymmetry of the error curve.
@@
 %when the optimal value is unknown.
-
+%
 The two error functions
@@
   \caption{
-    \label{fig:err_singlebin}
+    \label{fig:errsbin}
     Error, $E$, versus the
@@
     The results have been averaged over $1000$ independent runs.
+    %
+    \note{The figure is produced by \texttt{doc/fig/errsbin.gp}.
+      The data are produced by the script \texttt{invt2run.py},
+      and saved under \texttt{data/singlebin}
+      (see the \texttt{Makefile} there).
+    }
   }

### 2016-04-07 doc/invt.tex
improving figure errscansig.gp

@@
 Assuming the system is initially equilibrated
-under a constant updating magnitude, $\alpha_0$,
+under a constant updating magnitude,
+$\alpha_0$,
 we may use Eq. \eqref{eq:error_res1} with
@@
 a constant updating magnitude,
-$\alpha_0$.
+$\alpha_0 = 10^{-4}$.
 %
@@
     %
-    \note{The figure is produced by \texttt{doc/fig/errsbin.gp}.
-      The data are produced by the script \texttt{invt2run.py},
+    \note{The figure was produced by \texttt{doc/fig/errsbin.gp}.
+      The data were produced by the script \texttt{data/invt2run.py},
       and saved under \texttt{data/singlebin}
@@
 %
-As shown in Fig. \ref{fig:err_nnbr},
+As shown in Fig. \ref{fig:errnnbr},
 the normalized error, $(T + t_0) \, \Err$,
@@
   \caption{
-    \label{fig:err_nnbr}
+    \label{fig:errnnbr}
     Normalized error, $(T + t_0) \, E$,
@@
     The results have been averaged over $1000$ independent runs.
+    \note{This figure was produced by \texttt{doc/fig/errnnbr.gp}.
+      The data were produced by \texttt{data/invt2run.py},
+      saved under \texttt{data/nb0.24}
+      (see the \texttt{Makefile} there).
+    }
   }
@@
     with $\sigma = 10$.
+    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
+      The data were computed by the program \texttt{prog/predict},
+      see particularly the function \texttt{intq\_save()}
+      in \texttt{prog/intq.h}.
+      The data were saved under the directory \texttt{data/opta}.
+      See the \texttt{make} object \texttt{sigsingle}
+      in the \texttt{Makefile} there.
+    }
   }
@@
 As shown in Fig. \ref{fig:optacmp}(a),
-the optimal schedule has a more complex curvature
+the optimal schedule is a more complex curve
 than the optimized inverse-time schedule.
@@
 and
-$5.9 \times 10^{-7}$,
+$6 \times 10^{-7}$,
 respectively,
@@
 and
-$1.67 \times 10^{-4}$,
+$1.7 \times 10^{-4}$,
 respectively
 for the local sampling process,
-suggesting a slight edge of
+showing the superiority of
 the optimal schedule from Eq. \eqref{eq:q_opt}.
@@
     is free of the zigzaggedness shown in the kernel.
+    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
+      The data were computed by the program \texttt{prog/predict},
+      see particularly the function \texttt{intq\_save()}
+      in \texttt{prog/intq.h} for panel (a),
+      and the functions \texttt{savewin()}
+      and \texttt{savewinmat()}
+      in \texttt{prog/invt.h} for panel (b),
+      The data were saved under the directory \texttt{data/sinc}.
+      See the \texttt{make} object \texttt{sincsingle}
+      in the \texttt{Makefile} there.
+    }
   }
@@
 by the widest window for the shorter simulation
-of $T = 10^8$,
-while the minimum around $\sigma = 0$
-became more favorable as the simulation lengthens.
+of $T = 10^8$.
+%
+However, as the simulation lengthens,
+the minimum at $\sigma = 0$ became more favorable.
 %
@@
 %
-Further, in all cases the bandpass scheme
-appeared to be superior to the Gaussian updating scheme.
+Further, in all cases the bandpass updating scheme
+appeared to be superior to the Gaussian one.
 
@@
     $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
+    %
+    Here, we assumed $t_0 \equiv T/(e^{q(T)} - 1)$
+    for the optimal value of $q(T)$.
+    %
+    The points were obtained from averaging results
+    from 1000 independent simulations;
+    the curves are theoretical predictions.
+    %
+    \note{This figure was produced by
+      \texttt{doc/fig/errsigscan.gp}.
+      The data were saved under
+      \texttt{data/scan}.
+    }
   }
@@
       \nu_k
-      -
+      +
       (\nu_k - 1)
@@
 
-\note{
+\note{The following subsection is deleted in the public version.
 
@@
 =
-\frac { \nu_k }
-      { 2 \, \nu_k - 1 }
+\frac { \nu_k
         \,
@@
           \nu_k
-  -
+          +
           (\nu_k - 1) \,
           r^{2 \, \nu_k - 1}
-\right].
+        \right]
+      }
+      { 2 \, \nu_k - 1 }
+.
 $$
@@
     %
+    \note{This figure is produced by
+      \texttt{doc/fig/errcomp.gp}.
+    }
   }

### 2016-04-07 prog/invt.h

@@ -138,7 +138,7 @@ static double *geteigvals(int n,
 
 
 /* save the updating window function */
-__inline static int savewin(int winn, double *win,
+static int savewin(int winn, double *win,
     const char *fn)
 {
   FILE *fp;
@@ -160,7 +160,7 @@ __inline static int savewin(int winn, double *win,
 
 /* save the updating window function
  * in matrix form */
-__inline static int savewinmat(int winn, double *win,
+static int savewinmat(int winn, double *win,
     int n, int pbc, const char *fn)
 {
   FILE *fp;

### 2016-04-07 prog/predict.c

@@ -12,7 +12,7 @@
 static void invt_geterr(invtpar_t *m,
     const double *gamma)
 {
-  double T, qT = 0, err1, err2;
+  double T, t0, qT = 0, err1, err2;
   double *lambda;
   intq_t *intq;
 
@@ -41,10 +41,11 @@ static void invt_geterr(invtpar_t *m,
       lambda, gamma, m->verbose);
 
   /* save the optimal schedule to file */
-  intq_save(intq, m->c, m->fnalpha);
+  t0 = 2 * m->c / m->alpha0;
+  intq_save(intq, m->c, t0, m->fnalpha);
 
-  printf("c %g, err %g, sqr %g (invt), %g, sqr %g (exact), %s\n",
-      m->c, err1, err1 * err1, err2, err2 * err2, m->fnalpha);
+  printf("c %g, t0 %g, err %g, sqr %g (invt), %g, sqr %g (exact), %s\n",
+      m->c, t0, err1, err1 * err1, err2, err2 * err2, m->fnalpha);
 
   intq_close(intq);
   free(lambda);
@@ -184,7 +185,7 @@ static void invt_scan(invtpar_t *m,
         m->alpha_nint, NULL, m->n, NULL,
         lambda, gamma, m->verbose);
 
-    t0 = T / (exp(qT) - 1);
+    t0 = intq_estt0(T, qT);
     err2norm = err2 * sqrt(T + t0);
 
     /* print the scanning variable */

### 2016-04-12 doc/invt.tex
invt.tex: edit abstract, introduction, appendix

@@
   along a quantity of interest, $z$,
-  by constructing a bias potential that offsets
-  the potential of mean force (PMF) on the fly.
-  %
-  The two, however, differ by
+  by adaptively constructing
+  \note{``incrementally building up''?}
+  a bias potential to offset
+  the potential of mean force (PMF).
+  %
+  The two differ by
   the manner of updating the bias potential:
@@
   the WL algorithm updates only
+  the bias potential of
   the bin containing the current $z$,
   while metadynamics updates also
-  several neighboring bins according to
+  those of neighboring bins according to
   a Gaussian wave packet.
   %
-  The error of the resulting PMF
+  The error of the final bias potential,
+  hence that of the PMF,
   is determined by the updating schedule,
@@
   we give a method of computing the optimal schedule
-  for a general updating scheme
-  involving multiple bins.
+  for a general multiple-bin updating scheme.
   %
@@
   used in the WL algorithm
-  belongs to a class of multiple-bin updating schemes
-  optimal for asymptotic convergence,
-  and the optimal schedules of these schemes
-  are all given by the inverse-time formula.
+  belongs to a class of multiple-bin bandpass updating schemes
+  that are optimal for asymptotic convergence.
+  %
+  The optimal schedule of these schemes
+  is given by the inverse-time formula,
+  with a shifted origin of time such that
+  the initial updating magnitude is half of
+  the previous equilibrium value.
 \end{abstract}
@@
 Given a system,
-the problem is to compute,
+the task is to compute,
 via either Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
@@
 %
-This solution can be clumsy in practice
+This solution can be laborious % clumsy in practice
 and it often fails to converge quickly
@@
 because the system can be trapped
-in a local free energy minimum
-for a long time.
+in a local free energy minimum.
+%for a long time.
 
@@
 the method must be able to construct a
-bias potential that exactly cancels the PMF.
+bias potential that cancels the PMF.
 
@@
 techniques for this purpose,
-in which a bias potential is actively built up on-the-fly
+in which a bias potential is gradually built up
 during simulation.
@@
 These techniques employ updating loops
-that discourage future visits to configurations
+to discourage future visits to configurations
 with the current $z$ value
@@
 is updated frequently (usually every few MC or MD steps),
-which brings about a side effect
-of disrupting the underlying
+which % brings about a side effect of disrupting
+disrupts \note{``disturbs''?} the underlying
 equilibrium sampling dynamics\cite{laio2002}.
@@
 or the manner of reducing
-the updating magnitude\cite{liang2007,
+the updating magnitude,
+therefore,
+determines the precision of the final bias potential,
+hence that of the PMF\cite{liang2007,
 belardinelli2007, belardinelli2007jcp, belardinelli2008,
@@
 komura2012, caparica2012, caparica2014,
-barducci2008, dickson2011, dama2014},
-therefore,
-determines the precision of the final bias potential,
-hence that of the PMF.
+barducci2008, dickson2011, dama2014}.
 %
@@
 The method reproduces the inverse-time formula
-for the WL case.
+for the WL case,
+and gives the optimal value of the origin of time.
 %
 It also predicts, for metadynamics,
-a more algebraically complex schedule,
+a more complex schedule,
 which is sensitive to the simulation length
@@
 %
-Further, we show that
+Further, we generalize
 the single-bin updating scheme
 used by the WL algorithm
-can be generalized to class of multiple-bin updating schemes
-that are partially optimal for infinitely long simulations,
+to a class of multiple-bin updating schemes
+that are optimal for infinitely long simulations,
 and the inverse-time schedule is optimal for these schemes.
@@
 We present the analytical results in Sec. \ref{sec:theory},
-numerically verify some key aspects
+numerically verify some aspects
 in Sec. \ref{sec:results},
@@
       (see the \texttt{Makefile} there).
-    }
+    }%
   }
@@
     The results have been averaged over $1000$ independent runs.
+    %
     \note{This figure was produced by \texttt{doc/fig/errnnbr.gp}.
@@
     with $\sigma = 10$.
+    %
     \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
@@
       in the \texttt{Makefile} there.
-    }
+    }%
   }
@@
     is free of the zigzaggedness shown in the kernel.
+    %
     \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
@@
       \texttt{data/scan}.
-    }
+    }%
   }
@@
 %
-here we consider the inverse-time formula,
-Eq. \eqref{eq:alpha_invtlambda},
+here we consider the analytically simpler
+inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
 with $\lambda$ optimized variationally.
 %
-This approach allows to obtain
-an explicit formula for the error,
-which can be tested against simulation.
-%
-%Several previous studies\cite{marsili2006, barducci2008, dickson2011}
-%have been focused on analytically simpler schedules.
-%%
-%In a similar spirit,
-%we consider here
-%the inverse-time formula,
-%Eq. \eqref{eq:alpha_invtlambda},
-%%
-%%
-%We shall compute the error computed from
-%this variationally optimized schedule
-%and compare it with that from the optimal schedule.
-%%
-%We shall show that the optimal parameter, $\lambda$,
-%is sensitive to the simulation time, $T$,
-%which parallels the gradual shift of
-%the instantaneous eigenvalue,
-%$\lambda(t)$,
-%as discussed in Sec. \ref{sec:optschedule}.
+This approach allows us to compute the error explicitly.
 
@@
 only if $2\,\nu_k > 1$ for every $k$.
+%
 \note{At long times, $T \to \infty$, we get
@@
 as they decay slower than $1/T$.
-}
+}%
 %
@@
      { \lambda \, (2 \, \lambda_k - \lambda) }
-\ge 1,
+\ge 1
+,
 $$
@@
 the optimal scheme must satisfy
-$\lambda_2 = \dots = \lambda_n = \lambda$.
-%
-%Now a matrix with uniform eigenvalues
-%is a multiple of the identity matrix,
-%which corresponds to the single-bin updating scheme.
-%
-%The above result means that
-%the single-bin updating scheme is best
-%for the asymptotic convergence.
-%
-%This result is discussed in more details in
-This again reflects the asymptotic optimality
-of the single-bin scheme, as discussed in
-Sec. \ref{sec:optWL}.
+$\lambda_2 = \dots = \lambda_n$,
+corresponding to the single-bin scheme
+(cf. Sec. \ref{sec:optWL}).
 
@@
 %
-By Eq. \eqref{eq:alpha_invtlambda}, we have
+Assuming Eqs. \eqref{eq:error_res1} and \eqref{eq:t0_sinc},
+we get
+%
+\begin{equation}
+\Err_R
+=
+\sum_{k = 2}^n
+  \frac{ \Gamma_k \, \nu_k }
+       {        t_0   }
+  \left(
+      \frac{   t_0   }
+           { T + t_0 }
+   \right)^{ 2 \, \nu_k },
+\notag
+%\label{eq:error_res_invt}
+\end{equation}
+%
+with
 %
@@
 %
-so that Eq. \eqref{eq:error_res1} becomes
-%
-\begin{equation}
-\Err_R
-=
-\sum_{k = 2}^n
-  \frac{ \Gamma_k \, \nu_k }
-       {        t_0   }
-  \left(
-      \frac{   t_0   }
-           { T + t_0 }
-   \right)^{ 2 \, \nu_k },
-\label{eq:error_res_invt}
-\end{equation}
-%
-assuming that
-the system was initially equilibrated
-at a constant $\alpha_0$,
-such that the values of
-$\left\langle y_k^2(0) \right\rangle$
-can be computed from Eq. \eqref{eq:y2_eql},
-as well as Eq. \eqref{eq:t0_sinc}.
+from integrating Eq. \eqref{eq:alpha_invtlambda}.
+%
+%We assume that
+%the system was initially equilibrated
+%at a constant $\alpha_0$,
+%such that the values of
+%$\left\langle y_k^2(0) \right\rangle$
+%can be computed from Eq. \eqref{eq:y2_eql},
+%as well as Eq. \eqref{eq:t0_sinc}.
 %
@@
 \begin{equation}
-\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
+  \lambda
+  =
+  \frac{ 1 - 4 \, \mu_1 }
+       { 1 - 2 \, \mu_1 }
+  ,
 \label{eq:lambda_nn_perfect}
@@
     1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
-\right).
-\label{eq:error_nn_prefect}
+  \right)
+  .
+\notag
+%\label{eq:error_nn_perfect}
 \end{equation}
@@
   \sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } - 1 }.
-\label{eq:nnint_local_zeros_part2}
-\end{equation}
-We share distinguish two cases.
+               { 8 \, \mu_1  } - 1 }
+  .
+\label{neq:nnint_local_zeros_part2}
+\end{equation}
+%
+We shall distinguish two cases.
 \textbf{Case A.}
@@
 If $2 - 8 \, \mu_1 \le \lambda < 2$,
-the four zeros in Eq. \eqref{eq:nnint_local_zeros_part2}:
+the four zeros in Eq. \eqref{neq:nnint_local_zeros_part2}:
 %
@@
   \sqrt { 1 - \frac{ 2 - \lambda }
-                 { 8 \, \mu_1  } },
-\label{eq:nnint_local_zeros_part2b}
+                   { 8 \, \mu_1  } }
+  ,
+\notag
+%\label{eq:nnint_local_zeros_part2b}
 \end{equation}

### 2016-04-12 doc/invt.tex
invt.tex: editing conclusion

@@
 is the case of $t_0 = 0$.
+%
+In Sec. \ref{sec:finlencorr},
+we show that $t_0$ can be adjusted
+to take into account for the initial error.
 
@@
 %
-The update magnitudes are given by a matrix
-$\mathbf w$
-with elements $w_{ij}$.
+The relative update magnitudes, $w_{ij}$,
+constitute a matrix $\mathbf w$.
 %
@@
 resembles a transition matrix,
-although the matrix elements can be negative.
+although the elements can be negative.
 
@@
 Further, by Eq. \eqref{eq:w_detailedbalance},
-we can define a symmetric matrix $\hat{\mathbf w}$
-as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$
-that can be diagonalized\cite{vankampen}
-with a set of orthonormal eigenvectors:
-%
-$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
-%
-Thus,
-in terms of diagonalizing $\mathbf w$, we have
+we can diagonalize $\mathbf w$ with a set of
+eigenvectors
+%
 \begin{equation}
@@
 \end{equation}
-where the eigenvectors
-$$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$$
-satisfy the orthonormal conditions\cite{vankampen}:
+%
+satisfying the orthonormal conditions\cite{vankampen}:
 %
@@
 \end{align}
+%
+\note{We can define a symmetric matrix $\hat{\mathbf w}$
+as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$,
+which can be diagonalized\cite{vankampen}
+with a set of orthonormal eigenvectors:
+%
+$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
+%
+Then,
+in terms of diagonalizing $\mathbf w$, we have
+$$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$$
+satisfying Eqs. \eqref{eq:eig_orthonormal_cols}
+and \eqref{eq:eig_orthonormal_rows}.
+%
+}%
 
@@
     \kappa_k(t' - t) \, dt' = 0.
+  \tag{N1}
 \label{eq:optimal_mbin}
@@
 If all $\lambda_k$, hence all $u_k$, are the same,
-the above condition can be satisfied
-by Eq. \eqref{eq:ddu_eq_0},
-which is the single-bin case.
+the above condition can be satisfied with $\ddot u_k = 0$,
+which recovers the single-bin case.
 %
@@
       u_k\bigl( q(t) \bigr) \,
-      \ddot u_k\bigl( q(t) \bigr) = 0.
+      \ddot u_k\bigl( q(t) \bigr) = 0
+    .
+    \tag{N2}
     \label{eq:optimal_mbin1}
@@
 To find the eigenvectors,
-we note that for a periodic variable $\varphi_i$,
+we note that for a periodic variable, $\phi_i$,
 the out-of-boundary values are defined by
@@
 \begin{equation}
-  \varphi_i = \varphi_{i \pm n},
+  \phi_i = \phi_{i \pm n},
 \label{eq:phi_pbc}
@@
   \sum_{ j = 1 }^n
-    w_{ij} \, \varphi_j
+    w_{ij} \, \phi_j
   =
   \sum_{ j = 1 - n }^{ 2 \, n }
-    \mu_{i - j} \, \varphi_j
+    \mu_{i - j} \, \phi_j
   =
   \sum_{ l = -b }^{ b }
-    \mu_l \, \varphi_{ i - l}
+    \mu_l \, \phi_{ i - l}
   .
@@
   \sum_{j = 1}^n
-    w_{ij} \, \varphi_j
+    w_{ij} \, \phi_j
   &=
   \sum_{j = 1}^n
-    \mu_{i - j} \, \varphi_j
+    \mu_{i - j} \, \phi_j
   +
   \sum_{j = 1}^n
-    \mu_{i - j - n} \, \varphi_j
+    \mu_{i - j - n} \, \phi_j
   +
   \sum_{j = 1}^n
-    \mu_{i - j + n} \, \varphi_j
+    \mu_{i - j + n} \, \phi_j
   \\
@@
   \sum_{j = 1}^n
-    \mu_{i - j} \, \varphi_j
+    \mu_{i - j} \, \phi_j
   +
   \sum_{j = 1+n}^{2 \, n}
-    \mu_{i - l} \, \varphi_l
+    \mu_{i - l} \, \phi_l
   +
   \sum_{j = 1-n}^0
-    \mu_{i - j + n} \, \varphi_j
+    \mu_{i - j + n} \, \phi_j
   \\
@@
   \sum_{j = 1-n}^{2 \, n}
-    \mu_{i - j} \, \varphi_j
+    \mu_{i - j} \, \phi_j
   .
@@
   \sum_{ i = 1 }^n
-    \varphi_i \, w_{ij}
+    \phi_i \, w_{ij}
   =
   \sum_{ l = -b }^b
-    \varphi_{j - l} \, \mu_{-l}
+    \phi_{j - l} \, \mu_{-l}
   .
@@
 we may construct a set of orthonormal eigenvectors,
-$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
+$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
 as
@@
   =
-  \frac{1}{\sqrt n}
+  \frac 1 n
   \exp
@@
 we define the out-of-boundary values
-of $\varphi_j$ as
-%
-\begin{equation}
-  \varphi_j
+of $\phi_j$ as
+%
+\begin{equation}
+  \phi_j
   =
   \begin{dcases}
-    \varphi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
-    \varphi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
+    \phi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
+    \phi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
   \end{dcases}
@@
   \begin{aligned}
-    \sum_{j = 1}^n w_{ij} \, \varphi_j
+    \sum_{j = 1}^n w_{ij} \, \phi_j
     &=
     \sum_{j = 1}^n
-      \mu_{i - j} \, \varphi_j
+      \mu_{i - j} \, \phi_j
     +
     \sum_{j = 1}^n
-      \mu_{i + j - 1} \, \varphi_j
+      \mu_{i + j - 1} \, \phi_j
     +
     \sum_{j = 1}^n
-      \mu_{i + j - 2 \, n - 1} \, \varphi_j
+      \mu_{i + j - 2 \, n - 1} \, \phi_j
     \\
@@
     \sum_{j = 1}^n
-      \mu_{i - j} \, \varphi_j
+      \mu_{i - j} \, \phi_j
     +
     \sum_{l = 1 - n}^0
-      \mu_{i - l} \, \varphi_l
+      \mu_{i - l} \, \phi_l
     +
     \sum_{l = n + 1}^{ 2 \, n }
-      \mu_{i - l} \, \varphi_l
+      \mu_{i - l} \, \phi_l
     \\
@@
     \sum_{j = 1 - n}^{ 2 \, n}
-      \mu_{i - j} \, \varphi_j.
+      \mu_{i - j} \, \phi_j.
   \end{aligned}
@@
 Thus, we also have $n$ orthonormal eigenvectors,
-$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
+$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
 as
@@
 \begin{equation}
-\varphi^{(k)}_i
-=
-\varphi_{k i}
-=
-\sqrt{
-    \frac{ 2 - \delta_{k, 1} }
+  \phi^{(k)}_i
+  =
+  \phi_{k i}
+  =
+  \frac{ \sqrt{ 2 - \delta_{k, 1} } }
        {             n              }
-     }
   \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, \pi}
-            {                  n                         }.
+            {                    n                           }
+  ,
 \notag
@@
     \cos \frac{(k - 1)  \, l \pi}{n}
+  .
   \label{eq:wband_eigenvalue_refl}
@@
     \sum_{k = 1}^n
-    \varphi^{(k)}_i \, \varphi^{(k)}_j
-    =
-    \frac 2 n
+    \phi^{(k)}_i \, \phi^{(k)}_j
+    =
+    \frac 2 { n^2 }
     \sum_{k = 1}^n
@@
     -
-    \frac 1 n
-    =
+    \frac 1 { n^2 }
+    =
+    \frac 1 n \,
     \delta_{i, j}.
@@
     \lambda_{ 2 \, n + 1 - k } = \lambda_k.
-    \label{eq:lambda_reflection}
+    \tag{N3}
+    \label{neq:lambda_reflection}
   \end{align}
@@
   $$
-  where we have used Eq. \eqref{eq:lambda_reflection}
+  where we have used Eq. \eqref{neq:lambda_reflection}
   in the last step.
@@
     u_k(t) = c_k \left(t + t_{k0} \right).
-  \label{eq:ui_solution}
+  \tag{N4}
+  \label{neq:ui_solution}
   \end{equation}
@@
   $$
-  Taking the logarithm of Eq. \eqref{eq:ui_solution} yields
+  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
   $$
@@
 of the free parameter, $t_0$.
-
-Assuming the system is initially equilibrated
-under a constant updating magnitude,
-$\alpha_0$,
-we may use Eq. \eqref{eq:error_res1} with
+%
+%Assuming the system is initially equilibrated
+%under a constant updating magnitude,
+%$\alpha_0$,
+Assuming Eq. \eqref{eq:error_res1} with
 $$
@@
 =
-\int \frac{ dt } { t + t_0 }
+\int_0^T \frac{ dt } { t + t_0 }
 =
@@
 %
-Thus, from Eqs. \eqref{eq:error_split}
-and \eqref{eq:error_asym_sinc}, we have
+we have, from Eqs. \eqref{eq:error_split}
+and \eqref{eq:error_asym_sinc},
 %
@@
 %
-Adaptive FDS methods
+Adaptive FDS methods,
+including the WL algorithm and metadynamics,
 effectively sample a flat distribution
 by constructing a bias potential
-that offsets the PMF.
+to offset the PMF.
 %
@@
 
-The optimal schedule computed from our method
-recovers the known inverse-time formula,
-Eq. \eqref{eq:alpha_invt1},
-for the single-bin (WL) case.
-%
-The origin of time, $t_0$, is better changed
-to the value given by Eq. \eqref{eq:t0_sinc},
-such that the initial updating magnitude
-is half of the previous equilibrium value.
+Each adaptive FDS method defines an updating scheme,
+or mathematically an updating matrix,
+whose elements $w_{ij}$ specifies the
+relative magnitude of updating the bias potential
+at bin $i$ upon a visit to bin $j$.
+%
+Our method computes the optimal schedule from
+the eigenvalues of the updating matrix,
+and the integrals of the correlation functions
+the corresponding eigenmodes.
+%
+For the single-bin (WL) scheme,
+the optimal schedule from our method
+is given by inverse-time formula,
+Eq. \eqref{eq:alpha_invt1}.
+%
+The parameter, $t_0$, can be adjusted
+according to Eq. \eqref{eq:t0_sinc},
+to take into account the initial error.
+%such that the initial updating magnitude
+%is half of the previous equilibrium value.
 %
@@
 including the Gaussian one used in metadynamics,
-the optimal schedule does not have a closed form,
+the optimal schedule has no closed form,
 but is given by an equation of motion
@@
 the optimal schedule and error
-are sensitive to the simulation length.
-
-We have also shown that in the long time limit
-the single-bin scheme can be generalized
-to a class of bandpass updating schemes that are
-asymptotically optimal.
-%
-The bandpass updating schemes
-eliminate local noise in the bias potential
-without affecting the global convergence,
-and they appeared to outperform
-the Gaussian updating schemes.
-%
-Thus, the bandpass schemes
-may be useful in real metadynamics simulations.
-%
-%That is, the single-bin scheme
-%is asymptotically as efficient as
-%an equilibrium FDS simulation
-%with the exact bias potential.
+can be sensitive to the simulation length.
+
+We have also shown that
+the single-bin scheme is optimal
+in the long time limit
+without a priori assumption
+of the smoothness of the PMF.
+%
+More generally, with perfect filtering
+of certain eigenmodes, we can construct a
+class of multiple-bin bandpass updating schemes
+that are also optimal for asymptotic convergence.
+%
+The optimal schedule of these bandpass schemes
+is always given by the inverse-time formula.
+%
+They appeared to outperform similar Gaussian schemes
+in our model study,
+suggesting that they can be useful
+in real metadynamics simulations.
 
@@
   .
+\tag{N5}
 \label{neq:nnint_local_zeros_part2}

### 2016-04-13 doc/invt.tex
add checkeq.py

@@
   by adaptively constructing
-  \note{``incrementally building up''?}
+  \note{``incrementally building up''?}%
   a bias potential to offset
@@
   \dot u\bigl( q(t') \bigr) = c,
-\label{eq:ddu_eq_0}
+\notag
+%\label{eq:ddu_eq_0}
 \end{equation}
@@
     \tag{N2}
-    \label{eq:optimal_mbin1}
+    \label{neq:optimal_mbin1}
   \end{align}
@@
   = \delta(t - t'),
-  \label{eq:kappa_perfect}
+\notag
+%\label{eq:kappa_perfect}
 \end{equation}
@@
 \eqref{eq:eta_def},
+\eqref{eq:zz_perfect},
 and
@@
 %
-\note{To show Eq. \eqref{eq:kappa_perfect},
+\note{Proof.%Eq. \eqref{eq:kappa_perfect},
   $$
@@
   \right),
-\label{eq:T_nn}
+\notag
+%\label{eq:T_nn}
 \end{equation}
@@
 
-\subsubsection{\label{eq:eqlerr}
+\subsubsection{\label{sec:eqlerr}
 Equilibrium error under a constant updating magnitude
@@
   \sum_{ k = 2 }^n
-    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0.
-  \label{eq:error_eql}
+    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0
+  .
+\notag
+%\label{eq:error_eql}
 \end{align}
@@
   ,
-  \label{eq:alpha_approxinvt}
+\notag
+%\label{eq:alpha_approxinvt}
 \end{equation}
@@
   ,
-  \label{eq:dinvadt_limit2}
+\notag
+%\label{eq:dinvadt_limit2}
 \end{equation}
@@
   ,
-\label{eq:w_band_pbc}
+\notag
+%\label{eq:w_band_pbc}
 \end{equation}
@@
   \end{dcases}
-\label{eq:phi_refl}
+\notag
+%\label{eq:phi_refl}
 \end{equation}
@@
   which satisfies the reflective boundary condition,
-  Eq. \eqref{eq:phi_refl}.
+  %Eq. \eqref{eq:phi_refl}.
   %
@@
   .
-\label{eq:wnn_eigenvalue}
+\notag
+%\label{eq:wnn_eigenvalue}
 \end{equation}
@@
   .
-  \label{eq:mu_sinc_pbc}
+\notag
+%\label{eq:mu_sinc_pbc}
 \end{equation}

### 2016-04-14 doc/invt.tex
invt.tex: edit to Sec. II.C.3

@@
 The negative logarithm,
-$-\log p^*(z)$,
+$-\ln p^*(z)$,
 defines a free energy,
@@
 of multiple-bin updating schemes
+and propose a method of computing
+the optimal schedule
 in Secs. \ref{sec:multiple-bin}
@@
 Finally, we compare different updating schemes
+and show that the single-bin scheme
+and a class of generalized bandpass schemes
+are optimal for convergence
 in the long time limit
@@
 
-Given a system,
-consider the problem of computing
+Consider the problem of computing
 the distribution, $p^*_i$,
-along a discrete quantity $i$.
+along a discrete quantity, $i$,
+for a given system.
 %
@@
 %
-For example, $i$ can be the energy $E$
+For example, $i$ can be the energy, $E$,
 in a lattice spin model; or the temperature index
-in a simulated tempering simulation.
+in a simulated tempering simulation\cite{
+marinari1992, lyubartsev1992}.
 %
 For a continuous quantity, $z$,
-which is typical in molecular systems,
+which is typical in a molecular system,
 we can discretize $z$
@@
 %
-and to find out the global shape,
+and to find out the global shape
+of the PMF, $-\ln p^*_i$,
 it is often advantageous to carry out
 a biased sampling that targets
-a wider distribution $p_i$.
+a wider distribution, $p_i$.
 %
@@
 
-To do so, we need to introduce a bias potential $V_i$
-to modify the target distribution to
+To do so, we introduce a bias potential, $V_i$,
+which changes the equilibrium distribution to
 %
@@
 %
-Upon normalization, $\sum_{i = 1}^n \pi_i = 1$,
+Upon the normalization,
+$\sum_{i = 1}^n \pi_i = 1$,
 we get
@@
 Particularly,
-to achieve a flat distribution $\pi_i$,
+to make $\pi_i$ flat,
 the bias potential $V_i$
-must coincide with $\log p^*_i$,
+must coincide with $\ln p^*_i$,
 up to an additive constant.
+%
+In practice, $V_i$ may contain error,
+and thus $\pi_i$ need not
+coincide with the intended distribution $p_i$.
 
 Here we shall call a simulation that targets
-a nearly flat distribution\cite{dayal2004, trebst2004, singh2011},
+a nearly flat distribution\cite{
+dayal2004, trebst2004, singh2011},
 $p_i$,
@@
 referred to as the equilibrium FDS, % (E-FDS) methods hereinafter,
-estimates the bias potential, $V_i$,
-before simulation,
-and fixes it during simulation.
+uses a fixed bias potential, $V_i$,
+(which is estimated beforehand)
+during simulation.
 %
 Examples include the original
-entropic or multicanonical sampling\cite{berg1992, lee1993}
-and simulated tempering\cite{marinari1992, lyubartsev1992}
+entropic or multicanonical sampling\cite{
+berg1992, lee1993}
+and simulated tempering\cite{
+marinari1992, lyubartsev1992}
 with $i$ being the index of energy
@@
 %
-Using $H_i$ as $\pi_i$ in
+Using $H_i$ as an estimate of $\pi_i$ in
 Eq. \eqref{eq:pi_p_phi1}
@@
 +
-\log \frac{ H_i }
+\ln \frac{ H_i }
          { p_i }.
@@
   \begin{aligned}
-    \pi_i &\propto p^*_i \, e^{-V_i}, \\
-    p_i &\propto p^*_i \, e^{-V^\mathrm{corr}_i}.
+    H_i \approx \pi_i
+    &\propto p^*_i \, e^{-V_i},
+    \\
+    p_i
+    &\propto p^*_i \, e^{-V^\mathrm{corr}_i}.
   \end{aligned}
@@
 For an unknown system,
-the main difficulty of the process
-is that it is hard to guess a sufficiently accurate
+it is often difficult
+to guess a sufficiently accurate
 $V_i$ to start with.
@@
 %
-Further, a reasonably precise histogram
+\note{Further, a reasonably precise histogram
 requires an accumulation period
 longer than the autocorrelation time,
-which can also be hard to estimate beforehand.
+which can also be hard to estimate beforehand.}
 
@@
 Examples include
-the WL and metadynamics methods.
-%
-These methods update the bias potential more frequently
-and can easily achieve a flat distribution.
+the WL algorithm and metadynamics.
+%
+These methods update the bias potential
+incrementally in regular intervals
+and can easily achieve the target distribution.
 %
 However, they are less well understood theoretically,
-as the updating loops used there
+as the updating loops
 break the microscopic reversibility
@@
 reduce the updating magnitude
-over the course of simulation\cite{marsili2006,
+over the course of simulation\cite{
+marsili2006,
 liang2007,
-belardinelli2007, belardinelli2007jcp, belardinelli2008}.
+belardinelli2007, belardinelli2007jcp, belardinelli2008,
+barducci2008}.
 %
@@
 the updating is negligible and
-such an adaptive simulation is rather similar
-to the corresponding equilibrium simulation.
+the adaptive simulation tends to
+the corresponding equilibrium one.
 %
@@
 
-For later convenience, we introduce a shifted bias potential
+For later convenience,
+we introduce a shifted bias potential
 %
@@
 remain constant during the course of simulation,
-updates to $V_i$'s are equivalent to those to $v_i$'s.
+the updates to $v_i$'s are
+the same as the updates to $V_i$'s.
 %
@@
 According to Eq. \eqref{eq:pi_p_phi1},
-$v_i$ should approach a constant of $i$
-upon convergence (i.e., when $\pi_i \approx p_i$)
-of an adaptive FDS.
+$v_i$ should ideally be a constant of $i$
+such that $\pi_i \approx p_i$.
 
@@
 
+We shall mainly focus on the adaptive methods below.
+%
 In the WL algorithm\cite{wang2001, wang2001pre},
@@
   \delta_{i, \, i(t)}
-  \frac{ \alpha(t) } { p_i }.
+  \frac{ \alpha(t) } { p_i }
+  ,
 \label{eq:wl_update}
@@
 %
-We refer to the above scheme as the \emph{single-bin scheme},
+We refer to this scheme as the \emph{single-bin scheme},
 for it applies only to the bias potential
@@
   w_{i, \, i(t)}
-  \frac{ \alpha(t) } { p_i },
+  \frac{ \alpha(t) }
+       { p_{i(t)} },
   \label{eq:mbin_update}
@@
 where $w_{ij}$ are a set of coefficients satisfying
-$\sum_{j=1}^n w_{ij} = 1$.
-%
-We shall discuss other restrictions on $w_{ij}$
+the restrictions to be discussed
 in Sec. \ref{sec:updating-matrix}.
@@
 The Gaussian updating scheme
-used by metadynamics
+used in metadynamics
 is a multiple-bin scheme.
@@
 the distribution collected from
-the trajectory is identical to $p_i$ in long times.
+an adaptive FDS simulation
+is identical to $p_i$ in long times.
 %
@@
 Eq. \eqref{eq:pi_p_v} no longer holds
-at all times.
-%
-In other words,
-the error in $v_i(t)$
+at all times,
+and the error in $v_i(t)$
 depends on the updating magnitude\cite{
@@
 Thus, one has to decrease $\alpha$ over time
-to reduce the error of $v_i$,
-in order to get the unbiased distribution
-$\ln p_i^*$ via Eq. \eqref{eq:v_def}.
-
-In the original WL scheme,
-the updating magnitude $\alpha$ (or $\ln f$
-in the original paper) is kept as a constant
-for a period of time,
-referred to as a stage below.
-%
-The histogram collected in the stage is monitored.
+to improve the estimate of the distribution,
+$p^*_i$. %via Eq. \eqref{eq:v_def}.
+
+\subsubsection{Updating magnitude and the inverse-time formula}
+
+In the original WL papers\cite{
+wang2001, wang2001pre},
+the updating magnitude $\alpha$
+(or $\ln f$ there)
+is controlled stage-wise.
+%
+In each stage, $\alpha$
+is kept as a constant,
+and the histogram along $i$
+is collected and monitored.
 %
 Once the histogram is sufficiently flat,
-we are allowed to enter a new stage
-with $\alpha$
-reduced by a factor of one
-half\cite{wang2001, wang2001pre}.
+we switch to a new stage
+with $\alpha$ reduced by a factor of $1/2$\cite{
+wang2001, wang2001pre}.
 %
@@
 saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
-%
-
-\subsubsection{Updating magnitude and the inverse-time formula}
 
@@
 a better way
-of updating the schedule $\alpha(t)$
+of controlling the schedule $\alpha(t)$
 in late stages
@@
 %
-where $t$ is the number of steps
-from the beginning of the MC simulation,
-which shall be referred to as the ``time'' below.
+where $t$,
+referred to as the ``time'' below,
+is the number of steps
+from the beginning of the simulation.
 %
@@
 the optimal formula for $\alpha(t)$
-might be $C/t$ with a constant $C$
+might be $C/t$ with a $C$
 different from $1.0$\cite{
@@
 were developed for umbrella sampling and metadynamics,
-particularly the
-self-healing umbrella sampling
+particularly, the self-healing umbrella sampling
 and well-tempered metadynamics\cite{
@@
 %
-Our aim below is to the optimal schedule $\alpha(t)$,
+Our aim below is to find the optimal schedule $\alpha(t)$,
 for a general updating scheme.
@@
 \subsection{\label{sec:single-bin}
-Single-bin scheme}
+Single-bin updating scheme}
 
@@
 which is equal to $1.0$
-for the current bin $i(t)$
+for the bin $i(t)$ at time $t$
 or zero otherwise.
@@
 The initial states and the stochastic forces
-during the process may, however, be different.
+during the process may differ, however.
 
@@
 To proceed, we first observe that
-the average $\bar v = \sum_{i = 1}^n p_i \, v_i$
+the average
+$$
+\bar v = \sum_{i = 1}^n p_i \, v_i
+$$
 increases steadily over time:
@@
 %
-However, we expect the difference between $v_i$ and $\bar v$,
+However, we expect the difference %between $v_i$ and $\bar v$,
 %
@@
   \approx
-\frac{                 p_i (1 - x_i) }
-     { \sum_{ j = 1}^n p_j (1 - x_j) }
-\notag
-\\
-&\approx
-p_i \, \left(
-         1 - x_i + \sum_{j=1}^n p_j \, x_j
-       \right)
-=
   p_i \, (1 - x_i),
@@
 which follows directly from Eq. \eqref{eq:x_def}.
-%
+\note{For \eqref{eq:hdet_linearize}. Following the second step
+$$
+\langle h_i(t) \rangle
+\approx
+\frac{                 p_i (1 - x_i) }
+     { \sum_{ j = 1}^n p_j (1 - x_j) }
+.
+$$
+}%
 From Eqs.
@@
   =
-\sum_{i = 1}^n p_i \, \left\langle (v_i - \bar v)^2 \right\rangle,
+  \sum_{i = 1}^n p_i \, \left\langle (v_i - \bar v)^2 \right\rangle
+  .
 \label{eq:error_sum}
 \end{equation}
-where
-$\bar v = \sum_{i = 1}^n p_i v_i$.
 %
 Our aim is to find the $\alpha(t)$
-that minimizes Eq. \eqref{eq:error_sum}.
-%
-Fortunately, we find that it suffices to consider
-a simpler one-variable problem
-in this case.
+that minimizes the error $E$
+at the end of the simulation period $T$.
+%
+Fortunately, it suffices to consider
+a simpler one-variable problem.
 
@@
 \begin{equation}
-\dot x(t) = -\alpha(t) \, \lambda \left[ x(t) - \xi(t) \right],
+\dot x(t) = -\alpha(t) \, \left[ x(t) - \xi(t) \right],
 \label{eq:dxdt_alpha}
@@
   =
-\kappa(t - t').
+  \kappa(t - t')
+  .
 \notag
@@
 %
-Additionally, we require
-\begin{equation}
-  \lim_{t \rightarrow \pm\infty} \kappa(t) = 0.
+Additionally,
+we assume that $\kappa(t)$ is an even function and
+\begin{equation}
+  \lim_{t \rightarrow \infty} \kappa(t)
+  = 0
+  .
   \label{eq:kappa_limit}
@@
 \begin{equation}
-  \alpha(t) = \frac{1}{\lambda \, (t + t_0)}.
-\label{eq:alpha_opt}
+  \alpha(t) = \frac{ 1 }{ t + t_0 }.
+\label{eq:alpha_invt1}
 \end{equation}
@@
 =
-x(0) \, e^{-\lambda \, q(T)}
+x(0) \, e^{-q(T)}
 +
@@
 %
-where,
+where
 %
@@
 &\equiv
-e^{-\lambda \, q(T) + \lambda \, q'}.
+e^{q' - q(T)}.
 \label{eq:u_def}
@@
 \langle x^2(0) \,
-\rangle e^{-2 \, \lambda \, q(T)}
+\rangle e^{-2 \, q(T)}
 \notag
@@
 This is equivalent to variating $u$
-with fixed endpoint values,
-$u\bigl( q(0) \bigr)  = e^{-\lambda \, q(T)}$
+with fixed endpoints,
+$u\bigl( q(0) \bigr)  = e^{- q(T)}$
 and
@@
 of Eq. \eqref{eq:x2t_average} is fixed during
-the variation process.
+the process.
 %
@@
 where we have dropped the boundary terms
-by using Eq. \eqref{eq:kappa_limit}.
+by assuming Eq. \eqref{eq:kappa_limit}
+and a large $T$.
 %
@@
 $$
-e^{-\lambda \, q(T) + \lambda \, q(t')}
+e^{ q(t') - q(T) }
 =
@@
 with respect to $t'$
-yields Eq. \eqref{eq:alpha_opt}.
-
-\subsubsection{Optimal $\alpha(t)$ for the single-bin scheme}
-
-Going back to the problem of
-minimizing Eq. \eqref{eq:error_sum}
-for the single-bin scheme.
-%
-Note that Eq. \eqref{eq:dxdt_WL}
+yields Eq. \eqref{eq:alpha_invt1}.
+
+\subsubsection{Optimal schedule for the single-bin scheme}
+
+For the problem of
+minimizing the error defined in Eq. \eqref{eq:error_sum}
+for the single-bin scheme,
+we note that Eq. \eqref{eq:dxdt_WL}
 is of the form of Eq. \eqref{eq:dxdt_alpha}.
 %
-Thus, with $\lambda = 1$,
-Eq. \eqref{eq:alpha_opt}
-will optimize each term $\langle x_i^2 \rangle$,
-and hence the sum.
-%
-In other words, the optimal $\alpha(t)$
-is given by
-\begin{equation}
-\alpha(t)
-=
-\frac{    1     }
-     {  t + t_0 }
-,
-\label{eq:alpha_invt1}
-\end{equation}
-and Eq. \eqref{eq:alpha_invt}
-is the case of $t_0 = 0$.
+Thus,
+the optimal schedule $\alpha(t)$ is given by
+Eq. \eqref{eq:alpha_invt1}
+for it optimizes each term $\langle x_i^2 \rangle$,
+and hence the sum;
+Eq. \eqref{eq:alpha_invt}
+is the special case of $t_0 = 0$.
 %
@@
 we show that $t_0$ can be adjusted
-to take into account for the initial error.
+to accommodate the initial error.
 
 \subsection{\label{sec:multiple-bin}
-Multiple-bin scheme}
-
-We now consider the multiple-bin scheme,
-Eq. \eqref{eq:mbin_update}.
+Multiple-bin updating scheme}
+
+We now consider a multiple-bin updating scheme,
+given in Eq. \eqref{eq:mbin_update}.
 %
@@
 %
-The relative update magnitudes, $w_{ij}$,
-constitute a matrix $\mathbf w$.
+The relative updating magnitudes,
+$w_{ij}$'s,
+form a matrix $\mathbf w$.
 %
@@
 
-The optimization procedure is similar to the single-bin case,
-whereas the matrix nature brings about
-some technical complexities that require
-certain restrictions and approximations.
+The optimization procedure is similar to the single-bin case.
+However, we shall adopt some restrictions and approximations
+to simplify the discussion.
 
@@
 if $h_j(t)$ were the same as $p_j$,
-the rates of change $\dot v_i(t)$'s
-should be independent of $i$.
+the $\dot v_i(t)$ should be a constant independent of $i$.
 %
@@
 \begin{equation}
-  \sum_{j = 1}^n w_{ij} = 1.
+  \sum_{j = 1}^n w_{ij} = 1
+  .
 \label{eq:w_sumj}
@@
 %
-The above condition means that $(1, \dots, 1)^T$
+In other words, $(1, \dots, 1)^T$
 is a right eigenvector of $\mathbf w$
@@
 resembles a transition matrix,
-although the elements can be negative.
+except that the elements can be negative.
 
@@
 
-Further, by Eq. \eqref{eq:w_detailedbalance},
+By assuming Eq. \eqref{eq:w_detailedbalance},
 we can diagonalize $\mathbf w$ with a set of
@@
 \begin{equation}
-  \sum_{i = 1}^n \phi_{ki} \, w_{ij} = \lambda_k \, \phi_{kj},
+  \sum_{i = 1}^n \phi_{ki} \, w_{ij}
+  =
+  \lambda_k \, \phi_{kj}
+  ,
 \label{eq:eig_w}
@@
 \begin{align}
-\sum_{k = 1}^n \phi_{ki} \, \phi_{kj}
-&= \delta_{ij} \, p_i,
+  \sum_{k = 1}^n
+    \phi_{ki} \, \phi_{kj}
+  &=
+  \delta_{ij} \, p_i,
   \label{eq:eig_orthonormal_cols}
   \\
-\sum_{k = 1}^n \frac{ \phi_{ik} \, \phi_{jk} }{ p_k }
-&= \delta_{ij}.
+  \sum_{i = 1}^n
+    \frac{ \phi_{ki} \, \phi_{li} }
+         { p_i }
+  &=
+  \delta_{kl}
+  .
 \label{eq:eig_orthonormal_rows}
@@
 Particularly, there is always an eigenvector
-corresponding to the equilibrium distribution.
+corresponding to the target distribution, $\mathbf p$.
 %
@@
 \begin{equation}
-\phi_{1k} = p_k,
+  \phi_{1i} = p_i,
 \label{eq:eigenmode1}
@@
 %
-which satisfies Eq. \eqref{eq:eig_orthonormal_rows}.
-%
-The corresponding eigenvalue is $\lambda_1 = 1$.
+which satisfies Eq. \eqref{eq:eig_orthonormal_rows}
+for $k = l = 1$.
+%
+The corresponding eigenvalue is $\lambda_1 = 1$
+by Eq. \eqref{eq:w_balance}.
+%
+Further, for $k \ge 2$, we have
+%
+\begin{equation}
+  \sum_{ i = 1 }^n \phi_{ki}
+  =
+  \sum_{ i = 1 }^n
+    \frac{ \phi_{ki} \, \phi_{1i} }
+         { p_i }
+  =
+  0
+  .
+\label{eq:ortho1}
+\end{equation}
 
@@
 In a \emph{stable} updating scheme,
-no fluctuating mode can increase indefinitely,
-which means all eigenvalues $\lambda_k$ of $\mathbf w$
-must be nonnegative.
+no eigenmode grows indefinitely,
+which means all eigenvalues $\lambda_k$'s
+are nonnegative.
 
@@
 \begin{aligned}
-\dot x_i
+  \dot x_i(t)
   &= \alpha(t) \sum_{j=1}^n w_{ij}
-\left( \frac{ h_j } { p_j }  - 1 \right)
+  \left( \frac{ h_j(t) } { p_j }  - 1 \right)
   \\
@@
 we have expanded the right-hand side
-in linear terms of $x_j$.
+in linear terms of $x_j$
+on the second line.
 
@@
 Now by diagonalizing the matrix $\mathbf w$
-using the eigenvectors defined in Eqs. \eqref{eq:eig_w},
-\eqref{eq:eig_orthonormal_cols},
-and
-\eqref{eq:eig_orthonormal_rows},
+using the eigenvectors defined in Eqs.
+\eqref{eq:eig_w}-\eqref{eq:eig_orthonormal_rows},
 we get a set of decoupled equations
@@
   -\alpha(t) \, \lambda_k \, [y_k(t) - \theta_k(t)],
-\notag
-%\label{eq:yt_diffeq}
+\label{eq:yt_diffeq}
 \end{equation}
@@
 \begin{align}
-  y_k &= \sum_{j=1}^n \phi_{kj} \, x_j,
+  y_k(t)
+  &=
+  \sum_{j=1}^n
+    \phi_{kj} \, x_j(t)
+  ,
   \label{eq:y_def}
   \\
-  \theta_k &= \sum_{j=1}^n \phi_{kj} \frac{ \zeta_j}{ p_j}.
+  \theta_k(t)
+  &=
+  \sum_{j=1}^n
+    \phi_{kj} \frac{ \zeta_j(t) }
+                   { p_j        }
+  .
   \label{eq:eta_def}
@@
 
-The error function can be obtained
-from Eq. \eqref{eq:eig_orthonormal_cols}
-as
+The error function defined in Eq. \eqref{eq:error_sum}
+can be rewritten in terms of the eigenmodes
 \begin{align}
@@
   \sum_{i = 1}^n p_i \,
-                 \langle x_i^2 \rangle
+    \left\langle x_i^2 \right\rangle
   =
   \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \,
-                     \langle x_i \, x_j \rangle
-  =
-  \sum_{k = 1}^n \langle y_k^2 \rangle,
-  \label{eq:y2_sum}
+    \left\langle x_i \, x_j \right\rangle
+  =
+  \sum_{k = 1}^n
+    \left\langle
+      y_k^2
+    \right\rangle
+  ,
+\notag
 \end{align}
 %
-which is a superposition of the components of the modes.
+where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
 %
@@
 %
-Thus, we can start the last sum in Eq. \eqref{eq:y2_sum}
-from $k = 2$.
+Thus, we can start the last sum %in Eq. \eqref{eq:y2_sum}
+from $k = 2$,
+\begin{equation}
+  \Err
+  =
+  \sum_{k = 2}^n
+    \left\langle
+      y_k^2
+    \right\rangle
+  .
+\label{eq:y2_sum}
+\end{equation}
 
 In analogous to Eq. \eqref{eq:x2t_average},
-we have, from Eq. \eqref{eq:y2_sum},
-the error after a period of $T$,
+we can write,
+from Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
+the error after a period $T$,
 %
@@
 %
-which is the sum of the residual error
+as the sum of the residual error
+(from the decay of the initial error)
 %
@@
 %
-from the decay of the initial error,
 and the asymptotic error
@@
 %
-where,
+where
 \begin{align*}
@@
 Unfortunately, the above expression
-appears to be useless in a general setting.
+appears to be useless in general.
 }
@@
   -\lambda_k \, q(T) + \lambda_k \, q(t)
-  = \log c_k + \log\left( t + t_{k0} \right).
+  = \ln c_k + \ln \left( t + t_{k0} \right).
   $$
@@
       \left(
-        \log \frac { H_i }
+        \ln \frac { H_i }
                   { p_i }
@@
   \frac{ 1 } { \lambda }
-  \log\left(
+  \ln\left(
     \frac{ T + t_0 } { t_0 }

### 2016-04-14 doc/invt.tex
invt.tex: update to Sec. II.C

@@
   those of neighboring bins according to
-  a Gaussian wave packet.
+  a Gaussian window.
   %
@@
 while metadynamics adopts an extended
-Gaussian wave packet
+Gaussian window
 covering several neighboring bins.
@@
 which is sensitive to the simulation length
-and the width of the Gaussian wave packet.
+and the width of the Gaussian window.
 %
@@
 \begin{equation}
-  v_i \equiv V_i - \ln p^*_i + \ln p_i.
+  v_i
+  \equiv
+  V_i
+  +
+  \ln \frac { p_i }
+            { p^*_i }
+  .
   \label{eq:v_def}
@@
   \kappa_k(t) \approx \Gamma_k \, \delta(t)
+  ,
 \label{eq:kappa_delta}
@@
 %
-where,
+where
 \begin{equation}
@@
   once $\alpha(t)$ has dropped below a certain level.
-}
-%
+}%
 \note{
@@
   $$\,$$
-}
+}%
 In this way,
@@
 
-We evaluate the integrals of the autocorrelation functions,
+We can evaluate the integrals of the autocorrelation functions,
 $\Gamma_k$'s,
@@
 %
-In this case,
+To see this, we note that
 the bin index $i$ is an independent random variable
-at each time step $t$, and
-%
-\begin{equation}
+at each time step $t$ in this case, and
+%
+\begin{align}
   \left\langle
@@
   \right\rangle
-  =
-  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \, \delta(t - t'),
+  &=
+  \langle h_i(t) \, h_j(t') \rangle
+  -
+  \langle h_i(t) \rangle
+  \langle h_j(t') \rangle
+  \notag
+  \\
+  &=
+  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \,
+  \delta_{t, t'}
+  ,
   \label{eq:zz_perfect}
-\end{equation}
+\end{align}
 %
@@
   =
-  \sum_{j,l = 1}^n
+  \sum_{i,j = 1}^n
+  \frac{ \phi_{ki} } { p_i }
   \frac{ \phi_{kj} } { p_j }
-  \frac{ \phi_{kl} } { p_l }
-  \langle \zeta_j(t) \, \zeta_l(t') \rangle
-  %
-  = \delta(t - t'),
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  %
+  = \delta_{t, t'},
 \notag
@@
 where we have used
-Eqs. \eqref{eq:eig_orthonormal_rows},
-\eqref{eq:eigenmode1},
+Eqs. \eqref{eq:eig_orthonormal_rows}-\eqref{eq:ortho1},
 \eqref{eq:eta_def},
-\eqref{eq:zz_perfect},
 and
-$$
-  \sum_{j = 1}^n \phi_{kj}
-  =
-  \sum_{j = 1}^n \frac{ \phi_{kj} \, \phi_{1j} } { p_j }
-  =
-  \delta_{k1}
-  =
-  0.
-$$
+\eqref{eq:zz_perfect}.
 %
@@
   &=
-  \sum_{j,l = 1}^n
+  \sum_{i,j = 1}^n
+  \frac{ \phi_{ki} } { p_i }
   \frac{ \phi_{kj} } { p_j }
-  \frac{ \phi_{kl} } { p_l }
-  \langle \zeta_j(t) \, \zeta_l(t') \rangle
-  \\
-  %
-  &=
-  \sum_{j,l = 1}^n
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  \\
+  %
+  &=
+  \sum_{i,j = 1}^n
+  \frac{ \phi_{ki} } { p_i }
   \frac{ \phi_{kj} } { p_j }
-  \frac{ \phi_{kl} } { p_l }
   \left(
-    p_j \, \delta_{j, l} - p_j \, p_l
+    p_i \, \delta_{i, j} - p_i \, p_j
   \right)
-  \, \delta(t - t')
+  \, \delta_{t, t'}
   \\
@@
     -
+    \sum_{i = 1}^n \phi_{ki}
     \sum_{j = 1}^n \phi_{kj}
-    \sum_{l = 1}^n \phi_{kl}
   \right)
-  \, \delta(t - t')
-  \\
-  &=
-  (1 - 0 \cdot 0) \, \delta(t - t').
+  \, \delta_{t, t'}
+  \\
+  &=
+  (1 - 0 \cdot 0) \, \delta_{t, t'}.
   \end{aligned}
@@
 
-Another special case is the one-step process\cite{vankampen}:
-in each step, the system hops to the previous bin $i - 1$
-or the next bin $i + 1$ with equal probability $1/2$.
+Another special case is the
+one-dimensional one-step process\cite{vankampen},
+which models a local sampling process.
+In each step, the system hops to
+one of the two neighboring bins
+with equal probability $1/2$.
 %
@@
       \vdots & &  & & \vdots \\
-      0 & \dots & \frac 1 2 & 0  & \frac 1 2 \\
-      \frac {1 - \sigma} 2 & \dots & 0 & \frac 1 2 & \frac \sigma 2
+      \frac {1 - \sigma} 2 & \dots &  & \frac 1 2 & \frac \sigma 2
     \end{array}
@@
 %
-For a large class of updating schemes
+For many updating schemes
 (cf. Sec. \ref{sec:band-matrix}),
@@
 the matrices $\mathbf Q$ and $\mathbf w$
-share eigvenvectors, but not eigenvalues.
+share the eigenvectors, but not the eigenvalues.
 %
@@
 decays as $\kappa_k(0) \, \Lambda_k^t$,
-with the same-time value
-$\kappa_k(0) = 1$
-being the same as
-that in perfect sampling.
+with $\kappa_k(0) = 1$ being the same as
+the value in perfect sampling.
 %
@@
   =
-\frac{ 1 + \Lambda_k } { 1 - \Lambda_k }
-=
-\cot^2 \left[ \frac{ (k - 1) \, \pi } { 2 \, n } \right].
+  \frac{ 1 + \Lambda_k }
+       { 1 - \Lambda_k }
+  =
+  \cot^2 \left[
+    \frac{ (k - 1) \, g \, \pi }
+         { 2 \, n }
+  \right]
+  .
 \label{eq:Gamma_onestep}
@@
 %
-This case is intended to model a local sampling process.
 
 \subsection{\label{sec:mbin_opta}
-Optimal schedule for the multiple-bin updating scheme}
+Optimal schedule for multiple-bin updating schemes}
 
@@
 
-Before addressing the optimal $\alpha(t)$,
-we shall first consider a constant schedule
-$\alpha(t) \equiv \alpha_0$,
-and compute the equilibrium error, $\Err$,
-in this case.
+Before addressing the optimal schedule
+for a multiple-bin scheme,
+we shall first compute the equilibrium error
+under a constant schedule,
+$\alpha(t) = \alpha_0$.
 %
@@
   &
-  \stackrel{ T' \to \infty }
-  { =\joinrel=\joinrel=\joinrel= }
+  =
+  %\stackrel{ T' \to \infty }
+  %{ =\joinrel=\joinrel=\joinrel= }
   %
@@
     \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0
+  ,
+  \qquad
+  ( T' \to \infty )
   .
@@
 %
-That is, for $t < 0$, we assume
-$\alpha(t) = \alpha_0$, and
+That is, we assume
+$\alpha(t) = \alpha_0$
+for $t < 0$, and
 the system has entered the equilibrium
@@
 Note that the residual error
-depends on the schedule $\alpha(t)$
-only through endpoint value of
+depends on $\alpha(t)$ only through
 the integral $q(T) = \int_0^T \alpha(t) \, dt$.
 %
-Thus, if we variate the schedule under a fixed $q(T)$,
+Thus, if we variate $\alpha(t)$ under a fixed $q(T)$,
 this residual part can be ignored,
@@
 
-As shown above, with fixed endpoint value, $q(T)$,
+As shown above, with a fixed value of $q(T)$,
 the optimization of the total error is equivalent to
@@
 does not explicitly depend on time, $t$,
-the Hamiltonian,
+the Hamiltonian
 %
@@
        { \partial \dot q     }
-  \, \partial \dot q
+  \, \dot q
   -
@@
   c^2
+  ,
   \quad
@@
 %
-Integrating the square root of the left-hand side, we get
+Integrating the square root, we get
 %
@@
 %
-Since we only optimized the asymptotic error $E_A$,
-the parameter $q(T)$ can be further optimized
-such that total error, $E = E_R + E_A$, is minimal.
+Since we only optimized the asymptotic error, $E_A$,
+in th above,
+the parameter $q(T)$ can be further adjusted
+to minimize the total error, $E = E_R + E_A$.
 
 Unfortunately,
-Eq. \eqref{eq:q_opt} does not provide much information
-on the behavior of $\alpha(t)$,
-and the explicit solution
-can be formidable in the general case.
-%
-We can nonetheless obtain an expression regarding $\alpha(t)$
+Eq. \eqref{eq:q_opt} is not an explicit equation of $\alpha(t)$.
+%
+However,
 by differentiating Eq. \eqref{eq:Lagrangian_const},
-which yields
+we have
 %
@@
 %
-The expression in the middle is a weighted average
+The expression in the middle is an average
 of the eigenvalues, $\lambda_k$'s,
-and we shall refer to it as the instantaneous
-eigenvalue, $\lambda(t)$.
-%
-It gives the rate of change of
-the inverse updating magnitude.
-
-For the single-bin scheme, $\lambda_k = 1$,
-and the instantaneous eigenvalue is always unity.
-Thus,
+weighted by $\Gamma_k \, \lambda_k^2 \, u_k^2$.
+%
+We shall refer to the average as the instantaneous
+eigenvalue, $\lambda(t)$,
+which gives the rate of change of $\alpha^{-1}(t)$.
+
+\note{For the single-bin scheme, $\lambda_k = 1$,
+and $\lambda(t)$ is always unity.
+%
+So,
 $$
@@
 $t_0(t)$ is a slowly-varying function.
-
-By differentiating Eq. \eqref{eq:dinvadt},
-we can show that instantaneous eigenvalue,
-$\lambda(t)$, is an increasing function.
+}
+
+By further differentiating Eq. \eqref{eq:dinvadt},
+we can show that $\lambda(t)$
+is an increasing function.
 %
@@
   is nonnegative by the Cauchy-Schwarz inequality,
-  or by the nonpositiveness of the discriminant
+  or by the nonpositivity of the discriminant
   of the following nonnegative definite quadratic form of $X$:
@@
   Thus, $\dot \lambda \ge 0$, and $\lambda(t)$ is monotonically increasing.
-}
-%
-Below we shall study the initial and final values
-of $\lambda(t)$.
-%
-At the end of a simulation,
-we have $t = T$, and $u_k = 1$.
-%
-Thus,
-the instantaneous eigenvalue
+}%
+%
+Below we shall study the initial and final values.
+%
+At the end of the simulation,
+$t = T$, we have $u_k = 1$, and
 %
@@
 \end{equation}
-is an average of $\lambda_k$'s weighted
-by $\Gamma_k \, \lambda_k^2$,
-and it is independent of the simulation length, $T$.
-%
-Further, it often occurs that $\Gamma_k$
-increases with $\lambda_k$,
+is independent of the simulation length, $T$.
+%
+Further, $\Gamma_k$ often increases with $\lambda_k$,
 such that the relative weight $\Gamma_k \, \lambda_k^2$
@@
 It follows that $\lambda(T)$ will be dominated
-the larger values of $\lambda_k$'s.
+by the larger values of $\lambda_k$'s.
 
 By contrast,
-at the beginning of a simulation,
+at the beginning of the simulation,
 the relative weight, $\Gamma_k \, \lambda_k^2 \, u_k^2(0)$,
 is heavily influenced by the exponential factor
-$u_k^2(0) = e^{ -2 \, \lambda_k \, q(T) }$,
-and hence is sensitive to the simulation length, $T$.
+$u_k^2(0) = e^{ -2 \, \lambda_k \, q(T) }$.
+%
+Thus, $\lambda(0)$
+is sensitive to the simulation length, $T$.
 %
 A longer simulation means a smaller $\lambda(0)$,
-and that $\lambda(t)$ would be small
+and that $\lambda(t)$ would stay small
 for longer near the beginning of the simulation.
@@
 \subsubsection{\label{sec:measure_gamma}
-Measurement the integrals
+Measurement of the integrals
 of the autocorrelation functions}
@@
 
-The idea is to measure $\Gamma_k$'s
-via a simulation with a constant schedule,
+The idea is to run
+a simulation with a constant schedule,
 $\alpha(t) = \alpha_0$,
-using Eq. \eqref{eq:y2_eql}.
+and then use Eq. \eqref{eq:y2_eql}.
 %
@@
 we often do not know the intrinsic distribution, $p^*_i$,
-beforehand, and thus cannot compute
+in advance to compute
 the shifted bias potential, $v_i$,
@@
 %
-To circumvent the problem, we shall use
-the following modification:
+To circumvent the problem, we shall
+modify Eq. \eqref{eq:y2_eql} as
 %
@@
   &=
-(V_i - \bar V) + s_i
+  (V_i - \bar V) + s_i,
   \\
@@
 Finally, by using Eq. \eqref{eq:y2_eql},
-we get Eq. \eqref{eq:varY}.
-
-In practice, Eq. \eqref{eq:varY}
-may fail if some eigenvalue $\lambda_k$ is zero.
-%
-However, the corresponding modes can be ignored
+we reach Eq. \eqref{eq:varY}.
+
+\note{In practice, Eq. \eqref{eq:varY}
+in inapplicable to zero eigenvalues, $\lambda_k$'s.
+%
+However, these eigenmodes can be ignored
 as they do not contribute to the Lagrangian defined by
 Eq. \eqref{eq:error_asym_Lagrangian}.
+}
 
@@
 \item
-Given an updating matrix, $\mathbf w$,
+From the updating matrix, $\mathbf w$,
 compute the eigenvalues, $\lambda_k$'s.
@@
 updating magnitude, $\alpha_0$.
-Use Eq. \eqref{eq:varY} to compute $\Gamma_k$'s.
+
+\item
+Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}.
 
@@
 \item
-Compute the optimal schedule from
+Compute the optimal schedule by
 numerically integrating Eq. \eqref{eq:q_opt}.
@@
 \item
-Modify the value of $q(T)$ and repeat the above two steps,
+Modify the value of $q(T)$ and repeat the previous two steps,
 until the minimal error $E$ is found.
@@
   \node[draw, text width={width("Updating ")}]
-    (w) at (0, 3) {Updating \\matrix, $w$};
-
-  \node[draw, text width={width("System  11")}]
-    (sys) at (0, 1) {System \& \\sampling \\process};
+    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
+
+  \node[draw, text width={width("System and")}]
+    (sys) at (0, 1) {System and\\sampling \\process};
 
@@
   \node[draw, text width={width("Schedule ")}]
-    (alpha) at (5, 1) {Optimal schedule, \\ $\alpha(t)$};
+    (alpha) at (5.2, 1) {Optimal schedule, \\ $\alpha(t)$};
 
@@
 
-  \draw[->] (w) edge (lambda)
+  \node[inner sep=0, minimum size=0] (M1) at (3.7, 1) {};
+  \node[inner sep=0, minimum size=0] (M2) at (4.0, 1) {};
+  \node[inner sep=0, minimum size=0] (N2) at (4.0, -0.2) {};
+  \node[inner sep=0, minimum size=0] (N3) at (6.2, -0.2) {};
+  \node[inner sep=0, minimum size=0] (M3) at (6.2, 1) {};
+
+  \draw[->, >=latex, thick]
+    (w) edge (lambda)
     (w) edge (gamma)
     (sys) edge (gamma)
-            (lambda) edge (alpha)
-            (gamma) edge (alpha)
-            (T) edge (alpha)
     (alpha) edge (err);
+
+  \draw[->, >=latex, thick]
+    (lambda) -| (M1)
+    (gamma) -- (M1)
+    (T) -| (M1)
+    (M1) -- (alpha);
+
+  \draw[->, >=latex, thick]
+    (M2) -- (N2) -- (N3) -- (M3);
 
@@
 Many commonly-used updating schemes
-can be characterized by a rigid wave packet
-whose center drifts with the current bin.
+can be characterized by a rigid window
+centered on the current bin.
 %
@@
 $\mathbf w$,
-can be reduced to a convolution with the wave packet,
+can be reduced to a convolution with the window,
 or the updating kernel.
@@
 
-While the translational invariance of
-the updating scheme
-naturally suits a periodic variable\cite{dama2014},
-it can also be extended to a non-periodic variable
-by imposing the reflective boundary condition.
+\subsubsection{\label{sec:bandkernel}
+Updating kernel and eigenvalues}
+
+While translationally-invariant
+updating schemes
+naturally suit periodic variables\cite{
+dama2014},
+they can also be extended to non-periodic variables
+by imposing the reflective boundary condition\cite{
+bussi2006}.
 %
@@
 
-\subsubsection{\label{sec:bandkernel}
-Updating kernel and eigenvalues}
-
-It is natural to introduce a homogeneous updating scheme
+It is natural to introduce homogeneous updating schemes
 for a periodic variable\cite{dama2014}.
 %
-Consider the following matrix, $\mathbf w$,
+The updating matrix, $\mathbf w$,
+assume the following form
 %
@@
 %
-and $\mu_l = 0$ for $l > b$.
+and $\mu_l = 0$ for $|l| > b$.
 %
@@
 %
-If $b = n/2$, then $\mu_{-b}$ is dropped
-from Eq. \eqref{eq:mu_normalization}.
-%
-These numbers constitutes a \emph{updating kernel}.
-
-This above matrix, $\mathbf w$,
+If $b = n/2$, then $\mu_{-b}$ is removed 
+from the sum in Eq. \eqref{eq:mu_normalization}.
+%
+These numbers constitutes an \emph{updating kernel}.
+
+The above matrix, $\mathbf w$,
 being symmetric to the indices $i$ and $j$,
 satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
-for the flat distribution $p_i = 1/n$.
+for the flat distribution.
 %
 To find the eigenvectors,
-we note that for a periodic variable, $\phi_i$,
-the out-of-boundary values are defined by
+we define, for a periodic variable, $\phi_i$,
+out-of-boundary values by
 %
@@
 %
-We then have
+Then
 %
@@
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-as
+compatible with Eq. \eqref{eq:phi_pbc},
 \begin{equation}
@@
 %
-with the eigenvalues being
+where $\ii^2 = -1$, and
+the eigenvalues are
 %
@@
 %
-Again the updating kernel satisfies
+Again the updating kernel should satisfy
 Eqs. \eqref{eq:mu_normalization}
-and \eqref{eq:mu_symm}.
-%
-However,
+and \eqref{eq:mu_symm},
+but
 the cutoff $b$ only need to be less than $n$.
@@
 The last two terms
-on the right-hand side of \eqref{eq:w_band_refl} are added
+on the right-hand side of Eq. \eqref{eq:w_band_refl} are added
 to avoid unintended distortion\cite{dickson2011, mcgovern2013}
-to the equilibrium distribution $p_i = 1/n$\cite{bussi2006},
+to the equilibrium distribution\cite{bussi2006},
 such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
@@
   \frac{   1     }
-       { T + 2 / \alpha_0 }
+       { T + t_0 }
   \sum_{ k = 2 }^K
@@
 
-We first tested the theory using
-the inverse-time schedule,
+We first studied the inverse-time schedule,
 %
@@
 %the bias potential of a neighborhood of the current one
-%is updated with relative weight specified by a Gaussian wave packet.
+%is updated with relative weight specified by a Gaussian window.
 
@@
 suggesting that they can be useful
-in real metadynamics simulations.
+in high-precision metadynamics simulations.
 

### 2016-04-16 doc/invt.tex
add errscan.gp
invt.tex: edit to the first paragraph of Results

@@
 %
-However, they are less well understood theoretically,
-as the updating loops
-break the microscopic reversibility
+However,
+%they are less well understood theoretically, as
+the updating loops break the microscopic reversibility
 of the underlying sampling mechanism.
@@
 Thus, in late stages of a long simulation,
-the updating is negligible and
-the adaptive simulation tends to
-the corresponding equilibrium one.
-%
-If the bias potential were converged to the exact value,
-this particular equilibrium simulation can be thought as
-the optimal asymptotic limit of the adaptive counterpart,
+the updating becomes negligible and
+%the adaptive simulation tends to
+%the corresponding equilibrium one.
+%
+if the bias potential were converged to the exact value,
+this adaptive simulation tends to the ideal equilibrium simulation
+%particular equilibrium simulation can be thought as
+%the optimal asymptotic limit of the adaptive counterpart,
 delivering the maximal efficiency.
@@
 %
-where $\sigma_{ij}(t)$ is an even function of $t$
+where $\sigma_{ij}(t)$ is an even function
 that vanishes at large $t$.
 %
-More explicitly,
+\note{More explicitly,
 if the transition matrix of a period $\tau$
@@
 $
+}%
 
@@
 %
-\begin{equation}
+$
   \left\langle \xi(t) \, \xi(t') \right\rangle
@@
   .
-\notag
-%\label{eq:noise_correlation}
-\end{equation}
+$
 %
@@
 %
-\begin{equation}
+$
 q(t) \equiv \int_0^t \alpha(t') \, dt',
-\notag
-%\label{eq:q_def}
-\end{equation}
+$
 %
@@
 %
-\begin{equation}
-  \ddot u\bigl( q(t') \bigr) = 0,
-  \qquad
-  \mathrm{or}
-  \;\;
-  \dot u\bigl( q(t') \bigr) = c,
-\notag
-%\label{eq:ddu_eq_0}
-\end{equation}
+$
+\ddot u\bigl( q(t') \bigr) = 0
+,
+$
+or
+$
+\dot u\bigl( q(t') \bigr) = c
+,
+$
 %
@@
   \; \dot q^2( t ),
-  \label{eq:error_asym_Lagrangian}
+\notag
+%\label{eq:error_asym_Lagrangian}
 \end{align}
@@
 as they do not contribute to the Lagrangian defined by
-Eq. \eqref{eq:error_asym_Lagrangian}.
+Eq. \eqref{eq:Lagrangian_const}.
 }
@@
 Thus,
-the eigenvalues, $\lambda_k$'s, of $\mathbf w$
-can be readily computed and manipulated
-via the shape of the updating kernel.
+there is a one-to-one correspondence between
+the eigenvalues of $\mathbf w$
+and the updating kernel.
 
@@
 we define, for a periodic variable, $\phi_i$,
-out-of-boundary values by
+the out-of-boundary values by
 %
@@
 
-Similar to Eq. \eqref{eq:phi_pbc},
+Parallel to Eq. \eqref{eq:phi_pbc},
 we define the out-of-boundary values
-of $\phi_j$ as
-%
-\begin{equation}
-  \phi_j
+of $\phi_i$ as
+%
+\begin{equation}
+  \phi_i
   =
   \begin{dcases}
-    \phi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
-    \phi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
+    \phi_{ 1 - i }           & \mathrm{for \;} i \le 0, \\
+    \phi_{ 2 \, n + 1 - i }  & \mathrm{for \;} i > n,
   \end{dcases}
-\notag
-%\label{eq:phi_refl}
+\label{eq:phi_refl}
 \end{equation}
@@
   $$
-}
-%
+}%
 Thus, we also have $n$ orthonormal eigenvectors,
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-as
+compatible with Eq. \eqref{eq:phi_refl},
 %
@@
 %
-\note{We derive the above two equations below.
-
-  To simplify the calculation,
-  let us first consider the unnormalized eigenvectors,
+\note{Derivation.
+  First consider the unnormalized eigenvectors,
   $$
@@
   which satisfies the reflective boundary condition,
-  %Eq. \eqref{eq:phi_refl}.
-  %
-  So
+  Eq. \eqref{eq:phi_refl}.
+  %
+  So, by Eq. \eqref{eq:wmul_to_convol},
   $$
@@
 %
-Equation \eqref{eq:wband_eigenvalue}
-shows that the eigenvalues, $\lambda_k$,
-of the updating scheme
+In both cases,
+the eigenvalues, $\lambda_k$'s,
 is the cosine transform of
@@
 %
-So, for the updating scheme to be stable,
+For the updating scheme to be stable,
 the cosine transform of the kernel
@@
 %
-If this is not so,
+\note{If this is not so,
 we can modify the kernel
 using the technique described
-in Sec. \ref{sec:stabilize_wband}.
+in Sec. \ref{sec:stabilize_wband}.}
 
@@
 %
-Thus, for this updating matrix to be stable,
-we need
-$\min \lambda_k = \lambda_n > 0$,
+Thus, this updating scheme is stable if
+$\min \lambda_k = \lambda_n \ge 0$,
 or
 \begin{equation}
-  \mu_1 <
+  \mu_1 \le
   \begin{dcases}
     \frac 1 4
-    & \mathrm{periodic, \;} n \mathrm{ \; even,}
+    & \mathrm{periodic \; and \;} n \mathrm{ \; even,}
     \\
     \frac 1 4
-    \cos^{-2}\frac{  \pi   }
+    \sec^2
+    \left( \frac{  \pi   }
                 { 2 \, n }
+    \right)
     & \mathrm{otherwise}.
   \end{dcases}
-\label{eq:nn_stable}
+\notag
+%\label{eq:nn_stable}
 \end{equation}
@@
 
-The Gaussian updating kernel is commonly
-adopted by metadynamics simulations.
-%
-Below we show that the corresponding updating scheme
-is stable, albeit less effective
-in reducing short-wavelength noises.
-
-We start by defining
-a continuous limit
-for $b = n - 1 \gg 1$:
-$x = l \, \Delta x$,
+The Gaussian updating scheme is commonly
+used in metadynamics simulations.
+%
+Below we shall show its stability
+in the continuous limit.
+
+For $n \gg 1$,
+we define
+$x = l \, g \, \Delta x$,
 and
@@
 %
-Then
+Then,
+by assuming the largest $b \approx n/g$,
 Eq. \eqref{eq:wband_eigenvalue}
-can be approximated by an integral:
+can be approximated as an integral:
 %
@@
     \mu(x) \, \cos \bigl[ (k-1) \, x \bigr] \, dx,
-\label{eq:lambda_int}
+\notag
+%\label{eq:lambda_int}
 \end{equation}
@@
   1 = 2 \int_0^\pi \mu(x) \, dx.
-\label{eq:mx_normalization}
-\end{equation}
-
-Particularly,
-for the Gaussian updating kernel,
-if $\Delta x \ll \sigma \ll 1$,
+\notag
+%\label{eq:mx_normalization}
+\end{equation}
+
+For the Gaussian updating kernel of width $\sigma$
+with $\Delta x \ll \sigma \ll 1$,
 we can extended
 the upper limit of the integrals
-in Eqs. \eqref{eq:lambda_int}
-and \eqref{eq:mx_normalization}
 to infinity, and
@@
               { 2 \, \sigma^2 }
-    \right),
+      \right)
+  .
 \notag
@@
 %
-The eigenvalues are given by
+If follows that all eigenvalues are positive:
 %
@@
       \right]
-\notag
-\\
-&=
-\exp\left[
-      -\frac{ \pi^2 }{ 2 }
-      \left(
-        \frac{ k - 1 }
-             {   n   }
-      \right)^2
-      \left(
-        \frac{  \sigma }
-             { \Delta x }
-      \right)^2
-    \right]
+  %\notag
+  %\\
+  %&=
+  %\exp\left[
+  %      -\frac{ \pi^2 }{ 2 }
+  %      \left(
+  %        \frac{ k - 1 }
+  %             {   n   }
+  %      \right)^2
+  %      \left(
+  %        \frac{  \sigma }
+  %             { \Delta x }
+  %      \right)^2
+  %    \right]
   .
@@
 \end{align}
-%
-The eigenvalues are all positive,
-suggesting stability of the updating scheme.
 
@@
 
-A practical problem of using homogeneous updating scheme
-is the following.
+%A practical problem of using homogeneous updating scheme
+%is the following.
 %
 A homogeneous updating scheme
-specified by an arbitrary set of numbers
-$\mu_1, \dots, \mu_b$, ($b < n - 1$),
+specified by an arbitrary kernel
 is not necessarily stable
@@
 %
-For example, the nearest-neighbor updating scheme
-is unstable if Eq. \eqref{eq:nn_stable} is violated.
-%
-Even the Gaussian updating scheme can be unstable
+\note{For example,
+  %the nearest-neighbor updating scheme
+  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
+  %
+  %Even
+  the Gaussian updating scheme can be unstable
   if the updating kernel is truncated.
+}%
 %
@@
 the updating kernel
-to meet the stability criteria.
+to stabilize it.
 %
@@
 by a cosine transform,
-and thus the relationship can be readily inverted.
-%
-For the periodic case,
+and thus the relation can be readily inverted.
+%
+In the periodic case,
 the inversion formula is
@@
 
-For a non-periodic variable,
-the explicit inversion formula is
+In the non-periodic case,
+the formula is
 %
@@
 we have defined
-$\lambda_k \equiv \lambda_{n + 2 - k}$
+$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
 for $k = n + 2, \dots, 2 \, n$,
@@
   (-1)^{ n - 1 }
-  \left[
   \lambda_1
@@
   2 \, \sum_{ k = 2 }^{ n }
-      (-1)^{k - 1} \, \lambda_k
-  \right]
+      (-1)^{n - k} \, \lambda_k
   ,
@@
 
-The inversion formula is derived from the cosine transform.
+  %The inversion formula is derived from the cosine transform.
   %
@@
   \begin{align}
-    \lambda_{ 2 \, n + 1 - k } = \lambda_k.
+    \lambda_{ 2 \, n + 2 - k } = \lambda_k.
     \tag{N3}
@@
   \mu_0
-  &=
+  =
   \lambda_2,
-  \\
+  \qquad
   \mu_1
-  &=
+  =
   \frac{ \lambda_1 - \lambda_2 }
@@
   \mu_0
-  &=
-  \frac{ 1 } { 3 }
-  \left(
-    \lambda_1 + 2 \, \lambda_3
-  \right)
+  =
+  \frac{ \lambda_1 + 2 \, \lambda_3 } { 3 }
   ,
-  \\
+  \quad
   \mu_1
-  &=
-  \frac{ 1 } { 3 }
-  \left(
-    \frac 3 2 \lambda_2
-    -
-    \frac 3 2 \lambda_3
-  \right)
+  =
+  \frac{ \lambda_2 - \lambda_3 } { 2 }
   ,
-  \\
+  \quad
   \mu_2
-  &=
-  \frac{ 1 } { 3 }
-  \left(
-    \lambda_1
-    -
-    \frac 3 2 \lambda_2
-    +
-    \frac 1 2 \lambda_3
-  \right)
+  =
+  \frac { 2 \, \lambda_1 - 3 \, \lambda_2 + \lambda_3 } { 6 }
   .
@@
 
-Now given a updating kernel, $\mu_1, \dots, \mu_n$,
+Now given an updating kernel,
 we can compute the eigenvalues from
@@
 or \eqref{eq:mu_from_lambda_refl},
-we get a new updating kernel
-$\mu^\mathrm{new}_1, \dots, \mu^\mathrm{new}_{n - 1}$,
+we get a new updating kernel,
 which is stable by construction.
 %
-However, the new kernel
-is widened
-as it contains $n/g - 1$ instead of $b$ entries.
-If we truncate the new updating kernel
-at a radius of $b$ bins,
-the stability problem remains
-albeit, hopefully, becomes less severe
-(in terms of the most negative eigenvalue).
-%
-Thus, to obtain a stable updating kernel
-without increasing the radius $b$,
-we have to iterate the above process many times.
-The resulting algorithm is the following.
+However, the new kernel is widened
+as $\mu_l$ may be nonzero for $l > b$.
+%
+If we truncate the new kernel at $\mu_b$,
+the stability problem remains,
+although the negative eigenvalues
+often become smaller in magnitude.
+%
+Thus, to stabilize the updating kernel
+without increasing the width,
+we need to iterate the above process many times,
+resulting the following algorithm.
 
@@
   \item
-    Given a updating kernel $\mu_1, \dots, \mu_b$,
+    Given an updating kernel, $\mu_0, \dots, \mu_b$,
     compute the eigenvalues
-    $\lambda_0, \dots, \lambda_{n-1}$
+    $\lambda_1, \dots, \lambda_n$
     from Eq. \eqref{eq:wband_eigenvalue}.
@@
     Otherwise, replace negative eigenvalues by zeros,
-    and compute a new updating kernel,
-    $\mu_1, \dots, \mu_b$ from
+    and compute the new kernel,
+    $\mu_0, \dots, \mu_b$, from
     Eq. \eqref{eq:mu_from_lambda_pbc}
     or
-    \eqref{eq:mu_from_lambda_refl}.
-    Set $\mu_l = 0$ for $l > b$.
+    \eqref{eq:mu_from_lambda_refl}
+    with $\mu_l = 0$ for $l > b$.
     Go to step 1.
@@
 for a given updating scheme,
-can we also compare different updating schemes
-and pick the best one?
-%
-Below we address this question
-in the asymptotic, or long simulation time,
-limit.
+we shall turn to the problem of choosing
+optimal updating schemes below.
 
 \subsubsection{\label{sec:optWL}
-Asymptotic optimality of the single-bin (WL) scheme}
-
-We shall first show,
-directly from Eq. \eqref{eq:error_asym1},
+Asymptotic optimality of the single-bin scheme}
+
+We shall first show
 that the single-bin scheme is asymptotically optimal.
@@
 sharing the same set of eigenvectors,
-hence same $\Gamma_k$'s,
+hence the same $\Gamma_k$'s,
 but with different positive eigenvalues,
@@
 %
-This inequality follows from the non-negativity of
+\note{This inequality follows from the non-negativity of
 the following quadratic polynomial of $Y$,
@@
   A \, Y^2 + B \, Y + C
-%\ge 0,
+  ,
 $$
@@
 $B^2 - 4 \, A \, C$.
+}%
 %
@@
 only through the endpoint value, $q(T)$,
-which is fixed in the variation process.
+which is fixed in the variational process.
 %
 Thus, the inequality sets a lower bound
-of the error $\Err$
+for the asymptotic error $\Err$
 given in Eq. \eqref{eq:error_asym1}.
@@
 %
-Solving this set of equations yields
+Solving these equations yields
 %
@@
   \lambda_2 = \dots = \lambda_n.
-\end{equation}
-%
-In this case,
+  \notag
+\end{equation}
+%
+Then, by Eq. \eqref{eq:tk0},
 all $t_{k0}$'s also share the same value,
-$t_0$ by Eq. \eqref{eq:tk0},
+$t_0$,
 such that
@@
 and
-$c_k \propto 1/(T + t_0)$
+$c_k = 1/(T + t_0)$
 for $k \ge 2$.
@@
 %
-If we further assume that $\lambda_2 = 1$,
-the optimal schedule given by Eq. \eqref{eq:alpha_invtk}
+If we further assume $\lambda_2 = 1$,
+%the optimal schedule given by
+Eq. \eqref{eq:alpha_invtk}
 recovers Eq. \eqref{eq:alpha_invt1}.
@@
 
-Since $\lambda_1$ corresponds to the mode
-of a uniform shift of all $v_k$,
+Since the first eigenmode corresponds to
+a uniform shift of all $v_k$'s,
 we can freely set $\lambda_1$ to $\lambda_2$
@@
 %
-Now an updating matrix $\mathbf w$
+Now an updating matrix
 with identical eigenvalues
-is essentially a multiple of the identity matrix.
-%
-Thus, if no eigenvalue $\lambda_k$ is zero,
-then in terms of asymptotic convergence,
-the single-bin scheme adopted in the WL algorithm
-is most efficient.
+is a multiple of the identity matrix.
+%
+Thus,
+the single-bin scheme is most efficient
+in asymptotic convergence
+if no eigenvalue $\lambda_k$ is zero.
 
@@
 
-We can also generalize
+We can generalize
 the single-bin scheme to a class of
 perfect (brick-wall) \emph{bandpass} updating schemes
-that allow zero eigenvalues.
+to allow zero eigenvalues.
 %
@@
 if we can assume that the PMF is sufficiently smooth
-such that the error modes can be limited to $k \le K$.
+such that the eigenmodes can be limited to $k \le K$.
 %
@@
 The optimal schedule of these schemes
-remains to be the inverse-time schedule,
+remains to be the inverse-time one
 given by Eq. \eqref{eq:alpha_invt1}.
@@
 
-For a periodic variable, if we demand
+For a periodic variable, we demand
 %
@@
 $$
-for $K \ge 1$, then by using
+for $K \ge 1$.
+%
+Then by using
 Eq. \eqref{eq:mu_from_lambda_pbc},
@@
 Then Eq. \eqref{eq:lambdan} gives
-$\lambda_n = (-1)^{n+K-1}$,
+$\lambda_{n+1} = (-1)^{n+K-1}$,
 and from Eq. \eqref{eq:mu_from_lambda_refl},
@@
     {
-      n \, \sin \dfrac{ l \, \pi } { 2 \, n }
+      \sin \dfrac{ l \, \pi } { 2 \, n }
     }
@@
 \begin{aligned}
-  \lambda_l
+  \mu_l
   &=
@@
 
-In the above, we have been focusing on
+In the above, we have focused on
 the asymptotic error.
@@
 For a finite-length simulation,
-we should include the residual part,
-Eq. \eqref{eq:error_res}
+we should include the residual part
 and minimize the total error.
@@
 For the bandpass updating schemes,
-this allows to determine the optimal value
+this leads to the optimal value
 of the free parameter, $t_0$.
@@
 %
-The optimal error is reached at
+The minimum is reached at
 %
@@
 %
-This means that at the beginning of the schedule,
+At the beginning of the schedule,
 the updating magnitude is
@@
 %
-Interestingly, this factor of $1/2$
-happens to coincide with the suggested
+This factor $1/2$
+happens to coincide with the
 reduction factor of updating magnitude
-in the original WL algorithm\cite{wang2001, wang2001pre}
-at stage transitions.
+at stage transitions
+in the original WL algorithm\cite{
+wang2001, wang2001pre}.
 
@@
 
-As discussed in Sec. \ref{sec:FDS},
-the asymptotic limit of an adaptive FDS method
-is the corresponding equilibrium FDS method
-with the bias potential fixed to the exact value.
-%
-Such a simulation delivers the maximal efficiency.
-
-A practical question now arises.
-%
+%As discussed in Sec. \ref{sec:FDS},
+%the asymptotic limit of an adaptive FDS method
+%is the corresponding equilibrium FDS method
+%with the bias potential fixed to the exact value.
+%%
+%Such a simulation delivers the maximal efficiency.
+
 In late stages of an adaptive FDS simulation,
-if we are confident that the bias potential
-is sufficiently accurate,
-will we be better off switching to the equilibrium FDS method
-to prevent possible efficiency loss
-due to the adaptive updates?
-%
-Below, we shall show that this is unnecessary
-if the simulation uses the single-bin (WL) updating scheme
+the bias potential is often sufficiently precise.
+Some practitioners would then fix
+the bias potential and switch
+to the equilibrium method
+to prevent possible future efficiency loss
+due to the adaptive updates.
+%
+Below, we shall show that this practice is unnecessary
+if the simulation uses the single-bin updating scheme
 with the optimal schedule, Eq. \eqref{eq:alpha_invt1}.
@@
 That is, the bias potential obtained from
-an optimal single-bin scheme FDS simulation
+an optimal single-bin scheme adaptive simulation
 will be similar in precision to that from
-the ideal equilibrium FDS simulation
-in long times.
-%
-Thus, our demonstration below serves as an alternative proof
-of the optimality of the single-bin scheme.
+the ideal equilibrium simulation
+for long times.
+%
+%Thus, our demonstration below serves as an alternative proof
+%of the optimality of the single-bin scheme.
 
@@
 %
-The error is minimized
-when $v_i$'s assume the exact values,
-%
-and in this case,
-the error came from the second term
+If all $v_i$'s are exact,
+the error is minimized, and can be computed
+from the second term
 on the right-hand side of Eq. \eqref{eq:vcorr_equil},
@@
 
-To make better comparison with
-the error of the single-bin case,
-we shall rewrite the error as
-a sum over the eigenmodes.
-We define
+We can rewrite the error as
+a sum of the eigenmode as
+%
 \begin{align}
+  E
+  =
+  \sum_{ k = 2 }^n
+    \left\langle
+      \Theta_k^2
+    \right\rangle
+  ,
+\notag
+%\label{eq:error_sbin_equil}
+\end{align}
+%
+where
+%
+$
   \Theta_k
@@
   ,
-\notag
-\end{align}
-%
-with $\theta_k(t)$ defined by Eq. \eqref{eq:eta_def}
-and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
+$
+%
+with $\theta_k(t)$ defined by Eq. \eqref{eq:eta_def}.
+%and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
 %\begin{align}
@@
 %
-Then,
-%
-\begin{align}
-  E
-  =
-  \sum_{ k = 1 }^n
-    \left\langle
-      \Theta_k^2
-    \right\rangle.
-\notag
-%\label{eq:error_sbin_equil}
-\end{align}
-%
 \note{Because
@@
 \end{align*}
-}
-The sum can start from $k = 2$, for
-by Eq. \eqref{eq:eigenmode1}, we have
+}%
+The sum starts from $k = 2$, for
+by Eq. \eqref{eq:eigenmode1}, % we have
 $$
@@
 =
-\sum_{ i = 1 }^n h_i(t)
--
-\sum_{ i = 1 }^n p_i = 0.
+\sum_{ i = 1 }^n \left[ h_i(t) - p_i \right]
+= 0.
 $$
@@
   \frac{1}{T}
-  \int_{-\infty}^\infty
-  \kappa_n(\tau) \, d\tau
+  \sum_{\tau = -\infty}^\infty
+    \kappa_n(\tau)
   =
@@
 
-Now comparing this with the error of the single-bin scheme
-with the optimal schedule,
-Eq. \eqref{eq:error_singlebin},
-we find that the difference is negligible
+Now comparing this with the error of
+the optimal adaptive simulation,
+Eq. \eqref{eq:error_sinc} (with $K = n$),
+we find the difference to be negligible
 in the long time limit.
 
-Again, the above argument can be extended
+The above argument can be extended
 to the bandpass updating schemes.
 %
-If we are certain that
-the intrinsic PMF is sufficiently smooth,
-then we can explicitly filter some error modes
-in the histogram, and truncate the sum in
+If the intrinsic PMF is sufficiently smooth,
+we can explicitly filter out some error modes
+in the histogram,
+say those with $k >  K$,
+and truncate the sum in
 Eq. \eqref{eq:error_equil} at $k = K$.
@@
 
-\subsection{Test system and simulation protocol}
+We tested the above theory on a model system,
+which is to be described in Sec. \ref{sec:results_system}.
+%
+First, we verify the method of computing errors
+using an analytically simple schedule
+in Sec. \ref{sec:results_invt}.
+%
+We then study the optimal schedule from
+Eq. \eqref{eq:q_opt}
+in Sec. \ref{sec:results_optschedule}.
+%
+Finally, we compare the Gaussian and
+bandpass schemes of different widths
+in Sec. \ref{sec:results_cmpschemes}.
+
+\subsection{\label{sec:results_system}
+Test system and simulation protocol}
 
@@
 %
-Unless specified otherwise,
-we assume $i$ is a non-periodic variable
-under the reflective boundary condition.
-%
 %Since in the asymptotic regime,
@@
 %we also set the latter to be flat for convenience.
-For convenience,
-the intrinsic distribution, $p^*_i$,
-is a set to be flat as well.
-%
-Our goal is to observe how the fluctuating error
-of the bias potential defined in Eq. \eqref{eq:error_sum}
-depends on the schedule, $\alpha(t)$,
+The intrinsic distribution, $p^*_i$,
+is also assumed to be flat.
+%
+Unless specified otherwise,
+we assume $i$ to be non-periodic.
+%
+We wish to see how
+the schedule, $\alpha(t)$,
+affects the fluctuating error
+of the bias potential
 at the end of the simulation.
@@
 
-\subsection{Inverse-time schedule}
+\subsection{\label{sec:results_invt}
+Inverse-time schedule}
 
@@
 %
-where $\lambda$ is a parameter to be optimized
-with $t_0 = 2/(\lambda \, \alpha_0)$.
+where $\lambda$ is a free parameter,
+and $t_0 = 2/(\lambda \, \alpha_0)$.
 %
@@
 
-\subsubsection{Single-bin (WL) updating scheme}
+%\subsubsection{Single-bin (WL) updating scheme}
 
@@
 
-\subsubsection{Nearest-neighbor updating scheme}
+%\subsubsection{Nearest-neighbor updating scheme}
 
@@
 
-\subsection{Optimal schedule}
+\subsection{\label{sec:results_optschedule}
+Optimal schedule}
 
@@
 
-\subsection{Comparison of updating schemes}
+\subsection{\label{sec:results_cmpschemes}
+Comparison of updating schemes}
 
@@
 without a priori assumption
-of the smoothness of the PMF.
+on the smoothness of the PMF.
 %
@@
 
+Strictly speaking, our theory is valid
+only for long simulations.
+%
+This allows us to assume the sampling process
+to be a quasi-equilibrium one,
+and to represent the autocorrelation function
+of each eigenmode by a single number, $\Gamma_k$.
+%
+Although these assumptions appeared to work well
+on the model system,
+they may break down for relatively short simulations
+of glassy systems.
+
 \section{Acknowledgments}
@@
 optimal schedule from Eq. \eqref{eq:q_opt},
-%
-here we consider the analytically simpler
+here we consider the simpler
 inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
-with $\lambda$ optimized variationally.
-%
-This approach allows us to compute the error explicitly.
+%with a free parameter $\lambda$.
+%
+whose error can be computed explicitly.
 
@@
 For perfect sampling,
-we have, from Eq. \eqref{eq:Gamma_perfect},
-$\Gamma(p) = 1$, and
+$\Gamma(p) = 1$ from Eq. \eqref{eq:Gamma_perfect},
+and
 $$
@@
   \Err_A
-&=
-\frac{2 \, n}{\pi \, T}
-\int_0^{\pi/2}
-\frac{ \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2 }
-{ \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-\, dp
-\notag \\
+  %&=
+  %\frac{2 \, n}{\pi \, T}
+  %\int_0^{\pi/2}
+  %\frac{ \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2 }
+  %{ \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
+  %\, dp
+  %\notag \\
   &=
@@
     { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
-\right).
+  \right)
+.
 \end{aligned}
@@
   { 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
-\, \frac{dz}{z},
+  \, \frac{dz}{z}
+  ,
 \end{aligned}
@@
   \right)
-  .
+  ,
 \notag
@@
 %From Eq. \eqref{eq:error_nn}, it is clear that
-Clearly, the error
+which
 increases with the magnitude of $\mu_1$
@@
 %
-Note that we have ignored the residual error
+\note{Note that
+we have ignored the residual error
 and assumed $\lambda < 2 \, \min \lambda_k = 2 - 8 \, \mu_1$
@@
 $4 \, \mu_1 < 1$.
+}%
 \note{This is because,
@@
 $$
-}
+}%
 
@@
 Then
+$$
+  \Err_A
+  \propto
+  \frac{   2 \, n }
+       { \pi \, T }
+  \frac{             1             }
+       {  \lambda \, (2 - \lambda) }
+  .
+$$
+\note{
 $$
@@
   \right|_{ p \approx 0 }
-%\notag \\
-%&
+  \notag \\
+  &
   \approx
@@
 $$
+}%
 %

### 2016-04-16 doc/invt.tex
invt.tex: move a few sessions to Appendices

@@
 to minimize the total error, $E = E_R + E_A$.
+%
+A complete program of computing the optimal schedule
+is listed in Appendix \ref{sec:program}.
 
@@
 for longer near the beginning of the simulation.
-
-\subsubsection{\label{sec:measure_gamma}
-Measurement of the integrals
-of the autocorrelation functions}
-
-To use Eq. \eqref{eq:q_opt} in a real simulation,
-we need to know both $\lambda_k$'s and $\Gamma_k$'s.
-%
-While the updating scheme alone
-determines the eigenmodes and $\lambda_k$'s,
-the values of $\Gamma_k$'s depend also on the system
-and the underlying sampling process (MC or MD).
-%
-Below is a technique of measuring $\Gamma_k$'s.
-
-The idea is to run
-a simulation with a constant schedule,
-$\alpha(t) = \alpha_0$,
-and then use Eq. \eqref{eq:y2_eql}.
-%
-A practical difficulty is that
-we often do not know the intrinsic distribution, $p^*_i$,
-in advance to compute
-the shifted bias potential, $v_i$,
-from the original one, $V_i$
-(cf. Sec. \ref{sec:FDS}).
-%
-To circumvent the problem, we shall
-modify Eq. \eqref{eq:y2_eql} as
-%
-\begin{equation}
-  \operatorname{var} Y_k
-  =
-  \frac{1}{2}
-  \alpha \, \Gamma_k \, \lambda_k,
-\label{eq:varY}
-\end{equation}
-%
-where
-``$\operatorname{var}$''
-means the variance from time averaging,
-and
-%
-\begin{align}
-Y_k
-&=
-\sum_{ i = 1 }^n
-  \phi_{k i} \, (V_i - \bar V),
-\label{eq:Y_def}
-\\
-\bar V
-&=
-\sum_{ j = 1 }^n p_j \, V_j.
-\notag
-\end{align}
-
-To show Eq. \eqref{eq:varY},
-we first use Eq. \eqref{eq:v_def}
-to rewrite Eq. \eqref{eq:x_def}
-as
-$$
-\begin{aligned}
-  x_i
-  &=
-  (V_i - \bar V) + s_i,
-  \\
-  s_i
-  &=
-  \ln \frac{ p_i } { p^*_i }
-  -
-  \sum_{ j = 1 }^n
-    p_j \, \ln \frac{ p_j } { p^*_j }.
-\end{aligned}
-$$
-Thus, by Eqs. \eqref{eq:y_def} and \eqref{eq:Y_def},
-we have
-$$
-y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, s_i.
-$$
-Since the second term on the right-hand side
-is a constant, and the average of $y_k$ is zero,
-we have
-$$
-\operatorname{var} Y_k
-=
-\operatorname{var} y_k
-=
-\langle y_k^2 \rangle.
-$$
-Finally, by using Eq. \eqref{eq:y2_eql},
-we reach Eq. \eqref{eq:varY}.
-
-\note{In practice, Eq. \eqref{eq:varY}
-in inapplicable to zero eigenvalues, $\lambda_k$'s.
-%
-However, these eigenmodes can be ignored
-as they do not contribute to the Lagrangian defined by
-Eq. \eqref{eq:Lagrangian_const}.
-}
-
-\subsubsection{\label{sec:program}
-Program}
-
-We can now give a complete program of computing
-the optimal schedule for a given updating scheme.
-
-\begin{enumerate}
-
-\item
-From the updating matrix, $\mathbf w$,
-compute the eigenvalues, $\lambda_k$'s.
-
-\item
-Run a preliminary adaptive FDS simulation
-using the updating scheme with a constant
-updating magnitude, $\alpha_0$.
-
-\item
-Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}.
-
-\item
-Select an initial value of $q(T)$,
-e.g., $\ln(1 + \alpha_0 \, T)$.
-
-\item
-Compute the optimal schedule by
-numerically integrating Eq. \eqref{eq:q_opt}.
-
-\item
-Compute the error $E$ from
-Eqs. \eqref{eq:error_split},
-  \eqref{eq:error_asym1} and
-  \eqref{eq:error_res1}.
-
-\item
-Modify the value of $q(T)$ and repeat the previous two steps,
-until the minimal error $E$ is found.
-
-\end{enumerate}
-
-The dependence of variables are shown in Fig. \ref{fig:vardep}.
-
-\begin{figure}
-\begin{tikzpicture}
-  \node[draw, text width={width("Updating ")}]
-    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
-
-  \node[draw, text width={width("System and")}]
-    (sys) at (0, 1) {System and\\sampling \\process};
-
-  \node[draw, text width={width("Eigenvalues, ")}]
-    (lambda) at (2.5, 3) {Eigenvalues, \\$\lambda_2, \dots, \lambda_n$};
-
-  \node[draw, text width={width("Integrals, ")}]
-    (gamma) at (2.5, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
-
-  \node[draw, text width={width("Simulation ")}]
-    (T) at (2.5, -0.5) {Simulation \\time, $T$};
-
-  \node[draw, text width={width("Schedule ")}]
-    (alpha) at (5.2, 1) {Optimal schedule, \\ $\alpha(t)$};
-
-  \node[draw, text width={width("Error ")}]
-    (err) at (7, 1) {Error, \\ $\Err$};
-
-  \node[inner sep=0, minimum size=0] (M1) at (3.7, 1) {};
-  \node[inner sep=0, minimum size=0] (M2) at (4.0, 1) {};
-  \node[inner sep=0, minimum size=0] (N2) at (4.0, -0.2) {};
-  \node[inner sep=0, minimum size=0] (N3) at (6.2, -0.2) {};
-  \node[inner sep=0, minimum size=0] (M3) at (6.2, 1) {};
-
-  \draw[->, >=latex, thick]
-    (w) edge (lambda)
-    (w) edge (gamma)
-    (sys) edge (gamma)
-    (alpha) edge (err);
-
-  \draw[->, >=latex, thick]
-    (lambda) -| (M1)
-    (gamma) -- (M1)
-    (T) -| (M1)
-    (M1) -- (alpha);
-
-  \draw[->, >=latex, thick]
-    (M2) -- (N2) -- (N3) -- (M3);
-
-\end{tikzpicture}
-\caption{
-  \label{fig:vardep}
-  Dependence of variables in computing
-  the optimal schedule.
-}
-\end{figure}
 
@@
 %
-\note{If this is not so,
+If this is not so,
 we can modify the kernel
 using the technique described
-in Sec. \ref{sec:stabilize_wband}.}
+in Appendix \ref{sec:stabilize_wband}.
 
@@
 %
-\begin{equation}
+$
   1 = 2 \int_0^\pi \mu(x) \, dx.
-\notag
-%\label{eq:mx_normalization}
-\end{equation}
+$
 
@@
 
-\subsubsection{\label{sec:stabilize_wband}
-Inversion and Stabilization}
-
-%A practical problem of using homogeneous updating scheme
-%is the following.
-%
-A homogeneous updating scheme
-specified by an arbitrary kernel
-is not necessarily stable
-with all nonnegative eigenvalues.
-%
-%
-\note{For example,
-  %the nearest-neighbor updating scheme
-  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
-  %
-  %Even
-  the Gaussian updating scheme can be unstable
-  if the updating kernel is truncated.
-}%
-%
-Below we show how to minimally modify
-the updating kernel
-to stabilize it.
-%
-\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
-}
-
-The key is the observation of Eq. \eqref{eq:wband_eigenvalue}
-that the eigenvalues are related to the updating kernel
+\subsubsection{\label{sec:invert_wband}
+Inversion}
+
+Since the eigenvalues are related to the updating kernel
 by a cosine transform,
-and thus the relation can be readily inverted.
+Eq. \eqref{eq:wband_eigenvalue},
+the relation can be readily inverted.
 %
 In the periodic case,
-the inversion formula is
+we have
+%the inversion formula is
 %
@@
 
-Now given an updating kernel,
-we can compute the eigenvalues from
-Eq. \eqref{eq:wband_eigenvalue}.
-%
-By replacing negative eigenvalues with zeros and
-using Eq. \eqref{eq:mu_from_lambda_pbc}
-or \eqref{eq:mu_from_lambda_refl},
-we get a new updating kernel,
-which is stable by construction.
-%
-However, the new kernel is widened
-as $\mu_l$ may be nonzero for $l > b$.
-%
-If we truncate the new kernel at $\mu_b$,
-the stability problem remains,
-although the negative eigenvalues
-often become smaller in magnitude.
-%
-Thus, to stabilize the updating kernel
-without increasing the width,
-we need to iterate the above process many times,
-resulting the following algorithm.
-
-%
-\begin{enumerate}
-  \item
-    Given an updating kernel, $\mu_0, \dots, \mu_b$,
-    compute the eigenvalues
-    $\lambda_1, \dots, \lambda_n$
-    from Eq. \eqref{eq:wband_eigenvalue}.
-  \item
-    If all eigenvalues are nonnegative,
-    we are done, and exit the loop.
-  \item
-    Otherwise, replace negative eigenvalues by zeros,
-    and compute the new kernel,
-    $\mu_0, \dots, \mu_b$, from
-    Eq. \eqref{eq:mu_from_lambda_pbc}
-    or
-    \eqref{eq:mu_from_lambda_refl}
-    with $\mu_l = 0$ for $l > b$.
-    Go to step 1.
-\end{enumerate}
-
 \subsection{\label{sec:cmpschemes}
@@
 %
-At the beginning of the schedule,
+In Appendix \ref{sec:equilerr},
+we show that the above error tends
+to the error of the ideal equilibrium FDS simulation
+in the asymptotic limit.
+
+Note also that at the beginning of the schedule,
 the updating magnitude is
@@
 %has some $\lambda_k$'s close to $0$.
-
-\subsubsection{\label{sec:equilerr}
-Comparison to the error of the equilibrium FDS method}
-
-%As discussed in Sec. \ref{sec:FDS},
-%the asymptotic limit of an adaptive FDS method
-%is the corresponding equilibrium FDS method
-%with the bias potential fixed to the exact value.
-%%
-%Such a simulation delivers the maximal efficiency.
-
-In late stages of an adaptive FDS simulation,
-the bias potential is often sufficiently precise.
-Some practitioners would then fix
-the bias potential and switch
-to the equilibrium method
-to prevent possible future efficiency loss
-due to the adaptive updates.
-%
-Below, we shall show that this practice is unnecessary
-if the simulation uses the single-bin updating scheme
-with the optimal schedule, Eq. \eqref{eq:alpha_invt1}.
-%
-That is, the bias potential obtained from
-an optimal single-bin scheme adaptive simulation
-will be similar in precision to that from
-the ideal equilibrium simulation
-for long times.
-%
-%Thus, our demonstration below serves as an alternative proof
-%of the optimality of the single-bin scheme.
-
-The error of the equilibrium FDS method
-depends on the precision of the histogram.
-%
-If all $v_i$'s are exact,
-the error is minimized, and can be computed
-from the second term
-on the right-hand side of Eq. \eqref{eq:vcorr_equil},
-%
-\begin{align}
-  E
-  &=
-  \left\langle
-    \sum_{ i = 1 }^ n
-      p_i \,
-      \left(
-        \ln \frac { H_i }
-                  { p_i }
-      \right)^2
-  \right\rangle
-%\notag
-%\\
-  %&
-  \approx
-  \sum_{ i = 1 }^ n
-    p_i \,
-    \left\langle
-      \left(
-        \frac { H_i }
-              { p_i }
-        - 1
-      \right)^2
-    \right\rangle,
-\notag
-\end{align}
-where
-$H_i = \frac{1}{T} \sum_{t = 1}^T h_i(t)$
-with
-$h_i(t)$ defined by Eq. \eqref{eq:h_def}.
-
-We can rewrite the error as
-a sum of the eigenmode as
-%
-\begin{align}
-  E
-  =
-  \sum_{ k = 2 }^n
-    \left\langle
-      \Theta_k^2
-    \right\rangle
-  ,
-\notag
-%\label{eq:error_sbin_equil}
-\end{align}
-%
-where
-%
-$
-  \Theta_k
-  =
-  \frac{ 1 } { T }
-  \sum_{ t = 1 } ^ T
-    \theta_k( t )
-  ,
-$
-%
-with $\theta_k(t)$ defined by Eq. \eqref{eq:eta_def}.
-%and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
-%\begin{align}
-%\theta_k( t )
-%=
-%\sum_{ i = 1 }^ n
-%\phi_{k i} \left[
-%             \frac{ h_i(t) }
-%                  { p_i    }
-%             - 1
-%           \right],
-%\end{align}
-%and
-%\begin{align}
-%\sum_{ k = 1 }^n
-%  \phi_{k i} \, \phi_{k j}
-%  = p_i \, \delta_{i j}.
-%\end{align}
-%
-\note{Because
-%
-\begin{align*}
-\sum_{ k = 1 }^n
-  \Theta_k^2
-&=
-\frac{ 1 } { T^2 }
-\sum_{ t, t' = 1 }^T
-  \sum_{ k = 1 }^n
-    \theta_k( t ) \, \theta_k( t' )
-\\
-&=
-\frac{ 1 } { T^2 }
-\sum_{ t, t' = 1 }^T
-  \sum_{ i, j = 1 }^n
-    \left[
-      \frac{ h_i(t) }
-           { p_i    }
-      - 1
-    \right]
-    \left[
-      \frac{ h_j(t') }
-           { p_j     }
-      - 1
-    \right]
-    \sum_{ k = 1 }^n
-      \phi_{k \, i} \, \phi_{k \, j}
-\\
-&=
-\frac{ 1 } { T^2 }
-\sum_{ t, t' = 1 }^T
-  \sum_{ i, j = 1 }^n
-    \left[
-      \frac{ h_i(t) }
-           { p_i    }
-      - 1
-    \right]
-    \left[
-      \frac{ h_j(t') }
-           { p_j     }
-      - 1
-    \right]
-    p_i \, \delta_{i, j}
-\\
-&=
-\sum_{ i = 1 }^n
-\frac{ 1 } { T }
-\sum_{ t = 1 }^T
-    \left[
-      \frac{ h_i(t) }
-           { p_i    }
-      - 1
-    \right]
-\frac{ 1 } { T }
-\sum_{ t' = 1 }^T
-    \left[
-      \frac{ h_i(t') }
-           { p_i     }
-      - 1
-    \right]
-\\
-&=
-\sum_{ i = 1 }^n
-  p_i \,
-    \left(
-      \frac{ \bar h_i }
-           { p_i      }
-      - 1
-    \right)^2
-.
-\end{align*}
-}%
-The sum starts from $k = 2$, for
-by Eq. \eqref{eq:eigenmode1}, % we have
-$$
-\theta_1(t)
-=
-\sum_{ i = 1 }^n
-  \phi_{1i} \, \frac{ h_i(t) - p_i } { p_i }
-=
-\sum_{ i = 1 }^n \left[ h_i(t) - p_i \right]
-= 0.
-$$
-
-In the long-time limit,
-\begin{align*}
-  \left\langle
-    \Theta_k^2
-  \right\rangle
-  &=
-  \frac{1}{T^2}
-  \sum_{ t, t' = 1 }^T
-  \langle \theta_k(t) \, \theta_k(t') \rangle
-  %\\
-  %&\approx
-  %\frac{1}{T}
-  %\sum_{ \tau = -\infty }^{ \infty }
-  %\langle \theta_k(t) \, \theta_k(t + \tau) \rangle
-  %\\
-  %&=
-  =
-  \frac{1}{T}
-  \sum_{\tau = -\infty}^\infty
-    \kappa_n(\tau)
-  =
-  \frac{ \Gamma_k }{ T }
-.
-\end{align*}
-%
-Thus,
-\begin{equation}
-  E
-  =
-  \frac{ 1 } { T }
-  \sum_{ k = 2 }^n \Gamma_k
-  .
-\label{eq:error_equil}
-\end{equation}
-
-Now comparing this with the error of
-the optimal adaptive simulation,
-Eq. \eqref{eq:error_sinc} (with $K = n$),
-we find the difference to be negligible
-in the long time limit.
-
-The above argument can be extended
-to the bandpass updating schemes.
-%
-If the intrinsic PMF is sufficiently smooth,
-we can explicitly filter out some error modes
-in the histogram,
-say those with $k >  K$,
-and truncate the sum in
-Eq. \eqref{eq:error_equil} at $k = K$.
 
@@
 according to Eq. \eqref{eq:t0_sinc},
-to take into account the initial error.
+to accommodate the initial error.
 %such that the initial updating magnitude
@@
 \appendix
+
+\section{\label{sec:program}
+Program of computing the optimal schedule and error
+}
+
+Here we give a program of computing
+the optimal schedule and the error
+of a given adaptive FDS simulation.
+%
+Basically, we extract
+both $\lambda_i$'s and $\Gamma_i$'s
+from the input,
+and then feed them to Eq. \eqref{eq:q_opt}
+for the schedule.
+%
+The dependence of variables are shown in Fig. \ref{fig:vardep}.
+
+\begin{figure}[h]
+\begin{tikzpicture}
+  \node[draw, text width={width("Updating ")}]
+    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
+
+  \node[draw, text width={width("System and")}]
+    (sys) at (0, 1) {System and\\sampling \\process};
+
+  \node[draw, text width={width("Eigenvalues, ")}]
+    (lambda) at (2.5, 3) {Eigenvalues, \\$\lambda_2, \dots, \lambda_n$};
+
+  \node[draw, text width={width("Integrals, ")}]
+    (gamma) at (2.5, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
+
+  \node[draw, text width={width("Simulation ")}]
+    (T) at (2.5, -0.5) {Simulation \\time, $T$};
+
+  \node[draw, text width={width("Schedule ")}]
+    (alpha) at (5.2, 1) {Optimal schedule, \\ $\alpha(t)$};
+
+  \node[draw, text width={width("Error ")}]
+    (err) at (7, 1) {Error, \\ $\Err$};
+
+  \node[inner sep=0, minimum size=0] (M1) at (3.7, 1) {};
+  \node[inner sep=0, minimum size=0] (M2) at (4.0, 1) {};
+  \node[inner sep=0, minimum size=0] (N2) at (4.0, -0.2) {};
+  \node[inner sep=0, minimum size=0] (N3) at (6.2, -0.2) {};
+  \node[inner sep=0, minimum size=0] (M3) at (6.2, 1) {};
+
+  \draw[->, >=latex, thick]
+    (w) edge (lambda)
+    (w) edge (gamma)
+    (sys) edge (gamma)
+    (alpha) edge (err);
+
+  \draw[->, >=latex, thick]
+    (lambda) -| (M1)
+    (gamma) -- (M1)
+    (T) -| (M1)
+    (M1) -- (alpha);
+
+  \draw[->, >=latex, thick]
+    (M2) -- (N2) -- (N3) -- (M3);
+
+\end{tikzpicture}
+\caption{
+  \label{fig:vardep}
+  Dependence of variables in computing
+  the optimal schedule and the error.
+}
+\end{figure}
+
+%\subsection{\label{sec:measure_gamma}
+%Measurement of the integrals
+%of the autocorrelation functions}
+
+%To use Eq. \eqref{eq:q_opt} in a real simulation,
+%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
+%
+While the updating scheme alone
+determines the eigenmodes and $\lambda_k$'s,
+the values of $\Gamma_k$'s depend also on the system
+and the underlying sampling process (MC or MD).
+%
+Below is a technique of measuring $\Gamma_k$'s.
+
+The idea is to run
+a simulation with a constant schedule,
+$\alpha(t) = \alpha_0$,
+and then use Eq. \eqref{eq:y2_eql}.
+%
+A practical difficulty is that
+we often do not know the intrinsic distribution, $p^*_i$,
+in advance to compute
+the shifted bias potential, $v_i$,
+from the original one, $V_i$
+(cf. Sec. \ref{sec:FDS}).
+%
+To circumvent the problem, we shall
+modify Eq. \eqref{eq:y2_eql} as
+%
+\begin{equation}
+  \operatorname{var} Y_k
+  =
+  \frac{1}{2}
+  \alpha \, \Gamma_k \, \lambda_k,
+\label{eq:varY}
+\end{equation}
+%
+where
+``$\operatorname{var}$''
+means the variance from time averaging,
+and
+%
+\begin{align}
+Y_k
+&=
+\sum_{ i = 1 }^n
+  \phi_{k i} \, (V_i - \bar V),
+\label{eq:Y_def}
+\\
+\bar V
+&=
+\sum_{ j = 1 }^n p_j \, V_j.
+\notag
+\end{align}
+
+To show Eq. \eqref{eq:varY},
+we first use Eq. \eqref{eq:v_def}
+to rewrite Eq. \eqref{eq:x_def}
+as
+$$
+\begin{aligned}
+  x_i
+  &=
+  (V_i - \bar V) + s_i,
+  \\
+  s_i
+  &=
+  \ln \frac{ p_i } { p^*_i }
+  -
+  \sum_{ j = 1 }^n
+    p_j \, \ln \frac{ p_j } { p^*_j }.
+\end{aligned}
+$$
+Thus, by Eqs. \eqref{eq:y_def} and \eqref{eq:Y_def},
+we have
+$$
+y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, s_i.
+$$
+Since the second term on the right-hand side
+is a constant, and the average of $y_k$ is zero,
+we have
+$$
+\operatorname{var} Y_k
+=
+\operatorname{var} y_k
+=
+\langle y_k^2 \rangle.
+$$
+Finally, by using Eq. \eqref{eq:y2_eql},
+we reach Eq. \eqref{eq:varY}.
+
+\note{In practice, Eq. \eqref{eq:varY}
+in inapplicable to zero eigenvalues, $\lambda_k$'s.
+%
+However, these eigenmodes can be ignored
+as they do not contribute to the Lagrangian defined by
+Eq. \eqref{eq:Lagrangian_const}.
+}
+
+%\subsection{\label{sec:program_list}
+%Program
+%}
+
+The complete program is the following.
+
+\begin{enumerate}
+
+\item
+From the updating matrix, $\mathbf w$,
+compute the eigenvalues, $\lambda_k$'s.
+
+\item
+Run a preliminary adaptive FDS simulation
+using the updating scheme with a constant
+updating magnitude, $\alpha_0$.
+
+\item
+Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}.
+
+\item
+Select an initial value of $q(T)$,
+e.g., $\ln(1 + \alpha_0 \, T)$.
+
+\item
+Compute the optimal schedule by
+numerically integrating Eq. \eqref{eq:q_opt}.
+
+\item
+Compute the error $E$ from
+Eqs. \eqref{eq:error_split},
+  \eqref{eq:error_asym1} and
+  \eqref{eq:error_res1}.
+
+\item
+Modify the value of $q(T)$ and repeat the previous two steps,
+until the minimal error $E$ is found.
+
+\end{enumerate}
 
@@
   \right|_{ p \approx 0 }
-  \notag \\
-  &
+  %\notag \\
+  %&
   \approx
@@
 %
-The optimal $\lambda$ is therefore
+The optimal $\lambda$ is, therefore,
 %
@@
 
+\section{\label{sec:stabilize_wband}
+Stabilization of homogeneous updating schemes}
+
+%A practical problem of using homogeneous updating scheme
+%is the following.
+%
+A homogeneous updating scheme
+specified by an arbitrary kernel
+is not necessarily stable
+with all nonnegative eigenvalues.
+%
+%
+\note{For example,
+  %the nearest-neighbor updating scheme
+  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
+  %
+  %Even
+  the Gaussian updating scheme can be unstable
+  if the updating kernel is truncated.
+}%
+%
+Here we show how to minimally modify
+the updating kernel
+to stabilize it.
+%
+\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
+}
+
+Given an updating kernel,
+we can compute the eigenvalues from
+Eq. \eqref{eq:wband_eigenvalue}.
+%
+By replacing negative eigenvalues with zeros and
+using Eq. \eqref{eq:mu_from_lambda_pbc}
+or \eqref{eq:mu_from_lambda_refl},
+we get a new updating kernel,
+which is stable by construction.
+%
+However, the new kernel is widened
+as $\mu_l$ may be nonzero for $l > b$.
+%
+If we truncate the new kernel at $\mu_b$,
+the stability problem remains,
+although the negative eigenvalues
+often become smaller in magnitude.
+%
+Thus, to stabilize the updating kernel
+without increasing the width,
+we need to iterate the above process many times,
+resulting the following algorithm.
+
+%
+\begin{enumerate}
+  \item
+    Given an updating kernel, $\mu_0, \dots, \mu_b$,
+    compute the eigenvalues
+    $\lambda_1, \dots, \lambda_n$
+    from Eq. \eqref{eq:wband_eigenvalue}.
+  \item
+    If all eigenvalues are nonnegative,
+    we are done, and exit the loop.
+  \item
+    Otherwise, replace negative eigenvalues by zeros,
+    and compute the new kernel,
+    $\mu_0, \dots, \mu_b$, from
+    Eq. \eqref{eq:mu_from_lambda_pbc}
+    or
+    \eqref{eq:mu_from_lambda_refl}
+    with $\mu_l = 0$ for $l > b$.
+    Go to step 1.
+\end{enumerate}
+
+\section{\label{sec:equilerr}
+Comparison to the error of the equilibrium FDS simulation}
+
+%As discussed in Sec. \ref{sec:FDS},
+%the asymptotic limit of an adaptive FDS method
+%is the corresponding equilibrium FDS method
+%with the bias potential fixed to the exact value.
+%%
+%Such a simulation delivers the maximal efficiency.
+
+In late stages of an adaptive FDS simulation,
+the bias potential is often sufficiently precise.
+Some practitioners would then fix
+the bias potential and switch
+to the equilibrium method
+to prevent possible future efficiency loss
+due to the adaptive updates.
+%
+Below, we shall show that this practice is unnecessary
+if the simulation uses the single-bin updating scheme
+with the optimal schedule, Eq. \eqref{eq:alpha_invt1}.
+%
+That is, the bias potential obtained from
+an optimal single-bin scheme adaptive simulation
+will be similar in precision to that from
+the ideal equilibrium simulation
+for long times.
+%
+%Thus, our demonstration below serves as an alternative proof
+%of the optimality of the single-bin scheme.
+
+The error of the equilibrium FDS method
+depends on the precision of the histogram.
+%
+If all $v_i$'s are exact,
+the error is minimized, and can be computed
+from the second term
+on the right-hand side of Eq. \eqref{eq:vcorr_equil},
+%
+\begin{align}
+  E
+  &=
+  \left\langle
+    \sum_{ i = 1 }^ n
+      p_i \,
+      \left(
+        \ln \frac { H_i }
+                  { p_i }
+      \right)^2
+  \right\rangle
+%\notag
+%\\
+  %&
+  \approx
+  \sum_{ i = 1 }^ n
+    p_i \,
+    \left\langle
+      \left(
+        \frac { H_i }
+              { p_i }
+        - 1
+      \right)^2
+    \right\rangle,
+\notag
+\end{align}
+where
+$H_i = \frac{1}{T} \sum_{t = 1}^T h_i(t)$
+with
+$h_i(t)$ defined by Eq. \eqref{eq:h_def}.
+
+We can rewrite the error as
+a sum of the eigenmode as
+%
+\begin{align}
+  E
+  =
+  \sum_{ k = 2 }^n
+    \left\langle
+      \Theta_k^2
+    \right\rangle
+  ,
+\notag
+%\label{eq:error_sbin_equil}
+\end{align}
+%
+where
+%
+$
+  \Theta_k
+  =
+  \frac{ 1 } { T }
+  \sum_{ t = 1 } ^ T
+    \theta_k( t )
+  ,
+$
+%
+with $\theta_k(t)$ defined by Eq. \eqref{eq:eta_def}.
+%and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
+%\begin{align}
+%\theta_k( t )
+%=
+%\sum_{ i = 1 }^ n
+%\phi_{k i} \left[
+%             \frac{ h_i(t) }
+%                  { p_i    }
+%             - 1
+%           \right],
+%\end{align}
+%and
+%\begin{align}
+%\sum_{ k = 1 }^n
+%  \phi_{k i} \, \phi_{k j}
+%  = p_i \, \delta_{i j}.
+%\end{align}
+%
+\note{Because
+%
+\begin{align*}
+\sum_{ k = 1 }^n
+  \Theta_k^2
+&=
+\frac{ 1 } { T^2 }
+\sum_{ t, t' = 1 }^T
+  \sum_{ k = 1 }^n
+    \theta_k( t ) \, \theta_k( t' )
+\\
+&=
+\frac{ 1 } { T^2 }
+\sum_{ t, t' = 1 }^T
+  \sum_{ i, j = 1 }^n
+    \left[
+      \frac{ h_i(t) }
+           { p_i    }
+      - 1
+    \right]
+    \left[
+      \frac{ h_j(t') }
+           { p_j     }
+      - 1
+    \right]
+    \sum_{ k = 1 }^n
+      \phi_{k \, i} \, \phi_{k \, j}
+\\
+&=
+\frac{ 1 } { T^2 }
+\sum_{ t, t' = 1 }^T
+  \sum_{ i, j = 1 }^n
+    \left[
+      \frac{ h_i(t) }
+           { p_i    }
+      - 1
+    \right]
+    \left[
+      \frac{ h_j(t') }
+           { p_j     }
+      - 1
+    \right]
+    p_i \, \delta_{i, j}
+\\
+&=
+\sum_{ i = 1 }^n
+\frac{ 1 } { T }
+\sum_{ t = 1 }^T
+    \left[
+      \frac{ h_i(t) }
+           { p_i    }
+      - 1
+    \right]
+\frac{ 1 } { T }
+\sum_{ t' = 1 }^T
+    \left[
+      \frac{ h_i(t') }
+           { p_i     }
+      - 1
+    \right]
+\\
+&=
+\sum_{ i = 1 }^n
+  p_i \,
+    \left(
+      \frac{ \bar h_i }
+           { p_i      }
+      - 1
+    \right)^2
+.
+\end{align*}
+}%
+The sum starts from $k = 2$, for
+by Eq. \eqref{eq:eigenmode1}, % we have
+$$
+\theta_1(t)
+=
+\sum_{ i = 1 }^n
+  \phi_{1i} \, \frac{ h_i(t) - p_i } { p_i }
+=
+\sum_{ i = 1 }^n \left[ h_i(t) - p_i \right]
+= 0.
+$$
+
+In the long-time limit,
+\begin{align*}
+  \left\langle
+    \Theta_k^2
+  \right\rangle
+  &=
+  \frac{1}{T^2}
+  \sum_{ t, t' = 1 }^T
+  \langle \theta_k(t) \, \theta_k(t') \rangle
+  %\\
+  %&\approx
+  %\frac{1}{T}
+  %\sum_{ \tau = -\infty }^{ \infty }
+  %\langle \theta_k(t) \, \theta_k(t + \tau) \rangle
+  %\\
+  %&=
+  =
+  \frac{1}{T}
+  \sum_{\tau = -\infty}^\infty
+    \kappa_n(\tau)
+  =
+  \frac{ \Gamma_k }{ T }
+.
+\end{align*}
+%
+Thus,
+\begin{equation}
+  E
+  =
+  \frac{ 1 } { T }
+  \sum_{ k = 2 }^n \Gamma_k
+  .
+\label{eq:error_equil}
+\end{equation}
+
+Now comparing this with the error of
+the optimal adaptive simulation,
+Eq. \eqref{eq:error_sinc} (with $K = n$),
+we find the difference to be negligible
+in the long time limit.
+
+The above argument can be extended
+to the bandpass updating schemes.
+%
+If the intrinsic PMF is sufficiently smooth,
+we can explicitly filter out some error modes
+in the histogram,
+say those with $k >  K$,
+and truncate the sum in
+Eq. \eqref{eq:error_equil} at $k = K$.
+
 \bibliography{simul}

### 2016-04-18 doc/invt.tex
separate invt_notes.tex

@@
 \end{align}
-%
-\note{We can define a symmetric matrix $\hat{\mathbf w}$
-as $\hat w_{ij} = \sqrt{p_i/p_j} \, w_{ij}$,
-which can be diagonalized\cite{vankampen}
-with a set of orthonormal eigenvectors:
-%
-$\sum_{i = 1}^n \varphi_{ki} \, \hat w_{ij} = \lambda_k \, \varphi_{kj}$.
-%
-Then,
-in terms of diagonalizing $\mathbf w$, we have
-$$\phi_{ij} \equiv \sqrt{p_j} \, \varphi_{ij}$$
-satisfying Eqs. \eqref{eq:eig_orthonormal_cols}
-and \eqref{eq:eig_orthonormal_rows}.
-%
-}%
 
@@
 \end{align*}
-%
-\note{The extremal condition of $\Err_A$ is given by
-%
-\begin{align}
-  \sum_{k = 2}^n
-  \lambda_k \, u_k\bigl( q(t) \bigr)
-  \int_0^T
-    \ddot u_k \bigl( q(t') \bigr) \,
-    \kappa_k(t' - t) \, dt' = 0.
-  \tag{N1}
-\label{eq:optimal_mbin}
-\end{align}
-%
-If all $\lambda_k$, hence all $u_k$, are the same,
-the above condition can be satisfied with $\ddot u_k = 0$,
-which recovers the single-bin case.
-%
-Unfortunately, the above expression
-appears to be useless in general.
-}
 
@@
 %
-\note{The next order correction to Eq. \eqref{eq:kappa_delta}
-would lead to a correction to $\Err$
-that is about $\alpha_{\max}^2$ times as large,
-where $\alpha_{\max} = \max \alpha(t)$.
-%
-This is equivalent to saying that the Fourier transform
-  $$
-  \tilde \kappa_k(\omega) = \Gamma_k
-  $$
-  is a roughly constant.
-  %
-  For the next order correction we would have
-  $$
-  \tilde \kappa_k(\omega) = \Gamma_k + \Gamma^{(2)}_k \omega^2 + \dots,
-  $$
-  which leads to a correction to $\Err$
-  $$
-  \Gamma^{(2)}_k
-  \int_0^T \ddot u_k^2\bigl( q(t) \bigr) \, dt
-  \propto
-  \alpha^2 \, \Err.
-  $$
-  This correction is of order $\alpha^2$
-  compared to the main contribution in Eq. \eqref{eq:error_asym1}.
-  %
-  Thus, we can assume Eq. \eqref{eq:kappa_delta},
-  once $\alpha(t)$ has dropped below a certain level.
-}%
-\note{
-  The extremal property, Eq. \eqref{eq:optimal_mbin},
-  now becomes
-  %
-  \begin{align}
-    \sum_{k = 1}^n
-      \Gamma_k \, \lambda_k \,
-      u_k\bigl( q(t) \bigr) \,
-      \ddot u_k\bigl( q(t) \bigr) = 0
-    .
-    \tag{N2}
-    \label{neq:optimal_mbin1}
-  \end{align}
-  %
-  If we interpret $u_k$ as the position,
-  and $\Gamma_k \, \lambda_k$ as the mass,
-  then the right-hand side gives the virial.
-  %
-  Thus, the extremal condition demands
-  the total virial to be zero.
-  $$\,$$
-}%
 In this way,
@@
 for the $\langle y_k^2(0) \rangle$'s in Eq. \eqref{eq:error_res}.
-%
-\note{Another possible approximation
-  is to interpret $\Gamma_k$ as twice the autocorrelation time,
-  and the kernel is exponential
-  $$
-  \kappa_k(t) = \exp\left( - \frac{2}{\Gamma_k} |t| \right),
-  $$
-  then a slightly more accurate value is given by\cite{vankampen}
-  $$
-  \left\langle
-    y_k^2
-  \right\rangle
-  =
-  \frac{      ( \lambda_k \, \alpha_0 )^2     }
-       { \lambda_k \, \alpha_0 + 2 / \Gamma_k }.
-  $$
-  But it appears that the main error of Eq. \eqref{eq:y2_eql}
-  is an underestimate instead of an overestimate.
-  And a possible source of error might be from
-  converting a discrete sum to a continuous integral.
-}
 %
@@
   \notag
-  %\label{eq:q_opt}
 \end{equation}
@@
 is listed in Appendix \ref{sec:program}.
-
-Unfortunately,
-Eq. \eqref{eq:q_opt} is not an explicit equation of $\alpha(t)$.
-%
-However,
-by differentiating Eq. \eqref{eq:Lagrangian_const},
-we have
-%
-\begin{equation}
-  \frac{ d   }
-       { d t }
-  \left[
-    \frac{       1     }
-         { \alpha( t ) }
-  \right]
-  =
-  \frac{
-    \sum_{ k = 2 }^n
-      \Gamma_k \, \lambda_k^3
-      \, u_k^2 \bigl[ q(t) \bigr]
-  }
-  {
-    \sum_{ k = 2 }^n
-      \Gamma_k \, \lambda_k^2
-      \, u_k^2 \bigl[ q(t) \bigr]
-  }
-  \equiv
-  \lambda(t)
-  .
-  \label{eq:dinvadt}
-\end{equation}
-%
-The expression in the middle is an average
-of the eigenvalues, $\lambda_k$'s,
-weighted by $\Gamma_k \, \lambda_k^2 \, u_k^2$.
-%
-We shall refer to the average as the instantaneous
-eigenvalue, $\lambda(t)$,
-which gives the rate of change of $\alpha^{-1}(t)$.
-
-\note{For the single-bin scheme, $\lambda_k = 1$,
-and $\lambda(t)$ is always unity.
-%
-So,
-$$
-\frac { d   }
-      { d t }
-\left[
-  \frac{      1    }
-       { \alpha(t) }
-\right]
-=
-1,
-$$
-which again leads to Eq. \eqref{eq:alpha_invt1}.
-
-For a general multiple-bin scheme,
-we can characterize the solution of
-Eq. \eqref{eq:dinvadt} as
-\begin{equation}
-  \alpha(t)
-  =
-  \frac{                  1                     }
-       { \lambda(t) \, \bigl[ t + t_0(t) \bigr] }
-  ,
-\notag
-%\label{eq:alpha_approxinvt}
-\end{equation}
-%
-where
-$t_0(t)$ is a slowly-varying function.
-}
-
-By further differentiating Eq. \eqref{eq:dinvadt},
-we can show that $\lambda(t)$
-is an increasing function.
-%
-\note{Proof that $\lambda(t)$ is an increasing function.
-  Using $\dot u_k = \lambda_k \, u_k \, \alpha(t)$,
-  we have
-  $$
-  \begin{aligned}
-    \dot \lambda
-    &=
-    \frac
-    {
-      \sum_{k = 2}^n 2 \, \Gamma_k \, \lambda_k^3 \, u_k \, \dot u_k
-    }
-    {
-      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
-    }
-    -
-    \frac
-    {
-      \sum_{k = 2}^n \Gamma_k \, \lambda_k^3 \, u_k^2
-      \sum_{k = 2}^n 2 \, \Gamma_k \, \lambda_k^2 \, u_k \, \dot u_k
-    }
-    {
-      \left(
-        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
-      \right)^2
-    }
-    \\
-    &=
-    \frac
-    {
-      \sum_{k = 2}^n \Gamma_k \, \lambda_k^4 \, u_k^2
-      \,
-      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
-      -
-      \left(
-        \sum_{k = 2}^n \Gamma_k \, \lambda_k^3 \, u_k^2
-      \right)^2
-    }
-    {
-      \left(
-        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
-      \right)^2
-    }
-    2 \alpha(t).
-  \end{aligned}
-  $$
-  The numerator of the fraction of the last expression
-  is nonnegative by the Cauchy-Schwarz inequality,
-  or by the nonpositivity of the discriminant
-  of the following nonnegative definite quadratic form of $X$:
-  $$
-  \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2 \, (X - \lambda_k)^2 \ge 0.
-  $$
-  Thus, $\dot \lambda \ge 0$, and $\lambda(t)$ is monotonically increasing.
-}%
-%
-Below we shall study the initial and final values.
-%
-At the end of the simulation,
-$t = T$, we have $u_k = 1$, and
-%
-\begin{equation}
-  \lambda(T)
-  =
-  \frac{
-    \sum_{ k = 2 }^n
-      \Gamma_k \, \lambda_k^3
-  }
-  {
-    \sum_{ k = 2 }^n
-      \Gamma_k \, \lambda_k^2
-  }
-  ,
-\notag
-%\label{eq:dinvadt_limit2}
-\end{equation}
-is independent of the simulation length, $T$.
-%
-Further, $\Gamma_k$ often increases with $\lambda_k$,
-such that the relative weight $\Gamma_k \, \lambda_k^2$
-also increases with $\lambda_k$.
-%
-It follows that $\lambda(T)$ will be dominated
-by the larger values of $\lambda_k$'s.
-
-By contrast,
-at the beginning of the simulation,
-the relative weight, $\Gamma_k \, \lambda_k^2 \, u_k^2(0)$,
-is heavily influenced by the exponential factor
-$u_k^2(0) = e^{ -2 \, \lambda_k \, q(T) }$.
-%
-Thus, $\lambda(0)$
-is sensitive to the simulation length, $T$.
-%
-A longer simulation means a smaller $\lambda(0)$,
-and that $\lambda(t)$ would stay small
-for longer near the beginning of the simulation.
 
@@
   $$
-  which satisfies the reflective boundary condition,
-  Eq. \eqref{eq:phi_refl}.
+  which satisfies Eq. \eqref{eq:phi_refl}.
   %
@@
   with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue_refl}.
-
-  \hrulefill
-
-  Next, let us compute the normalization factor.
-  %
-  If $k = 1$, $\Phi^{(1)}_i = 1$, and
-  $
-  \sum_{i = 1}^n \left( \Phi^{(1)}_i \right)^2 = n.
-  $
-  For $k \ge 2$,
-  $$
-  \begin{aligned}
-    \sum_{i = 1}^n \cos^2 \left[\left(i - \frac1 2 \right) a\right]
-    &=
-    \frac n 2
-    +
-    \frac 1 2
-    \sum_{i = 1}^n \cos\left[(2 i - 1)\, a \right]
-    =
-    \frac n 2,
-  \end{aligned}
-  $$
-  where $a = \frac{k-1}{n} \pi$.
-  Thus, the normalization factor $\sqrt{(2 - \delta_{k1})/n}$
-  encompasses both cases.
-
-  \hrulefill
-
-  To show the column-wise orthogonality,
-  we need a lemma, for $a = q \, \pi/n$,
-  with $q$ a positive integer.
-  $$
-  \begin{aligned}
-  \sum_{k = 1}^n \cos[(k - 1) \, a]
-  &=
-  1 + \cos a + \dots + \cos[(n - 1) \, a]
-  \\
-  &=
-  \frac{
-        \sin\frac a 2
-      + \sin \left[ \left( n - \frac 1 2 \right) a \right]
-      }
-      {
-        2 \, \sin \frac a 2
-      }
-  \\
-  &=
-  \frac{ 1 - (-1)^q } { 2 }
-  = \operatorname{odd}(q),
-  \end{aligned}
-  $$
-  %
-  where, we have used
-  $\sin \left[ \left( n - \frac 1 2 \right) a \right]
-  = \sin \left( q \, \pi - \frac a 2 \right)
-  = -(-)^q\sin\frac a 2.$
-  %
-  and $\operatorname{odd}(q)$
-  yields $1$ for odd $q$ or $0$ otherwise.
-  %
-  Adding the case of $q = 0$, where the sum yields $n$,
-  we get
-  $$
-  \sum_{k = 1}^n
-    \cos\left[(k - 1) \, \frac { q \, \pi } { n }  \right]
-  = n \, \delta_{q, 0}
-  + \operatorname{odd}(q).
-  $$
-
-  Now, for the orthogonality,
-  $$
-  \begin{aligned}
-    \sum_{k = 1}^n
-    \Phi^{(k)}_i \, \Phi^{(k)}_j
-    &=
-    \frac 1 2
-    \sum_{k = 1}^n
-    \left[
-      \cos \tfrac{ (i - j) (k - 1) \, \pi }
-                 {         n              }
-      +
-      \cos \tfrac{ (i + j - 1) (k - 1) \, \pi }
-                 {             n              }
-    \right]
-    \\
-    &=
-    \frac 1 2
-    \left[
-      n \, \delta_{i, j}
-      +
-      \operatorname{odd}(i - j)
-      +
-      \operatorname{odd}(i + j - 1)
-    \right]
-    \\
-    &=
-    \frac n 2 \, \delta_{i, j}
-    + \frac 1 2.
-  \end{aligned}
-  $$
-  where we have used the fact
-  that $i - j$ shares the parity with $i + j$
-  in the last step.
-  %
-  So
-  $$
-    \sum_{k = 1}^n
-    \phi^{(k)}_i \, \phi^{(k)}_j
-    =
-    \frac 2 { n^2 }
-    \sum_{k = 1}^n
-    \Phi^{(k)}_i \, \Phi^{(k)}_j
-    -
-    \frac 1 { n^2 }
-    =
-    \frac 1 n \,
-    \delta_{i, j}.
-  $$
-}
-%
+}%
 
@@
   %
-  \begin{align}
-    \mu_{ 2 n - l } = \mu_l
-    \quad
-    \mathrm{for\;} l = 1, \dots, n - 1,
-  \notag
-  %\label{eq:mu_reflection}
-  \end{align}
+  $
+    \mu_{ 2 n - l } = \mu_l,
+  $
+  for
+  $l = 1, \dots, n - 1$,
   %
@@
   %
-  \begin{align}
-    \lambda_{ 2 \, n + 2 - k } = \lambda_k.
-    \tag{N3}
-    \label{neq:lambda_reflection}
-  \end{align}
+  $
+    \lambda_{ 2 \, n + 2 - k } = \lambda_k
+    .
+  $
   %
@@
   $$
-  This entails
-  $$
-  \begin{aligned}
-    \mu_l
-    &=
-    \frac{    1   }
-         { 2 \, n }
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \lambda_{ k + 1 } \,
-      \cos \frac{ k \, l \, \pi }
-                {      n        }
-              \\
-    &=
-    \frac{    1   }
-         { 2 \, n }
-    \left[
-      \lambda_1
-      +
-      (-1)^l \, \lambda_{n + 1}
-      +
-      2 \sum_{ k = 1 }^{ n - 1 }
-        \lambda_{ k + 1 } \,
-        \cos \frac{ k \, l \, \pi }
-                  {      n        }
-    \right],
-  \end{aligned}
-  $$
-  where we have used Eq. \eqref{neq:lambda_reflection}
-  in the last step.
-
-  However, we are usually given only $\lambda_1, \dots, \lambda_n$
-  without $\lambda_{n + 1}$.
-  %
-  Fortunately, we can deduce the latter from
-  the condition that $\mu_n = 0$, which means
-  $$
-  \begin{aligned}
-    0 = \mu_n
-    =
-    \frac{1}{n}
-    \left[
-      \frac{ \lambda_1 + (-1)^n \, \lambda_{n+1} }
-           {               2                     }
-      +
-      \sum_{ k = 1 }^{ n - 1 }
-      \lambda_{k+1} (-1)^k
-    \right].
-  \end{aligned}
-  $$
-  This allows us to solve for $\lambda_{ n + 1 }$,
-  yielding Eq. \eqref{eq:lambdan}
-
-  Some examples for checking.
-  For $n = 2$,
-  $\lambda_3
-  = \mu_0 - 2 \, \mu_1
-  = -(\lambda_1 - 2 \, \lambda_2)$,
-  and
-  $$
-  \begin{aligned}
-  \mu_0
-  =
-  \lambda_2,
-  \qquad
-  \mu_1
-  =
-  \frac{ \lambda_1 - \lambda_2 }
-       {           2           }
-  .
-  \end{aligned}
-  $$
-  For $n = 3$,
-  $\lambda_4
-  = \mu_0 - 2 \, \mu_1 + 2 \, \mu_2
-  = \lambda_1 - 2 \, \lambda_2 + 2 \, \lambda_3$,
-  and
-  $$
-  \begin{aligned}
-  \mu_0
-  =
-  \frac{ \lambda_1 + 2 \, \lambda_3 } { 3 }
-  ,
-  \quad
-  \mu_1
-  =
-  \frac{ \lambda_2 - \lambda_3 } { 2 }
-  ,
-  \quad
-  \mu_2
-  =
-  \frac { 2 \, \lambda_1 - 3 \, \lambda_2 + \lambda_3 } { 6 }
-  .
-  \end{aligned}
-  $$
+  This entails Eq. \eqref{eq:mu_from_lambda_refl}.
+  %
+  The value of $\lambda_{n + 1}$
+  can be deduced from
+  imposing $\mu_n = 0$
+  and solving Eq. \eqref{eq:mu_from_lambda_refl}
+  for $l = n$,
+  yielding Eq. \eqref{eq:lambdan}.
 }
@@
   Differentiating this with respect to $t$ yields
-  $$
-  \lambda_k \, \alpha(t)
-  =
-  \frac {     1      }
-        { t + t_{k0} }.
-  $$
+  Eq. \eqref{eq:alpha_invtk}.
 }
@@
     \lambda_1
-    + 2 (-\lambda_2 + \lambda_3 - \cdots + (-1)^K \lambda_{K+1})
+    - 2 \, \lambda_2
+    + 2 \, \lambda_3 - \cdots
+    + (-1)^{K} 2 \, \lambda_{K+1})
   \right]
@@
 $$
-
-\hrulefill
-
-Example 1. If $n = 3, K = 1$,
-then $\lambda_1 = \lambda_2 = 1$, $\lambda_3 = 0$,
-and $\lambda_4 = \lambda_1 - 2 \, \lambda_2 + 2 \, \lambda_3 = -1$.
-%
-We can verify that $\mu_0 = 1/3, \mu_1 = 1/2, \mu_2 = -1/6$.
-%
-Although $\mu_1 > \mu_0$, the system will not be trapped
-in the current bin because of the reflective boundary condition.
-%
-Upon a visit to the middle bin, $j = 2$,
-$v_i$ at bin $i = 1$ or $3$ is updated by $1/2 - 1/6 = 1/3$,
-same as the update on bin $2$.
-Upon a visit to bin $i = 1$, the three bins are updated
-by $1/2 + 1/3 = 5/6$, $1/2 - 1/6 = 1/3$, and $-1/6$,
-respectively.
-
-\hrulefill
-
-Example 2. If $n = 3, K = 2$,
-then $\lambda_1 = \lambda_2 = \lambda_3 = 1$, $\lambda_4 = 1$.
-We have $\mu_0 = 1, \mu_1 = \mu_2 = 0$,
-recovering the single-bin scheme.
-
-\hrulefill
-
-Example 3. If $n = 4, K = 1$,
-then $\lambda_1 = \lambda_2 = 1, \lambda_3 = \lambda_4 = 0$,
-$\lambda_5 = (-1)(\lambda_1 - 2 \, \lambda_2 + 2 \, \lambda_3 - 2 \, \lambda_4 = 1$.
-The kernel is given by
-$\mu_0 = 1/2, \mu_1 = \sqrt{2}/8, \mu_2 = 1/4, \mu_3 = -\sqrt{2}/8$.
-}
+}%
 
@@
 we chose $\alpha_0 = 0.0001$
-and ran the simulation for $t = 10^8$ steps.
+and ran the simulation for $T = 10^8$ steps.
 %
@@
 we chose the nearest-neighbor updating scheme
-described in Sec. \eqref{sec:nnscheme}.
+described in Sec. \ref{sec:nnscheme}.
 %
@@
 %
-The optimal $1/\lambda$ gradually increased
-with simulation length,
-which parallel the gradual shift of
-the instantaneous eigenvalue $\lambda(t)$
-as discussed in Sec. \ref{sec:optschedule}.
-%which is in agreement with
-%the discussion in Sec. \ref{sec:optlambda}.
+%The optimal $1/\lambda$ gradually increased
+%with simulation length,
+%as discussed in Sec. \ref{sec:optlambda}.
 %
@@
 %
-Below is a technique of measuring $\Gamma_k$'s.
-
-The idea is to run
+%Below is a technique of measuring $\Gamma_k$'s.
+%
+%The idea is to run
+To measure $\Gamma_k$'s, we can run
 a simulation with a constant schedule,
@@
 %
-A practical difficulty is that
+%A practical difficulty is that
+However,
 we often do not know the intrinsic distribution, $p^*_i$,
@@
 %
-To circumvent the problem, we shall
+To circumvent the problem, we
 modify Eq. \eqref{eq:y2_eql} as
@@
 
-\section{\label{sec:invt_schedule}
-Inverse-time schedules}
-
-As an alternative to the
-optimal schedule from Eq. \eqref{eq:q_opt},
-here we consider the simpler
-inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
-%with a free parameter $\lambda$.
-%
-whose error can be computed explicitly.
-
-\subsection{\label{sec:invt_error}
-Error
-}
-
-Using Eq. \eqref{eq:alpha_invtlambda}
-in Eq. \eqref{eq:error_asym1} yields
-%
-\begin{align}
-\Err_A
-&=
-\frac{    1    }
-     { T + t_0 }
-\sum_{k = 2}^n
-  \frac{ \Gamma_k \, \nu_k^2 }
-       {    2 \, \nu_k - 1   }
-\left[
-  1 - \left(
-        \frac {     t_0 }
-              { T + t_0 }
-      \right)^{ 2 \, \nu_k - 1 }
-\right],
-\label{eq:error_asym_invt}
-\end{align}
-%
-where $\nu_k \equiv \lambda_k / \lambda$.
-%
-The error is asymptotically minimal
-only if $2\,\nu_k > 1$ for every $k$.
-%
-\note{At long times, $T \to \infty$, we get
-$$
-\begin{aligned}
-  \Err_A
-  =
-  \sum_{k = 2}^n
-  \begin{dcases}
-    \frac{    1    }
-         { T + t_0 }
-    \frac{ \Gamma_k \, \nu_k^2 }
-         {   2 \, \nu_k - 1    }
-    &
-    \mathrm{if \;} 2 \, \nu_k > 1,
-    \\%[1em]
-    %
-    %
-    \frac{    \Gamma_k    }
-         { 4 \, (T + t_0) }
-    \ln \frac{ T + t_0 }
-             {   t_0   }
-    &
-    \mathrm{if \;} 2 \, \nu_k = 1,
-    \\
-    %
-    %
-    \frac{  t_0^{ 2 \, \nu_k  - 1}  }
-         { (T + t_0)^{ 2 \, \nu_k } }
-    \frac{ \Gamma_k \, \nu_k^2 }
-         {   1 - 2 \, \nu_k    }
-    &
-    \mathrm{if \;} 2 \, \nu_k < 1.
-  \end{dcases}
-\end{aligned}
-$$
-%
-The last two cases
-are asymptotically suboptimal
-as they decay slower than $1/T$.
-}%
-%
-In this case,
-we can drop the second term
-in the square brackets,
-and identify the optimal $\lambda$ as
-the smallest real root of
-%
-\begin{equation}
-  \sum_{k = 2}^n
-    \frac { \lambda_k - \lambda }
-          { \left(2 - \lambda/ \lambda_k \right)^2 }
-  = 0.
-\notag
-%\label{eq:optimal_lambda_approx}
-\end{equation}
-%
-Since
-$$
-\frac{ \nu_k^2        }
-     { 2 \, \nu_k - 1 }
-=
-\frac{ \lambda_k^2 }
-     { \lambda \, (2 \, \lambda_k - \lambda) }
-\ge 1
-,
-$$
-with the equality achieved only at $\lambda = \lambda_k$,
-the optimal scheme must satisfy
-$\lambda_2 = \dots = \lambda_n$,
-corresponding to the single-bin scheme
-(cf. Sec. \ref{sec:optWL}).
-
-For a finite-length simulation,
-we need to include the residual error, $\Err_R$.
-%
-Assuming Eqs. \eqref{eq:error_res1} and \eqref{eq:t0_sinc},
-we get
-%
-\begin{equation}
-\Err_R
-=
-\sum_{k = 2}^n
-  \frac{ \Gamma_k \, \nu_k }
-       {        t_0   }
-  \left(
-      \frac{   t_0   }
-           { T + t_0 }
-   \right)^{ 2 \, \nu_k },
-\notag
-%\label{eq:error_res_invt}
-\end{equation}
-%
-with
-%
-\begin{equation}
-  q(T)
-  =
-  \frac{ 1 } { \lambda }
-  \ln\left(
-    \frac{ T + t_0 } { t_0 }
-  \right)
-  ,
-\notag
-%\label{eq:qt_invtlambda}
-\end{equation}
-%
-from integrating Eq. \eqref{eq:alpha_invtlambda}.
-%
-%We assume that
-%the system was initially equilibrated
-%at a constant $\alpha_0$,
-%such that the values of
-%$\left\langle y_k^2(0) \right\rangle$
-%can be computed from Eq. \eqref{eq:y2_eql},
-%as well as Eq. \eqref{eq:t0_sinc}.
-%
-The total error is
-%
-\begin{align}
-\Err
-&=
-\Err_R + \Err_A
-\notag
-\\
-&=
-\sum_{ k = 2 }^n
-  \frac
-  {
-    \Gamma_k \, \nu_k \,
-    \left[
-      \nu_k
-      +
-      (\nu_k - 1)
-      \left(
-        \frac{ t_0 } { T + t_0 }
-      \right)^{ 2 \, \nu_k - 1 }
-    \right]
-  }
-  {
-    (T + t_0) \, (2 \, \nu_k - 1)
-  }
-  .
-\label{eq:error_invt}
-\end{align}
-%
-%where
-%$$
-%\Err^f_k
-%\equiv
-%\left\langle
-%  y_k^2
-%\right\rangle_{ \alpha(t) }
-%=
-%\frac{  \Gamma_k \, \nu_k   }
-%     {    2 \, (T + t_0)    }
-%,
-%$$
-%is the equilibrium, or saturated, error of mode $k$
-%at the terminal updating magnitude
-%of $\alpha(T) = 1/[\lambda (T + t_0)]$.
-
-%\note{Derivation of Eq. \eqref{eq:error_invt}:
-%$$
-%\begin{aligned}
-%\Err_R
-%&=
-%\sum_{ k = 2 }^n
-%  \Err^f_k \left( \frac{   t_0   }
-%                       { T + t_0 }
-%          \right)^{ 2 \nu_k - 1 }
-%\\
-%\Err_A
-%&=
-%\sum_{ k = 2 }^n
-%  \frac{ 2 \, \Err^f_k \, \nu_k }
-%       {    2 \, \nu_k - 1      }
-%  \left[
-%    1 -
-%    \left(
-%      \tfrac{   t_0   }
-%            { T + t_0 }
-%      \right)^{2 \, \nu_k - 1}
-%  \right]
-%\\
-%&=
-%\sum_{ k = 2 }^n
-%  \Err^f_k \,
-%  \left[
-%    1 -
-%    \left(
-%      \tfrac{   t_0   }
-%            { T + t_0 }
-%      \right)^{2 \, \nu_k - 1}
-%  \right]
-%  +
-%  \frac{   \Err^f_k     }
-%       { 2 \, \nu_k - 1 }
-%  \left[
-%    1 -
-%    \left(
-%      \tfrac{   t_0   }
-%           { T + t_0 }
-%      \right)^{2 \, \nu_k - 1}
-%  \right].
-%\end{aligned}
-%$$
-%Adding the two yields Eq. \eqref{eq:error_invt}.
-%}
-
-%We give several remarks on Eq. \eqref{eq:error_invt}.
-%
-
-%Clearly, the final error is always greater than
-%the saturated equilibrium error
-%at the terminal $\alpha(T)$:
-%$$
-%E \ge E_\mathrm{sat} \equiv \sum_{k = 2}^n E^f_k,
-%$$
-%because the second term in the braces
-%of Eq. \eqref{eq:error_invt}
-%is nonnegative no matter the sign of $2 \, \nu_k - 1$
-%(in fact, it is a decreasing function
-%that vanishes only at $\nu_k \to +\infty$).
-%%
-%For a set of nonidentical $\lambda_k$'s,
-%the equality is reached only at $T = 0$.
-
-\note{The following subsection is deleted in the public version.
-
-\subsection{\label{sec:optlambda}
-Optimal proportionality constant
-}
-
-Intuitively,
-we expect the optimal $\lambda$
-that minimizes the error of the inverse-time schedule,
-Eq. \eqref{eq:error_invt},
-to be a kind of time average of
-the instantaneous eigenvalue $\lambda(t)$
-in Sec. \ref{sec:optschedule}.
-%
-Thus, it should also decrease with
-the simulation length, $T$.
-%
-Below we show this is indeed so.
-
-Let us first rewrite Eq. \eqref{eq:error_invt} as
-%
-$$
-\Err
-=
-\sum_{ k = 2 }^n
-  \frac{ \Gamma_k  }
-       { (T + t_0) }
-  G\left( \nu_k, \frac{ t_0 } { T + t_0} \right),
-$$
-%
-where the $k$th component is proportional to
-%
-$$
-G\left( \nu_k, r \right)
-=
-\frac { \nu_k
-        \,
-        \left[
-          \nu_k
-          +
-          (\nu_k - 1) \,
-          r^{2 \, \nu_k - 1}
-        \right]
-      }
-      { 2 \, \nu_k - 1 }
-.
-$$
-%
-As shown in Fig. \ref{fig:err_component},
-the function $G(\nu_k, r)$ depends non-trivially on $\nu_k$.
-%
-For a fixed and sufficiently small $r = t_0 / (T + t_0)$,
-$G(\nu_k, r)$ has two local minima along $\nu_k$,
-one at $\nu_k = 0$,
-and the other shallower one around $\nu_k = 1$.
-%
-For a short simulation, or a large $r$,
-the minimal error may be achieved at $\nu_k \approx 0$,
-corresponding to a large value of $\lambda$
-($\lambda \gg \lambda_k$).
-%
-However, this minimum is unstable,
-for $G(\nu_k, r)$ increases rapidly
-with decreasing $r$, or increasing simulation length,
-under a fixed $\nu_k$.
-%
-In other words,
-asymptotically, the minimal error
-is always achieved at the other minimum,
-$\nu_k = 1$ or $\lambda = \lambda_k$.
-%
-In this case,
-$G(\nu_k, r) \to G(1, 0) = 2$
-and the total error is roughly twice $E_\mathrm{sat}$.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/errcomp.pdf}
-  }
-  \caption{
-    \label{fig:err_component}
-    Component of the error, $G(\nu_k, r)$,
-    versus $\nu_k = \lambda_k / \lambda$,
-    where $r = t_0 / (T + t_0)$.
-    %
-    \note{This figure is produced by
-      \texttt{doc/fig/errcomp.gp}.
-    }
-  }
-\end{center}
-\end{figure}
-
-For a set of nonidentical eigenvalues,
-$\lambda_2, \dots, \lambda_n$,
-it is impossible to achieve $\lambda = \lambda_k$
-for every $k$,
-and we shall argue that
-the optimal $\lambda$
-tends to be biased toward
-the low end of the spectrum
-in a long simulation.
-%
-From Fig. \ref{fig:err_component},
-it is seen that around the minimum at $\nu_k = 1$,
-the error function $G(\nu_k, r)$
-has a steeper slope for the $\nu_k < 1$
-(or $\lambda > \lambda_k$) side
-than for the $\nu_k > 1$ (or $\lambda < \lambda_k$) side.
-%
-This means a smaller $\lambda$ is preferred
-in terms of minimizing the error.
-%
-Further, the difference in slopes
-between the two sides increases
-with the simulation length, $T$
-(or as $r$ decreases).
-%
-We thus expect the optimal $\lambda$
-to shift toward the smaller values of $\lambda_k$'s,
-and ultimately to approach $\min \lambda_k$
-in the infinite $T$ limit.
-
-The shift of the optimal $\lambda$
-is also affected by the values of $\Gamma_k$'s.
-%
-If large $\lambda_k$'s are associated with
-large $\Gamma_k$'s,
-these modes weight more in the error function,
-and the above shift toward smaller $\lambda$
-slows down.
-%
-For example,
-in the nearest-neighbor or Gaussian updating scheme
-discussed in Sec. \ref{sec:band-matrix},
-the index $k$ represents the wave number
-(or the inverse wave length)
-and $\lambda_k$ decreases with $k$.
-%
-As shown in Eq. \eqref{eq:Gamma_onestep},
-in the one-step sampling process,
-$\Gamma_k$ also decreases with $k$,
-and thus, $\lambda_k$ and $\Gamma_k$
-are positively correlated.
-%
-Therefore, we expect the shift of the optimal $\lambda$
-in the above case
-to be slower than in that in the case of perfect sampling,
-where $\Gamma_k \equiv 1$ is a constant.
-%
-If $\Gamma_k$ decreases too quickly with $k$,
-the shift can be hard to detect
-[see Eq. \eqref{eq:lambda_nn_onestep} for an example].
-}
-
-\subsection{Nearest-neighbor updating scheme}
-
-For the nearest-neighbor updating scheme
-introduced in Sec. \ref{sec:nnscheme},
-we can find an analytical expression
-of the optimal value of $\lambda$
-in the limit of large $n$.
-%
-In this case,
-we convert Eq.
-\eqref{eq:error_asym_invt}
-%and
-%\eqref{eq:optimal_lambda_approx}
-to an integral,
-$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
-%
-\begin{align}
-\Err_A
-=
-\frac{2 \, n}{\pi \, T}
-\int_0^{\pi/2}
-  \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
-       {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-\, dp.
-\notag
-%\label{eq:error_nn_asym_int}
-\end{align}
-
-For perfect sampling,
-$\Gamma(p) = 1$ from Eq. \eqref{eq:Gamma_perfect},
-and
-$$
-\begin{aligned}
-  \Err_A
-  %&=
-  %\frac{2 \, n}{\pi \, T}
-  %\int_0^{\pi/2}
-  %\frac{ \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2 }
-  %{ \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-  %\, dp
-  %\notag \\
-  &=
-  \frac{n}{4 \, T}
-  \left(
-    \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
-    +
-    \frac{ \lambda }
-    { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
-  \right)
-.
-\end{aligned}
-$$
-\note{The integral is evaluated by contour integration.
-%
-It is somewhat more convenient to change variable $p \to \frac{ \pi } { 2 } - p$,
-and
-$$
-\begin{aligned}
-  \Err_A
-  &=
-  \frac{n}{\lambda \, T}
-  \frac{1}{2 \pi i}
-  \int_0^{2 \, \pi}
-  \frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
-  { 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
-  \, dp
-  \\
-  &=
-  \frac{n}{\lambda \, T}
-  \frac{1}{2 \pi i}
-  \oint
-  \frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
-  { 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
-  \, \frac{dz}{z}
-  ,
-\end{aligned}
-$$
-where the contour is the unit circle
-in the complex plane of $z$.
-%
-This integral has five poles, one at the origin,
-which produces a residue
-$\frac{2 - 4 \, \mu_1 - \lambda}{4}$
-(from series expansion of small $z$),
-and the other four at
-$$
-z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
-$$
-with two of them inside the unit circle:
-$$
-z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
-{\sqrt{8 \mu_1}}.
-$$
-Thus,
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{      n       }
-     { \lambda \, T }
-\left(
- \frac{ 2 - 4 \, \mu_1 - \lambda }
-      {          4               }
- +
- \sum_{z = z_{\pm} }
- \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
- { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
-\right)
-\\
-&=
-\frac{       n      }
-     { \lambda \, t }
-\left(
-  \frac{ 2 - 4 \, \mu_1 - \lambda }
-       {          4               }
- +
- \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
-      { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2                 }
-\right)
-\\
-&=
-\frac{       n      }
-     { \lambda \, T }
-\left(
-  \frac{ 2 - 4 \, \mu_1 - \lambda }
-       {          4               }
- +
- \frac{ (\lambda/2)^2 }
- { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
-\right).
-\end{aligned}
-$$
-}
-%
-Minimizing the function yields
-%
-\begin{equation}
-  \lambda
-  =
-  \frac{ 1 - 4 \, \mu_1 }
-       { 1 - 2 \, \mu_1 }
-  ,
-\label{eq:lambda_nn_perfect}
-\end{equation}
-%
-and the asymptotic error in the optimal case is given by
-%
-\begin{equation}
-  \Err_A
-  =
-  \frac{n}{T}
-  \left(
-    1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
-  \right)
-  ,
-\notag
-%\label{eq:error_nn_perfect}
-\end{equation}
-%
-%From Eq. \eqref{eq:error_nn}, it is clear that
-which
-increases with the magnitude of $\mu_1$
-no matter its sign,
-and the minimal value, obtained at $\mu_1 = 0$,
-corresponds to the single-bin scheme.
-%
-\note{Note that
-we have ignored the residual error
-and assumed $\lambda < 2 \, \min \lambda_k = 2 - 8 \, \mu_1$
-for all $k$ in the above solution.
-%
-Fortunately,
-the $\lambda$ from Eq. \eqref{eq:lambda_nn_perfect}
-satisfies this condition
-by the stability condition,
-$4 \, \mu_1 < 1$.
-}%
-\note{This is because,
-$$
-\frac{ 1 - 4 \, \mu_1 } { 1 - 2 \, \mu_1 } < 2 - 8 \, \mu_1
-\quad \leftrightarrow \quad
-1 < 2 - 4 \, \mu_1.
-$$
-}%
-
-For the one-step sampling process,
-we find from Eq. \eqref{eq:Gamma_onestep}
-that $\Gamma(p) = \cot^2 p$
-is peaked around $p \approx 0$,
-%
-Then
-$$
-  \Err_A
-  \propto
-  \frac{   2 \, n }
-       { \pi \, T }
-  \frac{             1             }
-       {  \lambda \, (2 - \lambda) }
-  .
-$$
-\note{
-$$
-\begin{aligned}
-  \Err_A
-  &
-  \propto
-  \frac{   2 \, n }
-       { \pi \, T }
-  \left.
-  \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
-       { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-  \right|_{ p \approx 0 }
-  %\notag \\
-  %&
-  \approx
-  \frac{   2 \, n }
-       { \pi \, T }
-  \frac{             1             }
-       {  \lambda \, (2 - \lambda) }
-  .
-\end{aligned}
-$$
-}%
-%
-The optimal $\lambda$ is, therefore,
-%
-\begin{equation}
-\lambda \approx 1.
-\label{eq:lambda_nn_onestep}
-\end{equation}
-%
-\note{We can be a bit more precise here.
-%
-From Eq. \eqref{eq:Gamma_onestep}, we get
-%
-$$
-\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
-$$
-%
-where we have introduced
-$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
-as a small parameter to avoid the divergence
-around the origin.
-%
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{   2 \, n }
-     { \pi \, t }
-\int_0^{\pi/2}
-    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
-         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-    \frac{ \cos^2 p }
-         { \sin^2 p + \delta^2 }
-\, dp
-\notag \\
-&
-\stackrel{    \delta \ll 1     }
-         { =\joinrel=\joinrel= }
-\frac{   2 \, n }
-     { \pi \, t }
-\frac{             1             }
-     {  \lambda \, (2 - \lambda) }
-\frac{    1   }
-     { \delta }
-.
-\end{aligned}
-$$
-The integral is evaluated as follows.
-%
-We again change variable $p \to \frac{ \pi } { 2 } - p$,
-and
-$$
-\begin{aligned}
-\Err_A
-&=
-\frac{n}{\lambda \, t}
-\frac{1}{2 \pi i}
-\int_0^{2 \, \pi}
-\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
-     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
-\cdot
-\frac{ \sin^2 p            }
-     { \cos^2 p + \delta^2 }
-\, dp
-\\
-&=
-\frac{n}{\lambda \, t}
-\frac{1}{2 \pi i}
-\oint
-\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
-     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
-\cdot
-\frac{ 4 - \left( z + \frac 1 z \right)^2 }
-     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
-\, \frac{dz}{z}
-\\
-&=
-\frac{n}{\lambda \, t}
-\frac{1}{2 \pi i}
-\oint
-\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
-     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
-\cdot
-\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
-     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
-\, \frac{ dz }{ z^3 }
-\\
-&=
-\frac{ n } { \lambda \, t }
-\sum_l \operatorname{Res}_l
-,
-\end{aligned}
-$$
-where the contour is the unit circle
-in the complex plane of $z$,
-and the integral is reduced to a sum of residues
-around the poles within.
-
-The poles of the integral come from three sources,
-namely,
-% 1.
-$z = 0$,
-% 2.
-the zeros of
-$(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2$,
-% 3.
-and the zeros of
-$( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2$.
-
-For the first pole around the origin,
-we need to expand the integrand as
-$$
-\frac{ A + B \, z^2 + \cdots }
-     {          z^3          },
-$$
-then the residue is $B$.
-%
-Since
-$$
-\begin{aligned}
-\left[ z^2 - \mu_1 (z^2 + 1)^2 \right]^2
-&
-\approx
-[ - \mu_1 + (1 - 2 \, \mu_1) \, z^2]^2
-\\
-&
-\approx
-\mu_1^2
-\left[
-  1 + \left(4 - \tfrac { 2  } { \mu_1 } \right) z^2
-\right],
-\\
-%
-4 \, z^2 - ( z^2 + 1 )^2
-&
-\approx
--( 1 - 2 \, z^2 ),
-\\
-%
-(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2
-&
-\approx
--2 \, \mu_1 \,
-\left[
-  1 + \left(
-        2 + \tfrac{ \lambda - 2 } { 2 \, \mu_1 }
-      \right)
-      \, z^2
-\right],
-\\
-%
-( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2
-&
-\approx
-1 + (2 + 4 \, \delta^2) \, z^2,
-\end{aligned}
-$$
-the integrand is
-$$
-\begin{aligned}
-\frac{ \mu_1 }
-     {   2   }
-\left[
-  1 + \left(
-        - \frac{ 2 + \lambda } { 2 \, \mu_1 }
-        - 2 - 4 \, \delta^2
-      \right) \, z^2
-\right].
-\end{aligned}
-$$
-and the residue is
-$$
-\operatorname{Res}_1
-=
-  - \frac{ 2 + \lambda } { 4 }
-  - \mu_1 \, \left( 1 + 2 \, \delta^2 \right).
-$$
-
-For the second part,
-we have four roots of
-$$
-(2 - \lambda) \, z^2 = 2 \, \mu_1 \, (z^2 + 1)^2,
-$$
-which are
-\begin{equation}
-  z
-  =
-  \pm
-  \sqrt { \frac{ 2 - \lambda }
-               { 8 \, \mu_1  } }
-  \pm
-  \sqrt { \frac{ 2 - \lambda }
-               { 8 \, \mu_1  } - 1 }
-  .
-\tag{N5}
-\label{neq:nnint_local_zeros_part2}
-\end{equation}
-%
-We shall distinguish two cases.
-\textbf{Case A.}
-If $\lambda < 2 - 8 \, \mu_1$,
-two of the roots lie in the unit circle, namely
-$$
-z
-=
-\pm
-\left(
-\sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } }
--
-\sqrt { \frac{ 2 - \lambda }
-             { 8 \, \mu_1  } - 1 }
-\right),
-$$
-which satisfy
-$$
-4 \, \mu_1 \, ( z^2 + 1)
-=
-2 - \lambda
--
-\sqrt{ ( 2 - \lambda ) ( 2 - \lambda - 8 \, \mu_1 ) }.
-$$
-Thus,
-$$
-\begin{aligned}
-\left[ z^2 - \mu_1 \, \left( z^2 + 1 \right)^2 \right]^2
-&=
-\left( z^2 - \tfrac{ 2 - \lambda } { 2 } z^2 \right)^2
-=
-\frac{ \lambda^2 } { 4 } z^4.
-\\
-%
-4 z^2 - \left( z^2 + 1 \right)^2
-&=
--\frac{ 2 - \lambda - 8 \, \mu_1  } { 2 \, \mu_1 } z^2,
-\\
-%
-\left( z^2 + 1 \right)^2 + 4 \, \delta^2 \, z^2
-&=
-\frac{ 2 - \lambda + 8 \, \mu_1 \, \delta^2 }
-     { 2 \, \mu_1 }
-     z^2,
-\\
-%
-(2 - \lambda) \, 2 \, z
--
-8 \, \mu_1 \, \left( z^2 + 1 \right) \, z
-&=
-2 \, z \, \left[2 - \lambda - 4 \,  \mu_1 \, (z^2 + 1) \right]
-\\
-&= 2 \, z \, \sqrt{ (2 - \lambda) ( 2 - \lambda - 8 \, \mu_1 ) },
-\end{aligned}
-$$
-where we have used the L'H\^{o}pital's rule
-and taken the derivative for the last factor.
-%
-Note that the two roots within the unit share the same
-residue and the sum for this part is
-%
-$$
-\operatorname{Res}_2
-=
--\frac{                 \lambda^2                     }
-      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
-\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }.
-$$
-%
-\textbf{Case B.}
-If $2 - 8 \, \mu_1 \le \lambda < 2$,
-the four zeros in Eq. \eqref{neq:nnint_local_zeros_part2}:
-%
-\begin{equation}
-  z
-  =
-  \pm
-  \sqrt { \frac{ 2 - \lambda }
-               { 8 \, \mu_1  } }
-  \pm i \,
-  \sqrt { 1 - \frac{ 2 - \lambda }
-                   { 8 \, \mu_1  } }
-  ,
-\notag
-%\label{eq:nnint_local_zeros_part2b}
-\end{equation}
-%
-all satisfy $|z| = 1$, and hence
-all lie on the border of the unit circle.
-%
-Thus, each contributes half of the residue there.
-%
-However, since the four zeros satisfy
-$$
-2 - \lambda - 4 \, \mu_1 \, ( z^2 + 1)
-=
-\pm i
-\sqrt{ ( 2 - \lambda ) ( 8 \, \mu_1 - 2 + \lambda ) },
-$$
-the sum of the four residues gives zero in this case:
-$$
-\operatorname{Res}_2 = 0.
-$$
-In summary,
-$$
-\operatorname{Res}_2
-=
-\begin{dcases}
--\frac{                 \lambda^2                     }
-      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
-\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }
-&
-\mathrm{if \;}
-\lambda < 2 - 8 \, \mu_1,
-\\
-0
-&
-\mathrm{otherwise}.
-\end{dcases}
-$$
-
-For the last part,
-we have four roots of
-$(z^2 + 1)^2 = - 4 \, \delta^2 \, z^2$,
-with two of them lying in the unit circle:
-$$
-z = \pm i \left( \sqrt{ 1 + \delta^2 } - \delta \right),
-$$
-satisfying
-$$
-z^2 + 1 + 2 \, \delta^2 = 2 \, \delta \, \sqrt{1 + \delta^2}.
-$$
-Then,
-$$
-\begin{aligned}
-z^2 - \mu_1 \, \left( z^2 + 1 \right)^2
-&= z^2 \, (1 + 4 \, \mu_1 \, \delta^2),
-\\
-4 \, z^2 - \left( z^2 + 1 \right)^2
-&=
-4 \, z^2 \, (1 + \delta^2)
-\\
-%
-(2 - \lambda) \, z^2
-- 2 \, \mu_1 \, \left( z^2 + 1 \right)^2
-&=
-\left(
-  2 - \lambda + 8 \, \mu_1 \, \delta^2
-\right) \, z^2
-\\
-%
-4 \, z \, \left( z^2 + 1 \right)
-+
-8 \, \delta^2 \, z
-&=
-4 \, z \, (z^2 + 1 + \delta^2)
-=
-8 \, z \, \delta \, \sqrt{ 1 + \delta^2 }.
-\end{aligned}
-$$
-Note that the above two roots share the same
-residue and the sum for this part is
-$$
-\operatorname{Res}_3
-=
-\frac{ \left( 1 + 4 \, \mu_1 \, \delta^2 \right)^2 }
-     {        2 + 8 \, \mu_1 \, \delta^2 - \lambda }
-\frac{ \sqrt{ 1 + \delta^2 } }
-     {        \delta         }.
-$$
-
-Now as $\delta \to 0$, the last contribution will dominate,
-and we have
-$$
-\Err_A
-\approx
-\frac{ n } { T }
-\frac{ (1 + 4 \, \mu_1 \, \delta^2)^2 }
-{ \lambda \, (2 + 8 \, \mu_1 \, \delta^2 - \lambda) }
-\frac{ 1 } { \delta }
-\approx
-\frac{ n } { T }
-\frac{ 1 }
-{ \lambda \, (2 - \lambda) }
-\frac{ 1 } { \delta },
-$$
-and the minimum is reached at $\lambda \approx 1$.
-
-The weakness of the above calculation is two fold.
-%
-First, it ignores the residual error.
-%
-Second, the simplified expression for the asymptotic error
-$E_A$ may also be misused.
-%
-Namely, with $\lambda \approx 1$,
-we cannot assume $\lambda < 2 \, \lambda_k$ in general.
-%
-So it only works for a very small value of $\mu_1$.
-}
-
-However, Eq. \eqref{eq:lambda_nn_onestep}
-is an approximation good only
-for sufficiently short simulations,
-since we have taken the limit of $n \to \infty$
-for a fixed $T$.
-%
-For very long simulations,
-the optimal $\lambda$
-will ultimately drift to
-the smallest eigenvalue,
-$\min \lambda_k = 1 - 4 \, \mu_1$.
-%as discussed in Sec. \ref{sec:optlambda}.
-
 \section{\label{sec:stabilize_wband}
@@
 specified by an arbitrary kernel
-is not necessarily stable
-with all nonnegative eigenvalues.
+is not necessarily stable,
+i.e., some eigenvalues may be negative.
 %
@@
 
+\section{\label{sec:invt_schedule}
+Inverse-time schedules}
+
+As an alternative to the
+optimal schedule from Eq. \eqref{eq:q_opt},
+here we consider the simpler
+inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
+%with a free parameter $\lambda$.
+%
+whose error can be computed explicitly.
+
+\subsection{\label{sec:invt_error}
+Error
+}
+
+Using Eq. \eqref{eq:alpha_invtlambda}
+in Eq. \eqref{eq:error_asym1} yields
+%
+\begin{align}
+\Err_A
+&=
+\frac{    1    }
+     { T + t_0 }
+\sum_{k = 2}^n
+  \frac{ \Gamma_k \, \nu_k^2 }
+       {    2 \, \nu_k - 1   }
+\left[
+  1 - \left(
+        \frac {     t_0 }
+              { T + t_0 }
+      \right)^{ 2 \, \nu_k - 1 }
+\right],
+\label{eq:error_asym_invt}
+\end{align}
+%
+where $\nu_k \equiv \lambda_k / \lambda$.
+%
+Assuming Eq. \eqref{eq:error_res1}
+for the residual error, and Eq. \eqref{eq:t0_sinc},
+we get
+%
+\begin{equation}
+\Err_R
+=
+\sum_{k = 2}^n
+  \frac{ \Gamma_k \, \nu_k }
+       {        t_0   }
+  \left(
+      \frac{   t_0   }
+           { T + t_0 }
+   \right)^{ 2 \, \nu_k },
+\notag
+%\label{eq:error_res_invt}
+\end{equation}
+%
+with
+$
+  q(T)
+  =
+  ( 1 / \lambda )
+  \ln\left(
+    1 + T / t_0
+  \right)
+$
+%
+from integrating Eq. \eqref{eq:alpha_invtlambda}.
+%
+%We assume that
+%the system was initially equilibrated
+%at a constant $\alpha_0$,
+%such that the values of
+%$\left\langle y_k^2(0) \right\rangle$
+%can be computed from Eq. \eqref{eq:y2_eql},
+%as well as Eq. \eqref{eq:t0_sinc}.
+%
+The total error is
+%
+\begin{align}
+\Err
+&=
+\Err_R + \Err_A
+\notag
+\\
+&=
+\sum_{ k = 2 }^n
+  \frac
+  {
+    \Gamma_k \, \nu_k \,
+    \left[
+      \nu_k
+      +
+      (\nu_k - 1)
+      \left(
+        \frac{ t_0 } { T + t_0 }
+      \right)^{ 2 \, \nu_k - 1 }
+    \right]
+  }
+  {
+    (T + t_0) \, (2 \, \nu_k - 1)
+  }
+  .
+\label{eq:error_invt}
+\end{align}
+%
+
+\subsection{Nearest-neighbor updating scheme}
+
+For the nearest-neighbor updating scheme
+introduced in Sec. \ref{sec:nnscheme},
+we can find an analytical expression
+of the optimal value of $\lambda$
+in the limit of large $n$.
+%
+In this case,
+we convert Eq.
+\eqref{eq:error_asym_invt}
+%and
+%\eqref{eq:optimal_lambda_approx}
+to an integral,
+$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
+%
+\begin{align}
+\Err_A
+=
+\frac{2 \, n}{\pi \, T}
+\int_0^{\pi/2}
+  \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
+       {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
+\, dp.
+\notag
+%\label{eq:error_nn_asym_int}
+\end{align}
+
+For perfect sampling,
+$\Gamma(p) = 1$ from Eq. \eqref{eq:Gamma_perfect},
+and
+$$
+\begin{aligned}
+  \Err_A
+  &=
+  \frac{n}{4 \, T}
+  \left(
+    \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
+    +
+    \frac{ \lambda }
+    { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
+  \right)
+.
+\end{aligned}
+$$
+%
+Minimizing the function yields
+%
+\begin{equation}
+  \lambda
+  =
+  \frac{ 1 - 4 \, \mu_1 }
+       { 1 - 2 \, \mu_1 }
+  ,
+\notag
+%\label{eq:lambda_nn_perfect}
+\end{equation}
+%
+and the asymptotic error in the optimal case is given by
+%
+\begin{equation}
+  \Err_A
+  =
+  \frac{n}{T}
+  \left(
+    1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
+  \right)
+  ,
+\notag
+%\label{eq:error_nn_perfect}
+\end{equation}
+%
+%From Eq. \eqref{eq:error_nn}, it is clear that
+which
+increases with the magnitude of $\mu_1$
+no matter its sign,
+and the minimal value, obtained at $\mu_1 = 0$,
+corresponds to the single-bin scheme.
+%
+
+For the one-step sampling process,
+we find from Eq. \eqref{eq:Gamma_onestep}
+that $\Gamma(p) = \cot^2 p$
+is peaked around $p \approx 0$,
+%
+Then
+$$
+  \Err_A
+  \propto
+  \frac{   2 \, n }
+       { \pi \, T }
+  \frac{             1             }
+       {  \lambda \, (2 - \lambda) }
+  .
+$$
+%
+The optimal $\lambda$ is, therefore,
+%
+\begin{equation}
+\lambda \approx 1.
+\label{eq:lambda_nn_onestep}
+\end{equation}
+%
+
+However, Eq. \eqref{eq:lambda_nn_onestep}
+is an approximation good only
+for sufficiently short simulations,
+since we have taken the limit of $n \to \infty$
+for a fixed $T$.
+%
+For very long simulations,
+the optimal $\lambda$
+will ultimately drift to
+the smallest eigenvalue,
+$\min \lambda_k = 1 - 4 \, \mu_1$.
+
 \bibliography{simul}

### 2016-04-19 doc/invt.tex
even sampling of the optimal schedule

@@
 the index of a small interval, or a bin,
-$(z, z + dz)$.
+$(z, z + \delta z)$.
 %
@@
 we set the bias potential to zero,
-$v_i \equiv 0$,
+$V_i \equiv 0$,
 %
@@
 we reset the origin of the time, $t$,
-and started the target schedule.
+and started the testing schedule.
 %specified by either
@@
 For the Gaussian updating scheme,
-the kernel is truncated at
-$\min\{10 \, \sigma, n - 1\}$
+the kernel was truncated at
+$b = \min\{10 \, \sigma, n - 1\}$
 and numerically stabilized
 using the technique
-in Sec. \ref{sec:stabilize_wband}.
+in Appendix \ref{sec:stabilize_wband}.
 
@@
 the underlying sampling process,
-we have devised two sampling processes,
+we tested two sampling processes,
 a global one and a local one.
 %
-Both cases were based on
-Metropolis algorithm\cite{metropolis1953, newman, frenkel,
+Both are based on
+Metropolis MC algorithm\cite{
+  metropolis1953, newman, frenkel,
   landau_binder}:
 %
-in each step, a new bin index $j$ was proposed
+in each step, a new bin index $j$ is proposed
 and then accepted with probability
@@
 $
-A(i \to j) = \min\{ 1, \exp(v_i - v_j) \}.
+A(i \to j) = \min\{ 1, \exp(V_i - V_j) \}.
 $
 However,
-in the first and global sampling process,
+in the first or global sampling process,
 we choose the destination $j$
-randomly out of the $n$ bins.
-%
-This sampling process
+uniformly out of the $n$ bins.
+%
+This process
 emulates the perfect sampling process
-discussed in Sec. \ref{sec:Gamma}
-in the long time limit.
-%
-In the second or local process,
-we limit the destination $j$
-to the two neighbors of $i$,
-$i \pm 1$, with equal probability $1/2$,
-to mimic the one-step sampling process
+in Sec. \ref{sec:Gamma} asymptotically.
+%
+In the second or local sampling process,
+$j$ is either $i - 1$ or $i + 1$
+(adjusted by the boundary condition)
+with equal probability $1/2$,
+which mimics the one-step sampling process
 in Sec. \ref{sec:Gamma}.
@@
 (cf. Appendix \ref{sec:invt_schedule}),
-and thus be tested against the theory.
+and tested against simulation.
 
@@
 For the single-bin updating scheme,
-we chose $\alpha_0 = 0.0001$
+we chose $\alpha_0 = 10^{-4}$
 and ran the simulation for $T = 10^8$ steps.
@@
 the optimal $\lambda$ occurred around $1.0$,
-verifying Eq. \eqref{eq:alpha_invt1}.
+verifying the optimality of Eq. \eqref{eq:alpha_invt1}.
 %
@@
 differed only by a multiplicative constant,
-again agreeing with the theory.
-%
-With identical eigenvalues,
-$\lambda_1 = \cdots = \lambda_n = 1$,
+which again agreed with the theory:
+%
+with identical eigenvalues,
+$\lambda_2 = \cdots = \lambda_n = 1$,
 the error in Eq. \eqref{eq:error_invt}
@@
     %
-    The results have been averaged over $1000$ independent runs.
+    The results were averaged over $1000$ independent runs.
     %
@@
 %
-We again find good agreement
-between simulation and analytical results.
-%
-A new feature is that the optimal $\lambda$
-depends on the simulation length, $T$.
-%
-As shown in Fig. \ref{fig:errnnbr},
-the normalized error, $(T + t_0) \, \Err$,
-hence the optimal value of $1/\lambda$,
-depended non-trivially on the simulation length, $T$,
-in the global sampling case.
+We choose $\mu_1 = 0.24$ to the smallest eigenvalue
+close to zero.
+%
+Again, simulation and theory agreed well,
+as shown in Fig. \ref{fig:errnnbr}.
+%
+The normalized error
+and the optimal value of $1/\lambda$,
+depended on the simulation length, $T$,
+in the global sampling case,
+because of the near zero eigenvalues,
+as discussed in Appendix \ref{sec:invt_schedule}.
 %
@@
 %
-For the local sampling case,
-the $T$-dependence was hardly visible
-because of the predominantly large $\Gamma_k$'s
-of the small $k$ modes.
+However, for the local sampling case,
+the $T$-dependence was suppressed
+because of the dominance of large $\Gamma_k$'s
+of the small-$k$ modes.
 %
@@
 %    %
-%    The results have been averaged over $1000$ independent runs.
+%    The results were averaged over $1000$ independent runs.
 %  }
@@
 with $\sigma = 10$,
-we computed the schedule,
+we computed the schedule for the local sampling process,
 and compared it with the inverse-time schedule
@@
 with the optimal $\lambda$.
+%
+As shown in Fig. \ref{fig:optacmp}(a),
+the optimal schedule is a more complex curve
+than the optimized inverse-time schedule.
+%
+The complexity of the optimal schedule
+corresponds to the non-exponential tail
+of the mass distribution function,
+as shown Fig. \ref{fig:optacmp}(b)
+[cf. Sec. \ref{sec:mass_distr}].
+%
+The errors in the two cases were
+$9.1 \times 10^{-5}$
+and
+$1.7 \times 10^{-4}$,
+respectively,
+%for the local sampling process,
+showing the superiority of
+the optimal schedule from Eq. \eqref{eq:q_opt}.
+%
+Besides, we verified the scaling property
+discussed in Sec. \ref{sec:mass_distr},
+by comparison with a simulation [Fig. \ref{fig:optacmp}(a)]
+of half length, $T' = 5 \times 10^7$
+and doubled initial updating magnitude,
+$\alpha_0 = 2 \times 10^{-4}$,
+which left the normalized mass distribution invariant,
+as shown in Fig. \ref{fig:optacmp}(b).
 
@@
     for the Gaussian updating scheme
-    with $\sigma = 10$.
+    with $\sigma = 10$
+    and the local updating process.
     %
@@
 
-As shown in Fig. \ref{fig:optacmp}(a),
-the optimal schedule is a more complex curve
-than the optimized inverse-time schedule.
-%
-In Fig. \ref{fig:optacmp}(b),
-we show that in the former case,
-the rate of change of $\alpha^{-1}(t)$
-increases gradually
-from $\lambda(0)$ to $\lambda(T) \approx 1$,
-as discussed in Sec. \ref{sec:optschedule}.
-%
-The errors in the two cases were
-$2.7 \times 10^{-7}$
-and
-$6 \times 10^{-7}$,
-respectively,
-for the global sampling process,
-%
-or
-$9.1 \times 10^{-5}$
-and
-$1.7 \times 10^{-4}$,
-respectively
-for the local sampling process,
-showing the superiority of
-the optimal schedule from Eq. \eqref{eq:q_opt}.
-
 On the other hand,
@@
 Eq. \eqref{eq:alpha_invt1},
-with the optimal $t_0$
-given by Eq. \eqref{eq:t0_sinc},
+with $t_0$ given by Eq. \eqref{eq:t0_sinc},
 as shown in Fig. \ref{fig:sinc}(a).
-
+%
 In Fig. \ref{fig:sinc}(b),
@@
 %
-The equivalent width, $\sigma$, is roughly given
+The equivalent width, $\sigma$, is roughly
+%
 \begin{equation}
@@
 \end{equation}
+%
 where $g = 2$ for the periodic case
@@
 %
-However, as shown in the inset of Fig. \ref{fig:sinc}(b),
+However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
 the ruggedness disappears
-in the actual updating matrix, Eq. \eqref{eq:w_band_refl},
+in the updating matrix by Eq. \eqref{eq:w_band_refl},
 as the above oscillatory term for $l = i - j$
-is canceled by either the term for $l = i + j - 1$
-or that for $l = i + j - 2 \, n - 1$,
+is canceled by the term for either $l = i + j - 1$
+or $l = i + j - 2 \, n - 1$
+with the opposite sign.
 
@@
     compared to a Gaussian with a similar width,
-    $\sigma = 5$.
+    $\sigma = 6$.
+    %
+    The local sampling process was used.
     The lines are a guide to the eyes.
@@
 %
-To illustrate this,
+As an example,
 we compare Gaussian updating schemes
-with different width, $\sigma$.
-%
-For the global sampling process,
-we see from Fig. \ref{fig:errsigscan}(a)
+with different widths, $\sigma$.
+%
+As shown in Fig. \ref{fig:errsigscan}(a)
 that the minimal error was achieved
 by the widest window for the shorter simulation
-of $T = 10^8$.
-%
-However, as the simulation lengthens,
+of $T = 10^8$,
+for the global sampling process.
+%
+As the simulation lengthens,
 the minimum at $\sigma = 0$ became more favorable.
@@
 However,
-the shift was hard to detect
-for the local sampling scheme,
+the minimum stayed at zero width,
 as shown in Fig. \ref{fig:errsigscan}(b).
@@
 In contrast,
-the normalized error of the bandpass updating scheme
-was insensitive to the simulation length,
+the normalized error of the bandpass updating schemes
+were insensitive to the simulation length,
 in agreement with Eq. \eqref{eq:error_sinc}.
 %
-Further, in all cases the bandpass updating scheme
-appeared to be superior to the Gaussian one.
+Besides, the bandpass updating schemes
+yielded smaller errors than
+the corresponding Gaussian ones.
 
@@
 
-%\subsection{\label{sec:measure_gamma}
-%Measurement of the integrals
-%of the autocorrelation functions}
+\subsection{\label{sec:measure_gamma}
+Measurement of the integrals
+of the autocorrelation functions}
 
@@
 \label{eq:Y_def}
-\\
+\end{align}
+%
+with
+$
 \bar V
-&=
+\equiv
 \sum_{ j = 1 }^n p_j \, V_j.
-\notag
-\end{align}
+$
 
@@
 as
-$$
-\begin{aligned}
+$
   x_i
-  &=
+  =
   (V_i - \bar V) + s_i,
-  \\
+$
+where
+$
   s_i
-  &=
+  \equiv
   \ln \frac{ p_i } { p^*_i }
@@
     p_j \, \ln \frac{ p_j } { p^*_j }.
-\end{aligned}
-$$
+$
 Thus, by Eqs. \eqref{eq:y_def} and \eqref{eq:Y_def},
@@
 
-%\subsection{\label{sec:program_list}
-%Program
-%}
+As an example, we used the method to compute
+the integrals, $\Gamma_k$'s, from an MD simulation,
+and compared them to those from
+the local MC sampling process
+[cf. Sec. \ref{sec:results_system}].
+%
+As shown in Fig. \ref{fig:gamcmp},
+the larger values of $\Gamma_k$'s
+with smaller $k$'s
+from the two sampling processes
+were almost proportional,
+suggesting similarity 
+of local sampling processes.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/gamcmp.pdf}
+  }
+  \caption{
+    \label{fig:gamcmp}
+    Integrals of the autocorrelation functions
+    from an MD simulation versus those
+    from the local MC simulation.
+    %
+    The MD simulation was ran for $10^{10}$ steps with
+    the step size $\Delta t_\mathrm{MD} = 0.005$.
+    %
+    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$,
+    and the temperature was $1.0$.
+    %
+    The velocity rescaling\cite{bussi2007} thermostat was
+    applied every MD step with the viscocity set to 20.
+    %
+    The reflective boundary condition was used.
+    %
+    \note{The figure was produced by \texttt{doc/fig/gamcmp.gp}.
+      Thee data were saved under \texttt{data/gamcmp}
+      (see the \texttt{Makefile} there).
+    }%
+  }
+\end{center}
+\end{figure}
+
+\subsection{\label{sec:program_list}
+Program
+}
 
@@
 %
-\note{For example,
+%\note{
+  %For example,
   %the nearest-neighbor updating scheme
@@
   %
-  %Even
+  Even
   the Gaussian updating scheme can be unstable
   if the updating kernel is truncated.
-}%
+%}%
 %
@@
 of the optimal value of $\lambda$
-in the limit of large $n$.
+in the large-$n$ limit.
 %
 In this case,
-we convert Eq.
-\eqref{eq:error_asym_invt}
-%and
-%\eqref{eq:optimal_lambda_approx}
-to an integral,
-$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
+we
+drop the second term in the square brackets
+of Eq. \eqref{eq:error_asym_invt},
+which vanishes asymptotically if $2 \, \nu_k > 1$,
+and convert the asymptotic error
+to an integral:
+%$\sum_i \to \frac{2 \, n}{\pi} \int_0^{\pi/2} dp$, and
 %

### 2016-04-19 prog/predict.c

@@ -42,7 +42,7 @@ static void invt_geterr(invtpar_t *m,
 
   /* save the optimal schedule to file */
   t0 = 2 * m->c / m->alpha0;
-  intq_save(intq, m->c, t0, m->fnalpha);
+  intq_save(intq, m->c, t0, m->alpha_resample, m->fnalpha);
 
   printf("c %g, t0 %g, err %g, sqr %g (invt), %g, sqr %g (exact), %s\n",
       m->c, t0, err1, err1 * err1, err2, err2 * err2, m->fnalpha);

### 2016-04-19 doc/invt.tex
add massq.gp

@@
 is listed in Appendix \ref{sec:program}.
+
+\subsubsection{\label{sec:mass_distr}
+Characterization of the optimal schedule}
+
+We can characterize the optimal schedule
+geometrically
+through the normalized mass distribution
+along $q$ as
+\begin{equation}
+  m(q)
+  =
+  \frac{
+    \sqrt{
+      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2(q)
+    }
+  }
+  {
+    \int_0^{ q(T) }
+      \sqrt{
+        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2(q')
+      }
+      \, d q'
+  }
+  ,
+\label{eq:mass_distr}
+\end{equation}
+%
+such that
+$\int_0^{q(T)} m(q) = 1$.
+%
+As shown in Fig. ,
+the distribution peaks at $q = q(T)$,
+and decreases with the difference $q(T) - q$.
+%
+According to Eq. \eqref{eq:q_opt},
+the ordinate of the distribution curve at $q$
+gives $1/(T\alpha) = 1/(T \dot q)$,
+and the area under the curve in the region $q' < q$
+gives $t/T$.
+%
+From this picture, we can draw several conclusions.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.9\linewidth]{fig/massq.pdf}
+  }
+  \caption{
+    \label{fig:massq}
+    Geometric characterization of the optimal schedule,
+    $\alpha(t)$,
+    via the normalized mass distribution, $m(q)$.
+    %
+    \note{The figure was produced by \texttt{doc/fig/massq.gp}.
+    }%
+  }
+\end{center}
+\end{figure}
+
+First,
+the optimal schedule has the following scaling property.
+%
+For two simulations of lengths $T_1$ and $T_2$,
+if $q(T_1) = q(T_2)$
+and they share the same $\lambda_k$'s and $\Gamma_k$'s,
+the updating magnitudes at the corresponding times,
+$t_1/T_1 = t_2/T_2$,
+can be related as $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
+%
+In other words,
+if $\alpha(t_1)$ is
+an optimal schedule of the simulation of length $T_1$,
+then
+$\frac{T_1}{T_2} \alpha\left( \frac{ T_1 } { T_2 } t_2 \right)$
+is an optimal schedule of the simulation of length $T_2$.
+%
+The asymptotic errors are related by
+$T_1 \, E_{A1} = T_2 \, E_{A2}$.
 
@@
 
+\subsection{\label{sec:deplength}
+Dependence on the simulation length}
+
+The parameter, $\lambda$, is equal to
+the difference
+$$
+  \lambda
+  =
+  \Delta \left( \frac{ 1 } { T \, \alpha } \right)
+  \equiv
+  \left.
+  \frac { 1 } { T \, \alpha(t) }
+  \right|_{t = 0}^{t = T}
+  .
+$$
+which corresponds to the height difference
+of the normalized mass distribution
+$
+\left. m(q) \right|_{q = 0}^{q = q(T)}
+$
+in the optimal schedule
+(cf. Sec. \ref{sec:mass_distr}).
+%
+Therefore, we expect the optimal $\lambda$
+to decrease with the span $q(T)$,
+hence the simulation length $T$,
+if the distribution has a long tail.
+
 \subsection{Nearest-neighbor updating scheme}

### 2016-04-20 doc/invt.tex
edit abstract

@@
 
-\title{Optimal updating factor in Wang-Landau and metadynamics simulations}
+\title{Optimal updating factor in adaptive flat-distribution-sampling simulations}
 
 \begin{abstract}
-  The Wang-Landau (WL) algorithm and metadynamics
-  are two closely-related techniques
-  in free energy calculations.
-  %
-  Both methods sample a flat distribution
-  along a quantity of interest, $z$,
-  by adaptively constructing
-  \note{``incrementally building up''?}%
-  a bias potential to offset
-  the potential of mean force (PMF).
-  %
-  The two differ by
-  the manner of updating the bias potential:
-  %
-  the WL algorithm updates only
-  the bias potential of
-  the bin containing the current $z$,
-  while metadynamics updates also
-  those of neighboring bins according to
-  a Gaussian window.
-  %
-  The error of the final bias potential,
-  hence that of the PMF,
-  is determined by the updating schedule,
-  or the updating magnitude
-  as a function of simulation time.
-  %
-  In the WL case,
-  the optimal schedule
-  is known to be given by
-  the inverse simulation time.
-  %
-  In this study,
-  we give a method of computing the optimal schedule
-  for a general multiple-bin updating scheme.
-  %
-  We show that the optimal schedule of metadynamics
-  depends on the simulation length
-  and the width of the Gaussian.
-  %is solved from the equation of motion
-  %of a free particle with a position-dependent mass.
+  We present a method of computing the optimal schedule
+  of the updating magnitude
+  for adaptive flat-distribution sampling methods,
+  such as the Wang-Landau (WL) algorithm and metadynamics.
+  %
+  These sampling methods
+  construct a bias potential to offset
+  the potential of mean force
+  by incremental updating loops.
+  %
+  The updating magnitude should decrease over time
+  to reduce the fluctuation of the bias potential,
+  and the optimal schedule optimizes
+  the rate of convergence.
+  %
+  We derive the optimal schedule from
+  an equation of motion
+  of a free particle with a position-dependent mass,
+  %
+  and show that
+  the initial updating magnitude is ideally half of
+  the previous equilibrium value.
   %
   Further,
-  the single-bin updating scheme
-  used in the WL algorithm
-  belongs to a class of multiple-bin bandpass updating schemes
-  that are optimal for asymptotic convergence.
-  %
-  The optimal schedule of these schemes
-  is given by the inverse-time formula,
-  with a shifted origin of time such that
-  the initial updating magnitude is half of
-  the previous equilibrium value.
+  we show that the WL updating scheme
+  belongs to a class of bandpass updating schemes
+  that are optimal for asymptotic convergence,
+  and the optimal schedule of these schemes
+  is given by a modified inverse-time formula.
+  %
 \end{abstract}
@@
 the optimal schedule
-for a general updating scheme,
+for a general adaptive flat-distribution sampling methods,
 including those used in WL and metadynamics simulations.
 %
-The method reproduces the inverse-time formula
-for the WL case,
-and gives the optimal value of the origin of time.
-%
-It also predicts, for metadynamics,
-a more complex schedule,
-which is sensitive to the simulation length
+In this method, the optimal schedule is derived
+from an equation of motion of a free particle,
+with the initial updating magnitude set to
+half of the previous equilibrium value.
+%
+The optimal schedule in the WL case
+is shown to be given by
+the inverse-time formula,
+with the origin of time modified
+to satisfy the above optimal initial condition.
+%
+For metadynamics, however,
+the optimal schedule is more complex,
+and it can be sensitive to the simulation length
 and the width of the Gaussian window.
@@
 Further, we generalize
-the single-bin updating scheme
-used by the WL algorithm
-to a class of multiple-bin updating schemes
+the WL updating scheme
+to a class of bandpass updating schemes
 that are optimal for infinitely long simulations,
-and the inverse-time schedule is optimal for these schemes.
+and the modified inverse-time schedule is optimal for these schemes.
 %
@@
 the index of a small interval, or a bin,
-$(z, z + \delta z)$.
+$(z, z + \Delta z)$.
 %
@@
 the optimal schedule of the updating magnitude
-for a general adaptive FDS simulation.
-%
-Adaptive FDS methods,
-including the WL algorithm and metadynamics,
+for a general adaptive FDS simulation,
+e.g., the WL algorithm or metadynamics.
+%
+Adaptive FDS methods
 effectively sample a flat distribution
-by constructing a bias potential
+by incrementally updating the bias potential
 to offset the PMF.
@@
 
-Each adaptive FDS method defines an updating scheme,
+Each adaptive FDS method is associated with
+an updating scheme,
 or mathematically an updating matrix,
@@
 the eigenvalues of the updating matrix,
-and the integrals of the correlation functions
-the corresponding eigenmodes.
+and the integrals of the autocorrelation functions
+of the eigenmodes.
 %
@@
     The velocity rescaling\cite{bussi2007} thermostat was
-    applied every MD step with the viscocity set to 20.
+    applied every MD step with viscocity 20.
     %

### 2016-04-20 prog/predict.c

@@ -12,7 +12,7 @@
 static void invt_geterr(invtpar_t *m,
     const double *gamma)
 {
-  double T, t0, qT = 0, err1, err2;
+  double T, c0, t0, qT = 0, err0, err1, err2;
   double *lambda;
   intq_t *intq;
 
@@ -30,9 +30,13 @@ static void invt_geterr(invtpar_t *m,
     savewinmat(m->winn, m->win, m->n, m->pbc, m->fnwinmat);
   }
 
+  c0 = m->c;
+  err0 = esterror_invt(T, m->c, m->alpha0, m->t0, m->n,
+      NULL, lambda, gamma);
+
   /* compute the optimal c and the error
    * for the inverse-time formula */
-  m->c = estbestc_invt(T, m->alpha0, m->n, lambda, gamma,
+  m->c = estbestc_invt(T, m->alpha0, 0, m->n, lambda, gamma,
       0, &err1, 0);
 
   /* compute the exact minimal error under the same condition */
@@ -44,8 +48,10 @@ static void invt_geterr(invtpar_t *m,
   t0 = 2 * m->c / m->alpha0;
   intq_save(intq, m->c, t0, m->alpha_resample, m->fnalpha);
 
-  printf("c %g, t0 %g, err %g, sqr %g (invt), %g, sqr %g (exact), %s\n",
-      m->c, t0, err1, err1 * err1, err2, err2 * err2, m->fnalpha);
+  printf("c %g, t0 %g, err %g, sqr %g (invt), "
+         "opt. c %g, err %g, sqr %g (opt. invt), %g, sqr %g (exact), %s\n",
+      c0, t0, err0, err0 * err0,
+      m->c, err1, err1 * err1, err2, err2 * err2, m->fnalpha);
 
   intq_close(intq);
   free(lambda);
@@ -80,7 +86,7 @@ static void invt_scanc(invtpar_t *m,
 
   for ( c = m->cmin; c < m->cmax + 0.001 * m->cdel; c += m->cdel ) {
     /* alpha(t) = c / (t + t0) */
-    err = esterror_invt(T, c, m->alpha0, m->n, NULL,
+    err = esterror_invt(T, c, m->alpha0, 0, m->n, NULL,
         lambda, gamma);
 
     /* final equilibrium value */
@@ -174,7 +180,7 @@ static void invt_scan(invtpar_t *m,
     }
 
     /* find the optimal c for the inverse-time schedule */
[…]

### 2016-04-20 doc/invt.tex
invt.tex: edits

@@
   \arraycolsep=3.6pt\def\arraystretch{1.4}
-  \mathbf Q
+  \mathbf A
   =
@@
 we can further assume that
-the matrices $\mathbf Q$ and $\mathbf w$
+the matrices $\mathbf A$ and $\mathbf w$
 share the eigenvectors, but not the eigenvalues.
 %
-The eigenvalues of $\mathbf Q$ are
+The eigenvalues of $\mathbf A$ are
 $$
@@
     {\mathcal L} \bigl[ q(t)\bigr]
-    \, dt,
-  \notag
-  %\label{eq:error_asym_action}
+    \, dt
+  ,
+%\notag
+\label{eq:error_asym_Lagrangian}
 \end{equation}
@@
 \notag
-%\label{eq:error_asym_Lagrangian}
 \end{align}
@@
 We can characterize the optimal schedule
-geometrically
-through the normalized mass distribution
-along $q$ as
-\begin{equation}
-  m(q)
+geometrically as follows.
+%
+First, we define a mass function
+$$
+  M(Q)
+  =
+  \sqrt{
+    \textstyle\sum_{ k = 2 }^n \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
+  }
+  .
+$$
+and a normalized mass distribution as
+\begin{equation}
+  m(Q)
   =
   \frac{
-    \sqrt{
-      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2(q)
-    }
+    M(Q)
   }
   {
-    \int_0^{ q(T) }
-      \sqrt{
-        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2(q')
-      }
-      \, d q'
+    \int_0^{ q(T) } M(Q') \, d Q'
   }
@@
 such that
-$\int_0^{q(T)} m(q) = 1$.
-%
-As shown in Fig. ,
-the distribution peaks at $q = q(T)$,
-and decreases with the difference $q(T) - q$.
+$\int_0^{q(T)} m(Q) \, dQ = 1$.
+%
+As shown in Fig. \ref{fig:massq},
+the distribution peaks at $Q = 0$,
+and decreases with $Q$.
 %
 According to Eq. \eqref{eq:q_opt},
-the ordinate of the distribution curve at $q$
+the ordinate of the distribution curve
 gives $1/(T\alpha) = 1/(T \dot q)$,
-and the area under the curve in the region $q' < q$
+and the area under the curve in the region $Q' < Q$
 gives $t/T$.
@@
     $\alpha(t)$,
-    via the normalized mass distribution, $m(q)$.
+    via the normalized mass distribution, $m(Q)$.
     %
@@
 \end{figure}
+
+For the asymptotic error, we have,
+from Eqs.
+\eqref{eq:error_asym_Lagrangian}
+and
+\eqref{eq:Lagrangian_const},
+%
+\begin{equation}
+  E_A = c^2 \, T
+  =
+  \frac 1 T
+  \left(
+    \int_0^{ q(T) } M(Q) \, dQ
+  \right)^2
+  .
+\label{eq:error_asym2}
+\end{equation}
+%
+The optimal value of $q(T)$ that minimizes
+the total error satisfies
+\begin{equation}
+  m\bigl( q(T) \bigr)
+  =
+  \frac{1} { T \, \alpha_0 / 2 }
+  ,
+\label{eq:opt_qT}
+\end{equation}
+i.e., the initial updating magnitude
+should be half of $\alpha_0$.
+
+\note{Let $Q_c \equiv q(T)$,
+$$
+\begin{aligned}
+  \frac{
+    \partial E_R
+  }
+  {
+    \partial q(T)
+  }
+  &=
+  -\alpha_0 \, M^2(Q_c)
+  ,
+  \\
+  \frac{
+    \partial E_A
+  }
+  {
+    \partial q(T)
+  }
+  &=
+  \frac 2 T \,
+  M(Q_c) \,
+  \int_0^{ Q_c } M(Q) \, dQ
+  .
+\end{aligned}
+$$
+Then solve $\partial (E_R + E_A) / \partial q(T) = 0$.
+}
 
@@
 
-  \node[draw, text width={width("System and")}]
-    (sys) at (0, 1) {System and\\sampling \\process};
-
-  \node[draw, text width={width("Eigenvalues, ")}]
-    (lambda) at (2.5, 3) {Eigenvalues, \\$\lambda_2, \dots, \lambda_n$};
+  \node[draw, text width={width("sampling")}]
+    (sys) at (0, 1) {System\\and\\sampling\\process};
+
+  \node[draw, text width={width("Eigenvalues,")}]
+    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_2, \dots, \lambda_n$};
 
   \node[draw, text width={width("Integrals, ")}]
-    (gamma) at (2.5, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
-
-  \node[draw, text width={width("Simulation ")}]
-    (T) at (2.5, -0.5) {Simulation \\time, $T$};
-
-  \node[draw, text width={width("Schedule ")}]
-    (alpha) at (5.2, 1) {Optimal schedule, \\ $\alpha(t)$};
-
-  \node[draw, text width={width("Error ")}]
-    (err) at (7, 1) {Error, \\ $\Err$};
-
-  \node[inner sep=0, minimum size=0] (M1) at (3.7, 1) {};
-  \node[inner sep=0, minimum size=0] (M2) at (4.0, 1) {};
-  \node[inner sep=0, minimum size=0] (N2) at (4.0, -0.2) {};
-  \node[inner sep=0, minimum size=0] (N3) at (6.2, -0.2) {};
-  \node[inner sep=0, minimum size=0] (M3) at (6.2, 1) {};
+    (gamma) at (2.2, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
+
+  \node[draw, text width={width("Simulation")}]
+    (T) at (2.2, -0.5) {Simulation\\time, $T$};
+
+  \node[draw, text width={width("Optimal")}]
+    (qT) at (4.7, 1) {Optimal\\ $q(T)$};
+
+  \node[draw, text width={width("Schedule,")}]
+    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};
+
+  \node[draw, text width={width("Error, $\Err$")}]
+    (err) at (7, 0) {Error, $\Err$};
+
+  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};
+
+  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};
+
+  %\node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
+  %\node[inner sep=0, minimum size=0] (R2) at (4.6, 0) {};
+  %\node[inner sep=0, minimum size=0] (R3) at (5.7, 1) {};
 
@@
     (w) edge (gamma)
-    (sys) edge (gamma)
-    (alpha) edge (err);
+    (sys) edge (gamma);
 
@@
     (T)      -| (M1)
-    (M1) -- (alpha);
+    (M1)     -- (qT);
 
   \draw[->, >=latex, thick]
-    (M2) -- (N2) -- (N3) -- (M3);
-
+    (qT) -- (N1)
+    (N1) |- (alpha);
+
+  \draw[->, >=latex, thick]
+    (N1) |- (err);
+
+  \draw[->, >=latex, thick]
+    (M1) edge[bend left=70] (N1);
+    %(R1) |- (R2) -| (R3);
 \end{tikzpicture}
@@
 \item
-Select an initial value of $q(T)$,
-e.g., $\ln(1 + \alpha_0 \, T)$.
+Adjust $q(T)$,
+to satisfy Eq. \eqref{eq:opt_qT}.
 
@@
 Eqs. \eqref{eq:error_split},
-  \eqref{eq:error_asym1} and
-  \eqref{eq:error_res1}.
-
-\item
-Modify the value of $q(T)$ and repeat the previous two steps,
-until the minimal error $E$ is found.
+  \eqref{eq:error_res1},
+  and
+  \eqref{eq:error_asym2}.
 

### 2016-04-21 doc/invt.tex
new routine to compute optimal q(T)
invt.tex: rewrite characterization of the optimal schedule

@@
 the optimal formula for $\alpha(t)$
-might be $C/t$ with a $C$
+might be $c/t$ with a $c$
 different from $1.0$\cite{
@@
   =
-    \textstyle\sum_{ k = 2 }^n
+    \sum_{ k = 2 }^n
       \Gamma_k \, \lambda_k^2
@@
   =
-  c^2
+  \left(
+    \frac C T
+  \right)^2
   ,
-  \quad
-  (c > 0).
   \label{eq:Lagrangian_const}
@@
 %
+for some positive $C$.
+%
 Integrating the square root, we get
@@
 \begin{equation}
+  \int_{ 0 }^{ q(T) }
+    \sqrt{
+      \textstyle\sum_{ k = 2 }^n
+        \Gamma_k \, \lambda_k^2
+        \, u_k^2( q' )
+    }
+    \;
+    d q'
+  =
+  C \, \frac t T
+  ,
+  \label{eq:q_opt}
+\end{equation}
+%
+where $c$ is to be determined by
+the boundary condition at $t = T$ as
+%
+\begin{equation}
+  C =
   \int_{ 0 }^{ q(t) }
@@
     d q'
-  =
-  c \, t
-  ,
-  \label{eq:q_opt}
-\end{equation}
-%
-where $c$ is to be determined by
-the boundary condition at $t = T$ as
-%
-\begin{equation}
-  c =
-  \frac 1 T
-  \int_{ 0 }^{ q(T) }
-    \sqrt{
-      \textstyle\sum_{ k = 2 }^n
-        \Gamma_k \, \lambda_k^2
-        \, u_k^2( q' )
-    }
-    \;
-    d q'
   .
@@
 \subsubsection{\label{sec:mass_distr}
-Characterization of the optimal schedule}
+Characterization of optimal schedules}
 
@@
 %
-First, we define a mass function
+First, we define the mass function as
 $$
@@
 $$
-and a normalized mass distribution as
+and the normalized mass distribution as
 \begin{equation}
@@
   }
+  =
+  \frac{
+    M(Q)
+  }
+  {
+    C
+  }
   ,
@@
 %
-As shown in Fig. \ref{fig:massq},
-the distribution peaks at $Q = 0$,
-and decreases with $Q$.
-%
-According to Eq. \eqref{eq:q_opt},
-the ordinate of the distribution curve
-gives $1/(T\alpha) = 1/(T \dot q)$,
-and the area under the curve in the region $Q' < Q$
-gives $t/T$.
-%
-From this picture, we can draw several conclusions.
+Clearly,
+the distribution decreases monotonically with $Q$.
 
@@
 
-For the asymptotic error, we have,
-from Eqs.
+Then, from Eq. \eqref{eq:q_opt},
+we find that for $Q = q(T) - q(t)$,
+%
+\begin{align}
+  m(Q)
+  &=
+  \frac{ 1 }
+       { T \, \alpha(t) }
+  ,
+\label{eq:mQ_invTa}
+  \\
+  \int_Q^{ q(T) }
+    m(Q') \, dQ'
+  &=
+  \frac t T
+  ,
+\label{eq:intmQ_tT}
+\end{align}
+%
+which means that each point $(t, \alpha(t))$
+on the schedule can be mapped
+to a point $(Q, m(Q))$ on the mass distribution,
+such that the ordinate of the latter curve
+gives $1/(T\alpha)$,
+and the area under the curve in the domain $[Q, q(T)]$
+is equal to $t/T$,
+as shown in Fig. \ref{fig:massq},
+
+This mapping allows us to classify all optimal schedules
+as follows.
+%
+We collect in each class optimal schedules
+computed from simulations with the same function $M(Q)$
+and cutoff value $q(T)$.
+%
+The former is determined by the values of
+$\lambda_k$'s and $\Gamma_k$'s,
+whereas the latter affects the distribution $m(Q)$
+via the normalization constant $C$.
+
+In each class, the updating magnitudes of two schedules
+of lengths $T_1$ and $T_2$
+at corresponding times $t_1/T_1 = t_2/T_2$
+are related by $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
+%
+Further, by Eqs.
 \eqref{eq:error_asym_Lagrangian}
@@
 \eqref{eq:Lagrangian_const},
-%
-\begin{equation}
-  E_A = c^2 \, T
+we have
+%
+\begin{equation}
+  E_A
+  =
+  \frac { C^2 } { T }
   =
@@
   \right)^2
+  ,
+\label{eq:error_asym2}
+\end{equation}
+%
+which means the asymptotic error
+is inversely proportional to the length
+for the same-class schedules,
+and we have the scaling relation
+$T_1 \, E_{A1} = T_2 \, E_{A2}$.
+
+%The optimal schedule has the following scaling property.
+%%
+%For two simulations of lengths $T_1$ and $T_2$,
+%if $q(T_1) = q(T_2)$
+%and they share the same $\lambda_k$'s and $\Gamma_k$'s,
+%the updating magnitudes at the corresponding times,
+%$t_1/T_1 = t_2/T_2$,
+%can be related as $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
+%%
+%In other words,
+%if $\alpha(t_1)$ is
+%an optimal schedule of the simulation of length $T_1$,
+%then
+%$\frac{T_1}{T_2} \alpha\left( \frac{ T_1 } { T_2 } t_2 \right)$
+%is an optimal schedule of the simulation of length $T_2$.
+%%
+%The asymptotic errors are related by
+%$T_1 \, E_{A1} = T_2 \, E_{A2}$.
+
+\subsubsection{\label{sec:optinitalpha}
+  Optimal initial updating magnitude
+}
+
+Let us now consider simulations in which
+the system is initially equilibrated at the same
+constant updating magnitude $\alpha$.
+%
+For these simulations,
+we can adjust $q(T)$
+to minimize the total error.
+%
+From Eqs. \eqref{eq:error_res1}
+and \eqref{eq:error_asym2},
+we find that
+the $q(T)$ that minimizes
+the total error satisfies
+\begin{equation}
+  m\bigl( q(T) \bigr)
+  =
+  \frac{1} { T \, (\alpha_0 / 2) }
   .
-\label{eq:error_asym2}
-\end{equation}
-%
-The optimal value of $q(T)$ that minimizes
-the total error satisfies
-\begin{equation}
-  m\bigl( q(T) \bigr)
-  =
-  \frac{1} { T \, \alpha_0 / 2 }
-  ,
 \label{eq:opt_qT}
 \end{equation}
-i.e., the initial updating magnitude
-should be half of $\alpha_0$.
-
+%
 \note{Let $Q_c \equiv q(T)$,
@@
 }
-
-First,
-the optimal schedule has the following scaling property.
-%
-For two simulations of lengths $T_1$ and $T_2$,
-if $q(T_1) = q(T_2)$
-and they share the same $\lambda_k$'s and $\Gamma_k$'s,
-the updating magnitudes at the corresponding times,
-$t_1/T_1 = t_2/T_2$,
-can be related as $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
-%
-In other words,
-if $\alpha(t_1)$ is
-an optimal schedule of the simulation of length $T_1$,
-then
-$\frac{T_1}{T_2} \alpha\left( \frac{ T_1 } { T_2 } t_2 \right)$
-is an optimal schedule of the simulation of length $T_2$.
-%
-The asymptotic errors are related by
-$T_1 \, E_{A1} = T_2 \, E_{A2}$.
+%
+Then, by Eqs. \eqref{eq:mQ_invTa} and \eqref{eq:intmQ_tT},
+we have
+\begin{equation}
+  \alpha( t = 0 )
+  =
+  \frac{ \alpha_0 }
+       { 2 }
+  ,
+\label{eq:half_alpha0}
+\end{equation}
+%
+i.e., the optimal initial updating magnitude
+is always half of the equilibration value.
+%
+This factor $1/2$
+happens to coincide with the
+reduction factor of updating magnitude
+at stage transitions
+in the original WL algorithm\cite{
+wang2001, wang2001pre}.
+
+Besides,
+for simulations with the same $\alpha_0$,
+Eq. \eqref{eq:opt_qT} shows that
+the optimal $q(T)$ increases with
+the simulation length $T$.
+%
+If $M(Q)$ decays rapidly,
+the distribution $m(Q)$ does not critically
+depend on the cutoff $q(T)$,
+as long as $q(T)$ is sufficiently large.
+%
+So the optimal schedules of these simulations
+approximately satisfy the scaling relations
+described in Sec. \ref{sec:mass_distr}.
+%
+If, however, $M(Q)$ has a long tail
+because of some near-zero $\lambda_k$'s,
+increasing the simulation length $T$
+generally alters the shape of the optimal schedule.
 
@@
 this leads to the optimal value
-of the free parameter, $t_0$.
-%
-%Assuming the system is initially equilibrated
-%under a constant updating magnitude,
-%$\alpha_0$,
-Assuming Eq. \eqref{eq:error_res1} with
-$$
-q(T)
-=
-\int_0^T \frac{ dt } { t + t_0 }
-=
-\ln \frac{ T + t_0 } { t_0 }.
-$$
-%
-we have, from Eqs. \eqref{eq:error_split}
-and \eqref{eq:error_asym_sinc},
-%
-\begin{equation}
-  E
-  =
-  E_R + E_A
-  =
-  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
-        {          ( T + t_0 )^2          }
-  \sum_{ k = 2 }^K \Gamma_k
-  .
-\notag
-%\label{eq:error_sinc0}
-\end{equation}
-%
-The minimum is reached at
+of the free parameter, $t_0$,
+which is, according to Eq. \eqref{eq:half_alpha0}
+%
+%%Assuming the system is initially equilibrated
+%%under a constant updating magnitude,
+%%$\alpha_0$,
+%Assuming Eq. \eqref{eq:error_res1} with
+%$$
+%q(T)
+%=
+%\int_0^T \frac{ dt } { t + t_0 }
+%=
+%\ln \frac{ T + t_0 } { t_0 }.
+%$$
+%%
+%we have, from Eqs. \eqref{eq:error_split}
+%and \eqref{eq:error_asym_sinc},
+%%
+%\begin{equation}
+%  E
+%  =
+%  E_R + E_A
+%  =
+%  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
+%        {          ( T + t_0 )^2          }
+%  \sum_{ k = 2 }^K \Gamma_k
+%  .
+%\notag
+%%\label{eq:error_sinc0}
+%\end{equation}
+%%
+%The minimum is reached at
 %
@@
 in the asymptotic limit.
-
-Note also that at the beginning of the schedule,
-the updating magnitude is
-$\alpha(0^+) = \alpha_0 / 2$,
-half of the previous equilibrium value.
-%
-This factor $1/2$
-happens to coincide with the
-reduction factor of updating magnitude
-at stage transitions
-in the original WL algorithm\cite{
-wang2001, wang2001pre}.
 
@@
 in the global sampling case,
-because of the near zero eigenvalues,
+because of the near-zero eigenvalues,
 as discussed in Appendix \ref{sec:invt_schedule}.
@@
 \item
-Adjust $q(T)$,
-to satisfy Eq. \eqref{eq:opt_qT}.
+Compute $q(T)$ from the root of Eq. \eqref{eq:opt_qT}.
 
@@
 \section{\label{sec:equilerr}
-Comparison to the error of the equilibrium FDS simulation}
+Error of the ideal equilibrium FDS simulation}
 
@@
 
-The error of the equilibrium FDS method
+The error of the equilibrium FDS simulation
 depends on the precision of the histogram.

### 2016-04-21 doc/invt.tex
invt.tex: small edits

@@
 of the free parameter, $t_0$,
-which is, according to Eq. \eqref{eq:half_alpha0}
+which is, according to Eq. \eqref{eq:half_alpha0},
 %
@@
 %
-and equilibrate the system under
-a constant updating magnitude,
-$\alpha_0 = 10^{-4}$.
+and equilibrate the system for $10^7$ steps
+under a constant updating magnitude,
+$\alpha_0 = 10^{-4}$,
+unless specified otherwise.
 %
@@
 
-For the single-bin updating scheme,
-we chose $\alpha_0 = 10^{-4}$
-and ran the simulation for $T = 10^8$ steps.
+For the single-bin updating scheme, we
+%we chose $\alpha_0 = 10^{-4}$ and
+ran the simulation for $T = 10^8$ steps.
 %
@@
 the numerical results of the terminal error agreed well with
-the analytical prediction by Eq. \eqref{eq:error_invt}.
+the analytical prediction.
 %
@@
 differed only by a multiplicative constant,
-which again agreed with the theory:
-%
-with identical eigenvalues,
-$\lambda_2 = \cdots = \lambda_n = 1$,
-the error in Eq. \eqref{eq:error_invt}
-depends on the sampling process only through
-the sum $\sum_{ k = 2 }^n \Gamma_k$,
-which gives rise to the above multiplicative constant.
+also in agreement with the theory
+(cf. Appendix \ref{sec:invt_singlebin}).
+%
 
@@
 %
-We choose $\mu_1 = 0.24$ to the smallest eigenvalue
+We chose $\mu_1 = 0.24$ to make the smallest eigenvalue
 close to zero.
@@
 %
-The normalized error
-and the optimal value of $1/\lambda$,
-depended on the simulation length, $T$,
-in the global sampling case,
-because of the near-zero eigenvalues,
-as discussed in Appendix \ref{sec:invt_schedule}.
-%
-%The optimal $1/\lambda$ gradually increased
-%with simulation length,
-%as discussed in Sec. \ref{sec:optlambda}.
-%
-However, for the local sampling case,
-the $T$-dependence was suppressed
-because of the dominance of large $\Gamma_k$'s
-of the small-$k$ modes.
-%
-The optimal $\lambda$ remained around $1.0$,
-as suggested by Eq. \eqref{eq:lambda_nn_onestep}.
+The optimal value of $1/\lambda$,
+increased with the simulation length, $T$,
+in the global sampling case
+(cf. Appendix \ref{sec:invt_deplength}).
+%
+However, the shift was suppressed
+for the local sampling case,
+with the optimal $\lambda$ remained around $1.0$
+[cf. Eq. \eqref{eq:lambda_nn_onestep}].
 
@@
 %
-The optimal schedule ensures fastest convergence
+The optimal schedule delivers fastest convergence
 of the bias potential.
@@
 an updating scheme,
-or mathematically an updating matrix,
-whose elements $w_{ij}$ specifies the
-relative magnitude of updating the bias potential
-at bin $i$ upon a visit to bin $j$.
+or mathematically an updating matrix.
+%whose elements $w_{ij}$ specifies the
+%relative magnitude of updating the bias potential
+%at bin $i$ upon a visit to bin $j$.
 %
@@
 the optimal schedule from our method
-is given by inverse-time formula,
+is given by the inverse-time formula,
 Eq. \eqref{eq:alpha_invt1}.
-%
-The parameter, $t_0$, can be adjusted
-according to Eq. \eqref{eq:t0_sinc},
-to accommodate the initial error.
-%such that the initial updating magnitude
-%is half of the previous equilibrium value.
 %
@@
 can be sensitive to the simulation length.
+%
+The initial updating magnitude is always
+half of the previous equilibrium value.
+% the value used during equilibration.
 
@@
 of certain eigenmodes, we can construct a
-class of multiple-bin bandpass updating schemes
+class of bandpass updating schemes
 that are also optimal for asymptotic convergence.
@@
 in our model study,
-suggesting that they can be useful
-in high-precision metadynamics simulations.
-
-Strictly speaking, our theory is valid
+and they can be useful
+for high-precision metadynamics simulations.
+%
+However, since the bandpass schemes
+update the bias potential globally,
+they can be more expensive than
+the Gaussian schemes.
+%
+Besides, the performance gain
+is relative small for a shorter simulation
+with a local sampling process such as MD.
+
+Strictly speaking, our method is valid
 only for long simulations.
@@
 Although these assumptions appeared to work well
-on the model system,
+on our test system,
 they may break down for relatively short simulations
@@
 %
-Basically, we extract
-both $\lambda_i$'s and $\Gamma_i$'s
+Basically, we need to extract
+$\lambda_i$'s and $\Gamma_i$'s
 from the input,
@@
   \node[draw, text width={width("Optimal")}]
-    (qT) at (4.7, 1) {Optimal\\ $q(T)$};
+    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};
 
@@
 
-  %\node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
-  %\node[inner sep=0, minimum size=0] (R2) at (4.6, 0) {};
-  %\node[inner sep=0, minimum size=0] (R3) at (5.7, 1) {};
+  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
+  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
+  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};
 
@@
     (T)      -| (M1)
-    (M1)     -- (qT);
+    (M1)     -- (R1)
+    (R1)     |- (qT);
 
   \draw[->, >=latex, thick]
-    (qT) -- (N1)
-    (N1) |- (alpha);
+    (qT)     -| (R3);
+
+  \draw[->, >=latex, thick]
+    (R3) -- (N1) |- (alpha);
 
@@
   \draw[->, >=latex, thick]
-    (M1) edge[bend left=70] (N1);
-    %(R1) |- (R2) -| (R3);
+    %(M1) edge[bend left=70] (N1);
+    (R1) |- (R2) -| (R3);
 \end{tikzpicture}
@@
         \right)^{ 2 \, \nu_k - 1 }
-\right],
+  \right]
+  ,
 \label{eq:error_asym_invt}
@@
 %
-Assuming Eq. \eqref{eq:error_res1}
-for the residual error, and Eq. \eqref{eq:t0_sinc},
+From Eq. \eqref{eq:error_res1},
 we get
@@
   =
+  \frac { \alpha_0 } { 2 }
   \sum_{k = 2}^n
-  \frac{ \Gamma_k \, \nu_k }
-       {        t_0   }
+  \Gamma_k \, \lambda_k \,
   \left(
@@
            { T + t_0 }
-   \right)^{ 2 \, \nu_k },
-\notag
-%\label{eq:error_res_invt}
-\end{equation}
-%
-with
-$
-  q(T)
-  =
-  ( 1 / \lambda )
-  \ln\left(
-    1 + T / t_0
-  \right)
-$
-%
-from integrating Eq. \eqref{eq:alpha_invtlambda}.
+  \right)^{ 2 \, \nu_k }
+  .
+\label{eq:error_res_invt}
+\end{equation}
+%
+The total error is the sum of the two.
+%with
+%$
+%  q(T)
+%  =
+%  ( 1 / \lambda )
+%  \ln\left(
+%    1 + T / t_0
+%  \right)
+%$
+%%
+%from integrating Eq. \eqref{eq:alpha_invtlambda}.
 %
@@
 %
-The total error is
-%
-\begin{align}
-\Err
-&=
-\Err_R + \Err_A
-\notag
-\\
-&=
-\sum_{ k = 2 }^n
-  \frac
-  {
-    \Gamma_k \, \nu_k \,
+%The total error is
+%%
+%\begin{align}
+%\Err
+%&=
+%\Err_R + \Err_A
+%\notag
+%\\
+%&=
+%\sum_{ k = 2 }^n
+%  \frac
+%  {
+%    \Gamma_k \, \nu_k \,
+%    \left[
+%      \nu_k
+%      +
+%      (\nu_k - 1)
+%      \left(
+%        \frac{ t_0 } { T + t_0 }
+%      \right)^{ 2 \, \nu_k - 1 }
+%    \right]
+%  }
+%  {
+%    (T + t_0) \, (2 \, \nu_k - 1)
+%  }
+%  .
+%%\notag
+%\label{eq:error_invt}
+%\end{align}
+
+\subsection{\label{sec:invt_singlebin}
+  Single-bin updating scheme
+}
+
+For the single-bin scheme, we have
+$\lambda_2 = \cdots = \lambda_n = 1$,
+and
+the errors
+%from Eqs. \eqref{eq:error_res_invt}
+%and \eqref{eq:error_asym_invt}
+depend on $\Gamma_k$'s only via the sum
+$\sum_{k = 2}^n \Gamma_k$,
+which serves as a multiplicative constant.
+
+In the asymptotic limit,
+we can focus on the asymptotic error
+and drop the second term in the square brackets
+of Eq. \eqref{eq:error_asym_invt}
+by assuming $\lambda < 2$.
+%
+Then,
+$$
+  E_A
+  =
+  \frac { 1 } { T + t_0 }
+  \frac {             1            }
+        { \lambda \, (2 - \lambda) }
+  \sum_{ k = 2 }^n \Gamma_k
+  ,
+$$
+is minimal at $\lambda = 1$.
+%
+\note{A higher-order correction:
+  $
+  \lambda = 1 -
   \left[
-      \nu_k
-      +
-      (\nu_k - 1)
-      \left(
-        \frac{ t_0 } { T + t_0 }
-      \right)^{ 2 \, \nu_k - 1 }
-    \right]
-  }
-  {
-    (T + t_0) \, (2 \, \nu_k - 1)
-  }
-  .
-\label{eq:error_invt}
-\end{align}
-%
-
-\subsection{\label{sec:deplength}
-Dependence on the simulation length}
-
-The parameter, $\lambda$, is equal to
-the difference
-$$
-  \lambda
-  =
-  \Delta \left( \frac{ 1 } { T \, \alpha } \right)
-  \equiv
-  \left.
-  \frac { 1 } { T \, \alpha(t) }
-  \right|_{t = 0}^{t = T}
-  .
-$$
-which corresponds to the height difference
-of the normalized mass distribution
+    1 - \frac 1 2 \alpha_0 \, (T+t_0)
+  \right] r \, \log r
+  ,
   $
-\left. m(q) \right|_{q = 0}^{q = q(T)}
-$
-in the optimal schedule
-(cf. Sec. \ref{sec:mass_distr}).
-%
-Therefore, we expect the optimal $\lambda$
-to decrease with the span $q(T)$,
-hence the simulation length $T$,
-if the distribution has a long tail.
-
-\subsection{Nearest-neighbor updating scheme}
+  where $r = t_0 / (T + t_0)$.
+}
+
+\subsection{\label{eq:invt_nn}
+Nearest-neighbor updating scheme}
 
@@
 of the optimal value of $\lambda$
-in the large-$n$ limit.
+in the large-$n$ and asymptotic limit.
 %
 In this case,
-we
-drop the second term in the square brackets
-of Eq. \eqref{eq:error_asym_invt},
-which vanishes asymptotically if $2 \, \nu_k > 1$,
-and convert the asymptotic error
-to an integral:
+if $2 \, \nu_k > 1$ for every $k$,
+the asymptotic error is given by the integral
 %$\sum_i \to \frac{2 \, n}{\pi} \int_0^{\pi/2} dp$, and
@@
   =
-\frac{2 \, n}{\pi \, T}
+  \frac{ 2 \, n  }
+       { \pi \, T }
   \int_0^{\pi/2}
@@
          {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-\, dp.
+  \, dp
+  .
 \notag
@@
 
+\subsection{\label{sec:invt_deplength}
+Dependence on the simulation length}
+
+For a finite-length simulation,
+we generally expect the optimal $\lambda$
+that minimizes the total error to
+decrease with the simulation length $T$.
+%
+From Eq. \eqref{eq:alpha_invt}, we have
+$$
+  \lambda
+  =
+  \Delta \left( \frac{ 1 } { T \, \alpha } \right)
+  \equiv
+  \left.
+  \frac { 1 } { T \, \alpha(t) }
+  \right|_{t = 0}^{t = T}
+  .
+$$
+This expression corresponds to the height difference
+of the mass distribution
+$
+\bigl. m(Q) \bigr|_{Q = q(T)}^{Q = 0}
+$
+in the optimal schedule
+(cf. Sec. \ref{sec:mass_distr}).
+%
+With a fixed $\alpha_0$,
+increasing the simulation length $T$
+also increases the optimal span $q(T)$ of the distribution
+(cf. Sec. \ref{sec:optinitalpha}),
+and thus reduces the height difference.
+%
+This effect is significant
+if the distribution $m(Q)$ has a long tail,
+or if some eigenvalues $\lambda_k$'s are near zero.
+%
+It is often suppressed in local sampling processes,
+because the eigenmodes with large $\lambda_k$'s
+are accompanied by large $\Gamma_k$'s,
+which shortens the distribution tail.
+
 \bibliography{simul}

### 2016-04-22 doc/invt.tex
small changes

@@
 
-We tested the above theory on a model system,
-which is to be described in Sec. \ref{sec:results_system}.
+We tested the above theory on a model system
+described in Sec. \ref{sec:results_system}.
 %
 First, we verify the method of computing errors
-using an analytically simple schedule
+%using an inverse-time schedule
 in Sec. \ref{sec:results_invt}.
@@
 Finally, we compare the Gaussian and
-bandpass schemes of different widths
+bandpass updating schemes %of different widths
 in Sec. \ref{sec:results_cmpschemes}.
@@
 
-We first studied the inverse-time schedule,
+We first tested the theory of computing errors
+on the simple inverse-time schedule,
 %
@@
 
-As an example of the multiple-bin schemes,
-we chose the nearest-neighbor updating scheme
-described in Sec. \ref{sec:nnscheme}.
-%
-We chose $\mu_1 = 0.24$ to make the smallest eigenvalue
+As an example of multiple-bin schemes,
+we tested the nearest-neighbor updating scheme
+described in Sec. \ref{sec:nnscheme},
+%
+with
+$\mu_1 = 0.24$ to make the smallest eigenvalue
 close to zero.

### 2016-04-22 doc/invt.tex
add new figure errinvt.gp

@@
 
-\title{Optimal updating factor in adaptive flat-distribution-sampling simulations}
+\title{Optimal updating factor in adaptive flat-distribution simulations}
 
@@
 \subsubsection{\label{sec:nnscheme}
-Nearest-neighbor updating scheme}
+Triple-bin updating scheme}
 
@@
 
-We first tested the theory of computing errors
+We tested the theory of computing errors
 on the simple inverse-time schedule,
@@
 where $\lambda$ is a free parameter,
-and $t_0 = 2/(\lambda \, \alpha_0)$.
+and $t_0 = 2/\alpha_0$.
 %
@@
 
-%\subsubsection{Single-bin (WL) updating scheme}
-
-For the single-bin updating scheme, we
-%we chose $\alpha_0 = 10^{-4}$ and
-ran the simulation for $T = 10^8$ steps.
-%
-As shown in Fig. \ref{fig:errsbin},
-the numerical results of the terminal error agreed well with
-the analytical prediction.
-%
-For both global and local sampling processes,
-the optimal $\lambda$ occurred around $1.0$,
-verifying the optimality of Eq. \eqref{eq:alpha_invt1}.
-%
-%Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
-%represents the relative updating magnitude.
-%%
-%Figure \ref{fig:errsbin} also shows
-%an asymmetry of the error curve.
-%A smaller than optimal value of
-%the updating magnitude, $1/\lambda$,
-%tends to increase the error more rapidly
-%than a larger than optimal value,
-%suggesting that the over-updating
-%is generally preferable to the under-updating
-%when the optimal value is unknown.
-%
-The two error functions
-differed only by a multiplicative constant,
-also in agreement with the theory
-(cf. Appendix \ref{sec:invt_singlebin}).
-%
+We performed FDS simulations
+using the single-bin updating scheme
+and the triple-bin updating scheme with $\mu_1 = 0.24$.
+%
+For each value of $\lambda$, we ran $1000$ independent simulations
+of $10^8$ steps, and averaged the results.
+%
+As shown in Fig. \ref{fig:errinvt},
+the errors from the simulations
+agreed well those from the theory.
 
@@
   \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsbin.pdf}
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/errinvt.pdf}
   }
   \caption{
-    \label{fig:errsbin}
-    Error, $E$, versus the
-    relative updating magnitude,
-    %proportionality constant,
-    $1/\lambda$,
+    \label{fig:errinvt}
+    Error $E$ versus $\lambda$
     in Eq. \eqref{eq:alpha_invtlambda}
-    for the single-bin (WL) scheme.
-    %
-    The results were averaged over $1000$ independent runs.
-    %
-    \note{The figure was produced by \texttt{doc/fig/errsbin.gp}.
-      The data were produced by the script \texttt{data/invt2run.py},
-      and saved under \texttt{data/singlebin}
+    for the inverse-time schedule.
+    %
+    The points are results from averaging over $1000$ independent runs;
+    the curves are theoretical predictions.
+    %
+    \note{The figure was produced by \texttt{doc/fig/errinvt.gp}.
+      The data were produced by the script \texttt{data/invtcscan.py},
+      and saved under \texttt{data/invt}
       (see the \texttt{Makefile} there).
@@
 
-%\subsubsection{Nearest-neighbor updating scheme}
-
-As an example of multiple-bin schemes,
-we tested the nearest-neighbor updating scheme
-described in Sec. \ref{sec:nnscheme},
-%
-with
-$\mu_1 = 0.24$ to make the smallest eigenvalue
-close to zero.
-%
-Again, simulation and theory agreed well,
-as shown in Fig. \ref{fig:errnnbr}.
-%
-The optimal value of $1/\lambda$,
-increased with the simulation length, $T$,
-in the global sampling case
-(cf. Appendix \ref{sec:invt_deplength}).
-%
-However, the shift was suppressed
-for the local sampling case,
-with the optimal $\lambda$ remained around $1.0$
-[cf. Eq. \eqref{eq:lambda_nn_onestep}].
-
-%To understand this dependence,
-%we observe that for shorter simulations,
-%long-wavelength modes
-%(with larger $\lambda_k$ and smaller $k$),
-%dominate the error function in Eq. \eqref{eq:error_invt}
-%because of the prefactor $\lambda_k \, \Gamma_k$,
-%which is inherited from
-%the equilibrium error under a fixed updating magnitude,
-%Eq. \eqref{eq:error_eql}.
-%%
-%For longer simulations,
-%short-wavelength modes
-%(with smaller $\lambda_k$ and larger $k$)
-%become more important,
-%for $\lambda_k$ also gives the decay rate
-%of the $k$th mode.
-%%and errors in smaller $\lambda_k$ modes
-%%are harder to damp out.
-%%
-%This drives the optimal $\lambda$ to shift toward
-%a smaller $\lambda_k$ value
-%as the simulation lengthens.
-%%
-%Note also that while increasing $\lambda$
-%also increases the error of the long-wavelength modes,
-%this rate of increase is much smaller,
-%as shown in the $\nu_k > 1$ branch
-%in Fig. \ref{fig:err_component}.
-%%
-%The shift of $\lambda$, however, is less noticeable
-%under the local sampling scheme,
-%because longer-wavelength modes more heavily
-%weighted by much larger values of $\Gamma_k$.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/errnnbr.pdf}
-  }
-  \caption{
-    \label{fig:errnnbr}
-    Normalized error, $(T + t_0) \, E$,
-    versus the %proportionality constant,
-    relative updating magnitude,
-    $1/\lambda$,
-    in Eq. \eqref{eq:alpha_invtlambda}
-    for the nearest-neighbor updating scheme with $\mu_1 = 0.24$.
-    %
-    The results were averaged over $1000$ independent runs.
-    %
-    \note{This figure was produced by \texttt{doc/fig/errnnbr.gp}.
-      The data were produced by \texttt{data/invt2run.py},
-      saved under \texttt{data/nb0.24}
-      (see the \texttt{Makefile} there).
-    }%
-  }
-\end{center}
-\end{figure}
-
-%\subsection{Gaussian (metadynamics) updating scheme}
-%
-%
-%The error of Gaussian updating scheme is similar to
-%that of the nearest neighbor scheme,
-%as shown in Fig \ref{fig:err_sig5}.
-%%
-%The dependency of the simulation length is more prominent.
-%%
-%Even for local sampling,
-%there is now a visible shift of the optimal $\lambda$.
-%
-%
-%\begin{figure}[h]
-%\begin{center}
-%  \makebox[\linewidth][c]{
-%    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsig5.pdf}
-%  }
-%  \caption{
-%    \label{fig:err_sig5}
-%    Normalized error, $(T + t_0) \, E$,
-%    versus the %proportionality constant,
-%    relative updating magnitude,
-%    $1/\lambda$,
-%    in Eq. \eqref{eq:alpha_invtlambda}
-%    for the Gaussian updating scheme with $\sigma = 5$ bins.
-%    %
-%    The results were averaged over $1000$ independent runs.
-%  }
-%\end{center}
-%\end{figure}
-%
+For the single-bin updating scheme,
+the optimal $\lambda$ occurred around $1.0$
+for both global and local sampling processes.
+%
+The sampling process affected the error function
+only by a multiplicative constant,
+%The error functions of the two sampling processes
+%differed only by a multiplicative constant,
+in agreement with the discussion in
+Appendix \ref{sec:invt_singlebin}.
+%
+For the triple-bin updating scheme,
+the optimal $\lambda$ was less than $1.0$
+for the global sampling process,
+although it remained around $1.0$
+for the local sampling process,
+as discussed in Appendix \ref{sec:invt_nn}.
+%
+Thus, the inverse-time formula,
+Eq. \eqref{eq:alpha_invt1},
+is optimal for the single-bin updating scheme,
+but not necessarily so for a general updating scheme.
 
@@
 \subsection{\label{sec:results_cmpschemes}
-Comparison of updating schemes}
+Optimal updating schemes}
 
@@
   %For example,
-  %the nearest-neighbor updating scheme
+  %the triple-bin updating scheme
   %is unstable if Eq. \eqref{eq:nn_stable} is violated.
@@
 and
-the errors
+$\Gamma_k$'s affect the errors
 %from Eqs. \eqref{eq:error_res_invt}
 %and \eqref{eq:error_asym_invt}
-depend on $\Gamma_k$'s only via the sum
-$\sum_{k = 2}^n \Gamma_k$,
-which serves as a multiplicative constant.
+only via the sum, $\sum_{k = 2}^n \Gamma_k$,
+as a multiplicative constant.
 
@@
 
-\subsection{\label{eq:invt_nn}
-Nearest-neighbor updating scheme}
-
-For the nearest-neighbor updating scheme
-introduced in Sec. \ref{sec:nnscheme},
+\subsection{\label{sec:invt_nn}
+Triple-bin updating scheme}
+
+For the triple-bin updating scheme,
+%introduced in Sec. \ref{sec:nnscheme},
 we can find an analytical expression
@@
 which
-increases with the magnitude of $\mu_1$
-no matter its sign,
-and the minimal value, obtained at $\mu_1 = 0$,
-corresponds to the single-bin scheme.
+is minimal at $\mu_1 = 0$, the single-bin case.
+%increases with the magnitude of $\mu_1$
+%no matter its sign,
+%and the minimal value, obtained at $\mu_1 = 0$,
+%corresponds to the single-bin scheme.
 %
@@
 that $\Gamma(p) = \cot^2 p$
-is peaked around $p \approx 0$,
+peaks around $p \approx 0$,
 %
@@
 %
-\begin{equation}
+$
+%\begin{equation}
   \lambda \approx 1.
-\label{eq:lambda_nn_onestep}
-\end{equation}
-%
-
-However, Eq. \eqref{eq:lambda_nn_onestep}
-is an approximation good only
+%\label{eq:lambda_nn_onestep}
+%\end{equation}
+$
+%
+
+\note{
+However, the above result %Eq. \eqref{eq:lambda_nn_onestep}
+is only good
 for sufficiently short simulations,
@@
 $\min \lambda_k = 1 - 4 \, \mu_1$.
-
-\subsection{\label{sec:invt_deplength}
-Dependence on the simulation length}
-
-For a finite-length simulation,
-we generally expect the optimal $\lambda$
-that minimizes the total error to
-decrease with the simulation length $T$.
-%
-From Eq. \eqref{eq:alpha_invt}, we have
-$$
-  \lambda
-  =
-  \Delta \left( \frac{ 1 } { T \, \alpha } \right)
-  \equiv
-  \left.
-  \frac { 1 } { T \, \alpha(t) }
-  \right|_{t = 0}^{t = T}
-  .
-$$
-This expression corresponds to the height difference
-of the mass distribution
-$
-\bigl. m(Q) \bigr|_{Q = q(T)}^{Q = 0}
-$
-in the optimal schedule
-(cf. Sec. \ref{sec:mass_distr}).
-%
-With a fixed $\alpha_0$,
-increasing the simulation length $T$
-also increases the optimal span $q(T)$ of the distribution
-(cf. Sec. \ref{sec:optinitalpha}),
-and thus reduces the height difference.
-%
-This effect is significant
-if the distribution $m(Q)$ has a long tail,
-or if some eigenvalues $\lambda_k$'s are near zero.
-%
-It is often suppressed in local sampling processes,
-because the eigenmodes with large $\lambda_k$'s
-are accompanied by large $\Gamma_k$'s,
-which shortens the distribution tail.
+}
 

### 2016-04-22 prog/predict.c

@@ -9,12 +9,43 @@
 
 
 
+static int save_xerr(invtpar_t *m, const char *fn,
+    const double *xerri, const double *xerrf, const double *xerrf2,
+    const double *lambda)
+{
+  FILE *fp;
+  int i, n = m->n;
+
+  if ( (fp = fopen(fn, "w")) == NULL ) {
+    fprintf(stderr, "cannot write %s\n", fn);
+    return -1;
+  }
+
+  fprintf(fp, "# %d\n", n);
+  for ( i = 1; i < n; i++ ) {
+    fprintf(fp, "%4d %18.8e %18.8e %18.8e %10.8f\n",
+        i, xerri[i], xerrf[i], xerrf2[i],
+        lambda[i]);
+  }
+
+  fclose(fp);
+  return 0;
+}
+
+
+
 static void invt_geterr(invtpar_t *m,
     const double *gamma)
 {
+  int n = m->n;
   double T, c0, t0, qT, inita, err0, err1, err2;
   double *lambda;
+  double *xerri, *xerrf, *xerrf_invt;
   intq_t *intq;
+
+  xnew(xerri, n);
+  xnew(xerrf, n);
+  xnew(xerrf_invt, n);
 
   T = (double) m->nsteps;
 
@@ -30,9 +61,12 @@ static void invt_geterr(invtpar_t *m,
     savewinmat(m->winn, m->win, m->n, m->pbc, m->fnwinmat);
   }
 
+  /* initial errors */
+  esterror_eql(m->alpha0, m->n, xerri, lambda, gamma);
+
   c0 = m->c;
   err0 = esterror_invt(T, m->c, m->alpha0, m->t0, m->n,
-      NULL, lambda, gamma);
+      xerrf_invt, lambda, gamma);
 
   /* compute the optimal c and the error
    * for the inverse-time formula */
@@ -42,7 +76,7 @@ static void invt_geterr(invtpar_t *m,
   /* compute the exact minimal error under the same condition */
   qT = m->qT;
   err2 = esterror_opt(T, m->alpha0, m->initalpha, &qT, m->qprec,
-      m->alpha_nint, &intq, m->n, NULL,
+      m->alpha_nint, &intq, m->n, xerrf,
       lambda, gamma, m->verbose);
   inita = intq_getinita(intq);
 
@@ -57,8 +91,13 @@ static void invt_geterr(invtpar_t *m,
       m->c, err1, err1 * err1,
[…]

### 2016-04-22 doc/invt.tex
invt.tex: move Gamma values to Appendix

@@
 the underlying sampling process
-affects the error only through the set of numbers, $\Gamma_i$'s.
-
-\subsubsection{\label{sec:Gamma}
-Integrals of the autocorrelation functions}
-
-We can evaluate the integrals of the autocorrelation functions,
-$\Gamma_k$'s,
-in two special cases.
-%
-First, if the sampling is perfect,
-Eq. \eqref{eq:kappa_delta} is exact, and
-%
-\begin{equation}
-  \Gamma_k = 1.
-\label{eq:Gamma_perfect}
-\end{equation}
-%
-To see this, we note that
-the bin index $i$ is an independent random variable
-at each time step $t$ in this case, and
-%
-\begin{align}
-  \left\langle
-    \zeta_i(t) \, \zeta_j(t')
-  \right\rangle
-  &=
-  \langle h_i(t) \, h_j(t') \rangle
-  -
-  \langle h_i(t) \rangle
-  \langle h_j(t') \rangle
-  \notag
-  \\
-  &=
-  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \,
-  \delta_{t, t'}
-  ,
-\label{eq:zz_perfect}
-\end{align}
-%
-where $\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
-in the continuous limit.
-%
-\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
-  with $t \ne t'$ (going back to the discrete time case),
-  $$
-  \begin{aligned}
-  \left\langle
-    \zeta_i(t) \, \zeta_j(t')
-  \right\rangle
-  &=
-  \left\langle
-    \zeta_i(t)
-  \right\rangle
-  \left\langle
-    \zeta_j(t')
-  \right\rangle
-  \\
-  &=
-  \bigl\langle h_i(t)  - \langle h_i(t)  \rangle \bigr\rangle
-  \bigl\langle h_j(t') - \langle h_j(t') \rangle \bigr\rangle
-  =
-  0.
-  \end{aligned}
-  $$
-  Next, for $t = t'$, we have
-  $$
-  \zeta_i(t)
-  = h_i(t) - \langle h_i(t) \rangle
-  = \delta_{i, i(t)} - p_i,
-  $$
-  and
-  $$
-  \begin{aligned}
-  \left\langle
-    \zeta_i(t) \, \zeta_j(t)
-  \right\rangle
-  &=
-  \left\langle
-    \delta_{i, i(t)} \, \delta_{j, i(t)}
-  \right\rangle
-  -
-  p_i \, p_j
-  \\
-  &=
-  \left\langle
-    \delta_{i, i(t)}
-  \right\rangle
-  \delta_{i, j}
-  -
-  p_i \, p_j
-  =
-  p_i \, \delta_{ij}
-  - p_i \, p_j.
-  \end{aligned}
-  $$
-  %
-  The average
-  $\left\langle \delta_{i, i(t)} \right\rangle$ is $p_i$,
-  because as $i(t)$ goes through $1$ to $n$,
-  $\delta_{i, i(t)}$ gives $1$
-  if $i(t)$ matches $i$ with probability $p_i$,
-  or $0$ otherwise.
-}
-%
-Thus,
-for $k \ge 2$, we get
-%
-\begin{equation}
-  \kappa_k(t - t')
-  =
-  \sum_{i,j = 1}^n
-  \frac{ \phi_{ki} } { p_i }
-  \frac{ \phi_{kj} } { p_j }
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  %
-  = \delta_{t, t'},
-\notag
-%\label{eq:kappa_perfect}
-\end{equation}
-%
-where we have used
-Eqs. \eqref{eq:eig_orthonormal_rows}-\eqref{eq:ortho1},
-\eqref{eq:eta_def},
-and
-\eqref{eq:zz_perfect}.
-%
-\note{Proof.%Eq. \eqref{eq:kappa_perfect},
-  $$
-  \begin{aligned}
-  \kappa_k(t - t')
-  &=
-  \sum_{i,j = 1}^n
-  \frac{ \phi_{ki} } { p_i }
-  \frac{ \phi_{kj} } { p_j }
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  \\
-  %
-  &=
-  \sum_{i,j = 1}^n
-  \frac{ \phi_{ki} } { p_i }
-  \frac{ \phi_{kj} } { p_j }
-  \left(
-    p_i \, \delta_{i, j} - p_i \, p_j
-  \right)
-  \, \delta_{t, t'}
-  \\
-  &=
-  \left(
-    \sum_{j = 1}^n
-    \frac{ \phi_{kj}^2 } { p_j }
-    -
-    \sum_{i = 1}^n \phi_{ki}
-    \sum_{j = 1}^n \phi_{kj}
-  \right)
-  \, \delta_{t, t'}
-  \\
-  &=
-  (1 - 0 \cdot 0) \, \delta_{t, t'}.
-  \end{aligned}
-  $$
-  where,
-  we have used
-  Eq. \eqref{eq:eta_def}
-  on the first line,
-  %
-  Eq. \eqref{eq:zz_perfect}
-  on the second line,
-  and
-  Eq. \eqref{eq:eig_orthonormal_rows}
-  on the last line.
-}
-
-Another special case is the
-one-dimensional one-step process\cite{vankampen},
-which models a local sampling process.
-In each step, the system hops to
-one of the two neighboring bins
-with equal probability $1/2$.
-%
-The transition matrix is
-%
-\begin{equation}
-  \arraycolsep=3.6pt\def\arraystretch{1.4}
-  \mathbf A
-  =
-  \left(
-    \begin{array}{cccccccc}
-      \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
-      \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
-      \vdots & &  & & \vdots \\
-      \frac {1 - \sigma} 2 & \dots &  & \frac 1 2 & \frac \sigma 2
-    \end{array}
-  \right),
-\notag
-%\label{eq:T_nn}
-\end{equation}
-%
-where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
-%
-For many updating schemes
-(cf. Sec. \ref{sec:band-matrix}),
-we can further assume that
-the matrices $\mathbf A$ and $\mathbf w$
-share the eigenvectors, but not the eigenvalues.
-%
-The eigenvalues of $\mathbf A$ are
-$$
-\Lambda_k
-=
-1 -
-2 \sin^2 \frac{ (k - 1) \, g \, \pi }
-              {       2 \, n        }
-,
-$$
-where
-$g = 2$ for a periodic variable, or $1$ otherwise.
-%Note that these $\Lambda_k$'s are
-%not to be confused with the eigenvalues of $\mathbf w$,
-%$\lambda_k$'s.
-%
-In this case,
-the autocorrelation function $\kappa_k(t)$
-decays as $\kappa_k(0) \, \Lambda_k^t$,
-with $\kappa_k(0) = 1$ being the same as
-the value in perfect sampling.
-%
-Thus,
-%$\Gamma_k$ is roughly twice the autocorrelation time, and
-from Eq. \eqref{eq:Gamma_sum}, we get
-%
-\begin{equation}
-  \Gamma_k
-  =
-  1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
-  =
-  \frac{ 1 + \Lambda_k }
-       { 1 - \Lambda_k }
-  =
-  \cot^2 \left[
-    \frac{ (k - 1) \, g \, \pi }
-         { 2 \, n }
-  \right]
-  .
-\label{eq:Gamma_onestep}
-\end{equation}
-%
+affects the error only through the set of numbers,
+$\Gamma_k$'s.
+%
+In Appendix \ref{sec:Gamma},
+we evaluate $\Gamma_k$'s in two special cases,
+and show how to measure them
+from an adaptive FDS simulation.
 
@@
 
-\section{\label{sec:program}
-Program of computing the optimal schedule and error
-}
-
-Here we give a program of computing
-the optimal schedule and the error
-of a given adaptive FDS simulation.
-%
-Basically, we need to extract
-$\lambda_i$'s and $\Gamma_i$'s
-from the input,
-and then feed them to Eq. \eqref{eq:q_opt}
-for the schedule.
-%
-The dependence of variables are shown in Fig. \ref{fig:vardep}.
-
-\begin{figure}[h]
-\begin{tikzpicture}
-  \node[draw, text width={width("Updating ")}]
-    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
-
-  \node[draw, text width={width("sampling")}]
-    (sys) at (0, 1) {System\\and\\sampling\\process};
-
-  \node[draw, text width={width("Eigenvalues,")}]
-    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_2, \dots, \lambda_n$};
-
-  \node[draw, text width={width("Integrals, ")}]
-    (gamma) at (2.2, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
-
-  \node[draw, text width={width("Simulation")}]
-    (T) at (2.2, -0.5) {Simulation\\time, $T$};
-
-  \node[draw, text width={width("Optimal")}]
-    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};
-
-  \node[draw, text width={width("Schedule,")}]
-    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};
-
-  \node[draw, text width={width("Error, $\Err$")}]
-    (err) at (7, 0) {Error, $\Err$};
-
-  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};
-
-  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};
-
-  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
-  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
-  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};
-
-  \draw[->, >=latex, thick]
-    (w) edge (lambda)
-    (w) edge (gamma)
-    (sys) edge (gamma);
-
-  \draw[->, >=latex, thick]
-    (lambda) -| (M1)
-    (gamma)  -- (M1)
-    (T)      -| (M1)
-    (M1)     -- (R1)
-    (R1)     |- (qT);
-
-  \draw[->, >=latex, thick]
-    (qT)     -| (R3);
-
-  \draw[->, >=latex, thick]
-    (R3) -- (N1) |- (alpha);
-
-  \draw[->, >=latex, thick]
-    (N1) |- (err);
-
-  \draw[->, >=latex, thick]
-    %(M1) edge[bend left=70] (N1);
-    (R1) |- (R2) -| (R3);
-\end{tikzpicture}
-\caption{
-  \label{fig:vardep}
-  Dependence of variables in computing
-  the optimal schedule and the error.
-}
-\end{figure}
-
-\subsection{\label{sec:measure_gamma}
-Measurement of the integrals
-of the autocorrelation functions}
+\section{\label{sec:Gamma}
+Integrals of the autocorrelation functions}
+
+Here, we evaluate the integrals of
+the autocorrelation functions,
+$\Gamma_k$'s,
+in two special cases
+in Secs. \ref{sec:Gamma_perfect}
+and \ref{sec:Gamma_onestep},
+and describe a general method of measuring them
+in Sec. \ref{sec:Gamma_measure}.
+
+\subsection{\label{sec:Gamma_perfect}
+Perfect sampling}
+
+First, if the sampling is perfect,
+Eq. \eqref{eq:kappa_delta} is exact, and
+%
+\begin{equation}
+  \Gamma_k = 1.
+\label{eq:Gamma_perfect}
+\end{equation}
+%
+To see this, we note that
+the bin index $i$ is an independent random variable
+at each time step $t$ in this case, and
+%
+\begin{align}
+  \left\langle
+    \zeta_i(t) \, \zeta_j(t')
+  \right\rangle
+  &=
+  \langle h_i(t) \, h_j(t') \rangle
+  -
+  \langle h_i(t) \rangle
+  \langle h_j(t') \rangle
+  \notag
+  \\
+  &=
+  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \,
+  \delta_{t, t'}
+  ,
+\label{eq:zz_perfect}
+\end{align}
+%
+where $\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
+in the continuous limit.
+%
+\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
+  with $t \ne t'$ (going back to the discrete time case),
+  $$
+  \begin{aligned}
+  \left\langle
+    \zeta_i(t) \, \zeta_j(t')
+  \right\rangle
+  &=
+  \left\langle
+    \zeta_i(t)
+  \right\rangle
+  \left\langle
+    \zeta_j(t')
+  \right\rangle
+  \\
+  &=
+  \bigl\langle h_i(t)  - \langle h_i(t)  \rangle \bigr\rangle
+  \bigl\langle h_j(t') - \langle h_j(t') \rangle \bigr\rangle
+  =
+  0.
+  \end{aligned}
+  $$
+  Next, for $t = t'$, we have
+  $$
+  \zeta_i(t)
+  = h_i(t) - \langle h_i(t) \rangle
+  = \delta_{i, i(t)} - p_i,
+  $$
+  and
+  $$
+  \begin{aligned}
+  \left\langle
+    \zeta_i(t) \, \zeta_j(t)
+  \right\rangle
+  &=
+  \left\langle
+    \delta_{i, i(t)} \, \delta_{j, i(t)}
+  \right\rangle
+  -
+  p_i \, p_j
+  \\
+  &=
+  \left\langle
+    \delta_{i, i(t)}
+  \right\rangle
+  \delta_{i, j}
+  -
+  p_i \, p_j
+  =
+  p_i \, \delta_{ij}
+  - p_i \, p_j.
+  \end{aligned}
+  $$
+  %
+  The average
+  $\left\langle \delta_{i, i(t)} \right\rangle$ is $p_i$,
+  because as $i(t)$ goes through $1$ to $n$,
+  $\delta_{i, i(t)}$ gives $1$
+  if $i(t)$ matches $i$ with probability $p_i$,
+  or $0$ otherwise.
+}
+%
+Thus,
+for $k \ge 2$, we get
+%
+\begin{equation}
+  \kappa_k(t - t')
+  =
+  \sum_{i,j = 1}^n
+  \frac{ \phi_{ki} } { p_i }
+  \frac{ \phi_{kj} } { p_j }
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  %
+  = \delta_{t, t'},
+\notag
+%\label{eq:kappa_perfect}
+\end{equation}
+%
+where we have used
+Eqs. \eqref{eq:eig_orthonormal_rows}-\eqref{eq:ortho1},
+\eqref{eq:eta_def},
+and
+\eqref{eq:zz_perfect}.
+%
+\note{Proof.%Eq. \eqref{eq:kappa_perfect},
+  $$
+  \begin{aligned}
+  \kappa_k(t - t')
+  &=
+  \sum_{i,j = 1}^n
+  \frac{ \phi_{ki} } { p_i }
+  \frac{ \phi_{kj} } { p_j }
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  \\
+  %
+  &=
+  \sum_{i,j = 1}^n
+  \frac{ \phi_{ki} } { p_i }
+  \frac{ \phi_{kj} } { p_j }
+  \left(
+    p_i \, \delta_{i, j} - p_i \, p_j
+  \right)
+  \, \delta_{t, t'}
+  \\
+  &=
+  \left(
+    \sum_{j = 1}^n
+    \frac{ \phi_{kj}^2 } { p_j }
+    -
+    \sum_{i = 1}^n \phi_{ki}
+    \sum_{j = 1}^n \phi_{kj}
+  \right)
+  \, \delta_{t, t'}
+  \\
+  &=
+  (1 - 0 \cdot 0) \, \delta_{t, t'}.
+  \end{aligned}
+  $$
+  where,
+  we have used
+  Eq. \eqref{eq:eta_def}
+  on the first line,
+  %
+  Eq. \eqref{eq:zz_perfect}
+  on the second line,
+  and
+  Eq. \eqref{eq:eig_orthonormal_rows}
+  on the last line.
+}
+
+\subsection{\label{sec:Gamma_onestep}
+One-step process}
+
+Another special case is the
+one-dimensional one-step process\cite{vankampen},
+which models a local sampling process.
+In each step, the system hops to
+one of the two neighboring bins
+with equal probability $1/2$.
+%
+The transition matrix is
+%
+\begin{equation}
+  \arraycolsep=3.6pt\def\arraystretch{1.4}
+  \mathbf A
+  =
+  \left(
+    \begin{array}{cccccccc}
+      \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
+      \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
+      \vdots & &  & & \vdots \\
+      \frac {1 - \sigma} 2 & \dots &  & \frac 1 2 & \frac \sigma 2
+    \end{array}
+  \right),
+\notag
+%\label{eq:T_nn}
+\end{equation}
+%
+where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
+%
+For many updating schemes
+(cf. Sec. \ref{sec:band-matrix}),
+we can further assume that
+the matrices $\mathbf A$ and $\mathbf w$
+share the eigenvectors, but not the eigenvalues.
+%
+The eigenvalues of $\mathbf A$ are
+$$
+\Lambda_k
+=
+1 -
+2 \sin^2 \frac{ (k - 1) \, g \, \pi }
+              {       2 \, n        }
+,
+$$
+where
+$g = 2$ for a periodic variable, or $1$ otherwise.
+%Note that these $\Lambda_k$'s are
+%not to be confused with the eigenvalues of $\mathbf w$,
+%$\lambda_k$'s.
+%
+In this case,
+the autocorrelation function $\kappa_k(t)$
+decays as $\kappa_k(0) \, \Lambda_k^t$,
+with $\kappa_k(0) = 1$ being the same as
+the value in perfect sampling.
+%
+Thus,
+%$\Gamma_k$ is roughly twice the autocorrelation time, and
+from Eq. \eqref{eq:Gamma_sum}, we get
+%
+\begin{equation}
+  \Gamma_k
+  =
+  1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
+  =
+  \frac{ 1 + \Lambda_k }
+       { 1 - \Lambda_k }
+  =
+  \cot^2 \left[
+    \frac{ (k - 1) \, g \, \pi }
+         { 2 \, n }
+  \right]
+  .
+\label{eq:Gamma_onestep}
+\end{equation}
+
+\subsection{\label{sec:Gamma_measure}
+Measurement via an adaptive FDS simulation
+}
 
@@
 %
-While the updating scheme alone
-determines the eigenmodes and $\lambda_k$'s,
-the values of $\Gamma_k$'s depend also on the system
-and the underlying sampling process (MC or MD).
-%
-%Below is a technique of measuring $\Gamma_k$'s.
-%
-%The idea is to run
-To measure $\Gamma_k$'s, we can run
+We can also
+use Eq. \eqref{eq:y2_eql}.
+measure $\Gamma_k$'s in
 a simulation with a constant schedule,
-$\alpha(t) = \alpha_0$,
-and then use Eq. \eqref{eq:y2_eql}.
+$\alpha(t) = \alpha_0$.
 %
@@
 
-\subsection{\label{sec:program_list}
-Program
-}
-
+\section{\label{sec:program}
+Program of computing the optimal schedule and error
+}
+
+Here we give a program of computing
+the optimal schedule and the error
+of a given adaptive FDS simulation.
+%
+Basically, we need to extract
+$\lambda_i$'s and $\Gamma_i$'s
+from the input,
+and then feed them to Eq. \eqref{eq:q_opt}
+for the schedule,
+as shown in Fig. \ref{fig:vardep}.
+%
+%The dependence of variables are shown in Fig. \ref{fig:vardep}.
+%
+While the updating scheme alone
+determines $\lambda_k$'s,
+$\Gamma_k$'s depend also on the system
+and the underlying sampling process (MC or MD),
+and need to be measured using the technique
+in Sec. \ref{sec:Gamma_measure}.
+%
 The complete program is the following.
+
+\begin{figure}[h]
+\begin{tikzpicture}
+  \node[draw, text width={width("Updating ")}]
+    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
+
+  \node[draw, text width={width("sampling")}]
+    (sys) at (0, 1) {System\\and\\sampling\\process};
+
+  \node[draw, text width={width("Eigenvalues,")}]
+    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_2, \dots, \lambda_n$};
+
+  \node[draw, text width={width("Integrals, ")}]
+    (gamma) at (2.2, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
+
+  \node[draw, text width={width("Simulation")}]
+    (T) at (2.2, -0.5) {Simulation\\time, $T$};
+
+  \node[draw, text width={width("Optimal")}]
+    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};
+
+  \node[draw, text width={width("Schedule,")}]
+    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};
+
+  \node[draw, text width={width("Error, $\Err$")}]
+    (err) at (7, 0) {Error, $\Err$};
+
+  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};
+
+  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};
+
+  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
+  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
+  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};
+
+  \draw[->, >=latex, thick]
+    (w) edge (lambda)
+    (w) edge (gamma)
+    (sys) edge (gamma);
+
+  \draw[->, >=latex, thick]
+    (lambda) -| (M1)
+    (gamma)  -- (M1)
+    (T)      -| (M1)
+    (M1)     -- (R1)
+    (R1)     |- (qT);
+
+  \draw[->, >=latex, thick]
+    (qT)     -| (R3);
+
+  \draw[->, >=latex, thick]
+    (R3) -- (N1) |- (alpha);
+
+  \draw[->, >=latex, thick]
+    (N1) |- (err);
+
+  \draw[->, >=latex, thick]
+    %(M1) edge[bend left=70] (N1);
+    (R1) |- (R2) -| (R3);
+\end{tikzpicture}
+\caption{
+  \label{fig:vardep}
+  Dependence of variables in computing
+  the optimal schedule and the error.
+}
+\end{figure}
 

### 2016-04-22 doc/invt.tex
invt.tex move sinc.gp

@@
 
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
+  }
+  \caption{
+    \label{fig:sinc}
+    Updating kernels of the bandpass updating schemes
+    compared to a Gaussian with a similar width.
+    %
+    The lines are a guide to the eyes.
+    Inset. Updating matrix in the non-periodic case.
+    %
+    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
+      The data were computed by the program \texttt{prog/predict},
+      see particularly the function \texttt{intq\_save()}
+      in \texttt{prog/intq.h} for panel (a),
+      and the functions \texttt{savewin()}
+      and \texttt{savewinmat()}
+      in \texttt{prog/invt.h} for panel (b),
+      The data were saved under the directory \texttt{data/sinc}.
+      See the \texttt{make} object \texttt{sincsingle}
+      in the \texttt{Makefile} there.
+    }%
+  }
+\end{center}
+\end{figure}
+
+In Fig. \ref{fig:sinc},
+we give examples of the updating kernels
+of the bandpass updating schemes,
+compared to a Gaussian
+with a similar width, $\sigma$,
+%
+\begin{equation}
+  \sigma
+  \approx
+  \frac
+  {
+    n
+  }
+  {
+    \sqrt{ 2 \, \pi } \, g \, K
+  }
+  .
+\label{eq:sigma_equiv}
+\end{equation}
+%
+%where $g = 2$ for the periodic case
+%or $g = 1$ otherwise.
+%
+The kernel in the non-periodic case is a zigzag
+because of the term $(-1)^{n+K+l-1}$
+in Eq. \eqref{eq:mu_sinc_refl}.
+%
+However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
+the ruggedness disappears
+in the updating matrix by Eq. \eqref{eq:w_band_refl},
+as the above oscillatory term for $l = i - j$
+is canceled by the term for either $l = i + j - 1$
+or $l = i + j - 2 \, n - 1$
+with the opposite sign.
+
 \subsubsection{\label{sec:finlencorr}
@@
 %
-For a finite-length simulation,
-we should include the residual part
-and minimize the total error.
-%
-For the bandpass updating schemes,
-this leads to the optimal value
-of the free parameter, $t_0$,
+By including the residual error
+and minimize the total,
+we get the optimal $t_0$,
 which is, according to Eq. \eqref{eq:half_alpha0},
@@
 
-Our test system is
-a model one-dimensional system
-of $n = 100$ bins,
-with a flat target distribution
-$p_i = 1/n$.
-%
-%Since in the asymptotic regime,
-%the distribution converges to $p_i$
-%no matter the intrinsic distribution, $p^*_i$,
-%we also set the latter to be flat for convenience.
-The intrinsic distribution, $p^*_i$,
-is also assumed to be flat.
-%
-Unless specified otherwise,
-we assume $i$ to be non-periodic.
-%
-We wish to see how
-the schedule, $\alpha(t)$,
-affects the fluctuating error
-of the bias potential
-at the end of the simulation.
-
+In our test system,
+$i$ is the index of a one-dimensional
+non-periodic variable,
+and $n = 100$.
+%
+Both the target and intrinsic distributions
+are assumed to be flat:
+$p^*_i = p_i = 1/n$.
+%
+%We wish to see how
+%the schedule, $\alpha(t)$,
+%affects the fluctuating error
+%of the bias potential
+%at the end of the simulation.
+%
 Initially,
@@
 we reset the origin of the time, $t$,
-and started the testing schedule.
+and start the testing schedule.
 %specified by either
@@
 
-Since the error depends on
-the underlying sampling process,
-we tested two sampling processes,
-a global one and a local one.
+%Since the error depends on
+%the underlying sampling process,
+We tested two sampling processes.
+%a global one and a local one.
 %
@@
 %
-This process
-emulates the perfect sampling process
-in Sec. \ref{sec:Gamma} asymptotically.
-%
 In the second or local sampling process,
@@
 (adjusted by the boundary condition)
-with equal probability $1/2$,
-which mimics the one-step sampling process
-in Sec. \ref{sec:Gamma}.
+with equal probability $1/2$.
+%
+Asymptotically,
+the above global and local sampling processes
+emulate the perfect and one-step processes
+discussed in Appendix \ref{sec:Gamma},
+respectively,
+and we shall use
+Eqs. \eqref{eq:Gamma_perfect}
+and \eqref{eq:Gamma_onestep}
+for the respective $\Gamma_k$'s.
 %
@@
 
-We now turn to the optimal schedule
-predicted by Eq. \eqref{eq:q_opt}.
-%
-For the Gaussian updating scheme
-with $\sigma = 10$,
-we computed the schedule for the local sampling process,
-and compared it with the inverse-time schedule
-Eq. \eqref{eq:alpha_invtlambda}
-with the optimal $\lambda$.
-%
-As shown in Fig. \ref{fig:optacmp}(a),
-the optimal schedule is a more complex curve
-than the optimized inverse-time schedule.
-%
-The complexity of the optimal schedule
-corresponds to the non-exponential tail
-of the mass distribution function,
-as shown Fig. \ref{fig:optacmp}(b)
-[cf. Sec. \ref{sec:mass_distr}].
-%
-The errors in the two cases were
-$9.1 \times 10^{-5}$
-and
-$1.7 \times 10^{-4}$,
-respectively,
-%for the local sampling process,
-showing the superiority of
-the optimal schedule from Eq. \eqref{eq:q_opt}.
-%
-Besides, we verified the scaling property
-discussed in Sec. \ref{sec:mass_distr},
-by comparison with a simulation [Fig. \ref{fig:optacmp}(a)]
-of half length, $T' = 5 \times 10^7$
-and doubled initial updating magnitude,
-$\alpha_0 = 2 \times 10^{-4}$,
-which left the normalized mass distribution invariant,
-as shown in Fig. \ref{fig:optacmp}(b).
+We computed the optimal schedule from Eq. \eqref{eq:q_opt}
+for the Gaussian (with $\sigma = 10$)
+and bandpass (with $K = 4$) updating schemes.
+%
+As shown in Fig. \ref{fig:optacmp},
+while the optimal schedules of the bandpass scheme
+were always given by the inverse-time formula,
+those of the Gaussian updating scheme was complex
+and dependent on the sampling process.
 
@@
       See the \texttt{make} object \texttt{sigsingle}
-      in the \texttt{Makefile} there.
-    }%
-  }
-\end{center}
-\end{figure}
-
-On the other hand,
-the optimal schedules of the bandpass updating schemes
-are always given by the inverse-time formula,
-Eq. \eqref{eq:alpha_invt1},
-with $t_0$ given by Eq. \eqref{eq:t0_sinc},
-as shown in Fig. \ref{fig:sinc}(a).
-%
-In Fig. \ref{fig:sinc}(b),
-we show that the updating kernels
-of the bandpass updating schemes
-are similar to Gaussian ones
-except the oscillatory tails.
-%
-The equivalent width, $\sigma$, is roughly
-%
-\begin{equation}
-  \sigma
-  \approx
-  \frac
-  {
-    n
-  }
-  {
-    \sqrt{ 2 \, \pi } \, g \, K
-  }
-  ,
-\label{eq:sigma_equiv}
-\end{equation}
-%
-where $g = 2$ for the periodic case
-or $g = 1$ otherwise.
-%
-The kernel is a zigzag in the non-periodic case
-because of the term $(-1)^{n+K+l-1}$
-in Eq. \eqref{eq:mu_sinc_refl}.
-%
-However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
-the ruggedness disappears
-in the updating matrix by Eq. \eqref{eq:w_band_refl},
-as the above oscillatory term for $l = i - j$
-is canceled by the term for either $l = i + j - 1$
-or $l = i + j - 2 \, n - 1$
-with the opposite sign.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
-  }
-  \caption{
-    \label{fig:sinc}
-    Bandpass updating schemes.
-    (a) Optimal schedule for a periodic variable
-    with $K = 3$,
-    and that for a non-periodic variable
-    with $K = 6$.
-    (b) Updating kernels
-    compared to a Gaussian with a similar width,
-    $\sigma = 6$.
-    %
-    The local sampling process was used.
-    The lines are a guide to the eyes.
-    Inset. Updating matrix in the non-periodic case
-    is free of the zigzaggedness shown in the kernel.
-    %
-    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h} for panel (a),
-      and the functions \texttt{savewin()}
-      and \texttt{savewinmat()}
-      in \texttt{prog/invt.h} for panel (b),
-      The data were saved under the directory \texttt{data/sinc}.
-      See the \texttt{make} object \texttt{sincsingle}
       in the \texttt{Makefile} there.
@@
 %
-We can also
-use Eq. \eqref{eq:y2_eql}.
+For an unknown system,
+we can use Eq. \eqref{eq:y2_eql} to
 measure $\Gamma_k$'s in
-a simulation with a constant schedule,
+an adaptive FDS simulation
+with a constant schedule,
 $\alpha(t) = \alpha_0$.
@@
 However,
-we often do not know the intrinsic distribution, $p^*_i$,
-in advance to compute
-the shifted bias potential, $v_i$,
+without the intrinsic distribution, $p^*_i$,
+we cannot compute the shifted bias potential, $v_i$,
 from the original one, $V_i$
@@
 and compared them to those from
-the local MC sampling process
-[cf. Sec. \ref{sec:results_system}].
+Eq. \eqref{eq:Gamma_onestep}.
+%the one-step sampling process.
 %
@@
     from an MD simulation versus those
-    from the local MC simulation.
+    from the one-step process.
     %
@@
 $\Gamma_k$'s depend also on the system
-and the underlying sampling process (MC or MD),
-and need to be measured using the technique
-in Sec. \ref{sec:Gamma_measure}.
-%
-The complete program is the following.
+and the underlying sampling process (MC or MD).
+%
+By using the technique in Sec. \ref{sec:Gamma_measure}
+to measure $\Gamma_k$'s,
+we have the following program.
 
@@
 Run a preliminary adaptive FDS simulation
-using the updating scheme with a constant
-updating magnitude, $\alpha_0$.
+%using the updating scheme
+with a constant updating magnitude, $\alpha_0$.
 

### 2016-04-22 doc/invt.tex
invtrun.py -> invtcscan.py, add log(c) scan with --x
improve the code for predicting error components

@@
 Since we only optimized the asymptotic error, $E_A$,
-in th above,
+in the above,
 the parameter $q(T)$ can be further adjusted
@@
   ,
-\label{eq:mass_distr}
+\notag
+%\label{eq:mass_distr}
 \end{equation}
@@
 %
-which means that each point $(t, \alpha(t))$
+which means that each point
+$\bigl(t, \alpha(t)\bigr)$
 on the schedule can be mapped
-to a point $(Q, m(Q))$ on the mass distribution,
+to a point
+$\bigl(Q, m(Q)\bigr)$
+on the mass distribution,
 such that the ordinate of the latter curve
@@
 %
-Thus, this updating scheme is stable if
+So this updating scheme is stable if
 $\min \lambda_k = \lambda_n \ge 0$,
@@
 %
-Then,
-by assuming the largest $b \approx n/g$,
+By assuming the largest $b \approx n/g$,
 Eq. \eqref{eq:wband_eigenvalue}
@@
 with $\Delta x \ll \sigma \ll 1$,
-we can extended
+we can extend
 the upper limit of the integrals
@@
 %
-\note{We add some steps for the above solution.
+\note{Derivation.
   Integrating $\dot u_k = c_k$ yields
@@
     u_k(t) = c_k \left(t + t_{k0} \right).
-  \tag{N4}
+  \tag{UK}
   \label{neq:ui_solution}
@@
 Then,
-Eq. \eqref{eq:error_singlebin} is changed to
-%
-\begin{equation}
-  E_A
-  =
-  \frac {       T     }
-        { (T + t_0)^2 }
-  \sum_{ k = 2 }^K
-    \Gamma_k.
-\label{eq:error_asym_sinc}
-\end{equation}
+we can change the upper limit of the sum in
+Eq. \eqref{eq:error_singlebin} to $K$.
+%
+%Then,
+%Eq. \eqref{eq:error_singlebin} is changed to
+%%
+%\begin{equation}
+%  E_A
+%  =
+%  \frac {       T     }
+%        { (T + t_0)^2 }
+%  \sum_{ k = 2 }^K
+%    \Gamma_k.
+%%\notag
+%\label{eq:error_asym_sinc}
+%\end{equation}
 %
@@
 In our test system,
-$i$ is the index of a one-dimensional
+$i$ represents a one-dimensional
 non-periodic variable,
@@
 where $\lambda$ is a free parameter,
-and $t_0 = 2/\alpha_0$.
+and $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
+%and $t_0 = 2/\alpha_0$.
 %
@@
 
-We performed FDS simulations
+We performed adaptive FDS simulations
 using the single-bin updating scheme
@@
     for the Gaussian
-    and the bandpass updating schemes.
+    and the bandpass updating schemes,
+    with $t_0$ given by Eq. \eqref{eq:t0_sinc}.
     %
@@
     $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
-    %
-    Here, we assumed $t_0 \equiv T/(e^{q(T)} - 1)$
-    for the optimal value of $q(T)$.
     %
@@
     %
-    The velocity rescaling\cite{bussi2007} thermostat was
-    applied every MD step with viscocity 20.
+    The velocity rescaling thermostat\cite{bussi2007} was
+    applied every MD step with viscosity $20$.
     %
@@
     \note{The figure was produced by \texttt{doc/fig/gamcmp.gp}.
-      Thee data were saved under \texttt{data/gamcmp}
+      The data were saved under \texttt{data/gamcmp}
       (see the \texttt{Makefile} there).
@@
   .
-\label{eq:error_res_invt}
+\notag
+%\label{eq:error_res_invt}
 \end{equation}
@@
 and
-$\Gamma_k$'s affect the errors
-%from Eqs. \eqref{eq:error_res_invt}
-%and \eqref{eq:error_asym_invt}
+$\Gamma_k$'s affect the error
 only via the sum, $\sum_{k = 2}^n \Gamma_k$,

### 2016-04-22 prog/invtpar.h

@@ -76,6 +76,8 @@ typedef struct {
   long nsteps; /* number of steps */
   long ntrials; /* number of trials */
 
+  char fnxerr[FILENAME_MAX]; /* file name for the error components */
+
 #ifdef SCAN /* for predict.c */
   int cscan;
   double cmin;
@@ -191,6 +193,8 @@ static void invtpar_init(invtpar_t *m)
   m->ntrials = 100;
   m->prog = "invt";
   m->verbose = 0;
+
+  m->fnxerr[0] = '\0';
 
 #ifdef SCAN /* for predict.c */
   m->cscan = 0;
@@ -450,6 +454,7 @@ static void invtpar_help(const invtpar_t *m)
   fprintf(stderr, "  --try=:        set the number of trials, default %ld\n", m->ntrials);
   fprintf(stderr, "  --nsteps=:     set the number of simulation steps, default %ld\n", m->nsteps);
   fprintf(stderr, "  --equil=:      set the number of equilibration steps, default %ld\n", m->nequil);
+  fprintf(stderr, "  --fnxerr=:     set the file name for the error components, default %s\n", m->fnxerr);
 #ifdef SCAN /* for predict.c */
   fprintf(stderr, "  --cmin=:       set the minimal c in c-scan, default %g\n", m->cmin);
   fprintf(stderr, "  --cmax=:       set the maximal c in c-scan, default %g\n", m->cmax);
@@ -850,6 +855,10 @@ static int invtpar_keymatch(invtpar_t *m,
   else if ( strcmpfuzzy(key, "verbose") == 0 )
   {
     m->verbose = invtpar_getbool(m, key, val);
+  }
+  else if ( strcmpfuzzy(key, "fnxerr") == 0 )
+  {
+    strcpy(m->fnxerr, val);
   }
 #ifdef SCAN /* for predict.c */
   /* c-scan paramerters */

### 2016-04-22 prog/predict.c

@@ -9,8 +9,10 @@
 
 
 
+/* save the error components to file */
 static int save_xerr(invtpar_t *m, const char *fn,
-    const double *xerri, const double *xerrf, const double *xerrf2,
+    const double *xerri, const double *xerrf,
+    const double *xerrf_r, const double *xerrf_a,
     const double *lambda)
 {
   FILE *fp;
@@ -23,8 +25,8 @@ static int save_xerr(invtpar_t *m, const char *fn,
 
   fprintf(fp, "# %d\n", n);
   for ( i = 1; i < n; i++ ) {
-    fprintf(fp, "%4d %18.8e %18.8e %18.8e %10.8f\n",
-        i, xerri[i], xerrf[i], xerrf2[i],
+    fprintf(fp, "%4d %18.8e %18.8e %18.8e %18.8e %18.8e\n",
+        i, xerri[i], xerrf[i], xerrf_r[i], xerrf_a[i],
         lambda[i]);
   }
 
@@ -40,12 +42,13 @@ static void invt_geterr(invtpar_t *m,
   int n = m->n;
   double T, c0, t0, qT, inita, err0, err1, err2;
   double *lambda;
-  double *xerri, *xerrf, *xerrf_invt;
+  double *xerri, *xerrf, *xerrf_r, *xerrf_a;
   intq_t *intq;
 
   xnew(xerri, n);
   xnew(xerrf, n);
-  xnew(xerrf_invt, n);
+  xnew(xerrf_r, n);
+  xnew(xerrf_a, n);
 
   T = (double) m->nsteps;
 
@@ -66,7 +69,7 @@ static void invt_geterr(invtpar_t *m,
 
   c0 = m->c;
   err0 = esterror_invt(T, m->c, m->alpha0, m->t0, m->n,
-      xerrf_invt, lambda, gamma);
+      NULL, lambda, gamma);
 
   /* compute the optimal c and the error
    * for the inverse-time formula */
@@ -76,8 +79,7 @@ static void invt_geterr(invtpar_t *m,
   /* compute the exact minimal error under the same condition */
   qT = m->qT;
   err2 = esterror_opt(T, m->alpha0, m->initalpha, &qT, m->qprec,
-      m->alpha_nint, &intq, m->n, xerrf,
-      lambda, gamma, m->verbose);
+      m->alpha_nint, &intq, m->n, lambda, gamma, m->verbose);
   inita = intq_getinita(intq);
 
[…]

### 2016-04-26 doc/invt.tex
k-limited optimal schedule
fix key legend

@@
 \subsection{\label{sec:results_invt}
-Inverse-time schedule}
+Modified inverse-time schedule}
 
 We tested the theory of computing errors
-on the simple inverse-time schedule,
+on the modified inverse-time schedule,
 %
@@
       in the \texttt{Makefile} there.
+    }%
+  }
+\end{center}
+\end{figure}
+
+For the Gaussian updating scheme,
+we listed the errors from different schedules,
+in Table \ref{tab:error_Gaussian}.
+%
+Clearly,
+the optimal schedules yielded the smallest errors.
+
+\begin{table}[h]
+  \caption{\label{tab:error_Gaussian}
+    Error of the Gaussian updating scheme of $\sigma = 10$.
+    $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
+    \note{ Use the simulation data.  }%
+  }
+  \setlength{\tabcolsep}{5pt}
+  \begin{tabular} { l | p{2cm} p{2.5cm} p{1.8cm} }
+    \hline
+      Schedule
+    &
+      Inverse-time, Eq. \eqref{eq:alpha_invt1}
+    &
+      Modified inverse-time,
+      Eq.~\eqref{eq:alpha_invtlambda} 
+      with optimized $\lambda$
+    &
+      Optimal,
+      Eq. \eqref{eq:q_opt}
+    \\
+    \hline
+    Global
+    &
+    $4.2 \times 10^{-6}$
+    &
+    $7.1 \times 10^{-7}$
+    &
+    $2.7 \times 10^{-7}$
+    \\
+    \hline
+    Local
+    &
+    $3.7 \times 10^{-4}$
+    &
+    $1.7 \times 10^{-4}$
+    &
+    $9.1 \times 10^{-5}$
+    \\
+    \hline
+  \end{tabular}
+\end{table}
+
+To verify Eq. \eqref{eq:half_alpha0},
+we computed the optimal schedules
+for different values of $q(T)$,
+and plotted the final error
+against the initial updating magnitude, $\alpha(0)$.
+%
+As shown in Fig. \ref{fig:erriascan},
+The minimal error is reached at
+$\alpha(0) = \alpha/2 = 5 \times 10^{-5}$.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/erriascan.pdf}
+  }
+  \caption{
+    \label{fig:erriascan}
+    Relative error
+    versus the initial updating magnitude, $\alpha(0)$,
+    for optimal schedules with different $q(T)$.
+    Assuming the Gaussian updating scheme
+    with $\sigma = 10$,
+    and $\alpha_0 = 10^{-4}$.
+    %
+    The points were obtained from averaging results
+    from 1000 independent simulations;
+    the curves are theoretical predictions.
+    %
+    \note{This figure was produced by
+      \texttt{doc/fig/erriascan.gp}.
+      The data were saved under
+      \texttt{data/iascan}.
     }%

### 2016-04-26 doc/invt.tex
new xerr.gp, edits

@@
 a nearly flat distribution\cite{
-dayal2004, trebst2004, singh2011},
+dayal2004, trebst2004, zhang2007, barducci2008, singh2011},
 $p_i$,
@@
 $V_i \equiv 0$,
-%
 and equilibrate the system for $10^7$ steps
 under a constant updating magnitude,
-$\alpha_0 = 10^{-4}$,
-unless specified otherwise.
+$\alpha_0 = 10^{-4}$.
 %
@@
 we reset the origin of the time, $t$,
-and start the testing schedule.
+and start the testing schedule
+for $T = 10^8$ steps.
+%
+We repeat the above process $1000$ times
+and average the results.
 %specified by either
@@
 %Eq. \eqref{eq:alpha_invtlambda}.
+Unless specified otherwise,
+the above parameters were used.
 
@@
 %
-For each value of $\lambda$, we ran $1000$ independent simulations
-of $10^8$ steps, and averaged the results.
+%For each value of $\lambda$,
+%we ran $1000$ independent simulations
+%of $10^8$ steps, and averaged the results.
 %
@@
     Error $E$ versus $\lambda$
-    in Eq. \eqref{eq:alpha_invtlambda}
-    for the inverse-time schedule.
+    for the modified inverse-time schedule,
+    Eq. \eqref{eq:alpha_invtlambda}.
     %
@@
 
-We computed the optimal schedule from Eq. \eqref{eq:q_opt}
-for the Gaussian (with $\sigma = 10$)
-and bandpass (with $K = 4$) updating schemes.
-%
-As shown in Fig. \ref{fig:optacmp},
-while the optimal schedules of the bandpass scheme
-were always given by the inverse-time formula,
-those of the Gaussian updating scheme was complex
-and dependent on the sampling process.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/optacmp.pdf}
-  }
-  \caption{
-    \label{fig:optacmp}
-    Optimal schedules from Eq. \eqref{eq:q_opt},
-    versus the optimized inverse-time schedule
-    Eq. \eqref{eq:alpha_invtlambda}
-    for the Gaussian updating scheme
-    with $\sigma = 10$
-    and the local updating process.
-    %
-    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h}.
-      The data were saved under the directory \texttt{data/opta}.
-      See the \texttt{make} object \texttt{sigsingle}
-      in the \texttt{Makefile} there.
-    }%
-  }
-\end{center}
-\end{figure}
-
-For the Gaussian updating scheme,
-we listed the errors from different schedules,
-in Table \ref{tab:error_Gaussian}.
-%
-Clearly,
-the optimal schedules yielded the smallest errors.
+To demonstrate the optimal schedule from Eq. \eqref{eq:q_opt},
+we used the Gaussian (with $\sigma = 10$) updating scheme
+as an example.
+%
+In Table \ref{tab:error_Gaussian},
+we showed that the optimal schedule
+produced smaller errors than
+the inverse-time schedules.
+%
+However, the optimal schedules
+decayed more slowly than
+the inverse-time formula, Eq. \eqref{eq:alpha_invt1},
+and depended on the sampling process,
+as shown in Fig. \ref{fig:optacmp}.
+%
+The impeded decay of the optimal schedule
+was necessary in order to
+damp out the short-wavelength eigenmodes
+(i.e., local noises),
+as shown in Fig. \ref{fig:xerr}.
+%
+If, however, we could ignore these local eigenmodes
+(by explicitly filtering out
+them in the biased potential),
+then we can modify the optimal schedule
+from Eq. \eqref{eq:q_opt}
+and limit the sum
+to the first few eigenmodes.
+%
+The resulting schedule
+would became similar to the inverse-time schedule,
+Eq. \eqref{eq:alpha_invt1},
+as shown in Fig. \ref{fig:optacmp}.
+%
+\note{A more thorough approach is
+to use the bandpass updating scheme
+to be discussed later.
+}
 
@@
 
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/optacmp.pdf}
+  }
+  \caption{
+    \label{fig:optacmp}
+    Optimal schedules from Eq. \eqref{eq:q_opt}
+    for the Gaussian updating scheme
+    with $\sigma = 10$.
+    %
+    The modified schedules, produced by
+    limiting the sum in Eq. \eqref{eq:q_opt}
+    to the first four terms,
+    became similar to the inverse-time schedule,
+    Eq. \eqref{eq:alpha_invt1}.
+    %
+    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
+      The data were computed by the program \texttt{prog/predict},
+      see particularly the function \texttt{intq\_save()}
+      in \texttt{prog/intq.h}.
+      The data were saved under the directory \texttt{data/opta}.
+    }%
+  }
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/xerr.pdf}
+  }
+  \caption{
+    \label{fig:xerr}
+    Error components for the Gaussian updating scheme
+    with $\sigma = 10$
+    from the optimal schedule
+    and the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
+    %
+    The latter schedule failed to damp out the errors
+    from the long-wavelength (local) eigenmodes
+    with $K \ge 5$.
+    %
+    For longer simulations,
+    the error components from the optimal schedule
+    would roughly maintain the same shape,
+    but ``intersect'' with the initial error curve
+    at a lower level (not shown).
+    %
+    \note{This figure was produced by \texttt{doc/fig/xerr.gp}.
+      The data were saved under the directory \texttt{data/xerr}.
+    }%
+  }
+\end{center}
+\end{figure}
+
 To verify Eq. \eqref{eq:half_alpha0},
@@
 As shown in Fig. \ref{fig:erriascan},
-The minimal error is reached at
+The minimal error was reached at
 $\alpha(0) = \alpha/2 = 5 \times 10^{-5}$.
@@
 In contrast,
-the normalized error of the bandpass updating schemes
+the normalized errors of the bandpass updating schemes
 were insensitive to the simulation length,
@@
 yielded smaller errors than
-the corresponding Gaussian ones.
+the corresponding Gaussian ones,
+although the differences tended to
+be small for shorter simulations
+with the local sampling process.
 
@@
 including the Gaussian one used in metadynamics,
-the optimal schedule has no closed form,
+the optimal schedule permits no simple closed form,
 but is given by an equation of motion
@@
 is always given by the inverse-time formula.
-%
-They appeared to outperform similar Gaussian schemes
+
+Although the bandpass updating schemes
+appeared to outperform similar Gaussian schemes
 in our model study,
-and they can be useful
-for high-precision metadynamics simulations.
-%
-However, since the bandpass schemes
-update the bias potential globally,
-they can be more expensive than
-the Gaussian schemes.
+they are global in nature
+and hence more expensive.
 %
@@
 with a local sampling process such as MD.
+%
+\note{
+One way to improve the Gaussian updating schemes
+is to use it with a hard cutoff of eigenmodes
+(as we did for Fig. \ref{fig:optacmp}).
+%
+Then the optimal schedule
+would be similar to the inverse-time one,
+and the error, with the short-wavelength eigenmodes filtered,
+would be close to that of the bandpass scheme.
+}
 
@@
 
-We thanks
-Dr. Y. Mei, Dr. J. Ma,
+We thank Dr. J. Ma, Dr. Y. Mei,
 J. Drake, O. Nassar, Dr. C. Lai and Dr. S. Ou
@@
 from the two sampling processes
-were almost proportional,
-suggesting similarity
-of local sampling processes.
+were almost proportional.
+%
+This suggests that
+the optimal schedule
+from simulations using local sampling processes
+would be similar in general properties.
 
@@
   the Gaussian updating scheme can be unstable
+  in practice
   if the updating kernel is truncated.
@@
 \section{\label{sec:invt_schedule}
-Inverse-time schedules}
+Modified inverse-time schedules}
 

### 2016-04-26 prog/predict.c

@@ -42,7 +42,7 @@ static void invt_geterr(invtpar_t *m,
     const double *gamma)
 {
   int n = m->n;
-  double T, c0, t0, qT, inita, err0, err1, err2;
+  double T, c0, c1, t0, qT, inita, err0, err1, err2;
   double *lambda;
   double *xerri, *xerrf, *xerrf_r, *xerrf_a;
   intq_t *intq;
@@ -75,10 +75,11 @@ static void invt_geterr(invtpar_t *m,
 
   /* compute the optimal c and the error
    * for the inverse-time formula */
-  m->c = estbestc_invt(T, m->alpha0, 0, m->n, lambda, gamma,
+  c1 = estbestc_invt(T, m->alpha0, 0, m->n, lambda, gamma,
       0, &err1, 0);
 
-  /* compute the exact minimal error under the same condition */
+  /* compute the minimal error from the optimal schedule
+   * under the same condition */
   qT = m->qT;
   err2 = esterror_opt(T, m->alpha0, m->initalpha, &qT, m->qprec,
       m->alpha_nint, &intq, m->n, m->kcutoff, m->pbc,
@@ -87,17 +88,24 @@ static void invt_geterr(invtpar_t *m,
 
   /* save the optimal schedule to file */
   t0 = 2 / m->alpha0;
-  intq_save(intq, m->c, t0, m->alpha_resample, m->fnalpha);
+  intq_save(intq, c1, t0, m->alpha_resample, m->fnalpha);
 
   printf("c %g, t0 %g, err %g, sqr %g (invt), "
          "opt. c %g, err %g, sqr %g (opt. invt), "
          "qT %g, a(0) %g, err %g, sqr %g (exact), %s\n",
       c0, t0, err0, err0 * err0,
-      m->c, err1, err1 * err1,
+      c1, err1, err1 * err1,
       qT, inita, err2, err2 * err2, m->fnalpha);
 
   if ( m->fnxerr[0] != '\0' ) {
-    intq_errcomp(intq, m->alpha0, qT, xerrf, xerrf_r, xerrf_a);
+    if ( m->opta ) {
+      /* optimal schedule */
+      intq_errcomp(intq, m->alpha0, qT, xerrf, xerrf_r, xerrf_a);
+    } else {
+      /* inverse-time schedule */
+      err0 = esterror_invt_x(T, m->c, m->alpha0, m->t0, m->n,
[…]

### 2016-04-27 doc/invt.tex
invt.tex: small edits

@@
 %
-The main difference between the two
+The main difference
 is that in the WL case,
@@
 is updated frequently (usually every few MC or MD steps),
-which % brings about a side effect of disrupting
+%which % brings about a side effect of disrupting
+and the fluctuation
 disrupts \note{``disturbs''?} the underlying
@@
 over the course of simulation
-to decrease the error.
+to decrease the effect.
 
@@
 For the WL algorithm, it is known
-that the optimal updating magnitude, $\ln f$,
+that the optimal schedule
 is given by the product of the
@@
 morozov2007, zhou2008}
-and the number of bins
 (with ``time'' being
@@
 the optimal schedule
-for a general adaptive flat-distribution sampling methods,
-including those used in WL and metadynamics simulations.
+for a general adaptive flat-distribution sampling method.
 %
@@
 to a class of bandpass updating schemes
-that are optimal for infinitely long simulations,
+that are optimal for long simulations,
 and the modified inverse-time schedule is optimal for these schemes.
@@
 are often improved
-by the histogram, $H_i$, along $i$ accumulated
+by the histogram, $H_i$, accumulated
 during simulation.
@@
 morozov2007, zhou2008, morozov2009}.
+%
+Here we will show that the ideal $c$
+is $1.0$, but the origin of $t$ should be shifted
+to accommodate the initial error.
 
@@
 or
-$
+\begin{equation}
   \dot u\bigl( q(t') \bigr) = c
   ,
-$
+\label{eq:dudt_const}
+\end{equation}
 %
@@
 no eigenmode grows indefinitely,
-which means all eigenvalues $\lambda_k$'s
-are nonnegative.
+which means that no eigenvalue $\lambda_k$
+can be negative.
 
@@
 the underlying sampling process
-affects the error only through the set of numbers,
+affects the error only through the numbers,
 $\Gamma_k$'s.
@@
 We can then use Eq. \eqref{eq:y2_eql}
-for the $\langle y_k^2(0) \rangle$'s in Eq. \eqref{eq:error_res}.
-%
-Then,
+for the $\langle y_k^2(0) \rangle$'s
+in Eq. \eqref{eq:error_res},
+and
 the residual error at the end of period $T$
@@
 does not explicitly depend on time, $t$,
-the Hamiltonian
+the Hamiltonian is conserved.
+%
+But the Hamiltonian
 %
@@
 %
-which coincides with the Lagrangian,
-is conserved.
+coincides with the Lagrangian.
 %
@@
 %
-for some positive $C$.
-%
-Integrating the square root, we get
-%
-\begin{equation}
+for some positive $C$,
+%
+%We may treat Eq. \eqref{eq:Lagrangian_const}
+as a generalization of Eq. \eqref{eq:dudt_const}.
+%
+By integrating the square root, we get
+%
+\begin{equation}
+  \int_{ 0 }^{ q(t) }
+    \sqrt{
+      \textstyle\sum_{ k = 2 }^n
+        \Gamma_k \, \lambda_k^2
+        \, u_k^2( q' )
+    }
+    \;
+    d q'
+  =
+  C \, \frac t T
+  ,
+  \label{eq:q_opt}
+\end{equation}
+%
+where $C$ is to be determined by
+the boundary condition at $t = T$ as
+%
+\begin{equation}
+  C =
   \int_{ 0 }^{ q(T) }
@@
     d q'
-  =
-  C \, \frac t T
-  ,
-  \label{eq:q_opt}
-\end{equation}
-%
-where $c$ is to be determined by
-the boundary condition at $t = T$ as
-%
-\begin{equation}
-  C =
-  \int_{ 0 }^{ q(t) }
-    \sqrt{
-      \textstyle\sum_{ k = 2 }^n
-        \Gamma_k \, \lambda_k^2
-        \, u_k^2( q' )
-    }
-    \;
-    d q'
   .
@@
 the parameter $q(T)$ can be further adjusted
-to minimize the total error, $E = E_R + E_A$.
+to minimize the total error, $E = E_R + E_A$
+(cf. Sec. \ref{sec:optinitalpha}).
 %
@@
 Clearly,
-the distribution decreases monotonically with $Q$.
+both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$.
 
@@
 which means the asymptotic error
-is inversely proportional to the length
+is inversely proportional to the simulation length
 for the same-class schedules,
-and we have the scaling relation
+i.e., we have the scaling relation
 $T_1 \, E_{A1} = T_2 \, E_{A2}$.
@@
 %
-Homogeneous updating schemes have significantly fewer
+Homogeneous updating schemes
+share the same set of eigenvectors
+but not the corresponding eigenvalues.
+%
+They have significantly fewer
 parameters than the general case,
@@
 If $b = n/2$, then $\mu_{-b}$ is removed
-from the sum in Eq. \eqref{eq:mu_normalization}.
+from the sum in Eq. \eqref{eq:mu_normalization}
+to avoid double counting.
 %
@@
 
-For the Gaussian updating kernel of width $\sigma$
+For the Gaussian updating kernel,
+if the width $\sigma$ satisifies
 with $\Delta x \ll \sigma \ll 1$,
@@
 the single-bin scheme is most efficient
-in asymptotic convergence
-if no eigenvalue $\lambda_k$ is zero.
+in asymptotic convergence.
+%if no eigenvalue $\lambda_k$ is zero.
 
@@
 For example,
-if we can assume that the PMF is sufficiently smooth
-such that the eigenmodes can be limited to $k \le K$.
-%
-Then we can construct an updating scheme with
+%if we can assume
+by assuming
+that the PMF is sufficiently smooth
+such that the eigenmodes can be limited to $k \le K$,
+%
+we can construct an updating scheme with
 $$
@@
 that the minimal error was achieved
-by the widest window for the shorter simulation
-of $T = 10^8$,
-for the global sampling process.
-%
-As the simulation lengthens,
-the minimum at $\sigma = 0$ became more favorable.
+by the widest window %for the shorter simulation of $T = 10^8$,
+for the global sampling process,
+%
+although
+the minimum at $\sigma = 0$ became more favorable
+as the simulation lengthens.
 %
 However,
+for the local sampling process,
 the minimum stayed at zero width,
@@
 optimal schedule from Eq. \eqref{eq:q_opt},
-here we consider the simpler
+here we consider the modified
 inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
@@
   ,
-\notag
-%\label{eq:lambda_nn_perfect}
+%\notag
+\label{eq:lambda_nn_perfect}
 \end{equation}
@@
 %
+Note that the value given by
+Eq. \eqref{eq:lambda_nn_perfect}
+was less than the observed optimal value
+from Fig. \ref{fig:errinvt},
+and the latter would decrease
+with the simulation length.
+%
 

### 2016-04-29 doc/invt.tex
add data for Table I
tikz version of Figure 1.

@@
 \usepackage{tikz}
+\usetikzlibrary{calc}
 \usepackage{hyperref}
@@
 \begin{equation}
-  p_i \, w_{ij} = p_j \, w_{ji}.
+  \rho_i \, w_{ij} = \rho_j \, w_{ji}.
   \label{eq:w_detailedbalance}
@@
 \begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.9\linewidth]{fig/massq.pdf}
-  }
-  \caption{
-    \label{fig:massq}
-    Geometric characterization of the optimal schedule,
-    $\alpha(t)$,
-    via the normalized mass distribution, $m(Q)$.
-    %
-    \note{The figure was produced by \texttt{doc/fig/massq.gp}.
-    }%
-  }
+  \begin{tikzpicture}[very thick, scale=2.5, every node/.style={scale=1.5}]
+    % distribution
+    \draw[domain=0:2.5, variable=\x, dashed, gray, ultra thick] 
+      plot({\x}, {1.5*exp(-0.5*\x*\x)});
+
+    % shade
+    \fill [gray!30!white, domain=0.7:2.2, variable=\x]
+      (0.7, 0) --
+      plot ({\x}, {1.5*exp(-0.5*\x*\x)}) 
+      -- (2.2, 0) -- cycle;
+
+    % thick curve
+    \draw[domain=0:2.2, variable=\x, ultra thick]
+      plot({\x}, {1.5*exp(-0.5*\x*\x)});
+
+    % label for the shaded area
+    \node[] at (1.1, 0.5) {$\dfrac{t}{T}$};
+
+    \draw[<->] (0.6, 0)  --
+               node[left, inner sep=1mm]
+               {$\dfrac{1}{T \, \alpha}$}
+               (0.6, 1.174);
+    % label for the ordinate
+    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
+
+    \draw[] (0.5, 1.174) -- (0.7, 1.1740);
+    \draw[->] (0, 0) -- node[below, inner sep=5mm] {$Q$} (2.8, 0);
+    \draw[->] (0, 0) -- node[above, rotate=90] {$M(Q)$} (0, 1.8);
+
+    % xtics
+    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
+    \draw[] (2.2, 0.1) -- (2.2, 0.0) node[below] {$q(T)$};
+  \end{tikzpicture}
+  %\makebox[\linewidth][c]{
+  %  \includegraphics[angle=0, width=0.9\linewidth]{fig/massq.pdf}
+  %}
+  %\caption{
+  %  \label{fig:massq}
+  %  Geometric characterization of the optimal schedule,
+  %  $\alpha(t)$,
+  %  via the normalized mass distribution, $m(Q)$.
+  %  %
+  %  \note{The figure was produced by \texttt{doc/fig/massq.gp}.
+  %  }%
+  %}
 \end{center}
@@
 the system is initially equilibrated at the same
-constant updating magnitude $\alpha$.
+constant updating magnitude $\alpha_0$.
 %
@@
 Then solve $\partial (E_R + E_A) / \partial q(T) = 0$.
-}
-%
-Then, by Eqs. \eqref{eq:mQ_invTa} and \eqref{eq:intmQ_tT},
-we have
+}%
+By Eqs. \eqref{eq:mQ_invTa} and \eqref{eq:intmQ_tT},
+we can rewrite the condition as
 \begin{equation}
@@
 %
+Naturally, $\rho_i = 1/n$ in this case.
+%
 An example is the Gaussian updating scheme
@@
 
-While translationally-invariant
-updating schemes
-naturally suit periodic variables\cite{
-dama2014},
-they can also be extended to non-periodic variables
-by imposing the reflective boundary condition\cite{
-bussi2006}.
-%
-For simplicity, we shall assume that the target
-distribution is flat, or $p_i = 1/n$, below.
-
-It is natural to introduce homogeneous updating schemes
+%While translationally-invariant
+%updating schemes
+%naturally suit periodic variables\cite{
+%dama2014},
+%they can also be extended to non-periodic variables
+%by imposing the reflective boundary condition\cite{
+%bussi2006}.
+%%
+%For simplicity, we shall assume that the target
+%distribution is flat, or $p_i = 1/n$, below.
+
+It is natural to first introduce homogeneous updating schemes
 for a periodic variable\cite{dama2014}.
@@
 and \eqref{eq:Gamma_onestep}
-for the respective $\Gamma_k$'s.
+for the respective $\Gamma_k$'s
+in the theoretical predictions.
 %
@@
     $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
-    \note{ Use the simulation data.  }%
+    \note{
+      The data were collected in
+      \texttt{data/tab1} by running
+      the program \texttt{invt}.
+    }%
   }
@@
     &
-    $4.2 \times 10^{-6}$
+    $4.4(1) \times 10^{-6}$
     &
@@
     &
-    $2.7 \times 10^{-7}$
+    $2.6(1) \times 10^{-7}$
     \\
@@
     &
-    $3.7 \times 10^{-4}$
+    $3.8(2) \times 10^{-4}$
     &
-    $1.7 \times 10^{-4}$
+    $1.8(1) \times 10^{-4}$
     &
-    $9.1 \times 10^{-5}$
+    $1.0(1) \times 10^{-4}$
     \\
@@
     would roughly maintain the same shape,
-    but ``intersect'' with the initial error curve
+    but join the initial error curve
     at a lower level (not shown).
@@
   =
-  (V_i - \bar V) + s_i,
+  (V_i - \bar V) + \chi_i,
 $
@@
 $
-  s_i
+  \chi_i
   \equiv
@@
 $$
-y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, s_i.
+y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, \chi_i.
 $$
@@
 and compared them to those from
-Eq. \eqref{eq:Gamma_onestep}.
-%the one-step sampling process.
+%Eq. \eqref{eq:Gamma_onestep}.
+the local MC sampling process.
 %
@@
     from an MD simulation versus those
-    from the one-step process.
-    %
-    The MD simulation was ran for $10^{10}$ steps with
+    from the local MC simulation.
+    %
+    Each simulation was ran for $10^{10}$ steps.
+    %
+    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$.
+    %
+    For the MD simulation,
     the step size $\Delta t_\mathrm{MD} = 0.005$.
     %
-    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$,
-    and the temperature was $1.0$.
-    %
     The velocity rescaling thermostat\cite{bussi2007} was
-    applied every MD step with viscosity $20$.
+    applied for the unit reduced temperature
+    every MD step with viscosity $20$.
     %

### 2016-05-02 doc/invt.tex
invt.tex: small changes
invt.tex: p/rho separation

@@
 
-\title{Optimal updating factor in adaptive flat-distribution simulations}
+\title{Optimal updating factor in adaptive flat-distribution-sampling simulations}
 
@@
   of the updating magnitude
-  for adaptive flat-distribution sampling methods,
+  for adaptive flat-distribution-sampling methods,
   such as the Wang-Landau (WL) algorithm and metadynamics.
@@
 
-To proceed, we first observe that
-the average
-$$
-\bar v = \sum_{i = 1}^n p_i \, v_i
-$$
-increases steadily over time:
-%
-\begin{equation}
-\frac{ d \bar v } { d t }
-=
-\sum_{i = 1}^n p_i \dot v_i
-=
-\alpha(t) \sum_{i = 1}^n h_i(t) = \alpha(t).
+To proceed, we first define an average\cite{
+dama2014},
+\begin{equation}
+  \bar v(t)
+  = \sum_{i = 1}^n \rho_i \, v_i(t)
+\label{eq:vbar_def}
+\end{equation}
+for some distribution,
+$\pmb \rho = (\rho_1, \dots, \rho_n)$.
+%
+This distribution $\pmb \rho$
+will be used to measure the error,
+and is independent of
+the target distribution $\mathbf p$.
+%
+For example,
+in well-tempered metadynamics,
+$\pmb\rho$ is always flat,
+while $\mathbf p$ is an attenuated version of
+$\mathbf p^*$\cite{
+barducci2008, dama2014}.
+%
+Then, we expect the difference %between $v_i$ and $\bar v$,
+%
+\begin{equation}
+  x_i \equiv v_i - \bar v
+  %=
+  %v_i - \sum_{j = 1}^n \rho_j \, v_j
+  ,
+\label{eq:x_def}
+\end{equation}
+%
+to be small in the asymptotic regime.
+%
+It follows that
+\begin{equation}
+  \sum_{i = 1}^n \rho_i \, x_i = 0
+  ,
+\label{eq:rhox_sum}
+\end{equation}
+and
+\begin{equation}
+  \frac{ d \bar v (t) }
+       { d t }
+  =
+  \alpha(t)
+  \sum_{ r = 1 }^n \frac{ h_r(t) } { p_r }
+  .
 \label{eq:dvbardt}
@@
 %
-However, we expect the difference %between $v_i$ and $\bar v$,
-%
-\begin{equation}
-  x_i \equiv v_i - \bar v = v_i - \sum_{j = 1}^n p_j \, v_j,
-  \label{eq:x_def}
-\end{equation}
-%
-to be small in the asymptotic regime.
 Expanding Eq. \eqref{eq:h_ave} in $x_i$ yields
@@
 \begin{align}
+  \frac{
     \langle h_i(t) \rangle
+  }
+  {
+    p_i
+  }
   &\approx
-  \frac{                 p_i \, e^{- x_i} }
+  \frac{                        e^{- x_i} }
        { \sum_{ j = 1}^n p_j \, e^{- x_j} }
   \approx
-  p_i \, (1 - x_i),
+  1 - x_i + \sum_{r = 1}^n p_r \, x_r
+  ,
 \label{eq:hdet_linearize}
@@
 %
-where we have used $\sum_{j=1}^n p_j = 1$,
-and
-%
-\begin{equation}
-  \sum_{i = 1}^n p_i \, x_i = 0,
-  \label{eq:px_sum}
-\end{equation}
-which follows directly from Eq. \eqref{eq:x_def}.
+where we have used $\sum_{j=1}^n p_j = 1$.
+%
 \note{For \eqref{eq:hdet_linearize}. Following the second step
@@
 \frac{                  p_i (1 - x_i) }
-     { \sum_{ j = 1}^n p_j (1 - x_j) }
+     { \sum_{ r = 1 }^n p_r (1 - x_r) }
+=
+\frac{                      p_i (1 - x_i) }
+     { 1 - \sum_{ r = 1 }^n p_r \, x_r }
 .
@@
 \eqref{eq:h_split},
-\eqref{eq:dvbardt}-\eqref{eq:hdet_linearize},
+\eqref{eq:x_def}-\eqref{eq:hdet_linearize},
 we get a set of decoupled equations
@@
   =
-  -\alpha(t) \, \left[ x_i(t) - \frac{ \zeta_i(t) } { p_i } \right].
+  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
+  ,
 \label{eq:dxdt_WL}
+\end{equation}
+where
+\begin{equation}
+  \xi_i(t)
+  =
+  \frac{ \zeta_i(t) }
+       { p_i }
+  -
+  \sum_{ r = 1 }^n
+    \rho_r
+    \frac{ \zeta_r(t) }
+         {     p_r    }
+  .
+\label{eq:xi_def}
 \end{equation}
@@
             {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left( \frac{ h_i } { p_i } - 1 \right)
+  \alpha(t) \left(
+    \frac{ h_i } { p_i }
+    -
+    \sum_{ r = 1 }^n
+      \rho_r \frac{ h_r } { p_r }
+  \right)
   \\
@@
             {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left( \frac{ \langle h_i \rangle } { p_i } + \frac{ \zeta_i } { p_i } - 1 \right)
-  \stackrel{\mathrm{Eq.\; \eqref{eq:hdet_linearize}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left[ -x_i(t) + \frac{ \zeta_i } { p_i } \right].
+  \alpha(t) \left(
+    \frac{ \langle h_i \rangle } { p_i }
+    -
+    \sum_{ r = 1 }^n
+      \rho_r \frac{ \langle h_r \rangle } { p_r }
+    +
+    \xi_i(t)
+  \right)
+  \\
+  &\stackrel{\mathrm{Eqs.\; \eqref{eq:rhox_sum} \;and\; \eqref{eq:hdet_linearize}}}
+            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \alpha(t) \left[ -x_i(t) + \xi_i(t) \right].
 \end{aligned}
@@
 
-The total error of the bias potential can be written in
-terms of $x_i$,
+We define
+the total error of the bias potential as
 \begin{equation}
@@
   =
-  \sum_{i = 1}^n p_i \, \langle x_i^2 \rangle
-  =
-  \sum_{i = 1}^n p_i \, \left\langle (v_i - \bar v)^2 \right\rangle
+  \sum_{i = 1}^n \rho_i \, \langle x_i^2 \rangle
+  =
+  \sum_{i = 1}^n \rho_i \, \left\langle (v_i - \bar v)^2 \right\rangle
   .
@@
 %
+for the distribution $\pmb \rho$
+introduced in Eq. \eqref{eq:vbar_def}.
+%
 It follows that
 \begin{equation}
-  \sum_{i = 1}^n p_i \, w_{ij}
-  =
-  \sum_{i = 1}^n p_j \, w_{ji}
-  = p_j,
+  \sum_{i = 1}^n \rho_i \, w_{ij}
+  =
+  \sum_{i = 1}^n \rho_j \, w_{ji}
+  =
+  \rho_j,
   \label{eq:w_balance}
@@
 %
-i.e., $\mathbf p$ is a left eigenvector of
+i.e., $\mathbf \rho$ is a left eigenvector of
 $\mathbf w$ with eigenvalue $1$.
@@
   &=
-  \delta_{ij} \, p_i,
+  \delta_{ij} \, \rho_i,
   \label{eq:eig_orthonormal_cols}
@@
     \frac{ \phi_{ki} \, \phi_{li} }
-         { p_i }
+         { \rho_i }
   &=
@@
 Particularly, there is always an eigenvector
-corresponding to the target distribution, $\mathbf p$.
+corresponding to the distribution, $\pmb \rho$.
 %
@@
 \begin{equation}
-  \phi_{1i} = p_i,
+  \phi_{1i} = \rho_i,
 \label{eq:eigenmode1}
@@
     \frac{ \phi_{ki} \, \phi_{1i} }
-         { p_i }
+         { \rho_i }
   =
@@
 %
-By using Eq. \eqref{eq:w_balance},
-we can show that
-Eqs. \eqref{eq:dvbardt} and \eqref{eq:px_sum}
-remain true, and
+By using Eqs. \eqref{eq:vbar_def}
+and \eqref{eq:w_balance},
+we can show that Eq. \eqref{eq:dvbardt}
+remains true,
+and by using Eqs.
+\eqref{eq:x_def},
+\eqref{eq:rhox_sum},
+\eqref{eq:hdet_linearize},
+\eqref{eq:xi_def},
+and \eqref{eq:w_sumj}, we get
 %
@@
   &= \alpha(t) \sum_{j=1}^n w_{ij}
-  \left( \frac{ h_j(t) } { p_j }  - 1 \right)
+  \left(
+    \frac{ h_j(t) } { p_j }
+    -
+    \sum_{r = 1}^n
+      \rho_r \frac{ h_r(t) } { p_r }
+  \right)
   \\
@@
   -\alpha(t) \sum_{j = 1}^n
-  w_{ij} \left[ x_j(t) - \frac{\zeta_j (t)}{p_j} \right],
+  w_{ij} \left[ x_j(t) - \xi_j (t) \right]
+  .
 \end{aligned}
 $$
-where
-we have expanded the right-hand side
-in linear terms of $x_j$
-on the second line.
+%where
+%we have expanded the right-hand side
+%in linear terms of $x_j$
+%on the second line.
 
@@
   \dot{ \bar v }
-  &=
-  \sum_{i = 1}^n p_i \, \dot v_i
+  &
+  \stackrel{\mathrm{Eq.\; \eqref{eq:vbar_def}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \sum_{i = 1}^n \rho_i \, \dot v_i
   %
@@
   %
-  \sum_{j = 1}^n \alpha(t) \, \frac{ h_j(t) } { p_j }
-                 \sum_{ i = 1 }^{n} p_i \, w_{ij}
+  \alpha(t)
+  \sum_{j = 1}^n \frac{ h_j } { p_j }
+                 \sum_{ i = 1 }^{n} \rho_i \, w_{ij}
   \\
@@
   {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \sum_{j = 1}^n \alpha(t) \, h_j(t)
-  =
-  \alpha(t).
+  \alpha(t)
+  \sum_{j = 1}^n \rho_j
+    \, \frac{ h_j } { p_j }
+  %
+  \stackrel{\mathrm{Eq.\; \eqref{eq:w_sumj}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \alpha(t)
+  \sum_{j = 1}^n w_{ij}
+  \sum_{r = 1}^n \rho_r
+    \, \frac{ h_r } { p_r }
+  .
 \end{aligned}
 $$
-Equation \eqref{eq:px_sum} follows directly from the definition
-Eq. \eqref{eq:x_def}.
 }
@@
   \sum_{j=1}^n
-    \phi_{kj} \frac{ \zeta_j(t) }
-                   { p_j        }
+    \phi_{kj} \, \xi_j(t)
   .
@@
   \equiv
-  \sum_{i = 1}^n p_i \,
+  \sum_{i = 1}^n \rho_i \,
     \left\langle x_i^2 \right\rangle
@@
 From Eq. \eqref{eq:eigenmode1},
-the first eigenmode, $y_1 = \sum_{j=1}^n p_j \, x_j$,
-vanishes identically by Eq. \eqref{eq:px_sum}.
+the first eigenmode,
+$y_1 = \sum_{j=1}^n \rho_j \, x_j$,
+vanishes identically by Eq. \eqref{eq:rhox_sum}.
 %
@@
   %}
-  %\caption{
-  %  \label{fig:massq}
-  %  Geometric characterization of the optimal schedule,
-  %  $\alpha(t)$,
-  %  via the normalized mass distribution, $m(Q)$.
-  %  %
-  %  \note{The figure was produced by \texttt{doc/fig/massq.gp}.
-  %  }%
-  %}
+  \caption{
+    \label{fig:massq}
+    Geometric characterization of the optimal schedule,
+    $\alpha(t)$,
+    via the normalized mass distribution, $m(Q)$.
+    %
+    \note{
+      The figure was produced by \texttt{tikz}.
+      % Gnuplot version by doc/fig/massq.gp
+    }%
+  }
 \end{center}
@@
 %
-Naturally, $\rho_i = 1/n$ in this case.
-%
 An example is the Gaussian updating scheme
 used in metadynamics.
-%
-Homogeneous updating schemes
-share the same set of eigenvectors
-but not the corresponding eigenvalues.
-%
-They have significantly fewer
-parameters than the general case,
-and the multiplication of the updating matrix,
-$\mathbf w$,
+
+By the translational invariance,
+homogeneous updating schemes
+satisfy detailed balance,
+Eq. \eqref{eq:w_detailedbalance},
+for the flat distribution, $\rho_i = 1/n$,
+i.e., the updating matrix, $\mathbf w$,
+is symmetric.
+%
+Besides, the updating matrices
+share the same set of eigenvectors,
+and thus are completely determined
+by the eigenvalues.
+%
+Further, the multiplication of $\mathbf w$
 can be reduced to a convolution with the window,
@@
     &
-    $4.4(1) \times 10^{-6}$
+    $4.36(9) \times 10^{-6}$
     &
-    $7.1 \times 10^{-7}$
+    $7.1(1) \times 10^{-7}$
     &
@@
     &
-    $3.8(2) \times 10^{-4}$
+    $3.83(7) \times 10^{-4}$
     &
-    $1.8(1) \times 10^{-4}$
+    $1.72(4) \times 10^{-4}$
     &
-    $1.0(1) \times 10^{-4}$
+    $9.1(2) \times 10^{-5}$
     \\
@@
     would roughly maintain the same shape,
-    but join the initial error curve
+    but meet the initial error curve
     at a lower level (not shown).
@@
 in Secs. \ref{sec:Gamma_perfect}
-and \ref{sec:Gamma_onestep},
+and \ref{sec:Gamma_onestep}
+in the asymptotic limit,
 and describe a general method of measuring them
@@
   \left\langle
-    \zeta_i(t) \, \zeta_j(t')
+    \xi_i(t) \, \xi_j(t')
   \right\rangle
-  &=
-  \langle h_i(t) \, h_j(t') \rangle
+  %&=
+  %\left\langle
+  %  \left(
+  %    \frac{ h_i(t) } { p_i }
+  %    -
+  %    \sum_{ r = 1 }^n
+  %      \rho_r \frac{ h_r(t) } { p_r }
+  %  \right)
+  %\right.
+  %\notag \\
+  %&\hphantom{==}\cdot
+  %\left.
+  %  \left(
+  %    \frac{ h_j(t') } { p_i }
+  %    -
+  %    \sum_{ s = 1 }^n
+  %      \rho_s \frac{ h_s(t') } { p_s }
+  %  \right)
+  %\right\rangle
+  %\notag
+  %\\
+  &=
+  \left(
+    \frac{ \delta_{ij} } { p_i }
     -
-  \langle h_i(t) \rangle
-  \langle h_j(t') \rangle
-  \notag
-  \\
-  &=
-  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \,
+    \frac{ \rho_i } { p_i }
+    -
+    \frac{ \rho_j } { p_j }
+    +
+    \sum_{r = 1}^n
+    \frac{ \rho_r^2 } { p_r }
+  \right) \,
   \delta_{t, t'}
@@
 %
-where $\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
-in the continuous limit.
+where
+we have assumed the asymptotic limit
+such that $\langle h_i(t) \rangle = p_i$,
+which leads to,
+by Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def},
+$$
+  \xi_i(t)
+  =
+  \frac{ h_i(t) } { p_i }
+  -
+  \sum_{ r = 1 }^n
+    \rho_r \frac{ h_r(t) } { p_r }
+  .
+$$
+%$\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
+%in the continuous limit.
 %
@@
   \left\langle
-    \zeta_i(t) \, \zeta_j(t')
+    \xi_i(t) \, \xi_j(t')
   \right\rangle
@@
   \left\langle
-    \zeta_i(t)
+    \xi_i(t)
   \right\rangle
   \left\langle
-    \zeta_j(t')
+    \xi_j(t')
   \right\rangle
-  \\
-  &=
-  \bigl\langle h_i(t)  - \langle h_i(t)  \rangle \bigr\rangle
-  \bigl\langle h_j(t') - \langle h_j(t') \rangle \bigr\rangle
   =
@@
   $$
-  \zeta_i(t)
-  = h_i(t) - \langle h_i(t) \rangle
-  = \delta_{i, i(t)} - p_i,
-  $$
-  and
-  $$
-  \begin{aligned}
-  \left\langle
-    \zeta_i(t) \, \zeta_j(t)
-  \right\rangle
-  &=
-  \left\langle
-    \delta_{i, i(t)} \, \delta_{j, i(t)}
-  \right\rangle
-  -
-  p_i \, p_j
-  \\
-  &=
-  \left\langle
-    \delta_{i, i(t)}
-  \right\rangle
-  \delta_{i, j}
-  -
-  p_i \, p_j
-  =
-  p_i \, \delta_{ij}
-  - p_i \, p_j.
-  \end{aligned}
-  $$
-  %
-  The average
-  $\left\langle \delta_{i, i(t)} \right\rangle$ is $p_i$,
-  because as $i(t)$ goes through $1$ to $n$,
-  $\delta_{i, i(t)}$ gives $1$
-  if $i(t)$ matches $i$ with probability $p_i$,
-  or $0$ otherwise.
+  \langle
+    h_i(t) \, h_j(t)
+  \rangle
+  = \delta_{ij} \, p_i
+  ,
+  $$
+  which leads to Eq. \eqref{eq:zz_perfect}.
 }
@@
   \sum_{i,j = 1}^n
-  \frac{ \phi_{ki} } { p_i }
-  \frac{ \phi_{kj} } { p_j }
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  %
-  = \delta_{t, t'},
+  \phi_{ki} \phi_{kj}
+  \langle \xi_i(t) \, \xi_j(t') \rangle
+  %
+  = \sum_{i = 1}^n \frac{ \phi_{ki}^2 } { p_i }
+  \delta_{t, t'},
 \notag
@@
 where we have used
-Eqs. \eqref{eq:eig_orthonormal_rows}-\eqref{eq:ortho1},
+Eqs. \eqref{eq:ortho1},
 \eqref{eq:eta_def},
@@
   \sum_{i,j = 1}^n
-  \frac{ \phi_{ki} } { p_i }
-  \frac{ \phi_{kj} } { p_j }
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  \phi_{ki}
+  \phi_{kj}
+  \langle \xi_i(t) \, \xi_j(t') \rangle
   \\
@@
   \sum_{i,j = 1}^n
-  \frac{ \phi_{ki} } { p_i }
-  \frac{ \phi_{kj} } { p_j }
+  \phi_{ki}
+  \phi_{kj}
   \left(
-    p_i \, \delta_{i, j} - p_i \, p_j
-  \right)
+    \frac{ \delta_{ij} } { p_i }
+    -
+    \frac{ \rho_i } { p_i }
+    -
+    \frac{ \rho_j } { p_j }
+    +
+    \sum_{r = 1}^n
+    \frac{ \rho_r^2 } { p_r }
+  \right) \,
   \, \delta_{t, t'}
-  \\
-  &=
-  \left(
-    \sum_{j = 1}^n
-    \frac{ \phi_{kj}^2 } { p_j }
-    -
-    \sum_{i = 1}^n \phi_{ki}
-    \sum_{j = 1}^n \phi_{kj}
-  \right)
-  \, \delta_{t, t'}
-  \\
-  &=
-  (1 - 0 \cdot 0) \, \delta_{t, t'}.
+  .
   \end{aligned}
@@
   Eq. \eqref{eq:zz_perfect}
-  on the second line,
-  and
+  on the second line.
+  %
+  With Eq. \eqref{eq:ortho1},
+  only the first term in the parentheses survives.
+}
+Particularly, if $p_i = \rho_i$, then, by
 Eq. \eqref{eq:eig_orthonormal_rows}
-  on the last line.
-}
+$$
+\kappa_k(t - t') = 1.
+$$
 
@@
 the matrices $\mathbf A$ and $\mathbf w$
-share the eigenvectors, but not the eigenvalues.
+share the eigenvectors,
+with $\rho_i = p_i = 1/n$.
+%but not the eigenvalues.
 %
@@
   =
-1 -
-2 \sin^2 \frac{ (k - 1) \, g \, \pi }
+  \cos \frac{ (k - 1) \, g \, \pi }
             {       2 \, n        }
@@
     \sum_{ i = 1 }^ n
-      p_i \,
+      \rho_i \,
       \left(
@@
                   { p_i }
+        -
+        \sum_{r = 1}^n \rho_r
+        \ln \frac { H_r }
+                  { p_r }
       \right)^2
   \right\rangle
-%\notag
-%\\
-  %&
+  \notag
+  \\
+  &
   \approx
   \sum_{ i = 1 }^ n
-    p_i \,
+    \rho_i \,
     \left\langle
@@
               { p_i }
-        - 1
+        -
+        \sum_{r = 1}^n \rho_r
+            \frac { H_r }
+                  { p_r }
       \right)^2
@@
     \sum_{ i, j = 1 }^n
-    \left[
-      \frac{ h_i(t) }
-           { p_i    }
-      - 1
-    \right]
-    \left[
-      \frac{ h_j(t') }
-           { p_j     }
-      - 1
-    \right]
+      \xi_i(t) \, \xi_j(t')
+      %\left[
+      %  \frac{ h_i(t) }
+      %       { p_i    }
+      %  - 1
+      %\right]
+      %\left[
+      %  \frac{ h_j(t') }
+      %       { p_j     }
+      %  - 1
+      %\right]
       \sum_{ k = 1 }^n
@@
     \sum_{ i, j = 1 }^n
-    \left[
-      \frac{ h_i(t) }
-           { p_i    }
-      - 1
-    \right]
-    \left[
-      \frac{ h_j(t') }
-           { p_j     }
-      - 1
-    \right]
-    p_i \, \delta_{i, j}
+      \xi_i(t) \, \xi_j(t') \,
+      \rho_i \, \delta_{i, j}
   \\
@@
   \frac{ 1 } { T }
-\sum_{ t = 1 }^T
-    \left[
-      \frac{ h_i(t) }
+  \sum_{ t = 1 }^T \xi_i(t)
+  \;
+  \frac{ 1 } { T }
+  \sum_{ t' = 1 }^T \xi_i(t')
+  \\
+  &=
+  \sum_{ i = 1 }^n
+    \rho_i \,
+      \left(
+        \frac{ H_i }
              { p_i }
-      - 1
-    \right]
-\frac{ 1 } { T }
-\sum_{ t' = 1 }^T
-    \left[
-      \frac{ h_i(t') }
+        -
+        \sum_{r = 1}^n \rho_r
+        \frac{ H_r }
+             { p_r }
+      \right)^2
+  ,
+\end{align*}
+where we have used
+$$
+  \frac{1}{T}
+  \sum_{t = 1}^T \xi_i(t)
+  =
+  \frac{ H_i }
        { p_i }
-      - 1
-    \right]
-\\
-&=
-\sum_{ i = 1 }^n
-  p_i \,
-    \left(
-      \frac{ \bar h_i }
-           { p_i      }
-      - 1
-    \right)^2
+  -
+  \sum_{r = 1}^n \rho_r
+  \frac{ H_r }
+       { p_r }
   .
-\end{align*}
+$$
 }%
@@
 \sum_{ i = 1 }^n
-  \phi_{1i} \, \frac{ h_i(t) - p_i } { p_i }
-=
-\sum_{ i = 1 }^n \left[ h_i(t) - p_i \right]
+  \phi_{1i} \, \left(
+    \frac{ h_i(t) } { p_i }
+    -
+    \sum_{r = 1}^n
+    \rho_r
+    \frac{ h_r(t) } { p_r }
+  \right)
 = 0.

### 2016-05-03 doc/invt.tex
invt.tex: editing gamcmp.gp and introduction

@@
 a distribution, $p^*(z)$,
-along a quantity of interest $z$.
-%
-The negative logarithm,
-$-\ln p^*(z)$,
-defines a free energy,
-or the potential of mean force (PMF).
-%
-Normally, the distribution $p^*(z)$ is
-limited in one or a few narrow $z$ regions.
-%
-Therefore, to capture the global shape
-of the PMF,
-one has to run multiple independent simulations
-under different umbrella or bias potentials
-along $z$\cite{mezei1987, berg1992, lee1993},
-and then piece together the entire curve.
-%
-This solution can be laborious % clumsy in practice
-and it often fails to converge quickly
-for a complex and glassy system,
-because the system can be trapped
-in a local free energy minimum.
-%for a long time.
-
-A better solution is to artificially flatten
-the target distribution along $z$\cite{mezei1987, berg1992, lee1993,
-wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
-%
-This alleviates the above problem of local trapping
-by forcing the system to travel more
-deliberately along $z$,
-which often represents
-a direction of slow motion.
-%
-To achieve a flat distribution, however,
-the method must be able to construct a
-bias potential that cancels the PMF.
+along a collective variable $z$.
+%
+However, $z$
+often represents a slow variable
+with a localized or multi-modal distribution, $p^*(z)$.
+%
+Thus,
+to capture the global shape of the free energy surface
+in terms of the potential of mean force (PMF), $-\ln p^*(z)$,
+an effective strategy is to sample a flat distribution
+along $z$ by introducing a bias potential that
+cancels the PMF.
+%The negative logarithm,
+%$-\ln p^*(z)$,
+%defines a free energy,
+%or the potential of mean force (PMF).
+%%
+%Normally, the distribution $p^*(z)$ is
+%limited in one or a few narrow $z$ regions.
+%%
+%Therefore, to capture the global shape
+%of the PMF,
+%one has to run multiple independent simulations
+%under different umbrella or bias potentials
+%along $z$\cite{mezei1987, berg1992, lee1993},
+%and then piece together the entire curve.
+%%
+%This solution can be laborious % clumsy in practice
+%and it often fails to converge quickly
+%for a complex and glassy system,
+%because the system can be trapped
+%in a local free energy minimum.
+%%for a long time.
+%
+%
+%
+%A better solution is to artificially flatten
+%the target distribution along $z$\cite{mezei1987, berg1992, lee1993,
+%wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
+%%
+%This alleviates the above problem of local trapping
+%by forcing the system to travel more
+%deliberately along $z$,
+%which often represents
+%a direction of slow motion.
+%%
+%To achieve a flat distribution, however,
+%the method must be able to construct a
+%bias potential that cancels the PMF.
 
@@
 are two widely-used and closely-related\cite{micheletti2004}
-techniques for this purpose,
-in which a bias potential is gradually built up
-during simulation.
-%
-These techniques employ updating loops
+flat-distribution-sampling (FDS) techniques
+that build the bias potential on the fly.
+%
+These techniques employ regular updating loops
 to discourage future visits to configurations
@@
 %
-The main difference
-is that in the WL case,
-each updating step is limited to the bin
-containing the current $z$,
-while metadynamics adopts an extended
-Gaussian window
-covering several neighboring bins.
-%
-The latter is more often used in MD simulations
-as it guarantees a smooth profile
-of the bias potential
-suitable for numerical differentiation
-in deriving the bias force.
-%
-However, in both cases, the bias potential
-is updated frequently (usually every few MC or MD steps),
-%which % brings about a side effect of disrupting
-and the fluctuation
-disrupts \note{``disturbs''?} the underlying
-equilibrium sampling dynamics\cite{laio2002}.
-%
-Thus, one often has to reduce
+%The main difference
+%is that in the WL case,
+%each updating step is limited to the bin
+%containing the current $z$,
+%while metadynamics adopts an extended
+%Gaussian window
+%covering several neighboring bins.
+%%
+%The latter is more often used in MD simulations
+%as it guarantees a smooth profile
+%of the bias potential
+%suitable for numerical differentiation
+%in deriving the bias force.
+%
+However,
+since the frequent updates to the bias potential
+%in both cases, the bias potential
+%is updated frequently (usually every few MC or MD steps),
+%and the fluctuation
+disrupts the underlying
+equilibrium sampling dynamics\cite{laio2002},
+one often has to gradually decrease
 the updating magnitude
-over the course of simulation
-to decrease the effect.
-
-The updating schedule,
-or the manner of reducing
+to reduce the effect.
+
+The schedule of reducing
 the updating magnitude,
@@
 the optimal schedule
-for a general adaptive flat-distribution sampling method.
-%
-In this method, the optimal schedule is derived
+for a general adaptive FDS method,
+including the WL algorithm and metadynamics.
+%
+We derive the optimal schedule
 from an equation of motion of a free particle,
-with the initial updating magnitude set to
+and show that the initial updating magnitude
+should be set to
 half of the previous equilibrium value.
@@
 the inverse-time formula,
-with the origin of time modified
-to satisfy the above optimal initial condition.
-%
-For metadynamics, however,
-the optimal schedule is more complex,
-and it can be sensitive to the simulation length
+with a modified origin of time.
+%
+However,
+the optimal schedule for metadynamics is more complex
+and sensitive to the simulation length
 and the width of the Gaussian window.
@@
 %metadynamics may deliver better performance.
+%
+These results may help practitioners
+plan and optimize FDS simulations.
 
@@
 In Sec. \ref{sec:background},
-we first review the basics of
-flat-distribution sampling
+we first review the basics of FDS
 and some known aspects on the optimal schedule.
 %
-This also helps fixing the notations.
+%This also helps fixing the notations.
 %
@@
 $p_i$,
-a flat-distribution sampling (FDS) simulation.
+an FDS simulation.
 %
@@
 If, however, we could ignore these local eigenmodes
-(by explicitly filtering out
-them in the biased potential),
+(by explicitly filtering
+them from the biased potential),
 then we can modify the optimal schedule
@@
     \label{fig:xerr}
-    Error components for the Gaussian updating scheme
+    Predicted error components for the Gaussian updating scheme
     with $\sigma = 10$
@@
 %
-The optimal schedule delivers fastest convergence
-of the bias potential.
+The optimal schedule delivers the fastest convergence
+of the bias potential,
+and thus helps minimize the computational effort
+of the adaptive FDS simulation.
 %They differ by the updating schemes,
@@
 This allows us to assume the sampling process
-to be a quasi-equilibrium one,
+to be a quasi-equilibrium one\cite{barducci2008, dama2014},
 and to represent the autocorrelation function
@@
 As an example, we used the method to compute
-the integrals, $\Gamma_k$'s, from an MD simulation,
-and compared them to those from
-%Eq. \eqref{eq:Gamma_onestep}.
-the local MC sampling process.
+the integrals, $\Gamma_k$'s, from
+a local MC simulation
+(cf. Sec. \ref{sec:results_system})
+and an MD simulation,
+and compared them with those from
+the one-step sampling process.
 %
 As shown in Fig. \ref{fig:gamcmp},
-the larger values of $\Gamma_k$'s
-with smaller $k$'s
-from the two sampling processes
-were almost proportional.
-%
-This suggests that
+the $\Gamma_k$'s of the small-$k$ eigenmodes
+from the three sampling processes
+were almost proportional,
+suggesting that
 the optimal schedule
-from simulations using local sampling processes
+for simulations using local sampling processes
 would be similar in general properties.
@@
     Integrals of the autocorrelation functions
-    from an MD simulation versus those
-    from the local MC simulation.
+    from a local MC simulation
+    and an MD simulation.
     %
@@
 from the second term
-on the right-hand side of Eq. \eqref{eq:vcorr_equil},
+on the right-hand side of Eq. \eqref{eq:vcorr_equil} as
 %

### 2016-05-05 doc/invt.tex
invt.tex: add the upper bound

@@
 on the simplest case of
-single-bin (WL) updating scheme.
-%
-We shall prove the optimality
+single-bin updating scheme
+used by the WL algorithm,
+by proving the optimality
 of the known inverse-time formula.
 %
-Next, we turn to the general case
-of multiple-bin updating schemes
-and propose a method of computing
-the optimal schedule
+Next, we extend the method to the general case
+of a multiple-bin updating scheme
 in Secs. \ref{sec:multiple-bin}
@@
 %
-A necessary condition to sample
+First, to sample
 the desired distribution
-$\mathbf p = (p_1, \dots, p_n)$
-is the following.
-%
-In Eq. \eqref{eq:vt_diffeq_mbin},
-if $h_j(t)$ were the same as $p_j$,
-the $\dot v_i(t)$ should be a constant independent of $i$.
+$\mathbf p = (p_1, \dots, p_n)$,
+the $\dot v_i(t)$
+in Eq. \eqref{eq:vt_diffeq_mbin}
+should be a constant independent of $i$,
+if $h_j(t)$ were the same as $p_j$.
 %
@@
 To simplify the ensuing discussion,
-we shall further impose the
-detailed balance condition:
-%
-\begin{equation}
-  \rho_i \, w_{ij} = \rho_j \, w_{ji}.
+we shall limit ourselves to $\mathbf w$'s
+that satisfy
+the detailed balance condition,
+%
+\begin{equation}
+  \rho_i \, w_{ij} = \rho_j \, w_{ji}
+  ,
   \label{eq:w_detailedbalance}
@@
 the residual error becomes negligible, and
+the $k$th component of the asymptotic error
+is given by
+%
+%\begin{align}
+%  \Err
+%  &=
+%  \Err_A
+%  =
+%  \int_0^{T'}
+%    \sum_{k = 2}^n
+%      \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
+%      e^{ 2 \, \lambda_k \, \alpha_0 \, (t - T') }
+%    \, dt
+%  \notag
+%  \\
+%  &
+%  =
+%  %\stackrel{ T' \to \infty }
+%  %{ =\joinrel=\joinrel=\joinrel= }
+%  %
+%  \sum_{ k = 2 }^n
+%    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0
+%  ,
+%  \qquad
+%  ( T' \to \infty )
+%  .
+%\notag
+%%\label{eq:error_eql}
+%\end{align}
+%
+%In terms of the components, we have
 %
 \begin{align}
-  \Err
+  \left\langle
+    y_k^2
+  \right\rangle
   &=
-  \Err_A
-  =
   \int_0^{T'}
-    \sum_{k = 2}^n
     \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
@@
   \\
-  &
-  =
-  %\stackrel{ T' \to \infty }
-  %{ =\joinrel=\joinrel=\joinrel= }
-  %
-  \sum_{ k = 2 }^n
+  &=
   \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0
@@
   .
-\notag
-%\label{eq:error_eql}
+\label{eq:y2_eql}
 \end{align}
-%
-In terms of the components, we have
-%
-\begin{equation}
-  \left\langle
-    y_k^2
-  \right\rangle
-  =
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0.
-  \label{eq:y2_eql}
-\end{equation}
 %
@@
 %
-for some positive $C$,
+for some positive constant $C$,
 %
@@
 %
-where $C$ is to be determined by
+where $C$ is determined by
 the boundary condition at $t = T$ as
@@
   .
-  \notag
+  \label{eq:mint}
 \end{equation}
@@
 First, we define the mass function as
-$$
+\begin{equation}
   M(Q)
@@
   .
-$$
+\label{eq:mass_func}
+\end{equation}
 and the normalized mass distribution as
@@
 %
-Clearly,
-both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$.
 
@@
 is inversely proportional to the simulation length
-for the same-class schedules,
+for schedules in the same class,
 i.e., we have the scaling relation
@@
   {
-    \partial q(T)
+    \partial Q_c
   }
@@
 $$
-Then solve $\partial (E_R + E_A) / \partial q(T) = 0$.
+Then solve $\partial (E_R + E_A) / \partial Q_c = 0$.
 }%
@@
 
-Besides,
-for simulations with the same $\alpha_0$,
+Now consider
+simulations with the same $\alpha_0$.
+%
+Since both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$,
 Eq. \eqref{eq:opt_qT} shows that
 the optimal $q(T)$ increases with
-the simulation length $T$.
-%
-If $M(Q)$ decays rapidly,
-the distribution $m(Q)$ does not critically
-depend on the cutoff $q(T)$,
-as long as $q(T)$ is sufficiently large.
-%
-So the optimal schedules of these simulations
-approximately satisfy the scaling relations
-described in Sec. \ref{sec:mass_distr}.
-%
-If, however, $M(Q)$ has a long tail
-because of some near-zero $\lambda_k$'s,
-increasing the simulation length $T$
-generally alters the shape of the optimal schedule.
+the simulation length $T$,
+more so if $M(Q)$ has a long tail.
+%
+This means that, if many $\lambda_k$'s are near zero,
+both the optimal schedule and
+normalized asymptotic error, $E_A \cdot T$,
+would be sensitive to the simulation length, $T$.
+%
+%If $M(Q)$ decays rapidly,
+%the distribution $m(Q)$ does not critically
+%depend on the cutoff $q(T)$,
+%once $q(T)$ is sufficiently large.
+%%
+%So the optimal schedules of these simulations
+%approximately satisfy the scaling relations
+%described in Sec. \ref{sec:mass_distr}.
+%%
+%If, however, $M(Q)$ has a long tail
+%because of some near-zero $\lambda_k$'s,
+%increasing the simulation length $T$
+%generally alters the shape of the optimal schedule.
 
@@
 
-The Gaussian updating scheme is commonly
-used in metadynamics simulations.
+We define a Gaussian updating scheme
+as an updating scheme with a
+Gaussian-shaped kernel.
+%
+This updating scheme is commonly
+used in metadynamics.
 %
@@
   .
-\label{eq:error_singlebin}
+\label{eq:error_asym_singlebin}
 \end{align}
@@
 Thus,
-the single-bin scheme is most efficient
+the corresponding updating scheme,
+the single-bin scheme, is most efficient
 in asymptotic convergence.
 %if no eigenvalue $\lambda_k$ is zero.
+%
+In Appendix \ref{sec:upperbound},
+we shall show that the opposite limit of
+a set of well-separated eigenvalues
+will provide an upper bound for
+the asymptotic error, Eq. \eqref{eq:error_asym_ubound}.
 
@@
 we can change the upper limit of the sum in
-Eq. \eqref{eq:error_singlebin} to $K$.
+Eq. \eqref{eq:error_asym_singlebin} to $K$:
 %
@@
 %Eq. \eqref{eq:error_singlebin} is changed to
-%%
-%\begin{equation}
-%  E_A
-%  =
-%  \frac {       T     }
-%        { (T + t_0)^2 }
-%  \sum_{ k = 2 }^K
-%    \Gamma_k.
-%%\notag
-%\label{eq:error_asym_sinc}
-%\end{equation}
+%
+\begin{equation}
+  E_A
+  =
+  \frac {       T     }
+        { (T + t_0)^2 }
+  \sum_{ k = 2 }^K
+    \Gamma_k.
+%\notag
+\label{eq:error_asym_sinc}
+\end{equation}
 %
@@
 the homogeneous case.
+%
+The definition of $K$ will be modified.
 
@@
 %
-and at this point
+and at this point, we get
+from Eqs. \eqref{eq:error_split},
+\eqref{eq:error_res1},
+\eqref{eq:tk0},
+\eqref{eq:error_asym_sinc},
 %
@@
     from the long-wavelength (local) eigenmodes
-    with $K \ge 5$.
+    with $k \ge 5$.
     %
@@
     the error components from the optimal schedule
-    would roughly maintain the same shape,
+    would roughly maintain the relatively flat profile,
     but meet the initial error curve
-    at a lower level (not shown).
+    at a lower level (not shown),
+    where $\lambda_k \approx 1/q(T)$
+    (cf. Appendix \ref{sec:upperbound}).
     %
@@
 First, if the sampling is perfect,
-Eq. \eqref{eq:kappa_delta} is exact, and
-%
-\begin{equation}
-  \Gamma_k = 1.
-\label{eq:Gamma_perfect}
-\end{equation}
-%
-To see this, we note that
+%Eq. \eqref{eq:kappa_delta} is exact, and
+%
+%\begin{equation}
+%  \Gamma_k = 1.
+%\label{eq:Gamma_perfect}
+%\end{equation}
+%
+%To see this, we note that
 the bin index $i$ is an independent random variable
-at each time step $t$ in this case, and
+at each time step $t$, and
 %
@@
   %
-  = \sum_{i = 1}^n \frac{ \phi_{ki}^2 } { p_i }
-  \delta_{t, t'},
+  =
+  \sum_{i = 1}^n \frac{ \phi_{ki}^2 } { p_i }
+  \, \delta_{t, t'},
 \notag
@@
 $$
+Thus, Eq. \eqref{eq:kappa_delta} is exact,
+with
+$$
+\Gamma_k
+=
+\sum_{ i = 1}^n \frac{ \phi_{ki}^2 } { p_i },
+$$
+or if $\mathbf p = \pmb \rho$,
+\begin{equation}
+  \Gamma_k = 1
+  .
+\label{eq:Gamma_perfect}
+\end{equation}
 
@@
 \item
-Compute $q(T)$ from the root of Eq. \eqref{eq:opt_qT}.
+Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
 
@@
 is not necessarily stable,
-i.e., some eigenvalues may be negative.
+i.e., some eigenvalues can be negative.
 %
@@
     If all eigenvalues are nonnegative,
-    we are done, and exit the loop.
+    we are done. % and exit the loop.
   \item
@@
 
+\section{\label{sec:upperbound}
+Asymptotic error for an updating scheme
+with a set of well-separated eigenvalues}
+
+We shall derive here an upper bound of the asymptotic error
+that can be associated to updating schemes
+with a set of well-separated eigenvalues,
+in parallel to the lower bound reachable by
+the single-bin scheme with a set of identical eigenvalues
+(cf. Sec. \ref{sec:optWL}).
+
+Starting from the definition Eq. \eqref{eq:mass_func}, we have
+$$
+  M(Q)
+  \le
+  \sum_{k = 2}^n
+  \sqrt{ \Gamma_k } \, \lambda_k \, e^{ -\lambda_k \, Q }
+  .
+$$
+Thus,
+it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
+\begin{equation}
+  C
+  \le
+  \sum_{k = 2}^n \sqrt{ \Gamma_k }
+  ,
+  \label{eq:mint_ubound}
+\end{equation}
+and the asymptotic error from Eq. \eqref{eq:error_asym2}
+is bounded as
+\begin{equation}
+  E_A
+  =
+  \frac{ C^2 } { T }
+  \le
+  \frac 1 T
+  \left(
+    \sum_{k=2}^n \sqrt{ \Gamma_k }
+  \right)^2
+  .
+  \label{eq:error_asym_ubound}
+\end{equation}
+
+Next, we show that this bound represents
+updating schemes with a set of very different eigenvalues.
+%
+First, observe that the summand of Eq. \eqref{eq:mass_func},
+$\Gamma_k \, \left( \lambda_k e^{-\lambda_k \, Q} \right)^2$
+is a function that peaks at $\lambda_k = 1/Q$,
+so that the sum is dominated by the term at $k = k(Q)$ with
+%
+\begin{equation}
+  \lambda_{k(Q)}
+  \approx
+  \frac 1 Q
+  .
+  \label{eq:kQ_def}
+\end{equation}
+%
+Thus,
+\begin{align}
+  M(Q)
+  &
+  \gtrsim
+  %\approx
+  \sqrt{ \Gamma_{ k(Q) } } \,
+  \lambda_{ k(Q) } \,
+  e^{ -\lambda_{ k(Q) } \, Q }
+\label{eq:MQ_lbound1}
+  \\
+  &
+  \approx
+  \frac{ \sqrt{ \Gamma_{ k(Q) } } }
+       { Q }
+  e^{ -\lambda_{ k(Q) } \, Q }
+  .
+\label{eq:MQ_lbound}
+\end{align}
+
+Now consider the $k$th error component, % of the asymptotic error,
+\begin{equation}
+  \left\langle
+    y_k^2
+  \right\rangle_A
+  =
+  \frac C T \,
+  \Gamma_k \, \lambda_k^2
+  \int_0^\infty
+    \frac{ e^{-2 \, \lambda_k \, Q} }
+         { M(Q) }
+    \, dQ
+  .
+\notag
+\end{equation}
+%
+Using Eq. \eqref{eq:MQ_lbound}, we have
+$$
+\begin{aligned}
+  \left\langle
+    y_k^2
+  \right\rangle_A
+  &
+  \lesssim
+  \frac C T \, \Gamma_k \, \lambda_k^2
+  \int_0^\infty
+    \frac{ Q } { \sqrt{ \Gamma_{ k(Q) } } }
+    e^{-\left[ 2\,\lambda_k - \lambda_k(Q) \right] Q }
+    \, dQ
+  \\
+  &
+  \approx
+  \frac C T \,
+  \frac{ \Gamma_k } { \sqrt{ \Gamma_{ k(Q^*) } } }
+  \left[
+    \frac{ \lambda_k } { 2 \, \lambda_k - \lambda_{ k(Q^*) } }
+  \right]^2
+  ,
+\end{aligned}
+$$
+where we have approximated $k(Q)$
+by its value around the peak of the integrand
+$$
+  Q^*
+  =
+  \frac{ 1 } { 2 \, \lambda_k - \lambda_{ k(Q^*) } }
+  .
+$$
+%
+Comparing this to Eq. \eqref{eq:kQ_def},
+we identify $k(Q^*) \approx k$, so
+\begin{equation}
+  \left\langle
+    y_k^2
+  \right\rangle_A
+  \lesssim
+  %\approx
+  \frac C T
+  \sqrt{ \Gamma_k }
+  .
+  \label{eq:y2_ubound}
+\end{equation}
+%
+This result shows that with a set of well-separated eigenvalues,
+the asymptotic error profile would be
+independent of $\lambda_k$'s and
+be flatter than the initial one from
+Eq. \eqref{eq:y2_eql}
+(cf. Fig. \ref{fig:xerr}
+for a Gaussian example).
+%
+Note that we can turn Eq, \eqref{eq:y2_ubound}
+to an inequality by choosing $k(Q) \to k$
+for Eq. \eqref{eq:MQ_lbound1}.
+%
+Then by summing over $k$
+and using Eq. \eqref{eq:mint_ubound},
+we recover Eq. \eqref{eq:error_asym_ubound}.
+
+For a finite-length simulation,
+the integrals $\int_0^{q(T)} dQ$
+can be extended to infinity
+only if $q(T) \gg Q^*$.
+%
+In other words,
+the above asymptotic limit fails around
+$Q^* \approx q(T)$, or equivalently
+around mode $k^* = k\bigl( q(T) \bigr)$,
+where $\lambda_{ k^* } \approx 1/q(T)$.
+%
+\note{This case also exemplifies
+  the sensitivity to the simulation length.
+  %
+  In this limit,
+  $M(Q) \propto 1/Q$ by Eq. \eqref{eq:MQ_lbound},
+  and $C \propto \ln\bigl[ q(T) / q_0 \bigr]$
+  for some $q_0$,
+  and we expect that with fixed $\alpha_0$,
+  $q(T)$ to increase almost linearly with $T$
+  by Eq. \eqref{eq:opt_qT}.
+  %
+  So the normalized asymptotic error,
+  $E_A \cdot T \propto \bigl( \ln[ q(T)/q_0 ] \bigr)^2 \sim (\ln T)^2$,
+  is sensitive to the simulation length,
+  and worse than the ideal case of a constant.
+}
+
 \section{\label{sec:equilerr}
@@
 Below, we shall show that this practice is unnecessary
-if the simulation uses the single-bin updating scheme
+if the simulation uses the single-bin or a bandpass updating scheme
 with the optimal schedule, Eq. \eqref{eq:alpha_invt1}.
@@
 
-For perfect sampling,
+For perfect sampling, with $\mathbf p = \pmb \rho$,
 $\Gamma(p) = 1$ from Eq. \eqref{eq:Gamma_perfect},

### 2016-05-06 doc/invt.tex
small edits

@@
-\documentclass[reprint, floatfix]{revtex4-1}
+\documentclass[preprint, floatfix]{revtex4-1}
 \usepackage{amsmath}
@@
 along $z$ by introducing a bias potential that
-cancels the PMF.
+cancels the PMF\cite{mezei1987, berg1992, lee1993,
+wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
 %The negative logarithm,

### 2016-05-09 doc/old/invt_annotatedV2.tex
changes according to Dr. Pettitt's modifications

--- 2016-05-05 doc/invt.tex
@@
   of the updating magnitude
-  for adaptive flat-distribution-sampling methods,
-  such as the Wang-Landau (WL) algorithm and metadynamics.
+  for the class of adaptive sampling methods including
+   the Wang-Landau (WL) algorithm and metadynamics.
   %
@@
   construct a bias potential to offset
-  the potential of mean force
-  by incremental updating loops.
+  the potential of mean force by adaptively updating a histogram.
   %
   The updating magnitude should decrease over time
-  to reduce the fluctuation of the bias potential,
-  and the optimal schedule optimizes
+  to help reduce the noise in the bias potential,
+  and the optimal schedule maximizes
   the rate of convergence.
@@
   and the optimal schedule of these schemes
-  is given by a modified inverse-time formula.
+  is a function of the inverse of the time or number of sampling steps.
   %
@@
 Given a system,
-the task is to compute,
-via either Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
-and molecular dynamics\cite{frenkel, karplus2002} (MD) simulations,
+often the task is to compute
 a distribution, $p^*(z)$,
-along a collective variable $z$.
+along a collective variable $z$ by sampling via either Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
+or molecular dynamics\cite{frenkel, karplus2002} (MD) simulations.
 %
@@
 often represents a slow variable
-with a localized or multi-modal distribution, $p^*(z)$.
+with a localized or multi-modal distribution.
 %
@@
 in terms of the potential of mean force (PMF), $-\ln p^*(z)$,
-an effective strategy is to sample a flat distribution
-along $z$ by introducing a bias potential that
-cancels the PMF.
+an effective strategy is to introduce a bias potential that
+cancels the PMF, which is equivalent to sampling a flat distribution
+along $z$ \cite{mezei1987, berg1992, lee1993,
+wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
 %The negative logarithm,
@@
 flat-distribution-sampling (FDS) techniques
-that build the bias potential on the fly.
-%
-These techniques employ regular updating loops
-to discourage future visits to configurations
-with the current $z$ value
-by elevating the bias potential there.
+that build the bias potential adaptively or on the fly.
+%
+These techniques employ regular updating of the bias potential
+to discourage future visits to the current configuration
+by incrementally elevating the bias potential there.
 %
@@
 However,
-since the frequent updates to the bias potential
+ frequent updates to the bias potential
 %in both cases, the bias potential
@@
 disrupts the underlying
-equilibrium sampling dynamics\cite{laio2002},
-one often has to gradually decrease
-the updating magnitude
-to reduce the effect.
+sampling\cite{laio2002},
+and near convergence
+one  has to gradually decrease
+the updating magnitude.
 
@@
 that the optimal schedule
-is given by the product of the
+is a function of the
 inverse time\cite{liang2007,
@@
 morozov2007, zhou2008}
-(with ``time'' being
-the number of simulation steps).
+or
+the number of simulation steps.
 %
@@
 is shown to be given by
-the inverse-time formula,
+an inverse-time formula,
 with a modified origin of time.
@@
 to a class of bandpass updating schemes
-that are optimal for long simulations,
-and the modified inverse-time schedule is optimal for these schemes.
+that are optimal for extensive sampling or long simulations,
+and the modified inverse-time schedule presented is optimal for these schemes.
 %
@@
 %
-These results may help practitioners
+These results allow one to better
 plan and optimize FDS simulations.
@@
 we first review the basics of FDS
-and some known aspects on the optimal schedule.
+and some known aspects of the optimal schedule.
 %
@@
 computing the optimal schedule
-on the simplest case of
+in the simple case of the
 single-bin updating scheme
@@
 are optimal for convergence
-in the long time limit
+in the long time limit or extensive sampling
 in Sec. \ref{sec:cmpschemes}.
@@
 For example, $i$ can be the energy, $E$,
-in a lattice spin model; or the temperature index
+in a lattice spin model or the temperature index
 in a simulated tempering simulation\cite{
@@
 
-For a large system,
-the distribution $p^*_i$ tends to
+For a large system at tempertures of interest,
+the distribution $p^*_i$ often tends to
 be localized,
 %
-and to find out the global shape
+and to explore the global shape
 of the PMF, $-\ln p^*_i$,
 it is often advantageous to carry out
-a biased sampling that targets
+biased sampling that targets
 a wider distribution, $p_i$.
@@
 %
-Particularly,
+In particular,
 to make $\pi_i$ flat,
@@
 %
-In practice, $V_i$ may contain error,
+In practice, $V_i$ may be non optimal or contain error,
 and thus $\pi_i$ need not
@@
 %
-The estimated bias potential, $V_i$,
+The estimated bias potentials at bin $i$, $V_i$,
 are often improved
-by the histogram, $H_i$, accumulated
+with histogram of the sampling, $H_i$, accumulated
 during simulation.
@@
 %
-These methods update the bias potential
-incrementally in regular intervals
-and can easily achieve the target distribution.
+These methods incrementally update the bias potential
+and can thus approximate the target distribution.
 %
@@
 %they are less well understood theoretically, as
-the updating loops break the microscopic reversibility
-of the underlying sampling mechanism.
+as mentioned, the updates  break the microscopic reversibility
+of the underlying sampling mechanism. Further failure to properly reduce the size of the histogram increments would lead to roughness or errors in the final estimate.
 %
@@
 To minimize this effect,
-many adaptive methods gradually
+many adaptive methods adopt a schedule to
 reduce the updating magnitude
@@
 Thus, in late stages of a long simulation,
-the updating becomes negligible and
+the updating becomes smaller and
 %the adaptive simulation tends to
@@
 if the bias potential were converged to the exact value,
-this adaptive simulation tends to the ideal equilibrium simulation
+this adaptive simulation tends to  ideal flat equilibrium sampling
 %particular equilibrium simulation can be thought as
@@
 
-For later convenience,
+For  convenience,
 we introduce a shifted bias potential
@@
 the bias potential, $v_i(t)$, is updated
-in each MC step $t$ as
+at each sampling step $t$ as
 %
@@
 (or $\ln f$ there)
-is controlled stage-wise.
-%
-In each stage, $\alpha$
-is kept as a constant,
+is controlled in stages.
+%
+At each stage, $\alpha$
+is kept  constant,
 and the histogram along $i$
@@
 %
-This scheme works well for early stages.
-%
-However, it tends to reduce $\alpha$
+While this scheme works well for early stages,
+%
+it tends to reduce $\alpha$
 too quickly in late stages, making the asymptotic error
@@
 There are, however, some reservation about
-the constant of proportionality:
+the constant of proportionality;
 the optimal formula for $\alpha(t)$
 might be $c/t$ with a $c$
-different from $1.0$\cite{
+different from $1$\cite{
 morozov2007, zhou2008, morozov2009}.
@@
 Here we will show that the ideal $c$
-is $1.0$, but the origin of $t$ should be shifted
+is unnity, but the origin of $t$ should be shifted
 to accommodate the initial error.
@@
 %
-Our aim below is to find the optimal schedule $\alpha(t)$,
+Our aim here is to find the optimal schedule $\alpha(t)$,
 for a general updating scheme.
@@
 
-We first approximate Eq. \eqref{eq:wl_update}
+We first approximate the finite difference Eq. \eqref{eq:wl_update}
 by a differential equation
@@
 Next, we split $h_i(t)$ into a deterministic part
-and a noise part,
+and a stochastic part,
 %
@@
 %
-For the noise part, we have
+For the random noise part, we have
 $\langle \zeta_i(t) \rangle = 0$,
@@
 %
-Fortunately, it suffices to consider
-a simpler one-variable problem.
+It suffices to consider
+a simple one-variable problem.
 
@@
 For example, for a white noise,
-$\kappa(t)$ is proportional to
-Dirac's $\delta$-function, $\delta(t)$.
+$\kappa(t)$ is proportional to a
+Dirac $\delta$-function, $\delta(t)$.
 %
@@
 We now fix the endpoint value $q(T)$
-and variate the curve, $q(t)$ ($0 < t < T$).
-%
-This is equivalent to variating $u$
+and vary the curve, $q(t)$ ($0 < t < T$).
+%
+This is equivalent to varying $u$
 with fixed endpoints,
@@
 $\mathbf p = (p_1, \dots, p_n)$,
-the $\dot v_i(t)$
+ $\dot v_i(t)$
 in Eq. \eqref{eq:vt_diffeq_mbin}
@@
 To simplify the ensuing discussion,
-we shall limit ourselves to $\mathbf w$'s
+we  limit ourselves to $\mathbf w$'s
 that satisfy
@@
 we can diagonalize $\mathbf w$ with a set of
-eigenvectors
+eigenvectors, $\phi_{ki}$,
 %
@@
 
-Particularly, there is always an eigenvector
+In particular, there is always an eigenvector
 corresponding to the distribution, $\pmb \rho$.
@@
 
-Now by diagonalizing the matrix $\mathbf w$
+Diagonalizing the matrix $\mathbf w$
 using the eigenvectors defined in Eqs.
@@
 The error function defined in Eq. \eqref{eq:error_sum}
-can be rewritten in terms of the eigenmodes
+can be rewritten in terms of the eigenmodes of $w$
 \begin{align}
@@
 
-In analogous to Eq. \eqref{eq:x2t_average},
-we can write,
-from Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
+In analogy to Eq. \eqref{eq:x2t_average},
+using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum}, we can write
 the error after a period $T$,
@@
 the underlying sampling process
-affects the error only through the numbers,
-$\Gamma_k$'s.
+affects the error only through the set of
+$\Gamma_k$.
 %
@@
 
-We shall use this result to model the initial error
+We will use this result to model the initial error
 before starting the schedule $\alpha(t)$ to be optimized.
@@
 for $t < 0$, and
-the system has entered the equilibrium
+the system has entered the sampling
 under a fixed $\alpha_0$ at $t = 0$.
@@
 %
-Thus, if we variate $\alpha(t)$ under a fixed $q(T)$,
+Thus, if we vary $\alpha(t)$ under a fixed $q(T)$,
 this residual part can be ignored,
@@
 %
-We can rewrite Eq. \eqref{eq:error_asym1} as an action
+We can rewrite Eq. \eqref{eq:error_asym1} much like an action
 %
@@
 with a position-dependent mass.
+
+%(Cheng: In the Hamiltonian or Newtonian version it seems that the position dependence of the mass could as easily be described by another conservative force at constant mass.)
 %
@@
 %
-These numbers constitutes an \emph{updating kernel}.
+These numbers constitute an \emph{updating kernel}.
 
 The above matrix, $\mathbf w$,
-being symmetric to the indices $i$ and $j$,
+being symmetric with respect to the indices $i$ and $j$,
 satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
@@
 but
-the cutoff $b$ only need to be less than $n$.
+the cutoff $b$ only needs to be less than $n$.
 %
@@
 %
-Below we shall show its stability
+Below we F show its stability
 in the continuous limit.
@@
 compared to a Gaussian
-with a similar width, $\sigma$,
+with a similar width, $\sigma$, (cheng, the inset in figure two should be replotted to fix the dynamic range of the gray scale.)
 %
@@
 
-In our test system,
+In our test system, (Cheng, we should give a  description of the test system)
 $i$ represents a one-dimensional
@@
 
-We tested the theory of computing errors
+We tested our analysis of computing errors
 on the modified inverse-time schedule,
@@
 Adaptive FDS methods
-effectively sample a flat distribution
-by incrementally updating the bias potential
-to offset the PMF.
+ incrementally update a bias potential
+to offset the PMF and thereby sample a flat distribution.
 %
@@
 in the long time limit
-without a priori assumption
-on the smoothness of the PMF.
+without a priori assumptions
+about the smoothness of the PMF.
 %
@@
 The optimal schedule of these bandpass schemes
-is always given by the inverse-time formula.
+is always given by an inverse-time formula.
 
@@
 
-Strictly speaking, our method is valid
-only for long simulations.
+Our method is most useful for long simulations to achieve relatively precise answers.
 %
@@
 they may break down for relatively short simulations
-of glassy systems.
+of glassy systems. Our immediate interest is to apply these relations to
+problems containing aqueous mixtures of proteins and nucleic acids.
 
@@
 for helpful discussions.
+We gratefully acknowledge the Robert A.
+Welch Foundation (H-0037), the National Science Foundation (CHE-1152876) and the National Institutes of Health (GM-037657) for partial support of this work.
+This research used computing resources of the National Science Foundation XSEDE grid.
 
@@
 ``$\operatorname{var}$''
-means the variance from time averaging,
+means the variance from the sampling (time) average,
 and
@@
 \section{\label{sec:program}
-Program of computing the optimal schedule and error
-}
-
-Here we give a program of computing
+Procedure for computing the optimal schedule and error
+}
+
+Here we give a procedure for computing
 the optimal schedule and the error
@@
 from the input,
-and then feed them to Eq. \eqref{eq:q_opt}
+and then apply them to Eq. \eqref{eq:q_opt}
 for the schedule,
@@
 to measure $\Gamma_k$'s,
-we have the following program.
+we have the following procedure.
 

### 2016-05-24 doc/notes/1p2T.tex
editing 1p2T.tex

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{hyperref}

\hypersetup{
    colorlinks,
    linkcolor={red!30!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\begin{document}

\title{Statistical Redundancy $1 + 2 \, \tau$}
\author{}
\date{\vspace{-10ex}}
\maketitle

\section{Introduction}

In molecular simulations, we compute trajectory averages.
%
The simulation engine
%running either Monte Carlo (MC) or molecular dynamics (MD),
generates a trajectory of molecular configurations,
and saves the trajectory frames in regular intervals.
%
From the trajectory,
we can compute the average of a quantity of interest, $x$,
%
\begin{equation}
  \bar x = S / n,
  \notag
  %\label{eq:xbar}
\end{equation}
%
and
%
\begin{equation}
  S = x_1 + \cdots + x_n,
  \notag
  %\label{eq:sumx}
\end{equation}
%
where $n$ is the number of saved frames,
and $x_i$ is the value of $x$ from frame $i$.

However, successive trajectory frames are usually correlated.
%
While this correlation does not affect the expected value
of the average, it affects the variance, which represents
the statistical redundancy of the trajectory.

There is, however, a simple formula
which accurately measures the statistical redundancy, $R$,
relative to a perfectly independent trajectory:
%
\begin{equation}
  R = 1 + 2 \, \tau,
  \label{eq:1p2T}
\end{equation}
%
where $\tau$ is the integral autocorrelation time measured
in multiples of frame separations.
%
In this article, we shall show why Eq. \eqref{eq:1p2T} is true.

\section{Proof}

We shall first sketch the proof of Eq. \eqref{eq:1p2T}.
%
For simplicity, we assume that
%
\begin{equation}
  \left\langle x_i \right\rangle = 0,
  \label{eq:ave_xi}
\end{equation}
%
and
%
\begin{equation}
  \left\langle x_i^2 \right\rangle = 1.
  \label{eq:var_xi}
\end{equation}
%
This can always be done by a proper scaling and translation.
%
Our argument has three steps.

\begin{enumerate}
\item
The integral autocorrelation time is given by
%
\begin{equation}
  \tau
  =
  \sum_{k = 1}^\infty c_k
  ,
  \label{eq:act}
\end{equation}
%
where $c_k$
is the normalized autocorrelation function with $c_0 = 1$, and
\begin{equation}
  c_k = \langle x_{i + k} \, x_i \rangle
  .
  \label{eq:ck_def}
\end{equation}

\item
The statistical redundancy is given by
%
\begin{equation}
  R
  = \sum_{k = -\infty}^\infty c_k
  = 1 + 2 \, \sum_{k = 1}^\infty c_k
  ,
  \label{eq:redund}
\end{equation}
%

\item
The desired result, Eq. \eqref{eq:1p2T},
follows from Eqs. \eqref{eq:act} and \eqref{eq:redund}.

\end{enumerate}

\section{Integral autocorrelation time}

We shall prove Eq. \eqref{eq:act} in this section
by building an underlying model for the correlation.
%
In each frame,
$x_{ i + 1 }$
can either inherit the previous value, $x_i$
with probability $c_1$,
or be an independent data point
with probability $c_0 - c_1 = 1 - c_1$.
%
In the former case,
$x_{i + 2}$ may hold on to the value of $x_i$,
with a smaller probability $c_2 \le c_1$,
and so on.
%
Generally,
the probability for $x_i$ to survive in the sequence for precisely
the next $k$ frames is given by
%
\begin{equation}
  p_k = c_k - c_{k + 1}
  .
  \notag
  %\label{eq:pk}
\end{equation}
%
So we can define the integral autocorrelation time as
the average of the life span, $k$,
%
$$
  \tau = \sum_{k = 0}^\infty k \, p_k.
$$

To show Eq. \eqref{eq:act} is true, we note that
$$
\begin{aligned}
\tau
&=
\sum_{ k = 0 }^\infty k \, c_k
- \sum_{ k = 0 }^\infty k \, c_{k + 1}
\\
&=
\sum_{ k = 1 }^\infty k \, c_k
- \sum_{ k = 0 }^\infty (k + 1 - 1) \, c_{k + 1}
\\
&=
\sum_{ k = 1 }^\infty k \, c_k
-
\sum_{ k = 0 }^\infty (k + 1) \, c_{ k + 1 }
+
\sum_{ k = 0 }^\infty c_{k + 1}
\\
&=
\sum_{ k = 1 }^\infty c_k,
\end{aligned}
$$

\section{\label{sec:redund}
Statistical redundancy}

In this section,
we shall prove Eq. \eqref{eq:redund}.
%
We shall first give an intuitive definition
of the statistical redundancy
and then show it is given by the variance.
%
Consider two trajectories generated by
different simulation methods with the same length.
%
The two simulations produce
$\bar x$ and $\bar x'$ for the average,
or
$S$ and $S'$ for the sum, respectively.
%
Since the two values are computed for the same system,
they share the same expectation.
%
But as the simulation methods differ, so are the variances.

We wish to obtain a better estimate of the sum, $\hat S$,
by a linear combination of $S$ and $S'$, i.e.,
\begin{equation}
  \hat S
  =
  c \, S + c' \, S',
  \notag
\end{equation}
%
with normalized coefficients
%
\begin{equation}
  c + c' = 1.
  \label{eq:csum}
\end{equation}
%
We wish to minimize the variance of $\hat S$ under the constraint
Eq. \eqref{eq:csum}.  Then the target function is
$$
\operatorname{var} \hat S - \lambda \, (c + c')
=
c^2 \, \operatorname{var} S
+
c'^2 \, \operatorname{var} S'
-
\lambda \, (c + c'),
$$
with $\lambda$ being a Lagrange multiplier.
%
Differentiating the expression with respect to $c$ and $c'$ yields
$$
2 \, c \, \operatorname{var} S
=
2 \, c' \, \operatorname{var} S'
=
\lambda,
$$
which shows that the coefficients of combination
are inversely proportional to the respective variances.
%
In other words, the variance serves as the discounting factor
for statistical redundancy in an optimal combination.

We can now show Eq. \eqref{eq:redund},
%
$$
\begin{aligned}
\operatorname{var} S
=
\left\langle S^2 \right\rangle
&=
\sum_{i = 1}^n
  \sum_{j = 1}^n
    \left\langle x_i \, x_j \right\rangle
\\
&=
\sum_{i = 1}^n
\left(
\sum_{j = 1}^n
c_{ |j - i| }
\right),
\end{aligned}
$$
%
where we have used Eq. \eqref{eq:ck_def}
in the last step.
%
For a very large $n$,
the inner sum is peaked around $j = i$,
and except for a few $i$ around $1$ or $n$,
we have
$$
\sum_{j = 1}^n c_{ | j - i | }
\approx
\cdots + c_2 + c_1 + 1 + c_1 + c_2 + \cdots,
$$
where the sum extends to infinity in both ends.
%
Thus,
\begin{equation}
  \operatorname{var} S
  =
  n \times ( 1 + 2 \, c_1 + 2 \, c_2 + \cdots ).
  \label{eq:varS}
\end{equation}
If the data points are independent, then $c_1 = c_2 = \cdots = 0$,
and this serves as the reference value
\begin{equation}
  \operatorname{var} S^\mathrm{perfect}
  =
  n
  .
\end{equation}
%
The ratio of the two yields the relative statistical redundancy,
$$
R \equiv
\frac{ \operatorname{var} S }
     { \operatorname{var} S^{ \mathrm{perfect} } }
=
1 + 2 \, c_1 + 2 \, c_2 + \cdots
,
$$
which is Eq. \eqref{eq:redund}.

\end{document}

### 2016-05-30 doc/invt.tex
small changes

--- 2016-05-05 doc/invt.tex
@@
 \usepackage{hyperref}
+\usepackage{setspace}
 
@@
   of the updating magnitude
-  for adaptive flat-distribution-sampling methods,
-  such as the Wang-Landau (WL) algorithm and metadynamics.
+  for the class of adaptive sampling methods including
+  the Wang-Landau (WL) algorithm and metadynamics.
   %
@@
   construct a bias potential to offset
-  the potential of mean force
-  by incremental updating loops.
+  the potential of mean force by histogram-based adaptive updates.
   %
   The updating magnitude should decrease over time
-  to reduce the fluctuation of the bias potential,
-  and the optimal schedule optimizes
+  to help reduce the noise in the bias potential,
+  and the optimal schedule maximizes
   the rate of convergence.
@@
   and the optimal schedule of these schemes
-  is given by a modified inverse-time formula.
+  is the inverse of the time or number of sampling steps with a proper shift.
   %
@@
 in terms of the potential of mean force (PMF), $-\ln p^*(z)$,
-an effective strategy is to sample a flat distribution
-along $z$ by introducing a bias potential that
-cancels the PMF.
+an effective strategy is to introduce a bias potential that
+cancels the PMF, which is equivalent to sampling a flat distribution
+along $z$\cite{mezei1987, berg1992, lee1993,
+wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
 %The negative logarithm,
@@
 However,
-since the frequent updates to the bias potential
+frequent updates to the bias potential
 %in both cases, the bias potential
@@
 are optimal for convergence
-in the long time limit
+in the long time or extensive sampling limit
 in Sec. \ref{sec:cmpschemes}.
@@
 
-For a large system,
-the distribution $p^*_i$ tends to
+For a large system at temperatures of interest,
+the distribution $p^*_i$ often tends to
 be localized,
 %
-and to find out the global shape
+and to explore the global shape
 of the PMF, $-\ln p^*_i$,
 it is often advantageous to carry out
-a biased sampling that targets
+biased sampling that targets
 a wider distribution, $p_i$.
@@
 %
-In practice, $V_i$ may contain error,
+In practice, $V_i$ may be non-optimal or contain error,
 and thus $\pi_i$ need not
@@
 %
-The estimated bias potential, $V_i$,
+The estimated bias potentials at bin $i$, $V_i$,
 are often improved
-by the histogram, $H_i$, accumulated
+with the histogram of the sampling, $H_i$, accumulated
 during simulation.
@@
 %
-These methods update the bias potential
-incrementally in regular intervals
-and can easily achieve the target distribution.
+These methods incrementally update the bias potential
+and can approximately sample the target distribution.
 %
@@
 To minimize this effect,
-many adaptive methods gradually
+many adaptive methods adopt a schedule to 
 reduce the updating magnitude
@@
 if the bias potential were converged to the exact value,
-this adaptive simulation tends to the ideal equilibrium simulation
+this adaptive FDS tends to the ideal equilibrium counterpart,
 %particular equilibrium simulation can be thought as
@@
 (or $\ln f$ there)
-is controlled stage-wise.
-%
-In each stage, $\alpha$
-is kept as a constant,
+is controlled in stages.
+%
+At each stage, $\alpha$
+is kept constant,
 and the histogram along $i$
@@
 Here we will show that the ideal $c$
-is $1.0$, but the origin of $t$ should be shifted
+is unity, but the origin of $t$ should be shifted
 to accommodate the initial error.
@@
 The error function defined in Eq. \eqref{eq:error_sum}
-can be rewritten in terms of the eigenmodes
+can be rewritten in terms of the eigenmodes of $\mathbf w$
 \begin{align}
@@
 the underlying sampling process
-affects the error only through the numbers,
+affects the error only through the
 $\Gamma_k$'s.
@@
 for $t < 0$, and
-the system has entered the equilibrium
+the system has entered the sampling equilibrium
 under a fixed $\alpha_0$ at $t = 0$.
@@
 %
-A complete program of computing the optimal schedule
-is listed in Appendix \ref{sec:program}.
+A complete procedure of computing the optimal schedule
+is listed in Appendix \ref{sec:procedure}.
 
@@
 %
-Below we shall show its stability
+Below we show its stability
 in the continuous limit.
@@
 with a similar width, $\sigma$,
+\note{(cheng, the inset in figure two should be replotted to fix the dynamic range of the gray scale.)}
 %
@@
 
-In our test system,
-$i$ represents a one-dimensional
-non-periodic variable,
+\note{(Cheng, we should give a  description of the test system)}
+Our test system is a one-dimensional model system, in which
+$i$ represents a non-periodic variable,
 and $n = 100$.
@@
 %
-We repeat the above process $1000$ times
+We repeat the above process $1000$ times (unless specified otherwise)
 and average the results.
@@
 %Eq. \eqref{eq:alpha_invtlambda}.
-Unless specified otherwise,
-the above parameters were used.
 
@@
 they may break down for relatively short simulations
-of glassy systems.
+of glassy systems.  Our immediate interest is to apply these relations to
+problems containing aqueous mixtures of proteins and nucleic acids.
 
@@
 
-We thank Dr. J. Ma, Dr. Y. Mei,
+We thank Dr. J. Ma, Dr. Y. Mei, J. Wang,
 J. Drake, O. Nassar, Dr. C. Lai and Dr. S. Ou
 for helpful discussions.
+We gratefully acknowledge the Robert A.
+Welch Foundation (H-0037), the National Science Foundation (CHE-1152876) and the National Institutes of Health (GM-037657) for partial support of this work.
+This research used computing resources of the National Science Foundation XSEDE grid.
 
@@
 
-\section{\label{sec:program}
-Program of computing the optimal schedule and error
-}
-
-Here we give a program of computing
+\section{\label{sec:procedure}
+Procedure for computing the optimal schedule and error
+}
+
+Here we give a procedure for computing
 the optimal schedule and the error
@@
 to measure $\Gamma_k$'s,
-we have the following program.
+we have the following procedure.
 
 \begin{figure}[h]
+\begin{singlespace}
+  \fontsize{10pt}{\baselineskip}
   \begin{tikzpicture}
@@
 \end{tikzpicture}
+\end{singlespace}
 \caption{

### 2016-06-09 doc/invt.tex
fix a scanave.py bug
invt.tex: add table of symbols

@@
 \usepackage{upgreek}
-\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage[table,usenames,dvipsnames]{xcolor}
 \usepackage{tikz}
@@
 \definecolor{DarkPurple}{RGB}{64,0,42}
+\definecolor{LightGray}{gray}{0.85}
 % annotation macros
@@
   .
-  \label{eq:eta_def}
+  \label{eq:theta_def}
 \end{align}
@@
     The points were obtained from averaging results
-    from 1000 independent simulations;
+    from over 10000 independent simulations;
     the curves are theoretical predictions.
@@
 Eqs. \eqref{eq:ortho1},
-\eqref{eq:eta_def},
+\eqref{eq:theta_def},
 and
@@
   we have used
-  Eq. \eqref{eq:eta_def}
+  Eq. \eqref{eq:theta_def}
   on the first line,
@@
 %
-with $\theta_k(t)$ defined by Eq. \eqref{eq:eta_def}.
+with $\theta_k(t)$ defined by Eq. \eqref{eq:theta_def}.
 %and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
@@
   \sum_{\tau = -\infty}^\infty
-    \kappa_n(\tau)
+    \kappa_k(\tau)
   =
@@
 
+\note{The table of symbols is listed in Table \ref{tab:symbols}.
+  \begin{table*}\footnotesize
+  \centering
+  \rowcolors{1}{white}{LightGray}
+  \setlength{\tabcolsep}{4pt} % column separation
+  %
+  \caption{\label{tab:symbols}Table of symbols.}
+  \begin{tabular}{l | p{12cm} }
+    Symbol          &   Description \\
+    \hline
+    $\mathbf{A}$    &   Transition matrix. \\
+    $\alpha$        &   Updating magnitude. \\
+    $b$             &   Cutoff of the kernel of homogeneous updating schemes. \\
+    $C$             &   $C \equiv \int_0^{ q(T) } M(Q) \, dQ$.  \\
+    $\delta$        &   Dirac's $\delta$-function. \\
+    $c$             &   Constant. \\
+    $\Err$          &   Error. \\
+    $\ln f$         &   Updating factor in the WL algorithm.  \\
+    $\phi_{ki}$     &   Eigenvectors of the updating matrix. \\
+    $g$             &   Degeneracy factor, $2$ for a periodic variable, or $1$ a non-periodic one. \\
+    $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
+    $h_i(t)$        &   Histogram.  \\
+    $H_i$           &   Average histogram.  \\
+    $i, j$          &   Bin indices. \\
+    $\lambda_k$     &   The $k$th eigenvalue of the updating matrix. \\
+    $k, l$          &   Mode indices. \\
+    $K$             &   Cutoff wave number of bandpass updating schemes.  \\
+    $\kappa_k(t)$   &   Autocorrelation function of mode $k$. \\
+    $M(Q)$          &   Mass function, Eq. \eqref{eq:mass_func}.   \\
+    $m(Q)$          &   Mass distribution,
+                        $m(Q) \equiv M(Q)/\int_0^{ q(T) } M(Q') \, dQ'$.  \\
+    $\mu_i$         &   Updating kernel of a homogeneous updating scheme. \\
+    $n$             &   Number of bins. \\
+    $\nu_k$         &   $\nu_k \equiv \lambda_k / \lambda$. \\
+    $p_i$           &   Distribution. \\
+    $\pi_i$         &   (Instantaneous) distribution, Eq. \eqref{eq:pi_p_phi} \\
+    $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\
+    $Q$             &   $q(T) - q(t)$.  \\
+    $\theta_k(t)$   &   Noise of mode $k$, Eq. \eqref{eq:theta_def}. \\
+    $\rho_i$        &   Distribution used to define the error. \\
+    $\sigma_{ij}(t)$   &   Autocorrelation function of $\zeta_i(t)$. \\
+    $\sigma$        &   Width of the Gaussian updating scheme. \\
+    $t, \tau$       &   Time. \\
+    $T$             &   Simulation length. \\
+    $u_i(q')$       &   $u_i(q') \equiv e^{\lambda \, [q - q(T)]}$. \\
+    $V_i(t)$        &   Original bias potential. \\
+    $v_i(t)$        &   Shifted bias potential. \\
+    $\mathbf w$     &   Updating matrix. \\
+    $x_i(t)$        &   Average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
+    $\xi_i(t)$      &   The noise, Eq. \eqref{eq:xi_def}. \\
+    $\chi_i$        &   $\chi_i \equiv \ln \frac{ p_i } { p^*_i } - \sum_{ j = 1 }^n p_j \, \ln \frac{ p_j } { p^*_j }$. \\
+    $y_k(t)$        &   Mode of the shifted bias potential, Eq. \eqref{eq:y_def}. \\
+    $Y_k$           &   Mode of the original bias potential, Eq. \eqref{eq:Y_def}. \\
+    $z$             &   Reaction coordinate. \\
+    $\zeta_i(t)$    &   The noise part of the instantaneous histogram, $h_i(t)$.
+  \end{tabular}
+  \end{table*}
+}
+
 \bibliography{simul}

### 2016-06-21 doc/invt.tex
small changes

@@
 dama2014},
+%
 \begin{equation}
@@
 It follows that
+%
 \begin{equation}
@@
 \end{equation}
+%
 and
+%
 \begin{equation}
@@
 the total error of the bias potential as
+%
 \begin{equation}
@@
 with a similar width, $\sigma$,
-\note{(cheng, the inset in figure two should be replotted to fix the dynamic range of the gray scale.)}
 %
@@
 
-\note{(Cheng, we should give a  description of the test system)}
 Our test system is a one-dimensional model system, in which
@@
 are assumed to be flat:
-$p^*_i = p_i = 1/n$.
+$p^*_i = p_i = 1/n$,
+so is the distribution $\rho_i$ used in Eqs.
+\eqref{eq:vbar_def} and \eqref{eq:error_sum}.
 %
@@
 %
-After this,
-we reset the origin of the time, $t$,
+At the end of the equilibration,
+the bias potential would deviate from the correct value
+and our aim to see how well
+the bias potential is restored
+at the end of the testing schedule.
+%
+After the equilibration,
+we reset the origin of the time, $t$, to zero,
 and start the testing schedule
@@
   }
-  \setlength{\tabcolsep}{5pt}
-  \begin{tabular} { l | p{2cm} p{2.5cm} p{1.8cm} }
+  \setlength{\tabcolsep}{2pt}
+  \renewcommand\arraystretch{1.2}
+  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
     \hline
@@
 they may break down for relatively short simulations
-of glassy systems.  Our immediate interest is to apply these relations to
-problems containing aqueous mixtures of proteins and nucleic acids.
+of glassy systems.
+%Our immediate interest is to apply these relations to
+%problems containing aqueous mixtures of proteins and nucleic acids.
 

### 2016-09-01 doc/invt.tex
small corrections

@@
 %
-There are, however, some reservation about
-the constant of proportionality;
-the optimal formula for $\alpha(t)$
-might be $c/t$ with a $c$
-different from $1$\cite{
-morozov2007, zhou2008, morozov2009}.
-%
-Here we will show that the ideal $c$
-is unity, but the origin of $t$ should be shifted
-to accommodate the initial error.
+Here we will re-derive this formula
+and suggest a slight modification
+whereby the origin of $t$ is shifted
+according to the initial error.
 
@@
 %
-At the end of the equilibration,
-the bias potential would deviate from the correct value
-and our aim to see how well
+Although the bias potential is correct initially,
+it would deviate from the correct value
+at the end of the equilibration.
+%
+Our aim is to see how well
 the bias potential is restored
@@
     \label{fig:errinvt}
-    Error $E$ versus $\lambda$
+    Error $E$ versus parameter $\lambda$
     for the modified inverse-time schedule,
@@
     %
-    The modified schedules, produced by
+    The mode-limited schedules were produced by
     limiting the sum in Eq. \eqref{eq:q_opt}
-    to the first four terms,
-    became similar to the inverse-time schedule,
+    to the first four terms.
+    They became similar to the inverse-time schedule,
     Eq. \eqref{eq:alpha_invt1}.
@@
 We thank Dr. J. Ma, Dr. Y. Mei, J. Wang,
-J. Drake, O. Nassar, Dr. C. Lai and Dr. S. Ou
+J. Drake, O. Nassar, Dr. C. Lai, Dr. S. Ou and D. Stuhlsatz
 for helpful discussions.
@@
   \sum_{ i = 1 }^n
+  \rho_i \,
   \frac{ 1 } { T }
@@
 if $2 \, \nu_k > 1$ for every $k$,
-the asymptotic error is given by the integral
+the asymptotic error is implicitly given by the integral
 %$\sum_i \to \frac{2 \, n}{\pi} \int_0^{\pi/2} dp$, and
@@
     $c$             &   Constant. \\
-    $\Err$          &   Error. \\
+    $\Err, \Err_R, \Err_A$          &   Error, residual error, asymptotic error. \\
     $\ln f$         &   Updating factor in the WL algorithm.  \\
     $\phi_{ki}$     &   Eigenvectors of the updating matrix. \\
-    $g$             &   Degeneracy factor, $2$ for a periodic variable, or $1$ a non-periodic one. \\
+    $g$             &   Degeneracy factor, $2$ for a periodic variable, or for $1$ a non-periodic one. \\
     $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
@@
     $p_i$           &   Distribution. \\
-    $\pi_i$         &   (Instantaneous) distribution, Eq. \eqref{eq:pi_p_phi} \\
+    $\pi_i$         &   (Instantaneous) distribution, Eq. \eqref{eq:pi_p_phi}. \\
     $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\

### 2016-12-23 doc/invt.tex
invt.tex: adding derivation from WL to invt

@@
 
+\author{Cheng Zhang}
+\author{B. Montgomery Pettitt}
+\email{bmpettitt@utmb.edu}
+\affiliation{
+Sealy Center for Structural Biology and Molecular Biophysics,
+The University of Texas Medical Branch,
+Galveston, Texas 77555-0304, USA}
+
 \begin{abstract}
@@
   of the updating magnitude
-  for the class of adaptive sampling methods including
+  for a class of free energy calculation methods including
   the Wang-Landau (WL) algorithm and metadynamics.
   %
-  These sampling methods
-  construct a bias potential to offset
+  These methods rely on adaptive construction of
+  a bias potential that offsets
   the potential of mean force by histogram-based adaptive updates.
@@
   The updating magnitude should decrease over time
-  to help reduce the noise in the bias potential,
-  and the optimal schedule maximizes
-  the rate of convergence.
-  %
-  We derive the optimal schedule from
+  to reduce the noise in the bias potential,
+  and one may speed up the convergence by choosing an optimal schedule
+  of the updating magnitude.
+  %
+  Here, we derive the optimal schedule from
   an equation of motion
-  of a free particle with a position-dependent mass,
-  %
-  and show that
+  of a free particle with a position-dependent mass.
+  %
+  Our results show that
   the initial updating magnitude is ideally half of
@@
   and the optimal schedule of these schemes
-  is the inverse of the time or number of sampling steps with a proper shift.
+  is the inverse number of sampling steps with a proper shift,
+  as given by the $1/t$ prescription.
   %
@@
 %
-However, $z$
-often represents a slow variable
-with a localized or multi-modal distribution.
+The logarithm, $\ln p^*(z)$,
+defines a free energy, or a potential of mean force (PMF).
+%
+Difficulty arises, however, when $z$
+represents a collective variable
+that tends to move locally in a multi-modal distribution.
 %
@@
 to capture the global shape of the free energy surface
-in terms of the potential of mean force (PMF), $-\ln p^*(z)$,
 an effective strategy is to introduce a bias potential that
-cancels the PMF, which is equivalent to sampling a flat distribution
+cancels the PMF, resulting sampling of a flat distribution
 along $z$\cite{mezei1987, berg1992, lee1993,
@@
 The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
-and meta\-dy\-nam\-ics\cite{huber1994, laio2002,
+and metadynamics\cite{huber1994, laio2002,
 laio2008, barducci2011, sutto2012}
@@
 %
-These techniques employ regular updating of the bias potential
+Both techniques employ regular updating of the bias potential
 to discourage future visits to the current configuration
@@
 %
-%The main difference
+The difference lies
+in the updating hill function,
+which takes the form of a localized
+$\delta$-function in the WL case,
+or that of a Gaussian
+in the metadynamics.
+%
 %is that in the WL case,
@@
 %
-However,
-frequent updates to the bias potential
+In both cases,
+regular updates to the bias potential
 %in both cases, the bias potential
@@
 %and the fluctuation
-disrupts the underlying
+disrupts the underlying equilibrium
 sampling\cite{laio2002},
-and near convergence
+and for convergence
 one  has to gradually decrease
@@
 the updating magnitude,
-therefore,
-determines the precision of the final bias potential,
+therefore, is critical in improving
+the precision of the final bias potential,
 hence that of the PMF\cite{liang2007,
@@
 or
-the number of simulation steps.
+the inverse number of simulation steps.
 %
@@
 to a class of bandpass updating schemes
-that are optimal for extensive sampling or long simulations,
-and the modified inverse-time schedule presented is optimal for these schemes.
+that are optimal for long simulations,
+and the above origin-modified inverse-time schedule
+turns out to be optimal for these schemes.
 %
@@
 %
-These results allow one to better
+These results may hopefully help one to better
 plan and optimize FDS simulations.
@@
 of a multiple-bin updating scheme
+that encompasses metadynamics
 in Secs. \ref{sec:multiple-bin}
@@
 of the PMF, $-\ln p^*_i$,
-it is often advantageous to carry out
+it is often advantageous to perform
 biased sampling that targets
@@
 %
-At each stage, $\alpha$
+In each stage, $\alpha$
 is kept constant,
@@
 according to the initial error.
+%
+Interestingly, we may reach the same result
+from optimally planning the simulation lengths in the stages,
+and the reduction factor, as shown in Appendix \ref{sec:wltoinvt}.
 
@@
 where the ensemble consists of many similar simulation copies
-that have experienced the same schedule $\alpha(t)$
-and have reached the same $v_i(t)$
-at time $t$.
+sharing the same schedule $\alpha(t)$
+the same $v_i(t)$ at time $t$.
 %
@@
 
-To proceed, we first define an average\cite{
+The bias potential, $v_i(t)$, following Eq. \eqref{eq:vt_diffeq}
+generally grows overtime.
+%
+While a uniform increase of $v_i(t)$ does not affect the 
+resulting distribution $\pi_i(t)$ as given by Eq. \eqref{eq:h_ave},
+%
+it is convenient to express the error of $v_i(t)$,
+from the difference between $v_i(t)$ and a baseline value, $\bar v(t)$,
+that represents the average growth\cite{
 dama2014},
+%
+\begin{equation}
+  x_i(t) \equiv v_i(t) - \bar v(t)
+  %=
+  %v_i - \sum_{j = 1}^n \rho_j \, v_j
+  ,
+\label{eq:x_def}
+\end{equation}
+%
+where
 %
@@
   = \sum_{i = 1}^n \rho_i \, v_i(t)
+  ,
 \label{eq:vbar_def}
@@
 $\pmb \rho = (\rho_1, \dots, \rho_n)$.
+%
+Apparently,
+%
+\begin{equation}
+  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
+  .
+\label{eq:rhox_sum}
+\end{equation}
 %
@@
 in well-tempered metadynamics,
-$\pmb\rho$ is always flat,
+$\pmb\rho$ is a flat distribution,
 while $\mathbf p$ is an attenuated version of
@@
 %
-Then, we expect the difference %between $v_i$ and $\bar v$,
-%
-\begin{equation}
-  x_i \equiv v_i - \bar v
-  %=
-  %v_i - \sum_{j = 1}^n \rho_j \, v_j
-  ,
-\label{eq:x_def}
-\end{equation}
-%
-to be small in the asymptotic regime.
-%
-It follows that
-%
-\begin{equation}
-  \sum_{i = 1}^n \rho_i \, x_i = 0
-  ,
-\label{eq:rhox_sum}
-\end{equation}
-%
-and
-%
+The rate of growth of $v(t)$ is given by
 \begin{equation}
@@
   \alpha(t)
-  \sum_{ r = 1 }^n \frac{ h_r(t) } { p_r }
+  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
   .
@@
 \end{equation}
+%
+
+In the asymptotic regime,
+we expect the deviation $x_i(t)$
+to be small.
 %
@@
 %
-\note{For \eqref{eq:hdet_linearize}. Following the second step
-$$
-\langle h_i(t) \rangle
+\note{
+The second step of Eq. \eqref{eq:hdet_linearize} follows
+$$
+\frac{ \langle h_i(t) \rangle }
+     { p_i }
 \approx
-\frac{                  p_i (1 - x_i) }
+\frac{                       1 - x_i  }
      { \sum_{ r = 1 }^n p_r (1 - x_r) }
 =
-\frac{                      p_i (1 - x_i) }
+\frac{                       1 - x_i  }
      { 1 - \sum_{ r = 1 }^n p_r \, x_r }
@@
 
-\subsubsection{Optimization of a one-variable problem}
+\subsubsection{\label{sec:onevar}
+Optimization of a one-variable problem}
 
@@
 \end{align}
+%
 At the end of period $T$, we have
+%
 \begin{align}
@@
 To simplify the ensuing discussion,
-we  limit ourselves to $\mathbf w$'s
+we  limit ourselves to matrices % $\mathbf w$'s
 that satisfy
@@
 %
-First, we define the mass function as
+First, we define a mass function as
 \begin{equation}
@@
 \end{equation}
-and the normalized mass distribution as
+and a normalized mass distribution as
 \begin{equation}
@@
     \draw[->] (0, 0) -- node[below, inner sep=5mm] {$Q$} (2.8, 0);
-    \draw[->] (0, 0) -- node[above, rotate=90] {$M(Q)$} (0, 1.8);
+    \draw[->] (0, 0) -- node[above, rotate=90] {$m(Q)$} (0, 1.8);
 
@@
 
+The above characterization allows us to determine
+the optimal initial updating magnitude.
+%
 Let us now consider simulations in which
@@
 Many commonly-used updating schemes
-can be characterized by a rigid window
+can be characterized by a hill function,
+which is a rigid window
 centered on the current bin.
@@
     \label{fig:errinvt}
-    Error $E$ versus parameter $\lambda$
+    Error $E$ versus the parameter $\lambda$
     for the modified inverse-time schedule,
@@
     the curves are theoretical predictions.
+    %
+    For the single-bin (WL) updating scheme,
+    the minimal should occur at $\lambda = 1$,
+    while it is not necessarily the case for a general updating scheme.
     %
@@
     &
-      Inverse-time, Eq. \eqref{eq:alpha_invt1}
+      Inverse-time, \newline
+      Eq. \eqref{eq:alpha_invt1}
     &
@@
     &
-      Optimal,
+      Optimal, \newline
       Eq. \eqref{eq:q_opt}
@@
     %
+    The minimal should always occur at
+    $\alpha(0) = \alpha_0/2 = 5 \times 10^{-5}$,
+    as predicted by Eq. \eqref{eq:half_alpha0}.
+    %
     \note{This figure was produced by
@@
 the optimal schedule of the updating magnitude
-for a general adaptive FDS simulation,
-e.g., the WL algorithm or metadynamics.
+for a general class of free energy calculation methods,
+which we refer to as adaptive flat-distribution sampling (FDS) methods,
+including the Wang-Landau (WL) algorithm or metadynamics.
 %
@@
 incrementally update a bias potential
-to offset the PMF and thereby sample a flat distribution.
+along a collective variable
+to offset the potential of mean force (PMF)
+and thereby sample a flat distribution.
 %
@@
 
-Each adaptive FDS method is associated with
-an updating scheme,
-or mathematically an updating matrix.
-%whose elements $w_{ij}$ specifies the
-%relative magnitude of updating the bias potential
-%at bin $i$ upon a visit to bin $j$.
+The optimal schedule depends on the updating hill function,
+which is limited to a single-bin in the WL case,
+but can be extended to several neighboring bins
+taking a Gaussian shape in the metadynamics case.
+%
+More generally,
+each adaptive FDS method is associated with
+an updating scheme
+that can be mathematically characterized
+by an updating matrix.
+%
+Each column of the matrix gives the normalized
+hill function for the bin corresponding to the column.
 %
@@
 including the Gaussian one used in metadynamics,
-the optimal schedule permits no simple closed form,
-but is given by an equation of motion
+the optimal schedule is not given by a simple closed form,
+but is implicitly given by an equation of motion
 of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
@@
 
-Our method is most useful for long simulations to achieve relatively precise answers.
+Our method is most useful for long simulations
+to achieve relatively precise results.
 %
@@
 \appendix
+
+\section{\label{sec:wltoinvt}
+Inverse-time schedule as the optimal limit of the WL algorithm}
+
+We may understand the inverse-time schedule %, Eq. \eqref{eq:},
+as an optimal limit of the stage-wise WL updating scheme.
+%
+Our calculation below will focus on the one-variable problem,
+as done in Sec. \ref{sec:onevar}.
+
+To begin, we split the entire simulation into $S$ stages,
+each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
+%
+We denote the number of steps in stage $s$ as $T_s$, and
+$\sum_{s=1}^S T_s = T$.
+%
+%\begin{equation}
+%  \sum_{s = 1}^S T_s = T
+%  ,
+%  \label{eq:sumT}
+%\end{equation}
+%
+We assume that
+the magnitude decreases geometrically,
+%
+\begin{equation}
+  \alpha_s = a \, r^s
+  ,
+  \label{eq:alphas}
+\end{equation}
+%
+with $0 < r < 1$.
+%
+%and that the system is properly equilibrated
+%under a constant magnitude initially.
+%
+We wish to find the best distribution of the simulation time
+into the $S$ stages
+as well as
+the optimal number of stages and $r$.
+
+If we denote the error at the end of stage $s$ as $E_s$,
+then from Eq. \eqref{eq:x2t_average}, we have
+%
+\begin{align}
+  E_s
+  &=
+  E_{s - 1} \, e^{-2 \, \alpha_s \, T_s }
+  \notag\\
+  &\phantom{=}
+  +
+  \int_0^{T_s} \int_0^{T_s}
+  \alpha_s^2 \,
+  e^{ -\alpha_s \, (T_s - t) }
+  e^{ -\alpha_s \, (T_s - t') }
+  \kappa( t - t' ) \, dt \, dt'
+  \notag
+  \\
+  &=
+  \left(
+    E_{s - 1} - \frac{ \Gamma \, \alpha_s } { 2 }
+  \right)
+  e^{ - 2 \, \alpha_s \, T_s }
+  +
+  \frac{ \Gamma \, \alpha_s } { 2 }
+  .
+  \label{eq:errorstage}
+\end{align}
+%
+where we have assumed
+$\kappa(t) = \Gamma \, \delta(t)$
+and
+$q(t) = \alpha_s \, t$.
+%
+Our aim is to minimize $E_S$
+of the final stage.
+%
+To do so, consider moving $\delta T$ simulation time from
+stage $s+1$ to stage $s$,
+then $E_{s+1}$ is affected as
+%
+\begin{align*}
+  \delta E_{s+1}
+  =
+\left[
+  2 \, \alpha_{s+1}
+  \left(
+    E_s - \tfrac{ \Gamma \, \alpha_{s+1} } { 2 }
+  \right)
+  +
+  \frac{ \partial E_s } { \partial T_s }
+\right]
+e^{-2 \, \alpha_{s+1} \, T_{s+1} }
+\,
+\delta T
+,
+\end{align*}
+%
+Since this variation does not directly affect the error of later stages,
+the above expression must be zero to minimize the final error, i.e,
+%
+\begin{align*}
+  2 \, \alpha_{s+1}
+  \left(
+    E_s - \frac{ \Gamma \, \alpha_{s+1} } { 2 }
+  \right)
+  &=
+  -\frac{ \partial E_s } { \partial T_s }
+  %\\
+  %&=
+  %2 \, \alpha_s \, e^{-2 \, \alpha_s \, T_s}
+  %\left(
+  %  E_{s-1} - \frac{ \Gamma \, \alpha_s } { 2 }
+  %\right)
+  %\\
+  =
+  2 \, \alpha_s \, \left(E_s - \frac{ \Gamma \, \alpha_s } { 2 }  \right)
+  ,
+\end{align*}
+%
+where we have used Eq. \eqref{eq:errorstage}
+in the second step.
+%
+Thus,
+\begin{equation}
+  E_s
+  =
+  \frac{ \Gamma } { 2 } \, ( \alpha_s + \alpha_{s+1} )
+  =
+  \frac{ \Gamma } { 2 } \, \alpha_s ( 1 + r )
+  .
+  \label{eq:Es}
+\end{equation}
+%
+Using Eqs. \eqref{eq:Es} and \eqref{eq:alphas}
+in Eq. \eqref{eq:errorstage}, we get
+%
+\begin{equation}
+  e^{-\alpha_s \, T_s } = r
+  .
+  \label{eq:exp_r}
+\end{equation}
+%
+
+To compute the optimal schedule,
+we map the simulation time $t$ to the stage number $s$ as
+%
+$$
+t = \sum_{s' = 1}^s T_{s'}
+= \frac{ 1 - r^s } { 1 - r } \, T_s
+= \frac{ \ln(1/r) } { 1 - r } \,
+\left(
+  \frac{1}{\alpha_s}
+  -
+  \frac{1}{a}
+\right)
+,
+$$
+where we have used Eq. \eqref{eq:exp_r} in the last step.
+%
+Thus,
+%
+\begin{equation}
+\alpha_s = \frac{1}{
+  \frac{1 -r }{ \ln(1/r) } t + \frac{1}{a}
+},
+\label{eq:alphas}
+\end{equation}
+%
+Thus, the final error is given by
+%
+\begin{align}
+  E_S
+  =
+  \frac{\Gamma}{2}
+  \frac{ 1 + r }
+  {
+    \frac{1}{a}
+    +
+    \frac{1-r}{\ln(1/r)} \,
+    T
+  }
+  ,
+  \notag
+\end{align}
+%
+which is minimized at $r \to 1$,
+for a large $T$.
+%
+If we identify the parameter $a$ as the initial updating magnitude $\alpha(t = 0)$,
+then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
+and the optimal $t_0 = 1/a$.
+%
+Thus, in this limit,
+the schedule given by Eq. \eqref{eq:alphas}
+is identical to that from Eq. \eqref{eq:alpha_invt1}.
 
@@
 However,
-without the intrinsic distribution, $p^*_i$,
-we cannot compute the shifted bias potential, $v_i$,
+we often do not know a priori the intrinsic distribution, $p^*_i$,
+and hence
+cannot compute the shifted bias potential, $v_i$,
 from the original one, $V_i$
@@
 with a set of well-separated eigenvalues,
-in parallel to the lower bound reachable by
+in parallel to the lower bound reached by
 the single-bin scheme with a set of identical eigenvalues
@@
   C
+  =
+  \int_0^\infty
+  M(q') \, dq'
   \le
@@
 \end{equation}
+%
+It is readily seen that this bound is always
+greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
+$$
+\frac{T}{(T+t_0)^2} \, \sum_{k=2}^n \Gamma_k
+\le
+\frac{1}{T} \sum_{k=2}^n \Gamma_k
+\le
+\frac{1}{T} \left( \sum_{k=2}^n \sqrt{ \Gamma_k } \right)^2
+.
+$$
 
@@
     $\rho_i$        &   Distribution used to define the error. \\
+    $s, S$          &   Stage index and the total number of stages in the Wang-Landau algorithm. \\
     $\sigma_{ij}(t)$   &   Autocorrelation function of $\zeta_i(t)$. \\

### 2016-12-23 prog/invtpar.h

@@ -1063,7 +1063,7 @@ static int invtpar_doargs(invtpar_t *m, int argc, char **argv)
         /* handle options that require an argument */
         q = p = argv[i] + j + 1;
         if ( *p != '\0' ) {
-          /* the argument follows immediately after the option
+          /* the argument follows the option immediately
            * e.g., -oa.dat */
           q = p;
         } else if ( ++i < argc ) {
@@ -1082,13 +1082,13 @@ static int invtpar_doargs(invtpar_t *m, int argc, char **argv)
         } else if ( ch == 'c' ) { /* constant for inverse time */
           m->c = atof(q);
         }
-        break; /* skip the rest of the characters in the option */
+        break; /* skip the remaining characters in the option */
       } else if ( ch == 'v' ) {
         m->verbose++;
       } else if ( ch == 'h' ) {
         invtpar_help(m);
       } else {
-        fprintf(stderr, "unknown option %s, j %d, ch %c\n", argv[i], j, ch);
+        fprintf(stderr, "Unknown option %s, j %d, ch %c\n", argv[i], j, ch);
         invtpar_help(m);
       }
     }

### 2017-01-06 doc/invt.tex
small edits

@@
 in the updating hill function,
-which takes the form of a localized
-$\delta$-function in the WL case,
+which takes the form of a discrete 
+$\delta$-function (or a boxcar function one histogram bin wide)
+in the WL case,
 or that of a Gaussian
@@
 the updating magnitude,
-therefore, is critical in improving
+therefore, critically affects
 the precision of the final bias potential,
@@
 %
-We derive the optimal schedule
+We will show that the optimal schedule can be derived
 from an equation of motion of a free particle,
-and show that the initial updating magnitude
+and the initial updating magnitude
 should be set to
@@
 %
-The optimal schedule in the WL case
-is shown to be given by
+Particularly, the optimal schedule in the WL case
+is given by
 an inverse-time formula,
@@
 %
-These results may hopefully help one to better
+These results may help to better
 plan and optimize FDS simulations.
@@
 For a continuous quantity, $z$,
-which is typical in a molecular system,
+which is often the case in a (classical) molecular system,
 we can discretize $z$
@@
 dayal2004, trebst2004, zhang2007, barducci2008, singh2011},
-$p_i$,
+$p_i$ constant for all $i$'s,
 an FDS simulation.
@@
 to guess a sufficiently accurate
-$V_i$ to start with.
+$V_i$ with which to start.
 %
@@
 %they are less well understood theoretically, as
-as mentioned, the updates  break the microscopic reversibility
-of the underlying sampling mechanism. Further failure to properly reduce the size of the histogram increments would lead to roughness or errors in the final estimate.
+as mentioned, the on-the-fly updates break the microscopic reversibility
+of the underlying sampling mechanism.
+Further, failure to properly reduce the size of the histogram increments
+would lead to roughness or errors in the final estimate.
 %
@@
 Thus, in late stages of a long simulation,
-the updating becomes smaller and
+the updating becomes negligible and
 %the adaptive simulation tends to
@@
 
-One can show that with a constant $\alpha(t) = \alpha > 0$,
+One can show that with a constant magnitude $\alpha(t) = \alpha > 0$,
 the distribution collected from
@@
 %
-Thus, one has to decrease $\alpha$ over time
+Thus, one has to reduce the magnitude $\alpha$ over time
 to improve the estimate of the distribution,
@@
   \dot v_i(t)
+  \equiv
+  \frac{ dv_i(t) } { dt }
   =
@@
 %
-where
-$\dot v_i(t) \equiv dv_i(t)/dt$,
-%
-and
+with
+%
 \begin{equation}
@@
 \end{equation}
-is the instantaneous histogram,
+%
+being the instantaneous histogram,
 which is equal to $1.0$
@@
 
-Next, we split $h_i(t)$ into a deterministic part
-and a stochastic part,
+Next, we split $h_i(t)$ into a deterministic averaged part
+and a stochastic part, $\zeta_i(t)$,
 %
@@
 %
-For the random noise part, we have
+For the random noise part,
+the average is zero by definition,
 $\langle \zeta_i(t) \rangle = 0$,
@@
 %
-While a uniform increase of $v_i(t)$ does not affect the 
+Since a uniform increase of $v_i(t)$ does not affect the
 resulting distribution $\pi_i(t)$ as given by Eq. \eqref{eq:h_ave},
 %
-it is convenient to express the error of $v_i(t)$,
+the real error in $v_i(t)$ is more readily defined
 from the difference between $v_i(t)$ and a baseline value, $\bar v(t)$,
@@
 %
-Apparently,
+So we have,
 %
@@
 \note{
-The second step of Eq. \eqref{eq:hdet_linearize} follows
+The second step of Eq. \eqref{eq:hdet_linearize} follows from
 $$
@@
 %
+The expression depends implicitly on the function $\alpha(t)$.
+%
 Our aim is to find the $\alpha(t)$
@@
 It suffices to consider
-a simple one-variable problem.
+a one-variable problem that
+minimizes each individual $\left\langle x_i^2 \right\rangle$.
 
@@
 
-Consider the following equation
+Consider the following stochastic differential equation
 of a single variable:
@@
 
+This is a functional of the curve $q(t)$.
+%
 We now fix the endpoint value $q(T)$
@@
 In Sec. \ref{sec:finlencorr},
-we show that $t_0$ can be adjusted
+we show that % $t_0$ can be adjusted
+one can further improve the schedule by adjusting $t_0$
 to accommodate the initial error.
@@
 The optimization procedure is similar to the single-bin case.
-However, we shall adopt some restrictions and approximations
+%
+However, we shall assume some restrictions and approximations
 to simplify the discussion.
@@
 in Eq. \eqref{eq:vt_diffeq_mbin}
-should be a constant independent of $i$,
+should be a constant of $i$,
 if $h_j(t)$ were the same as $p_j$.
@@
 resembles a transition matrix,
-except that the elements can be negative.
+except that some elements can be negative.
+%
+Below we shall take advantage of the analogy
+and borrow techniques used
+in studying transition matrices\cite{vankampen}.
 
@@
 using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum}, we can write
-the error after a period $T$,
+the error after a period $T$
+as a sum of two contributions,
 %
@@
 %
-as the sum of the residual error
+where the residual error
 (from the decay of the initial error)
+is given by
 %
@@
 and the asymptotic error
+is given by
 \begin{equation}
@@
 %
-where
+with
 \begin{align*}
@@
 In the asymptotic regime
-(i.e., if the simulation length, $T$, is sufficiently long),
+(i.e., if the simulation period, $T$, is sufficiently long),
 we can approximate $\kappa_k(t)$ as
@@
 affects the error only through the
-$\Gamma_k$'s.
+few numbers $\Gamma_k$'s.
 %
@@
 under a constant schedule,
-$\alpha(t) = \alpha_0$.
+$\alpha(t) = a_0$.
 %
@@
 %    \sum_{k = 2}^n
-%      \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
-%      e^{ 2 \, \lambda_k \, \alpha_0 \, (t - T') }
+%      \Gamma_k \, (\lambda_k \, a_0)^2 \,
+%      e^{ 2 \, \lambda_k \, a_0 \, (t - T') }
 %    \, dt
@@
 %  \sum_{ k = 2 }^n
-%    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0
+%    \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
 %  ,
@@
   \int_0^{T'}
-    \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
-      e^{ 2 \, \lambda_k \, \alpha_0 \, (t - T') }
+    \Gamma_k \, (\lambda_k \, a_0)^2 \,
+      e^{ 2 \, \lambda_k \, a_0 \, (t - T') }
     \, dt
@@
   &=
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0
+  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
   ,
@@
 That is, we assume
-$\alpha(t) = \alpha_0$
+$\alpha(t) = a_0$
 for $t < 0$, and
 the system has entered the sampling equilibrium
-under a fixed $\alpha_0$ at $t = 0$.
+under a fixed $a_0$ at $t = 0$.
 %
@@
   \sum_{k = 2}^n
-      \frac 1 2 \, \alpha_0 \, \Gamma_k \, \lambda_k \,
+      \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k \,
       e^{  - 2 \, \lambda_k \, q(T) }
@@
     \Gamma_k \, \dot u_k^2\bigl[ q(t) \bigr]
-  \notag
-  \\
-  &=
+  %\notag
+  %\\
+  %&=
+  =
   \sum_{ k = 2 }^n
@@
 %
-But the Hamiltonian
+But in this case the Hamiltonian
 %
@@
 %
-for some positive constant $C$,
-%
-%We may treat Eq. \eqref{eq:Lagrangian_const}
+for some positive constant $C$.
+%
+This serves
 as a generalization of Eq. \eqref{eq:dudt_const}.
@@
 in the above,
-the parameter $q(T)$ can be further adjusted
+the value of $q(T)$ can be further adjusted
 to minimize the total error, $E = E_R + E_A$
@@
 
-This mapping allows us to classify all optimal schedules
+This mapping allows us to classify optimal schedules
 as follows.
@@
 the system is initially equilibrated at the same
-constant updating magnitude $\alpha_0$.
+constant updating magnitude $a_0$.
 %
@@
   =
-  \frac{1} { T \, (\alpha_0 / 2) }
+  \frac{1} { T \, (a_0 / 2) }
   .
@@
   &=
-  -\alpha_0 \, M^2(Q_c)
+  -a_0 \, M^2(Q_c)
   ,
@@
   =
-  \frac{ \alpha_0 }
+  \frac{ a_0 }
        { 2 }
@@
 happens to coincide with the
-reduction factor of updating magnitude
+recommended reduction factor of updating magnitude
 at stage transitions
@@
 
-Now consider
-simulations with the same $\alpha_0$.
+Let us now consider
+simulations with the same $a_0$.
 %
@@
 Further, the multiplication of $\mathbf w$
-can be reduced to a convolution with the window,
+can be reduced to a convolution with the hill function,
 or the updating kernel.
@@
 The updating matrix, $\mathbf w$,
-assume the following form
+assumes the following form
 %
@@
 %
-Particularly,
-we shall only consider symmetric kernels
-that satisfy
+We shall further impose the symmetry
 %
@@
   +
-  \sum_{j = 1+n}^{2 \, n}
-    \mu_{i - l} \, \phi_l
+  \sum_{l = 1+n}^{2 \, n}
+    \mu_{i - l} \, \phi_{l - n}
   +
-  \sum_{j = 1-n}^0
-    \mu_{i - j + n} \, \phi_j
+  \sum_{l = 1-n}^0
+    \mu_{i - l} \, \phi_{l + n}
+  \\
+  &=
+  \sum_{j = 1}^n
+    \mu_{i - j} \, \phi_j
+  +
+  \sum_{l = 1+n}^{2 \, n}
+    \mu_{i - l} \, \phi_{l}
+  +
+  \sum_{l = 1-n}^0
+    \mu_{i - l} \, \phi_{l}
   \\
@@
 the eigenvalues, $\lambda_k$'s,
-is the cosine transform of
+are the cosine transform of
 the updating kernel.
@@
 So this updating scheme is stable if
-$\min \lambda_k = \lambda_n \ge 0$,
+$\min \lambda_k \ge 0$,
 or
@@
 
+\note{
+If the variable is periodic ($g = 2$) and $n$ is even,
+the minimum is achieved at $k = n/2$.
+
+If the variable is periodic ($g = 2$) but $n$ is odd,
+the minimum is achieved at $k = (n+1)/2$.
+$$
+\sin^2\left(
+  \frac{n-1}{2n} \pi
+\right)
+=
+\cos^2\left(
+  \frac{\pi}{2n}
+\right)
+.
+$$
+
+If the variable is nonperiodic ($g = 1$)
+the minimum is achieved at $k = n$.
+}
+
 \subsubsection{Gaussian updating scheme}
@@
 we define
-$x = l \, g \, \Delta x$,
+$x = l \, \Delta x$,
 and
@@
 where
-$\Delta x = \pi/n$
+$\Delta x = g \, \pi/n$
 is the bin size.
@@
 By assuming the largest $b \approx n/g$,
-Eq. \eqref{eq:wband_eigenvalue}
-can be approximated as an integral:
+we can approximate Eq. \eqref{eq:wband_eigenvalue}
+as an integral:
 %
@@
 $
+\note{
+\begin{align*}
+  \lambda_k
+  &\approx
+  \int_0^b \mu_l \, \cos[(k-1) \, x] \, dl
+  \\
+  &=
+  \int_0^b \mu(x) \, \cos[(k-1) \, x] \, \Delta x \, dl
+  \\
+  &=
+  \int_0^{b \, \Delta x} \mu(x) \, \cos[(k-1) \, x] \, dx
+\end{align*}
+The upper limit
+$b \, \Delta x \approx (n/g) \, (g\,\pi/n) = \pi$.
+}
 
 For the Gaussian updating kernel,
-if the width $\sigma$ satisifies
+if the width $\sigma$ is satisified
 with $\Delta x \ll \sigma \ll 1$,
@@
 a uniform shift of all $v_k$'s,
-we can freely set $\lambda_1$ to $\lambda_2$
+we can freely set $\lambda_1$ to be the same as $\lambda_2$
 without changing the nature of the updating scheme.
 %
-Now an updating matrix
+Note, an updating matrix
 with identical eigenvalues
@@
 %
-The definition of $K$ will be modified.
+The definition of $K$ will be modified
+for mathematical convenience.
 
@@
 By including the residual error
-and minimize the total,
+and minimizing the total,
 we get the optimal $t_0$,
@@
 %%under a constant updating magnitude,
-%%$\alpha_0$,
+%%$a_0$,
 %Assuming Eq. \eqref{eq:error_res1} with
@@
 %  =
-%  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
+%  \frac { \frac 1 2 a_0 \, t_0^2 + T }
 %        {          ( T + t_0 )^2          }
@@
   \frac{  2  }
-       { \alpha_0 }
+       { a_0 }
   ,
@@
 %For example, consider a system that is initially
-%prepared under a fixed $\alpha_0$
+%prepared under a fixed $a_0$
 %at the beginning of the schedule $\alpha(t)$.
@@
 $p^*_i = p_i = 1/n$,
-so is the distribution $\rho_i$ used in Eqs.
+so is the distribution $\rho_i$ in Eqs.
 \eqref{eq:vbar_def} and \eqref{eq:error_sum}.
@@
 under a constant updating magnitude,
-$\alpha_0 = 10^{-4}$.
+$a_0 = 10^{-4}$.
 %
@@
 Both are based on
-Metropolis MC algorithm\cite{
+a Metropolis MC algorithm\cite{
   metropolis1953, newman, frenkel,
@@
 and $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
-%and $t_0 = 2/\alpha_0$.
+%and $t_0 = 2/a_0$.
 %
@@
 but not necessarily so for a general updating scheme.
+%
 
@@
 The resulting schedule
-would became similar to the inverse-time schedule,
+would become similar to the inverse-time schedule,
 Eq. \eqref{eq:alpha_invt1},
@@
 The minimal error was reached at
-$\alpha(0) = \alpha/2 = 5 \times 10^{-5}$.
+$\alpha(0) = a_0/2 = 5 \times 10^{-5}$.
 
@@
     with $\sigma = 10$,
-    and $\alpha_0 = 10^{-4}$.
+    and $a_0 = 10^{-4}$.
     %
@@
     The minimal should always occur at
-    $\alpha(0) = \alpha_0/2 = 5 \times 10^{-5}$,
+    $\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
     as predicted by Eq. \eqref{eq:half_alpha0}.
@@
 which we refer to as adaptive flat-distribution sampling (FDS) methods,
-including the Wang-Landau (WL) algorithm or metadynamics.
+including the Wang-Landau (WL) algorithm and metadynamics.
 %
@@
 Our calculation below will focus on the one-variable problem,
-as done in Sec. \ref{sec:onevar}.
+as in Sec. \ref{sec:onevar}.
 
@@
 We assume that
-the magnitude decreases geometrically,
+the magnitude falls geometrically,
 %
@@
 then from Eq. \eqref{eq:x2t_average}, we have
+$q(t) = \alpha_s \, t$, and
 %
@@
 $\kappa(t) = \Gamma \, \delta(t)$
-and
-$q(t) = \alpha_s \, t$.
-%
-Our aim is to minimize $E_S$
-of the final stage.
-%
-To do so, consider moving $\delta T$ simulation time from
+in the second step.
+%
+Our aim is to minimize error
+of the final stage, $E_S$.
+%
+To do so, let us consider moving $\delta T$ simulation time from
 stage $s+1$ to stage $s$,
@@
 \delta T
-,
+.
 \end{align*}
@@
 \begin{equation}
-\alpha_s = \frac{1}{
-  \frac{1 -r }{ \ln(1/r) } t + \frac{1}{a}
-},
-\label{eq:alphas}
-\end{equation}
-%
-Thus, the final error is given by
+  \alpha_s =
+  \left[
+    \frac{1}{a} +
+    \frac{1 -r }{ \ln(1/r) } t
+  \right]^{-1}
+  ,
+  \notag
+  %\label{eq:alphas2}
+\end{equation}
+%
+and the final error is given by
 %
@@
 %
-which is minimized at $r \to 1$,
+which is minimized as $r \to 1$,
 for a large $T$.
@@
 the bin index $i$ is an independent random variable
-at each time step $t$, and
+at each time step $t$.
+%
+Further, in the asymptotic limit,
+$\langle h_i(t) \rangle = p_i$,
+we find from Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def} that
+$$
+  \xi_i(t)
+  =
+  \frac{ h_i(t) } { p_i }
+  -
+  \sum_{ r = 1 }^n
+    \rho_r \frac{ h_r(t) } { p_r }
+  ,
+$$
+%
+and
 %
@@
   \delta_{t, t'}
-  ,
+  .
 \label{eq:zz_perfect}
@@
 %
-where
-we have assumed the asymptotic limit
-such that $\langle h_i(t) \rangle = p_i$,
-which leads to,
-by Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def},
-$$
-  \xi_i(t)
-  =
-  \frac{ h_i(t) } { p_i }
-  -
-  \sum_{ r = 1 }^n
-    \rho_r \frac{ h_r(t) } { p_r }
-  .
-$$
 %$\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
@@
 with a constant schedule,
-$\alpha(t) = \alpha_0$.
+$\alpha(t) = a_0$.
 %
@@
 =
-\langle y_k^2 \rangle.
+\left\langle y_k^2 \right\rangle.
 $$
@@
 $\lambda_i$'s and $\Gamma_i$'s
-from the input,
+from the given system and updating scheme (input),
 and then apply them to Eq. \eqref{eq:q_opt}
@@
 %using the updating scheme
-with a constant updating magnitude, $\alpha_0$.
+with a constant updating magnitude, $a_0$.
 
@@
 By replacing negative eigenvalues with zeros and
-using Eq. \eqref{eq:mu_from_lambda_pbc}
+using the inversion formula,
+Eq. \eqref{eq:mu_from_lambda_pbc}
 or \eqref{eq:mu_from_lambda_refl},
@@
 although the negative eigenvalues
-often become smaller in magnitude.
+tend to be smaller in magnitude.
 %
@@
   \item
-    If all eigenvalues are nonnegative,
+    If all eigenvalues are nonnegative within the numerical error,
     we are done. % and exit the loop.
@@
   for some $q_0$,
-  and we expect that with fixed $\alpha_0$,
+  and we expect that with fixed $a_0$,
   $q(T)$ to increase almost linearly with $T$
@@
 where
+we have defined a time-averaged histogram
 $H_i = \frac{1}{T} \sum_{t = 1}^T h_i(t)$
-with
+with the instantaneous histogram 
 $h_i(t)$ defined by Eq. \eqref{eq:h_def}.
@@
 We can rewrite the error as
-a sum of the eigenmode as
+a sum of the eigenmodes as
 %
@@
 
-As an alternative to the
-optimal schedule from Eq. \eqref{eq:q_opt},
-here we consider the modified
-inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
+Here we compute the error of the schedule given by
+the modified inverse-time formula,
+Eq. \eqref{eq:alpha_invtlambda}.
+%As an alternative to the
+%optimal schedule from Eq. \eqref{eq:q_opt},
+%here we consider the modified
+%inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
 %with a free parameter $\lambda$.
 %
-whose error can be computed explicitly.
+%whose error can be computed explicitly.
 
 \subsection{\label{sec:invt_error}
-Error
+General results
 }
@@
 in Eq. \eqref{eq:error_asym1} yields
+the asymptotic error
 %
@@
 From Eq. \eqref{eq:error_res1},
-we get
+we get the residual error
 %
@@
   =
-  \frac { \alpha_0 } { 2 }
+  \frac { a_0 } { 2 }
   \sum_{k = 2}^n
@@
 %the system was initially equilibrated
-%at a constant $\alpha_0$,
+%at a constant $a_0$,
 %such that the values of
@@
   \left[
-    1 - \frac 1 2 \alpha_0 \, (T+t_0)
+    1 - \frac 1 2 a_0 \, (T+t_0)
   \right] r \, \log r
@@
 %
-and the asymptotic error in the optimal case is given by
+which shows that the original inverse-time formula,
+Eq. \eqref{eq:alpha_invt1},
+is not optimal in this case.
+%
+The asymptotic error is given by
 %
@@
 \note{The table of symbols is listed in Table \ref{tab:symbols}.
-  \begin{table*}\footnotesize
+  \begin{table*}
+  \footnotesize
   \centering
@@
     $C$             &   $C \equiv \int_0^{ q(T) } M(Q) \, dQ$.  \\
-    $\delta$        &   Dirac's $\delta$-function. \\
+    $\delta$        &   Dirac's or discrete $\delta$-function. \\
     $c$             &   Constant. \\

### 2017-01-06 doc/invt.tex
small edits to invt.tex

@@
-\documentclass[preprint, floatfix]{revtex4-1}
+\documentclass[preprint, superscriptaddress, floatfix]{revtex4-1}
 \usepackage{amsmath}
@@
 \author{Cheng Zhang}
+\author{Justin A. Drake}
+\affiliation{
+Sealy Center for Structural Biology and Molecular Biophysics,
+The University of Texas Medical Branch,
+Galveston, Texas 77555-0304, USA}
+\author{Jianpeng Ma}
+\affiliation{
+Department of Biochemistry and Molecular Biology,
+Baylor College of Medicine, Houston, Texas 77030, USA}
+\affiliation{
+Department of Bioengineering,
+Rice University, Houston, Texas 77005, USA}
 \author{B. Montgomery Pettitt}
@@
   %
-  Here, we derive the optimal schedule from
-  an equation of motion
-  of a free particle with a position-dependent mass.
-  %
-  Our results show that
-  the initial updating magnitude is ideally half of
+  Here, we will show that the optimal schedule
+  can be derived
+  from a mechanical analogy, in which
+  the schedule corresponds to
+  the velocity of a free particle with a position-dependent mass
+  and the final error serves as the action.
+  %
+  Therefore, the optimal schedule follows from
+  the equation of motion of the particle.
+  %
+  We will show that although
+  the optimal schedule depends on the updating scheme (WL or metadynamics),
+  the initial updating magnitude should always be half of
   the previous equilibrium value.
@@
   Further,
-  we show that the WL updating scheme
+  the WL updating scheme
   belongs to a class of bandpass updating schemes
@@
 The difference lies
-in the updating hill function,
+in the updating window function,
 which takes the form of a discrete 
@@
 %
-We will show that the optimal schedule can be derived
-from an equation of motion of a free particle,
-and the initial updating magnitude
-should be set to
-half of the previous equilibrium value.
-%
-Particularly, the optimal schedule in the WL case
-is given by
-an inverse-time formula,
-with a modified origin of time.
-%
-However,
+Our method maps the optimization problem to a mechanic one,
+in which the schedule becomes the velocity of a free particle,
+and the final error plays the role of the action.
+%
+In this way, the optimal schedule follows the equation of motion
+of the particle that minimizes the action.
+%
+We will show that the optimal schedule
+depends on how the bias potential is updated,
+and hence is different in the WL and metadynamics cases.
+%
+While the optimal schedule in the WL case
+is given by the known inverse-time formula,
 the optimal schedule for metadynamics is more complex
@@
 %
-Further, we generalize
+In either case, however,
+the initial updating magnitude
+should be set to
+half of the previous equilibrium value,
+which amounts to a shift of the origin of time in the WL case.
+%
+Further, we can generalize
 the WL updating scheme
@@
 that are optimal for long simulations,
-and the above origin-modified inverse-time schedule
+and the above origin-shifted inverse-time schedule
 turns out to be optimal for these schemes.
@@
 
-Here we shall call a simulation that targets
+Here we will call a simulation that targets
 a nearly flat distribution\cite{
@@
 
-We shall mainly focus on the adaptive methods below.
+We will mainly focus on the adaptive methods below.
 %
@@
 In this section,
-we shall derive the optimal $\alpha(t)$
+we will derive the optimal $\alpha(t)$
 for the single-bin scheme,
@@
 To do so,
-we shall first express the error of
+we will first express the error of
 the shifted bias potential, $v_i(t)$,
@@
 %
-%We shall further demand that
+%We will further demand that
 %%
@@
 %
-However, we shall assume some restrictions and approximations
+However, we will assume some restrictions and approximations
 to simplify the discussion.
@@
 %
-Below we shall take advantage of the analogy
+Below we will take advantage of the analogy
 and borrow techniques used
@@
 %
-We shall label this eigenvector as the first one,
+We will label this eigenvector as the first one,
 and
@@
 affects the error only through the
-few numbers $\Gamma_k$'s.
+few $\Gamma_k$'s.
 %
@@
 for a multiple-bin scheme,
-we shall first compute the equilibrium error
+we will first compute the equilibrium error
 under a constant schedule,
@@
 We can rewrite Eq. \eqref{eq:error_asym1} much like an action
+of particle whose position is given by $q(t)$:
 %
@@
 %
-where the Lagrangian,
+where the Lagrangian is
 %
@@
     \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
-  \; \dot q^2( t ),
+  \; \dot q^2( t )
+  .
 \notag
@@
 %
+As the schedule, $\dot q(t) = \alpha(t)$,
+corresponds to the velocity of the particle,
+the above Lagrangian contains only the kinetic energy.
+%
+Thus, the above mechanical analogy
 represents a free particle
-with a position-dependent mass.
+carrying a position-dependent mass.
 %
@@
 Many commonly-used updating schemes
-can be characterized by a hill function,
-which is a rigid window
+can be characterized by a rigid window
 centered on the current bin.
@@
 Besides, the updating matrices
-share the same set of eigenvectors,
+share the same set of eigenvectors
+(sines and cosines of different frequencies),
 and thus are completely determined
@@
 Further, the multiplication of $\mathbf w$
-can be reduced to a convolution with the hill function,
+can be reduced to a convolution with the window function,
 or the updating kernel.
@@
 %%
-%For simplicity, we shall assume that the target
+%For simplicity, we will assume that the target
 %distribution is flat, or $p_i = 1/n$, below.
@@
 %
-We shall further impose the symmetry
+We will further impose the symmetry
 %
@@
 These numbers constitute an \emph{updating kernel}.
-
-The above matrix, $\mathbf w$,
+%
+In this way, the matrix, $\mathbf w$,
 being symmetric with respect to the indices $i$ and $j$,
@@
 for the flat distribution.
-%
+
 To find the eigenvectors,
@@
 %
-Then
+Then, the multiplication of the matrix, $\mathbf w$,
+is equivalent to a convolution with the kernel:
 %
@@
 }
-
-Thus,
-we may construct a set of orthonormal eigenvectors,
+%
+As a result, the characteristic equation,
+$\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$,
+can be solved by discrete Fourier transform.
+%
+We may construct a set of orthonormal eigenvectors,
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-compatible with Eq. \eqref{eq:phi_pbc},
+compatible with the periodicity, Eq. \eqref{eq:phi_pbc},
+%
 \begin{equation}
@@
 the eigenvalues are
+the Fourier or cosine transform of the updating kernel:
 %
@@
   +
-  \mu_{ i + j - 2 n - 1 },
+  \mu_{ i + j - 2 n - 1 }
+  ,
   \label{eq:w_band_refl}
 \end{equation}
+%
+where the last two terms on the right hand side,
+representing two reflective mirrors at
+$j = 1/2$ and $j = n + 1/2$,
+are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
+to the equilibrium distribution\cite{bussi2006},
+such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
+are satisfied.
 %
@@
 the cutoff $b$ only needs to be less than $n$.
-%
-The last two terms
-on the right-hand side of Eq. \eqref{eq:w_band_refl} are added
-to avoid unintended distortion\cite{dickson2011, mcgovern2013}
-to the equilibrium distribution\cite{bussi2006},
-such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
-are satisfied.
 
@@
 %
-such that Eq. \eqref{eq:wmul_to_convol}
-still holds.
+such that the matrix multiplication is still equivalent to
+a convolution as Eq. \eqref{eq:wmul_to_convol}.
 %
@@
 }%
-Thus, we also have $n$ orthonormal eigenvectors,
+We note that Eq. \eqref{eq:phi_refl}
+defines a periodic variable of period $2 \, n$
+with a reflective symmetry around $i = 1/2$.
+%
+Accordingly, the eigenvectors,
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-compatible with Eq. \eqref{eq:phi_refl},
+must share the same periodicity,
+and be even functions about the axis $i = 1/2$:
 %
@@
        {             n              }
-  \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, \pi}
-            {                    n                           }
+  \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, 2 \, \pi}
+            {                    2 \, n                           }
   ,
@@
 %
-with eigenvalues
+The eigenvalues can be borrowed from
+Eq. \eqref{eq:wband_eigenvalue_pbc},
+with the periodic $n$ doubled:
 %
@@
     \mu_l
-    \cos \frac{(k - 1)  \, l \pi}{n}
+    \cos \frac{(k - 1)  \, l \, 2 \, \pi}{ 2 \, n }
   .
@@
 }%
-
-Note that Eq. \eqref{eq:wband_eigenvalue_refl}
-differs from Eq. \eqref{eq:wband_eigenvalue_pbc}
-by a factor of $2$
-in the argument of the cosine function.
-%
-We may unify the two as
+%
+We may unify Eqs. \eqref{eq:wband_eigenvalue_refl}
+and \eqref{eq:wband_eigenvalue_pbc} as
 %
@@
 for a given updating scheme,
-we shall turn to the problem of choosing
+we will turn to the problem of choosing
 optimal updating schemes below.
@@
 
-We shall first show
+We will first show
 that the single-bin scheme is asymptotically optimal.
@@
 In Appendix \ref{sec:upperbound},
-we shall show that the opposite limit of
+we will show that the opposite limit of
 a set of well-separated eigenvalues
@@
 respectively,
-and we shall use
+and we will use
 Eqs. \eqref{eq:Gamma_perfect}
@@
 
-The optimal schedule depends on the updating hill function,
+The optimal schedule depends on the updating window function,
 which is limited to a single-bin in the WL case,
@@
 Each column of the matrix gives the normalized
-hill function for the bin corresponding to the column.
+window function for the bin corresponding to the column.
 %
@@
 We thank Dr. J. Ma, Dr. Y. Mei, J. Wang,
-J. Drake, O. Nassar, Dr. C. Lai, Dr. S. Ou and D. Stuhlsatz
+O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz, C. Myers, and O. Samoylova
 for helpful discussions.
@@
 
-\section{\label{sec:wltoinvt}
+\section{\label{sec:invt_discussion}
+On the inverse-time schedule}
+
+\subsection{\label{sec:wltoinvt}
 Inverse-time schedule as the optimal limit of the WL algorithm}
@@
 \begin{equation}
-  \alpha_s = a \, r^s
+  \alpha_s = \alpha_0 \, r^s
   ,
@@
   -
-  \frac{1}{a}
+  \frac{1}{\alpha_0}
 \right)
@@
   \left[
-    \frac{1}{a} +
+    \frac{1}{\alpha_0} +
     \frac{1 -r }{ \ln(1/r) } t
@@
   {
-    \frac{1}{a}
+    \frac{1}{\alpha_0}
     +
@@
 %
-which is minimized as $r \to 1$,
+which is minimized as $r \to 1$
 for a large $T$.
 %
-If we identify the parameter $a$ as the initial updating magnitude $\alpha(t = 0)$,
+If we identify the parameter $\alpha_0$ as the initial updating magnitude $\alpha(t = 0)$,
 then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
-and the optimal $t_0 = 1/a$.
+and the optimal $t_0 = 1/\alpha_0$.
 %
@@
 is identical to that from Eq. \eqref{eq:alpha_invt1}.
+
+\subsection{\label{sec:adaptinvt}
+Inverse-time schedule from adaptive averaging
+}
+
+Result of the increase of the sample size.
+
+Final bias potential is already the average,
+No further trajectory averaging is needed.
+Fixed magnitude, improvable by trajectory averaging.
 
@@
 }
-Particularly, if $p_i = \rho_i$, then, by
+In particular, if $p_i = \rho_i$, then, by
 Eq. \eqref{eq:eig_orthonormal_rows}
@@
   \cos \frac{ (k - 1) \, g \, \pi }
-            {       2 \, n        }
+            {            n        }
   ,
@@
   =
-  \frac{1}{2}
-  \alpha \, \Gamma_k \, \lambda_k,
+  \frac{1}{2} \,
+  a_0 \, \Gamma_k \, \lambda_k,
 \label{eq:varY}
@@
 
-We shall derive here an upper bound of the asymptotic error
+We will derive here an upper bound of the asymptotic error
 that can be associated to updating schemes
@@
 %
-Below, we shall show that this practice is unnecessary
+Below, we will show that this practice is unnecessary
 if the simulation uses the single-bin or a bandpass updating scheme
@@
 \subsection{\label{sec:invt_error}
-General results
+General results on error
 }

### 2017-01-21 doc/invt.tex
invt.tex: improve introduction

@@
-\documentclass[preprint, superscriptaddress, floatfix]{revtex4-1}
+\documentclass[reprint, superscriptaddress, floatfix]{revtex4-1}
 \usepackage{amsmath}
@@
   %
-  Here, we will show that the optimal schedule
-  can be derived
-  from a mechanical analogy, in which
-  the schedule corresponds to
-  the velocity of a free particle with a position-dependent mass
-  and the final error serves as the action.
-  %
-  Therefore, the optimal schedule follows from
-  the equation of motion of the particle.
-  %
-  We will show that although
+  %Here, we will show that the optimal schedule
+  %can be derived
+  %from a mechanical analogy, in which
+  %the schedule corresponds to
+  %the velocity of a free particle with a position-dependent mass
+  %and the final error serves as the action.
+  %%
+  %Therefore, the optimal schedule follows from
+  %the equation of motion of the particle.
+  %
+  We will show that
   the optimal schedule depends on the updating scheme (WL or metadynamics),
-  the initial updating magnitude should always be half of
+  and the initial updating magnitude should be half of
   the previous equilibrium value.
@@
 Our method maps the optimization problem to a mechanic one,
-in which the schedule becomes the velocity of a free particle,
-and the final error plays the role of the action.
-%
-In this way, the optimal schedule follows the equation of motion
+in which the schedule plays the role of the velocity of
+a free particle with a position-dependent mass,
+and the final error becomes the action.
+%
+In this way, the optimal schedule
+can be deduced from the equation of motion
 of the particle that minimizes the action.
@@
 biased sampling that targets
-a wider distribution, $p_i$.
+a wider and smoother distribution, $p_i$.
 %
@@
   \pi_i \propto p^*_i \, e^{-V_i}.
-  \label{eq:pi_p_phi1}
+  \label{eq:pi_p_V1}
 \end{equation}
@@
   .
-  \label{eq:pi_p_phi}
-\end{equation}
-%
-In particular,
-to make $\pi_i$ flat,
-the bias potential $V_i$
-must coincide with $\ln p^*_i$,
-up to an additive constant.
-%
-In practice, $V_i$ may be non-optimal or contain error,
-and thus $\pi_i$ need not
-coincide with the intended distribution $p_i$.
-
-Here we will call a simulation that targets
-a nearly flat distribution\cite{
-dayal2004, trebst2004, zhang2007, barducci2008, singh2011},
-$p_i$ constant for all $i$'s,
-an FDS simulation.
-%
+  \label{eq:pi_p_V}
+\end{equation}
+%
+%In particular,
+%to make $\pi_i$ flat,
+%the bias potential $V_i$
+%must coincide with $\ln p^*_i$,
+%up to an additive constant.
+%%
+%In practice, $V_i$ may be non-optimal or contain error,
+%and thus $\pi_i$ need not
+%coincide with the intended distribution $p_i$.
+
+We wish the biased equilibrium distribution, $\pi_i$,
+to be the same as the intended distribution, $p_i$,
+which is usually flat,
+i.e., $p_i$ constant for all $i$'s,
+or nearly flat\cite{
+dayal2004, trebst2004, zhang2007, barducci2008, singh2011}.
+%
+We will generally call such a biased simulation
+a flat-distribution-sampling (FDS) simulation.
+%
+To fulfill this goal, the bias potential must satisfy
+%
+\begin{equation}
+  V_i = \ln p^*_i - \ln p_i + \mathrm{const.}
+  ,
+  \label{eq:Vi_target}
+\end{equation}
+%
+according to Eq. \eqref{eq:pi_p_V1}.
+%which allows us to deduce the PMF, $-\ln p^*_i$,
+%directly from the bias potential.
+%
+This relation thus furnishes a practical aim of convergence for FDS
+methods---adjusting the bias potential to make the
+observed distribution, which can be computed from the normalized histogram,
+match the intended one.
+%
+Once this aim is achieved,
+the PMF, $-\ln p^*_i$, can be naturally
+deduced from the bias potential. % $V_i$.
+
 Roughly speaking, there are two types of FDS methods.
@@
 Using $H_i$ as an estimate of $\pi_i$ in
-Eq. \eqref{eq:pi_p_phi1}
+Eq. \eqref{eq:pi_p_V1}
 allows us to correct $V_i$ as
@@
 }
-For an unknown system,
+However, for an unknown system,
 it is often difficult
@@
 %
-\note{Further, a reasonably precise histogram
-requires an accumulation period
-longer than the autocorrelation time,
-which can also be hard to estimate beforehand.}
-
-For this reason, the second type of methods,
-referred to as the adaptive FDS methods,
-were invented.
+%Another difficulty of this approach is that
+Besides, the effectiveness of this approach
+depends on the precision of the histogram, $H_i$,
+which in turn depends on the simulation length
+that must be decided in advance.
+%
+Thus, the rigidity of this framework disallows
+continual improvement of the bias potential, hence that of the PMF,
+as simulation lengthens.
+
+The second type of methods,
+which we will refer to as the adaptive FDS methods,
+address the above shortcomings
+by continually and incrementally update the target distribution.
+%For this reason, the second type of methods,
+%referred to as the adaptive FDS methods,
+%were invented.
 %
@@
 %
-These methods incrementally update the bias potential
-and can approximately sample the target distribution.
+%These methods incrementally update the bias potential
+%and can approximately sample the target distribution.
 %
@@
 of the underlying sampling mechanism.
-Further, failure to properly reduce the size of the histogram increments
+Further, failure to properly reduce the magnitude of
+the histogram-based increments
 would lead to roughness or errors in the final estimate.
@@
 %
-Thus, in late stages of a long simulation,
-the updating becomes negligible and
+%Thus, in late stages of a long simulation,
+%the updating becomes negligible and
 %the adaptive simulation tends to
@@
 %
-if the bias potential were converged to the exact value,
-this adaptive FDS tends to the ideal equilibrium counterpart,
+%if the bias potential were converged to the exact value,
+%this adaptive FDS tends to the ideal equilibrium counterpart,
 %particular equilibrium simulation can be thought as
 %the optimal asymptotic limit of the adaptive counterpart,
-delivering the maximal efficiency.
-
-For  convenience,
-we introduce a shifted bias potential
-%
-\begin{equation}
-  v_i
-  \equiv
-  V_i
-  +
-  \ln \frac { p_i }
-            { p^*_i }
-  .
-  \label{eq:v_def}
-\end{equation}
-%
-Since the last two terms of Eq. \eqref{eq:v_def}
-remain constant during the course of simulation,
-the updates to $v_i$'s are
-the same as the updates to $V_i$'s.
-%
-In terms of $v_i$'s, Eq. \eqref{eq:pi_p_phi}
-becomes
-%
-\begin{equation}
-  \pi_i
-  =
-  \frac{                p_i \, e^{-v_i} }
-       { \sum_{j = 1}^n p_j \, e^{-v_j} }
-  \propto
-  p_i \, e^{-v_i}.
-  \label{eq:pi_p_v}
-\end{equation}
-%
-According to Eq. \eqref{eq:pi_p_phi1},
-$v_i$ should ideally be a constant of $i$
-such that $\pi_i \approx p_i$.
+%
+For example, one can show that
+if the magnitude decreases according to
+the optimal inverse-time prescription\cite{
+belardinelli2007, belardinelli2007jcp, belardinelli2008,
+morozov2007, zhou2008, morozov2009,
+komura2012, caparica2012, caparica2014},
+the WL algorithm is equivalent to the ideal equilibrium counterpart
+in improving the precision of the bias potential,
+delivering the maximal efficiency
+(cf. Appendix \ref{sec:equilerr}).
+%the bias potential in the WL algorithm
+%will evolve similar to Eq. \eqref{eq:vcorr_equil}
+%with the precision of the histogram as simulation lengthens.
 
@@
 In the WL algorithm\cite{wang2001, wang2001pre},
-the bias potential, $v_i(t)$, is updated
+the bias potential, $V_i(t)$, is updated
 at each sampling step $t$ as
@@
 \begin{equation}
-  v_i(t+1)
-  =
-  v_i(t)
+  V_i(t+1)
+  =
+  V_i(t)
   +
@@
 However, as the above updates
-change $v_i(t)$ continuously,
-Eq. \eqref{eq:pi_p_v} no longer holds
+change $V_i(t)$ continuously,
+Eq. \eqref{eq:pi_p_V} no longer holds
 at all times,
-and the error in $v_i(t)$
+and the error in $V_i(t)$
 depends on the updating magnitude\cite{
@@
 %%
-%First, since $v_i(t)$ is updated continuously,
+%First, since $V_i(t)$ is updated continuously,
 %there is a random noise that is proportional
@@
 Interestingly, we may reach the same result
-from optimally planning the simulation lengths in the stages,
+from optimally planning the simulation lengths in the stages
 and the reduction factor, as shown in Appendix \ref{sec:wltoinvt}.
@@
 we will first express the error of
-the shifted bias potential, $v_i(t)$,
+the bias potential, $V_i(t)$,
 as a functional of $\alpha(t)$,
@@
 \begin{equation}
-  \dot v_i(t)
+  \dot V_i(t)
   \equiv
-  \frac{ dv_i(t) } { dt }
+  \frac{ dV_i(t) } { dt }
   =
@@
 sharing the same schedule $\alpha(t)$
-the same $v_i(t)$ at time $t$.
+the same bias potential at time $t$.
 %
@@
 For sufficiently small $\alpha(t)$,
-the $v_i$'s remain roughly the same for a short period.
+the bias potential remains roughly the same for a short period.
 %
@@
 the sampling process is approximately Markovian, % of a finite order,
-and we may assume Eq. \eqref{eq:pi_p_v}
+and we may assume Eq. \eqref{eq:pi_p_V}
 for the deterministic part
@@
   =
-  \frac{                p_i \, e^{-v_i(t)} }
-       { \sum_{j = 1}^n p_j \, e^{-v_j(t)} }.
+  \frac{                p^*_i \, e^{-V_i(t)} }
+       { \sum_{j = 1}^n p^*_j \, e^{-V_j(t)} }.
   \label{eq:h_ave}
@@
 
-\subsubsection{Linear approximation}
-
-The bias potential, $v_i(t)$, following Eq. \eqref{eq:vt_diffeq}
-generally grows overtime.
-%
-Since a uniform increase of $v_i(t)$ does not affect the
+\subsubsection{Error and linear approximation}
+
+Next, we need to define the error in the bias potential.
+%
+This is done by subtracting from the bias potential
+two contributions.
+%
+The first contribution comes from the target value,
+Eq. \eqref{eq:Vi_target}.
+%
+For this purpose,
+we introduce a shifted bias potential
+%
+\begin{equation}
+  v_i(t)
+  \equiv
+  V_i(t)
+  -
+  \ln \frac { p^*_i }
+            { p_i }
+  ,
+  \label{eq:v_def}
+\end{equation}
+%
+which should tend to a constant of $i$
+upon convergence.
+%Since the last two terms of Eq. \eqref{eq:v_def}
+%remain constant during the course of simulation,
+%the updates to $v_i$'s are
+%the same as the updates to $V_i$'s.
+%%
+%In terms of $v_i$'s, Eq. \eqref{eq:pi_p_V}
+%becomes
+%%
+%\begin{equation}
+%  \pi_i
+%  =
+%  \frac{                p_i \, e^{-v_i} }
+%       { \sum_{j = 1}^n p_j \, e^{-v_j} }
+%  \propto
+%  p_i \, e^{-v_i}.
+%  \label{eq:pi_p_v}
+%\end{equation}
+%%
+%According to Eq. \eqref{eq:pi_p_V1},
+
+Second, the bias potential following Eq. \eqref{eq:vt_diffeq}
+generally grows over time.
+%
+Since a uniform increase of $V_i(t)$, or, equivalently, that of $v_i(t)$,
+does not affect the
 resulting distribution $\pi_i(t)$ as given by Eq. \eqref{eq:h_ave},
 %
-the real error in $v_i(t)$ is more readily defined
+the real deviation is more readily defined
 from the difference between $v_i(t)$ and a baseline value, $\bar v(t)$,
@@
 %
-So we have,
-%
-\begin{equation}
-  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
-  .
-\label{eq:rhox_sum}
-\end{equation}
-%
-This distribution $\pmb \rho$
+In this way,
+the variables, $x_i(t)$'s,
+represent the error of the bias potential,
+and are expected to vanish on convergence.
+%
+
+Note that the distribution, $\pmb \rho$,
+introduced above
 will be used to measure the error,
@@
 %
-The rate of growth of $v(t)$ is given by
+From Eqs. \eqref{eq:x_def} and \eqref{eq:vbar_def},
+we have
+%
+\begin{equation}
+  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
+  .
+\label{eq:rhox_sum}
+\end{equation}
+%
+The average growth rate of $v(t)$ is given by
 \begin{equation}
@@
   \label{eq:y_def}
-  \\
+\end{align}
+and
+\begin{align}
   \theta_k(t)
@@
     \phi_{kj} \, \xi_j(t)
-  .
+  ,
   \label{eq:theta_def}
 \end{align}
+are the deviation and noise of mode $k$, respectively.
 
@@
 for a given updating scheme,
-we will turn to the problem of choosing
+we will turn to the problem of finding
 optimal updating schemes below.
@@
 %
-Note that the last expression of \eqref{eq:CSineq}
+Note that the last expression of Eq. \eqref{eq:CSineq}
 depends on the curve, $q(t)$ ($0 < t < T$),
@@
 %
+\note{Derivation.
+  Integrating $\dot u_k = c_k$ yields
+  \begin{equation}
+    u_k(t) = c_k \left(t + t_{k0} \right).
+  \tag{UK}
+  \label{neq:ui_solution}
+  \end{equation}
+  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
+  we have
+  $$
+  c_k = \frac{ 1 }{ T + t_{k0} },
+  \quad
+  \mathrm{and\;}
+  \frac{ t_{k0} } { T + t_{k0} }
+  =
+  e^{ -\lambda_k \, q(T) }.
+  $$
+  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
+  $$
+  -\lambda_k \, q(T) + \lambda_k \, q(t)
+  = \ln c_k + \ln \left( t + t_{k0} \right).
+  $$
+  Differentiating this with respect to $t$ yields
+  Eq. \eqref{eq:alpha_invtk}.
+}
+%
 Such a solution is possible only if
+all nontrivial eigenvalues are the same
 %
@@
 %
-If we further assume $\lambda_2 = 1$,
+If we further assume $\lambda_2 = \cdots = \lambda_n = 1$,
 %the optimal schedule given by
@@
 %
-\note{Derivation.
-  Integrating $\dot u_k = c_k$ yields
-  \begin{equation}
-    u_k(t) = c_k \left(t + t_{k0} \right).
-  \tag{UK}
-  \label{neq:ui_solution}
-  \end{equation}
-  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
-  we have
-  $$
-  c_k = \frac{ 1 }{ T + t_{k0} },
-  \quad
-  \mathrm{and\;}
-  \frac{ t_{k0} } { T + t_{k0} }
-  =
-  e^{ -\lambda_k \, q(T) }.
-  $$
-  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
-  $$
-  -\lambda_k \, q(T) + \lambda_k \, q(t)
-  = \ln c_k + \ln \left( t + t_{k0} \right).
-  $$
-  Differentiating this with respect to $t$ yields
-  Eq. \eqref{eq:alpha_invtk}.
-}
 
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-to allow zero eigenvalues.
+that ignore high-frequency eigenmodes.
+%
+Mathematically, these updating schemes will have
+zero eigenvalues for the ignored modes,
+and they are useful if we can assume that
+the ignored modes are absent from the PMF.
 %
@@
 %
-The definition of $K$ will be modified
+The definition of the cutoff $K$ will be modified
 for mathematical convenience.
@@
 of the bias potential,
-and thus helps minimize the computational effort
-of the adaptive FDS simulation.
+and thus helps minimize the effort
+of computing the PMF.
 %They differ by the updating schemes,
@@
 Besides, the performance gain
-is relative small for a shorter simulation
+is relatively small for a shorter simulation
 with a local sampling process such as MD.
@@
 
-Our method is most useful for long simulations
+Our method of analysis is most useful for long simulations
 to achieve relatively precise results.
@@
 
-\section{\label{sec:invt_discussion}
-On the inverse-time schedule}
-
-\subsection{\label{sec:wltoinvt}
+\section{\label{sec:wltoinvt}
 Inverse-time schedule as the optimal limit of the WL algorithm}
@@
 %
-$$
-t = \sum_{s' = 1}^s T_{s'}
+\begin{align*}
+  t
+  = \sum_{s' = 1}^s T_{s'}
   = \frac{ 1 - r^s } { 1 - r } \, T_s
-= \frac{ \ln(1/r) } { 1 - r } \,
+  = \frac{ 1 } { \gamma } \,
   \left(
@@
   ,
-$$
-where we have used Eq. \eqref{eq:exp_r} in the last step.
+\end{align*}
+where
+$\gamma = ( 1 - r ) / \ln(1/r)$,
+and we have used Eq. \eqref{eq:exp_r} in the last step.
 %
@@
   \alpha_s =
-  \left[
-    \frac{1}{\alpha_0} +
-    \frac{1 -r }{ \ln(1/r) } t
-  \right]^{-1}
+  \frac{ 1 }
+  { \gamma \, t + \alpha_0^{-1} }
   ,
-  \notag
-  %\label{eq:alphas2}
+  \label{eq:alphas_t}
 \end{equation}
@@
   \frac{ 1 + r }
-  {
-    \frac{1}{\alpha_0}
-    +
-    \frac{1-r}{\ln(1/r)} \,
-    T
-  }
+  { \gamma \, T + \alpha_0^{-1} }
   ,
@@
 which is minimized as $r \to 1$
-for a large $T$.
+for a large $T$
+(at this point $\gamma \to 1$).
 %
@@
 then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
-and the optimal $t_0 = 1/\alpha_0$.
+and the optimal $t_0 = \alpha_0^{-1}$.
 %
 Thus, in this limit,
-the schedule given by Eq. \eqref{eq:alphas}
+the schedule given by Eq. \eqref{eq:alphas_t}
 is identical to that from Eq. \eqref{eq:alpha_invt1}.
 
-\subsection{\label{sec:adaptinvt}
-Inverse-time schedule from adaptive averaging
+\section{\label{sec:adaptinvt}
+Inverse-time schedule as a result of increasing sample size
+in adaptive averaging
 }
@@
     $p_i$           &   Distribution. \\
-    $\pi_i$         &   (Instantaneous) distribution, Eq. \eqref{eq:pi_p_phi}. \\
+    $\pi_i$         &   (Instantaneous) distribution, Eq. \eqref{eq:pi_p_V}. \\
     $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\

### 2017-01-23 doc/invt.tex
invt.tex: adding new Appendix A

@@
 \begin{equation}
-  V_i = \ln p^*_i - \ln p_i + \mathrm{const.}
+  V_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
   ,
@@
 %
-This relation thus furnishes a practical aim of convergence for FDS
+This relation thus sets a practical goal for the convergence of FDS
 methods---adjusting the bias potential to make the
-observed distribution, which can be computed from the normalized histogram,
+observed distribution, or the normalized histogram,
 match the intended one.
@@
 deduced from the bias potential. % $V_i$.
+\note{
+  For example, in the context of simulated tempering\cite{zhang2007},
+  $p^*_i$ is replaced by $\ln Z_i$,
+  $p_i$ is the weight $\zeta_i$,
+  and the bias potential $V_i$ is $\ln (\tilde Z_i / \zeta_i)$.
+  Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
+  up to a constant shift.
+}
 
@@
 are often improved
-with the histogram of the sampling, $H_i$, accumulated
+with the normalized histogram of the sampling, $H_i$, accumulated
 during simulation.
@@
 Thus, the rigidity of this framework disallows
-continual improvement of the bias potential, hence that of the PMF,
+continuous improvement of the bias potential, hence that of the PMF,
 as simulation lengthens.
@@
 address the above shortcomings
-by continually and incrementally update the target distribution.
+by continuously and incrementally update the target distribution.
 %For this reason, the second type of methods,
@@
 %
-where $w_{ij}$ are a set of coefficients satisfying
+where the elements, $w_{ij}$'s, are a set of coefficients satisfying
 the restrictions to be discussed
@@
 
-\subsubsection{Differential equation}
-
-We first approximate the finite difference Eq. \eqref{eq:wl_update}
-by a differential equation
-%
-\begin{equation}
-  \dot V_i(t)
-  \equiv
-  \frac{ dV_i(t) } { dt }
-  =
-  \alpha(t) \, \frac{ h_i(t) } { p_i },
-  \label{eq:vt_diffeq}
-\end{equation}
-%
-with
-%
-\begin{equation}
-  h_i(t) = \delta_{i, i(t)}
-  ,
-  \label{eq:h_def}
-\end{equation}
-%
-being the instantaneous histogram,
-which is equal to $1.0$
-for the bin $i(t)$ at time $t$
-or zero otherwise.
-
-Next, we split $h_i(t)$ into a deterministic averaged part
-and a stochastic part, $\zeta_i(t)$,
-%
-\begin{equation}
-  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
-  \label{eq:h_split}
-\end{equation}
-%
-The deterministic part can be related
-to the ``ensemble average'' of $h_i$,
-where the ensemble consists of many similar simulation copies
-sharing the same schedule $\alpha(t)$
-the same bias potential at time $t$.
-%
-The initial states and the stochastic forces
-during the process may differ, however.
-
-For sufficiently small $\alpha(t)$,
-the bias potential remains roughly the same for a short period.
-%
-Then,
-the sampling process is approximately Markovian, % of a finite order,
-and we may assume Eq. \eqref{eq:pi_p_V}
-for the deterministic part
-%
-\begin{equation}
-  \langle h_i(t) \rangle
-  \approx
-  \pi_i(t)
-  =
-  \frac{                p^*_i \, e^{-V_i(t)} }
-       { \sum_{j = 1}^n p^*_j \, e^{-V_j(t)} }.
-  \label{eq:h_ave}
-\end{equation}
-%
-%
-%
-For the random noise part,
-the average is zero by definition,
-$\langle \zeta_i(t) \rangle = 0$,
-%
-%The noise is not necessarily white,
-and the autocorrelation functions
-depend only on the time difference
-%
-\begin{equation}
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  =
-  \sigma_{ij}(t - t'),
-  \notag
-  %\label{eq:zeta_zeta_correlation}
-\end{equation}
-%
-where $\sigma_{ij}(t)$ is an even function
-that vanishes at large $t$.
-%
-\note{More explicitly,
-if the transition matrix of a period $\tau$
-is $T^\tau_{ij}$,
-then
-$
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  =
-  \langle h_i(t) \, h_j(t') \rangle
-  -
-  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
-  =
-  T^{t - t'}_{ij} \pi_j - \pi_i \, \pi_j.
-$
-}%
-
-\subsubsection{Error and linear approximation}
-
-Next, we need to define the error in the bias potential.
+\subsubsection{Error}
+
+First, we need to define the error in the bias potential.
 %
@@
 upon convergence.
-%Since the last two terms of Eq. \eqref{eq:v_def}
-%remain constant during the course of simulation,
-%the updates to $v_i$'s are
-%the same as the updates to $V_i$'s.
+Note that the updates to $v_i$'s are
+the same as the updates to $V_i$'s,
+since the shift
+remains a constant during the course of simulation,
 %%
@@
 
-Second, the bias potential following Eq. \eqref{eq:vt_diffeq}
+Second, the bias potential following Eq. \eqref{eq:wl_update}
 generally grows over time.
@@
 does not affect the
-resulting distribution $\pi_i(t)$ as given by Eq. \eqref{eq:h_ave},
+resulting distribution $\pi_i(t)$,
 %
@@
 %
+We, therefore, define
+the error of the bias potential as
+%
+\begin{equation}
+  \Err
+  =
+  \sum_{i = 1}^n \rho_i \, \langle x_i^2 \rangle
+  =
+  \sum_{i = 1}^n \rho_i \, \left\langle (v_i - \bar v)^2 \right\rangle
+  .
+\label{eq:error_sum}
+\end{equation}
 
 Note that the distribution, $\pmb \rho$,
-introduced above
-will be used to measure the error,
-and is independent of
+introduced above is independent of
 the target distribution $\mathbf p$.
@@
 %
-The average growth rate of $v(t)$ is given by
-\begin{equation}
-  \frac{ d \bar v (t) }
-       { d t }
-  =
-  \alpha(t)
-  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
-  .
-\label{eq:dvbardt}
-\end{equation}
-%
-
-In the asymptotic regime,
-we expect the deviation $x_i(t)$
-to be small.
-%
-Expanding Eq. \eqref{eq:h_ave} in $x_i$ yields
+
+\subsubsection{Differential equation}
+
+To proceed, we will
+approximate the finite difference Eq. \eqref{eq:wl_update}
+by a differential equation
+%
+\begin{equation}
+  \dot V_i(t)
+  \equiv
+  \frac{ dV_i(t) } { dt }
+  =
+  \alpha(t) \, \frac{ h_i(t) } { p_i },
+  \label{eq:vt_diffeq}
+\end{equation}
+%
+with
+%
+$
+h_i(t) = \delta_{i, i(t)}
+$
+%\begin{equation}
+%  h_i(t) = \delta_{i, i(t)}
+%  ,
+%  \label{eq:h_def}
+%\end{equation}
+%
+being the instantaneous histogram,
+which is equal to $1.0$
+for the bin $i(t)$ at time $t$
+or zero otherwise.
+
+Next, we split $h_i(t)$ into a deterministic averaged part
+and a stochastic part, $\zeta_i(t)$,
+%
+\begin{equation}
+  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
+  \label{eq:h_split}
+\end{equation}
+%
+The deterministic part can be related
+to the ``ensemble average'' of $h_i$,
+where the ensemble consists of many similar simulation copies
+sharing the same schedule $\alpha(t)$
+the same bias potential at time $t$.
+%
+The initial states and the stochastic forces
+during the process may differ, however.
+
+For sufficiently small $\alpha(t)$,
+the bias potential remains roughly the same for a short period.
+%
+Then,
+the sampling process is approximately Markovian, % of a finite order,
+and we may assume Eq. \eqref{eq:pi_p_V}
+for the deterministic part
 %
 \begin{align}
-  \frac{
   \langle h_i(t) \rangle
-  }
-  {
-    p_i
-  }
   &\approx
-  \frac{                        e^{- x_i} }
-       { \sum_{ j = 1}^n p_j \, e^{- x_j} }
-  \approx
-  1 - x_i + \sum_{r = 1}^n p_r \, x_r
+  \pi_i(t)
+  =
+  \frac{                p_i \, e^{-x_i(t)} }
+       { \sum_{j = 1}^n p_j \, e^{-x_j(t)} }
+  \notag
+  \\
+  &\approx
+  p_i \,
+  \left(
+    1 - x_i + \sum_{r=1}^n p_r \, x_r
+  \right)
   ,
-\label{eq:hdet_linearize}
+  \label{eq:h_ave}
 \end{align}
 %
-where we have used $\sum_{j=1}^n p_j = 1$.
+where we have assumed the smallness
+of $x_i(t)$ in the linear expansion
+and used $\sum_{j=1}^n p_j = 1$
+on the second line.
 %
 \note{
-The second step of Eq. \eqref{eq:hdet_linearize} follows from
+The second step follows from
 $$
@@
 }%
+
+%
+For the random noise part,
+the average is zero by definition,
+$\langle \zeta_i(t) \rangle = 0$,
+%
+%The noise is not necessarily white,
+and the autocorrelation functions
+depend only on the time difference
+%
+\begin{equation}
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  =
+  \sigma_{ij}(t - t'),
+  \notag
+  %\label{eq:zeta_zeta_correlation}
+\end{equation}
+%
+where $\sigma_{ij}(t)$ is an even function
+that vanishes at large $t$.
+%
+\note{More explicitly,
+if the transition matrix of a period $\tau$
+is $T^\tau_{ij}$,
+then
+$
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  =
+  \langle h_i(t) \, h_j(t') \rangle
+  -
+  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
+  =
+  T^{t - t'}_{ij} \pi_j - \pi_i \, \pi_j.
+$
+}%
+
+The average growth rate of $v(t)$ is given by
+\begin{equation}
+  \frac{ d \bar v (t) }
+       { d t }
+  =
+  \alpha(t)
+  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
+  .
+\label{eq:dvbardt}
+\end{equation}
+%
+
 From Eqs.
-\eqref{eq:vt_diffeq},
-\eqref{eq:h_split},
-\eqref{eq:x_def}-\eqref{eq:hdet_linearize},
+\eqref{eq:x_def},
+\eqref{eq:rhox_sum}-\eqref{eq:dvbardt},
+%\eqref{eq:rhox_sum},
+%\eqref{eq:vt_diffeq},
+%\eqref{eq:h_split}
+%\eqref{eq:h_ave},
+%\eqref{eq:dvbardt},
 we get a set of decoupled equations
@@
   \\
-  &\stackrel{\mathrm{Eqs.\; \eqref{eq:rhox_sum} \;and\; \eqref{eq:hdet_linearize}}}
+  &\stackrel{\mathrm{Eqs.\; \eqref{eq:rhox_sum} \;and\; \eqref{eq:h_ave}}}
             {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
@@
 
-We define
-the total error of the bias potential as
-%
-\begin{equation}
-  \Err
-  =
-  \sum_{i = 1}^n \rho_i \, \langle x_i^2 \rangle
-  =
-  \sum_{i = 1}^n \rho_i \, \left\langle (v_i - \bar v)^2 \right\rangle
-  .
-\label{eq:error_sum}
-\end{equation}
-%
-The expression depends implicitly on the function $\alpha(t)$.
+The total error defined in Eq. \eqref{eq:error_sum}
+now depends implicitly on the function $\alpha(t)$
+via $x_i(t)$.
 %
 Our aim is to find the $\alpha(t)$
-that minimizes the error $E$
+that minimizes the error
 at the end of the simulation period $T$.
@@
 Then, Eq. \eqref{eq:vt_diffeq} is generalized to
-\begin{equation}
-  \dot v_i(t)
+%
+\begin{equation}
+  \dot V_i(t)
   =
@@
 $\mathbf p = (p_1, \dots, p_n)$,
- $\dot v_i(t)$
+ $\dot V_i(t)$
 in Eq. \eqref{eq:vt_diffeq_mbin}
@@
 %
-This allows $v_i(t)$'s to grow uniformly
+This allows the bias potential to grow uniformly
 in the asymptotic regime.
@@
 \eqref{eq:rhox_sum},
-\eqref{eq:hdet_linearize},
+\eqref{eq:h_ave},
 \eqref{eq:xi_def},
@@
 For the Gaussian updating kernel,
-if the width $\sigma$ is satisified
-with $\Delta x \ll \sigma \ll 1$,
+if the width $\sigma$ satisfies $\Delta x \ll \sigma \ll 1$,
 we can extend
@@
 Since the first eigenmode corresponds to
-a uniform shift of all $v_k$'s,
+a uniform shift of the bias potential,
 we can freely set $\lambda_1$ to be the same as $\lambda_2$
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-that ignore high-frequency eigenmodes.
+that ignore short-wave eigenmodes,
+which are typically local noises.
 %
@@
 %
-and at this point, we get
+and at this point, we get,
 from Eqs. \eqref{eq:error_split},
@@
 \eqref{eq:tk0},
+and
 \eqref{eq:error_asym_sinc},
@@
 %
-In Appendix \ref{sec:equilerr},
-we show that the above error tends
-to the error of the ideal equilibrium FDS simulation
-in the asymptotic limit.
+%In Appendix \ref{sec:equilerr},
+%we show that the above error tends
+%to the error of the ideal equilibrium FDS simulation
+%in the asymptotic limit.
 
@@
     In the latter case,
+    the equivalent width
     $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
@@
 
-Our method of analysis is most useful for long simulations
-to achieve relatively precise results.
+In this analysis, we have ignored the \emph{systematic} error\cite{zhou2005}
+caused by the updates in adaptive FDS methods.
+%
+This is a valid assumption in the asymptotic regime,
+where the random error is much greater than the systematic one.
+%
+%Our method of analysis is most useful for long simulations
+%to achieve relatively precise results.
 %
@@
 %
-Although these assumptions appeared to work well
-on our test system,
+Although this assumption appeared to work well
+on our test system for modeling long simulations,
 they may break down for relatively short simulations
@@
 We thank Dr. J. Ma, Dr. Y. Mei, J. Wang,
-O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz, C. Myers, and O. Samoylova
+O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
+Dr. C. Myers, and Dr. O. Samoylova
 for helpful discussions.
@@
 \appendix
+
+\section{\label{sec:equilerr}
+Inverse-time schedule as a result of increasing sample size
+in a runtime average
+}
+
+We can understand the optimality of the inverse-time schedule
+as a result of the stabilization of a runtime average
+with respect to the increase of sample size\cite{marsili2006, barducci2008}.
+%
+Consider a sample, $\{X^{(1)}, \dots, X^{(N)}\}$,
+with a dynamically growing size.
+%
+The runtime average
+%
+\begin{equation*}
+  {\bar X}^{(N)}
+  =
+  \frac{ X^{(1)} + \cdots + X^{(N)} } { N }
+  ,
+\end{equation*}
+%
+can be written as a recurrence relation regarding
+the addition of the last data point.
+%
+\begin{align}
+  {\bar X}^{(N)}
+  &=
+  {\bar X}^{(N-1)} + \frac{ X^{(N)} - {\bar X}^{(N-1)} } { N }
+  .
+  \label{eq:av_recur}
+\end{align}
+%
+If we interpret $X^{(N)} - {\bar X}^{(N-1)}$
+as the independent correction to the previous average,
+${\bar X}^{(N-1)}$, proposed by the latest data point, $X^{(N)}$,
+then this correction is absorbed into the average with
+a diminishing weight $1/N$ as the sample size, $N$, increases.
+
+Let us now consider a long equilibrium FDS simulation
+that consists of blocks of sub-simulations
+of $L$ steps.
+%
+If $X^{(N)}$ is
+the bias potential, $V_i^{(N)}$, computed
+from the $N$th sub-simulation,
+then ${\bar X}^{(N)}$ stands for the average bias potential,
+${\bar V}_i^{(N)}$,
+from the first $N$ sub-simulations,
+which is the best estimate at the end of
+the first $N$ sub-simulations.
+%
+If we wish to always use this best estimate, ${\bar V}_i^{(N-1)}$,
+from the previous $N-1$ sub-simulations
+as the bias potential for the $N$th sub-simulations,
+then this estimate can be updated according to Eq. \eqref{eq:av_recur}.
+%
+Assuming that the normalized histogram, $H_i^{(N)}$,
+from the $N$th sub-simulations is sufficiently precise,
+Eq. \eqref{eq:vcorr_equil} shows that the independent
+correction from the $N$th sub-simulation is
+%
+\begin{equation*}
+  \ln \frac{ H_i^{(N)} } { p_i }
+  \approx
+  \frac{ H_i^{(N)} } { p_i } - 1
+  ,
+\end{equation*}
+%
+where, the $-1$ on the right-hand side
+can be ignored as a uniform shift
+does not affect the distribution.
+%
+Thus, we may rewrite Eq. \eqref{eq:av_recur} as
+\begin{align}
+  {\bar V}_i^{(N)}
+  &=
+  {\bar V}_i^{(N-1)}
+  +
+  \frac{1}{N} \frac{ H_i^{(N)} }{ p_i }
+  \notag
+  \\
+  &=
+  {\bar V}_i^{(N-1)}
+  +
+  \sum_{l=1}^L \frac{1}{N \, L} \frac{ h_i^{(N)}(l) }{ p_i }
+  ,
+  \label{eq:vbar_update}
+\end{align}
+%
+where $h_i^{(N)}(l)$ stands for the instantaneous histogram
+at step $l$ of sub-simulation $N$,
+and it is related to the average $H_i^{(N)}$ as
+$H_i^{(N)} \equiv \frac{1}{L} \sum_{l=1}^L h_i^{(N)}(l)$.
+%
+Thus, Eq. \eqref{eq:vbar_update} shows that the best estimate
+or the runtime average of the bias potential can be computed
+through a step-by-step incremental updating scheme that is equivalent to
+the WL algorithm, Eq. \eqref{eq:wl_update},
+with the magnitude being
+%
+\begin{equation}
+  \alpha(t) = 1/(NL)
+  .
+  \label{eq:alpha_block}
+\end{equation}
+%
+In the long time limit,
+we may approximate the simulation time
+by the product of the number of sub-simulations
+and the number of steps in each sub-simulation, or $NL$,
+%
+and Eq. \eqref{eq:alpha_block} recovers Eq. \eqref{eq:alpha_invt}.
+
+The above demonstration has several practical implications.
+%
+First,
+since the bias potential is often sufficiently precise
+in late stages of a WL simulation,
+some practitioners would then fix
+the bias potential and switch
+to the equilibrium FDS method
+to prevent possible future efficiency loss
+due to the adaptive updates.
+%
+However, the above demonstration shows that
+this practice becomes unnecessary
+with the inverse-time schedule.
+%
+That is,
+the WL algorithm with the inverse-time schedule
+produces a bias potential that is similar in precision to that from 
+the ideal equilibrium simulation for long times.
+
+Second,
+an alternative way of improving the precision of the bias potential
+from a simulation using the WL algorithm or metadynamics
+with a fixed magnitude
+is to time average the bias potential\cite{zhou2005}.
+%
+This practice becomes redundant with the inverse-time schedule
+because the averaging of the bias potential
+has been implicitly performed by the schedule.
+%
+The inverse-time schedule approach is preferred, however,
+because with a decreasing updating magnitude,
+it minimizes the systematic error caused by
+the on-the-fly updating scheme\cite{zhou2005},
+which is ignored in this analysis.
+
+Third,
+it is possible to reformulate the WL algorithm
+with the inverse-time schedule
+such that the bias potential is computed as
+a weighted adaptive average of the histogram\cite{marsili2006, barducci2008},
+which is analogous to the adaptive bias force method\cite{darve2001, *darve2008}.
 
@@
 is identical to that from Eq. \eqref{eq:alpha_invt1}.
-
-\section{\label{sec:adaptinvt}
-Inverse-time schedule as a result of increasing sample size
-in adaptive averaging
-}
-
-Result of the increase of the sample size.
-
-Final bias potential is already the average,
-No further trajectory averaging is needed.
-Fixed magnitude, improvable by trajectory averaging.
 
@@
 %
+Thus, Eq. \eqref{eq:kappa_delta} is exact,
+with
+$
+\Gamma_k
+=
+\sum_{ i = 1}^n \phi_{ki}^2 / p_i.
+$
 \note{Proof.%Eq. \eqref{eq:kappa_perfect},
@@
 In particular, if $p_i = \rho_i$, then, by
-Eq. \eqref{eq:eig_orthonormal_rows}
-$$
-\kappa_k(t - t') = 1.
-$$
-Thus, Eq. \eqref{eq:kappa_delta} is exact,
-with
-$$
-\Gamma_k
-=
-\sum_{ i = 1}^n \frac{ \phi_{ki}^2 } { p_i },
-$$
-or if $\mathbf p = \pmb \rho$,
+Eq. \eqref{eq:eig_orthonormal_rows}, we have
+$\kappa_k(t - t') = \delta_{t, t'}$, or
 \begin{equation}
@@
 and hence
-cannot compute the shifted bias potential, $v_i$,
-from the original one, $V_i$
-(cf. Sec. \ref{sec:FDS}).
+cannot compute the deviation
+of the bias potential, $V_i$.
 %
@@
 \equiv
-\sum_{ j = 1 }^n p_j \, V_j.
+\sum_{ j = 1 }^n \rho_j \, V_j.
 $
@@
   \sum_{ j = 1 }^n
-    p_j \, \ln \frac{ p_j } { p^*_j }.
+    \rho_j \, \ln \frac{ p_j } { p^*_j }.
 $
@@
 }
-
-\section{\label{sec:equilerr}
-Error of the ideal equilibrium FDS simulation}
-
-%As discussed in Sec. \ref{sec:FDS},
-%the asymptotic limit of an adaptive FDS method
-%is the corresponding equilibrium FDS method
-%with the bias potential fixed to the exact value.
-%%
-%Such a simulation delivers the maximal efficiency.
-
-In late stages of an adaptive FDS simulation,
-the bias potential is often sufficiently precise.
-Some practitioners would then fix
-the bias potential and switch
-to the equilibrium method
-to prevent possible future efficiency loss
-due to the adaptive updates.
-%
-Below, we will show that this practice is unnecessary
-if the simulation uses the single-bin or a bandpass updating scheme
-with the optimal schedule, Eq. \eqref{eq:alpha_invt1}.
-%
-That is, the bias potential obtained from
-an optimal single-bin scheme adaptive simulation
-will be similar in precision to that from
-the ideal equilibrium simulation
-for long times.
-%
-%Thus, our demonstration below serves as an alternative proof
-%of the optimality of the single-bin scheme.
-
-The error of the equilibrium FDS simulation
-depends on the precision of the histogram.
-%
-If all $v_i$'s are exact,
-the error is minimized, and can be computed
-from the second term
-on the right-hand side of Eq. \eqref{eq:vcorr_equil} as
-%
-\begin{align}
-  E
-  &=
-  \left\langle
-    \sum_{ i = 1 }^ n
-      \rho_i \,
-      \left(
-        \ln \frac { H_i }
-                  { p_i }
-        -
-        \sum_{r = 1}^n \rho_r
-        \ln \frac { H_r }
-                  { p_r }
-      \right)^2
-  \right\rangle
-  \notag
-  \\
-  &
-  \approx
-  \sum_{ i = 1 }^ n
-    \rho_i \,
-    \left\langle
-      \left(
-        \frac { H_i }
-              { p_i }
-        -
-        \sum_{r = 1}^n \rho_r
-            \frac { H_r }
-                  { p_r }
-      \right)^2
-    \right\rangle,
-\notag
-\end{align}
-where
-we have defined a time-averaged histogram
-$H_i = \frac{1}{T} \sum_{t = 1}^T h_i(t)$
-with the instantaneous histogram
-$h_i(t)$ defined by Eq. \eqref{eq:h_def}.
-
-We can rewrite the error as
-a sum of the eigenmodes as
-%
-\begin{align}
-  E
-  =
-  \sum_{ k = 2 }^n
-    \left\langle
-      \Theta_k^2
-    \right\rangle
-  ,
-\notag
-%\label{eq:error_sbin_equil}
-\end{align}
-%
-where
-%
-$
-  \Theta_k
-  =
-  \frac{ 1 } { T }
-  \sum_{ t = 1 } ^ T
-    \theta_k( t )
-  ,
-$
-%
-with $\theta_k(t)$ defined by Eq. \eqref{eq:theta_def}.
-%and $\phi_{k i}$ by Eq. \eqref{eq:eig_orthonormal_cols}.
-%\begin{align}
-%\theta_k( t )
-%=
-%\sum_{ i = 1 }^ n
-%\phi_{k i} \left[
-%             \frac{ h_i(t) }
-%                  { p_i    }
-%             - 1
-%           \right],
-%\end{align}
-%and
-%\begin{align}
-%\sum_{ k = 1 }^n
-%  \phi_{k i} \, \phi_{k j}
-%  = p_i \, \delta_{i j}.
-%\end{align}
-%
-\note{Because
-%
-\begin{align*}
-  \sum_{ k = 1 }^n
-    \Theta_k^2
-  &=
-  \frac{ 1 } { T^2 }
-  \sum_{ t, t' = 1 }^T
-    \sum_{ k = 1 }^n
-      \theta_k( t ) \, \theta_k( t' )
-  \\
-  &=
-  \frac{ 1 } { T^2 }
-  \sum_{ t, t' = 1 }^T
-    \sum_{ i, j = 1 }^n
-      \xi_i(t) \, \xi_j(t')
-      %\left[
-      %  \frac{ h_i(t) }
-      %       { p_i    }
-      %  - 1
-      %\right]
-      %\left[
-      %  \frac{ h_j(t') }
-      %       { p_j     }
-      %  - 1
-      %\right]
-      \sum_{ k = 1 }^n
-        \phi_{k \, i} \, \phi_{k \, j}
-  \\
-  &=
-  \frac{ 1 } { T^2 }
-  \sum_{ t, t' = 1 }^T
-    \sum_{ i, j = 1 }^n
-      \xi_i(t) \, \xi_j(t') \,
-      \rho_i \, \delta_{i, j}
-  \\
-  &=
-  \sum_{ i = 1 }^n
-  \rho_i \,
-  \frac{ 1 } { T }
-  \sum_{ t = 1 }^T \xi_i(t)
-  \;
-  \frac{ 1 } { T }
-  \sum_{ t' = 1 }^T \xi_i(t')
-  \\
-  &=
-  \sum_{ i = 1 }^n
-    \rho_i \,
-      \left(
-        \frac{ H_i }
-             { p_i }
-        -
-        \sum_{r = 1}^n \rho_r
-        \frac{ H_r }
-             { p_r }
-      \right)^2
-  ,
-\end{align*}
-where we have used
-$$
-  \frac{1}{T}
-  \sum_{t = 1}^T \xi_i(t)
-  =
-  \frac{ H_i }
-       { p_i }
-  -
-  \sum_{r = 1}^n \rho_r
-  \frac{ H_r }
-       { p_r }
-  .
-$$
-}%
-The sum starts from $k = 2$, for
-by Eq. \eqref{eq:eigenmode1}, % we have
-$$
-\theta_1(t)
-=
-\sum_{ i = 1 }^n
-  \phi_{1i} \, \left(
-    \frac{ h_i(t) } { p_i }
-    -
-    \sum_{r = 1}^n
-    \rho_r
-    \frac{ h_r(t) } { p_r }
-  \right)
-= 0.
-$$
-
-In the long-time limit,
-\begin{align*}
-  \left\langle
-    \Theta_k^2
-  \right\rangle
-  &=
-  \frac{1}{T^2}
-  \sum_{ t, t' = 1 }^T
-  \langle \theta_k(t) \, \theta_k(t') \rangle
-  %\\
-  %&\approx
-  %\frac{1}{T}
-  %\sum_{ \tau = -\infty }^{ \infty }
-  %\langle \theta_k(t) \, \theta_k(t + \tau) \rangle
-  %\\
-  %&=
-  =
-  \frac{1}{T}
-  \sum_{\tau = -\infty}^\infty
-    \kappa_k(\tau)
-  =
-  \frac{ \Gamma_k }{ T }
-.
-\end{align*}
-%
-Thus,
-\begin{equation}
-  E
-  =
-  \frac{ 1 } { T }
-  \sum_{ k = 2 }^n \Gamma_k
-  .
-\label{eq:error_equil}
-\end{equation}
-
-Now comparing this with the error of
-the optimal adaptive simulation,
-Eq. \eqref{eq:error_sinc} (with $K = n$),
-we find the difference to be negligible
-in the long time limit.
-
-The above argument can be extended
-to the bandpass updating schemes.
-%
-If the intrinsic PMF is sufficiently smooth,
-we can explicitly filter out some error modes
-in the histogram,
-say those with $k >  K$,
-and truncate the sum in
-Eq. \eqref{eq:error_equil} at $k = K$.
 
@@
     $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
-    $h_i(t)$        &   Histogram.  \\
-    $H_i$           &   Average histogram.  \\
+    $h_i(t)$        &   Instantaneous histogram.  \\
+    $H_i$           &   Average normalized histogram.  \\
     $i, j$          &   Bin indices. \\
@@
     $\xi_i(t)$      &   The noise, Eq. \eqref{eq:xi_def}. \\
-    $\chi_i$        &   $\chi_i \equiv \ln \frac{ p_i } { p^*_i } - \sum_{ j = 1 }^n p_j \, \ln \frac{ p_j } { p^*_j }$. \\
+    $\chi_i$        &   $\chi_i \equiv \ln \frac{ p_i } { p^*_i } - \sum_{ j = 1 }^n \rho_j \, \ln \frac{ p_j } { p^*_j }$. \\
     $y_k(t)$        &   Mode of the shifted bias potential, Eq. \eqref{eq:y_def}. \\

### 2017-01-23 doc/invt.tex
small changes to invt.tex

@@
-\documentclass[reprint, superscriptaddress, floatfix]{revtex4-1}
+\documentclass[preprint, superscriptaddress, floatfix]{revtex4-1}
 \usepackage{amsmath}
@@
 along $z$\cite{mezei1987, berg1992, lee1993,
-wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
+wang2001, *wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
 %The negative logarithm,
@@
 %the target distribution along $z$\cite{mezei1987, berg1992, lee1993,
-%wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
+%wang2001, *wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
 %%
@@
 
-The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
+The Wang-Landau (WL) algorithm\cite{wang2001, *wang2001pre}
 and metadynamics\cite{huber1994, laio2002,
@@
 hence that of the PMF\cite{liang2007,
-belardinelli2007, belardinelli2007jcp, belardinelli2008,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 morozov2007, zhou2008, morozov2009,
-komura2012, caparica2012, caparica2014,
+komura2012, caparica2012, *caparica2014,
 barducci2008, dickson2011, dama2014}.
@@
 inverse time\cite{liang2007,
-belardinelli2007, belardinelli2007jcp, belardinelli2008,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 morozov2007, zhou2008}
@@
 liang2007,
-belardinelli2007, belardinelli2007jcp, belardinelli2008,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 barducci2008}.
@@
 the optimal inverse-time prescription\cite{
-belardinelli2007, belardinelli2007jcp, belardinelli2008,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 morozov2007, zhou2008, morozov2009,
-komura2012, caparica2012, caparica2014},
+komura2012, caparica2012, *caparica2014},
 the WL algorithm is equivalent to the ideal equilibrium counterpart
@@
 %
-In the WL algorithm\cite{wang2001, wang2001pre},
+In the WL algorithm\cite{wang2001, *wang2001pre},
 the bias potential, $V_i(t)$, is updated
@@
 In the original WL papers\cite{
-wang2001, wang2001pre},
+wang2001, *wang2001pre},
 the updating magnitude $\alpha$
@@
 with $\alpha$ reduced by a factor of $1/2$\cite{
-wang2001, wang2001pre}.
+wang2001, *wang2001pre}.
 %
@@
 too quickly in late stages, making the asymptotic error
-saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
+saturate\cite{
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008}.
 
@@
 This simple yet effective formula has attracted
-several studies\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008,
+several studies\cite{
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 morozov2007, zhou2008, morozov2009,
-komura2012, caparica2012, caparica2014}.
+komura2012, caparica2012, *caparica2014}.
 %
@@
   =
-  \sum_{i = 1}^n \rho_i \, \langle x_i^2 \rangle
-  =
-  \sum_{i = 1}^n \rho_i \, \left\langle (v_i - \bar v)^2 \right\rangle
+  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
   .
@@
   p_i \,
-  \left(
-    1 - x_i + \sum_{r=1}^n p_r \, x_r
-  \right)
+  \left[
+    1 - x_i(t) + \sum_{r=1}^n p_r \, x_r(t)
+  \right]
   ,
@@
 in the original WL algorithm\cite{
-wang2001, wang2001pre}.
+wang2001, *wang2001pre}.
 
@@
 Eq. \eqref{eq:wband_eigenvalue_pbc},
-with the periodic $n$ doubled:
+with the period $n$ doubled:
 %
@@
 %
-\note{This inequality follows from the non-negativity of
-the following quadratic polynomial of $Y$,
+%\note{
+This inequality follows from the non-negativity of
+the following quadratic polynomial of $\chi$,
 $$
@@
   dt \sum_{k = 2}^n \Gamma_k \,
-    \left( \dot u_k \, Y - c_k \right)^2
+    \left( \dot u_k \, \chi - c_k \right)^2
   \equiv
-  A \, Y^2 + B \, Y + C
+  A_2 \, \chi^2 + A_1 \, \chi + A_0
   ,
@@
 which implies a non-positive discriminant,
-$B^2 - 4 \, A \, C$.
-}%
+$A_1^2 - 4 \, A_0 \, A_2$.
+%}%
 %
@@
 
-Since the first eigenmode corresponds to
+Since the first eigenmode represents
 a uniform shift of the bias potential,
@@
 %
+This error decreases inversely with increasing simulation length,
+as one would expect from an equilibrium FDS simulation.
 %In Appendix \ref{sec:equilerr},
@@
 
-We thank Dr. J. Ma, Dr. Y. Mei, J. Wang,
+We thank Dr. Y. Mei, J. Wang,
 O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
@@
 with the inverse-time schedule
-such that the bias potential is computed as
-a weighted adaptive average of the histogram\cite{marsili2006, barducci2008},
+such that the bias potential is computed from
+a weighted runtime average of the histogram\cite{marsili2006, barducci2008},
 which is analogous to the adaptive bias force method\cite{darve2001, *darve2008}.
@@
 we find from Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def} that
-$$
+\begin{equation*}
   \xi_i(t)
@@
   ,
-$$
+\end{equation*}
 %
@@
 with
-$
-\Gamma_k
-=
-\sum_{ i = 1}^n \phi_{ki}^2 / p_i.
-$
+$\Gamma_k = \sum_{ i = 1}^n \phi_{ki}^2 / p_i$.
 \note{Proof.%Eq. \eqref{eq:kappa_perfect},
@@
 %
-\begin{align}
+\begin{align*}
 Y_k
@@
   \phi_{k i} \, (V_i - \bar V),
-\label{eq:Y_def}
-\end{align}
+%\label{eq:Y_def}
+\end{align*}
 %
@@
 To show Eq. \eqref{eq:varY},
-we first use Eq. \eqref{eq:v_def}
-to rewrite Eq. \eqref{eq:x_def}
-as
-$
-  x_i
-  =
-  (V_i - \bar V) + \chi_i,
-$
-where
-$
-  \chi_i
-  \equiv
-  \ln \frac{ p_i } { p^*_i }
-  -
-  \sum_{ j = 1 }^n
-    \rho_j \, \ln \frac{ p_j } { p^*_j }.
-$
-Thus, by Eqs. \eqref{eq:y_def} and \eqref{eq:Y_def},
-we have
-$$
-y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, \chi_i.
-$$
-Since the second term on the right-hand side
-is a constant, and the average of $y_k$ is zero,
+we first note that
+the difference, $V_i - {\bar V}$,
+differs from the actual deviation $x_i$ of bin $i$,
+as defined in Eq. \eqref{eq:x_def},
+only by a constant shift,
+$\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
+%
+This means that the two would share the same variance
+under time averaging.
+%
+It follows that the corresponding linear combinations,
+$Y_k$ and $y_k$, would also share the same variance.
+%
+But since the average of $y_k$ is zero,
 we have
@@
 
-For perfect sampling, with $\mathbf p = \pmb \rho$,
+For perfect sampling, if $\mathbf p = \pmb \rho$,
+we have
 $\Gamma(p) = 1$ from Eq. \eqref{eq:Gamma_perfect},
@@
     $\xi_i(t)$      &   The noise, Eq. \eqref{eq:xi_def}. \\
-    $\chi_i$        &   $\chi_i \equiv \ln \frac{ p_i } { p^*_i } - \sum_{ j = 1 }^n \rho_j \, \ln \frac{ p_j } { p^*_j }$. \\
     $y_k(t)$        &   Mode of the shifted bias potential, Eq. \eqref{eq:y_def}. \\

### 2017-01-24 doc/invt.tex
invt.tex: cleaning up Appendix F

@@
 %
-This is done by subtracting from the bias potential
-two contributions.
+This is done by removing two contributions from the bias potential.
 %
@@
 Note that the updates to $v_i$'s are
-the same as the updates to $V_i$'s,
+equivalent to the updates to $V_i$'s,
 since the shift
@@
 In this way,
-the variables, $x_i(t)$'s,
-represent the error of the bias potential,
-and are expected to vanish on convergence.
+the variable, $x_i(t)$,
+represents the error of the bias potential,
+and is expected to vanish on convergence.
 %
@@
 or zero otherwise.
+%
+The shifted bias potential, $v_i(t)$,
+follows the same differential equation
+as the shift is a constant.
 
@@
 The deterministic part can be related
-to the ``ensemble average'' of $h_i$,
+to the ``ensemble average'',
 where the ensemble consists of many similar simulation copies
@@
 
-The average growth rate of $v(t)$ is given by
-\begin{equation}
-  \frac{ d \bar v (t) }
-       { d t }
-  =
-  \alpha(t)
-  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
-  .
-\label{eq:dvbardt}
-\end{equation}
+%The average growth rate of $v(t)$ is given by
+%\begin{equation}
+%  \frac{ d \bar v (t) }
+%       { d t }
+%  =
+%  \alpha(t)
+%  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
+%  .
+%\label{eq:dvbardt}
+%\end{equation}
 %
@@
 From Eqs.
-\eqref{eq:x_def},
-\eqref{eq:rhox_sum}-\eqref{eq:dvbardt},
+\eqref{eq:v_def}-\eqref{eq:vbar_def},
+\eqref{eq:rhox_sum}-\eqref{eq:h_ave},
+%\eqref{eq:v_def},
+%\eqref{eq:x_def},
+%\eqref{eq:vbar_def},
 %\eqref{eq:rhox_sum},
@@
   \dot v_i - \dot{ \bar v }
-  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:dvbardt}}}
+  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:vbar_def}}}
             {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
@@
   \\
-  &\stackrel{\mathrm{Eq.\; \eqref{eq:h_split}}}
+  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_split} and \; \eqref{eq:h_ave}}}
             {=\joinrel=\joinrel=\joinrel=\joinrel=}
@@
 a one-variable problem that
-minimizes each individual $\left\langle x_i^2 \right\rangle$.
+minimizes each individual
+$\left\langle x_i^2(T) \right\rangle$.
 
@@
     \left\langle
-      y_k^2
+    y_k^2(T)
     \right\rangle
@@
 %
-More generally, with perfect filtering
-of certain eigenmodes, we can construct a
+However, if certain eigenmodes for local noises
+are expected to vanish in the PMF,
+we can construct a
 class of bandpass updating schemes
@@
 the difference, $V_i - {\bar V}$,
-differs from the actual deviation $x_i$ of bin $i$,
-as defined in Eq. \eqref{eq:x_def},
+differs from the actual deviation, $x_i$,
+defined in Eq. \eqref{eq:x_def},
 only by a constant shift,
@@
 %
-This means that the two would share the same variance
+Thus, the two share the same variance
 under time averaging.
@@
 
-Next, we show that this bound represents
+Next, we will show that this bound represents
 updating schemes with a set of very different eigenvalues.
 %
-First, observe that the summand of Eq. \eqref{eq:mass_func},
-$\Gamma_k \, \left( \lambda_k e^{-\lambda_k \, Q} \right)^2$
-is a function that peaks at $\lambda_k = 1/Q$,
-so that the sum is dominated by the term at $k = k(Q)$ with
-%
-\begin{equation}
-  \lambda_{k(Q)}
-  \approx
-  \frac 1 Q
-  .
-  \label{eq:kQ_def}
-\end{equation}
-%
-Thus,
+%First, observe that the summand of Eq. \eqref{eq:mass_func},
+%$\Gamma_k \, \left( \lambda_k e^{-\lambda_k \, Q} \right)^2$
+%is a function that peaks at $\lambda_k = 1/Q$,
+%so that the sum is dominated by the term at $k = k(Q)$ with
+%%
+%\begin{equation}
+%  \lambda_{k(Q)}
+%  \approx
+%  \frac 1 Q
+%  .
+%  \label{eq:kQ_def}
+%\end{equation}
+%
+%Thus,
+First, we have for any mode $k^*$ $(k^* \ge 2)$,
+%
 \begin{align}
@@
   &
-  \gtrsim
+  \ge
   %\approx
-  \sqrt{ \Gamma_{ k(Q) } } \,
-  \lambda_{ k(Q) } \,
-  e^{ -\lambda_{ k(Q) } \, Q }
+  \sqrt{ \Gamma_{ k^* } } \,
+  \lambda_{ k^* } \,
+  e^{ -\lambda_{ k^* } \, Q }
+  ,
 \label{eq:MQ_lbound1}
+\end{align}
+%
+which would approach an equality if the $k^*$ mode
+dominates the mass function, $M(Q)$,
+defined in Eq. \eqref{eq:mass_func}.
+%
+Then, for the $k$th component of the asymptotic error,
+we have from Eq. \eqref{eq:error_asym_Lagrangian}
+%
+\begin{align}
+  \left\langle
+  y_k^2(T)
+  \right\rangle_A
+  &=
+  \int_0^T
+  \Gamma_k \, \lambda_k^2 \, e^{2 \, \lambda_k \, [q(t) - q(T)] }
+  \, {\dot q}^2(t) \, dt
+  \notag
   \\
-  &
-  \approx
-  \frac{ \sqrt{ \Gamma_{ k(Q) } } }
-       { Q }
-  e^{ -\lambda_{ k(Q) } \, Q }
-  .
-\label{eq:MQ_lbound}
+  &=
+  \frac{ C } { T }
+  \, \Gamma_k \, \lambda_k^2
+  \int_0^{q(T)}
+  \frac{ e^{-2 \, \lambda_k \, Q } } { M(Q) }
+  \,  \, dQ
+  \notag
+  \\
+  &\le
+  \frac C T \,
+  \frac{ \Gamma_k \, \lambda_k^2 } { \sqrt{ \Gamma_{k^*} } \, \lambda_{k^*} }
+  \int_0^{q(T)}
+    e^{-(2 \, \lambda_k - \lambda_{k^*} ) \, Q}
+    \, dQ
+  \notag
+  \\
+  &\le
+  \frac C T \,
+  \cdot
+  \frac{ \Gamma_k }
+  { \sqrt{ \Gamma_{k^*} } }
+  \cdot
+  \frac{ \lambda_k^2 }
+  { \lambda_{k^*} \, (2 \, \lambda_k -\lambda_{k^*}) }
+  ,
+  \label{eq:y2_ubound_k}
 \end{align}
-
-Now consider the $k$th error component, % of the asymptotic error,
+%
+where we have
+used Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:MQ_lbound1}
+on the second and third lines, respectively,
+and extended the upper limit to infinity on the last line.
+%
+For a set of well-separated eigenvalues,
+$\lambda_k$ depends on the mode index $k$
+much more critically than $\Gamma_k$.
+%
+Thus, the tightest upper bound
+can be approximately obtained from minimizing the last factor
+on the right-hand side of Eq. \eqref{eq:y2_ubound_k},
+and the minimal is reached at $k^* = k$,
+when $\lambda_{k^*} = \lambda_k$.
+%
+So
+%
 \begin{equation}
   \left\langle
-    y_k^2
+  y_k^2(T)
   \right\rangle_A
-  =
-  \frac C T \,
-  \Gamma_k \, \lambda_k^2
-  \int_0^\infty
-    \frac{ e^{-2 \, \lambda_k \, Q} }
-         { M(Q) }
-    \, dQ
-  .
-\notag
-\end{equation}
-%
-Using Eq. \eqref{eq:MQ_lbound}, we have
-$$
-\begin{aligned}
-  \left\langle
-    y_k^2
-  \right\rangle_A
-  &
-  \lesssim
-  \frac C T \, \Gamma_k \, \lambda_k^2
-  \int_0^\infty
-    \frac{ Q } { \sqrt{ \Gamma_{ k(Q) } } }
-    e^{-\left[ 2\,\lambda_k - \lambda_k(Q) \right] Q }
-    \, dQ
-  \\
-  &
-  \approx
-  \frac C T \,
-  \frac{ \Gamma_k } { \sqrt{ \Gamma_{ k(Q^*) } } }
-  \left[
-    \frac{ \lambda_k } { 2 \, \lambda_k - \lambda_{ k(Q^*) } }
-  \right]^2
-  ,
-\end{aligned}
-$$
-where we have approximated $k(Q)$
-by its value around the peak of the integrand
-$$
-  Q^*
-  =
-  \frac{ 1 } { 2 \, \lambda_k - \lambda_{ k(Q^*) } }
-  .
-$$
-%
-Comparing this to Eq. \eqref{eq:kQ_def},
-we identify $k(Q^*) \approx k$, so
-\begin{equation}
-  \left\langle
-    y_k^2
-  \right\rangle_A
-  \lesssim
+  \le
   %\approx
@@
 %
+By summing over $k$
+and using Eq. \eqref{eq:mint_ubound},
+we recover Eq. \eqref{eq:error_asym_ubound}.
+
 This result shows that with a set of well-separated eigenvalues,
@@
 %
-Note that we can turn Eq, \eqref{eq:y2_ubound}
-to an inequality by choosing $k(Q) \to k$
-for Eq. \eqref{eq:MQ_lbound1}.
-%
-Then by summing over $k$
-and using Eq. \eqref{eq:mint_ubound},
-we recover Eq. \eqref{eq:error_asym_ubound}.
-
-For a finite-length simulation,
-the integrals $\int_0^{q(T)} dQ$
-can be extended to infinity
-only if $q(T) \gg Q^*$.
-%
-In other words,
-the above asymptotic limit fails around
-$Q^* \approx q(T)$, or equivalently
-around mode $k^* = k\bigl( q(T) \bigr)$,
-where $\lambda_{ k^* } \approx 1/q(T)$.
-%
-\note{This case also exemplifies
-  the sensitivity to the simulation length.
-  %
-  In this limit,
-  $M(Q) \propto 1/Q$ by Eq. \eqref{eq:MQ_lbound},
-  and $C \propto \ln\bigl[ q(T) / q_0 \bigr]$
-  for some $q_0$,
-  and we expect that with fixed $a_0$,
-  $q(T)$ to increase almost linearly with $T$
-  by Eq. \eqref{eq:opt_qT}.
-  %
-  So the normalized asymptotic error,
-  $E_A \cdot T \propto \bigl( \ln[ q(T)/q_0 ] \bigr)^2 \sim (\ln T)^2$,
-  is sensitive to the simulation length,
-  and worse than the ideal case of a constant.
-}
+Note, however, that to approach this upper bound
+the simulation has to be long enough
+such that the upper limit of the integral
+can be extended from $q(T)$ to infinity
+in the last step of Eq. \eqref{eq:y2_ubound_k}.
+%
+As a result,
+the flatter profile from Eq. \eqref{eq:y2_ubound}
+would stop at mode $k$,
+at which $\lambda_k \approx 1/q(T)$.
+%In other words,
+%the above asymptotic limit fails around
+%$Q^* \approx q(T)$, or equivalently
+%around mode $k^* = k\bigl( q(T) \bigr)$,
+%where $\lambda_{ k^* } \approx 1/q(T)$.
+%%
+%\note{This case also exemplifies
+%  the sensitivity to the simulation length.
+%  %
+%  In this limit,
+%  $M(Q) \propto 1/Q$ by Eq. \eqref{eq:MQ_lbound},
+%  and $C \propto \ln\bigl[ q(T) / q_0 \bigr]$
+%  for some $q_0$,
+%  and we expect that with fixed $a_0$,
+%  $q(T)$ to increase almost linearly with $T$
+%  by Eq. \eqref{eq:opt_qT}.
+%  %
+%  So the normalized asymptotic error,
+%  $E_A \cdot T \propto \bigl( \ln[ q(T)/q_0 ] \bigr)^2 \sim (\ln T)^2$,
+%  is sensitive to the simulation length,
+%  and worse than the ideal case of a constant.
+%}
 

### 2017-01-26 doc/invt.tex
clean up

@@
 a distribution, $p^*(z)$,
-along a collective variable $z$ by sampling via either Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
+along a collective variable $z$ by sampling via either Monte Carlo\cite{
+  frenkel, newman, landau_binder} (MC)
 or molecular dynamics\cite{frenkel, karplus2002} (MD) simulations.
@@
 cancels the PMF, resulting sampling of a flat distribution
-along $z$\cite{mezei1987, berg1992, lee1993,
-wang2001, *wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
+along $z$\cite{mezei1987, berg1992, *lee1993,
+wang2001, *wang2001pre,
+huber1994,
+*laio2002, *laio2008, *barducci2011, *sutto2012}.
 %The negative logarithm,
@@
 %under different umbrella or bias potentials
-%along $z$\cite{mezei1987, berg1992, lee1993},
+%along $z$\cite{mezei1987, berg1992, *lee1993},
 %and then piece together the entire curve.
@@
 %A better solution is to artificially flatten
-%the target distribution along $z$\cite{mezei1987, berg1992, lee1993,
-%wang2001, *wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
+%the target distribution along $z$\cite{mezei1987, berg1992, *lee1993,
+%wang2001, *wang2001pre,
+%huber1994,
+%*laio2002, *laio2008, *barducci2011, *sutto2012}.
 %%
@@
 The Wang-Landau (WL) algorithm\cite{wang2001, *wang2001pre}
-and metadynamics\cite{huber1994, laio2002,
-laio2008, barducci2011, sutto2012}
+and metadynamics\cite{huber1994,
+*laio2002, *laio2008, *barducci2011, *sutto2012}
 are two widely-used and closely-related\cite{micheletti2004}
@@
 which takes the form of a discrete
-$\delta$-function (or a boxcar function one histogram bin wide)
+$\delta$-function (a boxcar function one histogram bin wide)
 in the WL case,
@@
 disrupts the underlying equilibrium
-sampling\cite{laio2002},
+sampling\cite{zhou2005, zhou2008},
 and for convergence
@@
 the precision of the final bias potential,
-hence that of the PMF\cite{liang2007,
-belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
-morozov2007, zhou2008, morozov2009,
-komura2012, caparica2012, *caparica2014,
+hence that of the PMF\cite{
+belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
+liang2007,
+morozov2007, zhou2008,
+komura2012, *caparica2012, *caparica2014,
 barducci2008, dickson2011, dama2014}.
@@
 is a function of the
-inverse time\cite{liang2007,
-belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
+inverse time\cite{
+belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
+liang2007,
 morozov2007, zhou2008}
@@
 the inverse number of simulation steps.
+\note{
+  About the origin of the inverse-time formula.
+  Besides the Belardinelli paper,
+  we can also find it in Eq. (6) of reference \cite{liang2007},
+  if I am not too mistaken.
+  But I find the paper is a bit hard to read.
+  There, however, the formula is $t_0/t$ instead of $1/t$.
+}
 %
@@
 the initial updating magnitude
-should be set to
+can be optimally set to
 half of the previous equilibrium value,
@@
 in a simulated tempering simulation\cite{
-marinari1992, lyubartsev1992}.
+marinari1992, *lyubartsev1992}.
 %
@@
 or nearly flat\cite{
-dayal2004, trebst2004, zhang2007, barducci2008, singh2011}.
+dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}.
 %
@@
 \note{
-  For example, in the context of simulated tempering\cite{zhang2007},
+  For example, in the context of simulated tempering (Zhang and Ma, 2007),
   $p^*_i$ is replaced by $\ln Z_i$,
@@
 entropic or multicanonical sampling\cite{
-berg1992, lee1993}
+berg1992, *lee1993}
 and simulated tempering\cite{
-marinari1992, lyubartsev1992}
+marinari1992, *lyubartsev1992}
 with $i$ being the index of energy
@@
 marsili2006,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
 liang2007,
-belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 barducci2008}.
@@
 For example, one can show that
-if the magnitude decreases according to
+if the magnitude falls according to
 the optimal inverse-time prescription\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
-morozov2007, zhou2008, morozov2009,
-komura2012, caparica2012, *caparica2014},
+belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
+liang2007,
+morozov2007, zhou2008,
+komura2012, *caparica2012, *caparica2014},
 the WL algorithm is equivalent to the ideal equilibrium counterpart
@@
 depends on the updating magnitude\cite{
-  zhou2005, liang2005, laio2005, bussi2006, poulain2006, liang2007,
-  morozov2007, zhou2008, morozov2009, crespo2010,
-  atchade2011, fort2015}.
+  belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
+  zhou2005, zhou2008, morozov2007,
+  laio2005, bussi2006, poulain2006, liang2007,
+  crespo2010, *atchade2011, *fort2015}.
 %%
@@
 saturate\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008}.
+belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
 
@@
 several studies\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
 morozov2007, zhou2008, morozov2009,
-komura2012, caparica2012, *caparica2014}.
+komura2012, *caparica2012, *caparica2014}.
+%
+An interesting explanation\cite{
+  marsili2006, barducci2008}
+comes from the interpretation where
+the bias potential under the WL updating scheme,
+Eq. \eqref{eq:wl_update}
+as a runtime average from all data collected before time, $t$,
+so that the addition of a new data point only
+affects the average by a weight of $1/t$,
+which is the inverse of the sample size, $t$
+(cf. Appendix \ref{sec:equilerr}).
 %
@@
 the variable, $x_i(t)$,
-represents the error of the bias potential,
-and is expected to vanish on convergence.
-%
-We, therefore, define
-the error of the bias potential as
-%
-\begin{equation}
-  \Err
+represents the actual deviation of the bias potential, $V_i(t)$,
+and the total error is given by
+%
+\begin{equation}
+  \Err(t)
   =
@@
 sharing the same schedule $\alpha(t)$
-the same bias potential at time $t$.
+and the same bias potential at time $t$.
 %
@@
 }%
+Below, we will assume the noise is white,
+such that
+$$
+\sigma_{ij}(t) = \Gamma_0 \, \delta(t)
+$$
+for some $\Gamma_0$,
+where $\delta(t)$ is the Dirac's delta function.
+%
+This approximation is valid if the simulation period, $T$,
+is much longer than the autocorrelation time of $\sigma_{ij}(t)$,
+%
+and for a non-white noise, 
+the equivalent value of $\Gamma_0$ can be obtained
+from the integral of the autocorrelation function
+$\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$.
 
@@
 
-The total error defined in Eq. \eqref{eq:error_sum}
+The total error
+at the end of the simulation, $t = T$,
+as defined in Eq. \eqref{eq:error_sum},
 now depends implicitly on the function $\alpha(t)$
@@
 Our aim is to find the $\alpha(t)$
-that minimizes the error
-at the end of the simulation period $T$.
+that minimizes this error.
 %
@@
 %
-where $\xi(t)$ is a generalized noise
-that is invariant under time translation:
+where $\xi(t)$ is a white noise, such that
 %
@@
   =
-  \kappa(t - t')
+  \Gamma \, \delta(t - t')
   .
 $
-%
-Additionally,
-we assume that $\kappa(t)$ is an even function and
-\begin{equation}
-  \lim_{t \rightarrow \infty} \kappa(t)
-  = 0
-  .
-  \label{eq:kappa_limit}
-\end{equation}
-%
-For example, for a white noise,
-$\kappa(t)$ is proportional to a
-Dirac $\delta$-function, $\delta(t)$.
 %
@@
 \rangle e^{-2 \, q(T)}
-\notag
-\\
-&\qquad
 +
-\int_0^T \int_0^T
-  \dot u\bigl( q(t) \bigr) \,
-  \dot u\bigl( q(t') \bigr) \,
-  \kappa(t - t') \, dt' \, dt.
+\Gamma \int_0^T
+  {\dot u}^2\bigl( q(t) \bigr) \,
+  \, dt.
 \label{eq:x2t_average}
@@
 of Eq. \eqref{eq:x2t_average} is fixed during
-the process.
+the process,
+while the second term is the action of a free particle,
+which must move at a constant velocity, $c$:
 %
@@
 %
-By the Euler-Lagrange equation, we get
-$$
-\begin{aligned}
-0
-&=
-\frac{d}{dt} \int_0^T
-  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
-\\
-&= \int_0^T
-  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
-\end{aligned}
-$$
-where we have dropped the boundary terms
-by assuming Eq. \eqref{eq:kappa_limit}
-and a large $T$.
+%By the Euler-Lagrange equation, we get
+%$$
+%\begin{aligned}
+%0
+%&=
+%\frac{d}{dt} \int_0^T
+%  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
+%\\
+%&= \int_0^T
+%  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
+%\end{aligned}
+%$$
+%where we have dropped the boundary terms
+%by assuming Eq. \eqref{eq:kappa_limit}
+%and a large $T$.
 %
@@
 %
-In order for this to hold for any $t$,
-we must have
-%
-$
-\ddot u\bigl( q(t') \bigr) = 0
-,
-$
-or
+%In order for this to hold for any $t$,
+%we must have
+%%
+%$
+%\ddot u\bigl( q(t') \bigr) = 0
+%,
+%$
+%or
 \begin{equation}
   \dot u\bigl( q(t') \bigr) = c
-  ,
+  .
 \label{eq:dudt_const}
@@
 %
-where $c$ is a constant of $t'$.
-%
-Using Eq. \eqref{eq:u_def},
-we get
+Then by using Eq. \eqref{eq:u_def}, we have
 $$
@@
 Eq. \eqref{eq:alpha_invt1}
-for it optimizes each term $\langle x_i^2 \rangle$,
+for it optimizes each term $\langle x_i^2(T) \rangle$,
 and hence the sum;
@@
 We will use this result to model the initial error
-before starting the schedule $\alpha(t)$ to be optimized.
-%
-That is, we assume
-$\alpha(t) = a_0$
-for $t < 0$, and
-the system has entered the sampling equilibrium
-under a fixed $a_0$ at $t = 0$.
+before starting the target schedule, $\alpha(t)$.
+%
+That is, we assume that there is an equilibration phase
+for $t < 0$ during which the adaptive FDS sampling was performed
+under a constant updating magnitude
+$\alpha(t) = a_0$, and at the end of the equilibration, $t = 0$,
+the system has entered the equilibrium state of sampling.
 %
@@
 i.e., the optimal initial updating magnitude
-is always half of the equilibration value.
+is always half of the equilibrium value.
 %
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-that ignore short-wave eigenmodes,
+that ignore short-wavelength eigenmodes,
 which are typically local noises.
@@
 %
+This equilibration process models
+a stage of the WL algorithm or
+a regular metadynamics
+with a fixed updating magnitude.
+%
 Although the bias potential is correct initially,
@@
 We repeat the above process $1000$ times (unless specified otherwise)
-and average the results.
+and report the averaged results.
 %specified by either
@@
 the kernel was truncated at
-$b = \min\{10 \, \sigma, n - 1\}$
+$b = \min\{10 \, \sigma, n - 1\}$ bins
 and numerically stabilized
@@
 the errors from the simulations
-agreed well those from the theory.
+agreed well with those from the theory.
 
@@
 for the global sampling process,
-although it remained around $1.0$
+although it stayed around $1.0$
 for the local sampling process,
-as discussed in Appendix \ref{sec:invt_nn}.
-%
-Thus, the inverse-time formula,
+also in agreement with the discussion
+in Appendix \ref{sec:invt_nn}.
+%
+Thus, the inverse-time schedule,
 Eq. \eqref{eq:alpha_invt1},
@@
 To demonstrate the optimal schedule from Eq. \eqref{eq:q_opt},
-we used the Gaussian (with $\sigma = 10$) updating scheme
+we used the Gaussian (with $\sigma = 10$ bins) updating scheme
 as an example.
+%
+This updating scheme is typically used in metadynamics simulations.
 %
@@
 damp out the short-wavelength eigenmodes
-(i.e., local noises),
-as shown in Fig. \ref{fig:xerr}.
-%
-If, however, we could ignore these local eigenmodes
-(by explicitly filtering
-them from the biased potential),
-then we can modify the optimal schedule
-from Eq. \eqref{eq:q_opt}
-and limit the sum
+(i.e., local noises).
+%
+As shown in Fig. \ref{fig:xerr},
+the inverse-time schedule was effective in error reduction
+only for the first few eigenmodes,
+whose eigenvalues were around $1.0$.
+%
+We can, however, ignore these local eigenmodes
+by explicitly filtering them out from the final bias potential,
+and truncating the error sum in Eq. \eqref{eq:error_sum}
 to the first few eigenmodes.
 %
-The resulting schedule
-would become similar to the inverse-time schedule,
+Then, the corresponding optimal schedule
+from Eq. \eqref{eq:q_opt} with the truncated error definition
+became similar to the inverse-time schedule,
 Eq. \eqref{eq:alpha_invt1},
-as shown in Fig. \ref{fig:optacmp}.
+as exemplified in Fig. \ref{fig:optacmp}.
 %
@@
     limiting the sum in Eq. \eqref{eq:q_opt}
-    to the first four terms.
+    to the first four terms,
+    or the four longest wavelength modes.
+    %
     They became similar to the inverse-time schedule,
@@
     the error components from the optimal schedule
-    would roughly maintain the relatively flat profile,
+    would roughly maintain a relatively flat profile,
     but meet the initial error curve
@@
 
-To verify Eq. \eqref{eq:half_alpha0},
+An interesting prediction from the theory is Eq. \eqref{eq:half_alpha0},
+which states that the optimal schedule should always start from
+half of the magnitude from the previous equilibration run.
+%
+To verify this,
 we computed the optimal schedules
-for different values of $q(T)$,
+for different values of $q(T)$
+for the Gaussian updating scheme,
 and plotted the final error
@@
 As shown in Fig. \ref{fig:erriascan},
-The minimal error was reached at
-$\alpha(0) = a_0/2 = 5 \times 10^{-5}$.
+The minimal error was indeed reached at
+$\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
+although the dependence on $\alpha(0)$
+is generally weak.
+%
+This dependence becomes even weaker
+in the single-bin updating scheme,
+as it only amounts to a shift of the origin of
+simulation time in the inverse-time schedule,
+Eq. \eqref{eq:alpha_invt1}.
 
@@
 with different widths, $\sigma$.
+%
+In Fig. \ref{fig:errsigscan},
+we show the errors, $\Err$,
+normalized by the shifted simulation length, $T+t_0$,
+so that the normalized errors
+would stay constant in the single-bin updating scheme
+(corresponding to the limit of $\sigma \to 0$).
 %
@@
 %
-However, if certain eigenmodes for local noises
-are expected to vanish in the PMF,
+However, if some eigenmodes
+correspond to local noises
+that are expected to be absent from the PMF,
 we can construct a
@@
 
-In this analysis, we have ignored the \emph{systematic} error\cite{zhou2005}
+The analysis in this study has two limitations.
+%
+First, we have ignored the \emph{systematic}
+error\cite{zhou2005, zhou2008}
 caused by the updates in adaptive FDS methods.
 %
-This is a valid assumption in the asymptotic regime,
-where the random error is much greater than the systematic one.
+This is valid in the asymptotic regime,
+where the updating magnitude is small
+and the sampling is a quasi-equilibrium one\cite{barducci2008, dama2014}.
+%
+Then the random error is much greater than the systematic one,
+because the former is proportional to the square root of the updating magnitude
+whereas the latter is proportional to the magnitude itself.
 %
@@
 %
-This allows us to assume the sampling process
-to be a quasi-equilibrium one\cite{barducci2008, dama2014},
-and to represent the autocorrelation function
+Second, by using the white noise approximation,
+we have represented the autocorrelation function
 of each eigenmode by a single number, $\Gamma_k$.
@@
 as a result of the stabilization of a runtime average
-with respect to the increase of sample size\cite{marsili2006, barducci2008}.
+with respect to the increase of sample size\cite{
+  marsili2006, barducci2008}.
 %
@@
 with a fixed magnitude
-is to time average the bias potential\cite{zhou2005}.
+is to time average the bias potential\cite{
+  zhou2005, zhou2008}.
 %
@@
 it minimizes the systematic error caused by
-the on-the-fly updating scheme\cite{zhou2005},
+the on-the-fly updating scheme\cite{
+  zhou2005, zhou2008},
 which is ignored in this analysis.
@@
 such that the bias potential is computed from
-a weighted runtime average of the histogram\cite{marsili2006, barducci2008},
-which is analogous to the adaptive bias force method\cite{darve2001, *darve2008}.
+a weighted runtime average of the histogram\cite{
+  marsili2006, barducci2008},
+which is analogous to the adaptive bias force method\cite{
+  darve2001, *darve2008}.
 

### 2017-01-26 doc/invt.tex
invt.tex: editing changes

@@
   the optimal schedule depends on the updating scheme (WL or metadynamics),
-  and the initial updating magnitude should be half of
+  and the initial updating magnitude is ideally half of
   the previous equilibrium value.
@@
 disrupts the underlying equilibrium
-sampling\cite{zhou2005, zhou2008},
+sampling\cite{zhou2005, morozov2007, zhou2008},
 and for convergence
@@
 %
-We will show that the optimal schedule
-depends on how the bias potential is updated,
-and hence is different in the WL and metadynamics cases.
+As the mass is defined from the updating window function,
+the resulting optimal schedule
+is different in the WL and metadynamics cases.
 %
@@
 For a large system at temperatures of interest,
-the distribution $p^*_i$ often tends to
+the distribution, $p^*_i$, often tends to
 be localized,
@@
 which is usually flat,
-i.e., $p_i$ constant for all $i$'s,
+i.e., $p_i$ is constant for all $i$'s,
 or nearly flat\cite{
@@
 %
-We will generally call such a biased simulation
+We will broadly call such a biased simulation
 a flat-distribution-sampling (FDS) simulation.
 %
-To fulfill this goal, the bias potential must satisfy
+To achieve this distribution, the bias potential must satisfy
 %
@@
 %
-This relation thus sets a practical goal for the convergence of FDS
-methods---adjusting the bias potential to make the
+This relation thus sets a practical aim for the convergence of FDS
+methods---the bias potential must be adjusted to make the
 observed distribution, or the normalized histogram,
@@
   $p_i$ is the weight $\zeta_i$,
-  and the bias potential $V_i$ is $\ln (\tilde Z_i / \zeta_i)$.
+  and the bias potential, $V_i$, is $\ln (\tilde Z_i / \zeta_i)$.
   Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
@@
 address the above shortcomings
-by continuously and incrementally update the target distribution.
+by continuously and incrementally updating the target distribution.
 %For this reason, the second type of methods,
@@
 For example, one can show that
-if the magnitude falls according to
+if the magnitude is controlled by
 the optimal inverse-time prescription\cite{
@@
 
-One can show that with a constant magnitude $\alpha(t) = \alpha > 0$,
+One can show that with a constant magnitude,
+$\alpha(t) = \alpha > 0$,
 the distribution collected from
@@
   belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-  zhou2005, zhou2008, morozov2007,
+  zhou2005, morozov2007, zhou2008,
   laio2005, bussi2006, poulain2006, liang2007,
@@
 %
-Thus, one has to reduce the magnitude $\alpha$ over time
+Thus, one has to reduce the magnitude, $\alpha$, over time
 to improve the estimate of the distribution,
@@
 wang2001, *wang2001pre},
-the updating magnitude $\alpha$
+the updating magnitude, $\alpha$,
 (or $\ln f$ there)
@@
 a better way
-of controlling the schedule $\alpha(t)$
+of controlling the schedule, $\alpha(t)$,
 in late stages
@@
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-morozov2007, zhou2008, morozov2009,
+morozov2007, zhou2008,
 komura2012, *caparica2012, *caparica2014}.
@@
   marsili2006, barducci2008}
-comes from the interpretation where
-the bias potential under the WL updating scheme,
-Eq. \eqref{eq:wl_update}
-as a runtime average from all data collected before time, $t$,
+comes from the interpretation of
+the bias potential %under the WL updating scheme, Eq. \eqref{eq:wl_update},
+as a runtime average from all data collected so far,
 so that the addition of a new data point only
 affects the average by a weight of $1/t$,
-which is the inverse of the sample size, $t$
+which is the inverse of the current sample size, $t$
 (cf. Appendix \ref{sec:equilerr}).
@@
 %
-Our aim here is to find the optimal schedule $\alpha(t)$,
+Our aim here is to find the optimal schedule, $\alpha(t)$,
 for a general updating scheme.
@@
 %
-This is done by removing two contributions from the bias potential.
+This is done by deducting two contributions from the bias potential.
 %
@@
 does not affect the
-resulting distribution $\pi_i(t)$,
+resulting distribution, $\pi_i(t)$,
 %
@@
 for some distribution,
-$\pmb \rho = (\rho_1, \dots, \rho_n)$.
+$\pmb \rho = (\rho_1, \dots, \rho_n)$
+(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
 %
@@
 introduced above is independent of
-the target distribution $\mathbf p$.
+the target distribution, $\mathbf p$.
 %
@@
 \note{More explicitly,
-if the transition matrix of a period $\tau$
+if the transition matrix of a period, $\tau$,
 is $T^\tau_{ij}$,
@@
   =
-  T^{t - t'}_{ij} \pi_j - \pi_i \, \pi_j.
+  T^{t - t'}_{ij} \, \pi_j - \pi_i \, \pi_j.
 $
@@
 for some $\Gamma_0$,
-where $\delta(t)$ is the Dirac's delta function.
+and $\delta(t)$ is the Dirac's delta function.
 %
@@
 the equivalent value of $\Gamma_0$ can be obtained
-from the integral of the autocorrelation function
+from the integral of the autocorrelation function,
 $\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$.
@@
          {     p_r    }
-  .
+  ,
 \label{eq:xi_def}
 \end{equation}
+%
+is also a white noise as $\zeta_i(t)$.
+%
 \note{The derivation:
@@
 now depends implicitly on the function $\alpha(t)$
-via $x_i(t)$.
+via $x_i(T)$.
 %
@@
 %
-We now fix the endpoint value $q(T)$
+We now fix the endpoint value, $q(T)$,
 and vary the curve, $q(t)$ ($0 < t < T$).
@@
 %
-Then by using Eq. \eqref{eq:u_def}, we have
+Then, by using Eq. \eqref{eq:u_def}, we have
 $$
@@
 and differentiating both sides
-with respect to $t'$
-yields Eq. \eqref{eq:alpha_invt1}.
+with respect to $t'$,
+we get Eq. \eqref{eq:alpha_invt1}.
 
@@
 Thus,
-the optimal schedule $\alpha(t)$ is given by
+the optimal schedule, $\alpha(t)$, is given by
 Eq. \eqref{eq:alpha_invt1}
-for it optimizes each term $\langle x_i^2(T) \rangle$,
+for it optimizes each term, $\langle x_i^2(T) \rangle$,
 and hence the sum;
@@
 $w_{ij}$'s,
-form a matrix $\mathbf w$.
+form a matrix, $\mathbf w$.
 %
@@
 
-The matrix $\mathbf w$ is not arbitrary.
+The matrix, $\mathbf w$, is not arbitrary.
 %
 First, to sample
-the desired distribution
+the desired distribution,
 $\mathbf p = (p_1, \dots, p_n)$,
- $\dot V_i(t)$
+${\dot V}_i(t)$
 in Eq. \eqref{eq:vt_diffeq_mbin}
@@
 %
-for the distribution $\pmb \rho$
+for the distribution, $\pmb \rho$
 introduced in Eq. \eqref{eq:vbar_def}.
@@
 no eigenmode grows indefinitely,
-which means that no eigenvalue $\lambda_k$
+which means that no eigenvalue, $\lambda_k$,
 can be negative.
@@
 
-\note{The proof of Eq. \eqref{eq:dvbardt},
+\note{The derivative of $\dot {\bar v} = \alpha(t)$,
 $$
@@
 
-Diagonalizing the matrix $\mathbf w$
+Diagonalizing the matrix, $\mathbf w$,
 using the eigenvectors defined in Eqs.
@@
 \end{align}
-are the deviation and noise of mode $k$, respectively.
+are the deviation and noise of mode, $k$, respectively.
 
@@
 can be rewritten in terms of the eigenmodes of $\mathbf w$
+%
 \begin{align}
@@
 \begin{equation}
-  \Err
+  \Err(T)
   =
@@
 In analogy to Eq. \eqref{eq:x2t_average},
-using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum}, we can write
-the error after a period $T$
+using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
+we can write the error after a period, $T$,
 as a sum of two contributions,
@@
 \begin{equation}
-  \Err
-  =
-  \Err_R + \Err_A,
+  \Err(T)
+  =
+  \Err_R(T) + \Err_A(T),
   \label{eq:error_split}
@@
 \begin{equation}
-  \Err_R
+  \Err_R(T)
   =
@@
 is given by
-\begin{equation}
-  \Err_A
+%
+\begin{equation}
+  \Err_A(T)
   =
@@
 with
+%
 \begin{align*}
@@
 \begin{align}
-  \Err_A
+  \Err_A(T)
   &=
@@
 %
-After a long period $T'$,
+After a long period, $T'$,
 the residual error becomes negligible, and
@@
 for $t < 0$ during which the adaptive FDS sampling was performed
-under a constant updating magnitude
+under a constant updating magnitude,
 $\alpha(t) = a_0$, and at the end of the equilibration, $t = 0$,
-the system has entered the equilibrium state of sampling.
-%
-We can then use Eq. \eqref{eq:y2_eql}
+the system has entered the equilibrium state of sampling,
+in which the error of mode $k$ obeys Eq. \eqref{eq:y2_eql}.
+%
+If we use this result
 for the $\langle y_k^2(0) \rangle$'s
 in Eq. \eqref{eq:error_res},
-and
 the residual error at the end of period $T$
@@
 depends on $\alpha(t)$ only through
-the integral $q(T) = \int_0^T \alpha(t) \, dt$.
+the integral, $q(T) = \int_0^T \alpha(t) \, dt$.
 %
@@
 and the optimization can be done exclusively
-on the asymptotic part, $\Err_A$.
+on the asymptotic part, $\Err_A(T)$.
 
@@
 We can rewrite Eq. \eqref{eq:error_asym1} much like an action
-of particle whose position is given by $q(t)$:
-%
-\begin{equation}
-  \Err_A
+of a particle whose position is given by $q(t)$:
+%
+\begin{equation}
+  \Err_A(T)
   =
@@
 %
-As the schedule, $\dot q(t) = \alpha(t)$,
-corresponds to the velocity of the particle,
-the above Lagrangian contains only the kinetic energy.
-%
-Thus, the above mechanical analogy
-represents a free particle
+In this mechanic analogy,
+the schedule, $\dot q(t) = \alpha(t)$,
+corresponds to the velocity of the particle.
+%
+So the above Lagrangian contains only the kinetic energy,
+and, therefore, describes a free particle
 carrying a position-dependent mass.
@@
   \left(
-    \frac C T
+    \frac{C}{T}
   \right)^2
@@
 %
-for some positive constant $C$.
+for some positive constant, $C$.
 %
@@
 %
-Since we only optimized the asymptotic error, $E_A$,
+Since we only optimized the asymptotic error, $\Err_A(T)$,
 in the above,
 the value of $q(T)$ can be further adjusted
-to minimize the total error, $E = E_R + E_A$
+to minimize the total error, $\Err(T) = \Err_R(T) + \Err_A(T)$
 (cf. Sec. \ref{sec:optinitalpha}).
@@
 First, we define a mass function as
+%
 \begin{equation}
@@
 \end{equation}
+%
 and a normalized mass distribution as
+%
 \begin{equation}
@@
 %
-which means that each point
-$\bigl(t, \alpha(t)\bigr)$
+which means that each point,
+$\bigl(t, \alpha(t)\bigr)$,
 on the schedule can be mapped
-to a point
-$\bigl(Q, m(Q)\bigr)$
+to a point,
+$\bigl(Q, m(Q)\bigr)$,
 on the mass distribution,
@@
 gives $1/(T\alpha)$,
-and the area under the curve in the domain $[Q, q(T)]$
+and the area under the curve in the domain, $[Q, q(T)]$,
 is equal to $t/T$,
@@
 We collect in each class optimal schedules
-computed from simulations with the same function $M(Q)$
-and cutoff value $q(T)$.
+computed from simulations with the same function, $M(Q)$,
+and cutoff value of $q(T)$.
 %
@@
 $\lambda_k$'s and $\Gamma_k$'s,
-whereas the latter affects the distribution $m(Q)$
-via the normalization constant $C$.
+whereas the latter affects the distribution, $m(Q)$,
+via the normalization constant, $C$.
 
@@
 of lengths $T_1$ and $T_2$
-at corresponding times $t_1/T_1 = t_2/T_2$
+at corresponding times, $t_1/T_1 = t_2/T_2$,
 are related by $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
@@
 \begin{equation}
-  E_A
+  \Err_A(T)
   =
@@
 i.e., we have the scaling relation
-$T_1 \, E_{A1} = T_2 \, E_{A2}$.
+$T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2)$.
 
@@
 the system is initially equilibrated at the same
-constant updating magnitude $a_0$.
+constant updating magnitude, $a_0$.
 %
@@
   \frac{
-    \partial E_R
+    \partial \Err_R(T)
   }
@@
   \frac{
-    \partial E_A
+    \partial \Err_A(T)
   }
@@
 $$
-Then solve $\partial (E_R + E_A) / \partial Q_c = 0$.
+Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
 }%
@@
 %
-This factor $1/2$
+This factor, $1/2$,
 happens to coincide with the
@@
 both the optimal schedule and
-normalized asymptotic error, $E_A \cdot T$,
+normalized asymptotic error, $\Err_A(T) \cdot T$,
 would be sensitive to the simulation length, $T$.
@@
 %
-These numbers constitute an \emph{updating kernel}.
+These numbers constitute an \emph{updating kernel}\cite{bussi2006}.
 %
 In this way, the matrix, $\mathbf w$,
-being symmetric with respect to the indices $i$ and $j$,
+being symmetric with respect to the indices, $i$ and $j$,
 satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
@@
 The second step of Eq. \eqref{eq:wmul_to_convol}
-follows from the constraint $\mu_l = 0$ for $|l| > b$.
+follows from the constraint, $\mu_l = 0$, for $|l| > b$.
 
@@
 %
-Again the updating kernel should satisfy
+Again, the updating kernel should satisfy
 Eqs. \eqref{eq:mu_normalization}
@@
 must share the same periodicity,
-and be even functions about the axis $i = 1/2$:
+and be even functions about the axis, $i = 1/2$:
 %
@@
 Eq. \eqref{eq:wband_eigenvalue_pbc},
-with the period $n$ doubled:
+with the period, $n$, doubled:
 %
@@
 %
-where the degeneracy $g$ is $2$
+where the degeneracy, $g$, is $2$
 for the periodic case,
@@
 If $\mu_2 = \dots = \mu_b = 0$,
-the matrix $\mathbf w$ is tridiagonal,
+the matrix, $\mathbf w$, is tridiagonal,
 %%
@@
 %
-By assuming the largest $b \approx n/g$,
+By assuming the largest cutoff, $b \approx n/g$,
 we can approximate Eq. \eqref{eq:wband_eigenvalue}
@@
 \end{align*}
-The upper limit
+The upper limit,
 $b \, \Delta x \approx (n/g) \, (g\,\pi/n) = \pi$.
@@
 For the Gaussian updating kernel,
-if the width $\sigma$ satisfies $\Delta x \ll \sigma \ll 1$,
+if the width, $\sigma$, satisfies $\Delta x \ll \sigma \ll 1$,
 we can extend
@@
 %
-If follows that all eigenvalues are positive:
+If follows that all eigenvalues are positive\cite{bussi2006}:
 %
@@
 \end{align}
-to satisfy the constraint $\mu_n = 0$.
+to satisfy the constraint, $\mu_n = 0$.
 %
@@
 This inequality follows from the non-negativity of
-the following quadratic polynomial of $\chi$,
+the following quadratic polynomial,
 $$
@@
 $$
-which implies a non-positive discriminant,
-$A_1^2 - 4 \, A_0 \, A_2$.
+which must carry a non-positive discriminant:
+$A_1^2 - 4 \, A_0 \, A_2 \ge 0$.
 %}%
@@
 Thus, the inequality sets a lower bound
-for the asymptotic error $\Err$
+for the asymptotic error $\Err(T)$
 given in Eq. \eqref{eq:error_asym1}.
@@
 The asymptotic error is then
+%
 \begin{align}
-  E_A
+  \Err_A(T)
   =
@@
 %
-The definition of the cutoff $K$ will be modified
+The definition of the cutoff, $K$, will be modified
 for mathematical convenience.
@@
 The kernel in the non-periodic case is a zigzag
-because of the term $(-1)^{n+K+l-1}$
+because of the term, $(-1)^{n+K+l-1}$,
 in Eq. \eqref{eq:mu_sinc_refl}.
@@
 %\begin{equation}
-%  E
+%  \Err
 %  =
-%  E_R + E_A
+%  \Err_R + \Err_A
 %  =
@@
 \begin{equation}
-  E
+  \Err(T)
   =
@@
 $p^*_i = p_i = 1/n$,
-so is the distribution $\rho_i$ in Eqs.
+so is the distribution, $\rho_i$, in Eqs.
 \eqref{eq:vbar_def} and \eqref{eq:error_sum}.
@@
     \label{fig:errinvt}
-    Error $E$ versus the parameter $\lambda$
+    Error, $\Err$, versus the parameter, $\lambda$,
     for the modified inverse-time schedule,
@@
 Then, the corresponding optimal schedule
-from Eq. \eqref{eq:q_opt} with the truncated error definition
+from Eq. \eqref{eq:q_opt} that minimizes the truncated error
 became similar to the inverse-time schedule,
@@
 %
-To verify this,
+To verify this result,
 we computed the optimal schedules
@@
 %
-This dependence becomes even weaker
-in the single-bin updating scheme,
-as it only amounts to a shift of the origin of
+The weak dependence is best illustrated in
+the case of single-bin updating scheme,
+as the initial updating magnitude only
+affects the schedule by a shift of the origin of
 simulation time in the inverse-time schedule,
@@
 In Fig. \ref{fig:errsigscan},
-we show the errors, $\Err$,
+we present the errors, $\Err(T)$,
 normalized by the shifted simulation length, $T+t_0$,
-so that the normalized errors
-would stay constant in the single-bin updating scheme
-(corresponding to the limit of $\sigma \to 0$).
+so that they could
+stay constant in the limit of single-bin updating scheme
+($\sigma \to 0$).
 %
@@
     \label{fig:errsigscan}
-    Normalized error, $(T + t_0) \, E$,
+    Normalized error, $(T + t_0) \, \Err(T)$,
     versus the width, $\sigma$,
@@
 First, we have ignored the \emph{systematic}
-error\cite{zhou2005, zhou2008}
+error\cite{zhou2005, morozov2007, zhou2008}
 caused by the updates in adaptive FDS methods.
 %
-This is valid in the asymptotic regime,
-where the updating magnitude is small
-and the sampling is a quasi-equilibrium one\cite{barducci2008, dama2014}.
-%
-Then the random error is much greater than the systematic one,
-because the former is proportional to the square root of the updating magnitude
-whereas the latter is proportional to the magnitude itself.
+This is valid in the asymptotic regime:
+when the updating magnitude is small
+and the sampling is a quasi-equilibrium one\cite{
+  zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
+the random error becomes more important than the systematic one,
+because the former is roughly proportional to
+the square root of the updating magnitude\cite{
+  zhou2005, morozov2007, zhou2008, bussi2006}
+whereas the latter is proportional to
+the magnitude itself\cite{morozov2007}.
 %
@@
 We can understand the optimality of the inverse-time schedule
-as a result of the stabilization of a runtime average
-with respect to the increase of sample size\cite{
+as a natural consequence of the stabilization of a runtime average
+as the sample size increases\cite{
   marsili2006, barducci2008}.
@@
 can be written as a recurrence relation regarding
-the addition of the last data point.
+the addition of the last data point,
 %
@@
 
-Let us now consider a long equilibrium FDS simulation
-that consists of blocks of sub-simulations
+The structural similarity between the recurrence relation,
+Eq. \eqref{eq:av_recur},
+and the WL updating scheme, Eq. \eqref{eq:wl_update},
+suggests that the bias potential may represent
+a runtime average in a particular case.
+%
+To establish a mapping between the two,
+let us now consider a long equilibrium FDS simulation
+that consists of a sequence of sub-simulations
 of $L$ steps.
@@
 from the $N$th sub-simulation,
-then ${\bar X}^{(N)}$ stands for the average bias potential,
+then ${\bar X}^{(N)}$ means the average bias potential,
 ${\bar V}_i^{(N)}$,
-from the first $N$ sub-simulations,
-which is the best estimate at the end of
-the first $N$ sub-simulations.
-%
-If we wish to always use this best estimate, ${\bar V}_i^{(N-1)}$,
-from the previous $N-1$ sub-simulations
-as the bias potential for the $N$th sub-simulations,
-then this estimate can be updated according to Eq. \eqref{eq:av_recur}.
+from the first $N$ sub-simulations.
+%
+Since this average, ${\bar V}_i^{(N)}$,
+gives the best estimate at the end of
+the first $N$ sub-simulations,
+we should ideally use it as the bias potential
+for the next sub-simulation,
+until it is further updated by Eq. \eqref{eq:av_recur}
+at the end of the next sub-simulation.
 %
 Assuming that the normalized histogram, $H_i^{(N)}$,
-from the $N$th sub-simulations is sufficiently precise,
-Eq. \eqref{eq:vcorr_equil} shows that the independent
+from the $N$th sub-simulation is sufficiently precise,
+Eq. \eqref{eq:vcorr_equil}
+\big[with
+$V_i^\mathrm{corr} \to V_i^{(N)}$, $V_i \to {\bar V}_i^{(N-1)}$\big]
+shows that the independent
 correction from the $N$th sub-simulation is
@@
 %
-where, the $-1$ on the right-hand side
+where the ``$-1$'' on the right-hand side
 can be ignored as a uniform shift
@@
 at step $l$ of sub-simulation $N$,
-and it is related to the average $H_i^{(N)}$ as
+and its average is
 $H_i^{(N)} \equiv \frac{1}{L} \sum_{l=1}^L h_i^{(N)}(l)$.
 %
-Thus, Eq. \eqref{eq:vbar_update} shows that the best estimate
-or the runtime average of the bias potential can be computed
+Now Eq. \eqref{eq:vbar_update} shows that the best estimate
+of the bias potential from the runtime average can be computed
 through a step-by-step incremental updating scheme that is equivalent to
 the WL algorithm, Eq. \eqref{eq:wl_update},
-with the magnitude being
+with the updating magnitude being
 %
@@
 by the product of the number of sub-simulations
-and the number of steps in each sub-simulation, or $NL$,
+and the number of steps in each sub-simulation, or $N \, L$,
 %
@@
 
-The above demonstration has several practical implications.
+The above demonstration implies several connections
+between the WL algorithm under the optimal inverse-time schedule
+with several related FDS methods.
 %
 First,
-since the bias potential is often sufficiently precise
-in late stages of a WL simulation,
-some practitioners would then fix
-the bias potential and switch
-to the equilibrium FDS method
-to prevent possible future efficiency loss
-due to the adaptive updates.
-%
-However, the above demonstration shows that
-this practice becomes unnecessary
-with the inverse-time schedule.
-%
-That is,
-the WL algorithm with the inverse-time schedule
-produces a bias potential that is similar in precision to that from
-the ideal equilibrium simulation for long times.
+it establishes an efficiency equivalence
+between an adaptive WL simulation under the inverse-time schedule
+and the ideal equilibrium FDS counterpart
+with a nearly perfect bias potential.
+%
+The bias potentials from the two approaches
+would have similar precision at long times.
+%
+Thus, it is unnecessary to convert
+a sufficiently convergent adaptive FDS simulation
+to the corresponding equilibirium FDS method
+with a fixed bias potential just for better precision
+of the final PMF.
 
 Second,
-an alternative way of improving the precision of the bias potential
-from a simulation using the WL algorithm or metadynamics
-with a fixed magnitude
-is to time average the bias potential\cite{
-  zhou2005, zhou2008}.
-%
-This practice becomes redundant with the inverse-time schedule
-because the averaging of the bias potential
-has been implicitly performed by the schedule.
+it also relates a WL simulation under the inverse-time schedule
+to a time-averaged variant of the regular WL simulation.
+%
+In the latter, the updating magnitude is fixed,
+but the bias potential over the simulation coarse is averaged
+as the final result\cite{zhou2005}.
+%
+The above demonstration shows that the inverse-time schedule
+has already effectively implemented an implicit average
+so that no explicit averaging is needed in the former approaches.
 %
@@
 it is possible to reformulate the WL algorithm
-with the inverse-time schedule
+under the inverse-time schedule
+as an explicit adaptive average method
 such that the bias potential is computed from
@@
 To compute the optimal schedule,
-we map the simulation time $t$ to the stage number $s$ as
+we map the simulation time, $t$, to the stage number, $s$, as
 %
@@
 for a large $T$
-(at this point $\gamma \to 1$).
-%
-If we identify the parameter $\alpha_0$ as the initial updating magnitude $\alpha(t = 0)$,
+(at this point, $\gamma \to 1$).
+%
+\note{
+  For a large $T$, $E_S \approx \frac{\Gamma}{2T} \frac{1+r}{\gamma}$.
+  Define $\delta = 1 - r$,
+  we have $1/\gamma = 1+ \delta/2 + \delta^2/3 + \cdots$,
+  and
+  $$
+  \frac{1+r}{\gamma}
+  =
+  (2 - \delta)
+  \left(
+  1 + \frac{\delta}{2} + \frac{\delta^2}{3} + \cdots
+  \right)
+  =
+  2 + \sum_{n=1}^\infty \frac{n-1}{n(n+1)}\delta^n \ge 2
+  .
+  $$
+}
+%
+If we identify the parameter, $\alpha_0$,
+as the initial updating magnitude, $\alpha(t = 0)$,
 then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
@@
 %To see this, we note that
-the bin index $i$ is an independent random variable
+the bin index, $i$, is an independent random variable
 at each time step $t$.
@@
 one of the two neighboring bins
-with equal probability $1/2$.
+with equal probability, $1/2$.
 %
@@
 In this case,
-the autocorrelation function $\kappa_k(t)$
+the autocorrelation function, $\kappa_k(t)$,
 decays as $\kappa_k(0) \, \Lambda_k^t$,
@@
 
-  \node[draw, text width={width("Error, $\Err$")}]
-    (err) at (7, 0) {Error, $\Err$};
+    \node[draw, text width={width("Error,")}]
+    (err) at (7, 0) {Error, $\Err(T)$};
 
@@
 \item
-Compute the error $E$ from
+  Compute the error $\Err(T)$ from
 Eqs. \eqref{eq:error_split},
@@
     Given an updating kernel, $\mu_0, \dots, \mu_b$,
-    compute the eigenvalues
-    $\lambda_1, \dots, \lambda_n$
+    compute the eigenvalues,
+    $\lambda_1, \dots, \lambda_n$,
     from Eq. \eqref{eq:wband_eigenvalue}.
@@
 it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
+%
 \begin{equation}
@@
 \end{equation}
+%
 and the asymptotic error from Eq. \eqref{eq:error_asym2}
@@
 greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
+%
 $$
@@
 dominates the mass function, $M(Q)$,
-defined in Eq. \eqref{eq:mass_func}.
+as defined in Eq. \eqref{eq:mass_func}.
 %
@@
 For a set of well-separated eigenvalues,
-$\lambda_k$ depends on the mode index $k$
+$\lambda_k$ depends on the mode index, $k$,
 much more critically than $\Gamma_k$.
@@
 $$
-  E_A
+  \Err_A(T)
   =
@@
 \begin{align}
-  \Err_A
+  \Err_A(T)
   =
@@
 \begin{aligned}
-  \Err_A
+  \Err_A(T)
   &=
@@
 \begin{equation}
-  \Err_A
+  \Err_A(T)
   =
@@
 $$
-  \Err_A
+  \Err_A(T)
   \propto

### 2017-02-01 doc/invt.tex
invt.tex absorb Justin's modifications

@@
 Free energy calculation\cite{frenkel, newman} is a central theme
-in computational physics and chemistry.
+in computational physics and chemistry
+that can provide insight into an array of phenomena not easily studied
+with traditional experiments.
 %
@@
 %
-The logarithm, $\ln p^*(z)$,
-defines a free energy, or a potential of mean force (PMF).
-%
-Difficulty arises, however, when $z$
-represents a collective variable
-that tends to move locally in a multi-modal distribution.
+The dimensionless free energy, or potential of mean force (PMF),
+along the collective variable, $z$,
+is $-\ln p^*(z)$.
+%
+However, accurately estimating the PMF is often difficult
+when the collective variable is energetically restricted to local regions
+of the complex, multi-modal free energy surface.
 %
 Thus,
+to overcome these energetic barriers and
 to capture the global shape of the free energy surface
 an effective strategy is to introduce a bias potential that
-cancels the PMF, resulting sampling of a flat distribution
+cancels the PMF.
+%
+Ideally, then, sampling yields a flat distribution
 along $z$\cite{mezei1987, berg1992, *lee1993,
@@
 *laio2002, *laio2008, *barducci2011, *sutto2012}.
-%The negative logarithm,
-%$-\ln p^*(z)$,
-%defines a free energy,
-%or the potential of mean force (PMF).
-%%
-%Normally, the distribution $p^*(z)$ is
-%limited in one or a few narrow $z$ regions.
-%%
-%Therefore, to capture the global shape
-%of the PMF,
-%one has to run multiple independent simulations
-%under different umbrella or bias potentials
-%along $z$\cite{mezei1987, berg1992, *lee1993},
-%and then piece together the entire curve.
-%%
-%This solution can be laborious % clumsy in practice
-%and it often fails to converge quickly
-%for a complex and glassy system,
-%because the system can be trapped
-%in a local free energy minimum.
-%%for a long time.
-%
-%
+%
+The flat-distribution sampling can be used to accelerate
+sampling of slow transitions that otherwise would not readily
+be observed with traditional simulations.
 %
@@
 %
-Both techniques employ regular updating of the bias potential
-to discourage future visits to the current configuration
-by incrementally elevating the bias potential there.
+Both techniques regularly update the bias potential
+to discourage future visits to previously sampled configurations
+by incrementally elevating the bias potential.
+%
+The bias potential is associated with discrete values of $z$.
 %
@@
 $\delta$-function (a boxcar function one histogram bin wide)
-in the WL case,
+in WL,
 or that of a Gaussian
-in the metadynamics.
+in metadynamics.
 %
@@
 sampling\cite{zhou2005, morozov2007, zhou2008},
-and for convergence
+and for convergence of the estimated PMF
 one  has to gradually decrease
-the updating magnitude.
+the updating magnitude of the bias.
 
@@
 the precision of the final bias potential,
-hence that of the PMF\cite{
+and hence that of the PMF\cite{
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
@@
 %
-For the WL algorithm, it is known
-that the optimal schedule
+For the WL algorithm, the optimal updating schedule
 is a function of the
@@
 %
-The optimal schedule for the metadynamics case,
+The optimal schedule for metadynamics,
 however, is less studied.
@@
 %
-Our method maps the optimization problem to a mechanic one,
+Our method maps the optimization problem to a mechanical one,
 in which the schedule plays the role of the velocity of
@@
 %
-In both cases,
-the distribution is normalized as
+The distribution is normalized as
 $\sum_{i = 1}^n p^*_i = 1$.
@@
 %
-Upon the normalization,
+Upon normalization,
 $\sum_{i = 1}^n \pi_i = 1$,
@@
 We wish the biased equilibrium distribution, $\pi_i$,
-to be the same as the intended distribution, $p_i$,
-which is usually flat,
-i.e., $p_i$ is constant for all $i$'s,
+to converge to the target distribution, $p_i$,
+which is usually chosen to be flat,
+i.e. $p_i$ is constant for all $i$'s,
 or nearly flat\cite{
@@
 uses a fixed bias potential, $V_i$,
-(which is estimated beforehand)
+(estimated a priori)
 during simulation.
@@
 %Another difficulty of this approach is that
-Besides, the effectiveness of this approach
+The effectiveness of this approach
 depends on the precision of the histogram, $H_i$,
@@
 Thus, the rigidity of this framework disallows
-continuous improvement of the bias potential, hence that of the PMF,
+continuous, rapid improvement of the bias potential,
+and subsequently the PMF,
 as simulation lengthens.
@@
 address the above shortcomings
-by continuously and incrementally updating the target distribution.
+by continuously and incrementally updating the bias potential.
 %For this reason, the second type of methods,
@@
 However,
-%they are less well understood theoretically, as
-as mentioned, the on-the-fly updates break the microscopic reversibility
+the on-the-fly updates break the microscopic reversibility
 of the underlying sampling mechanism.
@@
 the histogram-based increments
-would lead to roughness or errors in the final estimate.
+would lead to roughness or errors in the final estimate of the PMF.
 %
@@
 in improving the precision of the bias potential,
-delivering the maximal efficiency
+thus allowing for maximum efficiency
 (cf. Appendix \ref{sec:equilerr}).
@@
 where $i(t)$ is the bin at step $t$,
+$\delta_{i, \, i(t)}$ is the Kronecker delta function,
+which is $1$ when $i = i(t)$ or $0$ otherwise,
 and $\alpha(t)$ is the updating magnitude.
@@
 an adaptive FDS simulation
-is identical to $p_i$ in long times.
+is identical to $p_i$ in the long time or asymptotic limit.
 %
@@
 change $V_i(t)$ continuously,
-Eq. \eqref{eq:pi_p_V} no longer holds
+the equilibrium distribution,
+Eq. \eqref{eq:pi_p_V}, no longer holds
 at all times,
@@
 the updating magnitude, $\alpha$,
-(or $\ln f$ there)
+(or $\ln f$ therein)
 is controlled in stages.
@@
 Note that the distribution, $\pmb \rho$,
-introduced above is independent of
-the target distribution, $\mathbf p$.
+introduced above to define the error
+is independent of
+the target distribution,
+$\mathbf p = (p_1, \dots, p_n)$.
 %
@@
 The deterministic part can be related
-to the ``ensemble average'',
+to an ``ensemble average'',
 where the ensemble consists of many similar simulation copies
@@
 \begin{equation}
-  \alpha(t) = \frac{ 1 }{ t + t_0 }.
+  \alpha(t) = \frac{ 1 }{ t + t_0 },
 \label{eq:alpha_invt1}
 \end{equation}
+%
+where $t_0$ is a constant.
 
@@
 in Eq. \eqref{eq:vt_diffeq_mbin}
-should be a constant of $i$,
+should be a constant of $i$
 if $h_j(t)$ were the same as $p_j$.
@@
 %
-for the distribution, $\pmb \rho$
+for the distribution, $\pmb \rho$,
 introduced in Eq. \eqref{eq:vbar_def}.
@@
 %
-i.e., $\mathbf \rho$ is a left eigenvector of
+i.e. $\mathbf \rho$ is a left eigenvector of
 $\mathbf w$ with eigenvalue $1$.
@@
 %
-By using Eqs. \eqref{eq:vbar_def}
-and \eqref{eq:w_balance},
-we can show that Eq. \eqref{eq:dvbardt}
-remains true,
-and by using Eqs.
+By using Eqs.
 \eqref{eq:x_def},
+\eqref{eq:vbar_def},
 \eqref{eq:rhox_sum},
@@
 \eqref{eq:xi_def},
-and \eqref{eq:w_sumj}, we get
+\eqref{eq:w_sumj},
+and \eqref{eq:w_balance} we get
 %
@@
 In the asymptotic regime
-(i.e., if the simulation period, $T$, is sufficiently long),
+(i.e. if the simulation period, $T$, is sufficiently long),
 we can approximate $\kappa_k(t)$ as
@@
 %
-In this mechanic analogy,
+In this mechanical analogy,
 the schedule, $\dot q(t) = \alpha(t)$,
@@
 for schedules in the same class,
-i.e., we have the scaling relation
+i.e. we have the scaling relation
 $T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2)$.
@@
 %
-i.e., the optimal initial updating magnitude
+i.e. the optimal initial updating magnitude
 is always half of the equilibrium value.
@@
 happens to coincide with the
-recommended reduction factor of updating magnitude
+recommended reduction factor of the updating magnitude
 at stage transitions
@@
 for the flat distribution, $\rho_i = 1/n$,
-i.e., the updating matrix, $\mathbf w$,
+i.e. the updating matrix, $\mathbf w$,
 is symmetric.
 %
-Besides, the updating matrices
+The updating matrices
 share the same set of eigenvectors
-(sines and cosines of different frequencies),
+(which here are sines and cosines of different frequencies),
 and thus are completely determined
@@
 which must carry a non-positive discriminant:
-$A_1^2 - 4 \, A_0 \, A_2 \ge 0$.
+$A_1^2 - 4 \, A_0 \, A_2 \le 0$.
 %}%
@@
 that ignore short-wavelength eigenmodes,
-which are typically local noises.
+which are typically local noise.
 %
@@
 The optimal schedule of these schemes
-remains to be the inverse-time one
+remains to be the inverse-time schedule
 given by Eq. \eqref{eq:alpha_invt1}.
@@
 %
-The kernel in the non-periodic case is a zigzag
+The kernel in the non-periodic case is a sawtooth 
 because of the term, $(-1)^{n+K+l-1}$,
@@
 
-In the above, we have focused on
+In the above section,
+we have focused on
 the asymptotic error.
@@
 By including the residual error
-and minimizing the total,
+and minimizing the total error,
 we get the optimal $t_0$,
@@
 %
-and at this point, we get,
-from Eqs. \eqref{eq:error_split},
+and from Eqs. \eqref{eq:error_split},
 \eqref{eq:error_res1},
@@
 %
-First, we verify the method of computing errors
+We verify the method of computing errors
 %using an inverse-time schedule
-in Sec. \ref{sec:results_invt}.
-%
-We then study the optimal schedule from
+in Sec. \ref{sec:results_invt}
+%
+and study the optimal schedule from
 Eq. \eqref{eq:q_opt}
@@
 $p^*_i = p_i = 1/n$,
-so is the distribution, $\rho_i$, in Eqs.
+along with the distribution, $\rho_i$, in Eqs.
 \eqref{eq:vbar_def} and \eqref{eq:error_sum}.
@@
 Although the bias potential is correct initially,
-it would deviate from the correct value
+it would deviate from this value
 at the end of the equilibration.
@@
 
-We tested our analysis of computing errors
+Using the model system above, we
+tested our analysis of computing errors
 on the modified inverse-time schedule,
@@
 is optimal for the single-bin updating scheme,
-but not necessarily so for a general updating scheme.
+but not necessarily so for a general multiple-bin updating scheme.
 %
@@
 damp out the short-wavelength eigenmodes
-(i.e., local noises).
+(i.e. local noise).
 %
@@
 As shown in Fig. \ref{fig:erriascan},
-The minimal error was indeed reached at
+the minimal error was indeed reached at
 $\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
@@
 %
-Besides, the bandpass updating schemes
+The bandpass updating schemes
 yielded smaller errors than
@@
 for a general class of free energy calculation methods,
-which we refer to as adaptive flat-distribution sampling (FDS) methods,
-including the Wang-Landau (WL) algorithm and metadynamics.
+which we refer to as adaptive flat-distribution-sampling (FDS) methods
+that include the Wang-Landau (WL) algorithm and metadynamics.
 %
@@
 %the update is limited to the current bin;
-%while in the metadynamics case,
+%while metadynamics,
 %the bias potential of a neighborhood of the current one
@@
 However, if some eigenmodes
-correspond to local noises
-that are expected to be absent from the PMF,
+correspond to local noise
+that is expected to be absent from the PMF,
 we can construct a
@@
 %
-Besides, the performance gain
-is relatively small for a shorter simulation
+However, the performance gain
+is relatively small for short simulations
 with a local sampling process such as MD.
@@
 can be ignored as a uniform shift
-does not affect the distribution.
+that does not affect the distribution.
 %
@@
 between the WL algorithm under the optimal inverse-time schedule
-with several related FDS methods.
+with a few related FDS methods.
 %
@@
 We assume that
-the magnitude falls geometrically,
+the magnitude follows geometrically,
 %
@@
 %
-Our aim is to minimize error
+Our aim is to minimize the error
 of the final stage, $E_S$.
 %
-To do so, let us consider moving $\delta T$ simulation time from
+To do so, let us consider shifting simulation time
+by $\delta T$ from
 stage $s+1$ to stage $s$,
@@
 is not necessarily stable,
-i.e., some eigenvalues can be negative.
+i.e. some eigenvalues can be negative.
 %
@@
   %
+%}%
 Even
-  the Gaussian updating scheme can be unstable
-  in practice
-  if the updating kernel is truncated.
-%}%
+the Gaussian updating scheme, which is stable in the continuous limit,
+can be unstable upon discretization
+during which the updating kernel is truncated.
 %
@@
 The optimal $\lambda$ is, therefore,
-%
-$
-%\begin{equation}
-  \lambda \approx 1.
-%\label{eq:lambda_nn_onestep}
-%\end{equation}
-$
+$\lambda \approx 1$.
 %

### 2017-02-01 doc/invt.tex
invt.tex: small edits

@@
   \left[
-    1 - x_i(t) + \sum_{r=1}^n p_r \, x_r(t)
+    1 - x_i(t) + \sum_{j = 1}^n p_j \, x_j(t)
   \right]
@@
 \frac{                       1 - x_i  }
-     { \sum_{ r = 1 }^n p_r (1 - x_r) }
+     { \sum_{ r = 1 }^n p_j (1 - x_j) }
 =
 \frac{                       1 - x_i  }
-     { 1 - \sum_{ r = 1 }^n p_r \, x_r }
+     { 1 - \sum_{ j = 1 }^n p_j \, x_j }
 .
@@
   -
-  \sum_{ r = 1 }^n
-    \rho_r
-    \frac{ \zeta_r(t) }
-         {     p_r    }
+  \sum_{ j = 1 }^n
+    \rho_j
+    \frac{ \zeta_j(t) }
+         {     p_j    }
   ,
@@
 %
-is also a white noise as $\zeta_i(t)$.
+is also a white noise as $\zeta_i(t)$,
+i.e. we may write
+$$
+  \langle \xi_i(t) \, \xi_i(t') \rangle = \Gamma_i \, \delta(t - t'),
+$$
+for some $\Gamma_i$.
 %
@@
     -
-    \sum_{ r = 1 }^n
-      \rho_r \frac{ h_r } { p_r }
+    \sum_{ j = 1 }^n
+      \rho_j \frac{ h_j } { p_j }
   \right)
   \\
-  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_split} and \; \eqref{eq:h_ave}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_split} \; and \; \eqref{eq:xi_def}}}
+            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
   \alpha(t) \left(
@@
     -
-    \sum_{ r = 1 }^n
-      \rho_r \frac{ \langle h_r \rangle } { p_r }
+    \sum_{ j = 1 }^n
+      \rho_j \frac{ \langle h_j \rangle } { p_j }
     +
@@
   \\
-  &\stackrel{\mathrm{Eqs.\; \eqref{eq:rhox_sum} \;and\; \eqref{eq:h_ave}}}
+  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_ave} \;and\; \eqref{eq:rhox_sum}}}
             {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
@@
   =
-\langle x^2(0) \,
-\rangle e^{-2 \, q(T)}
+  \langle x^2(0) \rangle \, e^{-2 \, q(T)}
   +
@@
 is the special case of $t_0 = 0$.
-%
-In Sec. \ref{sec:finlencorr},
-we show that % $t_0$ can be adjusted
-one can further improve the schedule by adjusting $t_0$
-to accommodate the initial error.
+
+%In Sec. \ref{sec:finlencorr},
+%we show that % $t_0$ can be adjusted
+%one can further improve the schedule by adjusting $t_0$
+%to accommodate the initial error.
+To find the optimal value of $t_0$,
+we use the optimal schedule in Eq. \eqref{eq:x2t_average},
+and
+$$
+\left\langle x^2(T) \right\rangle
+=
+\frac{ \left\langle x^2(0) \right\rangle \, t_0^2 }
+     { (t_0 + T)^2 }
++
+\frac{ \Gamma \, T}
+     { (t_0 + T)^2 }
+.
+$$
+Using this result for each $x_i$,
+we get the final total error, $E(T)$,
+as defined in Eq. \eqref{eq:error_sum},
+\begin{equation}
+  E(T)
+  =
+  \frac{ \sum_{i=1}^n \left\langle x_i^2(0) \right\rangle \, t_0^2 }
+       { (t_0 + T)^2 }
+  +
+  \frac{ \sum_{i=1}^n \Gamma_i \, T}
+       { (t_0 + T)^2 }
+  .
+  \label{eq:error_singlebin}
+\end{equation}
+%
+This error is minimized if
+%
+\begin{equation}
+t_0 = \frac{ \sum_{i = 1}^n \Gamma_i }
+           { \sum_{i = 1}^n \left\langle x_i^2(0) \right\rangle }
+           ,
+  \label{eq:t0_singlebin}
+\end{equation}
+%
+and at this point
+%
+\begin{equation}
+  E(T)
+  =
+  \frac{ 1 } { T + t_0 }
+  \sum_{i = 1}^n \Gamma_i
+  .
+  \label{eq:Emin_singlebin}
+\end{equation}
 
@@
 By using Eqs.
-\eqref{eq:x_def},
-\eqref{eq:vbar_def},
-\eqref{eq:rhox_sum},
-\eqref{eq:h_ave},
-\eqref{eq:xi_def},
-\eqref{eq:w_sumj},
-and \eqref{eq:w_balance} we get
+\eqref{eq:v_def}, \eqref{eq:vbar_def},
+\eqref{eq:vt_diffeq_mbin}-\eqref{eq:w_balance}
+we get
+%
+\begin{align*}
+  \dot{ \bar v }(t)
+  =
+  \alpha(t)
+  \sum_{r = 1}^n \rho_r
+    \, \frac{ h_r(t) } { p_r }
+  =
+  \alpha(t)
+  \sum_{j = 1}^n w_{ij}
+  \sum_{r = 1}^n \rho_r
+    \, \frac{ h_r(t) } { p_r }
+  ,
+\end{align*}
+%
+\note{Derivation.
+$$\begin{aligned}
+  \dot{ \bar v }
+  &
+  \stackrel{\mathrm{Eq.\; \eqref{eq:vbar_def}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \sum_{i = 1}^n \rho_i \, \dot v_i
+  %
+  \stackrel{\mathrm{Eq.\; \eqref{eq:vt_diffeq_mbin}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  %
+  \alpha(t)
+  \sum_{j = 1}^n \frac{ h_j } { p_j }
+                 \sum_{ i = 1 }^{n} \rho_i \, w_{ij}
+  \\
+  &
+  \stackrel{\mathrm{Eq.\; \eqref{eq:w_balance}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \alpha(t)
+  \sum_{j = 1}^n \rho_j
+    \, \frac{ h_j } { p_j }
+  %
+  \stackrel{\mathrm{Eq.\; \eqref{eq:w_sumj}}}
+  {=\joinrel=\joinrel=\joinrel=\joinrel=}
+  \alpha(t)
+  \sum_{j = 1}^n w_{ij}
+  \sum_{r = 1}^n \rho_r
+    \, \frac{ h_r } { p_r }
+  .
+\end{aligned}$$
+}
+%
+and by Eqs. \eqref{eq:x_def}, \eqref{eq:rhox_sum} and \eqref{eq:h_ave},
 %
@@
   -\alpha(t) \sum_{j = 1}^n
-  w_{ij} \left[ x_j(t) - \xi_j (t) \right]
+  w_{ij} \bigl[ x_j(t) - \xi_j (t) \bigr]
   .
@@
 %on the second line.
-
-\note{The derivative of $\dot {\bar v} = \alpha(t)$,
-$$
-\begin{aligned}
-  \dot{ \bar v }
-  &
-  \stackrel{\mathrm{Eq.\; \eqref{eq:vbar_def}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \sum_{i = 1}^n \rho_i \, \dot v_i
-  %
-  \stackrel{\mathrm{Eq.\; \eqref{eq:vt_diffeq_mbin}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  %
-  \alpha(t)
-  \sum_{j = 1}^n \frac{ h_j } { p_j }
-                 \sum_{ i = 1 }^{n} \rho_i \, w_{ij}
-  \\
-  &
-  \stackrel{\mathrm{Eq.\; \eqref{eq:w_balance}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t)
-  \sum_{j = 1}^n \rho_j
-    \, \frac{ h_j } { p_j }
-  %
-  \stackrel{\mathrm{Eq.\; \eqref{eq:w_sumj}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t)
-  \sum_{j = 1}^n w_{ij}
-  \sum_{r = 1}^n \rho_r
-    \, \frac{ h_r } { p_r }
-  .
-\end{aligned}
-$$
-}
 
@@
 at corresponding times, $t_1/T_1 = t_2/T_2$,
-are related by $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
+are related by $T_1 \, \alpha_1(t_1) = T_2 \, \alpha_2(t_2)$.
 %
@@
 %
-which means the asymptotic error
+which means that the asymptotic error
 is inversely proportional to the simulation length
-for schedules in the same class,
-i.e. we have the scaling relation
-$T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2)$.
+for schedules in the same class.
+%which can be rephrased as another scaling relation
+%$$
+%T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2).
+%$$
 
@@
 
-Let us now consider
-simulations with the same $a_0$.
+For simulations with the same $a_0$,
+the shape of the mass distribution, $m(Q)$,
+also determines how much the optimal schedule
+would depend on the simulation length, $T$.
 %
@@
 the optimal $q(T)$ increases with
-the simulation length $T$,
+the simulation length, $T$,
 more so if $M(Q)$ has a long tail.
 %
-This means that, if many $\lambda_k$'s are near zero,
+Thus, if $M(Q)$ decays slowly because of
+many near zero eigenvalues, $\lambda_k$'s (as in the metadynamics case),
 both the optimal schedule and
@@
 would be sensitive to the simulation length, $T$.
+%
+In contrast, if all eigenvalues are $1.0$
+as in the single-bin or WL updating scheme discussed previously,
+the optimal schedule is
+given by the universal inverse-time formula,
+Eq. \eqref{eq:alpha_invt1},
+and the normalized asymptotic error
+is roughly a constant
+[cf. Eq. \eqref{eq:error_singlebin}].
 %
@@
        { a_0 }
-  ,
+  .
 \label{eq:t0_sinc}
@@
 %
-and from Eqs. \eqref{eq:error_split},
+For the single-bin updating scheme,
+it recovers Eq. \eqref{eq:t0_singlebin}
+under the assumption of Eq. \eqref{eq:y2_eql}.
+%
+From Eqs. \eqref{eq:error_split},
 \eqref{eq:error_res1},
@@
 \eqref{eq:error_asym_sinc},
+we get
 %
@@
 %
-This error decreases inversely with increasing simulation length,
+which serves as a generalization of
+the result from the single-bin scheme, Eq. \eqref{eq:Emin_singlebin}.
+%
+This error decreases with increasing simulation length,
 as one would expect from an equilibrium FDS simulation.
@@
 
-\begin{table}[h]
+\begin{table}[h]\footnotesize
   \caption{\label{tab:error_Gaussian}
-    Error of the Gaussian updating scheme of $\sigma = 10$.
-    $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
-    \note{
+    Error of the Gaussian updating scheme of $\sigma = 10$
+    under different updating schedules.
+    In Eqs. \eqref{eq:alpha_invt1} and \eqref{eq:alpha_invtlambda},
+    $t_0$ is given by Eq. \eqref{eq:t0_sinc};
+    and in Eq. \eqref{eq:alpha_invtlambda},
+    the optimal value of $\lambda$ is used.
+    \note{\newline
       The data were collected in
@@
     &
-      Modified inverse-time,
+      Modified, \newline
       Eq.~\eqref{eq:alpha_invtlambda}
-      with optimized $\lambda$
     &
@@
 for helpful discussions.
-We gratefully acknowledge the Robert A.
-Welch Foundation (H-0037), the National Science Foundation (CHE-1152876) and the National Institutes of Health (GM-037657) for partial support of this work.
+%
+We gratefully acknowledge the Robert A. Welch Foundation (H-0037),
+the National Science Foundation (CHE-1152876)
+and the National Institutes of Health (GM-037657)
+for partial support of this work.
+%
+JM thanks support from the National Institutes of Health (R01-GM067801, R01-GM116280),
+the Welch Foundation (Q-1512),
+and National Science Foundation (Grant PHY-1427654).
+%
 This research used computing resources of the National Science Foundation XSEDE grid.
+%
+%
 

### 2017-02-01 doc/old/V4.1_jd/invtV4.1_jd.tex
invt.tex: add the Justin edit

--- 2017-02-01 doc/invt.tex
@@
 \definecolor{LightGray}{gray}{0.85}
-% annotation macros
 \newcommand{\repl}[2]{{\color{gray} [#1] }{\color{blue} #2}}
@@
   the Wang-Landau (WL) algorithm and metadynamics.
-  %
   These methods rely on adaptive construction of
@@
   the potential of mean force by histogram-based adaptive updates.
-  %
   The updating magnitude should decrease over time
@@
   of the updating magnitude.
-  %
-  %Here, we will show that the optimal schedule
-  %can be derived
-  %from a mechanical analogy, in which
-  %the schedule corresponds to
-  %the velocity of a free particle with a position-dependent mass
-  %and the final error serves as the action.
-  %%
-  %Therefore, the optimal schedule follows from
-  %the equation of motion of the particle.
-  %
   We will show that
   the optimal schedule depends on the updating scheme (WL or metadynamics),
-  and the initial updating magnitude is ideally half of
+  and the initial updating magnitude should be half of
   the previous equilibrium value.
-  %
   Further,
@@
   as given by the $1/t$ prescription.
-  %
 \end{abstract}
@@
 Free energy calculation\cite{frenkel, newman} is a central theme
-in computational physics and chemistry
-that can provide insight into an array of phenomena not easily studied
-with traditional experiments.
-%
+in computational physics and chemistry that can provide insight into
+an array of phenomena not easily studied with traditional experiments.
 Given a system,
@@
 a distribution, $p^*(z)$,
-along a collective variable $z$ by sampling via either Monte Carlo\cite{
-  frenkel, newman, landau_binder} (MC)
+along a collective variable $z$ by sampling with Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
 or molecular dynamics\cite{frenkel, karplus2002} (MD) simulations.
-%
-The dimensionless free energy, or potential of mean force (PMF),
-along the collective variable, $z$,
-is $-\ln p^*(z)$.
-%
-However, accurately estimating the PMF is often difficult
-when the collective variable is energetically restricted to local regions
-of the complex, multi-modal free energy surface.
-%
+The free energy, or potential of mean force (PMF), along the collective variable is proportional to $\ln p^*(z)$.
+However, accurately estimating the PMF is often difficult when the collective variable is energetically restricted to local regions of the complex, multi-modal free energy surface.
 Thus,
-to overcome these energetic barriers and
-to capture the global shape of the free energy surface
+to overcome these energetic barriers and to capture the global shape of the free energy surface
 an effective strategy is to introduce a bias potential that
-cancels the PMF.
-%
-Ideally, then, sampling yields a flat distribution
-along $z$\cite{mezei1987, berg1992, *lee1993,
-wang2001, *wang2001pre,
-huber1994,
-*laio2002, *laio2008, *barducci2011, *sutto2012}.
-%
-The flat-distribution sampling can be used to accelerate
-sampling of slow transitions that otherwise would not readily
-be observed with traditional simulations.
-%
-%A better solution is to artificially flatten
-%the target distribution along $z$\cite{mezei1987, berg1992, *lee1993,
-%wang2001, *wang2001pre,
-%huber1994,
-%*laio2002, *laio2008, *barducci2011, *sutto2012}.
-%%
-%This alleviates the above problem of local trapping
-%by forcing the system to travel more
-%deliberately along $z$,
-%which often represents
-%a direction of slow motion.
-%%
-%To achieve a flat distribution, however,
-%the method must be able to construct a
-%bias potential that cancels the PMF.
+cancels the PMF. Ideally, then, sampling yields a flat distribution
+along $z$\cite{mezei1987, berg1992, lee1993,
+wang2001, *wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
 
 The Wang-Landau (WL) algorithm\cite{wang2001, *wang2001pre}
-and metadynamics\cite{huber1994,
-*laio2002, *laio2008, *barducci2011, *sutto2012}
+and metadynamics\cite{huber1994, laio2002,
+laio2008, barducci2011, sutto2012}
 are two widely-used and closely-related\cite{micheletti2004}
 flat-distribution-sampling (FDS) techniques
-that build the bias potential adaptively or on the fly.
-%
+that build the bias potential associated with the collective variable adaptively, or on-the-fly, during simulation.
 Both techniques regularly update the bias potential
@@
 by incrementally elevating the bias potential.
-%
-The bias potential is associated with discrete values of $z$.
-%
 The difference lies
@@
 which takes the form of a discrete
-$\delta$-function (a boxcar function one histogram bin wide)
-in WL,
+$\delta$-function (or a boxcar function one histogram bin wide)
+in WL ,
 or that of a Gaussian
 in metadynamics.
-%
-%is that in the WL case,
-%each updating step is limited to the bin
-%containing the current $z$,
-%while metadynamics adopts an extended
-%Gaussian window
-%covering several neighboring bins.
-%%
-%The latter is more often used in MD simulations
-%as it guarantees a smooth profile
-%of the bias potential
-%suitable for numerical differentiation
-%in deriving the bias force.
-%
 In both cases,
 regular updates to the bias potential
-%in both cases, the bias potential
-%is updated frequently (usually every few MC or MD steps),
-%and the fluctuation
 disrupts the underlying equilibrium
-sampling\cite{zhou2005, morozov2007, zhou2008},
-and for convergence of the estimated PMF
+sampling\cite{laio2002},
+and for convergence of the PMF
 one  has to gradually decrease
@@
 therefore, critically affects
-the precision of the final bias potential,
-and hence that of the PMF\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-liang2007,
-morozov2007, zhou2008,
-komura2012, *caparica2012, *caparica2014,
+the precision of the final bias potential, and
+hence that of the PMF\cite{liang2007,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
+morozov2007, zhou2008, morozov2009,
+komura2012, caparica2012, *caparica2014,
 barducci2008, dickson2011, dama2014}.
-%
 For the WL algorithm, the optimal updating schedule
 is a function of the
-inverse time\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-liang2007,
+inverse time\cite{liang2007,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 morozov2007, zhou2008}
@@
 the inverse number of simulation steps.
-\note{
-  About the origin of the inverse-time formula.
-  Besides the Belardinelli paper,
-  we can also find it in Eq. (6) of reference \cite{liang2007},
-  if I am not too mistaken.
-  But I find the paper is a bit hard to read.
-  There, however, the formula is $t_0/t$ instead of $1/t$.
-}
-%
 The optimal schedule for metadynamics,
@@
 In this study,
-we present a method of computing
+we present a general formalism for computing
 the optimal schedule
-for a general adaptive FDS method,
+for a any adaptive FDS method,
 including the WL algorithm and metadynamics.
-%
 Our method maps the optimization problem to a mechanical one,
@@
 and the final error becomes the action.
-%
 In this way, the optimal schedule
@@
 of the particle that minimizes the action.
-%
-As the mass is defined from the updating window function,
-the resulting optimal schedule
-is different in the WL and metadynamics cases.
-%
+We show that the optimal schedule
+depends on how the bias potential is updated,
+and hence is different in the WL and metadynamics cases.
 While the optimal schedule in the WL case
@@
 and the width of the Gaussian window.
-%
 In either case, however,
 the initial updating magnitude
-can be optimally set to
+should be set to
 half of the previous equilibrium value,
 which amounts to a shift of the origin of time in the WL case.
-%
 Further, we can generalize
@@
 and the above origin-shifted inverse-time schedule
-turns out to be optimal for these schemes.
-%
-%For finite-length simulations, however,
-%metadynamics may deliver better performance.
-%
+turns out to be optimal for these schemes as well.
 These results may help to better
-plan and optimize FDS simulations.
+plan and optimize FDS simulations, which, at present, is done heuristically and through trial-and-error.
 
 The article is organized as follows.
-%
 We present the analytical results in Sec. \ref{sec:theory},
@@
 in the following order.
-%
 In Sec. \ref{sec:background},
@@
 and some known aspects of the optimal schedule.
-%
-%This also helps fixing the notations.
-%
 Then, in Sec. \ref{sec:single-bin},
@@
 of the known inverse-time formula.
-%
 Next, we extend the method to the general case
@@
 to \ref{sec:band-matrix}.
-%
 Finally, we compare different updating schemes
@@
 for a given system.
-%
-%The discreteness of $i$ is natural
-%in discrete models.
-%
 For example, $i$ can be the energy, $E$,
@@
 in a simulated tempering simulation\cite{
-marinari1992, *lyubartsev1992}.
-%
+marinari1992, lyubartsev1992}.
 For a continuous quantity, $z$,
@@
 $(z, z + \Delta z)$.
-%
 The distribution is normalized as
@@
 For a large system at temperatures of interest,
-the distribution, $p^*_i$, often tends to
+the distribution $p^*_i$ often tends to
 be localized,
-%
 and to explore the global shape
@@
 it is often advantageous to perform
-biased sampling that targets
-a wider and smoother distribution, $p_i$.
-%
-%Here, we refer to simulations that target
-%a flat or nearly flat distribution
-%as entropic or multicanonical sampling.
+biased sampling that generates a
+a wider, smoother target distribution, $p_i$.
 
@@
 which changes the equilibrium distribution to
-%
 \begin{equation}
@@
 \end{equation}
-%
 Upon normalization,
@@
 we get
-%
 \begin{equation}
@@
 \end{equation}
-%
-%In particular,
-%to make $\pi_i$ flat,
-%the bias potential $V_i$
-%must coincide with $\ln p^*_i$,
-%up to an additive constant.
-%%
-%In practice, $V_i$ may be non-optimal or contain error,
-%and thus $\pi_i$ need not
-%coincide with the intended distribution $p_i$.
 
@@
 to converge to the target distribution, $p_i$,
-which is usually chosen to be flat,
-i.e. $p_i$ is constant for all $i$'s,
-or nearly flat\cite{
-dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}.
-%
-We will broadly call such a biased simulation
+which is usually chosen to be flat
+(i.e. $p_i$ constant for all $i$'s).\cite{
+dayal2004, trebst2004, zhang2007, barducci2008, singh2011}.
+We will generally call such a biased simulation
 a flat-distribution-sampling (FDS) simulation.
-%
-To achieve this distribution, the bias potential must satisfy
-%
+To fulfill this goal, the bias potential must satisfy
 \begin{equation}
@@
 \end{equation}
-%
 according to Eq. \eqref{eq:pi_p_V1}.
-%which allows us to deduce the PMF, $-\ln p^*_i$,
-%directly from the bias potential.
-%
-This relation thus sets a practical aim for the convergence of FDS
-methods---the bias potential must be adjusted to make the
+This relation thus sets a practical goal for the convergence of FDS
+methods---adjusting the bias potential to make the
 observed distribution, or the normalized histogram,
 match the intended one.
-%
 Once this aim is achieved,
@@
 deduced from the bias potential. % $V_i$.
-\note{
-  For example, in the context of simulated tempering (Zhang and Ma, 2007),
-  $p^*_i$ is replaced by $\ln Z_i$,
-  $p_i$ is the weight $\zeta_i$,
-  and the bias potential, $V_i$, is $\ln (\tilde Z_i / \zeta_i)$.
-  Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
-  up to a constant shift.
-}
 
 Roughly speaking, there are two types of FDS methods.
-%
 The first type,
 referred to as the equilibrium FDS, % (E-FDS) methods hereinafter,
-uses a fixed bias potential, $V_i$,
-(estimated a priori)
+uses a fixed bias potential, $V_i$ (estimated a priori),
 during simulation.
-%
 Examples include the original
 entropic or multicanonical sampling\cite{
-berg1992, *lee1993}
+berg1992, lee1993}
 and simulated tempering\cite{
-marinari1992, *lyubartsev1992}
+marinari1992, lyubartsev1992}
 with $i$ being the index of energy
 and that of temperature, respectively.
-%
 The estimated bias potentials at bin $i$, $V_i$,
-are often improved
+can be improved
 with the normalized histogram of the sampling, $H_i$, accumulated
 during simulation.
-%
 Using $H_i$ as an estimate of $\pi_i$ in
@@
 allows us to correct $V_i$ as
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{This follows from
-  $$
-  \begin{aligned}
-    H_i \approx \pi_i
-    &\propto p^*_i \, e^{-V_i},
-    \\
-    p_i
-    &\propto p^*_i \, e^{-V^\mathrm{corr}_i}.
-  \end{aligned}
-  $$
-}
 However, for an unknown system,
@@
 $V_i$ with which to start.
-%
 Thus,
@@
 to achieve the desired flat distribution.
-%
-%Another difficulty of this approach is that
 The effectiveness of this approach
@@
 that must be decided in advance.
-%
 Thus, the rigidity of this framework disallows
-continuous, rapid improvement of the bias potential,
-and subsequently the PMF,
+continuous, rapid improvement of the bias potential, and subsequently the PMF,
 as simulation lengthens.
@@
 The second type of methods,
-which we will refer to as the adaptive FDS methods,
+referred herein as adaptive FDS methods,
 address the above shortcomings
-by continuously and incrementally updating the bias potential.
-%For this reason, the second type of methods,
-%referred to as the adaptive FDS methods,
-%were invented.
-%
+by continuously and incrementally updating the bias potential to accelerate convergence of the target distribution.
 Examples include
 the WL algorithm and metadynamics.
-%
-%These methods incrementally update the bias potential
-%and can approximately sample the target distribution.
-%
 However,
-the on-the-fly updates break the microscopic reversibility
+as mentioned, the on-the-fly updates break the microscopic reversibility
 of the underlying sampling mechanism.
@@
 would lead to roughness or errors in the final estimate of the PMF.
-%
-%Thus, for a conservative practitioner,
-%the E-FDS methods appear more appealing
-%once a sufficiently accurate bias potential is available.
-%
 To minimize this effect,
-many adaptive methods adopt a schedule to
+many adaptive FDS methods adopt a schedule to
 reduce the updating magnitude
@@
 marsili2006,
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
 liang2007,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
 barducci2008}.
-%
-%Thus, in late stages of a long simulation,
-%the updating becomes negligible and
-%the adaptive simulation tends to
-%the corresponding equilibrium one.
-%
-%if the bias potential were converged to the exact value,
-%this adaptive FDS tends to the ideal equilibrium counterpart,
-%particular equilibrium simulation can be thought as
-%the optimal asymptotic limit of the adaptive counterpart,
-%
 For example, one can show that
-if the magnitude is controlled by
+if the magnitude decreases according to
 the optimal inverse-time prescription\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-liang2007,
-morozov2007, zhou2008,
-komura2012, *caparica2012, *caparica2014},
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
+morozov2007, zhou2008, morozov2009,
+komura2012, caparica2012, *caparica2014},
 the WL algorithm is equivalent to the ideal equilibrium counterpart
-in improving the precision of the bias potential,
-thus allowing for maximum efficiency
+in improving the precision of the bias potential, thus
+allowing for maximal efficiency
 (cf. Appendix \ref{sec:equilerr}).
-%the bias potential in the WL algorithm
-%will evolve similar to Eq. \eqref{eq:vcorr_equil}
-%with the precision of the histogram as simulation lengthens.
 
@@
 We will mainly focus on the adaptive methods below.
-%
 In the WL algorithm\cite{wang2001, *wang2001pre},
@@
 at each sampling step $t$ as
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $i(t)$ is the bin at step $t$,
-$\delta_{i, \, i(t)}$ is the Kronecker delta function,
-which is $1$ when $i = i(t)$ or $0$ otherwise,
 and $\alpha(t)$ is the updating magnitude.
-%
 We refer to this scheme as the \emph{single-bin scheme},
@@
 at the current bin, $i(t)$.
-%
 In a more general \emph{multiple-bin scheme}, however,
 several neighboring bins are updated:
-%
 \begin{equation}
@@
 \end{equation}
-%
 where the elements, $w_{ij}$'s, are a set of coefficients satisfying
@@
 in Sec. \ref{sec:updating-matrix}.
-%
 The Gaussian updating scheme
@@
 
-One can show that with a constant magnitude,
-$\alpha(t) = \alpha > 0$,
-the distribution collected from
+One can show that with a constant magnitude $\alpha(t) = \alpha > 0$,
+the distribution, $\pi_i$, collected from
 an adaptive FDS simulation
-is identical to $p_i$ in the long time or asymptotic limit.
-%
+is identical to $p_i$ in the long time limit.
 However, as the above updates
 change $V_i(t)$ continuously,
-the equilibrium distribution,
-Eq. \eqref{eq:pi_p_V}, no longer holds
+the equilibrium distribution ()Eq. \eqref{eq:pi_p_V}) no longer holds
 at all times,
@@
 depends on the updating magnitude\cite{
-  belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-  zhou2005, morozov2007, zhou2008,
-  laio2005, bussi2006, poulain2006, liang2007,
-  crespo2010, *atchade2011, *fort2015}.
-%%
-%The source of the deviation is two-fold.
-%%
-%First, since $V_i(t)$ is updated continuously,
-%there is a random noise that is proportional
-%to $\sqrt \alpha$.
-%%
-%Second, there is a systematic error
-%that comes from the updating dynamics itself,
-%as it breaks the Markovian nature
-%that underlies Eq. \eqref{eq:pi_p_v}.
-%%
-%
-Thus, one has to reduce the magnitude, $\alpha$, over time
+  zhou2005, liang2005, laio2005, bussi2006, poulain2006, liang2007,
+  morozov2007, zhou2008, morozov2009, crespo2010,
+  atchade2011, fort2015}.
+Thus, one has to reduce the magnitude $\alpha$ over time
 to improve the estimate of the distribution,
@@
 wang2001, *wang2001pre},
-the updating magnitude, $\alpha$,
-(or $\ln f$ therein)
+the updating magnitude $\alpha$
+(or $\ln f$ referred therein)
 is controlled in stages.
-%
 In each stage, $\alpha$
@@
 is collected and monitored.
-%
 Once the histogram is sufficiently flat,
@@
 wang2001, *wang2001pre}.
-%
 While this scheme works well for early stages,
-%
 it tends to reduce $\alpha$
@@
 saturate\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
-
-%For the single-bin scheme
-%used in WL algorithm,
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008}.
+
 Later,
@@
 a better way
-of controlling the schedule, $\alpha(t)$,
+of controlling the schedule $\alpha(t)$
 in late stages
 is to follow the formula
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $t$,
@@
 from the beginning of the simulation.
-%
 This simple yet effective formula has attracted
 several studies\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-morozov2007, zhou2008,
-komura2012, *caparica2012, *caparica2014}.
-%
-An interesting explanation\cite{
-  marsili2006, barducci2008}
-comes from the interpretation of
-the bias potential %under the WL updating scheme, Eq. \eqref{eq:wl_update},
-as a runtime average from all data collected so far,
-so that the addition of a new data point only
-affects the average by a weight of $1/t$,
-which is the inverse of the current sample size, $t$
-(cf. Appendix \ref{sec:equilerr}).
-%
+belardinelli2007, *belardinelli2007jcp, *belardinelli2016, belardinelli2008,
+morozov2007, zhou2008, morozov2009,
+komura2012, caparica2012, *caparica2014}.
 Here we will re-derive this formula
@@
 according to the initial error.
-%
 Interestingly, we may reach the same result
@@
   marsili2006, barducci2008, dickson2011, dama2014}.
-%
-Our aim here is to find the optimal schedule, $\alpha(t)$,
-for a general updating scheme.
+Our aim here is to find the optimal schedule $\alpha(t)$,
+for a general, universal updating scheme.
 
@@
 Eq. \eqref{eq:wl_update}.
-%
 To do so,
@@
 First, we need to define the error in the bias potential.
-%
-This is done by deducting two contributions from the bias potential.
-%
+This is done by removing two contributions from the bias potential.
 The first contribution comes from the target value,
 Eq. \eqref{eq:Vi_target}.
-%
 For this purpose,
 we introduce a shifted bias potential
-%
 \begin{equation}
@@
 \end{equation}
-%
 which should tend to a constant of $i$
@@
 remains a constant during the course of simulation,
-%%
-%In terms of $v_i$'s, Eq. \eqref{eq:pi_p_V}
-%becomes
-%%
-%\begin{equation}
-%  \pi_i
-%  =
-%  \frac{                p_i \, e^{-v_i} }
-%       { \sum_{j = 1}^n p_j \, e^{-v_j} }
-%  \propto
-%  p_i \, e^{-v_i}.
-%  \label{eq:pi_p_v}
-%\end{equation}
-%%
-%According to Eq. \eqref{eq:pi_p_V1},
 
@@
 generally grows over time.
-%
 Since a uniform increase of $V_i(t)$, or, equivalently, that of $v_i(t)$,
 does not affect the
-resulting distribution, $\pi_i(t)$,
-%
+resulting distribution $\pi_i(t)$,
 the real deviation is more readily defined
@@
 dama2014},
-%
 \begin{equation}
   x_i(t) \equiv v_i(t) - \bar v(t)
-  %=
-  %v_i - \sum_{j = 1}^n \rho_j \, v_j
   ,
@@
 \end{equation}
-%
 where
-%
 \begin{equation}
@@
 for some distribution,
-$\pmb \rho = (\rho_1, \dots, \rho_n)$
-(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
-%
+$\pmb \rho = (\rho_1, \dots, \rho_n)$.
 In this way,
 the variable, $x_i(t)$,
-represents the actual deviation of the bias potential, $V_i(t)$,
-and the total error is given by
-%
-\begin{equation}
-  \Err(t)
+represents the error of the bias potential,
+and is expected to vanish on convergence.
+We, therefore, define
+the error of the bias potential as
+\begin{equation}
+  \Err
   =
@@
 Note that the distribution, $\pmb \rho$,
-introduced above to define the error
-is independent of
-the target distribution,
-$\mathbf p = (p_1, \dots, p_n)$.
-%
+introduced above is independent of
+the target distribution $\mathbf p$.
 For example,
@@
 barducci2008, dama2014}.
-%
 From Eqs. \eqref{eq:x_def} and \eqref{eq:vbar_def},
 we have
-%
 \begin{equation}
@@
 \end{equation}
-%
 
@@
 $
-%\begin{equation}
-%  h_i(t) = \delta_{i, i(t)}
-%  ,
-%  \label{eq:h_def}
-%\end{equation}
 %
@@
 or zero otherwise.
-%
 The shifted bias potential, $v_i(t)$,
@@
 and a stochastic part, $\zeta_i(t)$,
-%
 \begin{equation}
@@
 \end{equation}
-%
 The deterministic part can be related
@@
 where the ensemble consists of many similar simulation copies
-sharing the same schedule $\alpha(t)$
-and the same bias potential at time $t$.
-%
+sharing the same schedule $\alpha(t)$ and
+the same bias potential at time $t$.
 The initial states and the stochastic forces
@@
 the bias potential remains roughly the same for a short period.
-%
 Then,
@@
 for the deterministic part
-%
 \begin{align}
@@
 \end{align}
-%
 where we have assumed the smallness
@@
 on the second line.
-%
-\note{
-The second step follows from
-$$
-\frac{ \langle h_i(t) \rangle }
-     { p_i }
-\approx
-\frac{                       1 - x_i  }
-     { \sum_{ r = 1 }^n p_r (1 - x_r) }
-=
-\frac{                       1 - x_i  }
-     { 1 - \sum_{ r = 1 }^n p_r \, x_r }
-.
-$$
-}%
-
-%
+
 For the random noise part,
@@
 $\langle \zeta_i(t) \rangle = 0$,
-%
-%The noise is not necessarily white,
 and the autocorrelation functions
 depend only on the time difference
-%
-\begin{equation}
-  \left\langle
-    \zeta_i(t) \, \zeta_j(t')
-  \right\rangle
+\begin{equation}
+  \langle \zeta_i(t) \, \zeta_j(t') \rangle
   =
@@
   \notag
-  %\label{eq:zeta_zeta_correlation}
-\end{equation}
-%
+\end{equation}
 where $\sigma_{ij}(t)$ is an even function
 that vanishes at large $t$.
-%
-\note{More explicitly,
-if the transition matrix of a period, $\tau$,
-is $T^\tau_{ij}$,
-then
-$
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  =
-  \langle h_i(t) \, h_j(t') \rangle
-  -
-  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
-  =
-  T^{t - t'}_{ij} \, \pi_j - \pi_i \, \pi_j.
-$
-}%
-Below, we will assume the noise is white,
-such that
-$$
-\sigma_{ij}(t) = \Gamma_0 \, \delta(t)
-$$
-for some $\Gamma_0$,
-and $\delta(t)$ is the Dirac's delta function.
-%
-This approximation is valid if the simulation period, $T$,
-is much longer than the autocorrelation time of $\sigma_{ij}(t)$,
-%
-and for a non-white noise,
-the equivalent value of $\Gamma_0$ can be obtained
-from the integral of the autocorrelation function,
-$\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$.
-
-%The average growth rate of $v(t)$ is given by
-%\begin{equation}
-%  \frac{ d \bar v (t) }
-%       { d t }
-%  =
-%  \alpha(t)
-%  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
-%  .
-%\label{eq:dvbardt}
-%\end{equation}
-%
 
@@
 \eqref{eq:rhox_sum}-\eqref{eq:h_ave},
-%\eqref{eq:v_def},
-%\eqref{eq:x_def},
-%\eqref{eq:vbar_def},
-%\eqref{eq:rhox_sum},
-%\eqref{eq:vt_diffeq},
-%\eqref{eq:h_split}
-%\eqref{eq:h_ave},
-%\eqref{eq:dvbardt},
 we get a set of decoupled equations
-%
 \begin{equation}
@@
          {     p_r    }
-  ,
+  .
 \label{eq:xi_def}
 \end{equation}
-%
-is also a white noise as $\zeta_i(t)$.
-%
-\note{The derivation:
-$$
-\begin{aligned}
-  \dot x_i
-  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \dot v_i - \dot{ \bar v }
-  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:vbar_def}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left(
-    \frac{ h_i } { p_i }
-    -
-    \sum_{ r = 1 }^n
-      \rho_r \frac{ h_r } { p_r }
-  \right)
-  \\
-  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_split} and \; \eqref{eq:h_ave}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left(
-    \frac{ \langle h_i \rangle } { p_i }
-    -
-    \sum_{ r = 1 }^n
-      \rho_r \frac{ \langle h_r \rangle } { p_r }
-    +
-    \xi_i(t)
-  \right)
-  \\
-  &\stackrel{\mathrm{Eqs.\; \eqref{eq:rhox_sum} \;and\; \eqref{eq:h_ave}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left[ -x_i(t) + \xi_i(t) \right].
-\end{aligned}
-$$
-}
-
-The total error
-at the end of the simulation, $t = T$,
-as defined in Eq. \eqref{eq:error_sum},
+
+The total error defined in Eq. \eqref{eq:error_sum}
 now depends implicitly on the function $\alpha(t)$
-via $x_i(T)$.
-%
+via $x_i(t)$.
 Our aim is to find the $\alpha(t)$
-that minimizes this error.
-%
+that minimizes the error
+at the end of the simulation period $T$.
 It suffices to consider
@@
 of a single variable:
-%
 \begin{equation}
@@
 \end{equation}
-%
-where $\xi(t)$ is a white noise, such that
-%
+where $\xi(t)$ is a generalized noise
+that is invariant under time translation:
 $
@@
   =
-  \Gamma \, \delta(t - t')
+  \kappa(t - t')
   .
 $
-%
+Additionally,
+we assume that $\kappa(t)$ is an even function and
+\begin{equation}
+  \lim_{t \rightarrow \infty} \kappa(t)
+  = 0
+  .
+  \label{eq:kappa_limit}
+\end{equation}
+For example, for a white noise,
+$\kappa(t)$ is proportional to a
+Dirac $\delta$-function, $\delta(t)$.
 We wish to show that the $\alpha(t)$
@@
 is given by
-%
-\begin{equation}
-  \alpha(t) = \frac{ 1 }{ t + t_0 },
+\begin{equation}
+  \alpha(t) = \frac{ 1 }{ t + t_0 }.
 \label{eq:alpha_invt1}
 \end{equation}
-%
-where $t_0$ is a constant.
 
@@
 \notag
-%\label{eq:xt_solution}
 \end{equation}
@@
 \end{align}
-%
 At the end of period $T$, we have
-%
 \begin{align}
@@
 \rangle e^{-2 \, q(T)}
+\notag
+\\
+&\qquad
 +
-\Gamma \int_0^T
-  {\dot u}^2\bigl( q(t) \bigr) \,
-  \, dt.
+\int_0^T \int_0^T
+  \dot u\bigl( q(t) \bigr) \,
+  \dot u\bigl( q(t') \bigr) \,
+  \kappa(t - t') \, dt' \, dt.
 \label{eq:x2t_average}
@@
 This is a functional of the curve $q(t)$.
-%
-We now fix the endpoint value, $q(T)$,
+We now fix the endpoint value $q(T)$
 and vary the curve, $q(t)$ ($0 < t < T$).
-%
 This is equivalent to varying $u$
@@
 $u\bigl( q(T) \bigr) = 1$.
-%
 Note also that the first term on the right-hand side
 of Eq. \eqref{eq:x2t_average} is fixed during
-the process,
-while the second term is the action of a free particle,
-which must move at a constant velocity, $c$:
-%
-%We will further demand that
-%%
-%\begin{equation}
-%  \lim_{t \to \infty} q(t) \to \infty.
-%  \label{eq:qt_limit}
-%\end{equation}
-%%
-%Then, for a large $t$,
-%the second term on the right-hand side
-%of Eq. \eqref{eq:xt_solution} can be neglected, and
-%%
-%\begin{align}
-%\left\langle x^2(t) \right\rangle
-%%&=
-%%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
-%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
-%%\notag
-%%\\
-%&=
-%\int_0^t \int_0^t
-%  \dot u\bigl( q(t') \bigr) \,
-%  \dot u\bigl( q(t'') \bigr) \,
-%  \kappa(t' - t'') \, dt'' \, dt'.
-%\label{eq:x2t_average_asym}
-%\end{align}
-%
-%By the Euler-Lagrange equation, we get
-%$$
-%\begin{aligned}
-%0
-%&=
-%\frac{d}{dt} \int_0^T
-%  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
-%\\
-%&= \int_0^T
-%  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
-%\end{aligned}
-%$$
-%where we have dropped the boundary terms
-%by assuming Eq. \eqref{eq:kappa_limit}
-%and a large $T$.
-%
-%In fact, we can repeat the process $n$ times, and
-%$$
-%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
-%\qquad (n \ge 1)
-%$$
-%
-%In order for this to hold for any $t$,
-%we must have
-%%
-%$
-%\ddot u\bigl( q(t') \bigr) = 0
-%,
-%$
-%or
+the process.
+By the Euler-Lagrange equation, we get
+$$
+\begin{aligned}
+0
+&=
+\frac{d}{dt} \int_0^T
+  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
+\\
+&= \int_0^T
+  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
+\end{aligned}
+$$
+where we have dropped the boundary terms
+by assuming Eq. \eqref{eq:kappa_limit}
+and a large $T$.
+In order for this to hold for any $t$,
+we must have
+$
+\ddot u\bigl( q(t') \bigr) = 0
+,
+$
+or
 \begin{equation}
   \dot u\bigl( q(t') \bigr) = c
-  .
+  ,
 \label{eq:dudt_const}
 \end{equation}
-%
-Then, by using Eq. \eqref{eq:u_def}, we have
+where $c$ is a constant of $t'$.
+Using Eq. \eqref{eq:u_def},
+we get
 $$
@@
 where $t_0$ is a constant.
-%
 Taking the logarithm,
 and differentiating both sides
-with respect to $t'$,
-we get Eq. \eqref{eq:alpha_invt1}.
+with respect to $t'$
+yields Eq. \eqref{eq:alpha_invt1}.
 
@@
 is of the form of Eq. \eqref{eq:dxdt_alpha}.
-%
 Thus,
-the optimal schedule, $\alpha(t)$, is given by
+the optimal schedule $\alpha(t)$ is given by
 Eq. \eqref{eq:alpha_invt1}
-for it optimizes each term, $\langle x_i^2(T) \rangle$,
+for it optimizes each term $\langle x_i^2 \rangle$,
 and hence the sum;
@@
 is the special case of $t_0 = 0$.
-%
 In Sec. \ref{sec:finlencorr},
@@
 given in Eq. \eqref{eq:mbin_update}.
-%
 In this case,
@@
 but also those at a few neighboring bins, $i$'s.
-%
 The relative updating magnitudes,
 $w_{ij}$'s,
-form a matrix, $\mathbf w$.
-%
+form a matrix $\mathbf w$.
 Then, Eq. \eqref{eq:vt_diffeq} is generalized to
-%
 \begin{equation}
@@
 The optimization procedure is similar to the single-bin case.
-%
 However, we will assume some restrictions and approximations
@@
 
-The matrix, $\mathbf w$, is not arbitrary.
-%
+The matrix $\mathbf w$ is not arbitrary.
 First, to sample
-the desired distribution,
+the desired distribution
 $\mathbf p = (p_1, \dots, p_n)$,
-${\dot V}_i(t)$
+ $\dot V_i(t)$
 in Eq. \eqref{eq:vt_diffeq_mbin}
@@
 if $h_j(t)$ were the same as $p_j$.
-%
 This allows the bias potential to grow uniformly
 in the asymptotic regime.
-%
 Thus, by a proper scaling of $\alpha(t)$,
 we may write this condition as
-%
 \begin{equation}
@@
 \end{equation}
-%
 In other words, $(1, \dots, 1)^T$
@@
 with eigenvalue $1$.
-%
 Thus, the transpose $\mathbf w^T$
@@
 except that some elements can be negative.
-%
 Below we will take advantage of the analogy
@@
 the detailed balance condition,
-%
 \begin{equation}
@@
 \end{equation}
-%
-for the distribution, $\pmb \rho$,
+for the distribution $\pmb \rho$
 introduced in Eq. \eqref{eq:vbar_def}.
-%
 It follows that
-%
 \begin{equation}
@@
 \end{equation}
-%
-i.e. $\mathbf \rho$ is a left eigenvector of
+i.e., $\mathbf \rho$ is a left eigenvector of
 $\mathbf w$ with eigenvalue $1$.
@@
 eigenvectors, $\phi_{ki}$,
-%
 \begin{equation}
@@
 \end{equation}
-%
 satisfying the orthonormal conditions\cite{vankampen}:
-%
 \begin{align}
@@
 corresponding to the distribution, $\pmb \rho$.
-%
 We will label this eigenvector as the first one,
 and
-%
 \begin{equation}
@@
 \end{equation}
-%
 which satisfies Eq. \eqref{eq:eig_orthonormal_rows}
 for $k = l = 1$.
-%
 The corresponding eigenvalue is $\lambda_1 = 1$
 by Eq. \eqref{eq:w_balance}.
-%
 Further, for $k \ge 2$, we have
-%
 \begin{equation}
@@
 
-%It is worth pointing out
 Finally, there is a stability condition.
-%
 In a \emph{stable} updating scheme,
 no eigenmode grows indefinitely,
-which means that no eigenvalue, $\lambda_k$,
+which means that no eigenvalue $\lambda_k$
 can be negative.
@@
 We can now simplify Eq. \eqref{eq:vt_diffeq_mbin}.
-%
-By using Eqs.
+By using Eqs. \eqref{eq:vbar_def}
+and \eqref{eq:w_balance},
+we can show that Eq. \eqref{eq:dvbardt}
+remains true,
+and by using Eqs.
 \eqref{eq:x_def},
-\eqref{eq:vbar_def},
 \eqref{eq:rhox_sum},
@@
 \eqref{eq:xi_def},
-\eqref{eq:w_sumj},
-and \eqref{eq:w_balance} we get
-%
+and \eqref{eq:w_sumj}, we get
 $$
@@
 $$
-%where
-%we have expanded the right-hand side
-%in linear terms of $x_j$
-%on the second line.
-
-\note{The derivative of $\dot {\bar v} = \alpha(t)$,
-$$
-\begin{aligned}
-  \dot{ \bar v }
-  &
-  \stackrel{\mathrm{Eq.\; \eqref{eq:vbar_def}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \sum_{i = 1}^n \rho_i \, \dot v_i
-  %
-  \stackrel{\mathrm{Eq.\; \eqref{eq:vt_diffeq_mbin}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  %
-  \alpha(t)
-  \sum_{j = 1}^n \frac{ h_j } { p_j }
-                 \sum_{ i = 1 }^{n} \rho_i \, w_{ij}
-  \\
-  &
-  \stackrel{\mathrm{Eq.\; \eqref{eq:w_balance}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t)
-  \sum_{j = 1}^n \rho_j
-    \, \frac{ h_j } { p_j }
-  %
-  \stackrel{\mathrm{Eq.\; \eqref{eq:w_sumj}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t)
-  \sum_{j = 1}^n w_{ij}
-  \sum_{r = 1}^n \rho_r
-    \, \frac{ h_r } { p_r }
-  .
-\end{aligned}
-$$
-}
-
-Diagonalizing the matrix, $\mathbf w$,
+
+Diagonalizing the matrix $\mathbf w$
 using the eigenvectors defined in Eqs.
@@
 we get a set of decoupled equations
-%
 \begin{equation}
@@
 \end{equation}
-%
 where
@@
 \end{align}
-are the deviation and noise of mode, $k$, respectively.
+are the deviation and noise of mode $k$, respectively.
 
@@
 can be rewritten in terms of the eigenmodes of $\mathbf w$
-%
 \begin{align}
@@
 \end{align}
-%
 where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
-%
 From Eq. \eqref{eq:eigenmode1},
@@
 vanishes identically by Eq. \eqref{eq:rhox_sum}.
-%
 Thus, we can start the last sum %in Eq. \eqref{eq:y2_sum}
@@
 \begin{equation}
-  \Err(T)
+  \Err
   =
@@
 In analogy to Eq. \eqref{eq:x2t_average},
-using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
-we can write the error after a period, $T$,
+using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum}, we can write
+the error after a period $T$
 as a sum of two contributions,
-%
-\begin{equation}
-  \Err(T)
-  =
-  \Err_R(T) + \Err_A(T),
+\begin{equation}
+  \Err
+  =
+  \Err_R + \Err_A,
   \label{eq:error_split}
 \end{equation}
-%
 where the residual error
@@
 \begin{equation}
-  \Err_R(T)
+  \Err_R
   =
@@
 \begin{equation}
-  \Err_A(T)
+  \Err_A
   =
@@
 In the asymptotic regime
-(i.e. if the simulation period, $T$, is sufficiently long),
+(i.e., if the simulation period, $T$, is sufficiently long),
 we can approximate $\kappa_k(t)$ as
@@
 where
+%
 \begin{equation}
@@
   \notag
-  %\label{eq:Gamma_integral}
 \end{equation}
@@
 \end{equation}
-%
 Then, Eq. \eqref{eq:error_asym} is simplified as
-%
 \begin{align}
-  \Err_A(T)
+  \Err_A
   &=
@@
 \end{align}
-%
 In this way,
@@
 few $\Gamma_k$'s.
-%
 In Appendix \ref{sec:Gamma},
@@
 $\alpha(t) = a_0$.
-%
-After a long period, $T'$,
+After a long period $T'$,
 the residual error becomes negligible, and
@@
 is given by
-%
-%\begin{align}
-%  \Err
-%  &=
-%  \Err_A
-%  =
-%  \int_0^{T'}
-%    \sum_{k = 2}^n
-%      \Gamma_k \, (\lambda_k \, a_0)^2 \,
-%      e^{ 2 \, \lambda_k \, a_0 \, (t - T') }
-%    \, dt
-%  \notag
-%  \\
-%  &
-%  =
-%  %\stackrel{ T' \to \infty }
-%  %{ =\joinrel=\joinrel=\joinrel= }
-%  %
-%  \sum_{ k = 2 }^n
-%    \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
-%  ,
-%  \qquad
-%  ( T' \to \infty )
-%  .
-%\notag
-%%\label{eq:error_eql}
-%\end{align}
-%
-%In terms of the components, we have
-%
 \begin{align}
@@
 \end{align}
-%
 
 We will use this result to model the initial error
-before starting the target schedule, $\alpha(t)$.
-%
-That is, we assume that there is an equilibration phase
-for $t < 0$ during which the adaptive FDS sampling was performed
-under a constant updating magnitude,
-$\alpha(t) = a_0$, and at the end of the equilibration, $t = 0$,
-the system has entered the equilibrium state of sampling,
-in which the error of mode $k$ obeys Eq. \eqref{eq:y2_eql}.
-%
-If we use this result
+before starting the schedule $\alpha(t)$ to be optimized.
+That is, we assume
+$\alpha(t) = a_0$
+for $t < 0$, and
+the system has entered the sampling equilibrium
+under a fixed $a_0$ at $t = 0$.
+We can then use Eq. \eqref{eq:y2_eql}
 for the $\langle y_k^2(0) \rangle$'s
 in Eq. \eqref{eq:error_res},
+and
 the residual error at the end of period $T$
 is given by
-%
-\begin{equation}
-  \Err_R(T)
+\begin{equation}
+  \Err_R
   =
@@
 \end{equation}
-%
 Note that the residual error
 depends on $\alpha(t)$ only through
-the integral, $q(T) = \int_0^T \alpha(t) \, dt$.
-%
+the integral $q(T) = \int_0^T \alpha(t) \, dt$.
 Thus, if we vary $\alpha(t)$ under a fixed $q(T)$,
@@
 and the optimization can be done exclusively
-on the asymptotic part, $\Err_A(T)$.
+on the asymptotic part, $\Err_A$.
 
@@
 that of the asymptotic part.
-%
 We can rewrite Eq. \eqref{eq:error_asym1} much like an action
 of a particle whose position is given by $q(t)$:
-%
-\begin{equation}
-  \Err_A(T)
+\begin{equation}
+  \Err_A
   =
@@
   ,
-%\notag
 \label{eq:error_asym_Lagrangian}
 \end{equation}
-%
 where the Lagrangian is
@@
     \Gamma_k \, \dot u_k^2\bigl[ q(t) \bigr]
-  %\notag
-  %\\
-  %&=
   =
@@
 %
-In this mechanical analogy,
-the schedule, $\dot q(t) = \alpha(t)$,
-corresponds to the velocity of the particle.
-%
-So the above Lagrangian contains only the kinetic energy,
-and, therefore, describes a free particle
+As the schedule, $\dot q(t) = \alpha(t)$,
+corresponds to the velocity of the particle,
+the above Lagrangian contains only the kinetic energy.
+Thus, the above mechanical analogy
+represents a free particle
 carrying a position-dependent mass.
-%
 Since the Lagrangian
@@
 the Hamiltonian is conserved.
-%
 But in this case the Hamiltonian
-%
 \begin{equation}
@@
   \mathcal L,
-  %=
-  %\mathrm{const.}
   \notag
-  %\label{eq:error_asym_Hamiltonian}
-\end{equation}
-%
+\end{equation}
 coincides with the Lagrangian.
-%
 So we may set
-%
 \begin{equation}
@@
   \left(
-    \frac{C}{T}
+    \frac C T
   \right)^2
@@
 \end{equation}
-%
-for some positive constant, $C$.
-%
+for some positive constant $C$.
 This serves
 as a generalization of Eq. \eqref{eq:dudt_const}.
-%
 By integrating the square root, we get
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $C$ is determined by
 the boundary condition at $t = T$ as
-%
 \begin{equation}
@@
 \end{equation}
-%
 Thus, we have an implicit equation for $q(t)$,
@@
 $\alpha(t) = \dot q(t)$.
-%
-Since we only optimized the asymptotic error, $\Err_A(T)$,
+Since we only optimized the asymptotic error, $E_A$,
 in the above,
 the value of $q(T)$ can be further adjusted
-to minimize the total error, $\Err(T) = \Err_R(T) + \Err_A(T)$
+to minimize the total error, $E = E_R + E_A$
 (cf. Sec. \ref{sec:optinitalpha}).
-%
 A complete procedure of computing the optimal schedule
@@
 geometrically as follows.
-%
 First, we define a mass function as
-%
 \begin{equation}
@@
 \end{equation}
-%
 and a normalized mass distribution as
-%
 \begin{equation}
@@
 \notag
-%\label{eq:mass_distr}
-\end{equation}
-%
+\end{equation}
 such that
 $\int_0^{q(T)} m(Q) \, dQ = 1$.
-%
 
@@
   \begin{tikzpicture}[very thick, scale=2.5, every node/.style={scale=1.5}]
-    % distribution
     \draw[domain=0:2.5, variable=\x, dashed, gray, ultra thick]
@@
 
-    % shade
     \fill [gray!30!white, domain=0.7:2.2, variable=\x]
@@
 
-    % thick curve
     \draw[domain=0:2.2, variable=\x, ultra thick]
@@
 
-    % label for the shaded area
     \node[] at (1.1, 0.5) {$\dfrac{t}{T}$};
@@
                (0.6, 1.174);
-    % label for the ordinate
-    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
 
@@
 
-    % xtics
     \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
@@
   \end{tikzpicture}
-  %\makebox[\linewidth][c]{
-  %  \includegraphics[angle=0, width=0.9\linewidth]{fig/massq.pdf}
-  %}
   \caption{
@@
     via the normalized mass distribution, $m(Q)$.
-    %
-    \note{
-      The figure was produced by \texttt{tikz}.
-      % Gnuplot version by doc/fig/massq.gp
-    }%
   }
@@
 %
-which means that each point,
-$\bigl(t, \alpha(t)\bigr)$,
+which means that each point
+$\bigl(t, \alpha(t)\bigr)$
 on the schedule can be mapped
-to a point,
-$\bigl(Q, m(Q)\bigr)$,
+to a point
+$\bigl(Q, m(Q)\bigr)$
 on the mass distribution,
@@
 gives $1/(T\alpha)$,
-and the area under the curve in the domain, $[Q, q(T)]$,
+and the area under the curve in the domain $[Q, q(T)]$
 is equal to $t/T$,
@@
 as follows.
-%
 We collect in each class optimal schedules
-computed from simulations with the same function, $M(Q)$,
-and cutoff value of $q(T)$.
-%
+computed from simulations with the same function $M(Q)$
+and cutoff value $q(T)$.
 The former is determined by the values of
 $\lambda_k$'s and $\Gamma_k$'s,
-whereas the latter affects the distribution, $m(Q)$,
-via the normalization constant, $C$.
+whereas the latter affects the distribution $m(Q)$
+via the normalization constant $C$.
 
@@
 of lengths $T_1$ and $T_2$
-at corresponding times, $t_1/T_1 = t_2/T_2$,
+at corresponding times $t_1/T_1 = t_2/T_2$
 are related by $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
-%
 Further, by Eqs.
@@
 \begin{equation}
-  \Err_A(T)
+  E_A
   =
@@
 for schedules in the same class,
-i.e. we have the scaling relation
-$T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2)$.
-
-%The optimal schedule has the following scaling property.
-%%
-%For two simulations of lengths $T_1$ and $T_2$,
-%if $q(T_1) = q(T_2)$
-%and they share the same $\lambda_k$'s and $\Gamma_k$'s,
-%the updating magnitudes at the corresponding times,
-%$t_1/T_1 = t_2/T_2$,
-%can be related as $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
-%%
-%In other words,
-%if $\alpha(t_1)$ is
-%an optimal schedule of the simulation of length $T_1$,
-%then
-%$\frac{T_1}{T_2} \alpha\left( \frac{ T_1 } { T_2 } t_2 \right)$
-%is an optimal schedule of the simulation of length $T_2$.
-%%
-%The asymptotic errors are related by
-%$T_1 \, E_{A1} = T_2 \, E_{A2}$.
+i.e., we have the scaling relation
+$T_1 \, E_{A1} = T_2 \, E_{A2}$.
 
@@
 the optimal initial updating magnitude.
-%
 Let us now consider simulations in which
 the system is initially equilibrated at the same
-constant updating magnitude, $a_0$.
-%
+constant updating magnitude $a_0$.
 For these simulations,
@@
 to minimize the total error.
-%
 From Eqs. \eqref{eq:error_res1}
@@
 \end{equation}
-%
-\note{Let $Q_c \equiv q(T)$,
-$$
-\begin{aligned}
-  \frac{
-    \partial \Err_R(T)
-  }
-  {
-    \partial Q_c
-  }
-  &=
-  -a_0 \, M^2(Q_c)
-  ,
-  \\
-  \frac{
-    \partial \Err_A(T)
-  }
-  {
-    \partial Q_c
-  }
-  &=
-  \frac 2 T \,
-  M(Q_c) \,
-  \int_0^{ Q_c } M(Q) \, dQ
-  .
-\end{aligned}
-$$
-Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
-}%
 By Eqs. \eqref{eq:mQ_invTa} and \eqref{eq:intmQ_tT},
@@
 \end{equation}
-%
-i.e. the optimal initial updating magnitude
-is always half of the equilibrium value.
-%
-This factor, $1/2$,
-happens to coincide with the
+i.e., the optimal initial updating magnitude
+is always half of the equilibration value.
+This factor $1/2$
+is precisely the
 recommended reduction factor of the updating magnitude
@@
 simulations with the same $a_0$.
-%
 Since both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$,
@@
 more so if $M(Q)$ has a long tail.
-%
 This means that, if many $\lambda_k$'s are near zero,
 both the optimal schedule and
-normalized asymptotic error, $\Err_A(T) \cdot T$,
+normalized asymptotic error, $E_A \cdot T$,
 would be sensitive to the simulation length, $T$.
-%
-%If $M(Q)$ decays rapidly,
-%the distribution $m(Q)$ does not critically
-%depend on the cutoff $q(T)$,
-%once $q(T)$ is sufficiently large.
-%%
-%So the optimal schedules of these simulations
-%approximately satisfy the scaling relations
-%described in Sec. \ref{sec:mass_distr}.
-%%
-%If, however, $M(Q)$ has a long tail
-%because of some near-zero $\lambda_k$'s,
-%increasing the simulation length $T$
-%generally alters the shape of the optimal schedule.
 
@@
 centered on the current bin.
-%
 We call these schemes homogeneous,
 or translationally invariant.
-%
 An example is the Gaussian updating scheme
@@
 Eq. \eqref{eq:w_detailedbalance},
-for the flat distribution, $\rho_i = 1/n$,
-i.e. the updating matrix, $\mathbf w$,
+for the flat distribution, $\rho_i = 1/n$ -
+i.e., the updating matrix, $\mathbf w$,
 is symmetric.
-%
 The updating matrices
 share the same set of eigenvectors
-(which here are sines and cosines of different frequencies),
+(sines and cosines of different frequencies),
 and thus are completely determined
 by the eigenvalues.
-%
 Further, the multiplication of $\mathbf w$
@@
 or the updating kernel.
-%
 Thus,
@@
 
-%While translationally-invariant
-%updating schemes
-%naturally suit periodic variables\cite{
-%dama2014},
-%they can also be extended to non-periodic variables
-%by imposing the reflective boundary condition\cite{
-%bussi2006}.
-%%
-%For simplicity, we will assume that the target
-%distribution is flat, or $p_i = 1/n$, below.
-
 It is natural to first introduce homogeneous updating schemes
 for a periodic variable\cite{dama2014}.
-%
 The updating matrix, $\mathbf w$,
 assumes the following form
-%
 \begin{equation}
@@
 \notag
-%\label{eq:w_band_pbc}
-\end{equation}
-%
+\end{equation}
 where
@@
 are a set of numbers satisfying
-%
 \begin{equation}
@@
 \end{equation}
-%
 and $\mu_l = 0$ for $|l| > b$.
-%
 We will further impose the symmetry
-%
 \begin{equation}
@@
 \end{equation}
-%
 If $b = n/2$, then $\mu_{-b}$ is removed
@@
 to avoid double counting.
-%
-These numbers constitute an \emph{updating kernel}\cite{bussi2006}.
-%
+These numbers constitute an \emph{updating kernel}.
 In this way, the matrix, $\mathbf w$,
-being symmetric with respect to the indices, $i$ and $j$,
+being symmetric with respect to the indices $i$ and $j$,
 satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
@@
 the out-of-boundary values by
-%
 \begin{equation}
@@
 \end{equation}
-%
 such that $i \pm n$ lies in between $1$ and $n$.
-%
 Then, the multiplication of the matrix, $\mathbf w$,
 is equivalent to a convolution with the kernel:
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{The derivation of the first step of
-  Eq. \eqref{eq:wmul_to_convol}:
-$$
-\begin{aligned}
-  \sum_{j = 1}^n
-    w_{ij} \, \phi_j
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{j = 1}^n
-    \mu_{i - j - n} \, \phi_j
-  +
-  \sum_{j = 1}^n
-    \mu_{i - j + n} \, \phi_j
-  \\
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{l = 1+n}^{2 \, n}
-    \mu_{i - l} \, \phi_{l - n}
-  +
-  \sum_{l = 1-n}^0
-    \mu_{i - l} \, \phi_{l + n}
-  \\
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{l = 1+n}^{2 \, n}
-    \mu_{i - l} \, \phi_{l}
-  +
-  \sum_{l = 1-n}^0
-    \mu_{i - l} \, \phi_{l}
-  \\
-  &=
-  \sum_{j = 1-n}^{2 \, n}
-    \mu_{i - j} \, \phi_j
-  .
-\end{aligned}
-$$
-The second step of Eq. \eqref{eq:wmul_to_convol}
-follows from the constraint, $\mu_l = 0$, for $|l| > b$.
-
-Similarly,
-for the left eigenvector, we have
-$$
-  \sum_{ i = 1 }^n
-    \phi_i \, w_{ij}
-  =
-  \sum_{ l = -b }^b
-    \phi_{j - l} \, \mu_{-l}
-  .
-$$
-But due to the symmetry Eq. \eqref{eq:mu_symm},
-it is identical to \eqref{eq:wmul_to_convol}.
-}
-%
 As a result, the characteristic equation,
@@
 can be solved by discrete Fourier transform.
-%
 We may construct a set of orthonormal eigenvectors,
@@
 compatible with the periodicity, Eq. \eqref{eq:phi_pbc},
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $\ii^2 = -1$, and
@@
 the Fourier or cosine transform of the updating kernel:
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{Consider the unnormalized version
-  $$
-  \Phi^{(k)}_j =
-  \exp\left[
-    \frac{ ( k - 1 ) \, j \, 2 \, \pi }
-         {              n             }
-    \ii
-  \right]
-  ,
-  $$
-  we have
-  $$
-  \begin{aligned}
-  \sum_{j = 1}^n
-    w_{ij} \, \Phi^{(k)}_j
-  &=
-  \sum_{l = -b}^b
-    \mu_l \, \Phi^{(k)}_{i - l}
-  \\
-  &=
-  \mu_0 \, \Phi^{(k)}_i
-  +
-  \sum_{l = 1}^b
-    \mu_l \,
-    \left[ \Phi^{(k)}_{i - l} + \Phi^{(k)}_{i + l} \right]
-  \\
-  &=
-  \Phi^{(k)}_i \,
-  \left[
-    \mu_0
-    +
-    2 \sum_{l = 1}^b
-      \mu_l \, \cos
-      \frac{ (k - 1) \, l \, 2 \, \pi }
-           {            n             }
-  \right]
-  .
-  \end{aligned}
-  $$
-  The term in the square brackets is the eigenvalue given by
-  Eq. \eqref{eq:wband_eigenvalue_pbc}.
-}
-%
 Note, however, the two-fold degeneracy as
@@
 and change the form of the updating matrix to
-%
-%
 \begin{equation}
@@
 \end{equation}
-%
 where the last two terms on the right hand side,
@@
 are satisfied.
-%
-Again, the updating kernel should satisfy
+Again the updating kernel should satisfy
 Eqs. \eqref{eq:mu_normalization}
@@
 of $\phi_i$ as
-%
 \begin{equation}
@@
 \end{equation}
-%
 such that the matrix multiplication is still equivalent to
 a convolution as Eq. \eqref{eq:wmul_to_convol}.
-%
-\note{The derivation of Eq. \eqref{eq:wmul_to_convol}
-  in this case is similar,
-  $$
-  \begin{aligned}
-    \sum_{j = 1}^n w_{ij} \, \phi_j
-    &=
-    \sum_{j = 1}^n
-      \mu_{i - j} \, \phi_j
-    +
-    \sum_{j = 1}^n
-      \mu_{i + j - 1} \, \phi_j
-    +
-    \sum_{j = 1}^n
-      \mu_{i + j - 2 \, n - 1} \, \phi_j
-    \\
-    &=
-    \sum_{j = 1}^n
-      \mu_{i - j} \, \phi_j
-    +
-    \sum_{l = 1 - n}^0
-      \mu_{i - l} \, \phi_l
-    +
-    \sum_{l = n + 1}^{ 2 \, n }
-      \mu_{i - l} \, \phi_l
-    \\
-    &=
-    \sum_{j = 1 - n}^{ 2 \, n}
-      \mu_{i - j} \, \phi_j.
-  \end{aligned}
-  $$
-}%
 We note that Eq. \eqref{eq:phi_refl}
@@
 with a reflective symmetry around $i = 1/2$.
-%
 Accordingly, the eigenvectors,
@@
 must share the same periodicity,
-and be even functions about the axis, $i = 1/2$:
-%
+and be even functions about the axis $i = 1/2$:
 \begin{equation}
@@
 \notag
-%\label{eq:wband_eigenvector_refl}
-\end{equation}
-%
+\end{equation}
 The eigenvalues can be borrowed from
 Eq. \eqref{eq:wband_eigenvalue_pbc},
-with the period, $n$, doubled:
-%
+with the period $n$ doubled:
 \begin{align}
@@
 \end{align}
-%
-\note{Derivation.
-  First consider the unnormalized eigenvectors,
-  $$
-  \Phi^{(k)}_i
-  =
-  \cos \frac{ \left( i - \frac 1 2 \right) (k - 1) \, \pi}{n},
-  $$
-  which satisfies Eq. \eqref{eq:phi_refl}.
-  %
-  So, by Eq. \eqref{eq:wmul_to_convol},
-  $$
-  \begin{aligned}
-  \sum_{i = 1}^n
-    \Phi^{(k)}_i \, w_{ij}
-  &=
-  \sum_{l = -b}^b
-    \Phi^{(k)}_{j - l} \, \mu_l
-  \\
-  &=
-    \Phi^{(k)}_j \, \mu_0
-  + \sum_{l=0}^{b}
-    \mu_l \,
-    \left[
-      \Phi^{(k)}_{j-l}
-      +
-      \Phi^{(k)}_{j+l}
-    \right]
-  \\
-  &= \Phi^{(k)}_j \, \lambda_k,
-  \end{aligned}
-  $$
-  with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue_refl}.
-}%
-%
 We may unify Eqs. \eqref{eq:wband_eigenvalue_refl}
 and \eqref{eq:wband_eigenvalue_pbc} as
-%
 \begin{equation}
@@
 \end{equation}
-%
-where the degeneracy, $g$, is $2$
+where the degeneracy $g$ is $2$
 for the periodic case,
 or $1$ otherwise.
-%
 In both cases,
@@
 the updating kernel.
-%
 For the updating scheme to be stable,
@@
 must be nonnegative.
-%
 If this is not so,
@@
 in Appendix \ref{sec:stabilize_wband}.
-
 We discuss two special cases below.
@@
 If $\mu_2 = \dots = \mu_b = 0$,
-the matrix, $\mathbf w$, is tridiagonal,
-%%
-%\begin{equation}
-%\arraycolsep=3.6pt\def\arraystretch{1.4}
-%\mathbf w
-%=
-%\left(
-%  \begin{array}{cccccccc}
-%    1 - \mu_1   & \mu_1 & 0 & \dots & 0 \\
-%    \mu_1 & 1 - 2 \, \mu_1  & \mu_1 & \dots & 0 \\
-%    \vdots & &  & & \vdots \\
-%    0 & \dots & \mu_1 & 1 - 2 \, \mu_1  & \mu_1 \\
-%    0 & \dots & 0 & \mu_1 & 1 - \mu_1
-%  \end{array}
-%\right).
-%\label{eq:wnn}
-%\end{equation}
-%%
-%We have
+the matrix $\mathbf w$ is tridiagonal,
 and the eigenvalues are
@@
 \notag
-%\label{eq:wnn_eigenvalue}
-\end{equation}
-%
+\end{equation}
 So this updating scheme is stable if
@@
 \notag
-%\label{eq:nn_stable}
-\end{equation}
-
-\note{
-If the variable is periodic ($g = 2$) and $n$ is even,
-the minimum is achieved at $k = n/2$.
-
-If the variable is periodic ($g = 2$) but $n$ is odd,
-the minimum is achieved at $k = (n+1)/2$.
-$$
-\sin^2\left(
-  \frac{n-1}{2n} \pi
-\right)
-=
-\cos^2\left(
-  \frac{\pi}{2n}
-\right)
-.
-$$
-
-If the variable is nonperiodic ($g = 1$)
-the minimum is achieved at $k = n$.
-}
+\end{equation}
 
@@
 Gaussian-shaped kernel.
-%
 This updating scheme is commonly
 used in metadynamics.
-%
 Below we show its stability
@@
 is the bin size.
-%
-By assuming the largest cutoff, $b \approx n/g$,
+By assuming the largest $b \approx n/g$,
 we can approximate Eq. \eqref{eq:wband_eigenvalue}
 as an integral:
-%
 \begin{equation}
@@
 \notag
-%\label{eq:lambda_int}
-\end{equation}
-%
+\end{equation}
 with the normalization
-%
 $
@@
 $
-\note{
-\begin{align*}
-  \lambda_k
-  &\approx
-  \int_0^b \mu_l \, \cos[(k-1) \, x] \, dl
-  \\
-  &=
-  \int_0^b \mu(x) \, \cos[(k-1) \, x] \, \Delta x \, dl
-  \\
-  &=
-  \int_0^{b \, \Delta x} \mu(x) \, \cos[(k-1) \, x] \, dx
-\end{align*}
-The upper limit,
-$b \, \Delta x \approx (n/g) \, (g\,\pi/n) = \pi$.
-}
 
 For the Gaussian updating kernel,
-if the width, $\sigma$, satisfies $\Delta x \ll \sigma \ll 1$,
+if the width $\sigma$ satisfies $\Delta x \ll \sigma \ll 1$,
 we can extend
@@
 to infinity, and
-%
 \begin{equation}
@@
        { \sqrt{ 2 \pi \sigma^2 } }
-  %
   \exp\left(
@@
 \end{equation}
-%
-%
-If follows that all eigenvalues are positive\cite{bussi2006}:
-%
+If follows that all eigenvalues are positive:
 \begin{align}
@@
       \right]
-  %\notag
-  %\\
-  %&=
-  %\exp\left[
-  %      -\frac{ \pi^2 }{ 2 }
-  %      \left(
-  %        \frac{ k - 1 }
-  %             {   n   }
-  %      \right)^2
-  %      \left(
-  %        \frac{  \sigma }
-  %             { \Delta x }
-  %      \right)^2
-  %    \right]
   .
 \notag
-%\label{eq:lambda_Gaussian}
 \end{align}
-
-%However,
-%as $\sigma/\Delta x \gg 1$,
-%the smallest eigenvalue
-%%
-%$$
-%\lambda_n
-%\approx
-%\exp\left(
-%      -\frac{ n^2 \, \sigma^2 }
-%            {        2        }
-%    \right)
-%=
-%\exp\left[
-%      -\frac{ \pi^2 }{ 2 }
-%      \left(
-%        \frac{  \sigma }
-%             { \Delta x }
-%      \right)^2
-%    \right]
-%.
-%$$
-%%
-%is exponentially small.
-%%
-%This means that
-%the optimal value of $\lambda$
-%given by Eq. \eqref{eq:optimal_lambda_approx}
-%would be too small in practice,
-%for it requires $\lambda < 2 \, \lambda_n$.
-%%
-%In other words,
-%with a reasonable $\lambda$,
-%the error of the last few fluctuation modes
-%will always decay suboptimally,
-%accordingly to Eq. \eqref{eq:error_asym_invt}.
-%%
-%However, since these modes
-%represent short wavelength fluctuations,
-%and these modes may be in
-%the system under consideration.
-%%
-%Thus,
-%we can truncate the sum of the error function
-%in Eq. \eqref{eq:error_asym_invt}
-%at some $k_{\max}$,
-%which corresponds to a minimal length scale
-%$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
-%greater than $\sigma$.
-%%
-%Then,
-%$$
-%\lambda_{ k_{\max} }
-%\approx
-%\exp\left(
-%      -\frac{ k_{\max}^2 \, \sigma^2 }
-%            {           2            }
-%    \right)
-%=
-%\exp\left(
-%      -\frac{  \pi^2 \, \sigma^2    }
-%            {    2 \, l_{\min}^2  }
-%    \right).
-%$$
 
@@
 the relation can be readily inverted.
-%
 In the periodic case,
 we have
-%the inversion formula is
-%
 \begin{equation}
@@
 the formula is
-%
 \begin{align}
@@
 \end{align}
-%
 where
@@
 as well as
-%
 \begin{align}
@@
 \end{align}
-to satisfy the constraint, $\mu_n = 0$.
-%
-\note{Eq. \eqref{eq:lambdan}
-  ensures that the $\mu_n$
-  computed from Eq. \eqref{eq:mu_from_lambda_refl}
-  vanishes.
-
-  %The inversion formula is derived from the cosine transform.
-  %
-  If we define $\mu_n = 0$ and
-  %
-  $
-    \mu_{ 2 n - l } = \mu_l,
-  $
-  for
-  $l = 1, \dots, n - 1$,
-  %
-  then Eq. \eqref{eq:wband_eigenvalue_refl} can be rewritten as
-  %
-  \begin{align}
-    \lambda_{k+1}
-    =
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
-  \notag
-  %\label{eq:lambda_cosine_sum}
-  \end{align}
-  %
-  This formula for $\lambda_{k+1}$
-  can be readily extended to $k = 2 \, n - 1$,
-  and we have
-  %
-  $
-    \lambda_{ 2 \, n + 2 - k } = \lambda_k
-    .
-  $
-  %
-  Thus,
-  %
-  $$
-  \begin{aligned}
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \lambda_{ k + 1 } \,
-      \cos \frac{ k \, p \, \pi }
-                {      n        }
-    &=
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \sum_{ l = 0 }^{ 2 \, n - 1 }
-        \mu_l \,
-        \cos \frac{ k \, l \, \pi }
-                  {      n        }
-        \cos \frac{ k \, p \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \frac{ \mu_l } { 2 }
-      \sum_{ k = 0 }^{ 2 \, n - 1 }
-        \cos \frac{ k \, (p + l) \, \pi }
-                  {      n        }
-                  +
-        \cos \frac{ k \, (p - l) \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \mu_l \, n \left(
-        \delta_{ p + l - 2 \, n, 0 }
-        +
-        \delta_{ p - l, 0 }
-      \right)
-    \\
-    &=
-    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
-    =
-    2 \, n \, \mu_p.
-  \end{aligned}
-  $$
-  This entails Eq. \eqref{eq:mu_from_lambda_refl}.
-  %
-  The value of $\lambda_{n + 1}$
-  can be deduced from
-  imposing $\mu_n = 0$
-  and solving Eq. \eqref{eq:mu_from_lambda_refl}
-  for $l = n$,
-  yielding Eq. \eqref{eq:lambdan}.
-}
+to satisfy the constraint $\mu_n = 0$.
 
@@
 that the single-bin scheme is asymptotically optimal.
-%
 Consider a class of $\mathbf w$ matrices
@@
 $\lambda_k$'s.
-%
 Using the Cauchy-Schwarz inequality, we have,
 for any set of nonnegative numbers, $c_k$'s,
-%
 \begin{align}
@@
 \right)
-%
 \left(
@@
 \right)
-%
 \notag
@@
 \end{align}
-%
-%\note{
-This inequality follows from the non-negativity of
-the following quadratic polynomial,
-$$
-\int_0^T
-  dt \sum_{k = 2}^n \Gamma_k \,
-    \left( \dot u_k \, \chi - c_k \right)^2
-  \equiv
-  A_2 \, \chi^2 + A_1 \, \chi + A_0
-  ,
-$$
-which must carry a non-positive discriminant:
-$A_1^2 - 4 \, A_0 \, A_2 \le 0$.
-%}%
-%
 Note that the last expression of Eq. \eqref{eq:CSineq}
@@
 which is fixed in the variational process.
-%
 Thus, the inequality sets a lower bound
-for the asymptotic error $\Err(T)$
+for the asymptotic error $\Err$
 given in Eq. \eqref{eq:error_asym1}.
-%
 The equality is achieved
@@
 up to a multiplicative factor.
-%
 Solving these equations yields
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{Derivation.
-  Integrating $\dot u_k = c_k$ yields
-  \begin{equation}
-    u_k(t) = c_k \left(t + t_{k0} \right).
-  \tag{UK}
-  \label{neq:ui_solution}
-  \end{equation}
-  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
-  we have
-  $$
-  c_k = \frac{ 1 }{ T + t_{k0} },
-  \quad
-  \mathrm{and\;}
-  \frac{ t_{k0} } { T + t_{k0} }
-  =
-  e^{ -\lambda_k \, q(T) }.
-  $$
-  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
-  $$
-  -\lambda_k \, q(T) + \lambda_k \, q(t)
-  = \ln c_k + \ln \left( t + t_{k0} \right).
-  $$
-  Differentiating this with respect to $t$ yields
-  Eq. \eqref{eq:alpha_invtk}.
-}
-%
 Such a solution is possible only if
 all nontrivial eigenvalues are the same
-%
 \begin{equation}
@@
 \end{equation}
-%
 Then, by Eq. \eqref{eq:tk0},
@@
 for $k \ge 2$.
-%
 The asymptotic error is then
-%
 \begin{align}
-  \Err_A(T)
+  E_A
   =
@@
 \end{align}
-%
 If we further assume $\lambda_2 = \cdots = \lambda_n = 1$,
-%the optimal schedule given by
 Eq. \eqref{eq:alpha_invtk}
 recovers Eq. \eqref{eq:alpha_invt1}.
-%
 
@@
 without changing the nature of the updating scheme.
-%
 Note, an updating matrix
@@
 is a multiple of the identity matrix.
-%
 Thus,
@@
 the single-bin scheme, is most efficient
-in asymptotic convergence.
-%if no eigenvalue $\lambda_k$ is zero.
-%
+in terms of asymptotic convergence.
 In Appendix \ref{sec:upperbound},
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-that ignore short-wavelength eigenmodes,
+that ignore short-wave eigenmodes,
 which are typically local noise.
-%
 Mathematically, these updating schemes will have
@@
 the ignored modes are absent from the PMF.
-%
 For example,
-%if we can assume
 by assuming
@@
 such that the eigenmodes can be limited to $k \le K$,
-%
 we can construct an updating scheme with
@@
 $$
-%
 Then,
@@
 Eq. \eqref{eq:error_asym_singlebin} to $K$:
-%
-%Then,
-%Eq. \eqref{eq:error_singlebin} is changed to
-%
-\begin{equation}
-  \Err_A(T)
+\begin{equation}
+  E_A
   =
@@
     \Gamma_k.
-%\notag
 \label{eq:error_asym_sinc}
 \end{equation}
-%
-The optimal schedule of these schemes
-remains to be the inverse-time schedule
-given by Eq. \eqref{eq:alpha_invt1}.
-%
+The optimal schedule for these types of schemes
+is given by the inverse-time scheme (Eq. \eqref{eq:alpha_invt1}).
 Below, we construct these schemes for
 the homogeneous case.
-%
-The definition of the cutoff, $K$, will be modified
+The definition of the cutoff $K$ will be modified
 for mathematical convenience.
@@
 For a periodic variable, we demand
-%
 $$
@@
 for $K \ge 1$.
-%
 Then by using
@@
 \notag
-%\label{eq:mu_sinc_pbc}
-\end{equation}
-\note{Derivation.
-$$
-\begin{aligned}
-\mu_l
-&=
-\frac 1 n \sum_{k = 0}^{n-1} \lambda_{k+1} \cos \frac{ k \, l \, 2 \, \pi } { n }
-\\
-&=
-\frac{1}{n}
-\left(
-  1 +
-  \sum_{k=1}^K
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-  +
-  \sum_{k=n-K}^{n-1}
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-\right)
-\\
-&=
-\frac 1 n
-\sum_{k=-K}^K
-\cos \frac { k \, l \, 2 \, \pi } { n }
-=
-  \frac{
-    \sin
-    \dfrac{ (2 K + 1) \, l \, \pi }
-         {              n        }
-  }
-  {
-    n \, \sin \dfrac{ l \, \pi } { n }
-  }
-.
-\end{aligned}
-$$
-}
+\end{equation}
 
@@
 \end{equation}
-\note{Derivation.
-$$
-\begin{aligned}
-  \lambda_{n+1}
-  &=
-  (-1)^{n-1}
-  \left[
-    \lambda_1
-    - 2 \, \lambda_2
-    + 2 \, \lambda_3 - \cdots
-    + (-1)^{K} 2 \, \lambda_{K+1})
-  \right]
-  \\
-  &=
-  \begin{dcases}
-    (-1)^{n-1} & K \mathrm{\; even,} \\
-    (-1)^n     & K \mathrm{\; odd.}
-  \end{dcases}
-\end{aligned}
-$$
-So
-$$
-\begin{aligned}
-  \mu_l
-  &=
-  \frac{1}{2\,n}
-  \left[
-    1 +
-    2 \sum_{k=2}^{K+1}
-    \cos \frac { (k - 1) \, l \pi } { n }
-    +
-    (-1)^{n+K-1} (-1)^l
-  \right]
-  \\
-  &=
-  \frac{1}{2\,n}
-  \left[
-    (-1)^{n+K+l-1}
-    +
-    \sum_{k=-K}^{K}
-    \cos \frac { k \, l \pi } { n }
-  \right]
-  .
-\end{aligned}
-$$
-}%
 
@@
     compared to a Gaussian with a similar width.
-    %
     The lines are a guide to the eyes.
-    Inset. Updating matrix in the non-periodic case.
-    %
-    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h} for panel (a),
-      and the functions \texttt{savewin()}
-      and \texttt{savewinmat()}
-      in \texttt{prog/invt.h} for panel (b),
-      The data were saved under the directory \texttt{data/sinc}.
-      See the \texttt{make} object \texttt{sincsingle}
-      in the \texttt{Makefile} there.
-    }%
+    Inset. Updating matrix associated with the non-periodic case.
   }
@@
 with a similar width, $\sigma$,
-%
 \begin{equation}
@@
 \end{equation}
-%
-%where $g = 2$ for the periodic case
-%or $g = 1$ otherwise.
-%
-The kernel in the non-periodic case is a sawtooth 
-because of the term, $(-1)^{n+K+l-1}$,
+The kernel in the non-periodic case is sawtooth
+because of the term $(-1)^{n+K+l-1}$
 in Eq. \eqref{eq:mu_sinc_refl}.
-%
 However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
@@
 as the above oscillatory term for $l = i - j$
-is canceled by the term for either $l = i + j - 1$
+is canceled by the terms associated with $l = i + j - 1$
 or $l = i + j - 2 \, n - 1$
-with the opposite sign.
+with opposite sign.
 
@@
 
-In the above section,
-we have focused on
+In the above section, we have focused on
 the asymptotic error.
-%
 By including the residual error
@@
 which is, according to Eq. \eqref{eq:half_alpha0},
-%
-%%Assuming the system is initially equilibrated
-%%under a constant updating magnitude,
-%%$a_0$,
-%Assuming Eq. \eqref{eq:error_res1} with
-%$$
-%q(T)
-%=
-%\int_0^T \frac{ dt } { t + t_0 }
-%=
-%\ln \frac{ T + t_0 } { t_0 }.
-%$$
-%%
-%we have, from Eqs. \eqref{eq:error_split}
-%and \eqref{eq:error_asym_sinc},
-%%
-%\begin{equation}
-%  \Err
-%  =
-%  \Err_R + \Err_A
-%  =
-%  \frac { \frac 1 2 a_0 \, t_0^2 + T }
-%        {          ( T + t_0 )^2          }
-%  \sum_{ k = 2 }^K \Gamma_k
-%  .
-%\notag
-%%\label{eq:error_sinc0}
-%\end{equation}
-%%
-%The minimum is reached at
-%
 \begin{equation}
@@
 \end{equation}
-%
-and from Eqs. \eqref{eq:error_split},
+From Eqs. \eqref{eq:error_split}, the total error is
 \eqref{eq:error_res1},
@@
 \eqref{eq:error_asym_sinc},
-%
-\begin{equation}
-  \Err(T)
+\begin{equation}
+  E
   =
@@
 \end{equation}
-%
 This error decreases inversely with increasing simulation length,
 as one would expect from an equilibrium FDS simulation.
-%In Appendix \ref{sec:equilerr},
-%we show that the above error tends
-%to the error of the ideal equilibrium FDS simulation
-%in the asymptotic limit.
-
-%
-%For example, consider a system that is initially
-%prepared under a fixed $a_0$
-%at the beginning of the schedule $\alpha(t)$.
-%%
-%Then for shorter times, $T \approx 0$,
-%the error, dominated by the residual error
-%given by Eq. \eqref{eq:error_res1},
-%is clearly smaller if some eigenvalues
-%are less than $1.0$ (the single-bin case).
-%
-%In practice, real simulations
-%may have long achieved the desired precision
-%before reaching the asymptotic regime,
-%especially if the updating matrix
-%has some $\lambda_k$'s close to $0$.
 
@@
 described in Sec. \ref{sec:results_system}.
-%
 We verify the method of computing errors
-%using an inverse-time schedule
-in Sec. \ref{sec:results_invt}
-%
+in Sec. \ref{sec:results_invt}.
 and study the optimal schedule from
@@
 in Sec. \ref{sec:results_optschedule}.
-%
 Finally, we compare the Gaussian and
@@
 and $n = 100$.
-%
 Both the target and intrinsic distributions
@@
 $p^*_i = p_i = 1/n$,
-along with the distribution, $\rho_i$, in Eqs.
+along with the distribution $\rho_i$ in Eqs.
 \eqref{eq:vbar_def} and \eqref{eq:error_sum}.
-%
-%We wish to see how
-%the schedule, $\alpha(t)$,
-%affects the fluctuating error
-%of the bias potential
-%at the end of the simulation.
-%
 Initially,
@@
 $a_0 = 10^{-4}$.
-%
-This equilibration process models
-a stage of the WL algorithm or
-a regular metadynamics
-with a fixed updating magnitude.
-%
 Although the bias potential is correct initially,
@@
 at the end of the equilibration.
-%
 Our aim is to see how well
@@
 at the end of the testing schedule.
-%
 After the equilibration,
@@
 for $T = 10^8$ steps.
-%
 We repeat the above process $1000$ times (unless specified otherwise)
-and report the averaged results.
-%specified by either
-%Eq. \eqref{eq:q_opt}
-%or
-%Eq. \eqref{eq:alpha_invtlambda}.
+and average the results.
 
@@
 the kernel was truncated at
-$b = \min\{10 \, \sigma, n - 1\}$ bins
+$b = \min\{10 \, \sigma, n - 1\}$
 and numerically stabilized
@@
 
-%Since the error depends on
-%the underlying sampling process,
 We tested two sampling processes.
-%a global one and a local one.
-%
 Both are based on
@@
   landau_binder}:
-%
 in each step, a new bin index $j$ is proposed
 and then accepted with probability
-%
 $
@@
 uniformly out of the $n$ bins.
-%
 In the second or local sampling process,
@@
 with equal probability $1/2$.
-%
 Asymptotically,
@@
 in the theoretical predictions.
-%
 
@@
 
-Using the model system above, we
-tested our analysis of computing errors
+Using the model system above, we tested our analysis of computing errors
 on the modified inverse-time schedule,
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $\lambda$ is a free parameter,
 and $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
-%and $t_0 = 2/a_0$.
-%
 The error of this schedule
@@
 using the single-bin updating scheme
-and the triple-bin updating scheme with $\mu_1 = 0.24$.
-%
-%For each value of $\lambda$,
-%we ran $1000$ independent simulations
-%of $10^8$ steps, and averaged the results.
-%
+and the triple-bin (bandpass) updating scheme with $\mu_1 = 0.24$.
 As shown in Fig. \ref{fig:errinvt},
 the errors from the simulations
-agreed well with those from the theory.
+agreed well those from the theory.
 
@@
     \label{fig:errinvt}
-    Error, $\Err$, versus the parameter, $\lambda$,
+    Error $E$ versus the parameter $\lambda$
     for the modified inverse-time schedule,
     Eq. \eqref{eq:alpha_invtlambda}.
-    %
     The points are results from averaging over $1000$ independent runs;
     the curves are theoretical predictions.
-    %
     For the single-bin (WL) updating scheme,
@@
     while it is not necessarily the case for a general updating scheme.
-    %
-    \note{The figure was produced by \texttt{doc/fig/errinvt.gp}.
-      The data were produced by the script \texttt{data/invtcscan.py},
-      and saved under \texttt{data/invt}
-      (see the \texttt{Makefile} there).
-    }%
   }
@@
 for both global and local sampling processes.
-%
 The sampling process affected the error function
 only by a multiplicative constant,
-%The error functions of the two sampling processes
-%differed only by a multiplicative constant,
 in agreement with the discussion in
 Appendix \ref{sec:invt_singlebin}.
-%
 For the triple-bin updating scheme,
@@
 for the global sampling process,
-although it stayed around $1.0$
+although it remained around $1.0$
 for the local sampling process,
-also in agreement with the discussion
-in Appendix \ref{sec:invt_nn}.
-%
-Thus, the inverse-time schedule,
+as discussed in Appendix \ref{sec:invt_nn}.
+Thus, the inverse-time formula,
 Eq. \eqref{eq:alpha_invt1},
 is optimal for the single-bin updating scheme,
-but not necessarily so for a general multiple-bin updating scheme.
-%
+but not necessarily so for a general, multi-bin updating scheme.
 
@@
 To demonstrate the optimal schedule from Eq. \eqref{eq:q_opt},
-we used the Gaussian (with $\sigma = 10$ bins) updating scheme
+we used the Gaussian (with $\sigma = 10$) updating scheme
 as an example.
-%
-This updating scheme is typically used in metadynamics simulations.
-%
 In Table \ref{tab:error_Gaussian},
-we showed that the optimal schedule
+we show that the optimal schedule
 produced smaller errors than
-the inverse-time schedules.
-%
-However, the optimal schedules
+the inverse-time schedule.
+However, the optimal schedule
 decayed more slowly than
@@
 as shown in Fig. \ref{fig:optacmp}.
-%
 The impeded decay of the optimal schedule
@@
 damp out the short-wavelength eigenmodes
-(i.e. local noise).
-%
-As shown in Fig. \ref{fig:xerr},
-the inverse-time schedule was effective in error reduction
-only for the first few eigenmodes,
-whose eigenvalues were around $1.0$.
-%
-We can, however, ignore these local eigenmodes
-by explicitly filtering them out from the final bias potential,
-and truncating the error sum in Eq. \eqref{eq:error_sum}
+(i.e., local noise),
+as shown in Fig. \ref{fig:xerr}.
+If, however, we could ignore these local eigenmodes
+(by explicitly filtering
+them from the biased potential),
+we could modify the optimal schedule
+(Eq. \eqref{eq:q_opt})
+and limit the sum
 to the first few eigenmodes.
-%
-Then, the corresponding optimal schedule
-from Eq. \eqref{eq:q_opt} that minimizes the truncated error
-became similar to the inverse-time schedule,
+The resulting schedule
+would become similar to the inverse-time schedule,
 Eq. \eqref{eq:alpha_invt1},
-as exemplified in Fig. \ref{fig:optacmp}.
-%
-\note{A more thorough approach is
-to use the bandpass updating scheme
-to be discussed later.
-}
+as shown in Fig. \ref{fig:optacmp}.
 
@@
     $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
-    \note{
-      The data were collected in
-      \texttt{data/tab1} by running
-      the program \texttt{invt}.
-    }%
   }
@@
     with $\sigma = 10$.
-    %
     The mode-limited schedules were produced by
     limiting the sum in Eq. \eqref{eq:q_opt}
-    to the first four terms,
-    or the four longest wavelength modes.
-    %
+    to the first four terms.
     They became similar to the inverse-time schedule,
     Eq. \eqref{eq:alpha_invt1}.
-    %
-    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h}.
-      The data were saved under the directory \texttt{data/opta}.
-    }%
   }
@@
     and the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
-    %
     The latter schedule failed to damp out the errors
@@
     with $k \ge 5$.
-    %
     For longer simulations,
     the error components from the optimal schedule
-    would roughly maintain a relatively flat profile,
+    would roughly maintain the relatively flat profile,
     but meet the initial error curve
@@
     (cf. Appendix \ref{sec:upperbound}).
-    %
-    \note{This figure was produced by \texttt{doc/fig/xerr.gp}.
-      The data were saved under the directory \texttt{data/xerr}.
-    }%
   }
@@
 
-An interesting prediction from the theory is Eq. \eqref{eq:half_alpha0},
-which states that the optimal schedule should always start from
-half of the magnitude from the previous equilibration run.
-%
-To verify this result,
+To verify Eq. \eqref{eq:half_alpha0},
 we computed the optimal schedules
-for different values of $q(T)$
-for the Gaussian updating scheme,
+for different values of $q(T)$,
 and plotted the final error
 against the initial updating magnitude, $\alpha(0)$.
-%
 As shown in Fig. \ref{fig:erriascan},
-the minimal error was indeed reached at
-$\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
-although the dependence on $\alpha(0)$
-is generally weak.
-%
-The weak dependence is best illustrated in
-the case of single-bin updating scheme,
-as the initial updating magnitude only
-affects the schedule by a shift of the origin of
-simulation time in the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1}.
+the minimal error was reached at
+$\alpha(0) = a_0/2 = 5 \times 10^{-5}$.
 
@@
     and $a_0 = 10^{-4}$.
-    %
     The points were obtained from averaging results
@@
     the curves are theoretical predictions.
-    %
     The minimal should always occur at
@@
     as predicted by Eq. \eqref{eq:half_alpha0}.
-    %
-    \note{This figure was produced by
-      \texttt{doc/fig/erriascan.gp}.
-      The data were saved under
-      \texttt{data/iascan}.
-    }%
   }
@@
 and other parameters.
-%
 As an example,
@@
 with different widths, $\sigma$.
-%
-In Fig. \ref{fig:errsigscan},
-we present the errors, $\Err(T)$,
-normalized by the shifted simulation length, $T+t_0$,
-so that they could
-stay constant in the limit of single-bin updating scheme
-($\sigma \to 0$).
-%
 As shown in Fig. \ref{fig:errsigscan}(a)
@@
 for the global sampling process,
-%
 although
@@
 as the simulation lengthens.
-%
 However,
@@
 in agreement with Eq. \eqref{eq:error_sinc}.
-%
 The bandpass updating schemes
@@
     \label{fig:errsigscan}
-    Normalized error, $(T + t_0) \, \Err(T)$,
+    Normalized error, $(T + t_0) \, E$,
     versus the width, $\sigma$,
@@
     with $t_0$ given by Eq. \eqref{eq:t0_sinc}.
-    %
     In the latter case,
@@
     $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
-    %
     The points were obtained from averaging results
@@
     the curves are theoretical predictions.
-    %
-    \note{This figure was produced by
-      \texttt{doc/fig/errsigscan.gp}.
-      The data were saved under
-      \texttt{data/scan}.
-    }%
   }
@@
 for a general class of free energy calculation methods,
-which we refer to as adaptive flat-distribution-sampling (FDS) methods
+which we refer to as adaptive flat-distribution sampling (FDS) methods,
 that include the Wang-Landau (WL) algorithm and metadynamics.
-%
 Adaptive FDS methods
@@
 and thereby sample a flat distribution.
-%
 The optimal schedule delivers the fastest convergence
@@
 of computing the PMF.
-%They differ by the updating schemes,
-%which specify how the bias potential is modified
-%in each updating step.
-%%
-%In the single-bin or WL case,
-%the update is limited to the current bin;
-%while metadynamics,
-%the bias potential of a neighborhood of the current one
-%is updated with relative weight specified by a Gaussian window.
 
@@
 taking a Gaussian shape in the metadynamics case.
-%
 More generally,
@@
 by an updating matrix.
-%
 Each column of the matrix gives the normalized
 window function for the bin corresponding to the column.
-%
 Our method computes the optimal schedule from
@@
 of the eigenmodes.
-%
 For the single-bin (WL) scheme,
-the optimal schedule from our method
+the optimal schedule
 is given by the inverse-time formula,
 Eq. \eqref{eq:alpha_invt1}.
-%
 For a general multiple-bin scheme,
@@
 of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
-%
 In this case,
@@
 can be sensitive to the simulation length.
-%
 The initial updating magnitude is always
 half of the previous equilibrium value.
-% the value used during equilibration.
 
@@
 about the smoothness of the PMF.
-%
-However, if some eigenmodes
-correspond to local noise
-that is expected to be absent from the PMF,
+However, if certain eigenmodes associated with local noise
+are expected to vanish in the PMF,
 we can construct a
@@
 that are also optimal for asymptotic convergence.
-%
 The optimal schedule of these bandpass schemes
@@
 and hence more expensive.
-%
 However, the performance gain
@@
 with a local sampling process such as MD.
-%
-\note{
-One way to improve the Gaussian updating schemes
-is to use it with a hard cutoff of eigenmodes
-(as we did for Fig. \ref{fig:optacmp}).
-%
-Then the optimal schedule
-would be similar to the inverse-time one,
-and the error, with the short-wavelength eigenmodes filtered,
-would be close to that of the bandpass scheme.
-}
-
-The analysis in this study has two limitations.
-%
-First, we have ignored the \emph{systematic}
-error\cite{zhou2005, morozov2007, zhou2008}
+
+In this analysis, we have ignored the \emph{systematic} error\cite{zhou2005}
 caused by the updates in adaptive FDS methods.
-%
-This is valid in the asymptotic regime:
-when the updating magnitude is small
-and the sampling is a quasi-equilibrium one\cite{
-  zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
-the random error becomes more important than the systematic one,
-because the former is roughly proportional to
-the square root of the updating magnitude\cite{
-  zhou2005, morozov2007, zhou2008, bussi2006}
-whereas the latter is proportional to
-the magnitude itself\cite{morozov2007}.
-%
-%Our method of analysis is most useful for long simulations
-%to achieve relatively precise results.
-%
-Second, by using the white noise approximation,
-we have represented the autocorrelation function
+This is a valid assumption in the asymptotic regime,
+where the random error is much greater than the systematic one.
+This allows us to assume the sampling process
+to be a quasi-equilibrium one\cite{barducci2008, dama2014},
+and to represent the autocorrelation function
 of each eigenmode by a single number, $\Gamma_k$.
-%
 Although this assumption appeared to work well
@@
 of glassy systems.
-%Our immediate interest is to apply these relations to
-%problems containing aqueous mixtures of proteins and nucleic acids.
 
@@
 We can understand the optimality of the inverse-time schedule
-as a natural consequence of the stabilization of a runtime average
-as the sample size increases\cite{
-  marsili2006, barducci2008}.
-%
+as a result of the stabilization of a runtime average
+with respect to the increase of sample size\cite{marsili2006, barducci2008}.
 Consider a sample, $\{X^{(1)}, \dots, X^{(N)}\}$,
 with a dynamically growing size.
-%
 The runtime average
-%
 \begin{equation*}
@@
 \end{equation*}
-%
 can be written as a recurrence relation regarding
-the addition of the last data point,
-%
+the addition of the last data point.
 \begin{align}
@@
 \end{align}
-%
 If we interpret $X^{(N)} - {\bar X}^{(N-1)}$
@@
 
-The structural similarity between the recurrence relation,
-Eq. \eqref{eq:av_recur},
-and the WL updating scheme, Eq. \eqref{eq:wl_update},
-suggests that the bias potential may represent
-a runtime average in a particular case.
-%
-To establish a mapping between the two,
-let us now consider a long equilibrium FDS simulation
-that consists of a sequence of sub-simulations
+Let us now consider a long equilibrium FDS simulation
+that consists of blocks of sub-simulations
 of $L$ steps.
-%
 If $X^{(N)}$ is
@@
 from the $N$th sub-simulation,
-then ${\bar X}^{(N)}$ means the average bias potential,
+then ${\bar X}^{(N)}$ stands for the average bias potential,
 ${\bar V}_i^{(N)}$,
-from the first $N$ sub-simulations.
-%
-Since this average, ${\bar V}_i^{(N)}$,
-gives the best estimate at the end of
-the first $N$ sub-simulations,
-we should ideally use it as the bias potential
-for the next sub-simulation,
-until it is further updated by Eq. \eqref{eq:av_recur}
-at the end of the next sub-simulation.
-%
+from the first $N$ sub-simulations,
+which is the best estimate at the end of
+the first $N$ sub-simulations.
+If we wish to always use this best estimate, ${\bar V}_i^{(N-1)}$,
+from the previous $N-1$ sub-simulations
+as the bias potential for the $N$th sub-simulations,
+then this estimate can be updated according to Eq. \eqref{eq:av_recur}.
 Assuming that the normalized histogram, $H_i^{(N)}$,
-from the $N$th sub-simulation is sufficiently precise,
-Eq. \eqref{eq:vcorr_equil}
-\big[with
-$V_i^\mathrm{corr} \to V_i^{(N)}$, $V_i \to {\bar V}_i^{(N-1)}$\big]
-shows that the independent
+from the $N$th sub-simulations is sufficiently precise,
+Eq. \eqref{eq:vcorr_equil} shows that the independent
 correction from the $N$th sub-simulation is
-%
 \begin{equation*}
@@
 \end{equation*}
-%
-where the ``$-1$'' on the right-hand side
-can be ignored as a uniform shift
-that does not affect the distribution.
-%
+where, the $-1$ on the right-hand side
+can be ignored as a uniform shift that
+does not affect the distribution.
 Thus, we may rewrite Eq. \eqref{eq:av_recur} as
@@
 \end{align}
-%
 where $h_i^{(N)}(l)$ stands for the instantaneous histogram
 at step $l$ of sub-simulation $N$,
-and its average is
+and it is related to the average $H_i^{(N)}$ as
 $H_i^{(N)} \equiv \frac{1}{L} \sum_{l=1}^L h_i^{(N)}(l)$.
-%
-Now Eq. \eqref{eq:vbar_update} shows that the best estimate
-of the bias potential from the runtime average can be computed
+Thus, Eq. \eqref{eq:vbar_update} shows that the best estimate
+or the runtime average of the bias potential can be computed
 through a step-by-step incremental updating scheme that is equivalent to
 the WL algorithm, Eq. \eqref{eq:wl_update},
-with the updating magnitude being
-%
+with the magnitude being
 \begin{equation}
@@
 \end{equation}
-%
 In the long time limit,
@@
 by the product of the number of sub-simulations
-and the number of steps in each sub-simulation, or $N \, L$,
-%
+and the number of steps in each sub-simulation, or $NL$,
 and Eq. \eqref{eq:alpha_block} recovers Eq. \eqref{eq:alpha_invt}.
 
-The above demonstration implies several connections
-between the WL algorithm under the optimal inverse-time schedule
-with a few related FDS methods.
-%
+The above demonstration has several practical implications.
 First,
-it establishes an efficiency equivalence
-between an adaptive WL simulation under the inverse-time schedule
-and the ideal equilibrium FDS counterpart
-with a nearly perfect bias potential.
-%
-The bias potentials from the two approaches
-would have similar precision at long times.
-%
-Thus, it is unnecessary to convert
-a sufficiently convergent adaptive FDS simulation
-to the corresponding equilibirium FDS method
-with a fixed bias potential just for better precision
-of the final PMF.
+since the bias potential is often sufficiently precise
+in late stages of a WL simulation,
+some practitioners would then fix
+the bias potential and switch
+to the equilibrium FDS method
+to prevent possible future efficiency loss
+due to the adaptive updates.
+However, the above demonstration shows that
+this practice becomes unnecessary
+with the inverse-time schedule.
+That is,
+the WL algorithm with the inverse-time schedule
+produces a bias potential that is similar in precision to that from
+the ideal equilibrium simulation for long times.
 
 Second,
-it also relates a WL simulation under the inverse-time schedule
-to a time-averaged variant of the regular WL simulation.
-%
-In the latter, the updating magnitude is fixed,
-but the bias potential over the simulation coarse is averaged
-as the final result\cite{zhou2005}.
-%
-The above demonstration shows that the inverse-time schedule
-has already effectively implemented an implicit average
-so that no explicit averaging is needed in the former approaches.
-%
+an alternative way of improving the precision of the bias potential
+from a simulation using the WL algorithm or metadynamics
+with a fixed magnitude
+is to time average the bias potential\cite{zhou2005}.
+This practice becomes redundant with the inverse-time schedule
+because the averaging of the bias potential
+has been implicitly performed by the schedule.
 The inverse-time schedule approach is preferred, however,
@@
 it minimizes the systematic error caused by
-the on-the-fly updating scheme\cite{
-  zhou2005, zhou2008},
+the on-the-fly updating scheme\cite{zhou2005},
 which is ignored in this analysis.
@@
 it is possible to reformulate the WL algorithm
-under the inverse-time schedule
-as an explicit adaptive average method
+with the inverse-time schedule
 such that the bias potential is computed from
-a weighted runtime average of the histogram\cite{
-  marsili2006, barducci2008},
-which is analogous to the adaptive bias force method\cite{
-  darve2001, *darve2008}.
+a weighted runtime average of the histogram\cite{marsili2006, barducci2008},
+which is analogous to the adaptive bias force method\cite{darve2001, *darve2008}.
 
@@
 as an optimal limit of the stage-wise WL updating scheme.
-%
 Our calculation below will focus on the one-variable problem,
@@
 each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
-%
 We denote the number of steps in stage $s$ as $T_s$, and
 $\sum_{s=1}^S T_s = T$.
-%
-%\begin{equation}
-%  \sum_{s = 1}^S T_s = T
-%  ,
-%  \label{eq:sumT}
-%\end{equation}
-%
 We assume that
 the magnitude follows geometrically,
-%
 \begin{equation}
@@
 \end{equation}
-%
 with $0 < r < 1$.
-%
-%and that the system is properly equilibrated
-%under a constant magnitude initially.
-%
 We wish to find the best distribution of the simulation time
-into the $S$ stages
+across the $S$ stages
 as well as
@@
 $q(t) = \alpha_s \, t$, and
-%
 \begin{align}
@@
 \end{align}
-%
 where we have assumed
@@
 in the second step.
-%
 Our aim is to minimize the error
 of the final stage, $E_S$.
-%
-To do so, let us consider shifting simulation time
-by $\delta T$ from
+To do so, let us consider moving simulation by $\delta T$ time from
 stage $s+1$ to stage $s$,
 then $E_{s+1}$ is affected as
-%
 \begin{align*}
@@
 \end{align*}
-%
 Since this variation does not directly affect the error of later stages,
 the above expression must be zero to minimize the final error, i.e,
-%
 \begin{align*}
@@
   -\frac{ \partial E_s } { \partial T_s }
-  %\\
-  %&=
-  %2 \, \alpha_s \, e^{-2 \, \alpha_s \, T_s}
-  %\left(
-  %  E_{s-1} - \frac{ \Gamma \, \alpha_s } { 2 }
-  %\right)
-  %\\
   =
@@
 \end{align*}
-%
 where we have used Eq. \eqref{eq:errorstage}
 in the second step.
-%
 Thus,
@@
 \end{equation}
-%
 Using Eqs. \eqref{eq:Es} and \eqref{eq:alphas}
 in Eq. \eqref{eq:errorstage}, we get
-%
 \begin{equation}
@@
 \end{equation}
-%
 
 To compute the optimal schedule,
-we map the simulation time, $t$, to the stage number, $s$, as
-%
+we map the simulation time $t$ to the stage number $s$ as
 \begin{align*}
@@
 and we have used Eq. \eqref{eq:exp_r} in the last step.
-%
 Thus,
@@
 \end{align}
-%
 which is minimized as $r \to 1$
 for a large $T$
-(at this point, $\gamma \to 1$).
-%
-\note{
-  For a large $T$, $E_S \approx \frac{\Gamma}{2T} \frac{1+r}{\gamma}$.
-  Define $\delta = 1 - r$,
-  we have $1/\gamma = 1+ \delta/2 + \delta^2/3 + \cdots$,
-  and
-  $$
-  \frac{1+r}{\gamma}
-  =
-  (2 - \delta)
-  \left(
-  1 + \frac{\delta}{2} + \frac{\delta^2}{3} + \cdots
-  \right)
-  =
-  2 + \sum_{n=1}^\infty \frac{n-1}{n(n+1)}\delta^n \ge 2
-  .
-  $$
-}
-%
-If we identify the parameter, $\alpha_0$,
-as the initial updating magnitude, $\alpha(t = 0)$,
+(at this point $\gamma \to 1$).
+If we identify the parameter $\alpha_0$ as the initial updating magnitude $\alpha(t = 0)$,
 then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
 and the optimal $t_0 = \alpha_0^{-1}$.
-%
 Thus, in this limit,
@@
 First, if the sampling is perfect,
-%Eq. \eqref{eq:kappa_delta} is exact, and
-%
-%\begin{equation}
-%  \Gamma_k = 1.
-%\label{eq:Gamma_perfect}
-%\end{equation}
-%
-%To see this, we note that
-the bin index, $i$, is an independent random variable
+the bin index $i$ is an independent random variable
 at each time step $t$.
-%
 Further, in the asymptotic limit,
@@
 we find from Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def} that
+%
 \begin{equation*}
@@
   \right\rangle
-  %&=
-  %\left\langle
-  %  \left(
-  %    \frac{ h_i(t) } { p_i }
-  %    -
-  %    \sum_{ r = 1 }^n
-  %      \rho_r \frac{ h_r(t) } { p_r }
-  %  \right)
-  %\right.
-  %\notag \\
-  %&\hphantom{==}\cdot
-  %\left.
-  %  \left(
-  %    \frac{ h_j(t') } { p_i }
-  %    -
-  %    \sum_{ s = 1 }^n
-  %      \rho_s \frac{ h_s(t') } { p_s }
-  %  \right)
-  %\right\rangle
-  %\notag
-  %\\
   &=
@@
 \end{align}
-%
-%$\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
-%in the continuous limit.
-%
-\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
-  with $t \ne t'$ (going back to the discrete time case),
-  $$
-  \begin{aligned}
-  \left\langle
-    \xi_i(t) \, \xi_j(t')
-  \right\rangle
-  &=
-  \left\langle
-    \xi_i(t)
-  \right\rangle
-  \left\langle
-    \xi_j(t')
-  \right\rangle
-  =
-  0.
-  \end{aligned}
-  $$
-  Next, for $t = t'$, we have
-  $$
-  \langle
-    h_i(t) \, h_j(t)
-  \rangle
-  = \delta_{ij} \, p_i
-  ,
-  $$
-  which leads to Eq. \eqref{eq:zz_perfect}.
-}
-%
 Thus,
 for $k \ge 2$, we get
-%
 \begin{equation}
@@
   \langle \xi_i(t) \, \xi_j(t') \rangle
-  %
   =
@@
 \notag
-%\label{eq:kappa_perfect}
-\end{equation}
-%
+\end{equation}
 where we have used
@@
 \eqref{eq:zz_perfect}.
-%
 Thus, Eq. \eqref{eq:kappa_delta} is exact,
@@
 $\Gamma_k = \sum_{ i = 1}^n \phi_{ki}^2 / p_i$.
-\note{Proof.%Eq. \eqref{eq:kappa_perfect},
-  $$
-  \begin{aligned}
-  \kappa_k(t - t')
-  &=
-  \sum_{i,j = 1}^n
-  \phi_{ki}
-  \phi_{kj}
-  \langle \xi_i(t) \, \xi_j(t') \rangle
-  \\
-  %
-  &=
-  \sum_{i,j = 1}^n
-  \phi_{ki}
-  \phi_{kj}
-  \left(
-    \frac{ \delta_{ij} } { p_i }
-    -
-    \frac{ \rho_i } { p_i }
-    -
-    \frac{ \rho_j } { p_j }
-    +
-    \sum_{r = 1}^n
-    \frac{ \rho_r^2 } { p_r }
-  \right) \,
-  \, \delta_{t, t'}
-  .
-  \end{aligned}
-  $$
-  where,
-  we have used
-  Eq. \eqref{eq:theta_def}
-  on the first line,
-  %
-  Eq. \eqref{eq:zz_perfect}
-  on the second line.
-  %
-  With Eq. \eqref{eq:ortho1},
-  only the first term in the parentheses survives.
-}
 In particular, if $p_i = \rho_i$, then, by
@@
 one of the two neighboring bins
-with equal probability, $1/2$.
-%
+with equal probability $1/2$.
 The transition matrix is
-%
 \begin{equation}
@@
 \notag
-%\label{eq:T_nn}
-\end{equation}
-%
+\end{equation}
 where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
-%
 For many updating schemes
@@
 with $\rho_i = p_i = 1/n$.
-%but not the eigenvalues.
-%
 The eigenvalues of $\mathbf A$ are
@@
 $g = 2$ for a periodic variable, or $1$ otherwise.
-%Note that these $\Lambda_k$'s are
-%not to be confused with the eigenvalues of $\mathbf w$,
-%$\lambda_k$'s.
-%
 In this case,
-the autocorrelation function, $\kappa_k(t)$,
+the autocorrelation function $\kappa_k(t)$
 decays as $\kappa_k(0) \, \Lambda_k^t$,
@@
 Thus,
-%$\Gamma_k$ is roughly twice the autocorrelation time, and
+%
 from Eq. \eqref{eq:Gamma_sum}, we get
-%
 \begin{equation}
@@
 
-%To use Eq. \eqref{eq:q_opt} in a real simulation,
-%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
-%
 For an unknown system,
@@
 $\alpha(t) = a_0$.
-%
-%A practical difficulty is that
 However,
@@
 of the bias potential, $V_i$.
-%
 To circumvent the problem, we
 modify Eq. \eqref{eq:y2_eql} as
-%
 \begin{equation}
@@
 and
-%
 \begin{align*}
@@
   \phi_{k i} \, (V_i - \bar V),
-%\label{eq:Y_def}
 \end{align*}
-%
 with
@@
 $\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
-%
 Thus, the two share the same variance
 under time averaging.
-%
 It follows that the corresponding linear combinations,
 $Y_k$ and $y_k$, would also share the same variance.
-%
 But since the average of $y_k$ is zero,
@@
 we reach Eq. \eqref{eq:varY}.
-
-\note{In practice, Eq. \eqref{eq:varY}
-in inapplicable to zero eigenvalues, $\lambda_k$'s.
-%
-However, these eigenmodes can be ignored
-as they do not contribute to the Lagrangian defined by
-Eq. \eqref{eq:Lagrangian_const}.
-}
 
@@
 the one-step sampling process.
-%
 As shown in Fig. \ref{fig:gamcmp},
@@
     and an MD simulation.
-    %
     Each simulation was ran for $10^{10}$ steps.
-    %
     The bin size $\Delta z = 2\, \pi /n$ with $n = 100$.
-    %
     For the MD simulation,
     the step size $\Delta t_\mathrm{MD} = 0.005$.
-    %
     The velocity rescaling thermostat\cite{bussi2007} was
@@
     every MD step with viscosity $20$.
-    %
     The reflective boundary condition was used.
-    %
-    \note{The figure was produced by \texttt{doc/fig/gamcmp.gp}.
-      The data were saved under \texttt{data/gamcmp}
-      (see the \texttt{Makefile} there).
-    }%
   }
@@
 of a given adaptive FDS simulation.
-%
 Basically, we need to extract
@@
 as shown in Fig. \ref{fig:vardep}.
-%
-%The dependence of variables are shown in Fig. \ref{fig:vardep}.
-%
 While the updating scheme alone
@@
 and the underlying sampling process (MC or MD).
-%
 By using the technique in Sec. \ref{sec:Gamma_measure}
@@
 
-    \node[draw, text width={width("Error,")}]
-    (err) at (7, 0) {Error, $\Err(T)$};
+  \node[draw, text width={width("Error, $\Err$")}]
+    (err) at (7, 0) {Error, $\Err$};
 
@@
   \draw[->, >=latex, thick]
-    %(M1) edge[bend left=70] (N1);
     (R1) |- (R2) -| (R3);
@@
 Run a preliminary adaptive FDS simulation
-%using the updating scheme
 with a constant updating magnitude, $a_0$.
@@
 \item
-  Compute the error $\Err(T)$ from
+Compute the error $E$ from
 Eqs. \eqref{eq:error_split},
@@
 
-%A practical problem of using homogeneous updating scheme
-%is the following.
-%
 A homogeneous updating scheme
@@
 is not necessarily stable,
-i.e. some eigenvalues can be negative.
-%
-%
-%\note{
-  %For example,
-  %the triple-bin updating scheme
-  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
-  %
-%}%
-Even
-the Gaussian updating scheme, which is stable in the continuous limit,
-can be unstable upon discretization
-during which the updating kernel is truncated.
-%
+i.e., some eigenvalues can be negative.
 Here we show how to minimally modify
@@
 to stabilize it.
-%
-\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
-}
 
@@
 Eq. \eqref{eq:wband_eigenvalue}.
-%
 By replacing negative eigenvalues with zeros and
@@
 which is stable by construction.
-%
 However, the new kernel is widened
 as $\mu_l$ may be nonzero for $l > b$.
-%
 If we truncate the new kernel at $\mu_b$,
@@
 tend to be smaller in magnitude.
-%
 Thus, to stabilize the updating kernel
@@
 
-%
 \begin{enumerate}
@@
     Given an updating kernel, $\mu_0, \dots, \mu_b$,
-    compute the eigenvalues,
-    $\lambda_1, \dots, \lambda_n$,
+    compute the eigenvalues
+    $\lambda_1, \dots, \lambda_n$
     from Eq. \eqref{eq:wband_eigenvalue}.
@@
 it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
-%
 \begin{equation}
@@
 \end{equation}
-%
 and the asymptotic error from Eq. \eqref{eq:error_asym2}
@@
 \end{equation}
-%
 It is readily seen that this bound is always
 greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
-%
 $$
@@
 updating schemes with a set of very different eigenvalues.
-%
-%First, observe that the summand of Eq. \eqref{eq:mass_func},
-%$\Gamma_k \, \left( \lambda_k e^{-\lambda_k \, Q} \right)^2$
-%is a function that peaks at $\lambda_k = 1/Q$,
-%so that the sum is dominated by the term at $k = k(Q)$ with
-%%
-%\begin{equation}
-%  \lambda_{k(Q)}
-%  \approx
-%  \frac 1 Q
-%  .
-%  \label{eq:kQ_def}
-%\end{equation}
-%
-%Thus,
 First, we have for any mode $k^*$ $(k^* \ge 2)$,
-%
 \begin{align}
@@
   \ge
-  %\approx
   \sqrt{ \Gamma_{ k^* } } \,
@@
 \end{align}
-%
 which would approach an equality if the $k^*$ mode
 dominates the mass function, $M(Q)$,
-as defined in Eq. \eqref{eq:mass_func}.
+defined in Eq. \eqref{eq:mass_func}.
 %
@@
 we have from Eq. \eqref{eq:error_asym_Lagrangian}
-%
 \begin{align}
@@
 \end{align}
-%
 where we have
@@
 and extended the upper limit to infinity on the last line.
-%
 For a set of well-separated eigenvalues,
-$\lambda_k$ depends on the mode index, $k$,
+$\lambda_k$ depends on the mode index $k$
 much more critically than $\Gamma_k$.
-%
 Thus, the tightest upper bound
@@
   \le
-  %\approx
   \frac C T
@@
 \end{equation}
-%
 By summing over $k$
@@
 for a Gaussian example).
-%
 Note, however, that to approach this upper bound
@@
 in the last step of Eq. \eqref{eq:y2_ubound_k}.
-%
 As a result,
@@
 at which $\lambda_k \approx 1/q(T)$.
-%In other words,
-%the above asymptotic limit fails around
-%$Q^* \approx q(T)$, or equivalently
-%around mode $k^* = k\bigl( q(T) \bigr)$,
-%where $\lambda_{ k^* } \approx 1/q(T)$.
-%%
-%\note{This case also exemplifies
-%  the sensitivity to the simulation length.
-%  %
-%  In this limit,
-%  $M(Q) \propto 1/Q$ by Eq. \eqref{eq:MQ_lbound},
-%  and $C \propto \ln\bigl[ q(T) / q_0 \bigr]$
-%  for some $q_0$,
-%  and we expect that with fixed $a_0$,
-%  $q(T)$ to increase almost linearly with $T$
-%  by Eq. \eqref{eq:opt_qT}.
-%  %
-%  So the normalized asymptotic error,
-%  $E_A \cdot T \propto \bigl( \ln[ q(T)/q_0 ] \bigr)^2 \sim (\ln T)^2$,
-%  is sensitive to the simulation length,
-%  and worse than the ideal case of a constant.
-%}
 
@@
 Eq. \eqref{eq:alpha_invtlambda}.
-%As an alternative to the
-%optimal schedule from Eq. \eqref{eq:q_opt},
-%here we consider the modified
-%inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
-%with a free parameter $\lambda$.
-%
-%whose error can be computed explicitly.
 
@@
 the asymptotic error
-%
 \begin{align}
-  \Err_A(T)
+  \Err_A
   &=
@@
 \end{align}
-%
 where $\nu_k \equiv \lambda_k / \lambda$.
-%
 From Eq. \eqref{eq:error_res1},
 we get the residual error
-%
-\begin{equation}
-  \Err_R(T)
+\begin{equation}
+  \Err_R
   =
@@
 \notag
-%\label{eq:error_res_invt}
-\end{equation}
-%
+\end{equation}
 The total error is the sum of the two.
-%with
-%$
-%  q(T)
-%  =
-%  ( 1 / \lambda )
-%  \ln\left(
-%    1 + T / t_0
-%  \right)
-%$
-%%
-%from integrating Eq. \eqref{eq:alpha_invtlambda}.
-%
-%We assume that
-%the system was initially equilibrated
-%at a constant $a_0$,
-%such that the values of
-%$\left\langle y_k^2(0) \right\rangle$
-%can be computed from Eq. \eqref{eq:y2_eql},
-%as well as Eq. \eqref{eq:t0_sinc}.
-%
-%The total error is
-%%
-%\begin{align}
-%\Err
-%&=
-%\Err_R + \Err_A
-%\notag
-%\\
-%&=
-%\sum_{ k = 2 }^n
-%  \frac
-%  {
-%    \Gamma_k \, \nu_k \,
-%    \left[
-%      \nu_k
-%      +
-%      (\nu_k - 1)
-%      \left(
-%        \frac{ t_0 } { T + t_0 }
-%      \right)^{ 2 \, \nu_k - 1 }
-%    \right]
-%  }
-%  {
-%    (T + t_0) \, (2 \, \nu_k - 1)
-%  }
-%  .
-%%\notag
-%\label{eq:error_invt}
-%\end{align}
 
@@
 Then,
+%
 $$
-  \Err_A(T)
+  E_A
   =
@@
 is minimal at $\lambda = 1$.
-%
-\note{A higher-order correction:
-  $
-  \lambda = 1 -
-  \left[
-    1 - \frac 1 2 a_0 \, (T+t_0)
-  \right] r \, \log r
-  ,
-  $
-  where $r = t_0 / (T + t_0)$.
-}
 
@@
 For the triple-bin updating scheme,
-%introduced in Sec. \ref{sec:nnscheme},
 we can find an analytical expression
@@
 in the large-$n$ and asymptotic limit.
-%
 In this case,
@@
 the asymptotic error is implicitly given by the integral
-%$\sum_i \to \frac{2 \, n}{\pi} \int_0^{\pi/2} dp$, and
-%
 \begin{align}
-  \Err_A(T)
+  \Err_A
   =
@@
 \notag
-%\label{eq:error_nn_asym_int}
 \end{align}
@@
 \begin{aligned}
-  \Err_A(T)
+  \Err_A
   &=
@@
 $$
-%
 Minimizing the function yields
-%
 \begin{equation}
@@
   ,
-%\notag
 \label{eq:lambda_nn_perfect}
 \end{equation}
-%
 which shows that the original inverse-time formula,
@@
 is not optimal in this case.
-%
 The asymptotic error is given by
-%
-\begin{equation}
-  \Err_A(T)
+\begin{equation}
+  \Err_A
   =
@@
 \notag
-%\label{eq:error_nn_perfect}
-\end{equation}
-%
-%From Eq. \eqref{eq:error_nn}, it is clear that
+\end{equation}
 which
 is minimal at $\mu_1 = 0$, the single-bin case.
-%increases with the magnitude of $\mu_1$
-%no matter its sign,
-%and the minimal value, obtained at $\mu_1 = 0$,
-%corresponds to the single-bin scheme.
-%
 Note that the value given by
@@
 with the simulation length.
-%
 
@@
 Then
+%
 $$
-  \Err_A(T)
+  \Err_A
   \propto
@@
 $$
-%
 The optimal $\lambda$ is, therefore,
 $\lambda \approx 1$.
-%
-
-\note{
-However, the above result %Eq. \eqref{eq:lambda_nn_onestep}
-is only good
-for sufficiently short simulations,
-since we have taken the limit of $n \to \infty$
-for a fixed $T$.
-%
-For very long simulations,
-the optimal $\lambda$
-will ultimately drift to
-the smallest eigenvalue,
-$\min \lambda_k = 1 - 4 \, \mu_1$.
-}
-
-\note{The table of symbols is listed in Table \ref{tab:symbols}.
-  \begin{table*}
-  \footnotesize
-  \centering
-  \rowcolors{1}{white}{LightGray}
-  \setlength{\tabcolsep}{4pt} % column separation
-  %
-  \caption{\label{tab:symbols}Table of symbols.}
-  \begin{tabular}{l | p{12cm} }
-    Symbol          &   Description \\
-    \hline
-    $\mathbf{A}$    &   Transition matrix. \\
-    $\alpha$        &   Updating magnitude. \\
-    $b$             &   Cutoff of the kernel of homogeneous updating schemes. \\
-    $C$             &   $C \equiv \int_0^{ q(T) } M(Q) \, dQ$.  \\
-    $\delta$        &   Dirac's or discrete $\delta$-function. \\
-    $c$             &   Constant. \\
-    $\Err, \Err_R, \Err_A$          &   Error, residual error, asymptotic error. \\
-    $\ln f$         &   Updating factor in the WL algorithm.  \\
-    $\phi_{ki}$     &   Eigenvectors of the updating matrix. \\
-    $g$             &   Degeneracy factor, $2$ for a periodic variable, or for $1$ a non-periodic one. \\
-    $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
-    $h_i(t)$        &   Instantaneous histogram.  \\
-    $H_i$           &   Average normalized histogram.  \\
-    $i, j$          &   Bin indices. \\
-    $\lambda_k$     &   The $k$th eigenvalue of the updating matrix. \\
-    $k, l$          &   Mode indices. \\
-    $K$             &   Cutoff wave number of bandpass updating schemes.  \\
-    $\kappa_k(t)$   &   Autocorrelation function of mode $k$. \\
-    $M(Q)$          &   Mass function, Eq. \eqref{eq:mass_func}.   \\
-    $m(Q)$          &   Mass distribution,
-                        $m(Q) \equiv M(Q)/\int_0^{ q(T) } M(Q') \, dQ'$.  \\
-    $\mu_i$         &   Updating kernel of a homogeneous updating scheme. \\
-    $n$             &   Number of bins. \\
-    $\nu_k$         &   $\nu_k \equiv \lambda_k / \lambda$. \\
-    $p_i$           &   Distribution. \\
-    $\pi_i$         &   (Instantaneous) distribution, Eq. \eqref{eq:pi_p_V}. \\
-    $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\
-    $Q$             &   $q(T) - q(t)$.  \\
-    $\theta_k(t)$   &   Noise of mode $k$, Eq. \eqref{eq:theta_def}. \\
-    $\rho_i$        &   Distribution used to define the error. \\
-    $s, S$          &   Stage index and the total number of stages in the Wang-Landau algorithm. \\
-    $\sigma_{ij}(t)$   &   Autocorrelation function of $\zeta_i(t)$. \\
-    $\sigma$        &   Width of the Gaussian updating scheme. \\
-    $t, \tau$       &   Time. \\
-    $T$             &   Simulation length. \\
-    $u_i(q')$       &   $u_i(q') \equiv e^{\lambda \, [q - q(T)]}$. \\
-    $V_i(t)$        &   Original bias potential. \\
-    $v_i(t)$        &   Shifted bias potential. \\
-    $\mathbf w$     &   Updating matrix. \\
-    $x_i(t)$        &   Average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
-    $\xi_i(t)$      &   The noise, Eq. \eqref{eq:xi_def}. \\
-    $y_k(t)$        &   Mode of the shifted bias potential, Eq. \eqref{eq:y_def}. \\
-    $Y_k$           &   Mode of the original bias potential, Eq. \eqref{eq:Y_def}. \\
-    $z$             &   Reaction coordinate. \\
-    $\zeta_i(t)$    &   The noise part of the instantaneous histogram, $h_i(t)$.
-  \end{tabular}
-  \end{table*}
-}
 

### 2017-04-14 doc/invt.tex
add is2.c

@@
 a distribution, $p^*(z)$,
-along a collective variable $z$ by sampling via either Monte Carlo\cite{
+along a collective variable, $z$, by sampling via either Monte Carlo\cite{
   frenkel, newman, landau_binder} (MC)
@@
 %
-The bias potential is associated with discrete values of $z$.
+%The bias potential is associated with discrete values of $z$.
 %
@@
 in the updating window function,
-which takes the form of a discrete
+which specifies
+a neighborhood around the current value of $z$ on the free energy surface
+where the updating should occur
+as well as the relative updating magnitude.
+%
+The window function
+takes the form of a discrete
 $\delta$-function (a boxcar function one histogram bin wide)
@@
 the bias potential %under the WL updating scheme, Eq. \eqref{eq:wl_update},
-as a runtime average from all data collected so far,
+as a runtime average computed from all data points collected so far,
 so that the addition of a new data point only
@@
 $\mathbf p^*$\cite{
-barducci2008, dama2014}.
+  barducci2008, dama2014} ($\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$).
 %
@@
 This approximation is valid if the simulation period, $T$,
-is much longer than the autocorrelation time of $\sigma_{ij}(t)$,
-%
-and for a non-white noise,
+is much longer than the autocorrelation time of $\sigma_{ij}(t)$.
+%
+For non-white noise,
 the equivalent value of $\Gamma_0$ can be obtained
@@
 %
-is also a white noise as $\zeta_i(t)$,
+is also governed by white noise as is $\zeta_i(t)$,
 i.e. we may write
@@
 %
-where $\xi(t)$ is a white noise, such that
+where $\xi(t)$ is some white noise, such that
 %
@@
 By using Eqs.
-\eqref{eq:v_def}, \eqref{eq:vbar_def},
+\eqref{eq:v_def}, \eqref{eq:vbar_def}, and
 \eqref{eq:vt_diffeq_mbin}-\eqref{eq:w_balance}
@@
 Thus, if $M(Q)$ decays slowly because of
-many near zero eigenvalues, $\lambda_k$'s (as in the metadynamics case),
+many near zero eigenvalues, $\lambda_k$'s,
+as in the metadynamics case,
 both the optimal schedule and
@@
 This inequality follows from the non-negativity of
-the following quadratic polynomial,
+the quadratic polynomial,
 $$
@@
   zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
-the random error becomes more important than the systematic one,
-because the former is roughly proportional to
+the random error,
+which is roughly proportional to
 the square root of the updating magnitude\cite{
-  zhou2005, morozov2007, zhou2008, bussi2006}
-whereas the latter is proportional to
+  zhou2005, morozov2007, zhou2008, bussi2006},
+can easily outweigh
+the systematic error,
+which is roughly proportional to
 the magnitude itself\cite{morozov2007}.
@@
 The bias potentials from the two approaches
-would have similar precision at long times.
+would reach similar precision at long times.
 %
@@
 a sufficiently convergent adaptive FDS simulation
-to the corresponding equilibirium FDS method
+to the corresponding equilibrium FDS simulation
 with a fixed bias potential just for better precision
@@
 %
-In the latter, the updating magnitude is fixed,
-but the bias potential over the simulation coarse is averaged
+In the variant, the updating magnitude is fixed,
+but the bias potential over the simulation course is averaged
 as the final result\cite{zhou2005}.
@@
 has already effectively implemented an implicit average
-so that no explicit averaging is needed in the former approaches.
+so that no explicit averaging is needed in the former approach.
 %
@@
 the Gaussian updating scheme, which is stable in the continuous limit,
-can be unstable upon discretization
-during which the updating kernel is truncated.
+can be unstable
+when the updating kernel is truncated during discretization.
 %

### 2017-04-14 prog/invt.c

@@ -1,4 +1,5 @@
-/* main driver */
+/* main driver
+ * One-dimensional model system */
 
 
 

### 2017-04-14 doc/invt.tex
add the referee reports

@@
   \delta_{i, \, i(t)}
-  \frac{ \alpha(t) } { p_i }
+  \frac{ \alpha(t) } { p_{i(t)} }
   ,
@@
 
-By the translational invariance,
-homogeneous updating schemes
-satisfy detailed balance,
-Eq. \eqref{eq:w_detailedbalance},
-for the flat distribution, $\rho_i = 1/n$,
-i.e. the updating matrix, $\mathbf w$,
-is symmetric.
-%
-The updating matrices
-share the same set of eigenvectors
-(which here are sines and cosines of different frequencies),
-and thus are completely determined
-by the eigenvalues.
-%
-Further, the multiplication of $\mathbf w$
-can be reduced to a convolution with the window function,
-or the updating kernel.
-%
-Thus,
-there is a one-to-one correspondence between
-the eigenvalues of $\mathbf w$
-and the updating kernel.
-
 \subsubsection{\label{sec:bandkernel}
@@
 
-It is natural to first introduce homogeneous updating schemes
-for a periodic variable\cite{dama2014}.
-%
-The updating matrix, $\mathbf w$,
+%It is natural to first introduce homogeneous updating schemes
+For a periodic variable\cite{dama2014},
+the updating matrix, $\mathbf w$,
 assumes the following form
@@
 the updating kernel.
-%
+
 For the updating scheme to be stable,
@@
 
-\section{\label{sec:stabilize_wband}
-Stabilization of homogeneous updating schemes}
+\subsection{\label{sec:more_wband}
+Homogeneous updating schemes
+}
+
+Here, we present some detailed derivations
+on the homogeneous updating schemes.
+%
+By the translational invariance,
+homogeneous updating schemes
+satisfy detailed balance,
+Eq. \eqref{eq:w_detailedbalance},
+for the flat distribution, $\rho_i = 1/n$,
+i.e. the updating matrix, $\mathbf w$,
+is symmetric.
+%
+The updating matrices
+share the same set of eigenvectors
+(which here are sines and cosines of different frequencies),
+and thus are completely determined
+by the eigenvalues.
+%
+Further, the multiplication of $\mathbf w$
+can be reduced to a convolution with the window function,
+or the updating kernel.
+%
+Thus,
+there is a one-to-one correspondence between
+the eigenvalues of $\mathbf w$
+and the updating kernel.
+
+\subsection{\label{sec:stabilize_wband}
+Stabilization}
 

### 2017-04-14 prog/invt.h

@@ -158,6 +158,7 @@ static int savewin(int winn, double *win,
 }
 
 
+
 /* save the updating window function
  * in matrix form */
 static int savewinmat(int winn, double *win,

### 2017-04-14 prog/invtsamp.h

@@ -209,9 +209,6 @@ typedef struct {
   double thermdt;
   double expndt, sqrtdt;
   int pbc;
-#if 0
-  double dwa, dwb; /* potential parameters sin(x) (dwa - dwb * sin(x)) */
-#endif
   double *vb; /* point to the bias potential */
 } invtmd_t;
 
@@ -229,10 +226,6 @@ static void invtmd_init(invtmd_t *md, int n,
   md->expndt = exp( -0.25 * thermdt );
   md->sqrtdt = sqrt( 2 * tp * thermdt * 0.5 );
 
-#if 0
-  md->dwa = dwa;
-  md->dwb = dwb;
-#endif
   md->vb = vb;
   md->x = 2 * M_PI * rand01();
   md->v = randgaus();
@@ -245,19 +238,6 @@ static void invtmd_init(invtmd_t *md, int n,
   md->pbc = pbc;
 }
 
-
-
-#if 0
-/* compute the normal force */
-static void invtmd_force(invtmd_t *md)
-{
-  double s = sin(md->x), c = cos(md->x);
-  double a = md->dwa, b = md->dwb;
-
-  md->ep = s * (a - b * s);
-  md->f = c * (-a + 2 * b * s);
-}
-#endif
 
 
 /* add the bias force */

### 2017-04-14 prog/is2.c

@@ -1,6 +1,21 @@
+#include "invt.h"
 #include "is2.h"
+
+static int invt_is2_run(invtpar_t *m)
+{
+  mtscramble( clock() );
+
+  return 0;
+}
 
 int main(int argc, char **argv)
 {
+  invtpar_t m[1];
+
+  invtpar_init(m);
+  invtpar_doargs(m, argc, argv);
+  invtpar_dump(m);
+  invt_is2_run(m);
+  invtpar_finish(m);
   return 0;
 }

### 2017-04-14 doc/invt.tex
small edits to invt.tex

@@
 
-The Wang-Landau (WL) algorithm\cite{wang2001, *wang2001pre}
+Many efficient flat-distribution-sampling (FDS) techniques
+based on adaptive construction of the bias potential
+have been introduced,
+including the Wang-Landau (WL) algorithm\cite{
+  wang2001, *wang2001pre}
 and metadynamics\cite{huber1994,
-*laio2002, *laio2008, *barducci2011, *sutto2012}
-are two widely-used and closely-related\cite{micheletti2004}
-flat-distribution-sampling (FDS) techniques
-that build the bias potential adaptively or on the fly.
-%
-Both techniques regularly update the bias potential
+*laio2002, *laio2008, *barducci2011, *sutto2012},
+more commonly used for MC and MD simulations, respectively\cite{junghans2014}.
+%
+%flat-distribution-sampling (FDS) techniques
+%that build the bias potential adaptively or on the fly.
+%
+These techniques regularly update the bias potential
 to discourage future visits to previously sampled configurations
-by incrementally elevating the bias potential.
+by incrementally elevating the bias potential,
+%and they are closely-related\cite{junghans2014, micheletti2004}.
 %
@@
 %
-The difference lies
+A key difference lies
 in the updating window function,
@@
 %
-The window function
+Usually, the window function
 takes the form of a discrete
@@
 in metadynamics.
+%
+Although the assignment of the window functions to the two methods
+is not absolute\cite{micheletti2004, kim2006, *kim2007, junghans2014},
+below we will use the terms WL and metadynamics
+to represent FDS simulations using the
+$\delta$- and Gaussian window functions, respectively.
 %
@@
 %
-The optimal schedule for metadynamics,
-however, is less studied.
+%The optimal schedule for metadynamics,
+%however, is less studied.
 
@@
 the optimal schedule
-for a general adaptive FDS method,
-including the WL algorithm and metadynamics.
+for an adaptive FDS method
+with a general window function,
+including the $\delta$- and Gaussian one
+used in the WL algorithm and metadynamics.
 %
@@
 we get Eq. \eqref{eq:alpha_invt1}.
+%
+From the boundary condition at $t' = T$,
+we get $c = 1/(T + t_0)$,
+and from the case of $t' = 0$,
+we get $q(T) = \ln(1 + T/t_0)$.
+%
+Since the above optimization is performed
+under a fixed $q(T)$, below
+one will choose a proper value of $q(T)$,
+or, equivalently, of $t_0$, to complete the optimization.
 
@@
 %
-However, we will assume some restrictions and approximations
-to simplify the discussion.
+The key simplification comes from the
+projection of the bias potential to the $n$ eigenmodes
+of the updating matrix, $\mathbf w$.
+%However, we will assume some restrictions and approximations
+%to simplify the discussion.
 
@@
 
-The matrix, $\mathbf w$, is not arbitrary.
+The matrix, $\mathbf w$, is subject to three conditions.
 %
@@
 resembles a transition matrix,
-except that some elements can be negative.
+except that some elements can be negative
+in our cases.
 %
@@
 in studying transition matrices\cite{vankampen}.
-
-To simplify the ensuing discussion,
-we  limit ourselves to matrices % $\mathbf w$'s
+%
+The quoted results were unaffected by the allowance
+of negative elements into the matrix, $\mathbf w$.
+
+%To simplify the ensuing discussion,
+Second, we limit ourselves to matrices % $\mathbf w$'s
 that satisfy
@@
 %
-i.e. $\mathbf \rho$ is a left eigenvector of
+i.e. $\pmb \rho$ is a left eigenvector of
 $\mathbf w$ with eigenvalue $1$.
+%
+In the way,
+the error weighting distribution, $\pmb \rho$,
+restrains the form of the updating matrix, $\mathbf w$.
+%
+The condition, albeit somewhat artificial,
+is readily met in many updating schemes used in practice:
+most commonly, $\pmb \rho$ is a uniform distribution,
+and this condition requires the updating matrix, $\pmb w$, to be symmetric.
 
@@
 \subsubsection{\label{sec:eqlerr}
-Equilibrium error under a constant updating magnitude
+Modeling initial error from a simulation
+under a constant updating magnitude
 }

### 2017-04-17 doc/invt.tex
small edits invt.tex

@@
 %
-%This also helps fixing the notations.
-%
 Then, in Sec. \ref{sec:single-bin},
@@
 %
+The above two introductory sections help fix our notations,
+and an expert reader can readily skip them.
+%
 Next, we extend the method to the general case
@@
 
-For a large system at temperatures of interest,
-the distribution, $p^*_i$, often tends to
-be localized,
-%
-and to explore the global shape
-of the PMF, $-\ln p^*_i$,
+%For a large system at temperatures of interest,
+%the distribution, $p^*_i$, often tends to
+%be localized,
+%%
+%and to explore the global shape
+%of the PMF, $-\ln p^*_i$,
+%it is often advantageous to perform
+%biased sampling that targets
+%a wider and smoother distribution, $p_i$.
+%%
+%%Here, we refer to simulations that target
+%%a flat or nearly flat distribution
+%%as entropic or multicanonical sampling.
+%
+%To do so, we introduce a bias potential, $V_i$,
+If the distribution, $p^*_i$, is multimodal,
 it is often advantageous to perform
 biased sampling that targets
-a wider and smoother distribution, $p_i$.
-%
-%Here, we refer to simulations that target
-%a flat or nearly flat distribution
-%as entropic or multicanonical sampling.
-
-To do so, we introduce a bias potential, $V_i$,
-which changes the equilibrium distribution to
-%
-\begin{equation}
-  \pi_i \propto p^*_i \, e^{-V_i}.
-  \label{eq:pi_p_V1}
-\end{equation}
-%
-Upon normalization,
-$\sum_{i = 1}^n \pi_i = 1$,
-we get
+a smoother distribution, $p_i$,
+by introducing a bias potential, $V_i$,
+such that the biased equilibrium distribution is
 %
@@
   \pi_i
+  %=
+  %C_i \, p^*_i \, e^{-V_i}
   =
@@
        { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
+  \propto
+  p^*_i \, e^{-V_i}
   .
@@
 \end{equation}
+%
+%where the proportionality constant, $C_i$,
+%is found from the normalization condition,
+%$\sum_{i = 1}^n \pi_i = 1$,
+%in the second step.
 %
@@
 
-We wish the biased equilibrium distribution, $\pi_i$,
-to converge to the target distribution, $p_i$,
-which is usually chosen to be flat,
-i.e. $p_i$ is constant for all $i$'s,
+We will broadly call a biased simulation
+targeting a flat (i.e. $p_i \equiv 1/n$)
 or nearly flat\cite{
-dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}.
-%
-We will broadly call such a biased simulation
+dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}
+$p_i$
 a flat-distribution-sampling (FDS) simulation.
 %
-To achieve this distribution, the bias potential must satisfy
+For the biased distribution, $\pi_i$,
+to converge to $p_i$,
+the bias potential must satisfy
 %
@@
   V_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
-  ,
+  .
   \label{eq:Vi_target}
@@
 %
-according to Eq. \eqref{eq:pi_p_V1}.
+%according to Eq. \eqref{eq:pi_p_V}.
 %which allows us to deduce the PMF, $-\ln p^*_i$,
@@
 %
-This relation thus sets a practical aim for the convergence of FDS
-methods---the bias potential must be adjusted to make the
+In practice,
+the bias potential is adjusted to make the
 observed distribution, or the normalized histogram,
-match the intended one.
-%
-Once this aim is achieved,
+match the intended one,
+and once this is done,
 the PMF, $-\ln p^*_i$, can be naturally
 deduced from the bias potential. % $V_i$.
+%
+%This relation thus sets a practical aim for the convergence of FDS
+%methods---the bias potential must be adjusted to make the
+%observed distribution, or the normalized histogram,
+%match the intended one.
+%%
+%Once this aim is achieved,
+%the PMF, $-\ln p^*_i$, can be naturally
+%deduced from the bias potential. % $V_i$.
 \note{
@@
 and simulated tempering\cite{
-marinari1992, *lyubartsev1992}
-with $i$ being the index of energy
-and that of temperature, respectively.
+marinari1992, *lyubartsev1992}.
+%with $i$ being the index of energy
+%and that of temperature, respectively.
 %
@@
 Using $H_i$ as an estimate of $\pi_i$ in
-Eq. \eqref{eq:pi_p_V1}
+Eq. \eqref{eq:pi_p_V}
 allows us to correct $V_i$ as
@@
 }
-However, for an unknown system,
-it is often difficult
-to guess a sufficiently accurate
-$V_i$ with which to start.
-%
-Thus,
-it can take many iterative rounds of simulations
-to achieve the desired flat distribution.
-%
-%Another difficulty of this approach is that
-The effectiveness of this approach
-depends on the precision of the histogram, $H_i$,
-which in turn depends on the simulation length
-that must be decided in advance.
-%
-Thus, the rigidity of this framework disallows
-continuous, rapid improvement of the bias potential,
-and subsequently the PMF,
+%However, for an unknown system,
+%it is often difficult
+%to guess a sufficiently accurate
+%$V_i$ with which to start.
+%%
+%Thus,
+%it can take many iterative rounds of simulations
+%to achieve the desired flat distribution.
+%%
+%%Another difficulty of this approach is that
+%The effectiveness of this approach
+%depends on the precision of the histogram, $H_i$,
+%which in turn depends on the simulation length
+%that must be decided in advance.
+However, this formula often requires a long
+simulation period to accumulate a precise histogram, $H_i$,
+and thus it disallows
+continuous improvement of the bias potential
+(hence that of the PMF)
 as simulation lengthens.
@@
 The second type of methods,
-which we will refer to as the adaptive FDS methods,
+including the WL algorithm and metadynamics,
 address the above shortcomings
@@
 %
-Examples include
-the WL algorithm and metadynamics.
-%
-%These methods incrementally update the bias potential
-%and can approximately sample the target distribution.
-%
-However,
-the on-the-fly updates break the microscopic reversibility
-of the underlying sampling mechanism.
-Further, failure to properly reduce the magnitude of
-the histogram-based increments
-would lead to roughness or errors in the final estimate of the PMF.
-%
-%Thus, for a conservative practitioner,
-%the E-FDS methods appear more appealing
-%once a sufficiently accurate bias potential is available.
-%
-To minimize this effect,
-many adaptive methods adopt a schedule to
-reduce the updating magnitude
-over the course of simulation\cite{
-marsili2006,
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-liang2007,
-barducci2008}.
-%
-%Thus, in late stages of a long simulation,
-%the updating becomes negligible and
-%the adaptive simulation tends to
-%the corresponding equilibrium one.
-%
-%if the bias potential were converged to the exact value,
-%this adaptive FDS tends to the ideal equilibrium counterpart,
-%particular equilibrium simulation can be thought as
-%the optimal asymptotic limit of the adaptive counterpart,
-%
-For example, one can show that
-if the magnitude is controlled by
-the optimal inverse-time prescription\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-liang2007,
-morozov2007, zhou2008,
-komura2012, *caparica2012, *caparica2014},
-the WL algorithm is equivalent to the ideal equilibrium counterpart
-in improving the precision of the bias potential,
-thus allowing for maximum efficiency
-(cf. Appendix \ref{sec:equilerr}).
-%the bias potential in the WL algorithm
-%will evolve similar to Eq. \eqref{eq:vcorr_equil}
-%with the precision of the histogram as simulation lengthens.
+These methods, refered to as the adaptive FDS methods,
+will be discussed below.
+%%
+%%These methods incrementally update the bias potential
+%%and can approximately sample the target distribution.
+%%
+%However,
+%the on-the-fly updates break the microscopic reversibility
+%of the underlying sampling mechanism.
+%Further, failure to properly reduce the magnitude of
+%the histogram-based increments
+%would lead to roughness or errors in the final estimate of the PMF.
+%%
+%%Thus, for a conservative practitioner,
+%%the E-FDS methods appear more appealing
+%%once a sufficiently accurate bias potential is available.
+%%
+%To minimize this effect,
+%many adaptive methods adopt a schedule to
+%reduce the updating magnitude
+%over the course of simulation\cite{
+%marsili2006,
+%belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
+%liang2007,
+%barducci2008}.
+%%
+%%Thus, in late stages of a long simulation,
+%%the updating becomes negligible and
+%%the adaptive simulation tends to
+%%the corresponding equilibrium one.
+%%
+%%if the bias potential were converged to the exact value,
+%%this adaptive FDS tends to the ideal equilibrium counterpart,
+%%particular equilibrium simulation can be thought as
+%%the optimal asymptotic limit of the adaptive counterpart,
+%%
+%For example, one can show that
+%if the magnitude is controlled by
+%the optimal inverse-time prescription\cite{
+%belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
+%liang2007,
+%morozov2007, zhou2008,
+%komura2012, *caparica2012, *caparica2014},
+%the WL algorithm is equivalent to the ideal equilibrium counterpart
+%in improving the precision of the bias potential,
+%thus allowing for maximum efficiency
+%(cf. Appendix \ref{sec:equilerr}).
+%%the bias potential in the WL algorithm
+%%will evolve similar to Eq. \eqref{eq:vcorr_equil}
+%%with the precision of the histogram as simulation lengthens.
 
@@
 
-We will mainly focus on the adaptive methods below.
+%We will mainly focus on the adaptive methods below.
 %
@@
 so that the addition of a new data point only
-affects the average by a weight of $1/t$,
+affects the average by its due weight, $1/t$,
 which is the inverse of the current sample size, $t$
@@
 
-As Eq. \eqref{eq:alpha_invt} was being developed for the WL algorithm,
-several similar ways of
-reducing the updating magnitude
-were developed for umbrella sampling and metadynamics,
-particularly, the self-healing umbrella sampling
-and well-tempered metadynamics\cite{
-  marsili2006, barducci2008, dickson2011, dama2014}.
-%
+%As Eq. \eqref{eq:alpha_invt} was being developed for the WL algorithm,
+%several similar ways of
+%reducing the updating magnitude
+%were developed for umbrella sampling and metadynamics,
+%particularly, the self-healing umbrella sampling
+%and well-tempered metadynamics\cite{
+%  marsili2006, barducci2008, dickson2011, dama2014}.
+%%
 Our aim here is to find the optimal schedule, $\alpha(t)$,
@@
 %%
-%According to Eq. \eqref{eq:pi_p_V1},
+%According to Eq. \eqref{eq:pi_p_V},
 
@@
 =
-c \, (t' + t_0),
-$$
-where $t_0$ is a constant.
+c \, (t' + t_0)
+=
+\frac{ t' + t_0 }
+     { T  + t_0 }
+,
+$$
+where $t_0$ is a constant,
+and in the second step,
+the constant $c$ is determined
+from the boundary condition at $t' = T$.
 %
@@
 %
-From the boundary condition at $t' = T$,
-we get $c = 1/(T + t_0)$,
-and from the case of $t' = 0$,
-we get $q(T) = \ln(1 + T/t_0)$.
+Further, from the case of $t' = 0$,
+we get $q(T) = \ln\bigl[1 + (T/t_0)\bigr]$.
 %
 Since the above optimization is performed
-under a fixed $q(T)$, below
-one will choose a proper value of $q(T)$,
-or, equivalently, of $t_0$, to complete the optimization.
+under a fixed $q(T)$,
+we will complete the optimization
+by choosing a proper value of $q(T)$,
+or, equivalently, of $t_0$.
 

### 2017-04-18 doc/invt.tex
editing invt.tex

@@
 The above two introductory sections help fix our notations,
-and an expert reader can readily skip them.
+and an expert reader may wish to skip them.
 %
@@
 in Secs. \ref{sec:multiple-bin}
-to \ref{sec:band-matrix}.
+to \ref{sec:band-matrix},
+and present a general procedure of computing the optimal schedule
+in Sec. \ref{sec:procedure}.
 %
@@
 
-We will broadly call a biased simulation
-targeting a flat (i.e. $p_i \equiv 1/n$)
-or nearly flat\cite{
-dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}
-$p_i$
-a flat-distribution-sampling (FDS) simulation.
+We will broadly define
+a flat-distribution-sampling (FDS) simulation
+as a biased simulation
+targeting a flat $p_i$ (i.e. $p_i \equiv 1/n$)
+or nearly flat one\cite{
+dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}.
 %
 For the biased distribution, $\pi_i$,
-to converge to $p_i$,
+to converge to the desired $p_i$,
 the bias potential must satisfy
@@
 %
-In practice,
-the bias potential is adjusted to make the
-observed distribution, or the normalized histogram,
-match the intended one,
-and once this is done,
+This relation thus sets a practical goal for the convergence of FDS
+methods---the bias potential must be adjusted to make the
+observed distribution match the intended one.
+%%
+Once this is done,
 the PMF, $-\ln p^*_i$, can be naturally
 deduced from the bias potential. % $V_i$.
-%
-%This relation thus sets a practical aim for the convergence of FDS
-%methods---the bias potential must be adjusted to make the
-%observed distribution, or the normalized histogram,
-%match the intended one.
-%%
-%Once this aim is achieved,
-%the PMF, $-\ln p^*_i$, can be naturally
-%deduced from the bias potential. % $V_i$.
 \note{
@@
 
-Roughly speaking, there are two types of FDS methods.
+In terms of how one adjusts the bias potential,
+there are roughly two types of FDS methods.
 %
@@
 %
-The estimated bias potentials at bin $i$, $V_i$,
-are often improved
-with the normalized histogram of the sampling, $H_i$, accumulated
-during simulation.
+After a sufficiently long run, the bias potential, $V_i$,
+can often be improved
+with the normalized histogram, $H_i$, from the simulation.
 %
@@
 However, this formula often requires a long
-simulation period to accumulate a precise histogram, $H_i$,
-and thus it disallows
+simulation period to accumulate a precise histogram,
+and thus disallows
 continuous improvement of the bias potential
-(hence that of the PMF)
+%(hence that of the PMF)
 as simulation lengthens.
@@
 including the WL algorithm and metadynamics,
-address the above shortcomings
+addresses the above shortcomings
 by continuously and incrementally updating the bias potential.
@@
 %
-These methods, refered to as the adaptive FDS methods,
+These methods, referred to as the adaptive FDS methods,
 will be discussed below.
@@
 
-One can show that with a constant magnitude,
-$\alpha(t) = \alpha > 0$,
-the distribution collected from
-an adaptive FDS simulation
-is identical to $p_i$ in the long time or asymptotic limit.
-%
-However, as the above updates
-change $V_i(t)$ continuously,
-the equilibrium distribution,
-Eq. \eqref{eq:pi_p_V}, no longer holds
-at all times,
-and the error in $V_i(t)$
-depends on the updating magnitude\cite{
+%One can show that with a constant magnitude,
+%$\alpha(t) = \alpha > 0$,
+%the distribution collected from
+%an adaptive FDS simulation
+%is identical to $p_i$ in the long time or asymptotic limit.
+%
+%However, as the above updates
+%change $V_i(t)$ continuously,
+%the equilibrium distribution,
+%Eq. \eqref{eq:pi_p_V}, no longer holds
+%at all times,
+While adaptive FDS simulations
+can often sample the desired distribution, $p_i$,
+even with a constant magnitude,
+$\alpha(t) = \alpha$,
+frequent updates to the bias potential
+disrupt the underlying equilibrium sampling
+and introduce artificial errors that need to be
+removed by gradually decreasing the updating magnitude\cite{
   belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
@@
 %%
-%
-Thus, one has to reduce the magnitude, $\alpha$, over time
-to improve the estimate of the distribution,
-$p^*_i$. %via Eq. \eqref{eq:v_def}.
 
@@
 %used in WL algorithm,
-Later,
-it was found that
-a better way
+A more effective way
 of controlling the schedule, $\alpha(t)$,
@@
 is the number of steps
-from the beginning of the simulation.
-%
-This simple yet effective formula has attracted
-several studies\cite{
+from the beginning of the simulation\cite{
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
@@
 %
-An interesting explanation\cite{
+An interesting explanation of this result\cite{
   marsili2006, barducci2008}
@@
 %
-Here we will re-derive this formula
-and suggest a slight modification
-whereby the origin of $t$ is shifted
-according to the initial error.
-%
-Interestingly, we may reach the same result
+Alternatively, we may derive it
 from optimally planning the simulation lengths in the stages
-and the reduction factor, as shown in Appendix \ref{sec:wltoinvt}.
+and the reduction factor in the WL setup (cf. Appendix \ref{sec:wltoinvt}).
+
+%In the next section,
+%we will re-derive this formula
+%and suggest a slight modification
+%whereby the origin of $t$ is shifted
+%according to the initial error.
+%
 
@@
 %%
-Our aim here is to find the optimal schedule, $\alpha(t)$,
-for a general updating scheme.
+%In this study, we wish to find the optimal schedule, $\alpha(t)$,
+%for a general updating scheme,
+%including the Gaussian scheme used in metadynamics.
 
@@
 for the single-bin scheme,
-Eq. \eqref{eq:wl_update}.
-%
-To do so,
-we will first express the error of
+Eq. \eqref{eq:wl_update},
+as a preparation
+for the more general multiple-bin scheme.
+%
+We will suggest a slight modification of Eq. \eqref{eq:alpha_invt}
+whereby the origin of $t$ is shifted by an amount
+determined from the initial error.
+%
+We will first express the error of
 the bias potential, $V_i(t)$,
@@
 and then minimize it by variation.
-
-\subsubsection{Error}
-
-First, we need to define the error in the bias potential.
-%
-This is done by deducting two contributions from the bias potential.
+%
+
+\subsubsection{Error definition}
+
+First, we define the net error in the bias potential
+by deducting two contributions.
 %
@@
 
-Note that the distribution, $\pmb \rho$,
-introduced above to define the error
+Note that the error weighting distribution, $\pmb\rho$,
+defined above
 is independent of
@@
 %
-For example,
-in well-tempered metadynamics,
-$\pmb\rho$ is a flat distribution,
-while $\mathbf p$ is an attenuated version of
+In practice, $\pmb\rho$ is usually flat,
+i.e. $\rho_i \equiv 1/n$,
+even when $\mathbf p$ is not
+(e.g. in well-tempered metadynamics,
+$\mathbf p$ is an attenuated version of
 $\mathbf p^*$\cite{
-  barducci2008, dama2014} ($\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$).
+  barducci2008, dama2014}:
+$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$).
 %
@@
 %
+and we will assume it is white, i.e.
 %The noise is not necessarily white,
-and the autocorrelation functions
-depend only on the time difference
+%and the autocorrelation function
+%depend only on the time difference
 %
@@
   =
-  \sigma_{ij}(t - t'),
+  %\sigma_{ij}(t - t')
+  G^{(0)}_{ij} \, \delta(t - t')
+  ,
   \notag
@@
 %
-where $\sigma_{ij}(t)$ is an even function
-that vanishes at large $t$.
+for some value of $G^{(0)}_{ij}$,
+and $\delta(t)$ is the Dirac's delta function.
+%where $\sigma_{ij}(t)$ is an even function
+%that vanishes at large $t$.
+%
+This approximation is valid if the simulation period, $T$,
+is much longer than the autocorrelation time of noise.
+%
+For non-white noise,
+the equivalent value of $G^{(0)}_i$ can be approximately obtained
+from the integral,
+$\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$,
+or in the discrete time case,
+$\sum_{-\infty}^\infty \sigma_{ij}(t)$.
 %
@@
 }%
-Below, we will assume the noise is white,
-such that
-$$
-\sigma_{ij}(t) = \Gamma_0 \, \delta(t)
-$$
-for some $\Gamma_0$,
-and $\delta(t)$ is the Dirac's delta function.
-%
-This approximation is valid if the simulation period, $T$,
-is much longer than the autocorrelation time of $\sigma_{ij}(t)$.
-%
-For non-white noise,
-the equivalent value of $\Gamma_0$ can be obtained
-from the integral of the autocorrelation function,
-$\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$.
 
@@
 \end{equation}
+%
 where
+%
 \begin{equation}
@@
 $$
-  \langle \xi_i(t) \, \xi_i(t') \rangle = \Gamma_i \, \delta(t - t'),
-$$
-for some $\Gamma_i$.
+  \langle \xi_i(t) \, \xi_i(t') \rangle = G_i \, \delta(t - t'),
+$$
+for some $G_i$.
 %
@@
 that minimizes this error.
-%
-It suffices to consider
-a one-variable problem that
-minimizes each individual
-$\left\langle x_i^2(T) \right\rangle$.
-
-\subsubsection{\label{sec:onevar}
-Optimization of a one-variable problem}
-
-Consider the following stochastic differential equation
-of a single variable:
-%
-\begin{equation}
-\dot x(t) = -\alpha(t) \, \left[ x(t) - \xi(t) \right],
-\label{eq:dxdt_alpha}
-\end{equation}
-%
-where $\xi(t)$ is some white noise, such that
-%
-$
-  \left\langle \xi(t) \, \xi(t') \right\rangle
-  =
-  \Gamma \, \delta(t - t')
-  .
-$
-%
-We wish to show that the $\alpha(t)$
-of minimizing $\langle x^2(t) \rangle$ at long times
-is given by
-%
-\begin{equation}
-  \alpha(t) = \frac{ 1 }{ t + t_0 },
-\label{eq:alpha_invt1}
-\end{equation}
-%
-where $t_0$ is a constant.
-
-To do so, we first recall
-the formal solution of Eq. \eqref{eq:dxdt_alpha}:
-%
-\begin{equation}
-x(T)
-=
-x(0) \, e^{-q(T)}
+%%
+%It suffices to consider
+%a one-variable problem that
+%minimizes each individual
+%$\left\langle x_i^2(T) \right\rangle$.
+
+%\subsubsection{\label{sec:onevar}
+%Optimization of a one-variable problem}
+
+\subsubsection{Optimal schedule for the single-bin scheme}
+
+We recall that
+the formal solution of Eq. \eqref{eq:dxdt_WL} is
+%
+\begin{equation}
+x_i(T)
+=
+x_i(0) \, e^{-q(T)}
 +
 \int_0^T
-  \dot u\bigl( q(t) \bigr) \, \xi(t) \, dt,
+  \dot u\bigl( q(t) \bigr) \, \xi_i(t) \, dt,
 \notag
@@
 %
-At the end of period $T$, we have
+At the end of period $T$,
+the total error defined in Eq. \eqref{eq:error_sum}
+is given by
+%
+%\begin{align}
+%  &\left\langle x_i^2(T) \right\rangle
+%  =
+%  \langle x_i^2(0) \rangle \, e^{-2 \, q(T)}
+%  +
+%  G_i \int_0^T
+%    {\dot u}^2\bigl( q(t) \bigr) \,
+%    \, dt.
+%  \label{eq:x2t_average}
+%\end{align}
 %
 \begin{align}
-  &\left\langle x^2(T) \right\rangle
-  =
-  \langle x^2(0) \rangle \, e^{-2 \, q(T)}
+  E(T)
+  =
+  E(0) \, e^{-2 \, q(T)}
   +
-  \Gamma \int_0^T
+  \Gamma \,
+  \int_0^T
     {\dot u}^2\bigl( q(t) \bigr) \,
-    \, dt.
-  \label{eq:x2t_average}
+    \, dt
+  ,
+  \label{eq:ET_average}
 \end{align}
+where
+$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
 
@@
 %
-This is equivalent to varying $u$
+This is equivalent to varying the curve $u(q(t))$
 with fixed endpoints,
@@
 Note also that the first term on the right-hand side
-of Eq. \eqref{eq:x2t_average} is fixed during
+of Eq. \eqref{eq:ET_average} is fixed during
 the process,
-while the second term is the action of a free particle,
-which must move at a constant velocity, $c$:
+while the second term is the action of a free particle.
+Thus, the velocity of $u(q(t))$ should be a constant:
 %
@@
 \begin{equation}
-  \dot u\bigl( q(t') \bigr) = c
+  \dot u\bigl( q(t) \bigr) = c
   .
@@
 Then, by using Eq. \eqref{eq:u_def}, we have
-$$
-e^{ q(t') - q(T) }
-=
-c \, (t' + t_0)
-=
-\frac{ t' + t_0 }
+%
+\begin{equation}
+  e^{ q(t) - q(T) }
+  =
+  c \, (t + t_0)
+  =
+  \frac{ t + t_0 }
        { T + t_0 }
   ,
-$$
+\label{eq:expqt}
+\end{equation}
 where $t_0$ is a constant,
-and in the second step,
-the constant $c$ is determined
-from the boundary condition at $t' = T$.
+and the constant $c$ is determined
+from the boundary condition at $t = T$
+in the second step.
 %
@@
 and differentiating both sides
-with respect to $t'$,
-we get Eq. \eqref{eq:alpha_invt1}.
-%
-Further, from the case of $t' = 0$,
-we get $q(T) = \ln\bigl[1 + (T/t_0)\bigr]$.
-%
+with respect to $t$,
+we get the inverse-time schedule
+%
+\begin{equation}
+  \alpha(t) = \frac{ 1 }{ t + t_0 }
+  ,
+\label{eq:alpha_invt1}
+\end{equation}
+%
+and Eq. \eqref{eq:alpha_invt}
+is the special case of $t_0 = 0$.
+%
+
 Since the above optimization is performed
 under a fixed $q(T)$,
-we will complete the optimization
-by choosing a proper value of $q(T)$,
-or, equivalently, of $t_0$.
-
-\subsubsection{Optimal schedule for the single-bin scheme}
-
-For the problem of
-minimizing the error defined in Eq. \eqref{eq:error_sum}
-for the single-bin scheme,
-we note that Eq. \eqref{eq:dxdt_WL}
-is of the form of Eq. \eqref{eq:dxdt_alpha}.
-%
-Thus,
-the optimal schedule, $\alpha(t)$, is given by
-Eq. \eqref{eq:alpha_invt1}
-for it optimizes each term, $\langle x_i^2(T) \rangle$,
-and hence the sum;
-Eq. \eqref{eq:alpha_invt}
-is the special case of $t_0 = 0$.
-
+it can be improved
+by choosing a proper value of $q(T)$.
+%
+This is equivalent to choosing a value of $t_0$
+\big[because $q(T) = \ln\bigl[1 + (T/t_0)\bigr]$,
+following the $t = 0$ case of Eq. \eqref{eq:expqt}\big].
+%
 %In Sec. \ref{sec:finlencorr},
@@
 %to accommodate the initial error.
-To find the optimal value of $t_0$,
-we use the optimal schedule in Eq. \eqref{eq:x2t_average},
-and
-$$
-\left\langle x^2(T) \right\rangle
-=
-\frac{ \left\langle x^2(0) \right\rangle \, t_0^2 }
+%
+%and
+%$$
+%\left\langle x^2(T) \right\rangle
+%=
+%\frac{ \left\langle x^2(0) \right\rangle \, t_0^2 }
+%     { (t_0 + T)^2 }
+%+
+%\frac{ \Gamma \, T}
+%     { (t_0 + T)^2 }
+%.
+%$$
+%Using this result for each $x_i$,
+%we get the final total error, $E(T)$,
+By using the optimal schedule in Eq. \eqref{eq:ET_average},
+we find the final total error to be
+\begin{equation}
+  E(T)
+  =
+  \frac{ E(0) \, t_0^2 }
        { (t_0 + T)^2 }
@@
        { (t_0 + T)^2 }
-.
-$$
-Using this result for each $x_i$,
-we get the final total error, $E(T)$,
-as defined in Eq. \eqref{eq:error_sum},
-\begin{equation}
-  E(T)
-  =
-  \frac{ \sum_{i=1}^n \left\langle x_i^2(0) \right\rangle \, t_0^2 }
-       { (t_0 + T)^2 }
-  +
-  \frac{ \sum_{i=1}^n \Gamma_i \, T}
-       { (t_0 + T)^2 }
-  .
+  ,
   \label{eq:error_singlebin}
@@
 %
-This error is minimized if
-%
-\begin{equation}
-t_0 = \frac{ \sum_{i = 1}^n \Gamma_i }
-           { \sum_{i = 1}^n \left\langle x_i^2(0) \right\rangle }
+which is minimized if
+%
+\begin{equation}
+t_0 = \frac{ \Gamma }
+           { E(0) }
            ,
@@
   =
-  \frac{ 1 } { T + t_0 }
-  \sum_{i = 1}^n \Gamma_i
+  \frac{ \Gamma } { T + t_0 }
   .
@@
 The condition, albeit somewhat artificial,
-is readily met in many updating schemes used in practice:
+is readily satisfied in practice:
 most commonly, $\pmb \rho$ is a uniform distribution,
@@
 
-By assuming Eq. \eqref{eq:w_detailedbalance},
-we can diagonalize $\mathbf w$ with a set of
+With Eq. \eqref{eq:w_detailedbalance},
+the scaled matrix, $\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
+is symmetric.
+Thus, we can diagonalize $\mathbf w$ with a set of
 eigenvectors, $\phi_{ki}$,
@@
   y_k(t)
-  &=
+  &\equiv
   \sum_{j=1}^n
@@
   \theta_k(t)
-  &=
+  &\equiv
   \sum_{j=1}^n
@@
 The error function defined in Eq. \eqref{eq:error_sum}
-can be rewritten in terms of the eigenmodes of $\mathbf w$
+can be rewritten in terms of the eigenmodes of $\mathbf w$,
 %
@@
 
-In analogy to Eq. \eqref{eq:x2t_average},
+In analogy to Eq. \eqref{eq:ET_average},
 using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
@@
     \dot u_k\bigl( q(t') \bigr) \,
-    \kappa_k(t - t') \, dt \, dt',
-  \label{eq:error_asym}
+    \kappa_k(t - t') \, dt \, dt'
+  ,
+  \notag
+  %\label{eq:error_asym}
 \end{equation}
@@
 with
-%
-\begin{align*}
-  u_k(q')
-  &\equiv
-  e^{\lambda_k \, [q' - q(T)]},
-  \\
-  \kappa_k(t - t')
-  &\equiv
-  \left\langle
-    \theta_k(t) \, \theta_k(t')
-  \right\rangle.
-\end{align*}
-
-\subsubsection{White-noise approximation}
-
-In the asymptotic regime
-(i.e. if the simulation period, $T$, is sufficiently long),
-we can approximate $\kappa_k(t)$ as
-%
-\begin{equation}
-  \kappa_k(t) \approx \Gamma_k \, \delta(t)
-  ,
-\label{eq:kappa_delta}
-\end{equation}
-%
-where
-\begin{equation}
-  \Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt,
-  \notag
+$u_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}$,
+and
+$\kappa_k(t - t') \equiv \left\langle \theta_k(t) \, \theta_k(t') \right\rangle$.
+%
+%\begin{align*}
+%  u_k(q')
+%  &\equiv
+%  e^{\lambda_k \, [q' - q(T)]},
+%  \\
+%  \kappa_k(t - t')
+%  &\equiv
+%  \left\langle
+%    \theta_k(t) \, \theta_k(t')
+%  \right\rangle.
+%\end{align*}
+
+%\subsubsection{White-noise approximation}
+
+%In the asymptotic regime
+%(i.e. if the simulation period, $T$, is sufficiently long),
+For a long simulation,
+we can again approximate $\theta_k(t)$
+as white noise:
+$\kappa_k(t) \approx \Gamma_k \, \delta(t)$,
+%
+with
+%
+%\begin{equation}
+%  \kappa_k(t) \approx \Gamma_k \, \delta(t)
+%  ,
+%\label{eq:kappa_delta}
+%\end{equation}
+%
+%\begin{equation}
+$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
+%  \notag
 %\label{eq:Gamma_integral}
-\end{equation}
+%\end{equation}
 or in discrete time,
 \begin{equation}
-  \Gamma_k = \sum_{t = -\infty}^\infty \kappa_k(t).
+  \Gamma_k
+  = \sum_{t = -\infty}^\infty \kappa_k(t)
+  = \sum_{t = -\infty}^\infty \left\langle \theta_k(0) \, \theta_k(t) \right\rangle
+  .
   \label{eq:Gamma_sum}
@@
 %
-Then, Eq. \eqref{eq:error_asym} is simplified as
+Then, the asymptotic error is simplified as
 %
@@
 (cf. Sec. \ref{sec:optinitalpha}).
-%
-A complete procedure of computing the optimal schedule
-is listed in Appendix \ref{sec:procedure}.
 
@@
 as shown in Fig. \ref{fig:massq},
-
-This mapping allows us to classify optimal schedules
-as follows.
-%
-We collect in each class optimal schedules
-computed from simulations with the same function, $M(Q)$,
-and cutoff value of $q(T)$.
-%
-The former is determined by the values of
-$\lambda_k$'s and $\Gamma_k$'s,
-whereas the latter affects the distribution, $m(Q)$,
-via the normalization constant, $C$.
-
-In each class, the updating magnitudes of two schedules
-of lengths $T_1$ and $T_2$
-at corresponding times, $t_1/T_1 = t_2/T_2$,
-are related by $T_1 \, \alpha_1(t_1) = T_2 \, \alpha_2(t_2)$.
-%
-Further, by Eqs.
-\eqref{eq:error_asym_Lagrangian}
-and
-\eqref{eq:Lagrangian_const},
-we have
+%
+In terms of the mass function,
+the optimized asymptotic error is given by
 %
@@
 %
-which means that the asymptotic error
-is inversely proportional to the simulation length
-for schedules in the same class.
-%which can be rephrased as another scaling relation
-%$$
-%T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2).
-%$$
+which follows from Eqs.
+\eqref{eq:error_asym_Lagrangian}
+and
+\eqref{eq:Lagrangian_const}.
+%
+
+%This mapping allows us to classify optimal schedules
+%as follows.
+%%
+%We collect in each class optimal schedules
+%computed from simulations with the same function, $M(Q)$,
+%and cutoff value of $q(T)$.
+%%
+%The former is determined by the values of
+%$\lambda_k$'s and $\Gamma_k$'s,
+%whereas the latter affects the distribution, $m(Q)$,
+%via the normalization constant, $C$.
+%
+%
+%In each class, the updating magnitudes of two schedules
+%of lengths $T_1$ and $T_2$
+%at corresponding times, $t_1/T_1 = t_2/T_2$,
+%are related by $T_1 \, \alpha_1(t_1) = T_2 \, \alpha_2(t_2)$.
+%%
+%Equation \eqref{eq:error_asym2}
+%means that the asymptotic error
+%is inversely proportional to the simulation length
+%for schedules in the same class.
+%%which can be rephrased as another scaling relation
+%%$$
+%%T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2).
+%%$$
 
@@
 
-For simulations with the same $a_0$,
-the shape of the mass distribution, $m(Q)$,
-also determines how much the optimal schedule
-would depend on the simulation length, $T$.
-%
-Since both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$,
-Eq. \eqref{eq:opt_qT} shows that
-the optimal $q(T)$ increases with
-the simulation length, $T$,
-more so if $M(Q)$ has a long tail.
-%
-Thus, if $M(Q)$ decays slowly because of
-many near zero eigenvalues, $\lambda_k$'s,
-as in the metadynamics case,
-both the optimal schedule and
-normalized asymptotic error, $\Err_A(T) \cdot T$,
-would be sensitive to the simulation length, $T$.
-%
-In contrast, if all eigenvalues are $1.0$
-as in the single-bin or WL updating scheme discussed previously,
-the optimal schedule is
-given by the universal inverse-time formula,
-Eq. \eqref{eq:alpha_invt1},
-and the normalized asymptotic error
-is roughly a constant
-[cf. Eq. \eqref{eq:error_singlebin}].
+%For simulations with the same $a_0$,
+%the shape of the mass distribution, $m(Q)$,
+%also determines how much the optimal schedule
+%would depend on the simulation length, $T$.
+%%
+%Since both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$,
+%Eq. \eqref{eq:opt_qT} shows that
+%the optimal $q(T)$ increases with
+%the simulation length, $T$,
+%more so if $M(Q)$ has a long tail.
+%%
+%Thus, if $M(Q)$ decays slowly because of
+%many near zero eigenvalues, $\lambda_k$'s,
+%as in the metadynamics case,
+%both the optimal schedule and
+%normalized asymptotic error, $\Err_A(T) \cdot T$,
+%would be sensitive to the simulation length, $T$.
+%%
+%In contrast, if all eigenvalues are $1.0$
+%as in the single-bin or WL updating scheme discussed previously,
+%the optimal schedule is
+%given by the universal inverse-time formula,
+%Eq. \eqref{eq:alpha_invt1},
+%and the normalized asymptotic error
+%is roughly a constant
+%[cf. Eq. \eqref{eq:error_singlebin}].
 %
@@
 
-\subsubsection{\label{sec:bandkernel}
-Updating kernel and eigenvalues}
+%\subsubsection{\label{sec:bandkernel}
+%Updating kernel and eigenvalues}
 
@@
 %
-In this way, the matrix, $\mathbf w$,
-being symmetric with respect to the indices, $i$ and $j$,
-satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
-for the flat distribution.
-
-To find the eigenvectors,
-we define, for a periodic variable, $\phi_i$,
-the out-of-boundary values by
+As shown in Appendix \ref{sec:wband_eig},
+the multiplication of $\mathbf w$
+is equivalent to a convolution with the kernel, and
+the eigenvalues of $\mathbf w$ are
+the Fourier or cosine transform of the updating kernel:
+%
+\begin{equation}
+  \lambda_k
+  =
+  \mu_0
+  +
+  2 \,
+  \sum_{ l = 1 }^b
+  \mu_l
+  \cos
+  \frac{ (k - 1) \, l \, 2 \, \pi }
+       {            n             }
+  ,
+  \label{eq:wband_eigenvalue_pbc}
+\end{equation}
+%
+with a two-fold degeneracy,
+$\lambda_{n - k + 2} = \lambda_k$.
+
+For a non-periodic variable,
+we use the reflective boundary condition\cite{bussi2006},
+and change the form of the updating matrix to
+%
+%
+\begin{equation}
+  w_{ij}
+  =
+  \mu_{ i - j }
+  +
+  \mu_{ i + j - 1 }
+  +
+  \mu_{ i + j - 2 n - 1 }
+  ,
+  \label{eq:w_band_refl}
+\end{equation}
+%
+where the last two terms on the right hand side,
+representing two reflective mirrors at
+$j = 1/2$ and $j = n + 1/2$,
+are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
+to the equilibrium distribution\cite{bussi2006},
+such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
+are satisfied.
+%
+Again, the updating kernel should satisfy
+Eqs. \eqref{eq:mu_normalization}
+and \eqref{eq:mu_symm},
+but
+the cutoff $b$ only needs to be less than $n$.
+
+Note that the two reflective mirrors effectively
+define a periodic variable of period $2 \, n$.
+Thus, the eigenvalues can be borrowed from
+Eq. \eqref{eq:wband_eigenvalue_pbc},
+with the period, $n$, doubled:
+%
+\begin{align}
+  \lambda_k
+  &=
+  \mu_0
+  +
+  2
+  \sum_{l = 1}^b
+    \mu_l
+    \cos \frac{(k - 1)  \, l \, 2 \, \pi}{ 2 \, n }
+  .
+\label{eq:wband_eigenvalue_refl}
+\end{align}
+%
+We may unify Eqs. \eqref{eq:wband_eigenvalue_refl}
+and \eqref{eq:wband_eigenvalue_pbc} as
+%
+\begin{equation}
+  \lambda_k
+  =
+  \mu_0
+  +
+  2
+  \sum_{ l = 1 }^b
+    \mu_l \,
+    \cos \frac{ (k - 1) \, l \, g \, \pi }
+              {            n             }
+  ,
+  \label{eq:wband_eigenvalue}
+\end{equation}
+%
+where the degeneracy, $g$, is $2$
+for the periodic case,
+or $1$ otherwise.
+%
+In both cases,
+the eigenvalues, $\lambda_k$'s,
+are the cosine transform of
+the updating kernel.
+
+For the updating scheme to be stable,
+the cosine transform of the kernel
+must be nonnegative.
+%
+If this is not so,
+we can modify the kernel
+using the technique described
+in Appendix \ref{sec:stabilize_wband}.
+
+%\subsubsection{\label{sec:invert_wband}
+%Inversion}
+
+Since the eigenvalues are related to the updating kernel
+by a cosine transform,
+Eq. \eqref{eq:wband_eigenvalue},
+the relation can be readily inverted.
+%
+In the periodic case,
+we have
+%the inversion formula is
+%
+\begin{equation}
+  \mu_l
+  =
+  \frac 1 n
+  \sum_{ k = 1 }^n
+  \lambda_k
+  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
+            {            n             }
+  .
+\label{eq:mu_from_lambda_pbc}
+\end{equation}
+
+In the non-periodic case,
+the formula is
+%
+\begin{align}
+  \mu_l
+  =
+  \frac 1 { 2 \, n }
+  \sum_{ k = 1 }^{ 2 \, n }
+    \lambda_{ k } \,
+    \cos \frac{ (k - 1) \, l \, \pi }
+              {            n        }
+  ,
+\label{eq:mu_from_lambda_refl}
+\end{align}
+%
+where
+we have defined
+$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
+for $k = n + 2, \dots, 2 \, n$,
+as well as
+%
+\begin{align}
+  \lambda_{ n + 1 }
+  =
+  (-1)^{ n - 1 }
+  \lambda_1
+  +
+  2 \, \sum_{ k = 2 }^{ n }
+      (-1)^{n - k} \, \lambda_k
+  ,
+\label{eq:lambdan}
+\end{align}
+to satisfy the constraint, $\mu_n = 0$.
+%
+\note{Eq. \eqref{eq:lambdan}
+  ensures that the $\mu_n$
+  computed from Eq. \eqref{eq:mu_from_lambda_refl}
+  vanishes.
+
+  %The inversion formula is derived from the cosine transform.
+  %
+  If we define $\mu_n = 0$ and
+  %
+  $
+    \mu_{ 2 n - l } = \mu_l,
+  $
+  for
+  $l = 1, \dots, n - 1$,
+  %
+  then Eq. \eqref{eq:wband_eigenvalue_refl} can be rewritten as
+  %
+  \begin{align}
+    \lambda_{k+1}
+    =
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
+  \notag
+  %\label{eq:lambda_cosine_sum}
+  \end{align}
+  %
+  This formula for $\lambda_{k+1}$
+  can be readily extended to $k = 2 \, n - 1$,
+  and we have
+  %
+  $
+    \lambda_{ 2 \, n + 2 - k } = \lambda_k
+    .
+  $
+  %
+  Thus,
+  %
+  $$
+  \begin{aligned}
+    \sum_{ k = 0 }^{ 2 \, n - 1 }
+      \lambda_{ k + 1 } \,
+      \cos \frac{ k \, p \, \pi }
+                {      n        }
+    &=
+    \sum_{ k = 0 }^{ 2 \, n - 1 }
+      \sum_{ l = 0 }^{ 2 \, n - 1 }
+        \mu_l \,
+        \cos \frac{ k \, l \, \pi }
+                  {      n        }
+        \cos \frac{ k \, p \, \pi }
+                  {      n        }
+    \\
+    &=
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+      \frac{ \mu_l } { 2 }
+      \sum_{ k = 0 }^{ 2 \, n - 1 }
+        \cos \frac{ k \, (p + l) \, \pi }
+                  {      n        }
+                  +
+        \cos \frac{ k \, (p - l) \, \pi }
+                  {      n        }
+    \\
+    &=
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+      \mu_l \, n \left(
+        \delta_{ p + l - 2 \, n, 0 }
+        +
+        \delta_{ p - l, 0 }
+      \right)
+    \\
+    &=
+    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
+    =
+    2 \, n \, \mu_p.
+  \end{aligned}
+  $$
+  This entails Eq. \eqref{eq:mu_from_lambda_refl}.
+  %
+  The value of $\lambda_{n + 1}$
+  can be deduced from
+  imposing $\mu_n = 0$
+  and solving Eq. \eqref{eq:mu_from_lambda_refl}
+  for $l = n$,
+  yielding Eq. \eqref{eq:lambdan}.
+}
+
+\subsection{\label{sec:procedure}
+Procedure for computing the optimal schedule and error
+}
+
+Here we give a procedure for computing
+the optimal schedule and the error
+of a given adaptive FDS simulation.
+%
+Basically, we need to extract
+$\lambda_i$'s and $\Gamma_i$'s
+from the given system and updating scheme (input),
+and then apply them to Eq. \eqref{eq:q_opt}
+for the schedule,
+as shown in Fig. \ref{fig:vardep}.
+%
+%The dependence of variables are shown in Fig. \ref{fig:vardep}.
+%
+While the updating scheme alone
+determines $\lambda_k$'s,
+$\Gamma_k$'s depend also on the system
+and the underlying sampling process (MC or MD).
+%
+We have the following procedure.
+
+\begin{figure}[h]
+\begin{singlespace}
+  \fontsize{10pt}{\baselineskip}
+  \begin{tikzpicture}
+  \node[draw, text width={width("Updating ")}]
+    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
+
+  \node[draw, text width={width("sampling")}]
+    (sys) at (0, 1) {System\\and\\sampling\\process};
+
+  \node[draw, text width={width("Eigenvalues,")}]
+    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_2, \dots, \lambda_n$};
+
+  \node[draw, text width={width("Integrals, ")}]
+    (gamma) at (2.2, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
+
+  \node[draw, text width={width("Simulation")}]
+    (T) at (2.2, -0.5) {Simulation\\time, $T$};
+
+  \node[draw, text width={width("Optimal")}]
+    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};
+
+  \node[draw, text width={width("Schedule,")}]
+    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};
+
+    \node[draw, text width={width("Error,")}]
+    (err) at (7, 0) {Error, $\Err(T)$};
+
+  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};
+
+  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};
+
+  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
+  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
+  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};
+
+  \draw[->, >=latex, thick]
+    (w) edge (lambda)
+    (w) edge (gamma)
+    (sys) edge (gamma);
+
+  \draw[->, >=latex, thick]
+    (lambda) -| (M1)
+    (gamma)  -- (M1)
+    (T)      -| (M1)
+    (M1)     -- (R1)
+    (R1)     |- (qT);
+
+  \draw[->, >=latex, thick]
+    (qT)     -| (R3);
+
+  \draw[->, >=latex, thick]
+    (R3) -- (N1) |- (alpha);
+
+  \draw[->, >=latex, thick]
+    (N1) |- (err);
+
+  \draw[->, >=latex, thick]
+    %(M1) edge[bend left=70] (N1);
+    (R1) |- (R2) -| (R3);
+\end{tikzpicture}
+\end{singlespace}
+\caption{
+  \label{fig:vardep}
+  Dependence of variables in computing
+  the optimal schedule and the error.
+}
+\end{figure}
+
+\begin{enumerate}
+
+\item \label{item:lambda}
+From the updating matrix, $\mathbf w$,
+compute the eigenvalues, $\lambda_k$'s.
+
+\item
+Run a preliminary adaptive FDS simulation
+%using the updating scheme
+with a constant updating magnitude, $a_0$.
+
+\item \label{item:Gamma}
+Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}.
+
+\item \label{item:qT}
+Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
+
+\item \label{item:qopt}
+Compute the optimal schedule by
+numerically integrating Eq. \eqref{eq:q_opt}.
+
+\item
+  Compute the error $\Err(T)$ from
+Eqs. \eqref{eq:error_split},
+  \eqref{eq:error_res1},
+  and
+  \eqref{eq:error_asym2}.
+
+\end{enumerate}
+
+Several comments are in order.
+%
+In step \ref{item:lambda},
+one can use Eq. \eqref{eq:wband_eigenvalue}
+for a homogeneous updating scheme.
+For a discrete variable,
+the updating scheme should be stabilize
+using the procedure in Appendix \ref{sec:stabilize_wband}
+to ensure all eigenvalues are nonnegative.
+%
+In step \ref{item:Gamma},
+the details of measuring $\Gamma_k$'s
+can be found in Appendix \ref{sec:Gamma_measure}.
+
+In step \ref{item:qT},
+since $m(Q)$ is a decreasing function of $Q$,
+one can solve Eq. \eqref{eq:opt_qT}
+for a sufficiently small $a_0$
+by the bisection method.
+%
+Starting from the interval $[Q_L, Q_R]$,
+where $m(Q_L) > 2/(T\,a_0) > m(Q_R)$,
+one can iteratively refine the interval
+by substituting $Q_M = (Q_L + Q_R)/2$
+for either $Q_L$ and $Q_R$,
+depending on the sign of $m(Q_M) - 2/(T \, a_0)$.
+
+In step \ref{item:qopt},
+the integrals in Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}
+are done on an even grid of $Q \in [0, q(T)]$
+of $N$ points
+using the standard trapezoidal rule.
+%
+The curve $q(t)$ is then defined on the grid points,
+$(Q_i, t_i)$ ($i = 0, \dots, N$),
+which can be numerically differentiated
+to get the corresponding values of the updating magnitude:
+$\alpha_i = (Q_{i+1} - Q_{i-1}) / (t_{i+1} - t_{i-1})$
+(for the boundary values, we define
+$Q_{-1} = t_{-1} = 0, Q_{N+1} = q(T), t_{N+1} = T$).
+This results in a table of $(t_i, \alpha_i)$
+that can be used for linear interpolation
+to get the updating magnitude at any given time, $t$.
+
+\subsection{\label{sec:cmpschemes}
+Comparison of updating schemes}
+
+Now that we can compute the optimal schedule
+for a given updating scheme,
+we will turn to the problem of finding
+optimal updating schemes below.
+
+\subsubsection{\label{sec:optWL}
+Asymptotic optimality of the single-bin scheme}
+
+We will first show
+that the single-bin scheme is asymptotically optimal.
+%
+Consider a class of $\mathbf w$ matrices
+sharing the same set of eigenvectors,
+hence the same $\Gamma_k$'s,
+but with different positive eigenvalues,
+$\lambda_k$'s.
+%
+Using the Cauchy-Schwarz inequality, we have,
+for any set of nonnegative numbers, $c_k$'s,
+%
+\begin{align}
+&
+\left(
+  \int_0^T dt
+    \sum_{k = 2}^n
+      \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr)
+\right)
+%
+\left(
+  \int_0^T dt
+    \sum_{k = 2}^n
+      \Gamma_k \, c_k^2
+\right)
+%
+\notag
+\\
+&
+\qquad \qquad
+\ge
+\left(
+  \int_0^T dt
+    \sum_{k = 2}^n
+      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t) \bigr)
+\right)^2
+\notag
+\\
+&
+\qquad \qquad
+=
+\left(
+  \sum_{k = 2}^n \Gamma_k \, c_k
+    \left[
+      1 - e^{ -\lambda_k \, q(T) }
+    \right]
+\right)^2.
+\label{eq:CSineq}
+\end{align}
+%
+%\note{
+This inequality follows from the non-negativity of
+the quadratic polynomial,
+$$
+\int_0^T
+  dt \sum_{k = 2}^n \Gamma_k \,
+    \left( \dot u_k \, \chi - c_k \right)^2
+  \equiv
+  A_2 \, \chi^2 + A_1 \, \chi + A_0
+  ,
+$$
+which must carry a non-positive discriminant:
+$A_1^2 - 4 \, A_0 \, A_2 \le 0$.
+%}%
+%
+Note that the last expression of Eq. \eqref{eq:CSineq}
+depends on the curve, $q(t)$ ($0 < t < T$),
+only through the endpoint value, $q(T)$,
+which is fixed in the variational process.
+%
+Thus, the inequality sets a lower bound
+for the asymptotic error $\Err(T)$
+given in Eq. \eqref{eq:error_asym1}.
+%
+The equality is achieved
+if $\dot u_k\left( q(t) \right) = c_k$
+for all $k = 2, \dots, n$ at any $t$,
+up to a multiplicative factor.
+%
+Solving these equations yields
+%
+\begin{equation}
+  \alpha(t) = \frac{              1             }
+                   { \lambda_k \, (t + t_{k0} ) },
+\label{eq:alpha_invtk}
+\end{equation}
+with
+\begin{equation}
+  t_{k0} = \frac{             T            }
+                { e^{ \lambda_k q(T) } - 1 }.
+\label{eq:tk0}
+\end{equation}
+%
+\note{Derivation.
+  Integrating $\dot u_k = c_k$ yields
+  \begin{equation}
+    u_k(t) = c_k \left(t + t_{k0} \right).
+  \tag{UK}
+  \label{neq:ui_solution}
+  \end{equation}
+  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
+  we have
+  $$
+  c_k = \frac{ 1 }{ T + t_{k0} },
+  \quad
+  \mathrm{and\;}
+  \frac{ t_{k0} } { T + t_{k0} }
+  =
+  e^{ -\lambda_k \, q(T) }.
+  $$
+  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
+  $$
+  -\lambda_k \, q(T) + \lambda_k \, q(t)
+  = \ln c_k + \ln \left( t + t_{k0} \right).
+  $$
+  Differentiating this with respect to $t$ yields
+  Eq. \eqref{eq:alpha_invtk}.
+}
+%
+Such a solution is possible only if
+all nontrivial eigenvalues are equal
+%
+\begin{equation}
+  \lambda_2 = \dots = \lambda_n.
+  \notag
+\end{equation}
+%
+Then, by Eq. \eqref{eq:tk0},
+all $t_{k0}$'s also share the same value,
+$t_0$,
+such that
+$u_k = (t + t_0) / (T + t_0)$,
+and
+$c_k = 1/(T + t_0)$
+for $k \ge 2$.
+%
+The asymptotic error is then
+%
+\begin{align}
+  \Err_A(T)
+  =
+  \frac{       T     }
+       { (T + t_0)^2 }
+  \sum_{ k = 2 }^n
+    \Gamma_k
+  .
+\label{eq:error_asym_singlebin}
+\end{align}
+%
+If we further assume $\lambda_2 = \cdots = \lambda_n = 1$,
+%the optimal schedule given by
+Eq. \eqref{eq:alpha_invtk}
+recovers Eq. \eqref{eq:alpha_invt1}.
+%
+
+Since the first eigenmode represents
+a uniform shift of the bias potential,
+we can freely set $\lambda_1$ to be the same as $\lambda_2$
+without changing the nature of the updating scheme.
+%
+Note, an updating matrix
+with identical eigenvalues
+is a multiple of the identity matrix.
+%
+Thus,
+the corresponding updating scheme,
+the single-bin scheme, is most efficient
+in asymptotic convergence.
+%if no eigenvalue $\lambda_k$ is zero.
+%
+In Appendix \ref{sec:upperbound},
+we will show that the opposite limit of
+a set of well-separated eigenvalues
+will provide an upper bound for
+the asymptotic error, Eq. \eqref{eq:error_asym_ubound}.
+
+\subsubsection{\label{sec:optscheme}
+Bandpass updating schemes}
+
+We can generalize
+the single-bin scheme to a class of
+perfect (brick-wall) \emph{bandpass} updating schemes
+that ignore short-wavelength eigenmodes,
+which are typically local noise.
+%
+Mathematically, these updating schemes will have
+zero eigenvalues for the ignored modes,
+and they are useful if we can assume that
+the ignored modes are absent from the PMF.
+%
+For example,
+%if we can assume
+by assuming
+that the PMF is sufficiently smooth
+such that the eigenmodes can be limited to $k \le K$,
+%
+we can construct an updating scheme with
+$$
+\lambda_1 = \cdots = \lambda_K = 1,
+\mathrm{\; and \;}
+\lambda_{K+1} = \cdots = \lambda_n = 0.
+$$
+%
+Then,
+we can change the upper limit of the sum in
+Eq. \eqref{eq:error_asym_singlebin} to $K$:
+%
+%Then,
+%Eq. \eqref{eq:error_singlebin} is changed to
+%
+\begin{equation}
+  \Err_A(T)
+  =
+  \frac {       T     }
+        { (T + t_0)^2 }
+  \sum_{ k = 2 }^K
+    \Gamma_k.
+%\notag
+\label{eq:error_asym_sinc}
+\end{equation}
+%
+The optimal schedule of these schemes
+remains to be the inverse-time schedule
+given by Eq. \eqref{eq:alpha_invt1}.
+%
+Below, we construct these schemes for
+the homogeneous case.
+%
+The definition of the cutoff, $K$, will be modified
+for mathematical convenience.
+
+For a periodic variable, we demand
+%
+$$
+\begin{aligned}
+&
+\lambda_1 = \cdots = \lambda_{K+1}
+= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
+\\
+&
+\lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
+\end{aligned}
+$$
+for $K \ge 1$.
+%
+Then by using
+Eq. \eqref{eq:mu_from_lambda_pbc},
+we get
+\begin{equation}
+  \mu_l
+  =
+  \frac{
+    \sin
+    \dfrac{ (2 K + 1) \, l \, \pi }
+         {              n        }
+  }
+  {
+    n \, \sin \dfrac{ l \, \pi } { n }
+  }
+  .
+\notag
+%\label{eq:mu_sinc_pbc}
+\end{equation}
+\note{Derivation.
+$$
+\begin{aligned}
+\mu_l
+&=
+\frac 1 n \sum_{k = 0}^{n-1} \lambda_{k+1} \cos \frac{ k \, l \, 2 \, \pi } { n }
+\\
+&=
+\frac{1}{n}
+\left(
+  1 +
+  \sum_{k=1}^K
+  \cos \frac { k \, l \, 2 \, \pi } { n }
+  +
+  \sum_{k=n-K}^{n-1}
+  \cos \frac { k \, l \, 2 \, \pi } { n }
+\right)
+\\
+&=
+\frac 1 n
+\sum_{k=-K}^K
+\cos \frac { k \, l \, 2 \, \pi } { n }
+=
+  \frac{
+    \sin
+    \dfrac{ (2 K + 1) \, l \, \pi }
+         {              n        }
+  }
+  {
+    n \, \sin \dfrac{ l \, \pi } { n }
+  }
+.
+\end{aligned}
+$$
+}
+
+For a non-periodic variable, we demand
+$$
+\lambda_1 = \cdots = \lambda_{K+1} = 1,
+\qquad
+\lambda_{K+2} = \cdots = \lambda_n = 0.
+$$
+Then Eq. \eqref{eq:lambdan} gives
+$\lambda_{n+1} = (-1)^{n+K-1}$,
+and from Eq. \eqref{eq:mu_from_lambda_refl},
+we have
+\begin{equation}
+  \mu_l
+  =
+  \frac{1}{2 \, n}
+  \left[
+    (-1)^{n+K+l-1}
+    +
+    \frac{
+      \sin
+      \dfrac{ (2 K + 1) \, l \, \pi }
+           {         2 \, n        }
+    }
+    {
+      \sin \dfrac{ l \, \pi } { 2 \, n }
+    }
+  \right]
+  .
+\label{eq:mu_sinc_refl}
+\end{equation}
+\note{Derivation.
+$$
+\begin{aligned}
+  \lambda_{n+1}
+  &=
+  (-1)^{n-1}
+  \left[
+    \lambda_1
+    - 2 \, \lambda_2
+    + 2 \, \lambda_3 - \cdots
+    + (-1)^{K} 2 \, \lambda_{K+1})
+  \right]
+  \\
+  &=
+  \begin{dcases}
+    (-1)^{n-1} & K \mathrm{\; even,} \\
+    (-1)^n     & K \mathrm{\; odd.}
+  \end{dcases}
+\end{aligned}
+$$
+So
+$$
+\begin{aligned}
+  \mu_l
+  &=
+  \frac{1}{2\,n}
+  \left[
+    1 +
+    2 \sum_{k=2}^{K+1}
+    \cos \frac { (k - 1) \, l \pi } { n }
+    +
+    (-1)^{n+K-1} (-1)^l
+  \right]
+  \\
+  &=
+  \frac{1}{2\,n}
+  \left[
+    (-1)^{n+K+l-1}
+    +
+    \sum_{k=-K}^{K}
+    \cos \frac { k \, l \pi } { n }
+  \right]
+  .
+\end{aligned}
+$$
+}%
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
+  }
+  \caption{
+    \label{fig:sinc}
+    Updating kernels of the bandpass updating schemes
+    compared to a Gaussian with a similar width.
+    %
+    The lines are a guide to the eyes.
+    Inset. Updating matrix in the non-periodic case.
+    %
+    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
+      The data were computed by the program \texttt{prog/predict},
+      see particularly the function \texttt{intq\_save()}
+      in \texttt{prog/intq.h} for panel (a),
+      and the functions \texttt{savewin()}
+      and \texttt{savewinmat()}
+      in \texttt{prog/invt.h} for panel (b),
+      The data were saved under the directory \texttt{data/sinc}.
+      See the \texttt{make} object \texttt{sincsingle}
+      in the \texttt{Makefile} there.
+    }%
+  }
+\end{center}
+\end{figure}
+
+In Fig. \ref{fig:sinc},
+we give examples of the updating kernels
+of the bandpass updating schemes,
+compared to a Gaussian
+with a similar width, $\sigma$,
+%
+\begin{equation}
+  \sigma
+  \approx
+  \frac
+  {
+    n
+  }
+  {
+    \sqrt{ 2 \, \pi } \, g \, K
+  }
+  .
+\label{eq:sigma_equiv}
+\end{equation}
+%
+%where $g = 2$ for the periodic case
+%or $g = 1$ otherwise.
+%
+The kernel in the non-periodic case is a sawtooth
+because of the term, $(-1)^{n+K+l-1}$,
+in Eq. \eqref{eq:mu_sinc_refl}.
+%
+However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
+the ruggedness disappears
+in the updating matrix by Eq. \eqref{eq:w_band_refl},
+as the above oscillatory term for $l = i - j$
+is canceled by the term for either $l = i + j - 1$
+or $l = i + j - 2 \, n - 1$
+with the opposite sign.
+
+\subsubsection{\label{sec:finlencorr}
+Finite-length correction}
+
+In the above section,
+we have focused on
+the asymptotic error.
+%
+By including the residual error
+and minimizing the total error,
+we get the optimal $t_0$,
+which is, according to Eq. \eqref{eq:half_alpha0},
+%
+%%Assuming the system is initially equilibrated
+%%under a constant updating magnitude,
+%%$a_0$,
+%Assuming Eq. \eqref{eq:error_res1} with
+%$$
+%q(T)
+%=
+%\int_0^T \frac{ dt } { t + t_0 }
+%=
+%\ln \frac{ T + t_0 } { t_0 }.
+%$$
+%%
+%we have, from Eqs. \eqref{eq:error_split}
+%and \eqref{eq:error_asym_sinc},
+%%
+%\begin{equation}
+%  \Err
+%  =
+%  \Err_R + \Err_A
+%  =
+%  \frac { \frac 1 2 a_0 \, t_0^2 + T }
+%        {          ( T + t_0 )^2          }
+%  \sum_{ k = 2 }^K \Gamma_k
+%  .
+%\notag
+%%\label{eq:error_sinc0}
+%\end{equation}
+%%
+%The minimum is reached at
+%
+\begin{equation}
+  t_0
+  =
+  \frac{  2  }
+       { a_0 }
+  .
+\label{eq:t0_sinc}
+\end{equation}
+%
+For the single-bin updating scheme,
+it recovers Eq. \eqref{eq:t0_singlebin}
+under the assumption of Eq. \eqref{eq:y2_eql}.
+%
+From Eqs. \eqref{eq:error_split},
+\eqref{eq:error_res1},
+\eqref{eq:tk0},
+and
+\eqref{eq:error_asym_sinc},
+we get
+%
+\begin{equation}
+  \Err(T)
+  =
+  \frac{   1     }
+       { T + t_0 }
+  \sum_{ k = 2 }^K
+    \Gamma_k
+  .
+\label{eq:error_sinc}
+\end{equation}
+%
+which serves as a generalization of
+the result from the single-bin scheme, Eq. \eqref{eq:Emin_singlebin}.
+%
+This error decreases with increasing simulation length,
+as one would expect from an equilibrium FDS simulation.
+%In Appendix \ref{sec:equilerr},
+%we show that the above error tends
+%to the error of the ideal equilibrium FDS simulation
+%in the asymptotic limit.
+
+%
+%For example, consider a system that is initially
+%prepared under a fixed $a_0$
+%at the beginning of the schedule $\alpha(t)$.
+%%
+%Then for shorter times, $T \approx 0$,
+%the error, dominated by the residual error
+%given by Eq. \eqref{eq:error_res1},
+%is clearly smaller if some eigenvalues
+%are less than $1.0$ (the single-bin case).
+%
+%In practice, real simulations
+%may have long achieved the desired precision
+%before reaching the asymptotic regime,
+%especially if the updating matrix
+%has some $\lambda_k$'s close to $0$.
+
+\section{\label{sec:results}
+Numerical results}
+
+We tested the above theory on a model system
+described in Sec. \ref{sec:results_system}.
+%
+We verify the method of computing errors
+%using an inverse-time schedule
+in Sec. \ref{sec:results_invt}
+%
+and study the optimal schedule from
+Eq. \eqref{eq:q_opt}
+in Sec. \ref{sec:results_optschedule}.
+%
+Finally, we compare the Gaussian and
+bandpass updating schemes %of different widths
+in Sec. \ref{sec:results_cmpschemes}.
+
+\subsection{\label{sec:results_system}
+Test system and simulation protocol}
+
+Our test system is a one-dimensional model system, in which
+$i$ represents a non-periodic variable,
+and $n = 100$.
+%
+Both the target and intrinsic distributions
+are assumed to be flat:
+$p^*_i = p_i = 1/n$,
+along with the distribution, $\rho_i$, in Eqs.
+\eqref{eq:vbar_def} and \eqref{eq:error_sum}.
+%
+%We wish to see how
+%the schedule, $\alpha(t)$,
+%affects the fluctuating error
+%of the bias potential
+%at the end of the simulation.
+%
+Initially,
+we set the bias potential to zero,
+$V_i \equiv 0$,
+and equilibrate the system for $10^7$ steps
+under a constant updating magnitude,
+$a_0 = 10^{-4}$.
+%
+This equilibration process models
+a stage of the WL algorithm or
+a regular metadynamics
+with a fixed updating magnitude.
+%
+Although the bias potential is correct initially,
+it would deviate from this value
+at the end of the equilibration.
+%
+Our aim is to see how well
+the bias potential is restored
+at the end of the testing schedule.
+%
+After the equilibration,
+we reset the origin of the time, $t$, to zero,
+and start the testing schedule
+for $T = 10^8$ steps.
+%
+We repeat the above process $1000$ times (unless specified otherwise)
+and report the averaged results.
+%specified by either
+%Eq. \eqref{eq:q_opt}
+%or
+%Eq. \eqref{eq:alpha_invtlambda}.
+
+For the Gaussian updating scheme,
+the kernel was truncated at
+$b = \min\{10 \, \sigma, n - 1\}$ bins
+and numerically stabilized
+using the technique
+in Appendix \ref{sec:stabilize_wband}.
+
+%Since the error depends on
+%the underlying sampling process,
+We tested two sampling processes.
+%a global one and a local one.
+%
+Both are based on
+a Metropolis MC algorithm\cite{
+  metropolis1953, newman, frenkel,
+  landau_binder}:
+%
+in each step, a new bin index $j$ is proposed
+and then accepted with probability
+%
+$
+A(i \to j) = \min\{ 1, \exp(V_i - V_j) \}.
+$
+However,
+in the first or global sampling process,
+we choose the destination $j$
+uniformly out of the $n$ bins.
+%
+In the second or local sampling process,
+$j$ is either $i - 1$ or $i + 1$
+(adjusted by the boundary condition)
+with equal probability $1/2$.
+%
+Asymptotically,
+the above global and local sampling processes
+emulate the perfect and one-step processes
+discussed in Appendix \ref{sec:Gamma},
+respectively,
+and we will use
+Eqs. \eqref{eq:Gamma_perfect}
+and \eqref{eq:Gamma_onestep}
+for the respective $\Gamma_k$'s
+in the theoretical predictions.
+%
+
+\subsection{\label{sec:results_invt}
+Modified inverse-time schedule}
+
+Using the model system above, we
+tested our analysis of computing errors
+on the modified inverse-time schedule,
+%
+\begin{equation}
+\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
+\label{eq:alpha_invtlambda}
+\end{equation}
+%
+where $\lambda$ is a free parameter,
+and $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
+%and $t_0 = 2/a_0$.
+%
+The error of this schedule
+can be computed analytically
+(cf. Appendix \ref{sec:invt_schedule}),
+and tested against simulation.
+
+We performed adaptive FDS simulations
+using the single-bin updating scheme
+and the triple-bin updating scheme with $\mu_1 = 0.24$.
+%
+%For each value of $\lambda$,
+%we ran $1000$ independent simulations
+%of $10^8$ steps, and averaged the results.
+%
+As shown in Fig. \ref{fig:errinvt},
+the errors from the simulations
+agreed well with those from the theory.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/errinvt.pdf}
+  }
+  \caption{
+    \label{fig:errinvt}
+    Error, $\Err$, versus the parameter, $\lambda$,
+    for the modified inverse-time schedule,
+    Eq. \eqref{eq:alpha_invtlambda}.
+    %
+    The points are results from averaging over $1000$ independent runs;
+    the curves are theoretical predictions.
+    %
+    For the single-bin (WL) updating scheme,
+    the minimal should occur at $\lambda = 1$,
+    while it is not necessarily the case for a general updating scheme.
+    %
+    \note{The figure was produced by \texttt{doc/fig/errinvt.gp}.
+      The data were produced by the script \texttt{data/invtcscan.py},
+      and saved under \texttt{data/invt}
+      (see the \texttt{Makefile} there).
+    }%
+  }
+\end{center}
+\end{figure}
+
+For the single-bin updating scheme,
+the optimal $\lambda$ occurred around $1.0$
+for both global and local sampling processes.
+%
+The sampling process affected the error function
+only by a multiplicative constant,
+%The error functions of the two sampling processes
+%differed only by a multiplicative constant,
+in agreement with the discussion in
+Appendix \ref{sec:invt_singlebin}.
+%
+For the triple-bin updating scheme,
+the optimal $\lambda$ was less than $1.0$
+for the global sampling process,
+although it stayed around $1.0$
+for the local sampling process,
+also in agreement with the discussion
+in Appendix \ref{sec:invt_nn}.
+%
+Thus, the inverse-time schedule,
+Eq. \eqref{eq:alpha_invt1},
+is optimal for the single-bin updating scheme,
+but not necessarily so for a general multiple-bin updating scheme.
+%
+
+\subsection{\label{sec:results_optschedule}
+Optimal schedule}
+
+To demonstrate the optimal schedule from Eq. \eqref{eq:q_opt},
+we used the Gaussian (with $\sigma = 10$ bins) updating scheme
+as an example.
+%
+This updating scheme is typically used in metadynamics simulations.
+%
+In Table \ref{tab:error_Gaussian},
+we showed that the optimal schedule
+produced smaller errors than
+the inverse-time schedules.
+%
+However, the optimal schedules
+decayed more slowly than
+the inverse-time formula, Eq. \eqref{eq:alpha_invt1},
+and depended on the sampling process,
+as shown in Fig. \ref{fig:optacmp}.
+%
+The impeded decay of the optimal schedule
+was necessary in order to
+damp out the short-wavelength eigenmodes
+(i.e. local noise).
+%
+As shown in Fig. \ref{fig:xerr},
+the inverse-time schedule was effective in error reduction
+only for the first few eigenmodes,
+whose eigenvalues were around $1.0$.
+%
+We can, however, ignore these local eigenmodes
+by explicitly filtering them out from the final bias potential,
+and truncating the error sum in Eq. \eqref{eq:error_sum}
+to the first few eigenmodes.
+%
+Then, the corresponding optimal schedule
+from Eq. \eqref{eq:q_opt} that minimizes the truncated error
+became similar to the inverse-time schedule,
+Eq. \eqref{eq:alpha_invt1},
+as exemplified in Fig. \ref{fig:optacmp}.
+%
+\note{A more thorough approach is
+to use the bandpass updating scheme
+to be discussed later.
+}
+
+\begin{table}[h]\footnotesize
+  \caption{\label{tab:error_Gaussian}
+    Error of the Gaussian updating scheme of $\sigma = 10$
+    under different updating schedules.
+    In Eqs. \eqref{eq:alpha_invt1} and \eqref{eq:alpha_invtlambda},
+    $t_0$ is given by Eq. \eqref{eq:t0_sinc};
+    and in Eq. \eqref{eq:alpha_invtlambda},
+    the optimal value of $\lambda$ is used.
+    \note{\newline
+      The data were collected in
+      \texttt{data/tab1} by running
+      the program \texttt{invt}.
+    }%
+  }
+  \setlength{\tabcolsep}{2pt}
+  \renewcommand\arraystretch{1.2}
+  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
+    \hline
+      Schedule
+    &
+      Inverse-time, \newline
+      Eq. \eqref{eq:alpha_invt1}
+    &
+      Modified, \newline
+      Eq.~\eqref{eq:alpha_invtlambda}
+    &
+      Optimal, \newline
+      Eq. \eqref{eq:q_opt}
+    \\
+    \hline
+    Global
+    &
+    $4.36(9) \times 10^{-6}$
+    &
+    $7.1(1) \times 10^{-7}$
+    &
+    $2.6(1) \times 10^{-7}$
+    \\
+    \hline
+    Local
+    &
+    $3.83(7) \times 10^{-4}$
+    &
+    $1.72(4) \times 10^{-4}$
+    &
+    $9.1(2) \times 10^{-5}$
+    \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/optacmp.pdf}
+  }
+  \caption{
+    \label{fig:optacmp}
+    Optimal schedules from Eq. \eqref{eq:q_opt}
+    for the Gaussian updating scheme
+    with $\sigma = 10$.
+    %
+    The mode-limited schedules were produced by
+    limiting the sum in Eq. \eqref{eq:q_opt}
+    to the first four terms,
+    or the four longest wavelength modes.
+    %
+    They became similar to the inverse-time schedule,
+    Eq. \eqref{eq:alpha_invt1}.
+    %
+    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
+      The data were computed by the program \texttt{prog/predict},
+      see particularly the function \texttt{intq\_save()}
+      in \texttt{prog/intq.h}.
+      The data were saved under the directory \texttt{data/opta}.
+    }%
+  }
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/xerr.pdf}
+  }
+  \caption{
+    \label{fig:xerr}
+    Predicted error components for the Gaussian updating scheme
+    with $\sigma = 10$
+    from the optimal schedule
+    and the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
+    %
+    The latter schedule failed to damp out the errors
+    from the long-wavelength (local) eigenmodes
+    with $k \ge 5$.
+    %
+    For longer simulations,
+    the error components from the optimal schedule
+    would roughly maintain a relatively flat profile,
+    but meet the initial error curve
+    at a lower level (not shown),
+    where $\lambda_k \approx 1/q(T)$
+    (cf. Appendix \ref{sec:upperbound}).
+    %
+    \note{This figure was produced by \texttt{doc/fig/xerr.gp}.
+      The data were saved under the directory \texttt{data/xerr}.
+    }%
+  }
+\end{center}
+\end{figure}
+
+An interesting prediction from the theory is Eq. \eqref{eq:half_alpha0},
+which states that the optimal schedule should always start from
+half of the magnitude from the previous equilibration run.
+%
+To verify this result,
+we computed the optimal schedules
+for different values of $q(T)$
+for the Gaussian updating scheme,
+and plotted the final error
+against the initial updating magnitude, $\alpha(0)$.
+%
+As shown in Fig. \ref{fig:erriascan},
+the minimal error was indeed reached at
+$\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
+although the dependence on $\alpha(0)$
+is generally weak.
+%
+The weak dependence is best illustrated in
+the case of single-bin updating scheme,
+as the initial updating magnitude only
+affects the schedule by a shift of the origin of
+simulation time in the inverse-time schedule,
+Eq. \eqref{eq:alpha_invt1}.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/erriascan.pdf}
+  }
+  \caption{
+    \label{fig:erriascan}
+    Relative error
+    versus the initial updating magnitude, $\alpha(0)$,
+    for optimal schedules with different $q(T)$.
+    Assuming the Gaussian updating scheme
+    with $\sigma = 10$,
+    and $a_0 = 10^{-4}$.
+    %
+    The points were obtained from averaging results
+    from over 10000 independent simulations;
+    the curves are theoretical predictions.
+    %
+    The minimal should always occur at
+    $\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
+    as predicted by Eq. \eqref{eq:half_alpha0}.
+    %
+    \note{This figure was produced by
+      \texttt{doc/fig/erriascan.gp}.
+      The data were saved under
+      \texttt{data/iascan}.
+    }%
+  }
+\end{center}
+\end{figure}
+
+\subsection{\label{sec:results_cmpschemes}
+Optimal updating schemes}
+
+For a finite-length simulation,
+the optimal updating scheme
+depends on the simulation length, $T$,
+and other parameters.
+%
+As an example,
+we compare Gaussian updating schemes
+with different widths, $\sigma$.
+%
+In Fig. \ref{fig:errsigscan},
+we present the errors, $\Err(T)$,
+normalized by the shifted simulation length, $T+t_0$,
+so that they could
+stay constant in the limit of single-bin updating scheme
+($\sigma \to 0$).
+%
+As shown in Fig. \ref{fig:errsigscan}(a)
+that the minimal error was achieved
+by the widest window %for the shorter simulation of $T = 10^8$,
+for the global sampling process,
+%
+although
+the minimum at $\sigma = 0$ became more favorable
+as the simulation lengthens.
+%
+However,
+for the local sampling process,
+the minimum stayed at zero width,
+as shown in Fig. \ref{fig:errsigscan}(b).
+
+In contrast,
+the normalized errors of the bandpass updating schemes
+were insensitive to the simulation length,
+in agreement with Eq. \eqref{eq:error_sinc}.
+%
+The bandpass updating schemes
+yielded smaller errors than
+the corresponding Gaussian ones,
+although the differences tended to
+be small for shorter simulations
+with the local sampling process.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsigscan.pdf}
+  }
+  \caption{
+    \label{fig:errsigscan}
+    Normalized error, $(T + t_0) \, \Err(T)$,
+    versus the width, $\sigma$,
+    for the Gaussian
+    and the bandpass updating schemes,
+    with $t_0$ given by Eq. \eqref{eq:t0_sinc}.
+    %
+    In the latter case,
+    the equivalent width
+    $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
+    %
+    The points were obtained from averaging results
+    from 1000 independent simulations;
+    the curves are theoretical predictions.
+    %
+    \note{This figure was produced by
+      \texttt{doc/fig/errsigscan.gp}.
+      The data were saved under
+      \texttt{data/scan}.
+    }%
+  }
+\end{center}
+\end{figure}
+
+\section{\label{sec:conclusion}
+Conclusions and Discussions}
+
+In conclusion,
+we have proposed a method of computing
+the optimal schedule of the updating magnitude
+for a general class of free energy calculation methods,
+which we refer to as adaptive flat-distribution-sampling (FDS) methods
+that include the Wang-Landau (WL) algorithm and metadynamics.
+%
+Adaptive FDS methods
+incrementally update a bias potential
+along a collective variable
+to offset the potential of mean force (PMF)
+and thereby sample a flat distribution.
+%
+The optimal schedule delivers the fastest convergence
+of the bias potential,
+and thus helps minimize the effort
+of computing the PMF.
+%They differ by the updating schemes,
+%which specify how the bias potential is modified
+%in each updating step.
+%%
+%In the single-bin or WL case,
+%the update is limited to the current bin;
+%while metadynamics,
+%the bias potential of a neighborhood of the current one
+%is updated with relative weight specified by a Gaussian window.
+
+The optimal schedule depends on the updating window function,
+which is limited to a single-bin in the WL case,
+but can be extended to several neighboring bins
+taking a Gaussian shape in the metadynamics case.
+%
+More generally,
+each adaptive FDS method is associated with
+an updating scheme
+that can be mathematically characterized
+by an updating matrix.
+%
+Each column of the matrix gives the normalized
+window function for the bin corresponding to the column.
+%
+Our method computes the optimal schedule from
+the eigenvalues of the updating matrix,
+and the integrals of the autocorrelation functions
+of the eigenmodes.
+%
+For the single-bin (WL) scheme,
+the optimal schedule from our method
+is given by the inverse-time formula,
+Eq. \eqref{eq:alpha_invt1}.
+%
+For a general multiple-bin scheme,
+including the Gaussian one used in metadynamics,
+the optimal schedule is not given by a simple closed form,
+but is implicitly given by an equation of motion
+of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
+%
+In this case,
+the optimal schedule and error
+can be sensitive to the simulation length.
+%
+The initial updating magnitude is always
+half of the previous equilibrium value.
+% the value used during equilibration.
+
+We have also shown that
+the single-bin scheme is optimal
+in the long time limit
+without a priori assumptions
+about the smoothness of the PMF.
+%
+However, if some eigenmodes
+correspond to local noise
+that is expected to be absent from the PMF,
+we can construct a
+class of bandpass updating schemes
+that are also optimal for asymptotic convergence.
+%
+The optimal schedule of these bandpass schemes
+is always given by an inverse-time formula.
+
+Although the bandpass updating schemes
+appeared to outperform similar Gaussian schemes
+in our model study,
+they are global in nature
+and hence more expensive.
+%
+However, the performance gain
+is relatively small for short simulations
+with a local sampling process such as MD.
+%
+\note{
+One way to improve the Gaussian updating schemes
+is to use it with a hard cutoff of eigenmodes
+(as we did for Fig. \ref{fig:optacmp}).
+%
+Then the optimal schedule
+would be similar to the inverse-time one,
+and the error, with the short-wavelength eigenmodes filtered,
+would be close to that of the bandpass scheme.
+}
+
+The analysis in this study has two limitations.
+%
+First, we have ignored the \emph{systematic}
+error\cite{zhou2005, morozov2007, zhou2008}
+caused by the updates in adaptive FDS methods.
+%
+This is valid in the asymptotic regime:
+when the updating magnitude is small
+and the sampling is a quasi-equilibrium one\cite{
+  zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
+the random error,
+which is roughly proportional to
+the square root of the updating magnitude\cite{
+  zhou2005, morozov2007, zhou2008, bussi2006},
+can easily outweigh
+the systematic error,
+which is roughly proportional to
+the magnitude itself\cite{morozov2007}.
+%
+%Our method of analysis is most useful for long simulations
+%to achieve relatively precise results.
+%
+Second, by using the white noise approximation,
+we have represented the autocorrelation function
+of each eigenmode by a single number, $\Gamma_k$.
+%
+Although this assumption appeared to work well
+on our test system for modeling long simulations,
+they may break down for relatively short simulations
+of glassy systems.
+%Our immediate interest is to apply these relations to
+%problems containing aqueous mixtures of proteins and nucleic acids.
+
+\section{Acknowledgments}
+
+We thank Dr. Y. Mei, J. Wang,
+O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
+Dr. C. Myers, and Dr. O. Samoylova
+for helpful discussions.
+%
+We gratefully acknowledge the Robert A. Welch Foundation (H-0037),
+the National Science Foundation (CHE-1152876)
+and the National Institutes of Health (GM-037657)
+for partial support of this work.
+%
+JM thanks support from the National Institutes of Health (R01-GM067801, R01-GM116280),
+the Welch Foundation (Q-1512),
+and National Science Foundation (Grant PHY-1427654).
+%
+This research used computing resources of the National Science Foundation XSEDE grid.
+%
+%
+
+\appendix
+
+\section{\label{sec:equilerr}
+Inverse-time schedule as a result of increasing sample size
+in a runtime average
+}
+
+We can understand the optimality of the inverse-time schedule
+as a natural consequence of the stabilization of a runtime average
+as the sample size increases\cite{
+  marsili2006, barducci2008}.
+%
+Consider a sample, $\{X^{(1)}, \dots, X^{(N)}\}$,
+with a dynamically growing size.
+%
+The runtime average
+%
+\begin{equation*}
+  {\bar X}^{(N)}
+  =
+  \frac{ X^{(1)} + \cdots + X^{(N)} } { N }
+  ,
+\end{equation*}
+%
+can be written as a recurrence relation regarding
+the addition of the last data point,
+%
+\begin{align}
+  {\bar X}^{(N)}
+  &=
+  {\bar X}^{(N-1)} + \frac{ X^{(N)} - {\bar X}^{(N-1)} } { N }
+  .
+  \label{eq:av_recur}
+\end{align}
+%
+If we interpret $X^{(N)} - {\bar X}^{(N-1)}$
+as the independent correction to the previous average,
+${\bar X}^{(N-1)}$, proposed by the latest data point, $X^{(N)}$,
+then this correction is absorbed into the average with
+a diminishing weight $1/N$ as the sample size, $N$, increases.
+
+The structural similarity between the recurrence relation,
+Eq. \eqref{eq:av_recur},
+and the WL updating scheme, Eq. \eqref{eq:wl_update},
+suggests that the bias potential may represent
+a runtime average in a particular case.
+%
+To establish a mapping between the two,
+let us now consider a long equilibrium FDS simulation
+that consists of a sequence of sub-simulations
+of $L$ steps.
+%
+If $X^{(N)}$ is
+the bias potential, $V_i^{(N)}$, computed
+from the $N$th sub-simulation,
+then ${\bar X}^{(N)}$ means the average bias potential,
+${\bar V}_i^{(N)}$,
+from the first $N$ sub-simulations.
+%
+Since this average, ${\bar V}_i^{(N)}$,
+gives the best estimate at the end of
+the first $N$ sub-simulations,
+we should ideally use it as the bias potential
+for the next sub-simulation,
+until it is further updated by Eq. \eqref{eq:av_recur}
+at the end of the next sub-simulation.
+%
+Assuming that the normalized histogram, $H_i^{(N)}$,
+from the $N$th sub-simulation is sufficiently precise,
+Eq. \eqref{eq:vcorr_equil}
+\big[with
+$V_i^\mathrm{corr} \to V_i^{(N)}$, $V_i \to {\bar V}_i^{(N-1)}$\big]
+shows that the independent
+correction from the $N$th sub-simulation is
+%
+\begin{equation*}
+  \ln \frac{ H_i^{(N)} } { p_i }
+  \approx
+  \frac{ H_i^{(N)} } { p_i } - 1
+  ,
+\end{equation*}
+%
+where the ``$-1$'' on the right-hand side
+can be ignored as a uniform shift
+that does not affect the distribution.
+%
+Thus, we may rewrite Eq. \eqref{eq:av_recur} as
+\begin{align}
+  {\bar V}_i^{(N)}
+  &=
+  {\bar V}_i^{(N-1)}
+  +
+  \frac{1}{N} \frac{ H_i^{(N)} }{ p_i }
+  \notag
+  \\
+  &=
+  {\bar V}_i^{(N-1)}
+  +
+  \sum_{l=1}^L \frac{1}{N \, L} \frac{ h_i^{(N)}(l) }{ p_i }
+  ,
+  \label{eq:vbar_update}
+\end{align}
+%
+where $h_i^{(N)}(l)$ stands for the instantaneous histogram
+at step $l$ of sub-simulation $N$,
+and its average is
+$H_i^{(N)} \equiv \frac{1}{L} \sum_{l=1}^L h_i^{(N)}(l)$.
+%
+Now Eq. \eqref{eq:vbar_update} shows that the best estimate
+of the bias potential from the runtime average can be computed
+through a step-by-step incremental updating scheme that is equivalent to
+the WL algorithm, Eq. \eqref{eq:wl_update},
+with the updating magnitude being
+%
+\begin{equation}
+  \alpha(t) = \frac{1}{N \, L}
+  .
+  \label{eq:alpha_block}
+\end{equation}
+%
+In the long time limit,
+we may approximate the simulation time
+by the product of the number of sub-simulations
+and the number of steps in each sub-simulation, or $N \, L$,
+%
+and Eq. \eqref{eq:alpha_block} recovers Eq. \eqref{eq:alpha_invt}.
+
+The above demonstration implies several connections
+between the WL algorithm under the optimal inverse-time schedule
+with a few related FDS methods.
+%
+First,
+it establishes an efficiency equivalence
+between an adaptive WL simulation under the inverse-time schedule
+and the ideal equilibrium FDS counterpart
+with a nearly perfect bias potential.
+%
+The bias potentials from the two approaches
+would reach similar precision at long times.
+%
+Thus, it is unnecessary to convert
+a sufficiently convergent adaptive FDS simulation
+to the corresponding equilibrium FDS simulation
+with a fixed bias potential just for better precision
+of the final PMF.
+
+Second,
+it also relates a WL simulation under the inverse-time schedule
+to a time-averaged variant of the regular WL simulation.
+%
+In the variant, the updating magnitude is fixed,
+but the bias potential over the simulation course is averaged
+as the final result\cite{zhou2005}.
+%
+The above demonstration shows that the inverse-time schedule
+has already effectively implemented an implicit average
+so that no explicit averaging is needed in the former approach.
+%
+The inverse-time schedule approach is preferred, however,
+because with a decreasing updating magnitude,
+it minimizes the systematic error caused by
+the on-the-fly updating scheme\cite{
+  zhou2005, zhou2008},
+which is ignored in this analysis.
+
+Third,
+it is possible to reformulate the WL algorithm
+under the inverse-time schedule
+as an explicit adaptive average method
+such that the bias potential is computed from
+a weighted runtime average of the histogram\cite{
+  marsili2006, barducci2008},
+which is analogous to the adaptive bias force method\cite{
+  darve2001, *darve2008}.
+
+\section{\label{sec:wltoinvt}
+Inverse-time schedule as the optimal limit of the WL algorithm}
+
+We may understand the inverse-time schedule %, Eq. \eqref{eq:},
+as an optimal limit of the stage-wise WL updating scheme.
+%
+%Our calculation below will focus on the one-variable problem,
+%as in Sec. \ref{sec:onevar}.
+
+To begin, we split the entire simulation into $S$ stages,
+each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
+%
+We denote the number of steps in stage $s$ as $T_s$, and
+$\sum_{s=1}^S T_s = T$.
+%
+%\begin{equation}
+%  \sum_{s = 1}^S T_s = T
+%  ,
+%  \label{eq:sumT}
+%\end{equation}
+%
+We assume that
+the magnitude follows geometrically,
+%
+\begin{equation}
+  \alpha_s = \alpha_0 \, r^s
+  ,
+  \label{eq:alphas}
+\end{equation}
+%
+with $0 < r < 1$.
+%
+%and that the system is properly equilibrated
+%under a constant magnitude initially.
+%
+We wish to find the best distribution of the simulation time
+into the $S$ stages
+as well as
+the optimal number of stages and $r$.
+
+If we denote the error at the end of stage $s$ as $E_s$,
+then from Eq. \eqref{eq:ET_average}, we have
+$q(t) = \alpha_s \, t$,
+where $t$ denotes the number of steps in stage $s$.
+%
+Then
+%
+\begin{align}
+  E_s
+  &=
+  E_{s - 1} \, e^{-2 \, \alpha_s \, T_s }
+  +
+  \Gamma \, \int_0^{T_s}
+  \alpha_s^2 \,
+  e^{ -2 \, \alpha_s \, (T_s - t) }
+  \, dt
+  \notag
+  \\
+  &=
+  \left(
+    E_{s - 1} - \frac{ \Gamma \, \alpha_s } { 2 }
+  \right)
+  e^{ - 2 \, \alpha_s \, T_s }
+  +
+  \frac{ \Gamma \, \alpha_s } { 2 }
+  .
+  \label{eq:errorstage}
+\end{align}
+%
+Our aim is to minimize the error
+of the final stage, $E_S$.
+%
+To do so, let us consider shifting simulation time
+by $\delta T$ from
+stage $s+1$ to stage $s$,
+then $E_{s+1}$ is affected as
+%
+\begin{align*}
+  \delta E_{s+1}
+  =
+\left[
+  2 \, \alpha_{s+1}
+  \left(
+    E_s - \tfrac{ \Gamma \, \alpha_{s+1} } { 2 }
+  \right)
+  +
+  \frac{ \partial E_s } { \partial T_s }
+\right]
+e^{-2 \, \alpha_{s+1} \, T_{s+1} }
+\,
+\delta T
+.
+\end{align*}
+%
+Since this variation does not directly affect the error of later stages,
+the above expression must be zero to minimize the final error, i.e,
+%
+\begin{align*}
+  2 \, \alpha_{s+1}
+  \left(
+    E_s - \frac{ \Gamma \, \alpha_{s+1} } { 2 }
+  \right)
+  &=
+  -\frac{ \partial E_s } { \partial T_s }
+  %\\
+  %&=
+  %2 \, \alpha_s \, e^{-2 \, \alpha_s \, T_s}
+  %\left(
+  %  E_{s-1} - \frac{ \Gamma \, \alpha_s } { 2 }
+  %\right)
+  %\\
+  =
+  2 \, \alpha_s \, \left(E_s - \frac{ \Gamma \, \alpha_s } { 2 }  \right)
+  ,
+\end{align*}
+%
+where we have used Eq. \eqref{eq:errorstage}
+in the second step.
+%
+Thus,
+\begin{equation}
+  E_s
+  =
+  \frac{ \Gamma } { 2 } \, ( \alpha_s + \alpha_{s+1} )
+  =
+  \frac{ \Gamma } { 2 } \, \alpha_s ( 1 + r )
+  .
+  \label{eq:Es}
+\end{equation}
+%
+Using Eqs. \eqref{eq:Es} and \eqref{eq:alphas}
+in Eq. \eqref{eq:errorstage}, we get
+%
+\begin{equation}
+  e^{-\alpha_s \, T_s } = r
+  .
+  \label{eq:exp_r}
+\end{equation}
+%
+
+To compute the optimal schedule,
+we map the overall simulation time, $t^{(s)}$,
+to the stage number, $s$, as
+%
+\begin{align*}
+  t^{(s)}
+  = \sum_{s' = 1}^s T_{s'}
+  = \frac{ 1 - r^s } { 1 - r } \, T_s
+  = \frac{ 1 } { \gamma } \,
+  \left(
+    \frac{1}{\alpha_s}
+    -
+    \frac{1}{\alpha_0}
+  \right)
+  ,
+\end{align*}
+where
+$\gamma = ( 1 - r ) / \ln(1/r)$,
+and we have used Eq. \eqref{eq:exp_r} in the last step.
+%
+Thus,
+%
+\begin{equation}
+  \alpha_s =
+  \frac{ 1 }
+  { \gamma \, t^{(s)} + \alpha_0^{-1} }
+  ,
+  \label{eq:alphas_t}
+\end{equation}
+%
+and the final error is given by
+%
+\begin{align}
+  E_S
+  =
+  \frac{\Gamma}{2}
+  \frac{ 1 + r }
+  { \gamma \, T + \alpha_0^{-1} }
+  ,
+  \notag
+\end{align}
+%
+which is minimized as $r \to 1$
+for a large $T$
+(at this point, $\gamma \to 1$).
+%
+\note{
+  For a large $T$, $E_S \approx \frac{\Gamma}{2T} \frac{1+r}{\gamma}$.
+  Define $\delta = 1 - r$,
+  we have $1/\gamma = 1+ \delta/2 + \delta^2/3 + \cdots$,
+  and
+  $$
+  \frac{1+r}{\gamma}
+  =
+  (2 - \delta)
+  \left(
+  1 + \frac{\delta}{2} + \frac{\delta^2}{3} + \cdots
+  \right)
+  =
+  2 + \sum_{n=1}^\infty \frac{n-1}{n(n+1)}\delta^n \ge 2
+  .
+  $$
+}
+%
+If we identify the parameter, $\alpha_0$,
+as the initial updating magnitude, $\alpha(t^{(s)} = 0)$,
+then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
+and the optimal $t_0 = \alpha_0^{-1}$.
+%
+Thus, in this limit,
+the schedule given by Eq. \eqref{eq:alphas_t}
+is identical to that from Eq. \eqref{eq:alpha_invt1}.
+
+\section{\label{sec:Gamma}
+Integrals of the autocorrelation functions}
+
+Here, we evaluate the integrals of
+the autocorrelation functions,
+$\Gamma_k$'s,
+in two special cases
+in Secs. \ref{sec:Gamma_perfect}
+and \ref{sec:Gamma_onestep}
+in the asymptotic limit,
+and describe a general method of measuring them
+in Sec. \ref{sec:Gamma_measure}.
+
+\subsection{\label{sec:Gamma_perfect}
+Perfect sampling}
+
+First, if the sampling is perfect,
+%
+%\begin{equation}
+%  \Gamma_k = 1.
+%\label{eq:Gamma_perfect}
+%\end{equation}
+%
+%To see this, we note that
+the bin index, $i$, is an independent random variable
+at each time step $t$.
+%
+Further, in the asymptotic limit,
+$\langle h_i(t) \rangle = p_i$,
+we find from Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def} that
+\begin{equation*}
+  \xi_i(t)
+  =
+  \frac{ h_i(t) } { p_i }
+  -
+  \sum_{ r = 1 }^n
+    \rho_r \frac{ h_r(t) } { p_r }
+  ,
+\end{equation*}
+%
+and
+%
+\begin{align}
+  \left\langle
+    \xi_i(t) \, \xi_j(t')
+  \right\rangle
+  %&=
+  %\left\langle
+  %  \left(
+  %    \frac{ h_i(t) } { p_i }
+  %    -
+  %    \sum_{ r = 1 }^n
+  %      \rho_r \frac{ h_r(t) } { p_r }
+  %  \right)
+  %\right.
+  %\notag \\
+  %&\hphantom{==}\cdot
+  %\left.
+  %  \left(
+  %    \frac{ h_j(t') } { p_i }
+  %    -
+  %    \sum_{ s = 1 }^n
+  %      \rho_s \frac{ h_s(t') } { p_s }
+  %  \right)
+  %\right\rangle
+  %\notag
+  %\\
+  &=
+  \left(
+    \frac{ \delta_{ij} } { p_i }
+    -
+    \frac{ \rho_i } { p_i }
+    -
+    \frac{ \rho_j } { p_j }
+    +
+    \sum_{r = 1}^n
+    \frac{ \rho_r^2 } { p_r }
+  \right) \,
+  \delta_{t, t'}
+  .
+\label{eq:zz_perfect}
+\end{align}
+%
+%$\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
+%in the continuous limit.
+%
+\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
+  with $t \ne t'$ (going back to the discrete time case),
+  $$
+  \begin{aligned}
+  \left\langle
+    \xi_i(t) \, \xi_j(t')
+  \right\rangle
+  &=
+  \left\langle
+    \xi_i(t)
+  \right\rangle
+  \left\langle
+    \xi_j(t')
+  \right\rangle
+  =
+  0.
+  \end{aligned}
+  $$
+  Next, for $t = t'$, we have
+  $$
+  \langle
+    h_i(t) \, h_j(t)
+  \rangle
+  = \delta_{ij} \, p_i
+  ,
+  $$
+  which leads to Eq. \eqref{eq:zz_perfect}.
+}
+%
+Thus,
+for $k \ge 2$, we get
+%
+\begin{equation}
+  \kappa_k(t - t')
+  =
+  \sum_{i,j = 1}^n
+  \phi_{ki} \phi_{kj}
+  \langle \xi_i(t) \, \xi_j(t') \rangle
+  %
+  =
+  \sum_{i = 1}^n \frac{ \phi_{ki}^2 } { p_i }
+  \, \delta_{t, t'},
+\notag
+%\label{eq:kappa_perfect}
+\end{equation}
+%
+where we have used
+Eqs. \eqref{eq:ortho1},
+\eqref{eq:theta_def},
+and
+\eqref{eq:zz_perfect}.
+%
+Thus, the white-noise approximation is exact,
+with
+$\Gamma_k = \sum_{ i = 1}^n \phi_{ki}^2 / p_i$.
+\note{Proof.%Eq. \eqref{eq:kappa_perfect},
+  $$
+  \begin{aligned}
+  \kappa_k(t - t')
+  &=
+  \sum_{i,j = 1}^n
+  \phi_{ki}
+  \phi_{kj}
+  \langle \xi_i(t) \, \xi_j(t') \rangle
+  \\
+  %
+  &=
+  \sum_{i,j = 1}^n
+  \phi_{ki}
+  \phi_{kj}
+  \left(
+    \frac{ \delta_{ij} } { p_i }
+    -
+    \frac{ \rho_i } { p_i }
+    -
+    \frac{ \rho_j } { p_j }
+    +
+    \sum_{r = 1}^n
+    \frac{ \rho_r^2 } { p_r }
+  \right) \,
+  \, \delta_{t, t'}
+  .
+  \end{aligned}
+  $$
+  where,
+  we have used
+  Eq. \eqref{eq:theta_def}
+  on the first line,
+  %
+  Eq. \eqref{eq:zz_perfect}
+  on the second line.
+  %
+  With Eq. \eqref{eq:ortho1},
+  only the first term in the parentheses survives.
+}
+In particular, if $p_i = \rho_i$, then, by
+Eq. \eqref{eq:eig_orthonormal_rows}, we have
+$\kappa_k(t - t') = \delta_{t, t'}$, or
+\begin{equation}
+  \Gamma_k = 1
+  .
+\label{eq:Gamma_perfect}
+\end{equation}
+
+\subsection{\label{sec:Gamma_onestep}
+One-step process}
+
+Another special case is the
+one-dimensional one-step process\cite{vankampen},
+which models a local sampling process.
+In each step, the system hops to
+one of the two neighboring bins
+with equal probability, $1/2$.
+%
+The transition matrix is
+%
+\begin{equation}
+  \arraycolsep=3.6pt\def\arraystretch{1.4}
+  \mathbf A
+  =
+  \left(
+    \begin{array}{cccccccc}
+      \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
+      \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
+      \vdots & &  & & \vdots \\
+      \frac {1 - \sigma} 2 & \dots &  & \frac 1 2 & \frac \sigma 2
+    \end{array}
+  \right),
+\notag
+%\label{eq:T_nn}
+\end{equation}
+%
+where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
+%
+For many updating schemes
+(cf. Sec. \ref{sec:band-matrix}),
+we can further assume that
+the transition matrix $\mathbf A$ and
+the updating matrix $\mathbf w$
+share the eigenvectors,
+with $\rho_i = p_i = 1/n$.
+%but not the eigenvalues.
+%
+The eigenvalues of $\mathbf A$ are
+$$
+  \Lambda_k
+  =
+  \cos \frac{ (k - 1) \, g \, \pi }
+            {            n        }
+  ,
+$$
+where
+$g = 2$ for a periodic variable, or $1$ otherwise.
+%Note that these $\Lambda_k$'s are
+%not to be confused with the eigenvalues of $\mathbf w$,
+%$\lambda_k$'s.
+%
+In this case,
+the autocorrelation function, $\kappa_k(t)$,
+decays as $\kappa_k(0) \, \Lambda_k^t$,
+with $\kappa_k(0) = 1$ being the same as
+the value in perfect sampling.
+%
+Thus,
+%$\Gamma_k$ is roughly twice the autocorrelation time, and
+from Eq. \eqref{eq:Gamma_sum}, we get
+%
+\begin{equation}
+  \Gamma_k
+  =
+  1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
+  =
+  \frac{ 1 + \Lambda_k }
+       { 1 - \Lambda_k }
+  =
+  \cot^2 \left[
+    \frac{ (k - 1) \, g \, \pi }
+         { 2 \, n }
+  \right]
+  .
+\label{eq:Gamma_onestep}
+\end{equation}
+
+\subsection{\label{sec:Gamma_measure}
+Measurement via an adaptive FDS simulation
+}
+
+%To use Eq. \eqref{eq:q_opt} in a real simulation,
+%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
+%
+For an unknown system,
+we can use Eq. \eqref{eq:y2_eql} to
+measure $\Gamma_k$'s in
+an adaptive FDS simulation
+with a constant schedule,
+$\alpha(t) = a_0$.
+%
+%A practical difficulty is that
+However,
+we often do not know a priori the intrinsic distribution, $p^*_i$,
+and hence
+cannot compute the deviation
+of the bias potential, $V_i$.
+%
+To circumvent the problem, we
+modify Eq. \eqref{eq:y2_eql} as
+%
+\begin{equation}
+  \operatorname{var} Y_k
+  =
+  \frac{1}{2} \,
+  a_0 \, \Gamma_k \, \lambda_k,
+\label{eq:varY}
+\end{equation}
+%
+where
+``$\operatorname{var}$''
+means the variance from the sampling (time) average,
+and
+%
+\begin{align*}
+Y_k
+&=
+\sum_{ i = 1 }^n
+  \phi_{k i} \, (V_i - \bar V),
+%\label{eq:Y_def}
+\end{align*}
+%
+with
+$
+\bar V
+\equiv
+\sum_{ j = 1 }^n \rho_j \, V_j.
+$
+
+To show Eq. \eqref{eq:varY},
+we first note that
+the difference, $V_i - {\bar V}$,
+differs from the actual deviation, $x_i$,
+defined in Eq. \eqref{eq:x_def},
+only by a constant shift,
+$\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
+%
+Thus, the two share the same variance
+under time averaging.
+%
+It follows that the corresponding linear combinations,
+$Y_k$ and $y_k$, would also share the same variance.
+%
+But since the average of $y_k$ is zero,
+we have
+$$
+\operatorname{var} Y_k
+=
+\operatorname{var} y_k
+=
+\left\langle y_k^2 \right\rangle.
+$$
+Finally, by using Eq. \eqref{eq:y2_eql},
+we reach Eq. \eqref{eq:varY}.
+
+\note{In practice, Eq. \eqref{eq:varY}
+in inapplicable to zero eigenvalues, $\lambda_k$'s.
+%
+However, these eigenmodes can be ignored
+as they do not contribute to the Lagrangian defined by
+Eq. \eqref{eq:Lagrangian_const}.
+}
+
+As an example, we used the method to compute
+the integrals, $\Gamma_k$'s, from
+a local MC simulation
+(cf. Sec. \ref{sec:results_system})
+and an MD simulation,
+and compared them with those from
+the one-step sampling process.
+%
+As shown in Fig. \ref{fig:gamcmp},
+the $\Gamma_k$'s of the small-$k$ eigenmodes
+from the three sampling processes
+were almost proportional,
+suggesting that
+the optimal schedule
+for simulations using local sampling processes
+would be similar in general properties.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/gamcmp.pdf}
+  }
+  \caption{
+    \label{fig:gamcmp}
+    Integrals of the autocorrelation functions
+    from a local MC simulation
+    and an MD simulation.
+    %
+    Each simulation was ran for $10^{10}$ steps.
+    %
+    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$.
+    %
+    For the MD simulation,
+    the step size $\Delta t_\mathrm{MD} = 0.005$.
+    %
+    The velocity rescaling thermostat\cite{bussi2007} was
+    applied for the unit reduced temperature
+    every MD step with viscosity $20$.
+    %
+    The reflective boundary condition was used.
+    %
+    \note{The figure was produced by \texttt{doc/fig/gamcmp.gp}.
+      The data were saved under \texttt{data/gamcmp}
+      (see the \texttt{Makefile} there).
+    }%
+  }
+\end{center}
+\end{figure}
+
+\section{\label{sec:more_wband}
+Homogeneous updating schemes
+}
+
+Here, we give some detailed derivations
+on the homogeneous updating schemes.
+%
+By the translational invariance,
+homogeneous updating schemes
+satisfy detailed balance,
+Eq. \eqref{eq:w_detailedbalance},
+for the flat distribution, $\rho_i = 1/n$,
+i.e. the updating matrix, $\mathbf w$,
+is symmetric.
+%
+The updating matrices
+share the same set of eigenvectors
+(which here are sines and cosines of different frequencies),
+and thus are completely determined
+by the eigenvalues.
+%
+Further, the multiplication of $\mathbf w$
+can be reduced to a convolution with the window function,
+or the updating kernel.
+%
+Thus,
+there is a one-to-one correspondence between
+the eigenvalues of $\mathbf w$
+and the updating kernel.
+
+\subsection{\label{sec:wband_eig}
+Eigenvalues and eigenvectors}
+
+To find the eigenvectors for a periodic variable, $\phi_i$,
+we define the out-of-boundary values by
 %
@@
 %
-where $\ii^2 = -1$, and
-the eigenvalues are
-the Fourier or cosine transform of the updating kernel:
-%
-\begin{equation}
-  \lambda_k
-  =
-  \mu_0
-  +
-  2 \,
-  \sum_{ l = 1 }^b
-  \mu_l
-  \cos
-  \frac{ (k - 1) \, l \, 2 \, \pi }
-       {            n             }
-  .
-  \label{eq:wband_eigenvalue_pbc}
-\end{equation}
+where $\ii^2 = -1$.
+%
+Then one can readily verify that the eigenvalues are given by
+  Eq. \eqref{eq:wband_eigenvalue_pbc}.
 %
@@
 %
-Note, however, the two-fold degeneracy as
-$\lambda_{n - k + 2} = \lambda_k$.
-
-To adapt the above formulae to a non-periodic variable,
-we use the reflective boundary condition\cite{bussi2006},
-and change the form of the updating matrix to
-%
-%
-\begin{equation}
-  w_{ij}
-  =
-  \mu_{ i - j }
-  +
-  \mu_{ i + j - 1 }
-  +
-  \mu_{ i + j - 2 n - 1 }
-  ,
-  \label{eq:w_band_refl}
-\end{equation}
-%
-where the last two terms on the right hand side,
-representing two reflective mirrors at
-$j = 1/2$ and $j = n + 1/2$,
-are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
-to the equilibrium distribution\cite{bussi2006},
-such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
-are satisfied.
-%
-Again, the updating kernel should satisfy
-Eqs. \eqref{eq:mu_normalization}
-and \eqref{eq:mu_symm},
-but
-the cutoff $b$ only needs to be less than $n$.
-
-Parallel to Eq. \eqref{eq:phi_pbc},
+
+For a non-periodic variable,
 we define the out-of-boundary values
@@
 %
-The eigenvalues can be borrowed from
-Eq. \eqref{eq:wband_eigenvalue_pbc},
-with the period, $n$, doubled:
-%
-\begin{align}
-  \lambda_k
-  &=
-  \mu_0
-  +
-  2
-  \sum_{l = 1}^b
-    \mu_l
-    \cos \frac{(k - 1)  \, l \, 2 \, \pi}{ 2 \, n }
-  .
-\label{eq:wband_eigenvalue_refl}
-\end{align}
-%
+and the eigenvalues are given by
+  Eq. \eqref{eq:wband_eigenvalue_refl}.
+
 \note{Derivation.
@@
 %
-We may unify Eqs. \eqref{eq:wband_eigenvalue_refl}
-and \eqref{eq:wband_eigenvalue_pbc} as
-%
-\begin{equation}
-  \lambda_k
-  =
-  \mu_0
-  +
-  2
-  \sum_{ l = 1 }^b
-    \mu_l \,
-    \cos \frac{ (k - 1) \, l \, g \, \pi }
-              {            n             }
-  ,
-  \label{eq:wband_eigenvalue}
-\end{equation}
-%
-where the degeneracy, $g$, is $2$
-for the periodic case,
-or $1$ otherwise.
-%
-In both cases,
-the eigenvalues, $\lambda_k$'s,
-are the cosine transform of
-the updating kernel.
-
-For the updating scheme to be stable,
-the cosine transform of the kernel
-must be nonnegative.
-%
-If this is not so,
-we can modify the kernel
-using the technique described
-in Appendix \ref{sec:stabilize_wband}.
-
-We discuss two special cases below.
+
+\subsection{Two examples}
+
+We give two examples of the homogeneous updating schemes.
 
@@
 
-\subsubsection{\label{sec:invert_wband}
-Inversion}
-
-Since the eigenvalues are related to the updating kernel
-by a cosine transform,
-Eq. \eqref{eq:wband_eigenvalue},
-the relation can be readily inverted.
-%
-In the periodic case,
-we have
-%the inversion formula is
-%
-\begin{equation}
-  \mu_l
-  =
-  \frac 1 n
-  \sum_{ k = 1 }^n
-  \lambda_k
-  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
-            {            n             }
-  .
-\label{eq:mu_from_lambda_pbc}
-\end{equation}
-
-In the non-periodic case,
-the formula is
-%
-\begin{align}
-  \mu_l
-  =
-  \frac 1 { 2 \, n }
-  \sum_{ k = 1 }^{ 2 \, n }
-    \lambda_{ k } \,
-    \cos \frac{ (k - 1) \, l \, \pi }
-              {            n        }
-  ,
-\label{eq:mu_from_lambda_refl}
-\end{align}
-%
-where
-we have defined
-$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
-for $k = n + 2, \dots, 2 \, n$,
-as well as
-%
-\begin{align}
-  \lambda_{ n + 1 }
-  =
-  (-1)^{ n - 1 }
-  \lambda_1
-  +
-  2 \, \sum_{ k = 2 }^{ n }
-      (-1)^{n - k} \, \lambda_k
-  ,
-\label{eq:lambdan}
-\end{align}
-to satisfy the constraint, $\mu_n = 0$.
-%
-\note{Eq. \eqref{eq:lambdan}
-  ensures that the $\mu_n$
-  computed from Eq. \eqref{eq:mu_from_lambda_refl}
-  vanishes.
-
-  %The inversion formula is derived from the cosine transform.
-  %
-  If we define $\mu_n = 0$ and
-  %
-  $
-    \mu_{ 2 n - l } = \mu_l,
-  $
-  for
-  $l = 1, \dots, n - 1$,
-  %
-  then Eq. \eqref{eq:wband_eigenvalue_refl} can be rewritten as
-  %
-  \begin{align}
-    \lambda_{k+1}
-    =
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
-  \notag
-  %\label{eq:lambda_cosine_sum}
-  \end{align}
-  %
-  This formula for $\lambda_{k+1}$
-  can be readily extended to $k = 2 \, n - 1$,
-  and we have
-  %
-  $
-    \lambda_{ 2 \, n + 2 - k } = \lambda_k
-    .
-  $
-  %
-  Thus,
-  %
-  $$
-  \begin{aligned}
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \lambda_{ k + 1 } \,
-      \cos \frac{ k \, p \, \pi }
-                {      n        }
-    &=
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \sum_{ l = 0 }^{ 2 \, n - 1 }
-        \mu_l \,
-        \cos \frac{ k \, l \, \pi }
-                  {      n        }
-        \cos \frac{ k \, p \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \frac{ \mu_l } { 2 }
-      \sum_{ k = 0 }^{ 2 \, n - 1 }
-        \cos \frac{ k \, (p + l) \, \pi }
-                  {      n        }
-                  +
-        \cos \frac{ k \, (p - l) \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \mu_l \, n \left(
-        \delta_{ p + l - 2 \, n, 0 }
-        +
-        \delta_{ p - l, 0 }
-      \right)
-    \\
-    &=
-    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
-    =
-    2 \, n \, \mu_p.
-  \end{aligned}
-  $$
-  This entails Eq. \eqref{eq:mu_from_lambda_refl}.
-  %
-  The value of $\lambda_{n + 1}$
-  can be deduced from
-  imposing $\mu_n = 0$
-  and solving Eq. \eqref{eq:mu_from_lambda_refl}
-  for $l = n$,
-  yielding Eq. \eqref{eq:lambdan}.
-}
-
-\subsection{\label{sec:cmpschemes}
-Comparison of updating schemes}
-
-Now that we can compute the optimal schedule
-for a given updating scheme,
-we will turn to the problem of finding
-optimal updating schemes below.
-
-\subsubsection{\label{sec:optWL}
-Asymptotic optimality of the single-bin scheme}
-
-We will first show
-that the single-bin scheme is asymptotically optimal.
-%
-Consider a class of $\mathbf w$ matrices
-sharing the same set of eigenvectors,
-hence the same $\Gamma_k$'s,
-but with different positive eigenvalues,
-$\lambda_k$'s.
-%
-Using the Cauchy-Schwarz inequality, we have,
-for any set of nonnegative numbers, $c_k$'s,
-%
-\begin{align}
-&
-\left(
-  \int_0^T dt
-    \sum_{k = 2}^n
-      \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr)
-\right)
-%
-\left(
-  \int_0^T dt
-    \sum_{k = 2}^n
-      \Gamma_k \, c_k^2
-\right)
-%
-\notag
-\\
-&
-\qquad \qquad
-\ge
-\left(
-  \int_0^T dt
-    \sum_{k = 2}^n
-      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t) \bigr)
-\right)^2
-\notag
-\\
-&
-\qquad \qquad
-=
-\left(
-  \sum_{k = 2}^n \Gamma_k \, c_k
-    \left[
-      1 - e^{ -\lambda_k \, q(T) }
-    \right]
-\right)^2.
-\label{eq:CSineq}
-\end{align}
-%
-%\note{
-This inequality follows from the non-negativity of
-the quadratic polynomial,
-$$
-\int_0^T
-  dt \sum_{k = 2}^n \Gamma_k \,
-    \left( \dot u_k \, \chi - c_k \right)^2
-  \equiv
-  A_2 \, \chi^2 + A_1 \, \chi + A_0
-  ,
-$$
-which must carry a non-positive discriminant:
-$A_1^2 - 4 \, A_0 \, A_2 \le 0$.
-%}%
-%
-Note that the last expression of Eq. \eqref{eq:CSineq}
-depends on the curve, $q(t)$ ($0 < t < T$),
-only through the endpoint value, $q(T)$,
-which is fixed in the variational process.
-%
-Thus, the inequality sets a lower bound
-for the asymptotic error $\Err(T)$
-given in Eq. \eqref{eq:error_asym1}.
-%
-The equality is achieved
-if $\dot u_k\left( q(t) \right) = c_k$
-for all $k = 2, \dots, n$ at any $t$,
-up to a multiplicative factor.
-%
-Solving these equations yields
-%
-\begin{equation}
-  \alpha(t) = \frac{              1             }
-                   { \lambda_k \, (t + t_{k0} ) },
-\label{eq:alpha_invtk}
-\end{equation}
-with
-\begin{equation}
-  t_{k0} = \frac{             T            }
-                { e^{ \lambda_k q(T) } - 1 }.
-\label{eq:tk0}
-\end{equation}
-%
-\note{Derivation.
-  Integrating $\dot u_k = c_k$ yields
-  \begin{equation}
-    u_k(t) = c_k \left(t + t_{k0} \right).
-  \tag{UK}
-  \label{neq:ui_solution}
-  \end{equation}
-  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
-  we have
-  $$
-  c_k = \frac{ 1 }{ T + t_{k0} },
-  \quad
-  \mathrm{and\;}
-  \frac{ t_{k0} } { T + t_{k0} }
-  =
-  e^{ -\lambda_k \, q(T) }.
-  $$
-  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
-  $$
-  -\lambda_k \, q(T) + \lambda_k \, q(t)
-  = \ln c_k + \ln \left( t + t_{k0} \right).
-  $$
-  Differentiating this with respect to $t$ yields
-  Eq. \eqref{eq:alpha_invtk}.
-}
-%
-Such a solution is possible only if
-all nontrivial eigenvalues are the same
-%
-\begin{equation}
-  \lambda_2 = \dots = \lambda_n.
-  \notag
-\end{equation}
-%
-Then, by Eq. \eqref{eq:tk0},
-all $t_{k0}$'s also share the same value,
-$t_0$,
-such that
-$u_k = (t + t_0) / (T + t_0)$,
-and
-$c_k = 1/(T + t_0)$
-for $k \ge 2$.
-%
-The asymptotic error is then
-%
-\begin{align}
-  \Err_A(T)
-  =
-  \frac{       T     }
-       { (T + t_0)^2 }
-  \sum_{ k = 2 }^n
-    \Gamma_k
-  .
-\label{eq:error_asym_singlebin}
-\end{align}
-%
-If we further assume $\lambda_2 = \cdots = \lambda_n = 1$,
-%the optimal schedule given by
-Eq. \eqref{eq:alpha_invtk}
-recovers Eq. \eqref{eq:alpha_invt1}.
-%
-
-Since the first eigenmode represents
-a uniform shift of the bias potential,
-we can freely set $\lambda_1$ to be the same as $\lambda_2$
-without changing the nature of the updating scheme.
-%
-Note, an updating matrix
-with identical eigenvalues
-is a multiple of the identity matrix.
-%
-Thus,
-the corresponding updating scheme,
-the single-bin scheme, is most efficient
-in asymptotic convergence.
-%if no eigenvalue $\lambda_k$ is zero.
-%
-In Appendix \ref{sec:upperbound},
-we will show that the opposite limit of
-a set of well-separated eigenvalues
-will provide an upper bound for
-the asymptotic error, Eq. \eqref{eq:error_asym_ubound}.
-
-\subsubsection{\label{sec:optscheme}
-Bandpass updating schemes}
-
-We can generalize
-the single-bin scheme to a class of
-perfect (brick-wall) \emph{bandpass} updating schemes
-that ignore short-wavelength eigenmodes,
-which are typically local noise.
-%
-Mathematically, these updating schemes will have
-zero eigenvalues for the ignored modes,
-and they are useful if we can assume that
-the ignored modes are absent from the PMF.
-%
-For example,
-%if we can assume
-by assuming
-that the PMF is sufficiently smooth
-such that the eigenmodes can be limited to $k \le K$,
-%
-we can construct an updating scheme with
-$$
-\lambda_1 = \cdots = \lambda_K = 1,
-\mathrm{\; and \;}
-\lambda_{K+1} = \cdots = \lambda_n = 0.
-$$
-%
-Then,
-we can change the upper limit of the sum in
-Eq. \eqref{eq:error_asym_singlebin} to $K$:
-%
-%Then,
-%Eq. \eqref{eq:error_singlebin} is changed to
-%
-\begin{equation}
-  \Err_A(T)
-  =
-  \frac {       T     }
-        { (T + t_0)^2 }
-  \sum_{ k = 2 }^K
-    \Gamma_k.
-%\notag
-\label{eq:error_asym_sinc}
-\end{equation}
-%
-The optimal schedule of these schemes
-remains to be the inverse-time schedule
-given by Eq. \eqref{eq:alpha_invt1}.
-%
-Below, we construct these schemes for
-the homogeneous case.
-%
-The definition of the cutoff, $K$, will be modified
-for mathematical convenience.
-
-For a periodic variable, we demand
-%
-$$
-\begin{aligned}
-&
-\lambda_1 = \cdots = \lambda_{K+1}
-= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
-\\
-&
-\lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
-\end{aligned}
-$$
-for $K \ge 1$.
-%
-Then by using
-Eq. \eqref{eq:mu_from_lambda_pbc},
-we get
-\begin{equation}
-  \mu_l
-  =
-  \frac{
-    \sin
-    \dfrac{ (2 K + 1) \, l \, \pi }
-         {              n        }
-  }
-  {
-    n \, \sin \dfrac{ l \, \pi } { n }
-  }
-  .
-\notag
-%\label{eq:mu_sinc_pbc}
-\end{equation}
-\note{Derivation.
-$$
-\begin{aligned}
-\mu_l
-&=
-\frac 1 n \sum_{k = 0}^{n-1} \lambda_{k+1} \cos \frac{ k \, l \, 2 \, \pi } { n }
-\\
-&=
-\frac{1}{n}
-\left(
-  1 +
-  \sum_{k=1}^K
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-  +
-  \sum_{k=n-K}^{n-1}
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-\right)
-\\
-&=
-\frac 1 n
-\sum_{k=-K}^K
-\cos \frac { k \, l \, 2 \, \pi } { n }
-=
-  \frac{
-    \sin
-    \dfrac{ (2 K + 1) \, l \, \pi }
-         {              n        }
-  }
-  {
-    n \, \sin \dfrac{ l \, \pi } { n }
-  }
-.
-\end{aligned}
-$$
-}
-
-For a non-periodic variable, we demand
-$$
-\lambda_1 = \cdots = \lambda_{K+1} = 1,
-\qquad
-\lambda_{K+2} = \cdots = \lambda_n = 0.
-$$
-Then Eq. \eqref{eq:lambdan} gives
-$\lambda_{n+1} = (-1)^{n+K-1}$,
-and from Eq. \eqref{eq:mu_from_lambda_refl},
-we have
-\begin{equation}
-  \mu_l
-  =
-  \frac{1}{2 \, n}
-  \left[
-    (-1)^{n+K+l-1}
-    +
-    \frac{
-      \sin
-      \dfrac{ (2 K + 1) \, l \, \pi }
-           {         2 \, n        }
-    }
-    {
-      \sin \dfrac{ l \, \pi } { 2 \, n }
-    }
-  \right]
-  .
-\label{eq:mu_sinc_refl}
-\end{equation}
-\note{Derivation.
-$$
-\begin{aligned}
-  \lambda_{n+1}
-  &=
-  (-1)^{n-1}
-  \left[
-    \lambda_1
-    - 2 \, \lambda_2
-    + 2 \, \lambda_3 - \cdots
-    + (-1)^{K} 2 \, \lambda_{K+1})
-  \right]
-  \\
-  &=
-  \begin{dcases}
-    (-1)^{n-1} & K \mathrm{\; even,} \\
-    (-1)^n     & K \mathrm{\; odd.}
-  \end{dcases}
-\end{aligned}
-$$
-So
-$$
-\begin{aligned}
-  \mu_l
-  &=
-  \frac{1}{2\,n}
-  \left[
-    1 +
-    2 \sum_{k=2}^{K+1}
-    \cos \frac { (k - 1) \, l \pi } { n }
-    +
-    (-1)^{n+K-1} (-1)^l
-  \right]
-  \\
-  &=
-  \frac{1}{2\,n}
-  \left[
-    (-1)^{n+K+l-1}
-    +
-    \sum_{k=-K}^{K}
-    \cos \frac { k \, l \pi } { n }
-  \right]
-  .
-\end{aligned}
-$$
-}%
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
-  }
-  \caption{
-    \label{fig:sinc}
-    Updating kernels of the bandpass updating schemes
-    compared to a Gaussian with a similar width.
-    %
-    The lines are a guide to the eyes.
-    Inset. Updating matrix in the non-periodic case.
-    %
-    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h} for panel (a),
-      and the functions \texttt{savewin()}
-      and \texttt{savewinmat()}
-      in \texttt{prog/invt.h} for panel (b),
-      The data were saved under the directory \texttt{data/sinc}.
-      See the \texttt{make} object \texttt{sincsingle}
-      in the \texttt{Makefile} there.
-    }%
-  }
-\end{center}
-\end{figure}
-
-In Fig. \ref{fig:sinc},
-we give examples of the updating kernels
-of the bandpass updating schemes,
-compared to a Gaussian
-with a similar width, $\sigma$,
-%
-\begin{equation}
-  \sigma
-  \approx
-  \frac
-  {
-    n
-  }
-  {
-    \sqrt{ 2 \, \pi } \, g \, K
-  }
-  .
-\label{eq:sigma_equiv}
-\end{equation}
-%
-%where $g = 2$ for the periodic case
-%or $g = 1$ otherwise.
-%
-The kernel in the non-periodic case is a sawtooth
-because of the term, $(-1)^{n+K+l-1}$,
-in Eq. \eqref{eq:mu_sinc_refl}.
-%
-However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
-the ruggedness disappears
-in the updating matrix by Eq. \eqref{eq:w_band_refl},
-as the above oscillatory term for $l = i - j$
-is canceled by the term for either $l = i + j - 1$
-or $l = i + j - 2 \, n - 1$
-with the opposite sign.
-
-\subsubsection{\label{sec:finlencorr}
-Finite-length correction}
-
-In the above section,
-we have focused on
-the asymptotic error.
-%
-By including the residual error
-and minimizing the total error,
-we get the optimal $t_0$,
-which is, according to Eq. \eqref{eq:half_alpha0},
-%
-%%Assuming the system is initially equilibrated
-%%under a constant updating magnitude,
-%%$a_0$,
-%Assuming Eq. \eqref{eq:error_res1} with
-%$$
-%q(T)
-%=
-%\int_0^T \frac{ dt } { t + t_0 }
-%=
-%\ln \frac{ T + t_0 } { t_0 }.
-%$$
-%%
-%we have, from Eqs. \eqref{eq:error_split}
-%and \eqref{eq:error_asym_sinc},
-%%
-%\begin{equation}
-%  \Err
-%  =
-%  \Err_R + \Err_A
-%  =
-%  \frac { \frac 1 2 a_0 \, t_0^2 + T }
-%        {          ( T + t_0 )^2          }
-%  \sum_{ k = 2 }^K \Gamma_k
-%  .
-%\notag
-%%\label{eq:error_sinc0}
-%\end{equation}
-%%
-%The minimum is reached at
-%
-\begin{equation}
-  t_0
-  =
-  \frac{  2  }
-       { a_0 }
-  .
-\label{eq:t0_sinc}
-\end{equation}
-%
-For the single-bin updating scheme,
-it recovers Eq. \eqref{eq:t0_singlebin}
-under the assumption of Eq. \eqref{eq:y2_eql}.
-%
-From Eqs. \eqref{eq:error_split},
-\eqref{eq:error_res1},
-\eqref{eq:tk0},
-and
-\eqref{eq:error_asym_sinc},
-we get
-%
-\begin{equation}
-  \Err(T)
-  =
-  \frac{   1     }
-       { T + t_0 }
-  \sum_{ k = 2 }^K
-    \Gamma_k
-  .
-\label{eq:error_sinc}
-\end{equation}
-%
-which serves as a generalization of
-the result from the single-bin scheme, Eq. \eqref{eq:Emin_singlebin}.
-%
-This error decreases with increasing simulation length,
-as one would expect from an equilibrium FDS simulation.
-%In Appendix \ref{sec:equilerr},
-%we show that the above error tends
-%to the error of the ideal equilibrium FDS simulation
-%in the asymptotic limit.
-
-%
-%For example, consider a system that is initially
-%prepared under a fixed $a_0$
-%at the beginning of the schedule $\alpha(t)$.
-%%
-%Then for shorter times, $T \approx 0$,
-%the error, dominated by the residual error
-%given by Eq. \eqref{eq:error_res1},
-%is clearly smaller if some eigenvalues
-%are less than $1.0$ (the single-bin case).
-%
-%In practice, real simulations
-%may have long achieved the desired precision
-%before reaching the asymptotic regime,
-%especially if the updating matrix
-%has some $\lambda_k$'s close to $0$.
-
-\section{\label{sec:results}
-Numerical results}
-
-We tested the above theory on a model system
-described in Sec. \ref{sec:results_system}.
-%
-We verify the method of computing errors
-%using an inverse-time schedule
-in Sec. \ref{sec:results_invt}
-%
-and study the optimal schedule from
-Eq. \eqref{eq:q_opt}
-in Sec. \ref{sec:results_optschedule}.
-%
-Finally, we compare the Gaussian and
-bandpass updating schemes %of different widths
-in Sec. \ref{sec:results_cmpschemes}.
-
-\subsection{\label{sec:results_system}
-Test system and simulation protocol}
-
-Our test system is a one-dimensional model system, in which
-$i$ represents a non-periodic variable,
-and $n = 100$.
-%
-Both the target and intrinsic distributions
-are assumed to be flat:
-$p^*_i = p_i = 1/n$,
-along with the distribution, $\rho_i$, in Eqs.
-\eqref{eq:vbar_def} and \eqref{eq:error_sum}.
-%
-%We wish to see how
-%the schedule, $\alpha(t)$,
-%affects the fluctuating error
-%of the bias potential
-%at the end of the simulation.
-%
-Initially,
-we set the bias potential to zero,
-$V_i \equiv 0$,
-and equilibrate the system for $10^7$ steps
-under a constant updating magnitude,
-$a_0 = 10^{-4}$.
-%
-This equilibration process models
-a stage of the WL algorithm or
-a regular metadynamics
-with a fixed updating magnitude.
-%
-Although the bias potential is correct initially,
-it would deviate from this value
-at the end of the equilibration.
-%
-Our aim is to see how well
-the bias potential is restored
-at the end of the testing schedule.
-%
-After the equilibration,
-we reset the origin of the time, $t$, to zero,
-and start the testing schedule
-for $T = 10^8$ steps.
-%
-We repeat the above process $1000$ times (unless specified otherwise)
-and report the averaged results.
-%specified by either
-%Eq. \eqref{eq:q_opt}
-%or
-%Eq. \eqref{eq:alpha_invtlambda}.
-
-For the Gaussian updating scheme,
-the kernel was truncated at
-$b = \min\{10 \, \sigma, n - 1\}$ bins
-and numerically stabilized
-using the technique
-in Appendix \ref{sec:stabilize_wband}.
-
-%Since the error depends on
-%the underlying sampling process,
-We tested two sampling processes.
-%a global one and a local one.
-%
-Both are based on
-a Metropolis MC algorithm\cite{
-  metropolis1953, newman, frenkel,
-  landau_binder}:
-%
-in each step, a new bin index $j$ is proposed
-and then accepted with probability
-%
-$
-A(i \to j) = \min\{ 1, \exp(V_i - V_j) \}.
-$
-However,
-in the first or global sampling process,
-we choose the destination $j$
-uniformly out of the $n$ bins.
-%
-In the second or local sampling process,
-$j$ is either $i - 1$ or $i + 1$
-(adjusted by the boundary condition)
-with equal probability $1/2$.
-%
-Asymptotically,
-the above global and local sampling processes
-emulate the perfect and one-step processes
-discussed in Appendix \ref{sec:Gamma},
-respectively,
-and we will use
-Eqs. \eqref{eq:Gamma_perfect}
-and \eqref{eq:Gamma_onestep}
-for the respective $\Gamma_k$'s
-in the theoretical predictions.
-%
-
-\subsection{\label{sec:results_invt}
-Modified inverse-time schedule}
-
-Using the model system above, we
-tested our analysis of computing errors
-on the modified inverse-time schedule,
-%
-\begin{equation}
-\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
-\label{eq:alpha_invtlambda}
-\end{equation}
-%
-where $\lambda$ is a free parameter,
-and $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
-%and $t_0 = 2/a_0$.
-%
-The error of this schedule
-can be computed analytically
-(cf. Appendix \ref{sec:invt_schedule}),
-and tested against simulation.
-
-We performed adaptive FDS simulations
-using the single-bin updating scheme
-and the triple-bin updating scheme with $\mu_1 = 0.24$.
-%
-%For each value of $\lambda$,
-%we ran $1000$ independent simulations
-%of $10^8$ steps, and averaged the results.
-%
-As shown in Fig. \ref{fig:errinvt},
-the errors from the simulations
-agreed well with those from the theory.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/errinvt.pdf}
-  }
-  \caption{
-    \label{fig:errinvt}
-    Error, $\Err$, versus the parameter, $\lambda$,
-    for the modified inverse-time schedule,
-    Eq. \eqref{eq:alpha_invtlambda}.
-    %
-    The points are results from averaging over $1000$ independent runs;
-    the curves are theoretical predictions.
-    %
-    For the single-bin (WL) updating scheme,
-    the minimal should occur at $\lambda = 1$,
-    while it is not necessarily the case for a general updating scheme.
-    %
-    \note{The figure was produced by \texttt{doc/fig/errinvt.gp}.
-      The data were produced by the script \texttt{data/invtcscan.py},
-      and saved under \texttt{data/invt}
-      (see the \texttt{Makefile} there).
-    }%
-  }
-\end{center}
-\end{figure}
-
-For the single-bin updating scheme,
-the optimal $\lambda$ occurred around $1.0$
-for both global and local sampling processes.
-%
-The sampling process affected the error function
-only by a multiplicative constant,
-%The error functions of the two sampling processes
-%differed only by a multiplicative constant,
-in agreement with the discussion in
-Appendix \ref{sec:invt_singlebin}.
-%
-For the triple-bin updating scheme,
-the optimal $\lambda$ was less than $1.0$
-for the global sampling process,
-although it stayed around $1.0$
-for the local sampling process,
-also in agreement with the discussion
-in Appendix \ref{sec:invt_nn}.
-%
-Thus, the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1},
-is optimal for the single-bin updating scheme,
-but not necessarily so for a general multiple-bin updating scheme.
-%
-
-\subsection{\label{sec:results_optschedule}
-Optimal schedule}
-
-To demonstrate the optimal schedule from Eq. \eqref{eq:q_opt},
-we used the Gaussian (with $\sigma = 10$ bins) updating scheme
-as an example.
-%
-This updating scheme is typically used in metadynamics simulations.
-%
-In Table \ref{tab:error_Gaussian},
-we showed that the optimal schedule
-produced smaller errors than
-the inverse-time schedules.
-%
-However, the optimal schedules
-decayed more slowly than
-the inverse-time formula, Eq. \eqref{eq:alpha_invt1},
-and depended on the sampling process,
-as shown in Fig. \ref{fig:optacmp}.
-%
-The impeded decay of the optimal schedule
-was necessary in order to
-damp out the short-wavelength eigenmodes
-(i.e. local noise).
-%
-As shown in Fig. \ref{fig:xerr},
-the inverse-time schedule was effective in error reduction
-only for the first few eigenmodes,
-whose eigenvalues were around $1.0$.
-%
-We can, however, ignore these local eigenmodes
-by explicitly filtering them out from the final bias potential,
-and truncating the error sum in Eq. \eqref{eq:error_sum}
-to the first few eigenmodes.
-%
-Then, the corresponding optimal schedule
-from Eq. \eqref{eq:q_opt} that minimizes the truncated error
-became similar to the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1},
-as exemplified in Fig. \ref{fig:optacmp}.
-%
-\note{A more thorough approach is
-to use the bandpass updating scheme
-to be discussed later.
-}
-
-\begin{table}[h]\footnotesize
-  \caption{\label{tab:error_Gaussian}
-    Error of the Gaussian updating scheme of $\sigma = 10$
-    under different updating schedules.
-    In Eqs. \eqref{eq:alpha_invt1} and \eqref{eq:alpha_invtlambda},
-    $t_0$ is given by Eq. \eqref{eq:t0_sinc};
-    and in Eq. \eqref{eq:alpha_invtlambda},
-    the optimal value of $\lambda$ is used.
-    \note{\newline
-      The data were collected in
-      \texttt{data/tab1} by running
-      the program \texttt{invt}.
-    }%
-  }
-  \setlength{\tabcolsep}{2pt}
-  \renewcommand\arraystretch{1.2}
-  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
-    \hline
-      Schedule
-    &
-      Inverse-time, \newline
-      Eq. \eqref{eq:alpha_invt1}
-    &
-      Modified, \newline
-      Eq.~\eqref{eq:alpha_invtlambda}
-    &
-      Optimal, \newline
-      Eq. \eqref{eq:q_opt}
-    \\
-    \hline
-    Global
-    &
-    $4.36(9) \times 10^{-6}$
-    &
-    $7.1(1) \times 10^{-7}$
-    &
-    $2.6(1) \times 10^{-7}$
-    \\
-    \hline
-    Local
-    &
-    $3.83(7) \times 10^{-4}$
-    &
-    $1.72(4) \times 10^{-4}$
-    &
-    $9.1(2) \times 10^{-5}$
-    \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/optacmp.pdf}
-  }
-  \caption{
-    \label{fig:optacmp}
-    Optimal schedules from Eq. \eqref{eq:q_opt}
-    for the Gaussian updating scheme
-    with $\sigma = 10$.
-    %
-    The mode-limited schedules were produced by
-    limiting the sum in Eq. \eqref{eq:q_opt}
-    to the first four terms,
-    or the four longest wavelength modes.
-    %
-    They became similar to the inverse-time schedule,
-    Eq. \eqref{eq:alpha_invt1}.
-    %
-    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h}.
-      The data were saved under the directory \texttt{data/opta}.
-    }%
-  }
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/xerr.pdf}
-  }
-  \caption{
-    \label{fig:xerr}
-    Predicted error components for the Gaussian updating scheme
-    with $\sigma = 10$
-    from the optimal schedule
-    and the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
-    %
-    The latter schedule failed to damp out the errors
-    from the long-wavelength (local) eigenmodes
-    with $k \ge 5$.
-    %
-    For longer simulations,
-    the error components from the optimal schedule
-    would roughly maintain a relatively flat profile,
-    but meet the initial error curve
-    at a lower level (not shown),
-    where $\lambda_k \approx 1/q(T)$
-    (cf. Appendix \ref{sec:upperbound}).
-    %
-    \note{This figure was produced by \texttt{doc/fig/xerr.gp}.
-      The data were saved under the directory \texttt{data/xerr}.
-    }%
-  }
-\end{center}
-\end{figure}
-
-An interesting prediction from the theory is Eq. \eqref{eq:half_alpha0},
-which states that the optimal schedule should always start from
-half of the magnitude from the previous equilibration run.
-%
-To verify this result,
-we computed the optimal schedules
-for different values of $q(T)$
-for the Gaussian updating scheme,
-and plotted the final error
-against the initial updating magnitude, $\alpha(0)$.
-%
-As shown in Fig. \ref{fig:erriascan},
-the minimal error was indeed reached at
-$\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
-although the dependence on $\alpha(0)$
-is generally weak.
-%
-The weak dependence is best illustrated in
-the case of single-bin updating scheme,
-as the initial updating magnitude only
-affects the schedule by a shift of the origin of
-simulation time in the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1}.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/erriascan.pdf}
-  }
-  \caption{
-    \label{fig:erriascan}
-    Relative error
-    versus the initial updating magnitude, $\alpha(0)$,
-    for optimal schedules with different $q(T)$.
-    Assuming the Gaussian updating scheme
-    with $\sigma = 10$,
-    and $a_0 = 10^{-4}$.
-    %
-    The points were obtained from averaging results
-    from over 10000 independent simulations;
-    the curves are theoretical predictions.
-    %
-    The minimal should always occur at
-    $\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
-    as predicted by Eq. \eqref{eq:half_alpha0}.
-    %
-    \note{This figure was produced by
-      \texttt{doc/fig/erriascan.gp}.
-      The data were saved under
-      \texttt{data/iascan}.
-    }%
-  }
-\end{center}
-\end{figure}
-
-\subsection{\label{sec:results_cmpschemes}
-Optimal updating schemes}
-
-For a finite-length simulation,
-the optimal updating scheme
-depends on the simulation length, $T$,
-and other parameters.
-%
-As an example,
-we compare Gaussian updating schemes
-with different widths, $\sigma$.
-%
-In Fig. \ref{fig:errsigscan},
-we present the errors, $\Err(T)$,
-normalized by the shifted simulation length, $T+t_0$,
-so that they could
-stay constant in the limit of single-bin updating scheme
-($\sigma \to 0$).
-%
-As shown in Fig. \ref{fig:errsigscan}(a)
-that the minimal error was achieved
-by the widest window %for the shorter simulation of $T = 10^8$,
-for the global sampling process,
-%
-although
-the minimum at $\sigma = 0$ became more favorable
-as the simulation lengthens.
-%
-However,
-for the local sampling process,
-the minimum stayed at zero width,
-as shown in Fig. \ref{fig:errsigscan}(b).
-
-In contrast,
-the normalized errors of the bandpass updating schemes
-were insensitive to the simulation length,
-in agreement with Eq. \eqref{eq:error_sinc}.
-%
-The bandpass updating schemes
-yielded smaller errors than
-the corresponding Gaussian ones,
-although the differences tended to
-be small for shorter simulations
-with the local sampling process.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsigscan.pdf}
-  }
-  \caption{
-    \label{fig:errsigscan}
-    Normalized error, $(T + t_0) \, \Err(T)$,
-    versus the width, $\sigma$,
-    for the Gaussian
-    and the bandpass updating schemes,
-    with $t_0$ given by Eq. \eqref{eq:t0_sinc}.
-    %
-    In the latter case,
-    the equivalent width
-    $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
-    %
-    The points were obtained from averaging results
-    from 1000 independent simulations;
-    the curves are theoretical predictions.
-    %
-    \note{This figure was produced by
-      \texttt{doc/fig/errsigscan.gp}.
-      The data were saved under
-      \texttt{data/scan}.
-    }%
-  }
-\end{center}
-\end{figure}
-
-\section{\label{sec:conclusion}
-Conclusions and Discussions}
-
-In conclusion,
-we have proposed a method of computing
-the optimal schedule of the updating magnitude
-for a general class of free energy calculation methods,
-which we refer to as adaptive flat-distribution-sampling (FDS) methods
-that include the Wang-Landau (WL) algorithm and metadynamics.
-%
-Adaptive FDS methods
-incrementally update a bias potential
-along a collective variable
-to offset the potential of mean force (PMF)
-and thereby sample a flat distribution.
-%
-The optimal schedule delivers the fastest convergence
-of the bias potential,
-and thus helps minimize the effort
-of computing the PMF.
-%They differ by the updating schemes,
-%which specify how the bias potential is modified
-%in each updating step.
-%%
-%In the single-bin or WL case,
-%the update is limited to the current bin;
-%while metadynamics,
-%the bias potential of a neighborhood of the current one
-%is updated with relative weight specified by a Gaussian window.
-
-The optimal schedule depends on the updating window function,
-which is limited to a single-bin in the WL case,
-but can be extended to several neighboring bins
-taking a Gaussian shape in the metadynamics case.
-%
-More generally,
-each adaptive FDS method is associated with
-an updating scheme
-that can be mathematically characterized
-by an updating matrix.
-%
-Each column of the matrix gives the normalized
-window function for the bin corresponding to the column.
-%
-Our method computes the optimal schedule from
-the eigenvalues of the updating matrix,
-and the integrals of the autocorrelation functions
-of the eigenmodes.
-%
-For the single-bin (WL) scheme,
-the optimal schedule from our method
-is given by the inverse-time formula,
-Eq. \eqref{eq:alpha_invt1}.
-%
-For a general multiple-bin scheme,
-including the Gaussian one used in metadynamics,
-the optimal schedule is not given by a simple closed form,
-but is implicitly given by an equation of motion
-of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
-%
-In this case,
-the optimal schedule and error
-can be sensitive to the simulation length.
-%
-The initial updating magnitude is always
-half of the previous equilibrium value.
-% the value used during equilibration.
-
-We have also shown that
-the single-bin scheme is optimal
-in the long time limit
-without a priori assumptions
-about the smoothness of the PMF.
-%
-However, if some eigenmodes
-correspond to local noise
-that is expected to be absent from the PMF,
-we can construct a
-class of bandpass updating schemes
-that are also optimal for asymptotic convergence.
-%
-The optimal schedule of these bandpass schemes
-is always given by an inverse-time formula.
-
-Although the bandpass updating schemes
-appeared to outperform similar Gaussian schemes
-in our model study,
-they are global in nature
-and hence more expensive.
-%
-However, the performance gain
-is relatively small for short simulations
-with a local sampling process such as MD.
-%
-\note{
-One way to improve the Gaussian updating schemes
-is to use it with a hard cutoff of eigenmodes
-(as we did for Fig. \ref{fig:optacmp}).
-%
-Then the optimal schedule
-would be similar to the inverse-time one,
-and the error, with the short-wavelength eigenmodes filtered,
-would be close to that of the bandpass scheme.
-}
-
-The analysis in this study has two limitations.
-%
-First, we have ignored the \emph{systematic}
-error\cite{zhou2005, morozov2007, zhou2008}
-caused by the updates in adaptive FDS methods.
-%
-This is valid in the asymptotic regime:
-when the updating magnitude is small
-and the sampling is a quasi-equilibrium one\cite{
-  zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
-the random error,
-which is roughly proportional to
-the square root of the updating magnitude\cite{
-  zhou2005, morozov2007, zhou2008, bussi2006},
-can easily outweigh
-the systematic error,
-which is roughly proportional to
-the magnitude itself\cite{morozov2007}.
-%
-%Our method of analysis is most useful for long simulations
-%to achieve relatively precise results.
-%
-Second, by using the white noise approximation,
-we have represented the autocorrelation function
-of each eigenmode by a single number, $\Gamma_k$.
-%
-Although this assumption appeared to work well
-on our test system for modeling long simulations,
-they may break down for relatively short simulations
-of glassy systems.
-%Our immediate interest is to apply these relations to
-%problems containing aqueous mixtures of proteins and nucleic acids.
-
-\section{Acknowledgments}
-
-We thank Dr. Y. Mei, J. Wang,
-O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
-Dr. C. Myers, and Dr. O. Samoylova
-for helpful discussions.
-%
-We gratefully acknowledge the Robert A. Welch Foundation (H-0037),
-the National Science Foundation (CHE-1152876)
-and the National Institutes of Health (GM-037657)
-for partial support of this work.
-%
-JM thanks support from the National Institutes of Health (R01-GM067801, R01-GM116280),
-the Welch Foundation (Q-1512),
-and National Science Foundation (Grant PHY-1427654).
-%
-This research used computing resources of the National Science Foundation XSEDE grid.
-%
-%
-
-\appendix
-
-\section{\label{sec:equilerr}
-Inverse-time schedule as a result of increasing sample size
-in a runtime average
-}
-
-We can understand the optimality of the inverse-time schedule
-as a natural consequence of the stabilization of a runtime average
-as the sample size increases\cite{
-  marsili2006, barducci2008}.
-%
-Consider a sample, $\{X^{(1)}, \dots, X^{(N)}\}$,
-with a dynamically growing size.
-%
-The runtime average
-%
-\begin{equation*}
-  {\bar X}^{(N)}
-  =
-  \frac{ X^{(1)} + \cdots + X^{(N)} } { N }
-  ,
-\end{equation*}
-%
-can be written as a recurrence relation regarding
-the addition of the last data point,
-%
-\begin{align}
-  {\bar X}^{(N)}
-  &=
-  {\bar X}^{(N-1)} + \frac{ X^{(N)} - {\bar X}^{(N-1)} } { N }
-  .
-  \label{eq:av_recur}
-\end{align}
-%
-If we interpret $X^{(N)} - {\bar X}^{(N-1)}$
-as the independent correction to the previous average,
-${\bar X}^{(N-1)}$, proposed by the latest data point, $X^{(N)}$,
-then this correction is absorbed into the average with
-a diminishing weight $1/N$ as the sample size, $N$, increases.
-
-The structural similarity between the recurrence relation,
-Eq. \eqref{eq:av_recur},
-and the WL updating scheme, Eq. \eqref{eq:wl_update},
-suggests that the bias potential may represent
-a runtime average in a particular case.
-%
-To establish a mapping between the two,
-let us now consider a long equilibrium FDS simulation
-that consists of a sequence of sub-simulations
-of $L$ steps.
-%
-If $X^{(N)}$ is
-the bias potential, $V_i^{(N)}$, computed
-from the $N$th sub-simulation,
-then ${\bar X}^{(N)}$ means the average bias potential,
-${\bar V}_i^{(N)}$,
-from the first $N$ sub-simulations.
-%
-Since this average, ${\bar V}_i^{(N)}$,
-gives the best estimate at the end of
-the first $N$ sub-simulations,
-we should ideally use it as the bias potential
-for the next sub-simulation,
-until it is further updated by Eq. \eqref{eq:av_recur}
-at the end of the next sub-simulation.
-%
-Assuming that the normalized histogram, $H_i^{(N)}$,
-from the $N$th sub-simulation is sufficiently precise,
-Eq. \eqref{eq:vcorr_equil}
-\big[with
-$V_i^\mathrm{corr} \to V_i^{(N)}$, $V_i \to {\bar V}_i^{(N-1)}$\big]
-shows that the independent
-correction from the $N$th sub-simulation is
-%
-\begin{equation*}
-  \ln \frac{ H_i^{(N)} } { p_i }
-  \approx
-  \frac{ H_i^{(N)} } { p_i } - 1
-  ,
-\end{equation*}
-%
-where the ``$-1$'' on the right-hand side
-can be ignored as a uniform shift
-that does not affect the distribution.
-%
-Thus, we may rewrite Eq. \eqref{eq:av_recur} as
-\begin{align}
-  {\bar V}_i^{(N)}
-  &=
-  {\bar V}_i^{(N-1)}
-  +
-  \frac{1}{N} \frac{ H_i^{(N)} }{ p_i }
-  \notag
-  \\
-  &=
-  {\bar V}_i^{(N-1)}
-  +
-  \sum_{l=1}^L \frac{1}{N \, L} \frac{ h_i^{(N)}(l) }{ p_i }
-  ,
-  \label{eq:vbar_update}
-\end{align}
-%
-where $h_i^{(N)}(l)$ stands for the instantaneous histogram
-at step $l$ of sub-simulation $N$,
-and its average is
-$H_i^{(N)} \equiv \frac{1}{L} \sum_{l=1}^L h_i^{(N)}(l)$.
-%
-Now Eq. \eqref{eq:vbar_update} shows that the best estimate
-of the bias potential from the runtime average can be computed
-through a step-by-step incremental updating scheme that is equivalent to
-the WL algorithm, Eq. \eqref{eq:wl_update},
-with the updating magnitude being
-%
-\begin{equation}
-  \alpha(t) = 1/(NL)
-  .
-  \label{eq:alpha_block}
-\end{equation}
-%
-In the long time limit,
-we may approximate the simulation time
-by the product of the number of sub-simulations
-and the number of steps in each sub-simulation, or $N \, L$,
-%
-and Eq. \eqref{eq:alpha_block} recovers Eq. \eqref{eq:alpha_invt}.
-
-The above demonstration implies several connections
-between the WL algorithm under the optimal inverse-time schedule
-with a few related FDS methods.
-%
-First,
-it establishes an efficiency equivalence
-between an adaptive WL simulation under the inverse-time schedule
-and the ideal equilibrium FDS counterpart
-with a nearly perfect bias potential.
-%
-The bias potentials from the two approaches
-would reach similar precision at long times.
-%
-Thus, it is unnecessary to convert
-a sufficiently convergent adaptive FDS simulation
-to the corresponding equilibrium FDS simulation
-with a fixed bias potential just for better precision
-of the final PMF.
-
-Second,
-it also relates a WL simulation under the inverse-time schedule
-to a time-averaged variant of the regular WL simulation.
-%
-In the variant, the updating magnitude is fixed,
-but the bias potential over the simulation course is averaged
-as the final result\cite{zhou2005}.
-%
-The above demonstration shows that the inverse-time schedule
-has already effectively implemented an implicit average
-so that no explicit averaging is needed in the former approach.
-%
-The inverse-time schedule approach is preferred, however,
-because with a decreasing updating magnitude,
-it minimizes the systematic error caused by
-the on-the-fly updating scheme\cite{
-  zhou2005, zhou2008},
-which is ignored in this analysis.
-
-Third,
-it is possible to reformulate the WL algorithm
-under the inverse-time schedule
-as an explicit adaptive average method
-such that the bias potential is computed from
-a weighted runtime average of the histogram\cite{
-  marsili2006, barducci2008},
-which is analogous to the adaptive bias force method\cite{
-  darve2001, *darve2008}.
-
-\section{\label{sec:wltoinvt}
-Inverse-time schedule as the optimal limit of the WL algorithm}
-
-We may understand the inverse-time schedule %, Eq. \eqref{eq:},
-as an optimal limit of the stage-wise WL updating scheme.
-%
-Our calculation below will focus on the one-variable problem,
-as in Sec. \ref{sec:onevar}.
-
-To begin, we split the entire simulation into $S$ stages,
-each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
-%
-We denote the number of steps in stage $s$ as $T_s$, and
-$\sum_{s=1}^S T_s = T$.
-%
-%\begin{equation}
-%  \sum_{s = 1}^S T_s = T
-%  ,
-%  \label{eq:sumT}
-%\end{equation}
-%
-We assume that
-the magnitude follows geometrically,
-%
-\begin{equation}
-  \alpha_s = \alpha_0 \, r^s
-  ,
-  \label{eq:alphas}
-\end{equation}
-%
-with $0 < r < 1$.
-%
-%and that the system is properly equilibrated
-%under a constant magnitude initially.
-%
-We wish to find the best distribution of the simulation time
-into the $S$ stages
-as well as
-the optimal number of stages and $r$.
-
-If we denote the error at the end of stage $s$ as $E_s$,
-then from Eq. \eqref{eq:x2t_average}, we have
-$q(t) = \alpha_s \, t$, and
-%
-\begin{align}
-  E_s
-  &=
-  E_{s - 1} \, e^{-2 \, \alpha_s \, T_s }
-  \notag\\
-  &\phantom{=}
-  +
-  \int_0^{T_s} \int_0^{T_s}
-  \alpha_s^2 \,
-  e^{ -\alpha_s \, (T_s - t) }
-  e^{ -\alpha_s \, (T_s - t') }
-  \kappa( t - t' ) \, dt \, dt'
-  \notag
-  \\
-  &=
-  \left(
-    E_{s - 1} - \frac{ \Gamma \, \alpha_s } { 2 }
-  \right)
-  e^{ - 2 \, \alpha_s \, T_s }
-  +
-  \frac{ \Gamma \, \alpha_s } { 2 }
-  .
-  \label{eq:errorstage}
-\end{align}
-%
-where we have assumed
-$\kappa(t) = \Gamma \, \delta(t)$
-in the second step.
-%
-Our aim is to minimize the error
-of the final stage, $E_S$.
-%
-To do so, let us consider shifting simulation time
-by $\delta T$ from
-stage $s+1$ to stage $s$,
-then $E_{s+1}$ is affected as
-%
-\begin{align*}
-  \delta E_{s+1}
-  =
-\left[
-  2 \, \alpha_{s+1}
-  \left(
-    E_s - \tfrac{ \Gamma \, \alpha_{s+1} } { 2 }
-  \right)
-  +
-  \frac{ \partial E_s } { \partial T_s }
-\right]
-e^{-2 \, \alpha_{s+1} \, T_{s+1} }
-\,
-\delta T
-.
-\end{align*}
-%
-Since this variation does not directly affect the error of later stages,
-the above expression must be zero to minimize the final error, i.e,
-%
-\begin{align*}
-  2 \, \alpha_{s+1}
-  \left(
-    E_s - \frac{ \Gamma \, \alpha_{s+1} } { 2 }
-  \right)
-  &=
-  -\frac{ \partial E_s } { \partial T_s }
-  %\\
-  %&=
-  %2 \, \alpha_s \, e^{-2 \, \alpha_s \, T_s}
-  %\left(
-  %  E_{s-1} - \frac{ \Gamma \, \alpha_s } { 2 }
-  %\right)
-  %\\
-  =
-  2 \, \alpha_s \, \left(E_s - \frac{ \Gamma \, \alpha_s } { 2 }  \right)
-  ,
-\end{align*}
-%
-where we have used Eq. \eqref{eq:errorstage}
-in the second step.
-%
-Thus,
-\begin{equation}
-  E_s
-  =
-  \frac{ \Gamma } { 2 } \, ( \alpha_s + \alpha_{s+1} )
-  =
-  \frac{ \Gamma } { 2 } \, \alpha_s ( 1 + r )
-  .
-  \label{eq:Es}
-\end{equation}
-%
-Using Eqs. \eqref{eq:Es} and \eqref{eq:alphas}
-in Eq. \eqref{eq:errorstage}, we get
-%
-\begin{equation}
-  e^{-\alpha_s \, T_s } = r
-  .
-  \label{eq:exp_r}
-\end{equation}
-%
-
-To compute the optimal schedule,
-we map the simulation time, $t$, to the stage number, $s$, as
-%
-\begin{align*}
-  t
-  = \sum_{s' = 1}^s T_{s'}
-  = \frac{ 1 - r^s } { 1 - r } \, T_s
-  = \frac{ 1 } { \gamma } \,
-  \left(
-    \frac{1}{\alpha_s}
-    -
-    \frac{1}{\alpha_0}
-  \right)
-  ,
-\end{align*}
-where
-$\gamma = ( 1 - r ) / \ln(1/r)$,
-and we have used Eq. \eqref{eq:exp_r} in the last step.
-%
-Thus,
-%
-\begin{equation}
-  \alpha_s =
-  \frac{ 1 }
-  { \gamma \, t + \alpha_0^{-1} }
-  ,
-  \label{eq:alphas_t}
-\end{equation}
-%
-and the final error is given by
-%
-\begin{align}
-  E_S
-  =
-  \frac{\Gamma}{2}
-  \frac{ 1 + r }
-  { \gamma \, T + \alpha_0^{-1} }
-  ,
-  \notag
-\end{align}
-%
-which is minimized as $r \to 1$
-for a large $T$
-(at this point, $\gamma \to 1$).
-%
-\note{
-  For a large $T$, $E_S \approx \frac{\Gamma}{2T} \frac{1+r}{\gamma}$.
-  Define $\delta = 1 - r$,
-  we have $1/\gamma = 1+ \delta/2 + \delta^2/3 + \cdots$,
-  and
-  $$
-  \frac{1+r}{\gamma}
-  =
-  (2 - \delta)
-  \left(
-  1 + \frac{\delta}{2} + \frac{\delta^2}{3} + \cdots
-  \right)
-  =
-  2 + \sum_{n=1}^\infty \frac{n-1}{n(n+1)}\delta^n \ge 2
-  .
-  $$
-}
-%
-If we identify the parameter, $\alpha_0$,
-as the initial updating magnitude, $\alpha(t = 0)$,
-then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
-and the optimal $t_0 = \alpha_0^{-1}$.
-%
-Thus, in this limit,
-the schedule given by Eq. \eqref{eq:alphas_t}
-is identical to that from Eq. \eqref{eq:alpha_invt1}.
-
-\section{\label{sec:Gamma}
-Integrals of the autocorrelation functions}
-
-Here, we evaluate the integrals of
-the autocorrelation functions,
-$\Gamma_k$'s,
-in two special cases
-in Secs. \ref{sec:Gamma_perfect}
-and \ref{sec:Gamma_onestep}
-in the asymptotic limit,
-and describe a general method of measuring them
-in Sec. \ref{sec:Gamma_measure}.
-
-\subsection{\label{sec:Gamma_perfect}
-Perfect sampling}
-
-First, if the sampling is perfect,
-%Eq. \eqref{eq:kappa_delta} is exact, and
-%
-%\begin{equation}
-%  \Gamma_k = 1.
-%\label{eq:Gamma_perfect}
-%\end{equation}
-%
-%To see this, we note that
-the bin index, $i$, is an independent random variable
-at each time step $t$.
-%
-Further, in the asymptotic limit,
-$\langle h_i(t) \rangle = p_i$,
-we find from Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def} that
-\begin{equation*}
-  \xi_i(t)
-  =
-  \frac{ h_i(t) } { p_i }
-  -
-  \sum_{ r = 1 }^n
-    \rho_r \frac{ h_r(t) } { p_r }
-  ,
-\end{equation*}
-%
-and
-%
-\begin{align}
-  \left\langle
-    \xi_i(t) \, \xi_j(t')
-  \right\rangle
-  %&=
-  %\left\langle
-  %  \left(
-  %    \frac{ h_i(t) } { p_i }
-  %    -
-  %    \sum_{ r = 1 }^n
-  %      \rho_r \frac{ h_r(t) } { p_r }
-  %  \right)
-  %\right.
-  %\notag \\
-  %&\hphantom{==}\cdot
-  %\left.
-  %  \left(
-  %    \frac{ h_j(t') } { p_i }
-  %    -
-  %    \sum_{ s = 1 }^n
-  %      \rho_s \frac{ h_s(t') } { p_s }
-  %  \right)
-  %\right\rangle
-  %\notag
-  %\\
-  &=
-  \left(
-    \frac{ \delta_{ij} } { p_i }
-    -
-    \frac{ \rho_i } { p_i }
-    -
-    \frac{ \rho_j } { p_j }
-    +
-    \sum_{r = 1}^n
-    \frac{ \rho_r^2 } { p_r }
-  \right) \,
-  \delta_{t, t'}
-  .
-\label{eq:zz_perfect}
-\end{align}
-%
-%$\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
-%in the continuous limit.
-%
-\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
-  with $t \ne t'$ (going back to the discrete time case),
-  $$
-  \begin{aligned}
-  \left\langle
-    \xi_i(t) \, \xi_j(t')
-  \right\rangle
-  &=
-  \left\langle
-    \xi_i(t)
-  \right\rangle
-  \left\langle
-    \xi_j(t')
-  \right\rangle
-  =
-  0.
-  \end{aligned}
-  $$
-  Next, for $t = t'$, we have
-  $$
-  \langle
-    h_i(t) \, h_j(t)
-  \rangle
-  = \delta_{ij} \, p_i
-  ,
-  $$
-  which leads to Eq. \eqref{eq:zz_perfect}.
-}
-%
-Thus,
-for $k \ge 2$, we get
-%
-\begin{equation}
-  \kappa_k(t - t')
-  =
-  \sum_{i,j = 1}^n
-  \phi_{ki} \phi_{kj}
-  \langle \xi_i(t) \, \xi_j(t') \rangle
-  %
-  =
-  \sum_{i = 1}^n \frac{ \phi_{ki}^2 } { p_i }
-  \, \delta_{t, t'},
-\notag
-%\label{eq:kappa_perfect}
-\end{equation}
-%
-where we have used
-Eqs. \eqref{eq:ortho1},
-\eqref{eq:theta_def},
-and
-\eqref{eq:zz_perfect}.
-%
-Thus, Eq. \eqref{eq:kappa_delta} is exact,
-with
-$\Gamma_k = \sum_{ i = 1}^n \phi_{ki}^2 / p_i$.
-\note{Proof.%Eq. \eqref{eq:kappa_perfect},
-  $$
-  \begin{aligned}
-  \kappa_k(t - t')
-  &=
-  \sum_{i,j = 1}^n
-  \phi_{ki}
-  \phi_{kj}
-  \langle \xi_i(t) \, \xi_j(t') \rangle
-  \\
-  %
-  &=
-  \sum_{i,j = 1}^n
-  \phi_{ki}
-  \phi_{kj}
-  \left(
-    \frac{ \delta_{ij} } { p_i }
-    -
-    \frac{ \rho_i } { p_i }
-    -
-    \frac{ \rho_j } { p_j }
-    +
-    \sum_{r = 1}^n
-    \frac{ \rho_r^2 } { p_r }
-  \right) \,
-  \, \delta_{t, t'}
-  .
-  \end{aligned}
-  $$
-  where,
-  we have used
-  Eq. \eqref{eq:theta_def}
-  on the first line,
-  %
-  Eq. \eqref{eq:zz_perfect}
-  on the second line.
-  %
-  With Eq. \eqref{eq:ortho1},
-  only the first term in the parentheses survives.
-}
-In particular, if $p_i = \rho_i$, then, by
-Eq. \eqref{eq:eig_orthonormal_rows}, we have
-$\kappa_k(t - t') = \delta_{t, t'}$, or
-\begin{equation}
-  \Gamma_k = 1
-  .
-\label{eq:Gamma_perfect}
-\end{equation}
-
-\subsection{\label{sec:Gamma_onestep}
-One-step process}
-
-Another special case is the
-one-dimensional one-step process\cite{vankampen},
-which models a local sampling process.
-In each step, the system hops to
-one of the two neighboring bins
-with equal probability, $1/2$.
-%
-The transition matrix is
-%
-\begin{equation}
-  \arraycolsep=3.6pt\def\arraystretch{1.4}
-  \mathbf A
-  =
-  \left(
-    \begin{array}{cccccccc}
-      \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
-      \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
-      \vdots & &  & & \vdots \\
-      \frac {1 - \sigma} 2 & \dots &  & \frac 1 2 & \frac \sigma 2
-    \end{array}
-  \right),
-\notag
-%\label{eq:T_nn}
-\end{equation}
-%
-where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
-%
-For many updating schemes
-(cf. Sec. \ref{sec:band-matrix}),
-we can further assume that
-the matrices $\mathbf A$ and $\mathbf w$
-share the eigenvectors,
-with $\rho_i = p_i = 1/n$.
-%but not the eigenvalues.
-%
-The eigenvalues of $\mathbf A$ are
-$$
-  \Lambda_k
-  =
-  \cos \frac{ (k - 1) \, g \, \pi }
-            {            n        }
-  ,
-$$
-where
-$g = 2$ for a periodic variable, or $1$ otherwise.
-%Note that these $\Lambda_k$'s are
-%not to be confused with the eigenvalues of $\mathbf w$,
-%$\lambda_k$'s.
-%
-In this case,
-the autocorrelation function, $\kappa_k(t)$,
-decays as $\kappa_k(0) \, \Lambda_k^t$,
-with $\kappa_k(0) = 1$ being the same as
-the value in perfect sampling.
-%
-Thus,
-%$\Gamma_k$ is roughly twice the autocorrelation time, and
-from Eq. \eqref{eq:Gamma_sum}, we get
-%
-\begin{equation}
-  \Gamma_k
-  =
-  1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
-  =
-  \frac{ 1 + \Lambda_k }
-       { 1 - \Lambda_k }
-  =
-  \cot^2 \left[
-    \frac{ (k - 1) \, g \, \pi }
-         { 2 \, n }
-  \right]
-  .
-\label{eq:Gamma_onestep}
-\end{equation}
-
-\subsection{\label{sec:Gamma_measure}
-Measurement via an adaptive FDS simulation
-}
-
-%To use Eq. \eqref{eq:q_opt} in a real simulation,
-%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
-%
-For an unknown system,
-we can use Eq. \eqref{eq:y2_eql} to
-measure $\Gamma_k$'s in
-an adaptive FDS simulation
-with a constant schedule,
-$\alpha(t) = a_0$.
-%
-%A practical difficulty is that
-However,
-we often do not know a priori the intrinsic distribution, $p^*_i$,
-and hence
-cannot compute the deviation
-of the bias potential, $V_i$.
-%
-To circumvent the problem, we
-modify Eq. \eqref{eq:y2_eql} as
-%
-\begin{equation}
-  \operatorname{var} Y_k
-  =
-  \frac{1}{2} \,
-  a_0 \, \Gamma_k \, \lambda_k,
-\label{eq:varY}
-\end{equation}
-%
-where
-``$\operatorname{var}$''
-means the variance from the sampling (time) average,
-and
-%
-\begin{align*}
-Y_k
-&=
-\sum_{ i = 1 }^n
-  \phi_{k i} \, (V_i - \bar V),
-%\label{eq:Y_def}
-\end{align*}
-%
-with
-$
-\bar V
-\equiv
-\sum_{ j = 1 }^n \rho_j \, V_j.
-$
-
-To show Eq. \eqref{eq:varY},
-we first note that
-the difference, $V_i - {\bar V}$,
-differs from the actual deviation, $x_i$,
-defined in Eq. \eqref{eq:x_def},
-only by a constant shift,
-$\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
-%
-Thus, the two share the same variance
-under time averaging.
-%
-It follows that the corresponding linear combinations,
-$Y_k$ and $y_k$, would also share the same variance.
-%
-But since the average of $y_k$ is zero,
-we have
-$$
-\operatorname{var} Y_k
-=
-\operatorname{var} y_k
-=
-\left\langle y_k^2 \right\rangle.
-$$
-Finally, by using Eq. \eqref{eq:y2_eql},
-we reach Eq. \eqref{eq:varY}.
-
-\note{In practice, Eq. \eqref{eq:varY}
-in inapplicable to zero eigenvalues, $\lambda_k$'s.
-%
-However, these eigenmodes can be ignored
-as they do not contribute to the Lagrangian defined by
-Eq. \eqref{eq:Lagrangian_const}.
-}
-
-As an example, we used the method to compute
-the integrals, $\Gamma_k$'s, from
-a local MC simulation
-(cf. Sec. \ref{sec:results_system})
-and an MD simulation,
-and compared them with those from
-the one-step sampling process.
-%
-As shown in Fig. \ref{fig:gamcmp},
-the $\Gamma_k$'s of the small-$k$ eigenmodes
-from the three sampling processes
-were almost proportional,
-suggesting that
-the optimal schedule
-for simulations using local sampling processes
-would be similar in general properties.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/gamcmp.pdf}
-  }
-  \caption{
-    \label{fig:gamcmp}
-    Integrals of the autocorrelation functions
-    from a local MC simulation
-    and an MD simulation.
-    %
-    Each simulation was ran for $10^{10}$ steps.
-    %
-    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$.
-    %
-    For the MD simulation,
-    the step size $\Delta t_\mathrm{MD} = 0.005$.
-    %
-    The velocity rescaling thermostat\cite{bussi2007} was
-    applied for the unit reduced temperature
-    every MD step with viscosity $20$.
-    %
-    The reflective boundary condition was used.
-    %
-    \note{The figure was produced by \texttt{doc/fig/gamcmp.gp}.
-      The data were saved under \texttt{data/gamcmp}
-      (see the \texttt{Makefile} there).
-    }%
-  }
-\end{center}
-\end{figure}
-
-\section{\label{sec:procedure}
-Procedure for computing the optimal schedule and error
-}
-
-Here we give a procedure for computing
-the optimal schedule and the error
-of a given adaptive FDS simulation.
-%
-Basically, we need to extract
-$\lambda_i$'s and $\Gamma_i$'s
-from the given system and updating scheme (input),
-and then apply them to Eq. \eqref{eq:q_opt}
-for the schedule,
-as shown in Fig. \ref{fig:vardep}.
-%
-%The dependence of variables are shown in Fig. \ref{fig:vardep}.
-%
-While the updating scheme alone
-determines $\lambda_k$'s,
-$\Gamma_k$'s depend also on the system
-and the underlying sampling process (MC or MD).
-%
-By using the technique in Sec. \ref{sec:Gamma_measure}
-to measure $\Gamma_k$'s,
-we have the following procedure.
-
-\begin{figure}[h]
-\begin{singlespace}
-  \fontsize{10pt}{\baselineskip}
-  \begin{tikzpicture}
-  \node[draw, text width={width("Updating ")}]
-    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
-
-  \node[draw, text width={width("sampling")}]
-    (sys) at (0, 1) {System\\and\\sampling\\process};
-
-  \node[draw, text width={width("Eigenvalues,")}]
-    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_2, \dots, \lambda_n$};
-
-  \node[draw, text width={width("Integrals, ")}]
-    (gamma) at (2.2, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
-
-  \node[draw, text width={width("Simulation")}]
-    (T) at (2.2, -0.5) {Simulation\\time, $T$};
-
-  \node[draw, text width={width("Optimal")}]
-    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};
-
-  \node[draw, text width={width("Schedule,")}]
-    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};
-
-    \node[draw, text width={width("Error,")}]
-    (err) at (7, 0) {Error, $\Err(T)$};
-
-  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};
-
-  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};
-
-  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
-  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
-  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};
-
-  \draw[->, >=latex, thick]
-    (w) edge (lambda)
-    (w) edge (gamma)
-    (sys) edge (gamma);
-
-  \draw[->, >=latex, thick]
-    (lambda) -| (M1)
-    (gamma)  -- (M1)
-    (T)      -| (M1)
-    (M1)     -- (R1)
-    (R1)     |- (qT);
-
-  \draw[->, >=latex, thick]
-    (qT)     -| (R3);
-
-  \draw[->, >=latex, thick]
-    (R3) -- (N1) |- (alpha);
-
-  \draw[->, >=latex, thick]
-    (N1) |- (err);
-
-  \draw[->, >=latex, thick]
-    %(M1) edge[bend left=70] (N1);
-    (R1) |- (R2) -| (R3);
-\end{tikzpicture}
-\end{singlespace}
-\caption{
-  \label{fig:vardep}
-  Dependence of variables in computing
-  the optimal schedule and the error.
-}
-\end{figure}
-
-\begin{enumerate}
-
-\item
-From the updating matrix, $\mathbf w$,
-compute the eigenvalues, $\lambda_k$'s.
-
-\item
-Run a preliminary adaptive FDS simulation
-%using the updating scheme
-with a constant updating magnitude, $a_0$.
-
-\item
-Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}.
-
-\item
-Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
-
-\item
-Compute the optimal schedule by
-numerically integrating Eq. \eqref{eq:q_opt}.
-
-\item
-  Compute the error $\Err(T)$ from
-Eqs. \eqref{eq:error_split},
-  \eqref{eq:error_res1},
-  and
-  \eqref{eq:error_asym2}.
-
-\end{enumerate}
-
-\subsection{\label{sec:more_wband}
-Homogeneous updating schemes
-}
-
-Here, we present some detailed derivations
-on the homogeneous updating schemes.
-%
-By the translational invariance,
-homogeneous updating schemes
-satisfy detailed balance,
-Eq. \eqref{eq:w_detailedbalance},
-for the flat distribution, $\rho_i = 1/n$,
-i.e. the updating matrix, $\mathbf w$,
-is symmetric.
-%
-The updating matrices
-share the same set of eigenvectors
-(which here are sines and cosines of different frequencies),
-and thus are completely determined
-by the eigenvalues.
-%
-Further, the multiplication of $\mathbf w$
-can be reduced to a convolution with the window function,
-or the updating kernel.
-%
-Thus,
-there is a one-to-one correspondence between
-the eigenvalues of $\mathbf w$
-and the updating kernel.
-
 \subsection{\label{sec:stabilize_wband}

### 2017-04-19 doc/invt.tex
further small edits to invt.tex

@@
   We will show that
-  the optimal schedule depends on the updating scheme (WL or metadynamics),
+  the optimal schedule depends on the updating scheme
+  (such as the single-bin scheme commonly used in WL
+  or the Gaussian one popular in metadynamics),
   and the initial updating magnitude is ideally half of
@@
   Further,
-  the WL updating scheme
+  the single-bin updating scheme
   belongs to a class of bandpass updating schemes
@@
 when the collective variable is energetically restricted to local regions
-of the complex, multi-modal free energy surface.
+of the complex, multimodal free energy surface.
 %
@@
 *laio2002, *laio2008, *barducci2011, *sutto2012},
-more commonly used for MC and MD simulations, respectively\cite{junghans2014}.
+popular for MC and MD simulations, respectively\cite{junghans2014}.
 %
@@
 which specifies
-a neighborhood around the current value of $z$ on the free energy surface
+a neighborhood around the current value of
+$z$ on the free energy surface
 where the updating should occur
@@
 %
-Although the assignment of the window functions to the two methods
-is not absolute\cite{micheletti2004, kim2006, *kim2007, junghans2014},
-below we will use the terms WL and metadynamics
+Although the two window functions
+are not always associated with
+the two methods\cite{micheletti2004, kim2006, *kim2007, junghans2014},
+we will loosely use the terms WL and metadynamics
 to represent FDS simulations using the
@@
 %
-The above two introductory sections help fix our notations,
+The above two introductory sections help define our notations,
 and an expert reader may wish to skip them.
@@
 
-Note that the error weighting distribution, $\pmb\rho$,
-defined above
+Although the above error weighting distribution, $\pmb\rho$,
+is usually flat,
+i.e. $\rho_i \equiv 1/n$,
+in practice,
+a non-flat distribution
+is sometimes more convenient in analyzing
+some multiple-bin updating schemes later on.
+%
+We also note that $\pmb\rho$
 is independent of
 the target distribution,
-$\mathbf p = (p_1, \dots, p_n)$.
-%
-In practice, $\pmb\rho$ is usually flat,
-i.e. $\rho_i \equiv 1/n$,
-even when $\mathbf p$ is not
+$\mathbf p = (p_1, \dots, p_n)$
 (e.g. in well-tempered metadynamics,
@@
   barducci2008, dama2014}:
-$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$).
+$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
+whereas $\pmb\rho$ is flat).
 %
@@
 Next, we split $h_i(t)$ into a deterministic averaged part
-and a stochastic part, $\zeta_i(t)$,
+and a random fluctuation part, $\zeta_i(t)$,
 %
@@
 %
-For the random noise part,
+For the histogram fluctuation part,
 the average is zero by definition,
@@
 %
-and we will assume it is white, i.e.
+and we will approximate it as white noise, i.e.
 %The noise is not necessarily white,
@@
 %
-where
+where the modified histogram fluctuation
 %
@@
 \big[because $q(T) = \ln\bigl[1 + (T/t_0)\bigr]$,
-following the $t = 0$ case of Eq. \eqref{eq:expqt}\big].
-%
-%In Sec. \ref{sec:finlencorr},
-%we show that % $t_0$ can be adjusted
-%one can further improve the schedule by adjusting $t_0$
-%to accommodate the initial error.
-%
-%and
-%$$
-%\left\langle x^2(T) \right\rangle
-%=
-%\frac{ \left\langle x^2(0) \right\rangle \, t_0^2 }
-%     { (t_0 + T)^2 }
-%+
-%\frac{ \Gamma \, T}
-%     { (t_0 + T)^2 }
-%.
-%$$
-%Using this result for each $x_i$,
-%we get the final total error, $E(T)$,
+from the $t = 0$ case of Eq. \eqref{eq:expqt}\big].
+%
 By using the optimal schedule in Eq. \eqref{eq:ET_average},
@@
   =
-  \frac{ E(0) \, t_0^2 }
-       { (t_0 + T)^2 }
-  +
-  \frac{ \Gamma \, T}
+  \frac{ E(0) \, t_0^2 + \Gamma \, T}
        { (t_0 + T)^2 }
   ,
-  \label{eq:error_singlebin}
+  \notag
+  %\label{eq:error_singlebin}
 \end{equation}
@@
 
-We now consider a multiple-bin updating scheme,
-given in Eq. \eqref{eq:mbin_update}.
-%
-In this case,
-upon a visit to bin $j$,
-we update not only the bias potential at bin $j$,
-but also those at a few neighboring bins, $i$'s.
-%
-The relative updating magnitudes,
-$w_{ij}$'s,
-form a matrix, $\mathbf w$.
+We now turn to a general multiple-bin updating scheme.
+%
+By definition, Eq. \eqref{eq:mbin_update},
+a visit to bin $j$ in this case results in updates of the bias potential
+not only at bin $j$, but also at a few neighboring bins $i$'s.
 %
@@
 projection of the bias potential to the $n$ eigenmodes
-of the updating matrix, $\mathbf w$.
+of the updating matrix, $\mathbf w$,
+formed by the relative magnitudes, $w_{ij}$'s.
+%
 %However, we will assume some restrictions and approximations
@@
 %
-First, to sample
-the desired distribution,
+First, we have a fixed-point condition\cite{bussi2006, dama2014}.
+%
+To sample the desired distribution,
 $\mathbf p = (p_1, \dots, p_n)$,
@@
 and borrow techniques used
-in studying transition matrices\cite{vankampen}.
-%
-The quoted results were unaffected by the allowance
-of negative elements into the matrix, $\mathbf w$.
+in studying transition matrices\cite{vankampen}
+(we will only use results that are unaffected by the allowance
+of negative elements into $\mathbf w$).
 
@@
 %
-for the distribution, $\pmb \rho$,
+for the error weighting distribution, $\pmb \rho$,
 introduced in Eq. \eqref{eq:vbar_def}.
 %
-It follows that
-%
-\begin{equation}
-  \sum_{i = 1}^n \rho_i \, w_{ij}
-  =
-  \sum_{i = 1}^n \rho_j \, w_{ji}
-  =
-  \rho_j,
-  \label{eq:w_balance}
-\end{equation}
-%
-i.e. $\pmb \rho$ is a left eigenvector of
-$\mathbf w$ with eigenvalue $1$.
-%
-In the way,
-the error weighting distribution, $\pmb \rho$,
-restrains the form of the updating matrix, $\mathbf w$.
-%
-The condition, albeit somewhat artificial,
-is readily satisfied in practice:
-most commonly, $\pmb \rho$ is a uniform distribution,
-and this condition requires the updating matrix, $\pmb w$, to be symmetric.
-
-With Eq. \eqref{eq:w_detailedbalance},
+This artificial condition
+restrains the form of the updating matrix, $\mathbf w$,
+and it requires, for the commonly-used uniform distribution, $\pmb\rho$,
+a symmetric updating matrix, $\pmb w$.
+%
+More generally, only
 the scaled matrix, $\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
-is symmetric.
-Thus, we can diagonalize $\mathbf w$ with a set of
+needs to be symmetric,
+and the symmetry allows the diagonalization
+of $\mathbf w$ with a set of
 eigenvectors, $\phi_{ki}$,
@@
 
-In particular, there is always an eigenvector
-corresponding to the distribution, $\pmb \rho$.
+The above symmetry implies the existence of
+a left eigenvector, $\pmb \rho$,
+corresponding to the uniform right eigenvector
+associated with Eq. \eqref{eq:w_sumj},
+with eigenvalue $1$:
+%
+\begin{equation}
+  \sum_{i = 1}^n \rho_i \, w_{ij}
+  =
+  \sum_{i = 1}^n \rho_j \, w_{ji}
+  =
+  \rho_j
+  .
+  \label{eq:w_balance}
+\end{equation}
 %
@@
 %
-which satisfies Eq. \eqref{eq:eig_orthonormal_rows}
-for $k = l = 1$.
-%
-The corresponding eigenvalue is $\lambda_1 = 1$
-by Eq. \eqref{eq:w_balance}.
-%
-Further, for $k \ge 2$, we have
+with $\lambda_1 = 1$.
+%
+Clearly, Eq. \eqref{eq:eigenmode1}
+satisfies the normalization condition
+given by Eq. \eqref{eq:eig_orthonormal_rows}
+for $k = l = 1$,
+while the orthogonality for $k \ge 2$ demands
 %
@@
 
-\subsubsection{Linearization and decoupling to eigenmodes}
+\subsubsection{Eigenmode decomposition}
 
@@
 using the eigenvectors defined in Eqs.
-\eqref{eq:eig_w}-\eqref{eq:eig_orthonormal_rows},
-we get a set of decoupled equations
+\eqref{eq:eig_w}-\eqref{eq:eig_orthonormal_rows}
+yields a set of decoupled equations
 %
@@
 \end{align}
-are the deviation and noise of mode, $k$, respectively.
-
-\subsubsection{Error function}
+%
+are the deviation and noise (histogram fluctuation)
+of mode $k$, respectively.
+
+\subsubsection{Error functional}
 
@@
 from $k = 2$,
+%
 \begin{equation}
@@
 In analogy to Eq. \eqref{eq:ET_average},
-using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
+%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
 we can write the error after a period, $T$,
@@
 %
-\begin{equation}
+\begin{align}
   \Err_A(T)
-  =
-  \int_0^T \int_0^T
+  &=
+  \int_0^T
   \sum_{k = 2}^n
-    \dot u_k\bigl( q(t) \bigr) \,
-    \dot u_k\bigl( q(t') \bigr) \,
-    \kappa_k(t - t') \, dt \, dt'
+  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt
   ,
+\label{eq:error_asym}
+\end{align}
+%
+where we have approximated $\theta_k(t)$
+as white noise:
+%
+\begin{equation}
+  \kappa_k(t)
+  \equiv \langle \theta_k(0) \, \theta_k(t) \rangle
+  \approx \Gamma_k \, \delta(t)
+  .
   \notag
-  %\label{eq:error_asym}
-\end{equation}
-%
-with
-$u_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}$,
-and
-$\kappa_k(t - t') \equiv \left\langle \theta_k(t) \, \theta_k(t') \right\rangle$.
+  %\label{eq:kappa_delta}
+\end{equation}
+%
+%\begin{equation}
+%$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
+%  \notag
+%\label{eq:Gamma_integral}
+%\end{equation}
+
+For the theory to be applicable to
+a general sampling process,
+we can map the histogram fluctuation
+to the equivalent white noise, with
+\begin{equation}
+  \Gamma_k
+  = \sum_{t = -\infty}^\infty \kappa_k(t)
+  %= \sum_{t = -\infty}^\infty \left\langle \theta_k(0) \, \theta_k(t) \right\rangle
+  .
+  \label{eq:Gamma_sum}
+\end{equation}
+%
+In this way,
+the underlying sampling process
+affects the error only through the
+few $\Gamma_k$'s.
+%
+For example, we expect large values of $\Gamma_k$
+from eigenmodes corresponding to long-range fluctation
+in a local sampling process,
+especially for complex or glassy systems.
+%
+In Appendix \ref{sec:Gamma},
+we evaluate $\Gamma_k$'s in two special cases,
+and show how to measure them
+from an adaptive FDS simulation.
+%
+%\begin{equation}
+%  \Err_A(T)
+%  =
+%  \int_0^T \int_0^T
+%  \sum_{k = 2}^n
+%    \dot u_k\bigl( q(t) \bigr) \,
+%    \dot u_k\bigl( q(t') \bigr) \,
+%    \kappa_k(t - t') \, dt \, dt'
+%  ,
+%  \notag
+%  %\label{eq:error_asym0}
+%\end{equation}
+%%
+%with
+%$u_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}$,
+%and
+%$\kappa_k(t - t') \equiv \left\langle \theta_k(t) \, \theta_k(t') \right\rangle$.
 %
@@
 %(i.e. if the simulation period, $T$, is sufficiently long),
-For a long simulation,
-we can again approximate $\theta_k(t)$
-as white noise:
-$\kappa_k(t) \approx \Gamma_k \, \delta(t)$,
-%
-with
-%
-%\begin{equation}
-%  \kappa_k(t) \approx \Gamma_k \, \delta(t)
-%  ,
-%\label{eq:kappa_delta}
-%\end{equation}
-%
-%\begin{equation}
-$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
-%  \notag
-%\label{eq:Gamma_integral}
-%\end{equation}
-or in discrete time,
-\begin{equation}
-  \Gamma_k
-  = \sum_{t = -\infty}^\infty \kappa_k(t)
-  = \sum_{t = -\infty}^\infty \left\langle \theta_k(0) \, \theta_k(t) \right\rangle
-  .
-  \label{eq:Gamma_sum}
-\end{equation}
-%
-Then, the asymptotic error is simplified as
-%
-\begin{align}
-  \Err_A(T)
-  &=
-  \int_0^T
-  \sum_{k = 2}^n
-  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt.
-\label{eq:error_asym1}
-\end{align}
-%
-In this way,
-the underlying sampling process
-affects the error only through the
-few $\Gamma_k$'s.
-%
-In Appendix \ref{sec:Gamma},
-we evaluate $\Gamma_k$'s in two special cases,
-and show how to measure them
-from an adaptive FDS simulation.
+%
 
@@
 %
-If we use this result
+Using this result
 for the $\langle y_k^2(0) \rangle$'s
@@
 %
-Note that the residual error
+
+\subsubsection{\label{sec:optschedule}
+Optimal schedule}
+
+We will minimize the error under a fixed value of $q(T)$.
+%
+This allows us to focus on the asymptotic part
+for the residual error
 depends on $\alpha(t)$ only through
@@
 %
-Thus, if we vary $\alpha(t)$ under a fixed $q(T)$,
-this residual part can be ignored,
-and the optimization can be done exclusively
-on the asymptotic part, $\Err_A(T)$.
-
-\subsubsection{\label{sec:optschedule}
-Optimal schedule}
-
-As shown above, with a fixed value of $q(T)$,
-the optimization of the total error is equivalent to
-that of the asymptotic part.
-%
-We can rewrite Eq. \eqref{eq:error_asym1} much like an action
+We can rewrite Eq. \eqref{eq:error_asym} much like an action
 of a particle whose position is given by $q(t)$:
@@
 is equal to $t/T$,
-as shown in Fig. \ref{fig:massq},
-%
-In terms of the mass function,
-the optimized asymptotic error is given by
+as shown in Fig. \ref{fig:massq}.
+%
+The optimized asymptotic error
+can then be rewritten as
 %
@@
 are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
-to the equilibrium distribution\cite{bussi2006},
-such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
-are satisfied.
+to the equilibrium distribution\cite{bussi2006}.
+%such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
+%are satisfied.
 %
@@
     \mu_l \,
-    \cos \frac{ (k - 1) \, l \, g \, \pi }
-              {            n             }
+    \cos \frac{ (k - 1) \, l \, 2 \, \pi }
+              {            g \, n        }
   ,
@@
 %
-where the degeneracy, $g$, is $2$
-for the periodic case,
-or $1$ otherwise.
+where the $g = 1$ for the periodic case,
+or $2$ otherwise.
 %
@@
 
-For the updating scheme to be stable,
-the cosine transform of the kernel
-must be nonnegative.
-%
-If this is not so,
-we can modify the kernel
-using the technique described
-in Appendix \ref{sec:stabilize_wband}.
-
 %\subsubsection{\label{sec:invert_wband}
@@
 by a cosine transform,
-Eq. \eqref{eq:wband_eigenvalue},
-the relation can be readily inverted.
-%
-In the periodic case,
-we have
-%the inversion formula is
+we can invert the relation as
 %
@@
   =
-  \frac 1 n
-  \sum_{ k = 1 }^n
+  \frac { 1 } { g \, n }
+  \sum_{ k = 1 }^{ g \, n }
   \lambda_k
   \cos \frac{ (k - 1) \, l \, 2 \, \pi }
-            {            n             }
-  .
-\label{eq:mu_from_lambda_pbc}
-\end{equation}
-
+            {            g \, n        }
+  .
+\label{eq:mu_from_lambda}
+\end{equation}
+%
 In the non-periodic case,
-the formula is
-%
-\begin{align}
-  \mu_l
-  =
-  \frac 1 { 2 \, n }
-  \sum_{ k = 1 }^{ 2 \, n }
-    \lambda_{ k } \,
-    \cos \frac{ (k - 1) \, l \, \pi }
-              {            n        }
-  ,
-\label{eq:mu_from_lambda_refl}
-\end{align}
-%
-where
 we have defined
@@
 %
-\note{Eq. \eqref{eq:lambdan}
-  ensures that the $\mu_n$
-  computed from Eq. \eqref{eq:mu_from_lambda_refl}
-  vanishes.
-
-  %The inversion formula is derived from the cosine transform.
-  %
-  If we define $\mu_n = 0$ and
-  %
-  $
-    \mu_{ 2 n - l } = \mu_l,
-  $
-  for
-  $l = 1, \dots, n - 1$,
-  %
-  then Eq. \eqref{eq:wband_eigenvalue_refl} can be rewritten as
-  %
-  \begin{align}
-    \lambda_{k+1}
-    =
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
-  \notag
-  %\label{eq:lambda_cosine_sum}
-  \end{align}
-  %
-  This formula for $\lambda_{k+1}$
-  can be readily extended to $k = 2 \, n - 1$,
-  and we have
-  %
-  $
-    \lambda_{ 2 \, n + 2 - k } = \lambda_k
-    .
-  $
-  %
-  Thus,
-  %
-  $$
-  \begin{aligned}
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \lambda_{ k + 1 } \,
-      \cos \frac{ k \, p \, \pi }
-                {      n        }
-    &=
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \sum_{ l = 0 }^{ 2 \, n - 1 }
-        \mu_l \,
-        \cos \frac{ k \, l \, \pi }
-                  {      n        }
-        \cos \frac{ k \, p \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \frac{ \mu_l } { 2 }
-      \sum_{ k = 0 }^{ 2 \, n - 1 }
-        \cos \frac{ k \, (p + l) \, \pi }
-                  {      n        }
-                  +
-        \cos \frac{ k \, (p - l) \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \mu_l \, n \left(
-        \delta_{ p + l - 2 \, n, 0 }
-        +
-        \delta_{ p - l, 0 }
-      \right)
-    \\
-    &=
-    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
-    =
-    2 \, n \, \mu_p.
-  \end{aligned}
-  $$
-  This entails Eq. \eqref{eq:mu_from_lambda_refl}.
-  %
-  The value of $\lambda_{n + 1}$
-  can be deduced from
-  imposing $\mu_n = 0$
-  and solving Eq. \eqref{eq:mu_from_lambda_refl}
-  for $l = n$,
-  yielding Eq. \eqref{eq:lambdan}.
-}
+
+For the updating scheme to be stable,
+the cosine transform of the kernel
+must be nonnegative.
+%
+If this is not so,
+we can modify the kernel
+using the technique described
+in Appendix \ref{sec:stabilize_wband}.
 
@@
 \item \label{item:lambda}
-From the updating matrix, $\mathbf w$,
-compute the eigenvalues, $\lambda_k$'s.
+Compute the eigenvalues, $\lambda_k$'s
+from the updating matrix, $\mathbf w$,
+or from Eq. \eqref{eq:wband_eigenvalue}
+for a homogeneous updating scheme.
+%
+Note also that
+the updating scheme should be stabilized
+using the procedure in Appendix \ref{sec:stabilize_wband}
+to ensure all eigenvalues are nonnegative.
 
@@
 \item \label{item:Gamma}
-Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}.
+Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}
+    (cf. Appendix \ref{sec:Gamma_measure}).
 
@@
 Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
+%
+Since $m(Q)$ is a decreasing function of $Q$,
+the equation can be solved
+for a sufficiently small $a_0$
+by the bisection method.
+%
+We start from an interval $[Q_L, Q_R]$
+that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
+(e.g. we can choose $Q_L = 0$,
+  and repeatedly increase $Q_R$ until
+  the latter inequality is satisfied).
+%
+We can then iteratively refine the interval
+by substituting $Q_M = (Q_L + Q_R)/2$
+for either $Q_L$ and $Q_R$,
+depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
+until the width of the interval is less than
+the desired precision.
+%
+Then, $q(T)$ is given by $Q_M$.
 
@@
 Compute the optimal schedule by
-numerically integrating Eq. \eqref{eq:q_opt}.
+numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.
+%
+This can be done on an even grid of $N$ points
+at $q_{[i]} = \frac{i}{N} \, q(T)$
+($i = 0, 1, \dots, N$)
+using the standard trapezoidal rule\cite{press3rd}.
+%
+The curve $q(t)$ is then defined on the grid points,
+$(q_{[i]}, t_{[i]})$,
+%
+and the updating magnitude can be obtained from numerical differentiation:
+$\alpha_i = (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
+with
+$q_{[ -1]} \equiv q_{[0]} = 0$,
+$q_{[N+1]} \equiv q_{[N]} = q(T)$,
+$t_{[ -1]} \equiv t_{[0]} = 0$, and
+$t_{[N+1]} \equiv t_{[N]} = T$.
+%
+The resulting values of $(t_{[i]}, \alpha_{[i]})$
+can be used for linear interpolation
+to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.
 
@@
 \end{enumerate}
-
-Several comments are in order.
-%
-In step \ref{item:lambda},
-one can use Eq. \eqref{eq:wband_eigenvalue}
-for a homogeneous updating scheme.
-For a discrete variable,
-the updating scheme should be stabilize
-using the procedure in Appendix \ref{sec:stabilize_wband}
-to ensure all eigenvalues are nonnegative.
-%
-In step \ref{item:Gamma},
-the details of measuring $\Gamma_k$'s
-can be found in Appendix \ref{sec:Gamma_measure}.
-
-In step \ref{item:qT},
-since $m(Q)$ is a decreasing function of $Q$,
-one can solve Eq. \eqref{eq:opt_qT}
-for a sufficiently small $a_0$
-by the bisection method.
-%
-Starting from the interval $[Q_L, Q_R]$,
-where $m(Q_L) > 2/(T\,a_0) > m(Q_R)$,
-one can iteratively refine the interval
-by substituting $Q_M = (Q_L + Q_R)/2$
-for either $Q_L$ and $Q_R$,
-depending on the sign of $m(Q_M) - 2/(T \, a_0)$.
-
-In step \ref{item:qopt},
-the integrals in Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}
-are done on an even grid of $Q \in [0, q(T)]$
-of $N$ points
-using the standard trapezoidal rule.
-%
-The curve $q(t)$ is then defined on the grid points,
-$(Q_i, t_i)$ ($i = 0, \dots, N$),
-which can be numerically differentiated
-to get the corresponding values of the updating magnitude:
-$\alpha_i = (Q_{i+1} - Q_{i-1}) / (t_{i+1} - t_{i-1})$
-(for the boundary values, we define
-$Q_{-1} = t_{-1} = 0, Q_{N+1} = q(T), t_{N+1} = T$).
-This results in a table of $(t_i, \alpha_i)$
-that can be used for linear interpolation
-to get the updating magnitude at any given time, $t$.
 
@@
 for the asymptotic error $\Err(T)$
-given in Eq. \eqref{eq:error_asym1}.
+given in Eq. \eqref{eq:error_asym}.
 %
@@
 Then by using
-Eq. \eqref{eq:mu_from_lambda_pbc},
+Eq. \eqref{eq:mu_from_lambda},
 we get
@@
 $\lambda_{n+1} = (-1)^{n+K-1}$,
-and from Eq. \eqref{eq:mu_from_lambda_refl},
+and from Eq. \eqref{eq:mu_from_lambda},
 we have
@@
   {
-    n
+    g \, n
   }
   {
-    \sqrt{ 2 \, \pi } \, g \, K
+    \sqrt{ 2 \, \pi } \, 2 \, K
   }
@@
 %
-%where $g = 2$ for the periodic case
-%or $g = 1$ otherwise.
+%where $g = 1$ for the periodic case
+%or $g = 2$ otherwise.
 %
@@
 
-In the above section,
-we have focused on
-the asymptotic error.
-%
-By including the residual error
-and minimizing the total error,
-we get the optimal $t_0$,
-which is, according to Eq. \eqref{eq:half_alpha0},
-%
-%%Assuming the system is initially equilibrated
-%%under a constant updating magnitude,
-%%$a_0$,
-%Assuming Eq. \eqref{eq:error_res1} with
-%$$
-%q(T)
-%=
-%\int_0^T \frac{ dt } { t + t_0 }
-%=
-%\ln \frac{ T + t_0 } { t_0 }.
-%$$
-%%
-%we have, from Eqs. \eqref{eq:error_split}
-%and \eqref{eq:error_asym_sinc},
-%%
-%\begin{equation}
-%  \Err
-%  =
-%  \Err_R + \Err_A
-%  =
-%  \frac { \frac 1 2 a_0 \, t_0^2 + T }
-%        {          ( T + t_0 )^2          }
-%  \sum_{ k = 2 }^K \Gamma_k
-%  .
-%\notag
-%%\label{eq:error_sinc0}
-%\end{equation}
-%%
-%The minimum is reached at
+By including the residual error,
+we determine the optimal $t_0$
+that minimizes the total error.
+%
+According to Eq. \eqref{eq:half_alpha0}, we have
 %
@@
   t_0
+  =
+  \frac{        1      }
+       { \alpha(t = 0) }
   =
@@
 which serves as a generalization of
-the result from the single-bin scheme, Eq. \eqref{eq:Emin_singlebin}.
-%
-This error decreases with increasing simulation length,
+the result from the single-bin scheme,
+Eq. \eqref{eq:Emin_singlebin}
+[one can show that $\sum_{k=2}^n \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma$,
+following the steps similar to those leading to Eq. \eqref{eq:y2_sum}].
+%
+\note{The derivation.
+From Eq. \eqref{eq:eig_orthonormal_cols}, we have
+$\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle
+=\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle$,
+so
+\begin{align*}
+\sum_{i=1}^n \rho_i G_i
+&=\sum_{t=-\infty}^\infty\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle \\
+&=\sum_{t=-\infty}^\infty\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle
+=\sum_{k=2}^n \Gamma_k.
+\end{align*}
+}
+%
+This error is roughly proportional to the inverse simulation length,
 as one would expect from an equilibrium FDS simulation.
@@
 If we denote the error at the end of stage $s$ as $E_s$,
-then from Eq. \eqref{eq:ET_average}, we have
-$q(t) = \alpha_s \, t$,
-where $t$ denotes the number of steps in stage $s$.
-%
-Then
+then $q(t) = \alpha_s \, t$,
+where $t$ denotes the number of steps in stage $s$,
+and from Eq. \eqref{eq:ET_average}, we have
 %
@@
 as the initial updating magnitude, $\alpha(t^{(s)} = 0)$,
-then this result agrees with Eqs. \eqref{eq:half_alpha0} and \eqref{eq:t0_sinc},
-and the optimal $t_0 = \alpha_0^{-1}$.
-%
-Thus, in this limit,
-the schedule given by Eq. \eqref{eq:alphas_t}
+then this result agrees with Eq. \eqref{eq:t0_sinc},
+with $t_0 = \alpha_0^{-1}$,
+and the schedule given by Eq. \eqref{eq:alphas_t}
 is identical to that from Eq. \eqref{eq:alpha_invt1}.
@@
   =
-  \cos \frac{ (k - 1) \, g \, \pi }
-            {            n        }
+  \cos \frac{ (k - 1) \, 2 \, \pi }
+            {       g \, n        }
   ,
@@
 where
-$g = 2$ for a periodic variable, or $1$ otherwise.
+$g = 1$ for a periodic variable, or $2$ otherwise.
 %Note that these $\Lambda_k$'s are
@@
   \cot^2 \left[
-    \frac{ (k - 1) \, g \, \pi }
-         { 2 \, n }
+    \frac{ (k - 1) \, \pi }
+         {       g \, n   }
   \right]
@@
 
-Here, we give some detailed derivations
+Here, we give some details
 on the homogeneous updating schemes.
@@
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-must share the same periodicity,
-and be even functions about the axis, $i = 1/2$:
+share the same periodicity,
+and are even functions about the axis, $i = 1/2$:
 %
@@
 
-\subsection{Two examples}
-
-We give two examples of the homogeneous updating schemes.
+\note{On the inversion.
+Eq. \eqref{eq:lambdan}
+  ensures that the $\mu_n$
+  computed from Eq. \eqref{eq:mu_from_lambda}
+  vanishes.
+
+  %The inversion formula is derived from the cosine transform.
+  %
+  If we define $\mu_n = 0$ and
+  %
+  $
+    \mu_{ 2 n - l } = \mu_l,
+  $
+  for
+  $l = 1, \dots, n - 1$,
+  %
+  then Eq. \eqref{eq:wband_eigenvalue_refl} can be rewritten as
+  %
+  \begin{align}
+    \lambda_{k+1}
+    =
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
+  \notag
+  %\label{eq:lambda_cosine_sum}
+  \end{align}
+  %
+  This formula for $\lambda_{k+1}$
+  can be readily extended to $k = 2 \, n - 1$,
+  and we have
+  %
+  $
+    \lambda_{ 2 \, n + 2 - k } = \lambda_k
+    .
+  $
+  %
+  Thus,
+  %
+  $$
+  \begin{aligned}
+    \sum_{ k = 0 }^{ 2 \, n - 1 }
+      \lambda_{ k + 1 } \,
+      \cos \frac{ k \, p \, \pi }
+                {      n        }
+    &=
+    \sum_{ k = 0 }^{ 2 \, n - 1 }
+      \sum_{ l = 0 }^{ 2 \, n - 1 }
+        \mu_l \,
+        \cos \frac{ k \, l \, \pi }
+                  {      n        }
+        \cos \frac{ k \, p \, \pi }
+                  {      n        }
+    \\
+    &=
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+      \frac{ \mu_l } { 2 }
+      \sum_{ k = 0 }^{ 2 \, n - 1 }
+        \cos \frac{ k \, (p + l) \, \pi }
+                  {      n        }
+                  +
+        \cos \frac{ k \, (p - l) \, \pi }
+                  {      n        }
+    \\
+    &=
+    \sum_{ l = 0 }^{ 2 \, n - 1 }
+      \mu_l \, n \left(
+        \delta_{ p + l - 2 \, n, 0 }
+        +
+        \delta_{ p - l, 0 }
+      \right)
+    \\
+    &=
+    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
+    =
+    2 \, n \, \mu_p.
+  \end{aligned}
+  $$
+  This entails Eq. \eqref{eq:mu_from_lambda}.
+  %
+  The value of $\lambda_{n + 1}$
+  can be deduced from
+  imposing $\mu_n = 0$
+  and solving Eq. \eqref{eq:mu_from_lambda}
+  for $l = n$,
+  yielding Eq. \eqref{eq:lambdan}.
+}
+\subsection{Examples}
 
@@
   4 \, \mu_1 \sin^2
-  \frac{ (k - 1) \, g \, \pi }
-       {       2 \, n        }
+  \frac{ (k - 1) \, \pi }
+       {       g \, n   }
   .
@@
 \note{
-If the variable is periodic ($g = 2$) and $n$ is even,
-the minimum is achieved at $k = n/2$.
-
-If the variable is periodic ($g = 2$) but $n$ is odd,
+If the variable is periodic ($g = 1$) and $n$ is even,
+the minimum is achieved at $k = (n/2) + 1$.
+
+If the variable is periodic ($g = 1$) but $n$ is odd,
 the minimum is achieved at $k = (n+1)/2$.
@@
 \sin^2\left(
-  \frac{n-1}{2n} \pi
+  \frac{n-1}{2 \, n} \pi
 \right)
@@
 \cos^2\left(
-  \frac{\pi}{2n}
+  \frac{\pi}{2 \, n}
 \right)
@@
 
-If the variable is nonperiodic ($g = 1$)
-the minimum is achieved at $k = n$.
+If the variable is nonperiodic ($g = 2$)
+the minimum is achieved at $k = n$ (the maximal index).
 }
@@
 where
-$\Delta x = g \, \pi/n$
+$\Delta x = 2 \, \pi/(g \, n)$
 is the bin size.
 %
-By assuming the largest cutoff, $b \approx n/g$,
+By assuming the largest cutoff, $b \approx g \, n/2$,
 we can approximate Eq. \eqref{eq:wband_eigenvalue}
@@
 The upper limit,
-$b \, \Delta x \approx (n/g) \, (g\,\pi/n) = \pi$.
+$b \, \Delta x \approx (g\,n/2) \, [2\,\pi/(g\,n)] = \pi$.
 }
@@
 using the inversion formula,
-Eq. \eqref{eq:mu_from_lambda_pbc}
-or \eqref{eq:mu_from_lambda_refl},
+Eq. \eqref{eq:mu_from_lambda},
 we get a new updating kernel,
@@
     $\mu_0, \dots, \mu_b$, from
-    Eq. \eqref{eq:mu_from_lambda_pbc}
-    or
-    \eqref{eq:mu_from_lambda_refl}
+    Eq. \eqref{eq:mu_from_lambda},
     with $\mu_l = 0$ for $l > b$.
@@
 Using Eq. \eqref{eq:alpha_invtlambda}
-in Eq. \eqref{eq:error_asym1} yields
+in Eq. \eqref{eq:error_asym} yields
 the asymptotic error

### 2017-04-20 doc/invt.tex
small edits to invt.tex

@@
   \Err_R(T) + \Err_A(T),
-  \label{eq:error_split}
+  \label{eq:error_tot}
 \end{equation}
@@
 %
-where we have approximated $\theta_k(t)$
-as white noise:
-%
-\begin{equation}
-  \kappa_k(t)
-  \equiv \langle \theta_k(0) \, \theta_k(t) \rangle
-  \approx \Gamma_k \, \delta(t)
-  .
-  \notag
-  %\label{eq:kappa_delta}
+where
+%
+\begin{equation}
+  u_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
+  ,
+  \label{eq:uk_def}
+\end{equation}
+%
+and $\Gamma_k$ is computed from
+the autocorrelation function
+of the histogram fluctuation,
+$\kappa_k(t) \equiv \langle \theta_k(0) \, \theta_k(t) \rangle$,
+as:
+%
+\begin{equation}
+  \Gamma_k
+  = \sum_{t = -\infty}^\infty \kappa_k(t)
+  = \sum_{t = -\infty}^\infty \left\langle \theta_k(0) \, \theta_k(t) \right\rangle
+  .
+  \label{eq:Gamma_sum}
 \end{equation}
@@
 %\end{equation}
-
-For the theory to be applicable to
-a general sampling process,
-we can map the histogram fluctuation
-to the equivalent white noise, with
-\begin{equation}
-  \Gamma_k
-  = \sum_{t = -\infty}^\infty \kappa_k(t)
-  %= \sum_{t = -\infty}^\infty \left\langle \theta_k(0) \, \theta_k(t) \right\rangle
-  .
-  \label{eq:Gamma_sum}
-\end{equation}
+%
+Equation \eqref{eq:Gamma_sum}
+essentially maps the histogram fluctuation, $\theta_k(t)$,
+to the equivalent white noise
+so that the theory can be applicable to
+a general sampling process.
 %
@@
 affects the error only through the
-few $\Gamma_k$'s.
+few numbers, $\Gamma_k$'s.
 %
 For example, we expect large values of $\Gamma_k$
-from eigenmodes corresponding to long-range fluctation
-in a local sampling process,
-especially for complex or glassy systems.
+from eigenmodes corresponding to
+persistent long-range motion
+in sampling complex or glassy systems.
 %
@@
 %with
-%$u_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}$,
 %and
@@
 
-We will minimize the error under a fixed value of $q(T)$.
-%
-This allows us to focus on the asymptotic part
-for the residual error
-depends on $\alpha(t)$ only through
-the integral, $q(T) = \int_0^T \alpha(t) \, dt$.
+As in the single-bin case,
+by varying the schedule, $\alpha(t)$, under a fixed value of
+$q(T) = \int_0^T \alpha(t) \, dt$,
+we can focus on the asymptotic error,
+$\Err_A(T)$.
+%
+The value of $q(T)$ will be adjusted later
+to minimize the total error in Sec. \ref{sec:optinitalpha}).
 %
@@
   =
-  \sum_{ k = 2 }^n
-    \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
+  %\sum_{ k = 2 }^n
+  %  \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
+  %\; \dot q^2( t )
+  %\notag
+  %\\
+  %&=
+  M^2\bigl(q(T) - q(t) \bigr)
   \; \dot q^2( t )
@@
 %
+where we have
+used Eq. \eqref{eq:uk_def} and
+defined the (square-root) mass function as
+%
+\begin{equation}
+  M(Q)
+  \equiv
+  \sqrt{
+    \textstyle\sum_{ k = 2 }^n \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
+  }
+  .
+\label{eq:mass_func}
+\end{equation}
+%
 In this mechanical analogy,
 the schedule, $\dot q(t) = \alpha(t)$,
-corresponds to the velocity of the particle.
-%
-So the above Lagrangian contains only the kinetic energy,
-and, therefore, describes a free particle
-carrying a position-dependent mass.
-%
-Since the Lagrangian
-does not explicitly depend on time, $t$,
-the Hamiltonian is conserved.
-%
-But in this case the Hamiltonian
+corresponds to the velocity of the particle,
+%
+and the above Lagrangian, containing only the kinetic energy,
+describes a free particle
+with a position-dependent mass.
+%
+Besides, the Hamiltonian
+coincides with the Lagrangian:
 %
@@
   =
-  \mathcal L,
+  \mathcal L
+  .
   %=
@@
 %
-coincides with the Lagrangian.
-%
-So we may set
-%
-\begin{equation}
-  \mathcal L
-  =
-    \sum_{ k = 2 }^n
-      \Gamma_k \, \lambda_k^2
-      \, u_k^2 \bigl[ q(t) \bigr]
+Since the Lagrangian
+does not explicitly depend on time, $t$,
+the Hamiltonian is conserved,
+and we may set
+%
+\begin{equation}
+  \sqrt{ \mathcal H }
+  =
+  \sqrt{ \mathcal L }
+  =
+  M\bigl( q(T) - q(t) \bigr)
   \;
-  \dot q^2(t)
-  =
-  \left(
+  \dot q(t)
+  =
   \frac{C}{T}
-  \right)^2
   ,
@@
 %
-for some positive constant, $C$.
+for some positive $C$
 %
@@
 %
-By integrating the square root, we get
-%
-\begin{equation}
-  \int_{ 0 }^{ q(t) }
-    \sqrt{
-      \textstyle\sum_{ k = 2 }^n
-        \Gamma_k \, \lambda_k^2
-        \, u_k^2( q' )
-    }
+By integrating this equation of motion, we get
+%
+\begin{equation}
+  \int_{ q(T) - q(t) }^{ q(T) }
+    M(Q)
     \;
-    d q'
+    d Q
   =
@@
 %
-where $C$ is determined by
-the boundary condition at $t = T$ as
+and $C$ can be determined from
+the $t = T$ case as
 %
@@
   \int_{ 0 }^{ q(T) }
-    \sqrt{
-      \textstyle\sum_{ k = 2 }^n
-        \Gamma_k \, \lambda_k^2
-        \, u_k^2( q' )
-    }
+    M( Q )
     \;
-    d q'
+    d Q
   .
@@
 $\alpha(t) = \dot q(t)$.
-%
-Since we only optimized the asymptotic error, $\Err_A(T)$,
-in the above,
-the value of $q(T)$ can be further adjusted
-to minimize the total error, $\Err(T) = \Err_R(T) + \Err_A(T)$
-(cf. Sec. \ref{sec:optinitalpha}).
 
@@
 %
-First, we define a mass function as
-%
-\begin{equation}
-  M(Q)
-  =
-  \sqrt{
-    \textstyle\sum_{ k = 2 }^n \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
-  }
-  .
-\label{eq:mass_func}
-\end{equation}
-%
-and a normalized mass distribution as
+First, we define
+a normalized mass distribution as
 %
@@
     % label for the shaded area
-    \node[] at (1.1, 0.5) {$\dfrac{t}{T}$};
+    \node[] at (1.1, 0.40) {$\dfrac{t}{T}$};
 
@@
 The above characterization allows us to determine
-the optimal initial updating magnitude.
-%
-Let us now consider simulations in which
+the optimal initial updating magnitude,
+or, equivalently, the optimal value of $q(T)$
+that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
+%
+We will consider simulations in which
 the system is initially equilibrated at the same
 constant updating magnitude, $a_0$.
-%
-For these simulations,
-we can adjust $q(T)$
-to minimize the total error.
 %
@@
 we find that
-the $q(T)$ that minimizes
-the total error satisfies
+the optimal $q(T)$ satisfies
 \begin{equation}
@@
 
-Many commonly-used updating schemes
+Many commonly-used updating schemes,
+such as the Gaussian updating scheme
+used in metadynamics,
 can be characterized by a rigid window
@@
 %
-We call these schemes homogeneous,
-or translationally invariant.
-%
-An example is the Gaussian updating scheme
-used in metadynamics.
+For these translationally invariant or homogeneous
+schemes, the eigenvalues can be readily found
+from the shape of the window function.
+%
 
@@
 %
-where
-$\mu_{-b}, \dots, \mu_b$ ($b \le n/2$)
-are a set of numbers satisfying
+where the numbers,
+$\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
+characterizing the rigid window
+will be referred to as an \emph{updating kernel}\cite{bussi2006}.
+%
+To satisfy Eq. \eqref{eq:w_sumj},
+we impose the normalization condition
 %
@@
 %
-and $\mu_l = 0$ for $|l| > b$.
-%
-We will further impose the symmetry
-%
-\begin{equation}
-  \mu_i = \mu_{-i}
-  .
-\label{eq:mu_symm}
-\end{equation}
+with $\mu_l = 0$ for $|l| > b$.
 %
@@
 %
-These numbers constitute an \emph{updating kernel}\cite{bussi2006}.
+For simplicity, we also impose the symmetry
+%
+\begin{equation}
+  \mu_i = \mu_{-i}
+  .
+\label{eq:mu_symm}
+\end{equation}
 %
@@
 Eq. \eqref{eq:wband_eigenvalue_pbc},
-with the period, $n$, doubled:
-%
-\begin{align}
-  \lambda_k
-  &=
-  \mu_0
-  +
-  2
-  \sum_{l = 1}^b
-    \mu_l
-    \cos \frac{(k - 1)  \, l \, 2 \, \pi}{ 2 \, n }
-  .
-\label{eq:wband_eigenvalue_refl}
-\end{align}
-%
-We may unify Eqs. \eqref{eq:wband_eigenvalue_refl}
-and \eqref{eq:wband_eigenvalue_pbc} as
+with a doubled period:
 %
@@
 %
-where the $g = 1$ for the periodic case,
-or $2$ otherwise.
-%
-In both cases,
-the eigenvalues, $\lambda_k$'s,
-are the cosine transform of
-the updating kernel.
+where $g = 1$ or $2$
+in the periodic or non-periodic case, respectively.
 
@@
 \item
-  Compute the error $\Err(T)$ from
-Eqs. \eqref{eq:error_split},
+  Compute the total error $\Err(T)$ from
+Eqs. \eqref{eq:error_tot},
   \eqref{eq:error_res1},
@@
 %
-From Eqs. \eqref{eq:error_split},
+From Eqs. \eqref{eq:error_tot},
 \eqref{eq:error_res1},
@@
 and the eigenvalues are given by
-  Eq. \eqref{eq:wband_eigenvalue_refl}.
+  Eq. \eqref{eq:wband_eigenvalue}
+  with $g = 2$.
 
@@
   $$
-  with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue_refl}.
+  with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}
+  with $g = 2$.
 }%
@@
   %
-  then Eq. \eqref{eq:wband_eigenvalue_refl} can be rewritten as
+  then Eq. \eqref{eq:wband_eigenvalue} ($g = 2$)
+  can be rewritten as
   %
@@
 For $n \gg 1$,
-we define
-$x = l \, \Delta x$,
+we define an equivalent angular variable, $\varphi$,
+over $[0, 2 \, \pi/g]$,
+with bin size
+$\Delta \varphi = 2 \, \pi/(g \, n)$.
+The variable is related to discrete index as
+$\varphi = l \, \Delta \varphi$,
 and
-$\mu_l = \mu(x) \, \Delta x$,
-where
-$\Delta x = 2 \, \pi/(g \, n)$
-is the bin size.
+$\mu_l = \mu(\varphi) \, \Delta \varphi$.
 %
@@
   2 \int_0^\pi
-    \mu(x) \, \cos \bigl[ (k-1) \, x \bigr] \, dx,
+    \mu(\varphi) \, \cos \bigl[ (k-1) \, \varphi \bigr] \, d\varphi,
 \notag
@@
 $
-  1 = 2 \int_0^\pi \mu(x) \, dx.
+  1 = 2 \int_0^\pi \mu(\varphi) \, d\varphi.
 $
@@
   &\approx
-  \int_0^b \mu_l \, \cos[(k-1) \, x] \, dl
+  \int_0^b \mu_l \, \cos[(k-1) \, \varphi] \, dl
   \\
   &=
-  \int_0^b \mu(x) \, \cos[(k-1) \, x] \, \Delta x \, dl
+  \int_0^b \mu(\varphi) \, \cos[(k-1) \, \varphi] \, \Delta \varphi \, dl
   \\
   &=
-  \int_0^{b \, \Delta x} \mu(x) \, \cos[(k-1) \, x] \, dx
+  \int_0^{b \, \Delta \varphi} \mu(\varphi) \, \cos[(k-1) \, \varphi] \, d\varphi
 \end{align*}
 The upper limit,
-$b \, \Delta x \approx (g\,n/2) \, [2\,\pi/(g\,n)] = \pi$.
+$b \, \Delta \varphi \approx (g\,n/2) \, [2\,\pi/(g\,n)] = \pi$.
 }
@@
 For the Gaussian updating kernel,
-if the width, $\sigma$, satisfies $\Delta x \ll \sigma \ll 1$,
+if the width of the angular variable, $\sigma_\varphi$,
+satisfies $\Delta \varphi \ll \sigma_\varphi \ll \pi$,
 we can extend
@@
 \begin{equation}
-  \mu(x)
+  \mu(\varphi)
   =
   \frac{            1            }
-       { \sqrt{ 2 \pi \sigma^2 } }
+       { \sqrt{ 2 \pi \sigma_\varphi^2 } }
   %
   \exp\left(
-        -\frac{       x^2     }
-              { 2 \, \sigma^2 }
+        -\frac{   \varphi^2   }
+              { 2 \, \sigma_\varphi^2 }
       \right)
@@
   \exp\left[
-        -\frac{ (k - 1)^2 \, \sigma^2 }
+        -\frac{ (k - 1)^2 \, \sigma_\varphi^2 }
               {           2           }
@@
   %      \left(
-  %        \frac{  \sigma }
-  %             { \Delta x }
+  %        \frac{  \sigma_\varphi }
+  %             { \Delta \varphi }
   %      \right)^2
@@
 %However,
-%as $\sigma/\Delta x \gg 1$,
+%as $\sigma_\varphi/\Delta \varphi \gg 1$,
 %the smallest eigenvalue
@@
 %\exp\left(
-%      -\frac{ n^2 \, \sigma^2 }
+%      -\frac{ n^2 \, \sigma_\varphi^2 }
 %            {        2        }
@@
 %      \left(
-%        \frac{  \sigma }
+%        \frac{  \sigma_\varphi }
 %             { \Delta x }
@@
 %$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
-%greater than $\sigma$.
+%greater than $\sigma_\varphi$.
 %%
@@
 %\exp\left(
-%      -\frac{ k_{\max}^2 \, \sigma^2 }
+%      -\frac{ k_{\max}^2 \, \sigma_\varphi^2 }
 %            {           2            }
@@
 %\exp\left(
-%      -\frac{  \pi^2 \, \sigma^2    }
+%      -\frac{  \pi^2 \, \sigma_\varphi^2    }
 %            {    2 \, l_{\min}^2  }
@@
 (cf. Sec. \ref{sec:optWL}).
+%
+For example, the smaller eigenvalues of the Gaussian updating scheme
+are well separated because of the rapid exponential decay.
 
@@
   \setlength{\tabcolsep}{4pt} % column separation
-  %
-  \caption{\label{tab:symbols}Table of symbols.}
+  \caption{\label{tab:symbols}
+    Table of symbols.}
   \begin{tabular}{l | p{12cm} }
@@
     $\phi_{ki}$     &   Eigenvectors of the updating matrix. \\
-    $g$             &   Degeneracy factor, $2$ for a periodic variable, or for $1$ a non-periodic one. \\
+    $g$             &   $1$ for a periodic variable, or for $2$ a non-periodic one. \\
+    $G_i$           &   Integral of the autocorrelation function of histogram fluctuation at bin $i$. \\
     $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
@@
     $\sigma_{ij}(t)$   &   Autocorrelation function of $\zeta_i(t)$. \\
-    $\sigma$        &   Width of the Gaussian updating scheme. \\
+    $\sigma_\varphi$   &   Width of the Gaussian updating scheme in radians. \\
     $t, \tau$       &   Time. \\

### 2017-04-21 doc/invt.tex
add histogram flatness == error

@@
 of a multiple-bin updating scheme
-that encompasses metadynamics
-in Secs. \ref{sec:multiple-bin}
-to \ref{sec:band-matrix},
+that encompasses the Gaussian scheme used in metadynamics
+in Sec. \ref{sec:multiple-bin},
 and present a general procedure of computing the optimal schedule
@@
 can often be improved
-with the normalized histogram, $H_i$, from the simulation.
-%
-Using $H_i$ as an estimate of $\pi_i$ in
-Eq. \eqref{eq:pi_p_V}
-allows us to correct $V_i$ as
+with the normalized histogram, $H_i$,
+accumulated from the simulation as
+%
+%Using $H_i$ as an estimate of $\pi_i$ in
+%Eq. \eqref{eq:pi_p_V}
+%allows us to correct $V_i$ as
 %
@@
 However, this formula often requires a long
-simulation period to accumulate a precise histogram,
+accumulation period for a precise histogram,
 and thus disallows
@@
   .
-\label{eq:error_sum}
+\label{eq:error_def}
 \end{equation}
@@
 %
-where the modified histogram fluctuation
+where the shifted histogram fluctuation
 %
@@
   \xi_i(t)
-  =
+  \equiv
   \frac{ \zeta_i(t) }
@@
 i.e. we may write
-$$
-  \langle \xi_i(t) \, \xi_i(t') \rangle = G_i \, \delta(t - t'),
-$$
+\begin{equation}
+  \langle \xi_i(t) \, \xi_i(t') \rangle
+  = G_i \, \delta(t - t')
+  ,
+  \label{eq:G_def}
+\end{equation}
 for some $G_i$.
@@
 at the end of the simulation, $t = T$,
-as defined in Eq. \eqref{eq:error_sum},
+as defined in Eq. \eqref{eq:error_def},
 now depends implicitly on the function $\alpha(t)$
@@
 At the end of period $T$,
-the total error defined in Eq. \eqref{eq:error_sum}
+the total error defined in Eq. \eqref{eq:error_def}
 is given by
@@
 where
-$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
+\begin{equation}
+  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i.
+  \label{eq:Gamma_def}
+\end{equation}
 
@@
 we find the final total error to be
-\begin{equation}
-  E(T)
-  =
-  \frac{ E(0) \, t_0^2 + \Gamma \, T}
-       { (t_0 + T)^2 }
-  ,
-  \notag
-  %\label{eq:error_singlebin}
-\end{equation}
+%
+$E(T) = [ E(0) \, t_0^2 + \Gamma \, T ]/(t_0 + T)^2$,
 %
@@
 %
-and at this point
+and at this point, the error is given by
 %
@@
 \end{equation}
+%
+This error is roughly equal to the overall histogram flatness
+and thus it can be conveniently estimated\cite{zhou2005, zhou2008}
+(see also Appendix \ref{sec:flatness}).
 
@@
   -\alpha(t) \, \lambda_k \, [y_k(t) - \theta_k(t)],
-\label{eq:yt_diffeq}
+  \notag
+  %\label{eq:yt_diffeq}
 \end{equation}
@@
   ,
-  \label{eq:y_def}
+  %\label{eq:y_def}
 \end{align}
@@
 
-The error function defined in Eq. \eqref{eq:error_sum}
+The error function defined in Eq. \eqref{eq:error_def}
 can be rewritten in terms of the eigenmodes of $\mathbf w$,
@@
 
-\subsection{\label{sec:mbin_opta}
-Optimal schedule for multiple-bin updating schemes}
+%\subsection{\label{sec:mbin_opta}
+%Optimal schedule for multiple-bin updating schemes}
 
@@
 %
-for some positive $C$
+for some positive $C$.
 %
@@
 
-\subsection{\label{sec:band-matrix}
+\subsubsection{\label{sec:band-matrix}
 Homogeneous updating schemes}
@@
 
-\subsection{\label{sec:procedure}
+\subsubsection{\label{sec:procedure}
 Procedure for computing the optimal schedule and error
@@
 %
-which serves as a generalization of
-the result from the single-bin scheme,
-Eq. \eqref{eq:Emin_singlebin}
-[one can show that $\sum_{k=2}^n \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma$,
-following the steps similar to those leading to Eq. \eqref{eq:y2_sum}].
+In particular, for $K = n$, we have
+\begin{equation}
+  \sum_{k=2}^n \Gamma_k
+  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
+  ,
+  \notag
+  %\label{eq:Gamma_sum}
+\end{equation}
+which can be shown by following the steps
+similar to those leading to Eq. \eqref{eq:y2_sum},
+and the above error recovers the single-bin result,
+Eq. \eqref{eq:Emin_singlebin}.
+%
+%
 %
@@
 along with the distribution, $\rho_i$, in Eqs.
-\eqref{eq:vbar_def} and \eqref{eq:error_sum}.
+\eqref{eq:vbar_def} and \eqref{eq:error_def}.
 %
@@
 by explicitly filtering them out from the final bias potential,
-and truncating the error sum in Eq. \eqref{eq:error_sum}
+and truncating the error sum in Eq. \eqref{eq:error_def}
 to the first few eigenmodes.
@@
   +
-  \sum_{l=1}^L \frac{1}{N \, L} \frac{ h_i^{(N)}(l) }{ p_i }
+  \sum_{l=1}^L \frac{1}{N L} \frac{ h_i^{(N)}(l) }{ p_i }
   ,
@@
 \begin{equation}
-  \alpha(t) = \frac{1}{N \, L}
+  \alpha(t) = \frac{1}{N L}
   .
@@
 by the product of the number of sub-simulations
-and the number of steps in each sub-simulation, or $N \, L$,
+and the number of steps in each sub-simulation, or $N L$,
 %
@@
 is identical to that from Eq. \eqref{eq:alpha_invt1}.
+
+\section{\label{sec:flatness}
+Histogram flatness}
+
+Here we wish to show that for an optimal WL simulation
+(also for an FDS simulation with a perfect bias potential),
+the overall histogram flatness
+is roughly equal to the final error.
+%
+First, in the long time limit, the fluctuation of
+the instantaneous histogram is dominated by the random part,
+$h_i(t) \approx \zeta_i(t)$,
+and $\xi_i(t)$ defined in Eq. \eqref{eq:xi_def}
+will dominate the shifted instantaneous histogram
+%
+\begin{equation}
+\xi_i(t) \approx
+\frac{ h_i(t) } { p_i }
+- \sum_{r = 1}^n \rho_r \frac{ h_r(t) } { p_r }
+.
+\label{eq:xi_h}
+\end{equation}
+%
+If we define the overall or time-averaged histogram as
+\begin{equation}
+  H_i \equiv \frac{1}{T} \sum_{t=1}^T h_i(t)
+  ,
+  \label{eq:H_def}
+\end{equation}
+then the time average of the right-hand side of Eq. \eqref{eq:xi_h}
+represents the fractional deviation of the shifted histogram.
+%
+We way thus define the histogram flatness as
+%
+\begin{equation}
+  F
+  \equiv
+  \left\langle
+  \sum_{i = 1}^n \rho_i
+  \left(
+  \frac{ H_i } { p_i }
+  - \sum_{r = 1}^n \rho_r \frac{ H_r } { p_r }
+  \right)^2
+  \right\rangle
+  .
+  \label{eq:F_def}
+\end{equation}
+%
+By using Eqs. \eqref{eq:H_def}, \eqref{eq:xi_h},
+\eqref{eq:G_def}, and \eqref{eq:Gamma_def}, we have
+%
+\begin{align*}
+  F
+  %&\approx
+  %\sum_{i = 1}^n \rho_i
+  %\left\langle
+  %\left(
+  %\frac{1}{T}
+  %\sum_{t = 1}^T \xi_i(t)
+  %\right)^2
+  %\right\rangle
+  %\\
+  &\approx
+  \frac{1}{T^2}
+  \sum_{i = 1}^n \rho_i
+  \sum_{t, t' = 1}^T
+  \left\langle
+  \xi_i(t) \, \xi_i(t')
+  \right\rangle
+  \approx
+  \frac{1}{T}
+  \sum_{i = 1}^n \rho_i \, G_i
+  =
+  \frac{ \Gamma } { T }
+  .
+\end{align*}
+%
+Compare the last expression to Eq. \eqref{eq:Emin_singlebin},
+we conclude
+\begin{equation}
+  F \approx E(T)
+  .
+  \notag
+  %\label{eq:flatness}
+\end{equation}
+%and the histogram flatness
+%serves as a good estimate of the final error.
+Note that in some previous studies\cite{zhou2005, zhou2008},
+the histogram flatness, Eq. \eqref{eq:F_def},
+was previously used to define the error,
+and the above demonstration shows the definition
+is compatible with ours, Eq. \eqref{eq:error_def}, in the long-time limit.
+%
+\note{
+An interesting application\cite{zhou2008},
+of the above result is
+to set the updating magnitude by $F/\Gamma$
+for some empirical value [$\Gamma \approx 10$ there, see Eq. (8)]
+as a substitute of the inverse-time schedule.
+}
 
@@
 \begin{equation}
-  \phi_i = \phi_{i \pm n},
+  \phi_j = \phi_{j \pm n},
 \label{eq:phi_pbc}
@@
 %
-such that $i \pm n$ lies in between $1$ and $n$.
+such that $j \pm n$ lies in between $1$ and $n$.
 %
@@
 %
+Alternatively, to avoid complex numbers, we can use
+%
+\begin{equation}
+  \phi^{(k)}_j
+  =
+  \phi_{kj}
+  =
+  \frac{ \sqrt{ 2 - \delta_{k,0} } } { n }
+  %\overset{ \cos } { \sin }
+  {\cos \atop \sin}
+  \left[
+    \frac{ (k - 1) \, j \, 2 \, \pi }
+         {            n             }
+  \right]
+  ,
+  \notag
+\end{equation}
+%
+where the function takes the cosine form for $k \le n/2 + 1$,
+or the sine form otherwise.
+%
 Then one can readily verify that the eigenvalues are given by
@@
 we define the out-of-boundary values
-of $\phi_i$ as
-%
-\begin{equation}
-  \phi_i
+of $\phi_j$ as
+%
+\begin{equation}
+  \phi_j
   =
   \begin{dcases}
-    \phi_{ 1 - i }           & \mathrm{for \;} i \le 0, \\
-    \phi_{ 2 \, n + 1 - i }  & \mathrm{for \;} i > n,
+    \phi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
+    \phi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
   \end{dcases}
@@
 }%
-We note that Eq. \eqref{eq:phi_refl}
+%
+Since Eq. \eqref{eq:phi_refl}
 defines a periodic variable of period $2 \, n$
-with a reflective symmetry around $i = 1/2$.
-%
-Accordingly, the eigenvectors,
+with a reflective symmetry around $j = 1/2$,
+the eigenvectors,
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-share the same periodicity,
-and are even functions about the axis, $i = 1/2$:
-%
-\begin{equation}
-  \phi^{(k)}_i
-  =
-  \phi_{k i}
+should share the same periodicity,
+and be even functions about the same axis:
+%
+\begin{equation}
+  \phi^{(k)}_j
+  =
+  \phi_{k j}
   =
@@
        {             n              }
-  \cos \frac{ ( k - 1) \, \left( i - \frac 1 2 \right) \, 2 \, \pi}
+  \cos \left[
+       \frac{ ( k - 1) \, \left( j - \frac 1 2 \right) \, 2 \, \pi}
             {                    2 \, n                           }
+       \right]
   ,
@@
   $$
-  \Phi^{(k)}_i
-  =
-  \cos \frac{ \left( i - \frac 1 2 \right) (k - 1) \, \pi}{n},
+  \Phi^{(k)}_j
+  =
+  \cos \frac{ (k - 1) \, \left( j - \frac 1 2 \right) \, \pi }{n},
   $$
@@
   ,
-\label{eq:error_asym_invt}
+  \notag
+  %\label{eq:error_asym_invt}
 \end{align}
@@
 From Eq. \eqref{eq:error_res1},
-we get the residual error
+we get % the residual error
 %
@@
 $\Gamma_k$'s affect the error
-only via the sum, $\sum_{k = 2}^n \Gamma_k$,
-as a multiplicative constant.
-
-In the asymptotic limit,
-we can focus on the asymptotic error
-and drop the second term in the square brackets
-of Eq. \eqref{eq:error_asym_invt}
-by assuming $\lambda < 2$.
-%
-Then,
+only via the sum, $\Gamma = \sum_{k = 2}^n \Gamma_k$,
+as a multiple.
+%
+In the long time limit,
+we have for $\lambda < 2$,
 $$
@@
   \frac { 1 } { T + t_0 }
-  \frac {             1            }
+  \frac {         \Gamma           }
         { \lambda \, (2 - \lambda) }
-  \sum_{ k = 2 }^n \Gamma_k
-  ,
-$$
-is minimal at $\lambda = 1$.
+  ,
+$$
+which is minimal at $\lambda = 1$.
 %
@@
     $\xi_i(t)$      &   The noise, Eq. \eqref{eq:xi_def}. \\
-    $y_k(t)$        &   Mode of the shifted bias potential, Eq. \eqref{eq:y_def}. \\
-    $Y_k$           &   Mode of the original bias potential, Eq. \eqref{eq:Y_def}. \\
+    $y_k(t)$        &   Mode of the shifted bias potential. \\
+    $Y_k$           &   Mode of the original bias potential. \\
     $z$             &   Reaction coordinate. \\

### 2017-04-21 prog/metad.h

@@ -54,7 +54,8 @@ __inline static int metad_getindex(metad_t *metad, int x)
 
 
 
-/* decide if a transition from xold to xnew is to be accepted */
+/* decide if a transition from xold (index iold)
+ * to xnew (index iold) is to be accepted */
 static int metad_acc(metad_t *metad, int iold, int xnew,
     int *inew)
 {
@@ -69,6 +70,45 @@ static int metad_acc(metad_t *metad, int iold, int xnew,
   if ( dv <= 0 ) return 1;
   r = rand01();
   return r < exp(-dv);
+}
+
+
+
+/* compute the histogram flatness */
+__inline static double metad_hflatness(metad_t *metad)
+{
+  int i;
+  double hmin, hmax, hi;
+
+  hmin = hmax = metad->h[0];
+  for ( i = 1; i < metad->n; i++ ) {
+    hi = metad->h[i];
+    if ( hi < hmin ) {
+      hmin = hi;
+    } else if ( hi > hmax ) {
+      hmax = hi;
+    }
+  }
+
+  return 2 * (hmax - hmin) / (hmax + hmin + DBL_EPSILON);
+}
+
+
+
+/* check if histogram is flat enough to switch to
+ * a smaller updating magnitude */
+static int metad_wlcheck(metad_t *metad)
+{
+  double hflatness;
+  int i;
+
+  /* compute the histogram flatness */
+  hflatness = metad_hflatness(metad);
+  /* return if the histogram not flatness enough */
+  if ( hflatness > 0.2 ) return 0;
+
+  metad->a *= 0.5;
+  for ( i = 0; i < metad->n; i++ ) metad->h[i] = 0;
 }
 
 
@@ -101,8 +141,8 @@ static int metad_save(metad_t *metad, const char *fn)
     return -1;
   }
   metad_trimv(metad);
-  fprintf(fp, "# %d %d %d %d\n",
-      metad->n, metad->xmin, metad->xmax, metad->xdel);
+  fprintf(fp, "# %d %d %d %d %g\n",
+      metad->n, metad->xmin, metad->xmax, metad->xdel, metad->a);
   for ( i = 0; i < metad->n; i++ ) {
     fprintf(fp, "%d %g %g\n",
         metad->xmin + i * metad->xdel, metad->v[i], metad->h[i]);

### 2017-04-26 doc/invt.tex
update invt.c

@@
   and the optimal schedule of these schemes
-  is the inverse number of sampling steps with a proper shift,
+  is the inverse number of updating steps with a proper shift,
   as given by the $1/t$ prescription.
@@
 the two methods\cite{micheletti2004, kim2006, *kim2007, junghans2014},
-we will loosely use the terms WL and metadynamics
+we will colloquially use the terms WL and metadynamics
 to represent FDS simulations using the
-$\delta$- and Gaussian window functions, respectively.
+single-bin and Gaussian window functions, respectively.
 %
@@
 %
-For the WL algorithm, the optimal updating schedule
+For the single-bin (WL) case, the optimal updating schedule
 is a function of the
@@
 or
-the inverse number of simulation steps.
+the inverse number of updating steps.
 \note{
@@
 with a general window function,
-including the $\delta$- and Gaussian one
+including the single-bin and Gaussian one
 used in the WL algorithm and metadynamics.
@@
 %
-As the mass is defined from the updating window function,
-the resulting optimal schedule
-is different in the WL and metadynamics cases.
-%
-While the optimal schedule in the WL case
+The resulting optimal schedule
+depends on the updating window function
+through the mass function.
+%
+While the optimal schedule in the single-bin case
 is given by the known inverse-time formula,
-the optimal schedule for metadynamics is more complex
+that for Gaussian case is more complex
 and sensitive to the simulation length
@@
 half of the previous equilibrium value,
-which amounts to a shift of the origin of time in the WL case.
+which amounts to a shift of the origin of time
+in the single-bin case.
 %
 Further, we can generalize
-the WL updating scheme
+the single-bin updating scheme
 to a class of bandpass updating schemes
@@
 %
-In a more general \emph{multiple-bin scheme}, however,
+In a more general \emph{multiple-bin scheme},
+including the Gaussian updating scheme
+used in metadynamics,
 several neighboring bins are updated:
@@
 %
-The Gaussian updating scheme
-used in metadynamics
-is a multiple-bin scheme.
-
-%One can show that with a constant magnitude,
-%$\alpha(t) = \alpha > 0$,
-%the distribution collected from
-%an adaptive FDS simulation
-%is identical to $p_i$ in the long time or asymptotic limit.
-%
-%However, as the above updates
-%change $V_i(t)$ continuously,
-%the equilibrium distribution,
-%Eq. \eqref{eq:pi_p_V}, no longer holds
-%at all times,
+Note that if the updating occurs
+only every few sampling steps,
+then $t$ means the number of updating steps.
+
+\subsubsection{Updating magnitude and the inverse-time formula}
+
 While adaptive FDS simulations
-can often sample the desired distribution, $p_i$,
-even with a constant magnitude,
-$\alpha(t) = \alpha$,
+can often sample the desired distribution,
 frequent updates to the bias potential
@@
   crespo2010, *atchade2011, *fort2015}.
+%One can show that with a constant magnitude,
+%$\alpha(t) = \alpha > 0$,
+%the distribution collected from
+%an adaptive FDS simulation
+%is identical to $p_i$ in the long time or asymptotic limit.
+%
+%However, as the above updates
+%change $V_i(t)$ continuously,
+%the equilibrium distribution,
+%Eq. \eqref{eq:pi_p_V}, no longer holds
+%at all times,
 %%
@@
 
-\subsubsection{Updating magnitude and the inverse-time formula}
-
 In the original WL papers\cite{
 wang2001, *wang2001pre},
-the updating magnitude, $\alpha$,
+the updating magnitude, $\alpha(t)$,
 (or $\ln f$ therein)
@@
 %
-In each stage, $\alpha$
+In each stage, $\alpha(t)$
 is kept constant,
@@
 we switch to a new stage
-with $\alpha$ reduced by a factor of $1/2$\cite{
+with $\alpha(t)$ reduced by a factor of $1/2$\cite{
 wang2001, *wang2001pre}.
@@
 While this scheme works well for early stages,
-%
-it tends to reduce $\alpha$
+it tends to reduce the magnitude
 too quickly in late stages, making the asymptotic error
@@
 referred to as the ``time'' below,
-is the number of steps
+is the number of updating steps
 from the beginning of the simulation\cite{
@@
 %
-An interesting explanation of this result\cite{
+An intuitive explanation of this result\cite{
   marsili2006, barducci2008}
@@
 the equivalent value of $G^{(0)}_i$ can be approximately obtained
-from the integral,
-$\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$,
-or in the discrete time case,
-$\sum_{-\infty}^\infty \sigma_{ij}(t)$.
+%from the integral,
+%$\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$,
+%or in the discrete time case,
+$\sum_{t = -\infty}^\infty \sigma_{ij}(t)$.
 %
@@
 %\eqref{eq:dvbardt},
-we get a set of decoupled equations
+we get % a set of decoupled equations
 %
@@
   ,
-\label{eq:dxdt_WL}
+\label{eq:dxdt_singlebin}
 \end{equation}
@@
 %
-is also governed by white noise as is $\zeta_i(t)$,
-i.e. we may write
+is also governed by white noise as is $\zeta_i(t)$:
+%i.e. we may write
 \begin{equation}
@@
 We recall that
-the formal solution of Eq. \eqref{eq:dxdt_WL} is
+the formal solution of Eq. \eqref{eq:dxdt_singlebin} is
 %
@@
 
-Since the above optimization is performed
-under a fixed $q(T)$,
-it can be improved
-by choosing a proper value of $q(T)$.
+We now determine the optimal value of $q(T)$,
+which is fixed in the above process.
 %
@@
 %
-By using the optimal schedule in Eq. \eqref{eq:ET_average},
+By using Eq. \eqref{eq:alpha_invt} in Eq. \eqref{eq:ET_average},
 we find the final total error to be
@@
 %
-This error is roughly equal to the overall histogram flatness
-and thus it can be conveniently estimated\cite{zhou2005, zhou2008}
+This error is roughly equal to
+the overall histogram flatness\cite{zhou2005, zhou2008}
+and thus it can be conveniently estimated
 (see also Appendix \ref{sec:flatness}).
@@
 %
-The key simplification comes from the
+The key step comes from the
 projection of the bias potential to the $n$ eigenmodes
@@
 should be a constant of $i$
-if $h_j(t)$ were the same as $p_j$.
-%
-This allows the bias potential to grow uniformly
+if $h_j(t)$ were the same as $p_j$,
+i.e.
+$\sum_{j=1}^n w_{ij}$ should be a constant
+to allow the bias potential to grow uniformly
 in the asymptotic regime.
 %
-Thus, by a proper scaling of $\alpha(t)$,
+By a proper overall scaling of $\mathbf w$,
 we may write this condition as
@@
 %
-This artificial condition
-restrains the form of the updating matrix, $\mathbf w$,
-and it requires, for the commonly-used uniform distribution, $\pmb\rho$,
-a symmetric updating matrix, $\pmb w$.
-%
-More generally, only
+This requires the updating matrix, $\mathbf w$
+to be symmetric
+for a uniform $\pmb\rho$,
+or more generally,
 the scaled matrix, $\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
-needs to be symmetric,
-and the symmetry allows the diagonalization
+to be symmetric.
+%
+This symmetry allows the diagonalization
 of $\mathbf w$ with a set of
@@
 no eigenmode grows indefinitely,
-which means that no eigenvalue, $\lambda_k$,
+which means that, as we will show, no eigenvalue, $\lambda_k$,
 can be negative.
@@
 
-\subsubsection{Error functional}
-
-The error function defined in Eq. \eqref{eq:error_def}
+\subsubsection{Error}
+
+The error defined in Eq. \eqref{eq:error_def}
 can be rewritten in terms of the eigenmodes of $\mathbf w$,
@@
 
-\subsubsection{\label{sec:eqlerr}
-Modeling initial error from a simulation
-under a constant updating magnitude
-}
-
-Before addressing the optimal schedule
-for a multiple-bin scheme,
-we will first compute the equilibrium error
-under a constant schedule,
-$\alpha(t) = a_0$.
-%
-After a long period, $T'$,
-the residual error becomes negligible, and
-the $k$th component of the asymptotic error
-is given by
-%
+%\subsubsection{\label{sec:eqlerr}
+%Initial error
+%%from a simulation under a constant updating magnitude
+%}
+
+%Before addressing the optimal schedule
+%for a multiple-bin scheme,
+%we will first compute the equilibrium error
+%under a constant schedule,
+%$\alpha(t) = a_0$.
+%%
+%%In terms of the components, we have
+%%
 %\begin{align}
-%  \Err
+%  \left\langle
+%    y_k^2
+%  \right\rangle
 %  &=
-%  \Err_A
-%  =
 %  \int_0^{T'}
-%    \sum_{k = 2}^n
 %    \Gamma_k \, (\lambda_k \, a_0)^2 \,
@@
 %  \\
-%  &
-%  =
-%  %\stackrel{ T' \to \infty }
-%  %{ =\joinrel=\joinrel=\joinrel= }
-%  %
-%  \sum_{ k = 2 }^n
+%  &=
 %  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
@@
 %  .
-%\notag
-%%\label{eq:error_eql}
+%\label{eq:y2_eql}
 %\end{align}
-%
-%In terms of the components, we have
-%
-\begin{align}
-  \left\langle
-    y_k^2
-  \right\rangle
-  &=
+%%
+
+To model the initial error for Eq. \eqref{eq:error_res},
+we will assume that
+the adaptive FDS sampling
+was performed under a constant updating magnitude
+for sufficiently long.
+%
+Then, one can show that
+\begin{equation}
+  \left\langle y_k^2(0) \right\rangle = \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  ,
+  \label{eq:y2_eql}
+\end{equation}
+%
+\note{
+Consider a long period, $T'$, under a fixed schedule,
+$\alpha(t) = a_0$,
+the residual error becomes negligible, and
+the $k$th component of the asymptotic error
+is given by
+%
+\begin{align*}
+  \left\langle y_k^2(T') \right\rangle
+  =
   \int_0^{T'}
@@
     \, dt
+  \stackrel{ T' \to \infty }
+  { =\joinrel=\joinrel=\joinrel= }
+  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  .
 \notag
-  \\
-  &=
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
-  ,
-  \qquad
-  ( T' \to \infty )
-  .
-\label{eq:y2_eql}
-\end{align}
-%
-
-We will use this result to model the initial error
-before starting the target schedule, $\alpha(t)$.
-%
-That is, we assume that there is an equilibration phase
-for $t < 0$ during which the adaptive FDS sampling was performed
-under a constant updating magnitude,
-$\alpha(t) = a_0$, and at the end of the equilibration, $t = 0$,
-the system has entered the equilibrium state of sampling,
-in which the error of mode $k$ obeys Eq. \eqref{eq:y2_eql}.
-%
-Using this result
-for the $\langle y_k^2(0) \rangle$'s
-in Eq. \eqref{eq:error_res},
-the residual error at the end of period $T$
-is given by
+%\label{eq:error_eql}
+\end{align*}
+}
+%
+and
 %
@@
 %
+This expression excludes systematic bias
+and thus serves as a lower bound for a realistic simulation.
 
@@
     \draw[domain=0:2.5, variable=\x, dashed, gray, ultra thick]
-      plot({\x}, {1.5*exp(-0.5*\x*\x)});
+      plot({\x}, {sqrt(2*exp(-2*\x)+exp(-\x))});
 
@@
       (0.7, 0) --
-      plot ({\x}, {1.5*exp(-0.5*\x*\x)})
+      plot ({\x}, {sqrt(2*exp(-2*\x)+exp(-\x))})
       -- (2.2, 0) -- cycle;
@@
     \draw[domain=0:2.2, variable=\x, ultra thick]
-      plot({\x}, {1.5*exp(-0.5*\x*\x)});
+      plot({\x}, {sqrt(2*exp(-2*\x)+exp(-\x))});
 
@@
                {$\dfrac{1}{T \, \alpha}$}
-               (0.6, 1.174);
+               (0.6, 0.99488);
     % label for the ordinate
@@
 
-    \draw[] (0.5, 1.174) -- (0.7, 1.1740);
+    \draw[] (0.5, 0.99488) -- (0.7, 0.99488);
     \draw[->] (0, 0) -- node[below, inner sep=5mm] {$Q$} (2.8, 0);
-    \draw[->] (0, 0) -- node[above, rotate=90] {$m(Q)$} (0, 1.8);
+    \draw[->] (0, 0) -- node[above, rotate=90] {$m(Q)$} (0, 1.9);
 
@@
 i.e. the optimal initial updating magnitude
-is always half of the equilibrium value.
-%
-This factor, $1/2$,
+is always half of the equilibrium value.\footnote{This factor, $1/2$,
 happens to coincide with the
@@
 wang2001, *wang2001pre}.
+}
 
@@
 To satisfy Eq. \eqref{eq:w_sumj},
-we impose the normalization condition
+we impose the normalization
 %
@@
 %
-If $b = n/2$, then $\mu_{-b}$ is removed
+If $b = n/2$, $\mu_{-b}$ is removed
 from the sum in Eq. \eqref{eq:mu_normalization}
@@
 As shown in Appendix \ref{sec:wband_eig},
-the multiplication of $\mathbf w$
-is equivalent to a convolution with the kernel, and
+%the multiplication of $\mathbf w$
+%is equivalent to a convolution with the kernel, and
 the eigenvalues of $\mathbf w$ are
@@
   \frac{ (k - 1) \, l \, 2 \, \pi }
-       {            n             }
-  ,
-  \label{eq:wband_eigenvalue_pbc}
-\end{equation}
-%
-with a two-fold degeneracy,
-$\lambda_{n - k + 2} = \lambda_k$.
+       {            g \, n        }
+  ,
+  \label{eq:wband_eigenvalue}
+\end{equation}
+%
+where $g = 1$ here for the periodic variable.
 
@@
 Thus, the eigenvalues can be borrowed from
-Eq. \eqref{eq:wband_eigenvalue_pbc},
-with a doubled period:
-%
-\begin{equation}
-  \lambda_k
-  =
-  \mu_0
-  +
-  2
-  \sum_{ l = 1 }^b
-    \mu_l \,
-    \cos \frac{ (k - 1) \, l \, 2 \, \pi }
-              {            g \, n        }
-  ,
-  \label{eq:wband_eigenvalue}
-\end{equation}
-%
-where $g = 1$ or $2$
-in the periodic or non-periodic case, respectively.
+Eq. \eqref{eq:wband_eigenvalue},
+with a doubled period,
+i.e. $g$ is changed to $2$ in the non-periodic case.
 
@@
 
-Since the eigenvalues are related to the updating kernel
-by a cosine transform,
-we can invert the relation as
-%
-\begin{equation}
-  \mu_l
-  =
-  \frac { 1 } { g \, n }
-  \sum_{ k = 1 }^{ g \, n }
-  \lambda_k
-  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
-            {            g \, n        }
-  .
-\label{eq:mu_from_lambda}
-\end{equation}
-%
-In the non-periodic case,
-we have defined
-$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
-for $k = n + 2, \dots, 2 \, n$,
-as well as
-%
-\begin{align}
-  \lambda_{ n + 1 }
-  =
-  (-1)^{ n - 1 }
-  \lambda_1
-  +
-  2 \, \sum_{ k = 2 }^{ n }
-      (-1)^{n - k} \, \lambda_k
-  ,
-\label{eq:lambdan}
-\end{align}
-to satisfy the constraint, $\mu_n = 0$.
-%
-
-For the updating scheme to be stable,
-the cosine transform of the kernel
-must be nonnegative.
-%
-If this is not so,
-we can modify the kernel
-using the technique described
-in Appendix \ref{sec:stabilize_wband}.
-
 \subsubsection{\label{sec:procedure}
-Procedure for computing the optimal schedule and error
-}
-
-Here we give a procedure for computing
+Procedure of computing the optimal schedule and error
+}
+
+Here we give a procedure of computing
 the optimal schedule and the error
-of a given adaptive FDS simulation.
+for a given adaptive FDS simulation.
 %
@@
 
-\item \label{item:lambda}
+\item
 Compute the eigenvalues, $\lambda_k$'s
@@
 %
-Note also that
-the updating scheme should be stabilized
-using the procedure in Appendix \ref{sec:stabilize_wband}
-to ensure all eigenvalues are nonnegative.
+Note that
+the updating scheme should be stable
+with no negative eigenvalue
+(cf. the stabilization procedure in Appendix \ref{sec:stabilize_wband}).
 
@@
 
-\item \label{item:Gamma}
+\item
 Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}
@@
 
-\item \label{item:qT}
-Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
+\item
+Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}
+using e.g. the bisection method.
 %
 Since $m(Q)$ is a decreasing function of $Q$,
-the equation can be solved
-for a sufficiently small $a_0$
-by the bisection method.
-%
-We start from an interval $[Q_L, Q_R]$
+we can start from an interval $[Q_L, Q_R]$
 that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
 (e.g. we can choose $Q_L = 0$,
-  and repeatedly increase $Q_R$ until
+  and repeatedly double $Q_R$ until
   the latter inequality is satisfied).
@@
 until the width of the interval is less than
-the desired precision.
-%
-Then, $q(T)$ is given by $Q_M$.
-
-\item \label{item:qopt}
+the desired precision,
+%
+and set $q(T)$ to $Q_M$.
+
+\item
 Compute the optimal schedule by
@@
 The resulting values of $(t_{[i]}, \alpha_{[i]})$
-can be used for linear interpolation
+can be linear interpolated
 to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.
@@
 Using the Cauchy-Schwarz inequality, we have,
-for any set of nonnegative numbers, $c_k$'s,
+for any set of nonnegative numbers, $c_k$'s\footnote{This
+inequality follows from the nonnegativity of
+the quadratic polynomial,
+$$
+\int_0^T
+  dt \sum_{k = 2}^n \Gamma_k \,
+    \left( \dot u_k \, \chi - c_k \right)^2
+  \equiv
+  A_2 \, \chi^2 + A_1 \, \chi + A_0
+  ,
+$$
+which must carry a non-positive discriminant:
+$A_1^2 - 4 \, A_0 \, A_2 \le 0$.},
+%
 %
@@
 %
-%\note{
-This inequality follows from the non-negativity of
-the quadratic polynomial,
-$$
-\int_0^T
-  dt \sum_{k = 2}^n \Gamma_k \,
-    \left( \dot u_k \, \chi - c_k \right)^2
-  \equiv
-  A_2 \, \chi^2 + A_1 \, \chi + A_0
-  ,
-$$
-which must carry a non-positive discriminant:
-$A_1^2 - 4 \, A_0 \, A_2 \le 0$.
-%}%
-%
-Note that the last expression of Eq. \eqref{eq:CSineq}
+Note that the last expression %of Eq. \eqref{eq:CSineq}
 depends on the curve, $q(t)$ ($0 < t < T$),
@@
 Such a solution is possible only if
-all nontrivial eigenvalues are equal
-%
-\begin{equation}
-  \lambda_2 = \dots = \lambda_n.
-  \notag
-\end{equation}
+all nontrivial eigenvalues are equal,
+%
+%\begin{equation}
+$\lambda_2 = \dots = \lambda_n$.
+%  \notag
+%\end{equation}
 %
@@
 %
-Note, an updating matrix
+Since an updating matrix
 with identical eigenvalues
-is a multiple of the identity matrix.
-%
-Thus,
+is a multiple of the identity matrix,
+%
 the corresponding updating scheme,
@@
 %
-Mathematically, these updating schemes will have
-zero eigenvalues for the ignored modes,
-and they are useful if we can assume that
-the ignored modes are absent from the PMF.
+Mathematically, these updating schemes
+allow us to assign zero eigenvalues
+to modes that are likely absent from the PMF.
 %
@@
 \begin{aligned}
-&
-\lambda_1 = \cdots = \lambda_{K+1}
-= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
-\\
-&
+%&
+%\lambda_1 = \cdots = \lambda_{K+1}
+%= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
+%\\
+%&
 \lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
@@
 $$
-for $K \ge 1$.
+for $K \ge 1$,
+with the rest of the eigenvalues being $1$.
 %
@@
 %
+\note{From Eq. \eqref{eq:y2_eql}, we have
+$$E(0) = \frac{1}{2} a_0 \sum_{k=2}^n \Gamma_k = \frac{1}{2} \, a_0 \, \Gamma.$$
+}
 From Eqs. \eqref{eq:error_tot},
@@
   \notag
-  %\label{eq:Gamma_sum}
+  %\label{eq:Gamma_G_sum}
 \end{equation}
@@
 %differed only by a multiplicative constant,
-in agreement with the discussion in
-Appendix \ref{sec:invt_singlebin}.
+(cf. Appendix \ref{sec:invt_singlebin}).
 %
@@
 If we denote the error at the end of stage $s$ as $E_s$,
-then $q(t) = \alpha_s \, t$,
-where $t$ denotes the number of steps in stage $s$,
+then $q\left( \tau^{(s)} \right) = \alpha_s \, \tau^{(s)}$,
+where $\tau^{(s)}$ denotes the number of steps in stage $s$,
 and from Eq. \eqref{eq:ET_average}, we have
@@
   \alpha_s^2 \,
-  e^{ -2 \, \alpha_s \, (T_s - t) }
-  \, dt
+  e^{ -2 \, \alpha_s \, \left(T_s - \tau^{(s)}\right) }
+  \, d \tau^{(s)}
   \notag
@@
 To compute the optimal schedule,
-we map the overall simulation time, $t^{(s)}$,
+we map the overall simulation time, $t$,
 to the stage number, $s$, as
@@
 \begin{align*}
-  t^{(s)}
+  t
   = \sum_{s' = 1}^s T_{s'}
@@
   \frac{ 1 }
-  { \gamma \, t^{(s)} + \alpha_0^{-1} }
+  { \gamma \, t + \alpha_0^{-1} }
   ,
@@
 If we identify the parameter, $\alpha_0$,
-as the initial updating magnitude, $\alpha(t^{(s)} = 0)$,
+as the initial updating magnitude, $\alpha(t = 0)$,
 then this result agrees with Eq. \eqref{eq:t0_sinc},
@@
 %serves as a good estimate of the final error.
-Note that in some previous studies\cite{zhou2005, zhou2008},
-the histogram flatness, Eq. \eqref{eq:F_def},
-was previously used to define the error,
+Note that the expected histogram flatness
+was previously used to define the error\cite{zhou2005, zhou2008},
 and the above demonstration shows the definition
@@
 By the translational invariance,
-homogeneous updating schemes
+these schemes
 satisfy detailed balance,
@@
 Then one can readily verify that the eigenvalues are given by
-  Eq. \eqref{eq:wband_eigenvalue_pbc}.
-%
+  Eq. \eqref{eq:wband_eigenvalue},
+  and there is a two-fold degeneracy,
+  $\lambda_{n - k + 2} = \lambda_k$.
 \note{Consider the unnormalized version
@@
   The term in the square brackets is the eigenvalue given by
-  Eq. \eqref{eq:wband_eigenvalue_pbc}.
+  Eq. \eqref{eq:wband_eigenvalue}.
 }
@@
 }
+
 \subsection{Examples}
@@
 we define an equivalent angular variable, $\varphi$,
-over $[0, 2 \, \pi/g]$,
+over $[0, 2 \, \pi/g]$
 with bin size
 $\Delta \varphi = 2 \, \pi/(g \, n)$.
-The variable is related to discrete index as
+The variable is related to the bin index, $l$, as
 $\varphi = l \, \Delta \varphi$,
@@
 %
-By assuming the largest cutoff, $b \approx g \, n/2$,
+By assuming the largest bin cutoff, $b \approx g \, n/2$,
 we can approximate Eq. \eqref{eq:wband_eigenvalue}
@@
 
+\subsection{\label{sec:invert_wband}
+Inversion}
+
+Since the eigenvalues are related to the updating kernel
+by a cosine transform,
+we can invert the relation as
+%
+\begin{equation}
+  \mu_l
+  =
+  \frac { 1 } { g \, n }
+  \sum_{ k = 1 }^{ g \, n }
+  \lambda_k
+  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
+            {            g \, n        }
+  .
+\label{eq:mu_from_lambda}
+\end{equation}
+%
+In the non-periodic case,
+we have defined
+$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
+for $k = n + 2, \dots, 2 \, n$,
+as well as
+%
+\begin{align}
+  \lambda_{ n + 1 }
+  =
+  (-1)^{ n - 1 }
+  \lambda_1
+  +
+  2 \, \sum_{ k = 2 }^{ n }
+      (-1)^{n - k} \, \lambda_k
+  ,
+\label{eq:lambdan}
+\end{align}
+to satisfy the constraint, $\mu_n = 0$.
+%
+
 \subsection{\label{sec:stabilize_wband}
@@
 Using Eq. \eqref{eq:alpha_invtlambda}
-in Eq. \eqref{eq:error_asym} yields
-the asymptotic error
+in Eqs. \eqref{eq:error_asym}
+and \eqref{eq:error_res1}
+yields
 %
@@
   %\label{eq:error_asym_invt}
-\end{align}
-%
-where $\nu_k \equiv \lambda_k / \lambda$.
-%
-From Eq. \eqref{eq:error_res1},
-we get % the residual error
-%
-\begin{equation}
+  \\
   \Err_R(T)
-  =
+  &=
   \frac { a_0 } { 2 }
@@
   \right)^{ 2 \, \nu_k }
-  .
+  ,
   \notag
   %\label{eq:error_res_invt}
-\end{equation}
-%
-The total error is the sum of the two.
+\end{align}
+%
+where $\nu_k \equiv \lambda_k / \lambda$.
+%
 %with
@@
 %introduced in Sec. \ref{sec:nnscheme},
-we can find an analytical expression
-of the optimal value of $\lambda$
-in the large-$n$ and asymptotic limit.
-%
-In this case,
-if $2 \, \nu_k > 1$ for every $k$,
-the asymptotic error is implicitly given by the integral
+we will consider the large-$n$ limit.
+%
+If $2 \, \nu_k > 1$ for every $k$, we have
 %$\sum_i \to \frac{2 \, n}{\pi} \int_0^{\pi/2} dp$, and
@@
 \end{align}
-
-For perfect sampling, if $\mathbf p = \pmb \rho$,
-we have
-$\Gamma(p) = 1$ from Eq. \eqref{eq:Gamma_perfect},
+%
+For perfect sampling with $\mathbf p = \pmb \rho$,
+we have Eq. \eqref{eq:Gamma_perfect},
 and
@@
   \right)
-.
+,
 \end{aligned}
@@
 %
-Minimizing the function yields
+which is minimal at
 %
@@
 %
-which shows that the original inverse-time formula,
-Eq. \eqref{eq:alpha_invt1},
-is not optimal in this case.
-%
-The asymptotic error is given by
+and at this point,
 %
@@
   \right)
-  ,
-\notag
-%\label{eq:error_nn_perfect}
-\end{equation}
-%
-%From Eq. \eqref{eq:error_nn}, it is clear that
-which
-is minimal at $\mu_1 = 0$, the single-bin case.
-%increases with the magnitude of $\mu_1$
-%no matter its sign,
-%and the minimal value, obtained at $\mu_1 = 0$,
-%corresponds to the single-bin scheme.
+  .
+  \label{eq:error_nn_perfect}
+\end{equation}
+%
+This shows that the original inverse-time formula,
+Eq. \eqref{eq:alpha_invt1},
+is not optimal unless $\mu_1 = 0$,
+the single-bin case,
+and the latter minimizes the asymptotic error,
+Eq. \eqref{eq:error_nn_perfect}.
 %

### 2017-04-27 doc/invt.tex
update is2dos.ma

@@
   the optimal schedule depends on the updating scheme
-  (such as the single-bin scheme commonly used in WL
-  or the Gaussian one popular in metadynamics),
+  (such as the single-bin and Gaussian schemes commonly used in WL and metadynamics),
   and the initial updating magnitude is ideally half of

### 2017-04-27 prog/metad.h

@@ -18,10 +18,11 @@ typedef struct {
   double *win;
   double *lambda;
   double *costab; /* cosine transform coefficients */
-  double *u;
   double hflatness;
   double *tmat; /* n x n transition matrix */
   corr_t *corr;
+  double *vtmp;
+  double *hmod;
 } metad_t;
 
 
@@ -83,7 +84,8 @@ static metad_t *metad_open(int xmin, int xmax, int xdel,
       gaussig, okmax, win, winn);
   metad->costab = mkcostab(metad->n, metad->pbc);
   xnew(metad->tmat, metad->n * metad->n);
-  xnew(metad->u, metad->n);
+  xnew(metad->vtmp, metad->n);
+  xnew(metad->hmod, metad->n);
   return metad;
 }
 
@@ -93,9 +95,11 @@ static void metad_close(metad_t *metad)
 {
   free(metad->v);
   free(metad->h);
+  free(metad->win);
   free(metad->costab);
   free(metad->tmat);
-  free(metad->u);
+  free(metad->vtmp);
+  free(metad->hmod);
   free(metad);
 }
 
@@ -130,9 +134,8 @@ static int metad_acc(metad_t *metad, int iold, int xnew,
 }
 
 
-
-/* compute the histogram flatness */
-__inline static double metad_hflatness(metad_t *metad)
+/* compute the histogram flatness (Wang-Landau version) */
+__inline static double metad_hflatness_wl(metad_t *metad)
 {
   int i;
   double hmin, hmax, hi;
@@ -152,19 +155,42 @@ __inline static double metad_hflatness(metad_t *metad)
 
 
 
+/* compute the histogram flatness */
+__inline static double metad_hflatness(metad_t *metad)
+{
+  int i, k, n = metad->n;
+  double x, fl = 0, tot = 0, sqr = 0;
+  const double lamcut = 0.1;
+
+  for ( i = 0; i < n; i++ ) tot += metad->h[i];
+  getcosmodes(metad->h, n, metad->vtmp, metad->costab);
+  /* truncate modes with lambda */
+  for ( k = 1; k < n; k++ ) {
+    if ( metad->lambda[k] < lamcut ) break;
+    x = n * metad->vtmp[k] / tot;
+    fl += x * x;
+  }
+  /* compute the flitered histogram */
+  for ( ; k < n; k++ ) metad->vtmp[k] = 0;
+  fromcosmodes(metad->hmod, n, metad->vtmp, metad->costab);
[…]

### 2017-05-02 doc/invt.tex
small changes

@@
   is the inverse number of updating steps with a proper shift,
-  as given by the $1/t$ prescription.
+  as given by the inverse-time prescription.
   %
@@
 \begin{equation}
-  \alpha(t) = \frac{1}{t},
+  \alpha(t) = \frac{1}{t + 1},
   \label{eq:alpha_invt}
@@
 %
-where $t$,
+where $t + 1$,
 referred to as the ``time'' below,
@@
 so that the addition of a new data point only
-affects the average by its due weight, $1/t$,
-which is the inverse of the current sample size, $t$
+affects the average by its due weight,
+which is the inverse of the current sample size, $t + 1$
 (cf. Appendix \ref{sec:equilerr}).
@@
 and Eq. \eqref{eq:alpha_invt}
-is the special case of $t_0 = 0$.
+is the special case of $t_0 = 1$.
 %

### 2017-05-02 prog/aus/is2aus.c

@@ -1,5 +1,5 @@
 #define IS2_LB 5
-#define L (1 << IS2_LB)
+#define L IS2_L
 #include "is2.h"
 #include "util.h"
 
@@ -17,7 +17,7 @@ static void savehist(double *h, int n,
     tot += h[i];
   for ( i = 0; i < n; i++ ) {
     if ( h[i] <= 0 ) continue;
-    fprintf(fp, "%d %g %g\n", emin+i*de, h[i]/tot/de, h[i]);
+    fprintf(fp, "%g %g %g\n", emin+i*de, h[i]/tot/de, h[i]);
   }
   fclose(fp);
 }
@@ -111,8 +111,8 @@ static void is2_aus(is2_t *is, double Eave, double Esig,
     y2 = (y1 * y1 - 1);
     c2 += y2 / (Esig * Esig) / t;
     if ( t % 10000 == 0 )
-    printf("c1 %g, c2 %g, E %d, Eave %g, y %g, %g, acc %g%%\n",
-        c1, c2, is->E, Eave, y1, y2, 100.*nacc/t);
+      printf("c1 %g, c2 %g, E %d, Eave %g, y %g, %g, acc %g%%\n",
+          c1, c2, is->E, Eave, y1, y2, 100.*nacc/t);
     his[(is->E + 2 * is->n)/4] += 1;
   }
   savehist(his, is->n + 1, -2*IS2_N, 4, "is2aus.his");

### 2017-05-04 doc/invt.tex
implementing WL schedule for pt2aus

@@
 the precision of the final bias potential,
-and hence that of the PMF\cite{
+and hence that of the PMF\cite{laio2005, bussi2006, poulain2006,
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-liang2007,
+liang2007, min2007,
 morozov2007, zhou2008,
@@
 %
-For the single-bin (WL) case, the optimal updating schedule
+For the single-bin case, the optimal updating schedule
 is a function of the

### 2017-05-04 prog/aus/gaus.h

@@ -6,6 +6,7 @@
  * */
 
 
+#include "mmwl.h"
 
 typedef struct {
   int n;
@@ -16,8 +17,10 @@ typedef struct {
   double *lnz; /* partition function */
   double alpha; /* updating magnitude */
   double hflatness;
+  double t, t0;
   int invt; /* using 1/t schedule */
-  double *cnt, *tcnt;
+  mmwl_t *mmwl;
+  double *cnt;
   double *acc;
   int xmin, xmax, dx;
   int xn;
@@ -41,8 +44,8 @@ __inline static gaus_t *gaus_open(double xcmin, double xcmax,
   xnew(gaus->beta1, n);
   xnew(gaus->beta2, n);
   xnew(gaus->lnz, n);
+  xnew(gaus->mmwl, n);
   xnew(gaus->cnt, n);
-  xnew(gaus->tcnt, n);
   xnew(gaus->acc, n);
   delx = (xcmax - xcmin) / (n - 1);
   for ( i = 0; i < n; i++ ) {
@@ -51,8 +54,8 @@ __inline static gaus_t *gaus_open(double xcmin, double xcmax,
     gaus->beta1[i] = beta;
     gaus->beta2[i] = 0;
     gaus->lnz[i] = 0;
+    mmwl_init(gaus->mmwl + i, 1e-3);
     gaus->cnt[i] = 0;
-    gaus->tcnt[i] = 0;
     gaus->acc[i] = 0;
   }
   gaus->xmin = xmin;
@@ -68,6 +71,8 @@ __inline static gaus_t *gaus_open(double xcmin, double xcmax,
     gaus->htot[i] = 0;
   }
   gaus->alpha = 1;
+  gaus->t = 0;
+  gaus->t0 = 1;
   gaus->invt = 0;
   fprintf(stderr, "%d states Ec %g ... %g\n",
       n, gaus->ave[0], gaus->ave[n-1]);
@@ -84,8 +89,8 @@ __inline static void gaus_close(gaus_t *gaus)
   free(gaus->beta1);
   free(gaus->beta2);
   free(gaus->lnz);
+  free(gaus->mmwl);
   free(gaus->cnt);
-  free(gaus->tcnt);
   free(gaus->acc);
   free(gaus->hist);
   free(gaus->htot);
@@ -146,9 +151,9 @@ __inline static int gaus_savehist(gaus_t *gaus, const char *fn)
[…]

### 2017-05-04 prog/aus/pt2gaus.c

@@ -1,5 +1,4 @@
 #define POTTS2_LB 5
-#define L POTTS2_L
 #include "potts2.h"
 #include "util.h"
 #include "gaus.h"
@@ -97,10 +96,10 @@ static void potts2_gaus(potts2_t *pt,
   const double tp = 1.4;
   gaus_t *gaus;
 
-  mtscramble(clock());
+  //mtscramble(clock());
 
-  ecmin = -1.8 * pt->n;
-  ecmax = -0.8 * pt->n;
+  ecmin = -1.7 * pt->n;
+  ecmax = -0.9 * pt->n;
   //ecmin = -1.3 * pt->n;
   //ecmax = -1.25 * pt->n;
   //ecmin = -0.5 * pt->n;
@@ -125,8 +124,8 @@ static void potts2_gaus(potts2_t *pt,
       gaus_wlcheck(gaus, 0.1, 0.5, (double) t);
     }
     if ( t % 100000 == 0 ) {
-      printf("t %ld, flatness %g, alpha %g, invt %d\n",
-          t, gaus->hflatness, gaus->alpha, gaus->invt);
+      printf("t %ld, flatness %g, alpha %g, id %d, invt %d\n",
+          t, gaus->hflatness, gaus->alpha, id, gaus->invt);
       gaus_savehist(gaus, "pt2gaus.his");
     }
   }
@@ -147,7 +146,7 @@ int main(int argc, char **argv)
   if ( nsteps <= 0 )
     nsteps = (method == 0) ? 100000000L : 5000000L;
 
-  pt = potts2_open(L, q);
+  pt = potts2_open(POTTS2_L, q);
   potts2_gaus(pt, method, nsteps);
   potts2_close(pt);
   return 0;

### 2017-05-10 doc/invt.tex
small edits to invt.tex

@@
 \end{equation}
+%
+We may write the updating matrix in terms of $\phi_{ki}$'s as
+%
+\begin{equation}
+  w_{ij}
+  =
+  \frac{1}{\rho_i} \sum_{k=1}^n \lambda_k \, \phi_{ki} \, \phi_{kj}
+  .
+  \notag
+\end{equation}
 
@@
 that the PMF is sufficiently smooth
-such that the eigenmodes can be limited to $k \le K$,
+such that the eigenmodes can be limited to $k \le K + 1$,
 %
@@
 $$
-\lambda_1 = \cdots = \lambda_K = 1,
+\lambda_1 = \cdots = \lambda_{K+1} = 1,
 \mathrm{\; and \;}
-\lambda_{K+1} = \cdots = \lambda_n = 0.
-$$
-%
-Then,
-we can change the upper limit of the sum in
-Eq. \eqref{eq:error_asym_singlebin} to $K$:
-%
-%Then,
-%Eq. \eqref{eq:error_singlebin} is changed to
+\lambda_{K+2} = \cdots = \lambda_n = 0.
+$$
+%
+Then, the updating matrix is given by
+%
+$$w_{ij} = \sum_{k=1}^{K+1} \phi_{ki} \, \phi_{kj},$$
+%
+and Eq. \eqref{eq:error_asym_singlebin} is modified as
 %
@@
         { (T + t_0)^2 }
-  \sum_{ k = 2 }^K
+  \sum_{ k = 2 }^{K+1}
     \Gamma_k.
@@
 remains to be the inverse-time schedule
-given by Eq. \eqref{eq:alpha_invt1}.
-%
-Below, we construct these schemes for
-the homogeneous case.
-%
-The definition of the cutoff, $K$, will be modified
-for mathematical convenience.
-
-For a periodic variable, we demand
-%
-$$
-\begin{aligned}
-%&
-%\lambda_1 = \cdots = \lambda_{K+1}
-%= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
-%\\
-%&
-\lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
-\end{aligned}
-$$
-for $K \ge 1$,
-with the rest of the eigenvalues being $1$.
-%
-Then by using
-Eq. \eqref{eq:mu_from_lambda},
-we get
-\begin{equation}
-  \mu_l
-  =
-  \frac{
-    \sin
-    \dfrac{ (2 K + 1) \, l \, \pi }
-         {              n        }
-  }
-  {
-    n \, \sin \dfrac{ l \, \pi } { n }
-  }
-  .
-\notag
-%\label{eq:mu_sinc_pbc}
-\end{equation}
-\note{Derivation.
-$$
-\begin{aligned}
-\mu_l
-&=
-\frac 1 n \sum_{k = 0}^{n-1} \lambda_{k+1} \cos \frac{ k \, l \, 2 \, \pi } { n }
-\\
-&=
-\frac{1}{n}
-\left(
-  1 +
-  \sum_{k=1}^K
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-  +
-  \sum_{k=n-K}^{n-1}
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-\right)
-\\
-&=
-\frac 1 n
-\sum_{k=-K}^K
-\cos \frac { k \, l \, 2 \, \pi } { n }
-=
-  \frac{
-    \sin
-    \dfrac{ (2 K + 1) \, l \, \pi }
-         {              n        }
-  }
-  {
-    n \, \sin \dfrac{ l \, \pi } { n }
-  }
-.
-\end{aligned}
-$$
-}
-
-For a non-periodic variable, we demand
-$$
-\lambda_1 = \cdots = \lambda_{K+1} = 1,
-\qquad
-\lambda_{K+2} = \cdots = \lambda_n = 0.
-$$
-Then Eq. \eqref{eq:lambdan} gives
-$\lambda_{n+1} = (-1)^{n+K-1}$,
-and from Eq. \eqref{eq:mu_from_lambda},
-we have
-\begin{equation}
-  \mu_l
-  =
-  \frac{1}{2 \, n}
-  \left[
-    (-1)^{n+K+l-1}
-    +
-    \frac{
-      \sin
-      \dfrac{ (2 K + 1) \, l \, \pi }
-           {         2 \, n        }
-    }
-    {
-      \sin \dfrac{ l \, \pi } { 2 \, n }
-    }
-  \right]
-  .
-\label{eq:mu_sinc_refl}
-\end{equation}
-\note{Derivation.
-$$
-\begin{aligned}
-  \lambda_{n+1}
-  &=
-  (-1)^{n-1}
-  \left[
-    \lambda_1
-    - 2 \, \lambda_2
-    + 2 \, \lambda_3 - \cdots
-    + (-1)^{K} 2 \, \lambda_{K+1})
-  \right]
-  \\
-  &=
-  \begin{dcases}
-    (-1)^{n-1} & K \mathrm{\; even,} \\
-    (-1)^n     & K \mathrm{\; odd.}
-  \end{dcases}
-\end{aligned}
-$$
-So
-$$
-\begin{aligned}
-  \mu_l
-  &=
-  \frac{1}{2\,n}
-  \left[
-    1 +
-    2 \sum_{k=2}^{K+1}
-    \cos \frac { (k - 1) \, l \pi } { n }
-    +
-    (-1)^{n+K-1} (-1)^l
-  \right]
-  \\
-  &=
-  \frac{1}{2\,n}
-  \left[
-    (-1)^{n+K+l-1}
-    +
-    \sum_{k=-K}^{K}
-    \cos \frac { k \, l \pi } { n }
-  \right]
-  .
-\end{aligned}
-$$
-}%
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
-  }
-  \caption{
-    \label{fig:sinc}
-    Updating kernels of the bandpass updating schemes
-    compared to a Gaussian with a similar width.
-    %
-    The lines are a guide to the eyes.
-    Inset. Updating matrix in the non-periodic case.
-    %
-    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h} for panel (a),
-      and the functions \texttt{savewin()}
-      and \texttt{savewinmat()}
-      in \texttt{prog/invt.h} for panel (b),
-      The data were saved under the directory \texttt{data/sinc}.
-      See the \texttt{make} object \texttt{sincsingle}
-      in the \texttt{Makefile} there.
-    }%
-  }
-\end{center}
-\end{figure}
-
-In Fig. \ref{fig:sinc},
-we give examples of the updating kernels
-of the bandpass updating schemes,
-compared to a Gaussian
-with a similar width, $\sigma$,
-%
-\begin{equation}
-  \sigma
-  \approx
-  \frac
-  {
-    g \, n
-  }
-  {
-    \sqrt{ 2 \, \pi } \, 2 \, K
-  }
-  .
-\label{eq:sigma_equiv}
-\end{equation}
-%
-%where $g = 1$ for the periodic case
-%or $g = 2$ otherwise.
-%
-The kernel in the non-periodic case is a sawtooth
-because of the term, $(-1)^{n+K+l-1}$,
-in Eq. \eqref{eq:mu_sinc_refl}.
-%
-However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
-the ruggedness disappears
-in the updating matrix by Eq. \eqref{eq:w_band_refl},
-as the above oscillatory term for $l = i - j$
-is canceled by the term for either $l = i + j - 1$
-or $l = i + j - 2 \, n - 1$
-with the opposite sign.
-
-\subsubsection{\label{sec:finlencorr}
-Finite-length correction}
-
-By including the residual error,
-we determine the optimal $t_0$
-that minimizes the total error.
-%
-According to Eq. \eqref{eq:half_alpha0}, we have
+given by Eq. \eqref{eq:alpha_invt1},
+%
+with
 %
@@
        { a_0 }
-  .
+  ,
 \label{eq:t0_sinc}
@@
 %
+where we have used Eq. \eqref{eq:half_alpha0}
+in the second step.
+%
+\note{
 For the single-bin updating scheme,
@@
 %
-\note{From Eq. \eqref{eq:y2_eql}, we have
+From Eq. \eqref{eq:y2_eql}, we have
 $$E(0) = \frac{1}{2} a_0 \sum_{k=2}^n \Gamma_k = \frac{1}{2} \, a_0 \, \Gamma.$$
 }
+%
 From Eqs. \eqref{eq:error_tot},
@@
        { T + t_0 }
-  \sum_{ k = 2 }^K
+  \sum_{ k = 2 }^{ K + 1 }
     \Gamma_k
@@
 %
-In particular, for $K = n$, we have
+This result recovers Eq. \eqref{eq:Emin_singlebin}
+for the single-bin updating scheme:
+if we set $K + 1 = n$,
+%
 \begin{equation}
@@
 which can be shown by following the steps
-similar to those leading to Eq. \eqref{eq:y2_sum},
-and the above error recovers the single-bin result,
-Eq. \eqref{eq:Emin_singlebin}.
+similar to those leading to Eq. \eqref{eq:y2_sum}.
 %
@@
 }
+
+\subsection{\label{sec:invert_wband}
+Inversion}
+
+Since the eigenvalues are related to the updating kernel
+by a cosine transform,
+we can invert the relation as
+%
+\begin{equation}
+  \mu_l
+  =
+  \frac { 1 } { g \, n }
+  \sum_{ k = 1 }^{ g \, n }
+  \lambda_k
+  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
+            {            g \, n        }
+  .
+\label{eq:mu_from_lambda}
+\end{equation}
+%
+In the non-periodic case,
+we have defined
+$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
+for $k = n + 2, \dots, 2 \, n$,
+as well as
+%
+\begin{align}
+  \lambda_{ n + 1 }
+  =
+  (-1)^{ n - 1 }
+  \lambda_1
+  +
+  2 \, \sum_{ k = 2 }^{ n }
+      (-1)^{n - k} \, \lambda_k
+  ,
+\label{eq:lambdan}
+\end{align}
+to satisfy the constraint, $\mu_n = 0$.
+%
+
+\subsection{\label{sec:stabilize_wband}
+Stabilization}
+
+%A practical problem of using homogeneous updating scheme
+%is the following.
+%
+A homogeneous updating scheme
+specified by an arbitrary kernel
+is not necessarily stable,
+i.e. some eigenvalues can be negative.
+%
+%
+%\note{
+  %For example,
+  %the triple-bin updating scheme
+  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
+  %
+%}%
+Even
+the Gaussian updating scheme, which is stable in the continuous limit,
+can be unstable
+when the updating kernel is truncated during discretization.
+%
+Here we show how to minimally modify
+the updating kernel
+to stabilize it.
+%
+\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
+}
+
+Given an updating kernel,
+we can compute the eigenvalues from
+Eq. \eqref{eq:wband_eigenvalue}.
+%
+By replacing negative eigenvalues with zeros and
+using the inversion formula,
+Eq. \eqref{eq:mu_from_lambda},
+we get a new updating kernel,
+which is stable by construction.
+%
+However, the new kernel is widened
+as $\mu_l$ may be nonzero for $l > b$.
+%
+If we truncate the new kernel at $\mu_b$,
+the stability problem remains,
+although the negative eigenvalues
+tend to be smaller in magnitude.
+%
+Thus, to stabilize the updating kernel
+without increasing the width,
+we need to iterate the above process many times,
+resulting the following algorithm.
+
+%
+\begin{enumerate}
+  \item
+    Given an updating kernel, $\mu_0, \dots, \mu_b$,
+    compute the eigenvalues,
+    $\lambda_1, \dots, \lambda_n$,
+    from Eq. \eqref{eq:wband_eigenvalue}.
+  \item
+    If all eigenvalues are nonnegative within the numerical error,
+    we are done. % and exit the loop.
+  \item
+    Otherwise, replace negative eigenvalues by zeros,
+    and compute the new kernel,
+    $\mu_0, \dots, \mu_b$, from
+    Eq. \eqref{eq:mu_from_lambda},
+    with $\mu_l = 0$ for $l > b$.
+    Go to step 1.
+\end{enumerate}
 
@@
 
-\subsection{\label{sec:invert_wband}
-Inversion}
-
-Since the eigenvalues are related to the updating kernel
-by a cosine transform,
-we can invert the relation as
-%
+\subsubsection{\label{sec:homo_bandpass}
+Bandpass updating schemes}
+
+For the bandpass updating schemes,
+the eigenvalues are known and
+we can use the inversion formula, Eq. \eqref{eq:mu_from_lambda}
+to find the updating kernel.
+%
+Below we will slightly modify the definition of
+the mode cutoff, $K$,
+for mathematical convenience.
+
+For a periodic variable, we demand
+%
+$$
+\begin{aligned}
+%&
+%\lambda_1 = \cdots = \lambda_{K+1}
+%= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
+%\\
+%&
+\lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
+\end{aligned}
+$$
+for $K \ge 1$,
+with the rest of the eigenvalues being $1$.
+%
+Then by using
+Eq. \eqref{eq:mu_from_lambda},
+we get
 \begin{equation}
@@
   =
-  \frac { 1 } { g \, n }
-  \sum_{ k = 1 }^{ g \, n }
-  \lambda_k
-  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
-            {            g \, n        }
-  .
-\label{eq:mu_from_lambda}
-\end{equation}
-%
-In the non-periodic case,
-we have defined
-$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
-for $k = n + 2, \dots, 2 \, n$,
-as well as
-%
-\begin{align}
-  \lambda_{ n + 1 }
-  =
-  (-1)^{ n - 1 }
+  \frac{
+    \sin
+    \dfrac{ (2 K + 1) \, l \, \pi }
+         {              n        }
+  }
+  {
+    n \, \sin \dfrac{ l \, \pi } { n }
+  }
+  .
+\notag
+%\label{eq:mu_sinc_pbc}
+\end{equation}
+\note{Derivation.
+$$
+\begin{aligned}
+\mu_l
+&=
+\frac 1 n \sum_{k = 0}^{n-1} \lambda_{k+1} \cos \frac{ k \, l \, 2 \, \pi } { n }
+\\
+&=
+\frac{1}{n}
+\left(
+  1 +
+  \sum_{k=1}^K
+  \cos \frac { k \, l \, 2 \, \pi } { n }
+  +
+  \sum_{k=n-K}^{n-1}
+  \cos \frac { k \, l \, 2 \, \pi } { n }
+\right)
+\\
+&=
+\frac 1 n
+\sum_{k=-K}^K
+\cos \frac { k \, l \, 2 \, \pi } { n }
+=
+  \frac{
+    \sin
+    \dfrac{ (2 K + 1) \, l \, \pi }
+         {              n        }
+  }
+  {
+    n \, \sin \dfrac{ l \, \pi } { n }
+  }
+.
+\end{aligned}
+$$
+}
+
+For a non-periodic variable, we demand
+$$
+\lambda_1 = \cdots = \lambda_{K+1} = 1,
+\qquad
+\lambda_{K+2} = \cdots = \lambda_n = 0.
+$$
+Then Eq. \eqref{eq:lambdan} gives
+$\lambda_{n+1} = (-1)^{n+K-1}$,
+and from Eq. \eqref{eq:mu_from_lambda},
+we have
+\begin{equation}
+  \mu_l
+  =
+  \frac{1}{2 \, n}
+  \left[
+    (-1)^{n+K+l-1}
+    +
+    \frac{
+      \sin
+      \dfrac{ (2 K + 1) \, l \, \pi }
+           {         2 \, n        }
+    }
+    {
+      \sin \dfrac{ l \, \pi } { 2 \, n }
+    }
+  \right]
+  .
+\label{eq:mu_sinc_refl}
+\end{equation}
+\note{Derivation.
+$$
+\begin{aligned}
+  \lambda_{n+1}
+  &=
+  (-1)^{n-1}
+  \left[
     \lambda_1
+    - 2 \, \lambda_2
+    + 2 \, \lambda_3 - \cdots
+    + (-1)^{K} 2 \, \lambda_{K+1})
+  \right]
+  \\
+  &=
+  \begin{dcases}
+    (-1)^{n-1} & K \mathrm{\; even,} \\
+    (-1)^n     & K \mathrm{\; odd.}
+  \end{dcases}
+\end{aligned}
+$$
+So
+$$
+\begin{aligned}
+  \mu_l
+  &=
+  \frac{1}{2\,n}
+  \left[
+    1 +
+    2 \sum_{k=2}^{K+1}
+    \cos \frac { (k - 1) \, l \pi } { n }
     +
-  2 \, \sum_{ k = 2 }^{ n }
-      (-1)^{n - k} \, \lambda_k
-  ,
-\label{eq:lambdan}
-\end{align}
-to satisfy the constraint, $\mu_n = 0$.
-%
-
-\subsection{\label{sec:stabilize_wband}
-Stabilization}
-
-%A practical problem of using homogeneous updating scheme
-%is the following.
-%
-A homogeneous updating scheme
-specified by an arbitrary kernel
-is not necessarily stable,
-i.e. some eigenvalues can be negative.
-%
-%
-%\note{
-  %For example,
-  %the triple-bin updating scheme
-  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
-  %
-%}%
-Even
-the Gaussian updating scheme, which is stable in the continuous limit,
-can be unstable
-when the updating kernel is truncated during discretization.
-%
-Here we show how to minimally modify
-the updating kernel
-to stabilize it.
-%
-\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
-}
-
-Given an updating kernel,
-we can compute the eigenvalues from
-Eq. \eqref{eq:wband_eigenvalue}.
-%
-By replacing negative eigenvalues with zeros and
-using the inversion formula,
-Eq. \eqref{eq:mu_from_lambda},
-we get a new updating kernel,
-which is stable by construction.
-%
-However, the new kernel is widened
-as $\mu_l$ may be nonzero for $l > b$.
-%
-If we truncate the new kernel at $\mu_b$,
-the stability problem remains,
-although the negative eigenvalues
-tend to be smaller in magnitude.
-%
-Thus, to stabilize the updating kernel
-without increasing the width,
-we need to iterate the above process many times,
-resulting the following algorithm.
-
-%
-\begin{enumerate}
-  \item
-    Given an updating kernel, $\mu_0, \dots, \mu_b$,
-    compute the eigenvalues,
-    $\lambda_1, \dots, \lambda_n$,
-    from Eq. \eqref{eq:wband_eigenvalue}.
-  \item
-    If all eigenvalues are nonnegative within the numerical error,
-    we are done. % and exit the loop.
-  \item
-    Otherwise, replace negative eigenvalues by zeros,
-    and compute the new kernel,
-    $\mu_0, \dots, \mu_b$, from
-    Eq. \eqref{eq:mu_from_lambda},
-    with $\mu_l = 0$ for $l > b$.
-    Go to step 1.
-\end{enumerate}
+    (-1)^{n+K-1} (-1)^l
+  \right]
+  \\
+  &=
+  \frac{1}{2\,n}
+  \left[
+    (-1)^{n+K+l-1}
+    +
+    \sum_{k=-K}^{K}
+    \cos \frac { k \, l \pi } { n }
+  \right]
+  .
+\end{aligned}
+$$
+}%
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
+  }
+  \caption{
+    \label{fig:sinc}
+    Updating kernels of the bandpass updating schemes
+    compared to a Gaussian with a similar width.
+    %
+    The lines are a guide to the eyes.
+    Inset. Updating matrix in the non-periodic case.
+    %
+    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
+      The data were computed by the program \texttt{prog/predict},
+      see particularly the function \texttt{intq\_save()}
+      in \texttt{prog/intq.h} for panel (a),
+      and the functions \texttt{savewin()}
+      and \texttt{savewinmat()}
+      in \texttt{prog/invt.h} for panel (b),
+      The data were saved under the directory \texttt{data/sinc}.
+      See the \texttt{make} object \texttt{sincsingle}
+      in the \texttt{Makefile} there.
+    }%
+  }
+\end{center}
+\end{figure}
+
+In Fig. \ref{fig:sinc},
+we give examples of the updating kernels
+of the bandpass updating schemes,
+compared to a Gaussian
+with a similar width, $\sigma$,
+%
+\begin{equation}
+  \sigma
+  \approx
+  \frac
+  {
+    g \, n
+  }
+  {
+    \sqrt{ 2 \, \pi } \, 2 \, K
+  }
+  .
+\label{eq:sigma_equiv}
+\end{equation}
+%
+%where $g = 1$ for the periodic case
+%or $g = 2$ otherwise.
+%
+The kernel in the non-periodic case is a sawtooth
+because of the term, $(-1)^{n+K+l-1}$,
+in Eq. \eqref{eq:mu_sinc_refl}.
+%
+However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
+the ruggedness disappears
+in the updating matrix by Eq. \eqref{eq:w_band_refl},
+as the above oscillatory term for $l = i - j$
+is canceled by the term for either $l = i + j - 1$
+or $l = i + j - 2 \, n - 1$
+with the opposite sign.
 

### 2017-05-12 doc/invt.tex
small edits to invt.tex

@@
   \sum_{k = 1}^n
-    \phi_{ki} \, \phi_{kj}
+    \phi_{ki} \, \phi_{kj}^*
   &=
@@
   \sum_{i = 1}^n
-    \frac{ \phi_{ki} \, \phi_{li} }
+    \frac{ \phi_{ki} \, \phi_{li}^* }
          { \rho_i }
@@
   \delta_{kl}
-  .
+  ,
   \label{eq:eig_orthonormal_rows}
 \end{align}
+%
+where the superscript $^*$
+denotes the complex conjugate.
 
@@
   \sum_{ i = 1 }^n
-    \frac{ \phi_{ki} \, \phi_{1i} }
+    \frac{ \phi_{ki} \, \phi_{1i}^* }
          { \rho_i }
@@
   =
-  \frac{1}{\rho_i} \sum_{k=1}^n \lambda_k \, \phi_{ki} \, \phi_{kj}
+  \frac{1}{\rho_i} \sum_{k=1}^n
+  \lambda_k \, \phi_{ki}^* \, \phi_{kj}
   .
@@
   =
-  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \,
+  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj}^* \,
     \left\langle x_i \, x_j \right\rangle
@@
     \left\langle
-      y_k^2
+      \left| y_k \right|^2
     \right\rangle
@@
     \left\langle
-    y_k^2(T)
+      \left| y_k(T) \right|^2
     \right\rangle
@@
 of the histogram fluctuation,
-$\kappa_k(t) \equiv \langle \theta_k(0) \, \theta_k(t) \rangle$,
+$\kappa_k(t) \equiv \langle \theta_k^*(0) \, \theta_k(t) \rangle$,
 as:
@@
   = \sum_{t = -\infty}^\infty \kappa_k(t)
-  = \sum_{t = -\infty}^\infty \left\langle \theta_k(0) \, \theta_k(t) \right\rangle
+  = \sum_{t = -\infty}^\infty
+  \left\langle \theta_k^*(0) \, \theta_k(t) \right\rangle
   .
@@
 \begin{equation}
-  \left\langle y_k^2(0) \right\rangle = \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  \left\langle
+    \left| y_k(0) \right|^2
+  \right\rangle
+  = \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
   ,
@@
 schemes, the eigenvalues can be readily found
-from the shape of the window function.
-%
-
-%\subsubsection{\label{sec:bandkernel}
-%Updating kernel and eigenvalues}
-
-%While translationally-invariant
-%updating schemes
-%naturally suit periodic variables\cite{
-%dama2014},
-%they can also be extended to non-periodic variables
-%by imposing the reflective boundary condition\cite{
-%bussi2006}.
-%%
-%For simplicity, we will assume that the target
-%distribution is flat, or $p_i = 1/n$, below.
-
-%It is natural to first introduce homogeneous updating schemes
-For a periodic variable\cite{dama2014},
-the updating matrix, $\mathbf w$,
-assumes the following form
-%
-\begin{equation}
-  w_{ij}
-  =
-  \mu_{i-j}
-  +
-  \mu_{i-j+n}
-  +
-  \mu_{i-j-n}
-  ,
-\notag
-%\label{eq:w_band_pbc}
-\end{equation}
-%
-where the numbers,
-$\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
-characterizing the rigid window
-will be referred to as an \emph{updating kernel}\cite{bussi2006}.
-%
-To satisfy Eq. \eqref{eq:w_sumj},
-we impose the normalization
-%
-\begin{equation}
-  \mu_{-b} + \cdots + \mu_b = 1
-  ,
-\label{eq:mu_normalization}
-\end{equation}
-%
-with $\mu_l = 0$ for $|l| > b$.
-%
-If $b = n/2$, $\mu_{-b}$ is removed
-from the sum in Eq. \eqref{eq:mu_normalization}
-to avoid double counting.
-%
-For simplicity, we also impose the symmetry
-%
-\begin{equation}
-  \mu_i = \mu_{-i}
-  .
-\label{eq:mu_symm}
-\end{equation}
-%
-As shown in Appendix \ref{sec:wband_eig},
-%the multiplication of $\mathbf w$
-%is equivalent to a convolution with the kernel, and
-the eigenvalues of $\mathbf w$ are
-the Fourier or cosine transform of the updating kernel:
+from the window function ($\mu_0, \mu_1, \dots, \mu_b$)
+as a cosine transform,
 %
@@
 %
-where $g = 1$ here for the periodic variable.
-
-For a non-periodic variable,
-we use the reflective boundary condition\cite{bussi2006},
-and change the form of the updating matrix to
-%
-%
-\begin{equation}
-  w_{ij}
-  =
-  \mu_{ i - j }
-  +
-  \mu_{ i + j - 1 }
-  +
-  \mu_{ i + j - 2 n - 1 }
-  ,
-  \label{eq:w_band_refl}
-\end{equation}
-%
-where the last two terms on the right hand side,
-representing two reflective mirrors at
-$j = 1/2$ and $j = n + 1/2$,
-are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
-to the equilibrium distribution\cite{bussi2006}.
-%such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
-%are satisfied.
-%
-Again, the updating kernel should satisfy
-Eqs. \eqref{eq:mu_normalization}
-and \eqref{eq:mu_symm},
-but
-the cutoff $b$ only needs to be less than $n$.
-
-Note that the two reflective mirrors effectively
-define a periodic variable of period $2 \, n$.
-Thus, the eigenvalues can be borrowed from
-Eq. \eqref{eq:wband_eigenvalue},
-with a doubled period,
-i.e. $g$ is changed to $2$ in the non-periodic case.
-
-%\subsubsection{\label{sec:invert_wband}
-%Inversion}
+where $g = 1$ here for the periodic variable,
+or $2$ for a non-periodic one
+[cf. Appendix \ref{sec:more_wband}].
 
@@
 %
-$$w_{ij} = \sum_{k=1}^{K+1} \phi_{ki} \, \phi_{kj},$$
+$$w_{ij} = n \sum_{k=1}^{K+1} \phi_{ki}^* \, \phi_{kj},$$
 %
@@
   \phi_{ki} \phi_{kj}
-  \langle \xi_i(t) \, \xi_j(t') \rangle
+  \langle \xi_i^*(t) \, \xi_j(t') \rangle
   %
@@
 
-To find the eigenvectors for a periodic variable, $\phi_i$,
-we define the out-of-boundary values by
+%\subsubsection{\label{sec:bandkernel}
+%Updating kernel and eigenvalues}
+
+%While translationally-invariant
+%updating schemes
+%naturally suit periodic variables\cite{
+%dama2014},
+%they can also be extended to non-periodic variables
+%by imposing the reflective boundary condition\cite{
+%bussi2006}.
+%%
+%For simplicity, we will assume that the target
+%distribution is flat, or $p_i = 1/n$, below.
+
+It is natural to first introduce homogeneous updating schemes
+for a periodic variable\cite{dama2014},
+the updating matrix, $\mathbf w$,
+assumes the following form
+%
+\begin{equation}
+  w_{ij}
+  =
+  \mu_{i-j}
+  +
+  \mu_{i-j+n}
+  +
+  \mu_{i-j-n}
+  ,
+\notag
+%\label{eq:w_band_pbc}
+\end{equation}
+%
+where the numbers,
+$\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
+characterizing the rigid window
+will be referred to as an \emph{updating kernel}\cite{bussi2006}.
+%
+To satisfy Eq. \eqref{eq:w_sumj},
+we impose the normalization
+%
+\begin{equation}
+  \mu_{-b} + \cdots + \mu_b = 1
+  ,
+\label{eq:mu_normalization}
+\end{equation}
+%
+with $\mu_l = 0$ for $|l| > b$.
+%
+If $b = n/2$, $\mu_{-b}$ is removed
+from the sum in Eq. \eqref{eq:mu_normalization}
+to avoid double counting.
+%
+For simplicity, we also impose the symmetry
+%
+\begin{equation}
+  \mu_i = \mu_{-i}
+  .
+\label{eq:mu_symm}
+\end{equation}
+
+To find the eigenvectors,
+we define for a periodic variable, $\phi_i$,
+the out-of-boundary values by
 %
@@
   %\overset{ \cos } { \sin }
+  \,
   {\cos \atop \sin}
@@
 For a non-periodic variable,
-we define the out-of-boundary values
+we use the reflective boundary condition\cite{bussi2006},
+and change the form of the updating matrix to
+%
+%
+\begin{equation}
+  w_{ij}
+  =
+  \mu_{ i - j }
+  +
+  \mu_{ i + j - 1 }
+  +
+  \mu_{ i + j - 2 n - 1 }
+  ,
+  \label{eq:w_band_refl}
+\end{equation}
+%
+where the last two terms on the right hand side,
+representing two reflective mirrors at
+$j = 1/2$ and $j = n + 1/2$,
+are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
+to the equilibrium distribution\cite{bussi2006}.
+%such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
+%are satisfied.
+%
+Again, the updating kernel should satisfy
+Eqs. \eqref{eq:mu_normalization}
+and \eqref{eq:mu_symm},
+but
+the cutoff $b$ only needs to be less than $n$.
+
+Note that the two reflective mirrors effectively
+define a periodic variable of period $2 \, n$.
+Thus, the eigenvalues can be borrowed from
+Eq. \eqref{eq:wband_eigenvalue},
+with a doubled period,
+i.e. $g$ is changed to $2$ in the non-periodic case.
+
+%\subsubsection{\label{sec:invert_wband}
+%Inversion}
+
+%For a non-periodic variable,
+We define the out-of-boundary values
 of $\phi_j$ as
@@
 For the bandpass updating schemes,
-the eigenvalues are known and
-we can use the inversion formula, Eq. \eqref{eq:mu_from_lambda}
+the eigenvalues are known ($0$ or $1$) and
+we can use Eq. \eqref{eq:mu_from_lambda}
 to find the updating kernel.
@@
     \sin
-    \dfrac{ (2 K + 1) \, l \, \pi }
+    \frac{ (2 K + 1) \, l \, \pi }
          {              n        }
@@
   {
-    n \, \sin \dfrac{ l \, \pi } { n }
-  }
-  .
+    n \, \sin \frac{ l \, \pi } { n }
+  }
+  ,
 \notag
@@
 \end{equation}
+and $w_{ij} = \mu_{i-j}$.
+%
 \note{Derivation.
@@
   {
-    n \, \sin \dfrac{ l \, \pi } { n }
+    n \, \sin \frac{ l \, \pi } { n }
   }
@@
   =
-  \frac{1}{2 \, n}
-  \left[
-    (-1)^{n+K+l-1}
+  \frac{ (-1)^{n+K+l-1} } { 2 \, n }
     +
@@
     \sin
-      \dfrac{ (2 K + 1) \, l \, \pi }
+    \frac{ (2 K + 1) \, l \, \pi }
          {         2 \, n        }
@@
   {
-      \sin \dfrac{ l \, \pi } { 2 \, n }
-    }
-  \right]
+    2 \, n \, \sin \frac{ l \, \pi } { 2 \, n }
+  }
   .
@@
 \end{equation}
+%
+The updating matrix is given by
+%
+\begin{equation}
+  w_{ij}
+  =
+    \frac{
+      \sin
+      \frac{ (2 K + 1) \, (i-j) \, \pi }
+           {         2 \, n        }
+    }
+    {
+      2 \, n \, \sin \frac{ (i-j) \, \pi } { 2 \, n }
+    }
+    +
+    \frac{
+      \sin
+      \frac{ (2 K + 1) \, (i+j-1) \, \pi }
+           {         2 \, n        }
+    }
+    {
+      2 \, n \, \sin \frac{ (i+j-1) \, \pi } { 2 \, n }
+    }
+  .
+  \notag
+\end{equation}
+%
+Although the kernel is a sawtooth
+because of the term, $(-1)^{n+K+l-1}$,
+the ruggedness disappears
+in the updating matrix
+as the above oscillatory term for $l = i - j$
+is canceled by the term for either $l = i + j - 1$
+or $l = i + j - 2 \, n - 1$
+with the opposite sign.
+%
 \note{Derivation.
@@
 
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
-  }
-  \caption{
-    \label{fig:sinc}
-    Updating kernels of the bandpass updating schemes
-    compared to a Gaussian with a similar width.
-    %
-    The lines are a guide to the eyes.
-    Inset. Updating matrix in the non-periodic case.
-    %
-    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h} for panel (a),
-      and the functions \texttt{savewin()}
-      and \texttt{savewinmat()}
-      in \texttt{prog/invt.h} for panel (b),
-      The data were saved under the directory \texttt{data/sinc}.
-      See the \texttt{make} object \texttt{sincsingle}
-      in the \texttt{Makefile} there.
-    }%
-  }
-\end{center}
-\end{figure}
-
-In Fig. \ref{fig:sinc},
-we give examples of the updating kernels
-of the bandpass updating schemes,
-compared to a Gaussian
-with a similar width, $\sigma$,
-%
-\begin{equation}
-  \sigma
-  \approx
-  \frac
-  {
-    g \, n
-  }
-  {
-    \sqrt{ 2 \, \pi } \, 2 \, K
-  }
-  .
-\label{eq:sigma_equiv}
-\end{equation}
-%
-%where $g = 1$ for the periodic case
-%or $g = 2$ otherwise.
-%
-The kernel in the non-periodic case is a sawtooth
-because of the term, $(-1)^{n+K+l-1}$,
-in Eq. \eqref{eq:mu_sinc_refl}.
-%
-However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
-the ruggedness disappears
-in the updating matrix by Eq. \eqref{eq:w_band_refl},
-as the above oscillatory term for $l = i - j$
-is canceled by the term for either $l = i + j - 1$
-or $l = i + j - 2 \, n - 1$
-with the opposite sign.
+%\begin{figure}[h]
+%\begin{center}
+%  \makebox[\linewidth][c]{
+%    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
+%  }
+%  \caption{
+%    \label{fig:sinc}
+%    Updating kernels of the bandpass updating schemes
+%    compared to a Gaussian with a similar width.
+%    %
+%    The lines are a guide to the eyes.
+%    Inset. Updating matrix in the non-periodic case.
+%    %
+%    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
+%      The data were computed by the program \texttt{prog/predict},
+%      see particularly the function \texttt{intq\_save()}
+%      in \texttt{prog/intq.h} for panel (a),
+%      and the functions \texttt{savewin()}
+%      and \texttt{savewinmat()}
+%      in \texttt{prog/invt.h} for panel (b),
+%      The data were saved under the directory \texttt{data/sinc}.
+%      See the \texttt{make} object \texttt{sincsingle}
+%      in the \texttt{Makefile} there.
+%    }%
+%  }
+%\end{center}
+%\end{figure}
+
+%In Fig. \ref{fig:sinc},
+%we give examples of the updating kernels
+%of the bandpass updating schemes,
+%compared to a Gaussian
+%with a similar width, $\sigma$,
+%%
+%\begin{equation}
+%  \sigma
+%  \approx
+%  \frac
+%  {
+%    g \, n
+%  }
+%  {
+%    \sqrt{ 2 \, \pi } \, 2 \, K
+%  }
+%  .
+%\label{eq:sigma_equiv}
+%\end{equation}
+%%
+%%where $g = 1$ for the periodic case
+%%or $g = 2$ otherwise.
+%%
 

### 2017-05-25 doc/invt.tex
adding a few references

@@
   of the updating magnitude
-  for a class of free energy calculation methods including
+  for a class of flat-distribution sampling methods
+  for free energy calculation including
   the Wang-Landau (WL) algorithm and metadynamics.
@@
   The updating magnitude should decrease over time
-  to reduce the noise in the bias potential,
+  to reduce the error in the bias potential,
   and one may speed up the convergence by choosing an optimal schedule
@@
   We will show that
-  the optimal schedule depends on the updating scheme
-  (such as the single-bin and Gaussian schemes commonly used in WL and metadynamics),
-  and the initial updating magnitude is ideally half of
-  the previous equilibrium value.
+  the optimal schedule depends on the updating scheme.
+  %
+  While the asymptotically optimal schedule for
+  the single-bin updating scheme (commonly used in WL)
+  is given by inverse-time formula,
+  that for the Gaussian updating scheme (commonly used in metadynamics)
+  is often more complex.
+  %and the initial updating magnitude is ideally half of
+  %the previous equilibrium value.
   %
   Further,
-  the single-bin updating scheme
+  we show that the single-bin updating scheme
   belongs to a class of bandpass updating schemes
-  that are optimal for asymptotic convergence,
-  and the optimal schedule of these schemes
-  is the inverse number of updating steps with a proper shift,
-  as given by the inverse-time prescription.
+  that are optimal for asymptotic convergence.
+  %
+  These bandpass updating schemes aim at
+  partially flattening a few selected histogram modes
+  and their optimal schedule
+  is also given by the inverse-time prescription.
+  %
+  Several previous flat-distribution-sampling algorithms
+  can be cast as special cases of the bandpass updating schemes. 
   %
@@
 for an adaptive FDS method
-with a general window function,
+of a general updating scheme,
 including the single-bin and Gaussian one
@@
 %
-In either case, however,
-the initial updating magnitude
-can be optimally set to
+%In either case, however,
+Ideally,
+the optimal initial updating magnitude
+can be set to
 half of the previous equilibrium value,
@@
 %
-Further, we can generalize
+Further, we will show that
 the single-bin updating scheme
-to a class of bandpass updating schemes
-that are optimal for long simulations,
-and the above origin-shifted inverse-time schedule
-turns out to be optimal for these schemes.
-%
+is optimal for asymptotic convergence,
+and it can be generalized
+to a class of asymptotically optimal
+bandpass updating schemes.
+%
+The latter, however,
+allow one to selectively flatten only
+a few long-range histogram modes,
+and to avoid introducing local noise
+into the bias potential during the updating process.
+%
+The optimal schedule of bandpass updating schemes
+is also the above origin-shifted inverse-time schedule.
+%
+Interestingly, several previous FDS algorithms\cite{
+  langfeld2012, pellegrini2014,
+  neuhaus2006, *neuhaus2007}
+
 %For finite-length simulations, however,
@@
 %
-These results may help to better
-plan and optimize FDS simulations.
+%These results may help to better
+%plan and optimize FDS simulations.
 
@@
 as the sample size increases\cite{
-  marsili2006, barducci2008}.
+  marsili2006, barducci2008, robbins1951, pellegrini2014}.
 %

### 2017-05-26 doc/invt.tex
simplify invt.tex

@@
 }
+The same schedule also applies to a variant of
+the WL algorithm\cite{langfeld2012, pellegrini2014}.
 %
@@
   neuhaus2006, *neuhaus2007}
+can be cast as special cases of bandpass updating schemes.
 
@@
 %
-%The discreteness of $i$ is natural
-%in discrete models.
-%
-For example, $i$ can be the energy, $E$,
-in a lattice spin model or the temperature index
-in a simulated tempering simulation\cite{
-marinari1992, *lyubartsev1992}.
+%For example, $i$ can be the energy, $E$,
+%in a lattice spin model or the temperature index
+%in a simulated tempering simulation\cite{
+%marinari1992, *lyubartsev1992}.
 %
 For a continuous quantity, $z$,
-which is often the case in a (classical) molecular system,
-we can discretize $z$
+%which is often the case in a (classical) molecular system,
+we can discretize it
 such that the integer $i$ represents
@@
 %
-The distribution is normalized as
-$\sum_{i = 1}^n p^*_i = 1$.
+%The distribution is normalized as
+%$\sum_{i = 1}^n p^*_i = 1$.
 
@@
 %To do so, we introduce a bias potential, $V_i$,
-If the distribution, $p^*_i$, is multimodal,
-it is often advantageous to perform
-biased sampling that targets
-a smoother distribution, $p_i$,
-by introducing a bias potential, $V_i$,
-such that the biased equilibrium distribution is
-%
-\begin{equation}
-  \pi_i
-  %=
-  %C_i \, p^*_i \, e^{-V_i}
-  =
-  \frac{                p^*_i \, e^{-V_i} }
-       { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
-  \propto
-  p^*_i \, e^{-V_i}
-  .
-  \label{eq:pi_p_V}
-\end{equation}
+The challenge of sampling a multimodal distribution, $p^*_i$,
+with deep valley(s) can often be eased by
+biased sampling targeting a smoother distribution, $p_i$.
+%
 %
@@
 %coincide with the intended distribution $p_i$.
-
+%
 We will broadly define
@@
 or nearly flat one\cite{
-dayal2004, *trebst2004, *zhang2007, barducci2008, singh2011}.
-%
-For the biased distribution, $\pi_i$,
-to converge to the desired $p_i$,
-the bias potential must satisfy
+dayal2004, *trebst2004, barducci2008, singh2011}.
+%
+By introducing a bias potential, $V_i$,
+the biased equilibrium distribution becomes
+%
+\begin{equation}
+  \pi_i
+  %=
+  %C_i \, p^*_i \, e^{-V_i}
+  =
+  \frac{                p^*_i \, e^{-V_i} }
+       { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
+  \propto
+  p^*_i \, e^{-V_i}
+  ,
+  \label{eq:pi_p_V}
+\end{equation}
+%
+and our aim is to find a biased potential
+such that the resulting $\pi_i$ would coincide
+with the desired $p_i$.
+%
+Upon convergence, the bias potential would satisfy
 %
@@
   V_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
-  .
+  ,
   \label{eq:Vi_target}
@@
 %
-This relation thus sets a practical goal for the convergence of FDS
-methods---the bias potential must be adjusted to make the
-observed distribution match the intended one.
-%%
-Once this is done,
-the PMF, $-\ln p^*_i$, can be naturally
+and the PMF, $-\ln p^*_i$, can be readily 
 deduced from the bias potential. % $V_i$.
@@
 
-In terms of how one adjusts the bias potential,
-there are roughly two types of FDS methods.
-%
-The first type,
-referred to as the equilibrium FDS, % (E-FDS) methods hereinafter,
-uses a fixed bias potential, $V_i$,
-(estimated a priori)
-during simulation.
-%
-Examples include the original
-entropic or multicanonical sampling\cite{
-berg1992, *lee1993}
-and simulated tempering\cite{
-marinari1992, *lyubartsev1992}.
-%with $i$ being the index of energy
-%and that of temperature, respectively.
-%
-After a sufficiently long run, the bias potential, $V_i$,
-can often be improved
-with the normalized histogram, $H_i$,
+Many methods\cite{mezei1987, berg1992, *lee1993,
+wang2001, *wang2001pre, huber1994,
+*laio2002, *laio2008, *barducci2011, *sutto2012}
+have been developed to find the proper bias potential.
+%
+In a class of equilibrium FDS methods\cite{
+  mezei1987, berg1992, *lee1993, marinari1992, *lyubartsev1992},
+the bias potential, $V_i$, is
+estimated a priori and fixed
+during simulation,
+%
+and a correction to the bias potential
+is derived from the normalized histogram, $H_i$,
 accumulated from the simulation as
@@
 }
-%However, for an unknown system,
-%it is often difficult
-%to guess a sufficiently accurate
-%$V_i$ with which to start.
-%%
-%Thus,
-%it can take many iterative rounds of simulations
-%to achieve the desired flat distribution.
-%%
-%%Another difficulty of this approach is that
-%The effectiveness of this approach
-%depends on the precision of the histogram, $H_i$,
-%which in turn depends on the simulation length
-%that must be decided in advance.
 However, this formula often requires a long
@@
 
-The second type of methods,
-including the WL algorithm and metadynamics,
-addresses the above shortcomings
-by continuously and incrementally updating the bias potential.
-%For this reason, the second type of methods,
-%referred to as the adaptive FDS methods,
-%were invented.
-%
-These methods, referred to as the adaptive FDS methods,
-will be discussed below.
+An other class of adaptive FDS methods,
+including the WL algorithm\cite{wang2001, *wang2001pre}
+and metadynamics\cite{huber1994, laio2002, *laio2008, *barducci2011, *sutto2012},
+address the above shortcoming
+by continuously and incrementally updating the bias potential,
+and we will focus on these methods below.
 %%
@@
 %
-In the WL algorithm\cite{wang2001, *wang2001pre},
+In the original WL algorithm\cite{wang2001, *wang2001pre},
 the bias potential, $V_i(t)$, is updated
@@
   \frac{ \alpha(t) }
-       { p_{i(t)} },
+       { p_{i(t)} }
+  .
   \label{eq:mbin_update}
@@
 %
-where the elements, $w_{ij}$'s, are a set of coefficients satisfying
-the restrictions to be discussed
-in Sec. \ref{sec:updating-matrix}.
+The Gaussian updating scheme used in metadynamics
+is an example, with $w_{ij}$ being a Gaussian of $i-j$.
 %
@@
 \begin{equation}
-  \alpha(t) = \frac{1}{t + 1},
+  \alpha(t) = \frac{1}{t},
   \label{eq:alpha_invt}
@@
 %
-where $t + 1$,
+where $t$,
 referred to as the ``time'' below,
@@
 %
-An intuitive explanation of this result\cite{
+This can be derived from a general result\cite{
+  robbins1951, pellegrini2014}
+and an intuitive explanation\cite{
   marsili2006, barducci2008}
@@
 affects the average by its due weight,
-which is the inverse of the current sample size, $t + 1$
+the inverse of the current sample size, $t$
 (cf. Appendix \ref{sec:equilerr}).
 %
-Alternatively, we may derive it
+We may also derive it
 from optimally planning the simulation lengths in the stages
@@
 %
-We will suggest a slight modification of Eq. \eqref{eq:alpha_invt}
-whereby the origin of $t$ is shifted by an amount
-determined from the initial error.
+%We will suggest a slight modification of Eq. \eqref{eq:alpha_invt}
+%whereby the origin of $t$ is shifted by an amount
+%determined from the initial error.
 %
@@
 
-\subsubsection{Error definition}
-
-First, we define the net error in the bias potential
+\subsubsection{Error}
+
+We define the net error in the bias potential
 by deducting two contributions.
 %
-The first contribution comes from the target value,
-Eq. \eqref{eq:Vi_target}.
-%
-For this purpose,
+First, to deduct the target value given by
+Eq. \eqref{eq:Vi_target},
 we introduce a shifted bias potential
@@
 upon convergence.
-Note that the updates to $v_i$'s are
-equivalent to the updates to $V_i$'s,
+Note that the updates to $v_i(t)$ are
+equivalent to the updates to $V_i(t)$,
 since the shift
@@
 %
-the real deviation is more readily defined
-from the difference between $v_i(t)$ and a baseline value, $\bar v(t)$,
-that represents the average growth\cite{
+the real deviation, $x_i(t)$, of the bias potential
+depends only on the difference between $v_i(t)$
+and a baseline value, $\bar v(t)$,
+for average growth\cite{
 dama2014},
@@
 %
-In this way,
-the variable, $x_i(t)$,
-represents the actual deviation of the bias potential, $V_i(t)$,
-and the total error is given by
+The total error is given by
 %
@@
 
-Although the above error weighting distribution, $\pmb\rho$,
-is usually flat,
-i.e. $\rho_i \equiv 1/n$,
-in practice,
-a non-flat distribution
-is sometimes more convenient in analyzing
-some multiple-bin updating schemes later on.
-%
-We also note that $\pmb\rho$
-is independent of
+%Although the above error weighting distribution, $\pmb\rho$,
+%is usually flat,
+%i.e. $\rho_i \equiv 1/n$,
+%in practice,
+%a non-flat distribution
+%is sometimes more convenient in analyzing
+%some multiple-bin updating schemes later on.
+%
+Note that the distribution, $\pmb\rho$, is arbitrary
+and independent of
 the target distribution,
-$\mathbf p = (p_1, \dots, p_n)$
-(e.g. in well-tempered metadynamics,
+$\mathbf p = (p_1, \dots, p_n)$,
+although it is usually flat.\footnote{In
+well-tempered metadynamics,
 $\mathbf p$ is an attenuated version of
@@
 $\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
-whereas $\pmb\rho$ is flat).
+whereas $\pmb\rho$ is flat.}
 %
@@
 approximate the finite difference Eq. \eqref{eq:wl_update}
-by a differential equation
+by a differential equation\footnote{This
+approximation requires us to shift the origin of time by $1$ in some cases.}
 %
@@
 
-Next, we split $h_i(t)$ into a deterministic averaged part
+We can split $h_i(t)$ into a deterministic averaged part
 and a random fluctuation part, $\zeta_i(t)$,
@@
 For non-white noise,
-the equivalent value of $G^{(0)}_i$ can be approximately obtained
-%from the integral,
-%$\int_{-\infty}^\infty \sigma_{ij}(t) \, dt$,
-%or in the discrete time case,
-$\sum_{t = -\infty}^\infty \sigma_{ij}(t)$.
+the equivalent value of $G^{(0)}_i$ can be approximately obtained as
+$\sum_{t = -\infty}^\infty \langle \zeta_i(t) \, \zeta_j(0) \rangle$.
 %
@@
 
-The total error
-at the end of the simulation, $t = T$,
-as defined in Eq. \eqref{eq:error_def},
-now depends implicitly on the function $\alpha(t)$
-via $x_i(T)$.
-%
-Our aim is to find the $\alpha(t)$
-that minimizes this error.
-%%
-%It suffices to consider
-%a one-variable problem that
-%minimizes each individual
-%$\left\langle x_i^2(T) \right\rangle$.
-
-%\subsubsection{\label{sec:onevar}
-%Optimization of a one-variable problem}
-
-\subsubsection{Optimal schedule for the single-bin scheme}
-
-We recall that
-the formal solution of Eq. \eqref{eq:dxdt_singlebin} is
+The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
 %
@@
 \end{align}
-%
+
+\subsubsection{Optimal schedule}
+
 At the end of period $T$,
@@
 is given by
-%
-%\begin{align}
-%  &\left\langle x_i^2(T) \right\rangle
-%  =
-%  \langle x_i^2(0) \rangle \, e^{-2 \, q(T)}
-%  +
-%  G_i \int_0^T
-%    {\dot u}^2\bigl( q(t) \bigr) \,
-%    \, dt.
-%  \label{eq:x2t_average}
-%\end{align}
 %
@@
 
-This is a functional of the curve $q(t)$.
+The error depends implicitly on the curve, $\alpha(t)$,
+or, equivalently, the curve, $q(t)$,
+via $x_i(T)$, and
+our aim is to find the $q(t)$
+that minimizes this error.
 %
@@
 \end{equation}
-where $t_0$ is a constant,
-and the constant $c$ is determined
-from the boundary condition at $t = T$
+where $t_0$ is some constant,
+and the other constant, $c$, has been determined
+from the $t = T$ case
 in the second step.
@@
 and Eq. \eqref{eq:alpha_invt}
-is the special case of $t_0 = 1$.
+is the special case of $t_0 = 0$.
 %
@@
 or $2$ for a non-periodic one
-[cf. Appendix \ref{sec:more_wband}].
+(cf. Appendix \ref{sec:more_wband}).
 
@@
 
-Here we give a procedure of computing
+Based on the above results,
+we give a procedure of computing
 the optimal schedule and the error
@@
 %
-In Appendix \ref{sec:upperbound},
-we will show that the opposite limit of
-a set of well-separated eigenvalues
-will provide an upper bound for
-the asymptotic error, Eq. \eqref{eq:error_asym_ubound}.
+%In Appendix \ref{sec:upperbound},
+%we will show that the opposite limit of
+%a set of well-separated eigenvalues
+%will provide an upper bound for
+%the asymptotic error, Eq. \eqref{eq:error_asym_ubound}.
 
@@
 \subsection{\label{sec:results_invt}
-Modified inverse-time schedule}
+Scaled inverse-time schedule}
 
@@
 tested our analysis of computing errors
-on the modified inverse-time schedule,
+on the scaled inverse-time schedule,
 %
@@
 can be computed analytically
-(cf. Appendix \ref{sec:invt_schedule}),
+(cf. Appendix \ref{sec:sinvt_schedule}),
 and tested against simulation.
@@
 We performed adaptive FDS simulations
-using the single-bin updating scheme
-and the triple-bin updating scheme with $\mu_1 = 0.24$.
+using the single-bin updating scheme.
+%and the triple-bin updating scheme with $\mu_1 = 0.24$.
 %
@@
     Error, $\Err$, versus the parameter, $\lambda$,
-    for the modified inverse-time schedule,
+    for the single-bin updating scheme
+    using the scaled inverse-time schedule,
     Eq. \eqref{eq:alpha_invtlambda}.
@@
     the curves are theoretical predictions.
-    %
-    For the single-bin (WL) updating scheme,
-    the minimal should occur at $\lambda = 1$,
-    while it is not necessarily the case for a general updating scheme.
     %
@@
 The sampling process affected the error function
-only by a multiplicative constant,
-%The error functions of the two sampling processes
-%differed only by a multiplicative constant,
-(cf. Appendix \ref{sec:invt_singlebin}).
-%
-For the triple-bin updating scheme,
-the optimal $\lambda$ was less than $1.0$
-for the global sampling process,
-although it stayed around $1.0$
-for the local sampling process,
-also in agreement with the discussion
-in Appendix \ref{sec:invt_nn}.
-%
-Thus, the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1},
-is optimal for the single-bin updating scheme,
-but not necessarily so for a general multiple-bin updating scheme.
-%
+only by a multiplicative constant, $\Gamma$
+(cf. Appendix \ref{sec:sinvt_schedule}).
+%
+%For the triple-bin updating scheme,
+%the optimal $\lambda$ was less than $1.0$
+%for the global sampling process,
+%although it stayed around $1.0$
+%for the local sampling process,
+%also in agreement with the discussion
+%in Appendix \ref{sec:invt_nn}.
+%%
+%Thus, the inverse-time schedule,
+%Eq. \eqref{eq:alpha_invt1},
+%is optimal for the single-bin updating scheme,
+%but not necessarily so for a general multiple-bin updating scheme.
+%%
 
@@
 %
-As shown in Fig. \ref{fig:xerr},
-the inverse-time schedule was effective in error reduction
-only for the first few eigenmodes,
-whose eigenvalues were around $1.0$.
-%
-We can, however, ignore these local eigenmodes
-by explicitly filtering them out from the final bias potential,
-and truncating the error sum in Eq. \eqref{eq:error_def}
-to the first few eigenmodes.
-%
-Then, the corresponding optimal schedule
-from Eq. \eqref{eq:q_opt} that minimizes the truncated error
-became similar to the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1},
-as exemplified in Fig. \ref{fig:optacmp}.
-%
-\note{A more thorough approach is
-to use the bandpass updating scheme
-to be discussed later.
-}
+%As shown in Fig. \ref{fig:xerr},
+%the inverse-time schedule was effective in error reduction
+%only for the first few eigenmodes,
+%whose eigenvalues were around $1.0$.
+%%
+%We can, however, ignore these local eigenmodes
+%by explicitly filtering them out from the final bias potential,
+%and truncating the error sum in Eq. \eqref{eq:error_def}
+%to the first few eigenmodes.
+%%
+%Then, the corresponding optimal schedule
+%from Eq. \eqref{eq:q_opt} that minimizes the truncated error
+%became similar to the inverse-time schedule,
+%Eq. \eqref{eq:alpha_invt1},
+%as exemplified in Fig. \ref{fig:optacmp}.
+%%
 
@@
     &
-      Inverse-time, \newline
+      $1/t$, \newline
       Eq. \eqref{eq:alpha_invt1}
     &
-      Modified, \newline
+      Scaled $1/t$, \newline
       Eq.~\eqref{eq:alpha_invtlambda}
@@
 
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/xerr.pdf}
-  }
-  \caption{
-    \label{fig:xerr}
-    Predicted error components for the Gaussian updating scheme
-    with $\sigma = 10$
-    from the optimal schedule
-    and the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
-    %
-    The latter schedule failed to damp out the errors
-    from the long-wavelength (local) eigenmodes
-    with $k \ge 5$.
-    %
-    For longer simulations,
-    the error components from the optimal schedule
-    would roughly maintain a relatively flat profile,
-    but meet the initial error curve
-    at a lower level (not shown),
-    where $\lambda_k \approx 1/q(T)$
-    (cf. Appendix \ref{sec:upperbound}).
-    %
-    \note{This figure was produced by \texttt{doc/fig/xerr.gp}.
-      The data were saved under the directory \texttt{data/xerr}.
-    }%
-  }
-\end{center}
-\end{figure}
-
 An interesting prediction from the theory is Eq. \eqref{eq:half_alpha0},
@@
 although the dependence on $\alpha(0)$
-is generally weak.
-%
-The weak dependence is best illustrated in
+is generally weak.\footnote{The weak dependence is best illustrated in
 the case of single-bin updating scheme,
@@
 simulation time in the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1}.
+Eq. \eqref{eq:alpha_invt1}.}
 
@@
 
-We can understand the optimality of the inverse-time schedule
+The optimality of the inverse-time schedule can be considered
 as a natural consequence of the stabilization of a runtime average
 as the sample size increases\cite{
-  marsili2006, barducci2008, robbins1951, pellegrini2014}.
-%
-Consider a sample, $\{X^{(1)}, \dots, X^{(N)}\}$,
-with a dynamically growing size.
-%
-The runtime average
+  marsili2006, barducci2008}.
+%
+We first note that the normalized histogram, $H_i(t)$,
+is a runtime average of the instantaneous histogram, $h_i(t)$.
 %
 \begin{equation*}
-  {\bar X}^{(N)}
-  =
-  \frac{ X^{(1)} + \cdots + X^{(N)} } { N }
-  ,
+  H_i(t)
+  =
+  \frac{ h_i(1) + \cdots + h_i(t) } { t }
+  .
 \end{equation*}
 %
-can be written as a recurrence relation regarding
+This formula can written as a recurrence relation regarding
 the addition of the last data point,
@@
 \begin{align}
-  {\bar X}^{(N)}
+  H_i(t)
   &=
-  {\bar X}^{(N-1)} + \frac{ X^{(N)} - {\bar X}^{(N-1)} } { N }
-  .
+  H_i(t-1) + \frac{ h_i(t) - H_i(t-1) } { t }
+  ,
   \label{eq:av_recur}
@@
 %
-If we interpret $X^{(N)} - {\bar X}^{(N-1)}$
-as the independent correction to the previous average,
-${\bar X}^{(N-1)}$, proposed by the latest data point, $X^{(N)}$,
-then this correction is absorbed into the average with
-a diminishing weight $1/N$ as the sample size, $N$, increases.
-
-The structural similarity between the recurrence relation,
-Eq. \eqref{eq:av_recur},
-and the WL updating scheme, Eq. \eqref{eq:wl_update},
-suggests that the bias potential may represent
-a runtime average in a particular case.
-%
-To establish a mapping between the two,
-let us now consider a long equilibrium FDS simulation
-that consists of a sequence of sub-simulations
-of $L$ steps.
-%
-If $X^{(N)}$ is
-the bias potential, $V_i^{(N)}$, computed
-from the $N$th sub-simulation,
-then ${\bar X}^{(N)}$ means the average bias potential,
-${\bar V}_i^{(N)}$,
-from the first $N$ sub-simulations.
-%
-Since this average, ${\bar V}_i^{(N)}$,
-gives the best estimate at the end of
-the first $N$ sub-simulations,
-we should ideally use it as the bias potential
-for the next sub-simulation,
-until it is further updated by Eq. \eqref{eq:av_recur}
-at the end of the next sub-simulation.
-%
-Assuming that the normalized histogram, $H_i^{(N)}$,
-from the $N$th sub-simulation is sufficiently precise,
-Eq. \eqref{eq:vcorr_equil}
-\big[with
-$V_i^\mathrm{corr} \to V_i^{(N)}$, $V_i \to {\bar V}_i^{(N-1)}$\big]
-shows that the independent
-correction from the $N$th sub-simulation is
+%If we interpret $x(t) - X(t-1)$
+%as the independent correction to the previous average,
+%$X(t-1)$, proposed by the last data point, $x(t)$,
+%then this correction is absorbed into the average with
+%a diminishing weight $1/t$ as the sample size, $t$, increases.
+%
+which allows us to incrementally improve the estimated
+bias potential given by Eq. \eqref{eq:vcorr_equil}.
+%
+With
+$V_i^\mathrm{corr} \to V_i(t+1)$, $V_i \to V_i(1)$,
+we have
 %
 \begin{equation*}
-  \ln \frac{ H_i^{(N)} } { p_i }
+  V_i(t+1) = V_i(1) +
+  \ln \frac{ H_i(t) } { p_i }
   \approx
-  \frac{ H_i^{(N)} } { p_i } - 1
+  V_i(1) +
+  \frac{ H_i(t) } { p_i } - 1
   ,
 \end{equation*}
-%
-where the ``$-1$'' on the right-hand side
-can be ignored as a uniform shift
-that does not affect the distribution.
-%
-Thus, we may rewrite Eq. \eqref{eq:av_recur} as
+where we have assumed that
+the normalized histogram, $H_i(t)$,
+is sufficiently precise in the second step.
+%
+%where the ``$-1$'' on the right-hand side
+%can be ignored as a uniform shift
+%that does not affect the distribution.
+%
+Thus,
+%Since $H_i(t)$ is the average of the instantaneous histogram, $h_i(t)$,
+%we may apply Eq. \eqref{eq:av_recur}, and
+%
 \begin{align}
-  {\bar V}_i^{(N)}
+  V_i(t+1)
   &=
-  {\bar V}_i^{(N-1)}
+  V_i(t)
   +
-  \frac{1}{N} \frac{ H_i^{(N)} }{ p_i }
+  \frac{ H_i(t) - H_i(t-1) }{ p_i }
   \notag
@@
   &=
-  {\bar V}_i^{(N-1)}
+  V_i(t)
   +
-  \sum_{l=1}^L \frac{1}{N L} \frac{ h_i^{(N)}(l) }{ p_i }
-  ,
+  \frac{1}{t} \frac{ h_i(t) }{ p_i }
+  -\frac{1}{t} \frac{ H_i(t-1) }{ p_i }
+  .
   \label{eq:vbar_update}
@@
 %
-where $h_i^{(N)}(l)$ stands for the instantaneous histogram
-at step $l$ of sub-simulation $N$,
-and its average is
-$H_i^{(N)} \equiv \frac{1}{L} \sum_{l=1}^L h_i^{(N)}(l)$.
-%
-Now Eq. \eqref{eq:vbar_update} shows that the best estimate
-of the bias potential from the runtime average can be computed
-through a step-by-step incremental updating scheme that is equivalent to
-the WL algorithm, Eq. \eqref{eq:wl_update},
-with the updating magnitude being
-%
-\begin{equation}
-  \alpha(t) = \frac{1}{N L}
-  .
-  \label{eq:alpha_block}
-\end{equation}
-%
-In the long time limit,
-we may approximate the simulation time
-by the product of the number of sub-simulations
-and the number of steps in each sub-simulation, or $N L$,
-%
-and Eq. \eqref{eq:alpha_block} recovers Eq. \eqref{eq:alpha_invt}.
+Now if we drop the last term,
+which tends to a uniform shift of the bias potential
+in the asymptotic limit,
+Eq. \eqref{eq:vbar_update} recovers
+the single-bin updating scheme, Eq. \eqref{eq:wl_update},
+under
+the inverse-time schedule, Eq. \eqref{eq:alpha_invt}.
 
@@
 
-\subsubsection{\label{sec:nnscheme}
-Triple-bin updating scheme}
-
-If $\mu_2 = \dots = \mu_b = 0$,
-the matrix, $\mathbf w$, is tridiagonal,
-%%
-%\begin{equation}
-%\arraycolsep=3.6pt\def\arraystretch{1.4}
-%\mathbf w
-%=
-%\left(
-%  \begin{array}{cccccccc}
-%    1 - \mu_1   & \mu_1 & 0 & \dots & 0 \\
-%    \mu_1 & 1 - 2 \, \mu_1  & \mu_1 & \dots & 0 \\
-%    \vdots & &  & & \vdots \\
-%    0 & \dots & \mu_1 & 1 - 2 \, \mu_1  & \mu_1 \\
-%    0 & \dots & 0 & \mu_1 & 1 - \mu_1
-%  \end{array}
-%\right).
-%\label{eq:wnn}
-%\end{equation}
-%%
-%We have
-and the eigenvalues are
-\begin{equation}
-  \lambda_k
-  =
-  1 -
-  4 \, \mu_1 \sin^2
-  \frac{ (k - 1) \, \pi }
-       {       g \, n   }
-  .
-\notag
-%\label{eq:wnn_eigenvalue}
-\end{equation}
-%
-So this updating scheme is stable if
-$\min \lambda_k \ge 0$,
-or
-\begin{equation}
-  \mu_1 \le
-  \begin{dcases}
-    \frac 1 4
-    & \mathrm{periodic \; and \;} n \mathrm{ \; even,}
-    \\
-    \frac 1 4
-    \sec^2
-    \left( \frac{  \pi   }
-                { 2 \, n }
-    \right)
-    & \mathrm{otherwise}.
-  \end{dcases}
-\notag
-%\label{eq:nn_stable}
-\end{equation}
-
-\note{
-If the variable is periodic ($g = 1$) and $n$ is even,
-the minimum is achieved at $k = (n/2) + 1$.
-
-If the variable is periodic ($g = 1$) but $n$ is odd,
-the minimum is achieved at $k = (n+1)/2$.
-$$
-\sin^2\left(
-  \frac{n-1}{2 \, n} \pi
-\right)
-=
-\cos^2\left(
-  \frac{\pi}{2 \, n}
-\right)
-.
-$$
-
-If the variable is nonperiodic ($g = 2$)
-the minimum is achieved at $k = n$ (the maximal index).
-}
-
 \subsubsection{Gaussian updating scheme}
@@
 
-\section{\label{sec:upperbound}
-Asymptotic error for an updating scheme
-with a set of well-separated eigenvalues}
-
-We will derive here an upper bound of the asymptotic error
-that can be associated to updating schemes
-with a set of well-separated eigenvalues,
-in parallel to the lower bound reached by
-the single-bin scheme with a set of identical eigenvalues
-(cf. Sec. \ref{sec:optWL}).
-%
-For example, the smaller eigenvalues of the Gaussian updating scheme
-are well separated because of the rapid exponential decay.
-
-Starting from the definition Eq. \eqref{eq:mass_func}, we have
-$$
-  M(Q)
-  \le
-  \sum_{k = 2}^n
-  \sqrt{ \Gamma_k } \, \lambda_k \, e^{ -\lambda_k \, Q }
-  .
-$$
-Thus,
-it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
-%
-\begin{equation}
-  C
-  =
-  \int_0^\infty
-  M(q') \, dq'
-  \le
-  \sum_{k = 2}^n \sqrt{ \Gamma_k }
-  ,
-  \label{eq:mint_ubound}
-\end{equation}
-%
-and the asymptotic error from Eq. \eqref{eq:error_asym2}
-is bounded as
-\begin{equation}
-  \Err_A(T)
-  =
-  \frac{ C^2 } { T }
-  \le
-  \frac 1 T
-  \left(
-    \sum_{k=2}^n \sqrt{ \Gamma_k }
-  \right)^2
-  .
-  \label{eq:error_asym_ubound}
-\end{equation}
-%
-It is readily seen that this bound is always
-greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
-%
-$$
-\frac{T}{(T+t_0)^2} \, \sum_{k=2}^n \Gamma_k
-\le
-\frac{1}{T} \sum_{k=2}^n \Gamma_k
-\le
-\frac{1}{T} \left( \sum_{k=2}^n \sqrt{ \Gamma_k } \right)^2
-.
-$$
-
-Next, we will show that this bound represents
-updating schemes with a set of very different eigenvalues.
-%
-%First, observe that the summand of Eq. \eqref{eq:mass_func},
-%$\Gamma_k \, \left( \lambda_k e^{-\lambda_k \, Q} \right)^2$
-%is a function that peaks at $\lambda_k = 1/Q$,
-%so that the sum is dominated by the term at $k = k(Q)$ with
-%%
-%\begin{equation}
-%  \lambda_{k(Q)}
-%  \approx
-%  \frac 1 Q
-%  .
-%  \label{eq:kQ_def}
-%\end{equation}
-%
-%Thus,
-First, we have for any mode $k^*$ $(k^* \ge 2)$,
-%
-\begin{align}
-  M(Q)
-  &
-  \ge
-  %\approx
-  \sqrt{ \Gamma_{ k^* } } \,
-  \lambda_{ k^* } \,
-  e^{ -\lambda_{ k^* } \, Q }
-  ,
-\label{eq:MQ_lbound1}
-\end{align}
-%
-which would approach an equality if the $k^*$ mode
-dominates the mass function, $M(Q)$,
-as defined in Eq. \eqref{eq:mass_func}.
-%
-Then, for the $k$th component of the asymptotic error,
-we have from Eq. \eqref{eq:error_asym_Lagrangian}
-%
-\begin{align}
-  \left\langle
-  y_k^2(T)
-  \right\rangle_A
-  &=
-  \int_0^T
-  \Gamma_k \, \lambda_k^2 \, e^{2 \, \lambda_k \, [q(t) - q(T)] }
-  \, {\dot q}^2(t) \, dt
-  \notag
-  \\
-  &=
-  \frac{ C } { T }
-  \, \Gamma_k \, \lambda_k^2
-  \int_0^{q(T)}
-  \frac{ e^{-2 \, \lambda_k \, Q } } { M(Q) }
-  \,  \, dQ
-  \notag
-  \\
-  &\le
-  \frac C T \,
-  \frac{ \Gamma_k \, \lambda_k^2 } { \sqrt{ \Gamma_{k^*} } \, \lambda_{k^*} }
-  \int_0^{q(T)}
-    e^{-(2 \, \lambda_k - \lambda_{k^*} ) \, Q}
-    \, dQ
-  \notag
-  \\
-  &\le
-  \frac C T \,
-  \cdot
-  \frac{ \Gamma_k }
-  { \sqrt{ \Gamma_{k^*} } }
-  \cdot
-  \frac{ \lambda_k^2 }
-  { \lambda_{k^*} \, (2 \, \lambda_k -\lambda_{k^*}) }
-  ,
-  \label{eq:y2_ubound_k}
-\end{align}
-%
-where we have
-used Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:MQ_lbound1}
-on the second and third lines, respectively,
-and extended the upper limit to infinity on the last line.
-%
-For a set of well-separated eigenvalues,
-$\lambda_k$ depends on the mode index, $k$,
-much more critically than $\Gamma_k$.
-%
-Thus, the tightest upper bound
-can be approximately obtained from minimizing the last factor
-on the right-hand side of Eq. \eqref{eq:y2_ubound_k},
-and the minimal is reached at $k^* = k$,
-when $\lambda_{k^*} = \lambda_k$.
-%
-So
-%
-\begin{equation}
-  \left\langle
-  y_k^2(T)
-  \right\rangle_A
-  \le
-  %\approx
-  \frac C T
-  \sqrt{ \Gamma_k }
-  .
-  \label{eq:y2_ubound}
-\end{equation}
-%
-By summing over $k$
-and using Eq. \eqref{eq:mint_ubound},
-we recover Eq. \eqref{eq:error_asym_ubound}.
-
-This result shows that with a set of well-separated eigenvalues,
-the asymptotic error profile would be
-independent of $\lambda_k$'s and
-be flatter than the initial one from
-Eq. \eqref{eq:y2_eql}
-(cf. Fig. \ref{fig:xerr}
-for a Gaussian example).
-%
-Note, however, that to approach this upper bound
-the simulation has to be long enough
-such that the upper limit of the integral
-can be extended from $q(T)$ to infinity
-in the last step of Eq. \eqref{eq:y2_ubound_k}.
-%
-As a result,
-the flatter profile from Eq. \eqref{eq:y2_ubound}
-would stop at mode $k$,
-at which $\lambda_k \approx 1/q(T)$.
-%In other words,
-%the above asymptotic limit fails around
-%$Q^* \approx q(T)$, or equivalently
-%around mode $k^* = k\bigl( q(T) \bigr)$,
-%where $\lambda_{ k^* } \approx 1/q(T)$.
-%%
-%\note{This case also exemplifies
-%  the sensitivity to the simulation length.
-%  %
-%  In this limit,
-%  $M(Q) \propto 1/Q$ by Eq. \eqref{eq:MQ_lbound},
-%  and $C \propto \ln\bigl[ q(T) / q_0 \bigr]$
-%  for some $q_0$,
-%  and we expect that with fixed $a_0$,
-%  $q(T)$ to increase almost linearly with $T$
-%  by Eq. \eqref{eq:opt_qT}.
-%  %
-%  So the normalized asymptotic error,
-%  $E_A \cdot T \propto \bigl( \ln[ q(T)/q_0 ] \bigr)^2 \sim (\ln T)^2$,
-%  is sensitive to the simulation length,
-%  and worse than the ideal case of a constant.
-%}
-
-\section{\label{sec:invt_schedule}
-Modified inverse-time schedules}
-
-Here we compute the error of the schedule given by
-the modified inverse-time formula,
-Eq. \eqref{eq:alpha_invtlambda}.
-%As an alternative to the
-%optimal schedule from Eq. \eqref{eq:q_opt},
-%here we consider the modified
-%inverse-time formula, Eq. \eqref{eq:alpha_invtlambda},
-%with a free parameter $\lambda$.
-%
-%whose error can be computed explicitly.
-
-\subsection{\label{sec:invt_error}
-General results on error
-}
-
-Using Eq. \eqref{eq:alpha_invtlambda}
-in Eqs. \eqref{eq:error_asym}
-and \eqref{eq:error_res1}
-yields
+\section{\label{sec:sinvt_schedule}
+Scaled inverse-time schedule}
+
+The error of the scaled inverse-time schedule,
+Eq. \eqref{eq:alpha_invtlambda},
+can be found by using Eq. \eqref{eq:alpha_invtlambda}
+in Eqs. \eqref{eq:error_asym} and \eqref{eq:error_res1}:
 %
@@
 
-\subsection{\label{sec:invt_singlebin}
-  Single-bin updating scheme
-}
-
-For the single-bin scheme, we have
+In particular,
+for the single-bin scheme, we have
 $\lambda_2 = \cdots = \lambda_n = 1$,
@@
   where $r = t_0 / (T + t_0)$.
-}
-
-\subsection{\label{sec:invt_nn}
-Triple-bin updating scheme}
-
-For the triple-bin updating scheme,
-%introduced in Sec. \ref{sec:nnscheme},
-we will consider the large-$n$ limit.
-%
-If $2 \, \nu_k > 1$ for every $k$, we have
-%$\sum_i \to \frac{2 \, n}{\pi} \int_0^{\pi/2} dp$, and
-%
-\begin{align}
-  \Err_A(T)
-  =
-  \frac{ 2 \, n  }
-       { \pi \, T }
-  \int_0^{\pi/2}
-    \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
-         {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
-  \, dp
-  .
-\notag
-%\label{eq:error_nn_asym_int}
-\end{align}
-%
-For perfect sampling with $\mathbf p = \pmb \rho$,
-we have Eq. \eqref{eq:Gamma_perfect},
-and
-$$
-\begin{aligned}
-  \Err_A(T)
-  &=
-  \frac{n}{4 \, T}
-  \left(
-    \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
-    +
-    \frac{ \lambda }
-    { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
-  \right)
-,
-\end{aligned}
-$$
-%
-which is minimal at
-%
-\begin{equation}
-  \lambda
-  =
-  \frac{ 1 - 4 \, \mu_1 }
-       { 1 - 2 \, \mu_1 }
-  ,
-%\notag
-\label{eq:lambda_nn_perfect}
-\end{equation}
-%
-and at this point,
-%
-\begin{equation}
-  \Err_A(T)
-  =
-  \frac{n}{T}
-  \left(
-    1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
-  \right)
-  .
-  \label{eq:error_nn_perfect}
-\end{equation}
-%
-This shows that the original inverse-time formula,
-Eq. \eqref{eq:alpha_invt1},
-is not optimal unless $\mu_1 = 0$,
-the single-bin case,
-and the latter minimizes the asymptotic error,
-Eq. \eqref{eq:error_nn_perfect}.
-%
-Note that the value given by
-Eq. \eqref{eq:lambda_nn_perfect}
-was less than the observed optimal value
-from Fig. \ref{fig:errinvt},
-and the latter would decrease
-with the simulation length.
-%
-
-For the one-step sampling process,
-we find from Eq. \eqref{eq:Gamma_onestep}
-that $\Gamma(p) = \cot^2 p$
-peaks around $p \approx 0$,
-%
-Then
-$$
-  \Err_A(T)
-  \propto
-  \frac{   2 \, n }
-       { \pi \, T }
-  \frac{             1             }
-       {  \lambda \, (2 - \lambda) }
-  .
-$$
-%
-The optimal $\lambda$ is, therefore,
-$\lambda \approx 1$.
-%
-
-\note{
-However, the above result %Eq. \eqref{eq:lambda_nn_onestep}
-is only good
-for sufficiently short simulations,
-since we have taken the limit of $n \to \infty$
-for a fixed $T$.
-%
-For very long simulations,
-the optimal $\lambda$
-will ultimately drift to
-the smallest eigenvalue,
-$\min \lambda_k = 1 - 4 \, \mu_1$.
 }

### 2017-06-05 doc/invt.tex
solve a baseline shift problem in metad.h
invt.tex: shift the lambda index by 1

@@
 \newcommand{\ii}{\mathrm{i}}
+%bin average
+\newcommand{\bav}[1]{#1_\mathrm{av}}
 
@@
 %
-%To do so, we introduce a bias potential, $V_i$,
+%To do so, we introduce a bias potential, $u_i$,
 The challenge of sampling a multimodal distribution, $p^*_i$,
@@
 %to make $\pi_i$ flat,
-%the bias potential $V_i$
+%the bias potential $u_i$
 %must coincide with $\ln p^*_i$,
@@
 %%
-%In practice, $V_i$ may be non-optimal or contain error,
+%In practice, $u_i$ may be non-optimal or contain error,
 %and thus $\pi_i$ need not
@@
 %
-By introducing a bias potential, $V_i$,
+By introducing a bias potential, $u_i$,
 the biased equilibrium distribution becomes
@@
   %=
-  %C_i \, p^*_i \, e^{-V_i}
-  =
-  \frac{                p^*_i \, e^{-V_i} }
-       { \sum_{j = 1}^n p^*_j \, e^{-V_j} }
+  %C_i \, p^*_i \, e^{-u_i}
+  =
+  \frac{                p^*_i \, e^{-u_i} }
+       { \sum_{j = 1}^n p^*_j \, e^{-u_j} }
   \propto
-  p^*_i \, e^{-V_i}
+  p^*_i \, e^{-u_i}
   ,
@@
 \begin{equation}
-  V_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
+  u_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
   ,
@@
 and the PMF, $-\ln p^*_i$, can be readily
-deduced from the bias potential. % $V_i$.
+deduced from the bias potential. % $u_i$.
 \note{
@@
   $p_i$ is the weight $\zeta_i$,
-  and the bias potential, $V_i$, is $\ln (\tilde Z_i / \zeta_i)$.
+  and the bias potential, $u_i$, is $\ln (\tilde Z_i / \zeta_i)$.
   Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
@@
   mezei1987, berg1992, *lee1993, marinari1992, *lyubartsev1992},
-the bias potential, $V_i$, is
+the bias potential, $u_i$, is
 estimated a priori and fixed
@@
 %Eq. \eqref{eq:pi_p_V}
-%allows us to correct $V_i$ as
-%
-\begin{equation}
-  V^\mathrm{corr}_i
-  =
-  V_i
+%allows us to correct $u_i$ as
+%
+\begin{equation}
+  u^\mathrm{corr}_i
+  =
+  u_i
   +
@@
     H_i \approx \pi_i
-    &\propto p^*_i \, e^{-V_i},
+    &\propto p^*_i \, e^{-u_i},
     \\
     p_i
-    &\propto p^*_i \, e^{-V^\mathrm{corr}_i}.
+    &\propto p^*_i \, e^{-v^{0, \mathrm{corr}}_i}.
   \end{aligned}
@@
 In the original WL algorithm\cite{wang2001, *wang2001pre},
-the bias potential, $V_i(t)$, is updated
+the bias potential, $u_i(t)$, is updated
 at each sampling step $t$ as
@@
 \begin{equation}
-  V_i(t+1)
-  =
-  V_i(t)
+  u_i(t+1)
+  =
+  u_i(t)
   +
@@
 %However, as the above updates
-%change $V_i(t)$ continuously,
+%change $u_i(t)$ continuously,
 %the equilibrium distribution,
@@
 %%
-%First, since $V_i(t)$ is updated continuously,
+%First, since $u_i(t)$ is updated continuously,
 %there is a random noise that is proportional
@@
 We will first express the error of
-the bias potential, $V_i(t)$,
+the bias potential, $u_i(t)$,
 as a functional of $\alpha(t)$,
@@
   \equiv
-  V_i(t)
+  u_i(t)
   -
@@
 Note that the updates to $v_i(t)$ are
-equivalent to the updates to $V_i(t)$,
+equivalent to the updates to $u_i(t)$,
 since the shift
@@
 %
-Since a uniform increase of $V_i(t)$, or, equivalently, that of $v_i(t)$,
+Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
 does not affect the
@@
 depends only on the difference between $v_i(t)$
-and a baseline value, $\bar v(t)$,
+and a baseline value, $\bav{v}(t)$,
 for average growth\cite{
@@
 \begin{equation}
-  x_i(t) \equiv v_i(t) - \bar v(t)
+  x_i(t) \equiv v_i(t) - \bav{v}(t)
   %=
@@
 \begin{equation}
-  \bar v(t)
+  \bav{v}(t)
   = \sum_{i = 1}^n \rho_i \, v_i(t)
@@
 \begin{equation}
-  \dot V_i(t)
+  \dot u_i(t)
   \equiv
-  \frac{ dV_i(t) } { dt }
+  \frac{ d u_i(t) } { dt }
   =
@@
 %\begin{equation}
-%  \frac{ d \bar v (t) }
+%  \frac{ d \bav{v} (t) }
 %       { d t }
@@
             {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \dot v_i - \dot{ \bar v }
+  \dot v_i - \bav{\dot v}
   \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:vbar_def}}}
@@
 \int_0^T
-  \dot u\bigl( q(t) \bigr) \, \xi_i(t) \, dt,
+  \dot y\bigl( q(t) \bigr) \, \xi_i(t) \, dt,
 \notag
@@
 \begin{align}
-u(q')
+y(q')
 &\equiv
 e^{q' - q(T)}.
-\label{eq:u_def}
+\label{eq:y_def}
 \end{align}
@@
   \int_0^T
-    {\dot u}^2\bigl( q(t) \bigr) \,
+    {\dot y}^2\bigl( q(t) \bigr) \,
     \, dt
@@
 %
-This is equivalent to varying the curve $u(q(t))$
+This is equivalent to varying the curve $y(q(t))$
 with fixed endpoints,
-$u\bigl( q(0) \bigr)  = e^{- q(T)}$
+$y\bigl( q(0) \bigr)  = e^{- q(T)}$
 and
-$u\bigl( q(T) \bigr) = 1$.
+$y\bigl( q(T) \bigr) = 1$.
 %
@@
 while the second term is the action of a free particle.
-Thus, the velocity of $u(q(t))$ should be a constant:
+Thus, the velocity of $y(q(t))$ should be a constant:
 %
@@
 %%&=
-%%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
+%%\int_0^t \int_0^t \dot y(t') \, \dot y(t'')
 %%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
@@
 %\int_0^t \int_0^t
-%  \dot u\bigl( q(t') \bigr) \,
-%  \dot u\bigl( q(t'') \bigr) \,
+%  \dot y\bigl( q(t') \bigr) \,
+%  \dot y\bigl( q(t'') \bigr) \,
 %  \kappa(t' - t'') \, dt'' \, dt'.
@@
 %\frac{d}{dt} \int_0^T
-%  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
+%  \dot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
 %\\
 %&= \int_0^T
-%  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
+%  \ddot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
 %\end{aligned}
@@
 %$$
-%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
+%\int_0^t y^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
 %\qquad (n \ge 1)
@@
 %$
-%\ddot u\bigl( q(t') \bigr) = 0
+%\ddot y\bigl( q(t') \bigr) = 0
 %,
@@
 \begin{equation}
-  \dot u\bigl( q(t) \bigr) = c
+  \dot y\bigl( q(t) \bigr) = c
   .
-\label{eq:dudt_const}
-\end{equation}
-%
-Then, by using Eq. \eqref{eq:u_def}, we have
+\label{eq:dydt_const}
+\end{equation}
+%
+Then, by using Eq. \eqref{eq:y_def}, we have
 %
@@
 %
-Then, Eq. \eqref{eq:vt_diffeq} is generalized to
-%
-\begin{equation}
-  \dot V_i(t)
-  =
-  \alpha(t) \sum_{j=1}^n \, w_{ij} \frac{ h_j(t) } { p_j }.
-  \label{eq:vt_diffeq_mbin}
-\end{equation}
-
 The optimization procedure is similar to the single-bin case.
@@
 $\mathbf p = (p_1, \dots, p_n)$,
-${\dot V}_i(t)$
-in Eq. \eqref{eq:vt_diffeq_mbin}
+${\dot u}_i(t)$
+in Eq. \eqref{eq:ut_diffeq_mbin}
 should be a constant of $i$
@@
 \begin{align}
-  \sum_{k = 1}^n
+  \sum_{k = 0}^{n - 1}
     \phi_{ki} \, \phi_{kj}^*
@@
 %
-where the superscript $^*$
-denotes the complex conjugate.
+where the superscript $^*$ denotes the complex conjugate,
+and we have set the index of the first eigenmode to $0$
+instead of $1$ for later convenience.
 
@@
 \begin{equation}
-  \phi_{1i} = \rho_i,
-\label{eq:eigenmode1}
-\end{equation}
-%
-with $\lambda_1 = 1$.
-%
-Clearly, Eq. \eqref{eq:eigenmode1}
+  \phi_{0i} = \rho_i,
+\label{eq:eigenmode0}
+\end{equation}
+%
+with $\lambda_0 = 1$.
+%
+Clearly, Eq. \eqref{eq:eigenmode0}
 satisfies the normalization condition
 given by Eq. \eqref{eq:eig_orthonormal_rows}
-for $k = l = 1$,
-while the orthogonality for $k \ge 2$ demands
+for $k = l = 0$,
+while the orthogonality demands
 %
@@
   =
-  \sum_{ i = 1 }^n
-    \frac{ \phi_{ki} \, \phi_{1i}^* }
-         { \rho_i }
-  =
-  0
+  \delta_{k0}
   .
@@
 %
-We may write the updating matrix in terms of $\phi_{ki}$'s as
+We may write the updating matrix in $\phi_{ki}$'s
+as\cite{bussi2006}
 %
@@
   =
-  \frac{1}{\rho_i} \sum_{k=1}^n
+  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
   \lambda_k \, \phi_{ki}^* \, \phi_{kj}
@@
 which means that, as we will show, no eigenvalue, $\lambda_k$,
-can be negative.
+can be negative for $k \ge 1$.
 
@@
 
-We can now simplify Eq. \eqref{eq:vt_diffeq_mbin}.
-%
-By using Eqs.
-\eqref{eq:v_def}, \eqref{eq:vbar_def}, and
-\eqref{eq:vt_diffeq_mbin}-\eqref{eq:w_balance}
-we get
+We can further define
+a generalized Fourier transform, $\mathcal{F}$,
+for variable, $A_i$,
+%
+\begin{equation}
+  {\tilde A}_k
+  \equiv \mathcal{F}[A_i]_k
+  \equiv \sum_{i = 1}^n \phi_{ki} \, A_i
+  ,
+  \label{eq:gft_def}
+\end{equation}
+and the inverse transform is
+%
+\begin{equation}
+  A_i =
+  \frac{1}{\rho_i} \sum_{k = 1}^n \phi^*_{ki} \, \tilde{A}_k
+  .
+  \label{eq:gft_inv}
+\end{equation}
+%
+The multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
+can be diagonalized as
+%
+$$
+{\tilde v}_k(t + 1) =
+{\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
+\mathcal{F}\bigl[ h_i(t) / p_i \bigr]_k
+,
+$$
+or in the continuous-time form,
+%where
+%%
+%$$
+%\eta_i \equiv \frac{ h_i(t) } { p_i } - \sum_{j = 1}^n \rho_j \frac{ h_j(t) }{ p_j }.
+%$$
+%
+\begin{equation}
+  \dot{\tilde v}_k(t)
+  =
+  \alpha(t) \, \lambda_k \, \mathcal{F} \bigl[ h_j(t)/ p_i  \bigr]_k
+  .
+  \label{eq:ut_diffeq_mbin}
+\end{equation}
+
+In particular, for the first eigenmode,
+we have $\tilde{v}_0 \equiv v_\mathrm{av}$,
+and
 %
 \begin{align*}
-  \dot{ \bar v }(t)
+  \bav{\dot v}(t)
   =
@@
 %
-\note{Derivation.
-$$\begin{aligned}
-  \dot{ \bar v }
-  &
-  \stackrel{\mathrm{Eq.\; \eqref{eq:vbar_def}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \sum_{i = 1}^n \rho_i \, \dot v_i
-  %
-  \stackrel{\mathrm{Eq.\; \eqref{eq:vt_diffeq_mbin}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  %
-  \alpha(t)
-  \sum_{j = 1}^n \frac{ h_j } { p_j }
-                 \sum_{ i = 1 }^{n} \rho_i \, w_{ij}
-  \\
-  &
-  \stackrel{\mathrm{Eq.\; \eqref{eq:w_balance}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t)
-  \sum_{j = 1}^n \rho_j
-    \, \frac{ h_j } { p_j }
-  %
-  \stackrel{\mathrm{Eq.\; \eqref{eq:w_sumj}}}
-  {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t)
-  \sum_{j = 1}^n w_{ij}
-  \sum_{r = 1}^n \rho_r
-    \, \frac{ h_r } { p_r }
-  .
-\end{aligned}$$
-}
+from Eq. \eqref{eq:eigenmode0},
+%
 %
@@
 \begin{equation}
-  \dot y_k(t)
-  =
-  -\alpha(t) \, \lambda_k \, [y_k(t) - \theta_k(t)],
+  \dot{\tilde x}_k(t)
+  =
+  -\alpha(t) \, \lambda_k \,
+  \left[{\tilde x}_k(t) - {\tilde \xi}_k(t) \right]
+  .
   \notag
@@
 %
-where
-\begin{align}
-  y_k(t)
-  &\equiv
-  \sum_{j=1}^n
-    \phi_{kj} \, x_j(t)
-  ,
-  %\label{eq:y_def}
-\end{align}
-and
-\begin{align}
-  \theta_k(t)
-  &\equiv
-  \sum_{j=1}^n
-    \phi_{kj} \, \xi_j(t)
-  ,
-  \label{eq:theta_def}
-\end{align}
-%
-are the deviation and noise (histogram fluctuation)
-of mode $k$, respectively.
+%are the deviation and noise (histogram fluctuation)
+%of mode $k$, respectively.
 
@@
   =
-  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj}^* \,
+  \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
+   \phi_{ki} \, \phi_{kj}^* \,
     \left\langle x_i \, x_j \right\rangle
   =
-  \sum_{k = 1}^n
+  \sum_{k = 0}^{n - 1}
     \left\langle
-      \left| y_k \right|^2
+      \bigl| {\tilde x}_k^2 \bigr|
     \right\rangle
@@
 %
-From Eq. \eqref{eq:eigenmode1},
+From Eq. \eqref{eq:eigenmode0},
 the first eigenmode,
-$y_1 = \sum_{j=1}^n \rho_j \, x_j$,
+${\tilde x}_0 = \sum_{j=1}^n \rho_j \, x_j$,
 vanishes identically by Eq. \eqref{eq:rhox_sum}.
 %
-Thus, we can start the last sum %in Eq. \eqref{eq:y2_sum}
-from $k = 2$,
+Thus, we can start the last sum %in Eq. \eqref{eq:xt2_sum}
+from $k = 1$,
 %
@@
   =
-  \sum_{k = 2}^n
+  \sum_{k = 1}^{n - 1}
     \left\langle
-      \left| y_k(T) \right|^2
+      \bigl| {\tilde x}_k^2(T) \bigr|
     \right\rangle
   .
-\label{eq:y2_sum}
+\label{eq:xt2_sum}
 \end{equation}
@@
 In analogy to Eq. \eqref{eq:ET_average},
-%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:y2_sum},
+%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:xt2_sum},
 we can write the error after a period, $T$,
@@
   =
-  \sum_{k = 2}^n
-    \left\langle y_k^2(0) \right\rangle \,
+  \sum_{k = 1}^{n-1}
+    \left\langle
+      \left| {\tilde x}_k^2(0) \right|
+    \right\rangle \,
     e^{ - 2 \, \lambda_k  \, q(T) },
@@
   \int_0^T
-  \sum_{k = 2}^n
-  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt
+  \sum_{k = 1}^{n-1}
+  \Gamma_k \, \dot y_k^2\bigl( q(t) \bigr) \, dt
   ,
@@
 \begin{equation}
-  u_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
+  y_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
   ,
@@
 of the histogram fluctuation,
-$\kappa_k(t) \equiv \langle \theta_k^*(0) \, \theta_k(t) \rangle$,
+$\kappa_k(t) \equiv \langle {\tilde \xi}_k^*(0) \, {\tilde \xi}_k(t) \rangle$,
 as:
@@
   = \sum_{t = -\infty}^\infty
-  \left\langle \theta_k^*(0) \, \theta_k(t) \right\rangle
+  \left\langle {\tilde \xi}_k^*(0) \, {\tilde \xi}_k(t) \right\rangle
   .
@@
 Equation \eqref{eq:Gamma_sum}
-essentially maps the histogram fluctuation, $\theta_k(t)$,
+essentially maps the histogram fluctuation, ${\tilde \xi}_k(t)$,
 to the equivalent white noise
@@
 %  \int_0^T \int_0^T
-%  \sum_{k = 2}^n
-%    \dot u_k\bigl( q(t) \bigr) \,
-%    \dot u_k\bigl( q(t') \bigr) \,
+%  \sum_{k = 1}^{n-1}
+%    \dot y_k\bigl( q(t) \bigr) \,
+%    \dot y_k\bigl( q(t') \bigr) \,
 %    \kappa_k(t - t') \, dt \, dt'
@@
 %\begin{align*}
-%  u_k(q')
+%  y_k(q')
 %  &\equiv
@@
 %\begin{align}
-%  \left\langle
-%    y_k^2
-%  \right\rangle
+%  \left\langle \left|
+%    {\tilde x}_k^2
+%  \right| \right\rangle
 %  &=
@@
 %  .
-%\label{eq:y2_eql}
+%\label{eq:xt2_eql}
 %\end{align}
@@
   \left\langle
-    \left| y_k(0) \right|^2
+    \bigl| {\tilde x}_k^2(0) \bigr|
   \right\rangle
@@
   ,
-  \label{eq:y2_eql}
+  \label{eq:xt2_eql}
 \end{equation}
@@
 \begin{align*}
-  \left\langle y_k^2(T') \right\rangle
+  \left\langle \bigl|
+    {\tilde x}_k^2(T')
+  \bigr| \right\rangle
   =
@@
   =
-  \sum_{k = 2}^n
+  \sum_{k = 1}^{n - 1}
       \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k \,
@@
   &=
-  \sum_{ k = 2 }^n
-    \Gamma_k \, \dot u_k^2\bigl[ q(t) \bigr]
+  \sum_{ k = 1 }^{n - 1}
+    \Gamma_k \, {\dot y}_k^2\bigl[ q(t) \bigr]
   %\notag
@@
   =
-  %\sum_{ k = 2 }^n
-  %  \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
+  %\sum_{ k = 1 }^{n - 1}
+  %  \Gamma_k \, \lambda_k^2 \, y_k^2 \bigl[ q(t) \bigr]
   %\; \dot q^2( t )
@@
   \sqrt{
-    \textstyle\sum_{ k = 2 }^n \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
+    \textstyle\sum_{ k = 1 }^{n - 1}\Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
   }
@@
 This serves
-as a generalization of Eq. \eqref{eq:dudt_const}.
+as a generalization of Eq. \eqref{eq:dydt_const}.
 %
@@
   \mu_l
-  \cos
-  \frac{ (k - 1) \, l \, 2 \, \pi }
+  \cos\left(
+  \frac{ k \, l \, 2 \, \pi }
        {      g \, n        }
+  \right)
   ,
@@
 Basically, we need to extract
-$\lambda_i$'s and $\Gamma_i$'s
+$\lambda_k$'s and $\Gamma_k$'s
 from the given system and updating scheme (input),
@@
   \node[draw, text width={width("Eigenvalues,")}]
-    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_2, \dots, \lambda_n$};
-
-  \node[draw, text width={width("Integrals, ")}]
-    (gamma) at (2.2, 1) {Integrals, \\$\Gamma_2, \dots, \Gamma_n$};
+    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_1, \dots, \lambda_{n-1}$};
+
+  \node[draw, text width={width("Integrals, iii")}]
+    (gamma) at (2.1, 1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$};
 
@@
 \item
-Compute $\Gamma_k$'s from Eq. \eqref{eq:varY}
+Compute $\Gamma_k$'s from Eq. \eqref{eq:varXt}
     (cf. Appendix \ref{sec:Gamma_measure}).
@@
 Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}
-using e.g. the bisection method.
-%
-Since $m(Q)$ is a decreasing function of $Q$,
-we can start from an interval $[Q_L, Q_R]$
-that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
-(e.g. we can choose $Q_L = 0$,
-  and repeatedly double $Q_R$ until
-  the latter inequality is satisfied).
-%
-We can then iteratively refine the interval
-by substituting $Q_M = (Q_L + Q_R)/2$
-for either $Q_L$ and $Q_R$,
-depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
-until the width of the interval is less than
-the desired precision,
-%
-and set $q(T)$ to $Q_M$.
+    using e.g. the Newton-Raphson or bisection method\cite{press3rd}.
+%
+%Since $m(Q)$ is a decreasing function of $Q$,
+%we can start from an interval $[Q_L, Q_R]$
+%that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
+%(e.g. we can choose $Q_L = 0$,
+%  and repeatedly double $Q_R$ until
+%  the latter inequality is satisfied).
+%%
+%We can then iteratively refine the interval
+%by substituting $Q_M = (Q_L + Q_R)/2$
+%for either $Q_L$ and $Q_R$,
+%depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
+%until the width of the interval is less than
+%the desired precision,
+%%
+%and set $q(T)$ to $Q_M$.
 
@@
 %
-This can be done on an even grid of $N$ points
-at $q_{[i]} = \frac{i}{N} \, q(T)$
-($i = 0, 1, \dots, N$)
-using the standard trapezoidal rule\cite{press3rd}.
+%This can be done on an even grid of $N$ points
+%at $q_{[i]} = \frac{i}{N} \, q(T)$
+%($i = 0, 1, \dots, N$)
+We have used the standard trapezoidal rule\cite{press3rd},
+on $N+1$ grid points on $[0, q(T)]$ that are defined adaptively
+such that the resulting change of $t_{[i+1]}-t_{[i]}$
+is no less than $T/(2*N)$ and no greater than $2T/N$.
 %
@@
 and the updating magnitude can be obtained from numerical differentiation:
-$\alpha_i = (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
+$\alpha_i = 1 - \exp[-(q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})]$,
 with
@@
 \int_0^T
-  dt \sum_{k = 2}^n \Gamma_k \,
-    \left( \dot u_k \, \chi - c_k \right)^2
+  dt \sum_{k = 1}^{n-1} \Gamma_k \,
+    \left( {\dot y}_k \, \chi - c_k \right)^2
   \equiv
@@
   \int_0^T dt
-    \sum_{k = 2}^n
-      \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr)
+    \sum_{k = 1}^{n-1}
+      \Gamma_k \, {\dot y}_k^2\bigl( q(t) \bigr)
 \right)
@@
   \int_0^T dt
-    \sum_{k = 2}^n
+    \sum_{k = 1}^{n-1}
       \Gamma_k \, c_k^2
@@
   \int_0^T dt
-    \sum_{k = 2}^n
-      \Gamma_k \, c_k \, \dot u_k \, \bigl( q(t) \bigr)
+    \sum_{k = 1}^{n-1}
+      \Gamma_k \, c_k \, {\dot y}_k \, \bigl( q(t) \bigr)
 \right)^2
@@
 \left(
-  \sum_{k = 2}^n \Gamma_k \, c_k
+  \sum_{k = 1}^{n-1} \Gamma_k \, c_k
     \left[
@@
 The equality is achieved
-if $\dot u_k\left( q(t) \right) = c_k$
-for all $k = 2, \dots, n$ at any $t$,
+if $\dot y_k\left( q(t) \right) = c_k$
+for all $k > 0$ at any $t$,
 up to a multiplicative factor.
@@
 \note{Derivation.
-  Integrating $\dot u_k = c_k$ yields
-  \begin{equation}
-    u_k(t) = c_k \left(t + t_{k0} \right).
+  Integrating $\dot y_k = c_k$ yields
+  \begin{equation}
+    y_k(t) = c_k \left(t + t_{k0} \right).
   \tag{UK}
@@
   \end{equation}
-  Given that $u_k(T) = 1$ and $u_k(0) = e^{-\lambda_k \, q(T)}$,
+  Given that $y_k(T) = 1$ and $y_k(0) = e^{-\lambda_k \, q(T)}$,
   we have
@@
 %\begin{equation}
-$\lambda_2 = \dots = \lambda_n$.
+  $\lambda_1 = \dots = \lambda_{n-1}$.
 %  \notag
@@
 such that
-$u_k = (t + t_0) / (T + t_0)$,
+$y_k = (t + t_0) / (T + t_0)$,
 and
 $c_k = 1/(T + t_0)$
-for $k \ge 2$.
+for $k > 0$.
 %
@@
        { (T + t_0)^2 }
-  \sum_{ k = 2 }^n
+  \sum_{ k = 1 }^{n-1}
     \Gamma_k
@@
 %
-If we further assume $\lambda_2 = \cdots = \lambda_n = 1$,
+If we further assume $\lambda_1 = \cdots = \lambda_{n-1} = 1$,
 %the optimal schedule given by
@@
 a uniform shift of the bias potential,
-we can freely set $\lambda_1$ to be the same as $\lambda_2$
+we can freely set $\lambda_0$ to be the same as $\lambda_1$
 without changing the nature of the updating scheme.
@@
 that the PMF is sufficiently smooth
-such that the eigenmodes can be limited to $k \le K + 1$,
+such that the eigenmodes can be limited to $k \le K$,
 %
@@
 $$
-\lambda_1 = \cdots = \lambda_{K+1} = 1,
+\lambda_0 = \cdots = \lambda_{K-1} = 1,
 \mathrm{\; and \;}
-\lambda_{K+2} = \cdots = \lambda_n = 0.
+\lambda_K = \cdots = \lambda_{n-1} = 0.
 $$
@@
 %
-$$w_{ij} = n \sum_{k=1}^{K+1} \phi_{ki}^* \, \phi_{kj},$$
+$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki}^* \, \phi_{kj},$$
 %
@@
         { (T + t_0)^2 }
-  \sum_{ k = 2 }^{K+1}
+  \sum_{ k = 1 }^{K-1}
     \Gamma_k.
@@
 it recovers Eq. \eqref{eq:t0_singlebin}
-under the assumption of Eq. \eqref{eq:y2_eql}.
-%
-From Eq. \eqref{eq:y2_eql}, we have
-$$E(0) = \frac{1}{2} a_0 \sum_{k=2}^n \Gamma_k = \frac{1}{2} \, a_0 \, \Gamma.$$
+under the assumption of Eq. \eqref{eq:xt2_eql}.
+%
+From Eq. \eqref{eq:xt2_eql}, we have
+$$E(0) = \frac{1}{2} a_0 \sum_{k=1}^{n-1} \Gamma_k
+= \frac{1}{2} \, a_0 \, \Gamma.$$
 }
@@
        { T + t_0 }
-  \sum_{ k = 2 }^{ K + 1 }
+  \sum_{ k = 1 }^{ K - 1 }
     \Gamma_k
@@
 \begin{equation}
-  \sum_{k=2}^n \Gamma_k
+  \sum_{k=1}^{n-1} \Gamma_k
   = \sum_{i=1}^n \rho_i \, G_i = \Gamma
@@
 which can be shown by following the steps
-similar to those leading to Eq. \eqref{eq:y2_sum}.
+similar to those leading to Eq. \eqref{eq:xt2_sum}.
 %
@@
 $\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle
-=\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle$,
+=\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k^*(0) \rangle$,
 so
@@
 &=\sum_{t=-\infty}^\infty\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle \\
-&=\sum_{t=-\infty}^\infty\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle
-=\sum_{k=2}^n \Gamma_k.
+  &=\sum_{t=-\infty}^\infty\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k^*(0) \rangle
+  =\sum_{k=1}^{n-1} \Gamma_k.
 \end{align*}
@@
 we set the bias potential to zero,
-$V_i \equiv 0$,
+$u_i \equiv 0$,
 and equilibrate the system for $10^7$ steps
@@
 $
-A(i \to j) = \min\{ 1, \exp(V_i - V_j) \}.
+A(i \to j) = \min\{ 1, \exp(u_i - u_j) \}.
 $
@@
 
+\subsection{\label{sec:lj}Lennard-Jones}
+
+100 particle
+\begin{table}[h]\footnotesize
+  \caption{\label{tab:error_LJ}
+    Error of the LJ system
+    \note{\newline
+      The data were collected in
+      \texttt{data/tab1} by running
+      the program \texttt{invt}.
+    }%
+  }
+  \setlength{\tabcolsep}{2pt}
+  \renewcommand\arraystretch{1.2}
+  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
+    \hline
+      XXX
+    &
+    &
+    &
+    \\
+    \hline
+    Global
+    &
+    &
+    &
+    \\
+    \hline
+    Local
+    &
+    &
+    &
+    \\
+    \hline
+  \end{tabular}
+\end{table}
+
 \section{\label{sec:conclusion}
@@
 With
-$V_i^\mathrm{corr} \to V_i(t+1)$, $V_i \to V_i(1)$,
+$u^\mathrm{corr}_i \to u_i(t+1)$, $u_i \to u_i(1)$,
 we have
@@
 \begin{equation*}
-  V_i(t+1) = V_i(1) +
+  u_i(t+1) = u_i(1) +
   \ln \frac{ H_i(t) } { p_i }
   \approx
-  V_i(1) +
+  u_i(1) +
   \frac{ H_i(t) } { p_i } - 1
@@
 \begin{align}
-  V_i(t+1)
+  u_i(t+1)
   &=
-  V_i(t)
+  u_i(t)
   +
@@
   &=
-  V_i(t)
+  u_i(t)
   +
@@
 Thus,
-for $k \ge 2$, we get
+for $k > 0$, we get
 %
@@
 Eqs. \eqref{eq:ortho1},
-\eqref{eq:theta_def},
-and
 \eqref{eq:zz_perfect}.
@@
   we have used
-  Eq. \eqref{eq:theta_def}
-  on the first line,
-  %
   Eq. \eqref{eq:zz_perfect}
@@
 For an unknown system,
-we can use Eq. \eqref{eq:y2_eql} to
+we can use Eq. \eqref{eq:xt2_eql} to
 measure $\Gamma_k$'s in
@@
 cannot compute the deviation
-of the bias potential, $V_i$.
+of the bias potential, $u_i$.
 %
 To circumvent the problem, we
-modify Eq. \eqref{eq:y2_eql} as
-%
-\begin{equation}
-  \operatorname{var} Y_k
+modify Eq. \eqref{eq:xt2_eql} as
+%
+\begin{equation}
+  \operatorname{var} {\tilde X}_k
   =
@@
   a_0 \, \Gamma_k \, \lambda_k,
-\label{eq:varY}
+\label{eq:varXt}
 \end{equation}
@@
 \begin{align*}
-Y_k
+{\tilde X}_k
 &=
 \sum_{ i = 1 }^n
-  \phi_{k i} \, (V_i - \bar V),
-%\label{eq:Y_def}
+  \phi_{k i} \, (u_i - \bav{u}),
+%\label{eq:Xt_def}
 \end{align*}
@@
 $
-\bar V
+\bav{u}
 \equiv
-\sum_{ j = 1 }^n \rho_j \, V_j.
+\sum_{ j = 1 }^n \rho_j \, u_j.
 $
 
-To show Eq. \eqref{eq:varY},
+To show Eq. \eqref{eq:varXt},
 we first note that
-the difference, $V_i - {\bar V}$,
+the difference, $u_i - \bav{u}$,
 differs from the actual deviation, $x_i$,
@@
 It follows that the corresponding linear combinations,
-$Y_k$ and $y_k$, would also share the same variance.
-%
-But since the average of $y_k$ is zero,
+${\tilde X}_k$ and ${\tilde x}_k$, would also share the same variance.
+%
+But since the average of ${\tilde x}_k$ is zero,
 we have
 $$
-\operatorname{var} Y_k
-=
-\operatorname{var} y_k
-=
-\left\langle y_k^2 \right\rangle.
-$$
-Finally, by using Eq. \eqref{eq:y2_eql},
-we reach Eq. \eqref{eq:varY}.
-
-\note{In practice, Eq. \eqref{eq:varY}
+\operatorname{var} {\tilde X}_k
+=
+\operatorname{var} {\tilde x}_k
+=
+\left\langle {\tilde x}_k^2 \right\rangle.
+$$
+Finally, by using Eq. \eqref{eq:xt2_eql},
+we reach Eq. \eqref{eq:varXt}.
+
+\note{In practice, Eq. \eqref{eq:varXt}
 in inapplicable to zero eigenvalues, $\lambda_k$'s.
@@
   \exp
-  \left[
+  \left(
     \ii
-    \frac{ (k - 1) \, j \, 2 \, \pi }
+    \frac{ k \, j \, 2 \, \pi }
          {      n             }
-  \right]
+  \right)
   ,
@@
   {\cos \atop \sin}
-  \left[
-    \frac{ (k - 1) \, j \, 2 \, \pi }
+  \left(
+    \frac{ k \, j \, 2 \, \pi }
          {      n             }
-  \right]
+  \right)
   ,
@@
 %
-where the function takes the cosine form for $k \le n/2 + 1$,
+where the function takes the cosine form for $k \le n/2$,
 or the sine form otherwise.
@@
   and there is a two-fold degeneracy,
-  $\lambda_{n - k + 2} = \lambda_k$.
+  $\lambda_{n - k} = \lambda_k$.
 \note{Consider the unnormalized version
@@
   \exp\left[
-    \frac{ ( k - 1 ) \, j \, 2 \, \pi }
+    \frac{ k \, j \, 2 \, \pi }
          {      n             }
@@
       \mu_l \, \cos
-      \frac{ (k - 1) \, l \, 2 \, \pi }
+      \frac{ k \, l \, 2 \, \pi }
            {      n             }
@@
   =
-  \frac{ \sqrt{ 2 - \delta_{k, 1} } }
+  \frac{ \sqrt{ 2 - \delta_{k, 0} } }
        {             n              }
   \cos \left[
-       \frac{ ( k - 1) \, \left( j - \frac 1 2 \right) \, 2 \, \pi}
+       \frac{ k \, \left( j - \frac 1 2 \right) \, 2 \, \pi}
             {             2 \, n                           }
@@
   =
-  \cos \frac{ (k - 1) \, \left( j - \frac 1 2 \right) \, \pi }{n},
+  \cos \frac{ k \, \left( j - \frac 1 2 \right) \, \pi }{n},
   $$
@@
   \begin{align}
-    \lambda_{k+1}
+    \lambda_k
     =
     \sum_{ l = 0 }^{ 2 \, n - 1 }
-    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
+    \mu_l \, \cos \left(\frac{ k \, l \, \pi } { n } \right).
   \notag
@@
   %
-  This formula for $\lambda_{k+1}$
-  can be readily extended to $k = 2 \, n - 1$,
+  This formula for $\lambda_k$
+  can be readily extended to $k = 1 \, n - 1$,
   and we have
@@
   $
-    \lambda_{ 2 \, n + 2 - k } = \lambda_k
+    \lambda_{ 2 \, n - k } = \lambda_k
     .
@@
     \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \lambda_{ k + 1 } \,
+      \lambda_k \,
       \cos \frac{ k \, p \, \pi }
@@
   %
-  The value of $\lambda_{n + 1}$
+  The value of $\lambda_n$
   can be deduced from
@@
   \frac { 1 } { g \, n }
-  \sum_{ k = 1 }^{ g \, n }
+  \sum_{ k = 0 }^{ g \, n - 1 }
   \lambda_k
-  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
+  \cos \left(
+       \frac{ k \, l \, 2 \, \pi }
             {      g \, n        }
+  \right)
   .
@@
 we have defined
-$\lambda_k \equiv \lambda_{2 \, n + 2 - k}$
-for $k = n + 2, \dots, 2 \, n$,
+$\lambda_k \equiv \lambda_{2 \, n - k}$
+for $k = n + 1, \dots, 2 \, n - 1$,
 as well as
@@
 \begin{align}
-  \lambda_{ n + 1 }
+  \lambda_n
   =
   (-1)^{ n - 1 }
-  \lambda_1
+  \lambda_0
   +
-  2 \, \sum_{ k = 2 }^{ n }
-      (-1)^{n - k} \, \lambda_k
+  2 \, \sum_{ k = 1 }^{ n - 1 }
+      (-1)^{n - k - 1} \, \lambda_k
   ,
@@
     compute the eigenvalues,
-    $\lambda_1, \dots, \lambda_n$,
+    $\lambda_1, \dots, \lambda_{n-1}$,
     from Eq. \eqref{eq:wband_eigenvalue}.
@@
   2 \int_0^\pi
-    \mu(\varphi) \, \cos \bigl[ (k-1) \, \varphi \bigr] \, d\varphi,
+    \mu(\varphi) \, \cos (k \, \varphi) \, d\varphi,
 \notag
@@
   &\approx
-  \int_0^b \mu_l \, \cos[(k-1) \, \varphi] \, dl
+  \int_0^b \mu_l \, \cos(k \, \varphi) \, dl
   \\
   &=
-  \int_0^b \mu(\varphi) \, \cos[(k-1) \, \varphi] \, \Delta \varphi \, dl
+  \int_0^b \mu(\varphi) \, \cos(k \, \varphi) \, \Delta \varphi \, dl
   \\
   &=
-  \int_0^{b \, \Delta \varphi} \mu(\varphi) \, \cos[(k-1) \, \varphi] \, d\varphi
+  \int_0^{b \, \Delta \varphi} \mu(\varphi) \, \cos(k \, \varphi) \, d\varphi
 \end{align*}
@@
   &=
-  \exp\left[
-        -\frac{ (k - 1)^2 \, \sigma_\varphi^2 }
+  \exp\left(
+        -\frac{ k^2 \, \sigma_\varphi^2 }
               {           2             }
-      \right]
+      \right)
   %\notag
@@
 %&
-%\lambda_1 = \cdots = \lambda_{K+1}
+%\lambda_0 = \cdots = \lambda_{K-1}
 %= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
@@
 %&
-\lambda_{K+2} = \cdots = \lambda_{n-K} = 0,
+\lambda_{K} = \cdots = \lambda_{n-K} = 0,
 \end{aligned}
 $$
-for $K \ge 1$,
+for $K > 1$,
 with the rest of the eigenvalues being $1$.
@@
     \sin
-    \frac{ (2 K + 1) \, l \, \pi }
+    \frac{ (2 K - 1) \, l \, \pi }
          {              n        }
@@
 &=
-\frac 1 n \sum_{k = 0}^{n-1} \lambda_{k+1} \cos \frac{ k \, l \, 2 \, \pi } { n }
+\frac 1 n \sum_{k = 0}^{n-1} \lambda_k \cos \frac{ k \, l \, 2 \, \pi } { n }
 \\
@@
   1 +
-  \sum_{k=1}^K
+  \sum_{k=1}^{K-1}
   \cos \frac { k \, l \, 2 \, \pi } { n }
   +
-  \sum_{k=n-K}^{n-1}
+  \sum_{k=n-K+1}^{n-1}
   \cos \frac { k \, l \, 2 \, \pi } { n }
@@
 \frac 1 n
-\sum_{k=-K}^K
+\sum_{k=1-K}^{K-1}
 \cos \frac { k \, l \, 2 \, \pi } { n }
@@
 $$
-\lambda_1 = \cdots = \lambda_{K+1} = 1,
+\lambda_0 = \cdots = \lambda_{K-1} = 1,
 \qquad
-\lambda_{K+2} = \cdots = \lambda_n = 0.
+\lambda_K = \cdots = \lambda_{n-1} = 0.
 $$
 Then Eq. \eqref{eq:lambdan} gives
-$\lambda_{n+1} = (-1)^{n+K-1}$,
+$\lambda_n = (-1)^{n-K}$,
 and from Eq. \eqref{eq:mu_from_lambda},
@@
   =
-  \frac{ (-1)^{n+K+l-1} } { 2 \, n }
+  \frac{ (-1)^{n-K+l} } { 2 \, n }
     +
@@
     \sin
-    \frac{ (2 K + 1) \, l \, \pi }
+    \frac{ (2 K - 1) \, l \, \pi }
          {         2 \, n        }
@@
       \sin
-      \frac{ (2 K + 1) \, (i-j) \, \pi }
+      \frac{ (2 K - 1) \, (i-j) \, \pi }
            {         2 \, n        }
@@
       \sin
-      \frac{ (2 K + 1) \, (i+j-1) \, \pi }
+      \frac{ (2 K - 1) \, (i+j-1) \, \pi }
            {         2 \, n        }
@@
 Although the kernel is a sawtooth
-because of the term, $(-1)^{n+K+l-1}$,
+because of the term, $(-1)^{n-K+l}$,
 the ruggedness disappears
@@
 \begin{aligned}
-  \lambda_{n+1}
+  \lambda_{n}
   &=
@@
   \left[
-    \lambda_1
-    - 2 \, \lambda_2
-    + 2 \, \lambda_3 - \cdots
-    + (-1)^{K} 2 \, \lambda_{K+1})
+    \lambda_0
+    - 2 \, \lambda_1
+    + 2 \, \lambda_2 - \cdots
+    + (-1)^{K-1} 2 \, \lambda_{K-1})
   \right]
@@
   \begin{dcases}
-    (-1)^{n-1} & K \mathrm{\; even,} \\
-    (-1)^n     & K \mathrm{\; odd.}
+    (-1)^{n-1} & K \mathrm{\; odd,} \\
+    (-1)^n     & K \mathrm{\; even.}
   \end{dcases}
@@
     1 +
-    2 \sum_{k=2}^{K+1}
-    \cos \frac { (k - 1) \, l \pi } { n }
+    2 \sum_{k=1}^{K-1}
+    \cos \frac { k \, l \pi } { n }
     +
-    (-1)^{n+K-1} (-1)^l
+    (-1)^{n-K} (-1)^l
   \right]
@@
   \left[
-    (-1)^{n+K+l-1}
+    (-1)^{n-K+l}
     +
-    \sum_{k=-K}^{K}
+    \sum_{k=1-K}^{K-1}
     \cos \frac { k \, l \pi } { n }
@@
        { T + t_0 }
-  \sum_{k = 2}^n
+  \sum_{k = 1}^{n-1}
     \frac{ \Gamma_k \, \nu_k^2 }
@@
   \frac { a_0 } { 2 }
-  \sum_{k = 2}^n
+  \sum_{k = 1}^{n-1}
   \Gamma_k \, \lambda_k \,
@@
 %such that the values of
-%$\left\langle y_k^2(0) \right\rangle$
-%can be computed from Eq. \eqref{eq:y2_eql},
+%$\left\langle {\tilde x}_k^2(0) \right\rangle$
+%can be computed from Eq. \eqref{eq:xt2_eql},
 %as well as Eq. \eqref{eq:t0_sinc}.
@@
 %&=
-%\sum_{ k = 2 }^n
+%\sum_{ k = 1 }^{n-1}
 %  \frac
@@
 for the single-bin scheme, we have
-$\lambda_2 = \cdots = \lambda_n = 1$,
+$\lambda_1 = \cdots = \lambda_{n-1} = 1$,
 and
 $\Gamma_k$'s affect the error
-only via the sum, $\Gamma = \sum_{k = 2}^n \Gamma_k$,
+only via the sum, $\Gamma = \sum_{k = 1}^{n-1} \Gamma_k$,
 as a multiple.
@@
     $Q$             &   $q(T) - q(t)$.  \\
-    $\theta_k(t)$   &   Noise of mode $k$, Eq. \eqref{eq:theta_def}. \\
     $\rho_i$        &   Distribution used to define the error. \\
@@
     $T$             &   Simulation length. \\
-    $u_i(q')$       &   $u_i(q') \equiv e^{\lambda \, [q - q(T)]}$. \\
-    $V_i(t)$        &   Original bias potential. \\
+    $y_k(q')$       &   $y_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
+    $u_i(t)$        &   Original bias potential. \\
     $v_i(t)$        &   Shifted bias potential. \\
@@
     $\xi_i(t)$      &   The noise, Eq. \eqref{eq:xi_def}. \\
-    $y_k(t)$        &   Mode of the shifted bias potential. \\
-    $Y_k$           &   Mode of the original bias potential. \\
+    ${\tilde x}_k(t)$        &   Mode of the shifted bias potential. \\
+    ${\tilde X}_k$           &   Mode of the original bias potential. \\
     $z$             &   Reaction coordinate. \\

### 2017-06-06 doc/invt.tex
small changes

@@
 Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
-does not affect the
-resulting distribution, $\pi_i(t)$,
-%
-the real deviation, $x_i(t)$, of the bias potential
+does not affect the resulting distribution, $\pi_i(t)$,
+the real deviation of the bias potential
 depends only on the difference between $v_i(t)$
@@
 for average growth\cite{
-dama2014},
-%
-\begin{equation}
-  x_i(t) \equiv v_i(t) - \bav{v}(t)
-  %=
-  %v_i - \sum_{j = 1}^n \rho_j \, v_j
+dama2014}.
+%
+We define a bin-average-deducted bias potential as
+%
+\begin{equation}
+  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
   ,
@@
 %
-where
+where the bin average, $\bav{v}(t)$, is defined as
 %
@@
   \bav{v}(t)
-  = \sum_{i = 1}^n \rho_i \, v_i(t)
-  ,
-\label{eq:vbar_def}
+  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
+  ,
+\label{eq:vav_def}
 \end{equation}
@@
 %
-From Eqs. \eqref{eq:x_def} and \eqref{eq:vbar_def},
+From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
 we have
@@
 
-\subsubsection{Differential equation}
+\subsubsection{\label{sec:sbin_diffeq}
+Differential equation}
 
@@
   =
-  \alpha(t) \, \frac{ h_i(t) } { p_i },
+  \alpha(t) \, \frac{ h_i(t) } { p_i }
+  \equiv
+  \alpha(t) \, \eta_i(t)
+  ,
   \label{eq:vt_diffeq}
@@
 for the bin $i(t)$ at time $t$
-or zero otherwise.
+or zero otherwise,
+and we have defined a relative histogram
+$\eta_i(t) \equiv h_i(t) /p_i$
+in the last step.
 %
@@
 
-We can split $h_i(t)$ into a deterministic averaged part
+We now try to split the histogram into a deterministic averaged part
 and a random fluctuation part, $\zeta_i(t)$,
@@
 \begin{equation}
-  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
+  \eta_i(t) = \langle \eta_i(t) \rangle + \zeta_i(t).
   \label{eq:h_split}
@@
 \begin{align}
-  \langle h_i(t) \rangle
+  \langle \eta_i(t) \rangle
+  %&=
+  %\frac{ \langle h_i(t) } { p_i } \rangle
   &\approx
-  \pi_i(t)
-  =
-  \frac{                p_i \, e^{-x_i(t)} }
+  \frac{ \pi_i(t) } { p_i }
+  =
+  \frac{                       e^{-x_i(t)} }
        { \sum_{j = 1}^n p_j \, e^{-x_j(t)} }
@@
   &\approx
-  p_i \,
-  \left[
   1 - x_i(t) + \sum_{j = 1}^n p_j \, x_j(t)
-  \right]
-  ,
-  \label{eq:h_ave}
+  ,
+  \notag
+  %\label{eq:h_ave}
 \end{align}
@@
 }%
+%
+Deducting the bin average [similar to Eq. \eqref{eq:vav_def}],
+we have
+%
+\begin{align}
+  \langle \eta_i(t)  - \eta_\mathrm{av}(t) \rangle
+  \equiv
+  \langle \eta_i(t) \rangle -
+  \sum_{j=1}^n \rho_j \langle \eta_j(t) \rangle
+  &\approx
+  - x_i(t)
+  .
+  \label{eq:sh_ave}
+\end{align}
 
@@
 %
-and we will approximate it as white noise, i.e.
-%The noise is not necessarily white,
-%and the autocorrelation function
-%depend only on the time difference
-%
-\begin{equation}
-  \left\langle
-    \zeta_i(t) \, \zeta_j(t')
-  \right\rangle
-  =
-  %\sigma_{ij}(t - t')
-  G^{(0)}_{ij} \, \delta(t - t')
-  ,
-  \notag
-  %\label{eq:zeta_zeta_correlation}
-\end{equation}
-%
-for some value of $G^{(0)}_{ij}$,
-and $\delta(t)$ is the Dirac's delta function.
+and we will approximate it as white noise.
+%i.e.
+%%The noise is not necessarily white,
+%%and the autocorrelation function
+%%depend only on the time difference
+%%
+%\begin{equation}
+%  \left\langle
+%    \zeta_i(t) \, \zeta_j(t')
+%  \right\rangle
+%  =
+%  %\sigma_{ij}(t - t')
+%  G^{(0)}_{ij} \, \delta(t - t')
+%  ,
+%  \notag
+%  %\label{eq:zeta_zeta_correlation}
+%\end{equation}
+%%
+%for some value of $G^{(0)}_{ij}$,
 %where $\sigma_{ij}(t)$ is an even function
@@
 %
+The bin-average-deducted fluctuation
+%
+\begin{equation}
+  \xi_i(t)
+  \equiv
+  \zeta_i(t) - \zeta_\mathrm{av}(t)
+  \equiv
+  \zeta_i(t)
+  -
+  \sum_{ j = 1 }^n
+    \rho_j \, \zeta_j(t)
+  ,
+\label{eq:xi_def}
+\end{equation}
+%
+would also be governed by white noise as is $\zeta_i(t)$:
+%i.e. we may write
+\begin{equation}
+  \langle \xi_i(t) \, \xi_i(t') \rangle
+  = G_i \, \delta(t - t')
+  ,
+  \label{eq:G_def}
+\end{equation}
+for some $G_i$,
+and $\delta(t)$ is the Dirac's delta function.
+%
 For non-white noise,
-the equivalent value of $G^{(0)}_i$ can be approximately obtained as
-$\sum_{t = -\infty}^\infty \langle \zeta_i(t) \, \zeta_j(0) \rangle$.
+the equivalent value of $G_i$ can be approximately obtained as
+$\sum_{t = -\infty}^\infty \langle \xi_i(t) \, \xi_i(0) \rangle$.
 %
@@
 $
-  \langle \zeta_i(t) \, \zeta_j(t') \rangle
+  p_i p_j \langle \zeta_i(t) \, \zeta_j(t') \rangle
   =
@@
 
-%The average growth rate of $v(t)$ is given by
-%\begin{equation}
-%  \frac{ d \bav{v} (t) }
-%       { d t }
-%  =
-%  \alpha(t)
-%  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
-%  .
-%\label{eq:dvbardt}
-%\end{equation}
-%
-
 From Eqs.
-\eqref{eq:v_def}-\eqref{eq:vbar_def},
-\eqref{eq:rhox_sum}-\eqref{eq:h_ave},
+\eqref{eq:x_def}, \eqref{eq:vt_diffeq}-\eqref{eq:xi_def},
 %\eqref{eq:v_def},
 %\eqref{eq:x_def},
-%\eqref{eq:vbar_def},
+%\eqref{eq:vav_def},
 %\eqref{eq:rhox_sum},
@@
   =
+  \alpha(t) \, \left[ \eta_i(t) - \eta_\mathrm{av} \right]
+  \approx
   -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
   ,
-\label{eq:dxdt_singlebin}
-\end{equation}
-%
-where the shifted histogram fluctuation
-%
-\begin{equation}
-  \xi_i(t)
-  \equiv
-  \frac{ \zeta_i(t) }
-       { p_i }
-  -
-  \sum_{ j = 1 }^n
-    \rho_j
-    \frac{ \zeta_j(t) }
-         {     p_j    }
-  ,
-\label{eq:xi_def}
-\end{equation}
-%
-is also governed by white noise as is $\zeta_i(t)$:
-%i.e. we may write
-\begin{equation}
-  \langle \xi_i(t) \, \xi_i(t') \rangle
-  = G_i \, \delta(t - t')
-  ,
-  \label{eq:G_def}
-\end{equation}
-for some $G_i$.
-%
-\note{The derivation:
-$$
-\begin{aligned}
-  \dot x_i
-  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \dot v_i - \bav{\dot v}
-  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:vbar_def}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left(
-    \frac{ h_i } { p_i }
-    -
-    \sum_{ j = 1 }^n
-      \rho_j \frac{ h_j } { p_j }
-  \right)
-  \\
-  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_split} \; and \; \eqref{eq:xi_def}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left(
-    \frac{ \langle h_i \rangle } { p_i }
-    -
-    \sum_{ j = 1 }^n
-      \rho_j \frac{ \langle h_j \rangle } { p_j }
-    +
-    \xi_i(t)
-  \right)
-  \\
-  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_ave} \;and\; \eqref{eq:rhox_sum}}}
-            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
-  \alpha(t) \left[ -x_i(t) + \xi_i(t) \right].
-\end{aligned}
-$$
-}
-
-The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
+  \notag
+  %\label{eq:dxdt_singlebin}
+\end{equation}
+%
+and the formal solution is % of Eq. \eqref{eq:dxdt_singlebin} is
 %
@@
 for the error weighting distribution, $\pmb \rho$,
-introduced in Eq. \eqref{eq:vbar_def}.
+introduced in Eq. \eqref{eq:vav_def}.
 %
@@
   \sum_{k = 0}^{n - 1}
-    \phi_{ki} \, \phi_{kj}^*
+    \phi_{ki} \, \phi_{kj}
   &=
@@
   \sum_{i = 1}^n
-    \frac{ \phi_{ki} \, \phi_{li}^* }
+    \frac{ \phi_{ki} \, \phi_{li} }
          { \rho_i }
@@
 %
-where the superscript $^*$ denotes the complex conjugate,
-and we have set the index of the first eigenmode to $0$
+where we have set the index of the first eigenmode to $0$
 instead of $1$ for later convenience.
@@
   \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
-  \lambda_k \, \phi_{ki}^* \, \phi_{kj}
+  \lambda_k \, \phi_{ki} \, \phi_{kj}
   .
@@
 \end{equation}
+%
 and the inverse transform is
@@
   A_i =
-  \frac{1}{\rho_i} \sum_{k = 1}^n \phi^*_{ki} \, \tilde{A}_k
+  \frac{1}{\rho_i} \sum_{k = 1}^n \phi_{ki} \, \tilde{A}_k
   .
@@
 %
-The multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
-can be diagonalized as
+Using this definition,
+we can rewrite Eq. \eqref{eq:ortho1} as
+%
+\begin{equation}
+  \mathcal F[1]_k = \delta_{k0}
+  ,
+  \label{eq:ortho0}
+\end{equation}
+%
+and the bin average, Eq. \eqref{eq:vav_def} can be written as
+\begin{equation}
+  v_\mathrm{av} = \mathcal F[v_i]_0 = \tilde v_0
+  .
+  \label{eq:vav_vt0}
+\end{equation}
+
+We can now diagonalize
+the multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
+as
 %
@@
 {\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
-\mathcal{F}\bigl[ h_i(t) / p_i \bigr]_k
+{\tilde \eta}_k(t)
 ,
@@
 or in the continuous-time form,
-%where
-%%
-%$$
-%\eta_i \equiv \frac{ h_i(t) } { p_i } - \sum_{j = 1}^n \rho_j \frac{ h_j(t) }{ p_j }.
-%$$
 %
@@
   =
-  \alpha(t) \, \lambda_k \, \mathcal{F} \bigl[ h_j(t)/ p_i  \bigr]_k
+  \alpha(t) \, \lambda_k \, {\tilde \eta}_k(t)
   .
@@
 \end{equation}
-
-In particular, for the first eigenmode,
-we have $\tilde{v}_0 \equiv v_\mathrm{av}$,
-and
-%
-\begin{align*}
-  \bav{\dot v}(t)
-  =
-  \alpha(t)
-  \sum_{r = 1}^n \rho_r
-    \, \frac{ h_r(t) } { p_r }
-  =
-  \alpha(t)
-  \sum_{j = 1}^n w_{ij}
-  \sum_{r = 1}^n \rho_r
-    \, \frac{ h_r(t) } { p_r }
-  ,
-\end{align*}
-%
-from Eq. \eqref{eq:eigenmode0},
-%
-%
-and by Eqs. \eqref{eq:x_def}, \eqref{eq:rhox_sum} and \eqref{eq:h_ave},
+%
+Deducting the bin average, we get
 %
@@
 \begin{aligned}
-  \dot x_i(t)
-  &= \alpha(t) \sum_{j=1}^n w_{ij}
-  \left(
-    \frac{ h_j(t) } { p_j }
+  \dot {\tilde x}_k(t)
+  &= \alpha(t) \, \lambda_k
+  \left[
+    {\tilde \eta}_k(t)
     -
-    \sum_{r = 1}^n
-      \rho_r \frac{ h_r(t) } { p_r }
-  \right)
+    {\tilde \eta}_\mathrm{av}(t) \, \delta_{k0}
+  \right]
   \\
   &\approx
-  -\alpha(t) \sum_{j = 1}^n
-  w_{ij} \bigl[ x_j(t) - \xi_j (t) \bigr]
-  .
+  -\alpha(t) \, \lambda_k \,
+  \bigl[ {\tilde x}_k(t) - {\tilde \xi}_k (t) \bigr]
+  ,
 \end{aligned}
 $$
-%where
-%we have expanded the right-hand side
-%in linear terms of $x_j$
-%on the second line.
-
-Diagonalizing the matrix, $\mathbf w$,
-using the eigenvectors defined in Eqs.
-\eqref{eq:eig_w}-\eqref{eq:eig_orthonormal_rows}
-yields a set of decoupled equations
-%
-\begin{equation}
-  \dot{\tilde x}_k(t)
-  =
-  -\alpha(t) \, \lambda_k \,
-  \left[{\tilde x}_k(t) - {\tilde \xi}_k(t) \right]
-  .
-  \notag
-  %\label{eq:yt_diffeq}
-\end{equation}
-%
-%are the deviation and noise (histogram fluctuation)
-%of mode $k$, respectively.
+where we have used the linear approximation
+introduced in Sec. \ref{sec:sbin_diffeq}.
 
@@
   \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
-   \phi_{ki} \, \phi_{kj}^* \,
+   \phi_{ki} \, \phi_{kj} \,
     \left\langle x_i \, x_j \right\rangle
@@
     \left\langle
-      \bigl| {\tilde x}_k^2 \bigr|
+      {\tilde x}_k^2
     \right\rangle
@@
     \left\langle
-      \bigl| {\tilde x}_k^2(T) \bigr|
+      {\tilde x}_k^2(T)
     \right\rangle
@@
 of the histogram fluctuation,
-$\kappa_k(t) \equiv \langle {\tilde \xi}_k^*(0) \, {\tilde \xi}_k(t) \rangle$,
+$\kappa_k(t) \equiv \langle {\tilde \xi}_k(0) \, {\tilde \xi}_k(t) \rangle$,
 as:
@@
   = \sum_{t = -\infty}^\infty
-  \left\langle {\tilde \xi}_k^*(0) \, {\tilde \xi}_k(t) \right\rangle
+  \left\langle {\tilde \xi}_k(0) \, {\tilde \xi}_k(t) \right\rangle
   .
@@
   \left\langle
-    \bigl| {\tilde x}_k^2(0) \bigr|
+    {\tilde x}_k^2(0)
   \right\rangle
@@
 such that the resulting change of $t_{[i+1]}-t_{[i]}$
-is no less than $T/(2*N)$ and no greater than $2T/N$.
+is no less than $T/(2N)$ and no greater than $2T/N$.
 %
@@
 and the updating magnitude can be obtained from numerical differentiation:
-$\alpha_i = 1 - \exp[-(q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})]$,
+    $\alpha_i = [1 - e^{-\lambda_1 (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})}]/\lambda_1$,
 with
@@
 %
-$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki}^* \, \phi_{kj},$$
+$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj},$$
 %
@@
 $\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle
-=\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k^*(0) \rangle$,
+=\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k(0) \rangle$,
 so
@@
 &=\sum_{t=-\infty}^\infty\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle \\
-  &=\sum_{t=-\infty}^\infty\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k^*(0) \rangle
+  &=\sum_{t=-\infty}^\infty\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k(0) \rangle
   =\sum_{k=1}^{n-1} \Gamma_k.
@@
 along with the distribution, $\rho_i$, in Eqs.
-\eqref{eq:vbar_def} and \eqref{eq:error_def}.
+\eqref{eq:vav_def} and \eqref{eq:error_def}.
 %
@@
   \phi_{ki} \phi_{kj}
-  \langle \xi_i^*(t) \, \xi_j(t') \rangle
+  \langle \xi_i(t) \, \xi_j(t') \rangle
   %

### 2017-06-07 doc/invt.tex
small changes

@@
   =
-  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
+  \sum_{i = 1}^n \rho_i \,
+  \left\langle v_{*i}^2(t) \right\rangle
   .
@@
 %
-From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
-we have
-%
-\begin{equation}
-  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
-  .
-\label{eq:rhox_sum}
-\end{equation}
+%From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
+%we have
+%
+%\begin{equation}
+%  \sum_{i = 1}^n \rho_i \, v_{*i}(t) = 0
+%  .
+%\label{eq:rhox_sum}
+%\end{equation}
 %
@@
   \frac{ d u_i(t) } { dt }
-  =
+  \approx
   \alpha(t) \, \frac{ h_i(t) } { p_i }
@@
   ,
-  \label{eq:vt_diffeq}
+  \label{eq:ut_diffeq}
 \end{equation}
@@
 %
-$
-h_i(t) = \delta_{i, i(t)}
-$
+$h_i(t) = \delta_{i, i(t)}$
 %\begin{equation}
@@
 as the shift is a constant.
-
-We now try to split the histogram into a deterministic averaged part
+%
+Thus, we have
+%
+\begin{equation}
+  \dot v_{*i}(t)
+  \approx
+  \alpha(t) \, \eta_{*i}(t)
+  ,
+  \label{eq:vt_diffeq}
+\end{equation}
+%
+where we have defined the bin-average-deducted
+histogram, $\eta_{*i}(t)$, in a similar way to
+$v_{*i}(t)$ [cf. Eq. \eqref{eq:x_def}].
+
+We now split the histogram into
+a deterministic averaged part, $\langle \eta_i(t) \rangle$,
 and a random fluctuation part, $\zeta_i(t)$,
@@
 \begin{equation}
-  \eta_i(t) = \langle \eta_i(t) \rangle + \zeta_i(t).
+  \eta_i(t) =
+  \langle \eta_i(t) \rangle
+  +
+  \zeta_i(t)
+  .
   \label{eq:h_split}
@@
   =
-  \frac{                       e^{-x_i(t)} }
-       { \sum_{j = 1}^n p_j \, e^{-x_j(t)} }
+  \frac{                       e^{-v_{*i}(t)} }
+       { \sum_{j = 1}^n p_j \, e^{-v_{*j}(t)} }
   \notag
@@
   &\approx
-  1 - x_i(t) + \sum_{j = 1}^n p_j \, x_j(t)
+  1 - v_{*i}(t) + \sum_{j = 1}^n p_j \, v_{*j}(t)
   ,
@@
 where we have assumed the smallness
-of $x_i(t)$ in the linear expansion
-and used $\sum_{j=1}^n p_j = 1$
-on the second line.
+of $v_{*i}(t)$ in the linear expansion.
+%and used $\sum_{j=1}^n p_j = 1$
+%on the second line.
 %
@@
 \approx
-\frac{                       1 - x_i  }
-     { \sum_{ r = 1 }^n p_j (1 - x_j) }
-=
-\frac{                       1 - x_i  }
-     { 1 - \sum_{ j = 1 }^n p_j \, x_j }
+\frac{                       1 - v_{*i}  }
+     { \sum_{ r = 1 }^n p_j (1 - v_{*j}) }
+=
+\frac{                       1 - v_{*i}  }
+     { 1 - \sum_{ j = 1 }^n p_j \, v_{*j} }
 .
@@
 %
-Deducting the bin average [similar to Eq. \eqref{eq:vav_def}],
-we have
+Deducting the bin average, we get
 %
 \begin{align}
-  \langle \eta_i(t)  - \eta_\mathrm{av}(t) \rangle
+  \langle \eta_{*i}(t) \rangle
   \equiv
@@
   &\approx
-  - x_i(t)
+  - v_{*i}(t)
   .
@@
 
-%
-For the histogram fluctuation part,
-the average is zero by definition,
-$\langle \zeta_i(t) \rangle = 0$,
-%
-and we will approximate it as white noise.
-%i.e.
-%%The noise is not necessarily white,
-%%and the autocorrelation function
-%%depend only on the time difference
+We will approximate the histogram fluctuation part, $\zeta_i(t)$,
+as white noise,
+and assume its validity for long simulations.
+%
+The bin-average-deducted fluctuation, $\zeta_{*i}(t)$,
 %%
 %\begin{equation}
-%  \left\langle
-%    \zeta_i(t) \, \zeta_j(t')
-%  \right\rangle
-%  =
-%  %\sigma_{ij}(t - t')
-%  G^{(0)}_{ij} \, \delta(t - t')
+%  \zeta_{*i}(t)
+%  \equiv
+%  \zeta_i(t)
+%  -
+%  \sum_{ j = 1 }^n
+%    \rho_j \, \zeta_j(t)
 %  ,
 %  \notag
-%  %\label{eq:zeta_zeta_correlation}
+%  %\label{eq:xi_def}
 %\end{equation}
 %%
-%for some value of $G^{(0)}_{ij}$,
-%where $\sigma_{ij}(t)$ is an even function
-%that vanishes at large $t$.
-%
-This approximation is valid if the simulation period, $T$,
-is much longer than the autocorrelation time of noise.
-%
-The bin-average-deducted fluctuation
-%
-\begin{equation}
-  \xi_i(t)
-  \equiv
-  \zeta_i(t) - \zeta_\mathrm{av}(t)
-  \equiv
-  \zeta_i(t)
-  -
-  \sum_{ j = 1 }^n
-    \rho_j \, \zeta_j(t)
-  ,
-\label{eq:xi_def}
-\end{equation}
-%
-would also be governed by white noise as is $\zeta_i(t)$:
-%i.e. we may write
-\begin{equation}
-  \langle \xi_i(t) \, \xi_i(t') \rangle
+would also be governed by white noise as is $\zeta_i(t)$, i.e.,
+\begin{equation}
+  \langle \zeta_{*i}(t) \, \zeta_{*i}(t') \rangle
   = G_i \, \delta(t - t')
@@
 the equivalent value of $G_i$ can be approximately obtained as
-$\sum_{t = -\infty}^\infty \langle \xi_i(t) \, \xi_i(0) \rangle$.
+$\sum_{t = -\infty}^\infty \langle \zeta_{*i}(t) \, \zeta_{*i}(0) \rangle$.
 %
@@
 From Eqs.
-\eqref{eq:x_def}, \eqref{eq:vt_diffeq}-\eqref{eq:xi_def},
-%\eqref{eq:v_def},
-%\eqref{eq:x_def},
-%\eqref{eq:vav_def},
-%\eqref{eq:rhox_sum},
-%\eqref{eq:vt_diffeq},
-%\eqref{eq:h_split}
-%\eqref{eq:h_ave},
-%\eqref{eq:dvbardt},
+\eqref{eq:vt_diffeq}-\eqref{eq:sh_ave},
 we get % a set of decoupled equations
@@
 \begin{equation}
-  \dot x_i(t)
-  =
-  \alpha(t) \, \left[ \eta_i(t) - \eta_\mathrm{av} \right]
+  \dot v_{*i}(t)
+  =
+  \alpha(t) \, \left[ \langle \eta_{*i}(t) \rangle + \zeta_{*i}(t) \right]
   \approx
-  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
+  -\alpha(t) \, \left[ v_{*i}(t) - \zeta_{*i}(t) \right]
   ,
@@
 \begin{equation}
-x_i(T)
-=
-x_i(0) \, e^{-q(T)}
+  v_{*i}(T)
+=
+  v_{*i}(0) \, e^{-q(T)}
 +
 \int_0^T
-  \dot y\bigl( q(t) \bigr) \, \xi_i(t) \, dt,
+  \dot y\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
 \notag
@@
 or, equivalently, the curve, $q(t)$,
-via $x_i(T)$, and
+via $v_{*i}(T)$, and
 our aim is to find the $q(t)$
@@
 $\mathbf p = (p_1, \dots, p_n)$,
+the growth rate of the bias potential,
 ${\dot u}_i(t)$
-in Eq. \eqref{eq:ut_diffeq_mbin}
+in Eq. \eqref{eq:ut_diffeq_mbin},
 should be a constant of $i$
@@
   ,
-  \label{eq:gft_def}
+  \notag
+  %\label{eq:gft_def}
 \end{equation}
@@
   .
-  \label{eq:gft_inv}
+  \notag
+  %\label{eq:gft_inv}
 \end{equation}
@@
 we can rewrite Eq. \eqref{eq:ortho1} as
-%
-\begin{equation}
-  \mathcal F[1]_k = \delta_{k0}
-  ,
-  \label{eq:ortho0}
-\end{equation}
-%
-and the bin average, Eq. \eqref{eq:vav_def} can be written as
-\begin{equation}
-  v_\mathrm{av} = \mathcal F[v_i]_0 = \tilde v_0
-  .
-  \label{eq:vav_vt0}
-\end{equation}
-
-We can now diagonalize
+$\mathcal F[1]_k = \delta_{k0}$,
+and for the bin average of any quantity, $f$, we have
+$\bav{f}= \mathcal F[f_i]_0 = \tilde f_0$,
+and
+\begin{equation}
+  \tilde f_{*k}
+  = \tilde f_k - \bav{f} \, \delta_{k0}
+  =
+  \begin{dcases}
+    0           & k = 0, \\
+    \tilde f_k  & k > 0,
+  \end{dcases}
+  \label{eq:ftstar}
+\end{equation}
+which means that the Fourier transform of
+a bin-average-deducted quantity always
+has a vanishing first mode.
+
+We can now use Eq. \eqref{eq:eig_w} to diagonalize
 the multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
@@
 %
-Deducting the bin average, we get
-%
+Deducting the bin average yields
 $$
-\begin{aligned}
-  \dot {\tilde x}_k(t)
-  &= \alpha(t) \, \lambda_k
-  \left[
-    {\tilde \eta}_k(t)
-    -
-    {\tilde \eta}_\mathrm{av}(t) \, \delta_{k0}
-  \right]
-  \\
-  &\approx
+  \dot {\tilde v}_{*k}(t)
+  = \alpha(t) \, \lambda_k \, {\tilde \eta}_{*k}(t)
+  \approx
   -\alpha(t) \, \lambda_k \,
-  \bigl[ {\tilde x}_k(t) - {\tilde \xi}_k (t) \bigr]
-  ,
-\end{aligned}
+  \bigl[ {\tilde v}_{*k}(t) - {\tilde \zeta}_{*k} (t) \bigr]
+  ,
 $$
 where we have used the linear approximation
-introduced in Sec. \ref{sec:sbin_diffeq}.
+as in Sec. \ref{sec:sbin_diffeq}.
 
@@
 \begin{align}
-  \Err
-  \equiv
-  \sum_{i = 1}^n \rho_i \,
-    \left\langle x_i^2 \right\rangle
+  \Err(T)
   =
   \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
-   \phi_{ki} \, \phi_{kj} \,
-    \left\langle x_i \, x_j \right\rangle
+   \phi_{ki} \, \phi_{kj}
+    \left\langle v_{*i}(T) \, v_{*j}(T) \right\rangle
   =
@@
     \left\langle
-      {\tilde x}_k^2
+      {\tilde v}_{*k}^2(T)
     \right\rangle
@@
 %
-From Eq. \eqref{eq:eigenmode0},
-the first eigenmode,
-${\tilde x}_0 = \sum_{j=1}^n \rho_j \, x_j$,
-vanishes identically by Eq. \eqref{eq:rhox_sum}.
-%
-Thus, we can start the last sum %in Eq. \eqref{eq:xt2_sum}
+By Eq. \eqref{eq:ftstar},
+we can start the last sum %in Eq. \eqref{eq:xt2_sum}
 from $k = 1$,
@@
     \left\langle
-      {\tilde x}_k^2(T)
+      {\tilde v}_{*k}^2(T)
     \right\rangle
@@
     \left\langle
-      \left| {\tilde x}_k^2(0) \right|
+      {\tilde v}_{*k}^2(0)
     \right\rangle \,
@@
 of the histogram fluctuation,
-$\kappa_k(t) \equiv \langle {\tilde \xi}_k(0) \, {\tilde \xi}_k(t) \rangle$,
+$\kappa_k(t) \equiv \langle {\tilde \zeta}_{*k}(0)
+\, {\tilde \zeta}_{*k}(t) \rangle$,
 as:
@@
   = \sum_{t = -\infty}^\infty
-  \left\langle {\tilde \xi}_k(0) \, {\tilde \xi}_k(t) \right\rangle
+  \left\langle {\tilde \zeta}_{*k}(0)
+            \, {\tilde \zeta}_{*k}(t) \right\rangle
   .
@@
 %
-Equation \eqref{eq:Gamma_sum}
-essentially maps the histogram fluctuation, ${\tilde \xi}_k(t)$,
-to the equivalent white noise
+We may understand Eq. \eqref{eq:Gamma_sum}
+as an attempt of modeling
+a general histogram fluctuation, ${\tilde \zeta}_{*k}(t)$,
+as an equivalent white noise
 so that the theory can be applicable to
@@
 
-%\subsubsection{\label{sec:eqlerr}
-%Initial error
-%%from a simulation under a constant updating magnitude
-%}
+\subsubsection{\label{sec:eqlerr}
+Initial error
+%from a simulation under a constant updating magnitude
+}
 
@@
 %  \left\langle \left|
-%    {\tilde x}_k^2
+%    {\tilde v}_{*k}^2
 %  \right| \right\rangle
@@
 we will assume that
-the adaptive FDS sampling
-was performed under a constant updating magnitude
-for sufficiently long.
-%
-Then, one can show that
+the preliminary adaptive FDS sampling has been
+performed under a constant updating magnitude for $T_0$, steps.
+%
+If the starting bias potential is zero, $u_i = 0$,
+one can show that
 \begin{equation}
   \left\langle
-    {\tilde x}_k^2(0)
+    {\tilde v}_{*k}^2(0)
   \right\rangle
-  = \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
+  + \tilde u_k^2 \, e^{-2\, \lambda_k \, T_0}
+  ,
+  \label{eq:xt2_eql1}
+\end{equation}
+%
+and if no eigenvalue is zero,
+then we can ignore the second term
+in the case of $T_0 \to \infty$,
+%
+\begin{equation}
+  \left\langle
+    {\tilde v}_{*k}^2(0)
+  \right\rangle
+  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
   ,
@@
 \begin{align*}
-  \left\langle \bigl|
-    {\tilde x}_k^2(T')
-  \bigr| \right\rangle
-  =
-  \int_0^{T'}
+  \left\langle
+    {\tilde v}_{*k}^2(T_0)
+  \right\rangle
+  =
+  \int_0^{T_0}
     \Gamma_k \, (\lambda_k \, a_0)^2 \,
-      e^{ 2 \, \lambda_k \, a_0 \, (t - T') }
+      e^{ 2 \, \lambda_k \, a_0 \, (t - T_0) }
     \, dt
-  \stackrel{ T' \to \infty }
+  \stackrel{ T_0 \to \infty }
   { =\joinrel=\joinrel=\joinrel= }
@@
   \sum_{k = 1}^{n - 1}
-      \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k \,
+    \left\langle
+      {\tilde v}_{*k}^2(0)
+    \right\rangle
       e^{  - 2 \, \lambda_k \, q(T) }
@@
 %
-This expression excludes systematic bias
-and thus serves as a lower bound for a realistic simulation.
 
@@
 From Eq. \eqref{eq:eig_orthonormal_cols}, we have
-$\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle
-=\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k(0) \rangle$,
+$\sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle
+=\sum_{k=1}^{n-1} \langle {\tilde v}_{*k}(t) \, {\tilde v}_{*k}(0) \rangle$,
 so
@@
 \sum_{i=1}^n \rho_i G_i
-&=\sum_{t=-\infty}^\infty\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle \\
-  &=\sum_{t=-\infty}^\infty\sum_{k=1}^{n-1} \langle {\tilde x}_k(t) \, {\tilde x}_k(0) \rangle
+&=\sum_{t=-\infty}^\infty\sum_{i=1}^n
+  \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle \\
+  &=\sum_{t=-\infty}^\infty\sum_{k=1}^{n-1}
+  \langle {\tilde v}_{*k}(t) \, {\tilde v}_{*k}(0) \rangle
   =\sum_{k=1}^{n-1} \Gamma_k.
@@
 $h_i(t) \approx \zeta_i(t)$,
-and $\xi_i(t)$ defined in Eq. \eqref{eq:xi_def}
+and $\zeta_{*i}(t)$
 will dominate the shifted instantaneous histogram
@@
 \begin{equation}
-\xi_i(t) \approx
-\frac{ h_i(t) } { p_i }
-- \sum_{r = 1}^n \rho_r \frac{ h_r(t) } { p_r }
+\zeta_{*i}(t) \approx
+\eta_{*i}(t)
 .
@@
   %\frac{1}{T}
-  %\sum_{t = 1}^T \xi_i(t)
+  %\sum_{t = 1}^T \zeta_{*i}(t)
   %\right)^2
@@
   \left\langle
-  \xi_i(t) \, \xi_i(t')
+  \zeta_{*i}(t) \, \zeta_{*i}(t')
   \right\rangle
@@
 $\langle h_i(t) \rangle = p_i$,
-we find from Eqs. \eqref{eq:h_split} and \eqref{eq:xi_def} that
+we find from Eq. \eqref{eq:h_split} that
 \begin{equation*}
-  \xi_i(t)
+  \zeta_{*i}(t)
   =
@@
   \left\langle
-    \xi_i(t) \, \xi_j(t')
+    \zeta_{*i}(t) \, \zeta_{*j}(t')
   \right\rangle
@@
   \left\langle
-    \xi_i(t)
+    \zeta_{*i}(t)
   \right\rangle
   \left\langle
-    \xi_j(t')
+    \zeta_{*j}(t')
   \right\rangle
@@
   \phi_{ki} \phi_{kj}
-  \langle \xi_i(t) \, \xi_j(t') \rangle
+  \langle \zeta_{*i}(t) \, \zeta_{*j}(t') \rangle
   %
@@
   \phi_{kj}
-  \langle \xi_i(t) \, \xi_j(t') \rangle
+    \langle \zeta_{*i}(t) \, \zeta_{*j}(t') \rangle
   \\
@@
 we first note that
-the difference, $u_i - \bav{u}$,
-differs from the actual deviation, $x_i$,
+the difference, $u_{*i} = u_i - \bav{u}$,
+differs from the actual deviation, $v_{*i}$,
 defined in Eq. \eqref{eq:x_def},
@@
 It follows that the corresponding linear combinations,
-${\tilde X}_k$ and ${\tilde x}_k$, would also share the same variance.
-%
-But since the average of ${\tilde x}_k$ is zero,
+${\tilde V}_{*k}$ and ${\tilde v}_{*k}$, would also share the same variance.
+%
+But since the average of ${\tilde v}_{*k}$ is zero,
 we have
-$$
-\operatorname{var} {\tilde X}_k
-=
-\operatorname{var} {\tilde x}_k
-=
-\left\langle {\tilde x}_k^2 \right\rangle.
-$$
+\begin{equation*}
+\operatorname{var} {\tilde V}_{*k}
+=
+\operatorname{var} {\tilde v}_{*k}
+=
+\left\langle {\tilde v}_{*k}^2 \right\rangle.
+\end{equation*}
 Finally, by using Eq. \eqref{eq:xt2_eql},
@@
 %such that the values of
-%$\left\langle {\tilde x}_k^2(0) \right\rangle$
+%$\left\langle {\tilde v}_{*k}^2(0) \right\rangle$
 %can be computed from Eq. \eqref{eq:xt2_eql},
@@
     $\mathbf w$     &   Updating matrix. \\
-    $x_i(t)$        &   Average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
-    $\xi_i(t)$      &   The noise, Eq. \eqref{eq:xi_def}. \\
-    ${\tilde x}_k(t)$        &   Mode of the shifted bias potential. \\
-    ${\tilde X}_k$           &   Mode of the original bias potential. \\
+    $v_{*i}(t)$        &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
+    $\zeta_{*i}(t)$      &   The noise of histogram. \\
+    ${\tilde v}_{*k}(t)$        &   Mode of the shifted bias potential. \\
+    ${\tilde V}_{*k}$           &   Mode of the original bias potential. \\
     $z$             &   Reaction coordinate. \\

### 2017-06-08 doc/invt.tex
invt.tex: cut the results section

@@
   ,
-  \label{eq:v_def}
+  \notag
+  %\label{eq:v_def}
 \end{equation}
@@
 %
-We define a bin-average-deducted bias potential as
+We define a bin-average-shifted bias potential as
 %
@@
 %
-The total error is given by
-%
-\begin{equation}
-  \Err(t)
-  =
-  \sum_{i = 1}^n \rho_i \,
-  \left\langle v_{*i}^2(t) \right\rangle
-  .
-\label{eq:error_def}
-\end{equation}
-
-%Although the above error weighting distribution, $\pmb\rho$,
-%is usually flat,
-%i.e. $\rho_i \equiv 1/n$,
-%in practice,
-%a non-flat distribution
-%is sometimes more convenient in analyzing
-%some multiple-bin updating schemes later on.
+Below we will use the notations, $f_{*i}$ and $\bav{f}$,
+for an arbitrary variable, $f_i$.
 %
@@
 whereas $\pmb\rho$ is flat.}
+
+The total error is given by
+%
+\begin{equation}
+  \Err(t)
+  =
+  \sum_{i = 1}^n \rho_i \,
+  \left\langle v_{*i}^2(t) \right\rangle
+  .
+\label{eq:error_def}
+\end{equation}
+
+%Although the above error weighting distribution, $\pmb\rho$,
+%is usually flat,
+%i.e. $\rho_i \equiv 1/n$,
+%in practice,
+%a non-flat distribution
+%is sometimes more convenient in analyzing
+%some multiple-bin updating schemes later on.
+%
 %
@@
 being the instantaneous histogram,
-which is equal to $1.0$
+which is equal to $1$
 for the bin $i(t)$ at time $t$
@@
 %
-Thus, we have
+Deducting the bin average [cf. Eq. \eqref{eq:x_def}],
+we get
 %
@@
   \alpha(t) \, \eta_{*i}(t)
-  ,
+  .
   \label{eq:vt_diffeq}
 \end{equation}
-%
-where we have defined the bin-average-deducted
-histogram, $\eta_{*i}(t)$, in a similar way to
-$v_{*i}(t)$ [cf. Eq. \eqref{eq:x_def}].
-
-We now split the histogram into
+
+We can split the histogram into
 a deterministic averaged part, $\langle \eta_i(t) \rangle$,
@@
 The deterministic part can be related
-to an ``ensemble average'',
-where the ensemble consists of many similar simulation copies
+to an average in an ensemble consists of many similar simulation copies
 sharing the same schedule $\alpha(t)$
@@
 during the process may differ, however.
-
+%
 For sufficiently small $\alpha(t)$,
-the bias potential remains roughly the same for a short period.
-%
-Then,
-the sampling process is approximately Markovian, % of a finite order,
-and we may assume Eq. \eqref{eq:pi_p_V}
-for the deterministic part
+the bias potential remains roughly the same for a short period,
+and we may assume a quasi-equilibrium sampling process
+such that
+%with the deterministic part specified by Eq. \eqref{eq:pi_p_V}:
 %
@@
 of $v_{*i}(t)$ in the linear expansion.
-%and used $\sum_{j=1}^n p_j = 1$
-%on the second line.
 %
@@
 We will approximate the histogram fluctuation part, $\zeta_i(t)$,
-as white noise,
-and assume its validity for long simulations.
-%
-The bin-average-deducted fluctuation, $\zeta_{*i}(t)$,
-%%
-%\begin{equation}
-%  \zeta_{*i}(t)
-%  \equiv
-%  \zeta_i(t)
-%  -
-%  \sum_{ j = 1 }^n
-%    \rho_j \, \zeta_j(t)
-%  ,
-%  \notag
-%  %\label{eq:xi_def}
-%\end{equation}
-%%
-would also be governed by white noise as is $\zeta_i(t)$, i.e.,
+hence its bin-average-shifted version, $\zeta_{*i}(t)$,
+as white noise such that
 \begin{equation}
@@
 and $\delta(t)$ is the Dirac's delta function.
+%
+We expect the approximation to hold for simulation
+much longer than the autocorrelation time.
 %
@@
   .
-  \label{eq:w_balance}
+  \notag
+  %\label{eq:w_balance}
 \end{equation}
@@
 $\bav{f}= \mathcal F[f_i]_0 = \tilde f_0$,
-and
+so that
 \begin{equation}
@@
 \end{equation}
-which means that the Fourier transform of
-a bin-average-deducted quantity always
-has a vanishing first mode.
+%
+which means that the only the only effect
+of bin average deduction is to
+annihilate the first mode, $\tilde f_0$.
 
@@
 $$
-or in the continuous-time form,
+or in continuous time, we have, for $k > 0$,
 %
@@
   \alpha(t) \, \lambda_k \, {\tilde \eta}_k(t)
-  .
-  \label{eq:ut_diffeq_mbin}
-\end{equation}
-%
-Deducting the bin average yields
-$$
-  \dot {\tilde v}_{*k}(t)
-  = \alpha(t) \, \lambda_k \, {\tilde \eta}_{*k}(t)
   \approx
   -\alpha(t) \, \lambda_k \,
-  \bigl[ {\tilde v}_{*k}(t) - {\tilde \zeta}_{*k} (t) \bigr]
-  ,
-$$
-where we have used the linear approximation
+  \bigl[ {\tilde v}_k(t) - {\tilde \zeta}_k(t) \bigr]
+  ,
+  \label{eq:ut_diffeq_mbin}
+\end{equation}
+%
+where we have removed the bin average,
+and used the transformed version of Eq. \eqref{eq:sh_ave}
+[the subscript ``$_*$'' is dropped by Eq. \eqref{eq:ftstar}].
+In the second step, we have used the linear approximation
 as in Sec. \ref{sec:sbin_diffeq}.
@@
   =
-  \sum_{k = 0}^{n - 1}
+  \sum_{k = 1}^{n - 1}
     \left\langle
-      {\tilde v}_{*k}^2(T)
+      {\tilde v}_k^2(T)
     \right\rangle
@@
 %
-where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
-%
-By Eq. \eqref{eq:ftstar},
-we can start the last sum %in Eq. \eqref{eq:xt2_sum}
-from $k = 1$,
-%
-\begin{equation}
-  \Err(T)
-  =
-  \sum_{k = 1}^{n - 1}
-    \left\langle
-      {\tilde v}_{*k}^2(T)
-    \right\rangle
-  .
-\label{eq:xt2_sum}
-\end{equation}
+where we have used
+Eq. \eqref{eq:eig_orthonormal_cols}
+and drop the first term
+by using Eq. \eqref{eq:ftstar}
+in the second step.
 
@@
     \left\langle
-      {\tilde v}_{*k}^2(0)
+      {\tilde v}_k^2(0)
     \right\rangle \,
@@
 of the histogram fluctuation,
-$\kappa_k(t) \equiv \langle {\tilde \zeta}_{*k}(0)
-\, {\tilde \zeta}_{*k}(t) \rangle$,
+$\kappa_k(t) \equiv \bigl\langle {\tilde \zeta}_{*k}(0)
+\, {\tilde \zeta}_{*k}(t) \bigr\rangle$,
 as:
@@
 
-%Before addressing the optimal schedule
-%for a multiple-bin scheme,
-%we will first compute the equilibrium error
-%under a constant schedule,
-%$\alpha(t) = a_0$.
-%%
-%%In terms of the components, we have
-%%
-%\begin{align}
-%  \left\langle \left|
-%    {\tilde v}_{*k}^2
-%  \right| \right\rangle
-%  &=
-%  \int_0^{T'}
-%    \Gamma_k \, (\lambda_k \, a_0)^2 \,
-%      e^{ 2 \, \lambda_k \, a_0 \, (t - T') }
-%    \, dt
-%  \notag
-%  \\
-%  &=
-%  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
-%  ,
-%  \qquad
-%  ( T' \to \infty )
-%  .
-%\label{eq:xt2_eql}
-%\end{align}
-%%
-
 To model the initial error for Eq. \eqref{eq:error_res},
@@
   \left\langle
-    {\tilde v}_{*k}^2(0)
+    {\tilde v}_k^2(0)
   \right\rangle
   = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
-  + \tilde u_k^2 \, e^{-2\, \lambda_k \, T_0}
+  + \epsilon_k
   ,
@@
 %
+where
+$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$
 and if no eigenvalue is zero,
-then we can ignore the second term
+then we can ignore the $\epsilon_k$ term
 in the case of $T_0 \to \infty$,
@@
   \left\langle
-    {\tilde v}_{*k}^2(0)
+    {\tilde v}_k^2(0)
   \right\rangle
   = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
-  ,
+  .
   \label{eq:xt2_eql}
@@
 \note{
-Consider a long period, $T'$, under a fixed schedule,
+Consider a long period, $T_0$, under a fixed schedule,
 $\alpha(t) = a_0$,
@@
   \left\langle
-    {\tilde v}_{*k}^2(T_0)
+    {\tilde v}_k^2(T_0)
   \right\rangle
@@
 }
-%
-and
-%
-\begin{equation}
-  \Err_R(T)
-  =
-  \sum_{k = 1}^{n - 1}
-    \left\langle
-      {\tilde v}_{*k}^2(0)
-    \right\rangle
-      e^{  - 2 \, \lambda_k \, q(T) }
-  .
-  \label{eq:error_res1}
-\end{equation}
 %
@@
 the Hamiltonian is conserved,
-and we may set
+i.e., the asymptotic error grows
+linearly with time at a rate of $\mathcal L$.
+%
+So we may set
 %
@@
 \eqref{eq:Lagrangian_const}.
-%
-
-%This mapping allows us to classify optimal schedules
-%as follows.
-%%
-%We collect in each class optimal schedules
-%computed from simulations with the same function, $M(Q)$,
-%and cutoff value of $q(T)$.
-%%
-%The former is determined by the values of
-%$\lambda_k$'s and $\Gamma_k$'s,
-%whereas the latter affects the distribution, $m(Q)$,
-%via the normalization constant, $C$.
-%
-%
-%In each class, the updating magnitudes of two schedules
-%of lengths $T_1$ and $T_2$
-%at corresponding times, $t_1/T_1 = t_2/T_2$,
-%are related by $T_1 \, \alpha_1(t_1) = T_2 \, \alpha_2(t_2)$.
-%%
-%Equation \eqref{eq:error_asym2}
-%means that the asymptotic error
-%is inversely proportional to the simulation length
-%for schedules in the same class.
-%%which can be rephrased as another scaling relation
-%%$$
-%%T_1 \, \Err_{A1}(T_1) = T_2 \, \Err_{A2}(T_2).
-%%$$
-
-%The optimal schedule has the following scaling property.
-%%
-%For two simulations of lengths $T_1$ and $T_2$,
-%if $q(T_1) = q(T_2)$
-%and they share the same $\lambda_k$'s and $\Gamma_k$'s,
-%the updating magnitudes at the corresponding times,
-%$t_1/T_1 = t_2/T_2$,
-%can be related as $T_1 \, \alpha_1 = T_2 \, \alpha_2$.
-%%
-%In other words,
-%if $\alpha(t_1)$ is
-%an optimal schedule of the simulation of length $T_1$,
-%then
-%$\frac{T_1}{T_2} \alpha\left( \frac{ T_1 } { T_2 } t_2 \right)$
-%is an optimal schedule of the simulation of length $T_2$.
-%%
-%The asymptotic errors are related by
-%$T_1 \, E_{A1} = T_2 \, E_{A2}$.
 
@@
 
-The above characterization allows us to determine
-the optimal initial updating magnitude,
-or, equivalently, the optimal value of $q(T)$
+We now determine the optimal value of $q(T)$
 that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
@@
 We will consider simulations in which
-the system is initially equilibrated at the same
+the system is initially equilibrated at a
 constant updating magnitude, $a_0$.
 %
-From Eqs. \eqref{eq:error_res1}
+From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
 and \eqref{eq:error_asym2},
@@
 the optimal $q(T)$ satisfies
-\begin{equation}
-  m\bigl( q(T) \bigr)
-  =
-  \frac{1} { T \, (a_0 / 2) }
+%
+\begin{align}
+  a_0 \, M\bigl( q(T) \bigr)
+  +
+  \frac{ 2 }{M\bigl( q(T) \bigr)}
+  \sum_{k = 1}^{n-1}
+  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
+  \notag \\
+  = \frac{2}{T} \int_0^{q(T)} M(Q') \, dQ'
   .
 \label{eq:opt_qT}
-\end{equation}
-%
-\note{Let $Q_c \equiv q(T)$,
+\end{align}
+%
+%\begin{align}
+%  m\bigl( q(T) \bigr)
+%  +
+%  \frac{2}{a_0 \, C \, M\bigl( q(T) \bigr)}
+%  \sum_{k = 1}^{n-1}
+%  \lambda_k \, \tilde u_k^2 \,
+%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
+%  \\=
+%  \frac{1} { T \, (a_0 / 2) }
+%  .
+%\label{eq:opt_qT}
+%\end{align}
+%
+\note{Let $q_T \equiv q(T)$,
 $$
@@
   {
-    \partial Q_c
+    \partial q_T
   }
   &=
-  -a_0 \, M^2(Q_c)
+  -a_0 \, M^2(q_T)
+  -2
+  \sum_{k=1}^{n-1} \lambda_k \,
+  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
   ,
@@
   {
-    \partial Q_c
+    \partial q_T
   }
@@
   \frac 2 T \,
-  M(Q_c) \,
-  \int_0^{ Q_c } M(Q) \, dQ
+  M(q_T) \,
+  \int_0^{ q_T } M(Q) \, dQ
+  =
+  \frac{ 2 \, C } { T } \, M(q_T)
   .
@@
 Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
-}%
-By Eqs. \eqref{eq:mQ_invTa} and \eqref{eq:intmQ_tT},
-we can rewrite the condition as
+}
+%
+We can rewrite the condition in terms of
+the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
+\begin{align}
+  \alpha(t = 0)
+  =
+  \frac{1}{T \, m\bigl( q(T) \bigr) }
+  =
+  \frac{ a_0 } { 2 }
+  +
+  \sum_{k = 1}^{n-1}
+  \frac{
+    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
+  }{M^2\bigl( q(T) \bigr)}
+  .
+  \notag %\label{eq:opt_alpha0}
+\end{align}
+%
+If the initial bias, $\epsilon_k$, is negligible, then
+%
 \begin{equation}
@@
 i.e. the optimal initial updating magnitude
-is always half of the equilibrium value.\footnote{This factor, $1/2$,
+is half of the equilibrium value.\footnote{This factor, $1/2$,
 happens to coincide with the
@@
 Based on the above results,
-we give a procedure of computing
+we now give a procedure of computing
 the optimal schedule and the error
@@
 %
-We have the following procedure.
+We have
 
@@
 \item
-Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}
-    using e.g. the Newton-Raphson or bisection method\cite{press3rd}.
+Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
+\footnote{Our implementation used a variant of the Newton-Raphson method
+    with bracketing, and the initial value is set to $q(T) = \ln(1+T\,a_0/2)$,
+    but other methods might be more effective\cite{press3rd}.}
 %
@@
 %($i = 0, 1, \dots, N$)
-We have used the standard trapezoidal rule\cite{press3rd},
-on $N+1$ grid points on $[0, q(T)]$ that are defined adaptively
+We have used the trapezoidal rule\cite{press3rd}
+  on an adaptively defined grid on $[0, q(T)]$ of $N+1$ points
   such that the resulting change of $t_{[i+1]}-t_{[i]}$
@@
 and the updating magnitude can be obtained from numerical differentiation:
-    $\alpha_i = [1 - e^{-\lambda_1 (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})}]/\lambda_1$,
+    $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$
+    and
+    $\hat \alpha_i \equiv (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
 with
@@
 Eqs. \eqref{eq:error_tot},
-  \eqref{eq:error_res1},
+  \eqref{eq:error_res},
+  \eqref{eq:xt2_eql1},
   and
@@
 Numerical results}
-
-We tested the above theory on a model system
-described in Sec. \ref{sec:results_system}.
-%
-We verify the method of computing errors
-%using an inverse-time schedule
-in Sec. \ref{sec:results_invt}
-%
-and study the optimal schedule from
-Eq. \eqref{eq:q_opt}
-in Sec. \ref{sec:results_optschedule}.
-%
-Finally, we compare the Gaussian and
-bandpass updating schemes %of different widths
-in Sec. \ref{sec:results_cmpschemes}.
-
-\subsection{\label{sec:results_system}
-Test system and simulation protocol}
-
-Our test system is a one-dimensional model system, in which
-$i$ represents a non-periodic variable,
-and $n = 100$.
-%
-Both the target and intrinsic distributions
-are assumed to be flat:
-$p^*_i = p_i = 1/n$,
-along with the distribution, $\rho_i$, in Eqs.
-\eqref{eq:vav_def} and \eqref{eq:error_def}.
-%
-%We wish to see how
-%the schedule, $\alpha(t)$,
-%affects the fluctuating error
-%of the bias potential
-%at the end of the simulation.
-%
-Initially,
-we set the bias potential to zero,
-$u_i \equiv 0$,
-and equilibrate the system for $10^7$ steps
-under a constant updating magnitude,
-$a_0 = 10^{-4}$.
-%
-This equilibration process models
-a stage of the WL algorithm or
-a regular metadynamics
-with a fixed updating magnitude.
-%
-Although the bias potential is correct initially,
-it would deviate from this value
-at the end of the equilibration.
-%
-Our aim is to see how well
-the bias potential is restored
-at the end of the testing schedule.
-%
-After the equilibration,
-we reset the origin of the time, $t$, to zero,
-and start the testing schedule
-for $T = 10^8$ steps.
-%
-We repeat the above process $1000$ times (unless specified otherwise)
-and report the averaged results.
-%specified by either
-%Eq. \eqref{eq:q_opt}
-%or
-%Eq. \eqref{eq:alpha_invtlambda}.
-
-For the Gaussian updating scheme,
-the kernel was truncated at
-$b = \min\{10 \, \sigma, n - 1\}$ bins
-and numerically stabilized
-using the technique
-in Appendix \ref{sec:stabilize_wband}.
-
-%Since the error depends on
-%the underlying sampling process,
-We tested two sampling processes.
-%a global one and a local one.
-%
-Both are based on
-a Metropolis MC algorithm\cite{
-  metropolis1953, newman, frenkel,
-  landau_binder}:
-%
-in each step, a new bin index $j$ is proposed
-and then accepted with probability
-%
-$
-A(i \to j) = \min\{ 1, \exp(u_i - u_j) \}.
-$
-However,
-in the first or global sampling process,
-we choose the destination $j$
-uniformly out of the $n$ bins.
-%
-In the second or local sampling process,
-$j$ is either $i - 1$ or $i + 1$
-(adjusted by the boundary condition)
-with equal probability $1/2$.
-%
-Asymptotically,
-the above global and local sampling processes
-emulate the perfect and one-step processes
-discussed in Appendix \ref{sec:Gamma},
-respectively,
-and we will use
-Eqs. \eqref{eq:Gamma_perfect}
-and \eqref{eq:Gamma_onestep}
-for the respective $\Gamma_k$'s
-in the theoretical predictions.
-%
-
-\subsection{\label{sec:results_invt}
-Scaled inverse-time schedule}
-
-Using the model system above, we
-tested our analysis of computing errors
-on the scaled inverse-time schedule,
-%
-\begin{equation}
-\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
-\label{eq:alpha_invtlambda}
-\end{equation}
-%
-where $\lambda$ is a free parameter,
-and $t_0$ is given by Eq. \eqref{eq:t0_sinc}.
-%and $t_0 = 2/a_0$.
-%
-The error of this schedule
-can be computed analytically
-(cf. Appendix \ref{sec:sinvt_schedule}),
-and tested against simulation.
-
-We performed adaptive FDS simulations
-using the single-bin updating scheme.
-%and the triple-bin updating scheme with $\mu_1 = 0.24$.
-%
-%For each value of $\lambda$,
-%we ran $1000$ independent simulations
-%of $10^8$ steps, and averaged the results.
-%
-As shown in Fig. \ref{fig:errinvt},
-the errors from the simulations
-agreed well with those from the theory.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/errinvt.pdf}
-  }
-  \caption{
-    \label{fig:errinvt}
-    Error, $\Err$, versus the parameter, $\lambda$,
-    for the single-bin updating scheme
-    using the scaled inverse-time schedule,
-    Eq. \eqref{eq:alpha_invtlambda}.
-    %
-    The points are results from averaging over $1000$ independent runs;
-    the curves are theoretical predictions.
-    %
-    \note{The figure was produced by \texttt{doc/fig/errinvt.gp}.
-      The data were produced by the script \texttt{data/invtcscan.py},
-      and saved under \texttt{data/invt}
-      (see the \texttt{Makefile} there).
-    }%
-  }
-\end{center}
-\end{figure}
-
-For the single-bin updating scheme,
-the optimal $\lambda$ occurred around $1.0$
-for both global and local sampling processes.
-%
-The sampling process affected the error function
-only by a multiplicative constant, $\Gamma$
-(cf. Appendix \ref{sec:sinvt_schedule}).
-%
-%For the triple-bin updating scheme,
-%the optimal $\lambda$ was less than $1.0$
-%for the global sampling process,
-%although it stayed around $1.0$
-%for the local sampling process,
-%also in agreement with the discussion
-%in Appendix \ref{sec:invt_nn}.
-%%
-%Thus, the inverse-time schedule,
-%Eq. \eqref{eq:alpha_invt1},
-%is optimal for the single-bin updating scheme,
-%but not necessarily so for a general multiple-bin updating scheme.
-%%
-
-\subsection{\label{sec:results_optschedule}
-Optimal schedule}
-
-To demonstrate the optimal schedule from Eq. \eqref{eq:q_opt},
-we used the Gaussian (with $\sigma = 10$ bins) updating scheme
-as an example.
-%
-This updating scheme is typically used in metadynamics simulations.
-%
-In Table \ref{tab:error_Gaussian},
-we showed that the optimal schedule
-produced smaller errors than
-the inverse-time schedules.
-%
-However, the optimal schedules
-decayed more slowly than
-the inverse-time formula, Eq. \eqref{eq:alpha_invt1},
-and depended on the sampling process,
-as shown in Fig. \ref{fig:optacmp}.
-%
-The impeded decay of the optimal schedule
-was necessary in order to
-damp out the short-wavelength eigenmodes
-(i.e. local noise).
-%
-%As shown in Fig. \ref{fig:xerr},
-%the inverse-time schedule was effective in error reduction
-%only for the first few eigenmodes,
-%whose eigenvalues were around $1.0$.
-%%
-%We can, however, ignore these local eigenmodes
-%by explicitly filtering them out from the final bias potential,
-%and truncating the error sum in Eq. \eqref{eq:error_def}
-%to the first few eigenmodes.
-%%
-%Then, the corresponding optimal schedule
-%from Eq. \eqref{eq:q_opt} that minimizes the truncated error
-%became similar to the inverse-time schedule,
-%Eq. \eqref{eq:alpha_invt1},
-%as exemplified in Fig. \ref{fig:optacmp}.
-%%
-
-\begin{table}[h]\footnotesize
-  \caption{\label{tab:error_Gaussian}
-    Error of the Gaussian updating scheme of $\sigma = 10$
-    under different updating schedules.
-    In Eqs. \eqref{eq:alpha_invt1} and \eqref{eq:alpha_invtlambda},
-    $t_0$ is given by Eq. \eqref{eq:t0_sinc};
-    and in Eq. \eqref{eq:alpha_invtlambda},
-    the optimal value of $\lambda$ is used.
-    \note{\newline
-      The data were collected in
-      \texttt{data/tab1} by running
-      the program \texttt{invt}.
-    }%
-  }
-  \setlength{\tabcolsep}{2pt}
-  \renewcommand\arraystretch{1.2}
-  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
-    \hline
-      Schedule
-    &
-      $1/t$, \newline
-      Eq. \eqref{eq:alpha_invt1}
-    &
-      Scaled $1/t$, \newline
-      Eq.~\eqref{eq:alpha_invtlambda}
-    &
-      Optimal, \newline
-      Eq. \eqref{eq:q_opt}
-    \\
-    \hline
-    Global
-    &
-    $4.36(9) \times 10^{-6}$
-    &
-    $7.1(1) \times 10^{-7}$
-    &
-    $2.6(1) \times 10^{-7}$
-    \\
-    \hline
-    Local
-    &
-    $3.83(7) \times 10^{-4}$
-    &
-    $1.72(4) \times 10^{-4}$
-    &
-    $9.1(2) \times 10^{-5}$
-    \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/optacmp.pdf}
-  }
-  \caption{
-    \label{fig:optacmp}
-    Optimal schedules from Eq. \eqref{eq:q_opt}
-    for the Gaussian updating scheme
-    with $\sigma = 10$.
-    %
-    The mode-limited schedules were produced by
-    limiting the sum in Eq. \eqref{eq:q_opt}
-    to the first four terms,
-    or the four longest wavelength modes.
-    %
-    They became similar to the inverse-time schedule,
-    Eq. \eqref{eq:alpha_invt1}.
-    %
-    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
-      The data were computed by the program \texttt{prog/predict},
-      see particularly the function \texttt{intq\_save()}
-      in \texttt{prog/intq.h}.
-      The data were saved under the directory \texttt{data/opta}.
-    }%
-  }
-\end{center}
-\end{figure}
-
-An interesting prediction from the theory is Eq. \eqref{eq:half_alpha0},
-which states that the optimal schedule should always start from
-half of the magnitude from the previous equilibration run.
-%
-To verify this result,
-we computed the optimal schedules
-for different values of $q(T)$
-for the Gaussian updating scheme,
-and plotted the final error
-against the initial updating magnitude, $\alpha(0)$.
-%
-As shown in Fig. \ref{fig:erriascan},
-the minimal error was indeed reached at
-$\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
-although the dependence on $\alpha(0)$
-is generally weak.\footnote{The weak dependence is best illustrated in
-the case of single-bin updating scheme,
-as the initial updating magnitude only
-affects the schedule by a shift of the origin of
-simulation time in the inverse-time schedule,
-Eq. \eqref{eq:alpha_invt1}.}
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/erriascan.pdf}
-  }
-  \caption{
-    \label{fig:erriascan}
-    Relative error
-    versus the initial updating magnitude, $\alpha(0)$,
-    for optimal schedules with different $q(T)$.
-    Assuming the Gaussian updating scheme
-    with $\sigma = 10$,
-    and $a_0 = 10^{-4}$.
-    %
-    The points were obtained from averaging results
-    from over 10000 independent simulations;
-    the curves are theoretical predictions.
-    %
-    The minimal should always occur at
-    $\alpha(0) = a_0/2 = 5 \times 10^{-5}$,
-    as predicted by Eq. \eqref{eq:half_alpha0}.
-    %
-    \note{This figure was produced by
-      \texttt{doc/fig/erriascan.gp}.
-      The data were saved under
-      \texttt{data/iascan}.
-    }%
-  }
-\end{center}
-\end{figure}
-
-\subsection{\label{sec:results_cmpschemes}
-Optimal updating schemes}
-
-For a finite-length simulation,
-the optimal updating scheme
-depends on the simulation length, $T$,
-and other parameters.
-%
-As an example,
-we compare Gaussian updating schemes
-with different widths, $\sigma$.
-%
-In Fig. \ref{fig:errsigscan},
-we present the errors, $\Err(T)$,
-normalized by the shifted simulation length, $T+t_0$,
-so that they could
-stay constant in the limit of single-bin updating scheme
-($\sigma \to 0$).
-%
-As shown in Fig. \ref{fig:errsigscan}(a)
-that the minimal error was achieved
-by the widest window %for the shorter simulation of $T = 10^8$,
-for the global sampling process,
-%
-although
-the minimum at $\sigma = 0$ became more favorable
-as the simulation lengthens.
-%
-However,
-for the local sampling process,
-the minimum stayed at zero width,
-as shown in Fig. \ref{fig:errsigscan}(b).
-
-In contrast,
-the normalized errors of the bandpass updating schemes
-were insensitive to the simulation length,
-in agreement with Eq. \eqref{eq:error_sinc}.
-%
-The bandpass updating schemes
-yielded smaller errors than
-the corresponding Gaussian ones,
-although the differences tended to
-be small for shorter simulations
-with the local sampling process.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsigscan.pdf}
-  }
-  \caption{
-    \label{fig:errsigscan}
-    Normalized error, $(T + t_0) \, \Err(T)$,
-    versus the width, $\sigma$,
-    for the Gaussian
-    and the bandpass updating schemes,
-    with $t_0$ given by Eq. \eqref{eq:t0_sinc}.
-    %
-    In the latter case,
-    the equivalent width
-    $\sigma$ is given by Eq. \eqref{eq:sigma_equiv}.
-    %
-    The points were obtained from averaging results
-    from 1000 independent simulations;
-    the curves are theoretical predictions.
-    %
-    \note{This figure was produced by
-      \texttt{doc/fig/errsigscan.gp}.
-      The data were saved under
-      \texttt{data/scan}.
-    }%
-  }
-\end{center}
-\end{figure}
 
@@
 
-As an example, we used the method to compute
-the integrals, $\Gamma_k$'s, from
-a local MC simulation
-(cf. Sec. \ref{sec:results_system})
-and an MD simulation,
-and compared them with those from
-the one-step sampling process.
-%
-As shown in Fig. \ref{fig:gamcmp},
-the $\Gamma_k$'s of the small-$k$ eigenmodes
-from the three sampling processes
-were almost proportional,
-suggesting that
-the optimal schedule
-for simulations using local sampling processes
-would be similar in general properties.
-
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=0.95\linewidth]{fig/gamcmp.pdf}
-  }
-  \caption{
-    \label{fig:gamcmp}
-    Integrals of the autocorrelation functions
-    from a local MC simulation
-    and an MD simulation.
-    %
-    Each simulation was ran for $10^{10}$ steps.
-    %
-    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$.
-    %
-    For the MD simulation,
-    the step size $\Delta t_\mathrm{MD} = 0.005$.
-    %
-    The velocity rescaling thermostat\cite{bussi2007} was
-    applied for the unit reduced temperature
-    every MD step with viscosity $20$.
-    %
-    The reflective boundary condition was used.
-    %
-    \note{The figure was produced by \texttt{doc/fig/gamcmp.gp}.
-      The data were saved under \texttt{data/gamcmp}
-      (see the \texttt{Makefile} there).
-    }%
-  }
-\end{center}
-\end{figure}
-
 \section{\label{sec:more_wband}
@@
   =
-  \frac 1 n
-  \exp
-  \left(
-    \ii
-    \frac{ k \, j \, 2 \, \pi }
-         {      n             }
-  \right)
-  ,
-  \notag
-\end{equation}
-%
-where $\ii^2 = -1$.
-%
-Alternatively, to avoid complex numbers, we can use
-%
-\begin{equation}
-  \phi^{(k)}_j
-  =
-  \phi_{kj}
-  =
   \frac{ \sqrt{ 2 - \delta_{k,0} } } { n }
@@
   $\lambda_{n - k} = \lambda_k$.
-\note{Consider the unnormalized version
+\note{Consider the unnormalized exponential equivalent
   $$
@@
   ,
-  \label{eq:w_band_refl}
+  \notag
+  %\label{eq:w_band_refl}
 \end{equation}
@@
 
-\subsection{\label{sec:stabilize_wband}
-Stabilization}
-
-%A practical problem of using homogeneous updating scheme
-%is the following.
-%
-A homogeneous updating scheme
-specified by an arbitrary kernel
-is not necessarily stable,
-i.e. some eigenvalues can be negative.
-%
-%
-%\note{
-  %For example,
-  %the triple-bin updating scheme
-  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
-  %
-%}%
-Even
-the Gaussian updating scheme, which is stable in the continuous limit,
-can be unstable
-when the updating kernel is truncated during discretization.
-%
-Here we show how to minimally modify
-the updating kernel
-to stabilize it.
-%
-\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
-}
-
-Given an updating kernel,
-we can compute the eigenvalues from
-Eq. \eqref{eq:wband_eigenvalue}.
-%
-By replacing negative eigenvalues with zeros and
-using the inversion formula,
-Eq. \eqref{eq:mu_from_lambda},
-we get a new updating kernel,
-which is stable by construction.
-%
-However, the new kernel is widened
-as $\mu_l$ may be nonzero for $l > b$.
-%
-If we truncate the new kernel at $\mu_b$,
-the stability problem remains,
-although the negative eigenvalues
-tend to be smaller in magnitude.
-%
-Thus, to stabilize the updating kernel
-without increasing the width,
-we need to iterate the above process many times,
-resulting the following algorithm.
-
-%
-\begin{enumerate}
-  \item
-    Given an updating kernel, $\mu_0, \dots, \mu_b$,
-    compute the eigenvalues,
-    $\lambda_1, \dots, \lambda_{n-1}$,
-    from Eq. \eqref{eq:wband_eigenvalue}.
-  \item
-    If all eigenvalues are nonnegative within the numerical error,
-    we are done. % and exit the loop.
-  \item
-    Otherwise, replace negative eigenvalues by zeros,
-    and compute the new kernel,
-    $\mu_0, \dots, \mu_b$, from
-    Eq. \eqref{eq:mu_from_lambda},
-    with $\mu_l = 0$ for $l > b$.
-    Go to step 1.
-\end{enumerate}
-
 \subsection{Examples}
@@
 %
-If follows that all eigenvalues are positive\cite{bussi2006}:
+It follows that all eigenvalues are positive\cite{bussi2006}:
 %
@@
   .
-\notag
-%\label{eq:lambda_Gaussian}
+%\notag
+\label{eq:lambda_Gaussian}
 \end{align}
 
-%However,
-%as $\sigma_\varphi/\Delta \varphi \gg 1$,
-%the smallest eigenvalue
-%%
-%$$
-%\lambda_n
-%\approx
-%\exp\left(
-%      -\frac{ n^2 \, \sigma_\varphi^2 }
-%            {        2        }
-%    \right)
-%=
-%\exp\left[
-%      -\frac{ \pi^2 }{ 2 }
-%      \left(
-%        \frac{  \sigma_\varphi }
-%             { \Delta x }
-%      \right)^2
-%    \right]
-%.
-%$$
-%%
-%is exponentially small.
-%%
-%This means that
-%the optimal value of $\lambda$
-%given by Eq. \eqref{eq:optimal_lambda_approx}
-%would be too small in practice,
-%for it requires $\lambda < 2 \, \lambda_n$.
-%%
-%In other words,
-%with a reasonable $\lambda$,
-%the error of the last few fluctuation modes
-%will always decay suboptimally,
-%accordingly to Eq. \eqref{eq:error_asym_invt}.
-%%
-%However, since these modes
-%represent short wavelength fluctuations,
-%and these modes may be in
-%the system under consideration.
-%%
-%Thus,
-%we can truncate the sum of the error function
-%in Eq. \eqref{eq:error_asym_invt}
-%at some $k_{\max}$,
-%which corresponds to a minimal length scale
-%$l_{\min} = \Delta x \, (n /k_{\max}) = \pi/k_{\max}$
-%greater than $\sigma_\varphi$.
-%%
-%Then,
-%$$
-%\lambda_{ k_{\max} }
-%\approx
-%\exp\left(
-%      -\frac{ k_{\max}^2 \, \sigma_\varphi^2 }
-%            {           2            }
-%    \right)
-%=
-%\exp\left(
-%      -\frac{  \pi^2 \, \sigma_\varphi^2    }
-%            {    2 \, l_{\min}^2  }
-%    \right).
-%$$
+For a finite bin width,
+some eigenvalues can be negative.
+However, we can avoid the stability problem
+by starting with the eigenvalues given by Eq. \eqref{eq:lambda_Gaussian}
+and inversely computing the kernel using Eq. \eqref{eq:mu_from_lambda}.
 
@@
   .
-\label{eq:mu_sinc_refl}
+  \notag
+  %\label{eq:mu_sinc_refl}
 \end{equation}
@@
 
-\section{\label{sec:sinvt_schedule}
-Scaled inverse-time schedule}
-
-The error of the scaled inverse-time schedule,
-Eq. \eqref{eq:alpha_invtlambda},
-can be found by using Eq. \eqref{eq:alpha_invtlambda}
-in Eqs. \eqref{eq:error_asym} and \eqref{eq:error_res1}:
-%
-\begin{align}
-  \Err_A(T)
-  &=
-  \frac{    1    }
-       { T + t_0 }
-  \sum_{k = 1}^{n-1}
-    \frac{ \Gamma_k \, \nu_k^2 }
-         {    2 \, \nu_k - 1   }
-  \left[
-    1 - \left(
-          \frac {     t_0 }
-                { T + t_0 }
-        \right)^{ 2 \, \nu_k - 1 }
-  \right]
-  ,
-  \notag
-  %\label{eq:error_asym_invt}
-  \\
-  \Err_R(T)
-  &=
-  \frac { a_0 } { 2 }
-  \sum_{k = 1}^{n-1}
-  \Gamma_k \, \lambda_k \,
-  \left(
-      \frac{   t_0   }
-           { T + t_0 }
-  \right)^{ 2 \, \nu_k }
-  ,
-  \notag
-  %\label{eq:error_res_invt}
-\end{align}
-%
-where $\nu_k \equiv \lambda_k / \lambda$.
-%
-%with
-%$
-%  q(T)
-%  =
-%  ( 1 / \lambda )
-%  \ln\left(
-%    1 + T / t_0
-%  \right)
-%$
-%%
-%from integrating Eq. \eqref{eq:alpha_invtlambda}.
-%
-%We assume that
-%the system was initially equilibrated
-%at a constant $a_0$,
-%such that the values of
-%$\left\langle {\tilde v}_{*k}^2(0) \right\rangle$
-%can be computed from Eq. \eqref{eq:xt2_eql},
-%as well as Eq. \eqref{eq:t0_sinc}.
-%
-%The total error is
-%%
-%\begin{align}
-%\Err
-%&=
-%\Err_R + \Err_A
-%\notag
-%\\
-%&=
-%\sum_{ k = 1 }^{n-1}
-%  \frac
-%  {
-%    \Gamma_k \, \nu_k \,
-%    \left[
-%      \nu_k
-%      +
-%      (\nu_k - 1)
-%      \left(
-%        \frac{ t_0 } { T + t_0 }
-%      \right)^{ 2 \, \nu_k - 1 }
-%    \right]
-%  }
-%  {
-%    (T + t_0) \, (2 \, \nu_k - 1)
-%  }
-%  .
-%%\notag
-%\label{eq:error_invt}
-%\end{align}
-
-In particular,
-for the single-bin scheme, we have
-$\lambda_1 = \cdots = \lambda_{n-1} = 1$,
-and
-$\Gamma_k$'s affect the error
-only via the sum, $\Gamma = \sum_{k = 1}^{n-1} \Gamma_k$,
-as a multiple.
-%
-In the long time limit,
-we have for $\lambda < 2$,
-$$
-  \Err_A(T)
-  =
-  \frac { 1 } { T + t_0 }
-  \frac {         \Gamma           }
-        { \lambda \, (2 - \lambda) }
-  ,
-$$
-which is minimal at $\lambda = 1$.
-%
-\note{A higher-order correction:
-  $
-  \lambda = 1 -
-  \left[
-    1 - \frac 1 2 a_0 \, (T+t_0)
-  \right] r \, \log r
-  ,
-  $
-  where $r = t_0 / (T + t_0)$.
-}
+\subsection{\label{sec:stabilize_wband}
+Stabilization}
+
+%A practical problem of using homogeneous updating scheme
+%is the following.
+%
+A homogeneous updating scheme
+specified by an arbitrary kernel
+is not necessarily stable,
+i.e. some eigenvalues can be negative.
+%
+Here we show how to minimally modify
+the updating kernel
+to stabilize it.
+%
+\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
+}
+
+Given an updating kernel,
+we can compute the eigenvalues from
+Eq. \eqref{eq:wband_eigenvalue}.
+%
+By replacing negative eigenvalues with zeros and
+using the inversion formula,
+Eq. \eqref{eq:mu_from_lambda},
+we get a new updating kernel,
+which is stable by construction.
+%
+However, the new kernel is widened
+as $\mu_l$ may be nonzero for $l > b$.
+%
+If we truncate the new kernel at $\mu_b$,
+the stability problem remains,
+although the negative eigenvalues
+tend to be smaller in magnitude.
+%
+Thus, to stabilize the updating kernel
+without increasing the width,
+we need to iterate the above process many times,
+resulting the following algorithm.
+
+%
+\begin{enumerate}
+  \item
+    Given an updating kernel, $\mu_0, \dots, \mu_b$,
+    compute the eigenvalues,
+    $\lambda_1, \dots, \lambda_{n-1}$,
+    from Eq. \eqref{eq:wband_eigenvalue}.
+  \item
+    If all eigenvalues are nonnegative within the numerical error,
+    we are done. % and exit the loop.
+  \item
+    Otherwise, replace negative eigenvalues by zeros,
+    and compute the new kernel,
+    $\mu_0, \dots, \mu_b$, from
+    Eq. \eqref{eq:mu_from_lambda},
+    with $\mu_l = 0$ for $l > b$.
+    Go to step 1.
+\end{enumerate}
 
@@
     $\mathbf w$     &   Updating matrix. \\
-    $v_{*i}(t)$        &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
+    $v_{*i}(t)$        &   Bin-average-shifted bias potential, Eq. \eqref{eq:x_def}. \\
     $\zeta_{*i}(t)$      &   The noise of histogram. \\

### 2017-06-09 doc/invt.tex
intq.h: fixing bugs and improve stability

@@
   \equiv
-  \alpha(t) \, \eta_i(t)
+  \alpha(t) \, \xi_i(t)
   ,
@@
 and we have defined a relative histogram
-$\eta_i(t) \equiv h_i(t) /p_i$
+$\xi_i(t) \equiv h_i(t) /p_i$
 in the last step.
@@
   \approx
-  \alpha(t) \, \eta_{*i}(t)
+  \alpha(t) \, \xi_{*i}(t)
   .
@@
 We can split the histogram into
-a deterministic averaged part, $\langle \eta_i(t) \rangle$,
+a deterministic averaged part, $\langle \xi_i(t) \rangle$,
 and a random fluctuation part, $\zeta_i(t)$,
@@
 \begin{equation}
-  \eta_i(t) =
-  \langle \eta_i(t) \rangle
+  \xi_i(t) =
+  \langle \xi_i(t) \rangle
   +
@@
 The deterministic part can be related
-to an average in an ensemble consists of many similar simulation copies
+to an average in an ensemble consists of
+many copies of similar simulations
 sharing the same schedule $\alpha(t)$
@@
 \begin{align}
-  \langle \eta_i(t) \rangle
+  \langle \xi_i(t) \rangle
   %&=
@@
 \begin{align}
-  \langle \eta_{*i}(t) \rangle
+  \langle \xi_{*i}(t) \rangle
   \equiv
-  \langle \eta_i(t) \rangle -
-  \sum_{j=1}^n \rho_j \langle \eta_j(t) \rangle
+  \langle \xi_i(t) \rangle -
+  \sum_{j=1}^n \rho_j \langle \xi_j(t) \rangle
   &\approx
@@
 $\sum_{t = -\infty}^\infty \langle \zeta_{*i}(t) \, \zeta_{*i}(0) \rangle$.
-%
-\note{More explicitly,
-if the transition matrix of a period, $\tau$,
-is $T^\tau_{ij}$,
-then
-$
-  p_i p_j \langle \zeta_i(t) \, \zeta_j(t') \rangle
-  =
-  \langle h_i(t) \, h_j(t') \rangle
-  -
-  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
-  =
-  T^{t - t'}_{ij} \, \pi_j - \pi_i \, \pi_j.
-$
-}%
 
@@
   =
-  \alpha(t) \, \left[ \langle \eta_{*i}(t) \rangle + \zeta_{*i}(t) \right]
+  \alpha(t) \, \left[ \langle \xi_{*i}(t) \rangle + \zeta_{*i}(t) \right]
   \approx
@@
 %
-and the formal solution is % of Eq. \eqref{eq:dxdt_singlebin} is
+and the formal solution is
 %
@@
 %
-%We will further demand that
-%%
-%\begin{equation}
-%  \lim_{t \to \infty} q(t) \to \infty.
-%  \label{eq:qt_limit}
-%\end{equation}
-%%
-%Then, for a large $t$,
-%the second term on the right-hand side
-%of Eq. \eqref{eq:xt_solution} can be neglected, and
-%%
-%\begin{align}
-%\left\langle x^2(t) \right\rangle
-%%&=
-%%\int_0^t \int_0^t \dot y(t') \, \dot y(t'')
-%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
-%%\notag
-%%\\
-%&=
-%\int_0^t \int_0^t
-%  \dot y\bigl( q(t') \bigr) \,
-%  \dot y\bigl( q(t'') \bigr) \,
-%  \kappa(t' - t'') \, dt'' \, dt'.
-%\label{eq:x2t_average_asym}
-%\end{align}
-%
-%By the Euler-Lagrange equation, we get
-%$$
-%\begin{aligned}
-%0
-%&=
-%\frac{d}{dt} \int_0^T
-%  \dot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
-%\\
-%&= \int_0^T
-%  \ddot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
-%\end{aligned}
-%$$
-%where we have dropped the boundary terms
-%by assuming Eq. \eqref{eq:kappa_limit}
-%and a large $T$.
-%
-%In fact, we can repeat the process $n$ times, and
-%$$
-%\int_0^t y^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
-%\qquad (n \ge 1)
-%$$
-%
-%In order for this to hold for any $t$,
-%we must have
-%%
-%$
-%\ddot y\bigl( q(t') \bigr) = 0
-%,
-%$
-%or
 \begin{equation}
@@
 formed by the relative magnitudes, $w_{ij}$'s.
-%
-%However, we will assume some restrictions and approximations
-%to simplify the discussion.
 
@@
 %
-which means that the only the only effect
-of bin average deduction is to
+which means that the only effect
+of the bin-average deduction is to
 annihilate the first mode, $\tilde f_0$.
@@
 {\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
-{\tilde \eta}_k(t)
+{\tilde \xi}_k(t)
 ,
@@
   =
-  \alpha(t) \, \lambda_k \, {\tilde \eta}_k(t)
+  \alpha(t) \, \lambda_k \, {\tilde \xi}_k(t)
   \approx
@@
 of the histogram fluctuation,
-$\kappa_k(t) \equiv \bigl\langle {\tilde \zeta}_{*k}(0)
-\, {\tilde \zeta}_{*k}(t) \bigr\rangle$,
+$\kappa_k(t) \equiv \bigl\langle {\tilde \zeta}_k(0)
+\, {\tilde \zeta}_k(t) \bigr\rangle$,
 as:
@@
   = \sum_{t = -\infty}^\infty
-  \left\langle {\tilde \zeta}_{*k}(0)
-            \, {\tilde \zeta}_{*k}(t) \right\rangle
+  \left\langle {\tilde \zeta}_k(0)
+            \, {\tilde \zeta}_k(t) \right\rangle
   .
@@
 as an attempt of modeling
-a general histogram fluctuation, ${\tilde \zeta}_{*k}(t)$,
+a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
 as an equivalent white noise
@@
 In Appendix \ref{sec:Gamma},
-we evaluate $\Gamma_k$'s in two special cases,
-and show how to measure them
-from an adaptive FDS simulation.
-%
-%\begin{equation}
-%  \Err_A(T)
-%  =
-%  \int_0^T \int_0^T
-%  \sum_{k = 1}^{n-1}
-%    \dot y_k\bigl( q(t) \bigr) \,
-%    \dot y_k\bigl( q(t') \bigr) \,
-%    \kappa_k(t - t') \, dt \, dt'
-%  ,
-%  \notag
-%  %\label{eq:error_asym0}
-%\end{equation}
-%%
-%with
-%and
-%$\kappa_k(t - t') \equiv \left\langle \theta_k(t) \, \theta_k(t') \right\rangle$.
-%
-%\begin{align*}
-%  y_k(q')
-%  &\equiv
-%  e^{\lambda_k \, [q' - q(T)]},
-%  \\
-%  \kappa_k(t - t')
-%  &\equiv
-%  \left\langle
-%    \theta_k(t) \, \theta_k(t')
-%  \right\rangle.
-%\end{align*}
-
-%\subsubsection{White-noise approximation}
-
-%In the asymptotic regime
-%(i.e. if the simulation period, $T$, is sufficiently long),
-%
-
-%\subsection{\label{sec:mbin_opta}
-%Optimal schedule for multiple-bin updating schemes}
+we discuss several ways of estimating $\Gamma_k$'s.
+%
 
@@
   .
-\label{eq:mass_func}
+  \notag
+  %\label{eq:mass_func}
 \end{equation}
@@
 the Hamiltonian is conserved,
-i.e., the asymptotic error grows
-linearly with time at a rate of $\mathcal L$.
-%
-So we may set
+which means that the asymptotic error grows
+linearly with time at a rate of $\mathcal L$,
+and we may set
 %
@@
 %
-This serves
-as a generalization of Eq. \eqref{eq:dydt_const}.
+This serves as a generalization of Eq. \eqref{eq:dydt_const}.
 %
@@
 $\alpha(t) = \dot q(t)$.
+%
+The minimal asymptotic error is
+%
+\begin{equation}
+  \Err_A(T)
+  =
+  \mathcal L \, T
+  =
+  \frac { C^2 } { T }
+  =
+  \frac 1 T
+  \left(
+    \int_0^{ q(T) } M(Q) \, dQ
+  \right)^2
+  .
+\label{eq:error_asym2}
+\end{equation}
 
@@
 $\int_0^{q(T)} m(Q) \, dQ = 1$.
+%
+Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
+we find that for $Q = q(T) - q(t)$,
+%
+\begin{align}
+  m(Q)
+  &=
+  \frac{ 1 }
+       { T \, \alpha(t) }
+  ,
+\label{eq:mQ_invTa}
+  \\
+  \int_Q^{ q(T) }
+    m(Q') \, dQ'
+  &=
+  \frac t T
+  ,
+\label{eq:intmQ_tT}
+\end{align}
+%
+which means that each point,
+$\bigl(t, \alpha(t)\bigr)$,
+on the schedule can be mapped
+to a point,
+$\bigl(Q, m(Q)\bigr)$,
+on the mass distribution,
+such that the ordinate of the latter curve
+gives $1/(T\alpha)$,
+and the area under the curve in the domain, $[Q, q(T)]$,
+is equal to $t/T$,
+as shown in Fig. \ref{fig:massq}.
 %
@@
 
-Then, from Eq. \eqref{eq:q_opt},
-we find that for $Q = q(T) - q(t)$,
-%
-\begin{align}
-  m(Q)
-  &=
-  \frac{ 1 }
-       { T \, \alpha(t) }
-  ,
-\label{eq:mQ_invTa}
-  \\
-  \int_Q^{ q(T) }
-    m(Q') \, dQ'
-  &=
-  \frac t T
-  ,
-\label{eq:intmQ_tT}
-\end{align}
-%
-which means that each point,
-$\bigl(t, \alpha(t)\bigr)$,
-on the schedule can be mapped
-to a point,
-$\bigl(Q, m(Q)\bigr)$,
-on the mass distribution,
-such that the ordinate of the latter curve
-gives $1/(T\alpha)$,
-and the area under the curve in the domain, $[Q, q(T)]$,
-is equal to $t/T$,
-as shown in Fig. \ref{fig:massq}.
-%
-The optimized asymptotic error
-can then be rewritten as
-%
-\begin{equation}
-  \Err_A(T)
-  =
-  \frac { C^2 } { T }
-  =
-  \frac 1 T
-  \left(
-    \int_0^{ q(T) } M(Q) \, dQ
-  \right)^2
-  ,
-\label{eq:error_asym2}
-\end{equation}
-%
-which follows from Eqs.
-\eqref{eq:error_asym_Lagrangian}
-and
-\eqref{eq:Lagrangian_const}.
-
 \subsubsection{\label{sec:optinitalpha}
@@
 \begin{align}
-  a_0 \, M\bigl( q(T) \bigr)
-  +
-  \frac{ 2 }{M\bigl( q(T) \bigr)}
+  \frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
+  &+
   \sum_{k = 1}^{n-1}
@@
   \notag \\
-  = \frac{2}{T} \int_0^{q(T)} M(Q') \, dQ'
+  &= \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(Q') \, dQ'
   .
@@
 $\Gamma_k$'s depend also on the system
-and the underlying sampling process (MC or MD).
+and the underlying sampling process.
 %
@@
 
-\item
+\item \label{step:prerun}
 Run a preliminary adaptive FDS simulation
@@
 
-\item
-Compute $\Gamma_k$'s from Eq. \eqref{eq:varXt}
-    (cf. Appendix \ref{sec:Gamma_measure}).
-
-\item
+\item \label{step:gamma}
+Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
+    (cf. Appendix \ref{sec:Gamma_measure})
+    as well as the initial bias $\epsilon_k$.
+
+\item \label{step:qT}
 Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
-\footnote{Our implementation used a variant of the Newton-Raphson method
-    with bracketing, and the initial value is set to $q(T) = \ln(1+T\,a_0/2)$,
-    but other methods might be more effective\cite{press3rd}.}
+\footnote{We applied the Newton-Raphson method on $\ln q(T)$
+    with the initial value set to $q(T) = \ln(1+T\,a_0/2)$.
+    Other root-finding methods can be more effective\cite{press3rd}.}
 %
@@
 
-\item
+\item \label{step:alpha}
 Compute the optimal schedule by
@@
 \end{enumerate}
+
+The above procedure is still somewhat idealized.
+%
+The preliminary run in step \ref{step:prerun}
+needs to be sufficiently long to precisely
+estimate the $\Gamma_k$'s and the initial bias, $\epsilon_k$,
+which will be used to compute the optimal $q(T)$
+and schedule later.
 
@@
 \zeta_{*i}(t) \approx
-\eta_{*i}(t)
+\xi_{*i}(t)
 .
@@
     $\kappa_k(t)$   &   Autocorrelation function of mode $k$. \\
-    $M(Q)$          &   Mass function, Eq. \eqref{eq:mass_func}.   \\
+    $M(Q)$          &   Mass function.   \\
     $m(Q)$          &   Mass distribution,

### 2017-06-09 prog/lj/lj.c

@@ -8,6 +8,7 @@ double rcdef = 100.0;
 double tp = 3.0;
 int mcblk = 5;
 double delr = 0.01; /* spacing */
+long nstsave = 1000000; /* interval of saving data */
 const char *fnvbias = "vbias.dat";
 const char *fnvref = "vref.dat";
 const char *fnlog = "verr.log";
@@ -73,7 +74,7 @@ static double lj_metroblk(lj_t *lj, metad_t *metad)
 
 /* constant updating magnitude run
  * to compute the gamma values */
-static int gammrun(invtpar_t *m, metad_t *metad, lj_t *lj)
+static int gammrun(invtpar_t *m, metad_t *metad, lj_t *lj, long nsteps)
 {
   long t;
   int ir0, ir;
@@ -82,33 +83,36 @@ static int gammrun(invtpar_t *m, metad_t *metad, lj_t *lj)
   metad->a = m->alpha0;
   ir0 = dist01(metad, lj, &dr);
 
-  fprintf(stderr, "starting constant magnitude %g metadynamics run of %ld steps...\n",
-      metad->a, m->gam_nsteps);
-  for ( t = 1; t <= m->gam_nsteps; t++ ) {
+  fprintf(stderr, "starting gamma run of %ld steps (updating magnitude %g)...%20s\n",
+      nsteps, metad->a, "");
+  metad_varv_clear(metad);
+  metad_clearav(metad); /* clear the average correction data */
+  for ( t = 1; t <= nsteps; t++ ) {
     sacc += lj_metroblk(lj, metad);
     ir = dist01(metad, lj, &dr);
     metad_updatev(metad, ir);
+    metad_updateav(metad, ir); /* for the average histogram-corrected bias potential */
     metad->tmat[ir*metad->n + ir0] += 1;
     ir0 = ir;
-    if ( t % 10 == 0 )
-      metad_add_varv(metad);
-    if ( t % 10000 == 0 ) fprintf(stderr, "t %ld/%ld = %5.2f%%, acc %.2f%% \r", t, m->gam_nsteps, 100.*t/m->gam_nsteps, 100*sacc/t);
-    if ( t % 1000000 == 0 || t == m->gam_nsteps ) {
-      metad_getgamma_varv(metad, m->alpha0, "gamma.dat");
-      metad_getgamma_tmat(metad, 1, "tgamma.dat");
+    if ( t % 100 == 0 )
+      metad_varv_add(metad);
[…]

### 2017-06-09 doc/invt.tex
intq.h: logarithmic grid points

@@
 Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
-\footnote{We applied the Newton-Raphson method on $\ln q(T)$
-    with the initial value set to $q(T) = \ln(1+T\,a_0/2)$.
-    Other root-finding methods can be more effective\cite{press3rd}.}
+\footnote{We solved the equation for $\ln q(T)$ using
+    a variant of the Newton-Raphson method
+    with the bisection method as fallback.
+    The initial value set to $q(T) = \ln(1+T\,a_0/2)$.
+    Other general methods can be more effective\cite{press3rd}.}
 %
@@
 Compute the optimal schedule by
-numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.
-%
-%This can be done on an even grid of $N$ points
-%at $q_{[i]} = \frac{i}{N} \, q(T)$
-%($i = 0, 1, \dots, N$)
+numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.\footnote{
   We have used the trapezoidal rule\cite{press3rd}
   on an adaptively defined grid on $[0, q(T)]$ of $N+1$ points
-  such that the resulting change of $t_{[i+1]}-t_{[i]}$
+  (we chose $N=20000$)
+  such that the resulting change of $\Delta t_{[i]} \equiv t_{[i+1]}-t_{[i]}$
   is no less than $T/(2N)$ and no greater than $2T/N$.
   %
-The curve $q(t)$ is then defined on the grid points,
-$(q_{[i]}, t_{[i]})$,
-%
-and the updating magnitude can be obtained from numerical differentiation:
+  This requires a few (usually $2$ or $3$) rounds of iterations
+  to determine the constant $C$.
+  %
+  The first guess of $C$ is obtained from integrating
+  Eq. \eqref{eq:mint} on a uniform grid.
+  %
+  Then in each of the next few rounds, we determine the grid points
+  as follows.  For grid point $i = 0, 1, \dots, N - 1$,
+  we determine the stride
+  $\Delta q_{[i]} \equiv q_{[i+1]} - q_{[i]}$
+  by a few iterations.
+  Starting with the guess, $\bigl(q(T)-q_{[i]}\bigr)/(N-i)$,
+  if the resulting increment $\Delta t_{[i]}$
+  for grid point $i$ is less than $T/(2N)$,
+  we multiply $\Delta q_{[i]}$ by $3/2$
+  unless the resulting $q_{[i+1]} > q(T)$,
+  or if $\Delta t_{[i]} > 2T/N$,
+  we divide $\Delta q_{[i]}$ by $2/3$.
+  %
+  Once the grid points are defined,
+  we will determine the new value of $C$,
+  and if the new value differs from the old estimate
+  by more than $10\%$, we start another round of iteration.
+%
+%The curve $q(t)$ is then defined on the grid points,
+%$(q_{[i]}, t_{[i]})$,
+%
+The updating magnitude can then be obtained from numerical differentiation:
     $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$
@@
 with
-$q_{[ -1]} \equiv q_{[0]} = 0$,
-$q_{[N+1]} \equiv q_{[N]} = q(T)$,
-$t_{[ -1]} \equiv t_{[0]} = 0$, and
-$t_{[N+1]} \equiv t_{[N]} = T$.
+$q_{[ -1]} \equiv 0$,
+$q_{[N+1]} \equiv q(T)$,
+$t_{[ -1]} \equiv 0$, and
+$t_{[N+1]} \equiv T$.
 %
@@
 to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.
+}
 
@@
 needs to be sufficiently long to precisely
-estimate the $\Gamma_k$'s and the initial bias, $\epsilon_k$,
-which will be used to compute the optimal $q(T)$
-and schedule later.
+estimate the $\Gamma_k$'s and the initial bias, $\epsilon_k$
+(for which one may use the histogram-corrected bias potential).
+%
+However,
 

### 2017-06-09 doc/invt.tex
intq: logarithmic grid clean up

@@
     a variant of the Newton-Raphson method
-    with the bisection method as fallback.
-    The initial value set to $q(T) = \ln(1+T\,a_0/2)$.
-    Other general methods can be more effective\cite{press3rd}.}
+    with the bisection method as a fallback.
+    The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.}
 %
@@
   numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.\footnote{
-  We have used the trapezoidal rule\cite{press3rd}
-  on an adaptively defined grid on $[0, q(T)]$ of $N+1$ points
-  (we chose $N=20000$)
-  such that the resulting change of $\Delta t_{[i]} \equiv t_{[i+1]}-t_{[i]}$
-  is no less than $T/(2N)$ and no greater than $2T/N$.
-  %
-  This requires a few (usually $2$ or $3$) rounds of iterations
-  to determine the constant $C$.
-  %
-  The first guess of $C$ is obtained from integrating
-  Eq. \eqref{eq:mint} on a uniform grid.
-  %
-  Then in each of the next few rounds, we determine the grid points
-  as follows.  For grid point $i = 0, 1, \dots, N - 1$,
-  we determine the stride
-  $\Delta q_{[i]} \equiv q_{[i+1]} - q_{[i]}$
-  by a few iterations.
-  Starting with the guess, $\bigl(q(T)-q_{[i]}\bigr)/(N-i)$,
-  if the resulting increment $\Delta t_{[i]}$
-  for grid point $i$ is less than $T/(2N)$,
-  we multiply $\Delta q_{[i]}$ by $3/2$
-  unless the resulting $q_{[i+1]} > q(T)$,
-  or if $\Delta t_{[i]} > 2T/N$,
-  we divide $\Delta q_{[i]}$ by $2/3$.
-  %
-  Once the grid points are defined,
-  we will determine the new value of $C$,
-  and if the new value differs from the old estimate
-  by more than $10\%$, we start another round of iteration.
-%
-%The curve $q(t)$ is then defined on the grid points,
-%$(q_{[i]}, t_{[i]})$,
+  We have used the trapezoidal rule
+  on a $N+1$ point logarithm grid over $[0, q(T)]$:
+    $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/N}$,
+  where $i = 0, \dots, N$, and we chose $N=20000$.
+%
+The integration yields a curve $q(t)$ on the grid points,
+$(q_{[i]}, t_{[i]})$.
 %
@@
 that the PMF is sufficiently smooth
-such that the eigenmodes can be limited to $k \le K$,
+such that the eigenmodes can be limited to $k < K$,
 %
@@
 From Eqs. \eqref{eq:error_tot},
-\eqref{eq:error_res1},
+\eqref{eq:error_res},
 \eqref{eq:tk0},
@@
 for the single-bin updating scheme:
-if we set $K + 1 = n$,
+if we set $K = n$, then\note{
+  From Eq. \eqref{eq:eig_orthonormal_cols}, we have
+  $\sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle =\sum_{k=1}^{n-1} \langle {\tilde v}_k(t) \, {\tilde v}_k(0) \rangle.$
+  Then sum over $t$ from $\infty$ to $\infty$.
+}
 %
@@
 \end{equation}
-which can be shown by following the steps
-similar to those leading to Eq. \eqref{eq:xt2_sum}.
-%
-%
-%
-\note{The derivation.
-From Eq. \eqref{eq:eig_orthonormal_cols}, we have
-$\sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle
-=\sum_{k=1}^{n-1} \langle {\tilde v}_{*k}(t) \, {\tilde v}_{*k}(0) \rangle$,
-so
-\begin{align*}
-\sum_{i=1}^n \rho_i G_i
-&=\sum_{t=-\infty}^\infty\sum_{i=1}^n
-  \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle \\
-  &=\sum_{t=-\infty}^\infty\sum_{k=1}^{n-1}
-  \langle {\tilde v}_{*k}(t) \, {\tilde v}_{*k}(0) \rangle
-  =\sum_{k=1}^{n-1} \Gamma_k.
-\end{align*}
-}
 %

### 2017-06-09 prog/intq.h

@@ -74,6 +74,8 @@ static void intq_setqgrid(intq_t *intq, double qT)
   }
 }
 
+
+
 /* compute the integrand for `intq` */
 static double intq_getmassx(intq_t *intq, double dq,
     double *massK)
@@ -131,67 +133,6 @@ static double intq_getq(intq_t *intq, double qT)
 
 
 
-/* compute the optimal schedule, in terms of q(t)
- * the results are saved to tarr[0..m], qarr[0..m]
- * */
-static double intq_getq_adp(intq_t *intq, double qT)
-{
-  const double r = 1.5;
-  int i, j, round, m = intq->m;
-  double c0, c, qp, dq, yK = 0, yK1 = 0, dt, del, T = intq->T;
-
-  /* get the normalization constant */
-  c0 = intq_getq(intq, qT);
-  printf("c %g\n", c0);
-  intq->tarr[0] = 0;
-  intq->qarr[0] = 0;
-
-  /* use several rounds to for improved estimate of c0
-   * usually one or two rounds would suffice */
-  for ( round = 0; round < 10; round++ ) {
-    intq_getmassx(intq, intq->qarr[0] - qT, &yK);
-    for ( j = 1; j <= m; j++ ) {
-      del = T / m;
-      qp = intq->qarr[j-1];
-      dq = (qT - qp) / (m + 1 - j);
-      /* adjust the integration step size */
-      for ( i = 0; i < 100; i++ ) {
-        intq_getmassx(intq, qp + dq - qT, &yK1);
-        dt = c0 * (yK + yK1) * 0.5 * dq;
-        //printf("point j %d, i %d, yK %g, %g, dq %g, dt %g, del %g, slope dq/dt %g | q %g, t %g\n", j, i, yK, yK1, dq/(qT-qp), dt/T, del, dq/dt, qp/qT, intq->tarr[j-1]/T);
-        /* skip the adjustment for the last point
-         * or if the step is already too small */
-        if ( j == m || dq < 1e-3 * qT/m ) {
-          break;
-        } else if ( dt < del/2 && qp + 2*dq < qT ) {
-          dq *= r;
-        } else if ( dt > del*2 ) {
-          dq /= r;
-        } else {
-          break;
-        }
[…]

### 2017-06-13 doc/invt.tex
invt.tex: add LJ results

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{mathtools}
@@
 \begin{equation}
-  u^\mathrm{corr}_i
+  \hat u_i
   =
@@
     p_i
-    &\propto p^*_i \, e^{-v^{0, \mathrm{corr}}_i}.
+    &\propto p^*_i \, e^{-\hat u_i}.
   \end{aligned}
@@
 %
-Note that the distribution, $\pmb\rho$, is arbitrary
-and independent of
+Note that the distribution, $\pmb\rho$,
+can be independent of
 the target distribution,
 $\mathbf p = (p_1, \dots, p_n)$,
-although it is usually flat.\footnote{In
+and chosen for convenience.\footnote{In
 well-tempered metadynamics,
@@
 \end{equation}
-
-%Although the above error weighting distribution, $\pmb\rho$,
-%is usually flat,
-%i.e. $\rho_i \equiv 1/n$,
-%in practice,
-%a non-flat distribution
-%is sometimes more convenient in analyzing
-%some multiple-bin updating schemes later on.
-%
-%
-%From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
-%we have
-%
-%\begin{equation}
-%  \sum_{i = 1}^n \rho_i \, v_{*i}(t) = 0
-%  .
-%\label{eq:rhox_sum}
-%\end{equation}
-%
 
@@
 \item \label{step:qT}
-Compute the optimal $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
-\footnote{We solved the equation for $\ln q(T)$ using
-    a variant of the Newton-Raphson method
-    with the bisection method as a fallback.
-    The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.}
+Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
+%
+We implemented a variant of the Newton-Raphson method\cite{press3rd}
+(for $\ln q(T)$ to avoid negative values)
+with the bisection method\cite{press3rd} as a fallback.
+The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
 %
@@
 Compute the optimal schedule by
-  numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.\footnote{
-  We have used the trapezoidal rule
-  on a $N+1$ point logarithm grid over $[0, q(T)]$:
-    $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/N}$,
-  where $i = 0, \dots, N$, and we chose $N=20000$.
-%
-The integration yields a curve $q(t)$ on the grid points,
-$(q_{[i]}, t_{[i]})$.
-%
-The updating magnitude can then be obtained from numerical differentiation:
+  numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.
+%
+To handle Gaussian updating schemes more effectively,
+  we have chosen the $\mathcal N+1$ grid points over $[0, q(T)]$ to be given by
+  $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
+  where $i = 0, \dots, \mathcal N$, and we chose $\mathcal N=20000$.
+%
+The curve $q(t)$ is then integrated
+  using the trapezoidal rule\cite{press3rd},
+  resulting in a list of values, $\bigl(q_{[i]}, t_{[i]} \bigr)$.
+%
+We then compute the updating magnitude by numerical differentiation:
   $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$
@@
   to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.
-}
 
@@
 
-The above procedure is still somewhat idealized.
-%
-The preliminary run in step \ref{step:prerun}
-needs to be sufficiently long to precisely
-estimate the $\Gamma_k$'s and the initial bias, $\epsilon_k$
-(for which one may use the histogram-corrected bias potential).
-%
-However,
+The above procedure is not always necessary.
+%
+As shown below, for a class of bandpass updating schemes,
+  which resemble the Gaussian updating schemes,
+  the asymptotically optimal schedule
+  is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
+%
+For these updating schemes,
+  the preliminary run with a constant updating magntiude
+  can be replaced by the WL-style stagewise reduction
+  until the updating magnitude falls under
+  the value given the inverse-time prescription\cite{
+    belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
 
@@
 we will turn to the problem of finding
-optimal updating schemes below.
+optimal updating schemes.
 
@@
 
-100 particle
+Our first test system is a Lennard-Jones fluid of $N=100$ particles,
+and we wish to compute the PMF along the distance of two special particles, $r$.
+%
+The two special particles are mutually non-interacting
+but interact normally with other particles
+(they have placed on an axis parallel to a side of the cubic box),
+and the intrinsic distribution, $p^*$, is often referred to as
+the cavity distribution function, $y(r)$.
+%
+We will assume reduced units in this calculation
+%
+The reference values were pre-calculated from a long simulation.
+%
+We used the Metropolis algorithm for configuration sampling\cite{frenkel}.
+%
+Each step consists of five MC trials of displacing a random particle,
+one for one of the two special particles,
+and the rest four for the other particles.
+%
+The MC displacement along each dimension is uniform,
+with the size given by $0.1/\varrho^*$ to ensure the acceptance ratio is around 50\%,
+where $\varrho^*$ is the reduced number density.
+
+The PMF was calculated from $r = 0$ to half of the box dimension
+on a grid of size $\Delta r = 0.01$ in reduced units.
+%
+Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
+and the numbers of bins were $500$ and $250$, respectively.
+%
+We tested the single-bin, Gaussian, and bandpass updating schemes.
+%
+Each simulation had an ``equilibration'' phase
+of $10^7$ steps with a fixed updating magnitude, $a_0 = 10^{-4}$,
+followed by a production phase
+with updating magnitude given by
+either the inverse-time or optimal schedule.
+%
+Each simulation was repeated multiple times
+and the average are reported.
+%
+For multiple runs under the same condition,
+the optimal schedule was pre-computed from a set of fixed values of
+$\Gamma_k$ and $\epsilon_k$,
+although they can also be fairly accurately estimated from the data
+collected from the equilibration phase
+of each individual run.
+
+As shown in Table 1,
+both the inverse-time and optimal schedule reduced
+the error significantly.
+%
+The Gaussian scheme appeared to be superier
+to the single-bin scheme in the $\varrho^* = 0.1$ case,
+but not in the $\varrho^* = 0.8$ case,
+and in the latter case,
+an overly large width of the Gaussian
+was counterproductive.
+%
+
 \begin{table}[h]\footnotesize
   \caption{\label{tab:error_LJ}
-    Error of the LJ system
-    \note{\newline
-      The data were collected in
-      \texttt{data/tab1} by running
-      the program \texttt{invt}.
-    }%
+    Error of the LJ system.
+    The numbers in parentheses are predicted values;
+    $\Err_\mathrm{init.}$
+    denotes the error
+    at the end of the equilibration phase with a constant magnitude;
+    $\Err_\mathrm{final}$
+    denotes the error at the end of the production phase under
+    either the inverse-time ($1/t$) or optimal schedule (opt.).
   }
   \setlength{\tabcolsep}{2pt}
-  \renewcommand\arraystretch{1.2}
-  \begin{tabular} { l | p{2.2cm} p{2.4cm} p{2.0cm} }
+  \renewcommand\arraystretch{1.4}
+  \begin{tabular} { l c c c c }
     \hline
-      XXX
-    &
-    &
-    &
+    Scheme & Sched. &
+    $\Err_\mathrm{init.}$ &
+    $\Err_\mathrm{final}$
     \\
     \hline
-    Global
-    &
-    &
-    &
+    \multicolumn{4}{c}{$\varrho^* = 0.1$, $500$ bins} \\
+    \hline
+    single-bin & $1/t$
+    & $5.7(5.7)\times10^{-2}$
+    & $9.1(11.4)\times10^{-5}$
+    \\
+    %\hline
+    $\sigma=0.1$ & $1/t$
+    & $7.7(8.9)\times10^{-3}$
+    & $6.8(8.2)\times10^{-5}$
+    \\
+    %\hline
+    $\sigma=0.1$ & opt.
+    & $7.6\times10^{-3}$
+    & $3.2\times10^{-5}$
+    \\
+    %\hline
+    $\sigma=0.2$ & $1/t$
+    & $6.5\times10^{-3}$
+    & $4.0\times10^{-5}$
+    \\
+    %\hline
+    $\sigma=0.2$ & opt.
+    & $7.0\times10^{-3}$
+    & $2.5\times10^{-5}$
+    \\
+    %\hline
+    $\sigma=0.5$ & $1/t$
+    & $5.5\times10^{-3}$
+    & $3.2\times10^{-5}$
+    \\
+    %\hline
+    $\sigma=0.5$ & opt.
+    & $5.8\times10^{-3}$
+    & $2.6\times10^{-5}$
+    \\
+    %\hline
+    $K=20$ & $1/t$
+    & $8.1(9.0)\times10^{-3}$
+    & $1.6(1.8)\times10^{-5}$
     \\
     \hline
-    Local
-    &
-    &
-    &
+    \multicolumn{4}{c}{$\varrho^* = 0.8$, $250$ bins} \\
+    \hline
+    single-bin & $1/t$
+    & $1.3(1.4)$
+    & $2.7(2.7)\times10^{-3}$
+    \\
+    %\hline
+    $\sigma=0.1$ & $1/t$
+    & $1.3(1.3)$
+    & $2.9(2.8)\times10^{-3}$
+    \\
+    %\hline
+    $\sigma=0.1$ & opt.
+    & $1.2$
+    & $2.8\times10^{-3}$
+    \\
+    %\hline
+    $\sigma=0.2$ & $1/t$
+    & $1.3$
+    & $3.4\times10^{-3}$
+    \\
+    %\hline
+    $\sigma=0.2$ & opt.
+    & $1.2$
+    & $3.1\times10^{-3}$
+    \\
+    %\hline
+    $\sigma=0.5$ & $1/t$
+    & $0.95$
+    & $9.8\times10^{-3}$
+    \\
+    %\hline
+    $\sigma=0.5$ & opt.
+    & $1.0$
+    & $5.8\times10^{-3}$
+    \\
+    %\hline
+    $K=20$ & $1/t$
+    & $1.3(1.3)$
+    & $3.0(2.7)\times10^{-3}$
     \\
@@
 \section{\label{sec:equilerr}
-Inverse-time schedule as a result of increasing sample size
-in a runtime average
+Runtime average interpretation
+of the inverse-time schedule
 }
@@
 %
-We first note that the normalized histogram, $H_i(t)$,
-is a runtime average of the instantaneous histogram, $h_i(t)$.
+We first note that a runtime average
+\begin{equation}
+  X(t) = \frac{x(1) + \cdots + x(t)}{t}
+  .
+  \label{eq:avx}
+\end{equation}
+can be written as a recurrence relation
+\begin{equation}
+  X(t) = X(t - 1) + \frac{x(t) - X(t-1)}{t}
+  .
+  \label{eq:avx_recur}
+\end{equation}
+%
+Here $x(t) - X(t-1)$ serves as an independent correction
+from data point at time $t$ to the previous average value,
+and this correction is absorbed into the new value, $X(t)$,
+with a weight of $1/t$,
+which is the inverse sample size.
+
+We can use Eq. \eqref{eq:avx_recur} to incrementally
+improve the estimated
+bias potential given by Eq. \eqref{eq:vcorr_equil}.
+%
+Assuming the correction, $\hat u_i$, is done at every step,
+then
 %
 \begin{equation*}
-  H_i(t)
-  =
-  \frac{ h_i(1) + \cdots + h_i(t) } { t }
-  .
+  \hat u_i(t+1) = u_i(1) +
+  \ln \frac{ H_i(t) } { p_i }
+  \approx
+  u_i(1) +
+  \frac{ H_i(t) } { p_i } - 1
+  ,
 \end{equation*}
-%
-This formula can written as a recurrence relation regarding
-the addition of the last data point,
+where we have assumed that
+the normalized histogram, $H_i(t)$,
+is sufficiently precise in the second step.
+%
+Since the normalized histogram, $H_i(t)$,
+is an average of the instantaneous histogram, $h_i(t)$,
+$H_i(t) = \frac 1 t \sum_{\tau = 1}^t h_i(\tau)$,
+we have, from Eq. \eqref{eq:avx_recur},
 %
@@
   ,
-  \label{eq:av_recur}
+  \label{eq:avH_recur}
 \end{align}
-%
-%If we interpret $x(t) - X(t-1)$
-%as the independent correction to the previous average,
-%$X(t-1)$, proposed by the last data point, $x(t)$,
-%then this correction is absorbed into the average with
-%a diminishing weight $1/t$ as the sample size, $t$, increases.
-%
-which allows us to incrementally improve the estimated
-bias potential given by Eq. \eqref{eq:vcorr_equil}.
-%
-With
-$u^\mathrm{corr}_i \to u_i(t+1)$, $u_i \to u_i(1)$,
-we have
-%
-\begin{equation*}
-  u_i(t+1) = u_i(1) +
-  \ln \frac{ H_i(t) } { p_i }
-  \approx
-  u_i(1) +
-  \frac{ H_i(t) } { p_i } - 1
-  ,
-\end{equation*}
-where we have assumed that
-the normalized histogram, $H_i(t)$,
-is sufficiently precise in the second step.
-%
-%where the ``$-1$'' on the right-hand side
-%can be ignored as a uniform shift
-%that does not affect the distribution.
 %
@@
 \begin{align}
-  u_i(t+1)
+  \hat u_i(t+1)
   &=
-  u_i(t)
+  \hat u_i(t)
   +
@@
   &=
-  u_i(t)
+  \hat u_i(t)
   +
@@
   .
-  \label{eq:vbar_update}
+  \label{eq:vhat_update}
 \end{align}
@@
 in the asymptotic limit,
-Eq. \eqref{eq:vbar_update} recovers
+Eq. \eqref{eq:vhat_update} recovers
 the single-bin updating scheme, Eq. \eqref{eq:wl_update},
@@
 between an adaptive WL simulation under the inverse-time schedule
-and the ideal equilibrium FDS counterpart
-with a nearly perfect bias potential.
-%
-The bias potentials from the two approaches
-would reach similar precision at long times.
-%
-Thus, it is unnecessary to convert
-a sufficiently convergent adaptive FDS simulation
-to the corresponding equilibrium FDS simulation
-with a fixed bias potential just for better precision
+and one with a fixed and nearly perfect bias potential.
+%
+Since the bias potential in the former approach
+essentially follows the same time evolution
+for the correction,
+$\hat u_i(t)$, Eq. \eqref{eq:vhat_update},
+the two would reach similar precision at long times.
+%
+Thus, it is unnecessary to fix the bias potential
+after an adaptive FDS simulation has reached
+sufficiently convergent just for better precision
 of the final PMF.
@@
 Second,
+it shows close connection between
+the WL algorithm under the inverse-time schedule
+and self-healing umbrella sampling.
+In the latter,
+the bias potential is computed from
+a weighted runtime average of the histogram\cite{
+  marsili2006, barducci2008},
+which is analogous to the adaptive bias force method\cite{
+  darve2001, *darve2008}.
+
+Third,
 it also relates a WL simulation under the inverse-time schedule
@@
 which is ignored in this analysis.
-
-Third,
-it is possible to reformulate the WL algorithm
-under the inverse-time schedule
-as an explicit adaptive average method
-such that the bias potential is computed from
-a weighted runtime average of the histogram\cite{
-  marsili2006, barducci2008},
-which is analogous to the adaptive bias force method\cite{
-  darve2001, *darve2008}.
 

### 2017-06-13 doc/invt.tex
invt.tex: remove appendix on Gamma

@@
 $\mathbf p = (p_1, \dots, p_n)$,
-and chosen for convenience.\footnote{In
-well-tempered metadynamics,
+and chosen for convenience.\footnote{Alternatively,
+we may define $v_{*i}(t)$ such that
+$\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
+i.e.,
+$$
+v_{*i}(t) \to -\log\left(
+  \frac{ (p^*_i/p_i) \, e^{ -u_i(t) } } 
+  { \sum_{j=1}^n \rho_j \, (p^*_j/p_j) \, e^{ -u_j(t) } } 
+\right).
+$$
+One can readily show that the two definitions are equivalent
+for small $v_{*i}$.
+%
+In well-tempered metadynamics,
 $\mathbf p$ is an attenuated version of
@@
 approximate the finite difference Eq. \eqref{eq:wl_update}
-by a differential equation\footnote{This
-approximation requires us to shift the origin of time by $1$ in some cases.}
+by a differential equation\footnote{In
+going to the continuous time setup,
+we find it more convenient to shift the origin of time by $-1$,
+e.g. the sum $\sum_{i=1}^T$ is mapped to the integral $\int_0^T dt$.}
 %
@@
 or zero otherwise,
-and we have defined a relative histogram
-$\xi_i(t) \equiv h_i(t) /p_i$
-in the last step.
+and we have defined
+$\xi_i(t) \equiv h_i(t) /p_i$.
 %
@@
 The deterministic part can be related
-to an average in an ensemble consists of
+to an average in an ensemble consisting of
 many copies of similar simulations
@@
 %
-For non-white noise,
-the equivalent value of $G_i$ can be approximately obtained as
-$\sum_{t = -\infty}^\infty \langle \zeta_{*i}(t) \, \zeta_{*i}(0) \rangle$.
+%For non-white noise,
+%the equivalent value of $G_i$ can be approximately obtained as
+%$\sum_{t = -\infty}^\infty \langle \zeta_{*i}(t) \, \zeta_{*i}(0) \rangle$.
 
@@
 where $t_0$ is some constant,
-and the other constant, $c$, has been determined
+and $c$ has been determined
 from the $t = T$ case
@@
 %
-which is minimized if
+which is minimal at
 %
@@
 %
-and at this point, the error is given by
+and the minimum is
 %
@@
 This error is roughly equal to
-the overall histogram flatness\cite{zhou2005, zhou2008}
-and thus it can be conveniently estimated
-(see also Appendix \ref{sec:flatness}).
+the average histogram fluctuation\cite{zhou2005, zhou2008}
+(see also Appendix \ref{sec:hfluc}).
 
@@
 %
-This requires the updating matrix, $\mathbf w$
-to be symmetric
-for a uniform $\pmb\rho$,
-or more generally,
-the scaled matrix, $\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
-to be symmetric.
-%
-This symmetry allows the diagonalization
+This requires the scaled updating matrix,
+$\sqrt{ \rho_i/\rho_j } \, w_{ij}$,
+to be symmetric,
+and the symmetry allows the diagonalization
 of $\mathbf w$ with a set of
@@
 %
-We may write the updating matrix in $\phi_{ki}$'s
-as\cite{bussi2006}
-%
-\begin{equation}
-  w_{ij}
-  =
-  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
-  \lambda_k \, \phi_{ki} \, \phi_{kj}
-  .
-  \notag
-\end{equation}
+%We may write the updating matrix in $\phi_{ki}$'s
+%as\cite{bussi2006}
+%%
+%\begin{equation}
+%  w_{ij}
+%  =
+%  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
+%  \lambda_k \, \phi_{ki} \, \phi_{kj}
+%  .
+%  \notag
+%\end{equation}
 
@@
 $\mathcal F[1]_k = \delta_{k0}$,
-and for the bin average of any quantity, $f$, we have
-$\bav{f}= \mathcal F[f_i]_0 = \tilde f_0$,
-so that
+%
+and Eq. \eqref{eq:eigenmode0} means that
+the bin average of any quantity, $f$,
+is the first mode of the Fourier transform:
+$\bav{f}= \mathcal F[f_i]_0 = \tilde f_0$.
+%
+So
 \begin{equation}
@@
 which means that the only effect
-of the bin-average deduction is to
+of the bin-average deduction on $\tilde f_k$ is to
 annihilate the first mode, $\tilde f_0$.
@@
 and used the transformed version of Eq. \eqref{eq:sh_ave}
-[the subscript ``$_*$'' is dropped by Eq. \eqref{eq:ftstar}].
+[but the subscript ``$_*$'' is dropped by Eq. \eqref{eq:ftstar}].
 In the second step, we have used the linear approximation
@@
 %
-We may understand Eq. \eqref{eq:Gamma_sum}
-as an attempt of modeling
+Equation \eqref{eq:Gamma_sum} provides a way
+of approximately modeling
 a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
-as an equivalent white noise
-so that the theory can be applicable to
-a general sampling process.
-%
+as an equivalent white noise.
 In this way,
@@
 %
-For example, we expect large values of $\Gamma_k$
-from eigenmodes corresponding to
-persistent long-range motion
-in sampling complex or glassy systems.
-%
-In Appendix \ref{sec:Gamma},
-we discuss several ways of estimating $\Gamma_k$'s.
+In Appendix \ref{sec:Gamma_measure},
+we discuss a way of estimating $\Gamma_k$
+from an FDS simulation with a constant magnitude.
 %
@@
 the preliminary adaptive FDS sampling has been
-performed under a constant updating magnitude for $T_0$, steps.
+performed under a constant updating magnitude for $T_0$ steps.
 %
@@
   $\sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle =\sum_{k=1}^{n-1} \langle {\tilde v}_k(t) \, {\tilde v}_k(0) \rangle.$
-  Then sum over $t$ from $\infty$ to $\infty$.
+  Then sum over $t$ from $-\infty$ to $\infty$.
 }
@@
 
-Our first test system is a Lennard-Jones fluid of $N=100$ particles,
+Our test system is a Lennard-Jones fluid of $N=100$ particles,
 and we wish to compute the PMF along the distance of two special particles, $r$.
@@
 The two special particles are mutually non-interacting
-but interact normally with other particles
-(they have placed on an axis parallel to a side of the cubic box),
-and the intrinsic distribution, $p^*$, is often referred to as
-the cavity distribution function, $y(r)$.
-%
-We will assume reduced units in this calculation
-%
-The reference values were pre-calculated from a long simulation.
-%
+but they interact normally with other particles,\footnote{they
+have placed on an axis parallel to a side of the cubic box}
+and the intrinsic distribution, $p^*$, is
+the cavity distribution function, $y(r)$\cite{hansen}.
+%
+We will assume reduced units in this calculation.
+%
+The reference bias potential was pre-calculated from a long simulation.
+
 We used the Metropolis algorithm for configuration sampling\cite{frenkel}.
 %
-Each step consists of five MC trials of displacing a random particle,
-one for one of the two special particles,
-and the rest four for the other particles.
-%
-The MC displacement along each dimension is uniform,
+Each step consists of
+one MC trial for displacing one of the two special particles,
+and another four trials for displacing the other particles.
+%
+In each trial, the displacement along each dimension is uniform,
 with the size given by $0.1/\varrho^*$ to ensure the acceptance ratio is around 50\%,
@@
 %
-Each simulation had an ``equilibration'' phase
+Each simulation started with an ``equilibration'' phase
 of $10^7$ steps with a fixed updating magnitude, $a_0 = 10^{-4}$,
@@
 %
-Each simulation was repeated multiple times
-and the average are reported.
-%
-For multiple runs under the same condition,
+We repeated each simulation multiple times
+and report the average.
+%
+To make sure the same schedule for multiple runs under the same condition,
 the optimal schedule was pre-computed from a set of fixed values of
@@
 although they can also be fairly accurately estimated from the data
-collected from the equilibration phase
-of each individual run.
+collected from the equilibration phase of each individual run.
 
@@
 from data point at time $t$ to the previous average value,
-and this correction is absorbed into the new value, $X(t)$,
+and this correction is absorbed into the average
 with a weight of $1/t$,
@@
   \frac{ H_i(t) } { p_i } - 1
-  ,
+  .
 \end{equation*}
-where we have assumed that
-the normalized histogram, $H_i(t)$,
+%
+Here the normalized histogram, $H_i(t)$,
+is a runtime average of the instantaneous histogram, $h_i(t)$,
+%
+\begin{equation}
+  H_i(t) = \frac 1 t \sum_{\tau = 1}^t h_i(\tau),
+  \label{eq:hav_def}
+\end{equation}
+%
+and we have assumed that it
 is sufficiently precise in the second step.
 %
-Since the normalized histogram, $H_i(t)$,
-is an average of the instantaneous histogram, $h_i(t)$,
-$H_i(t) = \frac 1 t \sum_{\tau = 1}^t h_i(\tau)$,
-we have, from Eq. \eqref{eq:avx_recur},
+Applying Eq. \eqref{eq:avx_recur} to the histogram yields
 %
@@
   H_i(t-1) + \frac{ h_i(t) - H_i(t-1) } { t }
-  ,
+  .
   \label{eq:avH_recur}
@@
   +
-  \frac{1}{t} \frac{ h_i(t) }{ p_i }
-  -\frac{1}{t} \frac{ H_i(t-1) }{ p_i }
+  \frac{1}{t} \left[
+    \frac{ h_i(t) }{ p_i }
+   -\frac{ H_i(t-1) }{ p_i }
+  \right]
   .
@@
 %
-Now if we drop the last term,
-which tends to a uniform shift of the bias potential
-in the asymptotic limit,
-Eq. \eqref{eq:vhat_update} recovers
+In the long time limit, we may assume that $H_i(t-1) \to p_i$
+and the second term in the square bracket
+becomes a constant shift that can be safely dropped.
+%
+Then Eq. \eqref{eq:vhat_update} is identical to
 the single-bin updating scheme, Eq. \eqref{eq:wl_update},
-under
-the inverse-time schedule, Eq. \eqref{eq:alpha_invt}.
-
-The above demonstration implies several connections
-between the WL algorithm under the optimal inverse-time schedule
-with a few related FDS methods.
-%
-First,
-it establishes an efficiency equivalence
-between an adaptive WL simulation under the inverse-time schedule
-and one with a fixed and nearly perfect bias potential.
-%
-Since the bias potential in the former approach
-essentially follows the same time evolution
-for the correction,
-$\hat u_i(t)$, Eq. \eqref{eq:vhat_update},
-the two would reach similar precision at long times.
-%
-Thus, it is unnecessary to fix the bias potential
-after an adaptive FDS simulation has reached
-sufficiently convergent just for better precision
-of the final PMF.
-
-Second,
-it shows close connection between
-the WL algorithm under the inverse-time schedule
-and self-healing umbrella sampling.
-In the latter,
-the bias potential is computed from
-a weighted runtime average of the histogram\cite{
-  marsili2006, barducci2008},
-which is analogous to the adaptive bias force method\cite{
-  darve2001, *darve2008}.
-
-Third,
-it also relates a WL simulation under the inverse-time schedule
-to a time-averaged variant of the regular WL simulation.
-%
-In the variant, the updating magnitude is fixed,
-but the bias potential over the simulation course is averaged
-as the final result\cite{zhou2005}.
-%
-The above demonstration shows that the inverse-time schedule
-has already effectively implemented an implicit average
-so that no explicit averaging is needed in the former approach.
-%
-The inverse-time schedule approach is preferred, however,
-because with a decreasing updating magnitude,
-it minimizes the systematic error caused by
-the on-the-fly updating scheme\cite{
-  zhou2005, zhou2008},
-which is ignored in this analysis.
+under the inverse-time schedule, Eq. \eqref{eq:alpha_invt}.
+%
+In other words, the bias potential obtained from the latter approach
+represents a runtime average.
+%Thus, the bias potential from a WL simulation
+%(which typically uses the single-bin updating scheme)
+%under the inverse-time schedule
+%would reach a similar precision
+%as an ideal multicanonical simulation
+%with fixed and nearly perfect bias potential.
+%the precision at long times.
+%%
+%Thus, it is unnecessary to fix the bias potential
+%after an adaptive FDS simulation has reached
+%sufficiently convergent just for better precision
+%of the final PMF.
+
+%Second,
+%it shows close connection between
+%the WL algorithm under the inverse-time schedule
+%and self-healing umbrella sampling.
+%In the latter,
+%the bias potential is computed from
+%a weighted runtime average of the histogram\cite{
+%  marsili2006, barducci2008},
+%which is analogous to the adaptive bias force method\cite{
+%  darve2001, *darve2008}.
+%
+%Third,
+%it also relates a WL simulation under the inverse-time schedule
+%to a time-averaged variant of the regular WL simulation.
+%%
+%In the variant, the updating magnitude is fixed,
+%but the bias potential over the simulation course is averaged
+%as the final result\cite{zhou2005}.
+%%
+%The above demonstration shows that the inverse-time schedule
+%has already effectively implemented an implicit average
+%so that no explicit averaging is needed in the former approach.
+%%
+%The inverse-time schedule approach is preferred, however,
+%because with a decreasing updating magnitude,
+%it minimizes the systematic error caused by
+%the on-the-fly updating scheme\cite{
+%  zhou2005, zhou2008},
+%which is ignored in this analysis.
 
@@
 
-\section{\label{sec:flatness}
-Histogram flatness}
-
-Here we wish to show that for an optimal WL simulation
-(also for an FDS simulation with a perfect bias potential),
-the overall histogram flatness
+\section{\label{sec:hfluc}
+Histogram fluctuation}
+
+Here we wish to show that for an FDS simulation
+with a nearly-perfect bias potential
+(e.g. a long WL simulation under the single-bin updating scheme
+with the inverse-time schedule),
+the histogram fluctuation
 is roughly equal to the final error.
@@
 the instantaneous histogram is dominated by the random part,
-$h_i(t) \approx \zeta_i(t)$,
-and $\zeta_{*i}(t)$
-will dominate the shifted instantaneous histogram
-%
-\begin{equation}
-\zeta_{*i}(t) \approx
-\xi_{*i}(t)
-.
-\label{eq:xi_h}
-\end{equation}
-%
-If we define the overall or time-averaged histogram as
-\begin{equation}
-  H_i \equiv \frac{1}{T} \sum_{t=1}^T h_i(t)
-  ,
-  \label{eq:H_def}
-\end{equation}
-then the time average of the right-hand side of Eq. \eqref{eq:xi_h}
-represents the fractional deviation of the shifted histogram.
-%
-We way thus define the histogram flatness as
-%
-\begin{equation}
-  F
-  \equiv
+%
+\begin{equation}
+  \xi_i(t) \approx \zeta_i(t)
+  ,
+  \label{eq:xi_zeta}
+\end{equation}
+and we would have for the Fourier transform,
+$\tilde \xi_k(t) \approx \tilde \zeta_k(t).$
+%
+If we define the time average,
+$$
+\tilde \Xi_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde \xi_k(t)
+=\mathcal F\left[ \frac{ H_i } { p_i } \right]_k
+,
+$$
+then the histogram fluctuation of mode $k$ is given by
+$\left\langle \tilde \Xi_k^2(T) \right\rangle$,
+and from Eq. \eqref{eq:xi_zeta}, we have
+\begin{align*}
+\left\langle \tilde \Xi_k^2(T) \right\rangle
+&=
+\frac{1}{T}
+\sum_{t=-\infty}^\infty
 \left\langle
-  \sum_{i = 1}^n \rho_i
-  \left(
-  \frac{ H_i } { p_i }
-  - \sum_{r = 1}^n \rho_r \frac{ H_r } { p_r }
-  \right)^2
+  \tilde \xi_k(t) \, \tilde \xi_k(0)
 \right\rangle
-  .
-  \label{eq:F_def}
-\end{equation}
-%
-By using Eqs. \eqref{eq:H_def}, \eqref{eq:xi_h},
-\eqref{eq:G_def}, and \eqref{eq:Gamma_def}, we have
-%
-\begin{align*}
-  F
-  %&\approx
-  %\sum_{i = 1}^n \rho_i
-  %\left\langle
-  %\left(
-  %\frac{1}{T}
-  %\sum_{t = 1}^T \zeta_{*i}(t)
-  %\right)^2
-  %\right\rangle
-  %\\
+\\
 &\approx
-  \frac{1}{T^2}
-  \sum_{i = 1}^n \rho_i
-  \sum_{t, t' = 1}^T
+\frac{1}{T}
+\sum_{t=-\infty}^\infty
 \left\langle
-  \zeta_{*i}(t) \, \zeta_{*i}(t')
+  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
 \right\rangle
-  \approx
-  \frac{1}{T}
-  \sum_{i = 1}^n \rho_i \, G_i
-  =
-  \frac{ \Gamma } { T }
+=
+\frac{ \Gamma_k } { T }
 .
@@
 %
-Compare the last expression to Eq. \eqref{eq:Emin_singlebin},
+The total histogram fluctuation is
+$$
+F(T)
+=
+\sum_{k=1}^{n-1} \left\langle \tilde \Xi_k^2(T) \right\rangle
+=
+\frac{ \Gamma } { T }.
+$$
+%
+Comparing this to Eq. \eqref{eq:Emin_singlebin},
 we conclude
 \begin{equation}
-  F \approx E(T)
+  F(T+t_0) \approx E(T)
   .
   \notag
-  %\label{eq:flatness}
-\end{equation}
-%and the histogram flatness
+  %\label{eq:fluc}
+\end{equation}
+%and the histogram fluc
 %serves as a good estimate of the final error.
+%
+\note{
 Note that the expected histogram flatness
@@
 is compatible with ours, Eq. \eqref{eq:error_def}, in the long-time limit.
-%
-\note{
 An interesting application\cite{zhou2008},
@@
 
-\section{\label{sec:Gamma}
-Integrals of the autocorrelation functions}
-
-Here, we evaluate the integrals of
-the autocorrelation functions,
-$\Gamma_k$'s,
-in two special cases
-in Secs. \ref{sec:Gamma_perfect}
-and \ref{sec:Gamma_onestep}
-in the asymptotic limit,
-and describe a general method of measuring them
-in Sec. \ref{sec:Gamma_measure}.
-
-\subsection{\label{sec:Gamma_perfect}
-Perfect sampling}
-
-First, if the sampling is perfect,
-%
-%\begin{equation}
-%  \Gamma_k = 1.
-%\label{eq:Gamma_perfect}
-%\end{equation}
-%
-%To see this, we note that
-the bin index, $i$, is an independent random variable
-at each time step $t$.
-%
-Further, in the asymptotic limit,
-$\langle h_i(t) \rangle = p_i$,
-we find from Eq. \eqref{eq:h_split} that
+\section{\label{sec:Gamma_measure}
+Measurement of the integrals of the autocorrelation function
+}
+
+Here we give a method of measuring
+the integrals of the autocorrelation functions, $\Gamma_k$'s,
+in an adaptive FDS simulation
+with a constant updating magnitude.
+%
+This method has the advantage of not requiring
+explicit computation of 
+the autocorrelation functions of the eigenmodes.
+%
+The basic idea is to use Eq. \eqref{eq:xt2_eql}.
+%
+However,
+without knowing the intrinsic distribution, $p^*_i$,
+we cannot compute the shifted bias potential, $v_i$.
+%
+The problem can be avoided by replacing the shift
+by the time average of the bias potential.
+
+We first note that the difference, 
+$u_i(t) - v_i(t) = \ln(p^*_i/p_i)$, is a constant of time,
+so is its Fourier transform,
+$\tilde u_k(t) - \tilde v_k(t)$.
+%
+It follows that
+${\tilde u}_{k}$ and ${\tilde v}_{k}$ would also share the same variance.
+%
+But since the ensemble average of ${\tilde v}_{k}$
+is zero, we have
 \begin{equation*}
-  \zeta_{*i}(t)
-  =
-  \frac{ h_i(t) } { p_i }
-  -
-  \sum_{ r = 1 }^n
-    \rho_r \frac{ h_r(t) } { p_r }
-  ,
+  \operatorname{var} {\tilde u}_{k}
+  =
+  \operatorname{var} {\tilde v}_{k}
+  =
+  \left\langle {\tilde v}_{k}^2 \right\rangle
+  .
 \end{equation*}
 %
-and
-%
-\begin{align}
-  \left\langle
-    \zeta_{*i}(t) \, \zeta_{*j}(t')
-  \right\rangle
-  %&=
-  %\left\langle
-  %  \left(
-  %    \frac{ h_i(t) } { p_i }
-  %    -
-  %    \sum_{ r = 1 }^n
-  %      \rho_r \frac{ h_r(t) } { p_r }
-  %  \right)
-  %\right.
-  %\notag \\
-  %&\hphantom{==}\cdot
-  %\left.
-  %  \left(
-  %    \frac{ h_j(t') } { p_i }
-  %    -
-  %    \sum_{ s = 1 }^n
-  %      \rho_s \frac{ h_s(t') } { p_s }
-  %  \right)
-  %\right\rangle
-  %\notag
-  %\\
-  &=
-  \left(
-    \frac{ \delta_{ij} } { p_i }
-    -
-    \frac{ \rho_i } { p_i }
-    -
-    \frac{ \rho_j } { p_j }
-    +
-    \sum_{r = 1}^n
-    \frac{ \rho_r^2 } { p_r }
-  \right) \,
-  \delta_{t, t'}
-  .
-\label{eq:zz_perfect}
-\end{align}
-%
-%$\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
-%in the continuous limit.
-%
-\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
-  with $t \ne t'$ (going back to the discrete time case),
-  $$
-  \begin{aligned}
-  \left\langle
-    \zeta_{*i}(t) \, \zeta_{*j}(t')
-  \right\rangle
-  &=
-  \left\langle
-    \zeta_{*i}(t)
-  \right\rangle
-  \left\langle
-    \zeta_{*j}(t')
-  \right\rangle
-  =
-  0.
-  \end{aligned}
-  $$
-  Next, for $t = t'$, we have
-  $$
-  \langle
-    h_i(t) \, h_j(t)
-  \rangle
-  = \delta_{ij} \, p_i
-  ,
-  $$
-  which leads to Eq. \eqref{eq:zz_perfect}.
-}
-%
-Thus,
-for $k > 0$, we get
-%
-\begin{equation}
-  \kappa_k(t - t')
-  =
-  \sum_{i,j = 1}^n
-  \phi_{ki} \phi_{kj}
-  \langle \zeta_{*i}(t) \, \zeta_{*j}(t') \rangle
-  %
-  =
-  \sum_{i = 1}^n \frac{ \phi_{ki}^2 } { p_i }
-  \, \delta_{t, t'},
-\notag
-%\label{eq:kappa_perfect}
-\end{equation}
-%
-where we have used
-Eqs. \eqref{eq:ortho1},
-\eqref{eq:zz_perfect}.
-%
-Thus, the white-noise approximation is exact,
-with
-$\Gamma_k = \sum_{ i = 1}^n \phi_{ki}^2 / p_i$.
-\note{Proof.%Eq. \eqref{eq:kappa_perfect},
-  $$
-  \begin{aligned}
-  \kappa_k(t - t')
-  &=
-  \sum_{i,j = 1}^n
-  \phi_{ki}
-  \phi_{kj}
-    \langle \zeta_{*i}(t) \, \zeta_{*j}(t') \rangle
-  \\
-  %
-  &=
-  \sum_{i,j = 1}^n
-  \phi_{ki}
-  \phi_{kj}
-  \left(
-    \frac{ \delta_{ij} } { p_i }
-    -
-    \frac{ \rho_i } { p_i }
-    -
-    \frac{ \rho_j } { p_j }
-    +
-    \sum_{r = 1}^n
-    \frac{ \rho_r^2 } { p_r }
-  \right) \,
-  \, \delta_{t, t'}
-  .
-  \end{aligned}
-  $$
-  where,
-  we have used
-  Eq. \eqref{eq:zz_perfect}
-  on the second line.
-  %
-  With Eq. \eqref{eq:ortho1},
-  only the first term in the parentheses survives.
-}
-In particular, if $p_i = \rho_i$, then, by
-Eq. \eqref{eq:eig_orthonormal_rows}, we have
-$\kappa_k(t - t') = \delta_{t, t'}$, or
-\begin{equation}
-  \Gamma_k = 1
-  .
-\label{eq:Gamma_perfect}
-\end{equation}
-
-\subsection{\label{sec:Gamma_onestep}
-One-step process}
-
-Another special case is the
-one-dimensional one-step process\cite{vankampen},
-which models a local sampling process.
-In each step, the system hops to
-one of the two neighboring bins
-with equal probability, $1/2$.
-%
-The transition matrix is
-%
-\begin{equation}
-  \arraycolsep=3.6pt\def\arraystretch{1.4}
-  \mathbf A
-  =
-  \left(
-    \begin{array}{cccccccc}
-      \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
-      \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
-      \vdots & &  & & \vdots \\
-      \frac {1 - \sigma} 2 & \dots &  & \frac 1 2 & \frac \sigma 2
-    \end{array}
-  \right),
-\notag
-%\label{eq:T_nn}
-\end{equation}
-%
-where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
-%
-For many updating schemes
-(cf. Sec. \ref{sec:band-matrix}),
-we can further assume that
-the transition matrix $\mathbf A$ and
-the updating matrix $\mathbf w$
-share the eigenvectors,
-with $\rho_i = p_i = 1/n$.
-%but not the eigenvalues.
-%
-The eigenvalues of $\mathbf A$ are
-$$
-  \Lambda_k
-  =
-  \cos \frac{ (k - 1) \, 2 \, \pi }
-            {       g \, n        }
-  ,
-$$
-where
-$g = 1$ for a periodic variable, or $2$ otherwise.
-%Note that these $\Lambda_k$'s are
-%not to be confused with the eigenvalues of $\mathbf w$,
-%$\lambda_k$'s.
-%
-In this case,
-the autocorrelation function, $\kappa_k(t)$,
-decays as $\kappa_k(0) \, \Lambda_k^t$,
-with $\kappa_k(0) = 1$ being the same as
-the value in perfect sampling.
-%
-Thus,
-%$\Gamma_k$ is roughly twice the autocorrelation time, and
-from Eq. \eqref{eq:Gamma_sum}, we get
-%
-\begin{equation}
-  \Gamma_k
-  =
-  1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
-  =
-  \frac{ 1 + \Lambda_k }
-       { 1 - \Lambda_k }
-  =
-  \cot^2 \left[
-    \frac{ (k - 1) \, \pi }
-         {       g \, n   }
-  \right]
-  .
-\label{eq:Gamma_onestep}
-\end{equation}
-
-\subsection{\label{sec:Gamma_measure}
-Measurement via an adaptive FDS simulation
-}
-
-%To use Eq. \eqref{eq:q_opt} in a real simulation,
-%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
-%
-For an unknown system,
-we can use Eq. \eqref{eq:xt2_eql} to
-measure $\Gamma_k$'s in
-an adaptive FDS simulation
-with a constant schedule,
-$\alpha(t) = a_0$.
-%
-%A practical difficulty is that
-However,
-we often do not know a priori the intrinsic distribution, $p^*_i$,
-and hence
-cannot compute the deviation
-of the bias potential, $u_i$.
-%
-To circumvent the problem, we
-modify Eq. \eqref{eq:xt2_eql} as
-%
-\begin{equation}
-  \operatorname{var} {\tilde X}_k
+Finally, by using Eq. \eqref{eq:xt2_eql},
+we get
+%
+\begin{equation}
+  \operatorname{var} {\tilde u}_k
   =
@@
 %
-where
-``$\operatorname{var}$''
-means the variance from the sampling (time) average,
-and
-%
-\begin{align*}
-{\tilde X}_k
-&=
-\sum_{ i = 1 }^n
-  \phi_{k i} \, (u_i - \bav{u}),
-%\label{eq:Xt_def}
-\end{align*}
-%
-with
-$
-\bav{u}
-\equiv
-\sum_{ j = 1 }^n \rho_j \, u_j.
-$
-
-To show Eq. \eqref{eq:varXt},
-we first note that
-the difference, $u_{*i} = u_i - \bav{u}$,
-differs from the actual deviation, $v_{*i}$,
-defined in Eq. \eqref{eq:x_def},
-only by a constant shift,
-$\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
-%
-Thus, the two share the same variance
-under time averaging.
-%
-It follows that the corresponding linear combinations,
-${\tilde V}_{*k}$ and ${\tilde v}_{*k}$, would also share the same variance.
-%
-But since the average of ${\tilde v}_{*k}$ is zero,
-we have
-\begin{equation*}
-\operatorname{var} {\tilde V}_{*k}
-=
-\operatorname{var} {\tilde v}_{*k}
-=
-\left\langle {\tilde v}_{*k}^2 \right\rangle.
-\end{equation*}
-Finally, by using Eq. \eqref{eq:xt2_eql},
-we reach Eq. \eqref{eq:varXt}.
-
-\note{In practice, Eq. \eqref{eq:varXt}
-in inapplicable to zero eigenvalues, $\lambda_k$'s.
-%
-However, these eigenmodes can be ignored
-as they do not contribute to the Lagrangian defined by
-Eq. \eqref{eq:Lagrangian_const}.
-}
+where the variance can be computed from trajectory averages.
+%
+A practical limitation is that Eq. \eqref{eq:varXt}
+would fail if some eigenvalue, $\lambda_k$, is close to zero.
+So it is best used for the single-bin updating scheme.
 

### 2017-06-13 prog/lj/lj.c

@@ -94,7 +94,7 @@ static int gammrun(invtpar_t *m, metad_t *metad, lj_t *lj, long nsteps)
     metad_updateav(metad, ir); /* for the average histogram-corrected bias potential */
     metad->tmat[ir*metad->n + ir0] += 1;
     ir0 = ir;
-    if ( t % 100 == 0 )
+    if ( t % m->gam_nstave == 0 )
       metad_varv_add(metad);
     if ( t % 10000 == 0 ) fprintf(stderr, "t %ld/%ld = %5.2f%%, acc %.2f%% \r", t, nsteps, 100.*t/nsteps, 100*sacc/t);
     if ( t % nstsave == 0 || t == nsteps ) {
@@ -219,12 +219,12 @@ static int work(invtpar_t *m)
     && m->gammethod != GAMMETHOD_LOAD ) {
     prodrun(m, metad, lj, 1, m->nequil, "vg.dat", &hfl, &errc); /* equilibration */
     gammrun(m, metad, lj, m->gam_nsteps);
-    /* compute the optimal schedule based the histogram-corrected
-     * bias potential obtained from the gamma run */
-    metad_getalphaerr(metad, m->opta, (double) m->nsteps,
-        m->gammethod, m->fngamma, m->sampmethod,
-        metad->vc, m->alpha0, (double) m->nequil, &m->qT,
-        m->qprec, m->alpha_nint, "alpha_vc.dat");
+    ///* compute the optimal schedule based the histogram-corrected
+    // * bias potential obtained from the gamma run */
+    //metad_getalphaerr(metad, m->opta, (double) m->nsteps,
+    //    m->gammethod, m->fngamma, m->sampmethod,
+    //    metad->vc, m->alpha0, (double) m->nequil, &m->qT,
+    //    m->qprec, m->alpha_nint, "alpha_vc.dat");
   }
 
   /* compute the optimal schedule and the error */
@@ -302,6 +302,7 @@ int main(int argc, char **argv)
   m->n = 0;
   m->alpha0 = 1e-4; /* updating magnitude for equilibration and gamma run */
   m->gam_nsteps = 10000000L;
+  m->gam_nstave = 100;
   m->nequil = 10000000L;
   m->nsteps = 10000000L;
   m->ntrials = 1000;

### 2017-06-13 doc/invt.tex
add lj_xerr.gp

@@
 both the inverse-time and optimal schedule reduced
-the error significantly.
-%
-The Gaussian scheme appeared to be superier
+the error significantly,
+and the errors roughly matched the predicted values
+calculated from the precomputed $\Gamma_k$'s and $\epsilon_k$'s.
+%
+The Gaussian scheme appeared to be superior
 to the single-bin scheme in the $\varrho^* = 0.1$ case,
-but not in the $\varrho^* = 0.8$ case,
-and in the latter case,
-an overly large width of the Gaussian
-was counterproductive.
-%
+but not in the $\varrho^* = 0.8$ case.
+%
+In either case, there was an optimal width of Gaussian
+(which was $\sigma = 0.37$ for $\varrho^* = 0.1$
+or $0.07$ for $\varrho^* = 0.8$),
+an overly large width was counterproductive.
+%
+In Figure \ref{fig:lj_xerr},
+we plotted the error components against the eigenmode index
+for the Gaussian updating scheme of width $\sigma = 0.2$.
+%
+While the Gaussian updating scheme effectively suppressed
+the initial errors for short-range (large $k$) modes,
+it also slowed down the reduction of errors of these modes.
+%
+As a result, the final errors of the long-range (small $k$) modes
+were larger than those from the single-bin scheme.
+%
+For the Gaussian updating scheme
+the inverse-time schedule tended to reduce the updating magnitude too rapid
+leaving a hump on the error profile for the middle-range modes.
+%
+The optimal schedule improved over the inverse-time schedule
+and produced a flatter profile for the final error.
 
@@
     Error of the LJ system.
-    The numbers in parentheses are predicted values;
+    The numbers have been averaged over multiple runs,
+    and the numbers in parentheses are predicted values;
     $\Err_\mathrm{init.}$
@@
     either the inverse-time ($1/t$) or optimal schedule (opt.).
+    \note{The reference values can be found on the second line of \texttt{verr.log}.}
   }
@@
     $\sigma=0.1$ & opt.
-    & $7.6\times10^{-3}$
-    & $3.2\times10^{-5}$
+    & $7.6(8.9)\times10^{-3}$
+    & $3.2(3.9)\times10^{-5}$
     \\
@@
     $\sigma=0.2$ & $1/t$
-    & $6.5\times10^{-3}$
-    & $4.0\times10^{-5}$
+    & $6.5(7.8)\times10^{-3}$
+    & $4.0(4.8)\times10^{-5}$
     \\
@@
     $\sigma=0.2$ & opt.
-    & $7.0\times10^{-3}$
-    & $2.5\times10^{-5}$
+    & $7.0(7.8)\times10^{-3}$
+    & $2.5(2.8)\times10^{-5}$
     \\
@@
     $\sigma=0.5$ & $1/t$
-    & $5.5\times10^{-3}$
-    & $3.2\times10^{-5}$
+    & $5.5(6.6)\times10^{-3}$
+    & $3.2(3.8)\times10^{-5}$
     \\
@@
     $\sigma=0.5$ & opt.
-    & $5.8\times10^{-3}$
-    & $2.6\times10^{-5}$
+    & $5.8(6.6)\times10^{-3}$
+    & $2.6(3.0)\times10^{-5}$
     \\
@@
     $K=20$ & $1/t$
-    & $8.1(9.0)\times10^{-3}$
+    & $8.1(8.9)\times10^{-3}$
     & $1.6(1.8)\times10^{-5}$
@@
     $\sigma=0.1$ & opt.
-    & $1.2$
-    & $2.8\times10^{-3}$
+    & $1.2(1.3)$
+    & $2.8(2.8)\times10^{-3}$
     \\
@@
     $\sigma=0.2$ & $1/t$
-    & $1.3$
-    & $3.4\times10^{-3}$
+    & $1.3(1.2)$
+    & $3.4(3.2)\times10^{-3}$
     \\
@@
     $\sigma=0.2$ & opt.
-    & $1.2$
-    & $3.1\times10^{-3}$
+    & $1.2(1.2)$
+    & $3.1(2.9)\times10^{-3}$
     \\
@@
     $\sigma=0.5$ & $1/t$
-    & $0.95$
-    & $9.8\times10^{-3}$
+    & $0.9(0.9)$
+    & $9.8(8.0)\times10^{-3}$
     \\
@@
     $\sigma=0.5$ & opt.
-    & $1.0$
-    & $5.8\times10^{-3}$
+    & $1.0(0.9)$
+    & $5.8(5.4)\times10^{-3}$
     \\
@@
 \end{table}
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_xerr.pdf}
+  }
+  \caption{
+    \label{fig:lj_xerr}
+    Error components for the FDS simulation
+    using the Gaussian updating scheme with $\sigma = 0.2$.
+    The optimal schedule (opt.) is more effective
+    in reducing the errors of middle-range modes
+    than the inverse-time ($1/t$) schedule.
+    The points are joined as a guide to the eyes.
+  }
+\end{center}
+\end{figure}
 
@@
 \section{\label{sec:Gamma_measure}
-Measurement of the integrals of the autocorrelation function
+Measurement of the integrals of the autocorrelation functions of the eigenmodes
 }
@@
 
-It is natural to first introduce homogeneous updating schemes
-for a periodic variable\cite{dama2014},
+For a periodic variable\cite{dama2014},
 the updating matrix, $\mathbf w$,
@@
 
-\subsection{Examples}
-
-\subsubsection{Gaussian updating scheme}
+\subsection{Gaussian updating scheme}
 
@@
 
-\subsubsection{\label{sec:homo_bandpass}
+\subsection{\label{sec:homo_bandpass}
 Bandpass updating schemes}

### 2017-06-13 prog/metad.h

@@ -59,7 +59,7 @@ static void metad_prepwin(metad_t *metad,
   xnew(metad->win, n);
   if ( gaussig > 0 ) {
     mkgauswin(gaussig, n, pbc, metad->win, &metad->winn);
-  } else if ( kc >= 0 ) {
+  } else if ( kc > 0 ) {
     mksincwin(kc, n, pbc, metad->win, &metad->winn);
   } else {
     /* copy the user window */
@@ -351,7 +351,12 @@ __inline static double metad_getx(metad_t *metad, int i)
   }
 }
 
-/* compute the mode-trunction error */
+/* estimate the mode-trunction error
+ * This routine should work for the bandpass updating scheme
+ * for estimating the bias of ignoring certain Fourier modes
+ * Since the Gaussian updating scheme does not have
+ * a clear cutoff, the routine will only give rough
+ * estimate for the truncation error */
 __inline static double metad_errtrunc(metad_t *metad,
     double *v, int *kc, const char *fntrunc)
 {

### 2017-06-14 doc/invt.tex
invt.tex: completing the LJ results

@@
 %In either case, however,
-Ideally,
-the optimal initial updating magnitude
-can be set to
-half of the previous equilibrium value,
-which amounts to a shift of the origin of time
-in the single-bin case.
+%Ideally,
+%the optimal initial updating magnitude
+%can be set to
+%half of the previous equilibrium value,
+%which amounts to a shift of the origin of time
+%in the single-bin case.
 %
@@
 \begin{align}
-  \alpha(t = 0)
+  \alpha(0)
   =
@@
   .
-  \notag %\label{eq:opt_alpha0}
+  \label{eq:opt_alpha0}
 \end{align}
@@
   \alpha( t = 0 )
-  =
+  \approx
   \frac{ a_0 }
@@
 %using the updating scheme
-with a constant updating magnitude, $a_0$.
+with constant updating magnitude, $a_0$.
 
@@
 
-Our test system is a Lennard-Jones fluid of $N=100$ particles,
+Our test system is a Lennard-Jones (LJ) fluid of $N=100$ particles,
 and we wish to compute the PMF along the distance of two special particles, $r$.
@@
 (which was $\sigma = 0.37$ for $\varrho^* = 0.1$
-or $0.07$ for $\varrho^* = 0.8$),
-an overly large width was counterproductive.
-%
-In Figure \ref{fig:lj_xerr},
-we plotted the error components against the eigenmode index
+or $0.07$ for $\varrho^* = 0.8$ in this case),
+an overly large width could be counterproductive.
+
+In Fig. \ref{fig:lj_xerr},
+we show the error components against the eigenmode index
 for the Gaussian updating scheme of width $\sigma = 0.2$.
@@
 and produced a flatter profile for the final error.
+%
+It is also possible to roughly
+recover the single-bin result by using the correction formula
+based on the histogram, Eq. \eqref{eq:UHcorr}.
+%
+This correction usually improves the accuracy of the small $k$ modes
+but not necessarily for the large $k$ modes.
+
+In Fig. \ref{fig:lj_alpha}. 
+we show the optimal schedules in a few cases.
+%
+Generally, the optimal schedule
+for the $\varrho^* = 0.8$ case
+was closer to the inverse-time schedule
+than that for the $\varrho^* = 0.1$ case.
+%
+This was probably because in the former case
+the first mode dominates the total error,
+$\Gamma_1 \approx \Gamma$,
+and the schedule was roughly optimized for this mode,
+resulting in
+$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
+as $\lambda_1 \approx 1$
+[especially when width of the Gaussian is small,
+cf. Eq. \eqref{eq:lambda_Gaussian}].
+%
+Since we have chosen a long equilibration period
+for preliminary run with a constant updating magnitude,
+the starting magnitude was usually around $a_0/2 = 5\times10^{-5}$
+[cf. Eq. \eqref{eq:half_alpha0}].
+%
+But if the Gaussian is too wide,
+such as the $\sigma = 0.6$ case,
+the eigenvalues $\lambda_k$ of some modes
+became too small to damp out the initial bias,
+resulting in nonnegligible $\epsilon_k$,
+and much larger initial updating magnitude, $\alpha(0)$
+[cf. Eq. \eqref{eq:opt_alpha0}].
+%
+In this case, our method would fail
+because the resulting bias potential would no longer be small.
 
@@
     \hline
-    \multicolumn{4}{c}{$\varrho^* = 0.1$, $500$ bins} \\
+    \multicolumn{4}{c}{
+      $\varrho^* = 0.1$,
+      $500$ bins,
+      $\Gamma_1 \approx 111$,
+      $\Gamma \approx 1.1\times10^3$.
+    } \\
     \hline
@@
     \hline
-    \multicolumn{4}{c}{$\varrho^* = 0.8$, $250$ bins} \\
+    \multicolumn{4}{c}{
+      $\varrho^* = 0.8$,
+      $250$ bins,
+      $\Gamma_1 \approx 2.0\times10^4$,
+      $\Gamma \approx 2.7\times10^4$
+    } \\
     \hline
@@
     \label{fig:lj_xerr}
-    Error components for the FDS simulation
-    using the Gaussian updating scheme with $\sigma = 0.2$.
+    Error components for the FDS simulation on the LJ system
+    using the Gaussian updating scheme with width $\sigma = 0.2$.
+    %
     The optimal schedule (opt.) is more effective
@@
     than the inverse-time ($1/t$) schedule.
-    The points are joined as a guide to the eyes.
+    The lines are a guide to the eyes.
+  }
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_alpha.pdf}
+  }
+  \caption{
+    \label{fig:lj_alpha}
+    Optimal schedule of the Gaussian updating scheme
+    for the LJ system.
+    %
+    The starting updating magnitude was usually
+    around $a_0/2 = 5\times10^{-5}$
+    as given by Eq. \eqref{eq:half_alpha0}.
+    But an overly wide Gaussian window ($\sigma = 0.6$)
+    resulted in a much larger updating magnitude.
   }
@@
 The optimal schedule depends on the updating window function,
-which is limited to a single-bin in the WL case,
-but can be extended to several neighboring bins
-taking a Gaussian shape in the metadynamics case.
+which can be limited to a single-bin (as often in the WL case),
+or extended to several neighboring bins
+taking a Gaussian shape (as often in the metadynamics case).
 %
@@
 %
-For the single-bin (WL) scheme,
-the optimal schedule from our method
+For the single-bin scheme,
+we confirmed that the optimal schedule
 is given by the inverse-time formula,
@@
 For a general multiple-bin scheme,
-including the Gaussian one used in metadynamics,
+including the Gaussian one,
 the optimal schedule is not given by a simple closed form,
@@
 %
-The initial updating magnitude is always
-half of the previous equilibrium value.
+%The initial updating magnitude is always
+%half of the previous equilibrium value.
 % the value used during equilibration.
@@
 
-Although the bandpass updating schemes
-appeared to outperform similar Gaussian schemes
-in our model study,
-they are global in nature
-and hence more expensive.
-%
-However, the performance gain
-is relatively small for short simulations
-with a local sampling process such as MD.
-%
-\note{
-One way to improve the Gaussian updating schemes
-is to use it with a hard cutoff of eigenmodes
-(as we did for Fig. \ref{fig:optacmp}).
-%
-Then the optimal schedule
-would be similar to the inverse-time one,
-and the error, with the short-wavelength eigenmodes filtered,
-would be close to that of the bandpass scheme.
-}
+We believe that one of the main appeals
+of the Gaussian updating scheme
+is its ability to maintain a smooth bias potential,
+which is useful for MD applications.
+%
+However, it requires a careful choice of the width and
+a sufficiently long equilibration phase under constant updating magnitude.
+%
+For reasonable parameters,
+one may use the histogram correction, Eq. \eqref{eq:UHcorr},
+in a post-simulation analysis
+to recover the possible accuracy loss
+and obtain a bias potential similar to
+that from the optimized single-bin scheme.
+%%
+%However, the performance gain
+%is relatively small for short simulations
+%with a local sampling process such as MD.
 
@@
 %
-We first note that a runtime average
-\begin{equation}
-  X(t) = \frac{x(1) + \cdots + x(t)}{t}
-  .
-  \label{eq:avx}
-\end{equation}
-can be written as a recurrence relation
-\begin{equation}
-  X(t) = X(t - 1) + \frac{x(t) - X(t-1)}{t}
-  .
-  \label{eq:avx_recur}
-\end{equation}
-%
-Here $x(t) - X(t-1)$ serves as an independent correction
-from data point at time $t$ to the previous average value,
-and this correction is absorbed into the average
-with a weight of $1/t$,
-which is the inverse sample size.
-
-We can use Eq. \eqref{eq:avx_recur} to incrementally
-improve the estimated
-bias potential given by Eq. \eqref{eq:vcorr_equil}.
-%
-Assuming the correction, $\hat u_i$, is done at every step,
-then
+%We first note that a runtime average
+%\begin{equation}
+%  X(t) = \frac{x(1) + \cdots + x(t)}{t}
+%  .
+%  \label{eq:avx}
+%\end{equation}
+%can be written as a recurrence relation
+%\begin{equation}
+%  X(t) = X(t - 1) + \frac{x(t) - X(t-1)}{t}
+%  .
+%  \label{eq:avx_recur}
+%\end{equation}
+%%
+%Here $x(t) - X(t-1)$ serves as an independent correction
+%from data point at time $t$ to the previous average value,
+%and this correction is absorbed into the average
+%with a weight of $1/t$,
+%which is the inverse sample size.
+
+%We can use Eq. \eqref{eq:avx_recur} to incrementally
+%improve the estimated
+%bias potential given by Eq. \eqref{eq:vcorr_equil}.
+%
+Assuming that we can linearize the histogram correction
+in Eq. \eqref{eq:vcorr_equil} as
+$\ln x \approx x - 1$ % (for $x \approx 1$),
+and apply it to the instantaneous histogram, $h_i(t)$,
+then the independent estimate of bias potential in each step
+is given by
 %
 \begin{equation*}
-  \hat u_i(t+1) = u_i(1) +
-  \ln \frac{ H_i(t) } { p_i }
-  \approx
-  u_i(1) +
-  \frac{ H_i(t) } { p_i } - 1
-  .
+  \hat u_i(t) \approx u_i(t) + \frac{ h_i(t) } { p_i } - 1
+  .
+  \label{eq:uhatt}
 \end{equation*}
 %
-Here the normalized histogram, $H_i(t)$,
-is a runtime average of the instantaneous histogram, $h_i(t)$,
-%
-\begin{equation}
-  H_i(t) = \frac 1 t \sum_{\tau = 1}^t h_i(\tau),
-  \label{eq:hav_def}
-\end{equation}
-%
-and we have assumed that it
-is sufficiently precise in the second step.
-%
-Applying Eq. \eqref{eq:avx_recur} to the histogram yields
-%
+The time average
+$\hat U_i(t) \equiv (1/t) \sum_{\tau=1}^t \hat u_i(\tau)$
+is
 \begin{align}
-  H_i(t)
-  &=
-  H_i(t-1) + \frac{ h_i(t) - H_i(t-1) } { t }
-  .
-  \label{eq:avH_recur}
+  \hat U_i(t)
+  &\approx
+  \frac{1}{t} \sum_{\tau=1}^t
+  \left[
+    u_i(\tau) + \frac{ h_i(\tau) } { p_i } - 1
+  \right]
+  .
+  \label{eq:Uhatt}
 \end{align}
 %
-Thus,
-%Since $H_i(t)$ is the average of the instantaneous histogram, $h_i(t)$,
-%we may apply Eq. \eqref{eq:av_recur}, and
+For a fixed bias potential, the above formula
+recovers the linearize Eq. \eqref{eq:vcorr_equil}.
+%
+But with the single-bin updating scheme,
+Eq. \eqref{eq:wl_update}
+we have, up to a constant shift,
 %
 \begin{align}
-  \hat u_i(t+1)
-  &=
-  \hat u_i(t)
-  +
-  \frac{ H_i(t) - H_i(t-1) }{ p_i }
+  \hat U_i(t)
+  &\approx
+  \frac{1}{t} \sum_{\tau=1}^t
+  \left[
+    u_i(\tau) +
+    \frac{ u_i(\tau + 1) - u_i(\tau) } { \alpha(t) }
+  \right]
+  ,
   \notag
   \\
-  &=
-  \hat u_i(t)
+  &\approx
+  \Delta u_i
   +
-  \frac{1}{t} \left[
-    \frac{ h_i(t) }{ p_i }
-   -\frac{ H_i(t-1) }{ p_i }
+  \sum_{\tau=1}^t
+  \left[
+    1 + \frac{1}{\alpha(\tau - 1)} - \frac{1}{\alpha(\tau)}
   \right]
-  .
-  \label{eq:vhat_update}
+  \frac{ u_i(\tau) } {t}
+  ,
+  \label{eq:Uhat_sum}
 \end{align}
-%
-In the long time limit, we may assume that $H_i(t-1) \to p_i$
-and the second term in the square bracket
-becomes a constant shift that can be safely dropped.
-%
-Then Eq. \eqref{eq:vhat_update} is identical to
-the single-bin updating scheme, Eq. \eqref{eq:wl_update},
-under the inverse-time schedule, Eq. \eqref{eq:alpha_invt}.
-%
-In other words, the bias potential obtained from the latter approach
-represents a runtime average.
+where
+$$
+\Delta u_i \equiv \frac{u_i(t+1)}{t \, \alpha(t)}
+- \frac{u_i(1)}{t \, \alpha(0)}.
+$$
+%
+With the inverse-time schedule, Eq. \eqref{eq:alpha_invt},
+each term in the square brackets vanishes,
+and if we further define $\alpha(0) \to \infty$
+$$
+u_i(t+1) = \hat U_i(t),
+$$
+the bias potential represents
+the runtime average of the correction, $\hat u_i(t)$.
+%
+Another special case is $\alpha(\tau) = 0$ for $\tau = 0, 1, \dots$,
+then Eq. \eqref{eq:Uhat_sum} is reduced to a plain average
+$$
+\hat U_i(t) \approx \frac 1 t \sum_{\tau=1}^t u_i(\tau).
+$$
+where we have assumed an equilibrium condition
+such that $u_i(t+1) \approx u_i(1)$.
+
+For other updating schemes and schedules, we may still use
+Eq. \eqref{eq:Uhatt} to construct a correction:
+%
+\begin{align}
+  \hat U_i(t)
+  &\approx
+  U_i(t)
+  +
+  \frac{ H_i(t) } { p_i } - 1
+  \approx
+  U_i(t)
+  +
+  \ln \frac{ H_i(t) } { p_i }
+  ,
+  \label{eq:UHcorr}
+\end{align}
+%
+where
+\begin{align}
+  U_i(t)
+  &= \frac 1 t \sum_{\tau = 1}^t u_i(\tau)
+  ,
+  \notag
+  %\label{eq:uav_def}
+  \\
+  H_i(t)
+  &= \frac 1 t \sum_{\tau = 1}^t h_i(\tau)
+  .
+  \label{eq:hav_def}
+\end{align}
+%
 %Thus, the bias potential from a WL simulation

### 2017-06-14 prog/invt.h

@@ -729,7 +729,7 @@ __inline static int loadgamma(int n, double *gamma,
   FILE *fp;
   char buf[128];
   int i, id;
-  double x;
+  double x, sum = 0;
 
   if ( (fp = fopen(fn, "r")) == NULL ) {
     fprintf(stderr, "cannot open %s\n", fn);
@@ -758,9 +758,10 @@ __inline static int loadgamma(int n, double *gamma,
     }
     //fprintf(stderr, "gamma(%d) = %g -> %g\n", i, gamma[i], x);
     gamma[i] = x;
+    sum += x;
   }
-  fprintf(stderr, "loaded %d gamma values %g,...%g, from %s\n",
-      i-1, gamma[1], gamma[i-1], fn);
+  fprintf(stderr, "loaded %d gamma values %g,...%g, sum %g, from %s\n",
+      i-1, gamma[1], gamma[i-1], sum, fn);
   fclose(fp);
 
   return 0;

### 2017-06-14 doc/invt.tex
intq.h: change the definition of (K -> kc) for mode cutoff

@@
   crespo2010, *atchade2011, *fort2015}.
-%One can show that with a constant magnitude,
+%One can show that under constant magnitude,
 %$\alpha(t) = \alpha > 0$,
@@
 we discuss a way of estimating $\Gamma_k$
-from an FDS simulation with a constant magnitude.
+from an FDS simulation under constant magnitude.
 %
@@
 Run a preliminary adaptive FDS simulation
-%using the updating scheme
-with constant updating magnitude, $a_0$.
+under constant updating magnitude, $a_0$.
+%
+The updating magnitude should be sufficiently small
+such that the resulting error is no greater than
+multiple of $1.0$.
 
@@
 
-We used the Metropolis algorithm for configuration sampling\cite{frenkel}.
+We used the Metropolis algorithm\cite{frenkel, metropolis1953} for configuration sampling.
 %
@@
 Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
-and the numbers of bins were $500$ and $250$, respectively.
+and the numbers of bins were $n = 500$ and $250$, respectively.
 %
@@
 Each simulation started with an ``equilibration'' phase
-of $10^7$ steps with a fixed updating magnitude, $a_0 = 10^{-4}$,
+of $10^7$ steps under constant updating magnitude,
+$a_0 = 10^{-4}$,\footnote{The equivalent $\ln f = n \, a_0$ for the WL algorithm
+was $0.05$ and $0.025$, respectively.}
 followed by a production phase
@@
 While the Gaussian updating scheme effectively suppressed
-the initial errors for short-range (large $k$) modes,
+the initial errors for short-range or large-$k$ modes,
 it also slowed down the reduction of errors of these modes.
 %
-As a result, the final errors of the long-range (small $k$) modes
+As a result, the final errors of the long-range or small-$k$ modes
 were larger than those from the single-bin scheme.
 %
-For the Gaussian updating scheme
-the inverse-time schedule tended to reduce the updating magnitude too rapid
-leaving a hump on the error profile for the middle-range modes.
+For the Gaussian updating scheme,
+the inverse-time schedule tended to reduce the updating magnitude
+too quickly
+leaving a hump on the error profile for the middle-range eigenmodes.
 %
@@
 %
-It is also possible to roughly
+It was also possible to roughly
 recover the single-bin result by using the correction formula
@@
 %
-This correction usually improves the accuracy of the small $k$ modes
-but not necessarily for the large $k$ modes.
+This correction usually benefited the small-$k$ modes
+but not necessarily for the large-$k$ modes.
 
@@
 as $\lambda_1 \approx 1$
-[especially when width of the Gaussian is small,
-cf. Eq. \eqref{eq:lambda_Gaussian}].
+[cf. Eq. \eqref{eq:lambda_Gaussian}].
 %
 Since we have chosen a long equilibration period
-for preliminary run with a constant updating magnitude,
+for the preliminary run under constant updating magnitude,
 the starting magnitude was usually around $a_0/2 = 5\times10^{-5}$
@@
 %
-But if the Gaussian is too wide,
-such as the $\sigma = 0.6$ case,
+But if the Gaussian were too wide,
+e.g. if $\sigma = 0.6$,
 the eigenvalues $\lambda_k$ of some modes
 became too small to damp out the initial bias,
-resulting in nonnegligible $\epsilon_k$,
-and much larger initial updating magnitude, $\alpha(0)$
+resulting in much larger initial updating magnitude, $\alpha(0)$
 [cf. Eq. \eqref{eq:opt_alpha0}].
@@
 In this case, our method would fail
-because the resulting bias potential would no longer be small.
+because the resulting bias potential would no longer be small,
 
@@
     The numbers have been averaged over multiple runs,
-    and the numbers in parentheses are predicted values;
+    and the predicted values are shown in parentheses for comparison;
     $\Err_\mathrm{init.}$
     denotes the error
-    at the end of the equilibration phase with a constant magnitude;
+    at the end of the equilibration phase under constant magnitude;
     $\Err_\mathrm{final}$
@@
     either the inverse-time ($1/t$) or optimal schedule (opt.).
-    \note{The reference values can be found on the second line of \texttt{verr.log}.}
+    \note{The reference values can be found
+    on the second line of \texttt{verr.log}.}%
   }
@@
     $K=20$ & $1/t$
-    & $8.1(8.9)\times10^{-3}$
+    & $8.0(8.9)\times10^{-3}$
     & $1.6(1.8)\times10^{-5}$
@@
     & $1.3(1.3)$
-    & $3.0(2.7)\times10^{-3}$
+    & $3.1(2.7)\times10^{-3}$
     \\
@@
     Error components for the FDS simulation on the LJ system
-    using the Gaussian updating scheme with width $\sigma = 0.2$.
+    using the Gaussian updating scheme of width $\sigma = 0.2$.
     %
     The optimal schedule (opt.) is more effective
-    in reducing the errors of middle-range modes
-    than the inverse-time ($1/t$) schedule.
+    in reducing the errors of middle-range eigenmodes
+    than the inverse-time schedule ($1/t$).
     The lines are a guide to the eyes.
@@
     But an overly wide Gaussian window ($\sigma = 0.6$)
-    resulted in a much larger updating magnitude.
+    led to a much larger updating magnitude.
   }
@@
 %and that the system is properly equilibrated
-%under a constant magnitude initially.
+%under constant magnitude initially.
 %
@@
 in an adaptive FDS simulation
-with a constant updating magnitude.
+under constant updating magnitude.
 %
@@
 $u_i(t) - v_i(t) = \ln(p^*_i/p_i)$, is a constant of time,
-so is its Fourier transform,
+so is the Fourier transform,
 $\tilde u_k(t) - \tilde v_k(t)$.
@@
       \right)
-  %\notag
-  %\\
-  %&=
-  %\exp\left[
-  %      -\frac{ \pi^2 }{ 2 }
-  %      \left(
-  %        \frac{ k - 1 }
-  %             {   n   }
-  %      \right)^2
-  %      \left(
-  %        \frac{  \sigma_\varphi }
-  %             { \Delta \varphi }
-  %      \right)^2
-  %    \right]
-  .
+  =
+  \exp\left[
+        -\frac 1 2
+        \left(
+         \frac{ 2 \, \pi \, k \, \sigma_z }
+              { g \, n \, \Delta z  }
+        \right)^2
+      \right]
+  ,
 %\notag
@@
 \end{align}
+where $\sigma_z$ and $\Delta z$
+are the width of the Gaussian
+and the bin width
+in the original collective variable, $z$,
+respectively.
 

### 2017-06-14 doc/invt.tex
move old figures to fig1/

@@
 %
-The optimization procedure is similar to the single-bin case.
+Similar to Eq. \eqref{eq:ut_diffeq},
+we can approximate the updating scheme
+as a differential equation,
+%
+\begin{equation}
+  \dot u_i(t)
+  \approx
+  \alpha(t) \,
+  \sum_{j=1}^n w_{ij} \, \frac{ h_j(t) } { p_j }
+  .
+  \label{eq:ut_diffeq_mbin}
+\end{equation}
+%
+The optimization procedure is similar to
+the single-bin case.
 %
@@
 in studying transition matrices\cite{vankampen}
-(we will only use results that are unaffected by the allowance
-of negative elements into $\mathbf w$).
+(we will only use properties that are unaffected
+by the admission of negative elements into $\mathbf w$).
 
@@
   ,
-  \label{eq:ut_diffeq_mbin}
+  \label{eq:vt_diffeq_mbin}
 \end{equation}
@@
 as in Sec. \ref{sec:sbin_diffeq}.
+%
+Equation \eqref{eq:vt_diffeq_mbin}
+shows that the rate of error reduction
+is now mode-dependent,
+and the minimization of the final error
+involves a compromise of the individual modes.
 
@@
 
-\subsection{\label{sec:lj}Lennard-Jones}
-
-Our test system is a Lennard-Jones (LJ) fluid of $N=100$ particles,
+%\subsection{\label{sec:lj}Lennard-Jones}
+
+Our test system is a Lennard-Jones (LJ) fluid of $100$ particles,
 and we wish to compute the PMF along the distance of two special particles, $r$.
@@
 %
+The reduced temperature was $T^* = 3$.
+%
 We tested the single-bin, Gaussian, and bandpass updating schemes.
@@
 was $0.05$ and $0.025$, respectively.}
-followed by a production phase
+followed by a production phase of another $10^7$ steps
 with updating magnitude given by
@@
 $\Gamma_k$ and $\epsilon_k$,
-although they can also be fairly accurately estimated from the data
+although they can be fairly accurately estimated from the data
 collected from the equilibration phase of each individual run.
@@
 %
+In the latter case,
+the error was dominated by the $k=1$ eigenmode
+corresponding to the longest range motion, as $\Gamma_1 \approx \Gamma$,
+and the suppression of local noise
+had little effect on reducing the total error.
+%
 In either case, there was an optimal width of Gaussian
-(which was $\sigma = 0.37$ for $\varrho^* = 0.1$
-or $0.07$ for $\varrho^* = 0.8$ in this case),
 an overly large width could be counterproductive.
-
-In Fig. \ref{fig:lj_xerr},
-we show the error components against the eigenmode index
-for the Gaussian updating scheme of width $\sigma = 0.2$.
+%
+The optimal width was $\sigma = 0.37$ for $\varrho^* = 0.1$
+or $0.07$ for $\varrho^* = 0.8$ in this case,
+although it depended on the simulation length.
+
+To further investigate the behavior of the Gaussian updating scheme,
+we show in Fig. \ref{fig:lj_xerr}
+the error components, $\langle \tilde v_k^2 \rangle$,
+against the eigenmode index
+for the $\varrho^* = 0.1$ case
+with width $\sigma = 0.2$.
 %
@@
 the initial errors for short-range or large-$k$ modes,
-it also slowed down the reduction of errors of these modes.
+it also slowed down the error reduction of these modes
+[cf. Eq. \eqref{eq:vt_diffeq_mbin} and \eqref{eq:lambda_Gaussian}].
 %

### 2017-06-15 doc/invt.tex
small changes

@@
 the index of a small interval, or a bin,
-$(z, z + \Delta z)$.
+$(z, z + \delta z)$.
 %
@@
 $$
-v_{*i}(t) \to -\log\left(
+v_{*i}(t) \to -\ln\left(
   \frac{ (p^*_i/p_i) \, e^{ -u_i(t) } }
@@
 %
-%We may write the updating matrix in $\phi_{ki}$'s
-%as\cite{bussi2006}
-%%
-%\begin{equation}
-%  w_{ij}
-%  =
-%  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
-%  \lambda_k \, \phi_{ki} \, \phi_{kj}
-%  .
-%  \notag
-%\end{equation}
+We may reconstruct the updating matrix in $\phi_{ki}$'s
+as\cite{bussi2006}
+%
+\begin{equation}
+  w_{ij}
+  =
+  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
+  \lambda_k \, \phi_{ki} \, \phi_{kj}
+  .
+  \label{eq:w_from_phi}
+\end{equation}
 
@@
 of the histogram fluctuation,
-$\kappa_k(t) \equiv \bigl\langle {\tilde \zeta}_k(0)
+$\bigl\langle {\tilde \zeta}_k(0)
 \, {\tilde \zeta}_k(t) \bigr\rangle$,
@@
   \Gamma_k
-  = \sum_{t = -\infty}^\infty \kappa_k(t)
   = \sum_{t = -\infty}^\infty
@@
 %
-If the initial bias, $\epsilon_k$, is negligible, then
+If the initial bias, $\epsilon_k$, is negligible,
+which is typically true if no eigenvalue is close to $0$,
+then
 %
@@
 (cf. Appendix \ref{sec:Gamma_measure})
-    as well as the initial bias $\epsilon_k$.
+as well as the initial bias $\epsilon_k$.\footnote{The
+  value of $\tilde u_k$ can be estimated
+  from the histogram-corrected bias potential
+  by Eq. \eqref{eq:uhatav},
+  although it can overestimate
+  $\epsilon_k$ for large-$k$ modes.
+}
 
@@
     \Gamma_k
-  .
+  =
+  \frac{ \Gamma \, T     }
+       { (T + t_0)^2 }
+  ,
 \label{eq:error_asym_singlebin}
 \end{align}
+%
+where we have used
+%
+\begin{align}
+  \sum_{k=1}^{n-1} \Gamma_k
+  &=
+  \sum_{t=-\infty}^\infty \sum_{k=1}^{n-1} \langle {\tilde v}_k(t) \, {\tilde v}_k(0) \rangle
+  \notag
+  \\
+  &=
+  \sum_{t=-\infty}^\infty \sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle
+  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
+  .
+  \notag
+  %\label{eq:Gamma_G_sum}
+\end{align}
+%
 %
@@
 %
-Mathematically, these updating schemes
-allow us to assign zero eigenvalues
-to modes that are likely absent from the PMF.
+A bandpass updating scheme is essentially
+the single-bin updating scheme
+supplied with a perfect filter
+for the noise modes,
+and zero eigenvalues are assigned to the noise modes.
 %
 For example,
-%if we can assume
 by assuming
@@
 %
-Then, the updating matrix is given by
+Then, the updating matrix is, according to
+Eq. \eqref{eq:w_from_phi},
 %
 $$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj},$$
-%
-and Eq. \eqref{eq:error_asym_singlebin} is modified as
-%
-\begin{equation}
-  \Err_A(T)
-  =
-  \frac {       T     }
-        { (T + t_0)^2 }
-  \sum_{ k = 1 }^{K-1}
-    \Gamma_k.
-%\notag
-\label{eq:error_asym_sinc}
-\end{equation}
 %
@@
 %
-From Eqs. \eqref{eq:error_tot},
-\eqref{eq:error_res},
-\eqref{eq:tk0},
-and
-\eqref{eq:error_asym_sinc},
-we get
+%From Eqs. \eqref{eq:error_tot},
+%\eqref{eq:error_res},
+%\eqref{eq:tk0},
+%and
+%\eqref{eq:error_asym_sinc},
+%we get
+The final error is given by
 %
@@
     \Gamma_k
-  .
+  ,
 \label{eq:error_sinc}
@@
 %
-This result recovers Eq. \eqref{eq:Emin_singlebin}
-for the single-bin updating scheme:
-if we set $K = n$, then\note{
-  From Eq. \eqref{eq:eig_orthonormal_cols}, we have
-  $\sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle =\sum_{k=1}^{n-1} \langle {\tilde v}_k(t) \, {\tilde v}_k(0) \rangle.$
-  Then sum over $t$ from $-\infty$ to $\infty$.
-}
-%
-\begin{equation}
-  \sum_{k=1}^{n-1} \Gamma_k
-  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
-  ,
-  \notag
-  %\label{eq:Gamma_G_sum}
-\end{equation}
-%
-This error is roughly proportional to the inverse simulation length,
-as one would expect from an equilibrium FDS simulation.
+as a generalization of Eq. \eqref{eq:Emin_singlebin}
+for the single-bin updating scheme.
+
+Since the bandpass updating scheme is
+a straightforward generalization of
+the single-bin updating scheme,
+we can also borrow the WL algorithm
+to control the updating magnitude
+in initial stages of the simulation.
+%
+However, the histogram fluctuation should
+be limited to first few modes, $k = 1, \dots, K$,
+as the bandpass updating scheme
+only partially flattens the histogram
+of these modes (cf. Appendix \ref{sec:hfluc}).
+
+%
+%This error is roughly proportional to the inverse simulation length,
+%as one would expect from an equilibrium FDS simulation.
 %In Appendix \ref{sec:equilerr},
@@
 The PMF was calculated from $r = 0$ to half of the box dimension
-on a grid of size $\Delta r = 0.01$ in reduced units.
+on a grid of size $\delta r = 0.01$ in reduced units.
 %
@@
 %
-It was also possible to roughly
+Even without the optimal schedule,
+it was also possible to roughly
 recover the single-bin result by using the correction formula
-based on the histogram, Eq. \eqref{eq:UHcorr}.
-%
-This correction usually benefited the small-$k$ modes
+based on the histogram, Eq. \eqref{eq:uhatav}.
+%
+This correction usually improves
+the estimate for the small-$k$ modes
 but not necessarily for the large-$k$ modes.
@@
 For reasonable parameters,
-one may use the histogram correction, Eq. \eqref{eq:UHcorr},
+one may use the histogram correction,
+Eq. \eqref{eq:uhatav},
 in a post-simulation analysis
@@
 %
-\begin{equation*}
+\begin{equation}
   \hat u_i(t) \approx u_i(t) + \frac{ h_i(t) } { p_i } - 1
+  = u_i(t) + \xi_i(t) - 1
   .
   \label{eq:uhatt}
-\end{equation*}
-%
-The time average
-$\hat U_i(t) \equiv (1/t) \sum_{\tau=1}^t \hat u_i(\tau)$
-is
+\end{equation}
+%
+We define a time average for variable, $f$ as
+\begin{equation}
+\overline{f}(t) \equiv \frac{1}{t} \sum_{\tau=1}^t f(\tau)
+.
+  \label{eq:timeav}
+\end{equation}
+%
+Then after time averaging, Eq. \eqref{eq:uhatt} becomes
 \begin{align}
-  \hat U_i(t)
+  \overline{ \hat u_i }(t)
   &\approx
-  \frac{1}{t} \sum_{\tau=1}^t
-  \left[
-    u_i(\tau) + \frac{ h_i(\tau) } { p_i } - 1
-  \right]
-  .
-  \label{eq:Uhatt}
+  \overline{ u_i } (t)
+  + \overline{ \xi_i } (t) - 1
+  \approx
+  \overline{ u_i } (t)
+  + \ln \frac{ H_i(t) } { p_i }
+  ,
+  \label{eq:uhatav}
 \end{align}
 %
-For a fixed bias potential, the above formula
-recovers the linearize Eq. \eqref{eq:vcorr_equil}.
-%
-But with the single-bin updating scheme,
-Eq. \eqref{eq:wl_update}
+where we have introduced average histogram
+%
+\begin{align}
+  H_i(t)
+  &= \frac 1 t \sum_{\tau = 1}^t h_i(\tau)
+  = \overline{ \xi_i } (t) \, p_i
+  .
+  \label{eq:hav_def}
+\end{align}
+
+Equation \eqref{eq:uhatav} serves as
+a generalization of Eq. \eqref{eq:vcorr_equil}
+for FDS simulations under a variable bias potential
+and it can be used to improve the estimated PMF
+after simulation.
+%
+In particular, for the single-bin updating scheme,
+Eq. \eqref{eq:wl_update},
 we have, up to a constant shift,
@@
 \begin{align}
-  \hat U_i(t)
+  \overline{ \hat u_i}(t)
   &\approx
@@
 each term in the square brackets vanishes,
-and if we further define $\alpha(0) \to \infty$
-$$
-u_i(t+1) = \hat U_i(t),
-$$
-the bias potential represents
-the runtime average of the correction, $\hat u_i(t)$.
+and if we further define $\alpha(0) \to \infty$,
+then
+$$
+u_i(t+1) = \overline{ \hat u_i}(t),
+$$
+i.e. the bias potential in this case coincides with
+the runtime average of the correction, $\hat u_i(t)$
+at any step, $t$.
 %
 Another special case is $\alpha(\tau) = 0$ for $\tau = 0, 1, \dots$,
-then Eq. \eqref{eq:Uhat_sum} is reduced to a plain average
-$$
-\hat U_i(t) \approx \frac 1 t \sum_{\tau=1}^t u_i(\tau).
-$$
-where we have assumed an equilibrium condition
-such that $u_i(t+1) \approx u_i(1)$.
-
-For other updating schemes and schedules, we may still use
-Eq. \eqref{eq:Uhatt} to construct a correction:
-%
-\begin{align}
-  \hat U_i(t)
-  &\approx
-  U_i(t)
-  +
-  \frac{ H_i(t) } { p_i } - 1
-  \approx
-  U_i(t)
-  +
-  \ln \frac{ H_i(t) } { p_i }
-  ,
-  \label{eq:UHcorr}
-\end{align}
-%
-where
-\begin{align}
-  U_i(t)
-  &= \frac 1 t \sum_{\tau = 1}^t u_i(\tau)
-  ,
-  \notag
-  %\label{eq:uav_def}
-  \\
-  H_i(t)
-  &= \frac 1 t \sum_{\tau = 1}^t h_i(\tau)
-  .
-  \label{eq:hav_def}
-\end{align}
-%
+then Eq. \eqref{eq:Uhat_sum} is reduced to a plain average\cite{zhou2005}
+$$
+\overline{\hat u_i}(t) \approx \overline{ u_i }(t).
+$$
+if we can assume an equilibrium condition
+with $u_i(t+1) \approx u_i(1)$.
+
 %Thus, the bias potential from a WL simulation
@@
 %
-First, in the long time limit, the fluctuation of
+In the long time limit, the fluctuation of
 the instantaneous histogram is dominated by the random part,
@@
 \tilde \Xi_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde \xi_k(t)
-=\mathcal F\left[ \frac{ H_i } { p_i } \right]_k
-,
-$$
-then the histogram fluctuation of mode $k$ is given by
-$\left\langle \tilde \Xi_k^2(T) \right\rangle$,
-and from Eq. \eqref{eq:xi_zeta}, we have
+=\mathcal F\left[ \frac{ H_i(T) } { p_i } \right]_k
+,
+$$
+for the histogram fluctuation of mode $k$,
+then the total fluctuation is
+%
+\begin{align}
+F(T)
+&\equiv
+\sum_{k=1}^{n-1} \tilde \Xi_k^2(T)
+=
+\sum_{i=1}^{n} \rho_i \, \Xi_{*i}^2(T)
+  \notag
+  \\
+  &=
+\left\langle
+  \sum_{i=1}^n \rho_i
+  \left(
+  \frac{H_i(T)}{p_i}
+  -
+  \sum_{j=1}^n \rho_j \frac{ H_j(T) } {p_j}
+  \right)^2
+\right\rangle
+.
+  \label{eq:hfluc_def}
+\end{align}
+From Eq. \eqref{eq:xi_zeta}, we have
 \begin{align*}
@@
 The total histogram fluctuation is
-$$
-F(T)
-=
+\begin{equation}
+\langle F(T) \rangle
+\equiv
 \sum_{k=1}^{n-1} \left\langle \tilde \Xi_k^2(T) \right\rangle
@@
 \frac{ \Gamma } { T }.
-$$
+  \label{eq:hfluc}
+\end{equation}
 %
@@
 \begin{equation}
-  F(T+t_0) \approx E(T)
+  \langle F(T+t_0) \rangle \approx E(T)
   .
@@
 %serves as a good estimate of the final error.
+%
+
+The relation between error and histogram fluctuation
+supports the WL criterion for
+checking convergence at stage transitions.
+%
+For bandpass updating scheme,
+we may borrow this criterion
+in initial stages, with a modified definiton
+for the histogram fluctuation
+$$
+F_{<K}(T) = \sum_{k=1}^{K-1} \tilde \Xi_k^2(T),
+$$
+since the histogram-flatten
+is only partially applied to modes less $K$.
 %
@@
 So it is best used for the single-bin updating scheme.
+
+Alternatively,
+if the propagation along the collective-variable is Markovian,
+then in the long time limit, the evolution can be approximately
+described by a transition matrix, $\mathbf A$, with
+$$
+\langle h_i(t) \rangle = p_i,
+\quad
+\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, p_j
+\quad \mbox{(for $t \ge 0$)}
+.
+$$
+Further if the transition matrix satisfies detailed balance
+with a set of eigenvectors, $\psi_{li}$,
+that are orthonormal with respect to $\mathbf p$
+[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
+\eqref{eq:eig_orthonormal_rows}]
+then we have the decomposition
+$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / p_j$,
+with $\Lambda_l$ being the eigenvalue of mode $l$.
+%
+The autocorrelation function can then be computed as
+%
+\begin{align*}
+\left\langle
+  \tilde \xi_k(t) \, \tilde \xi_k(0)
+\right\rangle
+=
+\sum_{l = 0}^{n-1} \Lambda_l^t \,
+\left(
+  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ p_i }
+\right)^2
+.
+\end{align*}
+%
+In particular, if the eigenvectors, $\pmb \phi_k$
+and $\pmb \psi_k$, coincide and $\mathbf p = \pmb \rho$,
+we would have
+$\bigl\langle
+  \tilde \xi_k(t) \, \tilde \xi_k(0)
+\bigr\rangle
+=
+\Lambda_k^t,$
+and
+\begin{equation}
+  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty \left\langle
+    \tilde \xi_k(t) \, \tilde \xi_k(0)
+  \right\rangle = \frac{1+\Lambda_k}{1-\Lambda_k}
+  .
+  \label{eq:Gamma_est}
+\end{equation}
+
+The transition matrix method usually only for
+simple sampling processes without major barrier
+along the collective variable, $z$.
+%
+We may nonetheless estimate the magnitude of $\Gamma_k$
+in a few special cases.
+%
+If sampling is perfect, $\Lambda_k = \delta_{k, 0}$,
+then $\Gamma_k = 1$ for $k > 0$.
+%
+For a local sampling process,
+if we borrow \eqref{eq:lambda_Gaussian}
+for $\Lambda_k$,
+with $\sigma_z$ interpreted as the average move size along $z$,
+then
+$$
+\Gamma_k = \coth\left[
+  \left(
+    \frac{ \pi \, k \, \sigma_z } { g \, \Delta z}
+  \right)^2
+  \right].
+$$
+For small $k$, we have $\Gamma_k \propto \Delta z^2/(\sigma_z \, k)^2$,
+which means that a doubled sampling range, $\Delta z$,
+would make the histogram flattening
+at least four times as difficult,
+i.e. FDS methods are most effective
+for short ranges\cite{wang2001, *wang2001pre}.
 
@@
 with bin size
-$\Delta \varphi = 2 \, \pi/(g \, n)$.
+$\delta \varphi = 2 \, \pi/(g \, n)$.
 The variable is related to the bin index, $l$, as
-$\varphi = l \, \Delta \varphi$,
+$\varphi = l \, \delta \varphi$,
 and
-$\mu_l = \mu(\varphi) \, \Delta \varphi$.
+$\mu_l = \mu(\varphi) \, \delta \varphi$.
 %
@@
   &=
-  \int_0^b \mu(\varphi) \, \cos(k \, \varphi) \, \Delta \varphi \, dl
+  \int_0^b \mu(\varphi) \, \cos(k \, \varphi) \, \delta \varphi \, dl
   \\
   &=
-  \int_0^{b \, \Delta \varphi} \mu(\varphi) \, \cos(k \, \varphi) \, d\varphi
+  \int_0^{b \, \delta \varphi} \mu(\varphi) \, \cos(k \, \varphi) \, d\varphi
 \end{align*}
 The upper limit,
-$b \, \Delta \varphi \approx (g\,n/2) \, [2\,\pi/(g\,n)] = \pi$.
+$b \, \delta \varphi \approx (g\,n/2) \, [2\,\pi/(g\,n)] = \pi$.
 }
@@
 if the width of the angular variable, $\sigma_\varphi$,
-satisfies $\Delta \varphi \ll \sigma_\varphi \ll \pi$,
+satisfies $\delta \varphi \ll \sigma_\varphi \ll \pi$,
 we can extend
@@
          \frac{ 2 \, \pi \, k \, \sigma_z }
-              { g \, n \, \Delta z  }
+              { g \, \Delta z  }
         \right)^2
@@
 \end{align}
-where $\sigma_z$ and $\Delta z$
-are the width of the Gaussian
-and the bin width
+where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
+are the width of the Gaussian and span
 in the original collective variable, $z$,
@@
     $K$             &   Cutoff wave number of bandpass updating schemes.  \\
-    $\kappa_k(t)$   &   Autocorrelation function of mode $k$. \\
     $M(Q)$          &   Mass function.   \\

### 2017-06-16 doc/invt.tex
small changes

@@
 \usepackage{mathtools}
+\usepackage{array}
 \usepackage{upgreek}
@@
 \subsection{\label{sec:background}
-Background}
+Flat-distribution sampling}
 
 \subsubsection{\label{sec:FDS}
-Flat-distribution sampling}
+Bias potential}
 
@@
 \begin{equation}
-  M(Q)
+  M(q')
   \equiv
   \sqrt{
-    \textstyle\sum_{ k = 1 }^{n - 1}\Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, Q}
+    \textstyle\sum_{ k = 1 }^{n - 1}
+    \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, q'}
   }
@@
   \int_{ q(T) - q(t) }^{ q(T) }
-    M(Q)
+    M(q')
     \;
-    d Q
+    d q'
   =
@@
   \int_{ 0 }^{ q(T) }
-    M( Q )
+    M( q' )
     \;
-    d Q
+    d q'
   .
@@
   \left(
-    \int_0^{ q(T) } M(Q) \, dQ
+    \int_0^{ q(T) } M(q') \, d q'
   \right)^2
@@
 \begin{equation}
-  m(Q)
+  m(\bar q)
   =
   \frac{
-    M(Q)
+    M(\bar q)
   }
   {
-    \int_0^{ q(T) } M(Q') \, d Q'
+    \int_0^{ q(T) } M(q') \, d q'
   }
@@
   \frac{
-    M(Q)
+    M(\bar q)
   }
@@
 such that
-$\int_0^{q(T)} m(Q) \, dQ = 1$.
+$\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
 %
 Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
-we find that for $Q = q(T) - q(t)$,
+we find that for $\bar q(t) \equiv q(T) - q(t)$,
 %
 \begin{align}
-  m(Q)
+  m\left( \bar q(t) \right)
   &=
@@
   \\
-  \int_Q^{ q(T) }
-    m(Q') \, dQ'
+  \int_{\bar q(t)}^{ q(T) }
+    m(q') \, dq'
   &=
@@
 %
-which means that each point,
+Thus, each point on the schedule,
 $\bigl(t, \alpha(t)\bigr)$,
-on the schedule can be mapped
-to a point,
-$\bigl(Q, m(Q)\bigr)$,
+can be mapped to a point,
+$\bigl(\bar q, m(\bar q)\bigr)$,
 on the mass distribution,
 such that the ordinate of the latter curve
-gives $1/(T\alpha)$,
-and the area under the curve in the domain, $[Q, q(T)]$,
+is $1/(T\alpha)$,
+and the area under the curve in the domain, $[\bar q(t), q(T)]$,
 is equal to $t/T$,
 as shown in Fig. \ref{fig:massq}.
-%
 
@@
     \draw[] (0.5, 0.99488) -- (0.7, 0.99488);
-    \draw[->] (0, 0) -- node[below, inner sep=5mm] {$Q$} (2.8, 0);
-    \draw[->] (0, 0) -- node[above, rotate=90] {$m(Q)$} (0, 1.9);
+    \draw[->] (0, 0) -- node[below, inner sep=7mm] {$\bar q$} (2.8, 0);
+    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 1.9);
 
@@
     \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
+    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$q(T) - q(t)$};
     \draw[] (2.2, 0.1) -- (2.2, 0.0) node[below] {$q(T)$};
@@
     $\alpha(t)$,
-    via the normalized mass distribution, $m(Q)$.
+    via the normalized mass distribution, $m(\bar q)$.
     %
@@
 
+We may consider the schedule, $\alpha(t)$,
+as a variable transformation of
+the mass distribution, $m(\bar q)$.
+%much like the Legendre transform.
+%
+If we define
+$\bar q^* \equiv \bar q/q(T)$
+and $t^* \equiv t/T$
+such that both are limited in $[0, 1]$,
+as well as
+\begin{align*}
+  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
+  \\
+  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
+\end{align*}
+then it is clear that
+both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
+can describe the differential relation
+between $\bar q^*$ and $t^*$,
+but with the natural variables
+being $\bar q^*$ and $t^*$, respectively.
+%
+Further, the transformation is an involution:
+if we apply the same transform on $\alpha^*(t^*)$
+we would recover $m^*(\bar q^*)$.
+%
+A few examples are shown in Table \ref{tab:m_and_a}.
+%
+For example, in the single-bin scheme,
+the mass distribution is a single exponential decay,
+and we would have the inverse-time schedule.
+%
+On the other hand,
+if the mass distribution is inversely proportional to $\bar q$,
+which is approximately true for the Gaussian updating scheme
+at large $\bar q$,
+then we would have an exponentially decaying schedule.
+
+\begin{table}[h]\footnotesize
+  \caption{\label{tab:m_and_a}
+    Variable transformation between the
+    reduced schedule, $\alpha^*(t^*)$,
+    and the reduced mass distribution, $m^*(\bar q^*)$,
+    as defined in Eqs. \eqref{eq:mQ_invTa}
+    and \eqref{eq:intmQ_tT}.
+  }
+  \setlength{\tabcolsep}{3pt} % spacing between columns
+  \renewcommand\arraystretch{2.3} % spacing between rows
+  \begin{tabular} { c c c }
+    \hline
+    $m^*(\bar q^*) = -\dfrac{d t^*}{d\bar q^*}$ &
+    $\alpha^*(t^*) = -\dfrac{d\bar q^*}{d t^*}$ &
+    $\bar q^*(t^*)$
+    \\
+    \hline
+    $\dfrac{c \, a}{\left( {\bar q}^* + q^*_0 \right)^{a+1}}$ &
+    $\dfrac{c^{1/a} \, (1/a)}{\left( t^* + t^*_0 \right)^{(1/a)+1}}$ &
+    $\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a}$
+    %$\,\begin{aligned}
+    %  &a, q^*_0 > 0, \\
+    %  &\frac{1}{t^*_0} \equiv \left(1+\frac{1}{q^*_0}\right)^a - 1, \\
+    %  &c \equiv {q^*_0}^a(t^*_0 + 1)
+    %\end{aligned}$
+    \\
+    $\dfrac{1}{\gamma \, {\bar q}^* + \frac{\gamma}{e^\gamma - 1}}$ &
+    $\dfrac{\gamma}{1-e^{-\gamma}} \, e^{-\gamma \, t^*}$ &
+    $\dfrac{e^{-\gamma \, t^*} - e^{-\gamma}}{1-e^{-\gamma}}$
+    \\
+    $\dfrac{\gamma}{1 - e^{-\gamma}} e^{-\gamma \, {\bar q}^*}$ &
+    $\dfrac{1}{ \gamma \, t^* + \frac{\gamma}{e^\gamma - 1} }$ &
+    $\dfrac{\ln\left[(1-e^{-\gamma}) \, t^* + e^{-\gamma} \right]}{-\gamma}$
+    \\
+    \hline
+    \multicolumn{3}{c}{
+    Here,
+    $a, q^*_0, \gamma > 0$,
+    ${t^*_0}^{-1} \equiv \left(1+{q^*_0}^{-1}\right)^a - 1$,
+    and
+    $c \equiv {q^*_0}^a(t^*_0 + 1)$.
+    } \\
+    \hline
+  \end{tabular}
+\end{table}
+
 \subsubsection{\label{sec:optinitalpha}
@@
 \begin{align}
-  \frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
-  &+
-  \sum_{k = 1}^{n-1}
+  \mathcal Y\bigl( q(T) \bigr)
+  &\equiv
+  \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(q') \, dq'
+  -\frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
+  \notag \\
+  &
+  -\sum_{k = 1}^{n-1}
   \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
-  \notag \\
-  &= \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(Q') \, dQ'
+  =0
   .
@@
   M(q_T) \,
-  \int_0^{ q_T } M(Q) \, dQ
+  \int_0^{ q_T } M(q) \, dq
   =
@@
 %For simulations with the same $a_0$,
-%the shape of the mass distribution, $m(Q)$,
+%the shape of the mass distribution, $m(\bar q)$,
 %also determines how much the optimal schedule
@@
 %%
-%Since both $M(Q)$ and $m(Q)$ decrease monotonically with $Q$,
+%Since both $M(\bar q)$ and $m(\bar q)$ decrease monotonically with $\bar q$,
 %Eq. \eqref{eq:opt_qT} shows that
@@
 %the simulation length, $T$,
-%more so if $M(Q)$ has a long tail.
+%more so if $M(\bar q)$ has a long tail.
 %%
-%Thus, if $M(Q)$ decays slowly because of
+%Thus, if $M(\bar q)$ decays slowly because of
 %many near zero eigenvalues, $\lambda_k$'s,
@@
 %
-%If $M(Q)$ decays rapidly,
-%the distribution $m(Q)$ does not critically
+%If $M(\bar q)$ decays rapidly,
+%the distribution $m(\bar q)$ does not critically
 %depend on the cutoff $q(T)$,
@@
 %%
-%If, however, $M(Q)$ has a long tail
+%If, however, $M(\bar q)$ has a long tail
 %because of some near-zero $\lambda_k$'s,
@@
   2 \,
-  \sum_{ l = 1 }^b
-  \mu_l
+  \sum_{ i = 1 }^b
+  \mu_i
   \cos\left(
-  \frac{ k \, l \, 2 \, \pi }
+  \frac{ k \, i \, 2 \, \pi }
        {      g \, n        }
@@
 (cf. Appendix \ref{sec:more_wband}).
+
+In particular, for the Gaussian updating scheme,
+we have\cite{bussi2006}
+\begin{align}
+  \lambda_k
+  &=
+  \exp\left(
+        -\frac{ k^2 \, \sigma_\varphi^2 }
+              {           2             }
+      \right)
+  =
+  \exp\left[
+        -\frac 1 2
+        \left(
+         \frac{ 2 \, \pi \, k \, \sigma_z }
+              { g \, \Delta z  }
+        \right)^2
+      \right]
+  ,
+%\notag
+\label{eq:lambda_Gaussian}
+\end{align}
+where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
+are the width of the Gaussian and span
+in the original collective variable, $z$,
+respectively.
 
@@
   \node[draw, text width={width("Integrals, iii")}]
-    (gamma) at (2.1, 1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$};
+    (Gamma) at (2.1, 1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$};
 
@@
     (w) edge (lambda)
-    (w) edge (gamma)
-    (sys) edge (gamma);
+    (w) edge (Gamma)
+    (sys) edge (Gamma);
 
@@
     (lambda) -| (M1)
-    (gamma)  -- (M1)
+    (Gamma)  -- (M1)
     (T)      -| (M1)
@@
 
-\item \label{step:gamma}
+\item \label{step:Gamma}
 Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
@@
 We implemented a variant of the Newton-Raphson method\cite{press3rd}
-(for $\ln q(T)$ to avoid negative values)
-with the bisection method\cite{press3rd} as a fallback.
+with the bisection method\cite{press3rd}
+as a fallback.\footnote{
+The equation was solved for $\ln q(T)$
+to avoid negative values of $q(T)$,
+and we imposed a lower bound,
+$\bigl|\mathcal Y/q(T)\bigr|$, for
+$\mathcal Y'\bigl(q(T)\bigr)$
+in computing its change.
+}
 The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
 %
-%Since $m(Q)$ is a decreasing function of $Q$,
+%Since $m(\bar q)$ is a decreasing function of $\bar q$,
 %we can start from an interval $[Q_L, Q_R]$
@@
 
-As shown in Table 1,
+As shown in Table \ref{tab:lj_error},
 both the inverse-time and optimal schedule reduced
@@
 %
-The Gaussian scheme appeared to be superior
-to the single-bin scheme in the $\varrho^* = 0.1$ case,
+The Gaussian scheme produced less error than
+the single-bin scheme in the $\varrho^* = 0.1$ case,
 but not in the $\varrho^* = 0.8$ case.
 %
-In the latter case,
+In the low density case,
+when the local noise contributes significantly to the total error,
+the Gaussian scheme
+yielded less error than the single bin scheme,
+and the optimal schedule was more effective
+in reducing the error than the inverse-time schedule.
+%
+However, in the high-density case
 the error was dominated by the $k=1$ eigenmode
-corresponding to the longest range motion, as $\Gamma_1 \approx \Gamma$,
-and the suppression of local noise
-had little effect on reducing the total error.
-%
-In either case, there was an optimal width of Gaussian
-an overly large width could be counterproductive.
-%
-The optimal width was $\sigma = 0.37$ for $\varrho^* = 0.1$
+corresponding to the longest range motion,
+as $\Gamma_1 \approx \Gamma$,
+and the reduction of local noise
+had little effect on the total error.
+%
+The bandpass updating scheme behaved similarly,
+and in the low density case when
+local noise is important,
+this scheme (with the mode cutoff $K = 20$)
+was superior to the Gaussian counterpart.
+
+The Gaussian updating scheme was most effective
+at an optimal width, which
+was $\sigma = 0.37$ for $\varrho^* = 0.1$
 or $0.07$ for $\varrho^* = 0.8$ in this case,
 although it depended on the simulation length.
-
-To further investigate the behavior of the Gaussian updating scheme,
+%
+%
+To further investigate its behavior
 we show in Fig. \ref{fig:lj_xerr}
@@
 \begin{table}[h]\footnotesize
-  \caption{\label{tab:error_LJ}
-    Error of the LJ system.
+  \caption{\label{tab:lj_error}
+    Error of the bias potential under
+    the single-bin,
+    Gaussian (labeled by the width, $\sigma$),
+    and bandpass (labeled by the mode cutoff, $K$)
+    updating schemes
+    on the LJ system.
+    %
     The numbers have been averaged over multiple runs,
@@
     either the inverse-time ($1/t$) or optimal schedule (opt.).
+    %
     \note{The reference values can be found
@@
 %
-In the long time limit, the fluctuation of
+With a sufficiently accurate and stable bias potential,
+the fluctuation of
 the instantaneous histogram is dominated by the random part,
@@
 \end{equation}
-and we would have for the Fourier transform,
+and the same holds for the Fourier transform,
 $\tilde \xi_k(t) \approx \tilde \zeta_k(t).$
@@
 \begin{align}
-F(T)
+G(T)
 &\equiv
@@
   &=
-\left\langle
   \sum_{i=1}^n \rho_i
@@
   \right)^2
-\right\rangle
 .
@@
 From Eq. \eqref{eq:xi_zeta}, we have
+%
 \begin{align*}
-\left\langle \tilde \Xi_k^2(T) \right\rangle
+\bigl\langle \tilde \Xi_k^2(T) \bigr\rangle
 &=
@@
 \sum_{t=-\infty}^\infty
-\left\langle
+\bigl\langle
   \tilde \xi_k(t) \, \tilde \xi_k(0)
-\right\rangle
+\bigr\rangle
 \\
@@
 \sum_{t=-\infty}^\infty
-\left\langle
+\bigl\langle
   \tilde \zeta_k(t) \, \tilde \zeta_k(0)
-\right\rangle
+\bigr\rangle
 =
@@
 \begin{equation}
-\langle F(T) \rangle
+\bigl\langle G(T) \bigr\rangle
 \equiv
-\sum_{k=1}^{n-1} \left\langle \tilde \Xi_k^2(T) \right\rangle
+\sum_{k=1}^{n-1}
+  \bigl\langle
+    \tilde \Xi_k^2(T)
+  \bigr\rangle
 =
@@
 Comparing this to Eq. \eqref{eq:Emin_singlebin},
-we conclude
-\begin{equation}
-  \langle F(T+t_0) \rangle \approx E(T)
+we may rephrase the optimality of the inverse-time schedule
+as the inequality,
+\begin{equation}
+  E(T)
+  \ge
+  \bigl\langle G(T+t_0) \bigr\rangle
   .
@@
 \end{equation}
-%and the histogram fluc
+%and the histogram fluctuation
 %serves as a good estimate of the final error.
@@
 we may borrow this criterion
-in initial stages, with a modified definiton
+in initial stages, with a modified definition
 for the histogram fluctuation
 $$
-F_{<K}(T) = \sum_{k=1}^{K-1} \tilde \Xi_k^2(T),
-$$
-since the histogram-flatten
+G_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde \Xi_k^2(T),
+$$
+since the histogram flattening
 is only partially applied to modes less $K$.
+%
+One can alternatively define the histogram fluctuation
+from the largest mode
+$$
+\max_{k=1, \dots, K-1} \left| \, \tilde \Xi_k(T) \, \right|
+.
+$$
+For a local sampling process,
+we expect that the two definition would behave similarly.
 %
@@
 \section{\label{sec:Gamma_measure}
-Measurement of the integrals of the autocorrelation functions of the eigenmodes
+Integrals of the autocorrelation functions of the eigenmodes
 }
@@
 $$
-For small $k$, we have $\Gamma_k \propto \Delta z^2/(\sigma_z \, k)^2$,
+For small $k$,
+we have $\Gamma_k \propto \Delta z^2/(\sigma_z \, k)^2$,\cite{bussi2006}
 which means that a doubled sampling range, $\Delta z$,
@@
 for short ranges\cite{wang2001, *wang2001pre}.
+%
+Using finer bins would, however, not change this value.
+%
+For a modestly large $k$,
+$\Gamma_k \approx 1$.
+and by Eq. \eqref{eq:xt2_eql},
+we would expect a roughly flat profile
+for the single-bin updating scheme,
+cf. Fig. \ref{fig:lj_xerr}.
 
@@
 \begin{equation}
-  \mu_l
+  \mu_i
   =
@@
   \cos \left(
-       \frac{ k \, l \, 2 \, \pi }
+       \frac{ k \, i \, 2 \, \pi }
             {      g \, n        }
@@
     \mu(\varphi) \, \cos (k \, \varphi) \, d\varphi,
-\notag
-%\label{eq:lambda_int}
+\label{eq:lambda_int}
 \end{equation}
@@
 %
-$
+\begin{equation}
   1 = 2 \int_0^\pi \mu(\varphi) \, d\varphi.
-$
+\label{eq:mu_int}
+\end{equation}
 \note{
@@
 
-For the Gaussian updating kernel,
-if the width of the angular variable, $\sigma_\varphi$,
-satisfies $\delta \varphi \ll \sigma_\varphi \ll \pi$,
-we can extend
-the upper limit of the integrals
-to infinity, and
+If the width of the Gaussian updating kernel
+is very small compared to $\sigma_\varphi$,
+then we can extend the upper limit of the integral
+in Eq. \eqref{eq:mu_int} to $\infty$, and
 %
@@
   \frac{            1            }
-       { \sqrt{ 2 \pi \sigma_\varphi^2 } }
+       { \sqrt{ 2 \pi } \sigma_\varphi }
   %
@@
 %
-%
-It follows that all eigenvalues are positive\cite{bussi2006}:
-%
-\begin{align}
-  \lambda_k
-  &=
-  \exp\left(
-        -\frac{ k^2 \, \sigma_\varphi^2 }
-              {           2             }
-      \right)
-  =
-  \exp\left[
-        -\frac 1 2
-        \left(
-         \frac{ 2 \, \pi \, k \, \sigma_z }
-              { g \, \Delta z  }
-        \right)^2
-      \right]
-  ,
-%\notag
-\label{eq:lambda_Gaussian}
-\end{align}
-where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
-are the width of the Gaussian and span
-in the original collective variable, $z$,
-respectively.
-
-For a finite bin width,
-some eigenvalues can be negative.
-However, we can avoid the stability problem
-by starting with the eigenvalues given by Eq. \eqref{eq:lambda_Gaussian}
+Then, from Eq. \eqref{eq:wband_eigenvalue},
+we get Eq. \eqref{eq:lambda_Gaussian},
+which shows that all eigenvalues are positive.
+
+In general, some eigenvalues can turn negative
+because of truncation and discretization.
+%
+We can however maintain the stability of updating scheme
+by starting with Eq. \eqref{eq:lambda_Gaussian}
 and inversely computing the kernel using Eq. \eqref{eq:mu_from_lambda}.
+%
 
@@
     $b$             &   Cutoff of the kernel of homogeneous updating schemes. \\
-    $C$             &   $C \equiv \int_0^{ q(T) } M(Q) \, dQ$.  \\
+    $C$             &   $C \equiv \int_0^{ q(T) } M(q) \, dq$.  \\
     $\delta$        &   Dirac's or discrete $\delta$-function. \\
@@
     $K$             &   Cutoff wave number of bandpass updating schemes.  \\
-    $M(Q)$          &   Mass function.   \\
-    $m(Q)$          &   Mass distribution,
-                        $m(Q) \equiv M(Q)/\int_0^{ q(T) } M(Q') \, dQ'$.  \\
+    $M(q)$          &   Mass function.   \\
+    $m(q)$          &   Mass distribution,
+                        $m(q) \equiv M(q)/\int_0^{ q(T) } M(q') \, dq'$.  \\
     $\mu_i$         &   Updating kernel of a homogeneous updating scheme. \\
@@
     $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\
-    $Q$             &   $q(T) - q(t)$.  \\
+    $\bar q(t)$     &   $q(T) - q(t)$.  \\
     $\rho_i$        &   Distribution used to define the error. \\

### 2017-06-16 doc/invt.tex
invt.tex: make p_i and \rho_i same, p^*_i -> p_i

@@
 \usepackage{mathtools}
-\usepackage{array}
 \usepackage{upgreek}
@@
 often the task is to compute
-a distribution, $p^*(z)$,
+a distribution, $p(z)$,
 along a collective variable, $z$, by sampling via either Monte Carlo\cite{
@@
 along the collective variable, $z$,
-is $-\ln p^*(z)$.
+is $-\ln p(z)$.
 %
@@
 Consider the problem of computing
-the distribution, $p^*_i$,
+the distribution, $p_i$,
 along a discrete quantity, $i$,
@@
 %The distribution is normalized as
-%$\sum_{i = 1}^n p^*_i = 1$.
+%$\sum_{i = 1}^n p_i = 1$.
 
 %For a large system at temperatures of interest,
-%the distribution, $p^*_i$, often tends to
+%the distribution, $p_i$, often tends to
 %be localized,
@@
 %and to explore the global shape
-%of the PMF, $-\ln p^*_i$,
+%of the PMF, $-\ln p_i$,
 %it is often advantageous to perform
 %biased sampling that targets
-%a wider and smoother distribution, $p_i$.
+%a wider and smoother distribution, $\rho_i$.
 %%
@@
 %To do so, we introduce a bias potential, $u_i$,
-The challenge of sampling a multimodal distribution, $p^*_i$,
+The challenge of sampling a multimodal distribution, $p_i$,
 with deep valley(s) can often be eased by
-biased sampling targeting a smoother distribution, $p_i$.
+biased sampling targeting a smoother distribution, $\rho_i$.
 %
@@
 %the bias potential $u_i$
-%must coincide with $\ln p^*_i$,
+%must coincide with $\ln p_i$,
 %up to an additive constant.
@@
 %and thus $\pi_i$ need not
-%coincide with the intended distribution $p_i$.
+%coincide with the intended distribution $\rho_i$.
 %
@@
 as a biased simulation
-targeting a flat $p_i$ (i.e. $p_i \equiv 1/n$)
+targeting a flat $\rho_i$ (i.e. $\rho_i \equiv 1/n$)
 or nearly flat one\cite{
@@
   %=
-  %C_i \, p^*_i \, e^{-u_i}
-  =
-  \frac{                p^*_i \, e^{-u_i} }
-       { \sum_{j = 1}^n p^*_j \, e^{-u_j} }
+  %C_i \, p_i \, e^{-u_i}
+  =
+  \frac{                p_i \, e^{-u_i} }
+       { \sum_{j = 1}^n p_j \, e^{-u_j} }
   \propto
-  p^*_i \, e^{-u_i}
-  ,
-  \label{eq:pi_p_V}
+  p_i \, e^{-u_i}
+  ,
+  \notag
+  %\label{eq:pi_p_V}
 \end{equation}
@@
 such that the resulting $\pi_i$ would coincide
-with the desired $p_i$.
+with the desired $\rho_i$.
 %
@@
 \begin{equation}
-  u_i \to \ln p^*_i - \ln p_i + \mathrm{const.}
+  u_i \to \ln \frac{p_i}{\rho_i} + \mathrm{const.}
   ,
@@
 %according to Eq. \eqref{eq:pi_p_V}.
-%which allows us to deduce the PMF, $-\ln p^*_i$,
+%which allows us to deduce the PMF, $-\ln p_i$,
 %directly from the bias potential.
 %
-and the PMF, $-\ln p^*_i$, can be readily
+and the PMF, $-\ln p_i$, can be readily
 deduced from the bias potential. % $u_i$.
@@
   For example, in the context of simulated tempering (Zhang and Ma, 2007),
-  $p^*_i$ is replaced by $\ln Z_i$,
-  $p_i$ is the weight $\zeta_i$,
+  $p_i$ is replaced by $\ln Z_i$,
+  $\rho_i$ is the weight $\zeta_i$,
   and the bias potential, $u_i$, is $\ln (\tilde Z_i / \zeta_i)$.
@@
   \ln \frac{ H_i }
-           { p_i }.
+           { \rho_i }.
   \label{eq:vcorr_equil}
@@
 \note{This follows from
-  $$
-  \begin{aligned}
+  \begin{align*}
     H_i \approx \pi_i
-    &\propto p^*_i \, e^{-u_i},
-    \\
-    p_i
-    &\propto p^*_i \, e^{-\hat u_i}.
-  \end{aligned}
-  $$
+    &\propto p_i \, e^{-u_i},
+    \\
+    \rho_i
+    &\propto p_i \, e^{-\hat u_i}.
+  \end{align*}
 }
@@
   \delta_{i, \, i(t)}
-  \frac{ \alpha(t) } { p_{i(t)} }
+  \frac{ \alpha(t) } { \rho_{i(t)} }
   ,
@@
   \frac{ \alpha(t) }
-       { p_{i(t)} }
+       { \rho_{i(t)} }
   .
@@
 %an adaptive FDS simulation
-%is identical to $p_i$ in the long time or asymptotic limit.
+%is identical to $\rho_i$ in the long time or asymptotic limit.
 %
@@
   -
-  \ln \frac { p^*_i }
-            { p_i }
+  \ln \frac { p_i }
+            { \rho_i }
   ,
@@
 %  =
-%  \frac{                p_i \, e^{-v_i} }
-%       { \sum_{j = 1}^n p_j \, e^{-v_j} }
+%  \frac{                \rho_i \, e^{-v_i} }
+%       { \sum_{j = 1}^n \rho_j \, e^{-v_j} }
 %  \propto
-%  p_i \, e^{-v_i}.
+%  \rho_i \, e^{-v_i}.
 %  \label{eq:pi_p_v}
@@
 %
-We define a bin-average-shifted bias potential as
-%
-\begin{equation}
-  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
-  ,
-\label{eq:x_def}
-\end{equation}
-%
-where the bin average, $\bav{v}(t)$, is defined as
-%
-\begin{equation}
-  \bav{v}(t)
-  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
-  ,
-\label{eq:vav_def}
-\end{equation}
-for some distribution,
-$\pmb \rho = (\rho_1, \dots, \rho_n)$
-(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
-%
-Below we will use the notations, $f_{*i}$ and $\bav{f}$,
-for an arbitrary variable, $f_i$.
-%
-Note that the distribution, $\pmb\rho$,
-can be independent of
-the target distribution,
-$\mathbf p = (p_1, \dots, p_n)$,
-and chosen for convenience.\footnote{Alternatively,
+We define a bin-average-deducted bias potential
+as\footnote{Alternatively,
 we may define $v_{*i}(t)$ such that
@@
 v_{*i}(t) \to -\ln\left(
-  \frac{ (p^*_i/p_i) \, e^{ -u_i(t) } }
-  { \sum_{j=1}^n \rho_j \, (p^*_j/p_j) \, e^{ -u_j(t) } }
+  \frac{ (p_i/\rho_i) \, e^{ -u_i(t) } }
+  { \sum_{j=1}^n p_j \, e^{ -u_j(t) } }
 \right).
@@
 One can readily show that the two definitions are equivalent
-for small $v_{*i}$.
-%
-In well-tempered metadynamics,
-$\mathbf p$ is an attenuated version of
-$\mathbf p^*$\cite{
-  barducci2008, dama2014}:
-$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
-whereas $\pmb\rho$ is flat.}
-
+for small $v_{*i}$.}
+%
+\begin{equation}
+  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
+  ,
+\label{eq:x_def}
+\end{equation}
+%
+where the bin average, $\bav{v}(t)$, is defined as
+$\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$.
+%
+%\begin{equation}
+%  \bav{v}(t)
+%  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
+%  .
+%  \label{eq:vav_def}
+%\end{equation}
+%
+Below we will use the notations, $x_{*i}$ and $\bav{x}$,
+for an arbitrary variable, $x_i$.
+%
 The total error is given by
@@
   \approx
-  \alpha(t) \, \frac{ h_i(t) } { p_i }
+  \alpha(t) \, \frac{ h_i(t) } { \rho_i }
   \equiv
-  \alpha(t) \, \xi_i(t)
+  \alpha(t) \, f_i(t)
   ,
@@
 and we have defined
-$\xi_i(t) \equiv h_i(t) /p_i$.
+$f_i(t) \equiv h_i(t) /\rho_i$.
+%
+Clearly,
+%
+\begin{equation}
+  \bav{f}(t) = \sum_{i=1}^n h_i(t) = 1
+  ,
+  \label{eq:fav1}
+\end{equation}
 %
@@
   \approx
-  \alpha(t) \, \xi_{*i}(t)
+  \alpha(t) \, f_{*i}(t)
   .
@@
 We can split the histogram into
-a deterministic averaged part, $\langle \xi_i(t) \rangle$,
+a deterministic averaged part, $\langle f_i(t) \rangle$,
 and a random fluctuation part, $\zeta_i(t)$,
@@
 \begin{equation}
-  \xi_i(t) =
-  \langle \xi_i(t) \rangle
+  f_i(t) =
+  \langle f_i(t) \rangle
   +
@@
   .
-  \label{eq:h_split}
+  \notag
+  %\label{eq:h_split}
 \end{equation}
@@
 \begin{align}
-  \langle \xi_i(t) \rangle
+  \langle f_i(t) \rangle
   %&=
-  %\frac{ \langle h_i(t) } { p_i } \rangle
+  %\frac{ \langle h_i(t) } { \rho_i } \rangle
   &\approx
-  \frac{ \pi_i(t) } { p_i }
+  \frac{ \pi_i(t) } { \rho_i }
   =
   \frac{                          e^{-v_{*i}(t)} }
-       { \sum_{j = 1}^n p_j \, e^{-v_{*j}(t)} }
-  \notag
-  \\
-  &\approx
-  1 - v_{*i}(t) + \sum_{j = 1}^n p_j \, v_{*j}(t)
+       { \sum_{j = 1}^n \rho_j \, e^{-v_{*j}(t)} }
+  %\notag
+  %\\
+  %&
+  \approx
+  1 - v_{*i}(t)
   ,
@@
 \frac{ \langle h_i(t) \rangle }
-     { p_i }
+     { \rho_i }
 \approx
 \frac{                       1 - v_{*i}  }
-     { \sum_{ r = 1 }^n p_j (1 - v_{*j}) }
+     { \sum_{ r = 1 }^n \rho_j (1 - v_{*j}) }
 =
 \frac{                       1 - v_{*i}  }
-     { 1 - \sum_{ j = 1 }^n p_j \, v_{*j} }
+     { 1 - \sum_{ j = 1 }^n \rho_j \, v_{*j} }
 .
@@
 %
-Deducting the bin average, we get
+Deducting the bin average and using Eq. \eqref{eq:fav1}, we get
 %
 \begin{align}
-  \langle \xi_{*i}(t) \rangle
+  \langle f_{*i}(t) \rangle
   \equiv
-  \langle \xi_i(t) \rangle -
-  \sum_{j=1}^n \rho_j \langle \xi_j(t) \rangle
+  \langle f_i(t) \rangle - 1
   &\approx
@@
 We will approximate the histogram fluctuation part, $\zeta_i(t)$,
-hence its bin-average-shifted version, $\zeta_{*i}(t)$,
+hence its bin-average-deducted version, $\zeta_{*i}(t)$,
 as white noise such that
@@
   ,
-  \label{eq:G_def}
+  \notag
+  %\label{eq:G_def}
 \end{equation}
@@
   =
-  \alpha(t) \, \left[ \langle \xi_{*i}(t) \rangle + \zeta_{*i}(t) \right]
+  \alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \zeta_{*i}(t) \right]
   \approx
@@
 \end{align}
-where
-\begin{equation}
-  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i.
-  \label{eq:Gamma_def}
-\end{equation}
+where $\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
+%\begin{equation}
+%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i.
+%  \label{eq:Gamma_from_G}
+%\end{equation}
 
@@
 \begin{equation}
-t_0 = \frac{ \Gamma }
-           { E(0) }
+  t_0 = \frac{ \Gamma } { E(0) }
   ,
@@
   \alpha(t) \,
-  \sum_{j=1}^n w_{ij} \, \frac{ h_j(t) } { p_j }
+  \sum_{j=1}^n w_{ij} \, \frac{ h_j(t) } { \rho_j }
   .
@@
 To sample the desired distribution,
-$\mathbf p = (p_1, \dots, p_n)$,
+$\pmb\rho = (\rho_1, \dots, \rho_n)$,
 the growth rate of the bias potential,
@@
 should be a constant of $i$
-if $h_j(t)$ were the same as $p_j$,
+if $h_j(t)$ were the same as $\rho_j$,
 i.e.
@@
 that satisfy
-the detailed balance condition,
+the detailed balance condition for $\pmb\rho$,
 %
@@
   \rho_i \, w_{ij} = \rho_j \, w_{ji}
-  ,
+  .
   \label{eq:w_detailedbalance}
 \end{equation}
-%
-for the error weighting distribution, $\pmb \rho$,
-introduced in Eq. \eqref{eq:vav_def}.
 %
@@
 and Eq. \eqref{eq:eigenmode0} means that
-the bin average of any quantity, $f$,
+the bin average of any quantity, $x_i$,
 is the first mode of the Fourier transform:
-$\bav{f}= \mathcal F[f_i]_0 = \tilde f_0$.
+$\bav{x}= \mathcal F[x_i]_0 = \tilde x_0$.
 %
@@
 \begin{equation}
-  \tilde f_{*k}
-  = \tilde f_k - \bav{f} \, \delta_{k0}
+  \tilde x_{*k}
+  = \tilde x_k - \bav{x} \, \delta_{k0}
   =
@@
     0           & k = 0, \\
-    \tilde f_k  & k > 0,
+    \tilde x_k  & k > 0,
   \end{dcases}
-  \label{eq:ftstar}
+  \label{eq:xtstar}
 \end{equation}
@@
 which means that the only effect
-of the bin-average deduction on $\tilde f_k$ is to
-annihilate the first mode, $\tilde f_0$.
+of the bin-average deduction on $\tilde x_k$ is to
+annihilate the first mode, $\tilde x_0$.
 
@@
 {\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
-{\tilde \xi}_k(t)
+{\tilde f}_k(t)
 ,
@@
   =
-  \alpha(t) \, \lambda_k \, {\tilde \xi}_k(t)
+  \alpha(t) \, \lambda_k \, {\tilde f}_k(t)
   \approx
@@
 and used the transformed version of Eq. \eqref{eq:sh_ave}
-[but the subscript ``$_*$'' is dropped by Eq. \eqref{eq:ftstar}].
+[but the subscript ``$_*$'' is dropped by Eq. \eqref{eq:xtstar}].
 In the second step, we have used the linear approximation
@@
 and drop the first term
-by using Eq. \eqref{eq:ftstar}
+by using Eq. \eqref{eq:xtstar}
 in the second step.
@@
   = \sum_{t = -\infty}^\infty
-  \left\langle {\tilde \zeta}_k(0)
-            \, {\tilde \zeta}_k(t) \right\rangle
+  \bigl\langle {\tilde \zeta}_k(0)
+            \, {\tilde \zeta}_k(t) \bigr\rangle
   .
@@
 can be mapped to a point,
-$\bigl(\bar q, m(\bar q)\bigr)$,
+$\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
 on the mass distribution,
@@
   \caption{\label{tab:m_and_a}
-    Variable transformation between the
+    Examples of the variable transformation between the
     reduced schedule, $\alpha^*(t^*)$,
@@
     \hline
+    $\dfrac{1}{\gamma \, ({\bar q}^* + s_\gamma) }$ &
+    $(1 + s_\gamma) \, \gamma \, e^{-\gamma \, t^*}$ &
+    $(1 + s_\gamma) \, e^{-\gamma \, t^*} - s_\gamma$
+    \\
+    $(1+ s_\gamma) \, \gamma \, e^{-\gamma \, {\bar q}^*}$ &
+    $\dfrac{1}{ \gamma \, (t^* + s_\gamma) }$ &
+    $\dfrac{1}{\gamma} \ln\left(\dfrac{1 + s_\gamma}{t^* + s_\gamma} \right)$
+    \\
     $\dfrac{c \, a}{\left( {\bar q}^* + q^*_0 \right)^{a+1}}$ &
@@
     \\
-    $\dfrac{1}{\gamma \, {\bar q}^* + \frac{\gamma}{e^\gamma - 1}}$ &
-    $\dfrac{\gamma}{1-e^{-\gamma}} \, e^{-\gamma \, t^*}$ &
-    $\dfrac{e^{-\gamma \, t^*} - e^{-\gamma}}{1-e^{-\gamma}}$
-    \\
-    $\dfrac{\gamma}{1 - e^{-\gamma}} e^{-\gamma \, {\bar q}^*}$ &
-    $\dfrac{1}{ \gamma \, t^* + \frac{\gamma}{e^\gamma - 1} }$ &
-    $\dfrac{\ln\left[(1-e^{-\gamma}) \, t^* + e^{-\gamma} \right]}{-\gamma}$
-    \\
     \hline
-    \multicolumn{3}{c}{
+    \multicolumn{3}{p{8cm}}{
     Here,
     $a, q^*_0, \gamma > 0$,
+    and we have defined
+    $s_\gamma = 1/(e^\gamma - 1)$,
     ${t^*_0}^{-1} \equiv \left(1+{q^*_0}^{-1}\right)^a - 1$,
     and
-    $c \equiv {q^*_0}^a(t^*_0 + 1)$.
+    $c \equiv {q^*_0}^a(1 + t^*_0)$.
     } \\
@@
 However, the histogram fluctuation should
-be limited to first few modes, $k = 1, \dots, K$,
+be limited to first few modes,
 as the bandpass updating scheme
@@
 have placed on an axis parallel to a side of the cubic box}
-and the intrinsic distribution, $p^*$, is
+and the intrinsic distribution, $p(r)$, is
 the cavity distribution function, $y(r)$\cite{hansen}.
@@
 As shown in Table \ref{tab:lj_error},
-both the inverse-time and optimal schedule reduced
+all updating schemes
+with either the inverse-time or optimal schedule
+were able to reduced
 the error significantly,
 and the errors roughly matched the predicted values
-calculated from the precomputed $\Gamma_k$'s and $\epsilon_k$'s.
-%
-The Gaussian scheme produced less error than
+calculated from the precomputed
+$\Gamma_k$'s and $\epsilon_k$'s.
+%
+The Gaussian updating scheme produced less error than
 the single-bin scheme in the $\varrho^* = 0.1$ case,
@@
 %
-In the low density case,
-when the local noise contributes significantly to the total error,
-the Gaussian scheme
-yielded less error than the single bin scheme,
-and the optimal schedule was more effective
-in reducing the error than the inverse-time schedule.
-%
-However, in the high-density case
+The effectiveness of the Gaussian updating scheme
+in the low density case
+was probably because that the sampling along
+the collective variable was free from major barriers
+and thus the histogram fluctuation largely
+resulted from local noise.
+%
+%The optimal schedule also appeared to be more effective
+%in reducing the error than the inverse-time schedule.
+%%
+In this case, we could further reduced the error
+by using the bandpass updating scheme
+(with the mode cutoff $K = 20$).
+%
+On the other hand, in the high-density case,
 the error was dominated by the $k=1$ eigenmode
-corresponding to the longest range motion,
-as $\Gamma_1 \approx \Gamma$,
+corresponding to the longest range fluctuation,
+$\Gamma_1 \approx \Gamma$,
 and the reduction of local noise
 had little effect on the total error.
-%
-The bandpass updating scheme behaved similarly,
-and in the low density case when
-local noise is important,
-this scheme (with the mode cutoff $K = 20$)
-was superior to the Gaussian counterpart.
-
-The Gaussian updating scheme was most effective
-at an optimal width, which
-was $\sigma = 0.37$ for $\varrho^* = 0.1$
-or $0.07$ for $\varrho^* = 0.8$ in this case,
-although it depended on the simulation length.
-%
-%
-To further investigate its behavior
-we show in Fig. \ref{fig:lj_xerr}
-the error components, $\langle \tilde v_k^2 \rangle$,
-against the eigenmode index
-for the $\varrho^* = 0.1$ case
-with width $\sigma = 0.2$.
-%
-While the Gaussian updating scheme effectively suppressed
-the initial errors for short-range or large-$k$ modes,
-it also slowed down the error reduction of these modes
-[cf. Eq. \eqref{eq:vt_diffeq_mbin} and \eqref{eq:lambda_Gaussian}].
-%
-As a result, the final errors of the long-range or small-$k$ modes
-were larger than those from the single-bin scheme.
-%
-For the Gaussian updating scheme,
-the inverse-time schedule tended to reduce the updating magnitude
-too quickly
-leaving a hump on the error profile for the middle-range eigenmodes.
-%
-The optimal schedule improved over the inverse-time schedule
-and produced a flatter profile for the final error.
-%
-Even without the optimal schedule,
-it was also possible to roughly
-recover the single-bin result by using the correction formula
-based on the histogram, Eq. \eqref{eq:uhatav}.
-%
-This correction usually improves
-the estimate for the small-$k$ modes
-but not necessarily for the large-$k$ modes.
-
-In Fig. \ref{fig:lj_alpha}.
-we show the optimal schedules in a few cases.
-%
-Generally, the optimal schedule
-for the $\varrho^* = 0.8$ case
-was closer to the inverse-time schedule
-than that for the $\varrho^* = 0.1$ case.
-%
-This was probably because in the former case
-the first mode dominates the total error,
-$\Gamma_1 \approx \Gamma$,
-and the schedule was roughly optimized for this mode,
-resulting in
-$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
-as $\lambda_1 \approx 1$
-[cf. Eq. \eqref{eq:lambda_Gaussian}].
-%
-Since we have chosen a long equilibration period
-for the preliminary run under constant updating magnitude,
-the starting magnitude was usually around $a_0/2 = 5\times10^{-5}$
-[cf. Eq. \eqref{eq:half_alpha0}].
-%
-But if the Gaussian were too wide,
-e.g. if $\sigma = 0.6$,
-the eigenvalues $\lambda_k$ of some modes
-became too small to damp out the initial bias,
-resulting in much larger initial updating magnitude, $\alpha(0)$
-[cf. Eq. \eqref{eq:opt_alpha0}].
-%
-In this case, our method would fail
-because the resulting bias potential would no longer be small,
 
@@
 
+For the Gaussian updating scheme,
+there was an optimal width for maximum effectiveness.
+%
+The optimal width was $\sigma = 0.37$ for $\varrho^* = 0.1$
+or $0.07$ for $\varrho^* = 0.8$,
+and it also depended on the simulation length
+and other parameters.
+%
+Although widening the Gaussian can
+effectively suppress random error
+from local histogram fluctuation,
+it also increased the risk of including
+more systematic bias in middle-range modes,
+which can be difficult to remove
+even with the optimal schedule.
+%
+This is because the rate of error reduction
+is proportional to the eigenvalue, $\lambda_k$
+[cf. Eq. \eqref{eq:vt_diffeq_mbin}],
+which can decrease rapidly with the mode index $k$,
+cf. Eq. \eqref{eq:lambda_Gaussian}.
+
+To illustrate the mode dependence of the error,
+we show in Fig. \ref{fig:lj_xerr}
+the error components, $\langle \tilde v_k^2 \rangle$,
+against the eigenmode index, $k$,
+for the $\varrho^* = 0.1$ case
+with width $\sigma = 0.2$.
+%
+For modes $k=10$ to $25$,
+the final errors from the Gaussian updating scheme
+were greater than those from the single-bin updating scheme
+although the initial values were less.
+%
+For the Gaussian updating scheme,
+the inverse-time schedule tended to
+reduce the updating magnitude too quickly
+leaving a hump on the error profile
+for these middle-range eigenmodes.
+%
+The optimal schedule improved over the inverse-time schedule
+and produced a flatter profile for the final error.
+%
+As a result, the optimal schedule usually produced
+less error than the inverse-time schedule
+for the Gaussian updating scheme,
+as shown in Table \ref{tab:lj_error}.
+
+One way to overcome the systematic bias
+in the Gaussian updating scheme
+is to use the correction formula
+based on histogram, Eq. \eqref{eq:uhatav}.
+%
+As shown in Fig. \ref{fig:lj_xerr},
+even for the simulation using the non-optimal $1/t$ schedule,
+the correction was able to reproduce
+the flatter profile seen in the single-bin case.
+%
+However, this correction also introduced
+some local noise in the large $k$ modes,
+and it would be best used as a post-simulation
+analysis tool for simulations with detectable bias.
+%
+For example, for the simulation in the $\varrho^* = 0.8$ case
+with $\sigma = 0.5$ and the inverse-time schedule,
+the correction was able to lower
+the error from $9.8\times10^{-3}$
+(Table \ref{tab:lj_error}) to $3.3\times10^{-3}$.
+
 \begin{figure}[h]
@@
     \label{fig:lj_xerr}
-    Error components for the FDS simulation on the LJ system
-    using the Gaussian updating scheme of width $\sigma = 0.2$.
-    %
-    The optimal schedule (opt.) is more effective
+    Error components from FDS simulations
+    using the Gaussian updating scheme
+    of width $\sigma = 0.2$
+    on the LJ system at density $\varrho^* = 0.1$.
+    %
+    The optimal schedule (opt.) was more effective
     in reducing the errors of middle-range eigenmodes
     than the inverse-time schedule ($1/t$).
+    The ``corrected'' values were obtained from
+    Eq. \eqref{eq:uhatt}. 
     The lines are a guide to the eyes.
@@
 \end{figure}
+
+In Fig. \ref{fig:lj_alpha},
+we show the optimal schedules in a few cases.
+%
+Generally, the optimal schedule
+for the $\varrho^* = 0.8$ case
+was closer to the inverse-time schedule
+than that for the $\varrho^* = 0.1$ case.
+%
+This was probably because in the former case
+the first mode dominates the total error,
+$\Gamma_1 \approx \Gamma$,
+and the schedule was roughly optimized for this mode,
+resulting in
+$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
+as $\lambda_1 \approx 1$
+[cf. Eq. \eqref{eq:lambda_Gaussian}].
+%
+Since we have chosen a long equilibration period
+for the preliminary run under constant updating magnitude,
+the starting magnitude was around
+the ideal value of $a_0/2 = 5\times10^{-5}$,
+as given by Eq. \eqref{eq:half_alpha0}.
+%
+But if the Gaussian were too wide,
+e.g. if $\sigma = 0.6$,
+the eigenvalues $\lambda_k$ of some modes
+became too small to damp out the initial bias,
+resulting in much larger initial updating magnitude, $\alpha(0)$
+[cf. Eq. \eqref{eq:opt_alpha0}].
+%
+In this case, the schedule predicted by our method
+would be inaccurate because the resulting deviation in the bias potential
+was no longer small,
+and one may have to either decrease the width
+or to elongate the preliminary simulation under constant magnitude.
 
@@
 %
-For the single-bin scheme,
+The eigenvalues represent
+the relative rate of error reduction,
+and the autocorrelation integrals
+represent the intrinsic fluctuation
+of the histograms modes.
+%
+For the single-bin scheme with a set of uniform eigenvalues,
 we confirmed that the optimal schedule
@@
 including the Gaussian one,
-the optimal schedule is not given by a simple closed form,
+the errors were reduced at different paces in various eigenmodes,
+and the optimal schedule,
+representing a compromise of the eigenmodes,
+is usually not given by a simple closed form,
 but is implicitly given by an equation of motion
-of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
+of a free particle with a position-dependent mass,
+Eq. \eqref{eq:q_opt}.
 %
@@
 the optimal schedule and error
-can be sensitive to the simulation length.
+can be sensitive to the simulation length
+and other parameters.
 %
@@
 
-We have also shown that
-the single-bin scheme is optimal
+In comparing different updating schemes,
+we found that
+the single-bin updating scheme was optimal
+in error reduction
 in the long time limit
@@
 
-We believe that one of the main appeals
-of the Gaussian updating scheme
-is its ability to maintain a smooth bias potential,
-which is useful for MD applications.
+The Gaussian updating scheme
+has the appeal of being able to maintain the smoothness
+of the bias potential,
+which is important for stable MD applications.
 %
 However, it requires a careful choice of the width and
-a sufficiently long equilibration phase under constant updating magnitude.
+the updating magnitude should often be fixed at constant level
+for a sufficiently long period before
+a magnitude-decreasing schedule could start.
 %
@@
 in a post-simulation analysis
-to recover the possible accuracy loss
+to regain the possible accuracy loss
 and obtain a bias potential similar to
@@
 
-The analysis in this study has two limitations.
+The analysis in this study has several limitations.
 %
@@
 %
-This is valid in the asymptotic regime:
+This simplification is justified in the asymptotic regime:
 when the updating magnitude is small
@@
 \begin{equation}
-  \hat u_i(t) \approx u_i(t) + \frac{ h_i(t) } { p_i } - 1
-  = u_i(t) + \xi_i(t) - 1
+  \hat u_i(t) \approx u_i(t) + \frac{ h_i(t) } { \rho_i } - 1
+  = u_i(t) + f_i(t) - 1
   .
@@
 %
-We define a time average for variable, $f$ as
-\begin{equation}
-\overline{f}(t) \equiv \frac{1}{t} \sum_{\tau=1}^t f(\tau)
+Below we shall use capital letter to denote
+time average: for variable $x$, we define
+\begin{equation}
+  X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
   .
@@
 \begin{align}
-  \overline{ \hat u_i }(t)
+  \hat U_i(t)
   &\approx
-  \overline{ u_i } (t)
-  + \overline{ \xi_i } (t) - 1
+  U_i (t)
+  + F_i(t) - 1
   \approx
-  \overline{ u_i } (t)
-  + \ln \frac{ H_i(t) } { p_i }
-  ,
+  U_i(t)
+  + \ln \frac{ H_i(t) } { \rho_i }
+  .
   \label{eq:uhatav}
-\end{align}
-%
-where we have introduced average histogram
-%
-\begin{align}
-  H_i(t)
-  &= \frac 1 t \sum_{\tau = 1}^t h_i(\tau)
-  = \overline{ \xi_i } (t) \, p_i
-  .
-  \label{eq:hav_def}
 \end{align}
@@
 \begin{align}
-  \overline{ \hat u_i}(t)
+  \hat U_i(t)
   &\approx
@@
 $$
-u_i(t+1) = \overline{ \hat u_i}(t),
+u_i(t+1) = \hat U_i(t),
 $$
@@
 $$
-\overline{\hat u_i}(t) \approx \overline{ u_i }(t).
+\hat U_i(t) \approx U_i(t).
 $$
@@
 \begin{equation}
-  \xi_i(t) \approx \zeta_i(t)
+  f_i(t) \approx \zeta_i(t)
   ,
@@
 and the same holds for the Fourier transform,
-$\tilde \xi_k(t) \approx \tilde \zeta_k(t).$
+$\tilde f_k(t) \approx \tilde \zeta_k(t).$
 %
 If we define the time average,
-$$
-\tilde \Xi_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde \xi_k(t)
-=\mathcal F\left[ \frac{ H_i(T) } { p_i } \right]_k
-,
-$$
+%
+\begin{align*}
+\tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
+=\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
+,
+\end{align*}
+%
 for the histogram fluctuation of mode $k$,
@@
 &\equiv
-\sum_{k=1}^{n-1} \tilde \Xi_k^2(T)
-=
-\sum_{i=1}^{n} \rho_i \, \Xi_{*i}^2(T)
+\sum_{k=1}^{n-1} \tilde F_k^2(T)
+=
+\sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
   \notag
@@
   \left(
-  \frac{H_i(T)}{p_i}
+  \frac{H_i(T)}{\rho_i}
   -
-  \sum_{j=1}^n \rho_j \frac{ H_j(T) } {p_j}
+  1
   \right)^2
@@
 \begin{align*}
-\bigl\langle \tilde \Xi_k^2(T) \bigr\rangle
+\bigl\langle \tilde F_k^2(T) \bigr\rangle
 &=
@@
 \bigl\langle
-  \tilde \xi_k(t) \, \tilde \xi_k(0)
+  \tilde f_k(t) \, \tilde f_k(0)
 \bigr\rangle
@@
   \bigl\langle
-    \tilde \Xi_k^2(T)
+    \tilde F_k^2(T)
   \bigr\rangle
@@
 $$
-G_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde \Xi_k^2(T),
+G_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde F_k^2(T),
 $$
@@
 $$
-\max_{k=1, \dots, K-1} \left| \, \tilde \Xi_k(T) \, \right|
+\max_{k=1, \dots, K-1} \left| \, \tilde F_k(T) \, \right|
 .
@@
 However,
-without knowing the intrinsic distribution, $p^*_i$,
+without knowing the intrinsic distribution, $p_i$,
 we cannot compute the shifted bias potential, $v_i$.
@@
 We first note that the difference,
-$u_i(t) - v_i(t) = \ln(p^*_i/p_i)$, is a constant of time,
+$u_i(t) - v_i(t) = \ln(p_i/\rho_i)$, is a constant of time,
 so is the Fourier transform,
@@
 described by a transition matrix, $\mathbf A$, with
-$$
-\langle h_i(t) \rangle = p_i,
+\begin{align*}
+\langle h_i(t) \rangle = \rho_i,
 \quad
-\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, p_j
+\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, \rho_j
 \quad \mbox{(for $t \ge 0$)}
 .
-$$
+\end{align*}
 Further if the transition matrix satisfies detailed balance
 with a set of eigenvectors, $\psi_{li}$,
-that are orthonormal with respect to $\mathbf p$
+that are orthonormal with respect to $\pmb\rho$
 [cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
@@
 then we have the decomposition
-$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / p_j$,
+$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
 with $\Lambda_l$ being the eigenvalue of mode $l$.
@@
 \left\langle
-  \tilde \xi_k(t) \, \tilde \xi_k(0)
+  \tilde f_k(t) \, \tilde f_k(0)
 \right\rangle
@@
 \left(
-  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ p_i }
+  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
 \right)^2
@@
 In particular, if the eigenvectors, $\pmb \phi_k$
-and $\pmb \psi_k$, coincide and $\mathbf p = \pmb \rho$,
+and $\pmb \psi_k$, coincide,
 we would have
 $\bigl\langle
-  \tilde \xi_k(t) \, \tilde \xi_k(0)
+  \tilde f_k(t) \, \tilde f_k(0)
 \bigr\rangle
@@
   \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty \left\langle
-    \tilde \xi_k(t) \, \tilde \xi_k(0)
+    \tilde f_k(t) \, \tilde f_k(0)
   \right\rangle = \frac{1+\Lambda_k}{1-\Lambda_k}
@@
 then
-$$
+\begin{align*}
 \Gamma_k = \coth\left[
@@
   \right].
-$$
+\end{align*}
 For small $k$,
@@
 %For simplicity, we will assume that the target
-%distribution is flat, or $p_i = 1/n$, below.
+%distribution is flat, or $\rho_i = 1/n$, below.
 
@@
   Eq. \eqref{eq:wmul_to_convol}:
-$$
-\begin{aligned}
+\begin{align*}
   \sum_{j = 1}^n
@@
   .
-\end{aligned}
-$$
+\end{align*}
 The second step of Eq. \eqref{eq:wmul_to_convol}
@@
 for the left eigenvector, we have
-$$
+\begin{align*}
   \sum_{ i = 1 }^n
@@
   .
-$$
+\end{align*}
 But due to the symmetry Eq. \eqref{eq:mu_symm},
@@
 \note{Consider the unnormalized exponential equivalent
-  $$
+  \begin{align*}
   \Phi^{(k)}_j =
@@
   ,
-  $$
+  \end{align*}
   we have
-  $$
-  \begin{aligned}
+  \begin{align*}
   \sum_{j = 1}^n
@@
   .
-  \end{aligned}
-  $$
+  \end{align*}
   The term in the square brackets is the eigenvalue given by
@@
   in this case is similar,
-  $$
-  \begin{aligned}
+  \begin{align*}
     \sum_{j = 1}^n w_{ij} \, \phi_j
@@
       \mu_{i - j} \, \phi_j.
-  \end{aligned}
-  $$
+  \end{align*}
 }%
@@
   First consider the unnormalized eigenvectors,
-  $$
+  \begin{align*}
   \Phi^{(k)}_j
@@
   \cos \frac{ k \, \left( j - \frac 1 2 \right) \, \pi }{n},
-  $$
+  \end{align*}
   which satisfies Eq. \eqref{eq:phi_refl}.
@@
   So, by Eq. \eqref{eq:wmul_to_convol},
-  $$
-  \begin{aligned}
+  \begin{align*}
   \sum_{i = 1}^n
@@
   &= \Phi^{(k)}_j \, \lambda_k,
-  \end{aligned}
-  $$
+  \end{align*}
   with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}
@@
   %
-  $$
-  \begin{aligned}
+  \begin{align*}
     \sum_{ k = 0 }^{ 2 \, n - 1 }
@@
     2 \, n \, \mu_p.
-  \end{aligned}
-  $$
+  \end{align*}
   This entails Eq. \eqref{eq:mu_from_lambda}.
@@
 %
-$$
-\begin{aligned}
+\begin{align*}
 %&
@@
 \lambda_{K} = \cdots = \lambda_{n-K} = 0,
-\end{aligned}
-$$
+\end{align*}
 for $K > 1$,
@@
 \note{Derivation.
-$$
-\begin{aligned}
+\begin{align*}
 \mu_l
@@
 .
-\end{aligned}
-$$
+\end{align*}
 }
@@
 For a non-periodic variable, we demand
-$$
+\begin{align*}
 \lambda_0 = \cdots = \lambda_{K-1} = 1,
@@
 \lambda_K = \cdots = \lambda_{n-1} = 0.
-$$
+\end{align*}
 Then Eq. \eqref{eq:lambdan} gives
@@
 \note{Derivation.
-$$
-\begin{aligned}
+\begin{align*}
   \lambda_{n}
@@
   \end{dcases}
-\end{aligned}
-$$
+\end{align*}
 So
-$$
-\begin{aligned}
+\begin{align*}
   \mu_l
@@
   .
-\end{aligned}
-$$
+\end{align*}
 }%
@@
     $\nu_k$         &   $\nu_k \equiv \lambda_k / \lambda$. \\
-    $p_i$           &   Distribution. \\
-    $\pi_i$         &   (Instantaneous) distribution, Eq. \eqref{eq:pi_p_V}. \\
+    $\rho_i$        &   Flat sampling distribution. \\
+    $\pi_i$         &   (Instantaneous) distribution. \\
     $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\
     $\bar q(t)$     &   $q(T) - q(t)$.  \\
-    $\rho_i$        &   Distribution used to define the error. \\
     $s, S$          &   Stage index and the total number of stages in the Wang-Landau algorithm. \\
@@
     $\mathbf w$     &   Updating matrix. \\
-    $v_{*i}(t)$        &   Bin-average-shifted bias potential, Eq. \eqref{eq:x_def}. \\
+    $v_{*i}(t)$        &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
     $\zeta_{*i}(t)$      &   The noise of histogram. \\
-    ${\tilde v}_{*k}(t)$        &   Mode of the shifted bias potential. \\
-    ${\tilde V}_{*k}$           &   Mode of the original bias potential. \\
+    ${\tilde v}_k$        &   Mode of the shifted bias potential. \\
+    ${\tilde u}_k$           &   Mode of the original bias potential. \\
     $z$             &   Reaction coordinate. \\

### 2017-06-17 doc/response.tex
invt.tex: add back sec:seplambda

\documentclass{article}
\usepackage{amsmath}
\begin{document}

The old definition of $\rho_i$ and $p_i$ are assumed to be same now
to avoid unnecessary complexity, preserving the old $\rho_i$ only.
The old $p^*_i$ is replaced by the new $p_i$.
The old $V_i$ is replaced by $u_i$.
The old $x_i$ is replaced by $v_{*i}$.
The old $y_k$ is replaced by $\tilde v_k$.
The old $\theta_k$ is replaced by $\tilde f_k$.
The old $\zeta_i/p_i$ is replaced by $\zeta_i$.
The old $C$ is replaced by $C_M$.
The old degeneracy $g$ is replaced by $2/g$ for the multiple of period.
%
We have defined $f_i \equiv h_i/\rho_i$.
We changed $Q \to \bar q$ to imply the meaning of complement.

\end{document}

### 2017-06-17 doc/invt.tex
invt.tex: add back sec:seplambda

@@
 \begin{equation}
-  M(q')
+  M(\bar q)
   \equiv
@@
     \textstyle\sum_{ k = 1 }^{n - 1}
-    \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, q'}
-  }
-  .
-  \notag
-  %\label{eq:mass_func}
+    \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, \bar q}
+  }
+  .
+  \label{eq:mass_func}
 \end{equation}
@@
   =
-  \frac{C}{T}
+  \frac{C_M}{T}
   ,
@@
 %
-for some positive $C$.
+for some positive $C_M$.
 %
@@
   \int_{ q(T) - q(t) }^{ q(T) }
-    M(q')
+    M(\bar q)
     \;
-    d q'
-  =
-  C \, \frac t T
+    d \bar q
+  =
+  C_M \, \frac t T
   ,
@@
 %
-and $C$ can be determined from
+and $C_M$ can be determined from
 the $t = T$ case as
@@
 \begin{equation}
-  C =
+  C_M =
   \int_{ 0 }^{ q(T) }
-    M( q' )
+    M( \bar q )
     \;
-    d q'
+    d \bar q
   .
@@
   =
-  \frac { C^2 } { T }
+  \frac { C_M^2 } { T }
   =
@@
   \left(
-    \int_0^{ q(T) } M(q') \, d q'
+    \int_0^{ q(T) } M(\bar q) \, d \bar q
   \right)^2
@@
   {
-    C
+    C_M
   }
@@
   \int_{\bar q(t)}^{ q(T) }
-    m(q') \, dq'
+    m(q') \, d q'
   &=
@@
 which is approximately true for the Gaussian updating scheme
-at large $\bar q$,
+at large $\bar q$ (cf. Appendix \ref{sec:seplambda}),
 then we would have an exponentially decaying schedule.
@@
   &\equiv
-  \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(q') \, dq'
+  \frac{ M\bigl( q(T) \bigr) } { T }
+    \int_0^{q(T)} M(q') \, d q'
   -\frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
@@
 %  +
-%  \frac{2}{a_0 \, C \, M\bigl( q(T) \bigr)}
+%  \frac{2}{a_0 \, C_M \, M\bigl( q(T) \bigr)}
 %  \sum_{k = 1}^{n-1}
@@
   =
-  \frac{ 2 \, C } { T } \, M(q_T)
+  \frac{ 2 \, C_M } { T } \, M(q_T)
   .
@@
 then
-\begin{align*}
+\begin{align}
 \Gamma_k = \coth\left[
@@
   \right].
-\end{align*}
+\label{eq:Gamma_local}
+\end{align}
 For small $k$,
@@
 and by Eq. \eqref{eq:xt2_eql},
-we would expect a roughly flat profile
-for the single-bin updating scheme,
+we would expect a roughly flat tail
+for the single-bin updating scheme
+as $\langle \tilde v_k^2 \rangle \propto \Gamma_k$,
 cf. Fig. \ref{fig:lj_xerr}.
+
+\section{\label{sec:seplambda}
+Updating scheme
+with a set of well-separated eigenvalues}
+
+We will study updating schemes
+with a set of well-separated eigenvalues,
+as a model for the large $k$ eigenmodes
+of the Gaussian updating scheme.
+%
+We will show that the mass distribution will have
+an inverse-$\bar q$ profile,
+and derive a upper bound for the asymptotic error.
+
+Starting from the definition, Eq. \eqref{eq:mass_func}, we have
+$$
+  M(\bar q)
+  \le
+  \sum_{k = 2}^n
+  \sqrt{ \Gamma_k } \, \lambda_k \, e^{ -\lambda_k \, \bar q }
+  .
+$$
+Thus,
+it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
+%
+\begin{equation}
+  C_M
+  =
+  \int_0^\infty
+  M(\bar q) \, d \bar q
+  \le
+  \sum_{k = 1}^{n-1} \sqrt{ \Gamma_k }
+  ,
+  \label{eq:mint_ubound}
+\end{equation}
+%
+and the asymptotic error from Eq. \eqref{eq:error_asym2}
+is bounded as
+\begin{equation}
+  \Err_A(T)
+  =
+  \frac{ C_M^2 } { T }
+  \le
+  \frac 1 T
+  \left(
+    \sum_{k=1}^{n-1} \sqrt{ \Gamma_k }
+  \right)^2
+  .
+  \label{eq:error_asym_ubound}
+\end{equation}
+%
+It is readily seen that this bound is always
+greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
+%
+$$
+\frac{T}{(T+t_0)^2} \, \sum_{k=1}^{n-1} \Gamma_k
+\le
+\frac{1}{T} \sum_{k=1}^{n-1} \Gamma_k
+\le
+\frac{1}{T} \left( \sum_{k=1}^{n-1} \sqrt{ \Gamma_k } \right)^2
+.
+$$
+
+Next, we will show that this bound represents
+updating schemes with a set of very different eigenvalues.
+%
+First, we have for any mode $k^*$ $(k^* > 0)$,
+%
+\begin{align}
+  M(\bar q)
+  &
+  \ge
+  \sqrt{ \Gamma_{ k^* } } \,
+  \lambda_{ k^* } \,
+  e^{ -\lambda_{ k^* } \, \bar q }
+  \ge
+  \frac{ \sqrt{ \Gamma_{ k^* } } } { \bar q }
+  ,
+\label{eq:MQ_lbound1}
+\end{align}
+%
+where the first inequality
+would approach an equality if the $k^*$ mode
+dominates the mass function, $M(\bar q)$,
+and the second inequality
+represent the limit of $\lambda_{k^*} \to 1/\bar q$.
+%
+Then, for the $k$th component of the asymptotic error,
+we have from Eq. \eqref{eq:error_asym_Lagrangian}
+%
+\begin{align}
+  \left\langle
+  \tilde v_k^2(T)
+  \right\rangle_A
+  &=
+  \int_0^T
+  \Gamma_k \, \lambda_k^2 \, e^{2 \, \lambda_k \, [q(t) - q(T)] }
+  \, {\dot q}^2(t) \, dt
+  \notag
+  \\
+  &=
+  \frac{ C_M } { T }
+  \, \Gamma_k \, \lambda_k^2
+  \int_0^{q(T)}
+  \frac{ e^{-2 \, \lambda_k \, \bar q } } { M(\bar q) }
+  \,  \, d \bar q
+  \notag
+  \\
+  &\le
+  \frac{ C_M }{ T } \,
+  \frac{ \Gamma_k \, \lambda_k^2 } { \sqrt{ \Gamma_{k^*} } \, \lambda_{k^*} }
+  \int_0^{q(T)}
+    e^{-(2 \, \lambda_k - \lambda_{k^*} ) \, \bar q}
+    \, d \bar q
+  \notag
+  \\
+  &\le
+  \frac{ C_M }{ T } \,
+  \cdot
+  \frac{ \Gamma_k }
+  { \sqrt{ \Gamma_{k^*} } }
+  \cdot
+  \frac{ \lambda_k^2 }
+  { \lambda_{k^*} \, (2 \, \lambda_k -\lambda_{k^*}) }
+  ,
+  \label{eq:y2_ubound_k}
+\end{align}
+%
+where we have
+used Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:MQ_lbound1}
+on the second and third lines, respectively,
+and extended the upper limit to infinity on the last line.
+%
+For a set of well-separated eigenvalues,
+$\lambda_k$ depends on the mode index, $k$,
+much more critically than $\Gamma_k$.
+%
+Thus, the tightest upper bound
+can be approximately obtained from minimizing the last factor
+on the right-hand side of Eq. \eqref{eq:y2_ubound_k},
+and the minimal is reached at $k^* = k$,
+when $\lambda_{k^*} = \lambda_k$.
+%
+So
+%
+\begin{equation}
+  \left\langle
+  \tilde v_k^2(T)
+  \right\rangle_A
+  \le
+  %\approx
+  \frac{ C_M }{ T }
+  \sqrt{ \Gamma_k }
+  .
+  \label{eq:y2_ubound}
+\end{equation}
+%
+By summing over $k$
+and using Eq. \eqref{eq:mint_ubound},
+we recover Eq. \eqref{eq:error_asym_ubound}.
+
+From Eq. \eqref{eq:Gamma_local},
+we know that $\Gamma_k$ would quickly approach a constant.
+%
+Then Eq. \eqref{eq:y2_ubound} shows that
+the asymptotic error profile would also
+approach a plateau in this case
+(cf. Fig. \ref{fig:lj_xerr}
+for a Gaussian example).
+%
+However, to reach this upper bound
+the simulation has to be long enough
+such that the upper limit of the integral
+can be extended from $q(T)$ to infinity
+in the last step of Eq. \eqref{eq:y2_ubound_k}.
+%
+As a result,
+the flatter profile from Eq. \eqref{eq:y2_ubound}
+would stop at mode $k$,
+at which $\lambda_k \approx 1/q(T)$.
+
+Finally, from Eqs. \eqref{eq:MQ_lbound1} and \eqref{eq:Gamma_local}
+we know that for a large $\bar q$, which implies a sufficiently
+large matching mode $k^*$ with $\lambda_{k^*} \approx 1/\bar q$,
+the mass function would be roughly proportional to $1/\bar q$,
+and so is the mass distribution, i.e.
+\begin{equation}
+  m(\bar q) \propto \frac{ 1} { \bar q },
+  \qquad
+  \mbox{for large $\bar q$.}
+  \label{eq:m_invq}
+\end{equation}
 
@@
     $b$             &   Cutoff of the kernel of homogeneous updating schemes. \\
-    $C$             &   $C \equiv \int_0^{ q(T) } M(q) \, dq$.  \\
+    $C_M$           &   Integral of the mass function, $M(\bar q)$.  \\
     $\delta$        &   Dirac's or discrete $\delta$-function. \\

### 2017-06-17 doc/invt.tex
small changes

@@
 \begin{align}
-  E(T)
-  =
-  E(0) \, e^{-2 \, q(T)}
+  \Err(T)
+  =
+  \Err(0) \, e^{-2 \, q(T)}
   +
@@
 %
-$E(T) = [ E(0) \, t_0^2 + \Gamma \, T ]/(t_0 + T)^2$,
+$\Err(T) = [ \Err(0) \, t_0^2 + \Gamma \, T ]/(t_0 + T)^2$,
 %
@@
 \begin{equation}
-  t_0 = \frac{ \Gamma } { E(0) }
-  ,
-  \label{eq:t0_singlebin}
+  t_0 = \frac{ \Gamma } { \Err(0) }
+  ,
+  \notag
+  %\label{eq:t0_singlebin}
 \end{equation}
@@
 \begin{equation}
-  E(T)
+  \Err(T)
   =
@@
 \right)^2.
-\label{eq:CSineq}
+\notag
+%\label{eq:CSineq}
 \end{align}
@@
 For the single-bin updating scheme,
-it recovers Eq. \eqref{eq:t0_singlebin}
+it recovers $t_0 = \Gamma/\Err(0)$ %Eq. \eqref{eq:t0_singlebin}
 under the assumption of Eq. \eqref{eq:xt2_eql}.
@@
 From Eq. \eqref{eq:xt2_eql}, we have
-$$E(0) = \frac{1}{2} a_0 \sum_{k=1}^{n-1} \Gamma_k
+$$\Err(0) = \frac{1}{2} a_0 \sum_{k=1}^{n-1} \Gamma_k
 = \frac{1}{2} \, a_0 \, \Gamma.$$
@@
   ,
-\label{eq:error_sinc}
+  \notag
+  %\label{eq:error_sinc}
 \end{equation}
@@
   .
-  \label{eq:timeav}
+  \notag
+  %\label{eq:timeav}
 \end{equation}
@@
 .
-  \label{eq:hfluc_def}
+  \notag
+  %\label{eq:hfluc_def}
 \end{align}
@@
 =
-\frac{ \Gamma } { T }.
-  \label{eq:hfluc}
+\frac{ \Gamma } { T }
+  .
+  \notag
+  %\label{eq:hfluc}
 \end{equation}
@@
 \begin{equation}
-  E(T)
+  \Err(T)
   \ge
@@
   .
-  \label{eq:Gamma_est}
-\end{equation}
-
-The transition matrix method usually only for
+  \notag
+  %\label{eq:Gamma_est}
+\end{equation}
+
+The transition matrix method usually works only for
 simple sampling processes without major barrier
@@
     \mu(\varphi) \, \cos (k \, \varphi) \, d\varphi,
-\label{eq:lambda_int}
+  \notag
+  %\label{eq:lambda_int}
 \end{equation}

### 2017-06-17 doc/invt.tex
invt.tex: various corrections, remove Appendix B and stabilization

@@
 The optimal schedule of bandpass updating schemes
-is also the above origin-shifted inverse-time schedule.
+is also the inverse-time schedule.
 %
@@
 %
-The above two introductory sections help define our notations,
-and an expert reader may wish to skip them.
+%The above two introductory sections help define our notations,
+%and an expert reader may wish to skip them.
 %
@@
 that encompasses the Gaussian scheme used in metadynamics
-in Sec. \ref{sec:multiple-bin},
-and present a general procedure of computing the optimal schedule
-in Sec. \ref{sec:procedure}.
+in Sec. \ref{sec:multiple-bin}.
 %
@@
 are optimal for convergence
-in the long time or extensive sampling limit
+in the long time limit
 in Sec. \ref{sec:cmpschemes}.
@@
 %
-We may also derive it
-from optimally planning the simulation lengths in the stages
-and the reduction factor in the WL setup (cf. Appendix \ref{sec:wltoinvt}).
+%We may also derive it
+%from optimally planning the simulation lengths in the stages
+%and the reduction factor in the WL setup (cf. Appendix \ref{sec:wltoinvt}).
 
@@
 %
-where the bin average, $\bav{v}(t)$, is defined as
+where we have defined the bin average,
 $\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$.
@@
 Below we will use the notations, $x_{*i}$ and $\bav{x}$,
-for an arbitrary variable, $x_i$.
+for a general variable, $x_i$.
 %
@@
 \int_0^T
-  \dot y\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
+  \dot \vartheta\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
 \notag
@@
 \begin{align}
-y(q')
+\vartheta(q')
 &\equiv
 e^{q' - q(T)}.
-\label{eq:y_def}
+\label{eq:theta_def}
 \end{align}
@@
 the total error defined in Eq. \eqref{eq:error_def}
-is given by
+is a sum of a residual error (from the decay of the initial error)
+and an asymptotic error:
 %
@@
   =
-  \Err(0) \, e^{-2 \, q(T)}
-  +
-  \Gamma \,
-  \int_0^T
-    {\dot y}^2\bigl( q(t) \bigr) \,
-    \, dt
-  ,
-  \label{eq:ET_average}
+  \Err_R(T) + \Err_A(T)
+  ,
+  \label{eq:error_tot}
 \end{align}
-where $\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
+%
+where
+\begin{align}
+  \Err_R(T)
+  &= \Err(0) \, e^{-2 \, q(T)}
+  ,
+  \label{eq:ER_sbin}
+  \\
+  \Err_A(T)
+  &= \Gamma \, \int_0^T \dot \vartheta^2\bigl( q(t) \bigr) \, dt
+  ,
+  \label{eq:EA_sbin}
+\end{align}
+with
+$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
 %\begin{equation}
@@
 or, equivalently, the curve, $q(t)$,
-via $v_{*i}(T)$, and
-our aim is to find the $q(t)$
+via $v_{*i}(T)$.
+%
+If $\alpha(t)$ is a constant, $a_0$,
+so that $q(t) = a_0 \, t$,
+then after a long period,
+the error approach an equilibrium
+%
+\begin{equation}
+  \Err_\mathrm{equil.}
+  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
+  = \frac{1}{2} a_0 \Gamma
+  .
+  \label{eq:error_equil_sbin}
+\end{equation}
+%
+Our aim is to find the $q(t)$
 that minimizes this error.
@@
 %
-This is equivalent to varying the curve $y(q(t))$
+This is equivalent to varying the curve $\vartheta(q(t))$
 with fixed endpoints,
-$y\bigl( q(0) \bigr)  = e^{- q(T)}$
+$\vartheta\bigl( q(0) \bigr)  = e^{- q(T)}$
 and
-$y\bigl( q(T) \bigr) = 1$.
-%
-Note also that the first term on the right-hand side
-of Eq. \eqref{eq:ET_average} is fixed during
-the process,
-while the second term is the action of a free particle.
-Thus, the velocity of $y(q(t))$ should be a constant:
-%
-\begin{equation}
-  \dot y\bigl( q(t) \bigr) = c
-  .
-\label{eq:dydt_const}
-\end{equation}
-%
-Then, by using Eq. \eqref{eq:y_def}, we have
+$\vartheta\bigl( q(T) \bigr) = 1$.
+%
+Note also that the residual error
+is fixed during the process,
+while the asymptotic error is the action of a free particle.
+Thus, the velocity of $\vartheta(q(t))$ should be a constant:
+%
+\begin{equation}
+  \dot \vartheta\bigl( q(t) \bigr) = c
+  .
+\label{eq:dthetadt_const}
+\end{equation}
+%
+Then, by using Eq. \eqref{eq:theta_def}, we have
 %
@@
 \end{equation}
+%
 where $t_0$ is some constant,
@@
 We now determine the optimal value of $q(T)$,
-which is fixed in the above process.
+which is fixed in the above.
 %
@@
 %
-By using Eq. \eqref{eq:alpha_invt} in Eq. \eqref{eq:ET_average},
-we find the final total error to be
-%
-$\Err(T) = [ \Err(0) \, t_0^2 + \Gamma \, T ]/(t_0 + T)^2$,
-%
-which is minimal at
-%
-\begin{equation}
-  t_0 = \frac{ \Gamma } { \Err(0) }
+Using Eq. \eqref{eq:alpha_invt1} in
+Eqs. \eqref{eq:ER_sbin} and \eqref{eq:EA_sbin}
+yields
+\begin{align}
+  \Err_R(T)
+  &= \frac{ \Err(0) \, t_0^2 } { (T + t_0)^2 }
   ,
   \notag
-  %\label{eq:t0_singlebin}
-\end{equation}
-%
-and the minimum is
-%
+  %\label{eq:ER_sbin1}
+  \\
+  \Err_A(T)
+  &= \frac{ \Gamma \, T } { (T + t_0)^2 }
+  ,
+  \label{eq:EA_sbin1}
+\end{align}
+%
+and the total has a minimum of
 \begin{equation}
@@
   \frac{ \Gamma } { T + t_0 }
-  .
-  \label{eq:Emin_singlebin}
-\end{equation}
-%
-This error is roughly equal to
-the average histogram fluctuation\cite{zhou2005, zhou2008}
-(see also Appendix \ref{sec:hfluc}).
+  ,
+  \label{eq:Emin_sbin}
+\end{equation}
+%
+which is reached at $t_0 = \Gamma /\Err(0)$,
+or if we assume Eq. \eqref{eq:error_equil_sbin} for $\Err(0)$,
+%
+\begin{equation}
+  t_0 = \frac{ 2 } { a_0 } = \frac{1}{ \alpha(t = 0) }
+  .
+  \label{eq:t0_sbin}
+\end{equation}
+%
+The minimum error can be related to
+the expected histogram fluctuation\cite{zhou2005, zhou2008}
+(cf. Appendix \ref{sec:hfluc}).
 
@@
   \alpha(t) \,
+  \sum_{j=1}^n w_{ij} \, f_j(t)
+  =
+  \alpha(t) \,
   \sum_{j=1}^n w_{ij} \, \frac{ h_j(t) } { \rho_j }
@@
 %
-The key step comes from the
+The key simplification comes from the
 projection of the bias potential to the $n$ eigenmodes
@@
 \subsubsection{\label{sec:updating-matrix}
-Updating matrix}
-
-The matrix, $\mathbf w$, is subject to three conditions.
+Updating matrix and eigenmode decomposition}
+
+The matrix, $\mathbf w$, is subject to several conditions.
 %
@@
 
-%It is worth pointing out
-Finally, there is a stability condition.
-%
-In a \emph{stable} updating scheme,
-no eigenmode grows indefinitely,
-which means that, as we will show, no eigenvalue, $\lambda_k$,
-can be negative for $k \ge 1$.
-
-\subsubsection{Eigenmode decomposition}
-
 We can further define
 a generalized Fourier transform, $\mathcal{F}$,
-for variable, $A_i$,
-%
-\begin{equation}
-  {\tilde A}_k
-  \equiv \mathcal{F}[A_i]_k
-  \equiv \sum_{i = 1}^n \phi_{ki} \, A_i
+for variable, $x_i$,
+%
+\begin{equation}
+  {\tilde x}_k
+  \equiv \mathcal{F}[x_i]_k
+  \equiv \sum_{i = 1}^n \phi_{ki} \, x_i
   ,
@@
 and the inverse transform is
-%
-\begin{equation}
-  A_i =
-  \frac{1}{\rho_i} \sum_{k = 1}^n \phi_{ki} \, \tilde{A}_k
-  .
-  \notag
-  %\label{eq:gft_inv}
-\end{equation}
+$x_i = \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k / \rho_i$.
+%
+%\begin{equation}
+%  x_i = \frac{1}{\rho_i} \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k
+%  .
+%  \notag
+%  %\label{eq:gft_inv}
+%\end{equation}
 %
@@
 and Eq. \eqref{eq:eigenmode0} means that
-the bin average of any quantity, $x_i$,
-is the first mode of the Fourier transform:
-$\bav{x}= \mathcal F[x_i]_0 = \tilde x_0$.
-%
-So
+the bin average
+%of any quantity, $x_i$, is the first mode of the Fourier transform:
+$\bav{x}= \mathcal F[x_i]_0 = \tilde x_0$,
+%
+and
 \begin{equation}
@@
   \begin{dcases}
-    0           & k = 0, \\
-    \tilde x_k  & k > 0,
+    0           & k = 0
+    ,
+    \\
+    \tilde x_k  & k > 0
+    .
   \end{dcases}
@@
 %
-which means that the only effect
+Thus, the only effect
 of the bin-average deduction on $\tilde x_k$ is to
 annihilate the first mode, $\tilde x_0$.
+%
+It follows Parseval's theorem,
+\begin{equation}
+  \sum_{k=1}^{n-1} \tilde x_k \, \tilde y_k
+  =
+  \sum_{k=0}^{n-1} \tilde x_{*k} \, \tilde y_{*k}
+  =
+  \sum_{i=1}^n x_{*i} \, y_{*i}
+  \label{eq:parseval}
+\end{equation}
 
@@
 %
-where we have removed the bin average,
-and used the transformed version of Eq. \eqref{eq:sh_ave}
-[but the subscript ``$_*$'' is dropped by Eq. \eqref{eq:xtstar}].
-In the second step, we have used the linear approximation
+where in the second step,
+we have used Eq. \eqref{eq:xtstar} and
+the transformed version of Eq. \eqref{eq:sh_ave}
+by assuming the linear approximation
 as in Sec. \ref{sec:sbin_diffeq}.
-%
+
 Equation \eqref{eq:vt_diffeq_mbin}
 shows that the rate of error reduction
-is now mode-dependent,
-and the minimization of the final error
+is given by $\alpha(t) \, \lambda_k$.
+%
+Thus, a \emph{stable} updating scheme
+permits no negative eigenvalue, $\lambda_k$,
+for $k > 0$.
+%
+With a set of heterogeneous eigenvalues,
+the errors of the eigenmodes are reduced at different rates.
+%
+Thus, the optimization problem on the final error
+becomes more complicated and it often
 involves a compromise of the individual modes.
@@
 
-The error defined in Eq. \eqref{eq:error_def}
-can be rewritten in terms of the eigenmodes of $\mathbf w$,
-%
+By Eq. \eqref{eq:parseval},
+we can rewrite the error defined in Eq. \eqref{eq:error_def} as
+$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
+%
+%\begin{align}
+%  \Err(T)
+%  =
+%  \sum_{k = 1}^{n - 1}
+%    \left\langle
+%      {\tilde v}_k^2(T)
+%    \right\rangle
+%  .
+%\notag
+%\end{align}
+%
+As in Eq. \eqref{eq:error_tot},
+we can decompose the error as the sum
+of the residual error and the asymptotic error, with
 \begin{align}
-  \Err(T)
-  =
-  \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
-   \phi_{ki} \, \phi_{kj}
-    \left\langle v_{*i}(T) \, v_{*j}(T) \right\rangle
-  =
-  \sum_{k = 1}^{n - 1}
-    \left\langle
-      {\tilde v}_k^2(T)
-    \right\rangle
-  ,
-\notag
-\end{align}
-%
-where we have used
-Eq. \eqref{eq:eig_orthonormal_cols}
-and drop the first term
-by using Eq. \eqref{eq:xtstar}
-in the second step.
-
-In analogy to Eq. \eqref{eq:ET_average},
-%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:xt2_sum},
-we can write the error after a period, $T$,
-as a sum of two contributions,
-%
-\begin{equation}
-  \Err(T)
-  =
-  \Err_R(T) + \Err_A(T),
-  \label{eq:error_tot}
-\end{equation}
-%
-where the residual error
-(from the decay of the initial error)
-is given by
-%
-\begin{equation}
   \Err_R(T)
-  =
+  &=
   \sum_{k = 1}^{n-1}
@@
     \right\rangle \,
-    e^{ - 2 \, \lambda_k  \, q(T) },
+    e^{ - 2 \, \lambda_k  \, q(T) }
+  ,
   \label{eq:error_res}
-\end{equation}
-%
-and the asymptotic error
-is given by
-%
-\begin{align}
+  \\
   \Err_A(T)
@@
   \sum_{k = 1}^{n-1}
-  \Gamma_k \, \dot y_k^2\bigl( q(t) \bigr) \, dt
+  \Gamma_k \, \dot \vartheta_k^2\bigl( q(t) \bigr) \, dt
   ,
@@
 \begin{equation}
-  y_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
+  \vartheta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
   ,
@@
 
-\subsubsection{\label{sec:eqlerr}
-Initial error
-%from a simulation under a constant updating magnitude
-}
+%\subsubsection{\label{sec:eqlerr}
+%Initial error
+%%from a simulation under a constant updating magnitude
+%}
 
@@
 %
-If the starting bias potential is zero, $u_i = 0$,
+If the starting bias potential is zero,
 one can show that
@@
   \sum_{ k = 1 }^{n - 1}
-    \Gamma_k \, {\dot y}_k^2\bigl[ q(t) \bigr]
+    \Gamma_k \, {\dot \vartheta}_k^2\bigl[ q(t) \bigr]
   %\notag
@@
   %\sum_{ k = 1 }^{n - 1}
-  %  \Gamma_k \, \lambda_k^2 \, y_k^2 \bigl[ q(t) \bigr]
+  %  \Gamma_k \, \lambda_k^2 \, \vartheta_k^2 \bigl[ q(t) \bigr]
   %\; \dot q^2( t )
@@
 %
-This serves as a generalization of Eq. \eqref{eq:dydt_const}.
+This serves as a generalization of Eq. \eqref{eq:dthetadt_const}.
 %
@@
     \draw[] (0.5, 0.99488) -- (0.7, 0.99488);
-    \draw[->] (0, 0) -- node[below, inner sep=7mm] {$\bar q$} (2.8, 0);
+    \draw[->] (0, 0) -- (2.8, 0);
+    \node[] at (2.8, -0.2) {$\bar q$};
     \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 1.9);
@@
     \label{fig:massq}
-    Geometric characterization of the optimal schedule,
+    Geometric construction of the optimal schedule,
     $\alpha(t)$,
-    via the normalized mass distribution, $m(\bar q)$.
+    from the normalized mass distribution,
+    $m(\bar q)$.
     %
@@
 
-\begin{table}[h]\footnotesize
+\begin{table}[h]
   \caption{\label{tab:m_and_a}
@@
   \setlength{\tabcolsep}{3pt} % spacing between columns
-  \renewcommand\arraystretch{2.3} % spacing between rows
+  \renewcommand\arraystretch{2.0} % spacing between rows
   \begin{tabular} { c c c }
     \hline
-    $m^*(\bar q^*) = -\dfrac{d t^*}{d\bar q^*}$ &
-    $\alpha^*(t^*) = -\dfrac{d\bar q^*}{d t^*}$ &
+    $m^*(\bar q^*)$ &
+    $\alpha^*(t^*)$ &
     $\bar q^*(t^*)$
@@
     $\dfrac{c^{1/a} \, (1/a)}{\left( t^* + t^*_0 \right)^{(1/a)+1}}$ &
-    $\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a}$
+    %$\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a} - q^*_0$
+    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^{\frac 1 a} - 1\right]$
     %$\,\begin{aligned}
@@
     Here,
-    $a, q^*_0, \gamma > 0$,
+    $\gamma > 0$,
     and we have defined
-    $s_\gamma = 1/(e^\gamma - 1)$,
+    $s_\gamma \equiv 1/(e^\gamma - 1)$,
     ${t^*_0}^{-1} \equiv \left(1+{q^*_0}^{-1}\right)^a - 1$,
@@
 Note that
-the updating scheme should be stable
-with no negative eigenvalue
-(cf. the stabilization procedure in Appendix \ref{sec:stabilize_wband}).
+no eigenvalue can be negative for the updating scheme to be stable.
 
@@
 and we imposed a lower bound,
-$\bigl|\mathcal Y/q(T)\bigr|$, for
+$\bigl|\mathcal Y\bigl(q(T)\bigr)/q(T)\bigr|$, for
 $\mathcal Y'\bigl(q(T)\bigr)$
@@
   dt \sum_{k = 1}^{n-1} \Gamma_k \,
-    \left( {\dot y}_k \, \chi - c_k \right)^2
+    \left( {\dot \vartheta}_k \, \chi - c_k \right)^2
   \equiv
@@
     \sum_{k = 1}^{n-1}
-      \Gamma_k \, {\dot y}_k^2\bigl( q(t) \bigr)
+      \Gamma_k \, {\dot \vartheta}_k^2\bigl( q(t) \bigr)
 \right)
@@
     \sum_{k = 1}^{n-1}
-      \Gamma_k \, c_k \, {\dot y}_k \, \bigl( q(t) \bigr)
+      \Gamma_k \, c_k \, {\dot \vartheta}_k \, \bigl( q(t) \bigr)
 \right)^2
@@
 Thus, the inequality sets a lower bound
-for the asymptotic error $\Err(T)$
+for the asymptotic error $\Err_A(T)$
 given in Eq. \eqref{eq:error_asym}.
@@
 The equality is achieved
-if $\dot y_k\left( q(t) \right) = c_k$
+if $\dot \vartheta_k\left( q(t) \right) = c_k$
 for all $k > 0$ at any $t$,
@@
 Solving these equations yields
-%
-\begin{equation}
-  \alpha(t) = \frac{              1             }
-                   { \lambda_k \, (t + t_{k0} ) },
-\label{eq:alpha_invtk}
-\end{equation}
-with
-\begin{equation}
-  t_{k0} = \frac{             T            }
-                { e^{ \lambda_k q(T) } - 1 }.
-\label{eq:tk0}
-\end{equation}
-%
-\note{Derivation.
-  Integrating $\dot y_k = c_k$ yields
-  \begin{equation}
-    y_k(t) = c_k \left(t + t_{k0} \right).
-  \tag{UK}
-  \label{neq:ui_solution}
-  \end{equation}
-  Given that $y_k(T) = 1$ and $y_k(0) = e^{-\lambda_k \, q(T)}$,
-  we have
-  $$
-  c_k = \frac{ 1 }{ T + t_{k0} },
-  \quad
-  \mathrm{and\;}
-  \frac{ t_{k0} } { T + t_{k0} }
-  =
-  e^{ -\lambda_k \, q(T) }.
-  $$
-  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
-  $$
-  -\lambda_k \, q(T) + \lambda_k \, q(t)
-  = \ln c_k + \ln \left( t + t_{k0} \right).
-  $$
-  Differentiating this with respect to $t$ yields
-  Eq. \eqref{eq:alpha_invtk}.
-}
-%
-Such a solution is possible only if
-all nontrivial eigenvalues are equal,
-%
-%\begin{equation}
-  $\lambda_1 = \dots = \lambda_{n-1}$.
-%  \notag
-%\end{equation}
-%
-Then, by Eq. \eqref{eq:tk0},
-all $t_{k0}$'s also share the same value,
-$t_0$,
-such that
-$y_k = (t + t_0) / (T + t_0)$,
-and
-$c_k = 1/(T + t_0)$
-for $k > 0$.
-%
-The asymptotic error is then
-%
-\begin{align}
-  \Err_A(T)
-  =
-  \frac{       T     }
-       { (T + t_0)^2 }
-  \sum_{ k = 1 }^{n-1}
-    \Gamma_k
-  =
-  \frac{ \Gamma \, T     }
-       { (T + t_0)^2 }
-  ,
-\label{eq:error_asym_singlebin}
-\end{align}
-%
-where we have used
+a set of identical eigenvalues,
+$\lambda_1 = \cdots = \lambda_{n-1}$
+and identical $c_k$'s.
+
+Since an updating matrix
+with identical eigenvalues\footnote{Since
+  the eigenmode of $k = 0$ represents
+  a uniform shift of the bias potential,
+  we can freely set $\lambda_0$ to be
+  the same as $\lambda_1$
+  without changing the nature of
+  the updating scheme.}
+is a multiple of the identity matrix,
+%
+and if the eigenvalues are set to $1$,
+we recover the single-bin scheme.
+%
+This means that the single-bin scheme is most efficient
+in error reduction in the asymptotic limit.
+%
+In particular,
+with $\vartheta_k = (t+t_0)/(T+t_0)$,
+we can readily recover the optimal schedule,
+Eq. \eqref{eq:alpha_invt1}, with
+the asymptotic error being
+$\Err_A(T) = T / (T + t_0)^2 \sum_{ k = 1 }^{n-1} \Gamma_k$,
+%
+%\begin{align}
+%  \Err_A(T)
+%  =
+%  \frac{       T     }
+%       { (T + t_0)^2 }
+%  \sum_{ k = 1 }^{n-1}
+%    \Gamma_k
+%  ,
+%\end{align}
+%
+which is identical to Eq. \eqref{eq:EA_sbin1}
+if we notice [cf. Eq. \eqref{eq:parseval}]
+%
+%where we have used
 %
@@
   &=
-  \sum_{t=-\infty}^\infty \sum_{k=1}^{n-1} \langle {\tilde v}_k(t) \, {\tilde v}_k(0) \rangle
+  \sum_{t=-\infty}^\infty \sum_{k=1}^{n-1}
+  \bigl\langle {\tilde \zeta}_k(t) \, {\tilde \zeta}_k(0) \bigr\rangle
   \notag
@@
   &=
-  \sum_{t=-\infty}^\infty \sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle
+  \sum_{t=-\infty}^\infty \sum_{i=1}^n
+  \rho_i \bigl\langle \zeta_{*i}(t) \, \zeta_{*i}(0) \bigr\rangle
   = \sum_{i=1}^n \rho_i \, G_i = \Gamma
   .
-  \notag
-  %\label{eq:Gamma_G_sum}
+  %\notag
+  \label{eq:Gamma_G_sum}
 \end{align}
-%
-%
-If we further assume $\lambda_1 = \cdots = \lambda_{n-1} = 1$,
-%the optimal schedule given by
-Eq. \eqref{eq:alpha_invtk}
-recovers Eq. \eqref{eq:alpha_invt1}.
-%
-
-Since the first eigenmode represents
-a uniform shift of the bias potential,
-we can freely set $\lambda_0$ to be the same as $\lambda_1$
-without changing the nature of the updating scheme.
-%
-Since an updating matrix
-with identical eigenvalues
-is a multiple of the identity matrix,
-%
-the corresponding updating scheme,
-the single-bin scheme, is most efficient
-in asymptotic convergence.
-%if no eigenvalue $\lambda_k$ is zero.
-%
-%In Appendix \ref{sec:upperbound},
+%%
+%\begin{equation}
+%  \alpha(t) = \frac{              1             }
+%                   { \lambda_k \, (t + t_{k0} ) },
+%\label{eq:alpha_invtk}
+%\end{equation}
+%with
+%\begin{equation}
+%  t_{k0} = \frac{             T            }
+%                { e^{ \lambda_k q(T) } - 1 }.
+%\label{eq:tk0}
+%\end{equation}
+%%
+%\note{Derivation.
+%  Integrating $\dot \vartheta_k = c_k$ yields
+%  \begin{equation}
+%    \vartheta_k(t) = c_k \left(t + t_{k0} \right).
+%  \tag{UK}
+%  \label{neq:ui_solution}
+%  \end{equation}
+%  Given that $\vartheta_k(T) = 1$ and
+%  $\vartheta_k(0) = e^{-\lambda_k \, q(T)}$,
+%  we have
+%  $$
+%  c_k = \frac{ 1 }{ T + t_{k0} },
+%  \quad
+%  \mathrm{and\;}
+%  \frac{ t_{k0} } { T + t_{k0} }
+%  =
+%  e^{ -\lambda_k \, q(T) }.
+%  $$
+%  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
+%  $$
+%  -\lambda_k \, q(T) + \lambda_k \, q(t)
+%  = \ln c_k + \ln \left( t + t_{k0} \right).
+%  $$
+%  Differentiating this with respect to $t$ yields
+%  Eq. \eqref{eq:alpha_invtk}.
+%}
+%%
+%Such a solution is possible only if
+%all nontrivial eigenvalues are equal,
+%%
+%%\begin{equation}
+%  $\lambda_1 = \dots = \lambda_{n-1}$.
+%%  \notag
+%%\end{equation}
+%%
+%Then, by Eq. \eqref{eq:tk0},
+%all $t_{k0}$'s also share the same value,
+%$t_0$,
+%such that
+%$\vartheta_k = (t + t_0) / (T + t_0)$,
+%and
+%$c_k = 1/(T + t_0)$
+%for $k > 0$.
+%%
+%
+%
+
+%
+%
+%In Appendix \ref{sec:seplambda},
 %we will show that the opposite limit of
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-that ignore short-wavelength eigenmodes,
+that completely eliminates certain short-wavelength eigenmodes,
 which are typically local noise.
@@
 the single-bin updating scheme
-supplied with a perfect filter
-for the noise modes,
-and zero eigenvalues are assigned to the noise modes.
+followed by a perfect filter
+for the designated noise modes,
+and zero eigenvalues are assigned to these modes.
 %
@@
 %
-we can construct an updating scheme with
-$$
+we demand
+%
+\begin{equation}
   \lambda_0 = \cdots = \lambda_{K-1} = 1,
@@
   \lambda_K = \cdots = \lambda_{n-1} = 0.
-$$
+  \label{eq:lambda_bandpass}
+\end{equation}
 %
@@
 %
-$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj},$$
+$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj}.$$
 %
@@
 %
-with
-%
-\begin{equation}
-  t_0
-  =
-  \frac{        1      }
-       { \alpha(t = 0) }
-  =
-  \frac{  2  }
-       { a_0 }
-  ,
-\label{eq:t0_sinc}
-\end{equation}
-%
-where we have used Eq. \eqref{eq:half_alpha0}
-in the second step.
-%
-\note{
-For the single-bin updating scheme,
-it recovers $t_0 = \Gamma/\Err(0)$ %Eq. \eqref{eq:t0_singlebin}
-under the assumption of Eq. \eqref{eq:xt2_eql}.
-%
-From Eq. \eqref{eq:xt2_eql}, we have
-$$\Err(0) = \frac{1}{2} a_0 \sum_{k=1}^{n-1} \Gamma_k
-= \frac{1}{2} \, a_0 \, \Gamma.$$
-}
-%
-%From Eqs. \eqref{eq:error_tot},
-%\eqref{eq:error_res},
-%\eqref{eq:tk0},
-%and
-%\eqref{eq:error_asym_sinc},
-%we get
-The final error is given by
-%
-\begin{equation}
-  \Err(T)
-  =
-  \frac{   1     }
-       { T + t_0 }
-  \sum_{ k = 1 }^{ K - 1 }
-    \Gamma_k
-  ,
-  \notag
-  %\label{eq:error_sinc}
-\end{equation}
-%
-as a generalization of Eq. \eqref{eq:Emin_singlebin}
-for the single-bin updating scheme.
+with $t_0$ given by Eq. \eqref{eq:t0_sbin}.
+%
+%The asymptotic error is given by
+%%
+%\begin{equation}
+%  \Err_A(T)
+%  =
+%  \frac{   T     }
+%       { (T + t_0)^2 }
+%  \sum_{ k = 1 }^{ K - 1 }
+%    \Gamma_k
+%  ,
+%  \notag
+%  %\label{eq:error_sinc}
+%\end{equation}
+%%
+%as a generalization of Eq. \eqref{eq:Emin_sbin}
+%for the single-bin updating scheme.
 
@@
 time average: for variable $x$, we define
-\begin{equation}
-  X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
-  .
-  \notag
-  %\label{eq:timeav}
-\end{equation}
-%
-Then after time averaging, Eq. \eqref{eq:uhatt} becomes
+$X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)$.
+%\begin{equation}
+%  X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
+%  .
+%  \notag
+%  %\label{eq:timeav}
+%\end{equation}
+%
+Time averaging Eq. \eqref{eq:uhatt} yields
 \begin{align}
@@
 %
-Another special case is $\alpha(\tau) = 0$ for $\tau = 0, 1, \dots$,
-then Eq. \eqref{eq:Uhat_sum} is reduced to a plain average\cite{zhou2005}
-$$
-\hat U_i(t) \approx U_i(t).
-$$
+Another special case is $\alpha(\tau)$ is a constant for $\tau = 0, 1, \dots$,
+then Eq. \eqref{eq:Uhat_sum} is reduced to a plain average\cite{zhou2005}:
+$\hat U_i(t) \approx U_i(t)$,
 if we can assume an equilibrium condition
@@
 
-\section{\label{sec:wltoinvt}
-Inverse-time schedule as the optimal limit of the WL algorithm}
-
-We may understand the inverse-time schedule %, Eq. \eqref{eq:},
-as an optimal limit of the stage-wise WL updating scheme.
-%
-%Our calculation below will focus on the one-variable problem,
-%as in Sec. \ref{sec:onevar}.
-
-To begin, we split the entire simulation into $S$ stages,
-each with a fixed magnitude, $\alpha_s$ ($1 \le s \le S$).
-%
-We denote the number of steps in stage $s$ as $T_s$, and
-$\sum_{s=1}^S T_s = T$.
-%
-%\begin{equation}
-%  \sum_{s = 1}^S T_s = T
-%  ,
-%  \label{eq:sumT}
-%\end{equation}
-%
-We assume that
-the magnitude follows geometrically,
-%
-\begin{equation}
-  \alpha_s = \alpha_0 \, r^s
-  ,
-  \label{eq:alphas}
-\end{equation}
-%
-with $0 < r < 1$.
-%
-%and that the system is properly equilibrated
-%under constant magnitude initially.
-%
-We wish to find the best distribution of the simulation time
-into the $S$ stages
-as well as
-the optimal number of stages and $r$.
-
-If we denote the error at the end of stage $s$ as $E_s$,
-then $q\left( \tau^{(s)} \right) = \alpha_s \, \tau^{(s)}$,
-where $\tau^{(s)}$ denotes the number of steps in stage $s$,
-and from Eq. \eqref{eq:ET_average}, we have
-%
-\begin{align}
-  E_s
-  &=
-  E_{s - 1} \, e^{-2 \, \alpha_s \, T_s }
-  +
-  \Gamma \, \int_0^{T_s}
-  \alpha_s^2 \,
-  e^{ -2 \, \alpha_s \, \left(T_s - \tau^{(s)}\right) }
-  \, d \tau^{(s)}
-  \notag
-  \\
-  &=
-  \left(
-    E_{s - 1} - \frac{ \Gamma \, \alpha_s } { 2 }
-  \right)
-  e^{ - 2 \, \alpha_s \, T_s }
-  +
-  \frac{ \Gamma \, \alpha_s } { 2 }
-  .
-  \label{eq:errorstage}
-\end{align}
-%
-Our aim is to minimize the error
-of the final stage, $E_S$.
-%
-To do so, let us consider shifting simulation time
-by $\delta T$ from
-stage $s+1$ to stage $s$,
-then $E_{s+1}$ is affected as
-%
-\begin{align*}
-  \delta E_{s+1}
-  =
-\left[
-  2 \, \alpha_{s+1}
-  \left(
-    E_s - \tfrac{ \Gamma \, \alpha_{s+1} } { 2 }
-  \right)
-  +
-  \frac{ \partial E_s } { \partial T_s }
-\right]
-e^{-2 \, \alpha_{s+1} \, T_{s+1} }
-\,
-\delta T
-.
-\end{align*}
-%
-Since this variation does not directly affect the error of later stages,
-the above expression must be zero to minimize the final error, i.e,
-%
-\begin{align*}
-  2 \, \alpha_{s+1}
-  \left(
-    E_s - \frac{ \Gamma \, \alpha_{s+1} } { 2 }
-  \right)
-  &=
-  -\frac{ \partial E_s } { \partial T_s }
-  %\\
-  %&=
-  %2 \, \alpha_s \, e^{-2 \, \alpha_s \, T_s}
-  %\left(
-  %  E_{s-1} - \frac{ \Gamma \, \alpha_s } { 2 }
-  %\right)
-  %\\
-  =
-  2 \, \alpha_s \, \left(E_s - \frac{ \Gamma \, \alpha_s } { 2 }  \right)
-  ,
-\end{align*}
-%
-where we have used Eq. \eqref{eq:errorstage}
-in the second step.
-%
-Thus,
-\begin{equation}
-  E_s
-  =
-  \frac{ \Gamma } { 2 } \, ( \alpha_s + \alpha_{s+1} )
-  =
-  \frac{ \Gamma } { 2 } \, \alpha_s ( 1 + r )
-  .
-  \label{eq:Es}
-\end{equation}
-%
-Using Eqs. \eqref{eq:Es} and \eqref{eq:alphas}
-in Eq. \eqref{eq:errorstage}, we get
-%
-\begin{equation}
-  e^{-\alpha_s \, T_s } = r
-  .
-  \label{eq:exp_r}
-\end{equation}
-%
-
-To compute the optimal schedule,
-we map the overall simulation time, $t$,
-to the stage number, $s$, as
-%
-\begin{align*}
-  t
-  = \sum_{s' = 1}^s T_{s'}
-  = \frac{ 1 - r^s } { 1 - r } \, T_s
-  = \frac{ 1 } { \gamma } \,
-  \left(
-    \frac{1}{\alpha_s}
-    -
-    \frac{1}{\alpha_0}
-  \right)
-  ,
-\end{align*}
-where
-$\gamma = ( 1 - r ) / \ln(1/r)$,
-and we have used Eq. \eqref{eq:exp_r} in the last step.
-%
-Thus,
-%
-\begin{equation}
-  \alpha_s =
-  \frac{ 1 }
-  { \gamma \, t + \alpha_0^{-1} }
-  ,
-  \label{eq:alphas_t}
-\end{equation}
-%
-and the final error is given by
-%
-\begin{align}
-  E_S
-  =
-  \frac{\Gamma}{2}
-  \frac{ 1 + r }
-  { \gamma \, T + \alpha_0^{-1} }
-  ,
-  \notag
-\end{align}
-%
-which is minimized as $r \to 1$
-for a large $T$
-(at this point, $\gamma \to 1$).
-%
-\note{
-  For a large $T$, $E_S \approx \frac{\Gamma}{2T} \frac{1+r}{\gamma}$.
-  Define $\delta = 1 - r$,
-  we have $1/\gamma = 1+ \delta/2 + \delta^2/3 + \cdots$,
-  and
-  $$
-  \frac{1+r}{\gamma}
-  =
-  (2 - \delta)
-  \left(
-  1 + \frac{\delta}{2} + \frac{\delta^2}{3} + \cdots
-  \right)
-  =
-  2 + \sum_{n=1}^\infty \frac{n-1}{n(n+1)}\delta^n \ge 2
-  .
-  $$
-}
-%
-If we identify the parameter, $\alpha_0$,
-as the initial updating magnitude, $\alpha(t = 0)$,
-then this result agrees with Eq. \eqref{eq:t0_sinc},
-with $t_0 = \alpha_0^{-1}$,
-and the schedule given by Eq. \eqref{eq:alphas_t}
-is identical to that from Eq. \eqref{eq:alpha_invt1}.
-
 \section{\label{sec:hfluc}
@@
 Here we wish to show that for an FDS simulation
-with a nearly-perfect bias potential
-(e.g. a long WL simulation under the single-bin updating scheme
-with the inverse-time schedule),
+with a nearly-perfect bias potential,
 the histogram fluctuation
-is roughly equal to the final error.
+is roughly equal to the final error
+from a long FDS simulation under the single-bin updating scheme
+with the inverse-time schedule.
 %
@@
 \begin{align}
-G(T)
+  F^2(T)
   &\equiv
@@
   \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
-  \notag
-  \\
-  &=
-  \sum_{i=1}^n \rho_i
-  \left(
-  \frac{H_i(T)}{\rho_i}
-  -
-  1
-  \right)^2
+  =
+  \sum_{i=1}^n
+  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
   .
@@
 \end{align}
+%
 From Eq. \eqref{eq:xi_zeta}, we have
+%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
+%=
+%\sum_{t=-\infty}^\infty
+%\frac{
+%\bigl\langle
+%  \tilde f_k(t) \, \tilde f_k(0)
+%\bigr\rangle } { T }
+%\approx
+%\sum_{t=-\infty}^\infty
+%\frac{
+%\bigl\langle
+%  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
+%\bigr\rangle } { T }
+%=
+%\frac{ \Gamma_k } { T },$$
 %
@@
 \frac{ \Gamma_k } { T }
-.
+,
 \end{align*}
 %
-The total histogram fluctuation is
-\begin{equation}
-\bigl\langle G(T) \bigr\rangle
-\equiv
-\sum_{k=1}^{n-1}
-  \bigl\langle
-    \tilde F_k^2(T)
-  \bigr\rangle
-=
-\frac{ \Gamma } { T }
-  .
-  \notag
-  %\label{eq:hfluc}
-\end{equation}
-%
-Comparing this to Eq. \eqref{eq:Emin_singlebin},
+and the total histogram fluctuation is
+$\bigl\langle F^2(T) \bigr\rangle = \Gamma/T$.
+%
+Comparing this to Eq. \eqref{eq:Emin_sbin},
 we may rephrase the optimality of the inverse-time schedule
@@
   \ge
-  \bigl\langle G(T+t_0) \bigr\rangle
+  \bigl\langle F^2(T+t_0) \bigr\rangle
   .
@@
 for the histogram fluctuation
-$$
-G_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde F_k^2(T),
-$$
+$F^2_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde F_k^2(T),$
 since the histogram flattening
@@
 For a local sampling process,
-if we borrow \eqref{eq:lambda_Gaussian}
+if we borrow Eq. \eqref{eq:lambda_Gaussian}
 for $\Lambda_k$,
@@
   \left(
-    \frac{ \pi \, k \, \sigma_z } { g \, \Delta z}
+    \frac{ \pi \, k \, \sigma_z^\mathrm{(samp)} } { g \, \Delta z}
   \right)^2
@@
 For small $k$,
-we have $\Gamma_k \propto \Delta z^2/(\sigma_z \, k)^2$,\cite{bussi2006}
-which means that a doubled sampling range, $\Delta z$,
+we have $\Gamma_k \propto \Delta z^2/k^2$,\cite{bussi2006}
+which means that doubling the sampling range, $\Delta z$,
 would make the histogram flattening
@@
 %
-Using finer bins would, however, not change this value.
+%Using finer bins would, however, not change this value.
 %
 For a modestly large $k$,
-$\Gamma_k \approx 1$.
+$\Gamma_k \approx 1$,
 and by Eq. \eqref{eq:xt2_eql},
@@
 It is readily seen that this bound is always
-greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
+greater than the lower bound given by
+Eq. \eqref{eq:EA_sbin1} [cf. Eq. \eqref{eq:Gamma_G_sum}]:
 %
@@
 \begin{equation}
-  \phi_j = \phi_{j \pm n},
+  \phi_i = \phi_{i \pm n},
 \label{eq:phi_pbc}
@@
 %
-such that $j \pm n$ lies in between $1$ and $n$.
+such that $i \pm n$ lies in between $1$ and $n$.
 %
@@
 \begin{equation}
-  \phi^{(k)}_j
-  =
-  \phi_{kj}
+  \phi^{(k)}_i
+  =
+  \phi_{ki}
   =
@@
   \left(
-    \frac{ k \, j \, 2 \, \pi }
+    \frac{ k \, i \, 2 \, \pi }
          {      n             }
@@
   \begin{align*}
-  \Phi^{(k)}_j =
+  \Phi^{(k)}_i =
   \exp\left[
-    \frac{ k \, j \, 2 \, \pi }
+    \frac{ k \, i \, 2 \, \pi }
          {      n             }
@@
 representing two reflective mirrors at
-$j = 1/2$ and $j = n + 1/2$,
+$i = 1/2$ and $i = n + 1/2$,
 are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
@@
 %For a non-periodic variable,
-We define the out-of-boundary values
-of $\phi_j$ as
-%
-\begin{equation}
-  \phi_j
+We define the out-of-boundary values as
+%
+\begin{equation}
+  \phi_i
   =
   \begin{dcases}
-    \phi_{ 1 - j }           & \mathrm{for \;} j \le 0, \\
-    \phi_{ 2 \, n + 1 - j }  & \mathrm{for \;} j > n,
+    \phi_{ 1 - i }           & \mathrm{for \;} i \le 0, \\
+    \phi_{ 2 \, n + 1 - i }  & \mathrm{for \;} i > n,
   \end{dcases}
@@
 defines a periodic variable of period $2 \, n$
-with a reflective symmetry around $j = 1/2$,
+with a reflective symmetry around $i = 1/2$,
 the eigenvectors,
@@
 \begin{equation}
-  \phi^{(k)}_j
-  =
-  \phi_{k j}
+  \phi^{(k)}_i
+  =
+  \phi_{k i}
   =
@@
   \cos \left[
-       \frac{ k \, \left( j - \frac 1 2 \right) \, 2 \, \pi}
+       \frac{ k \, \left( i - \frac 1 2 \right) \, 2 \, \pi}
             {             2 \, n                           }
@@
 For $n \gg 1$,
-we define an equivalent angular variable, $\varphi$,
-over $[0, 2 \, \pi/g]$
+we define an equivalent continuous variable, $z$,
+over $[0, \Delta z]$
 with bin size
-$\delta \varphi = 2 \, \pi/(g \, n)$.
-The variable is related to the bin index, $l$, as
-$\varphi = l \, \delta \varphi$,
+$\delta z = \Delta z/(g \, n)$.
+The variable is related to the bin index, $i$, as
+$z = i \, \delta z$,
 and
-$\mu_l = \mu(\varphi) \, \delta \varphi$.
+$\mu_i = \mu(z) \, \delta z$.
 %
@@
   =
-  2 \int_0^\pi
-    \mu(\varphi) \, \cos (k \, \varphi) \, d\varphi,
+  \int_{-\Delta z/2}^{\Delta z/2}
+    \mu(z) \, \cos\left( \frac{ 2 \, \pi \, k \, z } { \Delta z} \right)
+    \, d z,
   \notag
@@
 with the normalization
-%
-\begin{equation}
-  1 = 2 \int_0^\pi \mu(\varphi) \, d\varphi.
-\label{eq:mu_int}
-\end{equation}
-\note{
-\begin{align*}
-  \lambda_k
-  &\approx
-  \int_0^b \mu_l \, \cos(k \, \varphi) \, dl
-  \\
-  &=
-  \int_0^b \mu(\varphi) \, \cos(k \, \varphi) \, \delta \varphi \, dl
-  \\
-  &=
-  \int_0^{b \, \delta \varphi} \mu(\varphi) \, \cos(k \, \varphi) \, d\varphi
-\end{align*}
-The upper limit,
-$b \, \delta \varphi \approx (g\,n/2) \, [2\,\pi/(g\,n)] = \pi$.
-}
-
-If the width of the Gaussian updating kernel
-is very small compared to $\sigma_\varphi$,
-then we can extend the upper limit of the integral
-in Eq. \eqref{eq:mu_int} to $\infty$, and
-%
-\begin{equation}
-  \mu(\varphi)
-  =
+$1 = \int_{-\Delta z/2}^{\Delta z/2} \mu(z) \, dz$.
+
+If the width of the Gaussian updating kernel, $\sigma_z$,
+is very small compared to $\Delta z$,
+then we can extend the limits of the integrals
+to infinity, and
+%
+\begin{equation}
+  \mu(z)
+  \approx
   \frac{            1            }
-       { \sqrt{ 2 \pi } \sigma_\varphi }
+       { \sqrt{ 2 \pi } \sigma_z }
   %
   \exp\left(
-        -\frac{   \varphi^2   }
-              { 2 \, \sigma_\varphi^2 }
+        -\frac{   z^2   }
+              { 2 \, \sigma_z^2 }
       \right)
@@
 which shows that all eigenvalues are positive.
-
-In general, some eigenvalues can turn negative
-because of truncation and discretization.
-%
-We can however maintain the stability of updating scheme
-by starting with Eq. \eqref{eq:lambda_Gaussian}
+%
+In practice, however, some eigenvalues can turn negative
+because of the truncation and discretization.
+%
+We can maintain the stability of the updating scheme
+by redefining the Gaussian scheme
+starting with Eq. \eqref{eq:lambda_Gaussian}
 and inversely computing the kernel using Eq. \eqref{eq:mu_from_lambda}.
@@
 For the bandpass updating schemes,
-the eigenvalues are known ($0$ or $1$) and
+the eigenvalues are either $0$ or $1$ and
 we can use Eq. \eqref{eq:mu_from_lambda}
@@
 %
-Below we will slightly modify the definition of
-the mode cutoff, $K$,
-for mathematical convenience.
-
-For a periodic variable, we demand
-%
-\begin{align*}
-%&
-%\lambda_0 = \cdots = \lambda_{K-1}
-%= \lambda_{n-K+1} = \cdots = \lambda_n = 1,
-%\\
-%&
-\lambda_{K} = \cdots = \lambda_{n-K} = 0,
-\end{align*}
-for $K > 1$,
-with the rest of the eigenvalues being $1$.
+For a periodic variable,
+to take into account the two-fold degeneracy,
+we will modify the mode cutoff such that
+$\lambda_{K} = \cdots = \lambda_{n-K} = 0$,
+and the rest of the eigenvalues are $1$.
 %
@@
 
-For a non-periodic variable, we demand
-\begin{align*}
-\lambda_0 = \cdots = \lambda_{K-1} = 1,
-\qquad
-\lambda_K = \cdots = \lambda_{n-1} = 0.
-\end{align*}
-Then Eq. \eqref{eq:lambdan} gives
-$\lambda_n = (-1)^{n-K}$,
-and from Eq. \eqref{eq:mu_from_lambda},
+For a non-periodic variable,
+we have $\lambda_n = (-1)^{n-K}$
+from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan}.
+Then, from Eq. \eqref{eq:mu_from_lambda},
 we have
@@
 %%
-
-\subsection{\label{sec:stabilize_wband}
-Stabilization}
-
-%A practical problem of using homogeneous updating scheme
-%is the following.
-%
-A homogeneous updating scheme
-specified by an arbitrary kernel
-is not necessarily stable,
-i.e. some eigenvalues can be negative.
-%
-Here we show how to minimally modify
-the updating kernel
-to stabilize it.
-%
-\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
-}
-
-Given an updating kernel,
-we can compute the eigenvalues from
-Eq. \eqref{eq:wband_eigenvalue}.
-%
-By replacing negative eigenvalues with zeros and
-using the inversion formula,
-Eq. \eqref{eq:mu_from_lambda},
-we get a new updating kernel,
-which is stable by construction.
-%
-However, the new kernel is widened
-as $\mu_l$ may be nonzero for $l > b$.
-%
-If we truncate the new kernel at $\mu_b$,
-the stability problem remains,
-although the negative eigenvalues
-tend to be smaller in magnitude.
-%
-Thus, to stabilize the updating kernel
-without increasing the width,
-we need to iterate the above process many times,
-resulting the following algorithm.
-
-%
-\begin{enumerate}
-  \item
-    Given an updating kernel, $\mu_0, \dots, \mu_b$,
-    compute the eigenvalues,
-    $\lambda_1, \dots, \lambda_{n-1}$,
-    from Eq. \eqref{eq:wband_eigenvalue}.
-  \item
-    If all eigenvalues are nonnegative within the numerical error,
-    we are done. % and exit the loop.
-  \item
-    Otherwise, replace negative eigenvalues by zeros,
-    and compute the new kernel,
-    $\mu_0, \dots, \mu_b$, from
-    Eq. \eqref{eq:mu_from_lambda},
-    with $\mu_l = 0$ for $l > b$.
-    Go to step 1.
-\end{enumerate}
 
@@
     $T$             &   Simulation length. \\
-    $y_k(q')$       &   $y_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
+    $\vartheta_k(q')$       &   $\vartheta_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
     $u_i(t)$        &   Original bias potential. \\

### 2017-06-20 doc/invt.tex
invt.tex: add polynomial basis

@@
 
-\title{Optimal updating factor in adaptive flat-distribution-sampling simulations}
+\title{Optimal updating magnitude in adaptive flat-distribution-sampling simulations}
 
@@
 and metadynamics\cite{huber1994,
-*laio2002, *laio2008, *barducci2011, *sutto2012},
-popular for MC and MD simulations, respectively\cite{junghans2014}.
+  *laio2002, *laio2008, *barducci2011, *sutto2012, micheletti2004}
+among others\cite{kim2006, *kim2007, junghans2014,
+  langfeld2012, pellegrini2014}.
 %
@@
 %
-Although the two window functions
-are not always associated with
-the two methods\cite{micheletti2004, kim2006, *kim2007, junghans2014},
-we will colloquially use the terms WL and metadynamics
-to represent FDS simulations using the
-single-bin and Gaussian window functions, respectively.
+We will refer to the two updating schemes
+as the single-bin
+and Gaussian updating schemes, respectively.
+%
+While the former is popular for MC simulations\cite{wang2001,
+  wang2001pre, kim2006, *kim2007},
+the latter is more popular for MD simulations
+as it provides a smoother bias potential
+that can be readily differentiated for force calculation.\cite{huber1994,
+  *laio2002, *laio2008, *barducci2011, *sutto2012, junghans2014}
 %
@@
 }
-The same schedule also applies to a variant of
-the WL algorithm\cite{langfeld2012, pellegrini2014}.
+%The same schedule also applies to a variant of
+%the WL algorithm\cite{langfeld2012, pellegrini2014}.
 %
@@
 %
-Our method maps the optimization problem to a mechanical one,
+We will map the optimization problem to a mechanical one,
 in which the schedule plays the role of the velocity of
@@
 The latter, however,
-allow one to selectively flatten only
+allow one to partially flatten
 a few long-range histogram modes,
@@
 %
-Interestingly, several previous FDS algorithms\cite{
-  langfeld2012, pellegrini2014,
-  neuhaus2006, *neuhaus2007}
-can be cast as special cases of bandpass updating schemes.
+A previous FDS algorithm\cite{langfeld2012, pellegrini2014}
+can be cast as a special case of the bandpass updating scheme,
+and a similarly application can help other FDS
+algorithms\cite{neuhaus2006, *neuhaus2007, zhu2012}
+adjust the parameters of the bias potential automatically.
 
@@
 
-\subsubsection{\label{sec:FDS}
-Bias potential}
+%\subsubsection{\label{sec:FDS}
+%Bias potential}
 
@@
 %To do so, we introduce a bias potential, $u_i$,
-The challenge of sampling a multimodal distribution, $p_i$,
-with deep valley(s) can often be eased by
-biased sampling targeting a smoother distribution, $\rho_i$.
+The difficulty of sampling a multimodal distribution
+with deep valley(s) can sometimes be eased by
+sampling a smoother biased distribution, $\rho_i$.
 %
@@
 %%with the precision of the histogram as simulation lengthens.
-
-\subsubsection{Updating schemes}
-
+%
+%
+%\subsubsection{Updating schemes}
+%
+%
+%
 %We will mainly focus on the adaptive methods below.
 %
-In the original WL algorithm\cite{wang2001, *wang2001pre},
+In the WL algorithm\cite{wang2001, *wang2001pre},
 the bias potential, $u_i(t)$, is updated
@@
 In a more general \emph{multiple-bin scheme},
-including the Gaussian updating scheme
-used in metadynamics,
-several neighboring bins are updated:
+%including the Gaussian updating scheme used in metadynamics,
+the updating also involves neighboring bins:
 %
@@
 
-\subsubsection{Updating magnitude and the inverse-time formula}
+%\subsubsection{Updating magnitude and the inverse-time formula}
 
@@
 We define a bin-average-deducted bias potential
-as\footnote{Alternatively,
+as\note{Alternatively,
 we may define $v_{*i}(t)$ such that
@@
   .
-\label{eq:ortho1}
+\label{eq:ortho0}
 \end{equation}
@@
 Using this definition,
-we can rewrite Eq. \eqref{eq:ortho1} as
+we can rewrite Eq. \eqref{eq:ortho0} as
 $\mathcal F[1]_k = \delta_{k0}$,
@@
   ,
-%\notag
-\label{eq:error_asym_Lagrangian}
+  \notag
+  %\label{eq:error_asym_Lagrangian}
 \end{equation}
@@
   .
-  \label{eq:mass_func}
+  \notag
+  %\label{eq:mass_func}
 \end{equation}
@@
 does not explicitly depend on time, $t$,
-the Hamiltonian is conserved,
+the Hamiltonian is conserved,\cite{}
 which means that the asymptotic error grows
@@
 which is approximately true for the Gaussian updating scheme
-at large $\bar q$ (cf. Appendix \ref{sec:seplambda}),
+at large $\bar q$,
 then we would have an exponentially decaying schedule.
@@
 (cf. Appendix \ref{sec:more_wband}).
-
-In particular, for the Gaussian updating scheme,
+%
+For the Gaussian updating scheme,
 we have\cite{bussi2006}
@@
   \lambda_k
-  &=
-  \exp\left(
-        -\frac{ k^2 \, \sigma_\varphi^2 }
-              {           2             }
-      \right)
   =
@@
 %
-The updating magnitude should be sufficiently small
+\note{The updating magnitude should be sufficiently small
   such that the resulting error is no greater than
-multiple of $1.0$.
+  multiple of $1.0$.}
 
@@
   .
-  %\notag
-  \label{eq:Gamma_G_sum}
+  \notag
+  %\label{eq:Gamma_G_sum}
 \end{align}
@@
 
-%
-%
-%In Appendix \ref{sec:seplambda},
-%we will show that the opposite limit of
-%a set of well-separated eigenvalues
-%will provide an upper bound for
-%the asymptotic error, Eq. \eqref{eq:error_asym_ubound}.
-
 \subsubsection{\label{sec:optscheme}
@@
 by assuming
-that the PMF is sufficiently smooth
-such that the eigenmodes can be limited to $k < K$,
+that the PMF is so smooth
+that it can be represented by the first $K$ modes,
 %
@@
 \begin{equation}
-  \lambda_0 = \cdots = \lambda_{K-1} = 1,
+  \lambda_1 = \cdots = \lambda_{K-1} = 1,
   \mathrm{\; and \;}
@@
 %
-Then, the updating matrix is, according to
-Eq. \eqref{eq:w_from_phi},
-%
-$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj}.$$
+%Then, the updating matrix is, according to
+%Eq. \eqref{eq:w_from_phi},
+%%
+%$$w_{ij} = \frac{1}{\rho_i} \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj}.$$
+%%
+The asymptotic error is then
+$\Err_A(T) = \sum_{ k = 1 }^{ K - 1 } \Gamma_k T/(T + t_0)^2$,
+which is less than that from Eq. \eqref{eq:EA_sbin1}.
 %
@@
 remains to be the inverse-time schedule
-given by Eq. \eqref{eq:alpha_invt1},
-%
+given by Eq. \eqref{eq:alpha_invt1}
 with $t_0$ given by Eq. \eqref{eq:t0_sbin}.
@@
 
-Since the bandpass updating scheme is
-a straightforward generalization of
+Since bandpass updating schemes
+straightforward generalizations of
 the single-bin updating scheme,
@@
 However, the histogram fluctuation should
-be limited to first few modes,
+be limited to first $K$ modes,
 as the bandpass updating scheme
-only partially flattens the histogram
-of these modes (cf. Appendix \ref{sec:hfluc}).
-
-%
-%This error is roughly proportional to the inverse simulation length,
-%as one would expect from an equilibrium FDS simulation.
-%In Appendix \ref{sec:equilerr},
-%we show that the above error tends
-%to the error of the ideal equilibrium FDS simulation
-%in the asymptotic limit.
-
-%
-%For example, consider a system that is initially
-%prepared under a fixed $a_0$
-%at the beginning of the schedule $\alpha(t)$.
-%%
-%Then for shorter times, $T \approx 0$,
-%the error, dominated by the residual error
-%given by Eq. \eqref{eq:error_res1},
-%is clearly smaller if some eigenvalues
-%are less than $1.0$ (the single-bin case).
-%
-%In practice, real simulations
-%may have long achieved the desired precision
-%before reaching the asymptotic regime,
-%especially if the updating matrix
-%has some $\lambda_k$'s close to $0$.
+only partially flattens these histogram
+modes (cf. Appendix \ref{sec:hfluc}).
+
+In addition to the application to homogeneous updating schemes
+(Appendix \ref{sec:homo_bandpass}),
+bandpass updating schemes can also be applied to
+a continuous variable, $z$,\footnote{
+  In the transition to a continuous variable,
+  we will map
+  $\rho_i$ to $\rho(z) \delta z$,
+  $\phi_{ki}$ to $\phi_k(z) \delta z$,
+  $w_{ij}$ to $w(z, z') \delta z$, etc.}
+with the eigenmodes derived from orthogonal polynomials, $R_k(z)$,
+satisfying
+\begin{align*}
+  \int R_k(z) R_{k'}(z) \rho(z) \, d z = \delta_{k, k'}
+  .
+\end{align*}
+%
+With $\phi_k(z) = R_k(z) \, \rho(z)$,
+we can readily satisfy Eqs. \eqref{eq:eig_orthonormal_cols}-\eqref{eq:ortho0},
+and construct the updating matrix by
+Eqs. \eqref{eq:w_from_phi} and \eqref{eq:lambda_bandpass} as
+\begin{align*}
+  w(z, z')
+  =
+  \sum_{k=1}^{K-1} \frac{ \phi_k(z) \, \phi_k(z') } { \rho(z) }
+  = \sum_{k=1}^{K-1} R_k(z) \, R_k(z') \, \rho(z'),
+\end{align*}
+%
+where we have dropped the $k=0$ mode
+that corresponds to the uniform shift of the bias potential.
+%
+In this way, the bias potential is a superposition of the polynomials
+$$
+u(z, t) = \sum_{k=1}^{K-1} a_k(t) \, R_k(z),
+$$
+with the coefficients updated as, according to Eq. \eqref{eq:mbin_update},
+$$
+a_k(t+1) = a_k(t) + \alpha(t) \, R_k\bigl( z(t) \bigr).
+$$
+
+For example, by assuming the Legendre polynomials\cite{arfken}
+on a flat distribution over $[z_c - \Delta z/2, z_c + \Delta z/2]$
+and $K = 2$,
+the bias potential contains only a linear mode for
+$R_1(z) = 2 \sqrt{3} (z - z_c)/\Delta z$,
+and we recover the algorithm by
+Langfeld, Lucini and Rago\cite{langfeld2012}.
+%
+Applied to the energy space, $z = E$,
+$a_1(t)$ represents the local or
+statistical temperature\cite{kim2006, *kim2007},
+and the inverse-time schedule
+was recommended for its update\cite{pellegrini2014}.
+
+Alternatively, by assuming the Hermite polynomials\cite{arfken}
+on a Gaussian distribution centered at $z_c$ with width $\sigma_z$
+and $K=3$,
+we have a quadratic bias potential
+as a linear combination of
+$R_1(z) = (z-z_c)/\sigma_z$
+and
+$R_2(z) = (z - z_c)^2 /\left(\sqrt 2 \, \sigma_z^2\right)$.
+%
+This bias potential was commonly empolyed
+in umbrella sampling (see e.g. Ref. \cite{zhu2012})
+and the method of Gaussian ensemble\cite{hetherington1987,
+*challa1988, *costeniuc2006, neuhaus2006, *neuhaus2007},
+which is helpful in overcoming the barrier
+in a first-order phase transition.
+%
+Initially, we can employ the WL strategy
+to control the updating magnitude
+with the histogram fluctuation redefined as
+$$
+\max\left\{
+  \left|
+    \frac{ \overline z - z_c } { \sigma_z }
+  \right|
+  ,
+  \left|
+    \frac{ \overline{ (z - z_c)^2 } - \sigma_z^2 }
+    { \sqrt 2 \, \sigma_z^2 }
+  \right|
+  \right\}
+  ,
+$$
+where the overline denotes the time average.
 
@@
 
-We used the Metropolis algorithm\cite{frenkel, metropolis1953} for configuration sampling.
+We used the Metropolis algorithm\cite{frenkel, metropolis1953}
+for configuration sampling.
 %
@@
 In each trial, the displacement along each dimension is uniform,
-with the size given by $0.1/\varrho^*$ to ensure the acceptance ratio is around 50\%,
+with the size given by $0.1/\varrho^*$
+to ensure the acceptance ratio is around 50\%,
 where $\varrho^*$ is the reduced number density.
 
-The PMF was calculated from $r = 0$ to half of the box dimension
+The PMF was calculated from $r = 0$ to half of the box size
+(also the cutoff distance of the potential)
 on a grid of size $\delta r = 0.01$ in reduced units.
@@
 %
-Generally, the optimal schedule
-for the $\varrho^* = 0.8$ case
-was closer to the inverse-time schedule
-than that for the $\varrho^* = 0.1$ case.
-%
-This was probably because in the former case
-the first mode dominates the total error,
-$\Gamma_1 \approx \Gamma$,
-and the schedule was roughly optimized for this mode,
-resulting in
-$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
-as $\lambda_1 \approx 1$
-[cf. Eq. \eqref{eq:lambda_Gaussian}].
+%Generally, the optimal schedule
+%for the $\varrho^* = 0.8$ case
+%was closer to the inverse-time schedule
+%than that for the $\varrho^* = 0.1$ case.
+%%
+%This was probably because in the former case
+%the first mode dominates the total error,
+%$\Gamma_1 \approx \Gamma$,
+%and the schedule was roughly optimized for this mode,
+%resulting in
+%$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
+%as $\lambda_1 \approx 1$
+%[cf. Eq. \eqref{eq:lambda_Gaussian}].
 %
@@
 
-The optimality of the inverse-time schedule can be considered
-as a natural consequence of the stabilization of a runtime average
-as the sample size increases\cite{
-  marsili2006, barducci2008}.
+We wish to show that with the inverse-time schedule,
+the bias potential under the single-bin updating scheme
+happens to coincide with a runtime average
+of the independent corrections.
+%as a natural consequence of the stabilization of a runtime average
+%as the sample size increases\cite{
+%  marsili2006, barducci2008}.
 %
@@
 %
-Below we shall use capital letter to denote
-time average: for variable $x$, we define
+Below we shall use capital letters to denote
+time averages, e.g. for variable $x$, we define
 $X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)$.
@@
   \right].
-\label{eq:Gamma_local}
+  \notag
+  %\label{eq:Gamma_local}
 \end{align}
@@
 cf. Fig. \ref{fig:lj_xerr}.
-
-\section{\label{sec:seplambda}
-Updating scheme
-with a set of well-separated eigenvalues}
-
-We will study updating schemes
-with a set of well-separated eigenvalues,
-as a model for the large $k$ eigenmodes
-of the Gaussian updating scheme.
-%
-We will show that the mass distribution will have
-an inverse-$\bar q$ profile,
-and derive a upper bound for the asymptotic error.
-
-Starting from the definition, Eq. \eqref{eq:mass_func}, we have
-$$
-  M(\bar q)
-  \le
-  \sum_{k = 2}^n
-  \sqrt{ \Gamma_k } \, \lambda_k \, e^{ -\lambda_k \, \bar q }
-  .
-$$
-Thus,
-it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
-%
-\begin{equation}
-  C_M
-  =
-  \int_0^\infty
-  M(\bar q) \, d \bar q
-  \le
-  \sum_{k = 1}^{n-1} \sqrt{ \Gamma_k }
-  ,
-  \label{eq:mint_ubound}
-\end{equation}
-%
-and the asymptotic error from Eq. \eqref{eq:error_asym2}
-is bounded as
-\begin{equation}
-  \Err_A(T)
-  =
-  \frac{ C_M^2 } { T }
-  \le
-  \frac 1 T
-  \left(
-    \sum_{k=1}^{n-1} \sqrt{ \Gamma_k }
-  \right)^2
-  .
-  \label{eq:error_asym_ubound}
-\end{equation}
-%
-It is readily seen that this bound is always
-greater than the lower bound given by
-Eq. \eqref{eq:EA_sbin1} [cf. Eq. \eqref{eq:Gamma_G_sum}]:
-%
-$$
-\frac{T}{(T+t_0)^2} \, \sum_{k=1}^{n-1} \Gamma_k
-\le
-\frac{1}{T} \sum_{k=1}^{n-1} \Gamma_k
-\le
-\frac{1}{T} \left( \sum_{k=1}^{n-1} \sqrt{ \Gamma_k } \right)^2
-.
-$$
-
-Next, we will show that this bound represents
-updating schemes with a set of very different eigenvalues.
-%
-First, we have for any mode $k^*$ $(k^* > 0)$,
-%
-\begin{align}
-  M(\bar q)
-  &
-  \ge
-  \sqrt{ \Gamma_{ k^* } } \,
-  \lambda_{ k^* } \,
-  e^{ -\lambda_{ k^* } \, \bar q }
-  \ge
-  \frac{ \sqrt{ \Gamma_{ k^* } } } { \bar q }
-  ,
-\label{eq:MQ_lbound1}
-\end{align}
-%
-where the first inequality
-would approach an equality if the $k^*$ mode
-dominates the mass function, $M(\bar q)$,
-and the second inequality
-represent the limit of $\lambda_{k^*} \to 1/\bar q$.
-%
-Then, for the $k$th component of the asymptotic error,
-we have from Eq. \eqref{eq:error_asym_Lagrangian}
-%
-\begin{align}
-  \left\langle
-  \tilde v_k^2(T)
-  \right\rangle_A
-  &=
-  \int_0^T
-  \Gamma_k \, \lambda_k^2 \, e^{2 \, \lambda_k \, [q(t) - q(T)] }
-  \, {\dot q}^2(t) \, dt
-  \notag
-  \\
-  &=
-  \frac{ C_M } { T }
-  \, \Gamma_k \, \lambda_k^2
-  \int_0^{q(T)}
-  \frac{ e^{-2 \, \lambda_k \, \bar q } } { M(\bar q) }
-  \,  \, d \bar q
-  \notag
-  \\
-  &\le
-  \frac{ C_M }{ T } \,
-  \frac{ \Gamma_k \, \lambda_k^2 } { \sqrt{ \Gamma_{k^*} } \, \lambda_{k^*} }
-  \int_0^{q(T)}
-    e^{-(2 \, \lambda_k - \lambda_{k^*} ) \, \bar q}
-    \, d \bar q
-  \notag
-  \\
-  &\le
-  \frac{ C_M }{ T } \,
-  \cdot
-  \frac{ \Gamma_k }
-  { \sqrt{ \Gamma_{k^*} } }
-  \cdot
-  \frac{ \lambda_k^2 }
-  { \lambda_{k^*} \, (2 \, \lambda_k -\lambda_{k^*}) }
-  ,
-  \label{eq:y2_ubound_k}
-\end{align}
-%
-where we have
-used Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:MQ_lbound1}
-on the second and third lines, respectively,
-and extended the upper limit to infinity on the last line.
-%
-For a set of well-separated eigenvalues,
-$\lambda_k$ depends on the mode index, $k$,
-much more critically than $\Gamma_k$.
-%
-Thus, the tightest upper bound
-can be approximately obtained from minimizing the last factor
-on the right-hand side of Eq. \eqref{eq:y2_ubound_k},
-and the minimal is reached at $k^* = k$,
-when $\lambda_{k^*} = \lambda_k$.
-%
-So
-%
-\begin{equation}
-  \left\langle
-  \tilde v_k^2(T)
-  \right\rangle_A
-  \le
-  %\approx
-  \frac{ C_M }{ T }
-  \sqrt{ \Gamma_k }
-  .
-  \label{eq:y2_ubound}
-\end{equation}
-%
-By summing over $k$
-and using Eq. \eqref{eq:mint_ubound},
-we recover Eq. \eqref{eq:error_asym_ubound}.
-
-From Eq. \eqref{eq:Gamma_local},
-we know that $\Gamma_k$ would quickly approach a constant.
-%
-Then Eq. \eqref{eq:y2_ubound} shows that
-the asymptotic error profile would also
-approach a plateau in this case
-(cf. Fig. \ref{fig:lj_xerr}
-for a Gaussian example).
-%
-However, to reach this upper bound
-the simulation has to be long enough
-such that the upper limit of the integral
-can be extended from $q(T)$ to infinity
-in the last step of Eq. \eqref{eq:y2_ubound_k}.
-%
-As a result,
-the flatter profile from Eq. \eqref{eq:y2_ubound}
-would stop at mode $k$,
-at which $\lambda_k \approx 1/q(T)$.
-
-Finally, from Eqs. \eqref{eq:MQ_lbound1} and \eqref{eq:Gamma_local}
-we know that for a large $\bar q$, which implies a sufficiently
-large matching mode $k^*$ with $\lambda_{k^*} \approx 1/\bar q$,
-the mass function would be roughly proportional to $1/\bar q$,
-and so is the mass distribution, i.e.
-\begin{equation}
-  m(\bar q) \propto \frac{ 1} { \bar q },
-  \qquad
-  \mbox{for large $\bar q$.}
-  \label{eq:m_invq}
-\end{equation}
 
@@
   \notag
-  %\label{eq:w_band_refl}
+  \label{eq:w_band_refl}
 \end{equation}
@@
 
+\subsection{\label{sec:invert_wband}
+Inversion}
+
+Since the eigenvalues are related to the updating kernel
+by a cosine transform,
+we can invert the relation as
+%
+\begin{equation}
+  \mu_i
+  =
+  \frac { 1 } { g \, n }
+  \sum_{ k = 0 }^{ g \, n - 1 }
+  \lambda_k
+  \cos \left(
+       \frac{ k \, i \, 2 \, \pi }
+            {      g \, n        }
+  \right)
+  .
+\label{eq:mu_from_lambda}
+\end{equation}
+%
+In the non-periodic case,
+we have defined
+$\lambda_k \equiv \lambda_{2 \, n - k}$
+for $k = n + 1, \dots, 2 \, n - 1$,
+as well as
+%
+\begin{align}
+  \lambda_n
+  =
+  (-1)^{ n - 1 }
+  \lambda_0
+  +
+  2 \, \sum_{ k = 1 }^{ n - 1 }
+      (-1)^{n - k - 1} \, \lambda_k
+  ,
+\label{eq:lambdan}
+\end{align}
+to satisfy the constraint, $\mu_n = 0$.
+%
+
 \note{On the inversion.
@@
   This formula for $\lambda_k$
-  can be readily extended to $k = 1 \, n - 1$,
+  can be readily extended to $k = 2 \, n - 1$,
   and we have
@@
 
-\subsection{\label{sec:invert_wband}
-Inversion}
-
-Since the eigenvalues are related to the updating kernel
-by a cosine transform,
-we can invert the relation as
-%
-\begin{equation}
-  \mu_i
-  =
-  \frac { 1 } { g \, n }
-  \sum_{ k = 0 }^{ g \, n - 1 }
-  \lambda_k
-  \cos \left(
-       \frac{ k \, i \, 2 \, \pi }
-            {      g \, n        }
-  \right)
-  .
-\label{eq:mu_from_lambda}
-\end{equation}
-%
-In the non-periodic case,
-we have defined
-$\lambda_k \equiv \lambda_{2 \, n - k}$
-for $k = n + 1, \dots, 2 \, n - 1$,
-as well as
-%
-\begin{align}
-  \lambda_n
-  =
-  (-1)^{ n - 1 }
-  \lambda_0
-  +
-  2 \, \sum_{ k = 1 }^{ n - 1 }
-      (-1)^{n - k - 1} \, \lambda_k
-  ,
-\label{eq:lambdan}
-\end{align}
-to satisfy the constraint, $\mu_n = 0$.
-%
-
 \subsection{Gaussian updating scheme}
@@
 the eigenvalues are either $0$ or $1$ and
-we can use Eq. \eqref{eq:mu_from_lambda}
+it is convenient to set $\lambda_0 = 1$.
+%
+We can use Eq. \eqref{eq:mu_from_lambda}
 to find the updating kernel.
 %
-For a periodic variable,
-to take into account the two-fold degeneracy,
+To take into account the two-fold degeneracy
+for a periodic variable,
 we will modify the mode cutoff such that
@@
 %
-Then by using
-Eq. \eqref{eq:mu_from_lambda},
-we get
+Then, $w_{ij} = \mu_{i-j}$, and
 \begin{equation}
@@
   }
-  ,
+  .
 \notag
@@
 \end{equation}
-and $w_{ij} = \mu_{i-j}$.
 %
@@
 we have $\lambda_n = (-1)^{n-K}$
-from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan}.
-Then, from Eq. \eqref{eq:mu_from_lambda},
-we have
+from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan},
+and
 \begin{equation}
@@
     }
-  .
+  ,
   \notag
@@
 %
-Although the kernel is a sawtooth
-because of the term, $(-1)^{n-K+l}$,
-the ruggedness disappears
-in the updating matrix
-as the above oscillatory term for $l = i - j$
-is canceled by the term for either $l = i + j - 1$
-or $l = i + j - 2 \, n - 1$
-with the opposite sign.
+which is free from the oscillatory term, $(-1)^{n-K+1}$,
+that appeared in the kernel
+[its contribution to $\mu_{i-j}$ in Eq. \eqref{eq:w_band_refl}
+is canceled by its contribution to either $\mu_{i+j-1}$
+or $\mu_{i+j-2n-1}$,
+whichever exists, of the opposite sign].
+%Although the kernel is a sawtooth
+%because of the term, $(-1)^{n-K+l}$,
+%the ruggedness disappears
+%in the updating matrix
+%as the above oscillatory term for $l = i - j$
+%is canceled by the term for either $l = i + j - 1$
+%or $l = i + j - 2 \, n - 1$
+%with the opposite sign.
 %

### 2017-06-20 doc/invt.tex
invt.tex: small edits

@@
 %
-In this way, the bias potential is a superposition of the polynomials
+This results in a polynomial bias potential
+that can be written as a superposition
 $$
-u(z, t) = \sum_{k=1}^{K-1} a_k(t) \, R_k(z),
+u(z, t) = \sum_{k=1}^{K-1} c_k(t) \, R_k(z),
 $$
@@
 $$
-a_k(t+1) = a_k(t) + \alpha(t) \, R_k\bigl( z(t) \bigr).
+c_k(t+1) = c_k(t) + \alpha(t) \, R_k\bigl( z(t) \bigr).
 $$
 
-For example, by assuming the Legendre polynomials\cite{arfken}
-on a flat distribution over $[z_c - \Delta z/2, z_c + \Delta z/2]$
-and $K = 2$,
+For example,
+to sample a flat distribution, $\rho(z)$,
+over $[z_c - \Delta z/2, z_c + \Delta z/2]$,
+we can use the Legendre polynomials\cite{arfken},
+and with $K = 2$,
 the bias potential contains only a linear mode for
 $R_1(z) = 2 \sqrt{3} (z - z_c)/\Delta z$,
-and we recover the algorithm by
+which recovers the algorithm by
 Langfeld, Lucini and Rago\cite{langfeld2012}.
@@
 Applied to the energy space, $z = E$,
-$a_1(t)$ represents the local or
+$c_1(t)$ represents the scaled local or
 statistical temperature\cite{kim2006, *kim2007},
@@
 
-Alternatively, by assuming the Hermite polynomials\cite{arfken}
-on a Gaussian distribution centered at $z_c$ with width $\sigma_z$
-and $K=3$,
+Alternatively,
+to sample a Gaussian distribution
+centered at $z_c$ with width $\sigma_z$
+we can use the Hermite polynomials\cite{arfken},
+and with $K=3$,
 we have a quadratic bias potential
-as a linear combination of
-$R_1(z) = (z-z_c)/\sigma_z$
+from a linear combination of
+$R_1(z) = (z - z_c)/\sigma_z$
 and
@@
 *challa1988, *costeniuc2006, neuhaus2006, *neuhaus2007},
-which is helpful in overcoming the barrier
-in a first-order phase transition.
-%
-Initially, we can employ the WL strategy
-to control the updating magnitude
+which is helpful in overcoming barriers
+in first-order phase transitions.
+%
+While the asymptotic updating magntiude
+would be given by the inverse-time formula,
+we can employ the WL strategy early in simulation
+to control the updating magnitude in stages
 with the histogram fluctuation redefined as
@@
 where the overline denotes the time average.
+
+With only two nontrivial modes,
+the above strategy usually only works
+for a Gaussian with a narrow width.
+%
+We can, however, use it as a building block
+to construct an expanded ensemble with
+multiple Gaussians centered at different places.
+%
+Such a construction can be realized via
+either the replica exchange
+(parallel tempering)\cite{swendsen1986,
+*geyer1991, *hukushima1996, *hansmann1997}
+or the simulated tempering\cite{marinari1992,
+  *lyubartsev1992, li2007,
+  park2007, *nguyen2013, *zhang2015st}
+framework.
+While the mutual exchange probabiliy
+can be readily derived in the former
+framework\cite{neuhaus2006, *neuhaus2007},
+the transition probability in the latter case is
+requires estimation of the partition function
+of the Gaussian ensemble
+$Z \equiv \int e^{-\beta \, E(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$,
+where $\mathbf x$ denotes the molecular coordinates,
+$E(\mathbf x)$ is the unbiased energy function,
+$\beta$ is the inverse temperature,
+and the collective variable, $z$,
+has also been written as a function of $\mathbf x$.
+%
+The transition probabiliy from
+the $i$th Gaussian to the $j$th is given by
+is given by
+\begin{align*}
+A(i\to j) =
+\min\left\{1,
+  \frac{Z^{(i)}}{Z^{j}}
+  \exp\left[u^{(i)}\bigl( z(t) \bigr)
+   - u^{(j)}\bigl( z(t) \bigr) \right]
+  \right\}
+  ,
+\end{align*}
+where $z(t)$ is the current value of $z$.
+%
+For a neighboring pair, $j = i\pm 1$,
+we can estimate the ratio $Z^{(i)}/Z^{(j)}$
+from the following approximate formula
+%
+\begin{align*}
+  \frac{Z_i}{Z_j}
+  =
+  \frac{\sigma_i}{\sigma_j}
+  \exp\left[
+    u^{(j)}\bigl( z^{(i,j)} \bigr)
+   -u^{(i)}\bigl( z^{(i,j)} \bigr)
+  \right]
+  ,
+\end{align*}
+%
+where
+$z^{(i,j)} \equiv
+\bigl( z_c^{(i)}/\sigma_z^{(i)} + z_c^{(j)}/\sigma_z^{(j)} \bigr) /
+\bigl(         1/\sigma_z^{(i)} +         1/\sigma_z^{(j)} \bigr)$.
 
@@
 
-We wish to show that with the inverse-time schedule,
-the bias potential under the single-bin updating scheme
-happens to coincide with a runtime average
+We wish to show the optimality of the inverse-time schedule
+for the single-bin updating scheme
+from the observation that
+it leads to a bias potential
+that coincides with a runtime average
 of the independent corrections.

### 2017-06-20 prog/aus/gaus.h

@@ -392,6 +392,7 @@ __inline static int gaus_wlcheckx(gaus_t *gaus,
 
 
 
+/* transition to a neighboring umbrella */
 __inline static int gaus_move(gaus_t *gaus, double x, int *id)
 {
   int acc = 0;

### 2017-06-21 doc/invt.tex
add update.sh for collecting LJ data

@@
 
-\subsubsection{\label{sec:optscheme}
+\subsubsection{\label{sec:bandpass}
 Bandpass updating schemes}
@@
 with the eigenmodes derived from orthogonal polynomials, $R_k(z)$,
-satisfying
+that satisfy
 \begin{align*}
-  \int R_k(z) R_{k'}(z) \rho(z) \, d z = \delta_{k, k'}
+  \int R_k(z) \, R_{k'}(z) \, \rho(z) \, d z = \delta_{k, k'}
   .
@@
 the bias potential contains only a linear mode for
-$R_1(z) = 2 \sqrt{3} (z - z_c)/\Delta z$,
+$R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$,
 which recovers the algorithm by
@@
 Applied to the energy space, $z = E$,
-$c_1(t)$ represents the scaled local or
+$c_1(t)/\sigma_z$ represents the scaled local or
 statistical temperature\cite{kim2006, *kim2007},
@@
 and
-$R_2(z) = (z - z_c)^2 /\left(\sqrt 2 \, \sigma_z^2\right)$.
+$R_2(z) = \bigl[(z - z_c)^2 - \sigma_z^2\bigr] /\left(\sqrt 2 \, \sigma_z^2\right)$.
 %
@@
 to control the updating magnitude in stages
-with the histogram fluctuation redefined as
-$$
-\max\left\{
-  \left|
-    \frac{ \overline z - z_c } { \sigma_z }
-  \right|
-  ,
-  \left|
-    \frac{ \overline{ (z - z_c)^2 } - \sigma_z^2 }
-    { \sqrt 2 \, \sigma_z^2 }
-  \right|
-  \right\}
-  ,
-$$
-where the overline denotes the time average.
-
-With only two nontrivial modes,
-the above strategy usually only works
-for a Gaussian with a narrow width.
-%
-We can, however, use it as a building block
-to construct an expanded ensemble with
-multiple Gaussians centered at different places.
-%
-Such a construction can be realized via
+with the histogram fluctuation redefined by
+Eq. \eqref{eq:maxFk}.
+%
+This method will be referred to as
+the adaptive Gaussian ensemble method below.
+
+The use of polynomial basis makes
+the bias potential smooth and differentiable,
+and hence suitable for MD applications.\cite{zhu2012}
+%
+However, since we only used two modes
+for controlling the mean and variance,
+the method only works
+for narrow distributions.
+%
+To sample a wider range,
+we can construct an expanded ensemble with
+multiple Gaussian distributions
+centered at different places via
 either the replica exchange
-(parallel tempering)\cite{swendsen1986,
+(or parallel tempering)\cite{swendsen1986,
 *geyer1991, *hukushima1996, *hansmann1997}
@@
 framework.
-While the mutual exchange probabiliy
+%
+While the mutual exchange probability
 can be readily derived in the former
 framework\cite{neuhaus2006, *neuhaus2007},
-the transition probability in the latter case is
+the transition probability in the latter case
 requires estimation of the partition function
 of the Gaussian ensemble
-$Z \equiv \int e^{-\beta \, E(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$,
+$Z \equiv \int e^{-\beta \, \mathcal U(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$,
 where $\mathbf x$ denotes the molecular coordinates,
-$E(\mathbf x)$ is the unbiased energy function,
+$\mathcal U(\mathbf x)$ is the unbiased potential energy function,
 $\beta$ is the inverse temperature,
@@
 %
-The transition probabiliy from
+The transition probability from
 the $i$th Gaussian to the $j$th is given by
@@
 \min\left\{1,
-  \frac{Z^{(i)}}{Z^{j}}
+  \frac{Z^{(i)}}{Z^{(j)}}
   \exp\left[u^{(i)}\bigl( z(t) \bigr)
@@
 For a neighboring pair, $j = i\pm 1$,
-we can estimate the ratio $Z^{(i)}/Z^{(j)}$
+we can estimate the ratio $Z^{(j)}/Z^{(i)}$
 from the following approximate formula
@@
 \begin{align*}
-  \frac{Z_i}{Z_j}
-  =
-  \frac{\sigma_i}{\sigma_j}
-  \exp\left[
-    u^{(j)}\bigl( z^{(i,j)} \bigr)
-   -u^{(i)}\bigl( z^{(i,j)} \bigr)
-  \right]
+  \ln\frac{ Z^{(j)} }{ Z^{(i)} }
+  \approx
+  \ln\frac{ \sigma_j }{ \sigma_i }
+   -u^{(j)}\bigl( z^{(i,j)} \bigr)
+   +u^{(i)}\bigl( z^{(i,j)} \bigr)
   ,
@@
 \bigl(         1/\sigma_z^{(i)} +         1/\sigma_z^{(j)} \bigr)$.
+\note{For implementation, we can further simplify
+\begin{align*}
+  &-u^{(j)}\bigl( z^{(i,j)} \bigr)
+   +u^{(i)}\bigl( z^{(i,j)} \bigr)
+   \\
+   &=
+   (c_1^{(i)} + c_1^{(j)}) \, \Delta z^*
+   +
+   (c_2^{(i)} - c_2^{(j)}) \, \frac{{\Delta z^*}^2 - 1} {\sqrt 2}
+   ,
+\end{align*}
+where $\Delta z^* = [z_c^{(j)} - z_c^{(i)}]/[\sigma_z^{(j)} + \sigma_z^{(i)}]$.
+}
 
@@
 
-%\subsection{\label{sec:lj}Lennard-Jones}
+\subsection{\label{sec:lj}
+Comparison of updating schemes on a Lennard-Jones fluid}
 
@@
 the starting magnitude was around
-the ideal value of $a_0/2 = 5\times10^{-5}$,
-as given by Eq. \eqref{eq:half_alpha0}.
+the ideal value of $a_0/2 = 5\times10^{-5}$
+given by Eq. \eqref{eq:half_alpha0}.
 %
@@
 In this case, the schedule predicted by our method
-would be inaccurate because the resulting deviation in the bias potential
+would be unusable because the resulting deviation in the bias potential
 was no longer small,
@@
 \end{figure}
+
+\subsection{\label{sec:potts}
+Adaptive Gaussian ensemble}
+
+As a proof of principle,
+we tested the adaptive Gaussian ensemble method
+introduced in Sec. \ref{sec:bandpass}
+on the two-dimensional $L\times L$ $10$-state
+Potts model\cite{wu1982, newman, wang2001, wang2001pre}
+along the energy space.
+%
+We used $L = 32$ here.
+%
+We used $26$ Gaussian distributions of width $L$,
+and spread them over $[-1.7L^2, -0.9L^2]$.
+%
+For configuration sampling,
+we used the Wolff algorithm\cite{wolff1989, newman}
+at temperature $c_1/\sigma_z$
+followed by a Metropolis acceptance step
+to correct for the quadratic potential, $R_2$.
+%
+The acceptance ratio for configuration moves
+was no less than 90\%.
 
@@
 from the largest mode
-$$
-\max_{k=1, \dots, K-1} \left| \, \tilde F_k(T) \, \right|
-.
-$$
-For a local sampling process,
-we expect that the two definition would behave similarly.
+%$\max_{k=1, \dots, K-1} \left| \, \tilde F_k(T) \, \right|$.
+\begin{align}
+  \max_{k=1, \dots, K-1} \left| \, \tilde F_k(T) \, \right|.
+  \label{eq:maxFk}
+\end{align}
+This definition can be readily extended
+to the continuous variable case, with
+$\tilde F_k(T) = \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)$.
+%For a local sampling process,
+%we expect that the two definition would behave similarly.
 %

### 2017-06-21 doc/invt.tex
add pt_hist.gp

@@
 This method will be referred to as
-the adaptive Gaussian ensemble method below.
+the adaptive Gaussian umbrella sampling below.
 
@@
 \bigl(         1/\sigma_z^{(i)} +         1/\sigma_z^{(j)} \bigr)$.
+%
+In the $1/t$ scheme,
+%
 \note{For implementation, we can further simplify
@@
 where $\Delta z^* = [z_c^{(j)} - z_c^{(i)}]/[\sigma_z^{(j)} + \sigma_z^{(i)}]$.
+%
+However, the $c_2$ term does not appear to improve histogram flatness. Ideas?
 }
@@
 \subsection{\label{sec:potts}
-Adaptive Gaussian ensemble}
+Adaptive Gaussian umbrella sampling}
 
 As a proof of principle,
-we tested the adaptive Gaussian ensemble method
+we tested the method of adaptive Gaussian umbrella sampling
 introduced in Sec. \ref{sec:bandpass}
-on the two-dimensional $L\times L$ $10$-state
+on the two-dimensional $L\times L$ ten-state
 Potts model\cite{wu1982, newman, wang2001, wang2001pre}
@@
 %
-We used $L = 32$ here.
-%
-We used $26$ Gaussian distributions of width $L$,
-and spread them over $[-1.7L^2, -0.9L^2]$.
+We used $L = 32$ and $26$ Gaussian distributions of
+the same width $\sigma_E = L$ over $[-1.7L^2, -0.9L^2]$
+(the spacing between neighboring Gaussian centers and the width
+were roughly the same).
 %
 For configuration sampling,
-we used the Wolff algorithm\cite{wolff1989, newman}
-at temperature $c_1/\sigma_z$
+although one can use the Metropolis algorithm,
+we found it more efficient to use
+the Wolff algorithm\cite{wolff1989, newman}
+at temperature $c_1/\sigma_E$
 followed by a Metropolis acceptance step
-to correct for the quadratic potential, $R_2$.
-%
-The acceptance ratio for configuration moves
-was no less than 90\%.
+to correct for the quadratic potential, $R_2$
+(the acceptance ratio was no less than 92\%).
+%
+The results after $10^5 L^2$ steps are shown in
+Fig. \ref{fig:pt_hist}.
+%
+The Gaussian histograms were evenly spaced with similar width as intended,
+and the overall energy histogram was roughly flat,
+as shown in \ref{fig:pt_hist}(a).
+%
+The parameter $c_1/\sigma_E$ roughly represents the local temperature,
+and it demonstrated
+the backbending behavior\cite{kim2006, *kim2007}
+typical in a first-order phase transitions,
+as shown in Fig. \ref{fig:pt_hist}(b).
+%
+The parameter $c_2$ represents the amount of ``pressure''
+to restrain the distribution width,
+and its value was maximal around the center critical region,
+as shown in Fig. \ref{fig:pt_hist}(c).
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/pt_hist.pdf}
+  }
+  \caption{
+    \label{fig:pt_hist}
+    Application of the adaptive Gaussian umbrella sampling
+    on the $32 \times 32$ ten state Potts model.
+    %
+    (a) Normalized histograms (solid lines) from adaptive Gaussian umbrella sampling
+    matched the intended distributions.
+    The overall histogram (dashed line) was roughly flat.
+    %
+    (b) The parameter, $c_1/\sigma_E$, roughly corresponding to
+    the statistical temperature, $\beta(E)$,
+    demonstrated the backbending behavior.
+    %
+    (c) The parameter $c_2$ for restraining the variance
+    was maximal around the critical region.
+  }
+\end{center}
+\end{figure}
 
@@
 }%
-
-%\begin{figure}[h]
-%\begin{center}
-%  \makebox[\linewidth][c]{
-%    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
-%  }
-%  \caption{
-%    \label{fig:sinc}
-%    Updating kernels of the bandpass updating schemes
-%    compared to a Gaussian with a similar width.
-%    %
-%    The lines are a guide to the eyes.
-%    Inset. Updating matrix in the non-periodic case.
-%    %
-%    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
-%      The data were computed by the program \texttt{prog/predict},
-%      see particularly the function \texttt{intq\_save()}
-%      in \texttt{prog/intq.h} for panel (a),
-%      and the functions \texttt{savewin()}
-%      and \texttt{savewinmat()}
-%      in \texttt{prog/invt.h} for panel (b),
-%      The data were saved under the directory \texttt{data/sinc}.
-%      See the \texttt{make} object \texttt{sincsingle}
-%      in the \texttt{Makefile} there.
-%    }%
-%  }
-%\end{center}
-%\end{figure}
-
-%In Fig. \ref{fig:sinc},
-%we give examples of the updating kernels
-%of the bandpass updating schemes,
-%compared to a Gaussian
-%with a similar width, $\sigma$,
-%%
-%\begin{equation}
-%  \sigma
-%  \approx
-%  \frac
-%  {
-%    g \, n
-%  }
-%  {
-%    \sqrt{ 2 \, \pi } \, 2 \, K
-%  }
-%  .
-%\label{eq:sigma_equiv}
-%\end{equation}
-%%
-%%where $g = 1$ for the periodic case
-%%or $g = 2$ otherwise.
-%%
 

### 2017-06-22 doc/invt.tex
invt.tex: small edits

@@
   .
-  \notag
-  %\label{eq:mass_func}
+  %\notag
+  \label{eq:mass_func}
 \end{equation}
@@
 which is approximately true for the Gaussian updating scheme
-at large $\bar q$,
+at large $\bar q$,\footnote{If
+  the magntidue of the eigenvalues, $\lambda_k$'s,
+  span a wide range, then for a large $\bar q$,
+  the sum in Eq. \eqref{eq:mass_func} will be dominated by
+  the largest term, where $\lambda_k \approx 1/\bar q$,
+  and $M(\bar q) \approx \sqrt{\Gamma_k}/\bar q$.
+  }
+  %
 then we would have an exponentially decaying schedule.
@@
 that the PMF is so smooth
-that it can be represented by the first $K$ modes,
-%
+that it can be represented by the first $K$
+modes,\footnote{As an estimate of error,
+  we note that truncating a linear or quadratic function
+  at mode $K$ (for a non-periodic variable)
+  results in an error roughly proportional to
+  $|\max_i u_i - \min_i u_i|^2/K^3$.}
 we demand
@@
 %for the single-bin updating scheme.
+%
 
@@
 $R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$,
-which recovers the algorithm by
-Langfeld, Lucini and Rago\cite{langfeld2012}.
+which recovers the Langfeld, Lucini and Rago (LLR)
+algorithm\cite{langfeld2012}.
 %
@@
 we can use the Hermite polynomials\cite{arfken},
-and with $K=3$,
-we have a quadratic bias potential
-from a linear combination of
+and with $K=3$.
+%
+This results in a quadratic bias potential
+that can be written as a linear combination of
 $R_1(z) = (z - z_c)/\sigma_z$
@@
 %
-This bias potential was commonly empolyed
+This form of bias potential was commonly employed
 in umbrella sampling (see e.g. Ref. \cite{zhu2012})
@@
 %
-While the asymptotic updating magntiude
+While the asymptotic updating magnitude
 would be given by the inverse-time formula,
@@
 %
-This method will be referred to as
-the adaptive Gaussian umbrella sampling below.
+We will refer to this method as
+adaptive Gaussian umbrella sampling below.
 
@@
 for controlling the mean and variance,
-the method only works
-for narrow distributions.
+the method can only be applied
+to a narrow distribution.
 %
@@
 we can construct an expanded ensemble with
-multiple Gaussian distributions
+multiple Gaussian umbrellas
 centered at different places via
@@
 As a proof of principle,
-we tested the method of adaptive Gaussian umbrella sampling
+we tested the adaptive Gaussian umbrella sampling method
 introduced in Sec. \ref{sec:bandpass}
@@
 Potts model\cite{wu1982, newman, wang2001, wang2001pre}
-along the energy space.
-%
-We used $L = 32$ and $26$ Gaussian distributions of
+in the energy space.
+%
+We spread the Gaussian umbrellas
 the same width $\sigma_E = L$ over $[-1.7L^2, -0.9L^2]$
-(the spacing between neighboring Gaussian centers and the width
-were roughly the same).
-%
-For configuration sampling,
-although one can use the Metropolis algorithm,
-we found it more efficient to use
-the Wolff algorithm\cite{wolff1989, newman}
+with a spacing of $2L$.
+%
+For configuration sampling for each umbrella,
+we used the Wolff algorithm\cite{wolff1989, newman}
+(which appeared to be more efficient than the Metropolis algorithm),
 at temperature $c_1/\sigma_E$
-followed by a Metropolis acceptance step
-to correct for the quadratic potential, $R_2$
-(the acceptance ratio was no less than 92\%).
-%
-The results after $10^5 L^2$ steps are shown in
-Fig. \ref{fig:pt_hist}.
+followed by a Metropolis step with
+acceptance probability
+$\min\bigl\{1, e^{c_2 [R_2(E) - R_2(E')]} \bigr\}$,
+whose average was no less than 92\%.
+%
+To ensure stability,
+we avoid transitions among umbrellas
+until the current umbrella has
+completed the WL stage-wise way of
+reducing the updating magnitude
+and started the inverse-time schedule.\cite{belardinelli2007,
+*belardinelli2007jcp, *belardinelli2008, *belardinelli2016}
+
+The results for the $L = 64$ case
+after $10^5 L^2$ steps are shown
+in Fig. \ref{fig:pt_hist}.
 %
@@
 and the overall energy histogram was roughly flat,
-as shown in \ref{fig:pt_hist}(a).
+as shown in Fig. \ref{fig:pt_hist}(a).
 %
@@
     \label{fig:pt_hist}
-    Application of the adaptive Gaussian umbrella sampling
-    on the $32 \times 32$ ten state Potts model.
+    Application of adaptive Gaussian umbrella sampling
+    on the $64 \times 64$ ten state Potts model.
     %
@@
 %
-More generally,
-each adaptive FDS method is associated with
-an updating scheme
-that can be mathematically characterized
-by an updating matrix.
-%
-Each column of the matrix gives the normalized
-window function for the bin corresponding to the column.
+We can generally characterize
+an updating scheme with an updating matrix,
+with each column being the normalized
+window function for the bin of the column.
 %
@@
 the eigenvalues of the updating matrix,
-and the integrals of the autocorrelation functions
+which represent the relative rates of error reduction,
+and the integrals of the autocorrelation functions,
+which represent the intrinsic histogram fluctuation,
 of the eigenmodes.
-%
-The eigenvalues represent
-the relative rate of error reduction,
-and the autocorrelation integrals
-represent the intrinsic fluctuation
-of the histograms modes.
 %
@@
 including the Gaussian one,
-the errors were reduced at different paces in various eigenmodes,
-and the optimal schedule,
-representing a compromise of the eigenmodes,
+the optimal schedule, having to balance
+the different rates of error reduction of the eigenmodes,
 is usually not given by a simple closed form,
@@
 and other parameters.
-%
-%The initial updating magnitude is always
-%half of the previous equilibrium value.
-% the value used during equilibration.
+
+In particular, the Gaussian updating scheme is often
+equipped with a set of rapidly diminishing eigenvalues.
+%
+While this feature helps suppressing
+the accumulation of random error and
+maintaining the smoothness of the bias potential,
+it also slows down the reduction
+of the initial bias, if any,
+in the middle-range eigenmodes.
+%
+Thus, one needs to carefully choose the width of the Gaussian
+and the optimal schedule of decreasing the updating magnitude,
+which sometimes needs to be fixed at a maximal acceptable level
+for a long period before any reduction can happen.
+%
+The possible efficiency loss in
+error reduction of the long- and middle-range modes
+can often be compensated
+by the histogram-based correction formula, Eq. \eqref{eq:uhatav},
+for post-simulation analysis,
+which ideally would remove the remaining bias
+and recover a bias potential with precision
+similar to that from the single-bin case.
 
@@
 without a priori assumptions
-about the smoothness of the PMF.
-%
-However, if some eigenmodes
-correspond to local noise
-that is expected to be absent from the PMF,
-we can construct a
-class of bandpass updating schemes
-that are also optimal for asymptotic convergence.
-%
-The optimal schedule of these bandpass schemes
+on the smoothness of the PMF.
+%
+This scheme can be generalized to
+a class of bandpass updating schemes,
+which are also asymptotically optimal,
+but completely filter the components
+in the noise modes.
+%
+The optimal schedule of the bandpass schemes
 is always given by an inverse-time formula.
-
-The Gaussian updating scheme
-has the appeal of being able to maintain the smoothness
-of the bias potential,
-which is important for stable MD applications.
-%
-However, it requires a careful choice of the width and
-the updating magnitude should often be fixed at constant level
-for a sufficiently long period before
-a magnitude-decreasing schedule could start.
-%
-For reasonable parameters,
-one may use the histogram correction,
-Eq. \eqref{eq:uhatav},
-in a post-simulation analysis
-to regain the possible accuracy loss
-and obtain a bias potential similar to
-that from the optimized single-bin scheme.
-%%
-%However, the performance gain
-%is relatively small for short simulations
-%with a local sampling process such as MD.
+%
+If the bandpass updating scheme was constructed
+from orthogonal polynomials,
+the updating scheme became a generalization
+of the LLR algorithm\cite{langfeld2012, *pellegrini2014},
+and it could be used for parameter adjustment
+in umbrella sampling with quadratic
+bias potentials\cite{neuhaus2006, *neuhaus2007, zhu2012}.
 

### 2017-06-23 doc/invt.tex
small changes

@@
 along $z$\cite{mezei1987, berg1992, *lee1993,
-wang2001, *wang2001pre,
+wang2001, wang2001pre,
 huber1994,
@@
 %the target distribution along $z$\cite{mezei1987, berg1992, *lee1993,
-%wang2001, *wang2001pre,
+%wang2001, wang2001pre,
 %huber1994,
@@
 including the Wang-Landau (WL) algorithm\cite{
-  wang2001, *wang2001pre}
+  wang2001, wang2001pre}
 and metadynamics\cite{huber1994,
@@
 Many methods\cite{mezei1987, berg1992, *lee1993,
-wang2001, *wang2001pre, huber1994,
+wang2001, wang2001pre, huber1994,
 *laio2002, *laio2008, *barducci2011, *sutto2012}
@@
 An other class of adaptive FDS methods,
-including the WL algorithm\cite{wang2001, *wang2001pre}
+including the WL algorithm\cite{wang2001, wang2001pre}
 and metadynamics\cite{huber1994, laio2002, *laio2008, *barducci2011, *sutto2012},
@@
 %
-In the WL algorithm\cite{wang2001, *wang2001pre},
+In the WL algorithm\cite{wang2001, wang2001pre},
 the bias potential, $u_i(t)$, is updated
@@
 
-In the original WL papers\cite{
-wang2001, *wang2001pre},
+In the original WL algorithm\cite{
+wang2001, wang2001pre},
 the updating magnitude, $\alpha(t)$,
@@
 with $\alpha(t)$ reduced by a factor of $1/2$\cite{
-wang2001, *wang2001pre}.
+wang2001, wang2001pre}.
 %
@@
 %
-$$
+\begin{equation}
   {\tilde v}_k(t + 1) =
@@
   ,
-$$
+  \label{eq:vkupdate}
+\end{equation}
+%
 or in continuous time, we have, for $k > 0$,
@@
 in the original WL algorithm\cite{
-wang2001, *wang2001pre}.
+wang2001, wang2001pre}.
 }
@@
 We then compute the updating magnitude by numerical differentiation:
-  $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$
+  $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$\footnote{This
+    discretization follows from runtime average interpretation
+    discussed in Appendix \ref{sec:equilerr}.}
   and
@@
 %
-For these updating schemes,
-  the preliminary run with a constant updating magntiude
-  can be replaced by the WL-style stagewise reduction
-  until the updating magnitude falls under
-  the value given the inverse-time prescription\cite{
-    belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
+%For these updating schemes,
+%  the preliminary run with a constant updating magntiude
+%  can be replaced by the WL-style stagewise reduction
+%  until the updating magnitude falls under
+%  the value given the inverse-time prescription\cite{
+%    belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
 
@@
 where we have dropped the $k=0$ mode
-that corresponds to the uniform shift of the bias potential.
-%
-This results in a polynomial bias potential
-that can be written as a superposition
-$$
-u(z, t) = \sum_{k=1}^{K-1} c_k(t) \, R_k(z),
-$$
+that corresponds to a uniform shift
+of the bias potential.
+%
+The resulting bias potential
+assumes a polynomial form
+and can be written as a superposition
+$u(z, t) = \sum_{k=1}^{K-1} c_k(t) \, R_k(z)$,
 with the coefficients updated as, according to Eq. \eqref{eq:mbin_update},
-$$
-c_k(t+1) = c_k(t) + \alpha(t) \, R_k\bigl( z(t) \bigr).
-$$
+\begin{equation}
+  c_k(t+1) = c_k(t) + \alpha(t) \, R_k\bigl( z(t) \bigr)
+  .
+  \label{eq:ckupdate}
+\end{equation}
 
@@
 $R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$,
-which recovers the Langfeld, Lucini and Rago (LLR)
-algorithm\cite{langfeld2012}.
-%
-Applied to the energy space, $z = E$,
-$c_1(t)/\sigma_z$ represents the scaled local or
-statistical temperature\cite{kim2006, *kim2007},
+with its coefficient, $c_1$, updated according to
+the difference between the average of $z$
+and the center $z_c$.
+%
+We thus recover the Langfeld-Lucini-Rago (LLR)
+algorithm\cite{langfeld2012},
 and the inverse-time schedule
-was recommended for its update\cite{pellegrini2014}.
+was recommended for optimal convergence of $c_1$\cite{pellegrini2014}.
 
@@
 centered at $z_c$ with width $\sigma_z$
-we can use the Hermite polynomials\cite{arfken},
-and with $K=3$.
-%
-This results in a quadratic bias potential
-that can be written as a linear combination of
+we can use the Hermite polynomials\cite{arfken}.
+%
+With $K = 3$, the bias potential is quadratic
+and is a linear combination of
 $R_1(z) = (z - z_c)/\sigma_z$
@@
 %
-This form of bias potential was commonly employed
+Note that quadratic bias potentials are commonly employed
 in umbrella sampling (see e.g. Ref. \cite{zhu2012})
@@
 %
-While the asymptotic updating magnitude
-would be given by the inverse-time formula,
-we can employ the WL strategy early in simulation
-to control the updating magnitude in stages
-with the histogram fluctuation redefined by
-Eq. \eqref{eq:maxFk}.
-%
-We will refer to this method as
-adaptive Gaussian umbrella sampling below.
+By Eq. \eqref{eq:ckupdate},
+$c_1$ and $c_2$ are updated
+according to the first and second
+central moments of $z$, respectively.
+%
+Below we will refer to this application
+as adaptive Gaussian umbrella sampling.
 
@@
 %
-However, since we only used two modes
-for controlling the mean and variance,
-the method can only be applied
-to a narrow distribution.
-%
-To sample a wider range,
-we can construct an expanded ensemble with
-multiple Gaussian umbrellas
+However,
+the distribution, $\rho_i$,
+has to be sufficiently narrow
+because of the limited number of modes
+in controlling the moments.
+%
+We can nonetheless sample a wider range
+by constructing a mixture or expanded ensemble
+from a superposition of multiple Gaussian umbrellas
 centered at different places via
-either the replica exchange
-(or parallel tempering)\cite{swendsen1986,
-*geyer1991, *hukushima1996, *hansmann1997}
-or the simulated tempering\cite{marinari1992,
+the frameworks of
+parallel\cite{swendsen1986,
+  *geyer1991, *hukushima1996, *hansmann1997, *sugita1999, *earl2005}
+and/or simulated tempering\cite{marinari1992,
   *lyubartsev1992, li2007,
-  park2007, *nguyen2013, *zhang2015st}
-framework.
+  park2007, *nguyen2013, *zhang2015st}.
 %
@@
 $Z \equiv \int e^{-\beta \, \mathcal U(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$,
-where $\mathbf x$ denotes the molecular coordinates,
-$\mathcal U(\mathbf x)$ is the unbiased potential energy function,
+where
 $\beta$ is the inverse temperature,
-and the collective variable, $z$,
-has also been written as a function of $\mathbf x$.
+$\mathcal U(\mathbf x)$ is the unbiased potential energy
+as a function of molecular coordinates, $\mathbf x$,
+and $z$ has also been written as a function of $\mathbf x$.
 %
 The transition probability from
-the $i$th Gaussian to the $j$th is given by
-is given by
+the $i$th umbrella to the $j$th is given by
 \begin{align*}
@@
 %
-For a neighboring pair, $j = i\pm 1$,
-we can estimate the ratio $Z^{(j)}/Z^{(i)}$
-from the following approximate formula
-%
+While the asymptotic updating magnitude
+would be given by the inverse-time formula,
+we can employ the WL strategy early in simulation
+to control the updating magnitude in stages
+with the histogram fluctuation
+redefined by Eq. \eqref{eq:maxFk}
+(including the intra-umbrella and inter-umbrella
+components).
+%
+The weighted histogram analysis method
+(WHAM)\cite{ferrenberg1988, *ferrenberg1989,
+*kumar1992, *roux1995, *bartels1997, *souaille2001,
+*kobrak2003, *gallicchio2005, *chodera2007, zhu2012, newman, frenkel,
+fenwick2008, *kim2011}
+can be used to recover the unbiased distribution
+after simulation.
+%
+
+\note{The
+  following method appeared to have some stability issues,
+  so we switched back to the double-WL approach,
+  which hopefully is stable.
+  %
+  The reason of instability might be
+  that the inter-umbrella fluctuation
+  and the intra-umbrella fluctuation are some coupled.
+  So the updating magnitude cannot be adjusted individually
+  for the each umbrella.
+  %
+  Following the method for weight determination
+  in simulated tempering\cite{park2007,
+  *nguyen2013, *zhang2015st},
+  we can estimate $Z^{(j)}/Z^{(i)}$
+  for a neighboring pair, $(i, j)$, as
+  %
+  \begin{align*}
+    &\ln\frac{ Z^{(j)} }{ Z^{(i)} }
+    \approx
+    \ln\frac{ \sigma_z^{(j)} }{ \sigma_z^{(i)} }
+    +
+    \frac{ \Delta z_c } { 2 }
+    \left(
+      \frac{ c_1^{(i)} } { \sigma_z^{(i) 2} }
+      +
+      \frac{ c_1^{(j)} } { \sigma_z^{(j) 2} }
+    \right)
+    +
+    \frac{ c_2^{(j)} - c_2^{(i)} } { \sqrt 2 }
+    \\
+    &-
+    \left(
+      \frac{ \sqrt 2 \, c_2^{(j)} - 1 }
+           { 2 \, \sigma_z^{(j)} }
+      -
+      \frac{ \sqrt 2 \, c_2^{(i)} - 1 }
+           { 2 \, \sigma_z^{(i)} }
+    \right)
+    \left(
+      \frac{ \Delta z_c^{2} } { 6 }
+      +
+      \frac{ \sigma_z^{(i) 2} + \sigma_z^{(j) 2} } { 2 }
+    \right)
+    ,
+  \end{align*}
+  %
+  where
+  $\Delta z_c \equiv z_c^{(j)} - z_c^{(i)}$.
+  For equal width, we have
   \begin{align*}
@@
     \approx
-  \ln\frac{ \sigma_j }{ \sigma_i }
-   -u^{(j)}\bigl( z^{(i,j)} \bigr)
-   +u^{(i)}\bigl( z^{(i,j)} \bigr)
-  ,
+    \frac{ \Delta z_c \, ( c_1^{(i)} + c_1^{(j)} ) } { 2 \, \sigma_z }
+    -
+    \left(
+      \frac{ c_2^{(j)} - c_2^{(i)} }
+           { 6 \sqrt 2 }
+    \right)
+    \left(
+      \frac{ \Delta z_c^{2} } { \sigma_z^2 }
+    \right)
+    .
   \end{align*}
-%
-where
-$z^{(i,j)} \equiv
-\bigl( z_c^{(i)}/\sigma_z^{(i)} + z_c^{(j)}/\sigma_z^{(j)} \bigr) /
-\bigl(         1/\sigma_z^{(i)} +         1/\sigma_z^{(j)} \bigr)$.
-%
-In the $1/t$ scheme,
-%
-\note{For implementation, we can further simplify
-\begin{align*}
-  &-u^{(j)}\bigl( z^{(i,j)} \bigr)
-   +u^{(i)}\bigl( z^{(i,j)} \bigr)
-   \\
-   &=
-   (c_1^{(i)} + c_1^{(j)}) \, \Delta z^*
-   +
-   (c_2^{(i)} - c_2^{(j)}) \, \frac{{\Delta z^*}^2 - 1} {\sqrt 2}
-   ,
-\end{align*}
-where $\Delta z^* = [z_c^{(j)} - z_c^{(i)}]/[\sigma_z^{(j)} + \sigma_z^{(i)}]$.
-%
-However, the $c_2$ term does not appear to improve histogram flatness. Ideas?
 }
@@
 in the low density case
-was probably because that the sampling along
+was probably because the sampling along
 the collective variable was free from major barriers
@@
     & $5.7(5.7)\times10^{-2}$
-    & $9.1(11.4)\times10^{-5}$
+    & $9.2(11.4)\times10^{-5}$
     \\
@@
     $\sigma=0.1$ & $1/t$
-    & $7.7(8.9)\times10^{-3}$
+    & $7.5(8.9)\times10^{-3}$
     & $6.8(8.2)\times10^{-5}$
@@
     $\sigma=0.1$ & opt.
-    & $7.6(8.9)\times10^{-3}$
-    & $3.2(3.9)\times10^{-5}$
+    & $7.5(8.9)\times10^{-3}$
+    & $3.3(3.9)\times10^{-5}$
     \\
@@
     $\sigma=0.2$ & $1/t$
-    & $6.5(7.8)\times10^{-3}$
-    & $4.0(4.8)\times10^{-5}$
+    & $6.6(7.8)\times10^{-3}$
+    & $3.9(4.8)\times10^{-5}$
     \\
@@
     $\sigma=0.2$ & opt.
-    & $7.0(7.8)\times10^{-3}$
-    & $2.5(2.8)\times10^{-5}$
+    & $6.7(7.8)\times10^{-3}$
+    & $2.4(2.9)\times10^{-5}$
     \\
@@
     $\sigma=0.5$ & $1/t$
-    & $5.5(6.6)\times10^{-3}$
-    & $3.2(3.8)\times10^{-5}$
+    & $5.4(6.6)\times10^{-3}$
+    & $3.3(3.8)\times10^{-5}$
     \\
@@
     & $5.8(6.6)\times10^{-3}$
-    & $2.6(3.0)\times10^{-5}$
+    & $2.7(3.0)\times10^{-5}$
     \\
@@
 and it would be best used as a post-simulation
-analysis tool for simulations with detectable bias.
+analysis tool for simulations
+with detectable residual bias.
 %
@@
 %
-In this case, the schedule predicted by our method
+In this case, the predicted schedule
 would be unusable because the resulting deviation in the bias potential
@@
     \label{fig:lj_alpha}
-    Optimal schedule of the Gaussian updating scheme
+    Optimal schedules of Gaussian updating schemes
     for the LJ system.
@@
 We spread the Gaussian umbrellas
-the same width $\sigma_E = L$ over $[-1.7L^2, -0.9L^2]$
-with a spacing of $2L$.
-%
-For configuration sampling for each umbrella,
+of the same width $\sigma_E = L$ over $[-1.8L^2, -0.8L^2]$
+with a spacing of $L$.
+%
+For simplicity,
+only transitions between neighboring umbrellas
+were used.
+%
+For configuration sampling in each umbrella,
 we used the Wolff algorithm\cite{wolff1989, newman}
-(which appeared to be more efficient than the Metropolis algorithm),
+(which is more efficient than the Metropolis algorithm)
 at temperature $c_1/\sigma_E$
@@
 %
-To ensure stability,
-we avoid transitions among umbrellas
-until the current umbrella has
-completed the WL stage-wise way of
-reducing the updating magnitude
-and started the inverse-time schedule.\cite{belardinelli2007,
-*belardinelli2007jcp, *belardinelli2008, *belardinelli2016}
+The updating magnitude started at $0.01$,
+and it underwent the WL procedure
+with the fluctuation threshold being $0.2$
+[using the definition of Eq. \eqref{eq:maxFk}]
+and the reduction factor being $1/2$
+until it fell below
+the value given the inverse-time schedule\cite{
+belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
+%
+This updating magnitude was
+used to update the values of both $\ln Z^{(i)}$'s
+for inter-umbrella transitions
+and $c_1^{(i)}$'s and $c_2^{(i)}$'s
+for intra-umbrella sampling.
+
+%To ensure stability,
+%we avoid transitions among umbrellas
+%until the current umbrella has
+%completed the WL stage-wise way of
+%reducing the updating magnitude
+%and started the inverse-time schedule.\cite{belardinelli2007,
+%*belardinelli2007jcp, *belardinelli2008, *belardinelli2016}
 
@@
 The Gaussian histograms were evenly spaced with similar width as intended,
-and the overall energy histogram was roughly flat,
+%and the overall energy histogram was roughly flat,
 as shown in Fig. \ref{fig:pt_hist}(a).
 %
-The parameter $c_1/\sigma_E$ roughly represents the local temperature,
+By WHAM, we found the critical temperature at $0.70161$
+in agreement with the previous study\cite{wang2001pre}.
+%
+The parameter, $c_1/\sigma_E$, roughly represents
+the local or statistical temperature,
 and it demonstrated
 the backbending behavior\cite{kim2006, *kim2007}
-typical in a first-order phase transitions,
+typical in a first-order phase transition,
 as shown in Fig. \ref{fig:pt_hist}(b).
 %
-The parameter $c_2$ represents the amount of ``pressure''
-to restrain the distribution width,
-and its value was maximal around the center critical region,
+The second term of the bias potential,
+$c_2 \, R_2(E) = \sqrt 2 \, c_2 \, [(E - E_c)^2/(2 \, \sigma_E^2) - \frac 1 2]$,
+provides a quadratic restraint on the distribution width.
+%
+The value of $\sqrt 2 \, c_2$ represents the amount of ``pressure''
+needed to restore the target Gaussian,
+$\rho(E) \propto \exp[-(E - E_c)^2/(2 \, \sigma_E^2)]$,
+and a value over $1.0$ suggests
+that the intrinsic distribution
+(or the density of states here)
+is locally convex,
+which was so in the critical region,
 as shown in Fig. \ref{fig:pt_hist}(c).
@@
     \label{fig:pt_hist}
-    Application of adaptive Gaussian umbrella sampling
-    on the $64 \times 64$ ten state Potts model.
-    %
-    (a) Normalized histograms (solid lines) from adaptive Gaussian umbrella sampling
-    matched the intended distributions.
-    The overall histogram (dashed line) was roughly flat.
+    Adaptive Gaussian umbrella sampling
+    on the $64 \times 64$ ten-state Potts model.
+    %
+    (a) Normalized histograms (solid lines)
+    and the reweighted distribution
+    at the critical temperature (dashed line).
     %
@@
     %
-    (c) The parameter $c_2$ for restraining the variance
-    was maximal around the critical region.
+    (c) The parameter, $\sqrt 2 \, c_2$, exceeds $1.0$
+    in the critical region, indicating local convexity
+    of the density of states.
   }
@@
 of glassy systems.
+%
+Finally, the adaptive Gaussian umbrella sampling
+approach still needs some refinement to improve
+its stability for general application.
+%
+We find that with slow sampling processes,
+the updating scheme may occationally overly update
+the parameter, $c_2$, leading to unreasonable values.
 %Our immediate interest is to apply these relations to
@@
 
-We wish to show the optimality of the inverse-time schedule
+The optimality of the inverse-time schedule
 for the single-bin updating scheme
-from the observation that
-it leads to a bias potential
-that coincides with a runtime average
+can be understood from the observation that
+the schedule allows the bias potential
+to coincide with a runtime average
 of the independent corrections.
@@
 and apply it to the instantaneous histogram, $h_i(t)$,
-then the independent estimate of bias potential in each step
-is given by
+the independent estimate of bias potential in each step
+would be given by
 %
@@
 \end{align}
-
-Equation \eqref{eq:uhatav} serves as
+%
+%Equation \eqref{eq:uhatav}
+This equation serves as
 a generalization of Eq. \eqref{eq:vcorr_equil}
 for FDS simulations under a variable bias potential
-and it can be used to improve the estimated PMF
+and we can use it to improve the estimated PMF
 after simulation.
@@
 where
-$$
-\Delta u_i \equiv \frac{u_i(t+1)}{t \, \alpha(t)}
-- \frac{u_i(1)}{t \, \alpha(0)}.
-$$
+$\Delta u_i \equiv u_i(t+1) /[t \, \alpha(t)]
+- u_i(1)/[t \, \alpha(0)]$.
+%$$
+%\Delta u_i \equiv \frac{u_i(t+1)}{t \, \alpha(t)}
+%- \frac{u_i(1)}{t \, \alpha(0)}.
+%$$
 %
 With the inverse-time schedule, Eq. \eqref{eq:alpha_invt},
-each term in the square brackets vanishes,
+each term in the sum vanishes,
 and if we further define $\alpha(0) \to \infty$,
@@
 i.e. the bias potential in this case coincides with
-the runtime average of the correction, $\hat u_i(t)$
-at any step, $t$.
-%
-Another special case is $\alpha(\tau)$ is a constant for $\tau = 0, 1, \dots$,
-then Eq. \eqref{eq:Uhat_sum} is reduced to a plain average\cite{zhou2005}:
+the runtime average of the correction, $\hat u_i(t)$,
+at every step, $t$.
+%
+On the other hand,
+with a constant $\alpha(\tau) = a_0$ for $\tau = 0, 1, \dots$,
+Eq. \eqref{eq:Uhat_sum} is approximately a plain average\cite{zhou2005}:
 $\hat U_i(t) \approx U_i(t)$,
@@
 with $u_i(t+1) \approx u_i(1)$.
-
-%Thus, the bias potential from a WL simulation
-%(which typically uses the single-bin updating scheme)
-%under the inverse-time schedule
-%would reach a similar precision
-%as an ideal multicanonical simulation
-%with fixed and nearly perfect bias potential.
-%the precision at long times.
-%%
-%Thus, it is unnecessary to fix the bias potential
-%after an adaptive FDS simulation has reached
-%sufficiently convergent just for better precision
-%of the final PMF.
-
-%Second,
-%it shows close connection between
-%the WL algorithm under the inverse-time schedule
-%and self-healing umbrella sampling.
-%In the latter,
-%the bias potential is computed from
-%a weighted runtime average of the histogram\cite{
-%  marsili2006, barducci2008},
-%which is analogous to the adaptive bias force method\cite{
-%  darve2001, *darve2008}.
-%
-%Third,
-%it also relates a WL simulation under the inverse-time schedule
-%to a time-averaged variant of the regular WL simulation.
-%%
-%In the variant, the updating magnitude is fixed,
-%but the bias potential over the simulation course is averaged
-%as the final result\cite{zhou2005}.
-%%
-%The above demonstration shows that the inverse-time schedule
-%has already effectively implemented an implicit average
-%so that no explicit averaging is needed in the former approach.
-%%
-%The inverse-time schedule approach is preferred, however,
-%because with a decreasing updating magnitude,
-%it minimizes the systematic error caused by
-%the on-the-fly updating scheme\cite{
-%  zhou2005, zhou2008},
-%which is ignored in this analysis.
+\note{This would be rather risky for metadynamics,
+because we know that the certain noise modes
+do not easily go away, although the plain average
+formula was also recommended for metadynamics.}
+
+We can also cast other updating relations,
+such as Eqs. \eqref{eq:wl_update}
+and \eqref{eq:vkupdate},
+as runtime averages,
+but with dynamic weights.
+%
+The weighted average
+%
+\begin{align*}
+  X_\omega(t)
+  =
+  \frac{
+    \omega(0) \, x(0) + \cdots + \omega(t) \, x(t)
+  }
+  {
+    \Omega(t)
+  }
+  ,
+\end{align*}
+%
+where
+$\Omega(t) \equiv \omega(0) + \cdots + \omega(t)$
+is the accumulated weight,
+is equivalent to the recurrence relation
+%
+\begin{align*}
+X_\omega(t) = X_\omega(t-1)
+  + \frac{\omega(t)}{\Omega(t)}
+  \Delta \, x(t)
+,
+\end{align*}
+where
+$\Delta x(t) \equiv x(t) - X_\omega(t-1)$
+is the proposed change from the previous average
+from the data point at step $t$.
+%
+The relative weight, $\omega(t)/\Omega(t)$,
+can be identified as the updating magnitude,
+\begin{equation}
+  \alpha^*(t) \equiv
+  \frac{ \omega(t) } {\Omega(t)}
+  =
+  1 - \frac{ \Omega(t-1) } {\Omega(t)}
+  .
+  \label{eq:alpha_from_Omega}
+\end{equation}
+%
+For a plain average, $\omega(t) = 1$, for $t > 0$,
+the updating magnitude would follow
+the inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
+and updating magnitudes above and below this value
+correspond to runtime averages that favor future and past data,
+respectively.
+%
+In the case of Eq. \eqref{eq:vkupdate},
+we can identify
+the bias potential $\tilde v_k(t)$
+as $x(t)$,
+$\tilde f_k(t)$ as $\Delta x(t)$
+by Eq. \eqref{eq:uhatt},
+and
+the mode-$k$ updating magnitude,
+$\alpha(t) \, \lambda_k$,
+as $\alpha^*(t)$.
+%
+This mapping shows that
+updating magnitude should not exceed $1.0$.
+%
+One way to satisfy this condition
+in converting the results
+from the continuous-time setup is
+to use Eq. \eqref{eq:alpha_from_Omega}
+with
+$\Omega(t) = \prod_{\tau=1}^t [1- \alpha^*(\tau)]^{-1}
+\approx \exp\left(\int_0^t \alpha^*(\tau) \, d\tau \right)$.
+%
 
@@
 i.e. FDS methods are most effective
-for short ranges\cite{wang2001, *wang2001pre}.
+for short ranges\cite{wang2001, wang2001pre}.
 %

### 2017-06-23 prog/aus/pt2gaus.c

@@ -16,18 +16,19 @@ enum { SAMP_METROPOLIS, SAMP_WOLFF };
 
 
 
-/* modified Metropolis algorithm */
+/* modified Metropolis algorithm
+ * p(E) ~ exp(-b1*(E-Eave)-b2*(E-Eave)^2/2) */
 __inline static int potts2_metro_mod(potts2_t *pt,
-    double c1, double c2, double Eave)
+    double b1, double b2, double Eave)
 {
   int id, h, sn, enew;
   double edev, dv;
 
   POTTS2_PICK(pt, id, h, sn);
   enew = pt->E + h;
-  /* compute the change of bias potential */
+  /* compute the change of the bias potential */
   edev = (enew + pt->E)*.5 - Eave;
-  dv = h * (c1 + c2 * edev);
+  dv = h * (b1 + b2 * edev);
   if ( dv <= 0 || rand01() < exp(-dv) ) {
     POTTS2_FLIP(pt, id, h, sn);
     return 1;
@@ -37,14 +38,15 @@ __inline static int potts2_metro_mod(potts2_t *pt,
 
 
 
-/* modified Wolff algorithm */
+/* modified Wolff algorithm
+ * p(E) ~ exp(-b1*(E-Eave)-b2*(E-Eave)^2/2) */
 __inline static int potts2_wolff_mod(potts2_t *pt,
-    double c1, double c2, double Eave)
+    double b1, double b2, double Eave)
 {
   int l = pt->l, n = pt->n, i, ix, iy, id, so, sn, cnt = 0, h = 0;
   double padd, enew, dv;
 
-  padd = 1 - exp(-c1);
+  padd = 1 - exp(-b1);
 
   /* randomly selected a seed */
   id = (int) ( rand01() * n );
@@ -69,7 +71,7 @@ __inline static int potts2_wolff_mod(potts2_t *pt,
   }
 
   enew = pt->E + h;
-  dv = h * c2 * ((enew + pt->E)*.5 - Eave);
+  dv = h * b2 * ((enew + pt->E)*.5 - Eave);
   if ( dv <= 0 || rand01() < exp(-dv) ) {
     pt->E += h;
     return 1;
@@ -105,15 +107,16 @@ static void potts2_gaus(potts2_t *pt,
 {
   long t, nstsave;
   int id, acc;
[…]

### 2017-06-23 prog/aus/reweight.c

@@ -1,0 +1,235 @@
+/* standalone histogram reweighting program */
+
+#include "util.h"
+
+#ifndef SQRT2
+#define SQRT2 1.4142135623730951
+#endif
+
+const char *fndat = "pt2gaus.dat";
+const char *fnhis = "pt2gaus.his";
+int n; /* number of Gaussians */
+double *ave, *sig, *c1, *c2, *lnz, *cnt;
+double xcmin, xcmax, delx;
+double *hist, *htot, *lng;
+int xmin, xmax, dx, xn;
+
+/* load the dat file */
+static int getdatinfo(const char *fn)
+{
+  FILE *fp;
+  int i, i1;
+  char buf[256];
+
+  if ((fp = fopen(fn, "r")) == NULL) {
+    fprintf(stderr, "cannot open %s\n", fn);
+    return -1;
+  }
+  fgets(buf, sizeof buf, fp);
+  sscanf(buf, "# %d", &n);
+  xnew(ave, n);
+  xnew(sig, n);
+  xnew(c1, n);
+  xnew(c2, n);
+  xnew(lnz, n);
+  xnew(cnt, n);
+  for ( i = 0; i < n; i++ ) {
+    fgets(buf, sizeof buf, fp);
+    sscanf(buf, "%d%lf%lf%lf%lf%lf", &i1, &ave[i], &sig[i], &c1[i], &c2[i], &lnz[i]);
+    cnt[i] = 0;
+  }
+  xcmin = ave[0];
+  delx = ave[1] - ave[0];
+  xcmax = ave[n-1];
+  fclose(fp);
+  return 0;
+}
+
+static int gethisinfo(const char *fn)
+{
+  FILE *fp;
+  int i, ix, ln = 1;
+  char buf[256];
+  double x, y, z, h;
+
+  if ((fp = fopen(fn, "r")) == NULL) {
+    fprintf(stderr, "cannot open %s\n", fn);
+    return -1;
+  }
+  fgets(buf, sizeof buf, fp);
+  sscanf(buf, "# %d %d %d %d", &n, &xn, &xmin, &dx);
+  xnew(hist, n * xn);
+  xnew(htot, xn);
+  xnew(lng, xn);
+  for ( ix = 0; ix < xn; ix++ ) htot[ix] = 0;
+
+  for ( i = 0; i < n; i++ ) {
+    fgets(buf, sizeof buf, fp);
+    ln++;
+    if ( buf[0] != '#' ) {
[…]

### 2017-06-24 doc/invt.tex
invt.tex: improve massq, V6.3

@@
 by a differential equation\footnote{In
-going to the continuous time setup,
+going to the continuous-time setup,
 we find it more convenient to shift the origin of time by $-1$,
@@
 %
-If the starting bias potential is zero,
+If the initial bias potential
+before the priliminary run is zero,
 one can show that
@@
   .
-  %=
-  %\mathrm{const.}
   \notag
@@
 is equal to $t/T$,
-as shown in Fig. \ref{fig:massq}.
+as shown in Fig. \ref{fig:massq}(a).
 
@@
 \begin{center}
-  \begin{tikzpicture}[very thick, scale=2.5, every node/.style={scale=1.5}]
+  \begin{tikzpicture}[thick, scale=1.25, every node/.style={scale=0.8}]
+    \node [] at (-0.3, 2.0) {(a)};
     % distribution
-    \draw[domain=0:2.5, variable=\x, dashed, gray, ultra thick]
-      plot({\x}, {sqrt(2*exp(-2*\x)+exp(-\x))});
+    \draw[domain=0:2.5, variable=\x, dashed, gray, very thick]
+      plot({\x}, {2*exp(-\x)});
 
@@
       (0.7, 0) --
-      plot ({\x}, {sqrt(2*exp(-2*\x)+exp(-\x))})
+      plot ({\x}, {2*exp(-\x)})
       -- (2.2, 0) -- cycle;
@@
     % thick curve
-    \draw[domain=0:2.2, variable=\x, ultra thick]
-      plot({\x}, {sqrt(2*exp(-2*\x)+exp(-\x))});
+    \draw[domain=0:2.2, variable=\x, very thick]
+      plot({\x}, {2*exp(-\x)});
 
     % label for the shaded area
-    \node[] at (1.1, 0.40) {$\dfrac{t}{T}$};
+    \node[] at (1.1, 0.36) {$\dfrac{t}{T}$};
 
@@
                {$\dfrac{1}{T \, \alpha}$}
-               (0.6, 0.99488);
+               (0.6, 0.99317);
     % label for the ordinate
@@
 
-    \draw[] (0.5, 0.99488) -- (0.7, 0.99488);
+    \draw[] (0.5, 0.99317) -- (0.7, 0.99317);
     \draw[->] (0, 0) -- (2.8, 0);
     \node[] at (2.8, -0.2) {$\bar q$};
-    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 1.9);
+    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 2.2);
 
@@
   \end{tikzpicture}
-  %\makebox[\linewidth][c]{
-  %  \includegraphics[angle=0, width=0.9\linewidth]{fig/massq.pdf}
-  %}
+%
+  \begin{tikzpicture}[thick, scale=1.25, every node/.style={scale=0.8}]
+    \node [] at (-0.3, 2.0) {(b)};
+    % distribution
+    \draw[domain=0:2.3, variable=\x, dashed, gray, very thick]
+      plot({\x}, {1/(\x+0.5)});
+
+    % shade
+    \fill [gray!30!white, domain=0.7:2.0, variable=\x]
+      (0.7, 0) --
+      plot ({\x}, {1/(\x+0.5)})
+      -- (2.0, 0) -- cycle;
+
+    % thick curve
+    \draw[domain=0:2.0, variable=\x, very thick]
+      plot({\x}, {1/(\x+0.5)});
+
+    % label for the shaded area
+    \node[] at (1.1, 0.40) {$\bar q(t)$};
+
+    \draw[<->] (0.6, 0)  --
+               node[left, inner sep=1mm]
+               {$\dfrac{1}{T \, m}$}
+               (0.6, 0.83333);
+    % label for the ordinate
+    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
+
+    \draw[] (0.5, 0.83333) -- (0.7, 0.83333);
+    \draw[->] (0, 0) -- (2.6, 0);
+    \node[] at (2.6, -0.2) {$t$};
+    \draw[->] (0, 0) -- node[above, rotate=90] {$\alpha(t)$} (0, 2.2);
+
+    % xtics
+    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
+    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$t$};
+    \draw[] (2.0, 0.1) -- (2.0, 0.0) node[below] {$T$};
+  \end{tikzpicture}
   \caption{
     \label{fig:massq}
-    Geometric construction of the optimal schedule,
+    (a) Geometric construction of the optimal schedule,
     $\alpha(t)$,
@@
     %
+    (b) Inverse construction of $m(\bar q)$
+    from the optimal schedule $\alpha(t)$.
+    %
     \note{
       The figure was produced by \texttt{tikz}.
-      % Gnuplot version by doc/fig/massq.gp
     }%
@@
 if we apply the same transform on $\alpha^*(t^*)$
-we would recover $m^*(\bar q^*)$.
+we would recover $m^*(\bar q^*)$,
+as shown in Fig. \ref{fig:massq}.
 %
@@
 at large $\bar q$,\footnote{If
-  the magntidue of the eigenvalues, $\lambda_k$'s,
+  the magnitude of the eigenvalues, $\lambda_k$'s,
   span a wide range, then for a large $\bar q$,
@@
   }
-  \setlength{\tabcolsep}{3pt} % spacing between columns
+  \setlength{\tabcolsep}{2pt} % spacing between columns
   \renewcommand\arraystretch{2.0} % spacing between rows
@@
     %$\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a} - q^*_0$
-    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^{\frac 1 a} - 1\right]$
+    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^{1/a} - 1\right]$
     %$\,\begin{aligned}
@@
 
-%For simulations with the same $a_0$,
-%the shape of the mass distribution, $m(\bar q)$,
-%also determines how much the optimal schedule
-%would depend on the simulation length, $T$.
-%%
-%Since both $M(\bar q)$ and $m(\bar q)$ decrease monotonically with $\bar q$,
-%Eq. \eqref{eq:opt_qT} shows that
-%the optimal $q(T)$ increases with
-%the simulation length, $T$,
-%more so if $M(\bar q)$ has a long tail.
-%%
-%Thus, if $M(\bar q)$ decays slowly because of
-%many near zero eigenvalues, $\lambda_k$'s,
-%as in the metadynamics case,
-%both the optimal schedule and
-%normalized asymptotic error, $\Err_A(T) \cdot T$,
-%would be sensitive to the simulation length, $T$.
-%%
-%In contrast, if all eigenvalues are $1.0$
-%as in the single-bin or WL updating scheme discussed previously,
-%the optimal schedule is
-%given by the universal inverse-time formula,
-%Eq. \eqref{eq:alpha_invt1},
-%and the normalized asymptotic error
-%is roughly a constant
-%[cf. Eq. \eqref{eq:error_singlebin}].
-%
-%If $M(\bar q)$ decays rapidly,
-%the distribution $m(\bar q)$ does not critically
-%depend on the cutoff $q(T)$,
-%once $q(T)$ is sufficiently large.
-%%
-%So the optimal schedules of these simulations
-%approximately satisfy the scaling relations
-%described in Sec. \ref{sec:mass_distr}.
-%%
-%If, however, $M(\bar q)$ has a long tail
-%because of some near-zero $\lambda_k$'s,
-%increasing the simulation length $T$
-%generally alters the shape of the optimal schedule.
-
 \subsubsection{\label{sec:band-matrix}
@@
 as shown in Fig. \ref{fig:vardep}.
-%
-%The dependence of variables are shown in Fig. \ref{fig:vardep}.
 %
@@
   $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
-  where $i = 0, \dots, \mathcal N$, and we chose $\mathcal N=20000$.
+  where $i = 0, \dots, \mathcal N$,\footnote{For
+    a mass distribution, $m(\bar q)$, that is
+    inversely proportional to $\bar q$,
+    this grid would make the integral,
+    or the difference, $t_{[i+1]} - t_{[i]}$,
+    roughly a constant.}
+  and we chose $\mathcal N=20000$.
 %
@@
 %For these updating schemes,
-%  the preliminary run with a constant updating magntiude
+%  the preliminary run with a constant updating magnitude
 %  can be replaced by the WL-style stagewise reduction
@@
 $$
-which must carry a non-positive discriminant:
-$A_1^2 - 4 \, A_0 \, A_2 \le 0$.},
+which implies a nonpositive discriminant,
+$A_1^2 - 4 \, A_0 \, A_2$.}
 %
@@
 %%
-%\begin{equation}
-%  \alpha(t) = \frac{              1             }
-%                   { \lambda_k \, (t + t_{k0} ) },
-%\label{eq:alpha_invtk}
-%\end{equation}
-%with
-%\begin{equation}
-%  t_{k0} = \frac{             T            }
-%                { e^{ \lambda_k q(T) } - 1 }.
-%\label{eq:tk0}
-%\end{equation}
-%%
-%\note{Derivation.
-%  Integrating $\dot \vartheta_k = c_k$ yields
-%  \begin{equation}
-%    \vartheta_k(t) = c_k \left(t + t_{k0} \right).
-%  \tag{UK}
-%  \label{neq:ui_solution}
-%  \end{equation}
-%  Given that $\vartheta_k(T) = 1$ and
-%  $\vartheta_k(0) = e^{-\lambda_k \, q(T)}$,
-%  we have
-%  $$
-%  c_k = \frac{ 1 }{ T + t_{k0} },
-%  \quad
-%  \mathrm{and\;}
-%  \frac{ t_{k0} } { T + t_{k0} }
-%  =
-%  e^{ -\lambda_k \, q(T) }.
-%  $$
-%  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
-%  $$
-%  -\lambda_k \, q(T) + \lambda_k \, q(t)
-%  = \ln c_k + \ln \left( t + t_{k0} \right).
-%  $$
-%  Differentiating this with respect to $t$ yields
-%  Eq. \eqref{eq:alpha_invtk}.
-%}
-%%
-%Such a solution is possible only if
-%all nontrivial eigenvalues are equal,
-%%
-%%\begin{equation}
-%  $\lambda_1 = \dots = \lambda_{n-1}$.
-%%  \notag
-%%\end{equation}
-%%
-%Then, by Eq. \eqref{eq:tk0},
-%all $t_{k0}$'s also share the same value,
-%$t_0$,
-%such that
-%$\vartheta_k = (t + t_0) / (T + t_0)$,
-%and
-%$c_k = 1/(T + t_0)$
-%for $k > 0$.
-%%
-%
-%
 
@@
   we will map
-  $\rho_i$ to $\rho(z) \delta z$,
-  $\phi_{ki}$ to $\phi_k(z) \delta z$,
-  $w_{ij}$ to $w(z, z') \delta z$, etc.}
+  $\rho_i$ to $\rho(z) \, \delta z$,
+  $\phi_{ki}$ to $\phi_k(z) \, \delta z$,
+  $w_{ij}$ to $w(z, z') \, \delta z$, etc.}
 with the eigenmodes derived from orthogonal polynomials, $R_k(z)$,
@@
 The two special particles are mutually non-interacting
-but they interact normally with other particles,\footnote{they
-have placed on an axis parallel to a side of the cubic box}
+but they interact normally with other particles,\footnote{They
+were placed on an axis parallel to a side of the cubic box}
 and the intrinsic distribution, $p(r)$, is
@@
 and the errors roughly matched the predicted values
-calculated from the precomputed
+calculated from the pre-computed
 $\Gamma_k$'s and $\epsilon_k$'s.
@@
     single-bin & $1/t$
-    & $1.3(1.4)$
-    & $2.7(2.7)\times10^{-3}$
+    & $1.4(1.4)$
+    & $3.4(2.7)\times10^{-3}$
     \\
@@
     & $1.3(1.3)$
-    & $2.9(2.8)\times10^{-3}$
+    & $3.0(2.8)\times10^{-3}$
     \\
@@
     & $1.2(1.3)$
-    & $2.8(2.8)\times10^{-3}$
+    & $3.0(2.8)\times10^{-3}$
     \\
@@
     & $1.3(1.2)$
-    & $3.4(3.2)\times10^{-3}$
+    & $3.5(3.2)\times10^{-3}$
     \\
@@
     $\sigma=0.5$ & $1/t$
-    & $0.9(0.9)$
-    & $9.8(8.0)\times10^{-3}$
+    & $1.0(0.9)$
+    & $9.7(8.0)\times10^{-3}$
     \\
@@
     & $1.0(0.9)$
-    & $5.8(5.4)\times10^{-3}$
+    & $5.9(5.4)\times10^{-3}$
     \\
@@
     $K=20$ & $1/t$
-    & $1.3(1.3)$
+    & $1.4(1.3)$
     & $3.1(2.7)\times10^{-3}$
@@
 less error than the inverse-time schedule
-for the Gaussian updating scheme,
 as shown in Table \ref{tab:lj_error}.
+%
+Although the difference was small in the reported cases,
+the gain of using the optimal schedule
+would increase with the simulation length.\footnote{
+  For example, for the $\varrho^* = 0.1$ case
+  with width $\sigma = 0.2$,
+  the predicted error of a simulation of $10^8$ steps
+  was $2.5\times10^{-5}$ for the inverse-time schedule
+  or $3.7\times10^{-6}$ for the optimal schedule.}
 
@@
 for the preliminary run under constant updating magnitude,
-the starting magnitude was around
+the initial magnitude was around
 the ideal value of $a_0/2 = 5\times10^{-5}$
@@
 of the same width $\sigma_E = L$ over $[-1.8L^2, -0.8L^2]$
-with a spacing of $L$.
+with an even spacing of $L$.
 %
@@
 only transitions between neighboring umbrellas
-were used.
+were performed.
 %
@@
 %
-By WHAM, we found the critical temperature at $0.70161$
-in agreement with the previous study\cite{wang2001pre}.
+By WHAM, we found the critical temperature at {\color{red}$0.70162$}
+in agreement with a previous study\cite{wang2001pre}.
 %
@@
 %
-The value of $\sqrt 2 \, c_2$ represents the amount of ``pressure''
+The coefficient, $\sqrt 2 \, c_2$, represents the amount of ``pressure''
 needed to restore the target Gaussian,
@@
 (or the density of states here)
-is locally convex,
+is locally convex or unstable,
 which was so in the critical region,
@@
 %
-While this feature helps suppressing
+While this feature helps suppress
 the accumulation of random error and
-maintaining the smoothness of the bias potential,
+maintain the smoothness of the bias potential,
 it also slows down the reduction
@@
 of the independent corrections.
-%as a natural consequence of the stabilization of a runtime average
-%as the sample size increases\cite{
-%  marsili2006, barducci2008}.
-%
-%We first note that a runtime average
-%\begin{equation}
-%  X(t) = \frac{x(1) + \cdots + x(t)}{t}
-%  .
-%  \label{eq:avx}
-%\end{equation}
-%can be written as a recurrence relation
-%\begin{equation}
-%  X(t) = X(t - 1) + \frac{x(t) - X(t-1)}{t}
-%  .
-%  \label{eq:avx_recur}
-%\end{equation}
-%%
-%Here $x(t) - X(t-1)$ serves as an independent correction
-%from data point at time $t$ to the previous average value,
-%and this correction is absorbed into the average
-%with a weight of $1/t$,
-%which is the inverse sample size.
-
-%We can use Eq. \eqref{eq:avx_recur} to incrementally
-%improve the estimated
-%bias potential given by Eq. \eqref{eq:vcorr_equil}.
-%
+
 Assuming that we can linearize the histogram correction
@@
 This definition can be readily extended
-to the continuous variable case, with
+to a continuous variable, $z$, with
 $\tilde F_k(T) = \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)$.
@@
 %
-The basic idea is to use Eq. \eqref{eq:xt2_eql}.
-%
-However,
-without knowing the intrinsic distribution, $p_i$,
-we cannot compute the shifted bias potential, $v_i$.
-%
-The problem can be avoided by replacing the shift
+The basic idea is to use Eq. \eqref{eq:xt2_eql},
+and to avoid the problem of not
+knowing the intrinsic distribution, $p_i$,
+hence the potential shift, $\ln(p_i/\rho_i)$,
+we should replace the shift
 by the time average of the bias potential.
@@
 Eigenvalues and eigenvectors}
-
-%\subsubsection{\label{sec:bandkernel}
-%Updating kernel and eigenvalues}
 

### 2017-06-26 doc/response.tex
invt.tex: various corrections, remove Appendix B and stabilization
absorbing Dr. Pettitt's changes

@@
-\documentclass{article}
+\documentclass[preprint, superscriptaddress, floatfix]{revtex4-1}
 \usepackage{amsmath}
@@
 The old $x_i$ is replaced by $v_{*i}$.
+The old $u$ is replaced by $\vartheta$.
 The old $y_k$ is replaced by $\tilde v_k$.
@@
 
+Added description for the duality between the mass distribution and schedule.
+
+Added a histogram correction formula.
+
+Improved Appendix A.
+
+Added an appendix for histogram fluctuation.
+
+\section{Reviewer \#1 Evaluations:}
+
+Recommendation: Major revision
+
+New Potential Energy Surface: No
+
+Reviewer \#1 (Comments to the Author):
+
+My review is attached.
+
+This manuscript proposes a method to compute the optimal schedule of the update magnitude for
+adaptive flat-distribution-sampling (FDS) free energy simulation methods, including the Wang-
+Landau (WL) algorithm and metadynamics. This method characterizes each adaptive FDS
+simulation by its updating scheme encoded in an updating matrix, and then computes the optimal
+schedule by solving a numerical variational problem, which uses the eigenvalues and
+eigenvectors of the updating as input along with other quantities to the variational equations to
+be minimized. The authors show that the optimal schedule for the single-bin (WL) scheme is
+given by the well known inverse-time formula. Making an interesting analogy, they point out
+that for a general multiple-bin scheme, the optical schedule is implicitly given by an equation of
+motion of a free particle with a position-dependent mass. They suggest a “bandpass” updating
+scheme which they show has better convergence than the metadynamics Gaussian updating
+scheme.
+
+There are several very interesting ideas introduced in this paper, and I recommend publication
+following revisions to address my questions and concerns.
+
+The manuscript is highly mathematical, and it is difficult to follow all the details without
+devoting a very large amount of time to studying it. The authors can do a better job of
+explaining their mathematical analysis of FDS updating schemes, by reorganizing their
+manuscript, moving some sections of the manuscript to the appendix and expanding the
+explanations, while moving one section which is currently in an appendix to the main text. The
+procedure for computing the optimal schedule and error in Appendix D should be moved to the
+main text and the steps involved explained in more detail in a single section. An explanation of
+how eq. 55 is solved for q(T) is needed, this can be placed in an appendix, and an explanation of
+how eq. 49 is numerically integrated is also needed. A clearer discussion of what autocorrelation
+functions are being calculated, and how they are used in the variational optimization is also
+needed.
+
+Response.
+We are grateful to the reviewer for the time and effort in reviewing the manuscript and we apologize for the lack of polishing and structure in the manuscript.
+We have move the old Appendix D to the main text (current section II.F)
+and added some technical details.
+In particular, q(T) is solved from the bisection method.
+For the optimal schedule,
+Eqs. (48) and (49) are integrated on an even grid of q,
+using the standard trapezoidal rule.
+We have added some details in the new section IIF.
+Thank you.
+
+My biggest concern with the paper relates to the numerical example. It seems that the single test
+system studied in detail is a completely flat “potential”. While it is interesting to note that
+different FDS schemes will exhibit different efficiencies on this “trivial” example, it doesn’t give
+the reader any feeling of the corresponding results on the simplest “non-trivial” 1D or 2D
+problem with multiple wells and barriers, which at least begins to approach problems that
+motivate the need for advanced sampling schemes. The authors should report the results of at
+least one example of this kind in addition to the test system reported. While I am not requiring
+the authors to analyze how their (bandpass) FDS scheme and analysis methods perform on more
+realistic potentials than 1D or 2D potential with multiple barriers, I would point out that a key
+issue is whether metadynamics or any FDS scheme can really perform well on problems where
+there is a free energy balance between stable basins separated by free energy barriers, and the
+balance involves a competition between the enthalpic stabilization of one basin and the entropic
+stabilization of the other. The authors may want to comment on whether they think FDS
+sampling schemes are well suited for these kinds of landscape problems.
+
+Some additional questions:
+1. During the derivations of Eq.(22) and Eq.(49), the end point value $q(T)$ is fixed. Does
+this mean that the optimal schedule of the update magnitude suggested by Eq.(22) or (49)
+is only for a subgroup simulations which have the same $q(T)$? Please clarify this issue or
+add the corresponding discussion to the manuscript.
+
+Response.  This is correct. The optimizations are done for a subgroup of simulations with the same value of $q(T)$, and the value of $q(T)$ has to be further determined to complete the problem.
+For the single-bin case, the latter step is equivalent to the best choice of $t_0$, the new Eq. (23).
+For the multiple-bin case, the optimal value of $q(T)$ is solved from Eq. (55).
+We have added a few comments in the two respective sections to clarify the procedure.
+We thank the referee to point out these ambiguities.
+
+2.How to determine the time of the preliminary adaptive FDS simulations with a constant
+updating magnitude?
+
+3.The study of the updating matrix borrows techniques used in studying transition matrices.
+Does this require all the elements in the updating matrix $w_{ij}$ to be nonnegative? If it is
+true, please mention this requirement after Eq.(28).
+
+Response: The elements in the updating matrix can be negative.
+Fortunately, the balance property, as well as a few ensuing properties, are unaffected by the allowance of the negative numbers.
+For example, the balance property follows directly from
+the detailed balance condition as demonstrated in Eq. (28).
+The existence of a set of orthonormal eigenvector follows from
+the symmetry of the scaled version of w:
+$sqrt(rho_i/rho_j) w_{ij}$.
+We have added a note to clarify this, as well as several one-line derivations in section IIC.1 to show that the quoted results are unaffected by the negative matrix elements.
+Thank you for pointing out this subtle issue.
+
+4. For Eq.(5): perhaps Change $p_i$ to $p_{i(t)}$.
+
+Response: Corrected.  Thank you.
+
+5. Eq. 11 expresses the total error in the cumulative bias potential as a weighted average
+over the errors in each bin, weighted by the density rho, which is the equilibrium
+eigenvector of the updating matrix w. Why is the equilibrium eigenvector of the
+updating matrix, the correct way to weight the errors in the cumulative bias in the bins?
+Can the authors provide a physical explanation of why this is the case, as well as a
+mathematical justification.
+
+Response:
+This is an astute observation, and we also find this setup somewhat awkward.
+While the equilibrium eigenvector rho of matrix w may not be the most straightforward way of weighting errors, it sometimes serves as a mathematically convenient way of doing so.
+By adopting a looser weighting function of the error, we can analyze a larger class of updating matrices with somewhat simpler formulas using properties of transition matrices.
+
+%For example, the matrix
+%     0.6  0.4  0
+%w =  0.2  0.6  0.2
+%     0    0.4  0.6
+%is not symmetric, yet it satisfies detailed balance for the left equilibrium vector (0.25, 0.5, 0.25).
+
+Another way to think about this is the following.
+Instead of thinking of the equilibrium eigenvector of the updating matrix has to be correct way of weighting the errors, we may think of the updating matrix must adjust its elements such that the correct weighting vector is its eigenvector.
+In practice, the vector rho is almost always the uniform vector 1/n,
+which is a fair way of weighting the matrix.
+We have added several comments in sections II.B.1 and II.C.1 to clarify this point.
+Thank you.
+
+\section{Reviewer \#2 Evaluations:}
+
+Recommendation: Major revision
+
+New Potential Energy Surface: No
+
+Reviewer \#2 (Comments to the Author):
+
+The authors present a method of computing the optimal schedule of the updating magnitude for free energy calculations such as Wang-Landau or metadynamics.
+The method maps that optimization problem to a mechanical one, with the update schedule taking the role of a velocity of a free particle and the error becomes the action.
+This an interesting twist leading to new insight in this plaguing problem.
+In general, I would like to see this published at some point, but there are a few major issues which, in my opinion, make it not publishable at this point.
+
+1) There seems to be some misconception about the relation between WL and metadynamics.
+I recommend to adapt the wording in order to give less of an impression that WL and metadynamics are different methods.
+a) The authors agree both methods are "closely related", referring to [9].
+In fact, it has been shown that both methods can be made identical by applying the same schedule of the updating magnitude, see C Junghans, D Perez, T Vogel, Journal of Chemical Theory and Computation (JCTC) 10 (5), 1843 (2014).
+At that point, the authors should also recognize the existence of Statistical Temperature Molecular Dynamics, see J.
+Kim, J.
+E.
+Straub, and T.
+Keyes, Phys.
+Rev.
+Lett.
+97, 050601 (2006); and J.
+Chem.
+Phys.
+126, 135101 (2007).
+b) It is not true that "the difference" [between WL and metadynamics] "lies in the updating window function".
+While a discrete delta-function is often used in practice, WL was introduced saying that any kernel function can be used for updates, and a Gaussian one was explicitly mentioned in the original WL papers.
+c) It is true that it is commonly accepted that the 1/t method is "optimal" for WL, so it should be for metadynamics.
+Given the equivalence of WL and metadynamics, it is not obvious that "the resulting optimal schedule is different in the WL and metadynamics cases." Rephrasing this, for example, into "depends on the updating window function applied in either method" could help to clarify this point.
+d) Given the vast number of publications in the metadynamics community, I would also not agree that "the optimal schedule for metadynamics is less studied."
+
+Response.
+We apologize for the inaccurate wording in framing WL and metadynamics.
+The more precise and technical terms of differentiating the updating schemes, as the referee kindly pointed out, should be the single-bin updating scheme and the Gaussian updating scheme.
+We have corrected the terms in the abstract,
+and added an explicit note in the introduction that the terms "WL" and "metadynamics" are used as pseudonyms in this study in the loose sense to refer the two updating schemes.
+For 1.a), we apologize for ignoring the references
+and we have added the suggested references to the paper.
+For 1.b), we have changed the phrase "the difference" to "a key difference."
+For 1.d), we have deleted the sentence. "the optimal schedule for metadynamics is less studied."
+
+2) The length of the paper, also pointed out by the editor.
+a) I recommend to drastically shorten section IIA, this is too detailed given that it is just a review and
+b) shorten section IIB, since the content is also largely known to the community.
+The authors should focus on sections IID-F, which contains the new developments.
+
+Response.  We agree and we apologize for the length of the manuscript.
+For 2.a), we have shorten section IIA and keeping only what is necessary for introducing the notations and key equations useful for later development.
+For 2.b), similarly, section IIB was indeed too verbose with many unnecessary repetitions. We have now shortened it, and the original sections IIB.3 and IIB.4 are merged.
+The purpose of section IIB is mainly to fix the notations and definitions.
+
+We have added a note in the introduction of section II to explain the introductory nature of the sections so that an expert reader can readily skip it.
+We have slightly expanded section IID-F to explain several ambiguities, and we have included a section on the practical procedure that was originally in the Appendices, as suggested by the reviewer.
+
+3) The choice of an artificial, "trivial" test system to demonstrate the advantage of the proposed method.
+a) The model system is artificial in the sense that "intrinsic distributions are assumed to be flat".
+So, where is the challenge?
+I understand that the theory section of the paper is very general and "generic", but numerical checks should be performed using known, well-studied, and challenging systems.
+The presented numerical checks are not very convincing this way.
+b) "the errors from the simulations agreed well with those from theory". In particular this is not a surprise for a system with intrinsically flat distributions.
+c) In general, all the focus is on minimizing the statistical errors.
+While that is "neat", it is most often not the main issue in a "real world" study and neither critical.
+WL is correctly applied to determine the weights for multicanonical simulations.
+Multicanonical sampling is statistically correct and unacceptably large errors in the weight function, both statistical and systematic, become evident immediately.
+So, minimizing the statistical errors is mostly not a major concern.
+Being able to move through the whole phase space is.
+The real question should rather be: Can this new method and insight help to study systems that were formerly inaccessible to FDS methods?
+This question is not answered.
+To the contrary, the authors themselves admit that their assumptions "may break down for relatively short simulations of glassy systems".
+Did the authors try a glassy system?
+
+Response.
+
+Point 3.a) Add an example.
+
+Point 3.b) This is a subtle and important point on correctly modeling the initial error.
+The challenge lies in the recovery from the error introduced during the equilibration phase.
+Even though the error is zero at the beginning, the equilibration phase with a constant updating magnitude will introduce some error that would become challenging to remove in the production run.
+The single-bin updating scheme does not differentiate the source of error, either systematic or random.
+However, for the multiple-bin updating scheme, like the Gaussian one, the random error built from the equilibration phase will be sufficiently smooth with very little minimal short-wave (local) error, because these local modes are coupled to very small eigenvalues of the updating matrix.
+Even in such case, Gaussian updating scheme will have trouble in removing these short-wave or not-so-short-wave error modes in the long run, because the small eigenvalues coupled to local modes also represent how quickly the errors in these modes (if any) would be removed.
+Thus, if the intrinsic PMF is not smooth enough (or equivalently if the Gaussian window is chosen too wide), the real error would be must larger than those reported in our trivial test system.
+
+Point 3.c)
+We believe that even for hard and glassy problems, the long-time correlation might have already been taken into account, at least partially, by the integrals of the autocorrelation functions defined here.
+Thus, even if the integrals, $Gamma_k$'s may take very large values and difficult to estimate accurately, we can still hopefully use this framework to analyze these challenging problem, at least for long simulations.
+The removal of the systematic error can often by addressed with the usual WL and metadynamics techniques.
+We agree with the reviewer that the failure of multicanonical sampling lies in its lack of responsiveness of the updating scheme to the bias potential.
+The final stages of the WL algorithm suffers from a similar problem of losing responsiveness, where, the updating magnitude is reduced too quickly, making the iterative updating scheme ineffective.
+Our belief is that these issues, even for hard problems, can be addressed in the present framework.
+A practically difficulty may lie in the attainment of the correlation integrals, $Gamma_k$'s, which reflect the difficulty of phase-space sampling.
+Similarly, the Gaussian updating scheme can be inefficient in removing local errors, if any, in the long run without a proper scheduling.
+The method we present here aims at improve the situation in improve long-term convergence of the Gaussian updating scheme.
+However, we do not imagine a dramatic improvement on previous methods for these very challenging problems, which are associated with very long correlation times and likely hard to simulate even with the perfect bias potential.
+Thus, to best simulate these system, new simulation methods.
+
+We have added some discussion in the section IIA to address this issue.
+
+Overall, while points 1) and 2) should be straightforward to revise, point 3) makes me believe that this paper, even though doubtlessly very interesting, will have a limited impact on the community in practice.
+
+\section{Dr. Pettitt's response}
+
+We thank the referees for their thoughtful critique of our work. We have taken their suggestions and have returned a much improved manuscript.
+
+Referee 1: We have worked to better explain our mathematical analysis for flat-distribution-sampling errors from updating schemes.
+We also moved the method of computing the optimal schedule from the appendix to the main text and moved the solution for q(t) to appendix as requested.
+We now include an analysis of the errors in two nontrivial problems, the potential of mean force for a dense Lennard-Jones fluid and the 64x64 10 -tate Potts model.
+
+1) q(t) is integral of the updating magnitude, α(t). We better explain the variational problem and solution of varying q(t) on the interval [0,T]
+
+2) In Appendix D, we discuss a way of estimating Γk from an FDS simulation under constant magnitude.
+
+3) We now discuss the consequences of the sign of the elements of w.
+
+4) Done as suggested.
+
+5) Our choice was one of convenience and we now better discuss and justify the choice. (CZ: We could provide more explanation in the text for this)
+
+Referee 2:
+
+1a) We had made distinctions between WL and metadynamics for convenience in our analysis but we take the referee’s point and emphasize the close relation between the methods to avoid any confusion.
+
+1b) True and we have made the wording clear that we will consider only single bin WL and multi bin metadynamics as practical cases.
+
+1c) We agree. Done
+
+1d) Corrected.
+
+2a) We have greatly shortened section IIA.
+
+2b) We have also shortened IIB as requested.
+
+3a) We have removed the ideal system example and replaced it with a calculation of the PMF in a Lennard-Jones system
+
+3b) Same as above.
+
+3c) We chose the Lennard-Jones fluid as a well know example.
+We also consider the 64x64 10-state Potts model.
+We will use the method in the future on more interesting (difficult) examples.
+Here we give the theory and practical, familiar examples.
+
 \end{document}

### 2017-06-26 doc/invt.tex
absorbing Dr. Pettitt's changes

@@
 \author{B. Montgomery Pettitt}
-\email{bmpettitt@utmb.edu}
+\email{mpettitt@utmb.edu}
 \affiliation{
@@
   the single-bin updating scheme (commonly used in WL)
-  is given by inverse-time formula,
+  is given by an inverse-time formula,
   that for the Gaussian updating scheme (commonly used in metadynamics)
@@
 when the collective variable is energetically restricted to local regions
-of the complex, multimodal free energy surface.
+of the complicated, multimodal free energy surface.
 %
@@
 be observed with traditional simulations.
-%
-%A better solution is to artificially flatten
-%the target distribution along $z$\cite{mezei1987, berg1992, *lee1993,
-%wang2001, wang2001pre,
-%huber1994,
-%*laio2002, *laio2008, *barducci2011, *sutto2012}.
-%%
-%This alleviates the above problem of local trapping
-%by forcing the system to travel more
-%deliberately along $z$,
-%which often represents
-%a direction of slow motion.
-%%
-%To achieve a flat distribution, however,
-%the method must be able to construct a
-%bias potential that cancels the PMF.
 
@@
 %
-%flat-distribution-sampling (FDS) techniques
-%that build the bias potential adaptively or on the fly.
-%
 These techniques regularly update the bias potential
 to discourage future visits to previously sampled configurations
-by incrementally elevating the bias potential,
-%and they are closely-related\cite{junghans2014, micheletti2004}.
-%
-%The bias potential is associated with discrete values of $z$.
+by incrementally elevating the bias potential.
 %
@@
 %
-Usually, the window function
+Often, the window function
 takes the form of a discrete
@@
 %
-%is that in the WL case,
-%each updating step is limited to the bin
-%containing the current $z$,
-%while metadynamics adopts an extended
-%Gaussian window
-%covering several neighboring bins.
-%%
-%The latter is more often used in MD simulations
-%as it guarantees a smooth profile
-%of the bias potential
-%suitable for numerical differentiation
-%in deriving the bias force.
-%
 In both cases,
 regular updates to the bias potential
-%in both cases, the bias potential
-%is updated frequently (usually every few MC or MD steps),
-%and the fluctuation
 disrupts the underlying equilibrium
@@
 %the WL algorithm\cite{langfeld2012, pellegrini2014}.
-%
-%The optimal schedule for metadynamics,
-%however, is less studied.
 
@@
 of a general updating scheme,
-including the single-bin and Gaussian one
-used in the WL algorithm and metadynamics.
+including the single-bin and Gaussian versions.
 %
@@
 The optimal schedule of bandpass updating schemes
-is also the inverse-time schedule.
+is also an inverse-time schedule.
 %
@@
 
-%For finite-length simulations, however,
-%metadynamics may deliver better performance.
-%
-%These results may help to better
-%plan and optimize FDS simulations.
-
 The article is organized as follows.
@@
 We present the analytical results in Sec. \ref{sec:theory},
-numerically verify some aspects
+provide numerical examples
 in Sec. \ref{sec:results},
@@
 In Sec. \ref{sec:background},
-we first review the basics of FDS
+we briefly review the basics of FDS
 and some known aspects of the optimal schedule.
@@
 Then, in Sec. \ref{sec:single-bin},
-we illustrate the method of
+we derive the method of
 computing the optimal schedule
@@
 %
-%The above two introductory sections help define our notations,
-%and an expert reader may wish to skip them.
-%
 Next, we extend the method to the general case
@@
 %which is often the case in a (classical) molecular system,
-we can discretize it
-such that the integer $i$ represents
+we can use the equivalent integer $i$ to represent
 the index of a small interval, or a bin,
@@
 as a biased simulation
-targeting a flat $\rho_i$ (i.e. $\rho_i \equiv 1/n$)
+targeting a flat $\rho_i$ (i.e. $\rho_i \equiv 1/n$ for $n$ bins)
 or nearly flat one\cite{
@@
 %
-and our aim is to find a biased potential
+and our aim is to find a bias potential
 such that the resulting $\pi_i$ would coincide
@@
 The Gaussian updating scheme used in metadynamics
-is an example, with $w_{ij}$ being a Gaussian of $i-j$.
+is an example, with $w_{ij}$ being a Gaussian centered on $j$.
 %
@@
 since the shift
-remains a constant during the course of simulation,
+remains a constant during the course of simulation.
 %%
@@
 %
-We define a bin-average-deducted bias potential
-as\note{Alternatively,
+We define the bias potential deviation from the
+bin average, $\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$,
+as
+%
+\begin{equation}
+  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
+  .
+\label{eq:x_def}
+\end{equation}
+%
+\note{Alternatively,
   we may define $v_{*i}(t)$ such that
@@
   One can readily show that the two definitions are equivalent
-for small $v_{*i}$.}
-%
-\begin{equation}
-  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
-  ,
-\label{eq:x_def}
-\end{equation}
-%
-where we have defined the bin average,
-$\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$.
-%
-%\begin{equation}
-%  \bav{v}(t)
-%  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
-%  .
-%  \label{eq:vav_def}
-%\end{equation}
+  for small $v_{*i}$.}%
 %
@@
 %
-The total error is given by
+The total error is then given by
 %
@@
 a deterministic averaged part, $\langle f_i(t) \rangle$,
-and a random fluctuation part, $\zeta_i(t)$,
+and a random fluctuating part, $\zeta_i(t)$,
 %
@@
 We will approximate the histogram fluctuation part, $\zeta_i(t)$,
-hence its bin-average-deducted version, $\zeta_{*i}(t)$,
+hence its deviation from the bin average, $\zeta_{*i}(t)$,
 as white noise such that
@@
 for some $G_i$,
-and $\delta(t)$ is the Dirac's delta function.
+and $\delta(t)$ is the Dirac delta function.
 %
@@
 the total error defined in Eq. \eqref{eq:error_def}
-is a sum of a residual error (from the decay of the initial error)
-and an asymptotic error:
+is a sum of a residual error, $E_R$ (from the decay of the initial error)
+and an asymptotic error, $E_A$:
 %
@@
 of the bin-average deduction on $\tilde x_k$ is to
-annihilate the first mode, $\tilde x_0$.
-%
-It follows Parseval's theorem,
+annihilate the $k=0$ mode.
+%
+We also have Parseval's theorem,
 \begin{equation}
@@
 %
-Thus, the optimization problem on the final error
+Thus, the optimization problem of the final error
 becomes more complicated and it often
@@
 %
-Besides, the Hamiltonian
-coincides with the Lagrangian:
+Here, the Hamiltonian coincides with the Lagrangian:
 %
@@
 does not explicitly depend on time, $t$,
-the Hamiltonian is conserved,\cite{}
+the Hamiltonian is conserved,
 which means that the asymptotic error grows
@@
 and $t^* \equiv t/T$
-such that both are limited in $[0, 1]$,
+such that both are limited to the interval $[0, 1]$,
 as well as
@@
 The resulting values of $(t_{[i]}, \alpha_{[i]})$
-  can be linear interpolated
+  can be linearly interpolated
   to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.
@@
 %
-and if the eigenvalues are set to $1$,
+if the eigenvalues are set to $1$,
 we recover the single-bin scheme.
@@
 The optimal schedule of these schemes
-remains to be the inverse-time schedule
+remains the inverse-time schedule
 given by Eq. \eqref{eq:alpha_invt1}
@@
 
-The use of polynomial basis makes
+The use of a polynomial basis makes
 the bias potential smooth and differentiable,
@@
 We can nonetheless sample a wider range
-by constructing a mixture or expanded ensemble
+by constructing a mixture\cite{shirts2017}
+or an expanded ensemble
 from a superposition of multiple Gaussian umbrellas
@@
 *kumar1992, *roux1995, *bartels1997, *souaille2001,
-*kobrak2003, *gallicchio2005, *chodera2007, zhu2012, newman, frenkel,
-fenwick2008, *kim2011}
+*kobrak2003, *gallicchio2005, *chodera2007,
+*fenwick2008, *kim2011, *shirts2008,
+zhu2012, newman, frenkel},
 can be used to recover the unbiased distribution
@@
     & $1.0(0.9)$
-    & $9.7(8.0)\times10^{-3}$
+    & $9.9(8.0)\times10^{-3}$
     \\
@@
 became too small to damp out the initial bias,
-resulting in much larger initial updating magnitude, $\alpha(0)$
+resulting in a much larger initial updating magnitude, $\alpha(0)$
 [cf. Eq. \eqref{eq:opt_alpha0}].
@@
 
-As a proof of principle,
+As another proof of principle,
 we tested the adaptive Gaussian umbrella sampling method
@@
 %
-By WHAM, we found the critical temperature at {\color{red}$0.70162$}
+By WHAM, we found the critical temperature at $0.70162$
 in agreement with a previous study\cite{wang2001pre}.
@@
 Although this assumption appeared to work well
-on our test system for modeling long simulations,
-they may break down for relatively short simulations
+on our test systems modeling long simulations,
+it may break down for relatively short simulations
 of glassy systems.
@@
 Finally, the adaptive Gaussian umbrella sampling
-approach still needs some refinement to improve
+approach requires refinement to improve
 its stability for general application.
@@
 We find that with slow sampling processes,
-the updating scheme may occationally overly update
+the updating scheme may occasionally overly update
 the parameter, $c_2$, leading to unreasonable values.
@@
 %
-For bandpass updating scheme,
+For the bandpass updating scheme,
 we may borrow this criterion
@@
 since the histogram flattening
-is only partially applied to modes less $K$.
+is only partially applied to modes less than $K$.
 %
@@
 the autocorrelation functions of the eigenmodes.
-%
-The basic idea is to use Eq. \eqref{eq:xt2_eql},
-and to avoid the problem of not
+
+The basic idea is to use Eq. \eqref{eq:xt2_eql};
+to avoid the problem of not
 knowing the intrinsic distribution, $p_i$,
@@
 by the time average of the bias potential.
-
+%
 We first note that the difference,

### 2017-06-26 doc/invt.tex
small text changes

@@
   %
-  Several previous flat-distribution-sampling algorithms
-  can be cast as special cases of the bandpass updating schemes.
+  Constructed from orthogonal polynomials,
+  bandpass updating schemes serve as generalizations
+  of the Langfeld-Lucini-Rago algorithm
+  and can be used for parameter tuning for the method of Gaussian ensemble
+  and umbrella sampling.
   %
@@
 sampling of slow transitions that otherwise would not readily
-be observed with traditional simulations.
+be observed in traditional simulations.
 
@@
   *laio2002, *laio2008, *barducci2011, *sutto2012, micheletti2004}
-among others\cite{kim2006, *kim2007, junghans2014,
-  langfeld2012, pellegrini2014}.
+among others\cite{kim2006, *kim2007, kim2010, junghans2014,
+  langfeld2012, pellegrini2014,
+  maragliano2006, *abrams2008,
+  zheng2010}.
 %
@@
 or that of a Gaussian
-in metadynamics.
+in metadynamics.\cite{junghans2014}
 %
@@
 %
-The latter, however,
-allow one to partially flatten
+The latter allow one to partially flatten
 a few long-range histogram modes,
@@
 %
-A previous FDS algorithm\cite{langfeld2012, pellegrini2014}
-can be cast as a special case of the bandpass updating scheme,
-and a similarly application can help other FDS
+Constructed from orthogonal polynomials,
+bandpass updating schemes serve as
+generalizations of the Langfeld-Lucini-Rago (LLR)
+algorithm\cite{langfeld2012, pellegrini2014},
+and can help other FDS
 algorithms\cite{neuhaus2006, *neuhaus2007, zhu2012}
@@
 
-%For a large system at temperatures of interest,
-%the distribution, $p_i$, often tends to
-%be localized,
-%%
-%and to explore the global shape
-%of the PMF, $-\ln p_i$,
-%it is often advantageous to perform
-%biased sampling that targets
-%a wider and smoother distribution, $\rho_i$.
-%%
-%%Here, we refer to simulations that target
-%%a flat or nearly flat distribution
-%%as entropic or multicanonical sampling.
-%
-%To do so, we introduce a bias potential, $u_i$,
 The difficulty of sampling a multimodal distribution
@@
 %
-%
-%where the proportionality constant, $C_i$,
-%is found from the normalization condition,
-%$\sum_{i = 1}^n \pi_i = 1$,
-%in the second step.
-%
-%In particular,
-%to make $\pi_i$ flat,
-%the bias potential $u_i$
-%must coincide with $\ln p_i$,
-%up to an additive constant.
-%%
-%In practice, $u_i$ may be non-optimal or contain error,
-%and thus $\pi_i$ need not
-%coincide with the intended distribution $\rho_i$.
 %
@@
 targeting a flat $\rho_i$ (i.e. $\rho_i \equiv 1/n$ for $n$ bins)
-or nearly flat one\cite{
+or a nearly flat one\cite{
 dayal2004, *trebst2004, barducci2008, singh2011}.
 %
-By introducing a bias potential, $u_i$,
-the biased equilibrium distribution becomes
+The biased equilibrium distribution
+under a bias potential, $u_i$, is given by
 %
@@
 
-An other class of adaptive FDS methods,
+Below we will focus on another class of adaptive FDS methods,
 including the WL algorithm\cite{wang2001, wang2001pre}
@@
 address the above shortcoming
-by continuously and incrementally updating the bias potential,
-and we will focus on these methods below.
-%%
-%%These methods incrementally update the bias potential
-%%and can approximately sample the target distribution.
-%%
-%However,
-%the on-the-fly updates break the microscopic reversibility
-%of the underlying sampling mechanism.
-%Further, failure to properly reduce the magnitude of
-%the histogram-based increments
-%would lead to roughness or errors in the final estimate of the PMF.
-%%
-%%Thus, for a conservative practitioner,
-%%the E-FDS methods appear more appealing
-%%once a sufficiently accurate bias potential is available.
-%%
-%To minimize this effect,
-%many adaptive methods adopt a schedule to
-%reduce the updating magnitude
-%over the course of simulation\cite{
-%marsili2006,
-%belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-%liang2007,
-%barducci2008}.
-%%
-%%Thus, in late stages of a long simulation,
-%%the updating becomes negligible and
-%%the adaptive simulation tends to
-%%the corresponding equilibrium one.
-%%
-%%if the bias potential were converged to the exact value,
-%%this adaptive FDS tends to the ideal equilibrium counterpart,
-%%particular equilibrium simulation can be thought as
-%%the optimal asymptotic limit of the adaptive counterpart,
-%%
-%For example, one can show that
-%if the magnitude is controlled by
-%the optimal inverse-time prescription\cite{
-%belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-%liang2007,
-%morozov2007, zhou2008,
-%komura2012, *caparica2012, *caparica2014},
-%the WL algorithm is equivalent to the ideal equilibrium counterpart
-%in improving the precision of the bias potential,
-%thus allowing for maximum efficiency
-%(cf. Appendix \ref{sec:equilerr}).
-%%the bias potential in the WL algorithm
-%%will evolve similar to Eq. \eqref{eq:vcorr_equil}
-%%with the precision of the histogram as simulation lengthens.
-%
+by continuously and incrementally updating the bias potential.
 %
 %\subsubsection{Updating schemes}
-%
-%
-%
-%We will mainly focus on the adaptive methods below.
 %
@@
 the bias potential, $u_i(t)$, is updated
-at each sampling step $t$ as
+at each step $t$ as
 %
@@
 and introduce artificial errors that need to be
-removed by gradually decreasing the updating magnitude\cite{
+reduced by gradually decreasing the updating magnitude\cite{
   belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
@@
   crespo2010, *atchade2011, *fort2015}.
-%One can show that under constant magnitude,
-%$\alpha(t) = \alpha > 0$,
-%the distribution collected from
-%an adaptive FDS simulation
-%is identical to $\rho_i$ in the long time or asymptotic limit.
-%
-%However, as the above updates
-%change $u_i(t)$ continuously,
-%the equilibrium distribution,
-%Eq. \eqref{eq:pi_p_V}, no longer holds
-%at all times,
-%%
-%The source of the deviation is two-fold.
-%%
-%First, since $u_i(t)$ is updated continuously,
-%there is a random noise that is proportional
-%to $\sqrt \alpha$.
-%%
-%Second, there is a systematic error
-%that comes from the updating dynamics itself,
-%as it breaks the Markovian nature
-%that underlies Eq. \eqref{eq:pi_p_v}.
-%%
 
@@
 %
-This can be derived from a general result\cite{
+This general result\cite{
   robbins1951, pellegrini2014}
-and an intuitive explanation\cite{
+has an intuitive explanation\cite{
   marsili2006, barducci2008}
-comes from the interpretation of
-the bias potential %under the WL updating scheme, Eq. \eqref{eq:wl_update},
+that interprets the bias potential
 as a runtime average computed from all data points collected so far,
-so that the addition of a new data point only
-affects the average by its due weight,
+which absorbs the contribution of the newest data point
+by its due weight,
 the inverse of the current sample size, $t$
@@
 since the shift
-remains a constant during the course of simulation.
+remains a constant during simulation.
 %%
@@
 Deducting the bin average and using Eq. \eqref{eq:fav1}, we get
+$\langle f_{*i}(t) \rangle \equiv \langle f_i(t) \rangle - 1 \approx - v_{*i}(t)$,
+and
 %
 \begin{align}
-  \langle f_{*i}(t) \rangle
-  \equiv
-  \langle f_i(t) \rangle - 1
+  f_{*i}(t)
   &\approx
   - v_{*i}(t)
+  +
+  \zeta_i(t)
   .
@@
 much longer than the autocorrelation time.
-%
-%For non-white noise,
-%the equivalent value of $G_i$ can be approximately obtained as
-%$\sum_{t = -\infty}^\infty \langle \zeta_{*i}(t) \, \zeta_{*i}(0) \rangle$.
 
 From Eqs.
-\eqref{eq:vt_diffeq}-\eqref{eq:sh_ave},
+\eqref{eq:vt_diffeq} and \eqref{eq:sh_ave},
 we get % a set of decoupled equations
@@
   \dot v_{*i}(t)
-  =
-  \alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \zeta_{*i}(t) \right]
+  %=
+  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \zeta_{*i}(t) \right]
   \approx
@@
 then after a long period,
-the error approach an equilibrium
+the error approach an equilibrium value
 %
@@
 %
-This is equivalent to varying the curve $\vartheta(q(t))$
+This is equivalent to varying the curve,
+$\vartheta\bigl( q(t) \bigr)$,
 with fixed endpoints,
@@
 %
-Note also that the residual error
+Since the residual error
 is fixed during the process,
-while the asymptotic error is the action of a free particle.
-Thus, the velocity of $\vartheta(q(t))$ should be a constant:
+the asymptotic error,
+resembling the action of a free particle,
+demands a constant velocity
 %
@@
 %
-Below we will take advantage of the analogy
-and borrow techniques used
-in studying transition matrices\cite{vankampen}
-(we will only use properties that are unaffected
-by the admission of negative elements into $\mathbf w$).
+Below we will %take advantage of the analogy and
+borrow techniques used
+in studying transition matrices\cite{vankampen},
+but limit ourselves to properties that are unaffected
+by the admission of negative elements into $\mathbf w$.
 
@@
 %
-We will label this eigenvector as the first one,
+We will label this eigenvector by $k = 0$,
 and
@@
 %
-Using this definition,
+Then,
 we can rewrite Eq. \eqref{eq:ortho0} as
@@
 %
-where in the second step,
+where
 we have used Eq. \eqref{eq:xtstar} and
 the transformed version of Eq. \eqref{eq:sh_ave}
-by assuming the linear approximation
-as in Sec. \ref{sec:sbin_diffeq}.
+in the second step.
 
@@
 becomes more complicated and it often
-involves a compromise of the individual modes.
+involves a compromise among eigenmodes.
 
@@
 %
-Equation \eqref{eq:Gamma_sum} provides a way
-of approximately modeling
+Equation \eqref{eq:Gamma_sum} represents
+a first approximation for modeling
 a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
@@
 we discuss a way of estimating $\Gamma_k$
-from an FDS simulation under constant magnitude.
+from an FDS simulation under a constant magnitude.
 %
@@
 If the initial bias potential
-before the priliminary run is zero,
+before the preliminary run is zero,
 one can show that
@@
 does not explicitly depend on time, $t$,
-the Hamiltonian is conserved,
+the Hamiltonian is conserved,\cite{goldstein, *landau_mechanics, *arnold}
 which means that the asymptotic error grows
-linearly with time at a rate of $\mathcal L$,
-and we may set
+linearly with time at a rate of $\mathcal L$.
+We may set
 %
@@
 %
-for some positive $C_M$.
-%
-This serves as a generalization of Eq. \eqref{eq:dthetadt_const}.
+for some positive $C_M$
+as a generalization of Eq. \eqref{eq:dthetadt_const}.
 %
@@
 %
-Thus, we have an implicit equation for $q(t)$,
+Thus, Eq. \eqref{eq:Lagrangian_const}
+gives an implicit equation for $q(t)$,
 and hence the optimal schedule,
@@
 \begin{center}
-  \begin{tikzpicture}[thick, scale=1.25, every node/.style={scale=0.8}]
+  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
     \node [] at (-0.3, 2.0) {(a)};
+
     % distribution
-    \draw[domain=0:2.5, variable=\x, dashed, gray, very thick]
+    \draw[domain=0:2.0, variable=\x, dashed, gray, very thick]
       plot({\x}, {2*exp(-\x)});
@@
     % shade
-    \fill [gray!30!white, domain=0.7:2.2, variable=\x]
+    \fill [gray!30!white, domain=0.7:1.609, variable=\x]
       (0.7, 0) --
       plot ({\x}, {2*exp(-\x)})
-      -- (2.2, 0) -- cycle;
+      -- (1.609, 0) -- cycle;
 
     % thick curve
-    \draw[domain=0:2.2, variable=\x, very thick]
+    \draw[domain=0:1.609, variable=\x, very thick]
       plot({\x}, {2*exp(-\x)});
@@
     % label for the shaded area
-    \node[] at (1.1, 0.36) {$\dfrac{t}{T}$};
+    \node[] at (1.1, 0.32) {$\dfrac{t}{T}$};
 
@@
     \draw[] (0.5, 0.99317) -- (0.7, 0.99317);
-    \draw[->] (0, 0) -- (2.8, 0);
-    \node[] at (2.8, -0.2) {$\bar q$};
+    \draw[->] (0, 0) -- (2.2, 0);
+    \node[] at (2.2, -0.2) {$\bar q$};
     \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 2.2);
@@
     \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$q(T) - q(t)$};
-    \draw[] (2.2, 0.1) -- (2.2, 0.0) node[below] {$q(T)$};
+    \draw[] (1.609, 0.1) -- (1.609, 0.0) node[below] {$q(T)$};
   \end{tikzpicture}
 %
-  \begin{tikzpicture}[thick, scale=1.25, every node/.style={scale=0.8}]
+  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
     \node [] at (-0.3, 2.0) {(b)};
@@
     % label for the shaded area
-    \node[] at (1.1, 0.40) {$\bar q(t)$};
+    \node[] at (1.1, 0.35) {$\bar q(t)$};
 
@@
 
-We may consider the schedule, $\alpha(t)$,
-as a variable transformation of
+The schedule, $\alpha(t)$,
+also defines a variable transformation from
 the mass distribution, $m(\bar q)$.
@@
 For example, in the single-bin scheme,
-the mass distribution is a single exponential decay,
+the mass distribution decays exponentially,
 and we would have the inverse-time schedule.
@@
 at large $\bar q$,\footnote{If
-  the magnitude of the eigenvalues, $\lambda_k$'s,
+  the magnitudes of the eigenvalues, $\lambda_k$'s,
   span a wide range, then for a large $\bar q$,
@@
 We will consider simulations in which
-the system is initially equilibrated at a
-constant updating magnitude, $a_0$.
+the system is initially equilibrated
+at a constant updating magnitude, $a_0$.
 %
@@
 %
-For these translationally invariant or homogeneous
-schemes, the eigenvalues can be readily found
+The eigenvalues of these translationally invariant or homogeneous
+schemes can be readily deduced
 from the window function ($\mu_0, \mu_1, \dots, \mu_b$)
@@
 where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
-are the width of the Gaussian and span
+are the width of the Gaussian and the span
 in the original collective variable, $z$,
@@
 Basically, we need to extract
-$\lambda_k$'s and $\Gamma_k$'s
+$\lambda_k$'s, $\Gamma_k$'s, and $\epsilon_k$'s
 from the given system and updating scheme (input),
@@
   \node[draw, text width={width("sampling")}]
-    (sys) at (0, 1) {System\\and\\sampling\\process};
+    (sys) at (0, 1.1) {System\\and\\sampling\\process};
 
@@
   \node[draw, text width={width("Integrals, iii")}]
-    (Gamma) at (2.1, 1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$};
+    (Gamma) at (2.1, 1.1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$\\and biases,\\$\epsilon_1, \dots, \epsilon_{n-1}$};
 
@@
   \label{fig:vardep}
-  Dependence of variables in computing
-  the optimal schedule and the error.
+  Computation of the optimal schedule and the error.
 }
@@
 Run a preliminary adaptive FDS simulation
-under constant updating magnitude, $a_0$.
+under a constant updating magnitude, $a_0$.
 %
@@
   by Eq. \eqref{eq:uhatav},
-  although it can overestimate
+  although it leads to overestimated
   $\epsilon_k$ for large-$k$ modes.
@@
 in computing its change.
+The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
 }
-The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
 %
@@
 To handle Gaussian updating schemes more effectively,
-  we have chosen the $\mathcal N+1$ grid points over $[0, q(T)]$ to be given by
+  we have chosen the $\mathcal N+1$ grid points
+  ($\mathcal N=20000$ here)
+  over $[0, q(T)]$ to be given by
   $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
-  where $i = 0, \dots, \mathcal N$,\footnote{For
+  where $i = 0, \dots, \mathcal N$.\footnote{For
     a mass distribution, $m(\bar q)$, that is
@@
     roughly a constant.}
-  and we chose $\mathcal N=20000$.
 %
@@
 Since an updating matrix
-with identical eigenvalues\footnote{Since
+with identical eigenvalues
+is a multiple of the identity matrix,\footnote{Since
   the eigenmode of $k = 0$ represents
@@
   the updating scheme.}
-is a multiple of the identity matrix,
-%
-if the eigenvalues are set to $1$,
-we recover the single-bin scheme.
-%
-This means that the single-bin scheme is most efficient
+we recover the single-bin scheme
+by setting the eigenvalues to $1$.
+%
+Thus, the single-bin scheme is most efficient
 in error reduction in the asymptotic limit.
@@
 the asymptotic error being
-$\Err_A(T) = T / (T + t_0)^2 \sum_{ k = 1 }^{n-1} \Gamma_k$,
+$\Err_A(T) = \sum_{ k = 1 }^{n-1} \Gamma_k \, T / (T + t_0)^2$,
 %
@@
 %
-which is identical to Eq. \eqref{eq:EA_sbin1}
-if we notice [cf. Eq. \eqref{eq:parseval}]
-%
-%where we have used
-%
-\begin{align}
-  \sum_{k=1}^{n-1} \Gamma_k
-  &=
-  \sum_{t=-\infty}^\infty \sum_{k=1}^{n-1}
-  \bigl\langle {\tilde \zeta}_k(t) \, {\tilde \zeta}_k(0) \bigr\rangle
-  \notag
-  \\
-  &=
-  \sum_{t=-\infty}^\infty \sum_{i=1}^n
-  \rho_i \bigl\langle \zeta_{*i}(t) \, \zeta_{*i}(0) \bigr\rangle
-  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
-  .
-  \notag
-  %\label{eq:Gamma_G_sum}
-\end{align}
-%%
+which recovers Eq. \eqref{eq:EA_sbin1}.\footnote{By
+Eq. \eqref{eq:parseval}, we have
+$\sum_{k=1}^{n-1} \bigl\langle {\tilde \zeta}_k(t) \, {\tilde \zeta}_k(0) \bigr\rangle
+= \sum_{i=1}^n \rho_i \bigl\langle \zeta_{*i}(t) \, \zeta_{*i}(0) \bigr\rangle$,
+and by summing $t$ from $-\infty$ to $\infty$, we get
+$\sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma$.}
 
@@
 the single-bin updating scheme
-followed by a perfect filter
+accompanied by a perfect filter
 for the designated noise modes,
-and zero eigenvalues are assigned to these modes.
+which are assigned with zero eigenvalues.
 %
@@
 the difference between the average of $z$
-and the center $z_c$.
-%
-We thus recover the Langfeld-Lucini-Rago (LLR)
+and the center, $z_c$.
+%
+We thus recover the LLR
 algorithm\cite{langfeld2012},
@@
 %
-Note that quadratic bias potentials are commonly employed
-in umbrella sampling (see e.g. Ref. \cite{zhu2012})
+Note that quadratic bias potentials are commonly
+employed\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007}
+in umbrella sampling\cite{torrie1974, *torrie1977}
 and the method of Gaussian ensemble\cite{hetherington1987,
-*challa1988, *costeniuc2006, neuhaus2006, *neuhaus2007},
+*challa1988, *costeniuc2006},
 which is helpful in overcoming barriers
@@
 the bias potential smooth and differentiable,
-and hence suitable for MD applications.\cite{zhu2012}
+and hence suitable for MD applications.
 %
@@
 parallel\cite{swendsen1986,
-  *geyer1991, *hukushima1996, *hansmann1997, *sugita1999, *earl2005}
+  *geyer1991, *hukushima1996, *hansmann1997, *sugita1999,
+  *earl2005, *zuckerman2011, *rauscher2009, kim2010}
 and/or simulated tempering\cite{marinari1992,
@@
 can be readily derived in the former
-framework\cite{neuhaus2006, *neuhaus2007},
+framework\cite{neuhaus2006, *neuhaus2007, kim2010},
 the transition probability in the latter case
@@
 of the Gaussian ensemble
-$Z \equiv \int e^{-\beta \, \mathcal U(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$,
-where
+$Z \equiv \int e^{-\beta \, \mathcal U(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$
+as the bias potential for inter-umbrella transitions.
+%
+Here,
 $\beta$ is the inverse temperature,
@@
   \frac{Z^{(i)}}{Z^{(j)}}
-  \exp\left[u^{(i)}\bigl( z(t) \bigr)
-   - u^{(j)}\bigl( z(t) \bigr) \right]
+  e^{
+    u^{(i)}\bigl( z(t), t \bigr)
+    - u^{(j)}\bigl( z(t), t \bigr)
+    }
   \right\}
@@
 redefined by Eq. \eqref{eq:maxFk}
-(including the intra-umbrella and inter-umbrella
-components).
-%
-The weighted histogram analysis method
+to encompass both intra-umbrella and inter-umbrella
+components.
+%
+After simulation,
+the weighted histogram analysis method
 (WHAM)\cite{ferrenberg1988, *ferrenberg1989,
@@
 zhu2012, newman, frenkel},
-can be used to recover the unbiased distribution
-after simulation.
+can be used to recover the unbiased distribution.
 %
@@
 \subsection{\label{sec:lj}
-Comparison of updating schemes on a Lennard-Jones fluid}
-
-Our test system is a Lennard-Jones (LJ) fluid of $100$ particles,
-and we wish to compute the PMF along the distance of two special particles, $r$.
+Comparison of updating schemes}
+
+Our first test system is
+a Lennard-Jones (LJ) fluid of $100$ particles,
+and we wish to compute the PMF along
+the distance, $r$, between two special particles.
 %
 The two special particles are mutually non-interacting
-but they interact normally with other particles,\footnote{They
-were placed on an axis parallel to a side of the cubic box}
+but they interact normally with other particles,\footnote{The
+two special particles were placed on an axis parallel to a side of the cubic box}
 and the intrinsic distribution, $p(r)$, is
@@
 %
-In each trial, the displacement along each dimension is uniform,
-with the size given by $0.1/\varrho^*$
-to ensure the acceptance ratio is around 50\%,
-where $\varrho^*$ is the reduced number density.
+In each trial,
+a random particle is displaced in each dimension
+according to a uniform distribution over $(-\varepsilon^*, \varepsilon^*)$.
+We used $\varepsilon^* = 0.1/\varrho^*$
+to ensure reasonable acceptance ratio,
+and $\varrho^*$ is the reduced number density.
 
@@
 Each simulation started with an ``equilibration'' phase
-of $10^7$ steps under constant updating magnitude,
+of $10^7$ steps under a constant updating magnitude,
 $a_0 = 10^{-4}$,\footnote{The equivalent $\ln f = n \, a_0$ for the WL algorithm
@@
 To make sure the same schedule for multiple runs under the same condition,
-the optimal schedule was pre-computed from a set of fixed values of
-$\Gamma_k$ and $\epsilon_k$,
-although they can be fairly accurately estimated from the data
-collected from the equilibration phase of each individual run.
+the optimal schedule was pre-computed from a set of fixed
+$\Gamma_k$'s and $\epsilon_k$'s,
+although ideally they should be estimated from the data
+collected from the equilibration phase in individual runs.
 
@@
 and the errors roughly matched the predicted values
-calculated from the pre-computed
+calculated using the pre-computed
 $\Gamma_k$'s and $\epsilon_k$'s.
@@
 The effectiveness of the Gaussian updating scheme
-in the low density case
+in the low-density case
 was probably because the sampling along
@@
 and thus the histogram fluctuation largely
-resulted from local noise.
+came from local noise.
 %
@@
     denotes the error
-    at the end of the equilibration phase under constant magnitude;
+    at the end of the equilibration phase under a constant magnitude;
     $\Err_\mathrm{final}$
@@
 [cf. Eq. \eqref{eq:vt_diffeq_mbin}],
-which can decrease rapidly with the mode index $k$,
+which decreases rapidly with the mode index $k$,
 cf. Eq. \eqref{eq:lambda_Gaussian}.
@@
 %
-Although the difference was small in the reported cases,
-the gain of using the optimal schedule
-would increase with the simulation length.\footnote{
+Although the gain of using the optimal schedule
+was modest in the reported cases,
+it would increase with the simulation length.\footnote{
   For example, for the $\varrho^* = 0.1$ case
@@
 Since we have chosen a long equilibration period
-for the preliminary run under constant updating magnitude,
+for the preliminary run under the constant updating magnitude, $a_0$,
 the initial magnitude was around
@@
 and one may have to either decrease the width
-or to elongate the preliminary simulation under constant magnitude.
+or to elongate the preliminary simulation
+under the constant magnitude.
 
@@
 %
+Initially, we set $c_1 = 1.4 \sigma_E$, and $c_2 = 0$.
+%
 For configuration sampling in each umbrella,
@@
 and it demonstrated
-the backbending behavior\cite{kim2006, *kim2007}
+the backbending behavior\cite{kim2006, *kim2007, kim2010}
 typical in a first-order phase transition,
@@
 (or the density of states here)
-is locally convex or unstable,
+is locally convex or unstable,\cite{kim2010}
 which was so in the critical region,
@@
 of the initial bias, if any,
-in the middle-range eigenmodes.
+in the short- and middle-range eigenmodes.
 %
@@
 which sometimes needs to be fixed at a maximal acceptable level
-for a long period before any reduction can happen.
+for a long time before any reduction can happen.
 %
@@
 the updating scheme became a generalization
-of the LLR algorithm\cite{langfeld2012, *pellegrini2014},
+of the LLR algorithm\cite{langfeld2012, pellegrini2014},
 and it could be used for parameter adjustment
-in umbrella sampling with quadratic
-bias potentials\cite{neuhaus2006, *neuhaus2007, zhu2012}.
-
-The analysis in this study has several limitations.
+in umbrella sampling with
+quadratic\cite{neuhaus2006, *neuhaus2007, zhu2012}
+or similar\cite{kim2010}
+bias potentials.
+
+This study has several limitations.
 %
@@
 Although this assumption appeared to work well
-on our test systems modeling long simulations,
+in our test systems modeling long simulations,
 it may break down for relatively short simulations
@@
 %
-We find that with slow sampling processes,
-the updating scheme may occasionally overly update
-the parameter, $c_2$, leading to unreasonable values.
+For example, we find that with slow sampling processes,
+the updating scheme might occasionally overly update
+the parameter, $c_2$, leading to unreasonable values,
+and the parameters used here might not be optimal.
 %Our immediate interest is to apply these relations to
@@
 %
-JM thanks support from the National Institutes of Health (R01-GM067801, R01-GM116280),
+JM thanks support from the National Institutes of Health
+(R01-GM067801, R01-GM116280),
 the Welch Foundation (Q-1512),
@@
 %
-This research used computing resources of the National Science Foundation XSEDE grid.
-%
+This research used computing resources of
+the National Science Foundation XSEDE grid.
 %
@@
 to coincide with a runtime average
-of the independent corrections.
+of independent corrections.
 
@@
 %
-Below we shall use capital letters to denote
+Below we will use capital letters to denote
 time averages, e.g. for variable $x$, we define
@@
 %Equation \eqref{eq:uhatav}
-This equation serves as
+This equation can also be derived\footnote{From Eq. \eqref{eq:uhatt},
+  we can construct an independent estimator
+  of the unbiased distribution as
+  $\hat p_i(t) \approx h_i(t) \, e^{ u_i(t) }$.
+  Upon time averaging, we get
+  $\rho_i e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
+from the self-healing umbrella sampling
+method\cite{marsili2006},
+and it serves as
 a generalization of Eq. \eqref{eq:vcorr_equil}
@@
 as runtime averages,
-but with dynamic weights.
+but with a dynamic weight.
 %
@@
 %
-For a plain average, $\omega(t) = 1$, for $t > 0$,
-the updating magnitude would follow
-the inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
+The inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
+follows from the plain average,
+$\omega(t) = 1$ (for $t > 0$),
 and updating magnitudes above and below this value
@@
 the bias potential $\tilde v_k(t)$
-as $x(t)$,
+as $X(t-1)$,
 $\tilde f_k(t)$ as $\Delta x(t)$
@@
 with
-$\Omega(t) = \prod_{\tau=1}^t [1- \alpha^*(\tau)]^{-1}
-\approx \exp\left(\int_0^t \alpha^*(\tau) \, d\tau \right)$.
-%
+\begin{align*}
+  \frac{ \Omega(t) } { \omega(0) }
+  = \prod_{\tau=1}^t \frac{1}{1- \alpha^*(\tau)}
+\approx \exp\left(\int_0^t \alpha^*(\tau) \, d\tau \right).
+\end{align*}
 
@@
 The relation between error and histogram fluctuation
-supports the WL criterion for
-checking convergence at stage transitions.
-%
-For the bandpass updating scheme,
-we may borrow this criterion
-in initial stages, with a modified definition
-for the histogram fluctuation
-$F^2_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde F_k^2(T),$
-since the histogram flattening
-is only partially applied to modes less than $K$.
-%
-One can alternatively define the histogram fluctuation
+supports the WL criterion for stage transitions.
+%
+%For the bandpass updating scheme,
+%we may borrow this criterion
+%in initial stages, with a modified definition
+%for the histogram fluctuation
+%$F^2_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde F_k^2(T),$
+%since the histogram flattening
+%is only partially applied to modes less than $K$.
+%
+For a bandpass updating scheme,
+we can define the histogram fluctuation
 from the largest mode
@@
 \begin{align}
-  \max_{k=1, \dots, K-1} \left| \, \tilde F_k(T) \, \right|.
+  \max_{1 \le k < K} \left| \, \tilde F_k(T) \, \right|
+  ,
   \label{eq:maxFk}
 \end{align}
-This definition can be readily extended
+%
+which allows straightforward extension
 to a continuous variable, $z$, with
@@
 in an adaptive FDS simulation
-under constant updating magnitude.
+under a constant updating magnitude.
 %
@@
 
-The basic idea is to use Eq. \eqref{eq:xt2_eql};
-to avoid the problem of not
+The idea is to use a modification
+of Eq. \eqref{eq:xt2_eql}
+that avoids the problem of not
 knowing the intrinsic distribution, $p_i$,
-hence the potential shift, $\ln(p_i/\rho_i)$,
-we should replace the shift
-by the time average of the bias potential.
-%
-We first note that the difference,
-$u_i(t) - v_i(t) = \ln(p_i/\rho_i)$, is a constant of time,
-so is the Fourier transform,
-$\tilde u_k(t) - \tilde v_k(t)$.
-%
-It follows that
-${\tilde u}_{k}$ and ${\tilde v}_{k}$ would also share the same variance.
+hence the potential shift,
+$u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
+%
+Since the shift,
+$u_i(t) - v_i(t)$,
+hence its Fourier transform,
+$\tilde u_k(t) - \tilde v_k(t)$,
+is a constant of time,
+${\tilde u}_{k}$ and ${\tilde v}_{k}$
+should share the same variance.
 %
@@
 is zero, we have
-\begin{equation*}
-  \operatorname{var} {\tilde u}_{k}
-  =
-  \operatorname{var} {\tilde v}_{k}
-  =
-  \left\langle {\tilde v}_{k}^2 \right\rangle
-  .
-\end{equation*}
-%
-Finally, by using Eq. \eqref{eq:xt2_eql},
+$\operatorname{var} {\tilde u}_{k} =
+ \operatorname{var} {\tilde v}_{k} =
+ \left\langle {\tilde v}_{k}^2 \right\rangle$,
+%\begin{equation*}
+%  \operatorname{var} {\tilde u}_{k}
+%  =
+%  \operatorname{var} {\tilde v}_{k}
+%  =
+%  \left\langle {\tilde v}_{k}^2 \right\rangle
+%  .
+%\end{equation*}
+%
+and by using Eq. \eqref{eq:xt2_eql},
 we get
@@
 %
-A practical limitation is that Eq. \eqref{eq:varXt}
+In practice, Eq. \eqref{eq:varXt}
 would fail if some eigenvalue, $\lambda_k$, is close to zero.
@@
 \begin{align*}
-\left\langle
+\bigl\langle
   \tilde f_k(t) \, \tilde f_k(0)
-\right\rangle
+\bigr\rangle
 =
@@
 \begin{equation}
-  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty \left\langle
+  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
+  \bigl\langle
     \tilde f_k(t) \, \tilde f_k(0)
-  \right\rangle = \frac{1+\Lambda_k}{1-\Lambda_k}
+  \bigr\rangle
+  = \frac{1+\Lambda_k}{1-\Lambda_k}
   .
@@
 
-The transition matrix method usually works only for
-simple sampling processes without major barrier
-along the collective variable, $z$.
-%
-We may nonetheless estimate the magnitude of $\Gamma_k$
-in a few special cases.
-%
-If sampling is perfect, $\Lambda_k = \delta_{k, 0}$,
-then $\Gamma_k = 1$ for $k > 0$.
-%
-For a local sampling process,
-if we borrow Eq. \eqref{eq:lambda_Gaussian}
+The transition matrix method
+tends to underestimate
+the first few $\Gamma_k$'s.
+%
+We will nonetheless use it as a model to
+study the behavior of $\Gamma_k$
+in a local sampling process.
+%
+If we borrow Eq. \eqref{eq:lambda_Gaussian}
 for $\Lambda_k$,
@@
 
-Here, we give some details
+Here, we give some mathematical details
 on the homogeneous updating schemes.
@@
 Further, the multiplication of $\mathbf w$
-can be reduced to a convolution with the window function,
-or the updating kernel.
+can be reduced to a convolution with the window function.
 %
@@
 the eigenvalues of $\mathbf w$
-and the updating kernel.
+and the window function.
 
 \subsection{\label{sec:wband_eig}
-Eigenvalues and eigenvectors}
+Eigenvalues and window function}
 
@@
 %
-where the numbers,
+where
 $\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
-characterizing the rigid window
-will be referred to as an \emph{updating kernel}\cite{bussi2006}.
+will be referred to as a window function.
 %
@@
 Then, the multiplication of the matrix, $\mathbf w$,
-is equivalent to a convolution with the kernel:
+is equivalent to a convolution with the window function:
 %
@@
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-compatible with the periodicity, Eq. \eqref{eq:phi_pbc},
+compatible with the periodicity
+required by Eq. \eqref{eq:phi_pbc},
 %
@@
 %
-Again, the updating kernel should satisfy
+Again, the window function should satisfy
 Eqs. \eqref{eq:mu_normalization}
@@
 i.e. $g$ is changed to $2$ in the non-periodic case.
-
-%\subsubsection{\label{sec:invert_wband}
-%Inversion}
 
@@
 
-\subsection{\label{sec:invert_wband}
-Inversion}
-
-Since the eigenvalues are related to the updating kernel
-by a cosine transform,
-we can invert the relation as
+%\subsection{\label{sec:invert_wband}
+%Inversion}
+
+We can recover the window function from the eigenvalues
+by inversely transforming Eq. \eqref{eq:wband_eigenvalue},
 %
@@
 as an updating scheme with a
-Gaussian-shaped kernel.
+Gaussian-shaped window function.
 %
@@
 
-If the width of the Gaussian updating kernel, $\sigma_z$,
+If the width of the Gaussian window, $\sigma_z$,
 is very small compared to $\Delta z$,
@@
 starting with Eq. \eqref{eq:lambda_Gaussian}
-and inversely computing the kernel using Eq. \eqref{eq:mu_from_lambda}.
+and inversely computing the window function
+using Eq. \eqref{eq:mu_from_lambda}.
 %
@@
 We can use Eq. \eqref{eq:mu_from_lambda}
-to find the updating kernel.
-%
-To take into account the two-fold degeneracy
-for a periodic variable,
-we will modify the mode cutoff such that
-$\lambda_{K} = \cdots = \lambda_{n-K} = 0$,
-and the rest of the eigenvalues are $1$.
-%
-Then, $w_{ij} = \mu_{i-j}$, and
-\begin{equation}
-  \mu_l
-  =
-  \frac{
+to find the window function.
+%
+For a periodic variable,
+we will take into account the two-fold degeneracy
+and modify Eq. \eqref{eq:lambda_bandpass}
+by setting $\lambda_{n-K-1}, \dots, \lambda_{n-1}$
+also to $1$.
+%
+Then, $\mu_l = \mathrm{sinr}(K, l, n)$ and $w_{ij} = \mu_{i-j}$,
+where
+$\mathrm{sinr}(K, l, n)
+  =
   \sin
@@
        {              n        }
-  }
-  {
-    n \, \sin \frac{ l \, \pi } { n }
-  }
-  .
-\notag
-%\label{eq:mu_sinc_pbc}
-\end{equation}
+  / \left( n \, \sin \frac{ l \, \pi } { n } \right)$.
+%\begin{equation}
+%  \mathrm{sinr}(K, l, n)
+%  =
+%  \frac{
+%    \sin
+%    \frac{ (2 K - 1) \, l \, \pi }
+%         {              n        }
+%  }
+%  {
+%    n \, \sin \frac{ l \, \pi } { n }
+%  }
+%  .
+%\notag
+%%\label{eq:mu_sinc_pbc}
+%\end{equation}
 %
@@
 from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan},
-and
-\begin{equation}
-  \mu_l
-  =
-  \frac{ (-1)^{n-K+l} } { 2 \, n }
-    +
-  \frac{
-    \sin
-    \frac{ (2 K - 1) \, l \, \pi }
-         {         2 \, n        }
-  }
-  {
-    2 \, n \, \sin \frac{ l \, \pi } { 2 \, n }
-  }
-  .
-  \notag
-  %\label{eq:mu_sinc_refl}
-\end{equation}
+and $\mu_l = (-1)^{n-K+l}/(2 \, n) + \mathrm{sinr}(K, l, 2 \, n)$.
+%\begin{equation}
+%  \mu_l
+%  =
+%  (-1)^{n-K+l} / ( 2 \, n )
+%    +
+%  \mathrm{sinr}(K, l, 2n)
+%  .
+%  \notag
+%  %\label{eq:mu_sinc_refl}
+%\end{equation}
 %
 The updating matrix is given by
-%
-\begin{equation}
-  w_{ij}
-  =
-    \frac{
-      \sin
-      \frac{ (2 K - 1) \, (i-j) \, \pi }
-           {         2 \, n        }
-    }
-    {
-      2 \, n \, \sin \frac{ (i-j) \, \pi } { 2 \, n }
-    }
-    +
-    \frac{
-      \sin
-      \frac{ (2 K - 1) \, (i+j-1) \, \pi }
-           {         2 \, n        }
-    }
-    {
-      2 \, n \, \sin \frac{ (i+j-1) \, \pi } { 2 \, n }
-    }
-  ,
-  \notag
-\end{equation}
-%
-which is free from the oscillatory term, $(-1)^{n-K+1}$,
-that appeared in the kernel
-[its contribution to $\mu_{i-j}$ in Eq. \eqref{eq:w_band_refl}
+$w_{ij} = \mathrm{sinr}(K, i-j, 2 \, n) + \mathrm{sinr}(K, i+j-1, 2 \, n)$.\footnote{$w_{ij}$
+  is free from the oscillatory term, $(-1)^{n-K+1}$,
+  that appeared in the window function, since
+  its contribution to $\mu_{i-j}$ in Eq. \eqref{eq:w_band_refl}
   is canceled by its contribution to either $\mu_{i+j-1}$
   or $\mu_{i+j-2n-1}$,
-whichever exists, of the opposite sign].
-%Although the kernel is a sawtooth
-%because of the term, $(-1)^{n-K+l}$,
-%the ruggedness disappears
-%in the updating matrix
-%as the above oscillatory term for $l = i - j$
-%is canceled by the term for either $l = i + j - 1$
-%or $l = i + j - 2 \, n - 1$
-%with the opposite sign.
+  whichever exists, of the opposite sign.}
 %
@@
     $\alpha$        &   Updating magnitude. \\
-    $b$             &   Cutoff of the kernel of homogeneous updating schemes. \\
+    $b$             &   Cutoff of the window function of homogeneous updating schemes. \\
     $C_M$           &   Integral of the mass function, $M(\bar q)$.  \\
@@
                         $m(q) \equiv M(q)/\int_0^{ q(T) } M(q') \, dq'$.  \\
-    $\mu_i$         &   Updating kernel of a homogeneous updating scheme. \\
+    $\mu_i$         &   Updating window function of a homogeneous updating scheme. \\
     $n$             &   Number of bins. \\

### 2017-06-27 doc/invt.tex
invtV7.1: small changes

@@
 we would recover $m^*(\bar q^*)$,
-as shown in Fig. \ref{fig:massq}.
+as shown in Fig. \ref{fig:massq}(b).
 %
@@
 If the initial bias, $\epsilon_k$, is negligible,
-which is typically true if no eigenvalue is close to $0$,
+which is often true if no eigenvalue is close to $0$,
 then
@@
 %
-The eigenvalues of these translationally invariant or homogeneous
+If we denote the relative updating magnitude on
+the $i$th bin from the center by $\mu_i$,
+the eigenvalues of these translationally invariant or homogeneous
 schemes can be readily deduced
@@
   although it leads to overestimated
-  $\epsilon_k$ for large-$k$ modes.
+  $\epsilon_k$'s for large-$k$ modes.
 }
@@
 in computing its change.
-The initial value is set to $q(T) = \ln(1+T\,a_0/2)$.
+The initial value was set to $q(T) = \ln(1+T\,a_0/2)$.
 }
@@
   $\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$\footnote{This
-    discretization follows from runtime average interpretation
+    discretization follows from the runtime average interpretation
     discussed in Appendix \ref{sec:equilerr}.}
@@
 %
-As shown below, for a class of bandpass updating schemes,
-  which resemble the Gaussian updating schemes,
+As we will show below, for a class of bandpass updating schemes,
   the asymptotically optimal schedule
@@
 is a multiple of the identity matrix,\footnote{Since
-  the eigenmode of $k = 0$ represents
+  the $k = 0$ eigenmode represents
   a uniform shift of the bias potential,
@@
   we note that truncating a linear or quadratic function
-  at mode $K$ (for a non-periodic variable)
+  at mode $K$ for a non-periodic variable
   results in an error roughly proportional to
@@
 
-Since bandpass updating schemes
+Since bandpass updating schemes are
 straightforward generalizations of
@@
 $a_0 = 10^{-4}$,\footnote{The equivalent $\ln f = n \, a_0$ for the WL algorithm
-was $0.05$ and $0.025$, respectively.}
+were $0.05$ and $0.025$ for $\varrho^* = 0.1$ and $\varrho^* = 0.8$, respectively.}
 followed by a production phase of another $10^7$ steps
@@
 $\Gamma_k$'s and $\epsilon_k$'s,
-although ideally they should be estimated from the data
+although ideally they could be estimated from the data
 collected from the equilibration phase in individual runs.
@@
 corresponding to the longest range fluctuation,
-$\Gamma_1 \approx \Gamma$,
+as $\Gamma_1 \approx \Gamma$,
 and the reduction of local noise
@@
 the final errors from the Gaussian updating scheme
-were greater than those from the single-bin updating scheme
+were greater than those from the single-bin updating scheme,
 although the initial values were less.
@@
 the inverse-time schedule tended to
-reduce the updating magnitude too quickly
+reduce the updating magnitude too quickly,
 leaving a hump on the error profile
@@
 As a result, the optimal schedule usually produced
-less error than the inverse-time schedule
+less error than the inverse-time schedule,
 as shown in Table \ref{tab:lj_error}.
@@
 in the Gaussian updating scheme
-is to use the correction formula
-based on histogram, Eq. \eqref{eq:uhatav}.
+is to use the histogram-based correction formula,
+Eq. \eqref{eq:uhatav}.
 %
@@
 and one may have to either decrease the width
-or to elongate the preliminary simulation
-under the constant magnitude.
+or to elongate the preliminary run.
 
@@
     as given by Eq. \eqref{eq:half_alpha0}.
-    But an overly wide Gaussian window ($\sigma = 0.6$)
+    But an overly wide window ($\sigma = 0.6$)
     led to a much larger updating magnitude.
@@
 %
+In this case, $\mathcal U(\mathbf x) = 0$
+and the intrinsic distribution is
+the density of states, $\ln g(E)$.
+%
 We spread the Gaussian umbrellas
@@
 %
-Initially, we set $c_1 = 1.4 \sigma_E$, and $c_2 = 0$.
+Initially, we set $c_1 = 1.4 \sigma_E$ and $c_2 = 0$.
 %
@@
 until it fell below
-the value given the inverse-time schedule\cite{
-belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
+the value given the inverse-time schedule.\footnote{We
+  slightly modified the switch criterion from the WL
+  to the inverse-time schedule\cite{belardinelli2007,
+  *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
+  Such a switch was only permitted
+  upon completion of a WL stage,
+  and the time, $t$, was interpreted as the number of updating steps
+  in this just completed WL stage.
+  If the switch was successful,
+  the above $t$ was used as the offset $t_0$
+  in the inverse-time regime.}
 %
@@
 The parameter, $c_1/\sigma_E$, roughly represents
-the local or statistical temperature,
+the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
 and it demonstrated
@@
     %
-    (b) The parameter, $c_1/\sigma_E$, roughly corresponding to
+    (b) The parameter, $c_1/\sigma_E$ (squares), roughly corresponding to
     the statistical temperature, $\beta(E)$,
+    (dashed line, estimated from the smoothed density of states)
     demonstrated the backbending behavior.
     %
-    (c) The parameter, $\sqrt 2 \, c_2$, exceeds $1.0$
+    (c) The parameter, $\sqrt 2 \, c_2$ (squares),
+    exceeded $1.0$
     in the critical region, indicating local convexity
@@
 We can generally characterize
-an updating scheme with an updating matrix,
+an updating scheme by an updating matrix,
 with each column being the normalized
@@
 and recover a bias potential with precision
-similar to that from the single-bin case.
+similar to that from the single-bin simulation.
 
@@
 which are also asymptotically optimal,
-but completely filter the components
-in the noise modes.
+but completely filter the noise modes.
 %
@@
 from orthogonal polynomials,
-the updating scheme became a generalization
+it became a generalization
 of the LLR algorithm\cite{langfeld2012, pellegrini2014},
-and it could be used for parameter adjustment
+and could be used for parameter adjustment
 in umbrella sampling with
@@
 %Equation \eqref{eq:uhatav}
-This equation can also be derived\footnote{From Eq. \eqref{eq:uhatt},
+This equation serves as
+a generalization of Eq. \eqref{eq:vcorr_equil}
+for FDS simulations under a variable bias potential
+and we can use it to improve the estimated PMF
+after simulation.\footnote{Equation \eqref{eq:uhatav}
+  can also be derived
+  from the self-healing umbrella sampling method\cite{marsili2006}.
+  From Eq. \eqref{eq:uhatt},
   we can construct an independent estimator
@@
   $\rho_i e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
-from the self-healing umbrella sampling
-method\cite{marsili2006},
-and it serves as
-a generalization of Eq. \eqref{eq:vcorr_equil}
-for FDS simulations under a variable bias potential
-and we can use it to improve the estimated PMF
-after simulation.
 %
@@
 as runtime averages,
-but with a dynamic weight.
+but with dynamic weights.
 %
@@
 The inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
-follows from the plain average,
+follows from the case of unweighted average,
 $\omega(t) = 1$ (for $t > 0$),
@@
 Since the shift,
-$u_i(t) - v_i(t)$,
 hence its Fourier transform,
@@
 %
-But since the ensemble average of ${\tilde v}_{k}$
+But as the ensemble average of ${\tilde v}_{k}$
 is zero, we have

### 2017-06-28 doc/old/V7.1/invtV7.1.tex
adding Dr. Pettitt's V7.1

--- 2017-06-26 doc/invt.tex
@@
 \newcommand{\ii}{\mathrm{i}}
-%bin average
 \newcommand{\bav}[1]{#1_\mathrm{av}}
@@
   of the updating magnitude.
-  %
-  %Here, we will show that the optimal schedule
-  %can be derived
-  %from a mechanical analogy, in which
-  %the schedule corresponds to
-  %the velocity of a free particle with a position-dependent mass
-  %and the final error serves as the action.
-  %%
-  %Therefore, the optimal schedule follows from
-  %the equation of motion of the particle.
-  %
   We will show that
   the optimal schedule depends on the updating scheme.
-  %
   While the asymptotically optimal schedule for
@@
   is often more complex.
-  %and the initial updating magnitude is ideally half of
-  %the previous equilibrium value.
-  %
   Further,
@@
   that are optimal for asymptotic convergence.
-  %
   These bandpass updating schemes aim at
@@
   is also given by the inverse-time prescription.
-  %
   Constructed from orthogonal polynomials,
@@
   of the Langfeld-Lucini-Rago algorithm
-  and can be used for parameter tuning for the method of Gaussian ensemble
+  and can be used for parameter tuning for the Gaussian ensemble method
   and umbrella sampling.
-  %
 \end{abstract}
@@
 with traditional experiments.
-%
 Given a system,
@@
 or molecular dynamics\cite{frenkel, karplus2002} (MD) simulations.
-%
 The dimensionless free energy, or potential of mean force (PMF),
@@
 is $-\ln p(z)$.
-%
 However, accurately estimating the PMF is often difficult
@@
 of the complicated, multimodal free energy surface.
-%
 Thus,
@@
 cancels the PMF.
-%
 Ideally, then, sampling yields a flat distribution
@@
 *laio2002, *laio2008, *barducci2011, *sutto2012}.
-%
 The flat-distribution sampling can be used to accelerate
@@
   zheng2010}.
-%
 These techniques regularly update the bias potential
@@
 by incrementally elevating the bias potential.
-%
 A key difference lies
@@
 as well as the relative updating magnitude.
-%
 Often, the window function
@@
 in metadynamics.\cite{junghans2014}
-%
 We will refer to the two updating schemes
@@
 and Gaussian updating schemes, respectively.
-%
 While the former is popular for MC simulations\cite{wang2001,
@@
   *laio2002, *laio2008, *barducci2011, *sutto2012, junghans2014}
-%
 In both cases,
@@
 barducci2008, dickson2011, dama2014}.
-%
 For the single-bin case, the optimal updating schedule
@@
 the inverse number of updating steps.
-\note{
-  About the origin of the inverse-time formula.
-  Besides the Belardinelli paper,
-  we can also find it in Eq. (6) of reference \cite{liang2007},
-  if I am not too mistaken.
-  But I find the paper is a bit hard to read.
-  There, however, the formula is $t_0/t$ instead of $1/t$.
-}
-%The same schedule also applies to a variant of
-%the WL algorithm\cite{langfeld2012, pellegrini2014}.
 
@@
 including the single-bin and Gaussian versions.
-%
 We will map the optimization problem to a mechanical one,
@@
 of the particle that minimizes the action.
-%
 The resulting optimal schedule
@@
 through the mass function.
-%
 While the optimal schedule in the single-bin case
@@
 and the width of the Gaussian window.
-%
-%In either case, however,
-%Ideally,
-%the optimal initial updating magnitude
-%can be set to
-%half of the previous equilibrium value,
-%which amounts to a shift of the origin of time
-%in the single-bin case.
-%
 Further, we will show that
@@
 bandpass updating schemes.
-%
 The latter allow one to partially flatten
@@
 into the bias potential during the updating process.
-%
 The optimal schedule of bandpass updating schemes
 is also an inverse-time schedule.
-%
 Constructed from orthogonal polynomials,
@@
 The article is organized as follows.
-%
 We present the analytical results in Sec. \ref{sec:theory},
@@
 and some known aspects of the optimal schedule.
-%
 Then, in Sec. \ref{sec:single-bin},
@@
 in Sec. \ref{sec:multiple-bin}.
-%
 Finally, we compare different updating schemes
@@
 
-%\subsubsection{\label{sec:FDS}
-%Bias potential}
-
 Consider the problem of computing
@@
 for a given system.
-%
-%For example, $i$ can be the energy, $E$,
-%in a lattice spin model or the temperature index
-%in a simulated tempering simulation\cite{
-%marinari1992, *lyubartsev1992}.
-%
 For a continuous quantity, $z$,
-%which is often the case in a (classical) molecular system,
 we can use the equivalent integer $i$ to represent
@@
 $(z, z + \delta z)$.
-%
-%The distribution is normalized as
-%$\sum_{i = 1}^n p_i = 1$.
 
@@
 sampling a smoother biased distribution, $\rho_i$.
-%
-%
 We will broadly define
@@
 dayal2004, *trebst2004, barducci2008, singh2011}.
-%
 The biased equilibrium distribution
 under a bias potential, $u_i$, is given by
-%
 \begin{equation}
   \pi_i
-  %=
-  %C_i \, p_i \, e^{-u_i}
   =
@@
   \notag
-  %\label{eq:pi_p_V}
-\end{equation}
-%
+\end{equation}
 and our aim is to find a bias potential
@@
 with the desired $\rho_i$.
-%
 Upon convergence, the bias potential would satisfy
-%
 \begin{equation}
@@
 \end{equation}
-%
-%according to Eq. \eqref{eq:pi_p_V}.
-%which allows us to deduce the PMF, $-\ln p_i$,
-%directly from the bias potential.
-%
 and the PMF, $-\ln p_i$, can be readily
 deduced from the bias potential. % $u_i$.
-\note{
-  For example, in the context of simulated tempering (Zhang and Ma, 2007),
-  $p_i$ is replaced by $\ln Z_i$,
-  $\rho_i$ is the weight $\zeta_i$,
-  and the bias potential, $u_i$, is $\ln (\tilde Z_i / \zeta_i)$.
-  Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
-  up to a constant shift.
-}
 
@@
 have been developed to find the proper bias potential.
-%
 In a class of equilibrium FDS methods\cite{
@@
 during simulation,
-%
 and a correction to the bias potential
@@
 accumulated from the simulation as
-%
-%Using $H_i$ as an estimate of $\pi_i$ in
-%Eq. \eqref{eq:pi_p_V}
-%allows us to correct $u_i$ as
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{This follows from
-  \begin{align*}
-    H_i \approx \pi_i
-    &\propto p_i \, e^{-u_i},
-    \\
-    \rho_i
-    &\propto p_i \, e^{-\hat u_i}.
-  \end{align*}
-}
 However, this formula often requires a long
@@
 continuous improvement of the bias potential
-%(hence that of the PMF)
 as simulation lengthens.
@@
 by continuously and incrementally updating the bias potential.
-%
-%\subsubsection{Updating schemes}
-%
 In the WL algorithm\cite{wang2001, wang2001pre},
@@
 at each step $t$ as
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $i(t)$ is the bin at step $t$,
@@
 and $\alpha(t)$ is the updating magnitude.
-%
 We refer to this scheme as the \emph{single-bin scheme},
@@
 at the current bin, $i(t)$.
-%
 In a more general \emph{multiple-bin scheme},
-%including the Gaussian updating scheme used in metadynamics,
 the updating also involves neighboring bins:
-%
 \begin{equation}
@@
 \end{equation}
-%
 The Gaussian updating scheme used in metadynamics
 is an example, with $w_{ij}$ being a Gaussian centered on $j$.
-%
 Note that if the updating occurs
@@
 then $t$ means the number of updating steps.
-
-%\subsubsection{Updating magnitude and the inverse-time formula}
 
@@
 is controlled in stages.
-%
 In each stage, $\alpha(t)$
@@
 wang2001, wang2001pre}.
-%
 While this scheme works well for early stages,
@@
 
-%For the single-bin scheme
-%used in WL algorithm,
 A more effective way
@@
 is to follow the formula
-%
 \begin{equation}
@@
 komura2012, *caparica2012, *caparica2014}.
-%
 This general result\cite{
@@
 (cf. Appendix \ref{sec:equilerr}).
-%
-%We may also derive it
-%from optimally planning the simulation lengths in the stages
-%and the reduction factor in the WL setup (cf. Appendix \ref{sec:wltoinvt}).
-
-%In the next section,
-%we will re-derive this formula
-%and suggest a slight modification
-%whereby the origin of $t$ is shifted
-%according to the initial error.
-%
-
-%As Eq. \eqref{eq:alpha_invt} was being developed for the WL algorithm,
-%several similar ways of
-%reducing the updating magnitude
-%were developed for umbrella sampling and metadynamics,
-%particularly, the self-healing umbrella sampling
-%and well-tempered metadynamics\cite{
-%  marsili2006, barducci2008, dickson2011, dama2014}.
-%%
-%In this study, we wish to find the optimal schedule, $\alpha(t)$,
-%for a general updating scheme,
-%including the Gaussian scheme used in metadynamics.
 
@@
 for the more general multiple-bin scheme.
-%
-%We will suggest a slight modification of Eq. \eqref{eq:alpha_invt}
-%whereby the origin of $t$ is shifted by an amount
-%determined from the initial error.
-%
 We will first express the error of
@@
 and then minimize it by variation.
-%
 
@@
 by deducting two contributions.
-%
 First, to deduct the target value given by
@@
 we introduce a shifted bias potential
-%
 \begin{equation}
@@
   \notag
-  %\label{eq:v_def}
-\end{equation}
-%
+\end{equation}
 which should tend to a constant of $i$
@@
 remains a constant during simulation.
-%%
-%In terms of $v_i$'s, Eq. \eqref{eq:pi_p_V}
-%becomes
-%%
-%\begin{equation}
-%  \pi_i
-%  =
-%  \frac{                \rho_i \, e^{-v_i} }
-%       { \sum_{j = 1}^n \rho_j \, e^{-v_j} }
-%  \propto
-%  \rho_i \, e^{-v_i}.
-%  \label{eq:pi_p_v}
-%\end{equation}
-%%
-%According to Eq. \eqref{eq:pi_p_V},
 
@@
 generally grows over time.
-%
 Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
@@
 dama2014}.
-%
 We define the bias potential deviation from the
@@
 as
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{Alternatively,
-  we may define $v_{*i}(t)$ such that
-  $\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
-  i.e.,
-  $$
-  v_{*i}(t) \to -\ln\left(
-    \frac{ (p_i/\rho_i) \, e^{ -u_i(t) } }
-    { \sum_{j=1}^n p_j \, e^{ -u_j(t) } }
-  \right).
-  $$
-  One can readily show that the two definitions are equivalent
-  for small $v_{*i}$.}%
-%
 Below we will use the notations, $x_{*i}$ and $\bav{x}$,
 for a general variable, $x_i$.
-%
 The total error is then given by
-%
 \begin{equation}
@@
 e.g. the sum $\sum_{i=1}^T$ is mapped to the integral $\int_0^T dt$.}
-%
 \begin{equation}
@@
 \end{equation}
-%
 with
-%
 $h_i(t) = \delta_{i, i(t)}$
-%\begin{equation}
-%  h_i(t) = \delta_{i, i(t)}
-%  ,
-%  \label{eq:h_def}
-%\end{equation}
-%
 being the instantaneous histogram,
@@
 $f_i(t) \equiv h_i(t) /\rho_i$.
-%
 Clearly,
-%
 \begin{equation}
@@
 \end{equation}
-%
 The shifted bias potential, $v_i(t)$,
@@
 as the shift is a constant.
-%
 Deducting the bin average [cf. Eq. \eqref{eq:x_def}],
 we get
-%
 \begin{equation}
@@
 and a random fluctuating part, $\zeta_i(t)$,
-%
 \begin{equation}
@@
   \notag
-  %\label{eq:h_split}
-\end{equation}
-%
+\end{equation}
 The deterministic part can be related
@@
 and the same bias potential at time $t$.
-%
 The initial states and the stochastic forces
 during the process may differ, however.
-%
 For sufficiently small $\alpha(t)$,
@@
 such that
-%with the deterministic part specified by Eq. \eqref{eq:pi_p_V}:
-%
 \begin{align}
   \langle f_i(t) \rangle
-  %&=
-  %\frac{ \langle h_i(t) } { \rho_i } \rangle
   &\approx
@@
        { \sum_{j = 1}^n \rho_j \, e^{-v_{*j}(t)} }
-  %\notag
-  %\\
-  %&
   \approx
@@
   \notag
-  %\label{eq:h_ave}
 \end{align}
-%
 where we have assumed the smallness
 of $v_{*i}(t)$ in the linear expansion.
-%
-\note{
-The second step follows from
-$$
-\frac{ \langle h_i(t) \rangle }
-     { \rho_i }
-\approx
-\frac{                       1 - v_{*i}  }
-     { \sum_{ r = 1 }^n \rho_j (1 - v_{*j}) }
-=
-\frac{                       1 - v_{*i}  }
-     { 1 - \sum_{ j = 1 }^n \rho_j \, v_{*j} }
-.
-$$
-}%
-%
 Deducting the bin average and using Eq. \eqref{eq:fav1}, we get
@@
 and
-%
 \begin{align}
@@
   \notag
-  %\label{eq:G_def}
 \end{equation}
@@
 and $\delta(t)$ is the Dirac delta function.
-%
 We expect the approximation to hold for simulation
@@
 we get % a set of decoupled equations
-%
 \begin{equation}
   \dot v_{*i}(t)
-  %=
-  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \zeta_{*i}(t) \right]
   \approx
@@
   \notag
-  %\label{eq:dxdt_singlebin}
-\end{equation}
-%
+\end{equation}
 and the formal solution is
-%
 \begin{equation}
@@
 \notag
-%\label{eq:xt_solution}
-\end{equation}
-%
+\end{equation}
 where
-%
 $
@@
 $
-%
 and
-%
 \begin{align}
@@
 and an asymptotic error, $E_A$:
-%
 \begin{align}
@@
 $\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
-%\begin{equation}
-%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i.
-%  \label{eq:Gamma_from_G}
-%\end{equation}
 
@@
 via $v_{*i}(T)$.
-%
 If $\alpha(t)$ is a constant, $a_0$,
@@
 the error approach an equilibrium value
-%
 \begin{equation}
@@
 \end{equation}
-%
 Our aim is to find the $q(t)$
 that minimizes this error.
-%
 We now fix the endpoint value, $q(T)$,
 and vary the curve, $q(t)$ ($0 < t < T$).
-%
 This is equivalent to varying the curve,
@@
 $\vartheta\bigl( q(T) \bigr) = 1$.
-%
 Since the residual error
@@
 demands a constant velocity
-%
 \begin{equation}
@@
 \end{equation}
-%
 Then, by using Eq. \eqref{eq:theta_def}, we have
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $t_0$ is some constant,
@@
 in the second step.
-%
 Taking the logarithm,
@@
 we get the inverse-time schedule
-%
 \begin{equation}
@@
 \end{equation}
-%
 and Eq. \eqref{eq:alpha_invt}
 is the special case of $t_0 = 0$.
-%
 
@@
 which is fixed in the above.
-%
 This is equivalent to choosing a value of $t_0$
@@
 from the $t = 0$ case of Eq. \eqref{eq:expqt}\big].
-%
 Using Eq. \eqref{eq:alpha_invt1} in
@@
   \notag
-  %\label{eq:ER_sbin1}
   \\
@@
 \end{align}
-%
 and the total has a minimum of
@@
 \end{equation}
-%
 which is reached at $t_0 = \Gamma /\Err(0)$,
 or if we assume Eq. \eqref{eq:error_equil_sbin} for $\Err(0)$,
-%
 \begin{equation}
@@
 \end{equation}
-%
 The minimum error can be related to
@@
 We now turn to a general multiple-bin updating scheme.
-%
 By definition, Eq. \eqref{eq:mbin_update},
@@
 not only at bin $j$, but also at a few neighboring bins $i$'s.
-%
 Similar to Eq. \eqref{eq:ut_diffeq},
@@
 as a differential equation,
-%
 \begin{equation}
@@
 \end{equation}
-%
 The optimization procedure is similar to
 the single-bin case.
-%
 The key simplification comes from the
@@
 The matrix, $\mathbf w$, is subject to several conditions.
-%
 First, we have a fixed-point condition\cite{bussi2006, dama2014}.
-%
 To sample the desired distribution,
@@
 in the asymptotic regime.
-%
 By a proper overall scaling of $\mathbf w$,
 we may write this condition as
-%
 \begin{equation}
@@
 in our cases.
-%
 Below we will %take advantage of the analogy and
@@
 
-%To simplify the ensuing discussion,
 Second, we limit ourselves to matrices % $\mathbf w$'s
@@
 the detailed balance condition for $\pmb\rho$,
-%
 \begin{equation}
@@
 \end{equation}
-%
 This requires the scaled updating matrix,
@@
 \end{align}
-%
 where we have set the index of the first eigenmode to $0$
@@
 with eigenvalue $1$:
-%
 \begin{equation}
@@
   \notag
-  %\label{eq:w_balance}
-\end{equation}
-%
+\end{equation}
 We will label this eigenvector by $k = 0$,
 and
-%
 \begin{equation}
@@
 \end{equation}
-%
 with $\lambda_0 = 1$.
-%
 Clearly, Eq. \eqref{eq:eigenmode0}
@@
 while the orthogonality demands
-%
 \begin{equation}
@@
 \end{equation}
-%
 We may reconstruct the updating matrix in $\phi_{ki}$'s
 as\cite{bussi2006}
-%
 \begin{equation}
@@
 for variable, $x_i$,
-%
 \begin{equation}
@@
   \notag
-  %\label{eq:gft_def}
-\end{equation}
-%
+\end{equation}
 and the inverse transform is
 $x_i = \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k / \rho_i$.
-%
-%\begin{equation}
-%  x_i = \frac{1}{\rho_i} \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k
-%  .
-%  \notag
-%  %\label{eq:gft_inv}
-%\end{equation}
-%
 Then,
@@
 $\mathcal F[1]_k = \delta_{k0}$,
-%
 and Eq. \eqref{eq:eigenmode0} means that
 the bin average
-%of any quantity, $x_i$, is the first mode of the Fourier transform:
 $\bav{x}= \mathcal F[x_i]_0 = \tilde x_0$,
-%
 and
@@
 \end{equation}
-%
 Thus, the only effect
@@
 annihilate the $k=0$ mode.
-%
 We also have Parseval's theorem,
@@
 \end{equation}
-%
 or in continuous time, we have, for $k > 0$,
-%
 \begin{equation}
@@
 is given by $\alpha(t) \, \lambda_k$.
-%
 Thus, a \emph{stable} updating scheme
@@
 for $k > 0$.
-%
 With a set of heterogeneous eigenvalues,
 the errors of the eigenmodes are reduced at different rates.
-%
 Thus, the optimization problem of the final error
@@
 $\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
-%
-%\begin{align}
-%  \Err(T)
-%  =
-%  \sum_{k = 1}^{n - 1}
-%    \left\langle
-%      {\tilde v}_k^2(T)
-%    \right\rangle
-%  .
-%\notag
-%\end{align}
-%
 As in Eq. \eqref{eq:error_tot},
@@
 \end{align}
-%
 where
-%
 \begin{equation}
@@
 \end{equation}
-%
 and $\Gamma_k$ is computed from
@@
 as:
-%
 \begin{equation}
@@
 \end{equation}
-%
-%\begin{equation}
-%$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
-%  \notag
-%\label{eq:Gamma_integral}
-%\end{equation}
-%
 Equation \eqref{eq:Gamma_sum} represents
@@
 few numbers, $\Gamma_k$'s.
-%
 In Appendix \ref{sec:Gamma_measure},
@@
 from an FDS simulation under a constant magnitude.
-%
-
-%\subsubsection{\label{sec:eqlerr}
-%Initial error
-%%from a simulation under a constant updating magnitude
-%}
 
@@
 performed under a constant updating magnitude for $T_0$ steps.
-%
 If the initial bias potential
@@
 in the case of $T_0 \to \infty$,
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{
-Consider a long period, $T_0$, under a fixed schedule,
-$\alpha(t) = a_0$,
-the residual error becomes negligible, and
-the $k$th component of the asymptotic error
-is given by
-%
-\begin{align*}
-  \left\langle
-    {\tilde v}_k^2(T_0)
-  \right\rangle
-  =
-  \int_0^{T_0}
-    \Gamma_k \, (\lambda_k \, a_0)^2 \,
-      e^{ 2 \, \lambda_k \, a_0 \, (t - T_0) }
-    \, dt
-  \stackrel{ T_0 \to \infty }
-  { =\joinrel=\joinrel=\joinrel= }
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
-  .
-\notag
-%\label{eq:error_eql}
-\end{align*}
-}
-%
 
@@
 $\Err_A(T)$.
-%
 The value of $q(T)$ will be adjusted later
 to minimize the total error in Sec. \ref{sec:optinitalpha}).
-%
 We can rewrite Eq. \eqref{eq:error_asym} much like an action
 of a particle whose position is given by $q(t)$:
-%
 \begin{equation}
@@
   \notag
-  %\label{eq:error_asym_Lagrangian}
-\end{equation}
-%
+\end{equation}
 where the Lagrangian is
-%
 \begin{align}
@@
     \Gamma_k \, {\dot \vartheta}_k^2\bigl[ q(t) \bigr]
-  %\notag
-  %\\
-  %&=
-  =
-  %\sum_{ k = 1 }^{n - 1}
-  %  \Gamma_k \, \lambda_k^2 \, \vartheta_k^2 \bigl[ q(t) \bigr]
-  %\; \dot q^2( t )
-  %\notag
-  %\\
-  %&=
+  =
   M^2\bigl(q(T) - q(t) \bigr)
@@
 defined the (square-root) mass function as
-%
 \begin{equation}
@@
   .
-  %\notag
   \label{eq:mass_func}
 \end{equation}
-%
 In this mechanical analogy,
@@
 corresponds to the velocity of the particle,
-%
 and the above Lagrangian, containing only the kinetic energy,
@@
 with a position-dependent mass.
-%
 Here, the Hamiltonian coincides with the Lagrangian:
-%
 \begin{equation}
@@
   \notag
-  %\label{eq:error_asym_Hamiltonian}
-\end{equation}
-%
+\end{equation}
 Since the Lagrangian
@@
 We may set
-%
 \begin{equation}
@@
 \end{equation}
-%
 for some positive $C_M$
 as a generalization of Eq. \eqref{eq:dthetadt_const}.
-%
 By integrating this equation of motion, we get
-%
 \begin{equation}
@@
 \end{equation}
-%
 and $C_M$ can be determined from
 the $t = T$ case as
-%
 \begin{equation}
@@
 \end{equation}
-%
 Thus, Eq. \eqref{eq:Lagrangian_const}
@@
 $\alpha(t) = \dot q(t)$.
-%
 The minimal asymptotic error is
-%
 \begin{equation}
@@
 geometrically as follows.
-%
 First, we define
 a normalized mass distribution as
-%
 \begin{equation}
@@
 \notag
-%\label{eq:mass_distr}
-\end{equation}
-%
+\end{equation}
 such that
 $\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
-%
 Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
 we find that for $\bar q(t) \equiv q(T) - q(t)$,
-%
 \begin{align}
@@
 \end{align}
-%
 Thus, each point on the schedule,
@@
 
-    % distribution
     \draw[domain=0:2.0, variable=\x, dashed, gray, very thick]
@@
 
-    % shade
     \fill [gray!30!white, domain=0.7:1.609, variable=\x]
@@
 
-    % thick curve
     \draw[domain=0:1.609, variable=\x, very thick]
@@
 
-    % label for the shaded area
     \node[] at (1.1, 0.32) {$\dfrac{t}{T}$};
@@
                (0.6, 0.99317);
-    % label for the ordinate
-    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
 
@@
   \end{tikzpicture}
-%
   \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
     \node [] at (-0.3, 2.0) {(b)};
-    % distribution
     \draw[domain=0:2.3, variable=\x, dashed, gray, very thick]
@@
 
-    % shade
     \fill [gray!30!white, domain=0.7:2.0, variable=\x]
@@
 
-    % thick curve
     \draw[domain=0:2.0, variable=\x, very thick]
@@
 
-    % label for the shaded area
     \node[] at (1.1, 0.35) {$\bar q(t)$};
@@
                (0.6, 0.83333);
-    % label for the ordinate
-    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
 
@@
     $m(\bar q)$.
-    %
     (b) Inverse construction of $m(\bar q)$
     from the optimal schedule $\alpha(t)$.
-    %
-    \note{
-      The figure was produced by \texttt{tikz}.
-    }%
   }
@@
 the mass distribution, $m(\bar q)$.
-%much like the Legendre transform.
-%
 If we define
@@
 being $\bar q^*$ and $t^*$, respectively.
-%
 Further, the transformation is an involution:
@@
 as shown in Fig. \ref{fig:massq}.
-%
 A few examples are shown in Table \ref{tab:m_and_a}.
-%
 For example, in the single-bin scheme,
@@
 and we would have the inverse-time schedule.
-%
 On the other hand,
@@
   }
-  %
 then we would have an exponentially decaying schedule.
@@
     $\dfrac{c^{1/a} \, (1/a)}{\left( t^* + t^*_0 \right)^{(1/a)+1}}$ &
-    %$\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a} - q^*_0$
     $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^{1/a} - 1\right]$
-    %$\,\begin{aligned}
-    %  &a, q^*_0 > 0, \\
-    %  &\frac{1}{t^*_0} \equiv \left(1+\frac{1}{q^*_0}\right)^a - 1, \\
-    %  &c \equiv {q^*_0}^a(t^*_0 + 1)
-    %\end{aligned}$
     \\
@@
 that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
-%
 We will consider simulations in which
@@
 at a constant updating magnitude, $a_0$.
-%
 From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
@@
 the optimal $q(T)$ satisfies
-%
 \begin{align}
@@
 \end{align}
-%
-%\begin{align}
-%  m\bigl( q(T) \bigr)
-%  +
-%  \frac{2}{a_0 \, C_M \, M\bigl( q(T) \bigr)}
-%  \sum_{k = 1}^{n-1}
-%  \lambda_k \, \tilde u_k^2 \,
-%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
-%  \\=
-%  \frac{1} { T \, (a_0 / 2) }
-%  .
-%\label{eq:opt_qT}
-%\end{align}
-%
-\note{Let $q_T \equiv q(T)$,
-$$
-\begin{aligned}
-  \frac{
-    \partial \Err_R(T)
-  }
-  {
-    \partial q_T
-  }
-  &=
-  -a_0 \, M^2(q_T)
-  -2
-  \sum_{k=1}^{n-1} \lambda_k \,
-  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
-  ,
-  \\
-  \frac{
-    \partial \Err_A(T)
-  }
-  {
-    \partial q_T
-  }
-  &=
-  \frac 2 T \,
-  M(q_T) \,
-  \int_0^{ q_T } M(q) \, dq
-  =
-  \frac{ 2 \, C_M } { T } \, M(q_T)
-  .
-\end{aligned}
-$$
-Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
-}
-%
 We can rewrite the condition in terms of
@@
 \end{align}
-%
 If the initial bias, $\epsilon_k$, is negligible,
@@
 then
-%
 \begin{equation}
@@
 \end{equation}
-%
 i.e. the optimal initial updating magnitude
@@
 centered on the current bin.
-%
 The eigenvalues of these translationally invariant or homogeneous
@@
 as a cosine transform,
-%
 \begin{equation}
@@
 \end{equation}
-%
 where $g = 1$ here for the periodic variable,
@@
 (cf. Appendix \ref{sec:more_wband}).
-%
 For the Gaussian updating scheme,
@@
   ,
-%\notag
 \label{eq:lambda_Gaussian}
@@
 for a given adaptive FDS simulation.
-%
 Basically, we need to extract
@@
 as shown in Fig. \ref{fig:vardep}.
-%
 While the updating scheme alone
@@
 and the underlying sampling process.
-%
 We have
@@
 for a homogeneous updating scheme.
-%
 Note that
@@
 under a constant updating magnitude, $a_0$.
-%
-\note{The updating magnitude should be sufficiently small
-  such that the resulting error is no greater than
-  multiple of $1.0$.}
 
@@
 Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
-%
 We implemented a variant of the Newton-Raphson method\cite{press3rd}
@@
 }
-%
-%Since $m(\bar q)$ is a decreasing function of $\bar q$,
-%we can start from an interval $[Q_L, Q_R]$
-%that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
-%(e.g. we can choose $Q_L = 0$,
-%  and repeatedly double $Q_R$ until
-%  the latter inequality is satisfied).
-%%
-%We can then iteratively refine the interval
-%by substituting $Q_M = (Q_L + Q_R)/2$
-%for either $Q_L$ and $Q_R$,
-%depending on the sign of $m(Q_M) - 2/(T \, a_0)$,
-%until the width of the interval is less than
-%the desired precision,
-%%
-%and set $q(T)$ to $Q_M$.
 
@@
   numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.
-%
 To handle Gaussian updating schemes more effectively,
@@
     roughly a constant.}
-%
 The curve $q(t)$ is then integrated
@@
   resulting in a list of values, $\bigl(q_{[i]}, t_{[i]} \bigr)$.
-%
 We then compute the updating magnitude by numerical differentiation:
@@
   $t_{[N+1]} \equiv T$.
-%
 The resulting values of $(t_{[i]}, \alpha_{[i]})$
@@
 The above procedure is not always necessary.
-%
 As shown below, for a class of bandpass updating schemes,
@@
   is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
-%
-%For these updating schemes,
-%  the preliminary run with a constant updating magnitude
-%  can be replaced by the WL-style stagewise reduction
-%  until the updating magnitude falls under
-%  the value given the inverse-time prescription\cite{
-%    belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
 
@@
 $\lambda_k$'s.
-%
 Using the Cauchy-Schwarz inequality, we have,
@@
 $A_1^2 - 4 \, A_0 \, A_2$.}
-%
-%
 \begin{align}
@@
 \notag
-%\label{eq:CSineq}
 \end{align}
-%
 Note that the last expression %of Eq. \eqref{eq:CSineq}
@@
 which is fixed in the variational process.
-%
 Thus, the inequality sets a lower bound
@@
 given in Eq. \eqref{eq:error_asym}.
-%
 The equality is achieved
@@
 up to a multiplicative factor.
-%
 Solving these equations yields
@@
 by setting the eigenvalues to $1$.
-%
 Thus, the single-bin scheme is most efficient
 in error reduction in the asymptotic limit.
-%
 In particular,
@@
 $\Err_A(T) = \sum_{ k = 1 }^{n-1} \Gamma_k \, T / (T + t_0)^2$,
-%
-%\begin{align}
-%  \Err_A(T)
-%  =
-%  \frac{       T     }
-%       { (T + t_0)^2 }
-%  \sum_{ k = 1 }^{n-1}
-%    \Gamma_k
-%  ,
-%\end{align}
-%
 which recovers Eq. \eqref{eq:EA_sbin1}.\footnote{By
@@
 which are typically local noise.
-%
 A bandpass updating scheme is essentially
@@
 which are assigned with zero eigenvalues.
-%
 For example,
@@
 we demand
-%
 \begin{equation}
@@
 \end{equation}
-%
-%Then, the updating matrix is, according to
-%Eq. \eqref{eq:w_from_phi},
-%%
-%$$w_{ij} = \frac{1}{\rho_i} \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj}.$$
-%%
 The asymptotic error is then
@@
 which is less than that from Eq. \eqref{eq:EA_sbin1}.
-%
 The optimal schedule of these schemes
@@
 with $t_0$ given by Eq. \eqref{eq:t0_sbin}.
-%
-%The asymptotic error is given by
-%%
-%\begin{equation}
-%  \Err_A(T)
-%  =
-%  \frac{   T     }
-%       { (T + t_0)^2 }
-%  \sum_{ k = 1 }^{ K - 1 }
-%    \Gamma_k
-%  ,
-%  \notag
-%  %\label{eq:error_sinc}
-%\end{equation}
-%%
-%as a generalization of Eq. \eqref{eq:Emin_sbin}
-%for the single-bin updating scheme.
-%
 
@@
 in initial stages of the simulation.
-%
 However, the histogram fluctuation should
@@
 \end{align*}
-%
 With $\phi_k(z) = R_k(z) \, \rho(z)$,
@@
 \end{align*}
-%
 where we have dropped the $k=0$ mode
@@
 of the bias potential.
-%
 The resulting bias potential
@@
 and the center, $z_c$.
-%
 We thus recover the LLR
@@
 we can use the Hermite polynomials\cite{arfken}.
-%
 With $K = 3$, the bias potential is quadratic
@@
 $R_2(z) = \bigl[(z - z_c)^2 - \sigma_z^2\bigr] /\left(\sqrt 2 \, \sigma_z^2\right)$.
-%
 Note that quadratic bias potentials are commonly
@@
 in first-order phase transitions.
-%
 By Eq. \eqref{eq:ckupdate},
@@
 central moments of $z$, respectively.
-%
 Below we will refer to this application
@@
 and hence suitable for MD applications.
-%
 However,
@@
 in controlling the moments.
-%
 We can nonetheless sample a wider range
@@
   park2007, *nguyen2013, *zhang2015st}.
-%
 While the mutual exchange probability
@@
 as the bias potential for inter-umbrella transitions.
-%
 Here,
@@
 and $z$ has also been written as a function of $\mathbf x$.
-%
 The transition probability from
@@
 where $z(t)$ is the current value of $z$.
-%
 While the asymptotic updating magnitude
@@
 components.
-%
 After simulation,
@@
 can be used to recover the unbiased distribution.
-%
-
-\note{The
-  following method appeared to have some stability issues,
-  so we switched back to the double-WL approach,
-  which hopefully is stable.
-  %
-  The reason of instability might be
-  that the inter-umbrella fluctuation
-  and the intra-umbrella fluctuation are some coupled.
-  So the updating magnitude cannot be adjusted individually
-  for the each umbrella.
-  %
-  Following the method for weight determination
-  in simulated tempering\cite{park2007,
-  *nguyen2013, *zhang2015st},
-  we can estimate $Z^{(j)}/Z^{(i)}$
-  for a neighboring pair, $(i, j)$, as
-  %
-  \begin{align*}
-    &\ln\frac{ Z^{(j)} }{ Z^{(i)} }
-    \approx
-    \ln\frac{ \sigma_z^{(j)} }{ \sigma_z^{(i)} }
-    +
-    \frac{ \Delta z_c } { 2 }
-    \left(
-      \frac{ c_1^{(i)} } { \sigma_z^{(i) 2} }
-      +
-      \frac{ c_1^{(j)} } { \sigma_z^{(j) 2} }
-    \right)
-    +
-    \frac{ c_2^{(j)} - c_2^{(i)} } { \sqrt 2 }
-    \\
-    &-
-    \left(
-      \frac{ \sqrt 2 \, c_2^{(j)} - 1 }
-           { 2 \, \sigma_z^{(j)} }
-      -
-      \frac{ \sqrt 2 \, c_2^{(i)} - 1 }
-           { 2 \, \sigma_z^{(i)} }
-    \right)
-    \left(
-      \frac{ \Delta z_c^{2} } { 6 }
-      +
-      \frac{ \sigma_z^{(i) 2} + \sigma_z^{(j) 2} } { 2 }
-    \right)
-    ,
-  \end{align*}
-  %
-  where
-  $\Delta z_c \equiv z_c^{(j)} - z_c^{(i)}$.
-  For equal width, we have
-  \begin{align*}
-    \ln\frac{ Z^{(j)} }{ Z^{(i)} }
-    \approx
-    \frac{ \Delta z_c \, ( c_1^{(i)} + c_1^{(j)} ) } { 2 \, \sigma_z }
-    -
-    \left(
-      \frac{ c_2^{(j)} - c_2^{(i)} }
-           { 6 \sqrt 2 }
-    \right)
-    \left(
-      \frac{ \Delta z_c^{2} } { \sigma_z^2 }
-    \right)
-    .
-  \end{align*}
-}
 
@@
 the distance, $r$, between two special particles.
-%
 The two special particles are mutually non-interacting
@@
 the cavity distribution function, $y(r)$\cite{hansen}.
-%
 We will assume reduced units in this calculation.
-%
 The reference bias potential was pre-calculated from a long simulation.
@@
 for configuration sampling.
-%
 Each step consists of
@@
 and another four trials for displacing the other particles.
-%
 In each trial,
@@
 We used $\varepsilon^* = 0.1/\varrho^*$
-to ensure reasonable acceptance ratio,
+to ensure a reasonable acceptance ratio,
 and $\varrho^*$ is the reduced number density.
@@
 on a grid of size $\delta r = 0.01$ in reduced units.
-%
 Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
 and the numbers of bins were $n = 500$ and $250$, respectively.
-%
 The reduced temperature was $T^* = 3$.
-%
 We tested the single-bin, Gaussian, and bandpass updating schemes.
-%
 Each simulation started with an ``equilibration'' phase
@@
 either the inverse-time or optimal schedule.
-%
 We repeated each simulation multiple times
 and report the average.
-%
 To make sure the same schedule for multiple runs under the same condition,
@@
 $\Gamma_k$'s and $\epsilon_k$'s.
-%
 The Gaussian updating scheme produced less error than
@@
 but not in the $\varrho^* = 0.8$ case.
-%
 The effectiveness of the Gaussian updating scheme
@@
 came from local noise.
-%
-%The optimal schedule also appeared to be more effective
-%in reducing the error than the inverse-time schedule.
-%%
 In this case, we could further reduced the error
@@
 (with the mode cutoff $K = 20$).
-%
 On the other hand, in the high-density case,
@@
     on the LJ system.
-    %
     The numbers have been averaged over multiple runs,
@@
     either the inverse-time ($1/t$) or optimal schedule (opt.).
-    %
-    \note{The reference values can be found
-    on the second line of \texttt{verr.log}.}%
   }
@@
     \\
-    %\hline
     $\sigma=0.1$ & $1/t$
@@
     \\
-    %\hline
     $\sigma=0.1$ & opt.
@@
     \\
-    %\hline
     $\sigma=0.2$ & $1/t$
@@
     \\
-    %\hline
     $\sigma=0.2$ & opt.
@@
     \\
-    %\hline
     $\sigma=0.5$ & $1/t$
@@
     \\
-    %\hline
     $\sigma=0.5$ & opt.
@@
     \\
-    %\hline
     $K=20$ & $1/t$
@@
 there was an optimal width for maximum effectiveness.
-%
 The optimal width was $\sigma = 0.37$ for $\varrho^* = 0.1$
@@
 and other parameters.
-%
 Although widening the Gaussian can
@@
 even with the optimal schedule.
-%
 This is because the rate of error reduction
@@
 with width $\sigma = 0.2$.
-%
 For modes $k=10$ to $25$,
@@
 although the initial values were less.
-%
 For the Gaussian updating scheme,
@@
 for these middle-range eigenmodes.
-%
 The optimal schedule improved over the inverse-time schedule
 and produced a flatter profile for the final error.
-%
 As a result, the optimal schedule usually produced
@@
 as shown in Table \ref{tab:lj_error}.
-%
 Although the gain of using the optimal schedule
@@
 based on histogram, Eq. \eqref{eq:uhatav}.
-%
 As shown in Fig. \ref{fig:lj_xerr},
@@
 the flatter profile seen in the single-bin case.
-%
 However, this correction also introduced
@@
 with detectable residual bias.
-%
 For example, for the simulation in the $\varrho^* = 0.8$ case
@@
     on the LJ system at density $\varrho^* = 0.1$.
-    %
     The optimal schedule (opt.) was more effective
@@
 we show the optimal schedules in a few cases.
-%
-%Generally, the optimal schedule
-%for the $\varrho^* = 0.8$ case
-%was closer to the inverse-time schedule
-%than that for the $\varrho^* = 0.1$ case.
-%%
-%This was probably because in the former case
-%the first mode dominates the total error,
-%$\Gamma_1 \approx \Gamma$,
-%and the schedule was roughly optimized for this mode,
-%resulting in
-%$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
-%as $\lambda_1 \approx 1$
-%[cf. Eq. \eqref{eq:lambda_Gaussian}].
-%
 Since we have chosen a long equilibration period
@@
 given by Eq. \eqref{eq:half_alpha0}.
-%
 But if the Gaussian were too wide,
@@
 [cf. Eq. \eqref{eq:opt_alpha0}].
-%
 In this case, the predicted schedule
@@
     for the LJ system.
-    %
     The starting updating magnitude was usually
@@
 in the energy space.
-%
 We spread the Gaussian umbrellas
@@
 with an even spacing of $L$.
-%
 For simplicity,
@@
 were performed.
-%
 Initially, we set $c_1 = 1.4 \sigma_E$, and $c_2 = 0$.
-%
 For configuration sampling in each umbrella,
@@
 whose average was no less than 92\%.
-%
 The updating magnitude started at $0.01$,
@@
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
-%
 This updating magnitude was
@@
 
-%To ensure stability,
-%we avoid transitions among umbrellas
-%until the current umbrella has
-%completed the WL stage-wise way of
-%reducing the updating magnitude
-%and started the inverse-time schedule.\cite{belardinelli2007,
-%*belardinelli2007jcp, *belardinelli2008, *belardinelli2016}
-
 The results for the $L = 64$ case
@@
 in Fig. \ref{fig:pt_hist}.
-%
 The Gaussian histograms were evenly spaced with similar width as intended,
-%and the overall energy histogram was roughly flat,
 as shown in Fig. \ref{fig:pt_hist}(a).
-%
 By WHAM, we found the critical temperature at $0.70162$
 in agreement with a previous study\cite{wang2001pre}.
-%
 The parameter, $c_1/\sigma_E$, roughly represents
@@
 as shown in Fig. \ref{fig:pt_hist}(b).
-%
 The second term of the bias potential,
@@
 provides a quadratic restraint on the distribution width.
-%
 The coefficient, $\sqrt 2 \, c_2$, represents the amount of ``pressure''
@@
     on the $64 \times 64$ ten-state Potts model.
-    %
     (a) Normalized histograms (solid lines)
@@
     at the critical temperature (dashed line).
-    %
     (b) The parameter, $c_1/\sigma_E$, roughly corresponding to
@@
     demonstrated the backbending behavior.
-    %
     (c) The parameter, $\sqrt 2 \, c_2$, exceeds $1.0$
@@
 that include the Wang-Landau (WL) algorithm and metadynamics.
-%
 Adaptive FDS methods
@@
 taking a Gaussian shape (as often in the metadynamics case).
-%
 We can generally characterize
@@
 window function for the bin of the column.
-%
 Our method computes the optimal schedule from
@@
 of the eigenmodes.
-%
 For the single-bin scheme with a set of uniform eigenvalues,
@@
 Eq. \eqref{eq:alpha_invt1}.
-%
 For a general multiple-bin scheme,
@@
 Eq. \eqref{eq:q_opt}.
-%
 In this case,
@@
 equipped with a set of rapidly diminishing eigenvalues.
-%
 While this feature helps suppress
@@
 in the short- and middle-range eigenmodes.
-%
 Thus, one needs to carefully choose the width of the Gaussian
@@
 for a long time before any reduction can happen.
-%
 The possible efficiency loss in
@@
 on the smoothness of the PMF.
-%
 This scheme can be generalized to
@@
 in the noise modes.
-%
 The optimal schedule of the bandpass schemes
 is always given by an inverse-time formula.
-%
 If the bandpass updating scheme was constructed
@@
 This study has several limitations.
-%
 First, we have ignored the \emph{systematic}
@@
 caused by the updates in adaptive FDS methods.
-%
 This simplification is justified in the asymptotic regime:
@@
 the magnitude itself\cite{morozov2007}.
-%
-%Our method of analysis is most useful for long simulations
-%to achieve relatively precise results.
-%
 Second, by using the white noise approximation,
@@
 of each eigenmode by a single number, $\Gamma_k$.
-%
 Although this assumption appeared to work well
@@
 of glassy systems.
-%
 Finally, the adaptive Gaussian umbrella sampling
@@
 its stability for general application.
-%
 For example, we find that with slow sampling processes,
@@
 and the parameters used here might not be optimal.
-%Our immediate interest is to apply these relations to
-%problems containing aqueous mixtures of proteins and nucleic acids.
 
@@
 for helpful discussions.
-%
 We gratefully acknowledge the Robert A. Welch Foundation (H-0037),
@@
 for partial support of this work.
-%
 JM thanks support from the National Institutes of Health
@@
 and National Science Foundation (Grant PHY-1427654).
-%
 This research used computing resources of
 the National Science Foundation XSEDE grid.
-%
 
@@
 would be given by
-%
 \begin{equation}
@@
 \end{equation}
-%
 Below we will use capital letters to denote
@@
 $X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)$.
-%\begin{equation}
-%  X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
-%  .
-%  \notag
-%  %\label{eq:timeav}
-%\end{equation}
-%
 Time averaging Eq. \eqref{eq:uhatt} yields
@@
 \end{align}
-%
-%Equation \eqref{eq:uhatav}
 This equation can also be derived\footnote{From Eq. \eqref{eq:uhatt},
@@
 after simulation.
-%
 In particular, for the single-bin updating scheme,
@@
 we have, up to a constant shift,
-%
 \begin{align}
@@
 - u_i(1)/[t \, \alpha(0)]$.
-%$$
-%\Delta u_i \equiv \frac{u_i(t+1)}{t \, \alpha(t)}
-%- \frac{u_i(1)}{t \, \alpha(0)}.
-%$$
-%
 With the inverse-time schedule, Eq. \eqref{eq:alpha_invt},
@@
 at every step, $t$.
-%
 On the other hand,
@@
 with $u_i(t+1) \approx u_i(1)$.
-\note{This would be rather risky for metadynamics,
-because we know that the certain noise modes
-do not easily go away, although the plain average
-formula was also recommended for metadynamics.}
 
@@
 but with a dynamic weight.
-%
 The weighted average
-%
 \begin{align*}
@@
 \end{align*}
-%
 where
@@
 is equivalent to the recurrence relation
-%
 \begin{align*}
@@
 from the data point at step $t$.
-%
 The relative weight, $\omega(t)/\Omega(t)$,
@@
 \end{equation}
-%
 The inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
@@
 respectively.
-%
 In the case of Eq. \eqref{eq:vkupdate},
@@
 as $\alpha^*(t)$.
-%
 This mapping shows that
 updating magnitude should not exceed $1.0$.
-%
 One way to satisfy this condition
@@
 with the inverse-time schedule.
-%
 With a sufficiently accurate and stable bias potential,
@@
 the instantaneous histogram is dominated by the random part,
-%
 \begin{equation}
@@
 $\tilde f_k(t) \approx \tilde \zeta_k(t).$
-%
 If we define the time average,
-%
 \begin{align*}
@@
 \end{align*}
-%
 for the histogram fluctuation of mode $k$,
 then the total fluctuation is
-%
 \begin{align}
@@
   \notag
-  %\label{eq:hfluc_def}
 \end{align}
-%
 From Eq. \eqref{eq:xi_zeta}, we have
-%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
-%=
-%\sum_{t=-\infty}^\infty
-%\frac{
-%\bigl\langle
-%  \tilde f_k(t) \, \tilde f_k(0)
-%\bigr\rangle } { T }
-%\approx
-%\sum_{t=-\infty}^\infty
-%\frac{
-%\bigl\langle
-%  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
-%\bigr\rangle } { T }
-%=
-%\frac{ \Gamma_k } { T },$$
-%
 \begin{align*}
@@
 \end{align*}
-%
 and the total histogram fluctuation is
 $\bigl\langle F^2(T) \bigr\rangle = \Gamma/T$.
-%
 Comparing this to Eq. \eqref{eq:Emin_sbin},
@@
   \notag
-  %\label{eq:fluc}
-\end{equation}
-%and the histogram fluctuation
-%serves as a good estimate of the final error.
-%
+\end{equation}
 
@@
 supports the WL criterion for stage transitions.
-%
-%For the bandpass updating scheme,
-%we may borrow this criterion
-%in initial stages, with a modified definition
-%for the histogram fluctuation
-%$F^2_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde F_k^2(T),$
-%since the histogram flattening
-%is only partially applied to modes less than $K$.
-%
 For a bandpass updating scheme,
@@
 from the largest mode
-%$\max_{k=1, \dots, K-1} \left| \, \tilde F_k(T) \, \right|$.
 \begin{align}
@@
 \end{align}
-%
 which allows straightforward extension
@@
 $\tilde F_k(T) = \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)$.
-%For a local sampling process,
-%we expect that the two definition would behave similarly.
-%
-\note{
-Note that the expected histogram flatness
-was previously used to define the error\cite{zhou2005, zhou2008},
-and the above demonstration shows the definition
-is compatible with ours, Eq. \eqref{eq:error_def}, in the long-time limit.
-An interesting application\cite{zhou2008},
-of the above result is
-to set the updating magnitude by $F/\Gamma$
-for some empirical value [$\Gamma \approx 10$ there, see Eq. (8)]
-as a substitute of the inverse-time schedule.
-}
 
@@
 under a constant updating magnitude.
-%
 This method has the advantage of not requiring
@@
 $u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
-%
 Since the shift,
@@
 should share the same variance.
-%
 But since the ensemble average of ${\tilde v}_{k}$
@@
  \left\langle {\tilde v}_{k}^2 \right\rangle$,
-%\begin{equation*}
-%  \operatorname{var} {\tilde u}_{k}
-%  =
-%  \operatorname{var} {\tilde v}_{k}
-%  =
-%  \left\langle {\tilde v}_{k}^2 \right\rangle
-%  .
-%\end{equation*}
-%
 and by using Eq. \eqref{eq:xt2_eql},
 we get
-%
 \begin{equation}
@@
 \end{equation}
-%
 where the variance can be computed from trajectory averages.
-%
 In practice, Eq. \eqref{eq:varXt}
@@
 with $\Lambda_l$ being the eigenvalue of mode $l$.
-%
 The autocorrelation function can then be computed as
-%
 \begin{align*}
@@
 \end{align*}
-%
 In particular, if the eigenvectors, $\pmb \phi_k$
@@
   \notag
-  %\label{eq:Gamma_est}
 \end{equation}
@@
 the first few $\Gamma_k$'s.
-%
 We will nonetheless use it as a model to
@@
 in a local sampling process.
-%
 If we borrow Eq. \eqref{eq:lambda_Gaussian}
@@
   \notag
-  %\label{eq:Gamma_local}
 \end{align}
@@
 for short ranges\cite{wang2001, wang2001pre}.
-%
-%Using finer bins would, however, not change this value.
-%
 For a modestly large $k$,
@@
 on the homogeneous updating schemes.
-%
 By the translational invariance,
@@
 is symmetric.
-%
 The updating matrices
@@
 by the eigenvalues.
-%
 Further, the multiplication of $\mathbf w$
 can be reduced to a convolution with the window function.
-%
 Thus,
@@
 
-%While translationally-invariant
-%updating schemes
-%naturally suit periodic variables\cite{
-%dama2014},
-%they can also be extended to non-periodic variables
-%by imposing the reflective boundary condition\cite{
-%bussi2006}.
-%%
-%For simplicity, we will assume that the target
-%distribution is flat, or $\rho_i = 1/n$, below.
-
 For a periodic variable\cite{dama2014},
@@
 assumes the following form
-%
 \begin{equation}
@@
 will be referred to as a window function.
-%
 To satisfy Eq. \eqref{eq:w_sumj},
 we impose the normalization
-%
 \begin{equation}
@@
 \end{equation}
-%
 with $\mu_l = 0$ for $|l| > b$.
-%
 If $b = n/2$, $\mu_{-b}$ is removed
@@
 to avoid double counting.
-%
 For simplicity, we also impose the symmetry
-%
 \begin{equation}
@@
 \end{equation}
-%
 such that $i \pm n$ lies in between $1$ and $n$.
-%
 Then, the multiplication of the matrix, $\mathbf w$,
 is equivalent to a convolution with the window function:
-%
 \begin{equation}
@@
 \end{equation}
-%
-\note{The derivation of the first step of
-  Eq. \eqref{eq:wmul_to_convol}:
-\begin{align*}
-  \sum_{j = 1}^n
-    w_{ij} \, \phi_j
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{j = 1}^n
-    \mu_{i - j - n} \, \phi_j
-  +
-  \sum_{j = 1}^n
-    \mu_{i - j + n} \, \phi_j
-  \\
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{l = 1+n}^{2 \, n}
-    \mu_{i - l} \, \phi_{l - n}
-  +
-  \sum_{l = 1-n}^0
-    \mu_{i - l} \, \phi_{l + n}
-  \\
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{l = 1+n}^{2 \, n}
-    \mu_{i - l} \, \phi_{l}
-  +
-  \sum_{l = 1-n}^0
-    \mu_{i - l} \, \phi_{l}
-  \\
-  &=
-  \sum_{j = 1-n}^{2 \, n}
-    \mu_{i - j} \, \phi_j
-  .
-\end{align*}
-The second step of Eq. \eqref{eq:wmul_to_convol}
-follows from the constraint, $\mu_l = 0$, for $|l| > b$.
-
-Similarly,
-for the left eigenvector, we have
-\begin{align*}
-  \sum_{ i = 1 }^n
-    \phi_i \, w_{ij}
-  =
-  \sum_{ l = -b }^b
-    \phi_{j - l} \, \mu_{-l}
-  .
-\end{align*}
-But due to the symmetry Eq. \eqref{eq:mu_symm},
-it is identical to \eqref{eq:wmul_to_convol}.
-}
-%
 As a result, the characteristic equation,
@@
 required by Eq. \eqref{eq:phi_pbc},
-%
 \begin{equation}
@@
   \frac{ \sqrt{ 2 - \delta_{k,0} } } { n }
-  %\overset{ \cos } { \sin }
   \,
@@
 \end{equation}
-%
 where the function takes the cosine form for $k \le n/2$,
 or the sine form otherwise.
-%
 Then one can readily verify that the eigenvalues are given by
@@
   $\lambda_{n - k} = \lambda_k$.
-\note{Consider the unnormalized exponential equivalent
-  \begin{align*}
-  \Phi^{(k)}_i =
-  \exp\left[
-    \frac{ k \, i \, 2 \, \pi }
-         {      n             }
-    \ii
-  \right]
-  ,
-  \end{align*}
-  we have
-  \begin{align*}
-  \sum_{j = 1}^n
-    w_{ij} \, \Phi^{(k)}_j
-  &=
-  \sum_{l = -b}^b
-    \mu_l \, \Phi^{(k)}_{i - l}
-  \\
-  &=
-  \mu_0 \, \Phi^{(k)}_i
-  +
-  \sum_{l = 1}^b
-    \mu_l \,
-    \left[ \Phi^{(k)}_{i - l} + \Phi^{(k)}_{i + l} \right]
-  \\
-  &=
-  \Phi^{(k)}_i \,
-  \left[
-    \mu_0
-    +
-    2 \sum_{l = 1}^b
-      \mu_l \, \cos
-      \frac{ k \, l \, 2 \, \pi }
-           {      n             }
-  \right]
-  .
-  \end{align*}
-  The term in the square brackets is the eigenvalue given by
-  Eq. \eqref{eq:wband_eigenvalue}.
-}
-%
 
@@
 to the equilibrium distribution\cite{bussi2006}.
-%such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
-%are satisfied.
-%
 Again, the window function should satisfy
@@
 
-%For a non-periodic variable,
 We define the out-of-boundary values as
-%
 \begin{equation}
@@
 \end{equation}
-%
 such that the matrix multiplication is still equivalent to
 a convolution as Eq. \eqref{eq:wmul_to_convol}.
-%
-\note{The derivation of Eq. \eqref{eq:wmul_to_convol}
-  in this case is similar,
-  \begin{align*}
-    \sum_{j = 1}^n w_{ij} \, \phi_j
-    &=
-    \sum_{j = 1}^n
-      \mu_{i - j} \, \phi_j
-    +
-    \sum_{j = 1}^n
-      \mu_{i + j - 1} \, \phi_j
-    +
-    \sum_{j = 1}^n
-      \mu_{i + j - 2 \, n - 1} \, \phi_j
-    \\
-    &=
-    \sum_{j = 1}^n
-      \mu_{i - j} \, \phi_j
-    +
-    \sum_{l = 1 - n}^0
-      \mu_{i - l} \, \phi_l
-    +
-    \sum_{l = n + 1}^{ 2 \, n }
-      \mu_{i - l} \, \phi_l
-    \\
-    &=
-    \sum_{j = 1 - n}^{ 2 \, n}
-      \mu_{i - j} \, \phi_j.
-  \end{align*}
-}%
-%
 Since Eq. \eqref{eq:phi_refl}
@@
 and be even functions about the same axis:
-%
 \begin{equation}
@@
 \notag
-%\label{eq:wband_eigenvector_refl}
-\end{equation}
-%
+\end{equation}
 and the eigenvalues are given by
@@
 
-\note{Derivation.
-  First consider the unnormalized eigenvectors,
-  \begin{align*}
-  \Phi^{(k)}_j
-  =
-  \cos \frac{ k \, \left( j - \frac 1 2 \right) \, \pi }{n},
-  \end{align*}
-  which satisfies Eq. \eqref{eq:phi_refl}.
-  %
-  So, by Eq. \eqref{eq:wmul_to_convol},
-  \begin{align*}
-  \sum_{i = 1}^n
-    \Phi^{(k)}_i \, w_{ij}
-  &=
-  \sum_{l = -b}^b
-    \Phi^{(k)}_{j - l} \, \mu_l
-  \\
-  &=
-    \Phi^{(k)}_j \, \mu_0
-  + \sum_{l=0}^{b}
-    \mu_l \,
-    \left[
-      \Phi^{(k)}_{j-l}
-      +
-      \Phi^{(k)}_{j+l}
-    \right]
-  \\
-  &= \Phi^{(k)}_j \, \lambda_k,
-  \end{align*}
-  with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}
-  with $g = 2$.
-}%
-%
-
-%\subsection{\label{sec:invert_wband}
-%Inversion}
-
 We can recover the window function from the eigenvalues
 by inversely transforming Eq. \eqref{eq:wband_eigenvalue},
-%
 \begin{equation}
@@
 \end{equation}
-%
 In the non-periodic case,
@@
 to satisfy the constraint, $\mu_n = 0$.
-%
-
-\note{On the inversion.
-Eq. \eqref{eq:lambdan}
-  ensures that the $\mu_n$
-  computed from Eq. \eqref{eq:mu_from_lambda}
-  vanishes.
-
-  %The inversion formula is derived from the cosine transform.
-  %
-  If we define $\mu_n = 0$ and
-  %
-  $
-    \mu_{ 2 n - l } = \mu_l,
-  $
-  for
-  $l = 1, \dots, n - 1$,
-  %
-  then Eq. \eqref{eq:wband_eigenvalue} ($g = 2$)
-  can be rewritten as
-  %
-  \begin{align}
-    \lambda_k
-    =
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-    \mu_l \, \cos \left(\frac{ k \, l \, \pi } { n } \right).
-  \notag
-  %\label{eq:lambda_cosine_sum}
-  \end{align}
-  %
-  This formula for $\lambda_k$
-  can be readily extended to $k = 2 \, n - 1$,
-  and we have
-  %
-  $
-    \lambda_{ 2 \, n - k } = \lambda_k
-    .
-  $
-  %
-  Thus,
-  %
-  \begin{align*}
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \lambda_k \,
-      \cos \frac{ k \, p \, \pi }
-                {      n        }
-    &=
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \sum_{ l = 0 }^{ 2 \, n - 1 }
-        \mu_l \,
-        \cos \frac{ k \, l \, \pi }
-                  {      n        }
-        \cos \frac{ k \, p \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \frac{ \mu_l } { 2 }
-      \sum_{ k = 0 }^{ 2 \, n - 1 }
-        \cos \frac{ k \, (p + l) \, \pi }
-                  {      n        }
-                  +
-        \cos \frac{ k \, (p - l) \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \mu_l \, n \left(
-        \delta_{ p + l - 2 \, n, 0 }
-        +
-        \delta_{ p - l, 0 }
-      \right)
-    \\
-    &=
-    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
-    =
-    2 \, n \, \mu_p.
-  \end{align*}
-  This entails Eq. \eqref{eq:mu_from_lambda}.
-  %
-  The value of $\lambda_n$
-  can be deduced from
-  imposing $\mu_n = 0$
-  and solving Eq. \eqref{eq:mu_from_lambda}
-  for $l = n$,
-  yielding Eq. \eqref{eq:lambdan}.
-}
 
@@
 Gaussian-shaped window function.
-%
 This updating scheme is commonly
 used in metadynamics.
-%
 Below we show its stability
@@
 $\mu_i = \mu(z) \, \delta z$.
-%
 By assuming the largest bin cutoff, $b \approx g \, n/2$,
@@
 as an integral:
-%
 \begin{equation}
@@
   \notag
-  %\label{eq:lambda_int}
-\end{equation}
-%
+\end{equation}
 with the normalization
@@
        { \sqrt{ 2 \pi } \sigma_z }
-  %
   \exp\left(
@@
 \end{equation}
-%
 Then, from Eq. \eqref{eq:wband_eigenvalue},
@@
 which shows that all eigenvalues are positive.
-%
 In practice, however, some eigenvalues can turn negative
 because of the truncation and discretization.
-%
 We can maintain the stability of the updating scheme
@@
 using Eq. \eqref{eq:mu_from_lambda}.
-%
 
@@
 it is convenient to set $\lambda_0 = 1$.
-%
 We can use Eq. \eqref{eq:mu_from_lambda}
 to find the window function.
-%
 For a periodic variable,
@@
 also to $1$.
-%
 Then, $\mu_l = \mathrm{sinr}(K, l, n)$ and $w_{ij} = \mu_{i-j}$,
@@
   / \left( n \, \sin \frac{ l \, \pi } { n } \right)$.
-%\begin{equation}
-%  \mathrm{sinr}(K, l, n)
-%  =
-%  \frac{
-%    \sin
-%    \frac{ (2 K - 1) \, l \, \pi }
-%         {              n        }
-%  }
-%  {
-%    n \, \sin \frac{ l \, \pi } { n }
-%  }
-%  .
-%\notag
-%%\label{eq:mu_sinc_pbc}
-%\end{equation}
-%
-\note{Derivation.
-\begin{align*}
-\mu_l
-&=
-\frac 1 n \sum_{k = 0}^{n-1} \lambda_k \cos \frac{ k \, l \, 2 \, \pi } { n }
-\\
-&=
-\frac{1}{n}
-\left(
-  1 +
-  \sum_{k=1}^{K-1}
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-  +
-  \sum_{k=n-K+1}^{n-1}
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-\right)
-\\
-&=
-\frac 1 n
-\sum_{k=1-K}^{K-1}
-\cos \frac { k \, l \, 2 \, \pi } { n }
-=
-  \frac{
-    \sin
-    \frac{ (2 K - 1) \, l \, \pi }
-         {              n        }
-  }
-  {
-    n \, \sin \frac{ l \, \pi } { n }
-  }
-.
-\end{align*}
-}
 
@@
 and $\mu_l = (-1)^{n-K+l}/(2 \, n) + \mathrm{sinr}(K, l, 2 \, n)$.
-%\begin{equation}
-%  \mu_l
-%  =
-%  (-1)^{n-K+l} / ( 2 \, n )
-%    +
-%  \mathrm{sinr}(K, l, 2n)
-%  .
-%  \notag
-%  %\label{eq:mu_sinc_refl}
-%\end{equation}
-%
 The updating matrix is given by
@@
   whichever exists, of the opposite sign.}
-%
-\note{Derivation.
-\begin{align*}
-  \lambda_{n}
-  &=
-  (-1)^{n-1}
-  \left[
-    \lambda_0
-    - 2 \, \lambda_1
-    + 2 \, \lambda_2 - \cdots
-    + (-1)^{K-1} 2 \, \lambda_{K-1})
-  \right]
-  \\
-  &=
-  \begin{dcases}
-    (-1)^{n-1} & K \mathrm{\; odd,} \\
-    (-1)^n     & K \mathrm{\; even.}
-  \end{dcases}
-\end{align*}
-So
-\begin{align*}
-  \mu_l
-  &=
-  \frac{1}{2\,n}
-  \left[
-    1 +
-    2 \sum_{k=1}^{K-1}
-    \cos \frac { k \, l \pi } { n }
-    +
-    (-1)^{n-K} (-1)^l
-  \right]
-  \\
-  &=
-  \frac{1}{2\,n}
-  \left[
-    (-1)^{n-K+l}
-    +
-    \sum_{k=1-K}^{K-1}
-    \cos \frac { k \, l \pi } { n }
-  \right]
-  .
-\end{align*}
-}%
-
-\note{The table of symbols is listed in Table \ref{tab:symbols}.
-  \begin{table*}
-  \footnotesize
-  \centering
-  \rowcolors{1}{white}{LightGray}
-  \setlength{\tabcolsep}{4pt} % column separation
-  \caption{\label{tab:symbols}
-    Table of symbols.}
-  \begin{tabular}{l | p{12cm} }
-    Symbol          &   Description \\
-    \hline
-    $\mathbf{A}$    &   Transition matrix. \\
-    $\alpha$        &   Updating magnitude. \\
-    $b$             &   Cutoff of the window function of homogeneous updating schemes. \\
-    $C_M$           &   Integral of the mass function, $M(\bar q)$.  \\
-    $\delta$        &   Dirac's or discrete $\delta$-function. \\
-    $c$             &   Constant. \\
-    $\Err, \Err_R, \Err_A$          &   Error, residual error, asymptotic error. \\
-    $\ln f$         &   Updating factor in the WL algorithm.  \\
-    $\phi_{ki}$     &   Eigenvectors of the updating matrix. \\
-    $g$             &   $1$ for a periodic variable, or for $2$ a non-periodic one. \\
-    $G_i$           &   Integral of the autocorrelation function of histogram fluctuation at bin $i$. \\
-    $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
-    $h_i(t)$        &   Instantaneous histogram.  \\
-    $H_i$           &   Average normalized histogram.  \\
-    $i, j$          &   Bin indices. \\
-    $\lambda_k$     &   The $k$th eigenvalue of the updating matrix. \\
-    $k, l$          &   Mode indices. \\
-    $K$             &   Cutoff wave number of bandpass updating schemes.  \\
-    $M(q)$          &   Mass function.   \\
-    $m(q)$          &   Mass distribution,
-                        $m(q) \equiv M(q)/\int_0^{ q(T) } M(q') \, dq'$.  \\
-    $\mu_i$         &   Updating window function of a homogeneous updating scheme. \\
-    $n$             &   Number of bins. \\
-    $\nu_k$         &   $\nu_k \equiv \lambda_k / \lambda$. \\
-    $\rho_i$        &   Flat sampling distribution. \\
-    $\pi_i$         &   (Instantaneous) distribution. \\
-    $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\
-    $\bar q(t)$     &   $q(T) - q(t)$.  \\
-    $s, S$          &   Stage index and the total number of stages in the Wang-Landau algorithm. \\
-    $\sigma_{ij}(t)$   &   Autocorrelation function of $\zeta_i(t)$. \\
-    $\sigma_\varphi$   &   Width of the Gaussian updating scheme in radians. \\
-    $t, \tau$       &   Time. \\
-    $T$             &   Simulation length. \\
-    $\vartheta_k(q')$       &   $\vartheta_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
-    $u_i(t)$        &   Original bias potential. \\
-    $v_i(t)$        &   Shifted bias potential. \\
-    $\mathbf w$     &   Updating matrix. \\
-    $v_{*i}(t)$     &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
-    $\zeta_{*i}(t)$ &   The noise of histogram. \\
-    ${\tilde v}_k$  &   Mode of the shifted bias potential. \\
-    ${\tilde u}_k$  &   Mode of the original bias potential. \\
-    $z$             &   Reaction coordinate. \\
-    $\zeta_i(t)$    &   The noise part of the instantaneous histogram, $h_i(t)$.
-  \end{tabular}
-  \end{table*}
-}
 

### 2017-06-28 prog/aus/reweight.c

@@ -225,7 +225,7 @@ static double seekcrit(const double *arr, int *x1, int *x2, double *shift)
 
 static int save(const char *fn, double bc, int x1, int x2)
 {
-  int ix;
+  int ix, i1, i0;
   FILE *fp;
   double x, beta, y, ys;
 
@@ -238,7 +238,12 @@ static int save(const char *fn, double bc, int x1, int x2)
   for ( ix = 0; ix < xn; ix++ ) {
     if ( htot[ix] <= 0 ) continue;
     x = xmin + ix * dx;
-    if ( ix < xn - 1 ) beta = (lngs[ix + 1] - lngs[ix])/dx;
+    i0 = ix - 1;
+    if ( i0 < 0 || htot[i0] <= 0 ) i0 = ix;
+    i1 = ix + 1;
+    if ( i1 >= xn || htot[i1] <= 0 ) i1 = ix;
+    if ( i0 < i1 ) beta = (lngs[i1] - lngs[i0])/(dx*(i1 - i0));
+    else beta = 0;
     y = exp(lng[ix] - bc*x);
     ys = exp(lngs[ix] - bc*x);
     fprintf(fp, "%g %g %g %g %g %g %g\n",

### 2017-08-07 doc/invt.tex
modify invt.tex

@@
   of the Langfeld-Lucini-Rago algorithm
-  and can be used for parameter tuning for the method of Gaussian ensemble
+  and can be used for parameter tuning for the Gaussian ensemble method
   and umbrella sampling.
@@
 %
-Further, we will show that
+We will also show that
 the single-bin updating scheme
@@
 into the bias potential during the updating process.
-%
-The optimal schedule of bandpass updating schemes
-is also an inverse-time schedule.
-%
-Constructed from orthogonal polynomials,
-bandpass updating schemes serve as
-generalizations of the Langfeld-Lucini-Rago (LLR)
-algorithm\cite{langfeld2012, pellegrini2014},
-and can help other FDS
-algorithms\cite{neuhaus2006, *neuhaus2007, zhu2012}
-adjust the parameters of the bias potential automatically.
+
+The bandpass updating schemes are convenient
+in that their optimal schedule is simply given by
+the inverse-time formula,
+and thus requires no extra calculation.
+%
+They can be constructed from any orthogonal basis\cite{bussi2006},
+and for a polynomial basis,
+this construction results in
+a class of adaptive FDS algorithms that
+update the bias potential based on the
+statistical moments of observed distribution
+as a generalization of the Langfeld-Lucini-Rago (LLR)
+algorithm\cite{langfeld2012, pellegrini2014}.
+%
+We show that a similar construction
+can help FDS algorithms with
+quadratic bias potentials\cite{neuhaus2006, *neuhaus2007, zhu2012}
+automatically adjust the parameters
+to reach a set of Gaussian distributions
+with desired means and variances.
 
@@
 %
-Finally, we compare different updating schemes
+We then compare different updating schemes
 and show that the single-bin scheme
@@
 in Sec. \ref{sec:cmpschemes}.
+%
+Finally,
+we use the bandpass updating schemes
+to build a class of adaptive FDS algorithms
+that based on the statistical moments.
 
@@
   -\alpha(t) \, \left[ v_{*i}(t) - \zeta_{*i}(t) \right]
-  ,
-  \notag
-  %\label{eq:dxdt_singlebin}
-\end{equation}
-%
-and the formal solution is
+  .
+  %\notag
+  \label{eq:dxdt_singlebin}
+\end{equation}
+%
+Without the random part, $\zeta_{*i}(t)$,
+this equation would drive
+the deviation of the bias potential, $v_{*i}(t)$,
+to $0$ at long times,
+implying the convergence.
+%
+The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
 %
@@
 
+We will first study the eigenvalue decomposition
+of the updating matrix, $\mathbf w$,
+which helps establish the convergence criterion
+and the decomposition of error into
+independent eigenmodes.
+%
 The matrix, $\mathbf w$, is subject to several conditions.
@@
   ,
-  \notag
-  %\label{eq:gft_def}
+  %\notag
+  \label{eq:gft_def}
 \end{equation}
@@
 and the inverse transform is
-$x_i = \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k / \rho_i$.
-%
-%\begin{equation}
-%  x_i = \frac{1}{\rho_i} \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k
-%  .
-%  \notag
-%  %\label{eq:gft_inv}
-%\end{equation}
+%$x_i = \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k / \rho_i$.
+%
+\begin{equation}
+  x_i = \frac{1}{\rho_i} \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k
+  .
+  %\notag
+  \label{eq:gft_inv}
+\end{equation}
 %
@@
 %
-We also have Parseval's theorem,
+From Eqs. \eqref{eq:gft_def} and \eqref{eq:gft_inv},
+we have Parseval's theorem,
 \begin{equation}
@@
   =
-  \sum_{i=1}^n x_{*i} \, y_{*i}
+  \sum_{i=1}^n \rho \, x_{*i} \, y_{*i}
+  .
   \label{eq:parseval}
 \end{equation}
+\note{Proof:
+$$
+  \sum_{k=0}^{n-1} \tilde x_{*k} \, \tilde y_{*k}
+  =
+  \sum_{k=0}^{n-1} \tilde x_{*k} \sum_{i=1}^n \phi_{ki} \, y_{*i}
+  =
+  \sum_{i=1}^n \left( \sum_{k=0}^{n-1} \tilde x_{*k} \, \phi_{ki} \right) \, y_{*i}
+  =
+  \sum_{i=1}^n \rho \, x_{*i} \, y_{*i}.
+$$
+}
 
@@
 we can rewrite the error defined in Eq. \eqref{eq:error_def} as
-$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
-%
-%\begin{align}
-%  \Err(T)
-%  =
-%  \sum_{k = 1}^{n - 1}
-%    \left\langle
-%      {\tilde v}_k^2(T)
-%    \right\rangle
-%  .
-%\notag
-%\end{align}
+a sum of the errors in individual modes\footnote{If
+the error weighting distribution differs from $\rho_i$
+in Eq. \eqref{eq:error_def},
+the error would contain quadratic cross terms among different eigenmodes.
+%
+While the solution for the optimal schedule retains the same form,
+the mass function defined in Eq. \eqref{eq:mass_func}
+will include the corresponding cross terms.
+}
+%$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
+%
+\begin{align}
+  \Err(T)
+  =
+  \sum_{k = 1}^{n - 1}
+    \left\langle
+      {\tilde v}_k^2(T)
+    \right\rangle
+  .
+  \notag
+\end{align}
 %
@@
 
-%\subsubsection{\label{sec:eqlerr}
-%Initial error
-%%from a simulation under a constant updating magnitude
-%}
-
-To model the initial error for Eq. \eqref{eq:error_res},
-we will assume that
-the preliminary adaptive FDS sampling has been
-performed under a constant updating magnitude for $T_0$ steps.
-%
-If the initial bias potential
-before the preliminary run is zero,
-one can show that
-\begin{equation}
-  \left\langle
-    {\tilde v}_k^2(0)
-  \right\rangle
-  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
-  + \epsilon_k
-  ,
-  \label{eq:xt2_eql1}
-\end{equation}
-%
-where
-$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$
-and if no eigenvalue is zero,
-then we can ignore the $\epsilon_k$ term
-in the case of $T_0 \to \infty$,
-%
-\begin{equation}
-  \left\langle
-    {\tilde v}_k^2(0)
-  \right\rangle
-  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
-  .
-  \label{eq:xt2_eql}
-\end{equation}
-%
-\note{
-Consider a long period, $T_0$, under a fixed schedule,
-$\alpha(t) = a_0$,
-the residual error becomes negligible, and
-the $k$th component of the asymptotic error
-is given by
-%
-\begin{align*}
-  \left\langle
-    {\tilde v}_k^2(T_0)
-  \right\rangle
-  =
-  \int_0^{T_0}
-    \Gamma_k \, (\lambda_k \, a_0)^2 \,
-      e^{ 2 \, \lambda_k \, a_0 \, (t - T_0) }
-    \, dt
-  \stackrel{ T_0 \to \infty }
-  { =\joinrel=\joinrel=\joinrel= }
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
-  .
-\notag
-%\label{eq:error_eql}
-\end{align*}
-}
-%
-
 \subsubsection{\label{sec:optschedule}
@@
 
-\subsubsection{\label{sec:mass_distr}
-Characterization of optimal schedules}
-
-We can characterize the optimal schedule
-geometrically as follows.
-%
-First, we define
-a normalized mass distribution as
-%
-\begin{equation}
-  m(\bar q)
-  =
-  \frac{
-    M(\bar q)
-  }
-  {
-    \int_0^{ q(T) } M(q') \, d q'
-  }
-  =
-  \frac{
-    M(\bar q)
-  }
-  {
-    C_M
-  }
-  ,
-\notag
-%\label{eq:mass_distr}
-\end{equation}
-%
-such that
-$\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
-%
-Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
-we find that for $\bar q(t) \equiv q(T) - q(t)$,
-%
-\begin{align}
-  m\left( \bar q(t) \right)
-  &=
-  \frac{ 1 }
-       { T \, \alpha(t) }
-  ,
-  \label{eq:mQ_invTa}
-  \\
-  \int_{\bar q(t)}^{ q(T) }
-    m(q') \, d q'
-  &=
-  \frac t T
-  ,
-  \label{eq:intmQ_tT}
-\end{align}
-%
-Thus, each point on the schedule,
-$\bigl(t, \alpha(t)\bigr)$,
-can be mapped to a point,
-$\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
-on the mass distribution,
-such that the ordinate of the latter curve
-is $1/(T\alpha)$,
-and the area under the curve in the domain, $[\bar q(t), q(T)]$,
-is equal to $t/T$,
-as shown in Fig. \ref{fig:massq}(a).
-
-\begin{figure}[h]
-\begin{center}
-  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
-    \node [] at (-0.3, 2.0) {(a)};
-
-    % distribution
-    \draw[domain=0:2.0, variable=\x, dashed, gray, very thick]
-      plot({\x}, {2*exp(-\x)});
-
-    % shade
-    \fill [gray!30!white, domain=0.7:1.609, variable=\x]
-      (0.7, 0) --
-      plot ({\x}, {2*exp(-\x)})
-      -- (1.609, 0) -- cycle;
-
-    % thick curve
-    \draw[domain=0:1.609, variable=\x, very thick]
-      plot({\x}, {2*exp(-\x)});
-
-    % label for the shaded area
-    \node[] at (1.1, 0.32) {$\dfrac{t}{T}$};
-
-    \draw[<->] (0.6, 0)  --
-               node[left, inner sep=1mm]
-               {$\dfrac{1}{T \, \alpha}$}
-               (0.6, 0.99317);
-    % label for the ordinate
-    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
-
-    \draw[] (0.5, 0.99317) -- (0.7, 0.99317);
-    \draw[->] (0, 0) -- (2.2, 0);
-    \node[] at (2.2, -0.2) {$\bar q$};
-    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 2.2);
-
-    % xtics
-    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
-    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$q(T) - q(t)$};
-    \draw[] (1.609, 0.1) -- (1.609, 0.0) node[below] {$q(T)$};
-  \end{tikzpicture}
-%
-  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
-    \node [] at (-0.3, 2.0) {(b)};
-    % distribution
-    \draw[domain=0:2.3, variable=\x, dashed, gray, very thick]
-      plot({\x}, {1/(\x+0.5)});
-
-    % shade
-    \fill [gray!30!white, domain=0.7:2.0, variable=\x]
-      (0.7, 0) --
-      plot ({\x}, {1/(\x+0.5)})
-      -- (2.0, 0) -- cycle;
-
-    % thick curve
-    \draw[domain=0:2.0, variable=\x, very thick]
-      plot({\x}, {1/(\x+0.5)});
-
-    % label for the shaded area
-    \node[] at (1.1, 0.35) {$\bar q(t)$};
-
-    \draw[<->] (0.6, 0)  --
-               node[left, inner sep=1mm]
-               {$\dfrac{1}{T \, m}$}
-               (0.6, 0.83333);
-    % label for the ordinate
-    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
-
-    \draw[] (0.5, 0.83333) -- (0.7, 0.83333);
-    \draw[->] (0, 0) -- (2.6, 0);
-    \node[] at (2.6, -0.2) {$t$};
-    \draw[->] (0, 0) -- node[above, rotate=90] {$\alpha(t)$} (0, 2.2);
-
-    % xtics
-    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
-    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$t$};
-    \draw[] (2.0, 0.1) -- (2.0, 0.0) node[below] {$T$};
-  \end{tikzpicture}
-  \caption{
-    \label{fig:massq}
-    (a) Geometric construction of the optimal schedule,
-    $\alpha(t)$,
-    from the normalized mass distribution,
-    $m(\bar q)$.
-    %
-    (b) Inverse construction of $m(\bar q)$
-    from the optimal schedule $\alpha(t)$.
-    %
-    \note{
-      The figure was produced by \texttt{tikz}.
-    }%
-  }
-\end{center}
-\end{figure}
-
-The schedule, $\alpha(t)$,
-also defines a variable transformation from
-the mass distribution, $m(\bar q)$.
-%much like the Legendre transform.
-%
-If we define
-$\bar q^* \equiv \bar q/q(T)$
-and $t^* \equiv t/T$
-such that both are limited to the interval $[0, 1]$,
-as well as
-\begin{align*}
-  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
-  \\
-  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
-\end{align*}
-then it is clear that
-both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
-can describe the differential relation
-between $\bar q^*$ and $t^*$,
-but with the natural variables
-being $\bar q^*$ and $t^*$, respectively.
-%
-Further, the transformation is an involution:
-if we apply the same transform on $\alpha^*(t^*)$
-we would recover $m^*(\bar q^*)$,
-as shown in Fig. \ref{fig:massq}(b).
-%
-A few examples are shown in Table \ref{tab:m_and_a}.
-%
-For example, in the single-bin scheme,
-the mass distribution decays exponentially,
-and we would have the inverse-time schedule.
-%
-On the other hand,
-if the mass distribution is inversely proportional to $\bar q$,
-which is approximately true for the Gaussian updating scheme
-at large $\bar q$,\footnote{If
-  the magnitudes of the eigenvalues, $\lambda_k$'s,
-  span a wide range, then for a large $\bar q$,
-  the sum in Eq. \eqref{eq:mass_func} will be dominated by
-  the largest term, where $\lambda_k \approx 1/\bar q$,
-  and $M(\bar q) \approx \sqrt{\Gamma_k}/\bar q$.
-  }
-  %
-then we would have an exponentially decaying schedule.
-
-\begin{table}[h]
-  \caption{\label{tab:m_and_a}
-    Examples of the variable transformation between the
-    reduced schedule, $\alpha^*(t^*)$,
-    and the reduced mass distribution, $m^*(\bar q^*)$,
-    as defined in Eqs. \eqref{eq:mQ_invTa}
-    and \eqref{eq:intmQ_tT}.
-  }
-  \setlength{\tabcolsep}{2pt} % spacing between columns
-  \renewcommand\arraystretch{2.0} % spacing between rows
-  \begin{tabular} { c c c }
-    \hline
-    $m^*(\bar q^*)$ &
-    $\alpha^*(t^*)$ &
-    $\bar q^*(t^*)$
-    \\
-    \hline
-    $\dfrac{1}{\gamma \, ({\bar q}^* + s_\gamma) }$ &
-    $(1 + s_\gamma) \, \gamma \, e^{-\gamma \, t^*}$ &
-    $(1 + s_\gamma) \, e^{-\gamma \, t^*} - s_\gamma$
-    \\
-    $(1+ s_\gamma) \, \gamma \, e^{-\gamma \, {\bar q}^*}$ &
-    $\dfrac{1}{ \gamma \, (t^* + s_\gamma) }$ &
-    $\dfrac{1}{\gamma} \ln\left(\dfrac{1 + s_\gamma}{t^* + s_\gamma} \right)$
-    \\
-    $\dfrac{c \, a}{\left( {\bar q}^* + q^*_0 \right)^{a+1}}$ &
-    $\dfrac{c^{1/a} \, (1/a)}{\left( t^* + t^*_0 \right)^{(1/a)+1}}$ &
-    %$\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a} - q^*_0$
-    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^{1/a} - 1\right]$
-    %$\,\begin{aligned}
-    %  &a, q^*_0 > 0, \\
-    %  &\frac{1}{t^*_0} \equiv \left(1+\frac{1}{q^*_0}\right)^a - 1, \\
-    %  &c \equiv {q^*_0}^a(t^*_0 + 1)
-    %\end{aligned}$
-    \\
-    \hline
-    \multicolumn{3}{p{8cm}}{
-    Here,
-    $\gamma > 0$,
-    and we have defined
-    $s_\gamma \equiv 1/(e^\gamma - 1)$,
-    ${t^*_0}^{-1} \equiv \left(1+{q^*_0}^{-1}\right)^a - 1$,
-    and
-    $c \equiv {q^*_0}^a(1 + t^*_0)$.
-    } \\
-    \hline
-  \end{tabular}
-\end{table}
+%\subsubsection{\label{sec:eqlerr}
+%Initial error
+%%from a simulation under a constant updating magnitude
+%}
 
@@
 %
-We will consider simulations in which
-the system is initially equilibrated
-at a constant updating magnitude, $a_0$.
+To model the initial error for Eq. \eqref{eq:error_res},
+we will assume that
+the system undergoes a preliminary FDS simulation
+under a constant updating magnitude, $a_0$,
+for $T_0$ steps.
+%
+If the initial bias potential
+before the preliminary run is zero,
+one can show that
+\begin{equation}
+  \left\langle
+    {\tilde v}_k^2(0)
+  \right\rangle
+  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
+  + \epsilon_k
+  ,
+  \label{eq:xt2_eql1}
+\end{equation}
+%
+where
+$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$
+and if no eigenvalue is zero,
+then we can ignore the $\epsilon_k$ term
+in the case of $T_0 \to \infty$,
+%
+\begin{equation}
+  \left\langle
+    {\tilde v}_k^2(0)
+  \right\rangle
+  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
+  .
+  \label{eq:xt2_eql}
+\end{equation}
+%
+\note{
+Consider a long period, $T_0$, under a fixed schedule,
+$\alpha(t) = a_0$,
+the residual error becomes negligible, and
+the $k$th component of the asymptotic error
+is given by
+%
+\begin{align*}
+  \left\langle
+    {\tilde v}_k^2(T_0)
+  \right\rangle
+  =
+  \int_0^{T_0}
+    \Gamma_k \, (\lambda_k \, a_0)^2 \,
+      e^{ 2 \, \lambda_k \, a_0 \, (t - T_0) }
+    \, dt
+  \stackrel{ T_0 \to \infty }
+  { =\joinrel=\joinrel=\joinrel= }
+  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  .
+\notag
+%\label{eq:error_eql}
+\end{align*}
+}
+%
+
 %
@@
 We can rewrite the condition in terms of
-the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
+the initial updating magnitude by Eq. \eqref{eq:Lagrangian_const} %\eqref{eq:mQ_invTa},
 \begin{align}
   \alpha(0)
-  =
-  \frac{1}{T \, m\bigl( q(T) \bigr) }
+  %=
+  %\frac{1}{T \, m\bigl( q(T) \bigr) }
   =
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-that completely eliminates certain short-wavelength eigenmodes,
-which are typically local noise.
-%
-A bandpass updating scheme is essentially
-the single-bin updating scheme
-accompanied by a perfect filter
-for the designated noise modes,
-which are assigned with zero eigenvalues.
-%
-For example,
-by assuming
-that the PMF is so smooth
-that it can be represented by the first $K$
-modes,\footnote{As an estimate of error,
+that completely discard certain short-wavelength eigenmodes
+that are deemed as local noise.
+%
+If the eigenmodes are sorted in descending order
+of the wavelength,
+and the PMF is sufficiently smooth
+to be representable by the first $K$
+eigenmodes,\footnote{As an estimate of error,
   we note that truncating a linear or quadratic function
@@
   $|\max_i u_i - \min_i u_i|^2/K^3$.}
-we demand
+we may construct an updating scheme
+from Eq. \eqref{eq:w_from_phi} with
 %
@@
 %%
+With the eigenvalues for the non-noise and noise modes being
+$1$ and $0$, respectively,
+a bandpass updating scheme is essentially
+the single-bin updating scheme
+accompanied by a perfect filter
+for the designated noise modes.
+%
 The asymptotic error is then
@@
 %
-The optimal schedule of these schemes
-remains the inverse-time schedule
-given by Eq. \eqref{eq:alpha_invt1}
-with $t_0$ given by Eq. \eqref{eq:t0_sbin}.
-%
-%The asymptotic error is given by
-%%
-%\begin{equation}
-%  \Err_A(T)
-%  =
-%  \frac{   T     }
-%       { (T + t_0)^2 }
-%  \sum_{ k = 1 }^{ K - 1 }
-%    \Gamma_k
-%  ,
-%  \notag
-%  %\label{eq:error_sinc}
-%\end{equation}
-%%
-%as a generalization of Eq. \eqref{eq:Emin_sbin}
-%for the single-bin updating scheme.
-%
-
-Since bandpass updating schemes are
-straightforward generalizations of
-the single-bin updating scheme,
-we can also borrow the WL algorithm
-to control the updating magnitude
-in initial stages of the simulation.
-%
-However, the histogram fluctuation should
-be limited to first $K$ modes,
-as the bandpass updating scheme
-only partially flattens these histogram
-modes (cf. Appendix \ref{sec:hfluc}).
+
+Comparing Eq. \eqref{eq:lambda_bandpass}
+with Eq. \eqref{eq:lambda_Gaussian},
+we find that the bandpass updating scheme serves
+as an idealization of the Gaussian updating scheme
+with the Gaussian spectrum of eigenvalues
+being replaced by a step function.
+%
+Although both updating schemes are able to remove local noise
+from the bias potential,
+the bandpass updating schemes have the advantage
+of allowing convenient control of the updating magnitude.
+%
+In the asymptotic regime,
+the optimal schedule is simply given
+by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
+%with $t_0$ given by Eq. \eqref{eq:t0_sbin},
+just like the single-bin scheme,
+and it is therefore unnecessary to go through
+the procedure in Sec. \ref{sec:procedure}
+for a general updating scheme.
+%
+In initial stages, when the inverse-time schedule
+may be inapplicable,
+we can instead use the stage-wise WL technique
+as long as the histogram fluctuation
+used for stage transitions
+is computed from the first $K$ modes
+(cf. Appendix \ref{sec:hfluc}).
+
+\subsection{\label{sec:bandpass_poly}
+Bandpass updating schemes constructed from polynomial bases}
 
@@
 was modest in the reported cases,
-it would increase with the simulation length.\footnote{
+it would increase with the simulation length.
+%
 For example, for the $\varrho^* = 0.1$ case
 with width $\sigma = 0.2$,
-  the predicted error of a simulation of $10^8$ steps
-  was $2.5\times10^{-5}$ for the inverse-time schedule
-  or $3.7\times10^{-6}$ for the optimal schedule.}
+if we increase the simulation length to $10^8$ steps,
+the error for the optimal schedule was
+$2.9\times10^{-6}$,
+which was much smaller than that
+for the inverse-time schedule
+$1.8\times10^{-5}$.
+\note{The predicted errors were
+$3.7\times10^{-6}$ and $1.8\times10^{-5}$, respectively.}
 
@@
 cf. Fig. \ref{fig:lj_xerr}.
+
+\section{\label{sec:schedule_geometry}
+Characterization of optimal schedules}
+
+We can characterize the optimal schedule
+geometrically as follows.
+%
+First, we define
+a normalized mass distribution as
+%
+\begin{equation}
+  m(\bar q)
+  =
+  \frac{
+    M(\bar q)
+  }
+  {
+    \int_0^{ q(T) } M(q') \, d q'
+  }
+  =
+  \frac{
+    M(\bar q)
+  }
+  {
+    C_M
+  }
+  ,
+\notag
+%\label{eq:mass_distr}
+\end{equation}
+%
+such that
+$\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
+%
+Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
+we find that for $\bar q(t) \equiv q(T) - q(t)$,
+%
+\begin{align}
+  m\left( \bar q(t) \right)
+  &=
+  \frac{ 1 }
+       { T \, \alpha(t) }
+  ,
+  \label{eq:mQ_invTa}
+  \\
+  \int_{\bar q(t)}^{ q(T) }
+    m(q') \, d q'
+  &=
+  \frac t T
+  ,
+  \label{eq:intmQ_tT}
+\end{align}
+%
+Thus, each point on the schedule,
+$\bigl(t, \alpha(t)\bigr)$,
+can be mapped to a point,
+$\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
+on the mass distribution,
+such that the ordinate of the latter curve
+is $1/(T\alpha)$,
+and the area under the curve in the domain, $[\bar q(t), q(T)]$,
+is equal to $t/T$,
+as shown in Fig. \ref{fig:massq}(a).
+
+\begin{figure}[h]
+\begin{center}
+  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
+    \node [] at (-0.3, 2.0) {(a)};
+
+    % distribution
+    \draw[domain=0:2.0, variable=\x, dashed, gray, very thick]
+      plot({\x}, {2*exp(-\x)});
+
+    % shade
+    \fill [gray!30!white, domain=0.7:1.609, variable=\x]
+      (0.7, 0) --
+      plot ({\x}, {2*exp(-\x)})
+      -- (1.609, 0) -- cycle;
+
+    % thick curve
+    \draw[domain=0:1.609, variable=\x, very thick]
+      plot({\x}, {2*exp(-\x)});
+
+    % label for the shaded area
+    \node[] at (1.1, 0.32) {$\dfrac{t}{T}$};
+
+    \draw[<->] (0.6, 0)  --
+               node[left, inner sep=1mm]
+               {$\dfrac{1}{T \, \alpha}$}
+               (0.6, 0.99317);
+    % label for the ordinate
+    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
+
+    \draw[] (0.5, 0.99317) -- (0.7, 0.99317);
+    \draw[->] (0, 0) -- (2.2, 0);
+    \node[] at (2.2, -0.2) {$\bar q$};
+    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 2.2);
+
+    % xtics
+    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
+    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$q(T) - q(t)$};
+    \draw[] (1.609, 0.1) -- (1.609, 0.0) node[below] {$q(T)$};
+  \end{tikzpicture}
+%
+  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
+    \node [] at (-0.3, 2.0) {(b)};
+    % distribution
+    \draw[domain=0:2.3, variable=\x, dashed, gray, very thick]
+      plot({\x}, {1/(\x+0.5)});
+
+    % shade
+    \fill [gray!30!white, domain=0.7:2.0, variable=\x]
+      (0.7, 0) --
+      plot ({\x}, {1/(\x+0.5)})
+      -- (2.0, 0) -- cycle;
+
+    % thick curve
+    \draw[domain=0:2.0, variable=\x, very thick]
+      plot({\x}, {1/(\x+0.5)});
+
+    % label for the shaded area
+    \node[] at (1.1, 0.35) {$\bar q(t)$};
+
+    \draw[<->] (0.6, 0)  --
+               node[left, inner sep=1mm]
+               {$\dfrac{1}{T \, m}$}
+               (0.6, 0.83333);
+    % label for the ordinate
+    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
+
+    \draw[] (0.5, 0.83333) -- (0.7, 0.83333);
+    \draw[->] (0, 0) -- (2.6, 0);
+    \node[] at (2.6, -0.2) {$t$};
+    \draw[->] (0, 0) -- node[above, rotate=90] {$\alpha(t)$} (0, 2.2);
+
+    % xtics
+    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
+    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$t$};
+    \draw[] (2.0, 0.1) -- (2.0, 0.0) node[below] {$T$};
+  \end{tikzpicture}
+  \caption{
+    \label{fig:massq}
+    (a) Geometric construction of the optimal schedule,
+    $\alpha(t)$,
+    from the normalized mass distribution,
+    $m(\bar q)$.
+    %
+    (b) Inverse construction of $m(\bar q)$
+    from the optimal schedule $\alpha(t)$.
+    %
+    \note{
+      The figure was produced by \texttt{tikz}.
+    }%
+  }
+\end{center}
+\end{figure}
+
+The schedule, $\alpha(t)$,
+also defines a variable transformation from
+the mass distribution, $m(\bar q)$.
+%much like the Legendre transform.
+%
+If we define
+$\bar q^* \equiv \bar q/q(T)$
+and $t^* \equiv t/T$
+such that both are limited to the interval $[0, 1]$,
+as well as
+\begin{align*}
+  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
+  \\
+  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
+\end{align*}
+then it is clear that
+both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
+can describe the differential relation
+between $\bar q^*$ and $t^*$,
+but with the natural variables
+being $\bar q^*$ and $t^*$, respectively.
+%
+Further, the transformation is an involution:
+if we apply the same transform on $\alpha^*(t^*)$
+we would recover $m^*(\bar q^*)$,
+as shown in Fig. \ref{fig:massq}(b).
+%
+A few examples are shown in Table \ref{tab:m_and_a}.
+%
+For example, in the single-bin scheme,
+the mass distribution decays exponentially,
+and we would have the inverse-time schedule.
+%
+On the other hand,
+if the mass distribution is inversely proportional to $\bar q$,
+which is approximately true for the Gaussian updating scheme
+at large $\bar q$,\footnote{If
+  the magnitudes of the eigenvalues, $\lambda_k$'s,
+  span a wide range, then for a large $\bar q$,
+  the sum in Eq. \eqref{eq:mass_func} will be dominated by
+  the largest term, where $\lambda_k \approx 1/\bar q$,
+  and $M(\bar q) \approx \sqrt{\Gamma_k}/\bar q$.
+  }
+  %
+then we would have an exponentially decaying schedule.
+
+\begin{table}[h]
+  \caption{\label{tab:m_and_a}
+    Examples of the variable transformation between the
+    reduced schedule, $\alpha^*(t^*)$,
+    and the reduced mass distribution, $m^*(\bar q^*)$,
+    as defined in Eqs. \eqref{eq:mQ_invTa}
+    and \eqref{eq:intmQ_tT}.
+  }
+  \setlength{\tabcolsep}{2pt} % spacing between columns
+  \renewcommand\arraystretch{2.0} % spacing between rows
+  \begin{tabular} { c c c }
+    \hline
+    $m^*(\bar q^*)$ &
+    $\alpha^*(t^*)$ &
+    $\bar q^*(t^*)$
+    \\
+    \hline
+    $\dfrac{1}{\gamma \, ({\bar q}^* + s_\gamma) }$ &
+    $(1 + s_\gamma) \, \gamma \, e^{-\gamma \, t^*}$ &
+    $(1 + s_\gamma) \, e^{-\gamma \, t^*} - s_\gamma$
+    \\
+    $(1+ s_\gamma) \, \gamma \, e^{-\gamma \, {\bar q}^*}$ &
+    $\dfrac{1}{ \gamma \, (t^* + s_\gamma) }$ &
+    $\dfrac{1}{\gamma} \ln\left(\dfrac{1 + s_\gamma}{t^* + s_\gamma} \right)$
+    \\
+    $\dfrac{c \, a}{\left( {\bar q}^* + q^*_0 \right)^{a+1}}$ &
+    $\dfrac{c^{\frac 1 a} / a}{\left( t^* + t^*_0 \right)^{\frac{1}{a}+1}}$ &
+    %$\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a} - q^*_0$
+    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^\frac{1}{a} - 1\right]$
+    %$\,\begin{aligned}
+    %  &a, q^*_0 > 0, \\
+    %  &\frac{1}{t^*_0} \equiv \left(1+\frac{1}{q^*_0}\right)^a - 1, \\
+    %  &c \equiv {q^*_0}^a(t^*_0 + 1)
+    %\end{aligned}$
+    \\
+    \hline
+    \multicolumn{3}{p{8cm}}{
+    Here,
+    $\gamma > 0$,
+    and we have defined
+    $s_\gamma \equiv 1/(e^\gamma - 1)$,
+    ${t^*_0}^{-1} \equiv \left(1+{q^*_0}^{-1}\right)^a - 1$,
+    and
+    $c \equiv {q^*_0}^a(1 + t^*_0)$.
+    } \\
+    \hline
+  \end{tabular}
+\end{table}
 

### 2017-08-12 doc/invt.tex
change invt.tex

@@
 and for convergence of the estimated PMF
-one has to gradually decrease
+one needs to gradually decrease
 the updating magnitude of the bias.
@@
 The schedule of reducing
-the updating magnitude,
-therefore, critically affects
+the updating magnitude
+can critically affect
 the precision of the final bias potential,
@@
 barducci2008, dickson2011, dama2014}.
+%
+While a slow reduction leads to large variation of the bias potential,
+a rapid reduction can converge the bias potential
+to a wrong value.
 %
@@
 can be deduced from the equation of motion
-of the particle that minimizes the action.
+that minimizes the action.
 %
@@
 %
-%In either case, however,
-%Ideally,
-%the optimal initial updating magnitude
-%can be set to
-%half of the previous equilibrium value,
-%which amounts to a shift of the origin of time
-%in the single-bin case.
-%
-We will also show that
+Further, we will compare
+the highest efficiency of error reduction
+of different updating schemes.
+%
+We will show that
 the single-bin updating scheme
-is optimal for asymptotic convergence,
+is optimal for long simulations,
 and it can be generalized
-to a class of asymptotically optimal
-bandpass updating schemes.
+to a class of bandpass updating schemes
+that are equally optimal.
 %
@@
 a few long-range histogram modes,
-and to avoid introducing local noise
-into the bias potential during the updating process.
+and to filter local noise
+during the updating process.
 
@@
 
-The difficulty of sampling a multimodal distribution
-with deep valley(s) can sometimes be eased by
-sampling a smoother biased distribution, $\rho_i$.
-%
+Sampling a multimodal distribution with deep valley(s)
+is often challenging in traditional simulations,
+and a common remedy is to apply a bias potential
+to sample a smoother distribution, $\rho_i$.
 %
@@
 %
-The biased equilibrium distribution
+Since the equilibrium distribution
 under a bias potential, $u_i$, is given by
@@
 %
-and our aim is to find a bias potential
-such that the resulting $\pi_i$ would coincide
+our aim is to find a bias potential that allows
+the resulting $\pi_i$ to coincide
 with the desired $\rho_i$.
@@
 %
-In a class of equilibrium FDS methods\cite{
+In an earlier class of equilibrium FDS methods\cite{
   mezei1987, berg1992, *lee1993, marinari1992, *lyubartsev1992},
@@
 
-Below we will focus on another class of adaptive FDS methods,
+Below we will focus on a more recent class of adaptive FDS methods,
 including the WL algorithm\cite{wang2001, wang2001pre}
 and metadynamics\cite{huber1994, laio2002, *laio2008, *barducci2011, *sutto2012},
-address the above shortcoming
+which address the above shortcoming
 by continuously and incrementally updating the bias potential.
@@
 %
-We refer to this scheme as the \emph{single-bin scheme},
+We refer to this scheme as
+the \emph{single-bin (updating) scheme},
 for it applies only to the bias potential
@@
 %
-In a more general \emph{multiple-bin scheme},
+In a more general
+\emph{multiple-bin (updating) scheme},
 %including the Gaussian updating scheme used in metadynamics,
@@
 The Gaussian updating scheme used in metadynamics
-is an example, with $w_{ij}$ being a Gaussian centered on $j$.
+is an example, with $w_{ij}$ being
+roughly a Gaussian centered on bin $j$.
 %
@@
   ,
-  \notag
-  %\label{eq:v_def}
+  \label{eq:v_def}
 \end{equation}
@@
 \begin{equation}
-  \bav{f}(t) = \sum_{i=1}^n h_i(t) = 1
+  \bav{f}(t)
+  = \sum_{i=1}^n \rho_i \cdot \frac{ h_i(t) } { \rho_i }
+  = \sum_{i=1}^n \delta_{i, i(t)}
+  = 1
   ,
@@
 %
-If $\alpha(t)$ is a constant, $a_0$,
-so that $q(t) = a_0 \, t$,
-then after a long period,
-the error approach an equilibrium value
-%
-\begin{equation}
-  \Err_\mathrm{equil.}
-  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
-  = \frac{1}{2} a_0 \Gamma
-  .
-  \label{eq:error_equil_sbin}
-\end{equation}
+%If $\alpha(t)$ is a constant, $a_0$,
+%so that $q(t) = a_0 \, t$,
+%then after a long period,
+%the error approach an equilibrium value
+%%
+%\begin{equation}
+%  \Err_\mathrm{equil.}
+%  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
+%  = \frac{1}{2} a_0 \Gamma
+%  .
+%  \label{eq:error_equil_sbin}
+%\end{equation}
 %
@@
 %
-We now fix the endpoint value, $q(T)$,
-and vary the curve, $q(t)$ ($0 < t < T$).
-%
-This is equivalent to varying the curve,
-$\vartheta\bigl( q(t) \bigr)$,
-with fixed endpoints,
+It is convenient to
+vary the curve, $q(t)$ ($0 < t < T$),
+under a fixed value of $q(T)$,
+since it allows us to focus on
+the asymptotic error, $E_A(T)$.
+%
+The contribution of the residual error
+will be considered later
+by adjusting the value of $q(T)$.
+%
+Note that the asymptotic error
+has the form of an action of a free particle
+at $\vartheta\bigl( q(t) \bigr)$,
+with the endpoints fixed at
 $\vartheta\bigl( q(0) \bigr)  = e^{- q(T)}$
@@
 %
-Since the residual error
-is fixed during the process,
-the asymptotic error,
-resembling the action of a free particle,
-demands a constant velocity
-%
-\begin{equation}
-  \dot \vartheta\bigl( q(t) \bigr) = c
+Thus, the velocity should be a constant
+%Since the residual error
+%is fixed during the process,
+%the asymptotic error,
+%resembling the action of a free particle,
+%demands a constant velocity
+%
+\begin{equation}
+  \frac{d}{dt} \vartheta\bigl( q(t) \bigr) = c
   .
@@
 %
-where $t_0$ is some constant,
-and $c$ has been determined
+where $t_0$ and $c$ are two constants,
+and the latter has been determined
 from the $t = T$ case
@@
 %
-which is reached at $t_0 = \Gamma /\Err(0)$,
-or if we assume Eq. \eqref{eq:error_equil_sbin} for $\Err(0)$,
+reachable at $t_0 = \Gamma /\Err(0)$.
+%
+If we assume that at the beginning, $t = 0$,
+the system has gone through a long simulation
+under a constant updating magnitude, $a_0$,
+then the initial error can be computed from Eq. \eqref{eq:EA_sbin}
+as
+\begin{equation}
+  \Err(0)
+  = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
+  = \frac{1}{2} a_0 \Gamma
+  ,
+  \label{eq:error_equil_sbin}
+\end{equation}
+and
 %
@@
 %
-The minimum error can be related to
-the expected histogram fluctuation\cite{zhou2005, zhou2008}
-(cf. Appendix \ref{sec:hfluc}).
+%The minimum error can be related to
+%the expected histogram fluctuation\cite{zhou2005, zhou2008}
+%(cf. Appendix \ref{sec:hfluc}).
 
@@
 %
-By a proper overall scaling of $\mathbf w$,
-we may write this condition as
+If the rate of growth is positive,
+we can, by a proper overall scaling of $\mathbf w$,
+write this condition as
 %
@@
 %
-where we have set the index of the first eigenmode to $0$
+where we have started the index of the eigenmodes at $0$
 instead of $1$ for later convenience.
@@
 \end{equation}
-%
-We may reconstruct the updating matrix in $\phi_{ki}$'s
-as\cite{bussi2006}
-%
-\begin{equation}
-  w_{ij}
-  =
-  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
-  \lambda_k \, \phi_{ki} \, \phi_{kj}
-  .
-  \label{eq:w_from_phi}
-\end{equation}
 
@@
 
-Equation \eqref{eq:vt_diffeq_mbin}
-shows that the rate of error reduction
-is given by $\alpha(t) \, \lambda_k$.
-%
-Thus, a \emph{stable} updating scheme
-permits no negative eigenvalue, $\lambda_k$,
+Equation \eqref{eq:vt_diffeq_mbin} shows that
+in the absence of the fluctuating term, $\tilde \zeta_k(t)$,
+the deviation of the bias potential would tend to zero
+at a rate of $\alpha(t) \, \lambda_k$.
+%
+Thus, the updating scheme is convergent
+only if no eigenvalue, $\lambda_k$, is negative
 for $k > 0$.
@@
 %
-In Appendix \ref{sec:Gamma_measure},
-we discuss a way of estimating $\Gamma_k$
-from an FDS simulation under a constant magnitude.
+%In Appendix \ref{sec:Gamma_measure},
+%we discuss a way of estimating $\Gamma_k$
+%from an FDS simulation under a constant magnitude.
 %
@@
 }
-
-\subsubsection{\label{sec:band-matrix}
-Homogeneous updating schemes}
-
-Many commonly-used updating schemes,
-such as the Gaussian updating scheme
-used in metadynamics,
-can be characterized by a rigid window
-centered on the current bin.
-%
-If we denote the relative updating magnitude on
-the $i$th bin from the center by $\mu_i$,
-the eigenvalues of these translationally invariant or homogeneous
-schemes can be readily deduced
-from the window function ($\mu_0, \mu_1, \dots, \mu_b$)
-as a cosine transform,
-%
-\begin{equation}
-  \lambda_k
-  =
-  \mu_0
-  +
-  2 \,
-  \sum_{ i = 1 }^b
-  \mu_i
-  \cos\left(
-  \frac{ k \, i \, 2 \, \pi }
-       {      g \, n        }
-  \right)
-  ,
-  \label{eq:wband_eigenvalue}
-\end{equation}
-%
-where $g = 1$ here for the periodic variable,
-or $2$ for a non-periodic one
-(cf. Appendix \ref{sec:more_wband}).
-%
-For the Gaussian updating scheme,
-we have\cite{bussi2006}
-\begin{align}
-  \lambda_k
-  =
-  \exp\left[
-        -\frac 1 2
-        \left(
-         \frac{ 2 \, \pi \, k \, \sigma_z }
-              { g \, \Delta z  }
-        \right)^2
-      \right]
-  ,
-%\notag
-\label{eq:lambda_Gaussian}
-\end{align}
-where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
-are the width of the Gaussian and the span
-in the original collective variable, $z$,
-respectively.
 
@@
 
+\subsubsection{\label{sec:Gaussian_scheme}
+Gaussian and homogeneous updating schemes}
+
+Perhaps the most commonly-used multiple-bin updating scheme in metadynamics
+is the Gaussian updating scheme,
+whose updating matrix can be characterized by
+a fixed Gaussian centered on the bin of visit,
+with some boundary adjustment
+to satisfy the constraints discussed
+in Sec. \ref{sec:updating-matrix},
+as shown in Fig. \ref{fig:mat}(a).
+%
+The eigenvalues of this scheme are given by\cite{bussi2006}
+\begin{align}
+  \lambda_k
+  =
+  \exp\left[
+        -\frac 1 2
+        \left(
+         \frac{ 2 \, \pi \, k \, \sigma_z }
+              { g \, \Delta z  }
+        \right)^2
+      \right]
+  ,
+%\notag
+\label{eq:lambda_Gaussian}
+\end{align}
+where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
+are the width of the Gaussian and the span
+in the original collective variable, $z$,
+respectively.
+
+\begin{figure}[h]
+\begin{center}
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
+  }
+  \caption{
+    \label{fig:mat}
+    Updating matrix, $w_{ij}$, for
+    a non-periodic variable ($n = 500$).
+    Here $j$ is the bin of the current collective variable,
+    and $i$ is the bin where the bias potential is updated.
+    %
+    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
+    (b) The bandpass updating scheme of $K = 20$.
+    %
+    Both updating schemes are homogeneous in that
+    they can be characterized a rigid window
+    centered on the current bin, $j$,
+    with some boundary corrections.
+  }
+\end{center}
+\end{figure}
+
+More generally, we can define a class of
+translationally invariant or \emph{homogeneous} updating schemes
+that can be characterized by a rigid window function
+with a floating center.
+%
+These homogeneous updating schemes share
+the same set of orthonormal eigenmodes, $\phi_{ki}$,
+(which are sines and cosines),
+and differ only by the eigenvalues.
+%
+By the completeness of the eigenmodes,
+we can readily reconstruct the updating matrix
+as\cite{bussi2006}
+%
+\begin{equation}
+  w_{ij}
+  =
+  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
+  \lambda_k \, \phi_{ki} \, \phi_{kj}
+  .
+  \label{eq:w_from_phi}
+\end{equation}
+%
+This formula allows us to define new updating schemes
+whose spectrum of eigenvalues are optimal
+for asymptotic convergence,
+which we will discuss below.
+%
+The result is a class of bandpass updating schemes,
+as shown in Fig. \ref{fig:mat}(b).
+
 \subsection{\label{sec:cmpschemes}
@@
 we will turn to the problem of finding
-optimal updating schemes.
+optimal updating schemes
+given that each updating scheme
+adopts the optimal schedule.
 
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-that completely discard certain short-wavelength eigenmodes
-that are deemed as local noise.
+that completely discard certain noise modes
+of short wavelengths.
 %
@@
   $|\max_i u_i - \min_i u_i|^2/K^3$.}
-we may construct an updating scheme
-from Eq. \eqref{eq:w_from_phi} with
-%
-\begin{equation}
-  \lambda_1 = \cdots = \lambda_{K-1} = 1,
+we may assign the first $K$ eigenmodes with eigenvalue $1$
+and the rest of eigenmodes with eigenvalue $0$:
+%
+\begin{equation}
+  \lambda_1 = \cdots = \lambda_{K-1} = 1
+  ,
   \mathrm{\; and \;}
-  \lambda_K = \cdots = \lambda_{n-1} = 0.
+  \lambda_K = \cdots = \lambda_{n-1} = 0
+  .
   \label{eq:lambda_bandpass}
 \end{equation}
+%
+We can then obtain the updating matrix
+from the eigenvalues by Eq. \eqref{eq:w_from_phi}
+(cf. Sec. \ref{sec:homo_bandpass} for the case
+of homogeneous updating schemes).
 %
@@
 %$$w_{ij} = \frac{1}{\rho_i} \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj}.$$
-%%
-With the eigenvalues for the non-noise and noise modes being
-$1$ and $0$, respectively,
-a bandpass updating scheme is essentially
-the single-bin updating scheme
-accompanied by a perfect filter
-for the designated noise modes.
-%
-The asymptotic error is then
-$\Err_A(T) = \sum_{ k = 1 }^{ K - 1 } \Gamma_k T/(T + t_0)^2$,
-which is less than that from Eq. \eqref{eq:EA_sbin1}.
-%
-
-Comparing Eq. \eqref{eq:lambda_bandpass}
-with Eq. \eqref{eq:lambda_Gaussian},
-we find that the bandpass updating scheme serves
-as an idealization of the Gaussian updating scheme
-with the Gaussian spectrum of eigenvalues
-being replaced by a step function.
-%
-Although both updating schemes are able to remove local noise
-from the bias potential,
-the bandpass updating schemes have the advantage
+%
+In effect, each updating step of the bandpass updating scheme
+is equivalent to a single-bin updating step,
+Eq. \eqref{eq:wl_update},
+followed by a filtering step
+that completely removes the components of the noise modes
+from the potential of mean force.
+%
+%The asymptotic error is then
+%$\Err_A(T) = \sum_{ k = 1 }^{ K - 1 } \Gamma_k T/(T + t_0)^2$,
+%which is less than that from Eq. \eqref{eq:EA_sbin1}.
+
+For homogeneous updating schemes,
+the bandpass updating scheme resembles Gaussian updating schemes,
+but contains long-range wiggles,
+as shown in Fig. \ref{fig:mat}(b).
+%
+Both updating scheme are able to deliver
+a smooth and differentiable potential of mean force,
+the former is expected to be superior in asymptotic convergence,
+%
+With a $0$-or-$1$ spectrum of eigenvalues,
+Eq. \eqref{eq:lambda_bandpass},
+instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
+the bandpass updating scheme has the advantage
 of allowing convenient control of the updating magnitude.
@@
 may be inapplicable,
-we can instead use the stage-wise WL technique
-as long as the histogram fluctuation
-used for stage transitions
-is computed from the first $K$ modes
+we can instead use the WL style stage-wise control
+for the updating magnitude.
+%
+In this case, the histogram fluctuation used for stage transitions
+should be computed from only the first $K$ modes,
+where the histogram flattening was performed
 (cf. Appendix \ref{sec:hfluc}).
@@
 
-In addition to the application to homogeneous updating schemes
-(Appendix \ref{sec:homo_bandpass}),
-bandpass updating schemes can also be applied to
-a continuous variable, $z$,\footnote{
-  In the transition to a continuous variable,
-  we will map
-  $\rho_i$ to $\rho(z) \, \delta z$,
+\subsubsection{Orthogonal polynomial basis}
+
+We can also construct the bandpass updating schemes
+from an orthogonal polynomial basis,
+which can simplify the functional form of the bias potential.
+%
+In discussing this application,
+it is more convenient to switch to a continuous-variable setup:
+%
+We will replace the discrete bin index by the continuous variable, $z$.
+%
+and map $\rho_i$ to $\rho(z) \, \delta z$,
   $\phi_{ki}$ to $\phi_k(z) \, \delta z$,
-  $w_{ij}$ to $w(z, z') \, \delta z$, etc.}
-with the eigenmodes derived from orthogonal polynomials, $R_k(z)$,
-that satisfy
+  $w_{ij}$ to $w(z, z') \, \delta z$, etc.
+%
+We define the eigenvectors by
+$\phi_k(z) = R_k(z) \, \rho(z)$,
+where $R_k(z)$ are a set of orthogonal polynomials
+satisfying
 \begin{align*}
@@
 %
-With $\phi_k(z) = R_k(z) \, \rho(z)$,
-we can readily satisfy Eqs. \eqref{eq:eig_orthonormal_cols}-\eqref{eq:ortho0},
-and construct the updating matrix by
-Eqs. \eqref{eq:w_from_phi} and \eqref{eq:lambda_bandpass} as
+These eigenvectors allow us to construct the updating matrix
+Eq. \eqref{eq:w_from_phi},
+%we can readily satisfy Eqs. \eqref{eq:eig_orthonormal_cols}-\eqref{eq:ortho0},
+with the eigenvalues given by \eqref{eq:lambda_bandpass}:
 \begin{align*}
@@
 \end{equation}
+
+\subsubsection{LLR algorithm}
 
@@
 was recommended for optimal convergence of $c_1$\cite{pellegrini2014}.
+
+\subsubsection{Adaptive Gaussian umbrella sampling}
 
@@
 of the Gaussian ensemble
-$Z \equiv \int e^{-\beta \, \mathcal U(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$
+$Z \equiv \int e^{-\beta \, \Phi(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$
 as the bias potential for inter-umbrella transitions.
@@
 $\beta$ is the inverse temperature,
-$\mathcal U(\mathbf x)$ is the unbiased potential energy
+$\Phi(\mathbf x)$ is the unbiased potential energy
 as a function of molecular coordinates, $\mathbf x$,
@@
 
+\subsubsection{System and simulation details}
+
 Our first test system is
@@
 %
+We will assume reduced units in the calculation.
+%
 The two special particles are mutually non-interacting
-but they interact normally with other particles,\footnote{The
-two special particles were placed on an axis parallel to a side of the cubic box}
-and the intrinsic distribution, $p(r)$, is
-the cavity distribution function, $y(r)$\cite{hansen}.
-%
-We will assume reduced units in this calculation.
+but they interact normally with other particles
+via the pair potential $u(r) = 4(r^{-12} - r^{-6})$.
+%
+The intrinsic distribution of distance, $p(r)$, is
+\begin{align*}
+  p(r) = \langle \delta(\mathbf r_{12} - \mathbf r) \rangle
+  =
+  \frac{V}{Z}
+  \int e^{-\Phi(\mathbf r^N)/(k_B T)}
+  d\mathbf r_3 \cdots d\mathbf r_N
+  ,
+\end{align*}
+where
+$\Phi(\mathbf r^{N}) = \sum_{j=3}^N\sum_{i=1}^{j-1} \phi(r_{ij})$,
+%
+and it is related the cavity distribution function, $y(r)$\cite{hansen},
+as $y(r) = \left(1 - \frac{1}{N}\right) \, V \, p(r)$.
+%
+In biased sampling, we fixed the two special particles on the $x$-axis,
+and varied their distance $r$ along the axis.
+%
+Further, we added a bias potential $k_B T \, u(r)$,
+which would ideally bias $p(r)$
+to a flat distribution along $r$.
 %
@@
 In each trial,
-a random particle is displaced in each dimension
+a random particle is displaced in each permissible dimension
 according to a uniform distribution over $(-\varepsilon^*, \varepsilon^*)$.
@@
 %
-We tested the single-bin, Gaussian, and bandpass updating schemes.
+We tested the single-bin, Gaussian,
+and homogeneous bandpass updating schemes.
 %
@@
 %
+The deviation of the bias potential was calculated from
+Eqs. \eqref{eq:v_def} and \eqref{eq:x_def},
+and decomposed to eigenmodes by
+Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl}.
+%
+The error was computed from Eq. \eqref{eq:error_def}.
+%
 We repeated each simulation multiple times
@@
 collected from the equilibration phase in individual runs.
+
+\subsubsection{Results}
 
@@
 %
-In this case, $\mathcal U(\mathbf x) = 0$
+In this case, $\Phi(\mathbf x) = 0$
 and the intrinsic distribution is
@@
 Dr. C. Myers, and Dr. O. Samoylova
-for helpful discussions.
+for helpful discussions,
+and the two anonymous reviewers for critical reading
+of earlier versions of the manuscript.
 %
@@
 Here, we give some mathematical details
-on the homogeneous updating schemes.
+on the homogeneous updating schemes,
+%
+The updating matrix of the updating schemes, $w_{ij}$,
+can be characterized by
+a fixed window function, $\mu_0, \mu_1, \dots, \mu_b$,
+that shifts with the bin $j$.
+%
+Here $\mu_l$ denotes the relative updating magnitude
+on the $l$th bin from the center
+(excluding the boundary adjustment).
+%
+We will show below that the eigenvalues of the updating matrix
+is related to the window function as a cosine transform.
 %
 By the translational invariance,
-these schemes
-satisfy detailed balance,
+these schemes satisfy detailed balance,
 Eq. \eqref{eq:w_detailedbalance},
@@
 Then one can readily verify that the eigenvalues are given by
-  Eq. \eqref{eq:wband_eigenvalue},
-  and there is a two-fold degeneracy,
+\begin{equation}
+  \lambda_k
+  =
+  \mu_0
+  +
+  2 \,
+  \sum_{ l = 1 }^b
+  \mu_l
+  \cos\left(
+  \frac{ k \, l \, 2 \, \pi }
+       {      g \, n        }
+  \right)
+  ,
+  \label{eq:wband_eigenvalue}
+\end{equation}
+%
+where $g = 1$ here, with a two-fold degeneracy,
+%for the periodic variable,
+%or $2$ for a non-periodic one
+%There is a two-fold degeneracy,
   $\lambda_{n - k} = \lambda_k$.
@@
        \right]
-  ,
-\notag
-%\label{eq:wband_eigenvector_refl}
-\end{equation}
-%
-and the eigenvalues are given by
-  Eq. \eqref{eq:wband_eigenvalue}
-  with $g = 2$.
+  .
+  %\notag
+  \label{eq:wband_eigenvector_refl}
+\end{equation}
+%
+%and the eigenvalues are given by
+%  Eq. \eqref{eq:wband_eigenvalue}
+%  with $g = 2$.
 
@@
 \subsection{\label{sec:homo_bandpass}
-Bandpass updating schemes}
+Homogeneous bandpass updating schemes}
 

### 2017-08-12 prog/invt.h

@@ -238,7 +238,7 @@ __inline static int savewin(double *win, int winn,
 /* save the updating window function
  * in matrix form */
 __inline static int savewinmat(double *win, int winn,
-    int n, int pbc, const char *fn)
+    int n, int pbc, const char *fn, int matrix)
 {
   FILE *fp;
   int i, j, ks[3], k, l;
@@ -261,6 +261,7 @@ __inline static int savewinmat(double *win, int winn,
         ks[1] = i + j - 1;
         ks[2] = i + j - 2 * n - 1;
       }
+      /* add contributions from ks[0], ks[1], ks[2] */
       for ( l = 0; l < 3; l++ ) {
         k = ks[l];
         if ( k < winn && k > -winn ) {
@@ -273,7 +274,11 @@ __inline static int savewinmat(double *win, int winn,
       }
       if ( y > ymax ) ymax = y;
       if ( y < ymin ) ymin = y;
-      fprintf(fp, "%9.6f ", y);
+      if ( matrix ) { /* matrix form */
+        fprintf(fp, "%9.6f ", y);
+      } else { /* plain form */
+        fprintf(fp, "%d %d %9.6f\n", i, j, y);
+      }
     }
     fprintf(fp, "\n");
   }
@@ -408,7 +413,7 @@ __inline static double *prepwin(double *lambda, int n,
   }
   if ( fnwinmat[0] != '\0' ) {
     /* save the n x n updating matrix */
-    savewinmat(win, *winn, n, pbc, fnwinmat);
+    savewinmat(win, *winn, n, pbc, fnwinmat, 0);
   }
   return win;
 }

### 2017-08-12 prog/metad.h

@@ -74,7 +74,7 @@ static void metad_prepwin(metad_t *metad,
   /* save the window kernel */
   savewin(metad->win, metad->winn, "ker.dat");
   /* save the n x n updating matrix */
-  savewinmat(metad->win, metad->winn, n, pbc, "win.dat");
+  savewinmat(metad->win, metad->winn, n, pbc, "win.dat", 0);
 }
 
 static metad_t *metad_open(int imin, int imax, int idel,

### 2017-08-15 doc/invt.tex
editing invt

@@
 %
-The dimensionless free energy, or potential of mean force (PMF),
+The distribution can be related to
+the dimensionless free energy, or potential of mean force (PMF),
 along the collective variable, $z$,
-is $-\ln p(z)$.
-%
-However, accurately estimating the PMF is often difficult
+as $-\ln p(z)$.
+%
+However, reliable estimation of the PMF is often difficult
 when the collective variable is energetically restricted to local regions
@@
 Thus,
-to overcome these energetic barriers and
-to capture the global shape of the free energy surface
+to overcome these energetic barriers in sampling,
+%and to capture the global shape of the free energy surface
 an effective strategy is to introduce a bias potential that
-cancels the PMF.
-%
-Ideally, then, sampling yields a flat distribution
-along $z$\cite{mezei1987, berg1992, *lee1993,
+offsets the PMF
+such that the distribution along $z$ is
+flat\cite{mezei1987, berg1992, *lee1993,
 wang2001, wang2001pre,
@@
 %
-The flat-distribution sampling can be used to accelerate
-sampling of slow transitions that otherwise would not readily
-be observed in traditional simulations.
+\note{The flat-distribution sampling
+allows one to more accurately capture the
+global shape of the free energy surface
+by improving the sampling around the rarely visited barrier region
+and accelerating transitions among free energy basins.}
 
@@
 %
-While a slow reduction leads to large variation of the bias potential,
+While a slow reduction of the updating magnitude
+leads to large variation of the bias potential,
 a rapid reduction can converge the bias potential
@@
 %
-For the single-bin case, the optimal updating schedule
-is a function of the
+For the single-bin case, it is known that
+the optimal updating schedule is given by the
 inverse time\cite{
@@
 the inverse number of updating steps.
-\note{
-  About the origin of the inverse-time formula.
-  Besides the Belardinelli paper,
-  we can also find it in Eq. (6) of reference \cite{liang2007},
-  if I am not too mistaken.
-  But I find the paper is a bit hard to read.
-  There, however, the formula is $t_0/t$ instead of $1/t$.
-}
-%The same schedule also applies to a variant of
-%the WL algorithm\cite{langfeld2012, pellegrini2014}.
 
 In this study,
-we present a method of computing
+we first present a method of computing
 the optimal schedule
@@
 %
-In this way, the optimal schedule
-can be deduced from the equation of motion
+This mapping allows the optimal schedule
+to be derived from the equation of motion
 that minimizes the action.
@@
 depends on the updating window function
-through the mass function.
+through the mass of the particle.
 %
@@
 and the width of the Gaussian window.
-%
-Further, we will compare
-the highest efficiency of error reduction
+
+We will then compare
+the efficiency of error reduction
 of different updating schemes.
 %
-We will show that
+We find that
 the single-bin updating scheme
-is optimal for long simulations,
-and it can be generalized
-to a class of bandpass updating schemes
+is optimal for long simulations
+without a priori assumption of the smoothness of the PMF.
+%
+More generally,
+it can be generalized to
+a class of bandpass updating schemes
 that are equally optimal.
 %
-The latter allow one to partially flatten
-a few long-range histogram modes,
-and to filter local noise
-during the updating process.
-
-The bandpass updating schemes are convenient
-in that their optimal schedule is simply given by
-the inverse-time formula,
-and thus requires no extra calculation.
-%
-They can be constructed from any orthogonal basis\cite{bussi2006},
-and for a polynomial basis,
-this construction results in
-a class of adaptive FDS algorithms that
-update the bias potential based on the
-statistical moments of observed distribution
+The bandpass updating schemes
+can avoid local noise
+by partially flattening a few long-range distribution modes.
+%
+The bandpass updating schemes
+are easier to use than the Gaussian updating scheme
+in that the optimal schedule of the former
+is simply given by the inverse-time formula,
+saving the effort of going through the general procedure described above.
+
+Finally, we present an application of
+the bandpass updating schemes
+by constructing it from
+a sequence of orthogonal polynomials.
+%
+The bias potential is a linear combination
+of orthogonal polynomials
+and the coefficients are adjusted on-the-fly
+based on the observed statistical moments.
+%
+This application can be regarded
 as a generalization of the Langfeld-Lucini-Rago (LLR)
@@
 %
-We show that a similar construction
-can help FDS algorithms with
-quadratic bias potentials\cite{neuhaus2006, *neuhaus2007, zhu2012}
-automatically adjust the parameters
-to reach a set of Gaussian distributions
+A similar construction with
+a quadratic bias potential\cite{neuhaus2006, *neuhaus2007, zhu2012}
+allows one to sample an array of Gaussian distributions
 with desired means and variances.
@@
 to build a class of adaptive FDS algorithms
-that based on the statistical moments.
+with quadratic bias potentials.
 
@@
 
-Sampling a multimodal distribution with deep valley(s)
-is often challenging in traditional simulations,
-and a common remedy is to apply a bias potential
-to sample a smoother distribution, $\rho_i$.
+Sampling a multimodal distribution with deep basin(s)
+is often challenging in regular MC and MD simulations.
+%
+To promote transitions among the basins,
+we can apply a bias potential to alter
+the intrinsic distribution to a smoother one, $\rho_i$.
 %
@@
 %
-our aim is to find a bias potential that allows
+we wish to find a bias potential that allows
 the resulting $\pi_i$ to coincide
@@
 %
-%according to Eq. \eqref{eq:pi_p_V}.
-%which allows us to deduce the PMF, $-\ln p_i$,
-%directly from the bias potential.
-%
 and the PMF, $-\ln p_i$, can be readily
 deduced from the bias potential. % $u_i$.
-\note{
-  For example, in the context of simulated tempering (Zhang and Ma, 2007),
-  $p_i$ is replaced by $\ln Z_i$,
-  $\rho_i$ is the weight $\zeta_i$,
-  and the bias potential, $u_i$, is $\ln (\tilde Z_i / \zeta_i)$.
-  Upon convergence, we have $\ln \tilde Z_i \to \ln Z_i$
-  up to a constant shift.
-}
 
@@
 *laio2002, *laio2008, *barducci2011, *sutto2012}
-have been developed to find the proper bias potential.
+have been developed to find the desired bias potential.
 %
@@
 accumulated from the simulation as
-%
-%Using $H_i$ as an estimate of $\pi_i$ in
-%Eq. \eqref{eq:pi_p_V}
-%allows us to correct $u_i$ as
 %
@@
 %
-The Gaussian updating scheme used in metadynamics
-is an example, with $w_{ij}$ being
-roughly a Gaussian centered on bin $j$.
+A common example is the Gaussian updating scheme used in metadynamics,
+with $w_{ij}$ being roughly a Gaussian centered on bin $j$.
 %
@@
 Once the histogram is sufficiently flat,
-we switch to a new stage
+we reset the histogram and start a new stage
 with $\alpha(t)$ reduced by a factor of $1/2$\cite{
@@
 %
-This general result\cite{
-  robbins1951, pellegrini2014}
-has an intuitive explanation\cite{
+This is a general result\cite{
+  robbins1951, pellegrini2014},
+and it has an intuitive explanation\cite{
   marsili2006, barducci2008}
@@
 for the more general multiple-bin scheme.
-%
-%We will suggest a slight modification of Eq. \eqref{eq:alpha_invt}
-%whereby the origin of $t$ is shifted by an amount
-%determined from the initial error.
 %
@@
 and a baseline value, $\bav{v}(t)$,
-for average growth\cite{
+for the average growth\cite{
 dama2014}.
@@
   $\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
-  i.e.,
+  i.e.
   $$
@@
   $$
-  One can readily show that the two definitions are equivalent
+  We can readily show that the two definitions are equivalent
   for small $v_{*i}$.}%
@@
   \int_0^T
-  \dot \vartheta\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
-\notag
-%\label{eq:xt_solution}
+    \dot{\vartheta}\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
+  \label{eq:xt_solution}
 \end{equation}
@@
 %
-$
-q(t) \equiv \int_0^t \alpha(t') \, dt',
-$
+$q(t) \equiv \int_0^t \alpha(t') \, dt'$,
 %
@@
 
-At the end of period $T$,
-the total error defined in Eq. \eqref{eq:error_def}
-is a sum of a residual error, $E_R$ (from the decay of the initial error)
+From Eqs. \eqref{eq:error_def} and \eqref{eq:xt_solution},
+we can write the total error at the end of period $T$
+as a sum of a residual error, $E_R$ (from the decay of the initial error)
 and an asymptotic error, $E_A$:
@@
   \Err_A(T)
+  &=
+  \sum_{i=1}^n \rho_i
+    \int_0^T\!\!\int_0^T
+    \langle
+      \zeta_{*i}(t) \zeta_{*i}(t')
+    \rangle
+    \dot \vartheta\bigl( q(t) \bigr)
+    \dot \vartheta\bigl( q(t') \bigr)
+    \, dt dt'
+  \notag \\
   &= \Gamma \, \int_0^T \dot \vartheta^2\bigl( q(t) \bigr) \, dt
@@
 with
-$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
+%$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
+$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i \approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \zeta_{*i}(t) \zeta_{*i}(0) \rangle \, dt$.
 %\begin{equation}
-%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i.
-%  \label{eq:Gamma_from_G}
+%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
+%  =
+%  \sum_{i=1}^n \rho_i \int_0^T
+%    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
+%  .
+%  \notag%\label{eq:Gamma_from_G}
 %\end{equation}
@@
 %
-If we assume that at the beginning, $t = 0$,
-the system has gone through a long simulation
-under a constant updating magnitude, $a_0$,
-then the initial error can be computed from Eq. \eqref{eq:EA_sbin}
-as
-\begin{equation}
-  \Err(0)
-  = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
-  = \frac{1}{2} a_0 \Gamma
-  ,
-  \label{eq:error_equil_sbin}
-\end{equation}
-and
-%
-\begin{equation}
-  t_0 = \frac{ 2 } { a_0 } = \frac{1}{ \alpha(t = 0) }
-  .
-  \label{eq:t0_sbin}
-\end{equation}
+%If we assume that at $t = 0$,
+%the system has just completed a long simulation
+%under a constant updating magnitude, $a_0$,
+%then the initial error can be computed from Eq. \eqref{eq:EA_sbin}
+%as
+%\begin{equation}
+%  \Err(0)
+%  = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
+%  = \frac{1}{2} a_0 \Gamma
+%  ,
+%  \label{eq:error_equil_sbin}
+%\end{equation}
+%and
+%%
+%\begin{equation}
+%  t_0 = \frac{ 2 } { a_0 } = \frac{1}{ \alpha(t = 0) }
+%  .
+%  \label{eq:t0_sbin}
+%\end{equation}
 %
@@
 \subsubsection{\label{sec:updating-matrix}
-Updating matrix and eigenmode decomposition}
-
-We will first study the eigenvalue decomposition
+Updating matrix and spectral decomposition}
+
+We will first review the spectral decomposition
 of the updating matrix, $\mathbf w$,
@@
   =
-  \sum_{i=1}^n \rho \, x_{*i} \, y_{*i}
+  \sum_{i=1}^n \rho_i \, x_{*i} \, y_{*i}
   .
@@
 \note{Proof:
-$$
+\begin{align*}
   \sum_{k=0}^{n-1} \tilde x_{*k} \, \tilde y_{*k}
-  =
+  &=
   \sum_{k=0}^{n-1} \tilde x_{*k} \sum_{i=1}^n \phi_{ki} \, y_{*i}
-  =
+  \\
+  &=
   \sum_{i=1}^n \left( \sum_{k=0}^{n-1} \tilde x_{*k} \, \phi_{ki} \right) \, y_{*i}
   =
-  \sum_{i=1}^n \rho \, x_{*i} \, y_{*i}.
-$$
+  \sum_{i=1}^n \rho_i \, x_{*i} \, y_{*i}.
+\end{align*}
 }
@@
 %
-With a set of heterogeneous eigenvalues,
-the errors of the eigenmodes are reduced at different rates.
-%
-Thus, the optimization problem of the final error
-becomes more complicated and it often
-involves a compromise among eigenmodes.
+%With a set of heterogeneous eigenvalues,
+%the errors of the eigenmodes are reduced at different rates.
+%Thus, the optimization problem of the final error
+%becomes more complicated and it often
+%involves a compromise among eigenmodes.
+%
+The formal solution is given by
+\begin{equation}
+  \tilde v_k(T)
+  =
+  \tilde v_k(0) \, e^{-\lambda_k \, q(T)}
+  +
+  \int_0^T
+    \dot{\vartheta}_k\bigl( q(t) \bigr) \, \tilde\zeta_k(t) \, dt
+  ,
+  \label{eq:xt_solution}
+\end{equation}
+where
+\begin{equation}
+  \vartheta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
+  .
+  \label{eq:uk_def}
+\end{equation}
 
@@
   &=
+  \sum_{k=1}^{n-1}
+  \int_0^T\!\!\int_0^T
+  \langle
+    \tilde\zeta_k(t)
+    \tilde\zeta_k(t')
+  \rangle
+  \vartheta_k\bigl(q(t)\bigr)
+  \vartheta_k\bigl(q(t')\bigr)
+  \, dtdt'
+  \notag \\
+  &=
   \int_0^T
@@
 %
-\begin{equation}
-  \vartheta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
-  ,
-  \label{eq:uk_def}
-\end{equation}
-%
-and $\Gamma_k$ is computed from
-the autocorrelation function
-of the histogram fluctuation,
-$\bigl\langle {\tilde \zeta}_k(0)
-\, {\tilde \zeta}_k(t) \bigr\rangle$,
-as:
+%$\Gamma_k$ is computed from
+%the autocorrelation function
+%of the histogram fluctuation,
+%$\bigl\langle {\tilde \zeta}_k(0)
+%\, {\tilde \zeta}_k(t) \bigr\rangle$,
+%as:
 %
@@
   \Gamma_k
-  = \sum_{t = -\infty}^\infty
-  \bigl\langle {\tilde \zeta}_k(0)
-            \, {\tilde \zeta}_k(t) \bigr\rangle
-  .
+  \approx
+  \int_{-\infty}^\infty
+  \bigl\langle
+    \tilde\zeta_k(t) \tilde\zeta_k(0)
+  \bigr\rangle
+  \, dt
+  \to
+  \sum_{t = -\infty}^\infty
+  \bigl\langle
+    \tilde\zeta_k(t) \tilde\zeta_k(0)
+  \bigr\rangle
+  ,
   \label{eq:Gamma_sum}
@@
 %
-%\begin{equation}
-%$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
-%  \notag
-%\label{eq:Gamma_integral}
-%\end{equation}
+with the second expression being
+the equivalent in the discrete-time setting.
 %
@@
 a first approximation for modeling
-a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
+a general fluctuation, ${\tilde \zeta}_k(t)$,
 as an equivalent white noise.
@@
 %
+Note that by Eq. \eqref{eq:parseval}, we have
+$\sum_{k=1}^{n-1} \bigl\langle {\tilde \zeta}_k(t) \, {\tilde \zeta}_k(0) \bigr\rangle
+= \sum_{i=1}^n \rho_i \bigl\langle \zeta_{*i}(t) \, \zeta_{*i}(0) \bigr\rangle,$
+and after summing $t$ from $-\infty$ to $\infty$,
+\begin{equation}
+  \sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma
+  .
+  \label{eq:Gammak_sum}
+\end{equation}
+This shows the equivalence of
+Eq. \eqref{eq:EA_sbin1} and \eqref{eq:error_asym}
+in the case of the single-bin updating scheme, with $\lambda_k \equiv 1$.
 %In Appendix \ref{sec:Gamma_measure},
@@
 The value of $q(T)$ will be adjusted later
-to minimize the total error in Sec. \ref{sec:optinitalpha}).
+to minimize the total error in Sec. \ref{sec:optinitalpha}.
 %
@@
 %
-for some positive $C_M$
+for some positive $C_M$.
+%
+Equation \eqref{eq:Lagrangian_const} serves
 as a generalization of Eq. \eqref{eq:dthetadt_const}.
@@
 %
+In Appendix \ref{sec:schedule_geometry},
+we give a geometric characterization of this solution.
+%
 The minimal asymptotic error is
@@
 To model the initial error for Eq. \eqref{eq:error_res},
-we will assume that
-the system undergoes a preliminary FDS simulation
+we will assume that at $t = 0$
+the system has undergone a preliminary FDS simulation
 under a constant updating magnitude, $a_0$,
@@
 i.e. the optimal initial updating magnitude
-is half of the equilibrium value.\footnote{This factor, $1/2$,
+is half of the previous equilibrium value\footnote{This factor, $1/2$,
 happens to coincide with the
@@
 }
+%[cf. Eq. \eqref{eq:t0_sbin}].
 
@@
 Based on the above results,
-we now give a procedure of computing
+we give a procedure of computing
 the optimal schedule and the error
@@
 
-\begin{figure}[h]
-\begin{singlespace}
-  \fontsize{10pt}{\baselineskip}
-  \begin{tikzpicture}
-  \node[draw, text width={width("Updating ")}]
-    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};
-
-  \node[draw, text width={width("sampling")}]
-    (sys) at (0, 1.1) {System\\and\\sampling\\process};
-
-  \node[draw, text width={width("Eigenvalues,")}]
-    (lambda) at (2.2, 3) {Eigenvalues,\\$\lambda_1, \dots, \lambda_{n-1}$};
-
-  \node[draw, text width={width("Integrals, iii")}]
-    (Gamma) at (2.1, 1.1) {Integrals, \\$\Gamma_1, \dots, \Gamma_{n-1}$\\and biases,\\$\epsilon_1, \dots, \epsilon_{n-1}$};
-
-  \node[draw, text width={width("Simulation")}]
-    (T) at (2.2, -0.5) {Simulation\\time, $T$};
-
-  \node[draw, text width={width("Optimal")}]
-    (qT) at (4.6, 1.4) {Optimal\\ $q(T)$};
-
-  \node[draw, text width={width("Schedule,")}]
-    (alpha) at (7, 2) {Optimal schedule,\\ $\alpha(t)$};
-
-    \node[draw, text width={width("Error,")}]
-    (err) at (7, 0) {Error, $\Err(T)$};
-
-  \node[inner sep=0, minimum size=0] (M1) at (3.4, 1) {};
-
-  \node[inner sep=0, minimum size=0] (N1) at (5.9, 1) {};
-
-  \node[inner sep=0, minimum size=0] (R1) at (3.6, 1) {};
-  \node[inner sep=0, minimum size=0] (R2) at (4.6, 0.6) {};
-  \node[inner sep=0, minimum size=0] (R3) at (5.6, 1) {};
-
-  \draw[->, >=latex, thick]
-    (w) edge (lambda)
-    (w) edge (Gamma)
-    (sys) edge (Gamma);
-
-  \draw[->, >=latex, thick]
-    (lambda) -| (M1)
-    (Gamma)  -- (M1)
-    (T)      -| (M1)
-    (M1)     -- (R1)
-    (R1)     |- (qT);
-
-  \draw[->, >=latex, thick]
-    (qT)     -| (R3);
-
-  \draw[->, >=latex, thick]
-    (R3) -- (N1) |- (alpha);
-
-  \draw[->, >=latex, thick]
-    (N1) |- (err);
-
-  \draw[->, >=latex, thick]
-    %(M1) edge[bend left=70] (N1);
-    (R1) |- (R2) -| (R3);
-\end{tikzpicture}
-\end{singlespace}
+\begin{figure}[h]\centering
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/vardep.pdf}
+  }
   \caption{
@@
 Compute the eigenvalues, $\lambda_k$'s
-from the updating matrix, $\mathbf w$,
-or from Eq. \eqref{eq:wband_eigenvalue}
-for a homogeneous updating scheme.
+from the updating matrix, $\mathbf w$.
 %
 Note that
-no eigenvalue can be negative for the updating scheme to be stable.
+no eigenvalue can be negative in a convergent updating scheme.
 
@@
 \item \label{step:qT}
-Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT}.
-%
-We implemented a variant of the Newton-Raphson method\cite{press3rd}
-with the bisection method\cite{press3rd}
-as a fallback.\footnote{
-The equation was solved for $\ln q(T)$
+Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT},
+using, e.g. the Newton-Raphson method\cite{press3rd}\footnote{In
+  our implementation, the equation was solved for $\ln q(T)$
   to avoid negative values of $q(T)$,
@@
   The initial value was set to $q(T) = \ln(1+T\,a_0/2)$.
-}
+  We also used the bisection method as a fallback.}
 %
@@
 $q_{[ -1]} \equiv 0$,
-  $q_{[N+1]} \equiv q(T)$,
+$q_{[\mathcal N+1]} \equiv q(T)$,
 $t_{[ -1]} \equiv 0$, and
-  $t_{[N+1]} \equiv T$.
+$t_{[\mathcal N+1]} \equiv T$.
 %
@@
 Compute the total error $\Err(T)$ from
-Eqs. \eqref{eq:error_tot},
+Eqs.
+\eqref{eq:error_tot},
 \eqref{eq:error_res},
@@
 
-The above procedure is not always necessary.
-%
-As we will show below, for a class of bandpass updating schemes,
-  the asymptotically optimal schedule
-  is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
+%The above procedure is not always necessary.
+%%
+%As we will show below, for a class of bandpass updating schemes,
+%  the asymptotically optimal schedule
+%  is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
 %
@@
 
-Perhaps the most commonly-used multiple-bin updating scheme in metadynamics
+A commonly-used multiple-bin updating scheme in metadynamics
 is the Gaussian updating scheme,
 whose updating matrix can be characterized by
-a fixed Gaussian centered on the bin of visit,
-with some boundary adjustment
-to satisfy the constraints discussed
-in Sec. \ref{sec:updating-matrix},
-as shown in Fig. \ref{fig:mat}(a).
+a fixed Gaussian centered on the bin of visit
+(with some boundary adjustment).
+%to satisfy the constraints discussed in Sec. \ref{sec:updating-matrix},
+%as shown in Fig. \ref{fig:mat}(a).
 %
@@
 
-\begin{figure}[h]
-\begin{center}
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
-  }
-  \caption{
-    \label{fig:mat}
-    Updating matrix, $w_{ij}$, for
-    a non-periodic variable ($n = 500$).
-    Here $j$ is the bin of the current collective variable,
-    and $i$ is the bin where the bias potential is updated.
-    %
-    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
-    (b) The bandpass updating scheme of $K = 20$.
-    %
-    Both updating schemes are homogeneous in that
-    they can be characterized a rigid window
-    centered on the current bin, $j$,
-    with some boundary corrections.
-  }
-\end{center}
-\end{figure}
-
 More generally, we can define a class of
-translationally invariant or \emph{homogeneous} updating schemes
-that can be characterized by a rigid window function
-with a floating center.
-%
-These homogeneous updating schemes share
-the same set of orthonormal eigenmodes, $\phi_{ki}$,
-(which are sines and cosines),
+translationally-invariant or \emph{homogeneous} updating schemes
+by a rigid window function with a floating center.
+%
+As shown in Appendix \ref{sec:homo},
+the homogeneous updating schemes share
+the same set of orthonormal eigenvectors, $\phi_{ki}$,
+which are cosine and sine functions,
 and differ only by the eigenvalues.
-%
-By the completeness of the eigenmodes,
-we can readily reconstruct the updating matrix
+
+Note that we can reconstruct the updating matrix
+from the eigenvalues and eigenvectors
 as\cite{bussi2006}
@@
 %
-This formula allows us to define new updating schemes
-whose spectrum of eigenvalues are optimal
-for asymptotic convergence,
-which we will discuss below.
-%
-The result is a class of bandpass updating schemes,
-as shown in Fig. \ref{fig:mat}(b).
+Thus, it is possible to improve the updating scheme
+by selecting a spectrum of eigenvalues
+optimal for asymptotic convergence.
+We will pursue this strategy below.
 
@@
 %
-Consider a class of $\mathbf w$ matrices
+Consider a class of updating matrices
 sharing the same set of eigenvectors,
@@
 
-Since an updating matrix
-with identical eigenvalues
-is a multiple of the identity matrix,\footnote{Since
-  the $k = 0$ eigenmode represents
+We can show that the above optimal solution for asymptotic convergence
+represents the single-bin updating scheme.
+%
+Since the $k = 0$ eigenmode represents
 a uniform shift of the bias potential,
-  we can freely set $\lambda_0$ to be
-  the same as $\lambda_1$
-  without changing the nature of
-  the updating scheme.}
-we recover the single-bin scheme
-by setting the eigenvalues to $1$.
-%
-Thus, the single-bin scheme is most efficient
-in error reduction in the asymptotic limit.
-%
-In particular,
-with $\vartheta_k = (t+t_0)/(T+t_0)$,
-we can readily recover the optimal schedule,
-Eq. \eqref{eq:alpha_invt1}, with
-the asymptotic error being
-$\Err_A(T) = \sum_{ k = 1 }^{n-1} \Gamma_k \, T / (T + t_0)^2$,
+we can set $\lambda_0 = \lambda_1$
+without changing the nature of the updating scheme.
+%
+By Eqs. \eqref{eq:w_from_phi} and
+\eqref{eq:eig_orthonormal_cols}, we have
+\begin{align*}
+  w_{ij}
+  = \frac{1}{\rho_i} \sum_{k=0}^{n-1} \lambda_1 \phi_{ki} \phi_{kj}
+  = \lambda_1 \, \delta_{ij}
+  ,
+\end{align*}
+and $\lambda_1$ must be $1$ by \eqref{eq:w_sumj}.
+%
+Thus, $w_{ij} = \delta_{ij}$
+is reduced to the updating matrix of the single-bin updating scheme.
+%
+This demonstrates the asymptotic optimality of
+the single-bin updating scheme.
+%
+%We can further verify that
+%$\vartheta_k = (t+t_0)/(T+t_0)$
+%in this case,
+%and the optimal schedule coincides with
+%Eq. \eqref{eq:alpha_invt1}.
+%
+%The asymptotic error can be written as
+%$\Err_A(T) = \sum_{ k = 1 }^{n-1} \Gamma_k \, T / (T + t_0)^2$,
+%which is also identical to Eq. \eqref{eq:EA_sbin1}.
 %
@@
 %
-which recovers Eq. \eqref{eq:EA_sbin1}.\footnote{By
-Eq. \eqref{eq:parseval}, we have
-$\sum_{k=1}^{n-1} \bigl\langle {\tilde \zeta}_k(t) \, {\tilde \zeta}_k(0) \bigr\rangle
-= \sum_{i=1}^n \rho_i \bigl\langle \zeta_{*i}(t) \, \zeta_{*i}(0) \bigr\rangle$,
-and by summing $t$ from $-\infty$ to $\infty$, we get
-$\sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma$.}
 
@@
 perfect (brick-wall) \emph{bandpass} updating schemes
-that completely discard certain noise modes
+that discard certain noise modes
 of short wavelengths.
@@
 to be representable by the first $K$
-eigenmodes,\footnote{As an estimate of error,
-  we note that truncating a linear or quadratic function
-  at mode $K$ for a non-periodic variable
-  results in an error roughly proportional to
-  $|\max_i u_i - \min_i u_i|^2/K^3$.}
+eigenmodes,
 we may assign the first $K$ eigenmodes with eigenvalue $1$
@@
 %
-We can then obtain the updating matrix
-from the eigenvalues by Eq. \eqref{eq:w_from_phi}
-(cf. Sec. \ref{sec:homo_bandpass} for the case
-of homogeneous updating schemes).
+The updating matrix can then be reconstructed
+from the eigenvalues by Eq. \eqref{eq:w_from_phi}.
+%(cf. Appendix \ref{sec:homo_bandpass} for the case of homogeneous updating schemes).
 %
@@
 followed by a filtering step
-that completely removes the components of the noise modes
-from the potential of mean force.
+that completely removes the components of the noise modes of $k \ge K$
+from the bias potential.
 %
@@
 
-For homogeneous updating schemes,
-the bandpass updating scheme resembles Gaussian updating schemes,
-but contains long-range wiggles,
-as shown in Fig. \ref{fig:mat}(b).
-%
-Both updating scheme are able to deliver
-a smooth and differentiable potential of mean force,
-the former is expected to be superior in asymptotic convergence,
-%
-With a $0$-or-$1$ spectrum of eigenvalues,
-Eq. \eqref{eq:lambda_bandpass},
-instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
-the bandpass updating scheme has the advantage
-of allowing convenient control of the updating magnitude.
-%
-In the asymptotic regime,
-the optimal schedule is simply given
-by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
-%with $t_0$ given by Eq. \eqref{eq:t0_sbin},
-just like the single-bin scheme,
-and it is therefore unnecessary to go through
-the procedure in Sec. \ref{sec:procedure}
-for a general updating scheme.
-%
-In initial stages, when the inverse-time schedule
-may be inapplicable,
-we can instead use the WL style stage-wise control
-for the updating magnitude.
-%
-In this case, the histogram fluctuation used for stage transitions
-should be computed from only the first $K$ modes,
-where the histogram flattening was performed
-(cf. Appendix \ref{sec:hfluc}).
+Applying this construction to the basis of cosine and sine functions,
+we get a homogeneous bandpass updating scheme
+that resembles the Gaussian updating scheme,
+but with better convergence in the asymptotic limit.
+%
+In Sec. \ref{sec:lj},
+we will further compare the performance of the two updating schemes
+in an example.
+%
+%While both updating scheme are able to deliver
+%a smooth and differentiable potential of mean force,
+%the bandpass updating scheme is superior in asymptotic convergence,
+%because of the set of clear-cut spectrum of eigenvalues.
+%Eq. \eqref{eq:lambda_bandpass},
+%instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
+%
+%\footnote{As an estimate of error,
+%  we note that truncating a linear or quadratic function
+%  at mode $K$ for a non-periodic variable
+%  results in an error roughly proportional to
+%  $|\max_i u_i - \min_i u_i|^2/K^3$.}
 
 \subsection{\label{sec:bandpass_poly}
-Bandpass updating schemes constructed from polynomial bases}
+Bandpass updating schemes constructed from orthogonal polynomial bases}
+
+In addition to the above application to
+homogeneous updating schemes,
+we will consider here bandpass updating schemes
+constructed from a sequence of orthogonal polynomials.
+%
+These updating schemes are more intuitive and easier to implement
+as the bias potential in this case is
+a low-order polynomial of the collective variable.
 
@@
 
-We can also construct the bandpass updating schemes
-from an orthogonal polynomial basis,
-which can simplify the functional form of the bias potential.
-%
-In discussing this application,
-it is more convenient to switch to a continuous-variable setup:
-%
-We will replace the discrete bin index by the continuous variable, $z$.
-%
-and map $\rho_i$ to $\rho(z) \, \delta z$,
-  $\phi_{ki}$ to $\phi_k(z) \, \delta z$,
-  $w_{ij}$ to $w(z, z') \, \delta z$, etc.
+It is convenient to
+replace the discrete bin index, $i$,
+by the continuous variable, $z$;
+%
+and we will replace $\rho_i$ by $\rho(z) \, \delta z$,
+  $\phi_{ki}$ by $\phi_k(z) \, \delta z$,
+  and
+  $w_{ij}$ by $w(z, z') \, \delta z$, etc.
 %
@@
 $\phi_k(z) = R_k(z) \, \rho(z)$,
-where $R_k(z)$ are a set of orthogonal polynomials
-satisfying
+where $R_k(z)$ is the $k$th member of an sequence of orthogonal polynomials
+that satisfy
 \begin{align*}
@@
 %
-These eigenvectors allow us to construct the updating matrix
-Eq. \eqref{eq:w_from_phi},
-%we can readily satisfy Eqs. \eqref{eq:eig_orthonormal_cols}-\eqref{eq:ortho0},
+We can readily verify that the eigenvectors satisfy
+Eqs. \eqref{eq:eig_orthonormal_cols}, \eqref{eq:eig_orthonormal_rows},
+and \eqref{eq:ortho0}.
+%
+\note{For $k = 0$, $R_0(z)$ must be a constant,
+  and by the orthonormal condition, we have $R_0(z) = 1$.
+  Thus, Eq. \eqref{eq:ortho0} is satisfied.}%
+%
+The eigenvectors allow us to construct the updating matrix
+from Eq. \eqref{eq:w_from_phi},
 with the eigenvalues given by \eqref{eq:lambda_bandpass}:
@@
 %
-The resulting bias potential
-assumes a polynomial form
-and can be written as a superposition
-$u(z, t) = \sum_{k=1}^{K-1} c_k(t) \, R_k(z)$,
-with the coefficients updated as, according to Eq. \eqref{eq:mbin_update},
+By Eq. \eqref{eq:mbin_update},
+the bias potential is updated as
+%
+$$
+u(z, t+1) = u(z, t)
++ \sum_{k=1}^{K-1} R_k(z) R_k\bigl( z(t) \bigr) \, \alpha(t).
+$$
+%
+We can write the bias potential
+as a superposition of the orthogonal polynomials
+%
+\begin{equation}
+  u(z, t) = \sum_{k=1}^{K-1} c_k(t) \, R_k(z),
+  \label{eq:uz_decomp}
+\end{equation}
+%
+with the coefficients updated as
 \begin{equation}
@@
 
+For $k > 0$, $R_k(z)$ is a polynomial of degree $k$ with zero expectation.
+%
+Thus, $R_k\bigl(z(t)\bigr)$ roughly
+represents the deviation of the $k$th statistical moment
+of the instantaneous distribution
+from its expectation.
+Since high-order moments are generally less reliable,
+this method is best used with a small $K$ for a narrow distribution.
+%
+However, by constructing a mixture or expanded ensemble,
+we can simultaneously sample multiple target distributions
+spread over a wider range, as exemplified below.
+
 \subsubsection{LLR algorithm}
@@
 For example,
-to sample a flat distribution, $\rho(z)$,
+to sample a flat distribution, $\rho(z) = 1/\Delta z$,
 over $[z_c - \Delta z/2, z_c + \Delta z/2]$,
-we can use the Legendre polynomials\cite{arfken},
-and with $K = 2$,
-the bias potential contains only a linear mode for
+we can use the Legendre polynomials\cite{arfken}.
+%
+If the interval is sufficiently small,
+we can limit the basis to a single linear mode,
 $R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$,
-with its coefficient, $c_1$, updated according to
-the difference between the average of $z$
-and the center, $z_c$.
-%
-We thus recover the LLR
-algorithm\cite{langfeld2012},
+with $K = 2$.
+%
+The bias potential is a linear function of $z$,
+and this updating scheme is essentially the LLR algorithm\cite{langfeld2012}
+for the single-interval case,
 and the inverse-time schedule
@@
 
-\subsubsection{Adaptive Gaussian umbrella sampling}
-
-Alternatively,
-to sample a Gaussian distribution
-centered at $z_c$ with width $\sigma_z$
-we can use the Hermite polynomials\cite{arfken}.
-%
-With $K = 3$, the bias potential is quadratic
-and is a linear combination of
-$R_1(z) = (z - z_c)/\sigma_z$
-and
-$R_2(z) = \bigl[(z - z_c)^2 - \sigma_z^2\bigr] /\left(\sqrt 2 \, \sigma_z^2\right)$.
-%
-Note that quadratic bias potentials are commonly
-employed\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007}
-in umbrella sampling\cite{torrie1974, *torrie1977}
+In real application, the algorithm is often applied simultaneously
+to an array of adjacent intervals
+to cover a wider range of the collective variable, $z$.
+%
+In this way, an attempt to move out of the current interval
+does not have to be rejected,
+but can be interpreted as a transition to
+another internal and accepted with a proper probability.
+
+\subsubsection{\label{sec:aus}
+Umbrella sampling with adaptive Gaussian ensembles}
+
+In analogy to the LLR algorithm,
+we will consider the case of sampling a Gaussian distribution
+$\rho(z) \propto \exp[-(z-z_c)^2/(2\sigma_z^2)]$,
+with a basis constructed from
+the Hermite polynomials\cite{arfken}.
+%
+As discussed above,
+the method is effective only to
+a local region of a multimodal distribution,
+and thus we will assume a small width, $\sigma_z$.
+%
+Because the Gaussian distribution has no hard boundaries
+as in the previous case,
+we need to use
+in addition to the linear mode,
+$R_1(z) = (z - z_c)/\sigma_z$,
+a quadratic one,
+$R_2(z) = \bigl[(z - z_c)^2 - \sigma_z^2\bigr] /\left(\sqrt 2 \, \sigma_z^2\right)$,
+to restrict the sampling region.
+%
+This results in a quadratic bias
+potential\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007},
+as commonly employed in umbrella sampling\cite{torrie1974, *torrie1977}
 and the method of Gaussian ensemble\cite{hetherington1987,
-*challa1988, *costeniuc2006},
-which is helpful in overcoming barriers
-in first-order phase transitions.
-%
-By Eq. \eqref{eq:ckupdate},
-$c_1$ and $c_2$ are updated
-according to the first and second
-central moments of $z$, respectively.
-%
-Below we will refer to this application
-as adaptive Gaussian umbrella sampling.
-
-The use of a polynomial basis makes
-the bias potential smooth and differentiable,
-and hence suitable for MD applications.
-%
-However,
-the distribution, $\rho_i$,
-has to be sufficiently narrow
-because of the limited number of modes
-in controlling the moments.
-%
-We can nonetheless sample a wider range
+*challa1988, *costeniuc2006}.
+
+Upon convergence,
+the parameters, $c_1$ and $c_2$,
+can be related to the slope and curvature of the PMF
+[cf. Eq. \eqref{eq:Vi_target}].
+%
+By Eq. \eqref{eq:uz_decomp}, we have
+$$
+u(z) = \frac{c_1}{\sigma_z} (z - z_c)
++ \sqrt 2 \, c_2 \left[ \frac{(z-z_c)^2}{2 \, \sigma_z^2} - \frac{1}{2} \right].
+$$
+Thus, we can interpret
+$c_1/\sigma_z$ as the average local mean force around $z_c$.
+%
+The coefficient
+$\sqrt{2} \, c_2$ represents the relative ``pressure''
+to restrain the width of the distribution to
+be the same as the target Gaussian,
+and it would exceed $1$ in a barrier region
+where the PMF is locally concave,
+$-d^2\ln p(z)/dz^2 < 0$.
+
+We can extend the sampling range
 by constructing a mixture\cite{shirts2017}
@@
   *geyer1991, *hukushima1996, *hansmann1997, *sugita1999,
-  *earl2005, *zuckerman2011, *rauscher2009, kim2010}
+  *earl2005, *zuckerman2011, *rauscher2009,
+  neuhaus2006, *neuhaus2007, kim2010}
 and/or simulated tempering\cite{marinari1992,
@@
 %
-While the mutual exchange probability
-can be readily derived in the former
-framework\cite{neuhaus2006, *neuhaus2007, kim2010},
-the transition probability in the latter case
-requires estimation of the partition function
-of the Gaussian ensemble
-$Z \equiv \int e^{-\beta \, \Phi(\mathbf x)-u(z(\mathbf x),t)} \, d\mathbf x$
-as the bias potential for inter-umbrella transitions.
-%
-Here,
-$\beta$ is the inverse temperature,
-$\Phi(\mathbf x)$ is the unbiased potential energy
-as a function of molecular coordinates, $\mathbf x$,
-and $z$ has also been written as a function of $\mathbf x$.
-%
-The transition probability from
+Here, we will focus on the latter framework,
+in which the transition probability from
 the $i$th umbrella to the $j$th is given by
-\begin{align*}
+%
+\begin{align}
 A(i\to j) =
@@
   ,
-\end{align*}
-where $z(t)$ is the current value of $z$.
-%
-While the asymptotic updating magnitude
-would be given by the inverse-time formula,
-we can employ the WL strategy early in simulation
-to control the updating magnitude in stages
-with the histogram fluctuation
-redefined by Eq. \eqref{eq:maxFk}
-to encompass both intra-umbrella and inter-umbrella
-components.
-%
+  \label{eq:transprob_st}
+\end{align}
+where $z(t)$ is the current value of $z$,
+and $Z^{(i)} \equiv \int e^{-\beta \, \Phi(\mathbf x)-u^{(i)}(z(\mathbf x),t)} \, d\mathbf x$
+is the partition function.
+%
+Here,
+$\beta = 1/(k_B T)$ is the inverse temperature,
+$\mathbf x$ denotes the internal degrees of freedoms,
+and
+$\Phi(\mathbf x)$ is the unbiased potential energy.
+%
+Note that the unknown parameters, $\ln Z^{(i)}$'s,
+serve as the bias potential
+for inter-umbrella transitions, and thus can be updated
+using the single-bin scheme,
+Eq. \eqref{eq:wl_update}\cite{li2007}.
+
+Since bandpass updating schemes are straightforward extension
+of the single-bin updating scheme,
+we can borrow the existing protocol\cite{belardinelli2007,
+  *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}
+to the optimally control updating magnitude.
+%
+We can adopt the WL stage-by-stage strategy in early stages,
+and switch to the inverse-time schedule later
+once the updating magnitude from the former approach
+falls under that from the latter.\footnote{We
+  slightly modified the transition criterion
+  from the WL to the inverse-time schedule.
+  Such a switch was only permitted
+  upon completion of a WL stage,
+  and the time, $t$, was interpreted as the number of updating steps
+  in this just completed WL stage.
+  If the switch was successful,
+  the above $t$ was used as the offset $t_0$
+  in the inverse-time regime.}
+  %
+Since the WL strategy uses the histogram flatness
+as a criterion for stage transitions,
+we will redefine the flatness
+from the fluctuation given by Eq. \eqref{eq:maxFk}
+based on the eigenmode decomposition
+(this allows a uniform definition for
+both the continuous intra-umbrella
+and the discrete inter-umbrella components,
+cf. Appendix \ref{sec:hfluc}).
+%
+%In the asymptotic regime, the optimal schedule
+%is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
+%%
+%In initial stages, when the inverse-time schedule
+%is still inapplicable,
+%we can instead use the WL style stage-wise strategy
+%to control the updating magnitude
+%for the bandpass scheme,
+%since it is a straightforward extension of the single-bin scheme.
+%%
+%However, the histogram fluctuation used for stage transitions
+%should be computed from only the first $K$ modes,
+%where the histogram flattening was performed
+
+We summarize the method as the following procedure.
+
+\begin{enumerate}
+
+\item
+  Set up an array of umbrellas covering the desired range,
+    e.g. we may set
+    $z_c^{(i)} = z_{\min} + i \Delta z$
+    with the same width $\sigma_z$.
+  %
+  For every $i$,
+    set $c_1^{(i)}$ to the estimated mean force,
+    and set $\ln Z^{(i)} = c_2^{(i)} = 0$.
+
+\item
+  Equilibrate the system at umbrella $i = 1$.
+  Repeat the following steps until the simulation ends.
+
+\item
+  Conduct a regular MC or MD step
+  in the current umbrella $i$ under bias potential $u^{(i)}(z, t)$.
+  Update $c_1^{(i)}$ and $c_2^{(i)}$ according to Eq. \eqref{eq:ckupdate},
+  and compute the fluctuations
+  $\tilde F_1^{(i)}$ and $\tilde F_2^{(i)}$
+  defined in Eq. \eqref{eq:Fk_def}.
+
+\item
+  Attempt a transition to another umbrella
+  with the acceptance probability given by Eq. \eqref{eq:transprob_st}.
+  Update $\ln Z^{(i)}$ at the current umbrella
+  by Eq. \eqref{eq:wl_update}.
+
+\item
+  Change the updating magnitude,
+  which will be used for both intra-umbrella
+  parameters, $c_k^{(i)}$,
+  and inter-umbrella parameters, $\ln Z^{(i)}$.
+  %
+
+\end{enumerate}
+
 After simulation,
@@
 
-\note{The
-  following method appeared to have some stability issues,
-  so we switched back to the double-WL approach,
-  which hopefully is stable.
-  %
-  The reason of instability might be
-  that the inter-umbrella fluctuation
-  and the intra-umbrella fluctuation are some coupled.
-  So the updating magnitude cannot be adjusted individually
-  for the each umbrella.
-  %
-  Following the method for weight determination
-  in simulated tempering\cite{park2007,
-  *nguyen2013, *zhang2015st},
-  we can estimate $Z^{(j)}/Z^{(i)}$
-  for a neighboring pair, $(i, j)$, as
-  %
-  \begin{align*}
-    &\ln\frac{ Z^{(j)} }{ Z^{(i)} }
-    \approx
-    \ln\frac{ \sigma_z^{(j)} }{ \sigma_z^{(i)} }
-    +
-    \frac{ \Delta z_c } { 2 }
-    \left(
-      \frac{ c_1^{(i)} } { \sigma_z^{(i) 2} }
-      +
-      \frac{ c_1^{(j)} } { \sigma_z^{(j) 2} }
-    \right)
-    +
-    \frac{ c_2^{(j)} - c_2^{(i)} } { \sqrt 2 }
-    \\
-    &-
-    \left(
-      \frac{ \sqrt 2 \, c_2^{(j)} - 1 }
-           { 2 \, \sigma_z^{(j)} }
-      -
-      \frac{ \sqrt 2 \, c_2^{(i)} - 1 }
-           { 2 \, \sigma_z^{(i)} }
-    \right)
-    \left(
-      \frac{ \Delta z_c^{2} } { 6 }
-      +
-      \frac{ \sigma_z^{(i) 2} + \sigma_z^{(j) 2} } { 2 }
-    \right)
-    ,
-  \end{align*}
-  %
-  where
-  $\Delta z_c \equiv z_c^{(j)} - z_c^{(i)}$.
-  For equal width, we have
-  \begin{align*}
-    \ln\frac{ Z^{(j)} }{ Z^{(i)} }
-    \approx
-    \frac{ \Delta z_c \, ( c_1^{(i)} + c_1^{(j)} ) } { 2 \, \sigma_z }
-    -
-    \left(
-      \frac{ c_2^{(j)} - c_2^{(i)} }
-           { 6 \sqrt 2 }
-    \right)
-    \left(
-      \frac{ \Delta z_c^{2} } { \sigma_z^2 }
-    \right)
-    .
-  \end{align*}
-}
-
 \section{\label{sec:results}
@@
 Our first test system is
-a Lennard-Jones (LJ) fluid of $100$ particles,
+a Lennard-Jones (LJ) fluid of $N = 100$ particles,
 and we wish to compute the PMF along
@@
 %
-The two special particles are mutually non-interacting
+The two special particles, labeled as $1$ and $2$,
+are mutually non-interacting
 but they interact normally with other particles
-via the pair potential $u(r) = 4(r^{-12} - r^{-6})$.
+via the pair potential $\varphi(r) = 4(r^{-12} - r^{-6})$.
 %
@@
   \frac{V}{Z}
-  \int e^{-\Phi(\mathbf r^N)/(k_B T)}
+  \int e^{-\beta \Phi(\mathbf r^N)}
   d\mathbf r_3 \cdots d\mathbf r_N
@@
 where
-$\Phi(\mathbf r^{N}) = \sum_{j=3}^N\sum_{i=1}^{j-1} \phi(r_{ij})$,
-%
-and it is related the cavity distribution function, $y(r)$\cite{hansen},
-as $y(r) = \left(1 - \frac{1}{N}\right) \, V \, p(r)$.
+$V$ is the volume,
+$\Phi(\mathbf r^{N}) = \sum_{j=3}^N\sum_{i=1}^{j-1} \varphi(r_{ij})$,
+and
+$Z = \int e^{\beta \Phi(\mathbf r^N)} d\mathbf r_1 \cdots d\mathbf r_N$.
+%
+This distribution is related the cavity distribution function
+as $y(r) = \left(1 - \frac{1}{N}\right) \, V \, p(r)$\cite{hansen}.
 %
@@
 %
-Further, we added a bias potential $k_B T \, u(r)$,
-which would ideally bias $p(r)$
-to a flat distribution along $r$.
+Further, we added a bias potential, $k_B T \, u(r)$,
+to sample a flat distribution along $r$.
 %
@@
 a random particle is displaced in each permissible dimension
-according to a uniform distribution over $(-\varepsilon^*, \varepsilon^*)$.
-We used $\varepsilon^* = 0.1/\varrho^*$
+according to a uniform distribution over $(-\varepsilon, \varepsilon)$.
+We used $\varepsilon = 0.1/\varrho$
 to ensure reasonable acceptance ratio,
-and $\varrho^*$ is the reduced number density.
+and $\varrho = N/V$ is the reduced number density.
 
@@
 %
-Two densities were tried, $\varrho^* = 0.1$ and $0.8$,
+Two densities were tried, $\varrho = 0.1$ and $0.8$,
 and the numbers of bins were $n = 500$ and $250$, respectively.
 %
-The reduced temperature was $T^* = 3$.
+The reduced temperature was $T = 3$.
 %
@@
 of $10^7$ steps under a constant updating magnitude,
-$a_0 = 10^{-4}$,\footnote{The equivalent $\ln f = n \, a_0$ for the WL algorithm
-were $0.05$ and $0.025$ for $\varrho^* = 0.1$ and $\varrho^* = 0.8$, respectively.}
+$a_0 = 10^{-4}$,\footnote{The
+  seemingly small updating magnitude
+  corresponds to a larger value of $\ln f = n \, a_0$
+  in the WL algorithm.
+  The $\ln f$ was $0.05$ for $\varrho = 0.1$,
+  or $0.025$ for $\varrho = 0.8$.}
 followed by a production phase of another $10^7$ steps
@@
 The Gaussian updating scheme produced less error than
-the single-bin scheme in the $\varrho^* = 0.1$ case,
-but not in the $\varrho^* = 0.8$ case.
+the single-bin scheme in the $\varrho = 0.1$ case,
+but not in the $\varrho = 0.8$ case.
 %
@@
     \multicolumn{4}{c}{
-      $\varrho^* = 0.1$,
+      $\varrho = 0.1$,
       $500$ bins,
@@
     \multicolumn{4}{c}{
-      $\varrho^* = 0.8$,
+      $\varrho = 0.8$,
       $250$ bins,
@@
 %
-The optimal width was $\sigma = 0.37$ for $\varrho^* = 0.1$
-or $0.07$ for $\varrho^* = 0.8$,
+The optimal width was $\sigma = 0.37$ for $\varrho = 0.1$
+or $0.07$ for $\varrho = 0.8$,
 and it also depended on the simulation length
@@
 against the eigenmode index, $k$,
-for the $\varrho^* = 0.1$ case
+for the $\varrho = 0.1$ case
 with width $\sigma = 0.2$.
@@
 %
-For example, for the $\varrho^* = 0.1$ case
+For example, for the $\varrho = 0.1$ case
 with width $\sigma = 0.2$,
-if we increase the simulation length to $10^8$ steps,
+if we extend the simulation to $10^8$ steps,
 the error for the optimal schedule was
@@
 %
-For example, for the simulation in the $\varrho^* = 0.8$ case
+For example, for the simulation in the $\varrho = 0.8$ case
 with $\sigma = 0.5$ and the inverse-time schedule,
@@
 
-\begin{figure}[h]
-\begin{center}
+\begin{figure}[h]\centering
   \makebox[\linewidth][c]{
@@
     of width $\sigma = 0.2$
-    on the LJ system at density $\varrho^* = 0.1$.
+    on the LJ system at density $\varrho = 0.1$.
     %
@@
   }
-\end{center}
 \end{figure}
@@
 %Generally, the optimal schedule
-%for the $\varrho^* = 0.8$ case
+%for the $\varrho = 0.8$ case
 %was closer to the inverse-time schedule
-%than that for the $\varrho^* = 0.1$ case.
+%than that for the $\varrho = 0.1$ case.
 %%
@@
 \subsection{\label{sec:potts}
-Adaptive Gaussian umbrella sampling}
+Umbrella sampling with adaptive Gaussian ensembles}
 
 As another proof of principle,
-we tested the adaptive Gaussian umbrella sampling method
-introduced in Sec. \ref{sec:bandpass}
+we tested the adaptive umbrella sampling method
+introduced in Sec. \ref{sec:aus}
 on the two-dimensional $L\times L$ ten-state
@@
 and the intrinsic distribution is
-the density of states, $\ln g(E)$.
+the density of states, $g(E)$.
 %
@@
 %
-Initially, we set $c_1 = 1.4 \sigma_E$ and $c_2 = 0$.
+Initially, we set
+$c_1 = \tilde \beta_c \, \sigma_E$
+for all umbrellas,
+with $\tilde \beta_c = 1.4$
+being the approximate inverse critical temperature.
 %
@@
 (which is more efficient than the Metropolis algorithm)
-at temperature $c_1/\sigma_E$
+at $\beta = c_1/\sigma_E$
 followed by a Metropolis step with
@@
 $\min\bigl\{1, e^{c_2 [R_2(E) - R_2(E')]} \bigr\}$,
-whose average was no less than 92\%.
-%
-The updating magnitude started at $0.01$,
-and it underwent the WL procedure
-with the fluctuation threshold being $0.2$
-[using the definition of Eq. \eqref{eq:maxFk}]
-and the reduction factor being $1/2$
-until it fell below
-the value given the inverse-time schedule.\footnote{We
-  slightly modified the switch criterion from the WL
-  to the inverse-time schedule\cite{belardinelli2007,
-  *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}.
-  Such a switch was only permitted
-  upon completion of a WL stage,
-  and the time, $t$, was interpreted as the number of updating steps
-  in this just completed WL stage.
-  If the switch was successful,
-  the above $t$ was used as the offset $t_0$
-  in the inverse-time regime.}
-%
-This updating magnitude was
-used to update the values of both $\ln Z^{(i)}$'s
-for inter-umbrella transitions
-and $c_1^{(i)}$'s and $c_2^{(i)}$'s
-for intra-umbrella sampling.
-
-%To ensure stability,
-%we avoid transitions among umbrellas
-%until the current umbrella has
-%completed the WL stage-wise way of
-%reducing the updating magnitude
-%and started the inverse-time schedule.\cite{belardinelli2007,
-%*belardinelli2007jcp, *belardinelli2008, *belardinelli2016}
+where $E$ and $E'$ are the energies of the old and new states.
+%
+The acceptance ratio was no less than 92\% in any umbrella.
+%
+In the initial WL stages,
+the initial updating magnitude was $0.01$;
+the fluctuation threshold for stage transition was $0.2$;
+and the reduction factor was $1/2$.
 
@@
 %
-By WHAM, we found the critical temperature at $0.70162$
+We computed the density of states by WHAM,
+and found the critical temperature to be $T_c = 0.70162$,
 in agreement with a previous study\cite{wang2001pre}.
 %
-The parameter, $c_1/\sigma_E$, roughly represents
+As discussed in Sec. \ref{sec:aus},
+$c_1/\sigma_E$ roughly represents
 the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
@@
 %
-The second term of the bias potential,
-$c_2 \, R_2(E) = \sqrt 2 \, c_2 \, [(E - E_c)^2/(2 \, \sigma_E^2) - \frac 1 2]$,
-provides a quadratic restraint on the distribution width.
-%
-The coefficient, $\sqrt 2 \, c_2$, represents the amount of ``pressure''
-needed to restore the target Gaussian,
-$\rho(E) \propto \exp[-(E - E_c)^2/(2 \, \sigma_E^2)]$,
-and a value over $1.0$ suggests
-that the intrinsic distribution
-(or the density of states here)
-is locally convex or unstable,\cite{kim2010}
-which was so in the critical region,
+The coefficient, $\sqrt 2 \, c_2$,
+was indeed above $1.0$
+in the critical region\cite{kim2010},
 as shown in Fig. \ref{fig:pt_hist}(c).
 
-\begin{figure}[h]
-\begin{center}
+\begin{figure}[h]\centering
   \makebox[\linewidth][c]{
@@
     \label{fig:pt_hist}
-    Adaptive Gaussian umbrella sampling
+    Umbrella sampling with adaptive quadratic bias potentials
     on the $64 \times 64$ ten-state Potts model.
@@
 and the integrals of the autocorrelation functions,
-which represent the intrinsic histogram fluctuation,
+which represent the intrinsic histogram fluctuation
 of the eigenmodes.
@@
 %
-Finally, the adaptive Gaussian umbrella sampling
+Finally, the adaptive umbrella sampling method introduced in Sec. \ref{sec:aus}
 approach requires refinement to improve
-its stability for general application.
+its general applicability.
 %
@@
 
-\section{\label{sec:hfluc}
-Histogram fluctuation}
-
-Here we wish to show that for an FDS simulation
-with a nearly-perfect bias potential,
-the histogram fluctuation
-is roughly equal to the final error
-from a long FDS simulation under the single-bin updating scheme
-with the inverse-time schedule.
-%
-With a sufficiently accurate and stable bias potential,
-the fluctuation of
-the instantaneous histogram is dominated by the random part,
-%
-\begin{equation}
-  f_i(t) \approx \zeta_i(t)
-  ,
-  \label{eq:xi_zeta}
-\end{equation}
-and the same holds for the Fourier transform,
-$\tilde f_k(t) \approx \tilde \zeta_k(t).$
-%
-If we define the time average,
-%
-\begin{align*}
-\tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
-=\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
-,
-\end{align*}
-%
-for the histogram fluctuation of mode $k$,
-then the total fluctuation is
-%
-\begin{align}
-  F^2(T)
-  &\equiv
-  \sum_{k=1}^{n-1} \tilde F_k^2(T)
-  =
-  \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
-  =
-  \sum_{i=1}^n
-  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
-  .
-  \notag
-  %\label{eq:hfluc_def}
-\end{align}
-%
-From Eq. \eqref{eq:xi_zeta}, we have
-%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
-%=
-%\sum_{t=-\infty}^\infty
-%\frac{
-%\bigl\langle
-%  \tilde f_k(t) \, \tilde f_k(0)
-%\bigr\rangle } { T }
-%\approx
-%\sum_{t=-\infty}^\infty
-%\frac{
-%\bigl\langle
-%  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
-%\bigr\rangle } { T }
-%=
-%\frac{ \Gamma_k } { T },$$
-%
-\begin{align*}
-\bigl\langle \tilde F_k^2(T) \bigr\rangle
-&=
-\frac{1}{T}
-\sum_{t=-\infty}^\infty
-\bigl\langle
-  \tilde f_k(t) \, \tilde f_k(0)
-\bigr\rangle
-\\
-&\approx
-\frac{1}{T}
-\sum_{t=-\infty}^\infty
-\bigl\langle
-  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
-\bigr\rangle
-=
-\frac{ \Gamma_k } { T }
-,
-\end{align*}
-%
-and the total histogram fluctuation is
-$\bigl\langle F^2(T) \bigr\rangle = \Gamma/T$.
-%
-Comparing this to Eq. \eqref{eq:Emin_sbin},
-we may rephrase the optimality of the inverse-time schedule
-as the inequality,
-\begin{equation}
-  \Err(T)
-  \ge
-  \bigl\langle F^2(T+t_0) \bigr\rangle
-  .
-  \notag
-  %\label{eq:fluc}
-\end{equation}
-%and the histogram fluctuation
-%serves as a good estimate of the final error.
-%
-
-The relation between error and histogram fluctuation
-supports the WL criterion for stage transitions.
-%
-%For the bandpass updating scheme,
-%we may borrow this criterion
-%in initial stages, with a modified definition
-%for the histogram fluctuation
-%$F^2_{<K}(T) \equiv \sum_{k=1}^{K-1} \tilde F_k^2(T),$
-%since the histogram flattening
-%is only partially applied to modes less than $K$.
-%
-For a bandpass updating scheme,
-we can define the histogram fluctuation
-from the largest mode
-%$\max_{k=1, \dots, K-1} \left| \, \tilde F_k(T) \, \right|$.
-\begin{align}
-  \max_{1 \le k < K} \left| \, \tilde F_k(T) \, \right|
-  ,
-  \label{eq:maxFk}
-\end{align}
-%
-which allows straightforward extension
-to a continuous variable, $z$, with
-$\tilde F_k(T) = \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)$.
-%For a local sampling process,
-%we expect that the two definition would behave similarly.
-%
-\note{
-Note that the expected histogram flatness
-was previously used to define the error\cite{zhou2005, zhou2008},
-and the above demonstration shows the definition
-is compatible with ours, Eq. \eqref{eq:error_def}, in the long-time limit.
-An interesting application\cite{zhou2008},
-of the above result is
-to set the updating magnitude by $F/\Gamma$
-for some empirical value [$\Gamma \approx 10$ there, see Eq. (8)]
-as a substitute of the inverse-time schedule.
-}
-
-\section{\label{sec:Gamma_measure}
-Integrals of the autocorrelation functions of the eigenmodes
-}
-
-Here we give a method of measuring
-the integrals of the autocorrelation functions, $\Gamma_k$'s,
-in an adaptive FDS simulation
-under a constant updating magnitude.
-%
-This method has the advantage of not requiring
-explicit computation of
-the autocorrelation functions of the eigenmodes.
-
-The idea is to use a modification
-of Eq. \eqref{eq:xt2_eql}
-that avoids the problem of not
-knowing the intrinsic distribution, $p_i$,
-hence the potential shift,
-$u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
-%
-Since the shift,
-hence its Fourier transform,
-$\tilde u_k(t) - \tilde v_k(t)$,
-is a constant of time,
-${\tilde u}_{k}$ and ${\tilde v}_{k}$
-should share the same variance.
-%
-But as the ensemble average of ${\tilde v}_{k}$
-is zero, we have
-$\operatorname{var} {\tilde u}_{k} =
- \operatorname{var} {\tilde v}_{k} =
- \left\langle {\tilde v}_{k}^2 \right\rangle$,
-%\begin{equation*}
-%  \operatorname{var} {\tilde u}_{k}
-%  =
-%  \operatorname{var} {\tilde v}_{k}
-%  =
-%  \left\langle {\tilde v}_{k}^2 \right\rangle
-%  .
-%\end{equation*}
-%
-and by using Eq. \eqref{eq:xt2_eql},
-we get
-%
-\begin{equation}
-  \operatorname{var} {\tilde u}_k
-  =
-  \frac{1}{2} \,
-  a_0 \, \Gamma_k \, \lambda_k,
-\label{eq:varXt}
-\end{equation}
-%
-where the variance can be computed from trajectory averages.
-%
-In practice, Eq. \eqref{eq:varXt}
-would fail if some eigenvalue, $\lambda_k$, is close to zero.
-So it is best used for the single-bin updating scheme.
-
-Alternatively,
-if the propagation along the collective-variable is Markovian,
-then in the long time limit, the evolution can be approximately
-described by a transition matrix, $\mathbf A$, with
-\begin{align*}
-\langle h_i(t) \rangle = \rho_i,
-\quad
-\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, \rho_j
-\quad \mbox{(for $t \ge 0$)}
-.
-\end{align*}
-Further if the transition matrix satisfies detailed balance
-with a set of eigenvectors, $\psi_{li}$,
-that are orthonormal with respect to $\pmb\rho$
-[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
-\eqref{eq:eig_orthonormal_rows}]
-then we have the decomposition
-$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
-with $\Lambda_l$ being the eigenvalue of mode $l$.
-%
-The autocorrelation function can then be computed as
-%
-\begin{align*}
-\bigl\langle
-  \tilde f_k(t) \, \tilde f_k(0)
-\bigr\rangle
-=
-\sum_{l = 0}^{n-1} \Lambda_l^t \,
-\left(
-  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
-\right)^2
-.
-\end{align*}
-%
-In particular, if the eigenvectors, $\pmb \phi_k$
-and $\pmb \psi_k$, coincide,
-we would have
-$\bigl\langle
-  \tilde f_k(t) \, \tilde f_k(0)
-\bigr\rangle
-=
-\Lambda_k^t,$
-and
-\begin{equation}
-  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
-  \bigl\langle
-    \tilde f_k(t) \, \tilde f_k(0)
-  \bigr\rangle
-  = \frac{1+\Lambda_k}{1-\Lambda_k}
-  .
-  \notag
-  %\label{eq:Gamma_est}
-\end{equation}
-
-The transition matrix method
-tends to underestimate
-the first few $\Gamma_k$'s.
-%
-We will nonetheless use it as a model to
-study the behavior of $\Gamma_k$
-in a local sampling process.
-%
-If we borrow Eq. \eqref{eq:lambda_Gaussian}
-for $\Lambda_k$,
-with $\sigma_z$ interpreted as the average move size along $z$,
-then
-\begin{align}
-\Gamma_k = \coth\left[
-  \left(
-    \frac{ \pi \, k \, \sigma_z^\mathrm{(samp)} } { g \, \Delta z}
-  \right)^2
-  \right].
-  \notag
-  %\label{eq:Gamma_local}
-\end{align}
-For small $k$,
-we have $\Gamma_k \propto \Delta z^2/k^2$,\cite{bussi2006}
-which means that doubling the sampling range, $\Delta z$,
-would make the histogram flattening
-at least four times as difficult,
-i.e. FDS methods are most effective
-for short ranges\cite{wang2001, wang2001pre}.
-%
-%Using finer bins would, however, not change this value.
-%
-For a modestly large $k$,
-$\Gamma_k \approx 1$,
-and by Eq. \eqref{eq:xt2_eql},
-we would expect a roughly flat tail
-for the single-bin updating scheme
-as $\langle \tilde v_k^2 \rangle \propto \Gamma_k$,
-cf. Fig. \ref{fig:lj_xerr}.
-
 \section{\label{sec:schedule_geometry}
@@
 
-\begin{figure}[h]
-\begin{center}
-  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
-    \node [] at (-0.3, 2.0) {(a)};
-
-    % distribution
-    \draw[domain=0:2.0, variable=\x, dashed, gray, very thick]
-      plot({\x}, {2*exp(-\x)});
-
-    % shade
-    \fill [gray!30!white, domain=0.7:1.609, variable=\x]
-      (0.7, 0) --
-      plot ({\x}, {2*exp(-\x)})
-      -- (1.609, 0) -- cycle;
-
-    % thick curve
-    \draw[domain=0:1.609, variable=\x, very thick]
-      plot({\x}, {2*exp(-\x)});
-
-    % label for the shaded area
-    \node[] at (1.1, 0.32) {$\dfrac{t}{T}$};
-
-    \draw[<->] (0.6, 0)  --
-               node[left, inner sep=1mm]
-               {$\dfrac{1}{T \, \alpha}$}
-               (0.6, 0.99317);
-    % label for the ordinate
-    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
-
-    \draw[] (0.5, 0.99317) -- (0.7, 0.99317);
-    \draw[->] (0, 0) -- (2.2, 0);
-    \node[] at (2.2, -0.2) {$\bar q$};
-    \draw[->] (0, 0) -- node[above, rotate=90] {$m(\bar q)$} (0, 2.2);
-
-    % xtics
-    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
-    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$q(T) - q(t)$};
-    \draw[] (1.609, 0.1) -- (1.609, 0.0) node[below] {$q(T)$};
-  \end{tikzpicture}
-%
-  \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.9}]
-    \node [] at (-0.3, 2.0) {(b)};
-    % distribution
-    \draw[domain=0:2.3, variable=\x, dashed, gray, very thick]
-      plot({\x}, {1/(\x+0.5)});
-
-    % shade
-    \fill [gray!30!white, domain=0.7:2.0, variable=\x]
-      (0.7, 0) --
-      plot ({\x}, {1/(\x+0.5)})
-      -- (2.0, 0) -- cycle;
-
-    % thick curve
-    \draw[domain=0:2.0, variable=\x, very thick]
-      plot({\x}, {1/(\x+0.5)});
-
-    % label for the shaded area
-    \node[] at (1.1, 0.35) {$\bar q(t)$};
-
-    \draw[<->] (0.6, 0)  --
-               node[left, inner sep=1mm]
-               {$\dfrac{1}{T \, m}$}
-               (0.6, 0.83333);
-    % label for the ordinate
-    %\node[fill=white] at (0.4, 0.7) {$\dfrac{1}{Ta}$};
-
-    \draw[] (0.5, 0.83333) -- (0.7, 0.83333);
-    \draw[->] (0, 0) -- (2.6, 0);
-    \node[] at (2.6, -0.2) {$t$};
-    \draw[->] (0, 0) -- node[above, rotate=90] {$\alpha(t)$} (0, 2.2);
-
-    % xtics
-    \draw[] (0.0, 0.1) -- (0.0, 0.0) node[below] {$0$};
-    \draw[] (0.7, 0.1) -- (0.7, 0.0) node[below] {$t$};
-    \draw[] (2.0, 0.1) -- (2.0, 0.0) node[below] {$T$};
-  \end{tikzpicture}
+\begin{figure}[h]\centering
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=\linewidth]{fig/massq.pdf}
+  }
   \caption{
@@
     from the optimal schedule $\alpha(t)$.
-    %
-    \note{
-      The figure was produced by \texttt{tikz}.
-    }%
   }
-\end{center}
 \end{figure}
@@
     \hline
+    $(1+ s_\gamma) \, \gamma \, e^{-\gamma \, {\bar q}^*}$ &
+    $\dfrac{1}{ \gamma \, (t^* + s_\gamma) }$ &
+    $\dfrac{1}{\gamma} \ln\left(\dfrac{1 + s_\gamma}{t^* + s_\gamma} \right)$
+    \\
     $\dfrac{1}{\gamma \, ({\bar q}^* + s_\gamma) }$ &
@@
     $(1 + s_\gamma) \, e^{-\gamma \, t^*} - s_\gamma$
-    \\
-    $(1+ s_\gamma) \, \gamma \, e^{-\gamma \, {\bar q}^*}$ &
-    $\dfrac{1}{ \gamma \, (t^* + s_\gamma) }$ &
-    $\dfrac{1}{\gamma} \ln\left(\dfrac{1 + s_\gamma}{t^* + s_\gamma} \right)$
     \\
@@
 
-\section{\label{sec:more_wband}
+\section{\label{sec:Gamma_measure}
+Integrals of the autocorrelation functions of the eigenmodes
+}
+
+Here we give a method of measuring
+the integrals of the autocorrelation functions, $\Gamma_k$'s,
+in an adaptive FDS simulation
+under a constant updating magnitude.
+%
+This method has the advantage of not requiring
+explicit computation of
+the autocorrelation functions of the eigenmodes.
+
+The idea is to use a modification
+of Eq. \eqref{eq:xt2_eql}
+that avoids the problem of not
+knowing the intrinsic distribution, $p_i$,
+hence the potential shift,
+$u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
+%
+Since the shift,
+hence its Fourier transform,
+$\tilde u_k(t) - \tilde v_k(t)$,
+is a constant of time,
+${\tilde u}_{k}$ and ${\tilde v}_{k}$
+should share the same variance.
+%
+But as the ensemble average of ${\tilde v}_{k}$
+is zero, we have
+$\operatorname{var} {\tilde u}_{k} =
+ \operatorname{var} {\tilde v}_{k} =
+ \left\langle {\tilde v}_{k}^2 \right\rangle$,
+%\begin{equation*}
+%  \operatorname{var} {\tilde u}_{k}
+%  =
+%  \operatorname{var} {\tilde v}_{k}
+%  =
+%  \left\langle {\tilde v}_{k}^2 \right\rangle
+%  .
+%\end{equation*}
+%
+and by using Eq. \eqref{eq:xt2_eql},
+we get
+%
+\begin{equation}
+  \operatorname{var} {\tilde u}_k
+  =
+  \frac{1}{2} \,
+  a_0 \, \Gamma_k \, \lambda_k,
+\label{eq:varXt}
+\end{equation}
+%
+where the variance can be computed from trajectory averages.
+%
+In practice, Eq. \eqref{eq:varXt}
+would fail if some eigenvalue, $\lambda_k$, is close to zero.
+So it is best used for the single-bin updating scheme.
+
+We can alternatively estimate the magnitude of $\Gamma_k$
+by the transition matrix method.
+%
+If the propagation along the collective-variable is Markovian,
+then in the long time limit, the evolution can be approximately
+described by a transition matrix, $\mathbf A$, with
+\begin{align*}
+\langle h_i(t) \rangle = \rho_i,
+\quad
+\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, \rho_j
+\quad \mbox{(for $t \ge 0$)}
+.
+\end{align*}
+Further if the transition matrix satisfies detailed balance
+with a set of eigenvectors, $\psi_{li}$,
+that are orthonormal with respect to $\pmb\rho$
+[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
+\eqref{eq:eig_orthonormal_rows}]
+then we have the decomposition
+$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
+with $\Lambda_l$ being the eigenvalue of mode $l$.
+%
+The autocorrelation function can then be computed as
+%
+\begin{align*}
+\bigl\langle
+  \tilde f_k(t) \, \tilde f_k(0)
+\bigr\rangle
+=
+\sum_{l = 0}^{n-1} \Lambda_l^t \,
+\left(
+  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
+\right)^2
+.
+\end{align*}
+%
+In particular, if the eigenvectors, $\pmb \phi_k$
+and $\pmb \psi_k$, coincide,
+we would have
+$\bigl\langle
+  \tilde f_k(t) \, \tilde f_k(0)
+\bigr\rangle
+=
+\Lambda_k^t,$
+and
+\begin{equation}
+  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
+  \bigl\langle
+    \tilde f_k(t) \, \tilde f_k(0)
+  \bigr\rangle
+  = \frac{1+\Lambda_k}{1-\Lambda_k}
+  .
+  \notag
+  %\label{eq:Gamma_est}
+\end{equation}
+
+The above transition matrix method
+tends to underestimate
+the first few $\Gamma_k$'s.
+%
+We will nonetheless use it as a model to
+study the behavior of $\Gamma_k$
+in a local sampling process.
+%
+If we borrow Eq. \eqref{eq:lambda_Gaussian}
+for $\Lambda_k$,
+with $\sigma_z$ interpreted as
+the average move size along $z$ during sampling,
+then
+\begin{align}
+\Gamma_k = \coth\left[
+  \left(
+    \frac{ \pi \, k \, \sigma_z^\mathrm{(samp)} } { g \, \Delta z}
+  \right)^2
+  \right].
+  \label{eq:Gamma_local}
+\end{align}
+For small $k$,
+we have $\Gamma_k \propto \Delta z^2/k^2$,\cite{bussi2006}
+which means that doubling the sampling range, $\Delta z$,
+would make the histogram flattening
+at least four times as difficult,
+i.e. FDS methods are most effective
+for short ranges\cite{wang2001, wang2001pre}.
+%
+%Using finer bins would, however, not change this value.
+%
+For a modestly large $k$,
+$\Gamma_k$ quickly approaches $1$.
+But since $\langle \tilde v_k^2 \rangle \propto \Gamma_k$,
+according to Eq. \eqref{eq:xt2_eql},
+the error profile for the single-bin updating scheme
+would have a roughly flat tail,
+as shown in Fig. \ref{fig:lj_xerr}.
+
+\section{\label{sec:homo}
 Homogeneous updating schemes
@@
 Here $\mu_l$ denotes the relative updating magnitude
-on the $l$th bin from the center
-(excluding the boundary adjustment).
-%
-We will show below that the eigenvalues of the updating matrix
-is related to the window function as a cosine transform.
-%
-By the translational invariance,
-these schemes satisfy detailed balance,
-Eq. \eqref{eq:w_detailedbalance},
-for the flat distribution, $\rho_i = 1/n$,
-i.e. the updating matrix, $\mathbf w$,
-is symmetric.
-%
-The updating matrices
-share the same set of eigenvectors
+(not adjusted by the boundary correction)
+on the $l$th bin from the center.
+%
+The translational invariance requires $\rho_i$ to be flat,
+and the updating matrix, $\mathbf w$,
+must be symmetric to satisfy detailed balance,
+Eq. \eqref{eq:w_detailedbalance}.
+%
+The updating matrices share the same set of eigenvectors
 (which here are sines and cosines of different frequencies),
@@
 %
-Thus,
-there is a one-to-one correspondence between
+We will show below that the eigenvalues of the updating matrix
+is related to the window function as a cosine transform.
+%
+Thus, there is a one-to-one correspondence between
 the eigenvalues of $\mathbf w$
@@
 For a periodic variable\cite{dama2014},
-the updating matrix, $\mathbf w$,
-assumes the following form
+the updating matrix, $w_{ij}$,
+must be a function of $(i-j)\mod n$,
+and we write it as
 %
@@
 %
-where
-$\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
-will be referred to as a window function.
+for some numbers $\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
+with $\mu_l = 0$ for $l > b$.
+%
+These numbers define a window function.
 %
@@
   \mu_{-b} + \cdots + \mu_b = 1
-  ,
+  .
 \label{eq:mu_normalization}
@@
 %
-with $\mu_l = 0$ for $|l| > b$.
-%
 If $b = n/2$, $\mu_{-b}$ is removed
-from the sum in Eq. \eqref{eq:mu_normalization}
+from the above sum in Eq. \eqref{eq:mu_normalization}
 to avoid double counting.
@@
 %
-\note{The derivation of the first step of
-  Eq. \eqref{eq:wmul_to_convol}:
-\begin{align*}
-  \sum_{j = 1}^n
-    w_{ij} \, \phi_j
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{j = 1}^n
-    \mu_{i - j - n} \, \phi_j
-  +
-  \sum_{j = 1}^n
-    \mu_{i - j + n} \, \phi_j
-  \\
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{l = 1+n}^{2 \, n}
-    \mu_{i - l} \, \phi_{l - n}
-  +
-  \sum_{l = 1-n}^0
-    \mu_{i - l} \, \phi_{l + n}
-  \\
-  &=
-  \sum_{j = 1}^n
-    \mu_{i - j} \, \phi_j
-  +
-  \sum_{l = 1+n}^{2 \, n}
-    \mu_{i - l} \, \phi_{l}
-  +
-  \sum_{l = 1-n}^0
-    \mu_{i - l} \, \phi_{l}
-  \\
-  &=
-  \sum_{j = 1-n}^{2 \, n}
-    \mu_{i - j} \, \phi_j
-  .
-\end{align*}
-The second step of Eq. \eqref{eq:wmul_to_convol}
-follows from the constraint, $\mu_l = 0$, for $|l| > b$.
-
-Similarly,
-for the left eigenvector, we have
-\begin{align*}
-  \sum_{ i = 1 }^n
-    \phi_i \, w_{ij}
-  =
-  \sum_{ l = -b }^b
-    \phi_{j - l} \, \mu_{-l}
-  .
-\end{align*}
-But due to the symmetry Eq. \eqref{eq:mu_symm},
-it is identical to \eqref{eq:wmul_to_convol}.
-}
-%
-As a result, the characteristic equation,
+The characteristic equation,
 $\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$,
-can be solved by discrete Fourier transform.
+can then be solved by discrete Fourier transform.
 %
@@
 %
-where $g = 1$ here, with a two-fold degeneracy,
+with $g = 1$.
+Note that there is a two-fold degeneracy,
 %for the periodic variable,
@@
 %
-\note{The derivation of Eq. \eqref{eq:wmul_to_convol}
-  in this case is similar,
-  \begin{align*}
-    \sum_{j = 1}^n w_{ij} \, \phi_j
-    &=
-    \sum_{j = 1}^n
-      \mu_{i - j} \, \phi_j
-    +
-    \sum_{j = 1}^n
-      \mu_{i + j - 1} \, \phi_j
-    +
-    \sum_{j = 1}^n
-      \mu_{i + j - 2 \, n - 1} \, \phi_j
-    \\
-    &=
-    \sum_{j = 1}^n
-      \mu_{i - j} \, \phi_j
-    +
-    \sum_{l = 1 - n}^0
-      \mu_{i - l} \, \phi_l
-    +
-    \sum_{l = n + 1}^{ 2 \, n }
-      \mu_{i - l} \, \phi_l
-    \\
-    &=
-    \sum_{j = 1 - n}^{ 2 \, n}
-      \mu_{i - j} \, \phi_j.
-  \end{align*}
-}%
-%
 Since Eq. \eqref{eq:phi_refl}
@@
 should share the same periodicity,
-and be even functions about the same axis:
+and be even functions about $i = 1/2$:
 %
@@
 \end{equation}
-%
-%and the eigenvalues are given by
-%  Eq. \eqref{eq:wband_eigenvalue}
-%  with $g = 2$.
-
-\note{Derivation.
-  First consider the unnormalized eigenvectors,
-  \begin{align*}
-  \Phi^{(k)}_j
-  =
-  \cos \frac{ k \, \left( j - \frac 1 2 \right) \, \pi }{n},
-  \end{align*}
-  which satisfies Eq. \eqref{eq:phi_refl}.
-  %
-  So, by Eq. \eqref{eq:wmul_to_convol},
-  \begin{align*}
-  \sum_{i = 1}^n
-    \Phi^{(k)}_i \, w_{ij}
-  &=
-  \sum_{l = -b}^b
-    \Phi^{(k)}_{j - l} \, \mu_l
-  \\
-  &=
-    \Phi^{(k)}_j \, \mu_0
-  + \sum_{l=0}^{b}
-    \mu_l \,
-    \left[
-      \Phi^{(k)}_{j-l}
-      +
-      \Phi^{(k)}_{j+l}
-    \right]
-  \\
-  &= \Phi^{(k)}_j \, \lambda_k,
-  \end{align*}
-  with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}
-  with $g = 2$.
-}%
-%
 
@@
 
-\note{On the inversion.
-Eq. \eqref{eq:lambdan}
-  ensures that the $\mu_n$
-  computed from Eq. \eqref{eq:mu_from_lambda}
-  vanishes.
-
-  %The inversion formula is derived from the cosine transform.
-  %
-  If we define $\mu_n = 0$ and
-  %
-  $
-    \mu_{ 2 n - l } = \mu_l,
-  $
-  for
-  $l = 1, \dots, n - 1$,
-  %
-  then Eq. \eqref{eq:wband_eigenvalue} ($g = 2$)
-  can be rewritten as
-  %
-  \begin{align}
-    \lambda_k
-    =
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-    \mu_l \, \cos \left(\frac{ k \, l \, \pi } { n } \right).
-  \notag
-  %\label{eq:lambda_cosine_sum}
-  \end{align}
-  %
-  This formula for $\lambda_k$
-  can be readily extended to $k = 2 \, n - 1$,
-  and we have
-  %
-  $
-    \lambda_{ 2 \, n - k } = \lambda_k
-    .
-  $
-  %
-  Thus,
-  %
-  \begin{align*}
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \lambda_k \,
-      \cos \frac{ k \, p \, \pi }
-                {      n        }
-    &=
-    \sum_{ k = 0 }^{ 2 \, n - 1 }
-      \sum_{ l = 0 }^{ 2 \, n - 1 }
-        \mu_l \,
-        \cos \frac{ k \, l \, \pi }
-                  {      n        }
-        \cos \frac{ k \, p \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \frac{ \mu_l } { 2 }
-      \sum_{ k = 0 }^{ 2 \, n - 1 }
-        \cos \frac{ k \, (p + l) \, \pi }
-                  {      n        }
-                  +
-        \cos \frac{ k \, (p - l) \, \pi }
-                  {      n        }
-    \\
-    &=
-    \sum_{ l = 0 }^{ 2 \, n - 1 }
-      \mu_l \, n \left(
-        \delta_{ p + l - 2 \, n, 0 }
-        +
-        \delta_{ p - l, 0 }
-      \right)
-    \\
-    &=
-    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
-    =
-    2 \, n \, \mu_p.
-  \end{align*}
-  This entails Eq. \eqref{eq:mu_from_lambda}.
-  %
-  The value of $\lambda_n$
-  can be deduced from
-  imposing $\mu_n = 0$
-  and solving Eq. \eqref{eq:mu_from_lambda}
-  for $l = n$,
-  yielding Eq. \eqref{eq:lambdan}.
-}
-
 \subsection{Gaussian updating scheme}
@@
 %
-Below we show its stability
-in the continuous limit.
-
-For $n \gg 1$,
-we define an equivalent continuous variable, $z$,
-over $[0, \Delta z]$
-with bin size
-$\delta z = \Delta z/(g \, n)$.
-The variable is related to the bin index, $i$, as
+Below we show its convergence
+in the continuous limit,
+and how to maintain the convergence
+on a finite grid.
+
+We recall that
+for continuous variable, $z$, over $[0, \Delta z]$
+the bin index, $i$, is related to $z$ as
 $z = i \, \delta z$,
+where
+$\delta z = \Delta z/(g \, n)$ is th bin size.
+
 and
@@
 %
+For $n \gg 1$,
 By assuming the largest bin cutoff, $b \approx g \, n/2$,
@@
 In practice, however, some eigenvalues can turn negative
-because of the truncation and discretization.
-%
-We can maintain the stability of the updating scheme
+when discretized on a finite grid.
+%
+We can maintain the convergence of the updating scheme
 by redefining the Gaussian scheme
-starting with Eq. \eqref{eq:lambda_Gaussian}
-and inversely computing the window function
-using Eq. \eqref{eq:mu_from_lambda}.
-%
+from the same spectrum of eigenvalues, Eq. \eqref{eq:lambda_Gaussian},
+and then inversely computing the window function
+using Eq. \eqref{eq:mu_from_lambda},
+or the updating matrix using Eq. \eqref{eq:w_from_phi}.
+%
+An example is shown in Fig. \ref{fig:mat}(a).
+
+\begin{figure}[h]\centering
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
+  }
+  \caption{
+    \label{fig:mat}
+    Updating matrices, $w_{ij}$, for
+    homogeneous updating schemes of
+    a non-periodic variable ($n = 500$).
+    Here $j$ is the bin of the current collective variable,
+    and $i$ is the bin where the bias potential is updated.
+    %
+    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
+    (b) The bandpass updating scheme of $K = 20$.
+    %
+    Both updating schemes
+    can be characterized by a rigid window
+    centered on the current bin, $j$,
+    with some boundary adjustments.
+  }
+\end{figure}
 
@@
 
-For the bandpass updating schemes,
+In a homogeneous bandpass updating scheme,
 the eigenvalues are either $0$ or $1$ and
@@
 %
-\note{Derivation.
-\begin{align*}
-\mu_l
-&=
-\frac 1 n \sum_{k = 0}^{n-1} \lambda_k \cos \frac{ k \, l \, 2 \, \pi } { n }
-\\
-&=
-\frac{1}{n}
-\left(
-  1 +
-  \sum_{k=1}^{K-1}
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-  +
-  \sum_{k=n-K+1}^{n-1}
-  \cos \frac { k \, l \, 2 \, \pi } { n }
-\right)
-\\
-&=
-\frac 1 n
-\sum_{k=1-K}^{K-1}
-\cos \frac { k \, l \, 2 \, \pi } { n }
-=
-  \frac{
-    \sin
-    \frac{ (2 K - 1) \, l \, \pi }
-         {              n        }
-  }
-  {
-    n \, \sin \frac{ l \, \pi } { n }
-  }
-.
-\end{align*}
-}
 
@@
 %
-\note{Derivation.
+The homogeneous bandpass updating scheme
+resembles the Gaussian updating scheme
+except for some long-range wiggles,
+as shown in Fig. \ref{fig:mat}(b).
+%
+
+\section{\label{sec:hfluc}
+Histogram fluctuation}
+
+We will first review the relation
+between the histogram fluctuation and the error
+in a long FDS simulation under the single-bin updating scheme.
+%
+We then give a measure of the overall histogram fluctuation
+that can be used in the WL-like criterion of stage transitions
+for the bandpass updating scheme based on orthogonal polynomials.
+%as a generalization of the WL algorithm for the single-bin scheme.
+
+\subsection{Histogram fluctuation and error}
+
+Let us consider an FDS simulation under a sufficiently accurate bias potential.
+The fluctuation of the instantaneous histogram
+is then dominated by the random part,
+%
+\begin{equation}
+  f_i(t) \approx \zeta_i(t)
+  ,
+  \label{eq:xi_zeta}
+\end{equation}
+and the same holds for the Fourier transform,
+$\tilde f_k(t) \approx \tilde \zeta_k(t).$
+%
+If we define the time average,
+%
 \begin{align*}
-  \lambda_{n}
+\tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
+=\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
+,
+\end{align*}
+%
+for the histogram fluctuation of mode $k$,
+then the total fluctuation is
+%
+\begin{align}
+  F^2(T)
+  &\equiv
+  \sum_{k=1}^{n-1} \tilde F_k^2(T)
+  =
+  \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
+  =
+  \sum_{i=1}^n
+  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
+  .
+  \notag
+  %\label{eq:hfluc_def}
+\end{align}
+%
+From Eq. \eqref{eq:xi_zeta}, we have
+%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
+%=
+%\sum_{t=-\infty}^\infty
+%\frac{
+%\bigl\langle
+%  \tilde f_k(t) \, \tilde f_k(0)
+%\bigr\rangle } { T }
+%\approx
+%\sum_{t=-\infty}^\infty
+%\frac{
+%\bigl\langle
+%  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
+%\bigr\rangle } { T }
+%=
+%\frac{ \Gamma_k } { T },$$
+%
+\begin{align*}
+\bigl\langle \tilde F_k^2(T) \bigr\rangle
 &=
-  (-1)^{n-1}
-  \left[
-    \lambda_0
-    - 2 \, \lambda_1
-    + 2 \, \lambda_2 - \cdots
-    + (-1)^{K-1} 2 \, \lambda_{K-1})
-  \right]
+\frac{1}{T}
+\sum_{t=-\infty}^\infty
+\bigl\langle
+  \tilde f_k(t) \, \tilde f_k(0)
+\bigr\rangle
 \\
-  &=
-  \begin{dcases}
-    (-1)^{n-1} & K \mathrm{\; odd,} \\
-    (-1)^n     & K \mathrm{\; even.}
-  \end{dcases}
+&\approx
+\frac{1}{T}
+\sum_{t=-\infty}^\infty
+\bigl\langle
+  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
+\bigr\rangle
+=
+\frac{ \Gamma_k } { T }
+,
 \end{align*}
-So
+%
+and the total histogram fluctuation is
+\begin{equation}
+  \bigl\langle F^2(T) \bigr\rangle
+  =
+  \sum_{k=1}^{n-1}
+  \bigl\langle \tilde F_k^2(T) \bigr\rangle
+  =
+  \sum_{k=1}^{n-1}
+  \frac{ \Gamma_k } { T }
+  =
+  \frac{ \Gamma } { T }
+  ,
+  \label{eq:F2sum}
+\end{equation}
+where we have used Eq. \eqref{eq:Gammak_sum}
+in the last step.
+%
+We may relate the histogram fluctuation
+to the minimal error under the single-bin updating scheme,
+Eq. \eqref{eq:Emin_sbin}, as
+%we may rephrase the optimality of the inverse-time schedule
+%as the inequality,
+\begin{equation}
+  \Err(T)
+  \ge
+  \bigl\langle F^2(T+t_0) \bigr\rangle
+  .
+  \notag
+  %\label{eq:fluc}
+\end{equation}
+%and the histogram fluctuation
+%serves as a good estimate of the final error.
+%
+
+\subsection{Overall histogram fluctuation}
+
+Equation \eqref{eq:F2sum} provides a way of
+defining the overall histogram fluctuation,
+hence the histogram flatness,
+which can be used in the WL algorithm
+as a criterion for transition to new stage with reduced updating magnitude.
+%
+This definition, however,
+diverges in the limit of $n \to \infty$,
+when the bin size becomes infinitely small
+and collective variable becomes continuous.
+%
+Thus, we find it more suitable to define the histogram fluctuation
+from the mode with the largest $|\tilde F_k(T)|$.
+%
+The two definitions are often congruent with each other
+for a local sampling process with a finite $n$
+because the major contribution to $\langle F^2(T) \rangle$
+often comes from the few longest-range modes [cf. Eq. \eqref{eq:Gamma_local}].
+%
+For the bandpass updating scheme,
+we need to further restrict the mode index by $k < K$
+since the histogram flattening applies
+only to the first $K$ modes.
+%
+Thus, the new definition of overall histogram fluctuation
+is given by
+\begin{align}
+  \max_{1 \le k < K} \left| \, \tilde F_k(T) \, \right|
+  .
+  \label{eq:maxFk}
+\end{align}
+
+If the eigenmodes are made of a sequence of orthogonal polynomials
+as in Sec. \ref{sec:bandpass_poly},
+we have
 \begin{align*}
-  \mu_l
-  &=
-  \frac{1}{2\,n}
-  \left[
-    1 +
-    2 \sum_{k=1}^{K-1}
-    \cos \frac { k \, l \pi } { n }
-    +
-    (-1)^{n-K} (-1)^l
-  \right]
-  \\
-  &=
-  \frac{1}{2\,n}
-  \left[
-    (-1)^{n-K+l}
-    +
-    \sum_{k=1-K}^{K-1}
-    \cos \frac { k \, l \pi } { n }
-  \right]
-  .
+\tilde f_k(t)
+=
+\int \phi_k(z) \, \frac{ \delta\bigl(z(t) - z\bigr) } { \rho(z) } \, dz
+=
+R_k\bigl( z(t) \bigr)
+,
 \end{align*}
-}%
+so that
+\begin{equation}
+  \tilde F_k(T)
+  =
+  \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)
+  =
+  \overline{
+    R_k\bigl( z(t) \bigr)
+  }
+  .
+  \label{eq:Fk_def}
+\end{equation}
 
@@
     $\sigma_{ij}(t)$   &   Autocorrelation function of $\zeta_i(t)$. \\
-    $\sigma_\varphi$   &   Width of the Gaussian updating scheme in radians. \\
     $t, \tau$       &   Time. \\

### 2017-08-17 doc/invt.tex
add getav.py

@@
 As shown in Table \ref{tab:lj_error},
-all updating schemes
-with either the inverse-time or optimal schedule
-were able to reduced
-the error significantly,
-and the errors roughly matched the predicted values
+both the optimal and inverse-time schedules
+were able to reduced the error significantly,
+suggesting the potential benefit of using a decreasing updating magnitude
+even with a sometimes sub-optimal schedule.
+%
+It is also reassuring to see that
+the errors roughly matched the predicted values
 calculated using the pre-computed
@@
 %
-The Gaussian updating scheme produced less error than
-the single-bin scheme in the $\varrho = 0.1$ case,
-but not in the $\varrho = 0.8$ case.
-%
-The effectiveness of the Gaussian updating scheme
-in the low-density case
-was probably because the sampling along
-the collective variable was free from major barriers
-and thus the histogram fluctuation largely
-came from local noise.
-%
-%The optimal schedule also appeared to be more effective
-%in reducing the error than the inverse-time schedule.
-%%
-In this case, we could further reduced the error
-by using the bandpass updating scheme
-(with the mode cutoff $K = 20$).
-%
-On the other hand, in the high-density case,
-the error was dominated by the $k=1$ eigenmode
-corresponding to the longest range fluctuation,
-as $\Gamma_1 \approx \Gamma$,
-and the reduction of local noise
-had little effect on the total error.
+Since the two schedules are identical
+in the single-bin and bandpass updating schemes,
+we will focus on the Gaussian updating scheme below
+in discussing the optimal schedule.
 
@@
 
-For the Gaussian updating scheme,
-there was an optimal width for maximum effectiveness.
-%
-The optimal width was $\sigma = 0.37$ for $\varrho = 0.1$
-or $0.07$ for $\varrho = 0.8$,
-and it also depended on the simulation length
-and other parameters.
-%
-Although widening the Gaussian can
-effectively suppress random error
-from local histogram fluctuation,
-it also increased the risk of including
-more systematic bias in middle-range modes,
-which can be difficult to remove
-even with the optimal schedule.
-%
-This is because the rate of error reduction
-is proportional to the eigenvalue, $\lambda_k$
-[cf. Eq. \eqref{eq:vt_diffeq_mbin}],
-which decreases rapidly with the mode index $k$,
-cf. Eq. \eqref{eq:lambda_Gaussian}.
-
-To illustrate the mode dependence of the error,
-we show in Fig. \ref{fig:lj_xerr}
-the error components, $\langle \tilde v_k^2 \rangle$,
-against the eigenmode index, $k$,
-for the $\varrho = 0.1$ case
-with width $\sigma = 0.2$.
-%
-For modes $k=10$ to $25$,
-the final errors from the Gaussian updating scheme
-were greater than those from the single-bin updating scheme,
-although the initial values were less.
-%
-For the Gaussian updating scheme,
-the inverse-time schedule tended to
-reduce the updating magnitude too quickly,
-leaving a hump on the error profile
-for these middle-range eigenmodes.
-%
-The optimal schedule improved over the inverse-time schedule
-and produced a flatter profile for the final error.
-%
-As a result, the optimal schedule usually produced
-less error than the inverse-time schedule,
-as shown in Table \ref{tab:lj_error}.
-%
-Although the gain of using the optimal schedule
+As shown in Fig. \ref{fig:lj_alpha},
+the optimal schedules for the Gaussian updating schemes
+differed from inverse-time one,
+and the former tended to be larger.
+%
+The optimal schedule was more effective
+than the sub-optimal inverse-time schedule
+in reducing the error,
+as shown in Table \eqref{tab:lj_error}.
+%
+Although the advantage of the former
 was modest in the reported cases,
@@
 %
-For example, for the $\varrho = 0.1$ case
-with width $\sigma = 0.2$,
+For example, for the case of
+$\varrho = 0.1$ and Gaussian width $\sigma = 0.2$,
 if we extend the simulation to $10^8$ steps,
@@
 
+To illustrate the mode dependence of the error,
+we show in Fig. \ref{fig:lj_xerr}
+the error components, $\langle \tilde v_k^2 \rangle$,
+against the eigenmode index, $k$,
+for the $\varrho = 0.1$ case
+with width $\sigma = 0.2$.
+%
+For modes $k=10$ to $25$,
+the final errors from the Gaussian updating scheme
+were greater than those from the single-bin updating scheme,
+although the initial values were less.
+%
+For the Gaussian updating scheme,
+the inverse-time schedule tended to
+reduce the updating magnitude too quickly,
+leaving a hump on the error profile
+for these middle-range eigenmodes.
+%
+The optimal schedule improved over the inverse-time schedule
+and produced a flatter profile for the final error.
+%
+As a result, the optimal schedule usually produced
+less error than the inverse-time schedule,
+as shown in Table \ref{tab:lj_error}.
+%
+
 One way to overcome the systematic bias
@@
 (Table \ref{tab:lj_error}) to $3.3\times10^{-3}$.
+
+\begin{figure}[h]\centering
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_alpha.pdf}
+  }
+  \caption{
+    \label{fig:lj_alpha}
+    Optimal schedules of Gaussian updating schemes
+    for the LJ system.
+    %
+    The starting updating magnitude was usually
+    around $a_0/2 = 5\times10^{-5}$
+    as given by Eq. \eqref{eq:half_alpha0}.
+    But an overly wide window ($\sigma = 0.6$)
+    led to a much larger updating magnitude.
+  }
+\end{figure}
+
+For the Gaussian updating scheme,
+there was an optimal width for maximum effectiveness.
+%
+The optimal width was $\sigma = 0.37$ for $\varrho = 0.1$
+or $0.07$ for $\varrho = 0.8$,
+and it also depended on the simulation length
+and other parameters.
+%
+Although widening the Gaussian can
+effectively suppress random error
+from local histogram fluctuation,
+it also increased the risk of including
+more systematic bias in middle-range modes,
+which can be difficult to remove
+even with the optimal schedule.
+%
+This is because the rate of error reduction
+is proportional to the eigenvalue, $\lambda_k$
+[cf. Eq. \eqref{eq:vt_diffeq_mbin}],
+which decreases rapidly with the mode index $k$,
+cf. Eq. \eqref{eq:lambda_Gaussian}.
 
@@
 
-In Fig. \ref{fig:lj_alpha},
-we show the optimal schedules in a few cases.
-%
-%Generally, the optimal schedule
-%for the $\varrho = 0.8$ case
-%was closer to the inverse-time schedule
-%than that for the $\varrho = 0.1$ case.
+The Gaussian updating scheme produced less error than
+the single-bin scheme in the $\varrho = 0.1$ case,
+but not in the $\varrho = 0.8$ case.
+%
+The effectiveness of the Gaussian updating scheme
+in the low-density case
+was probably because the sampling along
+the collective variable was free from major barriers
+and thus the histogram fluctuation largely
+came from local noise.
+%
+%The optimal schedule also appeared to be more effective
+%in reducing the error than the inverse-time schedule.
 %%
-%This was probably because in the former case
-%the first mode dominates the total error,
-%$\Gamma_1 \approx \Gamma$,
-%and the schedule was roughly optimized for this mode,
-%resulting in
-%$\alpha(t) \approx 1/[\lambda_1 (t + t_0)] \approx 1/(t+t_0)$
-%as $\lambda_1 \approx 1$
-%[cf. Eq. \eqref{eq:lambda_Gaussian}].
-%
-Since we have chosen a long equilibration period
-for the preliminary run under the constant updating magnitude, $a_0$,
-the initial magnitude was around
-the ideal value of $a_0/2 = 5\times10^{-5}$
-given by Eq. \eqref{eq:half_alpha0}.
-%
+In this case, we could further reduced the error
+by using the bandpass updating scheme
+(with the mode cutoff $K = 20$).
+%
+On the other hand, in the high-density case,
+the error was dominated by the $k=1$ eigenmode
+corresponding to the longest range fluctuation,
+as $\Gamma_1 \approx \Gamma$,
+and the reduction of local noise
+had little effect on the total error.
+
 But if the Gaussian were too wide,
@@
 
-\begin{figure}[h]\centering
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/lj_alpha.pdf}
-  }
-  \caption{
-    \label{fig:lj_alpha}
-    Optimal schedules of Gaussian updating schemes
-    for the LJ system.
-    %
-    The starting updating magnitude was usually
-    around $a_0/2 = 5\times10^{-5}$
-    as given by Eq. \eqref{eq:half_alpha0}.
-    But an overly wide window ($\sigma = 0.6$)
-    led to a much larger updating magnitude.
-  }
-\end{figure}
+The initial magnitude was, however, around
+the ideal value of $a_0/2 = 5\times10^{-5}$
+given by Eq. \eqref{eq:half_alpha0}.
+%
+Since we have chosen a long equilibration period
+for the preliminary run under the constant updating magnitude, $a_0$,
 

### 2017-08-17 prog/lj/getav.py

@@ -1,0 +1,26 @@
+#!/usr/bin/env python
+
+fn = "verr.log"
+
+class Ave:
+    def __init__(self):
+        self.cnt = 0
+        self.sx  = 0
+        self.sxx = 0
+
+    def add(self, x):
+        self.cnt += 1
+        self.sx += x
+        self.sxx += x * x
+
+    def getave(self):
+        return self.sx/(self.cnt+1e-300)
+
+av = Ave()
+
+for ln in open(fn).xreadlines():
+    if ln.startswith("#"): continue
+    x = float( ln.split()[1] )
+    av.add(x)
+
+print av.cnt, av.getave()

### 2017-08-18 doc/invt.tex
invt.tex: editing results

@@
   or $0.025$ for $\varrho = 0.8$.}
-followed by a production phase of another $10^7$ steps
+followed by a production phase of
+another $10^7$ steps (unless indicated otherwise)
 with updating magnitude given by
@@
 Eqs. \eqref{eq:v_def} and \eqref{eq:x_def},
-and decomposed to eigenmodes by
-Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl}.
-%
-The error was computed from Eq. \eqref{eq:error_def}.
+and the error from Eq. \eqref{eq:error_def}.
 %
@@
 
-To illustrate the mode dependence of the error,
-we show in Fig. \ref{fig:lj_xerr}
-the error components, $\langle \tilde v_k^2 \rangle$,
-against the eigenmode index, $k$,
-for the $\varrho = 0.1$ case
-with width $\sigma = 0.2$.
-%
-For modes $k=10$ to $25$,
-the final errors from the Gaussian updating scheme
-were greater than those from the single-bin updating scheme,
-although the initial values were less.
-%
-For the Gaussian updating scheme,
-the inverse-time schedule tended to
-reduce the updating magnitude too quickly,
-leaving a hump on the error profile
-for these middle-range eigenmodes.
-%
-The optimal schedule improved over the inverse-time schedule
-and produced a flatter profile for the final error.
-%
-As a result, the optimal schedule usually produced
-less error than the inverse-time schedule,
-as shown in Table \ref{tab:lj_error}.
-%
-
-One way to overcome the systematic bias
-in the Gaussian updating scheme
-is to use the histogram-based correction formula,
-Eq. \eqref{eq:uhatav}.
-%
-As shown in Fig. \ref{fig:lj_xerr},
-even for the simulation using the non-optimal $1/t$ schedule,
-the correction was able to reproduce
-the flatter profile seen in the single-bin case.
-%
-However, this correction also introduced
-some local noise in the large $k$ modes,
-and it would be best used as a post-simulation
-analysis tool for simulations
-with detectable residual bias.
-%
-For example, for the simulation in the $\varrho = 0.8$ case
-with $\sigma = 0.5$ and the inverse-time schedule,
-the correction was able to lower
-the error from $9.8\times10^{-3}$
-(Table \ref{tab:lj_error}) to $3.3\times10^{-3}$.
-
 \begin{figure}[h]\centering
@@
 
-For the Gaussian updating scheme,
-there was an optimal width for maximum effectiveness.
-%
-The optimal width was $\sigma = 0.37$ for $\varrho = 0.1$
-or $0.07$ for $\varrho = 0.8$,
-and it also depended on the simulation length
-and other parameters.
-%
-Although widening the Gaussian can
-effectively suppress random error
-from local histogram fluctuation,
-it also increased the risk of including
-more systematic bias in middle-range modes,
-which can be difficult to remove
-even with the optimal schedule.
-%
-This is because the rate of error reduction
-is proportional to the eigenvalue, $\lambda_k$
-[cf. Eq. \eqref{eq:vt_diffeq_mbin}],
-which decreases rapidly with the mode index $k$,
-cf. Eq. \eqref{eq:lambda_Gaussian}.
+For the same case,
+we further decomposed the error (after $10^7$ production steps)
+into the eigenmodes
+by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl},
+and ploted the error components, $\langle \tilde v_k^2 \rangle$,
+in Fig. \ref{fig:lj_xerr}.
+%
+We found that the inverse-time schedule tended to
+reduce the updating magnitude too quickly,
+leaving a hump on the error profile
+for the $k = 10$ to $25$ modes.
+%
+By a delayed reduction,
+the optimal schedule more effectively reduced
+the error in these modes,
+and produced a flatter profile for the final error.
 
@@
     of width $\sigma = 0.2$
-    on the LJ system at density $\varrho = 0.1$.
+    on the LJ system at density $\varrho = 0.1$
+    after $10^7$ production steps.
     %
@@
 
-The Gaussian updating scheme produced less error than
-the single-bin scheme in the $\varrho = 0.1$ case,
-but not in the $\varrho = 0.8$ case.
-%
-The effectiveness of the Gaussian updating scheme
-in the low-density case
-was probably because the sampling along
-the collective variable was free from major barriers
+Next, we will discuss the applicability of
+the Gaussian updating scheme.
+%
+As shown in Table \ref{tab:lj_error},
+the Gaussian updating schemes,
+or more generally the bandpass updating schemes,
+produced less error than the single-bin scheme 
+only in the low-density $\varrho = 0.1$ case,
+but not in the high-density $\varrho = 0.8$ case.
+%
+We believe that the difference lies in the sampling difficulty.
+%
+In the low-density case,
+the sampling along the collective variable
+was free from major barriers,
 and thus the histogram fluctuation largely
-came from local noise.
-%
-%The optimal schedule also appeared to be more effective
-%in reducing the error than the inverse-time schedule.
-%%
-In this case, we could further reduced the error
-by using the bandpass updating scheme
-(with the mode cutoff $K = 20$).
-%
-On the other hand, in the high-density case,
-the error was dominated by the $k=1$ eigenmode
-corresponding to the longest range fluctuation,
-as $\Gamma_1 \approx \Gamma$,
-and the reduction of local noise
+was largely local in nature.
+%
+In the high-density case, however,
+the error was dominated by the first few eigenmodes
+that correspond to the long-range fluctuations
+(as $\Lambda_1 \approx \Lambda$),
+and the reduction of local noise by multiple-bin schemes
 had little effect on the total error.
-
+%
+When the local noise are dominant,
+we found that the bandpass updating scheme
+(with the mode cutoff $K = 20$)
+was more effective than the Gaussian updating schemes
+as the expected.
+
+It is worth mentioning that the width $\sigma$
+of a Gaussian updating scheme
+can critially affect its effectiveness.
+%
+For example, the prediction showed that
+there was a optimal width
+around $\sigma = 0.37$ for $\varrho = 0.1$
+or $0.07$ for $\varrho = 0.8$,
+assuming the optimal schedule,
+and the optimal width also depended on the simulation length
+and other parameters.
+%
+We recall that a wider Gaussian means
+more middle-range eigemodes
+would have diminishing eigenvalues
+by Eq. \eqref{eq:lambda_Gaussian}.
+%
+Thus, although widening the Gaussian can
+effectively suppress random error
+from local histogram fluctuation,
+it also increased the risk of including 
+more systematic bias in middle-range modes,
+which can be difficult to remove
+as the rate of error reduction
+is also proportional to the eigenvalue, $\lambda_k$,
+by Eq. \eqref{eq:vt_diffeq_mbin}.
+
+An overly large width $\sigma$ can have a dramatic
+impact on the initial updating magnitude
+under the optimal schedule.
+%
+The initial updating magnitude, $\alpha(0)$,
+is usually around
+the ideal value of $a_0/2 = 5\times10^{-5}$
+given by Eq. \eqref{eq:half_alpha0},
+for a sufficiently long preliminary run
+under the constant updating magnitude, $a_0$,
+as shown in Fig. \ref{fig:lj_alpha}.
+%
 But if the Gaussian were too wide,
@@
 became too small to damp out the initial bias,
-resulting in a much larger initial updating magnitude, $\alpha(0)$
+resulting in a much larger $\alpha(0)$
 [cf. Eq. \eqref{eq:opt_alpha0}].
 %
-In this case, the predicted schedule
-would be unusable because the resulting deviation in the bias potential
-was no longer small,
-and one may have to either decrease the width
-or to elongate the preliminary run.
-
-The initial magnitude was, however, around
-the ideal value of $a_0/2 = 5\times10^{-5}$
-given by Eq. \eqref{eq:half_alpha0}.
-%
-Since we have chosen a long equilibration period
-for the preliminary run under the constant updating magnitude, $a_0$,
+In this case, the predicted optimal schedule
+becomes invalid the resulting deviation in the bias potential
+is no longer small.
+
+In practice, both the width and schedule
+of the Gaussian updating scheme can be sub-optimal.
+%
+We can use the histogram-based correction formula,
+Eq. \eqref{eq:uhatav}, after simulation
+to remove the remaining bias in PMF.
+%
+In Fig. \ref{fig:lj_xerr},
+we show a case where the correction produced
+the flatter profile seen in the single-bin case,
+for a simulation with a sub-optimal schedule.
+%
+However, this correction will also introduce
+some local noise in the large $k$ modes,
+although these noise are largely harmless
+in a post-simulation analysis, since
+the bias potential no longer need to be differentiated
+for force calculation.
+%
+%For example, for the simulation in the $\varrho = 0.8$ case
+%with $\sigma = 0.5$ and the inverse-time schedule,
+%the correction was able to lower
+%the error from $9.8\times10^{-3}$ to $3.3\times10^{-3}$.
 

### 2017-08-18 doc/invt.tex
invt.tex: small edits

@@
 
-The schedule of reducing
-the updating magnitude
-can critically affect
-the precision of the final bias potential,
+The schedule of reducing the updating magnitude
+can critically affect the precision of the final bias potential,
 and hence that of the PMF\cite{laio2005, bussi2006, poulain2006,
@@
 %
-While a slow reduction of the updating magnitude
-leads to large variation of the bias potential,
-a rapid reduction can converge the bias potential
-to a wrong value.
+Overly slow and rapid reduction of the updating magnitude
+can hurt the precision and accuracy of the bias potential.
 %
@@
 and sensitive to the simulation length
-and the width of the Gaussian window.
-
-We will then compare
+and the width of the Gaussian.
+
+We will then show by comparing
 the efficiency of error reduction
-of different updating schemes.
-%
-We find that
+of different updating schemes
+that
 the single-bin updating scheme
@@
 %
-More generally,
-it can be generalized to
+The same asymptotic optimality also applies to
 a class of bandpass updating schemes
-that are equally optimal.
-%
-The bandpass updating schemes
-can avoid local noise
-by partially flattening a few long-range distribution modes.
-%
-The bandpass updating schemes
-are easier to use than the Gaussian updating scheme
-in that the optimal schedule of the former
-is simply given by the inverse-time formula,
-saving the effort of going through the general procedure described above.
+that flatten only
+a few long-range distribution modes.
+%
+The bandpass updating schemes allow
+simpler control of the updating magnitude
+than the Gaussian updating scheme
+for the asymptotically optimal schedule of the former
+is given by the inverse-time formula.
 
@@
 the bandpass updating schemes
-by constructing it from
+by construction from
 a sequence of orthogonal polynomials.
 %
-The bias potential is a linear combination
-of orthogonal polynomials
-and the coefficients are adjusted on-the-fly
-based on the observed statistical moments.
-%
-This application can be regarded
+The bias potential in this case assumes
+a polynomial form,
+and coefficients adjusted on-the-fly
+to achieve the desired target distribution.
+%
+This method serves
 as a generalization of the Langfeld-Lucini-Rago (LLR)
-algorithm\cite{langfeld2012, pellegrini2014}.
-%
-A similar construction with
+algorithm\cite{langfeld2012, pellegrini2014},
+and we show that
+a similar construction with
 a quadratic bias potential\cite{neuhaus2006, *neuhaus2007, zhu2012}
 allows one to sample an array of Gaussian distributions
-with desired means and variances.
+with predefined means and variances.
 
@@
 a deterministic averaged part, $\langle f_i(t) \rangle$,
-and a random fluctuating part, $\zeta_i(t)$,
+and a random fluctuating part, $\xi_i(t)$,
 %
@@
   +
-  \zeta_i(t)
+  \xi_i(t)
   .
@@
 %
-\begin{align}
-  f_{*i}(t)
-  &\approx
-  - v_{*i}(t)
-  +
-  \zeta_i(t)
+\begin{equation}
+  f_{*i}(t) \approx - v_{*i}(t) + \xi_{*i}(t)
   .
   \label{eq:sh_ave}
-\end{align}
-
-We will approximate the histogram fluctuation part, $\zeta_i(t)$,
-hence its deviation from the bin average, $\zeta_{*i}(t)$,
+\end{equation}
+
+We will approximate the histogram fluctuation part, $\xi_i(t)$,
+hence its deviation from the bin average, $\xi_{*i}(t)$,
 as white noise such that
 \begin{equation}
-  \langle \zeta_{*i}(t) \, \zeta_{*i}(t') \rangle
+  \langle \xi_{*i}(t) \, \xi_{*i}(t') \rangle
   = G_i \, \delta(t - t')
@@
   %=
-  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \zeta_{*i}(t) \right]
+  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \xi_{*i}(t) \right]
   \approx
-  -\alpha(t) \, \left[ v_{*i}(t) - \zeta_{*i}(t) \right]
+  -\alpha(t) \, \left[ v_{*i}(t) - \xi_{*i}(t) \right]
   .
@@
 %
-Without the random part, $\zeta_{*i}(t)$,
+Without the random part, $\xi_{*i}(t)$,
 this equation would drive
@@
   \int_0^T
-    \dot{\vartheta}\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
+    \dot{\vartheta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
   \label{eq:xt_solution}
@@
     \langle
-      \zeta_{*i}(t) \zeta_{*i}(t')
+      \xi_{*i}(t) \xi_{*i}(t')
     \rangle
@@
 %$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
-$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i \approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \zeta_{*i}(t) \zeta_{*i}(0) \rangle \, dt$.
+$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
+\approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt$.
 %\begin{equation}
@@
   -\alpha(t) \, \lambda_k \,
-  \bigl[ {\tilde v}_k(t) - {\tilde \zeta}_k(t) \bigr]
+  \bigl[ {\tilde v}_k(t) - {\tilde \xi}_k(t) \bigr]
   ,
@@
 Equation \eqref{eq:vt_diffeq_mbin} shows that
-in the absence of the fluctuating term, $\tilde \zeta_k(t)$,
+in the absence of the fluctuating term, $\tilde \xi_k(t)$,
 the deviation of the bias potential would tend to zero
@@
   \int_0^T
-    \dot{\vartheta}_k\bigl( q(t) \bigr) \, \tilde\zeta_k(t) \, dt
-  ,
-  \label{eq:xt_solution}
+    \dot{\vartheta}_k\bigl( q(t) \bigr) \, \tilde\xi_k(t) \, dt
+  ,
+  \notag
+  %\label{eq:xt_solution}
 \end{equation}
@@
   \langle
-    \tilde\zeta_k(t)
-    \tilde\zeta_k(t')
+    \tilde\xi_k(t)
+    \tilde\xi_k(t')
   \rangle
@@
 %of the histogram fluctuation,
-%$\bigl\langle {\tilde \zeta}_k(0)
-%\, {\tilde \zeta}_k(t) \bigr\rangle$,
+%$\bigl\langle {\tilde \xi}_k(0)
+%\, {\tilde \xi}_k(t) \bigr\rangle$,
 %as:
@@
   \bigl\langle
-    \tilde\zeta_k(t) \tilde\zeta_k(0)
+    \tilde\xi_k(t) \tilde\xi_k(0)
   \bigr\rangle
@@
 a first approximation for modeling
-a general fluctuation, ${\tilde \zeta}_k(t)$,
+a general fluctuation, ${\tilde \xi}_k(t)$,
 as an equivalent white noise.
@@
 Note that by Eq. \eqref{eq:parseval}, we have
-$\sum_{k=1}^{n-1} \bigl\langle {\tilde \zeta}_k(t) \, {\tilde \zeta}_k(0) \bigr\rangle
-= \sum_{i=1}^n \rho_i \bigl\langle \zeta_{*i}(t) \, \zeta_{*i}(0) \bigr\rangle,$
+$\sum_{k=1}^{n-1} \bigl\langle {\tilde \xi}_k(t) \, {\tilde \xi}_k(0) \bigr\rangle
+= \sum_{i=1}^n \rho_i \bigl\langle \xi_{*i}(t) \, \xi_{*i}(0) \bigr\rangle,$
 and after summing $t$ from $-\infty$ to $\infty$,
@@
 \item
-Compute the eigenvalues, $\lambda_k$'s
+Compute the eigenvalues, $\lambda_k$'s,
 from the updating matrix, $\mathbf w$.
@@
 \item \label{step:qT}
-Compute the optimal value of $q(T)$ by solving Eq. \eqref{eq:opt_qT},
-using, e.g. the Newton-Raphson method\cite{press3rd}\footnote{In
-  our implementation, the equation was solved for $\ln q(T)$
+Compute the optimal value of $q(T)$
+by solving Eq. \eqref{eq:opt_qT}.\footnote{We
+  used a variant of the Newton-Raphson method\cite{press3rd}
+  with the bisection method as a fallback.
+  We solved the equation for $\ln q(T)$
   to avoid negative values of $q(T)$,
-  and we imposed a lower bound,
+  and imposed a lower bound,
   $\bigl|\mathcal Y\bigl(q(T)\bigr)/q(T)\bigr|$, for
@@
   in computing its change.
-  The initial value was set to $q(T) = \ln(1+T\,a_0/2)$.
-  We also used the bisection method as a fallback.}
+  The initial value was set to $q(T) = \ln(1+T\,a_0/2)$.}
 %
@@
 
-We can show that the above optimal solution for asymptotic convergence
+We can show that the above optimal solution
 represents the single-bin updating scheme.
@@
 %
-However, by constructing a mixture or expanded ensemble,
+However, by constructing a mixture\cite{shirts2017}
+or expanded ensemble\cite{swendsen1986,
+  *geyer1991, *hukushima1996, *hansmann1997, *sugita1999,
+  *earl2005, *zuckerman2011, *rauscher2009,
+  neuhaus2006, *neuhaus2007, kim2010,
+  marinari1992,
+  *lyubartsev1992, li2007,
+  park2007, *nguyen2013, *zhang2015st},
 we can simultaneously sample multiple target distributions
@@
 %
-This results in a quadratic bias
-potential\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007},
-as commonly employed in umbrella sampling\cite{torrie1974, *torrie1977}
+The bias potential is
+quadratic\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007},
+which is common in umbrella sampling\cite{torrie1974, *torrie1977}
 and the method of Gaussian ensemble\cite{hetherington1987,
@@
 We can extend the sampling range
-by constructing a mixture\cite{shirts2017}
-or an expanded ensemble
+by constructing an expanded ensemble
 from a superposition of multiple Gaussian umbrellas
@@
 \end{align}
-where $z(t)$ is the current value of $z$,
-and $Z^{(i)} \equiv \int e^{-\beta \, \Phi(\mathbf x)-u^{(i)}(z(\mathbf x),t)} \, d\mathbf x$
-is the partition function.
-%
-Here,
+where $z(t)$ is the current value of $z$.
+%
+The parameter, $Z^{(i)}$, is the inverse weight of the $i$th umbrella,
+in the expanded ensemble,
+and $\ln Z^{(i)}$ serves as the bias potential
+for inter-umbrella transitions.
+%
+We can be estimate $\ln Z^{(i)}$'s on-the-fly
+using the single-bin updating scheme,
+Eq. \eqref{eq:wl_update}\cite{li2007}.
+%
+Upon convergence, $Z^{(i)}$ should be proportional to
+the partition function,
+$\int e^{-\beta \, \Phi(\mathbf x)-u^{(i)}(z(\mathbf x),t)} \, d\mathbf x$,
+as in simulated tempering,
+where
 $\beta = 1/(k_B T)$ is the inverse temperature,
@@
 %
-Note that the unknown parameters, $\ln Z^{(i)}$'s,
-serve as the bias potential
-for inter-umbrella transitions, and thus can be updated
-using the single-bin scheme,
-Eq. \eqref{eq:wl_update}\cite{li2007}.
 
@@
 
-We can characterize the optimal schedule
-geometrically as follows.
-%
-First, we define
-a normalized mass distribution as
+Here we give a geometric characterization
+of the optimal schedule.
+%
+First, we define a normalized mass distribution as
 %
@@
   \frac t T
-  ,
+  .
   \label{eq:intmQ_tT}
@@
 %
-Thus, each point on the schedule,
+This allows each point on the schedule,
 $\bigl(t, \alpha(t)\bigr)$,
-can be mapped to a point,
+to be mapped to a point,
 $\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
@@
 Since the shift,
-hence its Fourier transform,
+hence its generalized Fourier transform,
 $\tilde u_k(t) - \tilde v_k(t)$,
@@
 ${\tilde u}_{k}$ and ${\tilde v}_{k}$
-should share the same variance.
-%
-But as the ensemble average of ${\tilde v}_{k}$
-is zero, we have
+should share the same variance,
 $\operatorname{var} {\tilde u}_{k} =
@@
  \left\langle {\tilde v}_{k}^2 \right\rangle$,
+where we have used the fact that
+the ensemble average of ${\tilde v}_{k}$
+is zero in the last step.
 %\begin{equation*}
@@
 %
-and by using Eq. \eqref{eq:xt2_eql},
-we get
+From Eq. \eqref{eq:xt2_eql}, we get
 %
@@
   \frac{1}{2} \,
-  a_0 \, \Gamma_k \, \lambda_k,
+  a_0 \, \Gamma_k \, \lambda_k
+  .
 \label{eq:varXt}
@@
 %
-where the variance can be computed from trajectory averages.
+The variance, $\tilde u_k$, can be computed
+by trajectory averaging.
 %
@@
 
-The above transition matrix method
+The above method
 tends to underestimate
@@
 the updating matrix, $w_{ij}$,
-must be a function of $(i-j)\mod n$,
+can only be a function of the difference of indices, $i-j$, modulo $n$,
 and we write it as
@@
 We recall that
-for continuous variable, $z$, over $[0, \Delta z]$
-the bin index, $i$, is related to $z$ as
+the bin index, $i$, can be related to
+the continuous variable $z$ as
 $z = i \, \delta z$,
 where
-$\delta z = \Delta z/(g \, n)$ is th bin size.
-
+$\delta z$ is the bin size.
+%
+We further define
+$\mu(z) \equiv \mu_i/\delta z$
 and
-$\mu_i = \mu(z) \, \delta z$.
+$\Delta z = b \delta z$
+for the largest bin cutoff $b = gn/2$.
 %
 For $n \gg 1$,
-By assuming the largest bin cutoff, $b \approx g \, n/2$,
 we can approximate Eq. \eqref{eq:wband_eigenvalue}
@@
   =
-  \int_{-\Delta z/2}^{\Delta z/2}
-    \mu(z) \, \cos\left( \frac{ 2 \, \pi \, k \, z } { \Delta z} \right)
+  \int_{-\Delta z}^{\Delta z}
+    \mu(z) \, \cos\left( \frac{ \pi \, k \, z } { \Delta z} \right)
     \, d z,
@@
 with the normalization
-$1 = \int_{-\Delta z/2}^{\Delta z/2} \mu(z) \, dz$.
+$1 = \int_{-\Delta z}^{\Delta z} \mu(z) \, dz$.
 
 If the width of the Gaussian window, $\sigma_z$,
-is very small compared to $\Delta z$,
+is much less than $\Delta z$,
 then we can extend the limits of the integrals
-to infinity, and
+to infinity, and normalize $\mu(z)$ as
 %
@@
 %
-In practice, however, some eigenvalues can turn negative
-when discretized on a finite grid.
+However,
+when discretized on a finite grid,
+some eigenvalues can turn negative.
 %
@@
 
-Let us consider an FDS simulation under a sufficiently accurate bias potential.
-The fluctuation of the instantaneous histogram
-is then dominated by the random part,
-%
-\begin{equation}
-  f_i(t) \approx \zeta_i(t)
-  ,
-  \label{eq:xi_zeta}
-\end{equation}
-and the same holds for the Fourier transform,
-$\tilde f_k(t) \approx \tilde \zeta_k(t).$
+Consider an ideal FDS simulation under the exact bias potential.
+%
+By Eq. \eqref{eq:sh_ave},
+the fluctuation of the instantaneous histogram
+is then given by the random part,
+%
+\begin{equation}
+  f_{*i}(t) = \xi_{*i}(t)
+  ,
+  \label{eq:f_xi}
+\end{equation}
+and
+%the same holds for the Fourier transform,
+$\tilde f_k(t) \approx \tilde \xi_k(t).$
 %
@@
 %
-From Eq. \eqref{eq:xi_zeta}, we have
+From Eq. \eqref{eq:f_xi}, we have
 %$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
@@
 %\bigl\langle
-%  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
+%  \tilde \xi_k(t) \, \tilde \xi_k(0)
 %\bigr\rangle } { T }
@@
 \bigl\langle
-  \tilde \zeta_k(t) \, \tilde \zeta_k(0)
+  \tilde \xi_k(t) \, \tilde \xi_k(0)
 \bigr\rangle
@@
   \Err(T)
-  \ge
-  \bigl\langle F^2(T+t_0) \bigr\rangle
+  \approx
+  \bigl\langle F^2(T) \bigr\rangle
   .
@@
 as in Sec. \ref{sec:bandpass_poly},
-we have
-\begin{align*}
-\tilde f_k(t)
-=
-\int \phi_k(z) \, \frac{ \delta\bigl(z(t) - z\bigr) } { \rho(z) } \, dz
-=
-R_k\bigl( z(t) \bigr)
-,
-\end{align*}
-so that
+then
+$\tilde f_k(t) = R_k\bigl( z(t) \bigr)$,
+%\begin{align*}
+%\tilde f_k(t)
+%=
+%\int \phi_k(z) \, \frac{ \delta\bigl(z(t) - z\bigr) } { \rho(z) } \, dz
+%=
+%R_k\bigl( z(t) \bigr)
+%,
+%\end{align*}
+and
 \begin{equation}
@@
     $s, S$          &   Stage index and the total number of stages in the Wang-Landau algorithm. \\
-    $\sigma_{ij}(t)$   &   Autocorrelation function of $\zeta_i(t)$. \\
+    $\sigma_{ij}(t)$   &   Autocorrelation function of $\xi_i(t)$. \\
     $t, \tau$       &   Time. \\
@@
     $v_{*i}(t)$     &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
-    $\zeta_{*i}(t)$ &   The noise of histogram. \\
+    $\xi_{*i}(t)$ &   The noise of histogram. \\
     ${\tilde v}_k$  &   Mode of the shifted bias potential. \\
@@
     $z$             &   Reaction coordinate. \\
-    $\zeta_i(t)$    &   The noise part of the instantaneous histogram, $h_i(t)$.
+    $\xi_i(t)$    &   The noise part of the instantaneous histogram, $h_i(t)$.
   \end{tabular}

### 2017-08-19 doc/invt.tex
invt.tex: editing appendix A

@@
   .
-  \label{eq:w_detailedbalance}
+  \notag
+  %\label{eq:w_detailedbalance}
 \end{equation}
@@
 %
-We then compute the updating magnitude by numerical differentiation:
-$\alpha_i = (1 - e^{-\lambda_1 \, \hat \alpha_i})/\lambda_1$\footnote{This
-  discretization follows from the runtime average interpretation
-  discussed in Appendix \ref{sec:equilerr}.}
-and
+We then compute the updating magnitude
+from (cf. Appendix \ref{sec:ave_us}):
+\begin{equation}
+  \alpha_i = \frac{ 1 - e^{-\lambda_1 \, \hat \alpha_i} } { \lambda_1 }
+  ,
+  \label{eq:alphai}
+\end{equation}
+where
 $\hat \alpha_i \equiv (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
@@
 homogeneous updating schemes,
-we will consider here bandpass updating schemes
+we will consider here a family of bandpass updating schemes
 constructed from a sequence of orthogonal polynomials.
@@
 we can simultaneously sample multiple target distributions
-spread over a wider range, as exemplified below.
+to cover a wider range, as exemplified below.
 
@@
 %
-In this way, an attempt to move out of the current interval
-does not have to be rejected,
-but can be interpreted as a transition to
-another internal and accepted with a proper probability.
+In this way, attempts to move out of the current interval
+are not necessarily rejected,
+but can be interpreted as transitions to
+another internal and accepted with proper probability.
 
@@
 \item
-  Change the updating magnitude,
+  Change the updating magnitude, $\alpha(t)$,
   which will be used for both intra-umbrella
@@
   and inter-umbrella parameters, $\ln Z^{(i)}$.
-  %
 
@@
 zhu2012, newman, frenkel},
-can be used to recover the unbiased distribution.
-%
+can be used to recover the unbiased distribution,
+%
+\begin{equation}
+  p(z) \, \delta z
+  =
+  \frac{
+    \sum_i n^{(i)}(z)
+  }
+  {
+    \sum_i N^{(i)} \, e^{-u^{(j)}(z)} / Z^{(i)}
+  }
+  ,
+  \notag
+\end{equation}
+%
+where $n^{(i)}(z)$
+the number of visits to the bin $(z-\delta/2, z+\delta/2)$
+in umbrella $i$,
+and the $N^{(i)}$ is the total number of visits
+to the umbrella $i$.
+%
+We can use the final converged value of $Z^{(i)}$
+without going through WHAM iteration.
 
@@
 \section{\label{sec:equilerr}
-Runtime average interpretation
+Runtime-average interpretation
 of the inverse-time schedule
@@
 
-The optimality of the inverse-time schedule
+We can understand
+the optimality of the inverse-time schedule
 for the single-bin updating scheme
-can be understood from the observation that
-the schedule allows the bias potential
-to coincide with a runtime average
-of independent corrections.
-
-Assuming that we can linearize the histogram correction
-in Eq. \eqref{eq:vcorr_equil} as
-$\ln x \approx x - 1$ % (for $x \approx 1$),
-and apply it to the instantaneous histogram, $h_i(t)$,
-the independent estimate of bias potential in each step
-would be given by
-%
-\begin{equation}
-  \hat u_i(t) \approx u_i(t) + \frac{ h_i(t) } { \rho_i } - 1
+by a runtime average interpretation.
+%
+We wish to show that the recurrence updating relation,
+Eq. \eqref{eq:wl_update},
+renders the bias potential
+as a runtime average with a dynamic weight
+determined by the updating schedule.
+%
+For the single-bin updating scheme,
+the inverse-time schedule is optimal
+because it corresponds to an unweighted average.
+
+\subsection{Runtime average and recurrence relation}
+
+Consider the weighted average
+%
+\begin{equation}
+  X_\omega(t)
+  =
+  \frac{
+    \omega(0) \, x(0) + \cdots + \omega(t) \, x(t)
+  }
+  {
+    \Omega(t)
+  }
+  ,
+  \label{eq:X_ave}
+\end{equation}
+%
+where
+$\Omega(t) \equiv \omega(0) + \cdots + \omega(t)$
+is the cumulative weight.
+%
+We can write the average as a recurrence relation
+%
+\begin{equation}
+  X_\omega(t)
+  = X_\omega(t-1)
+  +
+  \alpha(t)
+  \, [ x(t) - X_\omega(t-1) ]
+  ,
+  \label{eq:X_recur}
+\end{equation}
+%
+where the updating magnitude is defined as
+%
+%The relative weight, $\omega(t)/\Omega(t)$,
+%can be identified as the updating magnitude,
+\begin{equation}
+  \alpha(t) \equiv
+  \frac{ \omega(t) } {\Omega(t)}
+  =
+  1 - \frac{ \Omega(t-1) } {\Omega(t)}
+  .
+  \label{eq:alpha_from_Omega}
+\end{equation}
+%
+and $x(t) - X_\omega(t-1)$
+is the proposed change from the previous average
+from the data point at step $t$.
+
+Conversely, for a recurrence relation
+in the form of Eq. \eqref{eq:X_recur}
+with a given $\alpha(t)$, we can revert it
+to an average in the form of Eq. \eqref{eq:X_ave}
+with the weight given by
+\begin{align}
+  \frac{ \omega(t) } { \omega(0) }
+  =
+  \left(
+    \prod_{\tau=1}^{t-1} \frac{1}{1- \alpha(\tau)}
+  \right)
+  \frac{ \alpha(t) } { 1 - \alpha(t) }
+  .
+  \label{eq:omega_from_alpha}
+\end{align}
+
+%The inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
+For example,
+the unweighted average, $\omega(t) = \omega(0)/t_0$ (for $t > 0$),
+corresponds to the inverse-time updating magnitude,
+Eq. \eqref{eq:alpha_invt1}.
+%
+Updating magnitudes above and below this value
+correspond to runtime averages
+with growing and diminishing weights, respectively.
+
+\subsection{\label{sec:ave_us}
+Application to updating schemes}
+
+To apply the above formalism to updating schemes,
+we need a correction formula for the bias potential.
+%
+We will use a linearized version of Eq. \eqref{eq:vcorr_equil},
+$\hat u_i = u_i + \frac{ H_i } { \rho_i } - 1$.
+%
+%\begin{equation}
+%  \hat u_i = u_i + \frac{ H_i } { \rho_i } -1
+%  .
+%  \label{eq:vcorr_equil1}
+%\end{equation}
+%
+Using this formula to
+the instantaneous histogram, $h_i(t)$,
+we get an independent estimate
+of the bias potential at a single step $t$,
+%
+\begin{equation}
+  \hat u_i(t)
+  \approx u_i(t) + \frac{ h_i(t) } { \rho_i } - 1
   = u_i(t) + f_i(t) - 1
@@
 \end{equation}
-%
-Below we will use capital letters to denote
-time averages, e.g. for variable $x$, we define
-$X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)$.
+
+For the single-bin updating scheme, Eq. \eqref{eq:wl_update},
+we have
+\begin{align}
+  u_i(t+1)
+  &=
+  u_i(t) + \alpha(t) \, f_i(t)
+  \notag \\
+  &\approx
+  u_i(t) + \alpha(t) \, [\hat u_i(t) - u_i(t)] + \mathrm{const.}
+  ,
+  \label{eq:u_uhat}
+\end{align}
+where we have used Eq. \eqref{eq:uhatt}
+and ignored a constant shift $\alpha(t)$
+in the second step.
+%
+Comparing this to Eq. \eqref{eq:X_recur},
+we find that $u_i(t+1)$
+is equivalent to a runtime average $X_\omega(t)$
+of $\hat u_i(t)$
+with the weight given by Eq. \eqref{eq:omega_from_alpha}.
+%
+The inverse-time schedule is optimal
+because it corresponds to the unweighted average
+that uses the data most efficiently.
+
+Similarly, for the multiple-bin updating scheme,
+we have, from Eq. \eqref{eq:vkupdate},
+${\tilde u}_k(t + 1) = {\tilde u}_k(t) + \alpha(t) \, \lambda_k \, {\tilde f}_k(t)$,
+and from Eq. \eqref{eq:u_uhat},
+$\hat{\tilde u}_k(t) \approx {\tilde u}_k(t) + {\tilde f}_k(t)$,
+for $k > 0$.
+This means that
+$\tilde u_k(t+1)$ can also be cast
+as a runtime average of $\hat{\tilde u}_k(t)$,
+with a weight corresponding to the effective updating magnitude,
+$\lambda_k \, \alpha(t)$.
+
+The runtime average interpretation imposes
+an additional constraint on the updating magnitude.
+%
+By Eq. \eqref{eq:alpha_from_Omega},
+the effective updating magnitude, $\lambda_k \, \alpha(t)$,
+of any mode $k > 0$
+should never exceed $1.0$,
+which is not necessarily satisfied
+for the updating schedule
+derived from the continuous-time formalism.
+%
+An expedient fix,
+which we used in Eq. \eqref{eq:alphai},
+is to derive the effective updating magnitude
+from Eq. \eqref{eq:alpha_from_Omega}
+with
+\begin{align*}
+  \ln \frac{ \Omega(t-1) } { \Omega(t) }
+  &\approx
+  \frac{1}{t-t'} \ln \frac{ \Omega(t') } { \Omega(t) }
+  =
+  \sum_{\tau=t'+1}^t \frac{ \ln[1- \lambda_k \, \alpha(\tau)] } { t - t' }
+  \\
+  &\approx
+  -\lambda_k \cdot \frac{\int_{t'}^t \alpha(\tau) \, d\tau}{t-t'}
+  =
+  -\lambda_k \cdot \frac{q(t) - q(t')}{t- t'}
+  ,
+\end{align*}
+where $t$ and $t'$ are two consecutive grid points.
+%
+Using this formula to mode $k = 1$
+with the largest eigenvalue
+would guarantee $\lambda_k \, \alpha(t) \ge 1$
+for all nontrivial modes.
+
+\subsection{Histogram correction}
+
+We have shown above that
+for a general multiple-bin updating scheme,
+each mode of bias potential
+corresponds to a runtime average
+with a dynamic weight,
+which is not necessarily even to be most efficient.
+%
+Below we give a simple correction that
+could largely compensate the efficiency loss.
+%
+We will use capital letters to denote unweighted averages.
+%e.g. for variable $y$,
+%$Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t y(\tau)$.
 %\begin{equation}
-%  X(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
+%  Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
 %  .
@@
 %
-Time averaging Eq. \eqref{eq:uhatt} yields
+Then, by an unweighted average of Eq. \eqref{eq:uhatt}, we get
 \begin{align}
@@
 %
-%Equation \eqref{eq:uhatav}
-This equation serves as
+This correction serves as
 a generalization of Eq. \eqref{eq:vcorr_equil}
-for FDS simulations under a variable bias potential
-and we can use it to improve the estimated PMF
-after simulation.\footnote{Equation \eqref{eq:uhatav}
-  can also be derived
+for FDS simulations under
+a variable bias potential.\footnote{We can also
+  derive Eq. \eqref{eq:uhatav}
   from the self-healing umbrella sampling method\cite{marsili2006}.
-  From Eq. \eqref{eq:uhatt},
-  we can construct an independent estimator
-  of the unbiased distribution as
-  $\hat p_i(t) \approx h_i(t) \, e^{ u_i(t) }$.
-  Upon time averaging, we get
+  We can construct an independent estimator
+  of the unbiased distribution from Eq. \eqref{eq:Vi_target} as
+  $\hat p_i(t) \approx h_i(t) \, e^{ u_i(t) }$,
+  assuming a proper shift of $u_i(t)$.
+  Then after time averaging, we get
   $\rho_i e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
-%
-In particular, for the single-bin updating scheme,
-Eq. \eqref{eq:wl_update},
-we have, up to a constant shift,
-%
-\begin{align}
-  \hat U_i(t)
-  &\approx
-  \frac{1}{t} \sum_{\tau=1}^t
-  \left[
-    u_i(\tau) +
-    \frac{ u_i(\tau + 1) - u_i(\tau) } { \alpha(t) }
-  \right]
-  ,
-  \notag
-  \\
-  &\approx
-  \Delta u_i
-  +
-  \sum_{\tau=1}^t
-  \left[
-    1 + \frac{1}{\alpha(\tau - 1)} - \frac{1}{\alpha(\tau)}
-  \right]
-  \frac{ u_i(\tau) } {t}
-  ,
-  \label{eq:Uhat_sum}
-\end{align}
-where
-$\Delta u_i \equiv u_i(t+1) /[t \, \alpha(t)]
-- u_i(1)/[t \, \alpha(0)]$.
-%$$
-%\Delta u_i \equiv \frac{u_i(t+1)}{t \, \alpha(t)}
-%- \frac{u_i(1)}{t \, \alpha(0)}.
-%$$
-%
-With the inverse-time schedule, Eq. \eqref{eq:alpha_invt},
-each term in the sum vanishes,
-and if we further define $\alpha(0) \to \infty$,
-then
-$$
-u_i(t+1) = \hat U_i(t),
-$$
-i.e. the bias potential in this case coincides with
-the runtime average of the correction, $\hat u_i(t)$,
-at every step, $t$.
-%
-On the other hand,
-with a constant $\alpha(\tau) = a_0$ for $\tau = 0, 1, \dots$,
-Eq. \eqref{eq:Uhat_sum} is approximately a plain average\cite{zhou2005}:
-$\hat U_i(t) \approx U_i(t)$,
-if we can assume an equilibrium condition
-with $u_i(t+1) \approx u_i(1)$.
-\note{This would be rather risky for metadynamics,
-because we know that the certain noise modes
-do not easily go away, although the plain average
-formula was also recommended for metadynamics.}
-
-We can also cast other updating relations,
-such as Eqs. \eqref{eq:wl_update}
-and \eqref{eq:vkupdate},
-as runtime averages,
-but with dynamic weights.
-%
-The weighted average
-%
-\begin{align*}
-  X_\omega(t)
-  =
-  \frac{
-    \omega(0) \, x(0) + \cdots + \omega(t) \, x(t)
-  }
-  {
-    \Omega(t)
-  }
-  ,
-\end{align*}
-%
-where
-$\Omega(t) \equiv \omega(0) + \cdots + \omega(t)$
-is the accumulated weight,
-is equivalent to the recurrence relation
-%
-\begin{align*}
-X_\omega(t) = X_\omega(t-1)
-  + \frac{\omega(t)}{\Omega(t)}
-  \Delta \, x(t)
-,
-\end{align*}
-where
-$\Delta x(t) \equiv x(t) - X_\omega(t-1)$
-is the proposed change from the previous average
-from the data point at step $t$.
-%
-The relative weight, $\omega(t)/\Omega(t)$,
-can be identified as the updating magnitude,
-\begin{equation}
-  \alpha^*(t) \equiv
-  \frac{ \omega(t) } {\Omega(t)}
-  =
-  1 - \frac{ \Omega(t-1) } {\Omega(t)}
-  .
-  \label{eq:alpha_from_Omega}
-\end{equation}
-%
-The inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
-follows from the case of unweighted average,
-$\omega(t) = 1$ (for $t > 0$),
-and updating magnitudes above and below this value
-correspond to runtime averages that favor future and past data,
-respectively.
-%
-In the case of Eq. \eqref{eq:vkupdate},
-we can identify
-the bias potential $\tilde v_k(t)$
-as $X(t-1)$,
-$\tilde f_k(t)$ as $\Delta x(t)$
-by Eq. \eqref{eq:uhatt},
-and
-the mode-$k$ updating magnitude,
-$\alpha(t) \, \lambda_k$,
-as $\alpha^*(t)$.
-%
-This mapping shows that
-updating magnitude should not exceed $1.0$.
-%
-One way to satisfy this condition
-in converting the results
-from the continuous-time setup is
-to use Eq. \eqref{eq:alpha_from_Omega}
-with
-\begin{align*}
-  \frac{ \Omega(t) } { \omega(0) }
-  = \prod_{\tau=1}^t \frac{1}{1- \alpha^*(\tau)}
-\approx \exp\left(\int_0^t \alpha^*(\tau) \, d\tau \right).
-\end{align*}
 
@@
 
-Here, we give some mathematical details
-on the homogeneous updating schemes,
-%
-The updating matrix of the updating schemes, $w_{ij}$,
-can be characterized by
-a fixed window function, $\mu_0, \mu_1, \dots, \mu_b$,
-that shifts with the bin $j$.
-%
-Here $\mu_l$ denotes the relative updating magnitude
-(not adjusted by the boundary correction)
-on the $l$th bin from the center.
-%
-The translational invariance requires $\rho_i$ to be flat,
-and the updating matrix, $\mathbf w$,
-must be symmetric to satisfy detailed balance,
-Eq. \eqref{eq:w_detailedbalance}.
-%
-The updating matrices share the same set of eigenvectors
-(which here are sines and cosines of different frequencies),
-and thus are completely determined
-by the eigenvalues.
-%
-Further, the multiplication of $\mathbf w$
-can be reduced to a convolution with the window function.
-%
-We will show below that the eigenvalues of the updating matrix
-is related to the window function as a cosine transform.
-%
-Thus, there is a one-to-one correspondence between
-the eigenvalues of $\mathbf w$
-and the window function.
+Here, we discuss some details
+on the homogeneous updating schemes.
+%
+We can characterize
+the updating matrix, $w_{ij}$,
+of a homogeneous updating scheme
+by a rigid window function
+that shifts with the bin $j$ of visit
+(ignoring the boundary adjustment).
+%
+The window function is defined by
+the relative updating magnitude, $\mu_l$,
+on the $l$th bin from the central bin.
+%
+%The translational invariance requires $\rho_i$ to be flat,
+%and the updating matrix, $\mathbf w$,
+%must be symmetric to satisfy detailed balance,
+%Eq. \eqref{eq:w_detailedbalance}.
+%
+By the translational invariance,
+the updating matrices of homogeneous updating schemes
+share the same set of eigenvectors,
+which are sine and cosine functions.
+%
+The eigenvalues of the updating matrix
+are related to the window function as a cosine transform.
 
@@
 Eigenvalues and window function}
-
-%While translationally-invariant
-%updating schemes
-%naturally suit periodic variables\cite{
-%dama2014},
-%they can also be extended to non-periodic variables
-%by imposing the reflective boundary condition\cite{
-%bussi2006}.
-%%
-%For simplicity, we will assume that the target
-%distribution is flat, or $\rho_i = 1/n$, below.
 
@@
 %
-We can maintain the convergence of the updating scheme
-by redefining the Gaussian scheme
+To maintain the positivity of eigenvalues,
+we can redefine the Gaussian scheme
 from the same spectrum of eigenvalues, Eq. \eqref{eq:lambda_Gaussian},
@@
 %the same holds for the Fourier transform,
-$\tilde f_k(t) \approx \tilde \xi_k(t).$
+$\tilde f_k(t) = \tilde \xi_k(t).$
 %

### 2017-08-22 doc/invt.tex
invt.tex: small edits

@@
 \usepackage[table,usenames,dvipsnames]{xcolor}
-\usepackage{tikz}
-\usetikzlibrary{calc}
+%\usepackage{tikz}
+%\usetikzlibrary{calc}
 \usepackage{hyperref}
@@
   *laio2002, *laio2008, *barducci2011, *sutto2012, micheletti2004}
-among others\cite{kim2006, *kim2007, kim2010, junghans2014,
+among others\cite{yan2004, kim2006, *kim2007, kim2010, junghans2014,
   langfeld2012, pellegrini2014,
@@
 %
-Overly slow and rapid reduction of the updating magnitude
-can hurt the precision and accuracy of the bias potential.
+While an overly slow reduction of the updating magnitude
+increases the variance of the bias potential,
+an too rapid reduction often
+converges the bias potential to wrong value.
 %
@@
 to build a class of adaptive FDS algorithms
-with quadratic bias potentials.
+with polynomial bias potentials
+in Sec. \ref{sec:bandpass_poly}.
 
@@
 except that some elements can be negative
-in our cases.
+for our cases here.
 %
@@
 We then compute the updating magnitude
-from (cf. Appendix \ref{sec:ave_us}):
-\begin{equation}
-  \alpha_i = \frac{ 1 - e^{-\lambda_1 \, \hat \alpha_i} } { \lambda_1 }
+as (cf. Appendix \ref{sec:ave_us}):
+\begin{equation}
+  \alpha_{[i]}
+  =
+  \frac{1}{\lambda}
+  \left[
+    1
+    -
+    \exp\left(
+    -\lambda_1
+    \frac{ q_{[i+1]} - q_{[i-1]} }
+         { t_{[i+1]} - t_{[i-1]} }
+    \right)
+  \right]
   ,
@@
 \end{equation}
-where
-$\hat \alpha_i \equiv (q_{[i+1]} - q_{[i-1]}) / (t_{[i+1]} - t_{[i-1]})$,
 with
@@
   &
-\qquad \qquad
+  \qquad \qquad \qquad
   \ge
@@
   &
-\qquad \qquad
+  \qquad \qquad \qquad
   =
@@
 that discard certain noise modes
-of short wavelengths.
+of short wavelength.
 %
@@
 eigenmodes,
-we may assign the first $K$ eigenmodes with eigenvalue $1$
-and the rest of eigenmodes with eigenvalue $0$:
+we may assign the first $K$ eigenmodes eigenvalue $1$
+and the rest of eigenmodes eigenvalue $0$:
 %
@@
 
-Applying this construction to the basis of cosine and sine functions,
+We will consider two special cases of the bandpass updating schemes.
+%
+First,
+by applying this construction to the basis of cosine and sine functions,
 we get a homogeneous bandpass updating scheme
 that resembles the Gaussian updating scheme,
-but with better convergence in the asymptotic limit.
+but with better convergence in the asymptotic limit
+(cf. Appendix \ref{sec:homo_bandpass}).
 %
@@
 %  $|\max_i u_i - \min_i u_i|^2/K^3$.}
+%
+The second case is based on a family of orthogonal polynomials,
+which we will discuss below.
 
@@
 
-In addition to the above application to
-homogeneous updating schemes,
-we will consider here a family of bandpass updating schemes
+Here we will consider here a family of bandpass updating schemes
 constructed from a sequence of orthogonal polynomials.
@@
 \begin{align*}
-  \int R_k(z) \, R_{k'}(z) \, \rho(z) \, d z = \delta_{k, k'}
-  .
+  \int R_k(z) \, R_{k'}(z) \, \rho(z) \, d z
+  = \overline{ R_k(z) \, R_{k'}(z) }
+  = \delta_{k, k'}
+  ,
 \end{align*}
+%
+where the overline denotes the average for the distribution, $\rho(z)$.
+%
+The first few members are
+\begin{align*}
+  R_0(z) = 1,
+  R_1(z) = \frac{z - \overline{z}}{\sigma_z},
+  R_2(z) = \frac{ (z - \overline{z})^2 - \sigma_z^2 }
+  {\sqrt{\overline{\left[
+    (z - \overline z)^2 - \sigma_z^2
+  \right]^2 }}}
+  ,
+\end{align*}
+%
+where $\sigma_z \equiv \sqrt{\overline{\left(z - \overline z\right)^2}}$.
 %
@@
 Since high-order moments are generally less reliable,
-this method is best used with a small $K$ for a narrow distribution.
-%
-However, by constructing a mixture\cite{shirts2017}
-or expanded ensemble\cite{swendsen1986,
+this method is better used with a small $K$ for a narrow distribution.
+%
+However, by constructing
+an expanded ensemble or mixture\cite{swendsen1986,
   *geyer1991, *hukushima1996, *hansmann1997, *sugita1999,
@@
   *lyubartsev1992, li2007,
-  park2007, *nguyen2013, *zhang2015st},
+  park2007, *nguyen2013, *zhang2015st, shirts2017},
 we can simultaneously sample multiple target distributions
@@
 to sample a flat distribution, $\rho(z) = 1/\Delta z$,
-over $[z_c - \Delta z/2, z_c + \Delta z/2]$,
+over $\left[z_c - \frac 1 2 \Delta z, z_c + \frac 1 2 \Delta z\right]$,
 we can use the Legendre polynomials\cite{arfken}.
@@
 %
-The bias potential is a linear function of $z$,
-and this updating scheme is essentially the LLR algorithm\cite{langfeld2012}
+%The bias potential is a linear function of $z$,
+The updating scheme is essentially the LLR algorithm\cite{langfeld2012}
 for the single-interval case,
@@
 As discussed above,
-the method is effective only to
+the method is effective only for
 a local region of a multimodal distribution,
@@
   *belardinelli2007jcp, *belardinelli2008, *belardinelli2016}
-to the optimally control updating magnitude.
-%
-We can adopt the WL stage-by-stage strategy in early stages,
+to optimally control the updating magnitude.
+%
+We can adopt the WL stage-by-stage strategy early in simulation,
 and switch to the inverse-time schedule later
@@
   from the WL to the inverse-time schedule.
-  Such a switch was only permitted
-  upon completion of a WL stage,
-  and the time, $t$, was interpreted as the number of updating steps
+  We allowed such a switch only
+  upon the completion of a WL stage,
+  and interpreted $t$ as the number of updating steps
   in this just completed WL stage.
   If the switch was successful,
-  the above $t$ was used as the offset $t_0$
+  the above $t$ was used as the $t_0$
   in the inverse-time regime.}
@@
   p(z) \, \delta z
-  =
+  \propto
   \frac{
@@
 where $n^{(i)}(z)$
-the number of visits to the bin $(z-\delta/2, z+\delta/2)$
+the number of visits to the bin
+$\left(z-\frac{1}{2} \delta z, z + \frac{1}{2} \delta z\right)$
 in umbrella $i$,
@@
 %
-We can use the final converged value of $Z^{(i)}$
-without going through WHAM iteration.
+We can use the final converged values of $Z^{(i)}$'s
+without going through the usual WHAM iteration.
 
@@
 as $y(r) = \left(1 - \frac{1}{N}\right) \, V \, p(r)$\cite{hansen}.
-%
-In biased sampling, we fixed the two special particles on the $x$-axis,
-and varied their distance $r$ along the axis.
-%
-Further, we added a bias potential, $k_B T \, u(r)$,
-to sample a flat distribution along $r$.
-%
-The reference bias potential was pre-calculated from a long simulation.
 
@@
 In each trial,
-a random particle is displaced in each permissible dimension
+a random particle is displaced in each spatial dimension
 according to a uniform distribution over $(-\varepsilon, \varepsilon)$.
+%
+However, the two special particles are fixed on the $x$-axis,
+and are only displaced in this direction.
+%
 We used $\varepsilon = 0.1/\varrho$
@@
 
-The PMF was calculated from $r = 0$ to half of the box size
+We used adaptive FDS to sample a flat distribution along $r$,
+and calculated the PMF from $r = 0$ to half of the box size
 (also the cutoff distance of the potential)
@@
 The reduced temperature was $T = 3$.
+%
+The bias potential, $k_B T \, u(r)$,
+was initially set to zero and
+underwent the test updating scheme and schedule.
 %
@@
 Eqs. \eqref{eq:v_def} and \eqref{eq:x_def},
-and the error from Eq. \eqref{eq:error_def}.
+and the error from Eq. \eqref{eq:error_def},
+with the reference bias potential
+pre-calculated from a longer simulation.
 %
@@
 To make sure the same schedule for multiple runs under the same condition,
-the optimal schedule was pre-computed from a set of fixed
-$\Gamma_k$'s and $\epsilon_k$'s,
+the optimal schedule was pre-computed using a set of fixed
+$\Gamma_k$'s and $\epsilon_k$'s derived from the longer simulation,
 although ideally they could be estimated from the data
@@
 As shown in Table \ref{tab:lj_error},
-both the optimal and inverse-time schedules
-were able to reduced the error significantly,
-suggesting the potential benefit of using a decreasing updating magnitude
-even with a sometimes sub-optimal schedule.
-%
-It is also reassuring to see that
+both the optimal and sometimes sub-optimal inverse-time schedules
+were able to reduce the error significantly.
+%
+It is also reassuring that
 the errors roughly matched the predicted values
@@
 differed from inverse-time one,
-and the former tended to be larger.
+and the former tended to give larger updating magnitude.
 %
@@
 the error in these modes,
-and produced a flatter profile for the final error.
+and produced a flatter profile for the final error
+(the deviation from a perfectly flat profile
+was likely due to the error of estimated values of $\Gamma_k$'s).
 
@@
 was free from major barriers,
-and thus the histogram fluctuation largely
+and thus the histogram fluctuation
 was largely local in nature.
@@
 %
-When the local noise are dominant,
+When local noise is dominant,
 we found that the bandpass updating scheme
@@
 %
-For example, the prediction showed that
-there was a optimal width
-around $\sigma = 0.37$ for $\varrho = 0.1$
-or $0.07$ for $\varrho = 0.8$,
-assuming the optimal schedule,
-and the optimal width also depended on the simulation length
-and other parameters.
-%
-We recall that a wider Gaussian means
-more middle-range eigemodes
-would have diminishing eigenvalues
-by Eq. \eqref{eq:lambda_Gaussian}.
+With a overly wider Gaussian,
+the updating scheme would fail to
+capture finer features of the PMF.
+%
+From Eq. \eqref{eq:lambda_Gaussian},
+we know that increasing the width of the Gaussian
+would diminish the eigenvalue, $\lambda_k$,
+which is proportional to the rate of error reduction
+by Eq. \eqref{eq:vt_diffeq_mbin}.
 %
@@
 from local histogram fluctuation,
-it also increased the risk of including
-more systematic bias in middle-range modes,
-which can be difficult to remove
-as the rate of error reduction
-is also proportional to the eigenvalue, $\lambda_k$,
-by Eq. \eqref{eq:vt_diffeq_mbin}.
-
-An overly large width $\sigma$ can have a dramatic
-impact on the initial updating magnitude
+it also increased the risk of leaving
+more persistent bias in the middle-range modes.
+%
+For the $\varrho = 0.1$ case,
+the prediction showed that
+assuming the optimal schedule
+the optimal width was around $\sigma = 0.37$,
+and it would decrease with the simulation lengthen.
+%
+Similarly, the optimal mode cutoff, $K$,
+for the homogeneous bandpass updating scheme
+would increase with the simulation length.
+%
+For very long simulations, we expect both to reach
+the limit of single-bin updating scheme
+($\sigma \to 0$ or $K\to n$).
+
+An overly large width $\sigma$ can
+also induce a huge initial updating magnitude
 under the optimal schedule.
@@
 The initial updating magnitude, $\alpha(0)$,
-is usually around
+was usually around
 the ideal value of $a_0/2 = 5\times10^{-5}$
@@
 But if the Gaussian were too wide,
-e.g. if $\sigma = 0.6$,
+e.g. if $\sigma = 0.6$ in the $\varrho = 0.8$ case,
 the eigenvalues $\lambda_k$ of some modes
@@
 In practice, both the width and schedule
-of the Gaussian updating scheme can be sub-optimal.
+of the Gaussian updating scheme are often sub-optimal.
 %
@@
 In Fig. \ref{fig:lj_xerr},
-we show a case where the correction produced
+we show that the correction produced
 the flatter profile seen in the single-bin case,
@@
 %
-However, this correction will also introduce
+However, this correction also introduces
 some local noise in the large $k$ modes,
@@
     the statistical temperature, $\beta(E)$,
-    (dashed line, estimated from the smoothed density of states)
-    demonstrated the backbending behavior.
+    %(dashed line estimated from the smoothed density of states)
+    manifested the backbending behavior
+    typical in a first-order phase transition.
     %
@@
 
-The optimal schedule depends on the updating window function,
-which can be limited to a single-bin (as often in the WL case),
-or extended to several neighboring bins
-taking a Gaussian shape (as often in the metadynamics case).
-%
-We can generally characterize
-an updating scheme by an updating matrix,
-with each column being the normalized
-window function for the bin of the column.
+The optimal schedule depends on the updating scheme,
+which is mathematically characterized by an updating matrix,
+whose element, $w_{ij}$,
+represents the relative updating magnitude
+at bin $i$ when bin $j$ is visited.
 %
@@
 the eigenvalues of the updating matrix,
-which represent the relative rates of error reduction,
-and the integrals of the autocorrelation functions,
-which represent the intrinsic histogram fluctuation
+which reflect the relative rates of error reduction,
+and the intrinsic histogram fluctuations
 of the eigenmodes.
 %
-For the single-bin scheme with a set of uniform eigenvalues,
+For the single-bin scheme (as often used in the WL case),
+whose updating matrix is the identity matrix with identical eigenvalues,
 we confirmed that the optimal schedule
-is given by the inverse-time formula,
+is given by the known inverse-time formula,
 Eq. \eqref{eq:alpha_invt1}.
@@
 For a general multiple-bin scheme,
-including the Gaussian one,
+including the Gaussian one often used in metadynamics,
 the optimal schedule, having to balance
@@
 %
-In this case,
-the optimal schedule and error
-can be sensitive to the simulation length
-and other parameters.
-
-In particular, the Gaussian updating scheme is often
-equipped with a set of rapidly diminishing eigenvalues.
-%
-While this feature helps suppress
-the accumulation of random error and
-maintain the smoothness of the bias potential,
-it also slows down the reduction
-of the initial bias, if any,
-in the short- and middle-range eigenmodes.
-%
-Thus, one needs to carefully choose the width of the Gaussian
-and the optimal schedule of decreasing the updating magnitude,
-which sometimes needs to be fixed at a maximal acceptable level
-for a long time before any reduction can happen.
-%
-The possible efficiency loss in
-error reduction of the long- and middle-range modes
+%In this case,
+%the optimal schedule and error
+%can be sensitive to the simulation length
+%and other parameters.
+
+In particular,
+the Gaussian updating scheme has a nontrivial optimal schedule
+and its width needs to be chosen carefully.
+%
+With a set of rapidly diminishing eigenvalues,
+the Gaussian updating scheme suppresses
+the accumulation of local noise in the bias potential
+%maintain the smoothness of the bias potential,
+at the cost of slowing down the reduction
+of the error in the middle-range eigenmodes.
+%
+Fortunately, the efficiency loss in
+reducing the error of the long- and middle-range eigenmodes
 can often be compensated
 by the histogram-based correction formula, Eq. \eqref{eq:uhatav},
-for post-simulation analysis,
-which ideally would remove the remaining bias
-and recover a bias potential with precision
-similar to that from the single-bin simulation.
+which ideally would remove the persistent error
+in these eigenmodes.
+%and recover a bias potential
+%similar to that from a simulation
+%using the single-bin updating scheme.
 
 In comparing different updating schemes,
-we found that
-the single-bin updating scheme was optimal
-in error reduction
-in the long time limit
+we found
+the single-bin updating scheme to be optimal
+in error reduction in the long time limit
 without a priori assumptions
@@
 %
-The optimal schedule of the bandpass schemes
-is always given by an inverse-time formula.
-%
-If the bandpass updating scheme was constructed
-from orthogonal polynomials,
-it became a generalization
+The bandpass updating schemes allow convenient control
+for the updating magnitude
+as the asymptotically optimal schedule
+is simply given by an inverse-time formula.
+
+As an application,
+we study a particular class
+of the bandpass updating schemes
+constructed from a family of orthogonal polynomials,
+which allows the bias potential
+to be conveniently expressed as a low-order polynomial.
+%with the coefficients adjusted on-the-fly.
+%
+This construction serves as a generalization
 of the LLR algorithm\cite{langfeld2012, pellegrini2014},
-and could be used for parameter adjustment
+and hopefully it could be used for
+adaptive parameter adjustment
 in umbrella sampling with
 quadratic\cite{neuhaus2006, *neuhaus2007, zhu2012}
-or similar\cite{kim2010}
+or similar\cite{martin-mayor2007, *sergio2011, *persson2013, kim2010}
 bias potentials.
@@
 %
-Finally, the adaptive umbrella sampling method introduced in Sec. \ref{sec:aus}
-approach requires refinement to improve
+Finally, the adaptive umbrella sampling method
+introduced in Sec. \ref{sec:aus}
+requires refinement to improve
 its general applicability.
@@
 Runtime-average interpretation
-of the inverse-time schedule
 }
@@
 We will use a linearized version of Eq. \eqref{eq:vcorr_equil},
-$\hat u_i = u_i + \frac{ H_i } { \rho_i } - 1$.
-%
-%\begin{equation}
-%  \hat u_i = u_i + \frac{ H_i } { \rho_i } -1
-%  .
-%  \label{eq:vcorr_equil1}
-%\end{equation}
-%
-Using this formula to
+%$$\hat u_i = u_i + \frac{ H_i } { \rho_i } - 1.$$
+%
+\begin{equation}
+  \hat u_i = u_i + \frac{ H_i } { \rho_i } -1
+  .
+  \notag
+\end{equation}
+%
+Using this formula on
 the instantaneous histogram, $h_i(t)$,
-we get an independent estimate
-of the bias potential at a single step $t$,
+we get an independent estimate of the bias potential
+made by a single data point at step $t$,
 %
@@
   ,
-  \label{eq:u_uhat}
+  \notag
+  %\label{eq:u_uhat}
 \end{align}
 where we have used Eq. \eqref{eq:uhatt}
-and ignored a constant shift $\alpha(t)$
-in the second step.
+and ignored a uniform shift in the second step.
 %
@@
 we have, from Eq. \eqref{eq:vkupdate},
-${\tilde u}_k(t + 1) = {\tilde u}_k(t) + \alpha(t) \, \lambda_k \, {\tilde f}_k(t)$,
-and from Eq. \eqref{eq:u_uhat},
+${\tilde u}_k(t + 1) = {\tilde u}_k(t) + \lambda_k \, \alpha(t) \, {\tilde f}_k(t)$,
+and from Eq. \eqref{eq:uhatt},
 $\hat{\tilde u}_k(t) \approx {\tilde u}_k(t) + {\tilde f}_k(t)$,
@@
 with a weight corresponding to the effective updating magnitude,
-$\lambda_k \, \alpha(t)$.
+$\lambda_k \, \alpha(t)$ [instead of $\alpha(t)$].
+
+\subsection{Upper bound of the updating magnitude}
 
 The runtime average interpretation imposes
-an additional constraint on the updating magnitude.
+an upper bound on the updating magnitude.
 %
@@
   &\approx
-  \frac{1}{t-t'} \ln \frac{ \Omega(t') } { \Omega(t) }
-  =
-  \sum_{\tau=t'+1}^t \frac{ \ln[1- \lambda_k \, \alpha(\tau)] } { t - t' }
+  \frac{1}{t_+ - t_-} \ln \frac{ \Omega(t_-) } { \Omega(t_+) }
+  =
+  \!\!\sum_{\tau=t_- + 1}^{t_+}
+  \frac{ \ln[1- \lambda_k \, \alpha(\tau)] } { t_+ - t_- }
   \\
   &\approx
-  -\lambda_k \cdot \frac{\int_{t'}^t \alpha(\tau) \, d\tau}{t-t'}
-  =
-  -\lambda_k \cdot \frac{q(t) - q(t')}{t- t'}
+  -\frac{ \lambda_k \int_{t_-}^{t_+} \alpha(\tau) \, d\tau}{t_+ - t_-}
+  =
+  -\lambda_k \, \frac{q(t_+) - q(t_-)}{t_+ - t_-}
   ,
 \end{align*}
-where $t$ and $t'$ are two consecutive grid points.
+where $t_-$ and $t_+$ are two neighboring grid points
+around $t$.
 %
 Using this formula to mode $k = 1$
-with the largest eigenvalue
+of the largest eigenvalue
 would guarantee $\lambda_k \, \alpha(t) \ge 1$
@@
 with a dynamic weight,
-which is not necessarily even to be most efficient.
+which is not necessarily the optimal even weight.
 %
@@
 if the mass distribution is inversely proportional to $\bar q$,
-which is approximately true for the Gaussian updating scheme
+which is approximately the case for the Gaussian updating scheme
 at large $\bar q$,\footnote{If
-  the magnitudes of the eigenvalues, $\lambda_k$'s,
-  span a wide range, then for a large $\bar q$,
+  the eigenvalues, $\lambda_k$'s,
+  differ greatly in magnitude,
+  then for a large $\bar q$,
   the sum in Eq. \eqref{eq:mass_func} will be dominated by
@@
 would fail if some eigenvalue, $\lambda_k$, is close to zero.
-So it is best used for the single-bin updating scheme.
+So it is best used for the single-bin updating scheme.\footnote{Alternatively,
+  we can split the trajectory into blocks,
+  compute $\Gamma_k$ from the histogram fluctuation of each block
+  using Eq. \eqref{eq:Fk2_Gammak},
+  and then average the results.}
 
@@
 by a rigid window function
-that shifts with the bin $j$ of visit
+that shifts with a visit to bin $j$
 (ignoring the boundary adjustment).
@@
 where
-$\mathrm{sinr}(K, l, n)
-  =
+%$\mathrm{sinr}(K, l, n)
+%  =
+%  \sin
+%  \frac{ (2 K - 1) \, l \, \pi }
+%       {              n        }
+%  / \left( n \, \sin \frac{ l \, \pi } { n } \right)$.
+\begin{equation}
+  \mathrm{sinr}(K, l, n)
+  \equiv
+  \frac{
     \sin
@@
          {              n        }
-  / \left( n \, \sin \frac{ l \, \pi } { n } \right)$.
-%\begin{equation}
-%  \mathrm{sinr}(K, l, n)
-%  =
-%  \frac{
-%    \sin
-%    \frac{ (2 K - 1) \, l \, \pi }
-%         {              n        }
-%  }
-%  {
-%    n \, \sin \frac{ l \, \pi } { n }
-%  }
-%  .
-%\notag
-%%\label{eq:mu_sinc_pbc}
-%\end{equation}
+  }
+  {
+    n \, \sin \frac{ l \, \pi } { n }
+  }
+  .
+\notag
+%\label{eq:mu_sinc_pbc}
+\end{equation}
 %
@@
 for the bandpass updating scheme based on orthogonal polynomials.
-%as a generalization of the WL algorithm for the single-bin scheme.
 
@@
 
-Consider an ideal FDS simulation under the exact bias potential.
+Consider an FDS simulation
+under a nearly exact bias potential.
 %
@@
 the fluctuation of the instantaneous histogram
-is then given by the random part,
-%
-\begin{equation}
-  f_{*i}(t) = \xi_{*i}(t)
+is then dominated by the random part,
+%
+\begin{equation}
+  f_{*i}(t) \approx \xi_{*i}(t)
   ,
@@
 %the same holds for the Fourier transform,
-$\tilde f_k(t) = \tilde \xi_k(t).$
+$\tilde f_k(t) \approx \tilde \xi_k(t).$
 %
@@
 %
-\begin{align*}
+\begin{align}
   \bigl\langle \tilde F_k^2(T) \bigr\rangle
@@
   \bigr\rangle
-\\
+  \notag \\
   &\approx
@@
   ,
-\end{align*}
+  \label{eq:Fk2_Gammak}
+\end{align}
 %
@@
   \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)
-  =
-  \overline{
-    R_k\bigl( z(t) \bigr)
-  }
   .

### 2017-08-22 doc/invt.tex
absorb Justin's changes

@@
   %
-  We will show that
+  We show that
   the optimal schedule depends on the updating scheme.
@@
   While the asymptotically optimal schedule for
-  the single-bin updating scheme (commonly used in WL)
+  the single-bin updating scheme (commonly used in the WL algorithm)
   is given by an inverse-time formula,
@@
   These bandpass updating schemes aim at
-  partially flattening a few selected histogram modes
+  partially flattening a few long-range histogram modes
   and their optimal schedule
@@
   wang2001pre, kim2006, *kim2007},
-the latter is more popular for MD simulations
+the latter is commonly used for MD simulations
 as it provides a smoother bias potential
@@
 increases the variance of the bias potential,
-an too rapid reduction often
+a too rapid reduction often
 converges the bias potential to wrong value.
@@
 %
-The bandpass updating schemes allow
+Compared with the Gaussian updating scheme,
+Bandpass updating schemes have the advantage of allowing
 simpler control of the updating magnitude
-than the Gaussian updating scheme
-for the asymptotically optimal schedule of the former
-is given by the inverse-time formula.
+for asymptotic convergence.
 
@@
 a polynomial form,
-and coefficients adjusted on-the-fly
-to achieve the desired target distribution.
+and coefficients are adjusted on-the-fly
+to achieve a particular target distribution.
 %
@@
 single-bin updating scheme
-used by the WL algorithm,
+(as in the WL algorithm)
 by proving the optimality
@@
 %
-Next, we extend the method to the general case
+We extend the method to the general case
 of a multiple-bin updating scheme
@@
 %
-We then compare different updating schemes
+Then, we compare different updating schemes
 and show that the single-bin scheme
@@
 the distribution, $p_i$,
-along a discrete quantity, $i$,
+along a discrete collective variable, $i$,
 for a given system.
@@
 %
-For a continuous quantity, $z$,
+For a continuous collective variable, $z$,
 %which is often the case in a (classical) molecular system,
@@
 
-Sampling a multimodal distribution with deep basin(s)
+Sampling a multimodal distribution with deep basins
 is often challenging in regular MC and MD simulations.
@@
 %(hence that of the PMF)
-as simulation lengthens.
+as the simulation lengthens.
 
@@
 the bias potential, $u_i(t)$, is updated
-at each step $t$ as
+at each time step $t$ as
 %
@@
 %including the Gaussian updating scheme used in metadynamics,
-the updating also involves neighboring bins:
+neighboring bins are also updated:
 %
@@
 A common example is the Gaussian updating scheme used in metadynamics,
-with $w_{ij}$ being roughly a Gaussian centered on bin $j$.
+with $w_{ij}$ being roughly a Gaussian centered on bin $j = i(t)$.
 %
@@
 only every few sampling steps,
-then $t$ means the number of updating steps.
+then $t$ denotes the number of updating steps.
 
@@
 A more effective way
-of controlling the schedule, $\alpha(t)$,
+of controlling the schedule of the updating magnitude, $\alpha(t)$,
 in late stages
@@
   robbins1951, pellegrini2014},
-and it has an intuitive explanation\cite{
+and it has an intuitive explanation:\cite{
   marsili2006, barducci2008}
-that interprets the bias potential
-as a runtime average computed from all data points collected so far,
-which absorbs the contribution of the newest data point
-by its due weight,
+the bias potential undergoing the single-bin updating scheme
+can be mapped to a runtime average computed from all data points collected so far;
+upon the addition of a new data point,
+the average absorbs its contribution by the due weight,
 the inverse of the current sample size, $t$
@@
 In this section,
-we will derive the optimal $\alpha(t)$
+we derive the optimal $\alpha(t)$
 for the single-bin scheme,
@@
 %
-First, to deduct the target value given by
-Eq. \eqref{eq:Vi_target},
-we introduce a shifted bias potential
+First, we deduct the target value
+[Eq. \eqref{eq:Vi_target}]
+by introducing a shifted bias potential
 %
@@
 upon convergence.
-Note that the updates to $v_i(t)$ are
+Note that updates to $v_i(t)$ are
 equivalent to the updates to $u_i(t)$,
@@
 
-Second, the bias potential following Eq. \eqref{eq:wl_update}
+The bias potential following Eq. \eqref{eq:wl_update}
 generally grows over time.
@@
 %
-We define the bias potential deviation from the
-bin average, $\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$,
-as
+In the second deduction, we define
+the bin average as $\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$,
+and the deviation of the bias potential as
 %
@@
 %
-Below we will use the notations, $x_{*i}$ and $\bav{x}$,
+Below we will use the notations, $\bav{x}$ and $x_{*i}$,
 for a general variable, $x_i$.
@@
 
-To proceed, we will
-approximate the finite difference Eq. \eqref{eq:wl_update}
+To proceed, we
+approximate the finite difference equation, Eq. \eqref{eq:wl_update},
 by a differential equation\footnote{In
@@
 We will approximate the histogram fluctuation part, $\xi_i(t)$,
-hence its deviation from the bin average, $\xi_{*i}(t)$,
+and hence its deviation from the bin average, $\xi_{*i}(t)$,
 as white noise such that
@@
 \end{equation}
-for some $G_i$,
+for some constant $G_i$,
 and $\delta(t)$ is the Dirac delta function.
 %
-We expect the approximation to hold for simulation
+We expect the approximation to hold
+for simulations that are
 much longer than the autocorrelation time.
@@
 Without the random part, $\xi_{*i}(t)$,
-this equation would drive
-the deviation of the bias potential, $v_{*i}(t)$,
-to $0$ at long times,
-implying the convergence.
+this equation is convergent
+as the deviation of the bias potential, $v_{*i}(t)$,
+would tend to $0$ at long times.
 %
@@
 %
-and Eq. \eqref{eq:alpha_invt}
-is the special case of $t_0 = 0$.
+with Eq. \eqref{eq:alpha_invt} being
+the special case of $t_0 = 0$.
 %
@@
 We now determine the optimal value of $q(T)$,
-which is fixed in the above.
+which was previously held fixed.
 %
@@
 By definition, Eq. \eqref{eq:mbin_update},
-a visit to bin $j$ in this case results in updates of the bias potential
+a visit to bin $j = i(t)$ in this case results in updates of the bias potential
 not only at bin $j$, but also at a few neighboring bins $i$'s.
@@
 The key simplification comes from the
-projection of the bias potential to the $n$ eigenmodes
+projection of the bias potential to the eigenmodes
 of the updating matrix, $\mathbf w$,
@@
 except that some elements can be negative
-for our cases here.
+for cases considered here.
 %
@@
 %
-and Eq. \eqref{eq:eigenmode0} means that
+and Eq. \eqref{eq:eigenmode0} shows that
 the bin average
@@
   \bigl\langle
-    \tilde\xi_k(t) \tilde\xi_k(0)
+    \tilde\xi_k(t) \, \tilde\xi_k(0)
   \bigr\rangle
@@
 with the second expression being
-the equivalent in the discrete-time setting.
+the counterpart in the discrete-time setting.
 %
@@
 \subsubsection{\label{sec:procedure}
-Procedure of computing the optimal schedule and error
+Procedure for computing the optimal schedule and error
 }
@@
 Based on the above results,
-we give a procedure of computing
+we give a procedure for computing
 the optimal schedule and the error
@@
 %
-We have
+The procedure is given below.
 
@@
 whose updating matrix can be characterized by
-a fixed Gaussian centered on the bin of visit
-(with some boundary adjustment).
+a fixed Gaussian centered on the current bin
+(with some boundary adjustment),
 %to satisfy the constraints discussed in Sec. \ref{sec:updating-matrix},
-%as shown in Fig. \ref{fig:mat}(a).
+see Fig. \ref{fig:mat}(a) for an example.
 %
@@
 where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
-are the width of the Gaussian and the span
-in the original collective variable, $z$,
+are the width of the Gaussian and the range of
+the original collective variable, $z$,
 respectively.
 
-More generally, we can define a class of
-translationally-invariant or \emph{homogeneous} updating schemes
-by a rigid window function with a floating center.
+More generally, we can replace the Gaussian
+in the above updating scheme
+by a window function of different shape.
+%
+This results in a
+translationally-invariant or \emph{homogeneous} updating scheme,
+whose updating matrix can be
+characterized by a rigid window function with a floating center,
+see Fig. \ref{fig:mat}(b) for an example.
 %
@@
 optimal for asymptotic convergence.
-We will pursue this strategy below.
+We discuss this strategy in detail below.
 
@@
 for a given updating scheme,
-we will turn to the problem of finding
+we turn to the problem of finding
 optimal updating schemes
-given that each updating scheme
+assuming that each updating scheme
 adopts the optimal schedule.
@@
 
-We will first show
+We first show
 that the single-bin scheme is asymptotically optimal.
@@
 and the PMF is sufficiently smooth
-to be representable by the first $K$
+to be represented by the first $K$
 eigenmodes,
-we may assign the first $K$ eigenmodes eigenvalue $1$
-and the rest of eigenmodes eigenvalue $0$:
-%
-\begin{equation}
-  \lambda_1 = \cdots = \lambda_{K-1} = 1
+we may assign the eigenvalues of the first $K$ eigenmodes to $1$
+and the remaining eigenvalues to $0$:
+%
+\begin{equation}
+  \lambda_0 = \cdots = \lambda_{K-1} = 1
   ,
@@
 
-We will consider two special cases of the bandpass updating schemes.
+We will consider two special classes of bandpass updating schemes.
 %
@@
 In Sec. \ref{sec:lj},
-we will further compare the performance of the two updating schemes
+we will further compare the performance of
+the homogeneous bandpass and Gaussian updating schemes
 in an example.
@@
 %
-The second case is based on a family of orthogonal polynomials,
+The second class of bandpass updating schemes
+is constructed from a sequence of orthogonal polynomials,
 which we will discuss below.
@@
 
-Here we will consider here a family of bandpass updating schemes
+Here we consider a class of bandpass updating schemes
 constructed from a sequence of orthogonal polynomials.
 %
-These updating schemes are more intuitive and easier to implement
+These updating schemes are more intuitive
+than the homogeneous bandpass updating scheme
 as the bias potential in this case is
-a low-order polynomial of the collective variable.
+a simple polynomial of the collective variable.
 
@@
 
-It is convenient to
+For convenience, we
 replace the discrete bin index, $i$,
@@
 %
-and we will replace $\rho_i$ by $\rho(z) \, \delta z$,
-  $\phi_{ki}$ by $\phi_k(z) \, \delta z$,
+$\rho_i$ by $\rho(z) \, \delta z$;
+$\phi_{ki}$ by $\phi_k(z) \, \delta z$;
 and
@@
 $\phi_k(z) = R_k(z) \, \rho(z)$,
-where $R_k(z)$ is the $k$th member of an sequence of orthogonal polynomials
+where $R_k(z)$ is the $k$th member of a sequence of orthogonal polynomials
 that satisfy
@@
   =
-  \sum_{k=1}^{K-1} \frac{ \phi_k(z) \, \phi_k(z') } { \rho(z) }
-  = \sum_{k=1}^{K-1} R_k(z) \, R_k(z') \, \rho(z'),
+  \sum_{k=0}^{K-1} \frac{ \phi_k(z) \, \phi_k(z') } { \rho(z) }
+  =
+  \sum_{k=0}^{K-1} R_k(z) \, R_k(z') \, \rho(z')
+  .
 \end{align*}
 %
-where we have dropped the $k=0$ mode
-that corresponds to a uniform shift
-of the bias potential.
+%where we have dropped the $k=0$ mode
+%that corresponds to a uniform shift
+%of the bias potential.
 %
@@
 u(z, t+1) = u(z, t)
-+ \sum_{k=1}^{K-1} R_k(z) R_k\bigl( z(t) \bigr) \, \alpha(t).
++ \sum_{k=0}^{K-1} R_k(z) R_k\bigl( z(t) \bigr) \, \alpha(t).
 $$
@@
 \begin{equation}
-  u(z, t) = \sum_{k=1}^{K-1} c_k(t) \, R_k(z),
+  u(z, t) = \sum_{k=0}^{K-1} c_k(t) \, R_k(z),
   \label{eq:uz_decomp}
@@
 with the coefficients updated as
+%
 \begin{equation}
@@
 
-\subsubsection{LLR algorithm}
+\subsubsection{Langfeld-Lucini-Rago (LLR) algorithm}
 
@@
 If the interval is sufficiently small,
-we can limit the basis to a single linear mode,
-$R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$,
-with $K = 2$.
+we can use $K = 2$ to limit the basis to the first two modes,
+$R_0(z) = 1$ and $R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$.
 %
 %The bias potential is a linear function of $z$,
-The updating scheme is essentially the LLR algorithm\cite{langfeld2012}
+If we drop the $k = 0$ mode
+that corresponds to a uniform shift of the bias potential,
+the updating scheme is essentially the LLR algorithm\cite{langfeld2012}
 for the single-interval case,
@@
 was recommended for optimal convergence of $c_1$\cite{pellegrini2014}.
-
-In real application, the algorithm is often applied simultaneously
+%
+In practical applications, the algorithm is often applied simultaneously
 to an array of adjacent intervals
-to cover a wider range of the collective variable, $z$.
-%
-In this way, attempts to move out of the current interval
-are not necessarily rejected,
-but can be interpreted as transitions to
-another internal and accepted with proper probability.
+to cover a wide range over the collective variable, $z$.
+%
+%In this way, attempts to move out of the current interval
+%are not necessarily rejected,
+%but can be interpreted as transitions to
+%another internal and accepted with proper probability.
 
@@
 %
-The bias potential is
+The bias potential is thus
 quadratic\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007},
 which is common in umbrella sampling\cite{torrie1974, *torrie1977}
-and the method of Gaussian ensemble\cite{hetherington1987,
+and the Gaussian ensemble method\cite{hetherington1987,
 *challa1988, *costeniuc2006}.
@@
 $$
-u(z) = \frac{c_1}{\sigma_z} (z - z_c)
+u(z) = c_0 + \frac{c_1}{\sigma_z} (z - z_c)
 + \sqrt 2 \, c_2 \left[ \frac{(z-z_c)^2}{2 \, \sigma_z^2} - \frac{1}{2} \right].
@@
 by constructing an expanded ensemble
-from a superposition of multiple Gaussian umbrellas
-centered at different places via
-the frameworks of
+from a superposition of $m$ Gaussian umbrellas
+centered at different places.
+%
+The target distribution is a joint distribution
+of the discrete umbrella index, $i = 1, \dots, m$,
+and the continuous variable, $z$,
+$\rho(i, z) = \rho_i \, \rho^{(i)}(z)$,
+where the marginal inter-umbrella distribution is flat,
+$\rho_i = 1/m$,
+whereas the conditional intra-umbrella distribution, $\rho^{(i)}(z)$
+is the Gaussian given previously.
+%
+To sample this distribution, we give each umbrella
+an independent bias potential, $u^{(i)}(z)$,
+of the form of Eq. \eqref{eq:uz_decomp},
+and update coefficients as
+\begin{align*}
+  c_k^{(i)}(t+1) = c_k^{(i)}(t)
+  + m \, \alpha(t) \, R_k^{(i)}\bigl( z(t) \bigr) \, \delta_{i, i(t)}
+  ,
+\end{align*}
+%
+where the factor $m$ before $\alpha(t)$
+is due to the inverse of the marginal inter-umbrella distribution, $\rho_i$.
+%
+Note that since $R_0^{(i)}(z) = 1$,
+the parameter, $c_0^{(i)}$,
+serves as the bias potential
+for inter-umbrella transitions,
+and it is updated according to the single-bin scheme\cite{li2007}.
+%
+In the expanded ensemble,
+inter-umbrella transitions can be handled via the frameworks of
 parallel\cite{swendsen1986,
@@
 %
-Here, we will focus on the latter framework,
-in which the transition probability from
+Using the latter framework,
+the transition probability from
 the $i$th umbrella to the $j$th is given by
@@
 \min\left\{1,
-  \frac{Z^{(i)}}{Z^{(j)}}
   e^{
@@
 %
-The parameter, $Z^{(i)}$, is the inverse weight of the $i$th umbrella,
-in the expanded ensemble,
-and $\ln Z^{(i)}$ serves as the bias potential
-for inter-umbrella transitions.
-%
-We can be estimate $\ln Z^{(i)}$'s on-the-fly
-using the single-bin updating scheme,
-Eq. \eqref{eq:wl_update}\cite{li2007}.
-%
-Upon convergence, $Z^{(i)}$ should be proportional to
-the partition function,
-$\int e^{-\beta \, \Phi(\mathbf x)-u^{(i)}(z(\mathbf x),t)} \, d\mathbf x$,
-as in simulated tempering,
-where
-$\beta = 1/(k_B T)$ is the inverse temperature,
-$\mathbf x$ denotes the internal degrees of freedoms,
-and
-$\Phi(\mathbf x)$ is the unbiased potential energy.
-%
-
-Since bandpass updating schemes are straightforward extension
+%and upon convergence, $Z^{(i)}$ should be proportional to
+%the partition function,
+%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(i)}(z(\mathbf x),t)} \, d\mathbf x$,
+%as in simulated tempering,
+%where
+%$\beta = 1/(k_B T)$ is the inverse temperature,
+%$\mathbf x$ denotes the internal degrees of freedoms,
+%and
+%$\Phi(\mathbf x)$ is the unbiased potential energy.
+%
+
+Since bandpass updating schemes are a straightforward extension
 of the single-bin updating scheme,
@@
 We can adopt the WL stage-by-stage strategy early in simulation,
-and switch to the inverse-time schedule later
-once the updating magnitude from the former approach
-falls under that from the latter.\footnote{We
+and switch to the inverse-time schedule
+once the updating magnitude from the WL approach
+falls under the value given by the inverse-time formula.\footnote{We
   slightly modified the transition criterion
@@
     set $c_1^{(i)}$ to the estimated mean force,
-    and set $\ln Z^{(i)} = c_2^{(i)} = 0$.
+    and set $c_0^{(i)} = c_2^{(i)} = 0$.
 
@@
   in the current umbrella $i$ under bias potential $u^{(i)}(z, t)$.
-  Update $c_1^{(i)}$ and $c_2^{(i)}$ according to Eq. \eqref{eq:ckupdate},
-  and compute the fluctuations
-  $\tilde F_1^{(i)}$ and $\tilde F_2^{(i)}$
-  defined in Eq. \eqref{eq:Fk_def}.
+  Update the coefficients, $c_k^{(i)}$'s,
+  according to Eq. \eqref{eq:ckupdate}.
 
@@
   with the acceptance probability given by Eq. \eqref{eq:transprob_st}.
-  Update $\ln Z^{(i)}$ at the current umbrella
-  by Eq. \eqref{eq:wl_update}.
 
 \item
-  Change the updating magnitude, $\alpha(t)$,
-  which will be used for both intra-umbrella
-  parameters, $c_k^{(i)}$,
-  and inter-umbrella parameters, $\ln Z^{(i)}$.
+  Change the updating magnitude, $\alpha(t)$.
 
@@
   {
-    \sum_i N^{(i)} \, e^{-u^{(j)}(z)} / Z^{(i)}
+    \sum_i N^{(i)} \, e^{-u^{(j)}(z)}
   }
@@
 We can use the final converged values of $Z^{(i)}$'s
-without going through the usual WHAM iteration.
+without going through the usual WHAM iterations.
 
@@
 and we wish to compute the PMF along
-the distance, $r$, between two special particles.
+the distance, $r$, between two particular particles.
 %
@@
 %
-The two special particles, labeled as $1$ and $2$,
+These particles, labeled as $1$ and $2$,
 are mutually non-interacting
-but they interact normally with other particles
-via the pair potential $\varphi(r) = 4(r^{-12} - r^{-6})$.
-%
-The intrinsic distribution of distance, $p(r)$, is
+but do interact with other particles
+via the pair potential $\varphi(r) = 4 \, \left(r^{-12} - r^{-6}\right)$
+(truncated at half of the box size).
+%
+The PMF can be derived from
+the intrinsic distribution of the distance
+between the two tagged particles, $p(r)$,
+as $-\ln p(r)$,
+%
+Formally, we have
 \begin{align*}
@@
   \frac{V}{Z}
-  \int e^{-\beta \Phi(\mathbf r^N)}
+  \int e^{-\beta \, \Phi(\mathbf r^N)}
   d\mathbf r_3 \cdots d\mathbf r_N
@@
 where
-$V$ is the volume,
+$V$ is the reduced volume,
 $\Phi(\mathbf r^{N}) = \sum_{j=3}^N\sum_{i=1}^{j-1} \varphi(r_{ij})$,
@@
 %
-This distribution is related the cavity distribution function
-as $y(r) = \left(1 - \frac{1}{N}\right) \, V \, p(r)$\cite{hansen}.
+This distribution is proportional to the cavity distribution function\cite{hansen}.
+%as $y(r) = \left(1 - \frac{1}{N}\right) \, V \, p(r)$\cite{hansen}.
 
@@
 Each step consists of
-one MC trial for displacing one of the two special particles,
+one MC trial for displacing one of the two tagged particles,
 and another four trials for displacing the other particles.
@@
 %
-However, the two special particles are fixed on the $x$-axis,
+The two tagged particles are fixed on the $x$-axis,
 and are only displaced in this direction.
@@
 and calculated the PMF from $r = 0$ to half of the box size
-(also the cutoff distance of the potential)
+%(also the cutoff distance of the potential)
 on a grid of size $\delta r = 0.01$ in reduced units.
 %
-Two densities were tried, $\varrho = 0.1$ and $0.8$,
+Two densities were used, $\varrho = 0.1$ and $0.8$,
 and the numbers of bins were $n = 500$ and $250$, respectively.
@@
 %
-To make sure the same schedule for multiple runs under the same condition,
+To ensure the same schedule was used for multiple runs under the same condition,
 the optimal schedule was pre-computed using a set of fixed
@@
 As shown in Table \ref{tab:lj_error},
-both the optimal and sometimes sub-optimal inverse-time schedules
+both the optimal and potentially sub-optimal inverse-time schedules
 were able to reduce the error significantly.
 %
-It is also reassuring that
-the errors roughly matched the predicted values
+The errors roughly matched the predicted values
 calculated using the pre-computed
@@
 differed from inverse-time one,
-and the former tended to give larger updating magnitude.
+and the former tended to give a larger updating magnitude.
 %
@@
 For the same case,
-we further decomposed the error (after $10^7$ production steps)
+we further decomposed the error
 into the eigenmodes
@@
 and produced a flatter profile for the final error
-(the deviation from a perfectly flat profile
+(the remaining unevenness of the profile
 was likely due to the error of estimated values of $\Gamma_k$'s).
@@
     on the LJ system at density $\varrho = 0.1$
-    after $10^7$ production steps.
+    after $10^8$ production steps.
     %
@@
 
-Next, we will discuss the applicability of
+Next, we discuss the applicability of
 the Gaussian updating scheme.
@@
 that correspond to the long-range fluctuations
-(as $\Lambda_1 \approx \Lambda$),
+(as $\Gamma_1 \approx \Gamma$),
 and the reduction of local noise by multiple-bin schemes
@@
 was more effective than the Gaussian updating schemes
-as the expected.
+as expected.
 
@@
 %
-With a overly wider Gaussian,
+With a overly wide Gaussian,
 the updating scheme would fail to
@@
 effectively suppress random error
-from local histogram fluctuation,
+from local histogram fluctuations,
 it also increased the risk of leaving
@@
 the prediction showed that
-assuming the optimal schedule
+assuming the optimal schedule,
 the optimal width was around $\sigma = 0.37$,
@@
 An overly large width $\sigma$ can
-also induce a huge initial updating magnitude
+also induce a very large initial updating magnitude
 under the optimal schedule.
@@
 In this case, the predicted optimal schedule
-becomes invalid the resulting deviation in the bias potential
+becomes invalid as the resulting deviation in the bias potential
 is no longer small.
@@
 In Fig. \ref{fig:lj_xerr},
-we show that the correction produced
-the flatter profile seen in the single-bin case,
+we show that this correction produced
+the flatter profile seen in the single-bin case
 for a simulation with a sub-optimal schedule.
@@
 in a post-simulation analysis, since
-the bias potential no longer need to be differentiated
+the bias potential no longer needs to be differentiated
 for force calculation.
@@
 %
-In this case, $\Phi(\mathbf x) = 0$
-and the intrinsic distribution is
+In this case, % $\Phi(\mathbf x) = 0$ and
+the intrinsic distribution is
 the density of states, $g(E)$.
@@
 In conclusion,
-we have proposed a method of computing
+we have proposed a method for computing
 the optimal schedule of the updating magnitude
@@
 which is mathematically characterized by an updating matrix,
-whose element, $w_{ij}$,
-represents the relative updating magnitude
+with the element, $w_{ij}$,
+representing the relative updating magnitude
 at bin $i$ when bin $j$ is visited.
@@
 In a homogeneous bandpass updating scheme,
-the eigenvalues are either $0$ or $1$ and
-it is convenient to set $\lambda_0 = 1$.
+the eigenvalues are either $0$ or $1$.
+%and it is convenient to set $\lambda_0 = 1$.
 %

### 2017-08-22 prog/lj/lj.c

@@ -198,7 +198,7 @@ static int work(invtpar_t *m)
   double dr, errtrunc, hfl, errc;
   metad_t *metad;
 
-  //mtscramble(clock());
+  mtscramble(clock());
 
   lj = lj_open(np, rho, rcdef);
   lj_energy(lj);

### 2017-08-24 doc/invt.tex
pt: switch to all-to-all interumbrella transitions

@@
 $z$ on the free energy surface
-where the updating should occur
+%where the updating should occur
 as well as the relative updating magnitude.
@@
 $\delta$-function (a boxcar function one histogram bin wide)
-in WL,
+in the WL algorithm,
 or that of a Gaussian
@@
 %
-The resulting optimal schedule
-depends on the updating window function
-through the mass of the particle.
-%
-While the optimal schedule in the single-bin case
+%The resulting optimal schedule
+%depends on the updating window function
+%through the mass of the particle.
+%
+We found that
+while the optimal schedule in the single-bin case
 is given by the known inverse-time formula,
-that for Gaussian case is more complex
-and sensitive to the simulation length
-and the width of the Gaussian.
+that for Gaussian case is more complex.
+%and sensitive to the simulation length
+%and the width of the Gaussian.
 
@@
 the single-bin updating scheme
-is optimal for long simulations
-without a priori assumption of the smoothness of the PMF.
-%
-The same asymptotic optimality also applies to
+is optimal for very long simulations.
+%without a priori assumption of the smoothness of the PMF.
+%
+A similar but partial asymptotic optimality applies to
 a class of bandpass updating schemes
@@
 Compared with the Gaussian updating scheme,
-Bandpass updating schemes have the advantage of allowing
+bandpass updating schemes have the advantage of allowing
 simpler control of the updating magnitude
@@
 the bias potential, $u_i$, is
-estimated a priori and fixed
+estimated beforehand and fixed
 during simulation,
@@
 and metadynamics\cite{huber1994, laio2002, *laio2008, *barducci2011, *sutto2012},
-which address the above shortcoming
-by continuously and incrementally updating the bias potential.
+which
+continuously and incrementally update the bias potential.
 %
@@
 We refer to this scheme as
-the \emph{single-bin (updating) scheme},
+the \emph{single-bin} (updating) scheme,
 for it applies only to the bias potential
@@
 In a more general
-\emph{multiple-bin (updating) scheme},
+\emph{multiple-bin} (updating) scheme,
 %including the Gaussian updating scheme used in metadynamics,
@@
 the bias potential undergoing the single-bin updating scheme
-can be mapped to a runtime average computed from all data points collected so far;
+can be mapped to a cumulative moving average computed from all data points collected so far;
 upon the addition of a new data point,
@@
 This is equivalent to choosing a value of $t_0$
-\big[because $q(T) = \ln\bigl[1 + (T/t_0)\bigr]$,
+\big[because $q(T) = \ln\bigl(1 + \frac{T}{t_0}\bigr)$,
 from the $t = 0$ case of Eq. \eqref{eq:expqt}\big].
@@
   .
-  \label{eq:opt_alpha0}
+  \notag
+  %\label{eq:opt_alpha0}
 \end{align}
@@
 i.e. the optimal initial updating magnitude
-is half of the previous equilibrium value\footnote{This factor, $1/2$,
+is half of the previous equilibrium value.\footnote{This factor, $1/2$,
   happens to coincide with the
@@
   in the original WL algorithm\cite{
-wang2001, wang2001pre}.
-}
+  wang2001, wang2001pre}.  }
 %[cf. Eq. \eqref{eq:t0_sbin}].
@@
   by Eq. \eqref{eq:uhatav},
-  although it leads to overestimated
+  although it often leads to overestimated
   $\epsilon_k$'s for large-$k$ modes.
@@
   in computing its change.
-  The initial value was set to $q(T) = \ln(1+T\,a_0/2)$.}
+  The initial value was given by
+  $q(T) = \ln\left(1+\frac{1}{2}T\,a_0\right)$.}
 %
@@
 %
-To handle Gaussian updating schemes more effectively,
-we have chosen the $\mathcal N+1$ grid points
-($\mathcal N=20000$ here)
-over $[0, q(T)]$ to be given by
+%To handle Gaussian updating schemes more effectively,
+%we have chosen the $\mathcal N+1$ grid points
+%($\mathcal N=20000$ here)
+%over $[0, q(T)]$ to be given by
+In our implementation,
+the integration was performed using the trapezoidal rule\cite{press3rd}
+on the grid points,
 $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
-where $i = 0, \dots, \mathcal N$.\footnote{For
-  a mass distribution, $m(\bar q)$, that is
-  inversely proportional to $\bar q$,
+where $i = 0, \dots, \mathcal N$.\footnote{The
+  grid points are designed for the Gaussian updating scheme.
+  For a mass function, $M(\bar q)$, that is
+  inversely proportional to $\bar q$
+  (cf. Appendix \ref{sec:schedule_geometry}),
   this grid would make the integral,
   or the difference, $t_{[i+1]} - t_{[i]}$,
-  roughly a constant.}
-%
-The curve $q(t)$ is then integrated
-using the trapezoidal rule\cite{press3rd},
-resulting in a list of values, $\bigl(q_{[i]}, t_{[i]} \bigr)$.
-%
-We then compute the updating magnitude
-as (cf. Appendix \ref{sec:ave_us}):
+  roughly a constant.
+  We used $\mathcal N = 20000$.}
+%
+The resulting values, $\bigl(q_{[i]}, t_{[i]} \bigr)$,
+and used to compute the updating magnitude
+as (cf. Appendix \ref{sec:ave_ub}):
 \begin{equation}
@@
   =
-  \frac{1}{\lambda}
+  \frac{1}{\lambda_1}
   \left[
@@
 with
-$q_{[ -1]} \equiv 0$,
+$q_{[ -1]} \equiv t_{[-1]} \equiv 0$,
 $q_{[\mathcal N+1]} \equiv q(T)$,
-$t_{[ -1]} \equiv 0$, and
+and
 $t_{[\mathcal N+1]} \equiv T$.
 %
-The resulting values of $(t_{[i]}, \alpha_{[i]})$
-can be linearly interpolated
-to get the updating magnitude, $\alpha(t)$, at any simulation time, $t$.
+The values of $(t_{[i]}, \alpha_{[i]})$
+are linearly interpolated
+to get the updating magnitude, $\alpha(t)$, at any $t$.
 
@@
 whose updating matrix can be characterized by
-a fixed Gaussian centered on the current bin
+a fixed Gaussian function centered on the current bin
 (with some boundary adjustment),
@@
 
-More generally, we can replace the Gaussian
+More generally, we can replace the Gaussian function
 in the above updating scheme
@@
 %
-This results in a
-translationally-invariant or \emph{homogeneous} updating scheme,
-whose updating matrix can be
-characterized by a rigid window function with a floating center,
-see Fig. \ref{fig:mat}(b) for an example.
+We call the resulting updating scheme
+translationally-invariant or \emph{homogeneous}.
+%see Fig. \ref{fig:mat}(b) for an example.
+
+By morphing the window function without losing the translational invariance,
+we can design a homogeneous updating scheme
+that is more effective in error reduction
+than the Gaussian updating scheme.
+%
+%whose updating matrix can be
+%characterized by a rigid window function with a floating center,
 %
@@
 and differ only by the eigenvalues.
-
-Note that we can reconstruct the updating matrix
-from the eigenvalues and eigenvectors
-as\cite{bussi2006}
+%
+Thus, the window function is completely determined
+by the eigenvalues, and the updating matrix
+is given by\cite{bussi2006}
 %
@@
 %
-Thus, it is possible to improve the updating scheme
-by selecting a spectrum of eigenvalues
-optimal for asymptotic convergence.
 We discuss this strategy in detail below.
@@
 we turn to the problem of finding
-optimal updating schemes
-assuming that each updating scheme
-adopts the optimal schedule.
+the optimal updating scheme.
 
@@
 \end{align*}
-and $\lambda_1$ must be $1$ by \eqref{eq:w_sumj}.
+and $\lambda_1$ must be $1$ by Eq. \eqref{eq:w_sumj}.
 %
@@
 %
-This demonstrates the asymptotic optimality of
-the single-bin updating scheme.
+%This demonstrates the asymptotic optimality of
+%the single-bin updating scheme.
 %
@@
 %which is less than that from Eq. \eqref{eq:EA_sbin1}.
-
-We will consider two special classes of bandpass updating schemes.
+%
+Like the single-bin updating scheme,
+the bandpass updating scheme is also
+asymptotically optimal,
+but only for the first $K$ fluctuation modes.
+
+We will consider two special classes of bandpass updating schemes,
+homogeneous and polynomial bandpass updating schemes.
 %
@@
 that resembles the Gaussian updating scheme,
-but with better convergence in the asymptotic limit
-(cf. Appendix \ref{sec:homo_bandpass}).
+but with long-range wiggles
+[cf. Appendix \ref{sec:homo_bandpass}
+and Fig. \ref{fig:mat}(b)
+for an example].
 %
@@
 %
-The second class of bandpass updating schemes
-is constructed from a sequence of orthogonal polynomials,
-which we will discuss below.
+The polynomial bandpass updating schemes will be discussed below.
 
 \subsection{\label{sec:bandpass_poly}
-Bandpass updating schemes constructed from orthogonal polynomial bases}
+Polynomial bandpass updating scheme}
 
@@
 
-\subsubsection{\label{sec:aus}
-Umbrella sampling with adaptive Gaussian ensembles}
+\subsubsection{\label{sec:age}
+Adaptive Gaussian ensemble}
 
 In analogy to the LLR algorithm,
-we will consider the case of sampling a Gaussian distribution
-$\rho(z) \propto \exp[-(z-z_c)^2/(2\sigma_z^2)]$,
+we will consider the case of sampling
+a Gaussian distribution
+%$\rho(z) \propto \exp[-(z-z_c)^2/(2\sigma_z^2)]$,
+\begin{equation}
+  \rho(z)
+  =
+  \frac{1}{\sqrt{2\pi} \sigma_z}
+  \exp\left[ - \frac{(z-z_c)^2}{2\,\sigma_z^2} \right]
+  ,
+  \label{eq:rho_Gaussian}
+\end{equation}
+which is \emph{not} flat,
 with a basis constructed from
@@
 By Eq. \eqref{eq:uz_decomp}, we have
-$$
+\begin{equation}
   u(z) = c_0 + \frac{c_1}{\sigma_z} (z - z_c)
-+ \sqrt 2 \, c_2 \left[ \frac{(z-z_c)^2}{2 \, \sigma_z^2} - \frac{1}{2} \right].
-$$
+  + \sqrt 2 \, c_2 \left[ \frac{(z-z_c)^2}{2 \, \sigma_z^2} - \frac{1}{2} \right]
+  .
+  \label{eq:u_age}
+\end{equation}
 Thus, we can interpret
@@
 
+\subsubsection{\label{sec:st}
+Expanded ensemble}
+
 We can extend the sampling range
@@
 $\rho_i = 1/m$,
-whereas the conditional intra-umbrella distribution, $\rho^{(i)}(z)$
-is the Gaussian given previously.
+and the conditional intra-umbrella distribution, $\rho^{(i)}(z)$,
+is the Gaussian given in Eq. \eqref{eq:rho_Gaussian}. % previously.
 %
@@
 an independent bias potential, $u^{(i)}(z)$,
-of the form of Eq. \eqref{eq:uz_decomp},
-and update coefficients as
+of the form of Eq. \eqref{eq:u_age},
+and update the coefficients as
 \begin{align*}
   c_k^{(i)}(t+1) = c_k^{(i)}(t)
-  + m \, \alpha(t) \, R_k^{(i)}\bigl( z(t) \bigr) \, \delta_{i, i(t)}
-  ,
+  + \frac{ \alpha(t) } { \rho_i } \, R_k^{(i)}\bigl( z(t) \bigr) \, \delta_{i, i(t)}
+  .
 \end{align*}
 %
-where the factor $m$ before $\alpha(t)$
-is due to the inverse of the marginal inter-umbrella distribution, $\rho_i$.
-%
-Note that since $R_0^{(i)}(z) = 1$,
+%where the factor $m$ before $\alpha(t)$
+%is due to the inverse of the marginal inter-umbrella distribution, $\rho_i$.
+%
+Since $R_0^{(i)}(z) = 1$,
 the parameter, $c_0^{(i)}$,
@@
 and it is updated according to the single-bin scheme\cite{li2007}.
-%
+
 In the expanded ensemble,
@@
 and switch to the inverse-time schedule
-once the updating magnitude from the WL approach
+once the updating magnitude from the WL strategy
 falls under the value given by the inverse-time formula.\footnote{We
@@
   %
-Since the WL strategy uses the histogram flatness
-as a criterion for stage transitions,
-we will redefine the flatness
-from the fluctuation given by Eq. \eqref{eq:maxFk}
-based on the eigenmode decomposition
-(this allows a uniform definition for
-both the continuous intra-umbrella
-and the discrete inter-umbrella components,
-cf. Appendix \ref{sec:hfluc}).
+However, we need to redefine the measure of histogram fluctuation
+used in the WL strategy for stage transitions
+by Eq. \eqref{eq:F2st}
+to include the intra-umbrella fluctuations.
 %
@@
   Set up an array of umbrellas covering the desired range,
-    e.g. we may set
-    $z_c^{(i)} = z_{\min} + i \Delta z$
-    with the same width $\sigma_z$.
-  %
-  For every $i$,
-    set $c_1^{(i)}$ to the estimated mean force,
-    and set $c_0^{(i)} = c_2^{(i)} = 0$.
+    e.g. we may set, for $i = 1, \dots, m$,
+    $z_c^{(i)} = z_{\min} + (i - 1) \, \Delta z$,
+    the width, $\sigma_z^{(i)}$, to be a constant.
+    For the parameters, we can set
+    $c_1^{(i)}$ to the estimated mean force,
+    and $c_0^{(i)} = c_2^{(i)} = 0$.
 
@@
 \item
-  Conduct a regular MC or MD step
+  Perform a regular MC or MD step
   in the current umbrella $i$ under bias potential $u^{(i)}(z, t)$.
-  Update the coefficients, $c_k^{(i)}$'s,
-  according to Eq. \eqref{eq:ckupdate}.
 
@@
   Attempt a transition to another umbrella
-  with the acceptance probability given by Eq. \eqref{eq:transprob_st}.
+  with the acceptance probability given
+  by Eq. \eqref{eq:transprob_st}.
+  %
+  The above intra-umbrella sampling steps
+  and inter-umbrella transition steps
+  can also be applied in random order
+  with different frequencies.\cite{manousiouthakis1999}
+
+\item
+  Update the $c_k^{(i)}$'s
+  according to Eq. \eqref{eq:ckupdate}.
 
@@
 After simulation,
-the weighted histogram analysis method
+we can use the weighted histogram analysis method
 (WHAM)\cite{ferrenberg1988, *ferrenberg1989,
@@
 *fenwick2008, *kim2011, *shirts2008,
-zhu2012, newman, frenkel},
-can be used to recover the unbiased distribution,
+zhu2012, newman, frenkel}
+to recover the unbiased distribution,
 %
@@
 to the umbrella $i$.
-%
-We can use the final converged values of $Z^{(i)}$'s
-without going through the usual WHAM iterations.
 
@@
 %
-The PMF can be derived from
+The PMF is the negative logarithm of
 the intrinsic distribution of the distance
 between the two tagged particles, $p(r)$,
-as $-\ln p(r)$,
-%
-Formally, we have
-\begin{align*}
-  p(r) = \langle \delta(\mathbf r_{12} - \mathbf r) \rangle
-  =
-  \frac{V}{Z}
-  \int e^{-\beta \, \Phi(\mathbf r^N)}
-  d\mathbf r_3 \cdots d\mathbf r_N
-  ,
-\end{align*}
-where
-$V$ is the reduced volume,
-$\Phi(\mathbf r^{N}) = \sum_{j=3}^N\sum_{i=1}^{j-1} \varphi(r_{ij})$,
-and
-$Z = \int e^{\beta \Phi(\mathbf r^N)} d\mathbf r_1 \cdots d\mathbf r_N$.
-%
-This distribution is proportional to the cavity distribution function\cite{hansen}.
+which is proportional to the cavity distribution function\cite{hansen}.
+%
+%Formally, we have
+%\begin{align*}
+%  p(r) = \langle \delta(\mathbf r_{12} - \mathbf r) \rangle
+%  =
+%  \frac{V}{Z}
+%  \int e^{-\beta \, \Phi(\mathbf r^N)}
+%  d\mathbf r_3 \cdots d\mathbf r_N
+%  ,
+%\end{align*}
+%where
+%$V$ is the reduced volume,
+%$\Phi(\mathbf r^{N}) = \sum_{j=3}^N\sum_{i=1}^{j-1} \varphi(r_{ij})$,
+%and
+%$Z = \int e^{\beta \Phi(\mathbf r^N)} d\mathbf r_1 \cdots d\mathbf r_N$.
+%%
+%This distribution is proportional to the cavity distribution function\cite{hansen}.
 %as $y(r) = \left(1 - \frac{1}{N}\right) \, V \, p(r)$\cite{hansen}.
 
-We used the Metropolis algorithm\cite{frenkel, metropolis1953}
-for configuration sampling.
-%
-Each step consists of
-one MC trial for displacing one of the two tagged particles,
-and another four trials for displacing the other particles.
-%
-In each trial,
-a random particle is displaced in each spatial dimension
-according to a uniform distribution over $(-\varepsilon, \varepsilon)$.
-%
-The two tagged particles are fixed on the $x$-axis,
-and are only displaced in this direction.
-%
-We used $\varepsilon = 0.1/\varrho$
-to ensure reasonable acceptance ratio,
-and $\varrho = N/V$ is the reduced number density.
-
-We used adaptive FDS to sample a flat distribution along $r$,
-and calculated the PMF from $r = 0$ to half of the box size
-%(also the cutoff distance of the potential)
+We used adaptive FDS to calculate
+the PMF from $r = 0$ to half of the box size
 on a grid of size $\delta r = 0.01$ in reduced units.
 %
-Two densities were used, $\varrho = 0.1$ and $0.8$,
-and the numbers of bins were $n = 500$ and $250$, respectively.
-%
+The density was $\varrho = 0.1$.
+The numbers of bins was $n = 500$.
 The reduced temperature was $T = 3$.
 %
-The bias potential, $k_B T \, u(r)$,
+The bias potential, $u(r)$,
 was initially set to zero and
@@
 of $10^7$ steps under a constant updating magnitude,
-$a_0 = 10^{-4}$,\footnote{The
-  seemingly small updating magnitude
-  corresponds to a larger value of $\ln f = n \, a_0$
-  in the WL algorithm.
-  The $\ln f$ was $0.05$ for $\varrho = 0.1$,
-  or $0.025$ for $\varrho = 0.8$.}
+$a_0 = 10^{-4}$
+(corresponding to $\ln f = n \, a_0 = 0.05$ in the WL algorithm),
 followed by a production phase of
-another $10^7$ steps (unless indicated otherwise)
-with updating magnitude given by
+$10^8$ steps
+with the updating magnitude given by
 either the inverse-time or optimal schedule.
@@
 %
-To ensure the same schedule was used for multiple runs under the same condition,
+To ensure the same schedule was used for multiple runs,
 the optimal schedule was pre-computed using a set of fixed
-$\Gamma_k$'s and $\epsilon_k$'s derived from the longer simulation,
-although ideally they could be estimated from the data
-collected from the equilibration phase in individual runs.
+$\Gamma_k$'s and $\epsilon_k$'s derived from the longer simulation.
+
+We used the Metropolis algorithm\cite{frenkel, metropolis1953}
+for configuration sampling.
+%
+Each step consists of
+one MC trial for displacing one of the two tagged particles,
+and another four trials for displacing the other particles.
+%
+In each trial,
+a random particle is displaced in each spatial dimension
+according to a uniform distribution over $(-\varepsilon, \varepsilon)$.
+%
+The two tagged particles are fixed on the $x$-axis,
+and are only displaced in this direction.
+%
+We used $\varepsilon = 0.1/\varrho$
+to ensure reasonable acceptance ratio.
 
@@
 in the single-bin and bandpass updating schemes,
-we will focus on the Gaussian updating scheme below
+we will focus on the Gaussian updating scheme with width $\sigma = 0.2$ below
 in discussing the optimal schedule.
 
-\begin{table}[h]\footnotesize
+\begin{table}[h]
   \caption{\label{tab:lj_error}
-    Error of the bias potential under
+    Errors of the bias potential under
     the single-bin,
-    Gaussian (labeled by the width, $\sigma$),
-    and bandpass (labeled by the mode cutoff, $K$)
+    Gaussian (with width $\sigma = 0.2$),
+    and homogeneous bandpass (with mode cutoff $K = 20$)
     updating schemes
@@
     and the predicted values are shown in parentheses for comparison;
-    $\Err_\mathrm{init.}$
-    denotes the error
-    at the end of the equilibration phase under a constant magnitude;
-    $\Err_\mathrm{final}$
-    denotes the error at the end of the production phase under
-    either the inverse-time ($1/t$) or optimal schedule (opt.).
+    $\Err_\mathrm{init.}$ and $\Err_\mathrm{final}$
+    denote the errors at the beginning and end of the production phase, respectively.
     %
     \note{The reference values can be found
-    on the second line of \texttt{verr.log}.}%
+    on the second line of \texttt{verr.log}.
+    To check use \texttt{getav.py} under \texttt{prog/lj},
+    which will show the simulation and prediction values.}%
   }
@@
     \hline
-    Scheme & Sched. &
+    Scheme & Schedule &
     $\Err_\mathrm{init.}$ &
@@
     \\
+    %\hline
+    %\multicolumn{4}{c}{
+    %  $\varrho = 0.1$,
+    %  $500$ bins,
+    %  $\Gamma_1 \approx 111$,
+    %  $\Gamma \approx 1.1\times10^3$.
+    %} \\
     \hline
-    \multicolumn{4}{c}{
-      $\varrho = 0.1$,
-      $500$ bins,
-      $\Gamma_1 \approx 111$,
-      $\Gamma \approx 1.1\times10^3$.
-    } \\
-    \hline
-    single-bin & $1/t$
+    Single-bin & $1/t$
     & $5.7(5.7)\times10^{-2}$
-    & $9.2(11.4)\times10^{-5}$
+    & $9.0(11.4)\times10^{-6}$
     \\
     %\hline
-    $\sigma=0.1$ & $1/t$
-    & $7.5(8.9)\times10^{-3}$
-    & $6.8(8.2)\times10^{-5}$
+    Gaussian & $1/t$
+    & $6.7(7.8)\times10^{-3}$
+    & $1.8(2.6)\times10^{-5}$
     \\
     %\hline
-    $\sigma=0.1$ & opt.
-    & $7.5(8.9)\times10^{-3}$
-    & $3.3(3.9)\times10^{-5}$
+    Gaussian & optimal
+    & $6.7(7.8)\times10^{-3}$
+    & $2.9(3.7)\times10^{-6}$
     \\
     %\hline
-    $\sigma=0.2$ & $1/t$
-    & $6.6(7.8)\times10^{-3}$
-    & $3.9(4.8)\times10^{-5}$
-    \\
-    %\hline
-    $\sigma=0.2$ & opt.
-    & $6.7(7.8)\times10^{-3}$
-    & $2.4(2.9)\times10^{-5}$
-    \\
-    %\hline
-    $\sigma=0.5$ & $1/t$
-    & $5.4(6.6)\times10^{-3}$
-    & $3.3(3.8)\times10^{-5}$
-    \\
-    %\hline
-    $\sigma=0.5$ & opt.
-    & $5.8(6.6)\times10^{-3}$
-    & $2.7(3.0)\times10^{-5}$
-    \\
-    %\hline
-    $K=20$ & $1/t$
-    & $8.0(8.9)\times10^{-3}$
-    & $1.6(1.8)\times10^{-5}$
-    \\
-    \hline
-    \multicolumn{4}{c}{
-      $\varrho = 0.8$,
-      $250$ bins,
-      $\Gamma_1 \approx 2.0\times10^4$,
-      $\Gamma \approx 2.7\times10^4$
-    } \\
-    \hline
-    single-bin & $1/t$
-    & $1.4(1.4)$
-    & $3.4(2.7)\times10^{-3}$
-    \\
-    %\hline
-    $\sigma=0.1$ & $1/t$
-    & $1.3(1.3)$
-    & $3.0(2.8)\times10^{-3}$
-    \\
-    %\hline
-    $\sigma=0.1$ & opt.
-    & $1.2(1.3)$
-    & $3.0(2.8)\times10^{-3}$
-    \\
-    %\hline
-    $\sigma=0.2$ & $1/t$
-    & $1.3(1.2)$
-    & $3.5(3.2)\times10^{-3}$
-    \\
-    %\hline
-    $\sigma=0.2$ & opt.
-    & $1.2(1.2)$
-    & $3.1(2.9)\times10^{-3}$
-    \\
-    %\hline
-    $\sigma=0.5$ & $1/t$
-    & $1.0(0.9)$
-    & $9.9(8.0)\times10^{-3}$
-    \\
-    %\hline
-    $\sigma=0.5$ & opt.
-    & $1.0(0.9)$
-    & $5.9(5.4)\times10^{-3}$
-    \\
-    %\hline
-    $K=20$ & $1/t$
-    & $1.4(1.3)$
-    & $3.1(2.7)\times10^{-3}$
+    Bandpass & $1/t$
+    & $8.3(8.9)\times10^{-3}$
+    & $2.1(2.3)\times10^{-6}$
     \\
@@
 As shown in Fig. \ref{fig:lj_alpha},
-the optimal schedules for the Gaussian updating schemes
-differed from inverse-time one,
+the optimal schedule for the Gaussian updating scheme
+was more complex than the inverse-time one,
 and the former tended to give a larger updating magnitude.
 %
-The optimal schedule was more effective
-than the sub-optimal inverse-time schedule
-in reducing the error,
-as shown in Table \eqref{tab:lj_error}.
-%
-Although the advantage of the former
-was modest in the reported cases,
-it would increase with the simulation length.
-%
-For example, for the case of
-$\varrho = 0.1$ and Gaussian width $\sigma = 0.2$,
-if we extend the simulation to $10^8$ steps,
-the error for the optimal schedule was
-$2.9\times10^{-6}$,
-which was much smaller than that
-for the inverse-time schedule
-$1.8\times10^{-5}$.
-  \note{The predicted errors were
-  $3.7\times10^{-6}$ and $1.8\times10^{-5}$, respectively.}
+Since our equilibration run was sufficiently long,
+the initial updating magnitude, $\alpha(0)$,
+was around the ideal value of $a_0/2 = 5\times10^{-5}$
+given by Eq. \eqref{eq:half_alpha0}.
 
@@
     \label{fig:lj_alpha}
-    Optimal schedules of Gaussian updating schemes
-    for the LJ system.
-    %
-    The starting updating magnitude was usually
-    around $a_0/2 = 5\times10^{-5}$
-    as given by Eq. \eqref{eq:half_alpha0}.
-    But an overly wide window ($\sigma = 0.6$)
-    led to a much larger updating magnitude.
+    Optimal schedule of the Gaussian updating scheme.
   }
@@
 
-For the same case,
+The optimal schedule was more effective
+than the inverse-time schedule
+in reducing the error,
+as shown in Table \ref{tab:lj_error}.
+%
+For this case,
 we further decomposed the error
@@
 by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl},
-and ploted the error components, $\langle \tilde v_k^2 \rangle$,
+and plotted the error components, $\langle \tilde v_k^2 \rangle$,
 in Fig. \ref{fig:lj_xerr}.
@@
     \label{fig:lj_xerr}
-    Error components from FDS simulations
-    using the Gaussian updating scheme
-    of width $\sigma = 0.2$
-    on the LJ system at density $\varrho = 0.1$
+    Error components, $\tilde u_k^2$'s, of the PMF
     after $10^8$ production steps.
     %
-    The optimal schedule (opt.) was more effective
-    in reducing the errors of middle-range eigenmodes
+    The optimal schedule more effective
+    reduced the errors of middle-range eigenmodes
     than the inverse-time schedule ($1/t$).
     The ``corrected'' values were obtained from
-    Eq. \eqref{eq:uhatt}.
+    Eq. \eqref{eq:uhatav}.
     The lines are a guide to the eyes.
@@
 
-Next, we discuss the applicability of
-the Gaussian updating scheme.
-%
-As shown in Table \ref{tab:lj_error},
-the Gaussian updating schemes,
-or more generally the bandpass updating schemes,
-produced less error than the single-bin scheme
-only in the low-density $\varrho = 0.1$ case,
-but not in the high-density $\varrho = 0.8$ case.
-%
-We believe that the difference lies in the sampling difficulty.
-%
-In the low-density case,
+For the Gaussian updating scheme,
+the error depended not only the schedule
+but also the width of the Gaussian.
+%
+Although widening the Gaussian
+can suppress local noise,
+it can also mask finer features of the PMF.
+%it also diminishes the eigenvalues
+%[by Eq. \eqref{eq:lambda_Gaussian}]
+%of more middle-range modes
+%(which often represent finer features of the PMF)
+%and hence slows down the error reduction there
+%[by Eq. \eqref{eq:vt_diffeq_mbin}].
+%
+Thus, there is an optimal width that balances the two factors.
+%
+According to the analytical prediction
+(assuming the optimal schedule),
+the optimal width was around $\sigma = 0.29$ in this case,
+and it would decrease with the simulation length
+and approach the limit of single-bin updating scheme,
+$\sigma \to 0$, asymptotically.
+
+The homogeneous bandpass updating scheme
+yielded less error than the Gaussian updating scheme,
+as shown in Table \ref{tab:lj_error}.
+%
+Similar to the Gaussian case,
+the mode cutoff, $K$, has an optimal value,
+which was around $26$ in this case.
+%
+The optimal cutoff would increase with the simulation length
+and approach the limit of single-bin scheme, $K \to n$,
+asymptotically.
+
+The Gaussian and homogeneous bandpass updating schemes
+helped reduce the error in this case because
 the sampling along the collective variable
-was free from major barriers,
-and thus the histogram fluctuation
-was largely local in nature.
-%
-In the high-density case, however,
-the error was dominated by the first few eigenmodes
-that correspond to the long-range fluctuations
-(as $\Gamma_1 \approx \Gamma$),
-and the reduction of local noise by multiple-bin schemes
-had little effect on the total error.
-%
-When local noise is dominant,
-we found that the bandpass updating scheme
-(with the mode cutoff $K = 20$)
-was more effective than the Gaussian updating schemes
-as expected.
-
-It is worth mentioning that the width $\sigma$
-of a Gaussian updating scheme
-can critially affect its effectiveness.
-%
-With a overly wide Gaussian,
-the updating scheme would fail to
-capture finer features of the PMF.
-%
-From Eq. \eqref{eq:lambda_Gaussian},
-we know that increasing the width of the Gaussian
-would diminish the eigenvalue, $\lambda_k$,
-which is proportional to the rate of error reduction
-by Eq. \eqref{eq:vt_diffeq_mbin}.
-%
-Thus, although widening the Gaussian can
-effectively suppress random error
-from local histogram fluctuations,
-it also increased the risk of leaving
-more persistent bias in the middle-range modes.
-%
-For the $\varrho = 0.1$ case,
-the prediction showed that
-assuming the optimal schedule,
-the optimal width was around $\sigma = 0.37$,
-and it would decrease with the simulation lengthen.
-%
-Similarly, the optimal mode cutoff, $K$,
-for the homogeneous bandpass updating scheme
-would increase with the simulation length.
-%
-For very long simulations, we expect both to reach
-the limit of single-bin updating scheme
-($\sigma \to 0$ or $K\to n$).
-
-An overly large width $\sigma$ can
-also induce a very large initial updating magnitude
-under the optimal schedule.
-%
-The initial updating magnitude, $\alpha(0)$,
-was usually around
-the ideal value of $a_0/2 = 5\times10^{-5}$
-given by Eq. \eqref{eq:half_alpha0},
-for a sufficiently long preliminary run
-under the constant updating magnitude, $a_0$,
-as shown in Fig. \ref{fig:lj_alpha}.
-%
-But if the Gaussian were too wide,
-e.g. if $\sigma = 0.6$ in the $\varrho = 0.8$ case,
-the eigenvalues $\lambda_k$ of some modes
-became too small to damp out the initial bias,
-resulting in a much larger $\alpha(0)$
-[cf. Eq. \eqref{eq:opt_alpha0}].
-%
-In this case, the predicted optimal schedule
-becomes invalid as the resulting deviation in the bias potential
-is no longer small.
-
-In practice, both the width and schedule
-of the Gaussian updating scheme are often sub-optimal.
-%
-We can use the histogram-based correction formula,
+was free from major barriers
+such that the histogram fluctuation was mainly local in nature.
+%
+For similar simulations performed at a high density, $\varrho = 0.8$,
+we found the gain of using the above two updating schemes
+over the single-bin updating scheme, if any, to be negligible.
+
+In practice, updating schemes and schedules
+are often sub-optimal.
+%
+We can, however, use the histogram-based correction formula,
 Eq. \eqref{eq:uhatav}, after simulation
-to remove the remaining bias in PMF.
+to improve the estimated PMF.
 %
@@
 %
-However, this correction also introduces
+While this correction also introduces
 some local noise in the large $k$ modes,
-although these noise are largely harmless
-in a post-simulation analysis, since
+the noise is likely harmless
+in a post-simulation analysis, as
 the bias potential no longer needs to be differentiated
@@
 \subsection{\label{sec:potts}
-Umbrella sampling with adaptive Gaussian ensembles}
+Adaptive Gaussian ensemble}
 
 As another proof of principle,
-we tested the adaptive umbrella sampling method
-introduced in Sec. \ref{sec:aus}
+we tested the adaptive Gaussian ensemble method
+introduced in Sec. \ref{sec:age}
 on the two-dimensional $L\times L$ ten-state
@@
 %
-For simplicity,
-only transitions between neighboring umbrellas
-were performed.
+In attempting inter-umbrella transitions,
+the target umbrella was chosen from
+any of the other umbrellas.
 %
 Initially, we set
-$c_1 = \tilde \beta_c \, \sigma_E$
+$c_1^{(i)} = \tilde \beta_c \, \sigma_E$
 for all umbrellas,
@@
 (which is more efficient than the Metropolis algorithm)
-at $\beta = c_1/\sigma_E$
+at the inverse temperature, $\beta = c_1/\sigma_E$,
 followed by a Metropolis step with
@@
 the initial updating magnitude was $0.01$;
-the fluctuation threshold for stage transition was $0.2$;
+the fluctuation threshold for stage transitions
+was $F = 0.2$ [using the definition given by Eq. \eqref{eq:F2st}];
 and the reduction factor was $1/2$.
@@
 %
-The Gaussian histograms were evenly spaced with similar width as intended,
+The Gaussian histograms were evenly spaced with similar widths as intended,
 %and the overall energy histogram was roughly flat,
@@
 %
-As discussed in Sec. \ref{sec:aus},
+As discussed in Sec. \ref{sec:age},
 $c_1/\sigma_E$ roughly represents
 the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
-and it demonstrated
+and it manifested 
 the backbending behavior\cite{kim2006, *kim2007, kim2010}
@@
 was indeed above $1.0$
-in the critical region\cite{kim2010},
+in the unstable region,
+where $\beta(E)$ is increasing or, equivalently,
+$\ln g(E)$ is locally convex,
 as shown in Fig. \ref{fig:pt_hist}(c).
@@
     %
-    (b) The parameter, $c_1/\sigma_E$ (squares), roughly corresponding to
-    the statistical temperature, $\beta(E)$,
-    %(dashed line estimated from the smoothed density of states)
-    manifested the backbending behavior
-    typical in a first-order phase transition.
-    %
-    (c) The parameter, $\sqrt 2 \, c_2$ (squares),
-    exceeded $1.0$
-    in the critical region, indicating local convexity
-    of the density of states.
+    (b) $c_1^{(i)}/\sigma_E$, roughly corresponding to
+    the statistical temperature, $\beta(E)$.
+    %
+    (c) $\sqrt 2 \, c_2^{(i)}$
+    exceeded $1.0$ in the unstable region,
+    where $\beta(E)$ is increasing.
   }
@@
 the different rates of error reduction of the eigenmodes,
-is usually not given by a simple closed form,
+is usually not given in a simple closed form,
 but is implicitly given by an equation of motion
@@
 
-In particular,
-the Gaussian updating scheme has a nontrivial optimal schedule
-and its width needs to be chosen carefully.
-%
-With a set of rapidly diminishing eigenvalues,
-the Gaussian updating scheme suppresses
-the accumulation of local noise in the bias potential
-%maintain the smoothness of the bias potential,
-at the cost of slowing down the reduction
-of the error in the middle-range eigenmodes.
-%
-Fortunately, the efficiency loss in
-reducing the error of the long- and middle-range eigenmodes
-can often be compensated
-by the histogram-based correction formula, Eq. \eqref{eq:uhatav},
-which ideally would remove the persistent error
-in these eigenmodes.
-%and recover a bias potential
-%similar to that from a simulation
-%using the single-bin updating scheme.
+%In particular,
+%the Gaussian updating scheme has a nontrivial optimal schedule
+%and its width needs to be chosen carefully.
+%%
+%With a set of rapidly diminishing eigenvalues,
+%the Gaussian updating scheme suppresses
+%the accumulation of local noise in the bias potential
+%%maintain the smoothness of the bias potential,
+%at the cost of slowing down the reduction
+%of the error in the middle-range eigenmodes.
+%%
+%Fortunately, the efficiency loss in
+%reducing the error of the long- and middle-range eigenmodes
+%can often be compensated
+%by the histogram-based correction formula, Eq. \eqref{eq:uhatav},
+%which ideally would remove the persistent error
+%in these eigenmodes.
+%%and recover a bias potential
+%%similar to that from a simulation
+%%using the single-bin updating scheme.
 
@@
 the single-bin updating scheme to be optimal
-in error reduction in the long time limit
-without a priori assumptions
-on the smoothness of the PMF.
+in error reduction in the long time limit.
+%without a priori assumptions
+%on the smoothness of the PMF.
 %
@@
 a class of bandpass updating schemes,
-which are also asymptotically optimal,
-but completely filter the noise modes.
+which discard a few noise modes
+and are asymptotically optimal for the non-noise modes.
 %
@@
 %
-This simplification is justified in the asymptotic regime:
-when the updating magnitude is small
-and the sampling is a quasi-equilibrium one\cite{
+This is justified only in the asymptotic regime,
+where the updating magnitude is small
+and the sampling is a quasi-equilibrium one.\cite{
   zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
-the random error,
-which is roughly proportional to
-the square root of the updating magnitude\cite{
-  zhou2005, morozov2007, zhou2008, bussi2006},
-can easily outweigh
-the systematic error,
-which is roughly proportional to
-the magnitude itself\cite{morozov2007}.
-%
-%Our method of analysis is most useful for long simulations
-%to achieve relatively precise results.
-%
-Second, by using the white noise approximation,
-we have represented the autocorrelation function
-of each eigenmode by a single number, $\Gamma_k$.
-%
-Although this assumption appeared to work well
-in our test systems modeling long simulations,
-it may break down for relatively short simulations
-of glassy systems.
-%
-Finally, the adaptive umbrella sampling method
-introduced in Sec. \ref{sec:aus}
-requires refinement to improve
-its general applicability.
-%
-For example, we find that with slow sampling processes,
-the updating scheme might occasionally overly update
-the parameter, $c_2$, leading to unreasonable values,
-and the parameters used here might not be optimal.
+%the random error,
+%which is roughly proportional to
+%the square root of the updating magnitude\cite{
+%  zhou2005, morozov2007, zhou2008, bussi2006},
+%can easily outweigh
+%the systematic error,
+%which is roughly proportional to
+%the magnitude itself\cite{morozov2007}.
+%
+Second, the white noise approximation
+can break down for relatively short simulations.
+%
+Finally, the adaptive Gaussian ensemble method
+introduced in Sec. \ref{sec:age} can be refined
+to improve its general applicability and stability.
+%e.g. to prevent the updating scheme from
+%overly updating the parameters.
 %Our immediate interest is to apply these relations to
@@
 for the single-bin updating scheme
-by a runtime average interpretation.
+by a moving average interpretation.
 %
@@
 renders the bias potential
-as a runtime average with a dynamic weight
+as a cumulative moving average (CMA) with a dynamic weight
 determined by the updating schedule.
@@
 the inverse-time schedule is optimal
-because it corresponds to an unweighted average.
+because it corresponds to an equally-weighted average.
 
@@
 For example,
-the unweighted average, $\omega(t) = \omega(0)/t_0$ (for $t > 0$),
+the equally-weighted average, $\omega(t) = \omega(0)/t_0$ (for $t > 0$),
 corresponds to the inverse-time updating magnitude,
@@
 Updating magnitudes above and below this value
-correspond to runtime averages
-with growing and diminishing weights, respectively.
+are typical in CMAs of
+growing and diminishing weights, respectively.
 
@@
 the instantaneous histogram, $h_i(t)$,
-we get an independent estimate of the bias potential
-made by a single data point at step $t$,
+we get an independent estimate of the ideal bias potential
+from a single data point at step $t$,
 %
@@
 Comparing this to Eq. \eqref{eq:X_recur},
-we find that $u_i(t+1)$
-is equivalent to a runtime average $X_\omega(t)$
-of $\hat u_i(t)$
-with the weight given by Eq. \eqref{eq:omega_from_alpha}.
+we find that the time evolution of $u_i(t+1)$
+is the same as that of the CMA $X_\omega(t)$
+of $x(t) = \hat u_i(t)$.
+%
+Thus, the bias potential can be thought as
+a CMA with the weight given by Eq. \eqref{eq:omega_from_alpha}.
 %
 The inverse-time schedule is optimal
-because it corresponds to the unweighted average
-that uses the data most efficiently.
+because it corresponds to the equally-weighted average,
+which is most efficient statistically.
 
 Similarly, for the multiple-bin updating scheme,
-we have, from Eq. \eqref{eq:vkupdate},
+we have, from Eqs. \eqref{eq:v_def} and \eqref{eq:vkupdate},
 ${\tilde u}_k(t + 1) = {\tilde u}_k(t) + \lambda_k \, \alpha(t) \, {\tilde f}_k(t)$,
@@
 for $k > 0$.
-This means that
-$\tilde u_k(t+1)$ can also be cast
-as a runtime average of $\hat{\tilde u}_k(t)$,
-with a weight corresponding to the effective updating magnitude,
-$\lambda_k \, \alpha(t)$ [instead of $\alpha(t)$].
-
-\subsection{Upper bound of the updating magnitude}
-
-The runtime average interpretation imposes
+Thus,
+\begin{align*}
+  \tilde u_k(t+1) = \tilde u_k(t) + \lambda_k \, \alpha(t) \,
+  \left[
+    \hat{\tilde u}_k(t) - \tilde u_k(t)
+  \right]
+  ,
+\end{align*}
+%
+which means that
+$\tilde u_k(t+1)$ is equivalent to
+a CMA of $\hat{\tilde u}_k(t)$,
+and the weight is given by Eq. \eqref{eq:omega_from_alpha},
+with $\alpha(t)$ replaced by the effective updating magnitude,
+$\lambda_k \, \alpha(t)$.
+
+\subsection{\label{sec:ave_ub}
+Upper bound of the updating magnitude}
+
+The moving average interpretation imposes
 an upper bound on the updating magnitude.
@@
 of any mode $k > 0$
-should never exceed $1.0$,
-which is not necessarily satisfied
+should not exceed $1.0$.
+But this is not necessarily satisfied
 for the updating schedule
@@
 of the largest eigenvalue
-would guarantee $\lambda_k \, \alpha(t) \ge 1$
+guarantees $\lambda_k \, \alpha(t) \ge 1$
 for all nontrivial modes.
@@
 each mode of bias potential
-corresponds to a runtime average
-with a dynamic weight,
-which is not necessarily the optimal even weight.
-%
-Below we give a simple correction that
-could largely compensate the efficiency loss.
-%
-We will use capital letters to denote unweighted averages.
+corresponds to a CMA with a dynamic weight,
+which is not necessarily the optimal equal weight.
+%
+Below we give a simple correction
+to recover the equally-weighted average.
+%
+We will use capital letters to denote equally-weighted averages.
 %e.g. for variable $y$,
@@
 %
-Then, by an unweighted average of Eq. \eqref{eq:uhatt}, we get
+Then, by averaging both sides of Eq. \eqref{eq:uhatt}, we get
 \begin{align}
@@
   derive Eq. \eqref{eq:uhatav}
-  from the self-healing umbrella sampling method\cite{marsili2006}.
-  We can construct an independent estimator
-  of the unbiased distribution from Eq. \eqref{eq:Vi_target} as
+  from constructing an independent estimator
+  of the unbiased distribution\cite{marsili2006},
+  which is, by Eq. \eqref{eq:Vi_target},
   $\hat p_i(t) \approx h_i(t) \, e^{ u_i(t) }$,
@@
 would fail if some eigenvalue, $\lambda_k$, is close to zero.
-So it is best used for the single-bin updating scheme.\footnote{Alternatively,
-  we can split the trajectory into blocks,
-  compute $\Gamma_k$ from the histogram fluctuation of each block
-  using Eq. \eqref{eq:Fk2_Gammak},
-  and then average the results.}
-
-We can alternatively estimate the magnitude of $\Gamma_k$
-by the transition matrix method.
-%
-If the propagation along the collective-variable is Markovian,
-then in the long time limit, the evolution can be approximately
+So it is best used for the single-bin updating scheme.%\footnote{Alternatively,
+%  we can split the trajectory into blocks,
+%  compute $\Gamma_k$ from the histogram fluctuation of each block
+%  using Eq. \eqref{eq:Fk2_Gammak},
+%  and then average the results.}
+
+Below we use the transition matrix method
+to estimate $\Gamma_k$ in a model for local diffusion\cite{bussi2006}.
+%
+We assume that the propagation along the collective-variable is Markovian
+such that the evolution can be approximately
 described by a transition matrix, $\mathbf A$, with
@@
 \end{align*}
-Further if the transition matrix satisfies detailed balance
-with a set of eigenvectors, $\psi_{li}$,
-that are orthonormal with respect to $\pmb\rho$
+Further, if the transition matrix % satisfies detailed balance
+has a full set of orthonormal eigenvectors, $\psi_{li}$
+%that are orthonormal with respect to $\pmb\rho$
 [cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
-\eqref{eq:eig_orthonormal_rows}]
-then we have the decomposition
+\eqref{eq:eig_orthonormal_rows}],
+it can be decomposed as
 $A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
@@
 \end{equation}
-
-The above method
-tends to underestimate
-the first few $\Gamma_k$'s.
-%
-We will nonetheless use it as a model to
-study the behavior of $\Gamma_k$
-in a local sampling process.
-%
-If we borrow Eq. \eqref{eq:lambda_Gaussian}
-for $\Lambda_k$,
-with $\sigma_z$ interpreted as
+%
+To model a local sampling process,
+we can borrow Eq. \eqref{eq:lambda_Gaussian} for $\Lambda_k$,
+with $\sigma_z$ reinterpreted as
 the average move size along $z$ during sampling,
@@
   \right].
-  \label{eq:Gamma_local}
+  \notag
+  %\label{eq:Gamma_local}
 \end{align}
-For small $k$,
-we have $\Gamma_k \propto \Delta z^2/k^2$,\cite{bussi2006}
-which means that doubling the sampling range, $\Delta z$,
-would make the histogram flattening
-at least four times as difficult,
-i.e. FDS methods are most effective
-for short ranges\cite{wang2001, wang2001pre}.
+Thus, we have $\Gamma_k \propto \Delta z^2/k^2$
+for small $k$,\cite{bussi2006}
+whereas $\Gamma_k \to 1$ as $k$ increases.
+%
+However, this method tends to underestimate
+the first few $\Gamma_k$'s.
+%which means that doubling the sampling range, $\Delta z$,
+%would make the histogram flattening
+%at least four times as difficult,
+%i.e. FDS methods are most effective
+%for short ranges\cite{wang2001, wang2001pre}.
 %
@@
 %
-For a modestly large $k$,
-$\Gamma_k$ quickly approaches $1$.
-But since $\langle \tilde v_k^2 \rangle \propto \Gamma_k$,
-according to Eq. \eqref{eq:xt2_eql},
-the error profile for the single-bin updating scheme
-would have a roughly flat tail,
-as shown in Fig. \ref{fig:lj_xerr}.
 
@@
 by a rigid window function
-that shifts with a visit to bin $j$
+that shifts with the current bin $j$
 (ignoring the boundary adjustment).
 %
-The window function is defined by
+The unadjusted window function is defined by
 the relative updating magnitude, $\mu_l$,
@@
 share the same set of eigenvectors,
-which are sine and cosine functions.
-%
-The eigenvalues of the updating matrix
+and the eigenvalues
 are related to the window function as a cosine transform.
@@
 For a periodic variable\cite{dama2014},
-the updating matrix, $w_{ij}$,
+the element of the updating matrix, $w_{ij}$,
 can only be a function of the difference of indices, $i-j$, modulo $n$,
@@
 %
-for some numbers $\mu_{-b}, \dots, \mu_b$, ($b \le n/2$)
-with $\mu_l = 0$ for $l > b$.
-%
-These numbers define a window function.
-%
-To satisfy Eq. \eqref{eq:w_sumj},
-we impose the normalization
+for some numbers, $\mu_l$'s, which
+define an unadjusted window function.
+%
+For simplicity, we will assume that
+$\mu_l$'s are symmetric, $\mu_l = \mu_{-l}$,
+and they can be cutoff at $l = b$,
+where $b$ is the integral part of $(n-1)/2$,
+i.e. $\mu_l = 0$ for $l > b$.
+%
+\note{The cutoff $b$ is a bit tricky to define
+in order to encompass the cases of even and odd $n$.}%
+%
+To satisfy Eq. \eqref{eq:w_sumj}, we further impose
 %
@@
   .
-\label{eq:mu_normalization}
-\end{equation}
-%
-If $b = n/2$, $\mu_{-b}$ is removed
-from the above sum in Eq. \eqref{eq:mu_normalization}
-to avoid double counting.
-%
-For simplicity, we also impose the symmetry
-%
-\begin{equation}
-  \mu_i = \mu_{-i}
-  .
-\label{eq:mu_symm}
+  \notag
+  %\label{eq:mu_normalization}
 \end{equation}
@@
   .
-\label{eq:wmul_to_convol}
+  \notag
+  %\label{eq:wmul_to_convol}
 \end{equation}
@@
 with $g = 1$.
-Note that there is a two-fold degeneracy,
+Note the two-fold degeneracy,
 %for the periodic variable,
@@
 For a non-periodic variable,
-we use the reflective boundary condition\cite{bussi2006},
-and change the form of the updating matrix to
-%
-%
-\begin{equation}
-  w_{ij}
-  =
-  \mu_{ i - j }
-  +
-  \mu_{ i + j - 1 }
-  +
-  \mu_{ i + j - 2 n - 1 }
-  ,
-  \notag
-  \label{eq:w_band_refl}
-\end{equation}
-%
-where the last two terms on the right hand side,
-representing two reflective mirrors at
-$i = 1/2$ and $i = n + 1/2$,
-are added to avoid unintended distortion\cite{dickson2011, mcgovern2013}
-to the equilibrium distribution\cite{bussi2006}.
-%such that Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance}
-%are satisfied.
-%
-Again, the window function should satisfy
-Eqs. \eqref{eq:mu_normalization}
-and \eqref{eq:mu_symm},
-but
-the cutoff $b$ only needs to be less than $n$.
-
-Note that the two reflective mirrors effectively
-define a periodic variable of period $2 \, n$.
-Thus, the eigenvalues can be borrowed from
-Eq. \eqref{eq:wband_eigenvalue},
-with a doubled period,
-i.e. $g$ is changed to $2$ in the non-periodic case.
-
-%For a non-periodic variable,
-We define the out-of-boundary values as
-%
-\begin{equation}
-  \phi_i
-  =
-  \begin{dcases}
-    \phi_{ 1 - i }           & \mathrm{for \;} i \le 0, \\
-    \phi_{ 2 \, n + 1 - i }  & \mathrm{for \;} i > n,
-  \end{dcases}
-\label{eq:phi_refl}
-\end{equation}
-%
-such that the matrix multiplication is still equivalent to
-a convolution as Eq. \eqref{eq:wmul_to_convol}.
-%
-Since Eq. \eqref{eq:phi_refl}
-defines a periodic variable of period $2 \, n$
-with a reflective symmetry around $i = 1/2$,
-the eigenvectors,
-$\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-should share the same periodicity,
-and be even functions about $i = 1/2$:
+we will use the reflective boundary condition\cite{bussi2006},
+which sets up two reflective mirrors at
+$i = 1/2$ and $i = n + 1/2$.
+%
+Then the variable and the images in the two mirrors
+define an effective periodic variable of period $2 \, n$.
+%
+Thus, we can borrow the eigenvalues from
+Eq. \eqref{eq:wband_eigenvalue}
+with a doubled period, i.e. $g = 2$,
+and redefine $b = n - 1$
+(so $\mu_n = 0$).
+%
+The eigenvectors sharing the same periodicity
+and being even about $i = 1/2$ is given by
 %
@@
 \end{equation}
+%
+The updating matrix can be reconstructed
+by Eq. \eqref{eq:w_from_phi},
+%
+%
+\begin{equation}
+  w_{ij}
+  =
+  \mu_{ i - j }
+  +
+  \mu_{ i + j - 1 }
+  +
+  \mu_{ i + j - 2 n - 1 }
+  .
+  \notag
+  %\label{eq:w_band_refl}
+\end{equation}
+%
+%where the last two terms on the right hand side
+%help prevent unintended distortion\cite{dickson2011, mcgovern2013}
+%to the equilibrium distribution\cite{bussi2006}.
 
@@
 
-We define a Gaussian updating scheme
-as an updating scheme with a
-Gaussian-shaped window function.
+The Gaussian updating scheme
+is defined by a Gaussian-shaped $\mu_l$.
 %
@@
 and
-$\Delta z = b \delta z$
-for the largest bin cutoff $b = gn/2$.
+$\Delta z = b \, \delta z$
+with $b \approx g\,n/2$.
 %
@@
 
-If the width of the Gaussian window, $\sigma_z$,
+If the width of the Gaussian, $\sigma_z$,
 is much less than $\Delta z$,
@@
     \label{fig:mat}
-    Updating matrices, $w_{ij}$, for
+    Updating matrices for
     homogeneous updating schemes of
     a non-periodic variable ($n = 500$).
-    Here $j$ is the bin of the current collective variable,
-    and $i$ is the bin where the bias potential is updated.
+    Here $j$ is the current bin,
+    $i$ is the bin where updating occurs,
+    and $w_{ij}$ is the relative updating magnitude.
     %
@@
 Then, $\mu_l = \mathrm{sinr}(K, l, n)$ and $w_{ij} = \mu_{i-j}$,
-where
-%$\mathrm{sinr}(K, l, n)
-%  =
-%  \sin
-%  \frac{ (2 K - 1) \, l \, \pi }
-%       {              n        }
-%  / \left( n \, \sin \frac{ l \, \pi } { n } \right)$.
+where we have defined
 \begin{equation}
@@
 The updating matrix is given by
-$w_{ij} = \mathrm{sinr}(K, i-j, 2 \, n) + \mathrm{sinr}(K, i+j-1, 2 \, n)$.\footnote{$w_{ij}$
+$w_{ij} = \mathrm{sinr}(K, i-j, 2 \, n) + \mathrm{sinr}(K, i+j-1, 2 \, n)$.
+%
+\note{$w_{ij}$
   is free from the oscillatory term, $(-1)^{n-K+1}$,
   that appeared in the window function, since
-  its contribution to $\mu_{i-j}$ in Eq. \eqref{eq:w_band_refl}
+  its contribution to $\mu_{i-j}$ to the updating matrix
   is canceled by its contribution to either $\mu_{i+j-1}$
@@
 
-We will first review the relation
+Here we first review the relation
 between the histogram fluctuation and the error
@@
 We then give a measure of the overall histogram fluctuation
-that can be used in the WL-like criterion of stage transitions
-for the bandpass updating scheme based on orthogonal polynomials.
-
-\subsection{Histogram fluctuation and error}
+that can be used in the WL criterion of stage transitions
+for the polynomial bandpass updating schemes.
+
+%\subsection{Histogram fluctuation and error}
 
@@
   ,
-  \label{eq:Fk2_Gammak}
+  \notag
+  %\label{eq:Fk2_Gammak}
 \end{align}
@@
 
-\subsection{Overall histogram fluctuation}
-
-Equation \eqref{eq:F2sum} provides a way of
-defining the overall histogram fluctuation,
-hence the histogram flatness,
-which can be used in the WL algorithm
-as a criterion for transition to new stage with reduced updating magnitude.
-%
-This definition, however,
-diverges in the limit of $n \to \infty$,
-when the bin size becomes infinitely small
-and collective variable becomes continuous.
-%
-Thus, we find it more suitable to define the histogram fluctuation
-from the mode with the largest $|\tilde F_k(T)|$.
-%
-The two definitions are often congruent with each other
-for a local sampling process with a finite $n$
-because the major contribution to $\langle F^2(T) \rangle$
-often comes from the few longest-range modes [cf. Eq. \eqref{eq:Gamma_local}].
-%
-For the bandpass updating scheme,
-we need to further restrict the mode index by $k < K$
-since the histogram flattening applies
-only to the first $K$ modes.
-%
-Thus, the new definition of overall histogram fluctuation
-is given by
+%\subsection{Overall histogram fluctuation}
+
+The total histogram fluctuation given by Eq. \eqref{eq:F2sum}
+can be used in the WL algorithm
+as a criterion for stage transitions.
+%
+For a bandpass updating scheme,
+we need to further restrict the sum to the first $K$ modes,
+where the histogram flattening applies, i.e.
+$F^2(T) \equiv \sum_{k = 1}^{K-1} \tilde F_k^2(T)$.
+%
+%\begin{align}
+%  .
+%  \label{eq:maxFk}
+%\end{align}
+
+In particular,
+in sampling of the bivariate distribution,
+$\rho(i, z)$, in Sec. \ref{sec:st},
+the above definition becomes
 \begin{align}
-  \max_{1 \le k < K} \left| \, \tilde F_k(T) \, \right|
-  .
-  \label{eq:maxFk}
+  F^2
+  &=
+  \sum_{a = 0}^{n - 1}
+  \sideset{}{'}\sum_{k = 0}^{K - 1}
+  \tilde F_{a, k}^2
+  \notag \\
+  &=
+  \sum_{i=1}^n
+  \left[
+  \frac{ (H_i - \rho_i)^2 } { \rho_i }
+  +
+  \sum_{k = 1}^{K - 1}
+  \frac{ H_i^2 } { \rho_i }
+  \left|{\tilde F}_k^{(i)} \right|^2
+  \right]
+  ,
+  \label{eq:F2st}
 \end{align}
-
-If the eigenmodes are made of a sequence of orthogonal polynomials
-as in Sec. \ref{sec:bandpass_poly},
-then
-$\tilde f_k(t) = R_k\bigl( z(t) \bigr)$,
+where $a$ and $k$ are the indices of
+inter- and intra-umbrella fluctuation modes,
+and the primed sum excludes
+the term of $a = k = 0$;
+on the next line, $\tilde F_k^{(i)}$
+is the histogram fluctuation of the $k$th mode in umbrella $i$,
+which is the average of $R_k^{(i)}\bigl( z(t) \bigr)$
+for samples in umbrella $i$.
+%If the eigenmodes are made of a sequence of orthogonal polynomials
+%as in Sec. \ref{sec:bandpass_poly},
+%then
+%$\tilde f_k(t) = R_k\bigl( z(t) \bigr)$,
 %\begin{align*}
@@
 %\end{align*}
-and
-\begin{equation}
-  \tilde F_k(T)
-  =
-  \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)
-  .
-  \label{eq:Fk_def}
-\end{equation}
+%and
+%\begin{equation}
+%  \tilde F_k(T)
+%  =
+%  \frac 1 T \sum_{t = 1}^T R_k\bigl( z(t) \bigr)
+%  .
+%  \label{eq:Fk_def}
+%\end{equation}
 

### 2017-08-24 prog/invtpar.h

@@ -891,7 +891,7 @@ static void invtpar_dump(const invtpar_t *m)
       m->ntrials, m->n, m->c, m->t0, m->alpha0, m->qT, m->pbc,
       sampmethod_names[m->sampmethod][0], m->nsteps, m->nequil);
 
-  if ( m->gammethod != GAMMETHOD_NONE ) {
+  if ( m->gammethod != GAMMETHOD_NONE && m->gammethod != GAMMETHOD_LOAD ) {
     fprintf(stderr, "preliminary run: %ld steps, averaging every %d steps, method %s\n",
       m->gam_nsteps, m->gam_nstave, gammethod_names[m->gammethod][0]);
   } else {

### 2017-08-24 prog/lj/getav.py

@@ -16,11 +16,24 @@ class Ave:
     def getave(self):
         return self.sx/(self.cnt+1e-300)
 
-av = Ave()
+avf = Ave()
+avi = Ave()
 
+i = 0
+xiref = xfref = 0
 for ln in open(fn).xreadlines():
-    if ln.startswith("#"): continue
-    x = float( ln.split()[1] )
-    av.add(x)
+    i += 1
+    if ln.startswith("#"):
+        if i == 2:
+            arr = ln.split()
+            xiref = float( arr[-2] )
+            xfref = float( arr[-1] )
+        continue
+    arr = ln.split()
+    xf = float( arr[1] )
+    xi = float( arr[2] )
+    avf.add(xf)
+    avi.add(xi)
 
-print av.cnt, av.getave()
+print "count %s, init. %g(%g) final %g(%g)" % (
+        avf.cnt, avi.getave(), xiref, avf.getave(), xfref)

### 2017-08-24 prog/predict.c

@@ -217,8 +217,8 @@ static void invt_geterr(invtpar_t *m,
   intq_save(intq, c1, t0, m->alpha_resample, m->fnalpha);
 
   printf("c %g, t0 %g, err %g, sqr %g (invt), "
-         "opt. c %g, err %g, sqr %g (opt. invt), "
-         "qT %g, a(0) %g, err %g, sqr %g (exact), %s\n",
+         "opt. c %g, err %g, sqr %g (scl. invt), "
+         "qT %g, a(0) %g, err %g, sqr %g (opt.), %s\n",
       c0, t0, err0, err0 * err0,
       c1, err1, err1 * err1,
       qT, inita, err2, err2 * err2, m->fnalpha);

### 2017-08-25 doc/invt.tex
add scheme.pdf (in doc/fig/figs.tex)
invt.tex: V8.1

@@
 \subsubsection{\label{sec:Gaussian_scheme}
-Gaussian and homogeneous updating schemes}
+Gaussian updating scheme}
 
@@
 
+\begin{figure}[h]\centering
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
+  }
+  \caption{
+    \label{fig:mat}
+    Updating matrices for
+    homogeneous updating schemes of
+    a non-periodic variable ($n = 500$).
+    Here $j$ is the current bin,
+    $i$ is the bin where updating occurs,
+    and $w_{ij}$ is the relative updating magnitude.
+    %
+    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
+    (b) The bandpass updating scheme of $K = 20$.
+    %
+    Both updating schemes
+    can be characterized by a rigid window
+    centered on the current bin, $j$,
+    with some boundary adjustments.
+  }
+\end{figure}
+
+\subsubsection{\label{sec:homo_scheme}
+Homogeneous updating schemes}
+
 More generally, we can replace the Gaussian function
@@
 for a given updating scheme,
-we turn to the problem of finding
-the optimal updating scheme.
+we turn to the problem of optimizing
+the updating scheme.
 
@@
 %
+A summary of updating schemes is given in Fig. \ref{fig:scheme}.
+%
 First,
@@
 that resembles the Gaussian updating scheme,
-but with long-range wiggles
-[cf. Appendix \ref{sec:homo_bandpass}
-and Fig. \ref{fig:mat}(b)
-for an example].
+but with long-range wiggles,
+as shown in Fig. \ref{fig:mat}(b)
+(cf. Appendix \ref{sec:homo_bandpass}).
 %
@@
 The polynomial bandpass updating schemes will be discussed below.
+
+\begin{figure}[h]\centering
+  \makebox[\linewidth][c]{
+    \includegraphics[angle=0, width=1.0\linewidth]{fig/scheme.pdf}
+  }
+  \caption{
+    \label{fig:scheme}
+    Summary of updating schemes.
+    %
+    Only the Gaussian updating scheme
+    require explicit calculation for the optimal schedule,
+    because the optimal schedule of bandpass updating schemes
+    is always given by inverse-time formula.
+    %
+    The target distributions of all updating schemes are flat except
+    that for the adaptive Gaussian ensemble method,
+    which is a Gaussian distribution
+    or a mixture of Gaussian distributions.
+    \note{Do we need these descriptions, or the figure at all?}%
+  }
+\end{figure}
 
@@
 %
-For this case,
-we further decomposed the error
+We further decomposed the error
 into the eigenmodes
@@
 %
-According to the analytical prediction
-(assuming the optimal schedule),
-the optimal width was around $\sigma = 0.29$ in this case,
+According to the analytical prediction,
+the optimal width (assuming the optimal schedule)
+was around $\sigma = 0.29$ in this case,
 and it would decrease with the simulation length
@@
 In attempting inter-umbrella transitions,
-the target umbrella was chosen from
-any of the other umbrellas.
+the target umbrella is randomly chosen from
+either one of the nearest neighbors ($50\%$ chance),
+or any other umbrella.
+%
+\note{Here, we need to rerun simulations.}%
 %
@@
 We assume that the propagation along the collective-variable is Markovian
-such that the evolution can be approximately
+such that it can be approximately
 described by a transition matrix, $\mathbf A$, with
@@
 Further, if the transition matrix % satisfies detailed balance
-has a full set of orthonormal eigenvectors, $\psi_{li}$
+has a complete set of orthonormal eigenvectors, $\psi_{li}$
 %that are orthonormal with respect to $\pmb\rho$
@@
 For simplicity, we will assume that
-$\mu_l$'s are symmetric, $\mu_l = \mu_{-l}$,
+the $\mu_l$'s are symmetric, $\mu_l = \mu_{-l}$,
 and they can be cutoff at $l = b$,
@@
 The Gaussian updating scheme
-is defined by a Gaussian-shaped $\mu_l$.
+is defined by a Gaussian-shaped window function, $\mu_l$.
 %
@@
 or the updating matrix using Eq. \eqref{eq:w_from_phi}.
-%
-An example is shown in Fig. \ref{fig:mat}(a).
-
-\begin{figure}[h]\centering
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
-  }
-  \caption{
-    \label{fig:mat}
-    Updating matrices for
-    homogeneous updating schemes of
-    a non-periodic variable ($n = 500$).
-    Here $j$ is the current bin,
-    $i$ is the bin where updating occurs,
-    and $w_{ij}$ is the relative updating magnitude.
-    %
-    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
-    (b) The bandpass updating scheme of $K = 20$.
-    %
-    Both updating schemes
-    can be characterized by a rigid window
-    centered on the current bin, $j$,
-    with some boundary adjustments.
-  }
-\end{figure}
 
@@
   whichever exists, of the opposite sign.}
-%
-The homogeneous bandpass updating scheme
-resembles the Gaussian updating scheme
-except for some long-range wiggles,
-as shown in Fig. \ref{fig:mat}(b).
 %
@@
                         $m(q) \equiv M(q)/\int_0^{ q(T) } M(q') \, dq'$.  \\
-    $\mu_i$         &   Updating window function of a homogeneous updating scheme. \\
+    $\mu_i$         &   Unadjusted window function of a homogeneous updating scheme. \\
     $n$             &   Number of bins. \\
@@
     $\bar q(t)$     &   $q(T) - q(t)$.  \\
-    $s, S$          &   Stage index and the total number of stages in the Wang-Landau algorithm. \\
-    $\sigma_{ij}(t)$   &   Autocorrelation function of $\xi_i(t)$. \\
     $t, \tau$       &   Time. \\
@@
     $v_{*i}(t)$     &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
-    $\xi_{*i}(t)$ &   The noise of histogram. \\
+    $\xi_{*i}(t)$   &   Bin-average-deducted The noise of histogram. \\
     ${\tilde v}_k$  &   Mode of the shifted bias potential. \\

### 2017-08-25 prog/aus/gaus.h

@@ -469,18 +469,21 @@ __inline static int gaus_wlcheckx(gaus_t *gaus,
 
 
 /* transition to a neighboring umbrella */
-__inline static int gaus_move(gaus_t *gaus, double x, int *id)
+__inline static int gaus_move(gaus_t *gaus, double x, int *id, int local)
 {
   int acc = 0, n = gaus->n;
   int jd;
   double xi, xj, vi, vj, dv, sigi, sigj;
   mmwl_t *mm;
 
-  // jump to a nearest neighbor
-  //jd = (rand01() < 0.5) ? *id - 1 : *id + 1;
-  //if ( jd < 0 || jd >= n ) return 0;
-  // jump any other umbrella
-  jd = ( *id + 1 + (int) (rand01() * (n - 1)) ) % n; 
+  if ( local ) {
+    // jump to a nearest neighbor
+    jd = (rand01() < 0.5) ? *id - 1 : *id + 1;
+    if ( jd < 0 || jd >= n ) return 0;
+  } else {
+    // jump any other umbrella
+    jd = ( *id + 1 + (int) (rand01() * (n - 1)) ) % n; 
+  }
   sigi = gaus->sig[*id];
   sigj = gaus->sig[ jd];
   if ( gaus->lnzmethod == LNZ_AVE ) {

### 2017-08-28 doc/invt.tex
small changes

@@
   \Err_A(T)
-  &=
-  \sum_{k=1}^{n-1}
-  \int_0^T\!\!\int_0^T
-  \langle
-    \tilde\xi_k(t)
-    \tilde\xi_k(t')
-  \rangle
-  \vartheta_k\bigl(q(t)\bigr)
-  \vartheta_k\bigl(q(t')\bigr)
-  \, dtdt'
-  \notag \\
+  %&=
+  %\sum_{k=1}^{n-1}
+  %\int_0^T\!\!\int_0^T
+  %\langle
+  %  \tilde\xi_k(t)
+  %  \tilde\xi_k(t')
+  %\rangle
+  %\vartheta_k\bigl(q(t)\bigr)
+  %\vartheta_k\bigl(q(t')\bigr)
+  %\, dtdt'
+  %\notag \\
   &=
@@
 = \sum_{i=1}^n \rho_i \bigl\langle \xi_{*i}(t) \, \xi_{*i}(0) \bigr\rangle,$
-and after summing $t$ from $-\infty$ to $\infty$,
+and summing $t$ from $-\infty$ to $\infty$ yields,
 \begin{equation}
@@
 More generally, we can replace the Gaussian function
-in the above updating scheme
+in the Gaussian updating scheme
 by a window function of different shape.
@@
 %
-A summary of updating schemes is given in Fig. \ref{fig:scheme}.
+%A summary of updating schemes is given in Fig. \ref{fig:scheme}.
 %
@@
 we get a homogeneous bandpass updating scheme
+(cf. Appendix \ref{sec:homo_bandpass})
 that resembles the Gaussian updating scheme,
 but with long-range wiggles,
-as shown in Fig. \ref{fig:mat}(b)
-(cf. Appendix \ref{sec:homo_bandpass}).
+as shown in Fig. \ref{fig:mat}(b).
 %
@@
 
-\begin{figure}[h]\centering
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/scheme.pdf}
-  }
-  \caption{
-    \label{fig:scheme}
-    Summary of updating schemes.
-    %
-    Only the Gaussian updating scheme
-    require explicit calculation for the optimal schedule,
-    because the optimal schedule of bandpass updating schemes
-    is always given by inverse-time formula.
-    %
-    The target distributions of all updating schemes are flat except
-    that for the adaptive Gaussian ensemble method,
-    which is a Gaussian distribution
-    or a mixture of Gaussian distributions.
-    \note{Do we need these descriptions, or the figure at all?}%
-  }
-\end{figure}
+%\begin{figure}[h]\centering
+%  \makebox[\linewidth][c]{
+%    \includegraphics[angle=0, width=1.0\linewidth]{fig/scheme.pdf}
+%  }
+%  \caption{
+%    \label{fig:scheme}
+%    Summary of updating schemes.
+%    %
+%    Existing algorithms are shaded.
+%    %
+%    Only the Gaussian updating scheme
+%    require explicit calculation for the optimal schedule,
+%    because the optimal schedule of bandpass updating schemes
+%    is always given by inverse-time formula.
+%    %
+%    The target distributions of all updating schemes are flat except
+%    that for the adaptive Gaussian ensemble method,
+%    which is a Gaussian distribution
+%    or a mixture of Gaussian distributions.
+%    \note{Do we need these descriptions, or the figure at all?}%
+%  }
+%\end{figure}
 
@@
 from a superposition of $m$ Gaussian umbrellas
-centered at different places.
+centered at different places.\footnote{This sampling
+  algorithm should not be confused with the Gaussian updating scheme,
+  in which the bias potential is incrementally updated by Gaussian functions,
+  and the target distribution is flat.
+  However, in this case, the bias potential is quadratic,
+  and the target distribution of each umbrella is Gaussian.}
 %
@@
 and it is updated according to the single-bin scheme\cite{li2007}.
+%
+For a small $\sigma_z$, the converged
+$c_0^{(i)}$ as a function of $z_c^{(i)}$
+also serves as a good approximation of
+the negative PMF, $\ln p(z)$.\cite{maragliano2006, *abrams2008}
 
@@
   with the acceptance probability given
-  by Eq. \eqref{eq:transprob_st}.
-  %
-  The above intra-umbrella sampling steps
-  and inter-umbrella transition steps
-  can also be applied in random order
-  with different frequencies.\cite{manousiouthakis1999}
+    by Eq. \eqref{eq:transprob_st}.\cite{*[{Although
+    the intra-umbrella sampling step
+    and inter-umbrella transition step
+    are executed in sequential instead of random order,
+    the sampling scheme is still valid,
+    as shown in }] [{}] manousiouthakis1999}
 
@@
 by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl},
-and plotted the error components, $\langle \tilde v_k^2 \rangle$,
+and plotted the error components,
+$\langle \tilde v_k^2 \rangle$'s,
 in Fig. \ref{fig:lj_xerr}.
@@
 leaving a hump on the error profile
-for the $k = 10$ to $25$ modes.
+for the $k = 9$ to $31$ modes.
 %
@@
     \label{fig:lj_xerr}
-    Error components, $\tilde u_k^2$'s, of the PMF
-    after $10^8$ production steps.
+    Error components,
+    $\langle \tilde v_k^2 \rangle$'s,
+    of the PMF after $10^8$ production steps.
     %
@@
 %
-In attempting inter-umbrella transitions,
-the target umbrella is randomly chosen from
-either one of the nearest neighbors ($50\%$ chance),
-or any other umbrella.
-%
-\note{Here, we need to rerun simulations.}%
+For simplicity, only transitions between neighboring
+umbrellas were performed.\footnote{For
+  wider umbrellas, we might improve convergence
+  by mixing the nearest-neighbor inter-umbrella transitions
+  with some longer-range ones.}
+%In attempting inter-umbrella transitions,
+%the target umbrella is randomly chosen from
+%either one of the nearest neighbors ($50\%$ chance),
+%or any other umbrella.
+%
+%\note{Here, we need to rerun simulations.}%
 %
@@
 We computed the density of states by WHAM,
-and found the critical temperature to be $T_c = 0.70162$,
+and found the critical temperature to be $T_c = 0.70163$,
 in agreement with a previous study\cite{wang2001pre}.
 %
+Since umbrellas were relatively narrow,
+the logarithmic density of states, $\ln g(E)$,
+was roughly captured
+by $c_0^{(i)}$ as a function of $E_c^{(i)}$,
+as shown Fig. \ref{fig:pt_hist}(b).
+%
 As discussed in Sec. \ref{sec:age},
-$c_1/\sigma_E$ roughly represents
+$c_1^{(i)}/\sigma_E$ roughly represents
 the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
@@
 typical in a first-order phase transition,
-as shown in Fig. \ref{fig:pt_hist}(b).
-%
-The coefficient, $\sqrt 2 \, c_2$,
+as shown in Fig. \ref{fig:pt_hist}(c).
+%
+The coefficient, $\sqrt 2 \, c_2^{(i)}$,
 was indeed above $1.0$
@@
 $\ln g(E)$ is locally convex,
-as shown in Fig. \ref{fig:pt_hist}(c).
+as shown in Fig. \ref{fig:pt_hist}(d).
 
@@
     \label{fig:pt_hist}
-    Umbrella sampling with adaptive quadratic bias potentials
+    Sampling using adaptive Gaussian ensembles
     on the $64 \times 64$ ten-state Potts model.
@@
     and the reweighted distribution
-    at the critical temperature (dashed line).
-    %
-    (b) $c_1^{(i)}/\sigma_E$, roughly corresponding to
+    at the critical temperature, $\beta_c$ (dashed line).
+    %
+    (b) $c_0^{(i)}$ against $E_c^{(i)}$ (circles)
+    roughly agreed with the logarithmic density of states,
+    $\ln g(E)$ (dashed line).
+    In either case,
+    we have deducted the linear contribution, $\beta_c \, E$,
+    and applied a proper shift so as to normalize
+    the distribution at $\beta_c$.
+    %
+    (c) $c_1^{(i)}/\sigma_E$ roughly corresponded to
     the statistical temperature, $\beta(E)$.
     %
-    (c) $\sqrt 2 \, c_2^{(i)}$
+    (d) $\sqrt 2 \, c_2^{(i)}$
     exceeded $1.0$ in the unstable region,
@@
 %
-The optimal schedule delivers the fastest convergence
+The optimal schedule improves the convergence
 of the bias potential,
@@
 
-We thank Dr. Y. Mei, J. Wang,
+We thank Dr. Y. Mei, Dr. W. Tang, J. Wang,
 O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
@@
   derive Eq. \eqref{eq:uhatav}
-  from constructing an independent estimator
+  by forming an independent estimator
   of the unbiased distribution\cite{marsili2006},
@@
 we can redefine the Gaussian scheme
-from the same spectrum of eigenvalues, Eq. \eqref{eq:lambda_Gaussian},
+from the eigenvalues given by Eq. \eqref{eq:lambda_Gaussian},
 and then inversely computing the window function

### 2017-08-28 prog/aus/reweight.c

@@ -105,6 +105,7 @@ static double lnadd(double x, double y)
   return x + log(1 + exp(y));
 }
 
+/* non-iterative histogram reweighting */
 static void reweight(void)
 {
   int ix, i;
@@ -176,7 +177,7 @@ static int findpeak(const double *arr, double bc, int ileft, int iright, double 
 /* find the critical inverse temperature */
 static double seekcrit(const double *arr, int *x1, int *x2, double *shift)
 {
-  double bc, beta, w, sw, y1, y2, y, lns;
+  double bc, beta, w, sw, y1, y2, y, lns, lns2;
   int ix, ix1, ix2, il = -1, ir, im, t, x;
 
   bc = sw = 0;
@@ -197,7 +198,8 @@ static double seekcrit(const double *arr, int *x1, int *x2, double *shift)
   im = (il + ir)/2;
   printf("bc %g, il %d, ir %d, mid %d\n", bc, il, ir, im);
 
-  /* find the two density peaks */
+  /* find the two density peaks
+   * util the heights are equal */
   ix1 = il;
   ix2 = ir;
   for ( t = 0; t < 10; t++ ) {
@@ -220,7 +222,18 @@ static double seekcrit(const double *arr, int *x1, int *x2, double *shift)
     lns = lnadd(lns, y);
   }
   lns += log(dx);
-  printf("lns %g for normalization\n", lns);
+  /* normalize lnz at the critical temperature */
+  {
+    int i;
+    lns2 = LN0;
+    for ( i = 0; i < n; i++ ) {
+      y = lnz[i] - bc * ave[i] - lns;
+      lns2 = lnadd(lns2, y);
+      //printf("i %d, lnz %g, bc %g, ave %g, %g\n", i, lnz[i], bc, ave[i], y);
+    }
+    lns2 += log(delx);
+  }
+  printf("lns for normalization %g (lng), %g (lnz)\n", lns, lns + lns2);
   *shift = lns;
   return bc;
 }
@@ -268,8 +281,14 @@ int main(int argc, char **argv)
 
   if ( getdatinfo(fndat) != 0 ) return -1;
   if ( gethisinfo(fnhis) != 0 ) return -1;
+
+  /* non-iterative histogram reweighting */
   reweight();
+
+  /* smoothing lng to lngs */
   smooth(lngs, lng, nb);
+
+  /* looking for the critical temperature */
[…]

### 2017-08-29 doc/invt.tex
add proof for alpha(t) > 1/t

@@
 was more complex than the inverse-time one,
-and the former tended to give a larger updating magnitude.
+and the former tended to give a larger updating magnitude,
+in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
 %
@@
   ,
-\notag
-%\label{eq:mass_distr}
+  %\notag
+  \label{eq:mass_distr}
 \end{equation}
@@
   \int_{\bar q(t)}^{ q(T) }
-    m(q') \, d q'
+    m(\bar q) \, d \bar q
   &=
@@
 and the area under the curve in the domain, $[\bar q(t), q(T)]$,
+or the complementary cumulative distribution function of $m(\bar q)$,
 is equal to $t/T$,
@@
 \end{figure}
+
+We can use Eq. \eqref{eq:intmQ_tT} to show that
+for two simulations with equal lengths and initial updating magnitudes,
+the simulation under the Gaussian scheme with the optimal schedule
+has a larger updating magnitude than
+the value given by the inverse-time formula,
+which is optimal for the single-bin scheme.
+%
+From Eqs. \eqref{eq:mass_func} and \eqref{eq:mass_distr},
+we have
+\begin{align}
+  -\frac{ d \ln m(\bar q) }{ d \bar q }
+  %=
+  %-\frac{ dM^2(\bar q)/d\bar q } { 2 M^2(\bar q) }
+  =
+  \frac{ \sum_{k=1}^{n-1} \Gamma_k \, \lambda_k^3 \, e^{-2\lambda_k \bar q} }
+  { \sum_{k=1}^{n-1} \Gamma_k \, \lambda_k^2 \, e^{-2\lambda_k \bar q} }
+  \equiv
+  \bar\lambda(\bar q)
+  ,
+  \label{eq:dlnmdq}
+\end{align}
+which is a weighted average of $\lambda_k$'s
+with the weight being $\Gamma_k \, \lambda_k^2 \, e^{-2\lambda_k \bar q}$.
+%
+Let us use Eqs. \eqref{eq:intmQ_tT} and \eqref{eq:dlnmdq}
+to compare the time of
+reaching a certain updating magnitude, $\alpha_1$,
+in the two simulations:
+%
+\begin{align}
+  \frac{t}{T}
+  %=
+  %\int_{\bar q(t)}^{q(T)} m(\bar q) \, d\bar q
+  =
+  \int_{m( \bar q(t) )}^{m( q(T) )}
+  \!\!\frac{ m(\bar q) } { dm(\bar q)/d\bar q } \, d m
+  =
+  \int^{m( \bar q(t) )}_{m( q(T) )}
+  \!\!\frac{ d m } { \bar \lambda(\bar q) }
+  ,
+  \label{eq:t_int}
+\end{align}
+with the lower and upper limits,
+$m\bigl(q(T)\bigr) = 1/\bigl[T\alpha(0)\bigr]$ and
+$m\bigl(\bar q(t)\bigr) = 1/(T\alpha_1)$,
+being the same in both cases.
+%
+Since for the Gaussian scheme,
+we have $\bar \lambda(\bar q) \le 1$,
+by Eq. \eqref{eq:lambda_Gaussian},
+the integral in Eq. \eqref{eq:t_int}
+is no less than that for the single-bin scheme,
+i.e. it takes longer for the Gaussian scheme
+than the single-bin scheme
+to lower the updating magnitude
+to the same level.
 

### 2017-08-30 doc/invt.tex
small edits to invt.tex

@@
   %
-  The updating magnitude should decrease over time
-  to reduce the error in the bias potential,
-  and one may speed up the convergence by choosing an optimal schedule
-  of the updating magnitude.
-  %
-  %Here, we will show that the optimal schedule
-  %can be derived
-  %from a mechanical analogy, in which
-  %the schedule corresponds to
-  %the velocity of a free particle with a position-dependent mass
-  %and the final error serves as the action.
-  %%
-  %Therefore, the optimal schedule follows from
-  %the equation of motion of the particle.
+  The convergence of the bias potential can be accelerated
+  by decreasing the updating magnitude with an optimal schedule.
   %
@@
   we show that the single-bin updating scheme
-  belongs to a class of bandpass updating schemes
-  that are optimal for asymptotic convergence.
+  is optimal for very long simulations,
+  and it can be generalized to a class of bandpass updating schemes
+  that are similarly but partially optimal.
   %
@@
   and their optimal schedule
-  is also given by the inverse-time prescription.
+  is also given by the inverse-time formula.
   %
@@
 that can provide insight into an array of phenomena not easily studied
-with traditional experiments.
+with regular simulation methods.
+\note{``Regular simulation methods'' means MC and MD,
+and they are more pertinent to ``computational physics and chemistry''
+than ``traditional experiments.''}
 %
@@
 one needs to gradually decrease
-the updating magnitude of the bias.
+the updating magnitude of the bias potential.
 
@@
 While an overly slow reduction of the updating magnitude
-increases the variance of the bias potential,
+is ineffective in controlling the variance of the bias potential,
 a too rapid reduction often
-converges the bias potential to wrong value.
+converges the bias potential to a wrong value.
 %
 For the single-bin case, it is known that
-the optimal updating schedule is given by the
+the optimal updating magnitude is given by the
 inverse time\cite{
@@
 the optimal schedule
-for an adaptive FDS method
+for an adaptive FDS simulation
 of a general updating scheme,
@@
 %
-We found that
+We will show that
 while the optimal schedule in the single-bin case
@@
 simpler control of the updating magnitude
-for asymptotic convergence.
+for optimal convergence.
 
@@
 the bandpass updating schemes
-by construction from
+by a construction from
 a sequence of orthogonal polynomials.
@@
 and coefficients are adjusted on-the-fly
-to achieve a particular target distribution.
+to approximate a particular target distribution.
 %
@@
 \begin{align}
-  \alpha(0)
+  \alpha(t = 0)
   %=
@@
   }{M^2\bigl( q(T) \bigr)}
-  .
-  \notag
-  %\label{eq:opt_alpha0}
-\end{align}
-%
-If the initial bias, $\epsilon_k$, is negligible,
-which is often true if no eigenvalue is close to $0$,
-then
-%
-\begin{equation}
-  \alpha( t = 0 )
-  \approx
+  \ge
   \frac{ a_0 }
        { 2 }
-  ,
+  .
+  %\notag
   \label{eq:half_alpha0}
-\end{equation}
-%
-i.e. the optimal initial updating magnitude
+\end{align}
+%
+The equality in the last step is achievable
+if the initial bias, $\epsilon_k$, is negligible,
+which is often true if no eigenvalue is close to $0$,
+and in this limit, the optimal initial updating magnitude
 is half of the previous equilibrium value.\footnote{This factor, $1/2$,
@@
 where the PMF is locally concave,
-$-d^2\ln p(z)/dz^2 < 0$.
+i.e. $-d^2\ln p(z)/dz^2 < 0$.
 
@@
 \subsection{\label{sec:lj}
-Comparison of updating schemes}
+Comparison of homogeneous updating schemes}
+
+In this example, we wish to compare the single-bin,
+Gaussian (Sec. \ref{sec:Gaussian_scheme}),
+and homogeneous bandpass (Sec. \ref{sec:bandpass})
+updating schemes in a numerical example.
 
@@
 
-Our first test system is
+Our test system is
 a Lennard-Jones (LJ) fluid of $N = 100$ particles,
@@
 %
-The density was $\varrho = 0.1$.
+The reduced density was $\varrho = 0.1$.
 The numbers of bins was $n = 500$.
@@
 underwent the test updating scheme and schedule.
-%
-We tested the single-bin, Gaussian,
-and homogeneous bandpass updating schemes.
 %
@@
 followed by a production phase of
-$10^8$ steps
+$T = 10^8$ steps
 with the updating magnitude given by
@@
 We repeated each simulation multiple times
-and report the average.
+and report the averages.
 %
@@
 We used $\varepsilon = 0.1/\varrho$
-to ensure reasonable acceptance ratio.
+to ensure a reasonable acceptance ratio.
 
@@
 %
-Since the two schedules are identical
+Since the optimal and inverse-time schedules are identical
 in the single-bin and bandpass updating schemes,
@@
     Bandpass & $1/t$
-    & $8.3(8.9)\times10^{-3}$
-    & $2.1(2.3)\times10^{-6}$
+    & $7.8(8.9)\times10^{-3}$
+    & $2.0(2.3)\times10^{-6}$
     \\
@@
 was likely due to the error of estimated values of $\Gamma_k$'s).
+%
+However, the gain of using the optimal schedule
+diminished for shorter simulations,
+e.g. when the production phase was shortened to $T = 10^7$ steps,
+the final errors under the optimal and inverse-time schedules
+were $2.4\times 10^{-5}$ and $3.9\times 10^{-5}$, respectively.
 
@@
 we tested the adaptive Gaussian ensemble method
-introduced in Sec. \ref{sec:age}
+discussed in Secs. \ref{sec:age} and \ref{sec:st}
 on the two-dimensional $L\times L$ ten-state
@@
   by mixing the nearest-neighbor inter-umbrella transitions
-  with some longer-range ones.}
+  with some longer-range transitions.}
 %In attempting inter-umbrella transitions,
@@
 we used the Wolff algorithm\cite{wolff1989, newman}
-(which is more efficient than the Metropolis algorithm)
+(which was more efficient than the Metropolis algorithm)
 at the inverse temperature, $\beta = c_1/\sigma_E$,
@@
 the fluctuation threshold for stage transitions
-was $F = 0.2$ [using the definition given by Eq. \eqref{eq:F2st}];
+was $F = 0.5$ [using the definition given by Eq. \eqref{eq:F2st}];
 and the reduction factor was $1/2$.
@@
 We computed the density of states by WHAM,
-and found the critical temperature to be $T_c = 0.70163$,
+and found the critical temperature to be $T_c = 0.70161$,
 in agreement with a previous study\cite{wang2001pre}.
@@
     where $\beta(E)$ is increasing.
+    \note{When the new data come,
+    run the \texttt{update.sh},
+    copy the new value of the inverse critical temperature,
+    and the value in \texttt{lns for normalization \dots [value] (lnz)}
+    in the output.
+    Use these values to update \texttt{bc} and \texttt{shiftlnz}
+    in \texttt{pt\_hist.gp}, and run \texttt{gnuplot} again.  }%
   }
@@
 which discard a few noise modes
-and are asymptotically optimal for the non-noise modes.
+and are similarly optimal for the non-noise modes.
 %
@@
 for the updating magnitude
-as the asymptotically optimal schedule
+as the optimal schedule
 is simply given by an inverse-time formula.
@@
 %
-Second, the white noise approximation
-can break down for relatively short simulations.
+The white noise approximation
+can also be inaccurate for relatively short simulations.
+%
+For the Gaussian updating scheme,
+the calculation of the optimal schedule requires
+the estimation of a few parameters
+in a sufficiently long run
+under a constant updating magnitude,
+whereas the gain of using the optimal schedule
+can sometimes be rather modest,
+especially for relatively short simulations,
+even if the noise is largely local in nature.
 %
 Finally, the adaptive Gaussian ensemble method
-introduced in Sec. \ref{sec:age} can be refined
+introduced in Sec. \ref{sec:age} may still need refinement
 to improve its general applicability and stability.
@@
 for helpful discussions,
-and the two anonymous reviewers for critical reading
+and the anonymous reviewers for critical reading
 of earlier versions of the manuscript.
@@
 
-Conversely, for a recurrence relation
+Conversely, we can revert a recurrence relation
 in the form of Eq. \eqref{eq:X_recur}
-with a given $\alpha(t)$, we can revert it
+with a given $\alpha(t)$
 to an average in the form of Eq. \eqref{eq:X_ave}
@@
 We can use Eq. \eqref{eq:intmQ_tT} to show that
-for two simulations with equal lengths and initial updating magnitudes,
-the simulation under the Gaussian scheme with the optimal schedule
-has a larger updating magnitude than
-the value given by the inverse-time formula,
+for simulations with equal lengths and initial updating magnitudes,
+the optimal schedule for the Gaussian scheme
+always lies above the inverse-time schedule,
 which is optimal for the single-bin scheme.
 %
-From Eqs. \eqref{eq:mass_func} and \eqref{eq:mass_distr},
-we have
-\begin{align}
-  -\frac{ d \ln m(\bar q) }{ d \bar q }
-  %=
-  %-\frac{ dM^2(\bar q)/d\bar q } { 2 M^2(\bar q) }
-  =
-  \frac{ \sum_{k=1}^{n-1} \Gamma_k \, \lambda_k^3 \, e^{-2\lambda_k \bar q} }
-  { \sum_{k=1}^{n-1} \Gamma_k \, \lambda_k^2 \, e^{-2\lambda_k \bar q} }
-  \equiv
-  \bar\lambda(\bar q)
-  ,
-  \label{eq:dlnmdq}
-\end{align}
-which is a weighted average of $\lambda_k$'s
-with the weight being $\Gamma_k \, \lambda_k^2 \, e^{-2\lambda_k \bar q}$.
-%
-Let us use Eqs. \eqref{eq:intmQ_tT} and \eqref{eq:dlnmdq}
-to compare the time of
-reaching a certain updating magnitude, $\alpha_1$,
+%Let us use Eqs. \eqref{eq:intmQ_tT} and \eqref{eq:dlnmdq}
+Let us compare the times for the optimal
+schedules to lower the updating magnitude to a certain level,
+$\alpha_1$,
 in the two simulations:
@@
   \int_{m( \bar q(t) )}^{m( q(T) )}
-  \!\!\frac{ m(\bar q) } { dm(\bar q)/d\bar q } \, d m
+  \!\!\frac{ m(\bar q) } { dm/d\bar q } \, d m
   =
   \int^{m( \bar q(t) )}_{m( q(T) )}
-  \!\!\frac{ d m } { \bar \lambda(\bar q) }
+  \!\!\frac{ d m } { -\frac{d}{d\bar q}\ln m(\bar q) }
+  %=
+  %\int^{m( \bar q(t) )}_{m( q(T) )}
+  %\!\!\frac{ d m } { \bar \lambda(\bar q) }
   ,
@@
 $m\bigl(\bar q(t)\bigr) = 1/(T\alpha_1)$,
-being the same in both cases.
+being the same in the two cases.
+%
+But from Eqs. \eqref{eq:mass_func} and \eqref{eq:mass_distr},
+we have
+\begin{align}
+  -\frac{ d }{ d \bar q } \ln m(\bar q)
+  %=
+  %-\frac{ dM^2(\bar q)/d\bar q } { 2 M^2(\bar q) }
+  =
+  \frac{ \sum_{k=1}^{n-1} \Gamma_k \, \lambda_k^3 \, e^{-2\lambda_k \bar q} }
+  { \sum_{k=1}^{n-1} \Gamma_k \, \lambda_k^2 \, e^{-2\lambda_k \bar q} }
+  \equiv
+  \bar\lambda(\bar q)
+  ,
+  \notag
+  %\label{eq:dlnmdq}
+\end{align}
+which is a weighted average of $\lambda_k$'s
+with the weight being $\Gamma_k \, \lambda_k^2 \, e^{-2\lambda_k \bar q}$.
 %
 Since for the Gaussian scheme,
-we have $\bar \lambda(\bar q) \le 1$,
+we have $\bar \lambda(\bar q) \le 1$
 by Eq. \eqref{eq:lambda_Gaussian},

### 2017-09-01 doc/invt.tex
small edits to invt.tex

@@
   %
-  While the asymptotically optimal schedule for
+  While the optimal schedule for
   the single-bin updating scheme (commonly used in the WL algorithm)
@@
 and they are more pertinent to ``computational physics and chemistry''
-than ``traditional experiments.''}
+than ``traditional experiments.''}%
 %
@@
 The schedule of reducing the updating magnitude
-can critically affect the precision of the final bias potential,
+can critically affect the precision and accuracy
+of the final bias potential,
 and hence that of the PMF\cite{laio2005, bussi2006, poulain2006,
@@
 %
-We present the analytical results in Sec. \ref{sec:theory},
+We present the analytical results in Sec. \ref{sec:methods},
 provide numerical examples
@@
 
-\section{\label{sec:theory}
-Theory}
-
-We develop the theory
+\section{\label{sec:methods}
+Methods}
+
+We present the methods 
 in the following order.
@@
 where
-$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$
-and if no eigenvalue is zero,
-then we can ignore the $\epsilon_k$ term
-in the case of $T_0 \to \infty$,
+$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$.
+%
+If $\lambda_k$ is not close to zero, and $T_0 \to \infty$
+we can ignore the $\epsilon_k$ term, and
 %
@@
   \right\rangle
-  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
+  \approx
+  \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
   .
@@
 $\Gamma_k$'s depend also on the system
-and the underlying sampling process.
+and the underlying sampling method.
 %
@@
 In our implementation,
-the integration was performed using the trapezoidal rule\cite{press3rd}
+the integration is performed using the trapezoidal rule\cite{press3rd}
 on the grid points,
@@
 The resulting values, $\bigl(q_{[i]}, t_{[i]} \bigr)$,
-and used to compute the updating magnitude
+are used to compute the updating magnitude
 as (cf. Appendix \ref{sec:ave_ub}):
@@
 whose updating matrix can be characterized by
-a fixed Gaussian function centered on the current bin
+a rigid Gaussian function centered on the current bin
 (with some boundary adjustment),
@@
   =
-  \exp\left[
-        -\frac 1 2
-        \left(
-         \frac{ 2 \, \pi \, k \, \sigma_z }
-              { g \, \Delta z  }
-        \right)^2
-      \right]
+  \exp\left(
+        -
+        \frac{ \pi^2 \, \sigma_z^2 \, k^2 }
+             { 2 \, \Delta z^2 }
+      \right)
   ,
@@
 \end{align}
-where $\sigma_z$ and $\Delta z = z_{\max} - z_{\min}$
-are the width of the Gaussian and the range of
-the original collective variable, $z$,
-respectively.
+where $\sigma_z$ and $\Delta z = (z_{\max} - z_{\min}) \, g/2$
+are the width of the Gaussian and the effective span
+in terms of the continuous collective variable, $z$,
+respectively,
+with $g = 1$ if $z$ periodic or $2$ otherwise
+(cf. Appendix \ref{sec:Gaussian_math}).
 
@@
     (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
-    (b) The bandpass updating scheme of $K = 20$.
+    (b) The homogeneous bandpass updating scheme of $K = 20$.
     %
@@
 
-More generally, we can replace the Gaussian function
+More generally, we can replace the rigid Gaussian function
 in the Gaussian updating scheme
@@
 We call the resulting updating scheme
-translationally-invariant or \emph{homogeneous}.
+\emph{homogeneous} or
+translationally-invariant.
 %see Fig. \ref{fig:mat}(b) for an example.
 
-By morphing the window function without losing the translational invariance,
+By morphing the window function
+but preserving the translational invariance,
 we can design a homogeneous updating scheme
@@
 Since high-order moments are generally less reliable,
-this method is better used with a small $K$ for a narrow distribution.
+this method works better with a small $K$ for a narrow distribution.
 %
 However, by constructing
-an expanded ensemble or mixture\cite{swendsen1986,
+a mixture or expanded ensemble\cite{swendsen1986,
   *geyer1991, *hukushima1996, *hansmann1997, *sugita1999,
@@
 the parameters, $c_1$ and $c_2$,
-can be related to the slope and curvature of the PMF
+can be related to the local slope and curvature of the PMF
 [cf. Eq. \eqref{eq:Vi_target}].
@@
 Thus, we can interpret
-$c_1/\sigma_z$ as the average local mean force around $z_c$.
+$c_1/\sigma_z$ as the mean force around $z_c$.
 %
@@
 
-In this example, we wish to compare the single-bin,
+In this example, we will study the optimal schedule
+and compare the single-bin,
 Gaussian (Sec. \ref{sec:Gaussian_scheme}),
 and homogeneous bandpass (Sec. \ref{sec:bandpass})
-updating schemes in a numerical example.
+updating schemes.
 
@@
 in the single-bin and bandpass updating schemes,
-we will focus on the Gaussian updating scheme with width $\sigma = 0.2$ below
+we will assume the Gaussian updating scheme with width $\sigma = 0.2$ below
 in discussing the optimal schedule.
@@
     & $5.7(5.7)\times10^{-2}$
-    & $9.0(11.4)\times10^{-6}$
+    & $9.1(11.4)\times10^{-6}$
     \\
@@
     Bandpass & $1/t$
-    & $7.8(8.9)\times10^{-3}$
-    & $2.0(2.3)\times10^{-6}$
+    & $8.1(8.9)\times10^{-3}$
+    & $2.1(2.3)\times10^{-6}$
     \\
@@
 was more complex than the inverse-time one,
-and the former tended to give a larger updating magnitude,
+and the former lay above the latter,
 in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
@@
 was around the ideal value of $a_0/2 = 5\times10^{-5}$
-given by Eq. \eqref{eq:half_alpha0}.
+[cf. Eq. \eqref{eq:half_alpha0}].
 
@@
 
-The optimal schedule was more effective
-than the inverse-time schedule
-in reducing the error,
+The optimal schedule more effectively reduced
+the error than the inverse-time schedule,
 as shown in Table \ref{tab:lj_error}.
 %
-We further decomposed the error
-into the eigenmodes
-by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl},
-and plotted the error components,
+By decomposing the error
+into the components of the eigenmodes,
 $\langle \tilde v_k^2 \rangle$'s,
-in Fig. \ref{fig:lj_xerr}.
-%
-We found that the inverse-time schedule tended to
-reduce the updating magnitude too quickly,
+by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl}
+(Fig. \ref{fig:lj_xerr}),
+we found that the inverse-time schedule
+reduced the updating magnitude too quickly,
 leaving a hump on the error profile
@@
 the error in these modes,
-and produced a flatter profile for the final error
+and produced a flatter profile
 (the remaining unevenness of the profile
-was likely due to the error of estimated values of $\Gamma_k$'s).
+was likely due to the error of estimated values of
+$\Gamma_k$'s and $\epsilon_k$'s).
 %
@@
 the final errors under the optimal and inverse-time schedules
-were $2.4\times 10^{-5}$ and $3.9\times 10^{-5}$, respectively.
+became $2.4\times 10^{-5}$ and $3.9\times 10^{-5}$, respectively,
+which were more similar.
 
@@
 the optimal width (assuming the optimal schedule)
-was around $\sigma = 0.29$ in this case,
+was around $\sigma = 0.29$ in this case
+with the predicted minimal error being $3.3\times 10^{-6}$,
 and it would decrease with the simulation length
-and approach the limit of single-bin updating scheme,
+and approach the limit of the single-bin updating scheme,
 $\sigma \to 0$, asymptotically.
@@
 The homogeneous bandpass updating scheme
-yielded less error than the Gaussian updating scheme,
-as shown in Table \ref{tab:lj_error}.
-%
-Similar to the Gaussian case,
-the mode cutoff, $K$, has an optimal value,
+had a similar behavior.
+%
+The mode cutoff, $K$, also has an optimal value,
 which was around $26$ in this case.
 %
+The predicted minimal error was $2.0\times 10^{-6}$,
+  slightly less than the minimal error
+  in the previous Gaussian case.
+%
 The optimal cutoff would increase with the simulation length
-and approach the limit of single-bin scheme, $K \to n$,
+and approach the limit of the single-bin updating scheme, $K \to n$,
 asymptotically.
 
-The Gaussian and homogeneous bandpass updating schemes
-helped reduce the error in this case because
+Although the Gaussian and homogeneous bandpass updating schemes
+helped reduce the error in this case,
+it was only because
 the sampling along the collective variable
@@
 over the single-bin updating scheme, if any, to be negligible.
+%
+However, one may still prefer the multiple-bin updating schemes
+in MD simulations
+when a smooth bias potential is necessary in deriving the bias force.
 
@@
 (which was more efficient than the Metropolis algorithm)
-at the inverse temperature, $\beta = c_1/\sigma_E$,
+at the inverse temperature, $\beta = c_1^{(i)}/\sigma_E$,
 followed by a Metropolis step with
 acceptance probability
-$\min\bigl\{1, e^{c_2 [R_2(E) - R_2(E')]} \bigr\}$,
+$\min\bigl\{1, e^{c_2^{(i)} [R_2(E) - R_2(E')]} \bigr\}$,
 where $E$ and $E'$ are the energies of the old and new states.
@@
   \int_{m( \bar q(t) )}^{m( q(T) )}
-  \!\!\frac{ m(\bar q) } { dm/d\bar q } \, d m
+  \!\!\frac{ m(\bar q) } { \frac{dm(\bar q)}{d\bar q} } \, d m(\bar q)
   =
   \int^{m( \bar q(t) )}_{m( q(T) )}
-  \!\!\frac{ d m } { -\frac{d}{d\bar q}\ln m(\bar q) }
+  \!\!\frac{ d m } { -\frac{d\ln m(\bar q)}{d\bar q} }
   %=
@@
 
-\subsection{Gaussian updating scheme}
-
-The Gaussian updating scheme
+\subsection{\label{sec:Gaussian_math}
+Gaussian updating scheme}
+
+The Gaussian updating scheme introduced in Sec. \ref{sec:Gaussian_scheme}
 is defined by a Gaussian-shaped window function, $\mu_l$.
 %
-This updating scheme is commonly
-used in metadynamics.
+%This updating scheme is commonly
+%used in metadynamics.
 %

### 2017-09-02 doc/invt.tex
adding web

@@
   %
-  While the optimal schedule for
+  While the asymptotically optimal schedule for
   the single-bin updating scheme (commonly used in the WL algorithm)
@@
   and it can be generalized to a class of bandpass updating schemes
-  that are similarly but partially optimal.
+  that are similarly optimal.
   %

### 2017-09-02 prog/aus/potts2.h

@@ -80,7 +80,7 @@ __inline static void potts2_setuproba(double bet, unsigned *p)
 
 
 /* pick a random site, count neighbors with different spins */
-__inline static int potts2_pick(const potts2_t *pt, int *h, int *sn)
+__inline static int potts2_pick(const potts2_t *pt, int *sn, int *h)
 {
   int ix, ixp, ixm, iy, iyp, iym, id, s, so;
   int l = pt->l, n = pt->n;
@@ -104,8 +104,8 @@ __inline static int potts2_pick(const potts2_t *pt, int *h, int *sn)
 
 
 
-/* flip site id, with (-h) pt the energy before the flip */
-__inline static int potts2_flip(potts2_t *pt, int id, int h, int sn)
+/* flip site id, with h being the energy change */
+__inline static int potts2_flip(potts2_t *pt, int id, int sn, int h)
 {
   pt->s[id] = sn;
   return pt->E += h;
@@ -119,7 +119,7 @@ __inline static int potts2_flip(potts2_t *pt, int id, int h, int sn)
 #define POTTS2_L   (1 << POTTS2_LB)
 #define POTTS2_N   (POTTS2_L * POTTS2_L)
 
-#define POTTS2_GETH(pt, id, h, sn) { \
+#define POTTS2_GETH(pt, id, sn, h) { \
   unsigned ix, ixp, ixm, iy, iyp, iym; \
   int s, so; \
   ix = id % POTTS2_L; \
@@ -136,18 +136,18 @@ __inline static int potts2_flip(potts2_t *pt, int id, int h, int sn)
   h += ((s = pt->s[iym +  ix]) == so); h -= (s == sn); }
 #define POTTS2_IRND(pt, id)  id = mtrand() >> (32 - 2*POTTS2_LB);
 /* random picking */
-#define POTTS2_PICK(pt, id, h, sn) { POTTS2_IRND(pt, id); POTTS2_GETH(pt, id, h, sn); }
+#define POTTS2_PICK(pt, id, sn, h) { POTTS2_IRND(pt, id); POTTS2_GETH(pt, id, sn, h); }
 #define POTTS2_ISEQ(pt, id)  id = (id + 1) % POTTS2_N;
 /* sequential picking */
[…]

### 2017-09-02 web/htmlpack.py

@@ -1,0 +1,134 @@
+#!/usr/bin/env python
+
+
+import os, sys, glob, codecs
+
+
+
+rmcomments = False
+rmspaces = False
+
+
+
+def mkdic(fntrans):
+  fp = codecs.open(fntrans, "r", "utf-8")
+  stab = fp.read()
+  if stab[0] == u'\ufeff':
+    stab = stab[1:]
+  fp.close()
+  dic = {}
+  arr = stab.split('\n')
+  for i in range(len(arr)):
+    ln = arr[i].strip()
+    if ln.find('\t') < 0:
+      continue
+    a = ln.split('\t')
+    dic[a[0]] = a[1]
+  return dic
+
+
+
+def translate(src, dic):
+  ''' make a translation '''
+  if dic == None:
+    return
+
+  for i in range(len(src)):
+    if src[i].strip() == "":
+      continue
+    for k in dic:
+      src[i] = src[i].replace(k, dic[k])
+
+
+
+def loadsrc(fn):
+  ''' load a script or style sheet '''
+
+  try:
+    src = open(fn).readlines()
+  except:
+    print "cannot open", fn
+    raise
+  for i in range(len(src)):
+    ln = src[i].strip()
+    if ln == "":
+      src[i] = "" # remove blank lines
+    elif rmcomments:
+      if ln.startswith("/*"):
+        # remove block comments
+        for j in range(i, len(src)):
+          if src[j].strip().endswith("*/"):
+            break
+        for k in range(i, j+1):
+          src[k] = ""
+      else:
+        # remove line comments
+        p = src[i].rfind("//")
+        if p >= 0:
+          src[i] = src[i][:p] + "\n"
+
+  # remove spaces
+  if rmspaces:
+    for i in range(len(src)):
+      ln = src[i].strip()
+      if src[i].startswith("}") or ln == '"use strict";':
+        ln += "\n"
+      src[i] = ln
+
+  return ''.join(src)
+
+
+def htmlpack(fn, fntrans = None):
+  ''' embed scripts and stylesheets into the HTML file `fn` '''
+
+  p = fn.find("_pack.htm")
+  if p >= 0: fn = fn[:p] + fn[p+5:]
+  print "packing", fn
+
+  s = open(fn).readlines()
+  n = len(s)
+  for i in range(n):
+    ln = s[i].strip()
[…]

### 2017-09-02 web/js/cosmodes.js

@@ -1,0 +1,101 @@
+/* decomposition of histogram fluctuation into eigenmodes
+ * for translationally-invariant (homogeneous) updating schemes */
+
+
+"use strict";
+
+
+/* compute the eigenvectors,
+   costab = n * phi */
+function mkcostab(n, pbc)
+{
+  var i, k, costab, norm = 1.0, ang, y;
+
+  costab = new Array(n * n);
+
+  /* first row, uniform distribution
+   * costab(0, i) = n phi(0, i) = 1 */
+  for ( i = 0; i < n; i++ ) {
+    costab[i] = norm;
+  }
+
+  norm *= Math.sqrt(2);
+  if ( pbc ) {
+    /* periodic case, for k = 1, ..., n - 1
+     * in the paper, we use the exponential form
+     *   n phi(k, i) = exp(I i k 2 Pi / n)
+     * here, we use the equivalent cosine and sine form
+     *   n phi(k, i) = sqrt(2) cos(i k 2 Pi / n)
+     * for k <= n/2, or
+     *   n phi(k, i) = sqrt(2) sin(i k 2 Pi / n)
+     * for k > n/2 */
+    for ( k = 1; k < n; k++ ) {
+      ang = k * 2 * Math.PI / n;
+      for ( i = 0; i < n; i++ ) {
+        if ( k <= n/2 ) {
+          y = Math.cos(i * ang);
+        } else {
+          y = Math.sin(i * ang);
+        }
+        costab[k*n + i] = norm * y;
+      }
+    }
+  } else {
+    /* non-periodic case, for k = 1, ..., n - 1
+     * n phi(k, i) = sqrt(2) cos[k (i+1/2) Pi / n] */
+    for ( k = 1; k < n; k++ ) {
+      ang = k * Math.PI / n;
+      for ( i = 0; i < n; i++ ) {
+        costab[k*n + i] = norm * Math.cos((i + 0.5) * ang);
+      }
+    }
+  }
+
+  return costab;
+}
+
+
+
+/* decompose histogram visit to id, into eigenmodes, `u`
+ * the coefficients are pre-computed in costab */
+function getcosmodesh(id, n, u, costab)
+{
+  for ( var k = 0; k < n; k++ )
+    u[k] = costab[k*n + id] / n;
+}
+
+
+
+/* decompose a general fluctuation, `v`, into eigenmodes, `u`
+ * the coefficients are pre-computed in costab */
[…]

### 2017-09-02 web/js/mmwl.js

@@ -1,0 +1,75 @@
+/* Wang-Landau scheme for moments */
+
+
+
+"use strict";
+
+
+
+function MMWL(a)
+{
+  this.mm = [0, 0, 0];
+  this.fl = [0, 0, 0];
+  this.alphawl = a;
+  this.invt = false;
+  this.t0 = 1;
+}
+
+
+
+MMWL.prototype.add = function(y1, y2)
+{
+  this.mm[0] += 1;
+  this.mm[1] += y1;
+  this.mm[2] += y2;
+}
+
+
+
+MMWL.prototype.getalpha = function()
+{
+  return this.invt ? 1./(this.mm[0] + this.t0) : this.alphawl;
+}
+
+
+
+/* compute the fluctuation of the statistical moments
+ * return the maximum fluctuation */
+MMWL.prototype.calcfl = function(dovar)
+{
+  var t = this.mm[0], fl, fl2;
+  if ( t <= 0 ) return 99.0;
+  this.fl[1] = this.mm[1] / t;
+  this.fl[2] = this.mm[2] / t;
+  fl = Math.abs(this.fl[1]);
+  if ( dovar && (fl2 = Math.abs(this.fl[2])) > fl )
+    fl = fl2;
+  return fl;
+}
+
+
+
+/* switch to a stage with a reduced updating magnitude */
+MMWL.prototype.switch = function(magred)
+{
+  var t = this.mm[0];
+  this.alphawl *= magred;
+  if ( this.alphawl < 1./t ) {
+    this.t0 = t;
+    this.invt = true;
+  }
+  this.mm = [0, 0, 0];
+}
+
+
+MMWL.prototype.check = function(dovar, fl, magred)
+{
+  var flmax = this.calcfl(dovar);
+  if ( !this.invt && flmax < fl ) {
+    this.switch(magred);
+    return 1;
+  } else {
+    return 0;
+  }
+}
+

### 2017-09-02 doc/invt.tex
correct web

@@
 the fluctuation threshold for stage transitions
-was $F = 0.5$ [using the definition given by Eq. \eqref{eq:F2st}];
+was $F = 0.2$ [using the definition given by Eq. \eqref{eq:F2st}];
 and the reduction factor was $1/2$.

### 2017-09-02 web/js/age.js

@@ -192,7 +192,7 @@ AGE.prototype.switch = function(magred)
     this.invt = true;
     this.t0 = this.t;
   }
-  console.log("alpha " + this.alpha.wl + ", " + this.n/this.t + ", "
+  console.log("alpha " + this.alphawl + ", " + this.n/this.t + ", "
     + "fluc " + this.hfluc + ", invt " + this.invt);
   this.t = 0;
   for ( i = 0; i < n; i++ ) {
@@ -232,10 +232,13 @@ AGE.prototype.move = function(x, id, local)
   if ( local ) {
     // jump to a nearest neighbor
     jd = (rand01() < 0.5) ? id - 1 : id + 1;
-    if ( jd < 0 || jd >= n ) return 0;
+    if ( jd < 0 || jd >= n ) {
+      this.tacc = false;
+      return id;
+    }
   } else {
     // jump any other umbrella
-    jd = ( id + 1 + (int) (rand01() * (n - 1)) ) % n;
+    jd = ( id + 1 + Math.floor(rand01() * (n - 1)) ) % n;
   }
   sigi = this.sig[id];
   sigj = this.sig[jd];
@@ -254,6 +257,7 @@ AGE.prototype.move = function(x, id, local)
   vi = this.c1[ id] * xi + this.c2[ id] * (xi * xi - 1) / SQRT2;
   vj = this.c1[ jd] * xj + this.c2[ jd] * (xj * xj - 1) / SQRT2;
   dv += vj - vi;
+  //console.log(id, jd, x, vi, this.lnz[id], vj, this.lnz[jd], dv, local);
   this.tacc = ( dv <= 0 || rand01() < Math.exp(-dv) );
   if ( this.tacc ) {
     /* copy parameters to the new umbrella */

### 2017-09-07 doc/invt.tex
edits to invt.tex

@@
 \usepackage{setspace}
+
+\setcitestyle{super}
 
@@
   %
-  The convergence of the bias potential can be accelerated
+  The convergence of the bias potential can be improved
   by decreasing the updating magnitude with an optimal schedule.
@@
   the single-bin updating scheme (commonly used in the WL algorithm)
-  is given by an inverse-time formula,
+  is given by the known inverse-time formula,
   that for the Gaussian updating scheme (commonly used in metadynamics)
@@
 a distribution, $p(z)$,
-along a collective variable, $z$, by sampling via either Monte Carlo\cite{
-  frenkel, newman, landau_binder} (MC)
-or molecular dynamics\cite{frenkel, karplus2002} (MD) simulations.
-%
-The distribution can be related to
+along a collective variable, $z$, by sampling via either
+Monte Carlo (MC)\cite{frenkel, newman, landau_binder}
+or molecular dynamics (MD)\cite{frenkel, karplus2002}
+simulations.
+%
+The distribution is related to
 the dimensionless free energy, or potential of mean force (PMF),
@@
 such that the distribution along $z$ is
-flat\cite{mezei1987, berg1992, *lee1993,
+flat.\cite{mezei1987, berg1992, *lee1993,
 wang2001, wang2001pre,
 huber1994,
-*laio2002, *laio2008, *barducci2011, *sutto2012}.
+*laio2002, *laio2008, *barducci2011, *sutto2012}
 %
@@
 to discourage future visits to previously sampled configurations
-by incrementally elevating the bias potential.
+by incremental elevation of the bias potential.
 %
@@
 %
-Often, the window function
+The window function often
 takes the form of a discrete
@@
 %
-While the former is popular for MC simulations\cite{wang2001,
-  wang2001pre, kim2006, *kim2007},
+While the former is popular for MC simulations,\cite{wang2001,
+  wang2001pre, kim2006, *kim2007}
 the latter is commonly used for MD simulations
 as it provides a smoother bias potential
-that can be readily differentiated for force calculation.\cite{huber1994,
+that can be readily differentiated
+to derive the bias force.\cite{huber1994,
   *laio2002, *laio2008, *barducci2011, *sutto2012, junghans2014}
@@
 disrupts the underlying equilibrium
-sampling\cite{zhou2005, morozov2007, zhou2008},
-and for convergence of the estimated PMF
-one needs to gradually decrease
-the updating magnitude of the bias potential.
+sampling,\cite{zhou2005, morozov2007, zhou2008}
+and the convergence of the bias potential
+can benefit from a gradually decreasing
+updating magnitude of the bias potential.
 
 The schedule of reducing the updating magnitude
-can critically affect the precision and accuracy
+affects the precision and accuracy
 of the final bias potential,
-and hence that of the PMF\cite{laio2005, bussi2006, poulain2006,
+and hence that of the PMF.\cite{
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
 liang2007, min2007,
-morozov2007, zhou2008,
+morozov2007, morozov2009, zhou2008,
 komura2012, *caparica2012, *caparica2014,
-barducci2008, dickson2011, dama2014}.
+barducci2008, dickson2011, dama2014, poulain2006}
 %
@@
 liang2007,
-morozov2007, zhou2008}
+morozov2007, morozov2009, zhou2008}
 or
@@
 
-We will then show by comparing
-the efficiency of error reduction
+We will then show by comparing the errors
 of different updating schemes
-that
-the single-bin updating scheme
+that the single-bin updating scheme
 is optimal for very long simulations.
@@
 and coefficients are adjusted on-the-fly
-to approximate a particular target distribution.
+to approximate the target distribution.
 %
@@
 of a multiple-bin updating scheme
-that encompasses the Gaussian scheme used in metadynamics
+that encompasses the Gaussian updating scheme used in metadynamics
 in Sec. \ref{sec:multiple-bin}.
@@
 Then, we compare different updating schemes
-and show that the single-bin scheme
-and a class of generalized bandpass schemes
+and show that the single-bin updating scheme
+and a class of generalized bandpass updating schemes
 are optimal for convergence
@@
 and introduce artificial errors that need to be
-reduced by gradually decreasing the updating magnitude\cite{
+reduced by gradually decreasing the updating magnitude.\cite{
   belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-  zhou2005, morozov2007, zhou2008,
-  laio2005, bussi2006, poulain2006, liang2007,
-  crespo2010, *atchade2011, *fort2015}.
+  zhou2005, morozov2007, morozov2009, zhou2008, min2007,
+  poulain2006, liang2007,
+  crespo2010, *atchade2011, *fort2015}
 
@@
 
-%For the single-bin scheme
+%For the single-bin updating scheme
 %used in WL algorithm,
@@
 %
-where $t$,
+where $t$
 referred to as the ``time'' below,
 is the number of updating steps
-from the beginning of the simulation\cite{
+from the beginning of the simulation.\cite{
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-morozov2007, zhou2008,
-komura2012, *caparica2012, *caparica2014}.
+morozov2009, zhou2008,
+komura2012, *caparica2012, *caparica2014}
 %
@@
 we derive the optimal $\alpha(t)$
-for the single-bin scheme,
+for the single-bin updating scheme,
 Eq. \eqref{eq:wl_update},
@@
 and a baseline value, $\bav{v}(t)$,
-for the average growth\cite{
-dama2014}.
+for the average growth.\cite{
+dama2014}
 %
@@
   = 1
-  ,
+  .
   \label{eq:fav1}
@@
 Without the random part, $\xi_{*i}(t)$,
-this equation is convergent
-as the deviation of the bias potential, $v_{*i}(t)$,
-would tend to $0$ at long times.
+the deviation of the bias potential, $v_{*i}(t)$,
+would tend to $0$ at long times, reaching convergence.
 %
@@
   \sum_{i=1}^n \rho_i
-    \int_0^T\!\!\int_0^T
+    \int_0^T\!\!\!\int_0^T
+    \!\!
     \langle
@@
 %
-If the rate of growth is positive,
+If the growth rate is positive,\footnote{Although
+  in general the growth rate is not necessarily positive,
+  it is often the largest eigenvalue
+  for ``natural'' updating matrices considered
+  in this study.}
 we can, by a proper overall scaling of $\mathbf w$,
@@
   \Err_A(T)
-  %&=
-  %\sum_{k=1}^{n-1}
-  %\int_0^T\!\!\int_0^T
-  %\langle
-  %  \tilde\xi_k(t)
-  %  \tilde\xi_k(t')
-  %\rangle
-  %\vartheta_k\bigl(q(t)\bigr)
-  %\vartheta_k\bigl(q(t')\bigr)
-  %\, dtdt'
-  %\notag \\
+  &=
+  \sum_{k=1}^{n-1}
+  \int_0^T\!\!\!\int_0^T
+  \!\!
+  \langle
+    \tilde\xi_k(t)
+    \tilde\xi_k(t')
+  \rangle
+  \vartheta_k\bigl(q(t)\bigr)
+  \vartheta_k\bigl(q(t')\bigr)
+  \, dtdt'
+  \notag \\
   &=
@@
 The value of $q(T)$ will be adjusted later
-to minimize the total error in Sec. \ref{sec:optinitalpha}.
+to minimize the total error. % in Sec. \ref{sec:optinitalpha}.
 %
@@
 %
-In Appendix \ref{sec:schedule_geometry},
-we give a geometric characterization of this solution.
+This solution can be characterized geometrically,
+as shown in Appendix \ref{sec:schedule_geometry}.
 %
@@
     \int_0^{ q(T) } M(\bar q) \, d \bar q
-  \right)^2
+  \right)^{\!\!2}
   .
@@
 if the initial bias, $\epsilon_k$, is negligible,
-which is often true if no eigenvalue is close to $0$,
+%which is often true if no eigenvalue is close to $0$,
 and in this limit, the optimal initial updating magnitude
@@
 Compute the optimal value of $q(T)$
-by solving Eq. \eqref{eq:opt_qT}.\footnote{We
-  used a variant of the Newton-Raphson method\cite{press3rd}
-  with the bisection method as a fallback.
+by solving Eq. \eqref{eq:opt_qT}.\cite{*[{We
+  used a variant of the Newton-Raphson method,
+  see e.g. }]
+  [{, with the bisection method as a fallback.
   We solved the equation for $\ln q(T)$
@@
   The initial value was given by
-  $q(T) = \ln\left(1+\frac{1}{2}T\,a_0\right)$.}
+  $q(T) = \ln\left(1+\frac{1}{2}T\,a_0\right)$.}]
+  press3rd}
 %
@@
 Compute the optimal schedule by
-numerically integrating Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.
-%
-%To handle Gaussian updating schemes more effectively,
-%we have chosen the $\mathcal N+1$ grid points
-%($\mathcal N=20000$ here)
-%over $[0, q(T)]$ to be given by
-In our implementation,
-the integration is performed using the trapezoidal rule\cite{press3rd}
-on the grid points,
+numerically integrating
+Eqs. \eqref{eq:q_opt} and \eqref{eq:mint}.\footnote{We
+  used the trapezoidal rule for the integration.
+  The grid points were
   $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
-where $i = 0, \dots, \mathcal N$.\footnote{The
-  grid points are designed for the Gaussian updating scheme.
-  For a mass function, $M(\bar q)$, that is
+  where $i = 0, \dots, \mathcal N$
+  with $\mathcal N = 20000$.
+  %
+  The grid points were designed for long simulations
+  under the Gaussian updating scheme:
+  for a mass function, $M(\bar q)$, that is roughly
   inversely proportional to $\bar q$
   (cf. Appendix \ref{sec:schedule_geometry}),
-  this grid would make the integral,
-  or the difference, $t_{[i+1]} - t_{[i]}$,
-  roughly a constant.
-  We used $\mathcal N = 20000$.}
+  this grid makes the integral,
+  $t_{[i+1]} - t_{[i]}$,
+  approximately a constant.
   %
   The resulting values, $\bigl(q_{[i]}, t_{[i]} \bigr)$,
-are used to compute the updating magnitude
+  were used to compute the updating magnitude
   as (cf. Appendix \ref{sec:ave_ub}):
-\begin{equation}
+  $$%\begin{equation}
     \alpha_{[i]}
@@
     ,
-  \label{eq:alphai}
-\end{equation}
+    %\notag
+    %\label{eq:alphai}
+  $$%\end{equation}
   with
@@
   The values of $(t_{[i]}, \alpha_{[i]})$
-are linearly interpolated
+  were linearly interpolated
   to get the updating magnitude, $\alpha(t)$, at any $t$.
+  }
+%
 
@@
 \end{align}
-where $\sigma_z$ and $\Delta z = (z_{\max} - z_{\min}) \, g/2$
+where $\sigma_z$ and $\Delta z = (n \, \delta z) \, g/2$
 are the width of the Gaussian and the effective span
@@
 respectively,
-with $g = 1$ if $z$ periodic or $2$ otherwise
+with $g = 1$ if $z$ is periodic or $2$ otherwise
 (cf. Appendix \ref{sec:Gaussian_math}).
@@
 
+The optimal schedule of the Gaussian updating scheme
+generally has no simple closed-form expression.
+However, we can show that it always lies above
+the inverse-time schedule
+(cf. Appendix \ref{sec:schedule_geometry}).
+
 \subsubsection{\label{sec:homo_scheme}
@@
 we can design a homogeneous updating scheme
-that is more effective in error reduction
+that is more effective in error control
 than the Gaussian updating scheme.
@@
 %
-We discuss this strategy in detail below.
+%We discuss this strategy in detail below.
+As shown below,
+the optimized updating scheme will spare the need to
+go through the general procedure in Sec. \ref{sec:procedure}
+because the optimal schedule is given by
+the inverse-time formula.
 
 \subsection{\label{sec:cmpschemes}
-Comparison of updating schemes}
+Optimization of updating schemes}
 
@@
 \subsubsection{\label{sec:optWL}
-Asymptotic optimality of the single-bin scheme}
+Asymptotic optimality of the single-bin updating scheme}
 
 We first show
-that the single-bin scheme is asymptotically optimal.
-%
-Consider a class of updating matrices
-sharing the same set of eigenvectors,
+that the single-bin updating scheme is asymptotically optimal.
+%
+We wish to find a lower bound of the asymptotic error,
+$\Err_A(T)$, as defined in Eq. \eqref{eq:error_asym}
+for a class of updating matrices that
+share the same set of eigenvectors,
 hence the same $\Gamma_k$'s,
-but with different positive eigenvalues,
+but not the eigenvalues,
 $\lambda_k$'s.
 %
-Using the Cauchy-Schwarz inequality, we have,
-for any set of nonnegative numbers, $c_k$'s\footnote{This
+If all eigenvalues are positive,
+we can use the Cauchy-Schwarz inequality,\footnote{This
   inequality follows from the nonnegativity of
   the quadratic polynomial,
-$$
+  \begin{align*}
   \int_0^T
     dt \sum_{k = 1}^{n-1} \Gamma_k \,
-    \left( {\dot \vartheta}_k \, \chi - c_k \right)^2
+      ( {\dot \vartheta}_k \, \chi - 1 )^2
     \equiv
@@
     ,
-$$
+  \end{align*}
   which implies a nonpositive discriminant,
@@
   \left(
-    \int_0^T dt
+    \int_0^T \!\!dt
       \sum_{k = 1}^{n-1}
@@
   \left(
-    \int_0^T dt
+    \int_0^T \!\!dt
       \sum_{k = 1}^{n-1}
-        \Gamma_k \, c_k^2
+        \Gamma_k
   \right)
   %
-  \notag
-  \\
-  &
-  \qquad \qquad \qquad
+  \notag \\
+  & \qquad \qquad \qquad
   \ge
   \left(
-    \int_0^T dt
+    \int_0^T \!\!dt
       \sum_{k = 1}^{n-1}
-        \Gamma_k \, c_k \, {\dot \vartheta}_k \, \bigl( q(t) \bigr)
-  \right)^2
-  \notag
-  \\
-  &
-  \qquad \qquad \qquad
+        \Gamma_k \, {\dot \vartheta}_k \, \bigl( q(t) \bigr)
+  \right)^{\!\!2}
+  \notag \\
+  & \qquad \qquad \qquad
   =
   \left(
-    \sum_{k = 1}^{n-1} \Gamma_k \, c_k
+    \sum_{k = 1}^{n-1} \Gamma_k
       \left[
@@
       \right]
-  \right)^2.
+  \right)^{\!\!2}
+  .
   \notag
@@
 %
-Note that the last expression %of Eq. \eqref{eq:CSineq}
-depends on the curve, $q(t)$ ($0 < t < T$),
-only through the endpoint value, $q(T)$,
-which is fixed in the variational process.
-%
-Thus, the inequality sets a lower bound
-for the asymptotic error $\Err_A(T)$
-given in Eq. \eqref{eq:error_asym}.
-%
-The equality is achieved
-if $\dot \vartheta_k\left( q(t) \right) = c_k$
-for all $k > 0$ at any $t$,
-up to a multiplicative factor.
+In the long time limit, $q(T) \to \infty$,
+the factor $e^{-\lambda_k \, q(T)}$ is negligible,
+and the lower bound is independent of the $\lambda_k$'s:
+$\Err_A(T) \ge \sum_{k=1}^{n-1} \Gamma_k/T$.
+%Note that the last expression %of Eq. \eqref{eq:CSineq}
+%depends on the curve, $q(t)$ ($0 < t < T$),
+%only through the endpoint value, $q(T)$,
+%which is fixed in the variational process.
+%
+This inequality applies to any schedule
+and any updating scheme in the class,
+and the equality is achieved
+if $\dot \vartheta_k\left( q(t) \right) = \mathrm{constant}$
+for all $k > 0$ at any $t$.
 %
@@
 a set of identical eigenvalues,
-$\lambda_1 = \cdots = \lambda_{n-1}$
-and identical $c_k$'s.
+$\lambda_1 = \cdots = \lambda_{n-1}$.
 
@@
 \end{align*}
-and $\lambda_1$ must be $1$ by Eq. \eqref{eq:w_sumj}.
-%
-Thus, $w_{ij} = \delta_{ij}$
-is reduced to the updating matrix of the single-bin updating scheme.
+which is essentially the diagonal updating matrix
+that represents the single-bin updating scheme.
 %
@@
 We can generalize
-the single-bin scheme to a class of
+the single-bin updating scheme to a class of
 perfect (brick-wall) \emph{bandpass} updating schemes
@@
 If the eigenmodes are sorted in descending order
-of the wavelength,
+of the typical wavelength,
 and the PMF is sufficiently smooth
@@
 but only for the first $K$ fluctuation modes.
+
+A distinct advantage of the bandpass updating schemes
+is that they share with the single-bin updating scheme
+the asymptotically optimal schedule,
+the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
+%
+Thus, it is unnecessary to go through the general procedure
+described in Sec. \ref{sec:procedure}
+for these updating schemes.
 
@@
 In Sec. \ref{sec:lj},
-we will further compare the performance of
-the homogeneous bandpass and Gaussian updating schemes
+we will further compare
+the homogeneous bandpass updating scheme
+with the Gaussian updating scheme
 in an example.
@@
 %
-The polynomial bandpass updating schemes will be discussed below.
+The other class of polynomial bandpass updating schemes will be discussed below.
 
@@
 than the homogeneous bandpass updating scheme
-as the bias potential in this case is
+as the former expresses the bias potential as
 a simple polynomial of the collective variable.
@@
 replace the discrete bin index, $i$,
-by the continuous variable, $z$;
-%
-$\rho_i$ by $\rho(z) \, \delta z$;
-$\phi_{ki}$ by $\phi_k(z) \, \delta z$;
+by the continuous variable, $z$,
+%
+$\rho_i$ by $\rho(z) \, \delta z$,
+$\phi_{ki}$ by $\phi_k(z) \, \delta z$,
 and
@@
 %
-By Eq. \eqref{eq:mbin_update},
+Then, by Eq. \eqref{eq:mbin_update},
 the bias potential is updated as
@@
 the updating scheme is essentially the LLR algorithm\cite{langfeld2012}
-for the single-interval case,
+in the single-interval case,
 and the inverse-time schedule
@@
 %
-In practical applications, the algorithm is often applied simultaneously
+In practical applications,
+the LLR algorithm is often applied simultaneously
 to an array of adjacent intervals
@@
 we will consider the case of sampling
-a Gaussian distribution
+a \emph{non-flat} Gaussian distribution
 %$\rho(z) \propto \exp[-(z-z_c)^2/(2\sigma_z^2)]$,
@@
 \end{equation}
-which is \emph{not} flat,
-with a basis constructed from
-the Hermite polynomials\cite{arfken}.
+using a basis of the Hermite polynomials\cite{arfken}.
 %
@@
 for inter-umbrella transitions,
-and it is updated according to the single-bin scheme\cite{li2007}.
-%
-For a small $\sigma_z$, the converged
+and it is updated according to the single-bin updating scheme\cite{li2007}.
+%
+With a small width, $\sigma_z$, the converged
 $c_0^{(i)}$ as a function of $z_c^{(i)}$
@@
 
-Since bandpass updating schemes are a straightforward extension
+Since the bandpass updating schemes are a straightforward extension
 of the single-bin updating scheme,
@@
 %to control the updating magnitude
-%for the bandpass scheme,
-%since it is a straightforward extension of the single-bin scheme.
+%for the bandpass updating scheme,
+%since it is a straightforward extension of the single-bin updating scheme.
 %%
@@
 \item
-  Set up an array of umbrellas covering the desired range,
+  Set up an array of Gaussian umbrellas covering the desired range,
     e.g. we may set, for $i = 1, \dots, m$,
+    the mean,
     $z_c^{(i)} = z_{\min} + (i - 1) \, \Delta z$,
+    and
     the width, $\sigma_z^{(i)}$, to be a constant.
-    For the parameters, we can set
-    $c_1^{(i)}$ to the estimated mean force,
+    For the polynomial coefficients, we can set
+    $c_1^{(i)}/\sigma_z^{(i)}$ to the estimated mean force,
     and $c_0^{(i)} = c_2^{(i)} = 0$.
@@
   Update the $c_k^{(i)}$'s
-  according to Eq. \eqref{eq:ckupdate}.
+  according to Eq. \eqref{eq:ckupdate},
+  and increase $t$ by $1$.
 
@@
     Bandpass & $1/t$
-    & $8.1(8.9)\times10^{-3}$
-    & $2.1(2.3)\times10^{-6}$
+    & $8.0(8.9)\times10^{-3}$
+    & $2.0(2.3)\times10^{-6}$
     \\
@@
 can suppress local noise,
-it can also mask finer features of the PMF.
+it can also mask finer features of the PMF.\cite{laio2005}
 %it also diminishes the eigenvalues
@@
 %
-The acceptance ratio was no less than 92\% in any umbrella.
+The acceptance ratio was no less than 91\% in any umbrella.
 %
@@
 We computed the density of states by WHAM,
-and found the critical temperature to be $T_c = 0.70161$,
+and found the critical temperature to be $T_c = 0.70163$,
 in agreement with a previous study\cite{wang2001pre}.
@@
 %
-For the single-bin scheme (as often used in the WL case),
+For the single-bin updating scheme (as often used in the WL case),
 whose updating matrix is the identity matrix with identical eigenvalues,
@@
 the single-bin updating scheme to be optimal
-in error reduction in the long time limit.
+for the error control in the long time limit.
 %without a priori assumptions
@@
 and the sampling is a quasi-equilibrium one.\cite{
-  zhou2005, morozov2007, zhou2008, barducci2008, dama2014},
+  zhou2005, morozov2007, zhou2008, barducci2008, dama2014}
+Early in simulation, however,
+we still have to resort to the WL stage-by-stage
+technique even for the bandpss updating schemes.
 %the random error,
@@
 \section{\label{sec:equilerr}
-Runtime-average interpretation
+Moving average interpretation
 }
@@
 %
-We wish to show that the recurrence updating relation,
+Here, the bias potential undergoing
+the recurrence updating relation,
 Eq. \eqref{eq:wl_update},
-renders the bias potential
-as a cumulative moving average (CMA) with a dynamic weight
+is mapped to an equivalent moving average
+with a dynamic weight
 determined by the updating schedule.
@@
 
-\subsection{Runtime average and recurrence relation}
-
-Consider the weighted average
+\subsection{Moving average and recurrence relation}
+
+Consider the weighted cumulative moving average (CMA)
 %
@@
   1 - \frac{ \Omega(t-1) } {\Omega(t)}
-  .
+  ,
   \label{eq:alpha_from_Omega}
@@
 
-To apply the above formalism to updating schemes,
+To apply the above formalism to the updating schemes,
 we need a correction formula for the bias potential.
@@
 the instantaneous histogram, $h_i(t)$,
-we get an independent estimate of the ideal bias potential
+yields an independent estimate of the ideal bias potential
 from a single data point at step $t$,
@@
 An expedient fix,
-which we used in Eq. \eqref{eq:alphai},
+which we used in Sec. \ref{sec:procedure},
 is to derive the effective updating magnitude
 from Eq. \eqref{eq:alpha_from_Omega}
+for the $k = 1$ mode with the largest eigenvalue,
+i.e.
+$\alpha(t) = \lambda_1^{-1}[1-\Omega(t-1)/\Omega(t)]$,
 with
@@
 %
-Using this formula to mode $k = 1$
-of the largest eigenvalue
-guarantees $\lambda_k \, \alpha(t) \ge 1$
+This guarantees $\lambda_k \, \alpha(t) \le 1$
 for all nontrivial modes.
@@
 
-We have shown above that
-for a general multiple-bin updating scheme,
+For a general multiple-bin updating scheme,
 each mode of bias potential
@@
 
-We can use Eq. \eqref{eq:intmQ_tT} to show that
+We can show that
 for simulations with equal lengths and initial updating magnitudes,
-the optimal schedule for the Gaussian scheme
+the optimal schedule for the Gaussian updating scheme
 always lies above the inverse-time schedule,
-which is optimal for the single-bin scheme.
-%
-%Let us use Eqs. \eqref{eq:intmQ_tT} and \eqref{eq:dlnmdq}
-Let us compare the times for the optimal
-schedules to lower the updating magnitude to a certain level,
-$\alpha_1$,
-in the two simulations:
-%
-\begin{align}
-  \frac{t}{T}
-  %=
-  %\int_{\bar q(t)}^{q(T)} m(\bar q) \, d\bar q
-  =
-  \int_{m( \bar q(t) )}^{m( q(T) )}
-  \!\!\frac{ m(\bar q) } { \frac{dm(\bar q)}{d\bar q} } \, d m(\bar q)
-  =
-  \int^{m( \bar q(t) )}_{m( q(T) )}
-  \!\!\frac{ d m } { -\frac{d\ln m(\bar q)}{d\bar q} }
-  %=
-  %\int^{m( \bar q(t) )}_{m( q(T) )}
-  %\!\!\frac{ d m } { \bar \lambda(\bar q) }
-  ,
-  \label{eq:t_int}
-\end{align}
-with the lower and upper limits,
-$m\bigl(q(T)\bigr) = 1/\bigl[T\alpha(0)\bigr]$ and
-$m\bigl(\bar q(t)\bigr) = 1/(T\alpha_1)$,
-being the same in the two cases.
-%
-But from Eqs. \eqref{eq:mass_func} and \eqref{eq:mass_distr},
+which is optimal for the single-bin updating scheme.
+%
+From Eqs. \eqref{eq:mQ_invTa}, \eqref{eq:intmQ_tT},
+\eqref{eq:mass_distr}, and \eqref{eq:mass_func},
 we have
-\begin{align}
-  -\frac{ d }{ d \bar q } \ln m(\bar q)
-  %=
-  %-\frac{ dM^2(\bar q)/d\bar q } { 2 M^2(\bar q) }
+%
+\begin{align*}
+  %\frac{d\alpha^{-1}}{dt}
+  \frac{d}{dt}\left(\frac{1}{\alpha(t)}\right)
+  &=
+  \frac{ dm } { d(t/T) }
+  =
+  -\frac{dm}{m\,d\bar q}
+  =
+  -\frac{1}{2} \frac{ dM^2(\bar q)/d\bar q }
+  { M^2(\bar q) }
+  %\notag
+  \\
+  &
   =
@@
   ,
-  \notag
-  %\label{eq:dlnmdq}
-\end{align}
+\end{align*}
+%
 which is a weighted average of $\lambda_k$'s
@@
 %
-Since for the Gaussian scheme,
-we have $\bar \lambda(\bar q) \le 1$
-by Eq. \eqref{eq:lambda_Gaussian},
-the integral in Eq. \eqref{eq:t_int}
-is no less than that for the single-bin scheme,
-i.e. it takes longer for the Gaussian scheme
-than the single-bin scheme
-to lower the updating magnitude
-to the same level.
+For the Gaussian updating scheme,
+we have $\bar \lambda(\bar q) < 1$ for $\bar q > 0$
+by Eq. \eqref{eq:lambda_Gaussian}.
+%
+Thus the inverse updating magnitude
+increases more slowly
+in the Gaussian updating scheme
+than in the single-bin updating scheme,
+and the updating magnitude is larger in the former case.
 
@@
 %
-For example, in the single-bin scheme,
+For example, in the single-bin updating scheme,
 the mass distribution decays exponentially,
@@
     %$\left( \dfrac{c}{t^* + t^*_0} \right)^{1/a} - q^*_0$
-    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^\frac{1}{a} - 1\right]$
+    $q^*_0 \left[\left( \dfrac{1+t^*_0}{t^* + t^*_0} \right)^{\!\frac{1}{a}} \!- 1\right]$
     %$\,\begin{aligned}
@@
 
-The idea is to use a modification
-of Eq. \eqref{eq:xt2_eql}
-that avoids the problem of not
+We will use Eq. \eqref{eq:xt2_eql}
+but modify it to avoid the problem of not
 knowing the intrinsic distribution, $p_i$,
-hence the potential shift,
+hence the shift,
 $u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
@@
 The variance, $\tilde u_k$, can be computed
-by trajectory averaging.
+by trajectory averaging in a long simulation.
 %
@@
 
-Below we use the transition matrix method
-to estimate $\Gamma_k$ in a model for local diffusion\cite{bussi2006}.
+Next we will use a model to show that $\Gamma_k$
+assumes a large value for a small $k$,
+and then decays rapidly to $1$ as $k$ increases
+in a local sampling process.
 %
@@
 \end{align*}
-Further, if the transition matrix % satisfies detailed balance
-has a complete set of orthonormal eigenvectors, $\psi_{li}$
-%that are orthonormal with respect to $\pmb\rho$
-[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
-\eqref{eq:eig_orthonormal_rows}],
-it can be decomposed as
-$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
-with $\Lambda_l$ being the eigenvalue of mode $l$.
-%
-The autocorrelation function can then be computed as
-%
-\begin{align*}
-\bigl\langle
-  \tilde f_k(t) \, \tilde f_k(0)
-\bigr\rangle
-=
-\sum_{l = 0}^{n-1} \Lambda_l^t \,
-\left(
-  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
-\right)^2
-.
-\end{align*}
-%
-In particular, if the eigenvectors, $\pmb \phi_k$
-and $\pmb \psi_k$, coincide,
-we would have
+%
+We further assume that
+the transition matrix
+has a complete set of orthonormal eigenvectors,
+and the eigenvectors coincide with
+those of the updating matrix, $\phi_{ki}$'s.
+Then the transition matrix can be decomposed as
+$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \phi_{li} \, \phi_{lj} / \rho_j$,
+with $\Lambda_l$ being the eigenvalue of mode $l$,
+and
+%
 $\bigl\langle
@@
 =
-\Lambda_k^t,$
-and
-\begin{equation}
-  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
-  \bigl\langle
-    \tilde f_k(t) \, \tilde f_k(0)
-  \bigr\rangle
-  = \frac{1+\Lambda_k}{1-\Lambda_k}
-  .
-  \notag
-  %\label{eq:Gamma_est}
-\end{equation}
+\Lambda_k^t$.
+%If the transition matrix % satisfies detailed balance
+%has a complete set of orthonormal eigenvectors, $\psi_{li}$
+%%that are orthonormal with respect to $\pmb\rho$
+%[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
+%\eqref{eq:eig_orthonormal_rows}],
+%it can be decomposed as
+%$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
+%with $\Lambda_l$ being the eigenvalue of mode $l$.
+%%
+%The autocorrelation function can then be computed as
+%%
+%\begin{align*}
+%\bigl\langle
+%  \tilde f_k(t) \, \tilde f_k(0)
+%\bigr\rangle
+%=
+%\sum_{l = 0}^{n-1} \Lambda_l^t \,
+%\left(
+%  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
+%\right)^2
+%.
+%\end{align*}
+%%
+%In particular, if the eigenvectors, $\pmb \phi_k$
+%and $\pmb \psi_k$, coincide,
+%we would have
+%$\bigl\langle
+%  \tilde f_k(t) \, \tilde f_k(0)
+%\bigr\rangle
+%=
+%\Lambda_k^t$.
+%and
+%\begin{equation}
+%  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
+%  \bigl\langle
+%    \tilde f_k(t) \, \tilde f_k(0)
+%  \bigr\rangle
+%  = \frac{1+\Lambda_k}{1-\Lambda_k}
+%  .
+%  \notag
+%  %\label{eq:Gamma_est}
+%\end{equation}
 %
@@
 \begin{align}
-\Gamma_k = \coth\left[
-  \left(
-    \frac{ \pi \, k \, \sigma_z^\mathrm{(samp)} } { g \, \Delta z}
-  \right)^2
-  \right].
+  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
+  \bigl\langle
+    \tilde f_k(t) \, \tilde f_k(0)
+  \bigr\rangle
+  %= \frac{1+\Lambda_k}{1-\Lambda_k}
+%\Gamma_k
+  = \coth\left(
+  \frac{ \pi^2 \, k^2 \, {\sigma_z^\mathrm{(samp)}}^2 } { 4 \, \Delta z^2}
+  \right).
   \notag
@@
 %
-However, this method tends to underestimate
-the first few $\Gamma_k$'s.
+However, this model tends to underestimate
+the first few $\Gamma_k$'s in a typical simulation.
 %which means that doubling the sampling range, $\Delta z$,
@@
 the $\mu_l$'s are symmetric, $\mu_l = \mu_{-l}$,
-and they can be cutoff at $l = b$,
-where $b$ is the integral part of $(n-1)/2$,
-i.e. $\mu_l = 0$ for $l > b$.
+and they can be cutoff at $l = b$
+(i.e. $\mu_l = 0$ for $l > b$),
+where $b$ is the integral part of $(n-1)/2$.
 %
@@
 To satisfy Eq. \eqref{eq:w_sumj}, we further impose
-%
-\begin{equation}
-  \mu_{-b} + \cdots + \mu_b = 1
-  .
-  \notag
-  %\label{eq:mu_normalization}
-\end{equation}
+the normalization, $\sum_{l=-b}^b \mu_l = 1$.
+%
+%\begin{equation}
+%  \mu_{-b} + \cdots + \mu_b = 1
+%  .
+%  \notag
+%  %\label{eq:mu_normalization}
+%\end{equation}
 
@@
 the out-of-boundary values by
-%
-\begin{equation}
-  \phi_i = \phi_{i \pm n},
-\label{eq:phi_pbc}
-\end{equation}
+$\phi_i = \phi_{i \pm n}$,
 %
@@
 The characteristic equation,
-$\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$,
-can then be solved by discrete Fourier transform.
-%
-We may construct a set of orthonormal eigenvectors,
+$\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$
+can then be solved by discrete Fourier transform,
+with the orthogonal eigenvectors,
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-compatible with the periodicity
-required by Eq. \eqref{eq:phi_pbc},
+are given by
 %
@@
 %
-Then one can readily verify that the eigenvalues are given by
+%Then one can readily verify that
+The eigenvalues are given by
 \begin{equation}
@@
 with a doubled period, i.e. $g = 2$,
-and redefine $b = n - 1$
+and an increased cutoff $b = n - 1$
 (so $\mu_n = 0$).
@@
 The eigenvectors sharing the same periodicity
-and being even about $i = 1/2$ is given by
+and being even about $i = 1/2$ are given by
 %
@@
 and
-$\Delta z = b \, \delta z$
-with $b \approx g\,n/2$.
+$\Delta z = (n \, \delta z) \, g / 2$.
 %
@@
 %
-with the normalization
-$1 = \int_{-\Delta z}^{\Delta z} \mu(z) \, dz$.
+with the normalization,
+$\int_{-\Delta z}^{\Delta z} \mu(z) \, dz = 1$.
 
@@
 %
-To maintain the positivity of eigenvalues,
-we can redefine the Gaussian scheme
+To keep the eigenvalues positive,
+we can redefine the Gaussian updating scheme
 from the eigenvalues given by Eq. \eqref{eq:lambda_Gaussian},
@@
 Eq. \eqref{eq:Emin_sbin}, as
+  $\Err(T) \approx \bigl\langle F^2(T) \bigr\rangle$.
 %we may rephrase the optimality of the inverse-time schedule
 %as the inequality,
-\begin{equation}
-  \Err(T)
-  \approx
-  \bigl\langle F^2(T) \bigr\rangle
-  .
-  \notag
-  %\label{eq:fluc}
-\end{equation}
+%\begin{equation}
+%  \Err(T)
+%  \approx
+%  \bigl\langle F^2(T) \bigr\rangle
+%  .
+%  \notag
+%  %\label{eq:fluc}
+%\end{equation}
 %and the histogram fluctuation
@@
 
-The total histogram fluctuation given by Eq. \eqref{eq:F2sum}
-can be used in the WL algorithm
-as a criterion for stage transitions.
-%
-For a bandpass updating scheme,
-we need to further restrict the sum to the first $K$ modes,
+As an estimate of the error,
+the total histogram fluctuation given by Eq. \eqref{eq:F2sum}
+can be used in the WL criterion for stage transitions.
+%
+However, for a bandpass updating scheme,
+we should restrict the sum to the first $K$ modes,
 where the histogram flattening applies, i.e.

### 2017-09-07 prog/aus/pt2gaus.c

@@ -13,7 +13,7 @@ const char *fndat = "pt2gaus.dat";
 
 
 enum { SAMP_METROPOLIS, SAMP_WOLFF };
-
+const char *samp_names[] = {"Metropolis", "Wolff"};
 double localmovef = 1.0;
 
 
@@ -131,6 +131,9 @@ static void potts2_gaus(potts2_t *pt,
   }
   gaus = gaus_open(ecmin, ecmax, espacing, esig, lnzmethod,
       beta_c * esig, alpha0, -2*pt->n, 0, 1, 0);
+
+  fprintf(stderr, "samping-method %s, lnz-method %s, sig %g, spacing %g, localmove %g%%\n",
+      samp_names[sampmethod], lnz_names[lnzmethod], esig, espacing, localmovef*100.0);
 
   id = 0;
   sgn = 0;

### 2017-09-07 doc/invt.tex
invt.tex: remove second part of Appendix C, simplify web

@@
 While an overly slow reduction of the updating magnitude
-is ineffective in controlling the variance of the bias potential,
+is ineffective in restricting the variance of the bias potential,
 a too rapid reduction often
@@
 For the single-bin case, it is known that
-the optimal updating magnitude is given by the
+the optimal schedule is given by the
 inverse time\cite{
@@
 for an adaptive FDS simulation
-of a general updating scheme,
+under a general updating scheme,
 including the single-bin and Gaussian versions.
@@
 we reset the histogram and start a new stage
-with $\alpha(t)$ reduced by a factor of $1/2$\cite{
-wang2001, wang2001pre}.
+with $\alpha(t)$ reduced by a factor of $1/2$.\cite{
+  wang2001, wang2001pre}
 %
@@
 A more effective way
-of controlling the schedule of the updating magnitude, $\alpha(t)$,
-in late stages
-is to follow the formula
-%
-\begin{equation}
-  \alpha(t) = \frac{1}{t},
-  \label{eq:alpha_invt}
-\end{equation}
-%
-where $t$
-referred to as the ``time'' below,
-is the number of updating steps
-from the beginning of the simulation.\cite{
+of controlling the updating magnitude in late stages
+is to follow the inverse-time schedule,\cite{
   belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
@@
 %
-This is a general result\cite{
-  robbins1951, pellegrini2014},
+\begin{equation}
+  \alpha(t) = \frac{1}{t},
+  \label{eq:alpha_invt}
+\end{equation}
+%
+where $t$, referred to as the ``time'' below,
+is the number of updating steps
+from the beginning of the simulation.
+%
+This is a general result,\cite{robbins1951, pellegrini2014}
 and it has an intuitive explanation:\cite{
@@
 the bias potential undergoing the single-bin updating scheme
-can be mapped to a cumulative moving average computed from all data points collected so far;
+can be mapped to a cumulative moving average over all data points collected so far;
 upon the addition of a new data point,
@@
 upon convergence.
-Note that updates to $v_i(t)$ are
-equivalent to the updates to $u_i(t)$,
-since the shift
-remains a constant during simulation.
+%Note that $v_i(t)$ undergoes the same updating equation
+%as $u_i(t)$ for the shift remains a constant during simulation.
 %%
@@
 %
-Below we will use the notations, $\bav{x}$ and $x_{*i}$,
-for a general variable, $x_i$.
+Below we will use the notations, ``$\bav{}$'' and ``$_{*i}$,''
+for other quantities.
 %
@@
 %
-Clearly,
-%
-\begin{equation}
-  \bav{f}(t)
-  = \sum_{i=1}^n \rho_i \cdot \frac{ h_i(t) } { \rho_i }
-  = \sum_{i=1}^n \delta_{i, i(t)}
-  = 1
-  .
-  \label{eq:fav1}
-\end{equation}
-%
 The shifted bias potential, $v_i(t)$,
 follows the same differential equation
-as the shift is a constant.
+for the shift is a constant:
+$\dot v_i(t) \approx \alpha(t) \, f_i(t)$.
 %
@@
 %
-Deducting the bin average and using Eq. \eqref{eq:fav1}, we get
-$\langle f_{*i}(t) \rangle \equiv \langle f_i(t) \rangle - 1 \approx - v_{*i}(t)$,
+Deducting the bin average
+and noticing that
+$\bav{f}(t) = \sum_{i=1}^n \rho_i [h_i(t)/\rho_i] = \sum_{i=1}^n \delta_{i, i(t)} = 1$,
+we get
+$\langle f_{*i}(t) \rangle = \langle f_i(t) \rangle - 1 \approx - v_{*i}(t)$,
+%
+%\begin{align*}
+%  \langle f_{*i}(t) \rangle
+%  \equiv
+%  \langle f_i(t) \rangle -
+%  \bav{f}(t)
+%  =
+%  \langle f_i(t) \rangle - 1
+%  \approx
+%  - v_{*i}(t)
+%  ,
+%\end{align*}
+%
 and
@@
 for some constant $G_i$,
-and $\delta(t)$ is the Dirac delta function.
+and $\delta(t)$ is Dirac's delta function.
 %
@@
 The error depends implicitly on the curve, $\alpha(t)$,
-or, equivalently, the curve, $q(t)$,
-via $v_{*i}(T)$.
+or, equivalently, the curve, $q(t)$.
+%via $v_{*i}(T)$.
 %
@@
 a generalized Fourier transform, $\mathcal{F}$,
-for variable, $x_i$,
+for variable, $x_i$, as
 %
@@
 %
-and the inverse transform is
+and the inverse transform as
 %$x_i = \sum_{k = 0}^{n-1} \phi_{ki} \, \tilde{x}_k / \rho_i$.
@@
 \begin{align}
-  \mathcal Y\bigl( q(T) \bigr)
+  \eta\bigl( q(T) \bigr)
   &\equiv
@@
   \notag \\
-  &
+  &\quad
   -\sum_{k = 1}^{n-1}
@@
 We can rewrite the condition in terms of
-the initial updating magnitude by Eq. \eqref{eq:Lagrangian_const} %\eqref{eq:mQ_invTa},
+the initial updating magnitude
+by Eq. \eqref{eq:Lagrangian_const}, %\eqref{eq:mQ_invTa},
 \begin{align}
@@
 and in this limit, the optimal initial updating magnitude
-is half of the previous equilibrium value.\footnote{This factor, $1/2$,
+is half of the previous equilibrium value.\note{This factor, $1/2$,
   happens to coincide with the
@@
   in the original WL algorithm\cite{
-  wang2001, wang2001pre}.  }
+  wang2001, wang2001pre}.  }%
 %[cf. Eq. \eqref{eq:t0_sbin}].
@@
 by solving Eq. \eqref{eq:opt_qT}.\cite{*[{We
-  used a variant of the Newton-Raphson method,
-  see e.g. }]
-  [{, with the bisection method as a fallback.
+  used a variant of the Newton-Raphson method
+  [see e.g. }]
+  [{] with the bisection method as a fallback.
   We solved the equation for $\ln q(T)$
@@
   and imposed a lower bound,
-  $\bigl|\mathcal Y\bigl(q(T)\bigr)/q(T)\bigr|$, for
-  $\mathcal Y'\bigl(q(T)\bigr)$
+  $\bigl|\eta\bigl(q(T)\bigr)/q(T)\bigr|$, for
+  $\eta'\bigl(q(T)\bigr)$
   in computing its change.
@@
   The grid points were
-  $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/\mathcal N}$,
-  where $i = 0, \dots, \mathcal N$
-  with $\mathcal N = 20000$.
+  $q_{[i]} = \bigl(q(T) + 1\bigr) - \bigl(q(T)+1\bigr)^{1-i/N_g}$,
+  where $i = 0, \dots, N_g$
+  with $N_g = 20000$.
   %
@@
   $q_{[ -1]} \equiv t_{[-1]} \equiv 0$,
-  $q_{[\mathcal N+1]} \equiv q(T)$,
+  $q_{[N_g+1]} \equiv q(T)$,
   and
-  $t_{[\mathcal N+1]} \equiv T$.
+  $t_{[N_g+1]} \equiv T$.
   %
@@
 %
-The eigenvalues of this scheme are given by\cite{bussi2006}
+In terms of the continuous collective variable, $z = i \, \delta z$,
+the eigenvalues of this scheme are given by\cite{bussi2006}
 \begin{align}
@@
 \end{align}
-where $\sigma_z$ and $\Delta z = (n \, \delta z) \, g/2$
-are the width of the Gaussian and the effective span
-in terms of the continuous collective variable, $z$,
-respectively,
-with $g = 1$ if $z$ is periodic or $2$ otherwise
+where $\sigma_z$ is the width of the Gaussian
+and $\Delta z = n \, \delta z \, g/2$
+is the effective span
+with $g = 1$ for a periodic variable or $2$ otherwise
 (cf. Appendix \ref{sec:Gaussian_math}).
+
+The optimal schedule of the Gaussian updating scheme
+generally has no simple closed-form expression.
+However, we can show that it always lies above
+the inverse-time schedule
+(cf. Appendix \ref{sec:schedule_geometry}).
 
@@
 
-The optimal schedule of the Gaussian updating scheme
-generally has no simple closed-form expression.
-However, we can show that it always lies above
-the inverse-time schedule
-(cf. Appendix \ref{sec:schedule_geometry}).
-
 \subsubsection{\label{sec:homo_scheme}
@@
 we get a homogeneous bandpass updating scheme
-(cf. Appendix \ref{sec:homo_bandpass})
-that resembles the Gaussian updating scheme,
+(cf. Appendix \ref{sec:homo_bandpass}),
+which resembles the Gaussian updating scheme,
 but with long-range wiggles,
@@
 The eigenvectors allow us to construct the updating matrix
-from Eq. \eqref{eq:w_from_phi},
+from Eq. \eqref{eq:w_from_phi}
 with the eigenvalues given by \eqref{eq:lambda_bandpass}:
@@
 from its expectation.
-Since high-order moments are generally less reliable,
+Because high-order moments are generally unreliable,
 this method works better with a small $K$ for a narrow distribution.
@@
 to sample a flat distribution, $\rho(z) = 1/\Delta z$,
-over $\left[z_c - \frac 1 2 \Delta z, z_c + \frac 1 2 \Delta z\right]$,
+over $\left(z_c - \frac 1 2 \Delta z, z_c + \frac 1 2 \Delta z\right)$,
 we can use the Legendre polynomials\cite{arfken}.
@@
 and the inverse-time schedule
-was recommended for optimal convergence of $c_1$\cite{pellegrini2014}.
+was recommended for optimal convergence of $c_1$.\cite{pellegrini2014}
 %
@@
 
-We can extend the sampling range
+We can expand the sampling range
 by constructing an expanded ensemble
-from a superposition of $m$ Gaussian umbrellas
+from a superposition of $S$ Gaussian umbrellas
 centered at different places.\footnote{This sampling
@@
 The target distribution is a joint distribution
-of the discrete umbrella index, $i = 1, \dots, m$,
+of the discrete umbrella index, $s = 1, \dots, S$,
 and the continuous variable, $z$,
-$\rho(i, z) = \rho_i \, \rho^{(i)}(z)$,
+$\rho(s, z) = \rho_s \, \rho^{(s)}(z)$,
 where the marginal inter-umbrella distribution is flat,
-$\rho_i = 1/m$,
-and the conditional intra-umbrella distribution, $\rho^{(i)}(z)$,
+$\rho_s = 1/S$,
+and the conditional intra-umbrella distribution, $\rho^{(s)}(z)$,
 is the Gaussian given in Eq. \eqref{eq:rho_Gaussian}. % previously.
@@
 To sample this distribution, we give each umbrella
-an independent bias potential, $u^{(i)}(z)$,
+an independent bias potential, $u^{(s)}(z)$,
 of the form of Eq. \eqref{eq:u_age},
@@
 \begin{align*}
-  c_k^{(i)}(t+1) = c_k^{(i)}(t)
-  + \frac{ \alpha(t) } { \rho_i } \, R_k^{(i)}\bigl( z(t) \bigr) \, \delta_{i, i(t)}
-  .
+  c_k^{(s)}(t+1) = c_k^{(s)}(t)
+  + \frac{ \alpha(t) } { \rho_s } \, R_k^{(s)}\bigl( z(t) \bigr) \,
+    \delta_{s, s(t)}
+  ,
 \end{align*}
 %
-%where the factor $m$ before $\alpha(t)$
-%is due to the inverse of the marginal inter-umbrella distribution, $\rho_i$.
-%
-Since $R_0^{(i)}(z) = 1$,
-the parameter, $c_0^{(i)}$,
+where $s(t)$ is the umbrella at time $t$.
+%
+Since $R_0^{(s)}(z) = 1$,
+the parameter, $c_0^{(s)}$,
 serves as the bias potential
 for inter-umbrella transitions,
-and it is updated according to the single-bin updating scheme\cite{li2007}.
+and it is updated according to the single-bin updating scheme.\cite{li2007}
 %
 With a small width, $\sigma_z$, the converged
-$c_0^{(i)}$ as a function of $z_c^{(i)}$
+$c_0^{(s)}$ as a function of $z_c^{(s)}$
 also serves as a good approximation of
@@
 the transition probability from
-the $i$th umbrella to the $j$th is given by
+the $s$th umbrella to the $s'$th umbrella is given by
 %
 \begin{align}
-A(i\to j) =
+A(s \to s') =
 \min\left\{1,
-  e^{
-    u^{(i)}\bigl( z(t), t \bigr)
-    - u^{(j)}\bigl( z(t), t \bigr)
+  \frac{
+    \exp\bigl[ - u^{(s')}\bigl( z(t), t \bigr) \bigr]
   }
+  {
+    \exp\bigl[ - u^{(s)}\bigl( z(t), t \bigr) \bigr]
+  }
+  %e^{
+  %    u^{(s)}\bigl( z(t), t \bigr)
+  %  - u^{(s')}\bigl( z(t), t \bigr)
+  %}
   \right\}
@@
 %
-%and upon convergence, $Z^{(i)}$ should be proportional to
+%and upon convergence, $\exp c_1^{(s)}$ should be proportional to
 %the partition function,
-%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(i)}(z(\mathbf x),t)} \, d\mathbf x$,
+%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(s)}(z(\mathbf x),t)} \, d\mathbf x$,
 %as in simulated tempering,
@@
   Set up an array of Gaussian umbrellas covering the desired range,
-    e.g. we may set, for $i = 1, \dots, m$,
+    e.g. we may set, for $s = 1, \dots, S$,
     the mean,
-    $z_c^{(i)} = z_{\min} + (i - 1) \, \Delta z$,
+    $z_c^{(s)} = z_{\min} + \frac{s - 1}{S - 1} (z_{\max} - z_{\min})$,
     and
-    the width, $\sigma_z^{(i)}$, to be a constant.
+    the width, $\sigma_z^{(s)}$, to be a constant.
     For the polynomial coefficients, we can set
-    $c_1^{(i)}/\sigma_z^{(i)}$ to the estimated mean force,
-    and $c_0^{(i)} = c_2^{(i)} = 0$.
+    $c_1^{(s)}/\sigma_z^{(s)}$ to the estimated mean force,
+    and $c_0^{(s)} = c_2^{(s)} = 0$.
 
 \item
-  Equilibrate the system at umbrella $i = 1$.
+  Equilibrate the system at umbrella $s = 1$.
   Repeat the following steps until the simulation ends.
@@
   Perform a regular MC or MD step
-  in the current umbrella $i$ under bias potential $u^{(i)}(z, t)$.
+  in the current umbrella $s$ under bias potential $u^{(s)}(z, t)$.
 
@@
 \item
-  Update the $c_k^{(i)}$'s
+  Update the $c_k^{(s)}$'s
   according to Eq. \eqref{eq:ckupdate},
@@
   \frac{
-    \sum_i n^{(i)}(z)
+    \sum_{s=1}^S n^{(s)}(z)
   }
   {
-    \sum_i N^{(i)} \, e^{-u^{(j)}(z)}
+    \sum_{s=1}^S N^{(s)} \, e^{-u^{(s)}(z)}
   }
@@
 %
-where $n^{(i)}(z)$
+where $n^{(s)}(z)$
 the number of visits to the bin
 $\left(z-\frac{1}{2} \delta z, z + \frac{1}{2} \delta z\right)$
-in umbrella $i$,
-and the $N^{(i)}$ is the total number of visits
-to the umbrella $i$.
+in umbrella $s$,
+and the $N^{(s)}$ is the total number of visits
+to the umbrella $s$.
 
@@
 %
-Each step consists of
+Each step consisted of
 one MC trial for displacing one of the two tagged particles,
@@
 In each trial,
-a random particle is displaced in each spatial dimension
+a random particle was displaced in each spatial dimension
 according to a uniform distribution over $(-\varepsilon, \varepsilon)$.
 %
-The two tagged particles are fixed on the $x$-axis,
-and are only displaced in this direction.
+The two tagged particles were fixed on the $x$-axis
+and only displaced in this direction.
 %
@@
 both the optimal and potentially sub-optimal inverse-time schedules
-were able to reduce the error significantly.
+reduced the error significantly.
 %
@@
 (the remaining unevenness of the profile
-was likely due to the error of estimated values of
+might be due to errors of the estimated values of
 $\Gamma_k$'s and $\epsilon_k$'s).
@@
     %
-    The optimal schedule more effective
+    The optimal schedule more effectively
     reduced the errors of middle-range eigenmodes
@@
 The predicted minimal error was $2.0\times 10^{-6}$,
-  slightly less than the minimal error
-  in the previous Gaussian case.
+  slightly less than the corresponding value
+  in the Gaussian case.
 %
@@
 %
-For similar simulations performed at a high density, $\varrho = 0.8$,
+For similar simulations performed at a higher density, $\varrho = 0.8$,
 we found the gain of using the above two updating schemes
@@
 in MD simulations
-when a smooth bias potential is necessary in deriving the bias force.
-
-In practice, updating schemes and schedules
-are often sub-optimal.
-%
-We can, however, use the histogram-based correction formula,
-Eq. \eqref{eq:uhatav}, after simulation
+when a smooth bias potential is necessary
+for deriving the bias force.
+
+For simulations using sub-optimal
+updating schemes and schedules,
+we can use the histogram-based correction formula,
+Eq. \eqref{eq:uhatav},
 to improve the estimated PMF.
@@
 the noise is likely harmless
-in a post-simulation analysis, as
+for a post-simulation analysis, as
 the bias potential no longer needs to be differentiated
@@
 umbrellas were performed.\footnote{For
-  wider umbrellas, we might improve convergence
-  by mixing the nearest-neighbor inter-umbrella transitions
-  with some longer-range transitions.}
+  wider umbrellas,
+  one may consider
+  mixing the nearest-neighbor inter-umbrella transitions
+  with some longer-range transitions
+  to improve the convergence.}
 %In attempting inter-umbrella transitions,
@@
 Initially, we set
-$c_1^{(i)} = \tilde \beta_c \, \sigma_E$
+$c_1^{(s)} = \tilde \beta_c \, \sigma_E$
 for all umbrellas,
@@
 (which was more efficient than the Metropolis algorithm)
-at the inverse temperature, $\beta = c_1^{(i)}/\sigma_E$,
+at the inverse temperature, $\beta = c_1^{(s)}/\sigma_E$,
 followed by a Metropolis step with
 acceptance probability
-$\min\bigl\{1, e^{c_2^{(i)} [R_2(E) - R_2(E')]} \bigr\}$,
-where $E$ and $E'$ are the energies of the old and new states.
+$\min\bigl\{1, e^{c_2^{(s)} [R_2(E) - R_2(E')]} \bigr\}$,
+where $E$ and $E'$ are the energies of the old and new configurations.
 %
@@
 was roughly captured
-by $c_0^{(i)}$ as a function of $E_c^{(i)}$,
+by $c_0^{(s)}$ as a function of $E_c^{(s)}$,
 as shown Fig. \ref{fig:pt_hist}(b).
@@
 As discussed in Sec. \ref{sec:age},
-$c_1^{(i)}/\sigma_E$ roughly represents
+$c_1^{(s)}/\sigma_E$ roughly represents
 the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
@@
 %
-The coefficient, $\sqrt 2 \, c_2^{(i)}$,
+The coefficient, $\sqrt 2 \, c_2^{(s)}$,
 was indeed above $1.0$
@@
     %
-    (b) $c_0^{(i)}$ against $E_c^{(i)}$ (circles)
+    (b) $c_0^{(s)}$ against $E_c^{(s)}$ (circles)
     roughly agreed with the logarithmic density of states,
@@
     %
-    (c) $c_1^{(i)}/\sigma_E$ roughly corresponded to
+    (c) $c_1^{(s)}/\sigma_E$ roughly corresponded to
     the statistical temperature, $\beta(E)$.
     %
-    (d) $\sqrt 2 \, c_2^{(i)}$
+    (d) $\sqrt 2 \, c_2^{(s)}$
     exceeded $1.0$ in the unstable region,
@@
 As an application,
-we study a particular class
+we studied a particular class
 of the bandpass updating schemes
@@
 error\cite{zhou2005, morozov2007, zhou2008}
-caused by the updates in adaptive FDS methods.
-%
-This is justified only in the asymptotic regime,
+caused by the frequent updates to the bias potential.
+%
+While this is justified in the asymptotic regime,
 where the updating magnitude is small
-and the sampling is a quasi-equilibrium one.\cite{
+and the sampling is a quasi-equilibrium one,\cite{
   zhou2005, morozov2007, zhou2008, barducci2008, dama2014}
-Early in simulation, however,
-we still have to resort to the WL stage-by-stage
-technique even for the bandpss updating schemes.
+our analysis is inapplicable
+to early stages of a simulation
+and we still have to resort to the WL stage-by-stage
+technique even for the bandpass updating schemes.
 %the random error,
@@
 because it corresponds to the equally-weighted average,
-which is most efficient statistically.
+which is statistically most efficient.
 
@@
   $\rho_i e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
+%
+However, this is only an approximate result,
+and it does not remove the systematic bias
+caused by the adaptive updates.
 
@@
 \begin{align}
-  m\left( \bar q(t) \right)
+  m\bigl( \bar q(t) \bigr)
   &=
@@
 
-We can show that
+We can now show that
 for simulations with equal lengths and initial updating magnitudes,
@@
 On the other hand,
-if the mass distribution is inversely proportional to $\bar q$,
-which is approximately the case for the Gaussian updating scheme
-at large $\bar q$,\footnote{If
-  the eigenvalues, $\lambda_k$'s,
-  differ greatly in magnitude,
-  then for a large $\bar q$,
+the eigenvalues of the Gaussian updating scheme
+can differ greatly in magnitude
+[cf. Eq. \eqref{eq:lambda_Gaussian}].
+%
+Then, at a large $\bar q$,
 the sum in Eq. \eqref{eq:mass_func} will be dominated by
 the largest term, where $\lambda_k \approx 1/\bar q$,
-  and $M(\bar q) \approx \sqrt{\Gamma_k}/\bar q$.
-  }
-  %
-then we would have an exponentially decaying schedule.
+and $m(\bar q) \sim 1/\bar q$,
+and we would expect a roughly exponentially-decaying schedule.
 
@@
 We will use Eq. \eqref{eq:xt2_eql}
-but modify it to avoid the problem of not
-knowing the intrinsic distribution, $p_i$,
-hence the shift,
+but modify it to avoid the unknown
+intrinsic distribution, $p_i$,
+hence the unknown shift,
 $u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
 %
-Since the shift,
+Because the shift,
 hence its generalized Fourier transform,
@@
 ${\tilde u}_{k}$ and ${\tilde v}_{k}$
-should share the same variance,
-$\operatorname{var} {\tilde u}_{k} =
- \operatorname{var} {\tilde v}_{k} =
- \left\langle {\tilde v}_{k}^2 \right\rangle$,
-where we have used the fact that
-the ensemble average of ${\tilde v}_{k}$
-is zero in the last step.
+should share the same variance.
+%$\operatorname{var} {\tilde u}_{k} =
+% \operatorname{var} {\tilde v}_{k} =
+% \left\langle {\tilde v}_{k}^2 \right\rangle$,
 %\begin{equation*}
@@
 %  =
-%  \operatorname{var} {\tilde v}_{k}
-%  =
-%  \left\langle {\tilde v}_{k}^2 \right\rangle
 %  .
@@
 %
-From Eq. \eqref{eq:xt2_eql}, we get
+Then, from Eq. \eqref{eq:xt2_eql}, we get
 %
@@
   \operatorname{var} {\tilde u}_k
+  =
+  \operatorname{var} {\tilde v}_{k}
+  =
+  \left\langle {\tilde v}_{k}^2 \right\rangle
   =
@@
   a_0 \, \Gamma_k \, \lambda_k
-  .
+  ,
 \label{eq:varXt}
@@
 %
-The variance, $\tilde u_k$, can be computed
+where we have used the fact that the ensemble average of ${\tilde v}_{k}$
+is zero in the second step.
+%
+The variance, $\operatorname{var} \tilde u_k$, can be computed
 by trajectory averaging in a long simulation.
@@
 In practice, Eq. \eqref{eq:varXt}
-would fail if some eigenvalue, $\lambda_k$, is close to zero.
+would fail for a near zero eigenvalue.
 So it is best used for the single-bin updating scheme.%\footnote{Alternatively,
@@
 %  and then average the results.}
-
-Next we will use a model to show that $\Gamma_k$
-assumes a large value for a small $k$,
-and then decays rapidly to $1$ as $k$ increases
-in a local sampling process.
-%
-We assume that the propagation along the collective-variable is Markovian
-such that it can be approximately
-described by a transition matrix, $\mathbf A$, with
-\begin{align*}
-\langle h_i(t) \rangle = \rho_i,
-\quad
-\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, \rho_j
-\quad \mbox{(for $t \ge 0$)}
-.
-\end{align*}
-%
-We further assume that
-the transition matrix
-has a complete set of orthonormal eigenvectors,
-and the eigenvectors coincide with
-those of the updating matrix, $\phi_{ki}$'s.
-Then the transition matrix can be decomposed as
-$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \phi_{li} \, \phi_{lj} / \rho_j$,
-with $\Lambda_l$ being the eigenvalue of mode $l$,
-and
-%
-$\bigl\langle
-  \tilde f_k(t) \, \tilde f_k(0)
-\bigr\rangle
-=
-\Lambda_k^t$.
-%If the transition matrix % satisfies detailed balance
-%has a complete set of orthonormal eigenvectors, $\psi_{li}$
-%%that are orthonormal with respect to $\pmb\rho$
-%[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
-%\eqref{eq:eig_orthonormal_rows}],
-%it can be decomposed as
-%$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
-%with $\Lambda_l$ being the eigenvalue of mode $l$.
-%%
-%The autocorrelation function can then be computed as
-%%
-%\begin{align*}
-%\bigl\langle
-%  \tilde f_k(t) \, \tilde f_k(0)
-%\bigr\rangle
-%=
-%\sum_{l = 0}^{n-1} \Lambda_l^t \,
-%\left(
-%  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
-%\right)^2
-%.
-%\end{align*}
-%%
-%In particular, if the eigenvectors, $\pmb \phi_k$
-%and $\pmb \psi_k$, coincide,
-%we would have
-%$\bigl\langle
-%  \tilde f_k(t) \, \tilde f_k(0)
-%\bigr\rangle
-%=
-%\Lambda_k^t$.
-%and
-%\begin{equation}
-%  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
-%  \bigl\langle
-%    \tilde f_k(t) \, \tilde f_k(0)
-%  \bigr\rangle
-%  = \frac{1+\Lambda_k}{1-\Lambda_k}
-%  .
-%  \notag
-%  %\label{eq:Gamma_est}
-%\end{equation}
-%
-To model a local sampling process,
-we can borrow Eq. \eqref{eq:lambda_Gaussian} for $\Lambda_k$,
-with $\sigma_z$ reinterpreted as
-the average move size along $z$ during sampling,
-then
-\begin{align}
-  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
-  \bigl\langle
-    \tilde f_k(t) \, \tilde f_k(0)
-  \bigr\rangle
-  %= \frac{1+\Lambda_k}{1-\Lambda_k}
-%\Gamma_k
-  = \coth\left(
-  \frac{ \pi^2 \, k^2 \, {\sigma_z^\mathrm{(samp)}}^2 } { 4 \, \Delta z^2}
-  \right).
-  \notag
-  %\label{eq:Gamma_local}
-\end{align}
-Thus, we have $\Gamma_k \propto \Delta z^2/k^2$
-for small $k$,\cite{bussi2006}
-whereas $\Gamma_k \to 1$ as $k$ increases.
-%
-However, this model tends to underestimate
-the first few $\Gamma_k$'s in a typical simulation.
-%which means that doubling the sampling range, $\Delta z$,
-%would make the histogram flattening
-%at least four times as difficult,
-%i.e. FDS methods are most effective
-%for short ranges\cite{wang2001, wang2001pre}.
-%
-%Using finer bins would, however, not change this value.
-%
 
@@
 for some numbers, $\mu_l$'s, which
-define an unadjusted window function.
+define the unadjusted window function.
 %
@@
 $\pmb\phi^{(1)}, \dots, \pmb\phi^{(n)}$,
-are given by
+given by
 %
@@
 with $g = 1$.
-Note the two-fold degeneracy,
-%for the periodic variable,
-%or $2$ for a non-periodic one
+Note also the two-fold degeneracy,
 %There is a two-fold degeneracy,
@@
 Then the variable and the images in the two mirrors
-define an effective periodic variable of period $2 \, n$.
+can be mapped to an equivalent periodic variable of period $2 \, n$.
 %
@@
 
-The Gaussian updating scheme introduced in Sec. \ref{sec:Gaussian_scheme}
-is defined by a Gaussian-shaped window function, $\mu_l$.
+The Gaussian updating scheme (cf. in Sec. \ref{sec:Gaussian_scheme})
+is associated with a Gaussian-shaped window function, $\mu_l$.
 %
@@
 and
-$\Delta z = (n \, \delta z) \, g / 2$.
+$\Delta z \equiv (n \, \delta z) \, g / 2$.
 %
@@
 in sampling of the bivariate distribution,
-$\rho(i, z)$, in Sec. \ref{sec:st},
+$\rho(s, z)$, in Sec. \ref{sec:st},
 the above definition becomes
@@
   &=
-  \sum_{a = 0}^{n - 1}
+  \sum_{\nu = 0}^{S - 1}
   \sideset{}{'}\sum_{k = 0}^{K - 1}
-  \tilde F_{a, k}^2
+  \tilde F_{\nu, k}^2
   \notag \\
   &=
-  \sum_{i=1}^n
+  \sum_{s=1}^S
   \left[
-  \frac{ (H_i - \rho_i)^2 } { \rho_i }
+  \frac{ (H_s - \rho_s)^2 } { \rho_s }
   +
   \sum_{k = 1}^{K - 1}
-  \frac{ H_i^2 } { \rho_i }
-  \left|{\tilde F}_k^{(i)} \right|^2
+  \frac{ H_s^2 } { \rho_s }
+  \left|{\tilde F}_k^{(s)} \right|^2
   \right]
@@
 \end{align}
-where $a$ and $k$ are the indices of
+where $\nu$ and $k$ are the indices of
 inter- and intra-umbrella fluctuation modes,
 and the primed sum excludes
-the term of $a = k = 0$;
-on the next line, $\tilde F_k^{(i)}$
-is the histogram fluctuation of the $k$th mode in umbrella $i$,
-which is the average of $R_k^{(i)}\bigl( z(t) \bigr)$
-for samples in umbrella $i$.
+the term of $\nu = k = 0$;
+on the next line,
+$H_s$ is the normalized inter-umbrella histogram,
+and
+$\tilde F_k^{(s)}$
+is the histogram fluctuation of the $k$th mode in umbrella $s$,
+which is the average of $R_k^{(s)}\bigl( z(t) \bigr)$
+for samples in umbrella $s$.
 %If the eigenmodes are made of a sequence of orthogonal polynomials

### 2017-09-07 web/js/mmwl.js

@@ -1,75 +1,0 @@
-/* Wang-Landau scheme for moments */
-
-
-
-"use strict";
-
-
-
-function MMWL(a)
-{
-  this.mm = [0, 0, 0];
-  this.fl = [0, 0, 0];
-  this.alphawl = a;
-  this.invt = false;
-  this.t0 = 1;
-}
-
-
-
-MMWL.prototype.add = function(y1, y2)
-{
-  this.mm[0] += 1;
-  this.mm[1] += y1;
-  this.mm[2] += y2;
-}
-
-
-
-MMWL.prototype.getalpha = function()
-{
-  return this.invt ? 1./(this.mm[0] + this.t0) : this.alphawl;
-}
-
-
-
-/* compute the fluctuation of the statistical moments
- * return the maximum fluctuation */
-MMWL.prototype.calcfl = function(dovar)
-{
-  var t = this.mm[0], fl, fl2;
-  if ( t <= 0 ) return 99.0;
-  this.fl[1] = this.mm[1] / t;
-  this.fl[2] = this.mm[2] / t;
-  fl = Math.abs(this.fl[1]);
-  if ( dovar && (fl2 = Math.abs(this.fl[2])) > fl )
-    fl = fl2;
-  return fl;
-}
-
-
-
-/* switch to a stage with a reduced updating magnitude */
-MMWL.prototype.switch = function(magred)
-{
-  var t = this.mm[0];
-  this.alphawl *= magred;
-  if ( this.alphawl < 1./t ) {
-    this.t0 = t;
-    this.invt = true;
-  }
-  this.mm = [0, 0, 0];
-}
-
-
-MMWL.prototype.check = function(dovar, fl, magred)
-{
-  var flmax = this.calcfl(dovar);
-  if ( !this.invt && flmax < fl ) {
-    this.switch(magred);
-    return 1;
-  } else {
-    return 0;
-  }
-}
-

### 2017-09-07 web/js/pottsmain.js

@@ -33,7 +33,8 @@ var nstepspfmc = 100;  // number of steps per frame for MC
 
 var npulses = 0;
 var histplot = null;
-var vplot = null;
+var c1plot = null;
+var c2plot = null;
 
 function getparams()
 {
@@ -176,7 +177,9 @@ function paint()
   var l = potts.l, q = potts.q, id = 0;
   var s, colors = new Array(potts.q);
   for ( s = 0; s < potts.q; s++ ) {
-    colors[s] = getHueColor(1.0*s/potts.q, 0, 230);
+    // adjust the darkest according to the umbrella ID
+    var bright = 80 + Math.floor(170 * iage / (age.n - 1.0));
+    colors[s] = getHueColor(1.0*s/potts.q, 0, bright);
   }
 
   for ( var i = 0; i < l; i++ ) {
@@ -187,6 +190,38 @@ function paint()
       ctx.fillStyle = colors[s];
       ctx.fillRect(x, y, dx1, dy1)
     }
+  }
+}
+
+
+
+/* draw a color bar for the umbrella ID */
+function drawcolorbar(target)
+{
+  var c = grab(target);
+  var ctx = c.getContext("2d");
+  var width = c.width;
+  var height = c.height;
+
+  ctx.fillStyle = "#ffffff";
+  ctx.fillRect(0, 0, width, height);
+
+  var grd = ctx.createLinearGradient(0, 0, width, 0);
+  grd.addColorStop(0, "#505050");
+  grd.addColorStop(1, "#fafafa");
+
+  ctx.fillStyle = grd;
+  ctx.fillRect(0, 4, width, height-4);
+
+  // draw grids
+  if ( age ) {
+    var i, x;
+    for ( i = 0; i < age.n; i++ ) {
+      x = width * i / (age.n - 1.0);
+      drawLine(ctx, x, 4, x, height, "#808080", 1);
+    }
+    x = width * iage / (age.n - 1.0);
+    drawLine(ctx, x, 0, x, height, "#000000", 2);
   }
 }
 
@@ -229,7 +264,7 @@ function updatehistplot()
     dat += "\n";
   }
 
-  if ( histplot === null || 1 ) {
+  if ( histplot === null ) {
     var h = grab("animationbox").height / 2 - 5;
     var w = h * 3 / 2;
     var options = {
@@ -252,7 +287,7 @@ function updatehistplot()
 
 
 
-function updatevplot()
+function updatec1plot()
 {
   var i, j, ntp = age.n;
   // prepare the header
[…]

### 2017-09-09 doc/invt.tex
small edits to invt.tex

@@
 %and to capture the global shape of the free energy surface
-an effective strategy is to introduce a bias potential that
-offsets the PMF
+an effective strategy is to introduce
+a bias potential\cite{torrie1974, *torrie1977}
+that offsets the PMF
 such that the distribution along $z$ is
@@
 to discourage future visits to previously sampled configurations
-by incremental elevation of the bias potential.
+by incremental local elevation of the bias potential.
 %
@@
 and the convergence of the bias potential
-can benefit from a gradually decreasing
+can often benefit from a gradually decreasing
 updating magnitude of the bias potential.
@@
 belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
-liang2007, min2007,
+liang2007,
 morozov2007, morozov2009, zhou2008,
 komura2012, *caparica2012, *caparica2014,
+min2007,
 barducci2008, dickson2011, dama2014, poulain2006}
@@
 We will then show by comparing the errors
-of different updating schemes
+of simulations under different updating schemes
 that the single-bin updating scheme
@@
   \int_0^T
-    \dot{\vartheta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
+    \dot{\theta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
   \label{eq:xt_solution}
@@
 \begin{align}
-  \vartheta(q')
+  \theta(q')
   &\equiv
@@
     \rangle
-    \dot \vartheta\bigl( q(t) \bigr)
-    \dot \vartheta\bigl( q(t') \bigr)
+    \dot \theta\bigl( q(t) \bigr)
+    \dot \theta\bigl( q(t') \bigr)
     \, dt dt'
   \notag \\
-  &= \Gamma \, \int_0^T \dot \vartheta^2\bigl( q(t) \bigr) \, dt
+  &= \Gamma \, \int_0^T \dot \theta^2\bigl( q(t) \bigr) \, dt
   ,
@@
 has the form of an action of a free particle
-at $\vartheta\bigl( q(t) \bigr)$,
+at $\theta\bigl( q(t) \bigr)$,
 with the endpoints fixed at
-$\vartheta\bigl( q(0) \bigr)  = e^{- q(T)}$
+$\theta\bigl( q(0) \bigr)  = e^{- q(T)}$
 and
-$\vartheta\bigl( q(T) \bigr) = 1$.
+$\theta\bigl( q(T) \bigr) = 1$.
 %
@@
 \begin{equation}
-  \frac{d}{dt} \vartheta\bigl( q(t) \bigr) = c
+  \frac{d}{dt} \theta\bigl( q(t) \bigr) = c
   .
@@
   \int_0^T
-    \dot{\vartheta}_k\bigl( q(t) \bigr) \, \tilde\xi_k(t) \, dt
-  ,
-  \notag
-  %\label{eq:xt_solution}
+    \dot{\theta}_k\bigl( q(t) \bigr) \, \tilde\xi_k(t) \, dt
+  ,
+  \label{eq:xt_solution}
 \end{equation}
@@
 \begin{equation}
-  \vartheta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
+  \theta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
   .
@@
   \rangle
-  \vartheta_k\bigl(q(t)\bigr)
-  \vartheta_k\bigl(q(t')\bigr)
+  \theta_k\bigl(q(t)\bigr)
+  \theta_k\bigl(q(t')\bigr)
   \, dtdt'
@@
   \sum_{k = 1}^{n-1}
-  \Gamma_k \, \dot \vartheta_k^2\bigl( q(t) \bigr) \, dt
+  \Gamma_k \, \dot \theta_k^2\bigl( q(t) \bigr) \, dt
   ,
@@
   \sum_{ k = 1 }^{n - 1}
-    \Gamma_k \, {\dot \vartheta}_k^2\bigl[ q(t) \bigr]
+    \Gamma_k \, {\dot \theta}_k^2\bigl[ q(t) \bigr]
   %\notag
@@
   %\sum_{ k = 1 }^{n - 1}
-  %  \Gamma_k \, \lambda_k^2 \, \vartheta_k^2 \bigl[ q(t) \bigr]
+  %  \Gamma_k \, \lambda_k^2 \, \theta_k^2 \bigl[ q(t) \bigr]
   %\; \dot q^2( t )
@@
 does not explicitly depend on time, $t$,
-the Hamiltonian is conserved,\cite{goldstein, *landau_mechanics, *arnold}
+the Hamiltonian is conserved,\cite{goldstein, *landau_mechanics, *arnold, *jose}
 which means that the asymptotic error grows
@@
 before the preliminary run is zero,
-one can show that
-\begin{equation}
-  \left\langle
-    {\tilde v}_k^2(0)
-  \right\rangle
-  = \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
-  + \epsilon_k
-  ,
-  \label{eq:xt2_eql1}
-\end{equation}
+we have, from Eq. \eqref{eq:xt_solution} [with $T\to T_0, q(t) \to a_0 \, t$],
+%
+\begin{align}
+  \left\langle {\tilde v}_k(0) \right\rangle
+  &=
+  \tilde u_k \, e^{-\lambda_k \, a_0 \, T_0}
+  ,
+  \notag \\
+  \operatorname{var} \tilde v_k(0)
+  &=
+  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  \left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
+  ,
+  \label{eq:xtvar}
+  \\
+  \left\langle \tilde v_k^2(0) \right\rangle
+  &=
+  %\operatorname{var} \tilde v_k(0) + \zeta_k^2
+  %=
+  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  +
+  \epsilon_k \, e^{-2 \, \lambda_k \, a_0 \, T_0}
+  .
+  \label{eq:xt2_eql}
+\end{align}
 %
 where
-$\epsilon_k \equiv \tilde u_k^2 \, e^{-2\, \lambda_k \, a_0 \, T_0}$.
-%
-If $\lambda_k$ is not close to zero, and $T_0 \to \infty$
-we can ignore the $\epsilon_k$ term, and
-%
-\begin{equation}
-  \left\langle
-    {\tilde v}_k^2(0)
-  \right\rangle
-  \approx
-  \frac 1 2 \, a_0 \, \Gamma_k \, \lambda_k
-  .
-  \label{eq:xt2_eql}
-\end{equation}
-%
-\note{
-Consider a long period, $T_0$, under a fixed schedule,
-$\alpha(t) = a_0$,
-the residual error becomes negligible, and
-the $k$th component of the asymptotic error
-is given by
-%
-\begin{align*}
-  \left\langle
-    {\tilde v}_k^2(T_0)
-  \right\rangle
-  =
-  \int_0^{T_0}
-    \Gamma_k \, (\lambda_k \, a_0)^2 \,
-      e^{ 2 \, \lambda_k \, a_0 \, (t - T_0) }
-    \, dt
-  \stackrel{ T_0 \to \infty }
-  { =\joinrel=\joinrel=\joinrel= }
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
-  .
-\notag
-%\label{eq:error_eql}
-\end{align*}
-}
-%
-
-%
-From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
+$\epsilon_k \equiv \left(\tilde u_k^2 - \frac{1}{2} \Gamma_k \, \lambda_k \, a_0\right) \, e^{-2\,\lambda_k a_0 T_0}$.
+
+%
+From Eqs. \eqref{eq:error_res},
+\eqref{eq:xt2_eql},
 and \eqref{eq:error_asym2},
@@
   =0
-  .
+  ,
 \label{eq:opt_qT}
 \end{align}
+where
 %
@@
 %
-\note{Let $q_T \equiv q(T)$,
-$$
-\begin{aligned}
+\note{We can write
+\begin{align*}
+  \left\langle
+    {\tilde v}_k^2(0)
+  \right\rangle
+  &=
+  \frac { a_0 \, \Gamma_k \, \lambda_k } { 2 }
+  +
+  \Delta_k
+  e^{-2\, \lambda_k a_0 T_0}
+  ,
+\end{align*}
+%
+where $\Delta_k \equiv \tilde u_k^2 - \frac{a_0 \lambda_k \Gamma_k}{2}$.
+%
+Let $q_T \equiv q(T)$,
+\begin{align*}
   \frac{
@@
   &=
+  -2\sum_{k=1}^{n-1} \lambda_k
+  \left\langle
+    \tilde v_k^2(0)
+  \right\rangle \,
+  e^{-2 \, \lambda_k \, q_T}
+  \\
+  &=
   -a_0 \, M^2(q_T)
@@
   \sum_{k=1}^{n-1} \lambda_k \,
-  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
+  \Delta_k e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
   ,
@@
   .
-\end{aligned}
-$$
-Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
+\end{align*}
+Then solve $\partial (\Err_R + \Err_A) / \partial q_T = 0$,
+\begin{align*}
+  \frac{ M\bigl( q_T \bigr) } { T } C_M
+  -\frac{a_0}{2} \, M^2\bigl( q_T \bigr)
+  -\sum_{k = 1}^{n-1}
+  \lambda_k \, \Delta_k \, e^{-2\lambda_k \, (q_T + a_0 T_0)}
+  =0
+  .
+\end{align*}
 }
@@
 \end{align}
+%
 %
@@
 \eqref{eq:error_res},
-\eqref{eq:xt2_eql1},
+\eqref{eq:xt2_eql},
 and
@@
 We wish to find a lower bound of the asymptotic error,
-$\Err_A(T)$, as defined in Eq. \eqref{eq:error_asym}
+$\Err_A(T)$ [given by Eq. \eqref{eq:error_asym}],
 for a class of updating matrices that
@@
   inequality follows from the nonnegativity of
-  the quadratic polynomial,
+  the quadratic polynomial of $y$,
   \begin{align*}
-  \int_0^T
-    dt \sum_{k = 1}^{n-1} \Gamma_k \,
-      ( {\dot \vartheta}_k \, \chi - 1 )^2
+  \int_0^T \!\!dt \sum_{k = 1}^{n-1} \Gamma_k \,
+      ( y - {\dot \theta}_k )^2
     \equiv
-    A_2 \, \chi^2 + A_1 \, \chi + A_0
+    A_2 \, y^2 + A_1 \, y + A_0
     ,
   \end{align*}
-  which implies a nonpositive discriminant,
-  $A_1^2 - 4 \, A_0 \, A_2$.}
+  which implies that the discriminant is nonpositive,
+  $A_1^2 - 4 \, A_0 \, A_2 \le 0$.}
 %
@@
       \sum_{k = 1}^{n-1}
-        \Gamma_k \, {\dot \vartheta}_k^2\bigl( q(t) \bigr)
+        \Gamma_k \, {\dot \theta}_k^2\bigl( q(t) \bigr)
   \right)
@@
   \notag \\
-  & \qquad \qquad \qquad
+  & %\qquad \qquad \qquad
   \ge
-  \left(
+  \Biggl[
     \int_0^T \!\!dt
       \sum_{k = 1}^{n-1}
-        \Gamma_k \, {\dot \vartheta}_k \, \bigl( q(t) \bigr)
-  \right)^{\!\!2}
-  \notag \\
-  & \qquad \qquad \qquad
-  =
-  \left(
+        \Gamma_k \, {\dot \theta}_k \, \bigl( q(t) \bigr)
+  \Biggr]^{\!2}
+  %\notag \\
+  %& \qquad \qquad \qquad
+  \!\!=
+  \Biggl\{
     \sum_{k = 1}^{n-1} \Gamma_k
-      \left[
+      \Bigl[
         1 - e^{ -\lambda_k \, q(T) }
-      \right]
-  \right)^{\!\!2}
+      \Bigr]
+  \Biggr\}^{\!2}
   .
@@
 and the equality is achieved
-if $\dot \vartheta_k\left( q(t) \right) = \mathrm{constant}$
+if $\dot \theta_k\bigl( q(t) \bigr) = \mathrm{constant}$
 for all $k > 0$ at any $t$.
@@
 %We can further verify that
-%$\vartheta_k = (t+t_0)/(T+t_0)$
+%$\theta_k = (t+t_0)/(T+t_0)$
 %in this case,
@@
 and the inverse-time schedule
-was recommended for optimal convergence of $c_1$.\cite{pellegrini2014}
+was recommended for the optimal convergence of $c_1$.\cite{pellegrini2014}
 %
@@
 to an array of adjacent intervals
-to cover a wide range over the collective variable, $z$.
+to cover a wide range of the collective variable, $z$.
 %
@@
 quadratic\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007},
-which is common in umbrella sampling\cite{torrie1974, *torrie1977}
+which is common in umbrella sampling
 and the Gaussian ensemble method\cite{hetherington1987,
@@
 from a superposition of $S$ Gaussian umbrellas
-centered at different places.\footnote{This sampling
-  algorithm should not be confused with the Gaussian updating scheme,
-  in which the bias potential is incrementally updated by Gaussian functions,
-  and the target distribution is flat.
-  However, in this case, the bias potential is quadratic,
-  and the target distribution of each umbrella is Gaussian.}
-%
-The target distribution is a joint distribution
-of the discrete umbrella index, $s = 1, \dots, S$,
+centered at different places.
+%
+The expanded ensemble treats
+the umbrella index, $s = 1, \dots, S$,
+also as a random variable,
+and the biased sampling targets a joint distribution
+of the umbrella index
 and the continuous variable, $z$,
-$\rho(s, z) = \rho_s \, \rho^{(s)}(z)$,
-where the marginal inter-umbrella distribution is flat,
+$\rho(s, z) = \rho_s \, \rho^{(s)}(z)$.
+%
+Here,
+the marginal inter-umbrella distribution is flat,
 $\rho_s = 1/S$,
 and the conditional intra-umbrella distribution, $\rho^{(s)}(z)$,
-is the Gaussian given in Eq. \eqref{eq:rho_Gaussian}. % previously.
+is the Gaussian given in Eq. \eqref{eq:rho_Gaussian}.
+\footnote{This method
+  should not be confused with the Gaussian updating scheme.
+  %
+  In the former case, the bias potential is quadratic
+  and the target distribution is Gaussian in each umbrella,
+  although the inter-umbrella distribution, $\rho_s$, is flat.
+  %
+  In the latter case, the bias potential is a sum of
+  Gaussian functions, and the target distribution
+  is flat.}
 %
@@
 %
-With a small width, $\sigma_z$, the converged
+With a small width, $\sigma_z^{(s)}$, the converged
 $c_0^{(s)}$ as a function of $z_c^{(s)}$
@@
 the negative PMF, $\ln p(z)$.\cite{maragliano2006, *abrams2008}
+%
+Thus, we have\cite{park2007, *nguyen2013, *zhang2015st}
+\begin{equation}
+  c_0^{(s+1)} - c_0^{(s)}
+  \approx
+  \frac{1}{2}\left(
+    \frac{ c_1^{(s+1)} } { \sigma_z^{(s+1)} }
+    +
+    \frac{ c_1^{(s)} } { \sigma_z^{(s)} }
+  \right)
+  \left(
+  z_c^{(s+1)} - z_c^{(s)}
+  \right)
+  .
+  \label{eq:c0_est}
+\end{equation}
 
@@
     and
-    the width, $\sigma_z^{(s)}$, to be a constant.
+    the width, $\sigma_z^{(s)}$, to be a constant, $\sigma_z$.
     For the polynomial coefficients, we can set
-    $c_1^{(s)}/\sigma_z^{(s)}$ to the estimated mean force,
-    and $c_0^{(s)} = c_2^{(s)} = 0$.
+    $c_1^{(s)}/\sigma_z^{(s)}$ to some estimated mean force, $f_z$,
+    $c_0^{(s)} = f_z \bigl[ z_c^{(s)} - \frac{1}{2}(z_{\min} + z_{\max}) \bigr]$
+    by Eq. \eqref{eq:c0_est},
+    and $c_2^{(s)} = 0$.
 
@@
   with the acceptance probability given
-    by Eq. \eqref{eq:transprob_st}.\cite{*[{Although
-    the intra-umbrella sampling step
-    and inter-umbrella transition step
-    are executed in sequential instead of random order,
-    the sampling scheme is still valid,
-    as shown in }] [{}] manousiouthakis1999}
+    by Eq. \eqref{eq:transprob_st}.\cite{*[{The
+    intra-umbrella sampling step and
+    the inter-umbrella transition step
+    can be executed at different frequencies
+    without affecting the correctness of the algorithm.
+    Although some prefers mixing the two steps
+    in a random order,
+    the sequential execution adopted here
+    is also valid,
+    as discussed in }] [{}]
+    manousiouthakis1999}
 
@@
   wider umbrellas,
-  one may consider
-  mixing the nearest-neighbor inter-umbrella transitions
+  one may mix the nearest-neighbor inter-umbrella transitions
   with some longer-range transitions
@@
 $c_1^{(s)} = \tilde \beta_c \, \sigma_E$
-for all umbrellas,
+for all umbrellas
 with $\tilde \beta_c = 1.4$
@@
 %
-However, this is only an approximate result,
-and it does not remove the systematic bias
-caused by the adaptive updates.
+However, it does not remove the systematic bias
+caused by the adaptive updates,
+and thus still requires a sufficiently small updating magnitude.
 
@@
 
-We will use Eq. \eqref{eq:xt2_eql}
-but modify it to avoid the unknown
-intrinsic distribution, $p_i$,
-hence the unknown shift,
-$u_i(t) - v_i(t) = \ln(p_i/\rho_i)$.
-%
-Because the shift,
-hence its generalized Fourier transform,
-$\tilde u_k(t) - \tilde v_k(t)$,
-is a constant of time,
+Because the difference between $u_i$ and $v_i$,
+is a constant of time [cf. Eq. \eqref{eq:v_def}],
 ${\tilde u}_{k}$ and ${\tilde v}_{k}$
-should share the same variance.
+share the same variance.
+%
 %$\operatorname{var} {\tilde u}_{k} =
@@
 %
-Then, from Eq. \eqref{eq:xt2_eql}, we get
+Then, from Eq. \eqref{eq:xtvar}, we get
 %
@@
   =
-  \left\langle {\tilde v}_{k}^2 \right\rangle
-  =
-  \frac{1}{2} \,
-  a_0 \, \Gamma_k \, \lambda_k
-  ,
+  \frac{1}{2} \, \Gamma_k \, \lambda_k \, a_0
+  \left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
+  .
 \label{eq:varXt}
@@
 %
-where we have used the fact that the ensemble average of ${\tilde v}_{k}$
-is zero in the second step.
+%where we have used the fact that the ensemble average of ${\tilde v}_{k}$
+%is zero in the second step.
 %
@@
 %
-In practice, Eq. \eqref{eq:varXt}
-would fail for a near zero eigenvalue.
+In practice, Eq. \eqref{eq:varXt} would fail for a near zero eigenvalue.
+%
 So it is best used for the single-bin updating scheme.%\footnote{Alternatively,
@@
     $T$             &   Simulation length. \\
-    $\vartheta_k(q')$       &   $\vartheta_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
+    $\theta_k(q')$       &   $\theta_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
     $u_i(t)$        &   Original bias potential. \\

### 2017-09-09 web/js/age.js

@@ -4,10 +4,6 @@
 
 "use strict";
 
-/* values for lnzmethod */
-//enum { LNZ_WL, LNZ_AVE };
-//const char *lnz_names[2] = {"WL", "Ave"};
-
 var SQRT2 = 1.4142135623730951;
 
 function MMWL()
@@ -37,20 +33,19 @@ MMWL.prototype.calcfl = function()
 }
 
 
-function AGE(xcmin, xcmax, delx, sig, lnzmethod,
-  c1, alpha0, xmin, xmax, dx, pbc)
+function AGE(xcmin, xcmax, delx, sig,
+  c1, alpha0, xmin, xmax, dx)
 {
   var i, xn, n;
 
   n = Math.floor((xcmax - xcmin) / delx) + 1;
   this.n = n;
-  this.lnzmethod = lnzmethod;
   this.alpha0 = alpha0;
   this.ave = new Array(n);
   this.sig = new Array(n);
+  this.c0 = new Array(n);
   this.c1 = new Array(n);
   this.c2 = new Array(n);
-  this.lnz = new Array(n);
   this.mmwl = new Array(n);
   this.cnt = new Array(n);
   this.acc = new Array(n);
@@ -58,9 +53,9 @@ function AGE(xcmin, xcmax, delx, sig, lnzmethod,
   for ( i = 0; i < n; i++ ) {
     this.ave[i] = xcmin + i * delx;
     this.sig[i] = sig;
+    this.c0[i] = (i - (n - 1)*0.5) * c1 / sig * delx;
     this.c1[i] = c1;
     this.c2[i] = 0;
-    this.lnz[i] = (i - (n - 1)*0.5) * c1 / sig * delx;
     this.mmwl[i] = new MMWL();
     this.cnt[i] = 0;
     this.acc[i] = 0;
@@ -115,7 +110,7 @@ AGE.prototype.add = function(i, x, acc)
   this.acc[i] += acc;
   var j = Math.floor( (x - this.xmin) / this.dx );
   this.hist[i][j] += 1;
-  this.lnz[i] += alpha;
+  this.c0[i] += alpha;
 }
 
 
@@ -207,14 +202,14 @@ AGE.prototype.move = function(x, id, local)
   }
   sigi = this.sig[id];
   sigj = this.sig[jd];
-  dv = this.lnz[jd] - this.lnz[id];
+  dv = this.c0[jd] - this.c0[id];
[…]

### 2017-09-12 doc/invt.tex
metad.h correct the initial error

@@
   \!\!=
-  \Biggl\{
+  \Biggl[
     \sum_{k = 1}^{n-1} \Gamma_k
-      \Bigl[
+      \Bigl(
         1 - e^{ -\lambda_k \, q(T) }
-      \Bigr]
-  \Biggr\}^{\!2}
+      \Bigr)
+  \Biggr]^{\!2}
+  \!\!
   .
@@
   w_{ij}
-  = \frac{1}{\rho_i} \sum_{k=0}^{n-1} \lambda_1 \phi_{ki} \phi_{kj}
+  = \frac{1}{\rho_i} \sum_{k=0}^{n-1} \lambda_1 \, \phi_{ki} \, \phi_{kj}
   = \lambda_1 \, \delta_{ij}
@@
   R_1(z) = \frac{z - \overline{z}}{\sigma_z},
-  R_2(z) = \frac{ (z - \overline{z})^2 - \sigma_z^2 }
-  {\sqrt{\overline{\left[
-    (z - \overline z)^2 - \sigma_z^2
-  \right]^2 }}}
-  ,
+  R_2(z) = \frac{ (z - \overline{z})^2 - \sigma_z^2 } {\kappa_z},
 \end{align*}
 %
-where $\sigma_z \equiv \sqrt{\overline{\left(z - \overline z\right)^2}}$.
+where $\sigma_z \equiv \sqrt{\overline{\left(z - \overline z\right)^2}}$,
+and
+$\kappa_z \equiv \sqrt{\overline{\left[  (z - \overline z)^2 - \sigma_z^2  \right]^2 }}$.
 %
@@
 
-As shown in Table \ref{tab:lj_error},
-both the optimal and potentially sub-optimal inverse-time schedules
-reduced the error significantly.
+Both the optimal and potentially sub-optimal inverse-time schedules
+reduced the error significantly,
+as shown in Table \ref{tab:lj_error}.
 %
@@
     & $6.7(7.8)\times10^{-3}$
-    & $1.8(2.6)\times10^{-5}$
+    & $1.8(2.2)\times10^{-5}$
     \\
@@
     & $6.7(7.8)\times10^{-3}$
-    & $2.9(3.7)\times10^{-6}$
+    & $2.9(3.6)\times10^{-6}$
     \\
@@
 %
-Since our equilibration run was sufficiently long,
+Since our equilibration phase was sufficiently long,
 the initial updating magnitude, $\alpha(0)$,
@@
 we found that the inverse-time schedule
-reduced the updating magnitude too quickly,
-leaving a hump on the error profile
-for the $k = 9$ to $31$ modes.
+failed to reduce to the errors in the $k = 9$ to $31$ modes adequately.
 %
@@
 and produced a flatter profile
-(the remaining unevenness of the profile
+(The remaining unevenness of the profile
 might be due to errors of the estimated values of
-$\Gamma_k$'s and $\epsilon_k$'s).
+$\Gamma_k$'s and $\epsilon_k$'s.)
 %

### 2017-09-12 prog/intq.h

@@ -76,7 +76,8 @@ static void intq_setqgrid(intq_t *intq, double qT)
 
 
 
-/* compute the integrand for `intq` */
+/* compute the integrand for `intq`
+ * return the square-root mass function */
 static double intq_getmassx(intq_t *intq, double dq,
     double *massK)
 {
@@ -89,7 +90,7 @@ static double intq_getmassx(intq_t *intq, double dq,
     y = intq->gamma[k] * lam * lam * exp( 2 * lam * dq );
     mass += y;
     if ( kc <= 0 || k < kc || (pbc && k > n - kc) ) {
-      *massK += y; /* for mode limited optimal schedule */
+      *massK += y; /* for mode-limited optimal schedule */
     }
   }
 
@@ -242,7 +243,8 @@ static double intq_reserr(intq_t *intq, double a0, const double *xerr, double qT
 
 /* compute the error components */
 __inline static double intq_errcomp(intq_t *intq, double a0,
-    double qT, double *xerr, double *xerr_r, double *xerr_a)
+    double qT, const double *xerr0,
+    double *xerr, double *xerr_r, double *xerr_a)
 {
   int k, j, m = intq->m, n = intq->n;
   double err_r, err_a, err;
@@ -255,7 +257,9 @@ __inline static double intq_errcomp(intq_t *intq, double a0,
     gam = intq->gamma[k];
 
     /* residual error */
-    err_r = 0.5 * a0 * gam * lam * exp(-2.0 * lam * qT);
+    err_r = 0.5 * a0 * gam * lam;
+    if ( xerr0 != NULL ) err_r += xerr0[k];
+    err_r *= exp(-2.0 * lam * qT);
 
     /* asymptotic error */
     err_a = 0;
@@ -834,10 +838,11 @@ static double esterror_opt(double T, double a0,
 
 /* return the square-root error from the optimal schedule */
 static double esterror_optx(double T, double a0,
-    double *xerr, double *qT, double qprec,
+    double *xerr0, double *qT, double qprec,
     int m, intq_t **intq_ptr,
     int n, int kc, int pbc,
     const double *lambda, const double *gamma,
+    double *xerr,
     int verbose)
 {
   intq_t *intq;
@@ -847,10 +852,15 @@ static double esterror_optx(double T, double a0,
 
   /* compute the optimal schedule and error */
[…]

### 2017-09-12 doc/invt.tex
fit the c0 shift in reweight.c and pt_hist.gp

@@
 %
-Below we will use the notations, ``$\bav{}$'' and ``$_{*i}$,''
+Below we will use the notations ``$\bav{}$'' and ``$_{*i}$''
 for other quantities.
@@
 %
-For sufficiently small $\alpha(t)$,
+For a sufficiently small $\alpha(t)$,
 the bias potential remains roughly the same for a short period,
@@
 We will approximate the histogram fluctuation part, $\xi_i(t)$,
-and hence its deviation from the bin average, $\xi_{*i}(t)$,
+%and hence its deviation from the bin average, $\xi_{*i}(t)$,
 as white noise such that
@@
 We expect the approximation to hold
-for simulations that are
-much longer than the autocorrelation time.
+for simulations much longer than the autocorrelation time.
 
@@
 we can write the total error at the end of period $T$
-as a sum of a residual error, $E_R$ (from the decay of the initial error)
-and an asymptotic error, $E_A$:
+as a sum of a residual error, $E_R(T)$ (from the decay of the initial error)
+and an asymptotic error, $E_A(T)$:
 %
@@
 %
-Our aim is to find the $q(t)$
+We wish to find the $q(t)$
 that minimizes this error.
@@
 This is equivalent to choosing a value of $t_0$
-\big[because $q(T) = \ln\bigl(1 + \frac{T}{t_0}\bigr)$,
-from the $t = 0$ case of Eq. \eqref{eq:expqt}\big].
+because $q(T) = \ln\bigl(1 + \frac{T}{t_0}\bigr)$
+from the $t = 0$ case of Eq. \eqref{eq:expqt}.
 %
@@
 
-We now turn to a general multiple-bin updating scheme.
+We now turn to the general multiple-bin updating scheme.
 %
@@
   ,
-  \notag \\
-  \operatorname{var} \tilde v_k(0)
-  &=
-  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
-  \left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
-  ,
-  \label{eq:xtvar}
+  \label{eq:xt_eql}
   \\
+  %\operatorname{var} \tilde v_k(0)
+  %&=
+  %\frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
+  %\left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
+  %,
+  %\label{eq:xtvar}
+  %\\
   \left\langle \tilde v_k^2(0) \right\rangle
@@
   \epsilon_k \, e^{-2 \, \lambda_k \, a_0 \, T_0}
-  .
+  ,
   \label{eq:xt2_eql}
@@
 \begin{align}
-  \eta\bigl( q(T) \bigr)
+  \Delta\bigl( q(T) \bigr)
   &\equiv
@@
   and imposed a lower bound,
-  $\bigl|\eta\bigl(q(T)\bigr)/q(T)\bigr|$, for
-  $\eta'\bigl(q(T)\bigr)$
+  $\bigl|\Delta\bigl(q(T)\bigr)/q(T)\bigr|$, for
+  $\Delta'\bigl(q(T)\bigr)$
   in computing its change.
@@
 we can design a homogeneous updating scheme
-that is more effective in error control
+that is more convenient in error control
 than the Gaussian updating scheme.
@@
 As shown below,
-the optimized updating scheme will spare the need to
-go through the general procedure in Sec. \ref{sec:procedure}
-because the optimal schedule is given by
-the inverse-time formula.
+the optimized updating scheme
+allows the optimal schedule to coincide with
+the inverse-time formula,
+and thus spare the need to
+go through the general procedure in Sec. \ref{sec:procedure}.
 
@@
 \end{align*}
-which is essentially the diagonal updating matrix
-that represents the single-bin updating scheme.
+which is a multiple of the identity updating matrix
+that defines the single-bin updating scheme.
 %
@@
 followed by a filtering step
-that completely removes the components of the noise modes of $k \ge K$
+that eliminates the components of the noise modes of $k \ge K$
 from the bias potential.
@@
 %
-These updating schemes are more intuitive
+The resulting polynomial bandpass updating schemes are more intuitive
 than the homogeneous bandpass updating scheme
-as the former expresses the bias potential as
+in that the former express the bias potential as
 a simple polynomial of the collective variable.
@@
 \begin{align*}
-  R_0(z) = 1,
-  R_1(z) = \frac{z - \overline{z}}{\sigma_z},
+  R_0(z) = 1, \;
+  R_1(z) = \frac{z - \overline{z}}{\sigma_z}, \;
   R_2(z) = \frac{ (z - \overline{z})^2 - \sigma_z^2 } {\kappa_z},
@@
 %
-with the coefficients updated as
+with the variable coefficients updated as
 %
@@
 to an array of adjacent intervals
-to cover a wide range of the collective variable, $z$.
+to cover a wide range of the collective variable, $z$.\cite{langfeld2012, pellegrini2014}
 %
@@
 as in the previous case,
-we need to use
+we need to use,
 in addition to the linear mode,
@@
 %
-The bias potential is thus
-quadratic\cite{maragliano2006, *abrams2008, zhu2012, neuhaus2006, *neuhaus2007},
-which is common in umbrella sampling
+The bias potential is thus quadratic,
+which is common in umbrella sampling\cite{maragliano2006, *abrams2008, zhu2012}
 and the Gaussian ensemble method\cite{hetherington1987,
-*challa1988, *costeniuc2006}.
-
-Upon convergence,
-the parameters, $c_1$ and $c_2$,
-can be related to the local slope and curvature of the PMF
-[cf. Eq. \eqref{eq:Vi_target}].
-%
-By Eq. \eqref{eq:uz_decomp}, we have
-\begin{equation}
-  u(z) = c_0 + \frac{c_1}{\sigma_z} (z - z_c)
-  + \sqrt 2 \, c_2 \left[ \frac{(z-z_c)^2}{2 \, \sigma_z^2} - \frac{1}{2} \right]
-  .
-  \label{eq:u_age}
-\end{equation}
-Thus, we can interpret
-$c_1/\sigma_z$ as the mean force around $z_c$.
-%
-The coefficient
-$\sqrt{2} \, c_2$ represents the relative ``pressure''
-to restrain the width of the distribution to
-be the same as the target Gaussian,
-and it would exceed $1$ in a barrier region
-where the PMF is locally concave,
-i.e. $-d^2\ln p(z)/dz^2 < 0$.
+*challa1988, *costeniuc2006, neuhaus2006, *neuhaus2007}.
 
@@
 To sample this distribution, we give each umbrella
-an independent bias potential, $u^{(s)}(z)$,
-of the form of Eq. \eqref{eq:u_age},
+an independent bias potential of the form of Eq. \eqref{eq:uz_decomp},
+%
+\begin{equation}
+  %u^{(s)}(z)
+  u(s, z)
+  = \hat c_0^{(s)}
+  + \frac{c_1^{(s)}}{\sigma_z^{(s)}} \bigl(z - z_c^{(s)}\bigr)
+  + \sqrt 2 \, c_2^{(s)} \frac{\bigl(z-z_c^{(s)}\bigr)^2}
+  {2 \, \bigl( \sigma_z^{(s)} \bigr)^2}
+  ,
+  \label{eq:u_age}
+\end{equation}
+%
+with $\hat c_0^{(s)} \equiv c_0^{(s)} - c_2^{(s)}/\sqrt 2$,
 and update the coefficients as
-\begin{align*}
+%
+\begin{align}
   c_k^{(s)}(t+1) = c_k^{(s)}(t)
@@
   ,
-\end{align*}
+  \label{eq:ck_age_update}
+\end{align}
 %
@@
 %
-Since $R_0^{(s)}(z) = 1$,
-the parameter, $c_0^{(s)}$,
-serves as the bias potential
-for inter-umbrella transitions,
-and it is updated according to the single-bin updating scheme.\cite{li2007}
-%
-With a small width, $\sigma_z^{(s)}$, the converged
-$c_0^{(s)}$ as a function of $z_c^{(s)}$
-also serves as a good approximation of
-the negative PMF, $\ln p(z)$.\cite{maragliano2006, *abrams2008}
-%
-Thus, we have\cite{park2007, *nguyen2013, *zhang2015st}
-\begin{equation}
-  c_0^{(s+1)} - c_0^{(s)}
-  \approx
-  \frac{1}{2}\left(
-    \frac{ c_1^{(s+1)} } { \sigma_z^{(s+1)} }
-    +
-    \frac{ c_1^{(s)} } { \sigma_z^{(s)} }
-  \right)
-  \left(
-  z_c^{(s+1)} - z_c^{(s)}
-  \right)
-  .
-  \label{eq:c0_est}
-\end{equation}
 
@@
 %
-%and upon convergence, $\exp c_1^{(s)}$ should be proportional to
-%the partition function,
-%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(s)}(z(\mathbf x),t)} \, d\mathbf x$,
-%as in simulated tempering,
-%where
-%$\beta = 1/(k_B T)$ is the inverse temperature,
-%$\mathbf x$ denotes the internal degrees of freedoms,
-%and
-%$\Phi(\mathbf x)$ is the unbiased potential energy.
-%
+In this case,
+the parameter, $c_0^{(s)}$,
+serves as the bias potential
+for inter-umbrella transitions,
+and it is updated by Eq. \eqref{eq:ck_age_update},
+much like in the single-bin updating scheme.\cite{li2007}
 
@@
 
+Upon convergence,
+the parameters, $c_k^{(s)}$,
+can be approximately related to the PMF.
+By Eqs. \eqref{eq:Vi_target} and \eqref{eq:rho_Gaussian}, we have
+\begin{align*}
+  u(s, z)
+  \approx
+  \ln p(z)
+  +
+  \frac{ \bigl(z - z_c^{(s)} \bigr)^2 }
+  { 2 \, \bigl( \sigma_z^{(s)} \bigr)^2 }
+  +
+  \mathrm{const.}
+\end{align*}
+%
+Comparing this to Eq. \eqref{eq:u_age}, we get
+\begin{align*}
+  \hat c_0^{(s)}
+  &\approx \ln p\bigl(z_c^{(s)}\bigr) + \mathrm{const.}
+  ,
+  \\
+  c_1^{(s)}
+  &\approx \left. \frac{d \ln p(z) } { d z } \right|_{z = z_c^{(s)}}
+  ,
+  \\
+  \frac{
+    \sqrt 2 \, c_2^{(s)} - 1
+  } { \bigl( \sigma_c^{(s)} \bigr)^2 }
+  &\approx \left. \frac{d^2 \ln p(z) } { d z^2 } \right|_{z = z_c^{(s)}}
+  ,
+\end{align*}
+%
+Thus,
+the converged $c_0^{(s)}$ as a function of $z_c^{(s)}$
+approximates the negative PMF, $\ln p(z)$;\cite{maragliano2006, *abrams2008}
+$c_1^{(s)}/\sigma_z^{(s)}$
+approximates the mean force at $z = z_c^{(s)}$.
+%
+and $\sqrt{2} \, c_2^{(s)}$,
+serving as the relative ``pressure''
+to restrain the width of the distribution,
+would exceed $1$ in a barrier region
+where the PMF is locally concave,
+i.e. $-d^2\ln p(z)/dz^2 < 0$.
+%
+%
+We also have approximately\cite{park2007, *nguyen2013, *zhang2015st}
+\begin{equation}
+  \hat c_0^{(s+1)} - \hat c_0^{(s)}
+  \approx
+  \frac{1}{2}\left(
+    \frac{ c_1^{(s+1)} } { \sigma_z^{(s+1)} }
+    +
+    \frac{ c_1^{(s)} } { \sigma_z^{(s)} }
+  \right)
+  \left(
+  z_c^{(s+1)} - z_c^{(s)}
+  \right)
+  .
+  \label{eq:c0_est}
+\end{equation}
+%
+%and upon convergence, $\exp c_1^{(s)}$ should be proportional to
+%the partition function,
+%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(s)}(z(\mathbf x),t)} \, d\mathbf x$,
+%as in simulated tempering,
+%where
+%$\beta = 1/(k_B T)$ is the inverse temperature,
+%$\mathbf x$ denotes the internal degrees of freedoms,
+%and
+%$\Phi(\mathbf x)$ is the unbiased potential energy.
+%
+
 We summarize the method as the following procedure.
@@
     For the polynomial coefficients, we can set
-    $c_1^{(s)}/\sigma_z^{(s)}$ to some estimated mean force, $f_z$,
-    $c_0^{(s)} = f_z \bigl[ z_c^{(s)} - \frac{1}{2}(z_{\min} + z_{\max}) \bigr]$
+    $c_1^{(s)}/\sigma_z^{(s)}$ to some estimated mean force, $\eta_z$,
+    $c_0^{(s)} = \eta_z \bigl[ z_c^{(s)} - \frac{1}{2}(z_{\min} + z_{\max}) \bigr]$
     by Eq. \eqref{eq:c0_est},
@@
 with the reference bias potential
-pre-calculated from a longer simulation.
+pre-calculated from a longer simulation
+under the single-bin updating scheme.
 %
@@
 By a delayed reduction,
-the optimal schedule more effectively reduced
-the error in these modes,
+the optimal schedule reduced
+more error in these modes,
 and produced a flatter profile
@@
 %
-Then, from Eq. \eqref{eq:xtvar}, we get
+Then, from Eqs. \eqref{eq:xt_eql} and \eqref{eq:xt2_eql}, we have
 %
@@
   \operatorname{var} {\tilde v}_{k}
-  =
+  \approx
   \frac{1}{2} \, \Gamma_k \, \lambda_k \, a_0
-  \left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
-  .
+  %\left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
+  ,
 \label{eq:varXt}
@@
 %
+for a sufficiently long period, $T_0$.
 %where we have used the fact that the ensemble average of ${\tilde v}_{k}$
@@
 %
-In practice, Eq. \eqref{eq:varXt} would fail for a near zero eigenvalue.
-%
-So it is best used for the single-bin updating scheme.%\footnote{Alternatively,
+Note that Eq. \eqref{eq:varXt} would fail for a near zero, $\lambda_k$,
+and it is best used for the single-bin updating scheme.%\footnote{Alternatively,
 %  we can split the trajectory into blocks,
@@
 and modify Eq. \eqref{eq:lambda_bandpass}
-by setting $\lambda_{n-K-1}, \dots, \lambda_{n-1}$
+by setting $\lambda_{n-K+1}, \dots, \lambda_{n-1}$
 also to $1$.

### 2017-09-13 doc/invt.tex
allow checkeq.py to check duplicated equation labels

@@
 Below we will use the notations ``$\bav{}$'' and ``$_{*i}$''
-for other quantities.
+for other variables.
 %
@@
 
-We will approximate the histogram fluctuation part, $\xi_i(t)$,
+We will approximate the histogram fluctuation part, $\xi_{*i}(t)$,
 %and hence its deviation from the bin average, $\xi_{*i}(t)$,
 as white noise such that
+%
 \begin{equation}
@@
   ,
-  \notag
-  %\label{eq:G_def}
-\end{equation}
+  \label{eq:Gi_def}
+\end{equation}
+%
 for some constant $G_i$,
@@
     \dot{\theta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
-  \label{eq:xt_solution}
+  \label{eq:vi_solution}
 \end{equation}
@@
 
-From Eqs. \eqref{eq:error_def} and \eqref{eq:xt_solution},
+From Eqs. \eqref{eq:error_def},
+\eqref{eq:vi_solution}, and
+\eqref{eq:Gi_def},
 we can write the total error at the end of period $T$
@@
 %
-where $t_0$ and $c$ are two constants,
-and the latter has been determined
+where $t_0$ is some constant,
+and the previous constant, $c$, has been determined
 from the $t = T$ case
@@
   ,
-  \notag
+  &
+  %\notag
   %\label{eq:ER_sbin1}
@@
   ,
-  \label{eq:EA_sbin1}
+  %\label{eq:EA_sbin1}
 \end{align}
@@
 %
-We will label this eigenvector by $k = 0$,
+We will label this eigenvector by index $k = 0$,
 and
@@
 %
-The formal solution is given by
+The formal solution of Eq. \eqref{eq:vt_diffeq_mbin}
+is given by
 \begin{equation}
@@
   ,
-  \label{eq:xt_solution}
+  \label{eq:vk_solution}
 \end{equation}
@@
 This shows the equivalence of
-Eq. \eqref{eq:EA_sbin1} and \eqref{eq:error_asym}
+Eq. \eqref{eq:EA_sbin} and \eqref{eq:error_asym}
 in the case of the single-bin updating scheme, with $\lambda_k \equiv 1$.
@@
 before the preliminary run is zero,
-we have, from Eq. \eqref{eq:xt_solution} [with $T\to T_0, q(t) \to a_0 \, t$],
+we have, from Eq. \eqref{eq:vk_solution} [with $T\to T_0, q(t) \to a_0 \, t$],
 %
@@
 was roughly captured
-by $c_0^{(s)}$ as a function of $E_c^{(s)}$,
+by $\hat c_0^{(s)}$ as a function of $E_c^{(s)}$,
 as shown Fig. \ref{fig:pt_hist}(b).
@@
     %
-    (b) $c_0^{(s)}$ against $E_c^{(s)}$ (circles)
+    (b) $\hat c_0^{(s)}$ against $E_c^{(s)}$ (circles)
     roughly agreed with the logarithmic density of states,
@@
   zhou2005, morozov2007, zhou2008, barducci2008, dama2014}
-our analysis is inapplicable
-to early stages of a simulation
+it is invalid in
+early stages of a simulation
 and we still have to resort to the WL stage-by-stage
-technique even for the bandpass updating schemes.
+technique even for the single-bin and bandpass updating schemes.
 %the random error,
@@
 whereas the gain of using the optimal schedule
-can sometimes be rather modest,
-especially for relatively short simulations,
-even if the noise is largely local in nature.
+can sometimes be rather modest
+for relatively short simulations,
+even if the histogram fluctuation is largely
+local noise.
 %
@@
 than in the single-bin updating scheme,
-and the updating magnitude is larger in the former case.
+resulting in a larger updating magnitude
+in the former case.
 
@@
 
-Here we first review the relation
-between the histogram fluctuation and the error
-in a long FDS simulation under the single-bin updating scheme.
-%
-We then give a measure of the overall histogram fluctuation
-that can be used in the WL criterion of stage transitions
-for the polynomial bandpass updating schemes.
+Here we first relate the error
+in a long FDS simulation under the single-bin updating scheme
+to a measure of histogram fluctuation,
+which can be used in the WL criterion of stage transitions.
+%
+We then modify the measure
+such that it applies to
+the polynomial bandpass updating schemes.
 

### 2017-09-14 doc/invt.tex
improve web

@@
 While an overly slow reduction of the updating magnitude
-is ineffective in restricting the variance of the bias potential,
+can be ineffective in restricting the variance of the bias potential,
 a too rapid reduction often
@@
 %
-Below we will use the notations ``$\bav{}$'' and ``$_{*i}$''
-for other variables.
-%
 The total error is then given by
@@
 \end{equation}
+%
+Below we will use the notations ``$\bav{}$'' and ``$_{*i}$''
+for other variables.
 
@@
 by a differential equation\footnote{In
-going to the continuous-time setup,
-we find it more convenient to shift the origin of time by $-1$,
+  switching to the continuous-time setup,
+  we find it convenient to shift the origin of time by $-1$,
   e.g. the sum $\sum_{i=1}^T$ is mapped to the integral $\int_0^T dt$.}
@@
 we find that
-the optimal $q(T)$ satisfies
+the optimal $q(T)$ that minimizes $\Err_R(T) + \Err_A(T)$
+satisfies
 %
@@
   }{M^2\bigl( q(T) \bigr)}
-  \ge
+  \stackrel{\epsilon_k \to 0}{=\joinrel=\joinrel=}
   \frac{ a_0 }
@@
 and in this limit, the optimal initial updating magnitude
-is half of the previous equilibrium value.\note{This factor, $1/2$,
+is half of the value
+in the preliminary run.\note{This factor, $1/2$,
   happens to coincide with the
@@
 generally has no simple closed-form expression.
-However, we can show that it always lies above
-the inverse-time schedule
-(cf. Appendix \ref{sec:schedule_geometry}).
+Nonetheless, we can show that it always lies above
+the inverse-time schedule,
+and may assume a roughly exponentially-decaying form
+initially (cf. Appendix \ref{sec:schedule_geometry}).
 
@@
     \label{fig:mat}
-    Updating matrices for
+    Updating matrices for two
     homogeneous updating schemes of
@@
     \sum_{k = 1}^{n-1} \Gamma_k
-      \Bigl(
+      \bigl(
         1 - e^{ -\lambda_k \, q(T) }
-      \Bigr)
+      \bigr)
   \Biggr]^{\!2}
@@
 %
-The first few members are
+The first few members are\footnote{The
+  orthogonal polynomials share the same form
+  for different target distributions,
+  but the parameters,
+  $\overline z$,
+  $\sigma_z$,
+  $\kappa_z$, etc.,
+  are determined by the target distribution, $\rho(z)$.}
 \begin{align*}
@@
 We can readily verify that the eigenvectors satisfy
-Eqs. \eqref{eq:eig_orthonormal_cols}, \eqref{eq:eig_orthonormal_rows},
-and \eqref{eq:ortho0}.
+Eqs. \eqref{eq:eig_orthonormal_cols}-\eqref{eq:ortho0}.
 %
@@
 
-For $k > 0$, $R_k(z)$ is a polynomial of degree $k$ with zero expectation.
-%
-Thus, $R_k\bigl(z(t)\bigr)$ roughly
+Note that although the target distribution, $\rho(z)$,
+does not explicitly appear in the bias potential, Eq. \eqref{eq:uz_decomp},
+or the updating scheme, Eq. \eqref{eq:ckupdate},
+it affects both via the parameters,
+$\overline z$,
+$\sigma_z$,
+$\kappa_z$, etc.
+%
+For $k > 0$, $R_k(z)$ is a polynomial of degree $k$ with zero expectation,
+%
+and $R_k\bigl(z(t)\bigr)$ roughly
 represents the deviation of the $k$th statistical moment
@@
   %
-  In the former case, the bias potential is quadratic
-  and the target distribution is Gaussian in each umbrella,
-  although the inter-umbrella distribution, $\rho_s$, is flat.
+  In the former case, the bias potential, $u^{(s)}(z)$,
+  is quadratic
+  and the conditional intra-umbrella target distribution, $\rho^{(s)}(z)$,
+  is Gaussian,
+  although the marginal inter-umbrella distribution, $\rho_s$, is flat.
   %
@@
 \begin{equation}
-  %u^{(s)}(z)
-  u(s, z)
+  u^{(s)}(z)
+  %u(s, z)
   = \hat c_0^{(s)}
@@
 the parameters, $c_k^{(s)}$,
-can be approximately related to the PMF.
+can be related to the PMF.
 By Eqs. \eqref{eq:Vi_target} and \eqref{eq:rho_Gaussian}, we have
 \begin{align*}
-  u(s, z)
+  u^{(s)}(z)
+  %u(s, z)
   \approx
@@
 Comparing this to Eq. \eqref{eq:u_age}, we get
-\begin{align*}
+\begin{align}
   \hat c_0^{(s)}
@@
   ,
-  \\
+  \label{eq:c0hat_limit} \\
   c_1^{(s)}
@@
   ,
-  \\
+  \label{eq:c1_limit} \\
   \frac{
@@
   &\approx \left. \frac{d^2 \ln p(z) } { d z^2 } \right|_{z = z_c^{(s)}}
-  ,
-\end{align*}
+  .
+  \label{eq:c2_limit}
+\end{align}
 %
 Thus,
-the converged $c_0^{(s)}$ as a function of $z_c^{(s)}$
+$\hat c_0^{(s)}$ as a function of $z_c^{(s)}$
 approximates the negative PMF, $\ln p(z)$;\cite{maragliano2006, *abrams2008}
@@
     For the polynomial coefficients, we can set
-    $c_1^{(s)}/\sigma_z^{(s)}$ to some estimated mean force, $\eta_z$,
+    $c_1^{(s)}/\sigma_z^{(s)}$ to some estimated mean force, $\eta_z$;
+    $c_2^{(s)} = 0$;
+    and
     $c_0^{(s)} = \eta_z \bigl[ z_c^{(s)} - \frac{1}{2}(z_{\min} + z_{\max}) \bigr]$
-    by Eq. \eqref{eq:c0_est},
-    and $c_2^{(s)} = 0$.
+    by Eq. \eqref{eq:c0_est}.
 
@@
   Perform a regular MC or MD step
-  in the current umbrella $s$ under bias potential $u^{(s)}(z, t)$.
+  in the current umbrella $s$ under
+  the bias potential $u^{(s)}(z, t)$.
 
@@
 After simulation,
-we can use the weighted histogram analysis method
-(WHAM)\cite{ferrenberg1988, *ferrenberg1989,
+we can use the multiple histogram method\cite{ferrenberg1988, *ferrenberg1989,
 *kumar1992, *roux1995, *bartels1997, *souaille2001,
@@
   ,
-  \notag
+  \label{eq:WHAM}
 \end{equation}
@@
 %
-These particles, labeled as $1$ and $2$,
+The two particular particles, labeled as $1$ and $2$,
 are mutually non-interacting
@@
 
-We used adaptive FDS to calculate
+We used biased sampling to calculate
 the PMF from $r = 0$ to half of the box size
@@
 %
-Since the optimal and inverse-time schedules are identical
-in the single-bin and bandpass updating schemes,
-we will assume the Gaussian updating scheme with width $\sigma = 0.2$ below
-in discussing the optimal schedule.
 
@@
 the optimal schedule for the Gaussian updating scheme
-was more complex than the inverse-time one,
-and the former lay above the latter,
+with width $\sigma = 0.2$
+was more complex than the inverse-time one
+(We chose the Gaussian updating scheme here because
+the optimal and inverse-time schedules were identical
+in the single-bin and bandpass updating schemes.)
+%
+The optimal schedule lay above the inverse-time schedule,
+and its initial decay was roughly exponential,
 in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
@@
     \label{fig:lj_alpha}
-    Optimal schedule of the Gaussian updating scheme.
+    Optimal schedule of a Gaussian updating scheme.
   }
@@
 The optimal schedule more effectively reduced
-the error than the inverse-time schedule,
+the error than the inverse-time schedule
+for the Gaussian updating scheme,
 as shown in Table \ref{tab:lj_error}.
@@
 By a delayed reduction,
-the optimal schedule reduced
+the optimal schedule removed 
 more error in these modes,
@@
 For the Gaussian updating scheme,
-the error depended not only the schedule
-but also the width of the Gaussian.
+the error depended not only on the schedule
+but also on the width of the Gaussian.
 %
@@
 We spread the Gaussian umbrellas
-of the same width $\sigma_E = L$ over $[-1.8L^2, -0.8L^2]$
+of the same width $\sigma_E^{(s)} = L$ over $[-1.8L^2, -0.8L^2]$
 with an even spacing of $L$.
@@
   wider umbrellas,
-  one may mix the nearest-neighbor inter-umbrella transitions
+  one may need to mix the nearest-neighbor inter-umbrella transitions
   with some longer-range transitions
@@
 Initially, we set
-$c_1^{(s)} = \tilde \beta_c \, \sigma_E$
+$c_1^{(s)} = \tilde \beta_c \, \sigma_E^{(s)}$
 for all umbrellas
@@
 (which was more efficient than the Metropolis algorithm)
-at the inverse temperature, $\beta = c_1^{(s)}/\sigma_E$,
+at the inverse temperature, $\beta = c_1^{(s)}/\sigma_E^{(s)}$,
 followed by a Metropolis step with
@@
 %
-We computed the density of states by WHAM,
+We computed the density of states by Eq. \eqref{eq:WHAM},
 and found the critical temperature to be $T_c = 0.70163$,
@@
 %
-Since umbrellas were relatively narrow,
+Since Gaussian umbrellas were sufficiently narrow,
+we could readily verify
+Eqs. \eqref{eq:c0hat_limit}-\eqref{eq:c2_limit}:
 the logarithmic density of states, $\ln g(E)$,
@@
 by $\hat c_0^{(s)}$ as a function of $E_c^{(s)}$,
-as shown Fig. \ref{fig:pt_hist}(b).
-%
-As discussed in Sec. \ref{sec:age},
-$c_1^{(s)}/\sigma_E$ roughly represents
+as shown Fig. \ref{fig:pt_hist}(b);
+%
+$c_1^{(s)}/\sigma_E^{(s)}$ roughly representing
 the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
-and it manifested
-the backbending behavior\cite{kim2006, *kim2007, kim2010}
+manifested the backbending behavior\cite{kim2006, *kim2007, kim2010}
 typical in a first-order phase transition,
-as shown in Fig. \ref{fig:pt_hist}(c).
-%
-The coefficient, $\sqrt 2 \, c_2^{(s)}$,
+as shown in Fig. \ref{fig:pt_hist}(c);
+%
+$\sqrt 2 \, c_2^{(s)}$,
 was indeed above $1.0$
@@
     %
-    (b) $\hat c_0^{(s)}$ against $E_c^{(s)}$ (circles)
+    (b) $\hat c_0^{(s)} = c_0^{(s)} - c_2^{(s)}/\sqrt 2$
+    against $E_c^{(s)}$ (circles)
     roughly agreed with the logarithmic density of states,
@@
 %
-The optimal schedule improves the convergence
+The optimal schedule can improve the convergence
 of the bias potential,
-and thus helps minimize the effort
+and thus reduce the effort
 of computing the PMF.
@@
 %
-This scheme can be generalized to
+The single-bin updating scheme can be generalized to
 a class of bandpass updating schemes,
@@
 %
-This construction serves as a generalization
+This construction leads to a generalization
 of the LLR algorithm\cite{langfeld2012, pellegrini2014},
@@
   zhou2005, morozov2007, zhou2008, barducci2008, dama2014}
-it is invalid in
+it is inadequate in
 early stages of a simulation
-and we still have to resort to the WL stage-by-stage
-technique even for the single-bin and bandpass updating schemes.
+(This is why we still have to resort to the WL stage-by-stage
+technique even for the single-bin and bandpass updating schemes.)
 %the random error,
@@
 which is the average of $R_k^{(s)}\bigl( z(t) \bigr)$
-for samples in umbrella $s$.
+for the data points in umbrella $s$.
 %If the eigenmodes are made of a sequence of orthogonal polynomials

### 2017-09-14 prog/mtrand.h

@@ -218,5 +218,13 @@ __inline static int mtload(const char *fn, unsigned long seed)
 }
 
 
+/* randomly accept by probability exp(-de) */
+__inline int metroacc(double de)
+{
+  double r;
+  if ( de <= 0 ) return 1;
+  r = rand01();
+  return r <= exp(-de);
+}
 
 #endif

### 2017-09-14 doc/invt.tex
improve reweight.c and web

@@
 \item \label{step:Gamma}
-Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
+Estimate $\Gamma_k$'s
 (cf. Appendix \ref{sec:Gamma_measure})
-as well as the initial bias $\epsilon_k$.\footnote{The
+and the initial bias $\epsilon_k$.\footnote{The
   value of $\tilde u_k$ can be estimated
@@
 the largest term, where $\lambda_k \approx 1/\bar q$,
-and $m(\bar q) \sim 1/\bar q$,
+and $m(\bar q) \propto M(\bar q) \propto 1/\bar q$,
 and we would expect a roughly exponentially-decaying schedule.
@@
   \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
-  .
+  ,
   \notag
@@
 \end{align}
+%
+where we have used Eq. \eqref{eq:parseval} in the last step.
 %

### 2017-09-14 web/js/age.js

@@ -51,6 +51,7 @@ function AGE(xcmin, xcmax, delx, sig,
   this.acc = new Array(n);
   this.hfl = new Array(n);
   this.delx = delx;
+  this.bc = c1/sig;
   for ( i = 0; i < n; i++ ) {
     this.ave[i] = xcmin + i * delx;
     this.sig[i] = sig;
@@ -225,9 +226,7 @@ var LN0 = -1000000;
 function lnadd(x, y)
 {
   if ( x < y ) { // swap x and y
-    var z = y;
-    y = x;
-    x = z;
+    var z = y; y = x; x = z;
   }
   y -= x;
   return x + Math.log(1 + Math.exp(y));
@@ -283,12 +282,25 @@ function findpeak(arr, bc, ileft, iright, xmin, dx)
 /* find the critical inverse temperature */
 AGE.prototype.seekcrit = function()
 {
-  var bc, beta, w, sw, y1, y2, y, lns, lns2;
-  var ix, ix1, ix2, il = -1, ir, im, t, x, ret;
+  var bc, beta, y1, y2, y, lns, lns2;
+  var i, ix, ix1, ix2, il = -1, ir, im, t, x, ret;
   var xn = this.xn, n = this.n, dx = this.dx, xmin = this.xmin;
   var arr = this.lng;
 
-  bc = sw = 0;
+  // since c1/sigma is roughly beta, we will estimate the critical temperature
+  // at a point where c1/sigma rises quickest
+  var dbmax = 0;
+  for ( i = 0; i < n - 1; i++ ) {
+    var b0 = this.c1[i]/this.sig[i];
+    var b1 = this.c1[i+1]/this.sig[i+1];
+    var db = (b1 - b0)/(this.ave[i+1] - this.ave[i]);
+    if ( db > dbmax ) {
+      dbmax = db;
+      bc = b0;
+      im = Math.floor((this.ave[i] - this.xmin)/this.dx);
+    }
+  }
+
   for ( ix = 0; ix < xn - 1; ix++ ) {
     if ( arr[ix] <= LN0 || arr[ix+1] < LN0 ) {
       continue;
@@ -297,20 +309,9 @@ AGE.prototype.seekcrit = function()
     } else {
       ir = ix;
     }
-    if ( this.htot[ix] > 0 && this.htot[ix+1] > 0 ) {
-      beta = (arr[ix + 1] - arr[ix]) / dx;
[…]

### 2017-09-14 doc/invt.tex
correct a mistake for the homogeneous bandpass window function

@@
 %
-Then, $\mu_l = \mathrm{sinr}(K, l, n)$ and $w_{ij} = \mu_{i-j}$,
+Then, $\mu_l = \operatorname{sinr}(K, l, n)$ for $l \le b$
+and $w_{ij} = \operatorname{sinr}(K, i-j, n)$,
 where we have defined
 \begin{equation}
-  \mathrm{sinr}(K, l, n)
+  \operatorname{sinr}(K, l, n)
   \equiv
+  \begin{dcases}
     \frac{
@@
     }
-  .
+    & l \ne 0, \\
+    \frac{2K-1}{n} & l = 0. 
+  \end{dcases}
 \notag
@@
 from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan},
-and $\mu_l = (-1)^{n-K+l}/(2 \, n) + \mathrm{sinr}(K, l, 2 \, n)$.
+and $\mu_l = (-1)^{n-K+l}/(2 \, n) + \operatorname{sinr}(K, l, 2 \, n)$
+for $l \le b$.
 %\begin{equation}
@@
 %    +
-%  \mathrm{sinr}(K, l, 2n)
+%  \operatorname{sinr}(K, l, 2n)
 %  .
@@
 The updating matrix is given by
-$w_{ij} = \mathrm{sinr}(K, i-j, 2 \, n) + \mathrm{sinr}(K, i+j-1, 2 \, n)$.
+$w_{ij} = \operatorname{sinr}(K, i-j, 2 \, n) + \operatorname{sinr}(K, i+j-1, 2 \, n)$.
 %

### 2017-09-14 prog/invt.h

@@ -264,6 +264,7 @@ __inline static int savewinmat(double *win, int winn,
       /* add contributions from ks[0], ks[1], ks[2] */
       for ( l = 0; l < 3; l++ ) {
         k = ks[l];
+        if ( pbc && k * 2 == -n ) continue;
         if ( k < winn && k > -winn ) {
           if ( k >= 0 ) {
             y += win[k];
@@ -393,6 +394,7 @@ __inline static double *prepwin(double *lambda, int n,
 {
   int i;
 
+  for ( i = 0; i < n; i++ ) win[i] = 0;
   if ( gaussig > 0 ) {
     mkgauswin(gaussig, n, pbc, win, winn);
   } else if ( kc > 0 ) {

### 2017-09-15 doc/reviewers/replyV2.tex
histogram-based estimate of gamma

\documentclass{article}
\begin{document}

We thank the referees for their thoughtful critique of our work. We have taken their suggestions and have returned a much improved manuscript.

\section{Referee 1}

We have worked to better explain our mathematical analysis for flat-distribution-sampling errors from updating schemes. We also moved the method of computing the optimal schedule from the appendix to the main text and explained the solution of $q(T)$ as well as the integration scheme for the mass function as requested. We now include an analysis of the errors in the problem of computing the potential of mean force for a dense Lennard-Jones fluid and a nontrivial application of the bandpass updating scheme on the $64\times64$ 10-state Potts model.

\begin{enumerate}

\item
In Appendix C, we have improved the discussion the method of estimating $\Gamma_k$ from an FDS simulation under constant magnitude.
This method does not require explicit computation
of the autocorrelation functions.
It only requires accumulation of data for
the Fourier transformed bias potential,
just like what we usually do for computing the average and variance of the potential energy.
Thank you.

\item
As the reviewer correctly stated, the computed optimal schedule is indeed only for a subgroup simulations sharing the same value of $q(T)$.
We now better explain the variational problem and its solution of varying $q(t)$ on the interval $[0,T]$ at the beginning of Sec II C 3.
Thank you.

\item
We now discuss the effect of the time of the preliminary run with a constant magnitude in the Results section via an example.
Indeed, a short preliminary run can lead to unreasonably high initial updating magnitude, and thus should be avoided.
Thank you for pointing this out.

\item
We now discuss the consequences of the sign of the elements of $\mathbf w$.
%
The elements can be negative, but we use only a subset of properties of the transition matrix theory that are unaffected by the admission of negative elements.
Thanks.

\item
(Change $p_i$ to $p_{i(t)}$.)
Done as suggested.
Thanks.

\item
We believe that the error weighting distribution is rather arbitrary, and our choice was one of convenience and we now better discuss and justify the choice.
For simplicity, we now assume that the target distribution, $p_i$, and error weighting distribution, $\rho_i$, are the same.
The target distribution, $p_i$, hopefully serves as a natural and convenient choice for error weights.
Thank you.

\end{enumerate}

\section{Referee 2}

1a) We had made distinctions between WL and metadynamics for convenience in our analysis,
but we now take the referee's point and emphasize the close relation and distinction between the methods to avoid any confusion. Thank you for the advice.

1b) True and we have made the wording clear that we will study only single-bin and multiple-bin updating schemes instead of WL and metadynamics. Thanks.

1c) We agree. Done. Thanks.

1d) Corrected. Thanks.

2a) We have greatly shortened section IIA. Thanks.

2b) We have also shortened IIB as requested. Thanks.

3a) We have removed the ideal system example and replaced it with a calculation of the PMF in a Lennard-Jones system. Thank you for the suggestion.

3b) Same as above.

3c) As suggested by the reviewer, aside from the statistical error, the systematic error can also play a key role in the optimal schedule and the final error.  This is particular so for certain multiple-bin updating scheme, where the systematic bias can be nonnegligible even after a long simulation time. We have reflected on this point and made a few corrections in the procedure of computing the optimal schedule, in particular Eq. (40) in estimating the effect of the initial bias. Mathematically, we believe that the framework still works, but we need to more carefully choose the value of $q(T)$ in the second step of the optimization to ensure sufficient reduction of the systematic bias.  We now also illustrate this point on the Lennard-Jones fluid test case, and emphasize the optimal schedule calculated here only applies to long simulations.

We agree that the effective exploration of the phase space can be a more important problem. To illustrate potential applications in this area, we now show that the bandpass updating scheme can be constructed from a orthogonal polynomial basis, serving as a generalization of the Langfeld-Lucini-Rago algorithm, which is modification of the WL and has found successful applications in quantum field theory. We wish to show that similar modifications are possible. To this end, we have considered the $64\times64$ 10-state Potts model to illustrate the application of bandpass updating schemes in parameter tuning of the Gaussian ensemble method. Hopefully we may use the method in the future on more interesting (difficult) examples. Here we give the theory and practical, familiar examples. Thank you.

\end{document}

### 2017-09-15 prog/cmvar.h

@@ -57,7 +57,7 @@ __inline static void cmvar_close(cmvar_t *cm)
 }
 
 /* deposit a vector into the mode */
-__inline static void cmvar_add(cmvar_t *cm, double *v)
+__inline static void cmvar_add(cmvar_t *cm, const double *v)
 {
   double uave, du;
   long cnt = cm->cnt;

### 2017-09-15 prog/is2.c

@@ -83,7 +83,7 @@ static int gammrun(invtpar_t *m, metad_t *metad, is2_t *is)
     metad->tmat[icur*metad->n + iold] += 1;
     iold = icur;
     if ( t % m->gam_nstave == 0 )
-      metad_varv_add(metad);
+      metad_varv_add(metad, metad->v);
   }
 
   /* estimate gamma */

### 2017-09-15 doc/invt.tex
histogram-based estimate of gamma
clean up age code, c2 = 1/sqrt(2) initially

@@
 asymptotically optimal,
-but only for the first $K$ fluctuation modes.
+but only up to the first $K$ fluctuation modes.
 
@@
     $c_1^{(s)}/\sigma_z^{(s)}$ to some estimated mean force, $\eta_z$;
-    $c_2^{(s)} = 0$;
+    $c_2^{(s)} = 1/\sqrt 2$;
     and

### 2017-09-15 prog/aus/age.h

@@ -6,7 +6,6 @@
 
 
 #include "mtrand.h"
-#include "mmwl.h"
 
 
 #ifndef SQRT2
@@ -18,12 +17,12 @@ typedef struct {
   double *ave;
   double *sig;
   double *c0, *c1, *c2;
+  double (*mm)[3];
   double alphawl; /* updating magnitude */
   double hfluc, flfr[3];
   double t, t0;
   int invt; /* using 1/t schedule */
   double alpha0; /* initial updating magnitude */
-  mmwl_t *mmwl;
   double *cnt;
   double *acc;
   int xmin, xmax, dx;
@@ -50,7 +49,7 @@ __inline static age_t *age_open(double xcmin, double xcmax,
   xnew(age->c0, n);
   xnew(age->c1, n);
   xnew(age->c2, n);
-  xnew(age->mmwl, n);
+  xnew(age->mm, n);
   xnew(age->cnt, n);
   xnew(age->acc, n);
   xnew(age->hfl, n);
@@ -59,8 +58,10 @@ __inline static age_t *age_open(double xcmin, double xcmax,
     age->sig[i] = sig;
     age->c0[i] = (i - (n - 1)*0.5) * c1 / sig * delx;
     age->c1[i] = c1;
-    age->c2[i] = 0;
-    mmwl_init(age->mmwl + i, age->alpha0);
+    age->c2[i] = SQRT2*0.5;
+    age->mm[i][0] = 0;
+    age->mm[i][1] = 0;
+    age->mm[i][2] = 0;
     age->cnt[i] = 0;
     age->acc[i] = 0;
     age->hfl[i] = 0;
@@ -96,7 +97,7 @@ __inline static void age_close(age_t *age)
   free(age->c0);
   free(age->c1);
   free(age->c2);
-  free(age->mmwl);
+  free(age->mm);
   free(age->cnt);
   free(age->acc);
   free(age->hist);
@@ -131,8 +132,7 @@ __inline static void age_save(age_t *age, const char *fn)
 {
   int i, n = age->n;
   FILE *fp;
-  mmwl_t *mm;
-  double alpha, tot = 0, acc = 0;
+  double alpha, tot = 0, acc = 0, mcnt;
 
   if ( (fp = fopen(fn, "w")) == NULL ) {
     fp = stderr;
[…]

### 2017-09-15 web/js/pottsmain.js

@@ -26,7 +26,7 @@ var timer_interval = 100; // in milliseconds
 var potts_timer = null;
 var mc_algorithm = "Metropolis";
 
-var blocksizemc = 10; // number of steps for a block in the Metropolis algorithm
+var metroblk = 10; // number of steps for a block in the Metropolis algorithm
 var nstepspsmc = 1000; // number of steps per second for MC
 var nstepspfmc = 100;  // number of steps per frame for MC
 
@@ -59,7 +59,7 @@ function getparams()
 
   mc_algorithm = grab("mc_algorithm").value;
 
-  blocksizemc = get_int("blocksizemc", 10);
+  metroblk = get_int("metroblk", 10);
   nstepspsmc = get_int("nstepspersecmc", 10000);
   nstepspfmc = nstepspsmc * timer_interval / 1000;
   npulses = 0;
@@ -381,7 +381,7 @@ function updatec0plot()
     var w = h * 3 / 2;
     var options = {
       xlabel: "<small>Energy</small>",
-      ylabel: "<small><i>c</i><sub>0</sub> - <i>c</i><sub>2</sub>/&radic;2 - <i>&beta;<sub>c</sub>E<sub>c</sub></i></small>",
+      ylabel: "<small><i>c</i><sub>0</sub> - <i>c</i><sub>2</sub>/&radic;2 - <i>&beta;<sub>c</sub> E<sub>c</sub></i></small>",
       axisLabelFontSize: 10,
       xRangePad: 2,
       drawPoints: true,
@@ -408,10 +408,10 @@ function pulse()
     var Eave = age.ave[iage];
     var acc = 0;
     if ( mc_algorithm === "Metropolis" ) {
-      for ( jstep = 0; jstep < blocksizemc; jstep++ ) {
+      for ( jstep = 0; jstep < metroblk; jstep++ ) {
         acc += potts_metro_mod(beta1, beta2, Eave);
       }
-      acc = 1.0*acc/blocksizemc;
+      acc = 1.0*acc/metroblk;
     } else if ( mc_algorithm === "Wolff" ) {
       acc = potts_wolff_mod(beta1, beta2, Eave);
     }
@@ -428,8 +428,8 @@ function pulse()
     + ", E/N " + roundto(1.0*potts.E/potts.n, 2) + ", data " + age.cnt[iage]
[…]

### 2017-09-15 doc/invt.tex
lj.c estimate of Gamma_k for multiple-bin updating scheme

@@
 a rigid Gaussian function centered on the current bin
-(with some boundary adjustment),
+(with some boundary adjustment).
 %to satisfy the constraints discussed in Sec. \ref{sec:updating-matrix},
-see Fig. \ref{fig:mat}(a) for an example.
+%see Fig. \ref{fig:mat}(a) for an example.
 %
@@
 
-\begin{figure}[h]\centering
-  \makebox[\linewidth][c]{
-    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
-  }
-  \caption{
-    \label{fig:mat}
-    Updating matrices for two
-    homogeneous updating schemes of
-    a non-periodic variable ($n = 500$).
-    Here $j$ is the current bin,
-    $i$ is the bin where updating occurs,
-    and $w_{ij}$ is the relative updating magnitude.
-    %
-    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
-    (b) The homogeneous bandpass updating scheme of $K = 20$.
-    %
-    Both updating schemes
-    can be characterized by a rigid window
-    centered on the current bin, $j$,
-    with some boundary adjustments.
-  }
-\end{figure}
+%\begin{figure}[h]\centering
+%  \makebox[\linewidth][c]{
+%    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
+%  }
+%  \caption{
+%    \label{fig:mat}
+%    Updating matrices for two
+%    homogeneous updating schemes of
+%    a non-periodic variable ($n = 500$).
+%    Here $j$ is the current bin,
+%    $i$ is the bin where updating occurs,
+%    and $w_{ij}$ is the relative updating magnitude.
+%    %
+%    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
+%    (b) The homogeneous bandpass updating scheme of $K = 20$.
+%    %
+%    Both updating schemes
+%    can be characterized by a rigid window
+%    centered on the current bin, $j$,
+%    with some boundary adjustments.
+%  }
+%\end{figure}
 
@@
 which resembles the Gaussian updating scheme,
-but with long-range wiggles,
-as shown in Fig. \ref{fig:mat}(b).
+but with long-range wiggles.
+%as shown in Fig. \ref{fig:mat}(b).
 %
@@
 Comparing this to Eq. \eqref{eq:u_age}, we get
+\begin{subequations}
 \begin{align}
@@
   ,
-  \label{eq:c0hat_limit} \\
+  %\label{eq:c0hat_limit}
+  \\
   c_1^{(s)}
@@
   ,
-  \label{eq:c1_limit} \\
+  %\label{eq:c1_limit}
+  \\
   \frac{
@@
   .
-  \label{eq:c2_limit}
+  %\label{eq:c2_limit}
 \end{align}
+\label{eq:ck_limit}
+\end{subequations}
 %
@@
 we could readily verify
-Eqs. \eqref{eq:c0hat_limit}-\eqref{eq:c2_limit}:
+Eqs. \eqref{eq:ck_limit}:
 the logarithmic density of states, $\ln g(E)$,
@@
 %
-$c_1^{(s)}/\sigma_E^{(s)}$ roughly representing
+$c_1^{(s)}/\sigma_E^{(s)}$, roughly representing
 the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
@@
     %
-    (a) Normalized histograms (solid lines)
-    and the reweighted distribution
-    at the critical temperature, $\beta_c$ (dashed line).
+    (a) Normalized histograms.
     %
@@
     and applied a proper shift so as to normalize
-    the distribution at $\beta_c$.
-    %
-    (c) $c_1^{(s)}/\sigma_E$ roughly corresponded to
-    the statistical temperature, $\beta(E)$.
+    the distribution at the critical temperature, $\beta_c$.
+    %
+    (c) $c_1^{(s)}/\sigma_E$ (circles), roughly corresponding to
+    the statistical temperature, $\beta(E) = d\ln g(E)/dE$;
+    $\beta_c$ (dashed line).
     %
@@
   assuming a proper shift of $u_i(t)$.
-  Then after time averaging, we get
-  $\rho_i e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
+  Time averaging the equation yields
+  $\rho_i \, e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
 %
@@
 
-Here we give a method of measuring
+Here we give two methods of estimating
 the integrals of the autocorrelation functions, $\Gamma_k$'s,
@@
 %
-This method has the advantage of not requiring
+They have the advantage of not requiring
 explicit computation of
@@
 
-Because the difference between $u_i$ and $v_i$,
+First, because the difference between $u_i$ and $v_i$,
 is a constant of time [cf. Eq. \eqref{eq:v_def}],
@@
 for a sufficiently long period, $T_0$.
-%where we have used the fact that the ensemble average of ${\tilde v}_{k}$
-%is zero in the second step.
-%
-The variance, $\operatorname{var} \tilde u_k$, can be computed
-by trajectory averaging in a long simulation.
-%
-Note that Eq. \eqref{eq:varXt} would fail for a near zero, $\lambda_k$,
-and it is best used for the single-bin updating scheme.%\footnote{Alternatively,
-%  we can split the trajectory into blocks,
-%  compute $\Gamma_k$ from the histogram fluctuation of each block
-%  using Eq. \eqref{eq:Fk2_Gammak},
-%  and then average the results.}
+%
+We can compute the variance, $\operatorname{var} \tilde u_k$,
+by trajectory averaging in a long simulation,
+and use Eq. \eqref{eq:varXt} to estimate $\Gamma_k$.
+%
+However, this method would fail for a near zero, $\lambda_k$,
+and it is best used for the single-bin updating scheme,
+where $\lambda_k \equiv 1$.
+
+To use the method in a simulation under a multiple-bin updating scheme,
+we can construct a virtual bias potential, $u_i^\mathrm{vir}$,
+that emulates the bias potential in a simulation
+under the single-bin updating scheme.
+%
+This virtual bias potential can be constructed as (cf. Appendix \ref{sec:ave_us}),
+\begin{equation}
+  u_i^\mathrm{vir}(t+1)
+  =
+  u_i^\mathrm{vir}(t)
+  +
+  a_0 \, [\hat u_i(t) - u_i^\mathrm{vir}(t)]
+  ,
+  \notag
+\end{equation}
+%
+where $\hat u_i(t)$ is given by Eq. \eqref{eq:uhatt}.
+%
+We can then apply Eq. \eqref{eq:varXt} to $u_i^\mathrm{vir}$
+with the effective $\lambda_k \equiv 1$.
+
+In this method, the updating magnitude, $a_0$,
+needs to be sufficiently small, such that
+$a_0 < 1/\left(\max_k \Gamma_k \right)$.
+
+Alternatively, we can estimate $\Gamma_k$ using Eq. \eqref{eq:Fk2_Gammak}.
+%
+We can split the trajectory into blocks of $T_\mathrm{blk}$ steps,
+compute the histogram fluctuation, $\tilde F_k(T_\mathrm{blk})$,
+and estimate $\Gamma_k$ from the average of
+$T_\mathrm{blk} \, \tilde F_k^2(T_\mathrm{blk})$
+over the blocks.
+%
+For this method to work,
+the trajectory block should be large enough
+to cover at least a few autocorrelation times (which is roughly $\Gamma_k/2$),
+but also sufficiently small
+to avoid the artificial histogram flattening
+caused by the frequent non-equilibrium updates.
+Thus, we require $\max_k \Gamma_k < T_\mathrm{blk} < 1/a_0$
+(We used $T_\mathrm{blk} = 0.1/a_0$.)
 
@@
   ,
-  \notag
-  %\label{eq:Fk2_Gammak}
+  %\notag
+  \label{eq:Fk2_Gammak}
 \end{align}

### 2017-09-15 prog/metad.h

@@ -504,9 +504,9 @@ __inline static void metad_getgamma_varv(metad_t *metad, double alpha0,
 
   cmvar_get(metad->cm);
   for ( i = 1; i < n; i++ ) {
-    double lam = metad->lambda[i];
-    if ( lam < 0.1 ) lam = 0.1;
-    gamma[i] = metad->cm->uvar[i]*2/alpha0/lam;
+    //double lam = metad->lambda[i];
+    //if ( lam < 0.1 ) lam = 0.1;
+    gamma[i] = metad->cm->uvar[i]*2/alpha0; // /lam;
   }
   if ( fn != NULL ) metad_savegamma(metad, gamma, 1, fn);
 }
@@ -763,6 +763,22 @@ static double metad_geterrav(metad_t *metad, double *erc)
   *erc = metad_geterr(metad, metad->vcorr);
   return metad_geterr(metad, metad->v);
 }
+
+/* update the WEIGHTED average histogram-corrected bias potential */
+__inline static void metad_vav_wupdate(metad_t *metad, int i, double a)
+{
+  int j, n = metad->n;
+  double v0, v1, vold, vnew;
+
+  v0 = metad_getave(metad, metad->v);
+  v1 = metad_getave(metad, metad->vav);
+  for ( j = 0; j < n; j++ ) {
+    vold = metad->vav[j] - v1;
+    vnew = metad->v[j] - v0 + (i == j ? n : 0) - 1;
+    metad->vav[j] = vold + a * (vnew - vold);
+  }
+}
+
 
 __inline static void metad_hblk_clear(metad_t *metad)
 {

### 2017-09-18 doc/invt.tex
add ljgam.c for estimating Gamma in a simulation with constant updating magnitude

@@
 the optimal schedule was pre-computed using a set of fixed
-$\Gamma_k$'s and $\epsilon_k$'s derived from the longer simulation.
+$\Gamma_k$'s and $\epsilon_k$'s derived
+from a longer simulation under
+a smaller updating magnitude, $a_0 = 10^{-5}$.
+\note{Maybe using a simulation with constant updating magnitude
+using the block histogram method?}
 
@@
 %
+The initial errors
+at a relatively large updating magnitude
+were generally underestimated,
+suggesting that
+the regular updating might have caused the sampling
+to deviate from the quasi-equilibrium one
+as assumed in our theory.
 
@@
     the single-bin,
-    Gaussian (with width $\sigma = 0.2$),
-    and homogeneous bandpass (with mode cutoff $K = 20$)
+    Gaussian (with width $\sigma = 0.29$),
+    and homogeneous bandpass (with mode cutoff $K = 26$)
     updating schemes
@@
     Single-bin & $1/t$
-    & $5.7(5.7)\times10^{-2}$
-    & $9.1(11.4)\times10^{-6}$
+    & $5.7(4.4)\times10^{-2}$
+    & $9.1(8.8)\times10^{-6}$
     \\
@@
     Gaussian & $1/t$
-    & $6.7(7.8)\times10^{-3}$
-    & $1.8(2.2)\times10^{-5}$
+    & $6.5(5.9)\times10^{-3}$
+    & $1.4(1.4)\times10^{-5}$
     \\
@@
     Gaussian & optimal
-    & $6.7(7.8)\times10^{-3}$
-    & $2.9(3.6)\times10^{-6}$
+    & $6.0(5.9)\times10^{-3}$
+    & $2.6(2.6)\times10^{-6}$
     \\
@@
     Bandpass & $1/t$
-    & $8.0(8.9)\times10^{-3}$
-    & $2.0(2.3)\times10^{-6}$
+    & $8.0(7.6)\times10^{-3}$
+    & $1.7(1.6)\times10^{-6}$
     \\
@@
 the optimal schedule for the Gaussian updating scheme
-with width $\sigma = 0.2$
-was more complex than the inverse-time one
-(We chose the Gaussian updating scheme here because
+with width $\sigma = 0.29$
+was more complex than the inverse-time one.
+%
+We chose the Gaussian updating scheme here because
 the optimal and inverse-time schedules were identical
-in the single-bin and bandpass updating schemes.)
+in the single-bin and bandpass updating schemes.
 %
@@
 we found that the inverse-time schedule
-failed to reduce to the errors in the $k = 9$ to $31$ modes adequately.
+failed to reduce to the errors in the $k = 7$ to $21$ modes adequately.
 %
@@
 more error in these modes,
-and produced a flatter profile
-(The remaining unevenness of the profile
-might be due to errors of the estimated values of
-$\Gamma_k$'s and $\epsilon_k$'s.)
+and produced a flatter profile.
+%(The remaining unevenness of the profile
+%might be due to errors of the estimated values of
+%$\Gamma_k$'s and $\epsilon_k$'s.)
 %
@@
 the final errors under the optimal and inverse-time schedules
-became $2.4\times 10^{-5}$ and $3.9\times 10^{-5}$, respectively,
+became $2.1\times 10^{-5}$ and $3.3\times 10^{-5}$, respectively,
 which were more similar.
@@
 the optimal width (assuming the optimal schedule)
-was around $\sigma = 0.29$ in this case
-with the predicted minimal error being $3.3\times 10^{-6}$,
+was $\sigma = 0.29$, which we used in the simulations,
 and it would decrease with the simulation length
@@
 The mode cutoff, $K$, also has an optimal value,
-which was around $26$ in this case.
-%
-The predicted minimal error was $2.0\times 10^{-6}$,
-  slightly less than the corresponding value
-  in the Gaussian case.
+which was $26$ according to the prediction.
+%
+The minimal error of the bandpass case
+was slightly less than the corresponding value
+in the Gaussian case, as shown in Table \ref{tab:lj_error}.
 %
@@
 We computed the density of states by Eq. \eqref{eq:WHAM},
-and found the critical temperature to be $T_c = 0.70163$,
+and found the critical temperature to be $T_c = 0.70161$,
 in agreement with a previous study\cite{wang2001pre}.
@@
 
-In this method, the updating magnitude, $a_0$,
-needs to be sufficiently small, such that
-$a_0 < 1/\left(\max_k \Gamma_k \right)$.
+The above method
+requires a sufficiently small
+the updating magnitude,
+$a_0 \ll 2/\left(\max_k \Gamma_k \right)$,
+and it tends to overestimate the value of $\Gamma_k$,
+since Eq. \eqref{eq:varXt}
+does not take into account
+the non-equilibrium sampling effect
+caused by frequent updating.
 
@@
 caused by the frequent non-equilibrium updates.
-Thus, we require $\max_k \Gamma_k < T_\mathrm{blk} < 1/a_0$
-(We used $T_\mathrm{blk} = 0.1/a_0$.)
+%
+Thus, we require $\max_k \Gamma_k/2 \ll T_\mathrm{blk} \ll 1/a_0$.
 

### 2017-09-18 prog/lj/ljgam.c

@@ -1,0 +1,188 @@
+#include "lj.h"
+#include "../ave.h"
+#include <time.h>
+
+int np = 100;
+double rho = 0.1;
+double rcdef = 100.0;
+double tp = 3.0;
+int mcblk = 5;
+int hblksz = 0;
+double delr = 0.01; /* spacing */
+long nstsave = 1000000; /* interval of saving data */
+const char *fnvbias = "vbias.dat";
+const char *fnvref = "vref.dat";
+const char *fnlog = "verr.log";
+const char *fnxerr = "xerr.dat";
+
+static void ljpar_help(void)
+{
+  fprintf(stderr, "  --np=:      number of Lennard Jones particles, default %d\n", np);
+  fprintf(stderr, "  --rho=:     density of Lennard Jones particles, default %g\n", rho);
+  fprintf(stderr, "  --rc=:      distance cutoff, default %g\n", rcdef);
+  fprintf(stderr, "  --tp=:      temperature, default %g\n", tp);
+  fprintf(stderr, "  --dr=:      bin width, default %g\n", delr);
+  fprintf(stderr, "  --blk=:     number of steps in a Monte Carlo, default %d\n", mcblk);
+  fprintf(stderr, "  --hblk=:    number of steps in a histogram block, default %d\n", hblksz);
+  fprintf(stderr, "  --vref=:    reference bias potential, default %s\n", fnvref);
+}
+
+static int ljpar_keymatch(invtpar_t *m, const char *key, const char *val)
+{
+  if ( strcmpfuzzy(key, "np") == 0 ) {
+    np = invtpar_getint(m, key, val);
+  } else if ( strcmpfuzzy(key, "rho") == 0 ) {
+    rho = invtpar_getdouble(m, key, val);
+  } else if ( strcmpfuzzy(key, "rc") == 0
+           || strcmpfuzzy(key, "rcutoff") == 0 ) {
+    rcdef = invtpar_getdouble(m, key, val);
+  } else if ( strcmpfuzzy(key, "tp") == 0
+           || strcmpfuzzy(key, "temp") == 0 ) {
+    tp = invtpar_getdouble(m, key, val);
+  } else if ( strcmpfuzzy(key, "dr") == 0
[…]

### 2017-09-19 doc/invt.tex
change sig 0.29 to 0.28

@@
     the single-bin,
-    Gaussian (with width $\sigma = 0.29$),
+    Gaussian (with width $\sigma = 0.28$),
     and homogeneous bandpass (with mode cutoff $K = 26$)
@@
     Single-bin & $1/t$
-    & $5.7(4.4)\times10^{-2}$
-    & $9.1(8.8)\times10^{-6}$
+    & $5.7(4.5)\times10^{-2}$
+    & $9.1(9.1)\times10^{-6}$
     \\
@@
     Gaussian & $1/t$
-    & $6.5(5.9)\times10^{-3}$
+    & $6.5(6.2)\times10^{-3}$
     & $1.4(1.4)\times10^{-5}$
@@
     Gaussian & optimal
-    & $6.0(5.9)\times10^{-3}$
-    & $2.6(2.6)\times10^{-6}$
+    & $6.0(6.2)\times10^{-3}$
+    & $2.7(2.7)\times10^{-6}$
     \\
@@
     Bandpass & $1/t$
-    & $8.0(7.6)\times10^{-3}$
-    & $1.7(1.6)\times10^{-6}$
+    & $8.0(7.9)\times10^{-3}$
+    & $1.7(1.7)\times10^{-6}$
     \\
@@
 the optimal schedule for the Gaussian updating scheme
-with width $\sigma = 0.29$
+with width $\sigma = 0.28$
 was more complex than the inverse-time one.
@@
 the optimal width (assuming the optimal schedule)
-was $\sigma = 0.29$, which we used in the simulations,
+was $\sigma = 0.28$, which we used in the simulations,
 and it would decrease with the simulation length

### 2017-09-20 doc/fig/bak/scheme.tex
small changes

\newcommand{\note}[1]{{\footnotesize\textrm{#1}}}

\begin{tikzpicture}[scale=1.0, every node/.style={rounded corners=0.1cm, align=center}]\sffamily
  \node[]
    (H) at (-1.6, 3.5)
    {Homogeneous\\
    \note{Sec. II\,C\,7,}\\
    \note{Appendix D}.
    };

  \node[draw, fill=white!90!black]
    (G) at (-1.6, 1.0)
    {Gaussian\\
    (metadynamics)\\
    \note{Sec. II\,C\,6,}\\
    \note{Appendix D\,2,}\\
    \note{Fig. 2(a)}.
    };

  \node[draw, minimum width=6.3cm, minimum height=5cm]
    (Hbox) at (0.0, 2) {};

  \node[draw, minimum width=8cm, minimum height=5.8cm]
    (Bbox) at (3.8, 2.1) {};

  \node[draw, minimum width=3.2cm, minimum height=4.8cm]
    (HBbox) at (1.5, 2) {};

  \node[draw, text width={width("Single-bin ABCD")}, fill=white!90!black]
    (sbin) at (1.5, 0.7)
    {Single-bin\\
    (Wang-Landau)\\
    \note{Sec. II\,A}.
    };

  \node[text width={width("Homogeneous AAAAAA")}]
    (HB) at (1.45, 3.0)
    {Homogeneous\\bandpass\\
    \note{Fig. 2(b),}\\
    \note{Secs. II\,D\,2 and III\,A,}\\
    \note{Appendix D\,3}.
    };

  \node[] (B) at (5.5, 4.5)
    {Bandpass \;
    \note{Sec. II\,D\,2}.
    };

  \node[draw, minimum width=4.3cm, minimum height=4.5cm]
    (PBbox) at (5.5, 1.8) {};

  \node[] (PB) at (5.5, 3.6)
    {Polynomial bandpass\\
    \note{Sec. II\,E}.
    };

  \node[draw, fill=white!90!black]
    (LLR) at (5.5, 2.3)
    {Langfeld-Lucini-Rago\\
    \note{Sec. II\,E\,2}.
    };

  \node[draw, text width={width("Adaptive Gaussian ABCD")}]
    (AGE) at (5.5, 0.7)
    {Adaptive Gaussian\\ensemble\\
    \note{Secs. II\,E\,3, II\,E\,4, and III\,B}.
    };

\end{tikzpicture}

### 2017-09-20 doc/invt.tex
small changes

@@
 Many efficient flat-distribution-sampling (FDS) techniques
-based on adaptive construction of the bias potential
-have been introduced,
-including the Wang-Landau (WL) algorithm\cite{
+are based on adaptive construction of the bias potential,
+e.g. the Wang-Landau (WL) algorithm\cite{
   wang2001, wang2001pre}
@@
 Compared with the Gaussian updating scheme,
-bandpass updating schemes have the advantage of allowing
-simpler control of the updating magnitude
-for optimal convergence.
+bandpass updating schemes allow
+a simple protocol to optimally control the updating magnitude,
+which is borrow from the single-bin updating scheme.
 
@@
 we briefly review the basics of FDS
-and some known aspects of the optimal schedule.
+and introduce our notations.
 %
@@
   ,
-  \notag
-  %\label{eq:pi_p_V}
+  \label{eq:pi_p_V}
 \end{equation}
@@
 We refer to this scheme as
-the \emph{single-bin} (updating) scheme,
+the \emph{single-bin} updating scheme,
 for it applies only to the bias potential
@@
 In a more general
-\emph{multiple-bin} (updating) scheme,
+\emph{multiple-bin} updating scheme,
 %including the Gaussian updating scheme used in metadynamics,
-neighboring bins are also updated:
+the bias potential at neighboring bins is also updated:
 %
@@
 disrupt the underlying equilibrium sampling
-and introduce artificial errors that need to be
+and introduce artificial errors, which can be
 reduced by gradually decreasing the updating magnitude.\cite{
@@
 In this section,
-we derive the optimal $\alpha(t)$
+we derive the optimal schedule, $\alpha(t)$,
 for the single-bin updating scheme,
@@
 as a preparation
-for the more general multiple-bin scheme.
+for the more general multiple-bin case.
 %
@@
 
-We can split the histogram into
+We can split the $f_i(t)$ into
 a deterministic averaged part, $\langle f_i(t) \rangle$,
@@
 and we may assume a quasi-equilibrium sampling process
+[cf. Eq. \eqref{eq:pi_p_V}]
 such that
@@
 where we have assumed the smallness
-of $v_{*i}(t)$ in the linear expansion.
+of $v_{*i}(t)$ in the linear expansion,
+and used Eq. \eqref{eq:x_def} in the last step.
 %
@@
 .
+$$
+But
+$$
+\sum_{ j = 1 }^n \rho_j \, v_{*j}
+=
+\sum_{ j = 1 }^n \rho_j \, v_j - \sum_{j=1}^n \bav{v}  = 0.
 $$
@@
 
-From Eqs. \eqref{eq:error_def},
-\eqref{eq:vi_solution}, and
-\eqref{eq:Gi_def},
+From Eqs. \eqref{eq:error_def} and
+\eqref{eq:vi_solution},
 we can write the total error at the end of period $T$
@@
   &= \Gamma \, \int_0^T \dot \theta^2\bigl( q(t) \bigr) \, dt
-  ,
+  .
   \label{eq:EA_sbin}
 \end{align}
-with
+In Eq. \eqref{eq:EA_sbin},
+we have used Eq. \eqref{eq:Gi_def} and introduced
 %$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
@@
   in general the growth rate is not necessarily positive,
-  it is often the largest eigenvalue
+  it is often the largest, hence positive, eigenvalue
   for ``natural'' updating matrices considered
@@
 for $k = l = 0$,
-while the orthogonality demands
+while the orthogonality between this eigenvector and others demands
 %
@@
 We can now use Eq. \eqref{eq:eig_w} to diagonalize
-the multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
+the multiple-bin updating scheme,
+Eq. \eqref{eq:mbin_update} [using $u_i(t)$ instead of $v_i(t)$],
 as
@@
 as an equivalent white noise.
+%
 In this way,
@@
 Eq. \eqref{eq:EA_sbin} and \eqref{eq:error_asym}
-in the case of the single-bin updating scheme, with $\lambda_k \equiv 1$.
+in the case of the single-bin updating scheme, where $\lambda_k \equiv 1$.
 %In Appendix \ref{sec:Gamma_measure},
@@
   =0
-  ,
+  .
   \label{eq:opt_qT}
 \end{align}
-where
+%where
 %
@@
   by Eq. \eqref{eq:uhatav},
-  although it often leads to overestimated
+  although the approach often overestimates the
   $\epsilon_k$'s for large-$k$ modes.
@@
 %
-In terms of the continuous collective variable, $z = i \, \delta z$,
+In terms of the continuous collective variable, $z$,
 the eigenvalues of this scheme are given by\cite{bussi2006}
@@
 we can design a homogeneous updating scheme
-that is more convenient in error control
+that is more convenient for error control
 than the Gaussian updating scheme.
@@
 that discard certain noise modes
-of short wavelength.
+of short wavelengths.
 %
@@
   %
-  In the former case, the bias potential, $u^{(s)}(z)$,
+  In this case, the bias potential, $u^{(s)}(z)$,
   is quadratic
@@
   %
-  In the latter case, the bias potential is a sum of
+  In the Gaussian updating scheme, the bias potential is a sum of
   Gaussian functions, and the target distribution
@@
 %
-Using the latter framework,
+In the latter framework,
 the transition probability from
@@
 approximates the negative PMF, $\ln p(z)$;\cite{maragliano2006, *abrams2008}
+%
 $c_1^{(s)}/\sigma_z^{(s)}$
-approximates the mean force at $z = z_c^{(s)}$.
-%
-and $\sqrt{2} \, c_2^{(s)}$,
+approximates the mean force at $z = z_c^{(s)}$;
+%
+$\sqrt{2} \, c_2^{(s)}$,
 serving as the relative ``pressure''
@@
     without affecting the correctness of the algorithm.
-    Although some prefers mixing the two steps
+    Although a random mixing of the two steps
+    is sometimes preferred,
     in a random order,
@@
 %
-To ensure the same schedule was used for multiple runs,
-the optimal schedule was pre-computed using a set of fixed
-$\Gamma_k$'s and $\epsilon_k$'s derived
-from a longer simulation under
-a smaller updating magnitude, $a_0 = 10^{-5}$.
-\note{Maybe using a simulation with constant updating magnitude
-using the block histogram method?}
+For the Gaussian updating scheme,
+we used a set of fixed reference values of $\Gamma_k$'s\footnote{We
+  derived the reference $\Gamma_k$'s
+  from a long equilibrium simulation under the reference bias 
+  potential using the block histogram method with $T_\mathrm{blk} = 10^5$ steps
+  (cf. Appendix \ref{sec:Gamma_measure}).
+  We also assumed the reference bias potential in calculating $\epsilon_k$'s.}
+to ensure the same optimal schedule for multiple runs.
 
@@
 %
-The errors roughly matched the predicted values
-calculated using the pre-computed
-$\Gamma_k$'s and $\epsilon_k$'s.
-%
-The initial errors
-at a relatively large updating magnitude
-were generally underestimated,
-suggesting that
-the regular updating might have caused the sampling
-to deviate from the quasi-equilibrium one
-as assumed in our theory.
+While the final errors agreed well with
+the predicted values (which were calculated using the reference
+$\Gamma_k$'s and $\tilde u_k$'s),
+there were considerable derivations
+in the initial errors.
+%
+This shows that
+the assumptions used in our methods,
+including that on quasi-equilibrium sampling,
+could fail with a large updating magnitude.
 
@@
     Gaussian & $1/t$
-    & $6.5(6.2)\times10^{-3}$
-    & $1.4(1.4)\times10^{-5}$
+    & $6.3(6.2)\times10^{-3}$
+    & $1.5(1.4)\times10^{-5}$
     \\
@@
     Gaussian & optimal
-    & $6.0(6.2)\times10^{-3}$
-    & $2.7(2.7)\times10^{-6}$
+    & $6.1(6.2)\times10^{-3}$
+    & $2.9(2.7)\times10^{-6}$
     \\
@@
     Bandpass & $1/t$
-    & $8.0(7.9)\times10^{-3}$
+    & $8.5(8.0)\times10^{-3}$
     & $1.7(1.7)\times10^{-6}$
@@
 and its initial decay was roughly exponential,
-in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
+where were in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
 %
@@
 the optimal schedule removed
-more error in these modes,
+more errors in these modes,
 and produced a flatter profile.
@@
     we have deducted the linear contribution, $\beta_c \, E$,
-    and applied a proper shift so as to normalize
+    and applied a shift so as to normalize
     the distribution at the critical temperature, $\beta_c$.
@@
 %
-The bandpass updating schemes allow convenient control
+The bandpass updating schemes allow simpler control
 for the updating magnitude
-as the optimal schedule
-is simply given by an inverse-time formula.
+as the asymptotically optimal schedule
+is given by an inverse-time formula.
 
@@
 of the bandpass updating schemes
-constructed from a family of orthogonal polynomials,
-which allows the bias potential
-to be conveniently expressed as a low-order polynomial.
+constructed from a family of orthogonal polynomials.
+%
+This construction, 
+as a generalization of the LLR algorithm\cite{langfeld2012, pellegrini2014},
+delegates the biased sampling to a local umbrella
+with a simple polynomial bias potential,
+and applies the flat-distribution sampling
+to the umbrellas located at different places.
+%
 %with the coefficients adjusted on-the-fly.
 %
-This construction leads to a generalization
-of the LLR algorithm\cite{langfeld2012, pellegrini2014},
-and hopefully it could be used for
+While our example is limited to a quadratic bias potential\cite{neuhaus2006, *neuhaus2007, zhu2012},
+this approach may also be used for
 adaptive parameter adjustment
-in umbrella sampling with
-quadratic\cite{neuhaus2006, *neuhaus2007, zhu2012}
-or similar\cite{martin-mayor2007, *sergio2011, *persson2013, kim2010}
-bias potentials.
+for bias potentials of different
+forms\cite{martin-mayor2007, *sergio2011, *persson2013, kim2010}.
 
@@
 and $x(t) - X_\omega(t-1)$
-is the proposed change from the previous average
-from the data point at step $t$.
+is the change from the previous average
+proposed by the data point at step $t$.
 
@@
 Comparing this to Eq. \eqref{eq:X_recur},
-we find that the time evolution of $u_i(t+1)$
-is the same as that of the CMA $X_\omega(t)$
-of $x(t) = \hat u_i(t)$.
+we find that $u_i(t+1)$ follows the same time evolution
+as the CMA $X_\omega(t)$ of $x(t) = \hat u_i(t)$.
 %
@@
 For a general multiple-bin updating scheme,
-each mode of bias potential
+each mode of the bias potential
 corresponds to a CMA with a dynamic weight,
@@
 caused by the adaptive updates,
-and thus still requires a sufficiently small updating magnitude.
+and thus still requires the updating magnitude
+to be sufficiently small.
 
@@
 We can now show that
-for simulations with equal lengths and initial updating magnitudes,
+for simulations with equal lengths and
+equal initial updating magnitudes,
 the optimal schedule for the Gaussian updating scheme
@@
 the largest term, where $\lambda_k \approx 1/\bar q$,
-and $m(\bar q) \propto M(\bar q) \propto 1/\bar q$,
-and we would expect a roughly exponentially-decaying schedule.
+and $M(\bar q) \propto \sqrt{\Gamma_k}/\bar q$.
+%
+In a local sampling process,
+$\Gamma_k$ often rapidly tends to a constant
+with increasing $k$.
+%
+Thus, we have approximately 
+$m(\bar q) \propto M(\bar q) \propto 1/\bar q$,
+which gives a roughly exponentially-decaying schedule.
 
@@
 we can construct a virtual bias potential, $u_i^\mathrm{vir}$,
-that emulates the bias potential in a simulation
-under the single-bin updating scheme.
-%
-This virtual bias potential can be constructed as (cf. Appendix \ref{sec:ave_us}),
+that emulates the bias potential in the single-bin case
+as (cf. Appendix \ref{sec:ave_us}):
+%
 \begin{equation}
@@
 $a_0 \ll 2/\left(\max_k \Gamma_k \right)$,
-and it tends to overestimate the value of $\Gamma_k$,
-since Eq. \eqref{eq:varXt}
-does not take into account
-the non-equilibrium sampling effect
-caused by frequent updating.
+and it tends to overestimate the value of $\Gamma_k$
+(possibly because of the non-equilibrium sampling effect
+caused by frequent updating).
 
@@
 %
-We can split the trajectory into blocks of $T_\mathrm{blk}$ steps,
-compute the histogram fluctuation, $\tilde F_k(T_\mathrm{blk})$,
+By splitting the trajectory into blocks of $T_\mathrm{blk}$ steps,
+we can compute the histogram fluctuation, $\tilde F_k(T_\mathrm{blk})$,
 and estimate $\Gamma_k$ from the average of
@@
 %
-For this method to work,
-the trajectory block should be large enough
+The trajectory block should be large enough
 to cover at least a few autocorrelation times (which is roughly $\Gamma_k/2$),
 but also sufficiently small
+when there are dynamic updates to the bias potential
 to avoid the artificial histogram flattening
@@
 
-Here, we discuss some details
+Here, we discuss some mathematical details
 on the homogeneous updating schemes.
@@
 can only be a function of the difference of indices, $i-j$, modulo $n$,
-and we write it as
+and
 %
@@
 the $\mu_l$'s are symmetric, $\mu_l = \mu_{-l}$,
-and they can be cutoff at $l = b$
+and can be cutoff at $l = b$
 (i.e. $\mu_l = 0$ for $l > b$),
@@
 To find the eigenvectors,
-we define for a periodic variable, $\phi_i$,
-the out-of-boundary values by
+we define a periodic variable, $\phi_i$,
+with the out-of-boundary values given by
 $\phi_i = \phi_{i \pm n}$,
@@
 The characteristic equation,
-$\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$
+$\mathbf w \, \pmb\phi = \lambda \, \pmb\phi$,
 can then be solved by discrete Fourier transform,
@@
 the continuous variable $z$ as
-$z = i \, \delta z$,
+$z = z_0 + i \, \delta z$,
 where
@@
     \, d z,
-  \notag
-  %\label{eq:lambda_int}
+  \label{eq:lambda_int}
 \end{equation}
@@
 %
-Then, from Eq. \eqref{eq:wband_eigenvalue},
+Then, from Eq. \eqref{eq:lambda_int},
 we get Eq. \eqref{eq:lambda_Gaussian},
@@
 is then dominated by the random part,
-%
-\begin{equation}
-  f_{*i}(t) \approx \xi_{*i}(t)
-  ,
+$f_{*i}(t) \approx \xi_{*i}(t)$, and
+%
+\begin{equation}
+  \tilde f_k(t) \approx \tilde \xi_k(t)
+  .
   \label{eq:f_xi}
 \end{equation}
-and
 %the same holds for the Fourier transform,
-$\tilde f_k(t) \approx \tilde \xi_k(t).$
 %
@@
 %
-where we have used Eq. \eqref{eq:parseval} in the last step.
+where we have used Eq. \eqref{eq:parseval} in the second step.
 %

### 2017-09-26 doc/invt.tex
add catxerr.py to combine xerr.dat from different directories

@@
 can often sample the desired distribution,
-frequent updates to the bias potential
-disrupt the underlying equilibrium sampling
-and introduce artificial errors, which can be
-reduced by gradually decreasing the updating magnitude.\cite{
+the frequent updates cause the bias potential to fluctuate constantly,
+and disrupt the underlying equilibrium sampling.
+%
+To reduce the sampling error,
+the updating magnitude is often decreased gradually.\cite{
   belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
@@
 
-We can split the $f_i(t)$ into
-a deterministic averaged part, $\langle f_i(t) \rangle$,
-and a random fluctuating part, $\xi_i(t)$,
-%
-\begin{equation}
-  f_i(t) =
-  \langle f_i(t) \rangle
+We can split the $f_{*i}(t)$ into
+a deterministic averaged part, $\langle f_{*i}(t) \rangle$,
+and a random fluctuating part, $\xi_{*i}(t)$,
+%
+\begin{equation}
+  f_{*i}(t) =
+  \langle f_{*i}(t) \rangle
   +
-  \xi_i(t)
+  \xi_{*i}(t)
   .
@@
   -\alpha(t) \, \left[ v_{*i}(t) - \xi_{*i}(t) \right]
-  .
-  %\notag
-  \label{eq:dxdt_singlebin}
-\end{equation}
-%
-Without the random part, $\xi_{*i}(t)$,
-the deviation of the bias potential, $v_{*i}(t)$,
-would tend to $0$ at long times, reaching convergence.
-%
-The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
+  ,
+  \notag
+  %\label{eq:dxdt_singlebin}
+\end{equation}
+%
+and the formal solution is % of Eq. \eqref{eq:dxdt_singlebin} is
 %
@@
 \end{align}
+%
+Thus, without the random part, $\xi_{*i}(t)$,
+the deviation of the bias potential, $v_{*i}(t)$,
+would tend to $0$ at long times, reaching convergence,
+if $\alpha(t)$ decays sufficiently slowly,
+i.e. $q(T) \to \infty$ as $T \to \infty$.
 
@@
 %$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
-$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
-\approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt$.
-%\begin{equation}
-%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
-%  =
-%  \sum_{i=1}^n \rho_i \int_0^T
-%    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
-%  .
-%  \notag%\label{eq:Gamma_from_G}
-%\end{equation}
+%$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
+%\approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt$.
+\begin{equation}
+  \Gamma
+  \equiv
+  \sum_{i=1}^n \rho_i \, G_i
+  \approx
+  \sum_{i=1}^n \rho_i \int_0^T
+    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
+  .
+  \notag%\label{eq:Gamma_from_G}
+\end{equation}
 
@@
   {\tilde f}_k(t)
-  ,
+  .
   \label{eq:vkupdate}
@@
 %
-or in continuous time, we have, for $k > 0$,
+In continuous time, we have, for $k > 0$,
 %
@@
 where
-we have used Eq. \eqref{eq:xtstar} and
-the transformed version of Eq. \eqref{eq:sh_ave}
+we have Fourier transformed Eq. \eqref{eq:sh_ave}
+and used Eq. \eqref{eq:xtstar}
 in the second step.
@@
 the Hamiltonian is conserved,\cite{goldstein, *landau_mechanics, *arnold, *jose}
-which means that the asymptotic error grows
-linearly with time at a rate of $\mathcal L$.
+i.e. the asymptotic error grows
+linearly with time at a rate of
+$\mathcal L = \mathcal H$.
+%
 We may set
@@
   \frac{ M\bigl( q(T) \bigr) } { T }
-    \int_0^{q(T)} M(q') \, d q'
+    \int_0^{q(T)} \!\!\! M(\bar q) \, d\bar q
   -\frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
@@
 the initial updating magnitude
-by Eq. \eqref{eq:Lagrangian_const}, %\eqref{eq:mQ_invTa},
+by Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:mint}, %\eqref{eq:mQ_invTa},
 \begin{align}
@@
     \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
-  }{M^2\bigl( q(T) \bigr)}
+  }{
+    M^2\bigl( q(T) \bigr)
+  }
   \stackrel{\epsilon_k \to 0}{=\joinrel=\joinrel=}
@@
 In terms of the continuous collective variable, $z$,
-the eigenvalues of this scheme are given by\cite{bussi2006}
+the eigenvalues of the updating matrix are given by\cite{bussi2006}
 \begin{align}
@@
 does not explicitly appear in the bias potential, Eq. \eqref{eq:uz_decomp},
-or the updating scheme, Eq. \eqref{eq:ckupdate},
+or in the updating scheme, Eq. \eqref{eq:ckupdate},
 it affects both via the parameters,
@@
 %
-In practical applications,
+In practice,
 the LLR algorithm is often applied simultaneously
@@
     To check use \texttt{getav.py} under \texttt{prog/lj},
-    which will show the simulation and prediction values.}%
+    which will show the simulation and prediction values.
+    Alternatively, run \texttt{data/lj/update.sh} and copy the values
+    on the last few lines.}%
   }
@@
     Gaussian & optimal
-    & $6.1(6.2)\times10^{-3}$
-    & $2.9(2.7)\times10^{-6}$
+    & $6.3(6.2)\times10^{-3}$
+    & $2.7(2.7)\times10^{-6}$
     \\
@@
     Bandpass & $1/t$
-    & $8.5(8.0)\times10^{-3}$
+    & $8.4(8.0)\times10^{-3}$
     & $1.7(1.7)\times10^{-6}$
@@
 the final errors under the optimal and inverse-time schedules
-became $2.1\times 10^{-5}$ and $3.3\times 10^{-5}$, respectively,
+became $2.2\times 10^{-5}$ and $3.4\times 10^{-5}$, respectively,
 which were more similar.

### 2017-09-26 prog/lj/catxerr.py

@@ -1,0 +1,96 @@
+#!/usr/bin/env python
+
+''' combine xerr.dat from different directories
+Example:
+  python catxerr.py drun14/xerr.dat drun14a/xerr.dat drun14b/xerr.dat -o xerr_c.dat
+'''
+
+import sys, os
+
+class XErr:
+    def __init__(self, fn):
+        s = open(fn).readlines()
+        self.tag = s[0]
+        nsamp = int( s[1][1:].strip() )
+        dat = s[2:]
+        self.n = n = len(dat)
+        self.sx = [[0, 0, 0, 0] for i in range(n)]
+        self.sxx = [[0, 0, 0, 0] for i in range(n)]
+        self.ref = [0]*n
+        self.efref = [0]*n
+        self.eiref = [0]*n
+        for i in range(n):
+            ln = dat[i].strip()
+            a = [float(x) for x in ln.split()]
+            self.ref[i] = ref = a[9]
+            self.efref[i] = a[10]
+            self.eiref[i] = a[11]
+            for k in range(4):
+                a[2*k+2] += ref # average
+                self.sx[i][k] = a[2*k+2] * nsamp
+                # a[2*k+1] is the variance
+                self.sxx[i][k] = (a[2*k+1] + a[2*k+2]**2) * nsamp
+        print "loaded %s samples from %s" % (nsamp, fn)
+        self.nsamp = nsamp
+
+    def add(self, fn):
+        s = open(fn).readlines()
+        tag = s[0]
+        if tag != self.tag:
+            print "file %s has incompatible tagline\n%s%s" % (fn, tag, self.tag)
+        nsamp = int( s[1][1:].strip() )
+        dat = s[2:]
+        n = len(dat)
+        if n != self.n:
+            print "file %s has incompatible n %d != %d" % (fn, n, self.n)
+        for i in range(n):
+            ln = dat[i].strip()
+            a = [float(x) for x in ln.split()]
+            ref = a[9]
+            for k in range(4):
+                a[2*k+2] += ref
+                self.sx[i][k] += a[2*k+2] * nsamp
+                self.sxx[i][k] += (a[2*k+1] + a[2*k+2]**2) * nsamp
+        self.nsamp += nsamp
+        print "added %s samples (now %s) from %s" % (nsamp, self.nsamp, fn)
+
+    def save(self, fn):
+        fp = open(fn, "w")
[…]

### 2017-09-26 prog/lj/getav.py

@@ -39,5 +39,5 @@ for ln in open(fn).xreadlines():
     avf.add(xf)
     avi.add(xi)
 
-print "count %s, init. %g(%g) final %g(%g)" % (
+print "%s: count %s, init. %g(%g) final %g(%g)" % (fn,
         avf.cnt, avi.getave(), xiref, avf.getave(), xfref)

### 2017-09-26 prog/lj/lj.c

@@ -353,6 +353,8 @@ static int work(invtpar_t *m)
       if ( i != 0 ) goto TRIAL_END;
       metad_save_cmvar(metad, cmi, cmf, cci, ccf, "new_xerr.dat");
       //printf("please check new_xerr.dat...\n"); getchar();
+    } else {
+      metad_save_cmvar(metad, cmi, cmf, cci, ccf, "xerr0.dat");
     }
     alpha0 = m->opta ? metad->intq->aarr[0] : metad->a/2;
     fprintf(stderr, "starting %ld metadynamics run of %ld/%ld steps..., a %g (%g), err %g -> %g, loadxerr %d\n",

### 2017-10-03 doc/invt.tex
small changes

@@
   of the updating magnitude
-  for a class of flat-distribution sampling methods
-  for free energy calculation including
-  the Wang-Landau (WL) algorithm and metadynamics.
+  for a class of free energy methods based on flat-distribution sampling,
+  including the Wang-Landau (WL) algorithm and metadynamics.
   %
@@
   a bias potential that offsets
-  the potential of mean force by histogram-based adaptive updates.
+  the potential of mean force by histogram-based updates.
   %
@@
   %
-  We show that
-  the optimal schedule depends on the updating scheme.
-  %
-  While the asymptotically optimal schedule for
+  We show that while the asymptotically optimal schedule for
   the single-bin updating scheme (commonly used in the WL algorithm)
@@
   %
-  Further,
-  we show that the single-bin updating scheme
+  We further show that the single-bin updating scheme
   is optimal for very long simulations,
@@
   %
-  These bandpass updating schemes aim at
-  partially flattening a few long-range histogram modes
+  These bandpass updating schemes target only
+  a few long-range distribution modes
   and their optimal schedule
@@
   Constructed from orthogonal polynomials,
-  bandpass updating schemes serve as generalizations
-  of the Langfeld-Lucini-Rago algorithm
-  and can be used for parameter tuning for the Gaussian ensemble method
-  and umbrella sampling.
+  bandpass updating schemes generalize
+  the WL and Langfeld-Lucini-Rago algorithms
+  as an automatic parameter tuning scheme for umbrella sampling.
   %
@@
 
-Free energy calculation\cite{frenkel, newman} is a central theme
+Free energy calculation\cite{frenkel, newman, *landau_binder} is a central theme
 in computational physics and chemistry
@@
 with regular simulation methods.
-\note{``Regular simulation methods'' means MC and MD,
-and they are more pertinent to ``computational physics and chemistry''
-than ``traditional experiments.''}%
 %
@@
 along a collective variable, $z$, by sampling via either
-Monte Carlo (MC)\cite{frenkel, newman, landau_binder}
+Monte Carlo (MC)\cite{frenkel, newman, *landau_binder}
 or molecular dynamics (MD)\cite{frenkel, karplus2002}
@@
 the dimensionless free energy, or potential of mean force (PMF),
-along the collective variable, $z$,
+%along the collective variable, $z$,
 as $-\ln p(z)$.
@@
 %
-These techniques regularly update the bias potential
-to discourage future visits to previously sampled configurations
-by incremental local elevation of the bias potential.
+These techniques
+discourage future visits to sampled configurations
+by incrementally elevating the bias potential
+around the current value of $z$.
 %
@@
 in the updating window function,
-which specifies
-a neighborhood around the current value of
-$z$ on the free energy surface
-%where the updating should occur
-as well as the relative updating magnitude.
+which specifies,
+apart from on overall updating magnitude,
+the change to the bias potential upon each visit.
 %
@@
 as it provides a smoother bias potential
-that can be readily differentiated
-to derive the bias force.\cite{huber1994,
+suitable for deriving the bias force.\cite{huber1994,
   *laio2002, *laio2008, *barducci2011, *sutto2012, junghans2014}
@@
 and the convergence of the bias potential
-can often benefit from a gradually decreasing
-updating magnitude of the bias potential.
+often benefits from a gradually decreasing
+updating magnitude.
 
@@
 affects the precision and accuracy
-of the final bias potential,
+of the bias potential,
 and hence that of the PMF.\cite{
@@
 %
-While an overly slow reduction of the updating magnitude
-can be ineffective in restricting the variance of the bias potential,
+While an overly slow reduction
+is ineffective in restricting the variance of the bias potential,
 a too rapid reduction often
@@
 a free particle with a position-dependent mass,
-and the final error becomes the action.
+and the final error that of the action.
 %
@@
 
-We will then show by comparing the errors
-of simulations under different updating schemes
-that the single-bin updating scheme
-is optimal for very long simulations.
+We will then show that
+among different updating schemes,
+the single-bin updating scheme
+is optimal in error minimization for very long simulations.
 %without a priori assumption of the smoothness of the PMF.
@@
 a class of bandpass updating schemes
-that flatten only
+that target only
 a few long-range distribution modes.
@@
 bandpass updating schemes allow
-a simple protocol to optimally control the updating magnitude,
-which is borrow from the single-bin updating scheme.
+a simple protocol
+(borrowed from the single-bin updating scheme)
+to optimally control the updating magnitude.
 
@@
 This method serves
-as a generalization of the Langfeld-Lucini-Rago (LLR)
-algorithm\cite{langfeld2012, pellegrini2014},
+as a generalization of the WL and Langfeld-Lucini-Rago (LLR)
+algorithms\cite{langfeld2012, pellegrini2014},
 and we show that
@@
 and the PMF, $-\ln p_i$, can be readily
-deduced from the bias potential. % $u_i$.
+deduced from $u_i$.
 
@@
 %
-and a correction to the bias potential
+and a correction
 is derived from the normalized histogram, $H_i$,
@@
 Note that the asymptotic error
-has the form of an action of a free particle
+has the form of the action of a free particle
 at $\theta\bigl( q(t) \bigr)$,
@@
 
-\subsubsection{Langfeld-Lucini-Rago (LLR) algorithm}
+\subsubsection{\label{sec:LLR}
+Langfeld-Lucini-Rago (LLR) algorithm}
 
@@
 we can use $K = 2$ to limit the basis to the first two modes,
-$R_0(z) = 1$ and $R_1(z) = 2 \sqrt{3} \, (z - z_c)/\Delta z$.
+$R_0(z) = 1$ and $R_1(z) = \sqrt{12} \, (z - z_c)/\Delta z$
+(with $\sigma_z = \Delta z/\sqrt{12}$).
 %
@@
 was recommended for the optimal convergence of $c_1$.\cite{pellegrini2014}
-%
+
 In practice,
-the LLR algorithm is often applied simultaneously
+the above algorithm is often applied simultaneously
 to an array of adjacent intervals
 to cover a wide range of the collective variable, $z$.\cite{langfeld2012, pellegrini2014}
+%
+In this expanded ensemble setup,
+the updating equation for $c_0$ is no longer trivial,
+but corresponds to the WL way of constructing the bias potential
+for transitions among different intervals.
+%
+The WL algorithm is in fact the $K = 1$ case.
+%
+However, the LLR algorithm estimates $c_0$ from
+an approximate relation similar to Eq. \eqref{eq:c0_est} below.
 %
@@
 
-In analogy to the LLR algorithm,
+In analogy to the above algorithm,
 we will consider the case of sampling
-a \emph{non-flat} Gaussian distribution
+a Gaussian distribution
 %$\rho(z) \propto \exp[-(z-z_c)^2/(2\sigma_z^2)]$,
@@
 $\hat c_0^{(s)}$ as a function of $z_c^{(s)}$
-approximates the negative PMF, $\ln p(z)$;\cite{maragliano2006, *abrams2008}
+approximates the negative PMF, $\ln p(z)$,\cite{maragliano2006, *abrams2008}
 %
 $c_1^{(s)}/\sigma_z^{(s)}$
-approximates the mean force at $z = z_c^{(s)}$;
-%
-$\sqrt{2} \, c_2^{(s)}$,
+approximates the mean force at $z = z_c^{(s)}$,
+%
+and $\sqrt{2} \, c_2^{(s)}$,
 serving as the relative ``pressure''
@@
 We also have approximately\cite{park2007, *nguyen2013, *zhang2015st}
+%
 \begin{equation}
@@
 %
-%and upon convergence, $\exp c_1^{(s)}$ should be proportional to
-%the partition function,
-%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(s)}(z(\mathbf x),t)} \, d\mathbf x$,
-%as in simulated tempering,
-%where
-%$\beta = 1/(k_B T)$ is the inverse temperature,
-%$\mathbf x$ denotes the internal degrees of freedoms,
-%and
-%$\Phi(\mathbf x)$ is the unbiased potential energy.
-%
+Note that the LLR algorithm (cf. Sec. \ref{sec:LLR}) implicitly uses
+a similar relation for the estimation of $c_0^{(s)}$,
+with $\hat c_0^{(s)}$ replaced by $c_0^{(s)}$
+(for lack of $c_2^{(s)}$ in that case).
 
@@
 *fenwick2008, *kim2011, *shirts2008,
-zhu2012, newman, frenkel}
+zhu2012, newman, *landau_binder, frenkel}
 to recover the unbiased distribution,
@@
     and the predicted values are shown in parentheses for comparison;
-    $\Err_\mathrm{init.}$ and $\Err_\mathrm{final}$
+    $\Err_\mathrm{initial}$ and $\Err_\mathrm{final}$
     denote the errors at the beginning and end of the production phase, respectively.
@@
     Scheme & Schedule &
-    $\Err_\mathrm{init.}$ &
+    $\Err_\mathrm{initial}$ &
     $\Err_\mathrm{final}$
@@
 and its initial decay was roughly exponential,
-where were in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
+in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
 %
@@
 the initial updating magnitude, $\alpha(0)$,
-was around the ideal value of $a_0/2 = 5\times10^{-5}$
+was around the ideal value of $a_0/2$
 [cf. Eq. \eqref{eq:half_alpha0}].
@@
 $\langle \tilde v_k^2 \rangle$'s,
-by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl}
+using Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl}
 (Fig. \ref{fig:lj_xerr}),
@@
     $\langle \tilde v_k^2 \rangle$'s,
-    of the PMF after $10^8$ production steps.
+    of the PMF.
     %
@@
 was free from major barriers
-such that the histogram fluctuation was mainly local in nature.
+such that the histogram fluctuation was largely local in nature.
 %
@@
 %
-However, one may still prefer the multiple-bin updating schemes
-in MD simulations
-when a smooth bias potential is necessary
-for deriving the bias force.
+However,
+the multiple-bin updating schemes have advantages in MD simulations
+when a smooth bias potential is needed for calculating the force.
 
@@
 In Fig. \ref{fig:lj_xerr},
-we show that this correction produced
-the flatter profile seen in the single-bin case
-for a simulation with a sub-optimal schedule.
+we show that
+for a simulation with a sub-optimal schedule
+this correction produced
+the flatter profile seen in the single-bin case.
 %
@@
 on the two-dimensional $L\times L$ ten-state
-Potts model\cite{wu1982, newman, wang2001, wang2001pre}
+Potts model\cite{wu1982, newman, *landau_binder, wang2001, wang2001pre}
 in the energy space.
@@
 For configuration sampling in each umbrella,
-we used the Wolff algorithm\cite{wolff1989, newman}
+we used the Wolff algorithm\cite{wolff1989, newman, *landau_binder}
 (which was more efficient than the Metropolis algorithm)
@@
 $\min\bigl\{1, e^{c_2^{(s)} [R_2(E) - R_2(E')]} \bigr\}$,
-where $E$ and $E'$ are the energies of the old and new configurations.
+where $E$ and $E'$ are the energies of
+the old and new configurations, respectively.
 %
@@
 We computed the density of states by Eq. \eqref{eq:WHAM},
-and found the critical temperature to be $T_c = 0.70161$,
+and found the critical temperature
+that makes the two distribution peaks of equal height
+to be $T_c = 0.70161$,
 in agreement with a previous study\cite{wang2001pre}.
@@
 the optimal schedule of the updating magnitude
-for a general class of free energy calculation methods,
-which we refer to as adaptive flat-distribution-sampling (FDS) methods
-that include the Wang-Landau (WL) algorithm and metadynamics.
-%
-Adaptive FDS methods
+for a general class of free energy calculation methods
+based on flat-distribution sampling,
+including the Wang-Landau (WL) algorithm and metadynamics.
+%
+These methods
 incrementally update a bias potential
@@
 For the single-bin updating scheme (as often used in the WL case),
-whose updating matrix is the identity matrix with identical eigenvalues,
+whose updating matrix is the identity matrix with equal eigenvalues,
 we confirmed that the optimal schedule
@@
 the different rates of error reduction of the eigenmodes,
-is usually not given in a simple closed form,
+usually has no simple closed-form expression,
 but is implicitly given by an equation of motion
@@
 a class of bandpass updating schemes,
-which discard a few noise modes
-and are similarly optimal for the non-noise modes.
+which discard a few noise eigenmodes
+and are similarly optimal for the non-noise eigenmodes.
 %
@@
 This construction, 
-as a generalization of the LLR algorithm\cite{langfeld2012, pellegrini2014},
+as a generalization of
+the WL and LLR\cite{langfeld2012, pellegrini2014} algorithm,
 delegates the biased sampling to a local umbrella
@@
   zhou2005, morozov2007, zhou2008, barducci2008, dama2014}
-it is inadequate in
-early stages of a simulation
-(This is why we still have to resort to the WL stage-by-stage
-technique even for the single-bin and bandpass updating schemes.)
+it is inadequate in early stages of a simulation.
+%(This is why we still have to resort to the WL stage-by-stage
+%technique even for the single-bin and bandpass updating schemes.)
 %the random error,
@@
 whereas the gain of using the optimal schedule
-can sometimes be rather modest
-for relatively short simulations,
+is sometimes rather modest
+for a relatively short simulation,
 even if the histogram fluctuation is largely
@@
 Finally, the adaptive Gaussian ensemble method
-introduced in Sec. \ref{sec:age} may still need refinement
+introduced in Sec. \ref{sec:age} can be refined
 to improve its general applicability and stability.
@@
 
-First, because the difference between $u_i$ and $v_i$,
+First, because the difference between $u_i$ and $v_i$
 is a constant of time [cf. Eq. \eqref{eq:v_def}],
@@
 
-The above method
-requires a sufficiently small
-the updating magnitude,
+The above method requires a sufficiently small updating magnitude,
 $a_0 \ll 2/\left(\max_k \Gamma_k \right)$,
 and it tends to overestimate the value of $\Gamma_k$
-(possibly because of the non-equilibrium sampling effect
-caused by frequent updating).
+(possibly because of the deviation from
+the quasi-equilibrium sampling).
 
@@
 We then modify the measure
-such that it applies to
+to include intra-umbrella components for
 the polynomial bandpass updating schemes.
@@
 the fluctuation of the instantaneous histogram
-is then dominated by the random part,
+is dominated by the random part,
 $f_{*i}(t) \approx \xi_{*i}(t)$, and
@@
 for the histogram fluctuation of mode $k$,
-then the total fluctuation is
+the total fluctuation is given by
 %
@@
   \bigl\langle \tilde F_k^2(T) \bigr\rangle
-  &=
+  &\approx
   \frac{1}{T}
@@
 As an estimate of the error,
-the total histogram fluctuation given by Eq. \eqref{eq:F2sum}
+the above total histogram fluctuation
 can be used in the WL criterion for stage transitions.

### 2017-10-05 doc/invt.tex
absorb Justin's changes

@@
 \begin{equation}
-  \langle \xi_{*i}(t) \, \xi_{*i}(t') \rangle
+  \bigl\langle \xi_{*i}(t) \, \xi_{*i}(t') \bigr\rangle
   = G_i \, \delta(t - t')
@@
     \!\!
-    \langle
+    \bigl\langle
       \xi_{*i}(t) \xi_{*i}(t')
-    \rangle
+    \bigr\rangle
+    \,
     \dot \theta\bigl( q(t) \bigr)
     \dot \theta\bigl( q(t') \bigr)
-    \, dt dt'
+    \, dt \, dt'
   \notag \\
@@
   \sum_{i=1}^n \rho_i \int_0^T
-    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
+    \bigl\langle \xi_{*i}(t) \xi_{*i}(0) \bigr\rangle
+    \, dt
   .
@@
 %
-with Eq. \eqref{eq:alpha_invt} being
+and Eq. \eqref{eq:alpha_invt} is
 the special case of $t_0 = 0$.
@@
 of $\mathbf w$ with a set of
-eigenvectors, $\phi_{ki}$,
+eigenvectors, $\phi_{ki}$, as
 %
@@
 %
-satisfying the orthonormal conditions\cite{vankampen}:
+with the $\phi_{ki}$'s satisfying the orthonormal conditions\cite{vankampen}:
 %
@@
   \delta_{kl}
-  ,
+  .
   \label{eq:eig_orthonormal_rows}
@@
 %
-where we have started the index of the eigenmodes at $0$
+Here we have started the index of the eigenmodes at $0$
 instead of $1$ for later convenience.
@@
 the multiple-bin updating scheme,
-Eq. \eqref{eq:mbin_update} [using $u_i(t)$ instead of $v_i(t)$],
+Eq. \eqref{eq:mbin_update}
+[using $v_i(t)$ instead of $u_i(t)$],
 as
@@
   \!\!
-  \langle
+  \bigl\langle
     \tilde\xi_k(t)
     \tilde\xi_k(t')
-  \rangle
+  \bigr\rangle \,
   \theta_k\bigl(q(t)\bigr)
   \theta_k\bigl(q(t')\bigr)
-  \, dtdt'
+  \, dt \, dt'
   \notag \\
@@
 %
-We can rewrite Eq. \eqref{eq:error_asym} much like an action
+We can rewrite Eq. \eqref{eq:error_asym} much like the action
 of a particle whose position is given by $q(t)$:
@@
   }
-  \stackrel{\epsilon_k \to 0}{=\joinrel=\joinrel=}
+  %\stackrel{\epsilon_k \to 0}{=\joinrel=\joinrel=}
+  \xrightarrow{\epsilon_k \to 0}
   \frac{ a_0 }
@@
 %
-The equality in the last step is achievable
-if the initial bias, $\epsilon_k$, is negligible,
-%which is often true if no eigenvalue is close to $0$,
-and in this limit, the optimal initial updating magnitude
-is half of the value
-in the preliminary run.\note{This factor, $1/2$,
+If the initial bias, $\epsilon_k$, is negligible,
+the optimal initial updating magnitude
+will reach the ideal limit given by
+half of the updating magnitude in the preliminary run.
+
+\note{This factor, $1/2$,
   happens to coincide with the
@@
   under the Gaussian updating scheme:
-  for a mass function, $M(\bar q)$, that is roughly
-  inversely proportional to $\bar q$
+  for a mass function, $M(\bar q)$,
+  that is roughly proportional to $1/\bar q$
   (cf. Appendix \ref{sec:schedule_geometry}),
@@
 The first few members are\footnote{The
-  orthogonal polynomials share the same form
+  orthogonal polynomials share the same forms
   for different target distributions,
@@
 from its expectation.
-Because high-order moments are generally unreliable,
+As high-order moments are generally unreliable,
 this method works better with a small $K$ for a narrow distribution.
@@
 and the inverse-time schedule
-was recommended for the optimal convergence of $c_1$.\cite{pellegrini2014}
+has been recommended for the optimal convergence of $c_1$.\cite{pellegrini2014}
 
@@
 %
-In this expanded ensemble setup,
+Note that in this expanded ensemble setup,
 the updating equation for $c_0$ is no longer trivial,
-but corresponds to the WL way of constructing the bias potential
-for transitions among different intervals.
-%
-The WL algorithm is in fact the $K = 1$ case.
+but corresponds to a single-bin updating scheme of
+constructing the bias potential for transitions among different intervals.
+%
+The WL algorithm is in fact recovered by the $K = 1$ case.
 %
@@
 \end{equation}
-using a basis of the Hermite polynomials\cite{arfken}.
+using a basis of Hermite polynomials\cite{arfken}.
 %
@@
 %
-where $s(t)$ is the umbrella at time $t$.
+where $s(t)$ denotes the umbrella index at time $t$.
 %
@@
   %
-However, we need to redefine the measure of histogram fluctuation
+However, for this application,
+it is natural to redefine the measure of histogram fluctuation
 used in the WL strategy for stage transitions
@@
 Note that the LLR algorithm (cf. Sec. \ref{sec:LLR}) implicitly uses
-a similar relation for the estimation of $c_0^{(s)}$,
-with $\hat c_0^{(s)}$ replaced by $c_0^{(s)}$
-(for lack of $c_2^{(s)}$ in that case).
+a similar relation for the estimation of $c_0^{(s)}$
+(with $\hat c_0^{(s)}$ replaced by $c_0^{(s)}$
+for lack of $c_2^{(s)}$ in that case).
 
@@
   Perform a regular MC or MD step
-  in the current umbrella $s$ under
+  in the current umbrella, $s = s(t)$, under
   the bias potential $u^{(s)}(z, t)$.
@@
     is sometimes preferred,
-    in a random order,
     the sequential execution adopted here
@@
 The reduced density was $\varrho = 0.1$.
-The numbers of bins was $n = 500$.
+The number of bins was $n = 500$.
 The reduced temperature was $T = 3$.
@@
 For the Gaussian updating scheme,
-we used a set of fixed reference values of $\Gamma_k$'s\footnote{We
-  derived the reference $\Gamma_k$'s
+we used a set of fixed reference values of $\Gamma_k$'s\footnote{For
+  accuracy, we derived the reference $\Gamma_k$'s
   from a long equilibrium simulation under the reference bias 
@@
 $\Gamma_k$'s and $\tilde u_k$'s),
-there were considerable derivations
+there were considerable deviations
 in the initial errors.
@@
     Gaussian & $1/t$
-    & $6.3(6.2)\times10^{-3}$
-    & $1.5(1.4)\times10^{-5}$
+    & $6.2(6.2)\times10^{-3}$
+    & $1.46(1.44)\times10^{-5}$
     \\
@@
     Gaussian & optimal
-    & $6.3(6.2)\times10^{-3}$
-    & $2.7(2.7)\times10^{-6}$
+    & $6.4(6.2)\times10^{-3}$
+    & $2.72(2.74)\times10^{-6}$
     \\
@@
     Bandpass & $1/t$
-    & $8.4(8.0)\times10^{-3}$
-    & $1.7(1.7)\times10^{-6}$
+    & $8.3(8.0)\times10^{-3}$
+    & $1.69(1.67)\times10^{-6}$
     \\
@@
 %
-However, the gain of using the optimal schedule
+However, the gain in error reduction of using the optimal schedule
 diminished for shorter simulations,
@@
 For similar simulations performed at a higher density, $\varrho = 0.8$,
-we found the gain of using the above two updating schemes
-over the single-bin updating scheme, if any, to be negligible.
+we found that the above two updating schemes
+did not improve the convergence of the bias potential significantly
+compared to the single-bin updating scheme.
 %
@@
 We spread the Gaussian umbrellas
-of the same width $\sigma_E^{(s)} = L$ over $[-1.8L^2, -0.8L^2]$
+of the same width, $\sigma_E^{(s)} = L$, over $[-1.8L^2, -0.8L^2]$
 with an even spacing of $L$.
@@
 the single-bin updating scheme to be optimal
-for the error control in the long time limit.
+for error control in the long time limit.
 %without a priori assumptions
@@
 as a generalization of
-the WL and LLR\cite{langfeld2012, pellegrini2014} algorithm,
+the WL and LLR\cite{langfeld2012, pellegrini2014} algorithms,
 delegates the biased sampling to a local umbrella
@@
 For the Gaussian updating scheme,
-the calculation of the optimal schedule requires
-the estimation of a few parameters
-in a sufficiently long run
-under a constant updating magnitude,
-whereas the gain of using the optimal schedule
+the optimal schedule requires
+a long preliminary run under a constant updating magnitude,
+to extract a few parameters,
+while the gain in error reduction
 is sometimes rather modest
-for a relatively short simulation,
-even if the histogram fluctuation is largely
-local noise.
+for a relatively short simulation.
 %
@@
 To apply the above formalism to the updating schemes,
-we need a correction formula for the bias potential.
+we will need a correction formula for the bias potential.
 %
@@
 because it corresponds to the equally-weighted average,
-which is statistically most efficient.
+which is statistically most efficient, at least asymptotically.
 
@@
 from Eq. \eqref{eq:alpha_from_Omega}
-for the $k = 1$ mode with the largest eigenvalue,
+for the $k = 1$ mode, which has the largest eigenvalue,
 i.e.
@@
   \!\!\sum_{\tau=t_- + 1}^{t_+}
-  \frac{ \ln[1- \lambda_k \, \alpha(\tau)] } { t_+ - t_- }
+  \frac{ \ln[1- \lambda_1 \, \alpha(\tau)] } { t_+ - t_- }
   \\
   &\approx
-  -\frac{ \lambda_k \int_{t_-}^{t_+} \alpha(\tau) \, d\tau}{t_+ - t_-}
-  =
-  -\lambda_k \, \frac{q(t_+) - q(t_-)}{t_+ - t_-}
+  -\frac{ \lambda_1 \int_{t_-}^{t_+} \alpha(\tau) \, d\tau}{t_+ - t_-}
+  =
+  -\lambda_1 \, \frac{q(t_+) - q(t_-)}{t_+ - t_-}
   ,
@@
 
-\subsection{Histogram correction}
+\subsection{Histogram-based correction}
 
@@
 %
-Below we give a simple correction
-to recover the equally-weighted average.
-%
-We will use capital letters to denote equally-weighted averages.
-%e.g. for variable $y$,
-%$Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t y(\tau)$.
-%\begin{equation}
-%  Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
-%  .
-%  \notag
-%  %\label{eq:timeav}
-%\end{equation}
-%
-Then, by averaging both sides of Eq. \eqref{eq:uhatt}, we get
+We can, however, recover the equally-weighted average
+by averaging both sides of Eq. \eqref{eq:uhatt}:
+%
 \begin{align}
@@
   + \ln \frac{ H_i(t) } { \rho_i }
-  .
+  ,
   \label{eq:uhatav}
@@
 %
+where the capital letters denote the equally-weighted averages.
+%
 This correction serves as
-a generalization of Eq. \eqref{eq:vcorr_equil}
+an approximate generalization of Eq. \eqref{eq:vcorr_equil}
 for FDS simulations under
@@
 %
-Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
+Then, from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
 we find that for $\bar q(t) \equiv q(T) - q(t)$,
@@
   \int_{\bar q(t)}^{ q(T) }
-    m(\bar q) \, d \bar q
+    \!\! m(\bar q) \, d \bar q
   &=
@@
 %
-This allows each point on the schedule,
+The above two equations map each point on the schedule,
 $\bigl(t, \alpha(t)\bigr)$,
-to be mapped to a point,
+to a point,
 $\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
-on the mass distribution,
+on the mass distribution
 such that the ordinate of the latter curve
 is $1/(T\alpha)$,
-and the area under the curve in the domain, $[\bar q(t), q(T)]$,
+and the area under the curve between $\bar q(t)$ and $q(T)$,
 or the complementary cumulative distribution function of $m(\bar q)$,
@@
     (b) Inverse construction of $m(\bar q)$
-    from the optimal schedule $\alpha(t)$.
+    from the optimal schedule, $\alpha(t)$.
   }
@@
 equal initial updating magnitudes,
-the optimal schedule for the Gaussian updating scheme
+the optimal schedule for a Gaussian updating scheme
 always lies above the inverse-time schedule,
@@
 %
-For the Gaussian updating scheme,
+For a Gaussian updating scheme,
 we have $\bar \lambda(\bar q) < 1$ for $\bar q > 0$
@@
 Further, the transformation is an involution:
-if we apply the same transform on $\alpha^*(t^*)$
+if we apply the same transform on $\alpha^*(t^*)$,
 we would recover $m^*(\bar q^*)$,
@@
 $a_0 \ll 2/\left(\max_k \Gamma_k \right)$,
-and it tends to overestimate the value of $\Gamma_k$
-(possibly because of the deviation from
-the quasi-equilibrium sampling).
+and it tends to overestimate $\Gamma_k$
+(possibly because the sampling often
+deviates from the presumed quasi-equilibrium limit).
 
@@
 %
-By splitting the trajectory into blocks of $T_\mathrm{blk}$ steps,
+By breaking the trajectory into blocks of $T_\mathrm{blk}$ steps,
 we can compute the histogram fluctuation, $\tilde F_k(T_\mathrm{blk})$,
@@
 %
-The trajectory block should be large enough
+The trajectory block should be long enough
 to cover at least a few autocorrelation times (which is roughly $\Gamma_k/2$),
-but also sufficiently small
-when there are dynamic updates to the bias potential
+but not too long
+when dynamic updates to the bias potential exist
 to avoid the artificial histogram flattening
@@
 %
-with $g = 1$.
+with $g = 1$ here.
 Note also the two-fold degeneracy,
@@
     }
-    & l \ne 0, \\
-    \frac{2K-1}{n} & l = 0.
+    & \text{for $l \ne 0$}, \\
+    \frac{2K-1}{n}
+    & \text{for $l = 0$}.
   \end{dcases}
@@
   ,
-  \label{eq:F2sum}
+  \notag
+  %\label{eq:F2sum}
 \end{equation}

### 2017-10-20 doc/invt.tex
small changes in the proof

@@
   colorlinks,
-  linkcolor={red!30!black},
+  linkcolor={blue!80!black},
   citecolor={green!20!black},
@@
 \definecolor{DarkGreen}{RGB}{0,64,0}
-\definecolor{DarkPurple}{RGB}{64,0,42}
+\definecolor{DarkPurple}{RGB}{128,0,84}
 \definecolor{LightGray}{gray}{0.85}
@@
 \newcommand{\note}[1]{{\color{DarkGreen}\footnotesize \textsc{Note.} #1}}
+\newcommand{\notep}[1]{{\color{DarkPurple}\footnotesize \textsc{Note.} #1}}
 \newcommand{\answer}[1]{{\color{DarkBlue}\footnotesize \textsc{Answer.} #1}}
@@
 
-\title{Optimal updating magnitude in adaptive flat-distribution-sampling simulations}
+\title{Optimal updating magnitude in adaptive flat-distribution sampling}
+
+\notep{Deleted ``simulations'' to make the title shorter.}
 
@@
 \begin{abstract}
-  We present a method of computing the optimal schedule
+  We present a study on the optimization
   of the updating magnitude
@@
   including the Wang-Landau (WL) algorithm and metadynamics.
+  \notep{Deleted the ``optimal schedule'' because
+  we also studied the optimization of the updating scheme,
+  which is equivalent to the \emph{relative} updating magnitudes.}%
   %
@@
   The convergence of the bias potential can be improved
-  by decreasing the updating magnitude with an optimal schedule.
+  by a decreasing updating magnitude with an optimal schedule.
+  \notep{Fixed the dangling gerund, ``decreasing''.}%
   %
@@
   Constructed from orthogonal polynomials,
-  bandpass updating schemes generalize
+  the bandpass updating schemes generalize
   the WL and Langfeld-Lucini-Rago algorithms
@@
 Given a system,
-often the task is to compute
+we wish to compute
 a distribution, $p(z)$,
@@
 simulations.
+\notep{The word ``often'' has been used too often in this article.}%
 %
@@
 Thus,
-to overcome these energetic barriers in sampling,
-%and to capture the global shape of the free energy surface
-an effective strategy is to introduce
-a bias potential\cite{torrie1974, *torrie1977}
+we can overcome, at least partially, these energetic barriers in sampling
+by introducing a bias potential\cite{torrie1974, *torrie1977}
 that offsets the PMF
@@
 *laio2002, *laio2008, *barducci2011, *sutto2012}
-%
-\note{The flat-distribution sampling
-allows one to more accurately capture the
-global shape of the free energy surface
-by improving the sampling around the rarely visited barrier region
-and accelerating transitions among free energy basins.}
+\notep{The subject of the main clause should have been ``we'' in the previous version.}%
 
@@
 are based on adaptive construction of the bias potential,
-e.g. the Wang-Landau (WL) algorithm\cite{
+e.g., the Wang-Landau (WL) algorithm\cite{
   wang2001, wang2001pre}
@@
 which specifies,
-apart from on overall updating magnitude,
-the change to the bias potential upon each visit.
+apart from an overall updating magnitude,
+the relative change to the bias potential upon each visit.
 %
@@
 a polynomial form,
-and coefficients are adjusted on-the-fly
-to approximate the target distribution.
+and we will adjust the coefficients on-the-fly
+to better approximate the target distribution.
+\notep{Fixed the dangling infinitive, ``to approximate''.}%
 %
@@
 %
-We present the analytical results in Sec. \ref{sec:methods},
+We present the methods in Sec. \ref{sec:methods},
 provide numerical examples
@@
 as a biased simulation
-targeting a flat $\rho_i$ (i.e. $\rho_i \equiv 1/n$ for $n$ bins)
+targeting a flat $\rho_i$ (i.e., $\rho_i \equiv 1/n$ for $n$ bins)
 or a nearly flat one\cite{
@@
 %
-Upon convergence, the bias potential would satisfy
-%
+Such a bias potential would satisfy
+%
+\notep{We haven't defined the adaptive FDS, so we can't say ``upon convergence''.}
 \begin{equation}
@@
 and the PMF, $-\ln p_i$, can be readily
-deduced from $u_i$.
+deduced from the $u_i$.
 
@@
 To reduce the sampling error,
-the updating magnitude is often decreased gradually.\cite{
+one needs to gradually decrease the updating magnitude.\cite{
   belardinelli2007, *belardinelli2007jcp, *belardinelli2008, *belardinelli2016,
   zhou2005, morozov2007, morozov2009, zhou2008, min2007,
+  komura2012, *caparica2012, *caparica2014,
   poulain2006, liang2007,
   crespo2010, *atchade2011, *fort2015}
+\notep{Fixed the dangling infinitive, ``to reduce\dots''}
 
@@
 
-The bias potential following Eq. \eqref{eq:wl_update}
+Next, we observe that the bias potential following Eq. \eqref{eq:wl_update}
 generally grows over time.
@@
   $\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
-  i.e.
+  i.e.,
   $$
@@
   we find it convenient to shift the origin of time by $-1$,
-  e.g. the sum $\sum_{i=1}^T$ is mapped to the integral $\int_0^T dt$.}
+  e.g., the sum $\sum_{i=1}^T$ is mapped to the integral $\int_0^T dt$.}
 %
@@
 for the bin $i(t)$ at time $t$
-or zero otherwise,
+or $0$ otherwise,
 and we have defined
@@
 follows the same differential equation
-for the shift is a constant:
+since the shift is a constant:
 $\dot v_i(t) \approx \alpha(t) \, f_i(t)$.
@@
 would tend to $0$ at long times, reaching convergence,
-if $\alpha(t)$ decays sufficiently slowly,
-i.e. $q(T) \to \infty$ as $T \to \infty$.
+if $\alpha(t)$ decays sufficiently slowly
+such that $q(T) \to \infty$ as $T \to \infty$.
 
@@
     \bigl\langle
-      \xi_{*i}(t) \xi_{*i}(t')
+      \xi_{*i}(t) \, \xi_{*i}(t')
     \bigr\rangle
@@
   \approx
-  \sum_{i=1}^n \rho_i \int_0^T
+  \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty}
     \bigl\langle \xi_{*i}(t) \xi_{*i}(0) \bigr\rangle
@@
 It is convenient to
-vary the curve, $q(t)$ ($0 < t < T$),
+vary the curve, $q(0 \le t \le T)$,
 under a fixed value of $q(T)$,
@@
 will be considered later
-by adjusting the value of $q(T)$.
+in adjusting the value of $q(T)$.
 %
@@
 %
-By definition, Eq. \eqref{eq:mbin_update},
-a visit to bin $j = i(t)$ in this case results in updates of the bias potential
+By Eq. \eqref{eq:mbin_update},
+a visit to bin $j = i(t)$ in this case results in updates to the bias potential
 not only at bin $j$, but also at a few neighboring bins $i$'s.
@@
 $\pmb\rho = (\rho_1, \dots, \rho_n)$,
-the growth rate of the bias potential,
+we require the growth rate of the bias potential,
 ${\dot u}_i(t)$
 in Eq. \eqref{eq:ut_diffeq_mbin},
-should be a constant of $i$
+to be a constant of $i$
 if $h_j(t)$ were the same as $\rho_j$,
-i.e.
+\notep{Fixed the dangling-infinitive problem.}%
+i.e.,
 $\sum_{j=1}^n w_{ij}$ should be a constant
@@
 %
-and
+so that
 \begin{equation}
@@
   \bigl\langle
-    \tilde\xi_k(t)
+    \tilde\xi_k(t) \,
     \tilde\xi_k(t')
@@
 The value of $q(T)$ will be adjusted later
-to minimize the total error. % in Sec. \ref{sec:optinitalpha}.
+in minimizing the total error. % in Sec. \ref{sec:optinitalpha}.
+\notep{Fixed the dangling infinitive ``to minimize''.}
 %
@@
 the Hamiltonian is conserved,\cite{goldstein, *landau_mechanics, *arnold, *jose}
-i.e. the asymptotic error grows
+i.e., the asymptotic error grows
 linearly with time at a rate of
@@
   +
-  \epsilon_k \, e^{-2 \, \lambda_k \, a_0 \, T_0}
+  \epsilon_k
   ,
@@
   -\sum_{k = 1}^{n-1}
-  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
-  =0
+  \lambda_k \, \epsilon_k \, e^{-2 \, \lambda_k \, q(T)}
+  = 0
   .
@@
   +
-  \Delta_k
-  e^{-2\, \lambda_k a_0 T_0}
+  \epsilon_k
+  e^{-2\, \lambda_k \, a_0 \, T_0}
   ,
@@
 %
-where $\Delta_k \equiv \tilde u_k^2 - \frac{a_0 \lambda_k \Gamma_k}{2}$.
+where $\epsilon_k \equiv \tilde u_k^2 - \frac{a_0 \lambda_k \Gamma_k}{2}$.
 %
@@
   \sum_{k=1}^{n-1} \lambda_k \,
-  \Delta_k e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
+  \epsilon_k e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
   ,
@@
   -\sum_{k = 1}^{n-1}
-  \lambda_k \, \Delta_k \, e^{-2\lambda_k \, (q_T + a_0 T_0)}
+  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, (q_T + a_0 T_0)}
   =0
@@
   \frac{
-    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
+    \lambda_k \, \epsilon_k \, e^{-2 \, \lambda_k \, q(T)}
   }{
@@
   used a variant of the Newton-Raphson method
-  [see e.g. }]
+  [see e.g., }]
   [{] with the bisection method as a fallback.
@@
 %that satisfies $m(Q_L) > 2/(T\,a_0) > m(Q_R)$
-%(e.g. we can choose $Q_L = 0$,
+%(e.g., we can choose $Q_L = 0$,
 %  and repeatedly double $Q_R$ until
@@
   were linearly interpolated
-  to get the updating magnitude, $\alpha(t)$, at any $t$.
+  for the updating magnitude, $\alpha(t)$, at any $t$.
   }
@@
 generally has no simple closed-form expression.
-Nonetheless, we can show that it always lies above
+Nonetheless, we can show that it usually lies above
 the inverse-time schedule,
@@
 %We discuss this strategy in detail below.
-As shown below,
-the optimized updating scheme
-allows the optimal schedule to coincide with
-the inverse-time formula,
-and thus spare the need to
-go through the general procedure in Sec. \ref{sec:procedure}.
+%As shown below,
+%the optimized updating scheme
+%allows the optimal schedule to coincide with
+%the inverse-time formula,
+%and thus spare the need to
+%go through the general procedure in Sec. \ref{sec:procedure}.
 
@@
 $\Err_A(T) \ge \sum_{k=1}^{n-1} \Gamma_k/T$.
-%Note that the last expression %of Eq. \eqref{eq:CSineq}
-%depends on the curve, $q(t)$ ($0 < t < T$),
-%only through the endpoint value, $q(T)$,
-%which is fixed in the variational process.
 %
@@
 
+We can further construct an optimal \emph{time-dependent} updating scheme
+by a superposition of single-bin updating schemes.
+%
+Because each eigenmode evolves independently
+with an effective updating magnitude, $\lambda_k \, \alpha(t)$,
+as shown in Eq. \eqref{eq:vt_diffeq_mbin},
+the ideal error reduction requires
+a time-dependent eigenvalue, $\lambda_k(t)$,
+(hence a time-dependent updating scheme)
+that allows the effective updating magnitude
+to follow the eigenmode-specific optimal schedule,
+$\lambda_k(t) \, \alpha(t) = 1/(t + t_{0k})$,
+with $t_{0k} = \Gamma_k / \left\langle \tilde v_k^2(0) \right\rangle$,
+which is borrowed from Eq. \eqref{eq:alpha_invt1} for the single-bin case.
+%
+In long times, however, this time-dependent optimal solution
+is reduced to the single-bin updating scheme under the inverse-time schedule
+as the differences among the $t_{0k}$'s become unimportant to the schedule.
+
+\notep{This new paragraph gives a theoretically optimal updating scheme for a finite simulation.}
+
 \subsubsection{\label{sec:bandpass}
@@
 eigenmodes,
-we may assign the eigenvalues of the first $K$ eigenmodes to $1$
+we may set the eigenvalues of the first $K$ eigenmodes to $1$
 and the remaining eigenvalues to $0$:
+\notep{With ``assign'', we may need to say assign $1$ to the eigenvalues.}%
 %
@@
 from the eigenvalues by Eq. \eqref{eq:w_from_phi}.
-%(cf. Appendix \ref{sec:homo_bandpass} for the case of homogeneous updating schemes).
-%
-%Then, the updating matrix is, according to
-%Eq. \eqref{eq:w_from_phi},
-%%
-%$$w_{ij} = \frac{1}{\rho_i} \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj}.$$
 %
@@
 We will consider two special classes of bandpass updating schemes,
-homogeneous and polynomial bandpass updating schemes.
+the homogeneous and polynomial bandpass updating schemes.
 %
@@
 in an example.
-%
-%While both updating scheme are able to deliver
-%a smooth and differentiable potential of mean force,
-%the bandpass updating scheme is superior in asymptotic convergence,
-%because of the set of clear-cut spectrum of eigenvalues.
-%Eq. \eqref{eq:lambda_bandpass},
-%instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
-%
-%\footnote{As an estimate of error,
-%  we note that truncating a linear or quadratic function
-%  at mode $K$ for a non-periodic variable
-%  results in an error roughly proportional to
-%  $|\max_i u_i - \min_i u_i|^2/K^3$.}
 %
@@
 \subsection{\label{sec:bandpass_poly}
-Polynomial bandpass updating scheme}
+Polynomial bandpass updating schemes}
 
@@
 %
-The resulting polynomial bandpass updating schemes are more intuitive
-than the homogeneous bandpass updating scheme
-in that the former express the bias potential as
-a simple polynomial of the collective variable.
+The resulting polynomial bandpass updating schemes
+simplify the bias potential as a polynomial of the collective variable.
 
@@
 from Eq. \eqref{eq:w_from_phi}
-with the eigenvalues given by \eqref{eq:lambda_bandpass}:
+with the eigenvalues given by Eq. \eqref{eq:lambda_bandpass}:
 \begin{align*}
@@
 In practice,
-the above algorithm is often applied simultaneously
+we often apply the above algorithm simultaneously
 to an array of adjacent intervals
 to cover a wide range of the collective variable, $z$.\cite{langfeld2012, pellegrini2014}
-%
-Note that in this expanded ensemble setup,
+\notep{Fixed the dangling infinitive, ``to cover''.}%
+%
+Note that in this expanded ensemble setup of multiple intervals or multiple local umbrellas,
 the updating equation for $c_0$ is no longer trivial,
-but corresponds to a single-bin updating scheme of
-constructing the bias potential for transitions among different intervals.
+but corresponds to a single-bin updating scheme
+for the inter-umbrella component of the bias potential.
 %
@@
 In the expanded ensemble,
-inter-umbrella transitions can be handled via the frameworks of
+the inter-umbrella transitions can be handled via the frameworks of
 parallel\cite{swendsen1986,
@@
 for inter-umbrella transitions,
-and it is updated by Eq. \eqref{eq:ck_age_update},
-much like in the single-bin updating scheme.\cite{li2007}
+and its updating equation, Eq. \eqref{eq:ck_age_update},
+resembles that in the single-bin updating scheme.\cite{li2007}
 
@@
   slightly modified the transition criterion
-  from the WL to the inverse-time schedule.
+  from the WL stages to the inverse-time schedule.
   We allowed such a switch only
@@
 to include the intra-umbrella fluctuations.
-%
-%In the asymptotic regime, the optimal schedule
-%is simply given by the inverse-time schedule, Eq. \eqref{eq:alpha_invt1}.
-%%
-%In initial stages, when the inverse-time schedule
-%is still inapplicable,
-%we can instead use the WL style stage-wise strategy
-%to control the updating magnitude
-%for the bandpass updating scheme,
-%since it is a straightforward extension of the single-bin updating scheme.
-%%
-%However, the histogram fluctuation used for stage transitions
-%should be computed from only the first $K$ modes,
-%where the histogram flattening was performed
 
@@
   ,
-  %\label{eq:c0hat_limit}
+  \label{eq:c0hat_limit}
   \\
-  c_1^{(s)}
+  \frac{ c_1^{(s)} }{ \sigma_c^{(s)} }
   &\approx \left. \frac{d \ln p(z) } { d z } \right|_{z = z_c^{(s)}}
   ,
-  %\label{eq:c1_limit}
+  \label{eq:c1_limit}
   \\
-  \frac{
-    \sqrt 2 \, c_2^{(s)} - 1
-  } { \bigl( \sigma_c^{(s)} \bigr)^2 }
+  \frac{ \sqrt 2 \, c_2^{(s)} - 1 }{ \bigl( \sigma_c^{(s)} \bigr)^2 }
   &\approx \left. \frac{d^2 \ln p(z) } { d z^2 } \right|_{z = z_c^{(s)}}
   .
-  %\label{eq:c2_limit}
+  \label{eq:c2_limit}
 \end{align}
@@
 where the PMF is locally concave,
-i.e. $-d^2\ln p(z)/dz^2 < 0$.
-%
-%
-We also have approximately\cite{park2007, *nguyen2013, *zhang2015st}
+i.e., $-d^2\ln p(z)/dz^2 < 0$.
+%
+From Eqs. \eqref{eq:c0hat_limit} and \eqref{eq:c1_limit},
+we also have approximately\cite{park2007, *nguyen2013, *zhang2015st}
 %
@@
 a similar relation for the estimation of $c_0^{(s)}$
-(with $\hat c_0^{(s)}$ replaced by $c_0^{(s)}$
+(which is equivalent to $\hat c_0^{(s)}$ here
 for lack of $c_2^{(s)}$ in that case).
@@
   Set up an array of Gaussian umbrellas covering the desired range,
-    e.g. we may set, for $s = 1, \dots, S$,
+    e.g., we may set, for $s = 1, \dots, S$,
     the mean,
@@
   Update the $c_k^{(s)}$'s
-  according to Eq. \eqref{eq:ckupdate},
+  according to Eq. \eqref{eq:ck_age_update},
   and increase $t$ by $1$.
@@
 where $n^{(s)}(z)$
-the number of visits to the bin
+is the number of visits to the bin
 $\left(z-\frac{1}{2} \delta z, z + \frac{1}{2} \delta z\right)$
@@
 and the $N^{(s)}$ is the total number of visits
-to the umbrella $s$.
+to umbrella $s$.
 
@@
 The PMF is the negative logarithm of
-the intrinsic distribution of the distance
-between the two tagged particles, $p(r)$,
+the intrinsic distribution of the distance,
+%between the two tagged particles,
+$p(r)$,
 which is proportional to the cavity distribution function\cite{hansen}.
@@
 we found that the inverse-time schedule
-failed to reduce to the errors in the $k = 7$ to $21$ modes adequately.
+failed to reduce the errors in the $k = 7$ to $21$ modes adequately.
 %
@@
 diminished for shorter simulations,
-e.g. when the production phase was shortened to $T = 10^7$ steps,
+e.g., when the production phase was shortened to $T = 10^7$ steps,
 the final errors under the optimal and inverse-time schedules
@@
     Eq. \eqref{eq:uhatav}.
-    The lines are a guide to the eyes.
+    The lines are a guide for the eye.
   }
@@
 In this case, % $\Phi(\mathbf x) = 0$ and
-the intrinsic distribution is
+the intrinsic distribution is proportional to
 the density of states, $g(E)$.
@@
     %
-    (c) $c_1^{(s)}/\sigma_E$ (circles), roughly corresponding to
+    (c) $c_1^{(s)}/\sigma_E$ (circles), as an approximation of
     the statistical temperature, $\beta(E) = d\ln g(E)/dE$;
@@
 In conclusion,
-we have proposed a method for computing
-the optimal schedule of the updating magnitude
+we have studied the optimization
+of the updating magnitude
 for a general class of free energy calculation methods
@@
 %
-The optimal schedule can improve the convergence
+The optimal updating magnitude can improve the convergence
 of the bias potential,
@@
 
-The optimal schedule depends on the updating scheme,
+The optimal schedule of the overall updating magnitude depends on the updating scheme,
 which is mathematically characterized by an updating matrix,
@@
 as the asymptotically optimal schedule
-is given by an inverse-time formula.
+is given by the inverse-time formula.
 
@@
 to the umbrellas located at different places.
-%
-%with the coefficients adjusted on-the-fly.
 %
@@
 the optimal schedule requires
-a long preliminary run under a constant updating magnitude,
-to extract a few parameters,
+a long preliminary run for parameter estimation,
 while the gain in error reduction
@@
 to improve its general applicability and stability.
-%e.g. to prevent the updating scheme from
+%e.g., to prevent the updating scheme from
 %overly updating the parameters.
@@
 
-We thank Dr. Y. Mei, Dr. W. Tang, J. Wang,
-O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Stuhlsatz,
+We thank Dr. Y. Mei, Dr. W. Tang, J. Wang, Dr. M. W. Deem, Dr. K.-Y. Wong,
+O. Nassar, Dr. C. Lai, Dr. S. Ou, D. Seckfort, %D. Stuhlsatz,
 Dr. C. Myers, and Dr. O. Samoylova
@@
 %The inverse-time schedule, Eq. \eqref{eq:alpha_invt1},
-For example,
+In particular,
 the equally-weighted average, $\omega(t) = \omega(0)/t_0$ (for $t > 0$),
@@
 %
-Updating magnitudes above and below this value
-are typical in CMAs of
-growing and diminishing weights, respectively.
+%Updating magnitudes above and below this value
+%are typical in CMAs of
+%growing and diminishing weights, respectively.
 
@@
 we find that $u_i(t+1)$ follows the same time evolution
-as the CMA $X_\omega(t)$ of $x(t) = \hat u_i(t)$.
+as the CMA, $X_\omega(t)$, of $x(t) = \hat u_i(t)$.
 %
@@
 and from Eq. \eqref{eq:uhatt},
-$\hat{\tilde u}_k(t) \approx {\tilde u}_k(t) + {\tilde f}_k(t)$,
+$\tilde{\hat u}_k(t) \approx {\tilde u}_k(t) + {\tilde f}_k(t)$,
 for $k > 0$.
@@
 \begin{align*}
-  \tilde u_k(t+1) = \tilde u_k(t) + \lambda_k \, \alpha(t) \,
+  \tilde u_k(t+1) \approx \tilde u_k(t) + \lambda_k \, \alpha(t) \,
   \left[
-    \hat{\tilde u}_k(t) - \tilde u_k(t)
+    \tilde{\hat u}_k(t) - \tilde u_k(t)
   \right]
@@
 $\tilde u_k(t+1)$ is equivalent to
-a CMA of $\hat{\tilde u}_k(t)$,
+a CMA of $\tilde{\hat u}_k(t)$,
 and the weight is given by Eq. \eqref{eq:omega_from_alpha},
@@
 for the $k = 1$ mode, which has the largest eigenvalue,
-i.e.
+i.e.,
 $\alpha(t) = \lambda_1^{-1}[1-\Omega(t-1)/\Omega(t)]$,
@@
 %
-However, this method would fail for a near zero, $\lambda_k$,
+However, this method would fail for a near zero $\lambda_k$,
 and it is best used for the single-bin updating scheme,
@@
 when dynamic updates to the bias potential exist
-to avoid the artificial histogram flattening
-caused by the frequent non-equilibrium updates.
+so that the natural histogram flattening is not significantly accelerated
+by the frequent non-equilibrium updates.
+\notep{Fixed the dangling infinitive ``to avoid''.}
 %
@@
 and can be cutoff at $l = b$
-(i.e. $\mu_l = 0$ for $l > b$),
+(i.e., $\mu_l = 0$ for $l > b$),
 where $b$ is the integral part of $(n-1)/2$.
@@
 we define a periodic variable, $\phi_i$,
-with the out-of-boundary values given by
-$\phi_i = \phi_{i \pm n}$,
-%
-such that $i \pm n$ lies in between $1$ and $n$.
-%
-Then, the multiplication of the matrix, $\mathbf w$,
+%with the out-of-boundary values given by
+%$\phi_i = \phi_{i \pm n}$,
+%such that $i \pm n$ lies in between $1$ and $n$.
+%Then,
+so that
+the multiplication of the matrix, $\mathbf w$,
 is equivalent to a convolution with the window function:
@@
 Eq. \eqref{eq:wband_eigenvalue}
-with a doubled period, i.e. $g = 2$,
+with a doubled period, i.e., $g = 2$,
 and an increased cutoff $b = n - 1$
@@
 we should restrict the sum to the first $K$ modes,
-where the histogram flattening applies, i.e.
+where the histogram flattening applies, i.e.,
 $F^2(T) \equiv \sum_{k = 1}^{K-1} \tilde F_k^2(T)$.
